\bibbook{the third epistle of\\\bibtitlemain{john}}{iii. john}{III John}{3Jo}
\vs 3Jo 1:1 The elder unto the wellbeloved Gaius, whom I love in the truth.
\fn{in...: or, truly.}
\vs 3Jo 1:2 Beloved, I wish above all things that thou mayest prosper and be in health, even as thy soul prospereth.
\fn{wish: or, pray.}
\vs 3Jo 1:3 For I rejoiced greatly, when the brethren came and testified of the truth that is in thee, even as thou walkest in the truth.
\vs 3Jo 1:4 I have no greater joy than to hear that my children walk in truth.
\vs 3Jo 1:5 Beloved, thou doest faithfully whatsoever thou doest to the brethren, and to strangers;
\vs 3Jo 1:6 Which have borne witness of thy charity before the church: whom if thou bring forward on their journey after a godly sort, thou shalt do well:
\fn{after...: Gr. worthy of God.}
\vs 3Jo 1:7 Because that for his name's sake they went forth, taking nothing of the Gentiles.
\vs 3Jo 1:8 We therefore ought to receive such, that we might be fellowhelpers to the truth.
\fn{fellowhelpers: or, joint-labourers.}
\vs 3Jo 1:9 I wrote unto the church: but Diotrephes, who loveth to have the preeminence among them, receiveth us not.
\vs 3Jo 1:10 Wherefore, if I come, I will remember his deeds which he doeth, prating against us with malicious words: and not content therewith, neither doth he himself receive the brethren, and forbiddeth them that would, and casteth \bibemph{them} out of the church.
\vs 3Jo 1:11 Beloved, follow not that which is evil, but that which is good. He that doeth good is of God: but he that doeth evil hath not seen God.
\vs 3Jo 1:12 Demetrius hath good report of all \bibemph{men}, and of the truth itself: yea, and we \bibemph{also} bear record; and ye know that our record is true.
\vs 3Jo 1:13 I had many things to write, but I will not with ink and pen write unto thee:
\vs 3Jo 1:14 But I trust I shall shortly see thee, and we shall speak face to face. Peace \bibemph{be} to thee. \bibemph{Our} friends salute thee. Greet the friends by name.
\fn{face to...: Gr. mouth to mouth.}
