\bibbook{\bibtitlemain{ecclesiastes;}\\[3pt]\Large or, the preacher}{ecclesiastes}{Ecclesiastes}{Ecc}
\vs Ecc 1:1 The words of the Preacher, the son of David, king in Jerusalem.
\vs Ecc 1:2 Vanity of vanities, saith the Preacher, vanity of vanities; all \bibemph{is} vanity.
\vs Ecc 1:3 What profit hath a man of all his labour which he taketh under the sun?
\vs Ecc 1:4 \bibemph{One} generation passeth away, and \bibemph{another} generation cometh: but the earth abideth for ever.
\vs Ecc 1:5 The sun also ariseth, and the sun goeth down, and hasteth to his place where he arose.
\fn{hasteth: Heb. panteth.}
\vs Ecc 1:6 The wind goeth toward the south, and turneth about unto the north; it whirleth about continually, and the wind returneth again according to his circuits.
\vs Ecc 1:7 All the rivers run into the sea; yet the sea \bibemph{is} not full; unto the place from whence the rivers come, thither they return again.
\fn{return...: Heb. return to go.}
\vs Ecc 1:8 All things \bibemph{are} full of labour; man cannot utter \bibemph{it}: the eye is not satisfied with seeing, nor the ear filled with hearing.
\vs Ecc 1:9 The thing that hath been, it \bibemph{is that} which shall be; and that which is done \bibemph{is} that which shall be done: and \bibemph{there is} no new \bibemph{thing} under the sun.
\vs Ecc 1:10 Is there \bibemph{any} thing whereof it may be said, See, this \bibemph{is} new? it hath been already of old time, which was before us.
\vs Ecc 1:11 \bibemph{There is} no remembrance of former \bibemph{things}; neither shall there be \bibemph{any} remembrance of \bibemph{things} that are to come with \bibemph{those} that shall come after.
\vs Ecc 1:12 I the Preacher was king over Israel in Jerusalem.
\vs Ecc 1:13 And I gave my heart to seek and search out by wisdom concerning all \bibemph{things} that are done under heaven: this sore travail hath God given to the sons of man to be exercised therewith.
\fn{to be...: or, to afflict them.}
\vs Ecc 1:14 I have seen all the works that are done under the sun; and, behold, all \bibemph{is} vanity and vexation of spirit.
\vs Ecc 1:15 \bibemph{That which is} crooked cannot be made straight: and that which is wanting cannot be numbered.
\fn{that which is wanting: Heb. defect.}
\vs Ecc 1:16 I communed with mine own heart, saying, Lo, I am come to great estate, and have gotten more wisdom than all \bibemph{they} that have been before me in Jerusalem: yea, my heart had great experience of wisdom and knowledge.
\fn{had...: Heb. had seen much.}
\vs Ecc 1:17 And I gave my heart to know wisdom, and to know madness and folly: I perceived that this also is vexation of spirit.
\vs Ecc 1:18 For in much wisdom \bibemph{is} much grief: and he that increaseth knowledge increaseth sorrow.
\vs Ecc 2:1 I said in mine heart, Go to now, I will prove thee with mirth, therefore enjoy pleasure: and, behold, this also \bibemph{is} vanity.
\vs Ecc 2:2 I said of laughter, \bibemph{It is} mad: and of mirth, What doeth it?
\vs Ecc 2:3 I sought in mine heart to give myself unto wine, yet acquainting mine heart with wisdom; and to lay hold on folly, till I might see what \bibemph{was} that good for the sons of men, which they should do under the heaven all the days of their life.
\fn{to give...: Heb. to draw my flesh with wine. all...: Heb. the number of the days of their life.}
\vs Ecc 2:4 I made me great works; I builded me houses; I planted me vineyards:
\vs Ecc 2:5 I made me gardens and orchards, and I planted trees in them of all \bibemph{kind of} fruits:
\vs Ecc 2:6 I made me pools of water, to water therewith the wood that bringeth forth trees:
\vs Ecc 2:7 I got \bibemph{me} servants and maidens, and had servants born in my house; also I had great possessions of great and small cattle above all that were in Jerusalem before me:
\fn{servants born...: Heb. sons of my house.}
\vs Ecc 2:8 I gathered me also silver and gold, and the peculiar treasure of kings and of the provinces: I gat me men singers and women singers, and the delights of the sons of men, \bibemph{as} musical instruments, and that of all sorts.
\fn{musical...: Heb. musical instrument and instruments.}
\vs Ecc 2:9 So I was great, and increased more than all that were before me in Jerusalem: also my wisdom remained with me.
\vs Ecc 2:10 And whatsoever mine eyes desired I kept not from them, I withheld not my heart from any joy; for my heart rejoiced in all my labour: and this was my portion of all my labour.
\vs Ecc 2:11 Then I looked on all the works that my hands had wrought, and on the labour that I had laboured to do: and, behold, all \bibemph{was} vanity and vexation of spirit, and \bibemph{there was} no profit under the sun.
\vs Ecc 2:12 And I turned myself to behold wisdom, and madness, and folly: for what \bibemph{can} the man \bibemph{do} that cometh after the king? \bibemph{even} that which hath been already done.
\fn{even...: or, in those things which have been already done.}
\vs Ecc 2:13 Then I saw that wisdom excelleth folly, as far as light excelleth darkness.
\fn{that...: Heb. that there is an excellency in wisdom more than in folly, etc.}
\vs Ecc 2:14 The wise man's eyes \bibemph{are} in his head; but the fool walketh in darkness: and I myself perceived also that one event happeneth to them all.
\vs Ecc 2:15 Then said I in my heart, As it happeneth to the fool, so it happeneth even to me; and why was I then more wise? Then I said in my heart, that this also \bibemph{is} vanity.
\fn{happeneth even...: Heb. happeneth to me, even to me.}
\vs Ecc 2:16 For \bibemph{there is} no remembrance of the wise more than of the fool for ever; seeing that which now \bibemph{is} in the days to come shall all be forgotten. And how dieth the wise \bibemph{man}? as the fool.
\vs Ecc 2:17 Therefore I hated life; because the work that is wrought under the sun \bibemph{is} grievous unto me: for all \bibemph{is} vanity and vexation of spirit.
\vs Ecc 2:18 Yea, I hated all my labour which I had taken under the sun: because I should leave it unto the man that shall be after me.
\fn{taken: Heb. laboured.}
\vs Ecc 2:19 And who knoweth whether he shall be a wise \bibemph{man} or a fool? yet shall he have rule over all my labour wherein I have laboured, and wherein I have shewed myself wise under the sun. This \bibemph{is} also vanity.
\vs Ecc 2:20 Therefore I went about to cause my heart to despair of all the labour which I took under the sun.
\vs Ecc 2:21 For there is a man whose labour \bibemph{is} in wisdom, and in knowledge, and in equity; yet to a man that hath not laboured therein shall he leave it \bibemph{for} his portion. This also \bibemph{is} vanity and a great evil.
\fn{leave: Heb. give.}
\vs Ecc 2:22 For what hath man of all his labour, and of the vexation of his heart, wherein he hath laboured under the sun?
\vs Ecc 2:23 For all his days \bibemph{are} sorrows, and his travail grief; yea, his heart taketh not rest in the night. This is also vanity.
\vs Ecc 2:24 \bibemph{There is} nothing better for a man, \bibemph{than} that he should eat and drink, and \bibemph{that} he should make his soul enjoy good in his labour. This also I saw, that it \bibemph{was} from the hand of God.
\fn{should make...: or, delight his senses.}
\vs Ecc 2:25 For who can eat, or who else can hasten \bibemph{hereunto}, more than I?
\vs Ecc 2:26 For \bibemph{God} giveth to a man that \bibemph{is} good in his sight wisdom, and knowledge, and joy: but to the sinner he giveth travail, to gather and to heap up, that he may give to \bibemph{him that is} good before God. This also \bibemph{is} vanity and vexation of spirit.
\fn{in his...: Heb. before him.}
\vs Ecc 3:1 To every \bibemph{thing there is} a season, and a time to every purpose under the heaven:
\vs Ecc 3:2 A time to be born, and a time to die; a time to plant, and a time to pluck up \bibemph{that which is} planted;
\fn{to be...: Heb. to bear.}
\vs Ecc 3:3 A time to kill, and a time to heal; a time to break down, and a time to build up;
\vs Ecc 3:4 A time to weep, and a time to laugh; a time to mourn, and a time to dance;
\vs Ecc 3:5 A time to cast away stones, and a time to gather stones together; a time to embrace, and a time to refrain from embracing;
\fn{to refrain from: Heb. to be far from.}
\vs Ecc 3:6 A time to get, and a time to lose; a time to keep, and a time to cast away;
\fn{get: or, seek.}
\vs Ecc 3:7 A time to rend, and a time to sew; a time to keep silence, and a time to speak;
\vs Ecc 3:8 A time to love, and a time to hate; a time of war, and a time of peace.
\vs Ecc 3:9 What profit hath he that worketh in that wherein he laboureth?
\vs Ecc 3:10 I have seen the travail, which God hath given to the sons of men to be exercised in it.
\vs Ecc 3:11 He hath made every \bibemph{thing} beautiful in his time: also he hath set the world in their heart, so that no man can find out the work that God maketh from the beginning to the end.
\vs Ecc 3:12 I know that \bibemph{there is} no good in them, but for \bibemph{a man} to rejoice, and to do good in his life.
\vs Ecc 3:13 And also that every man should eat and drink, and enjoy the good of all his labour, it \bibemph{is} the gift of God.
\vs Ecc 3:14 I know that, whatsoever God doeth, it shall be for ever: nothing can be put to it, nor any thing taken from it: and God doeth \bibemph{it}, that \bibemph{men} should fear before him.
\vs Ecc 3:15 That which hath been is now; and that which is to be hath already been; and God requireth that which is past.
\fn{that which is past: Heb. that which is driven away.}
\vs Ecc 3:16 And moreover I saw under the sun the place of judgment, \bibemph{that} wickedness \bibemph{was} there; and the place of righteousness, \bibemph{that} iniquity \bibemph{was} there.
\vs Ecc 3:17 I said in mine heart, God shall judge the righteous and the wicked: for \bibemph{there is} a time there for every purpose and for every work.
\vs Ecc 3:18 I said in mine heart concerning the estate of the sons of men, that God might manifest them, and that they might see that they themselves are beasts.
\fn{that God...: or, that they might clear God, and see, etc.}
\vs Ecc 3:19 For that which befalleth the sons of men befalleth beasts; even one thing befalleth them: as the one dieth, so dieth the other; yea, they have all one breath; so that a man hath no preeminence above a beast: for all \bibemph{is} vanity.
\vs Ecc 3:20 All go unto one place; all are of the dust, and all turn to dust again.
\vs Ecc 3:21 Who knoweth the spirit of man that goeth upward, and the spirit of the beast that goeth downward to the earth?
\fn{of man: Heb. of the sons of man. goeth upward: Heb. is ascending, etc.}
\vs Ecc 3:22 Wherefore I perceive that \bibemph{there is} nothing better, than that a man should rejoice in his own works; for that \bibemph{is} his portion: for who shall bring him to see what shall be after him?
\vs Ecc 4:1 So I returned, and considered all the oppressions that are done under the sun: and behold the tears of \bibemph{such as were} oppressed, and they had no comforter; and on the side of their oppressors \bibemph{there was} power; but they had no comforter.
\fn{side: Heb. hand.}
\vs Ecc 4:2 Wherefore I praised the dead which are already dead more than the living which are yet alive.
\vs Ecc 4:3 Yea, better \bibemph{is he} than both they, which hath not yet been, who hath not seen the evil work that is done under the sun.
\vs Ecc 4:4 Again, I considered all travail, and every right work, that for this a man is envied of his neighbour. This \bibemph{is} also vanity and vexation of spirit.
\fn{every...: Heb. all the rightness of work. for this...: Heb. this is the envy of a man from his neighbour.}
\vs Ecc 4:5 The fool foldeth his hands together, and eateth his own flesh.
\vs Ecc 4:6 Better \bibemph{is} an handful \bibemph{with} quietness, than both the hands full \bibemph{with} travail and vexation of spirit.
\vs Ecc 4:7 Then I returned, and I saw vanity under the sun.
\vs Ecc 4:8 There is one \bibemph{alone}, and \bibemph{there is} not a second; yea, he hath neither child nor brother: yet \bibemph{is there} no end of all his labour; neither is his eye satisfied with riches; neither \bibemph{saith he}, For whom do I labour, and bereave my soul of good? This \bibemph{is} also vanity, yea, it \bibemph{is} a sore travail.
\vs Ecc 4:9 Two \bibemph{are} better than one; because they have a good reward for their labour.
\vs Ecc 4:10 For if they fall, the one will lift up his fellow: but woe to him \bibemph{that is} alone when he falleth; for \bibemph{he hath} not another to help him up.
\vs Ecc 4:11 Again, if two lie together, then they have heat: but how can one be warm \bibemph{alone}?
\vs Ecc 4:12 And if one prevail against him, two shall withstand him; and a threefold cord is not quickly broken.
\vs Ecc 4:13 Better \bibemph{is} a poor and a wise child than an old and foolish king, who will no more be admonished.
\fn{who...: Heb. who knoweth not to be admonished.}
\vs Ecc 4:14 For out of prison he cometh to reign; whereas also \bibemph{he that is} born in his kingdom becometh poor.
\vs Ecc 4:15 I considered all the living which walk under the sun, with the second child that shall stand up in his stead.
\vs Ecc 4:16 \bibemph{There is} no end of all the people, \bibemph{even} of all that have been before them: they also that come after shall not rejoice in him. Surely this also \bibemph{is} vanity and vexation of spirit.
\vs Ecc 5:1 Keep thy foot when thou goest to the house of God, and be more ready to hear, than to give the sacrifice of fools: for they consider not that they do evil.
\vs Ecc 5:2 Be not rash with thy mouth, and let not thine heart be hasty to utter \bibemph{any} thing before God: for God \bibemph{is} in heaven, and thou upon earth: therefore let thy words be few.
\fn{thing: or, word.}
\vs Ecc 5:3 For a dream cometh through the multitude of business; and a fool's voice \bibemph{is known} by multitude of words.
\vs Ecc 5:4 When thou vowest a vow unto God, defer not to pay it; for \bibemph{he hath} no pleasure in fools: pay that which thou hast vowed.
\vs Ecc 5:5 Better \bibemph{is it} that thou shouldest not vow, than that thou shouldest vow and not pay.
\vs Ecc 5:6 Suffer not thy mouth to cause thy flesh to sin; neither say thou before the angel, that it \bibemph{was} an error: wherefore should God be angry at thy voice, and destroy the work of thine hands?
\vs Ecc 5:7 For in the multitude of dreams and many words \bibemph{there are} also \bibemph{divers} vanities: but fear thou God.
\vs Ecc 5:8 If thou seest the oppression of the poor, and violent perverting of judgment and justice in a province, marvel not at the matter: for \bibemph{he that is} higher than the highest regardeth; and \bibemph{there be} higher than they.
\fn{at the... Heb. at the will, or, purpose.}
\vs Ecc 5:9 Moreover the profit of the earth is for all: the king \bibemph{himself} is served by the field.
\vs Ecc 5:10 He that loveth silver shall not be satisfied with silver; nor he that loveth abundance with increase: this \bibemph{is} also vanity.
\vs Ecc 5:11 When goods increase, they are increased that eat them: and what good \bibemph{is there} to the owners thereof, saving the beholding \bibemph{of them} with their eyes?
\vs Ecc 5:12 The sleep of a labouring man \bibemph{is} sweet, whether he eat little or much: but the abundance of the rich will not suffer him to sleep.
\vs Ecc 5:13 There is a sore evil \bibemph{which} I have seen under the sun, \bibemph{namely}, riches kept for the owners thereof to their hurt.
\vs Ecc 5:14 But those riches perish by evil travail: and he begetteth a son, and \bibemph{there is} nothing in his hand.
\vs Ecc 5:15 As he came forth of his mother's womb, naked shall he return to go as he came, and shall take nothing of his labour, which he may carry away in his hand.
\vs Ecc 5:16 And this also \bibemph{is} a sore evil, \bibemph{that} in all points as he came, so shall he go: and what profit hath he that hath laboured for the wind?
\vs Ecc 5:17 All his days also he eateth in darkness, and \bibemph{he hath} much sorrow and wrath with his sickness.
\vs Ecc 5:18 Behold \bibemph{that} which I have seen: \bibemph{it is} good and comely \bibemph{for one} to eat and to drink, and to enjoy the good of all his labour that he taketh under the sun all the days of his life, which God giveth him: for it \bibemph{is} his portion.
\fn{It is good...: Heb. there is a good which is comely, etc. all the days: Heb. the number of the days.}
\vs Ecc 5:19 Every man also to whom God hath given riches and wealth, and hath given him power to eat thereof, and to take his portion, and to rejoice in his labour; this \bibemph{is} the gift of God.
\vs Ecc 5:20 For he shall not much remember the days of his life; because God answereth \bibemph{him} in the joy of his heart.
\fn{For...: or, Though he give not much, yet he remembereth, etc.}
\vs Ecc 6:1 There is an evil which I have seen under the sun, and it \bibemph{is} common among men:
\vs Ecc 6:2 A man to whom God hath given riches, wealth, and honour, so that he wanteth nothing for his soul of all that he desireth, yet God giveth him not power to eat thereof, but a stranger eateth it: this \bibemph{is} vanity, and it \bibemph{is} an evil disease.
\vs Ecc 6:3 If a man beget an hundred \bibemph{children}, and live many years, so that the days of his years be many, and his soul be not filled with good, and also \bibemph{that} he have no burial; I say, \bibemph{that} an untimely birth \bibemph{is} better than he.
\vs Ecc 6:4 For he cometh in with vanity, and departeth in darkness, and his name shall be covered with darkness.
\vs Ecc 6:5 Moreover he hath not seen the sun, nor known \bibemph{any thing}: this hath more rest than the other.
\vs Ecc 6:6 Yea, though he live a thousand years twice \bibemph{told}, yet hath he seen no good: do not all go to one place?
\vs Ecc 6:7 All the labour of man \bibemph{is} for his mouth, and yet the appetite is not filled.
\fn{appetite: Heb. soul.}
\vs Ecc 6:8 For what hath the wise more than the fool? what hath the poor, that knoweth to walk before the living?
\vs Ecc 6:9 Better \bibemph{is} the sight of the eyes than the wandering of the desire: this \bibemph{is} also vanity and vexation of spirit.
\fn{than...: Heb. than the walking of the soul.}
\vs Ecc 6:10 That which hath been is named already, and it is known that it \bibemph{is} man: neither may he contend with him that is mightier than he.
\vs Ecc 6:11 Seeing there be many things that increase vanity, what \bibemph{is} man the better?
\vs Ecc 6:12 For who knoweth what \bibemph{is} good for man in \bibemph{this} life, all the days of his vain life which he spendeth as a shadow? for who can tell a man what shall be after him under the sun?
\fn{all...: Heb. the number of the days of the life of his vanity.}
\vs Ecc 7:1 A good name \bibemph{is} better than precious ointment; and the day of death than the day of one's birth.
\vs Ecc 7:2 \bibemph{It is} better to go to the house of mourning, than to go to the house of feasting: for that \bibemph{is} the end of all men; and the living will lay \bibemph{it} to his heart.
\vs Ecc 7:3 Sorrow \bibemph{is} better than laughter: for by the sadness of the countenance the heart is made better.
\fn{Sorrow: or, Anger.}
\vs Ecc 7:4 The heart of the wise \bibemph{is} in the house of mourning; but the heart of fools \bibemph{is} in the house of mirth.
\vs Ecc 7:5 \bibemph{It is} better to hear the rebuke of the wise, than for a man to hear the song of fools.
\vs Ecc 7:6 For as the crackling of thorns under a pot, so \bibemph{is} the laughter of the fool: this also \bibemph{is} vanity.
\fn{crackling: Heb. sound.}
\vs Ecc 7:7 Surely oppression maketh a wise man mad; and a gift destroyeth the heart.
\vs Ecc 7:8 Better \bibemph{is} the end of a thing than the beginning thereof: \bibemph{and} the patient in spirit \bibemph{is} better than the proud in spirit.
\vs Ecc 7:9 Be not hasty in thy spirit to be angry: for anger resteth in the bosom of fools.
\vs Ecc 7:10 Say not thou, What is \bibemph{the cause} that the former days were better than these? for thou dost not enquire wisely concerning this.
\fn{wisely: Heb. out of wisdom.}
\vs Ecc 7:11 Wisdom \bibemph{is} good with an inheritance: and \bibemph{by it there is} profit to them that see the sun.
\fn{good...: or, as good as an inheritance, yea, better too.}
\vs Ecc 7:12 For wisdom \bibemph{is} a defence, \bibemph{and} money \bibemph{is} a defence: but the excellency of knowledge \bibemph{is, that} wisdom giveth life to them that have it.
\fn{defence: Heb. shadow.}
\vs Ecc 7:13 Consider the work of God: for who can make \bibemph{that} straight, which he hath made crooked?
\vs Ecc 7:14 In the day of prosperity be joyful, but in the day of adversity consider: God also hath set the one over against the other, to the end that man should find nothing after him.
\fn{set: Heb. made.}
\vs Ecc 7:15 All \bibemph{things} have I seen in the days of my vanity: there is a just \bibemph{man} that perisheth in his righteousness, and there is a wicked \bibemph{man} that prolongeth \bibemph{his life} in his wickedness.
\vs Ecc 7:16 Be not righteous over much; neither make thyself over wise: why shouldest thou destroy thyself?
\fn{destroy...: Heb. be desolate?.}
\vs Ecc 7:17 Be not over much wicked, neither be thou foolish: why shouldest thou die before thy time?
\fn{before...: Heb. not in thy time?.}
\vs Ecc 7:18 \bibemph{It is} good that thou shouldest take hold of this; yea, also from this withdraw not thine hand: for he that feareth God shall come forth of them all.
\vs Ecc 7:19 Wisdom strengtheneth the wise more than ten mighty \bibemph{men} which are in the city.
\vs Ecc 7:20 For \bibemph{there is} not a just man upon earth, that doeth good, and sinneth not.
\vs Ecc 7:21 Also take no heed unto all words that are spoken; lest thou hear thy servant curse thee:
\fn{take...: Heb. give not thine heart.}
\vs Ecc 7:22 For oftentimes also thine own heart knoweth that thou thyself likewise hast cursed others.
\vs Ecc 7:23 All this have I proved by wisdom: I said, I will be wise; but it \bibemph{was} far from me.
\vs Ecc 7:24 That which is far off, and exceeding deep, who can find it out?
\vs Ecc 7:25 I applied mine heart to know, and to search, and to seek out wisdom, and the reason \bibemph{of things}, and to know the wickedness of folly, even of foolishness \bibemph{and} madness:
\fn{I applied...: Heb. I and mine heart compassed.}
\vs Ecc 7:26 And I find more bitter than death the woman, whose heart \bibemph{is} snares and nets, \bibemph{and} her hands \bibemph{as} bands: whoso pleaseth God shall escape from her; but the sinner shall be taken by her.
\fn{whoso...: Heb. he that is good before God.}
\vs Ecc 7:27 Behold, this have I found, saith the preacher, \bibemph{counting} one by one, to find out the account:
\fn{counting...: or, weighing one thing after another, to find out the reason.}
\vs Ecc 7:28 Which yet my soul seeketh, but I find not: one man among a thousand have I found; but a woman among all those have I not found.
\vs Ecc 7:29 Lo, this only have I found, that God hath made man upright; but they have sought out many inventions.
\vs Ecc 8:1 Who \bibemph{is} as the wise \bibemph{man}? and who knoweth the interpretation of a thing? a man's wisdom maketh his face to shine, and the boldness of his face shall be changed.
\fn{the boldness: Heb. the strength.}
\vs Ecc 8:2 I \bibemph{counsel thee} to keep the king's commandment, and \bibemph{that} in regard of the oath of God.
\vs Ecc 8:3 Be not hasty to go out of his sight: stand not in an evil thing; for he doeth whatsoever pleaseth him.
\vs Ecc 8:4 Where the word of a king \bibemph{is, there is} power: and who may say unto him, What doest thou?
\vs Ecc 8:5 Whoso keepeth the commandment shall feel no evil thing: and a wise man's heart discerneth both time and judgment.
\fn{shall...: Heb. shall know.}
\vs Ecc 8:6 Because to every purpose there is time and judgment, therefore the misery of man \bibemph{is} great upon him.
\vs Ecc 8:7 For he knoweth not that which shall be: for who can tell him when it shall be?
\fn{when...: or, how it shall be?.}
\vs Ecc 8:8 \bibemph{There is} no man that hath power over the spirit to retain the spirit; neither \bibemph{hath he} power in the day of death: and \bibemph{there is} no discharge in \bibemph{that} war; neither shall wickedness deliver those that are given to it.
\fn{discharge: or, casting off weapons.}
\vs Ecc 8:9 All this have I seen, and applied my heart unto every work that is done under the sun: \bibemph{there is} a time wherein one man ruleth over another to his own hurt.
\vs Ecc 8:10 And so I saw the wicked buried, who had come and gone from the place of the holy, and they were forgotten in the city where they had so done: this \bibemph{is} also vanity.
\vs Ecc 8:11 Because sentence against an evil work is not executed speedily, therefore the heart of the sons of men is fully set in them to do evil.
\vs Ecc 8:12 Though a sinner do evil an hundred times, and his \bibemph{days} be prolonged, yet surely I know that it shall be well with them that fear God, which fear before him:
\vs Ecc 8:13 But it shall not be well with the wicked, neither shall he prolong \bibemph{his} days, \bibemph{which are} as a shadow; because he feareth not before God.
\vs Ecc 8:14 There is a vanity which is done upon the earth; that there be just \bibemph{men}, unto whom it happeneth according to the work of the wicked; again, there be wicked \bibemph{men}, to whom it happeneth according to the work of the righteous: I said that this also \bibemph{is} vanity.
\vs Ecc 8:15 Then I commended mirth, because a man hath no better thing under the sun, than to eat, and to drink, and to be merry: for that shall abide with him of his labour the days of his life, which God giveth him under the sun.
\vs Ecc 8:16 When I applied mine heart to know wisdom, and to see the business that is done upon the earth: (for also \bibemph{there is that} neither day nor night seeth sleep with his eyes:)
\vs Ecc 8:17 Then I beheld all the work of God, that a man cannot find out the work that is done under the sun: because though a man labour to seek \bibemph{it} out, yet he shall not find \bibemph{it}; yea further; though a wise \bibemph{man} think to know \bibemph{it}, yet shall he not be able to find \bibemph{it}.
\vs Ecc 9:1 For all this I considered in my heart even to declare all this, that the righteous, and the wise, and their works, \bibemph{are} in the hand of God: no man knoweth either love or hatred \bibemph{by} all \bibemph{that is} before them.
\fn{I considered...: Heb. I gave, or, set to my heart.}
\vs Ecc 9:2 All \bibemph{things come} alike to all: \bibemph{there is} one event to the righteous, and to the wicked; to the good and to the clean, and to the unclean; to him that sacrificeth, and to him that sacrificeth not: as \bibemph{is} the good, so \bibemph{is} the sinner; \bibemph{and} he that sweareth, as \bibemph{he} that feareth an oath.
\vs Ecc 9:3 This \bibemph{is} an evil among all \bibemph{things} that are done under the sun, that \bibemph{there is} one event unto all: yea, also the heart of the sons of men is full of evil, and madness \bibemph{is} in their heart while they live, and after that \bibemph{they go} to the dead.
\vs Ecc 9:4 For to him that is joined to all the living there is hope: for a living dog is better than a dead lion.
\vs Ecc 9:5 For the living know that they shall die: but the dead know not any thing, neither have they any more a reward; for the memory of them is forgotten.
\vs Ecc 9:6 Also their love, and their hatred, and their envy, is now perished; neither have they any more a portion for ever in any \bibemph{thing} that is done under the sun.
\vs Ecc 9:7 Go thy way, eat thy bread with joy, and drink thy wine with a merry heart; for God now accepteth thy works.
\vs Ecc 9:8 Let thy garments be always white; and let thy head lack no ointment.
\vs Ecc 9:9 Live joyfully with the wife whom thou lovest all the days of the life of thy vanity, which he hath given thee under the sun, all the days of thy vanity: for that \bibemph{is} thy portion in \bibemph{this} life, and in thy labour which thou takest under the sun.
\fn{Live...: Heb. See, or, Enjoy life.}
\vs Ecc 9:10 Whatsoever thy hand findeth to do, do \bibemph{it} with thy might; for \bibemph{there is} no work, nor device, nor knowledge, nor wisdom, in the grave, whither thou goest.
\vs Ecc 9:11 I returned, and saw under the sun, that the race \bibemph{is} not to the swift, nor the battle to the strong, neither yet bread to the wise, nor yet riches to men of understanding, nor yet favour to men of skill; but time and chance happeneth to them all.
\vs Ecc 9:12 For man also knoweth not his time: as the fishes that are taken in an evil net, and as the birds that are caught in the snare; so \bibemph{are} the sons of men snared in an evil time, when it falleth suddenly upon them.
\vs Ecc 9:13 This wisdom have I seen also under the sun, and it \bibemph{seemed} great unto me:
\vs Ecc 9:14 \bibemph{There was} a little city, and few men within it; and there came a great king against it, and besieged it, and built great bulwarks against it:
\vs Ecc 9:15 Now there was found in it a poor wise man, and he by his wisdom delivered the city; yet no man remembered that same poor man.
\vs Ecc 9:16 Then said I, Wisdom \bibemph{is} better than strength: nevertheless the poor man's wisdom \bibemph{is} despised, and his words are not heard.
\vs Ecc 9:17 The words of wise \bibemph{men are} heard in quiet more than the cry of him that ruleth among fools.
\vs Ecc 9:18 Wisdom \bibemph{is} better than weapons of war: but one sinner destroyeth much good.
\vs Ecc 10:1 Dead flies cause the ointment of the apothecary to send forth a stinking savour: \bibemph{so doth} a little folly him that is in reputation for wisdom \bibemph{and} honour.
\fn{Dead...: Heb. Flies of death.}
\vs Ecc 10:2 A wise man's heart \bibemph{is} at his right hand; but a fool's heart at his left.
\vs Ecc 10:3 Yea also, when he that is a fool walketh by the way, his wisdom faileth \bibemph{him}, and he saith to every one \bibemph{that} he \bibemph{is} a fool.
\fn{his...: Heb. his heart.}
\vs Ecc 10:4 If the spirit of the ruler rise up against thee, leave not thy place; for yielding pacifieth great offences.
\vs Ecc 10:5 There is an evil \bibemph{which} I have seen under the sun, as an error \bibemph{which} proceedeth from the ruler:
\fn{from: Heb. from before.}
\vs Ecc 10:6 Folly is set in great dignity, and the rich sit in low place.
\fn{in great...: Heb. in great heights.}
\vs Ecc 10:7 I have seen servants upon horses, and princes walking as servants upon the earth.
\vs Ecc 10:8 He that diggeth a pit shall fall into it; and whoso breaketh an hedge, a serpent shall bite him.
\vs Ecc 10:9 Whoso removeth stones shall be hurt therewith; \bibemph{and} he that cleaveth wood shall be endangered thereby.
\vs Ecc 10:10 If the iron be blunt, and he do not whet the edge, then must he put to more strength: but wisdom \bibemph{is} profitable to direct.
\vs Ecc 10:11 Surely the serpent will bite without enchantment; and a babbler is no better.
\fn{a babbler: Heb. the master of the tongue.}
\vs Ecc 10:12 The words of a wise man's mouth \bibemph{are} gracious; but the lips of a fool will swallow up himself.
\fn{gracious: Heb. grace.}
\vs Ecc 10:13 The beginning of the words of his mouth \bibemph{is} foolishness: and the end of his talk \bibemph{is} mischievous madness.
\fn{his talk: Heb. his mouth.}
\vs Ecc 10:14 A fool also is full of words: a man cannot tell what shall be; and what shall be after him, who can tell him?
\fn{is full...: Heb. multiplieth words.}
\vs Ecc 10:15 The labour of the foolish wearieth every one of them, because he knoweth not how to go to the city.
\vs Ecc 10:16 Woe to thee, O land, when thy king \bibemph{is} a child, and thy princes eat in the morning!
\vs Ecc 10:17 Blessed \bibemph{art} thou, O land, when thy king \bibemph{is} the son of nobles, and thy princes eat in due season, for strength, and not for drunkenness!
\vs Ecc 10:18 By much slothfulness the building decayeth; and through idleness of the hands the house droppeth through.
\vs Ecc 10:19 A feast is made for laughter, and wine maketh merry: but money answereth all \bibemph{things}.
\fn{maketh...: Heb. maketh glad the life.}
\vs Ecc 10:20 Curse not the king, no not in thy thought; and curse not the rich in thy bedchamber: for a bird of the air shall carry the voice, and that which hath wings shall tell the matter.
\fn{thought: or, conscience.}
\vs Ecc 11:1 Cast thy bread upon the waters: for thou shalt find it after many days.
\fn{upon...: Heb. upon the face of the waters.}
\vs Ecc 11:2 Give a portion to seven, and also to eight; for thou knowest not what evil shall be upon the earth.
\vs Ecc 11:3 If the clouds be full of rain, they empty \bibemph{themselves} upon the earth: and if the tree fall toward the south, or toward the north, in the place where the tree falleth, there it shall be.
\vs Ecc 11:4 He that observeth the wind shall not sow; and he that regardeth the clouds shall not reap.
\vs Ecc 11:5 As thou knowest not what \bibemph{is} the way of the spirit, \bibemph{nor} how the bones \bibemph{do grow} in the womb of her that is with child: even so thou knowest not the works of God who maketh all.
\vs Ecc 11:6 In the morning sow thy seed, and in the evening withhold not thine hand: for thou knowest not whether shall prosper, either this or that, or whether they both \bibemph{shall be} alike good.
\fn{shall prosper: Heb. shall be right.}
\vs Ecc 11:7 Truly the light \bibemph{is} sweet, and a pleasant \bibemph{thing it is} for the eyes to behold the sun:
\vs Ecc 11:8 But if a man live many years, \bibemph{and} rejoice in them all; yet let him remember the days of darkness; for they shall be many. All that cometh \bibemph{is} vanity.
\vs Ecc 11:9 Rejoice, O young man, in thy youth; and let thy heart cheer thee in the days of thy youth, and walk in the ways of thine heart, and in the sight of thine eyes: but know thou, that for all these \bibemph{things} God will bring thee into judgment.
\vs Ecc 11:10 Therefore remove sorrow from thy heart, and put away evil from thy flesh: for childhood and youth \bibemph{are} vanity.
\fn{sorrow: or, anger.}
\vs Ecc 12:1 Remember now thy Creator in the days of thy youth, while the evil days come not, nor the years draw nigh, when thou shalt say, I have no pleasure in them;
\vs Ecc 12:2 While the sun, or the light, or the moon, or the stars, be not darkened, nor the clouds return after the rain:
\vs Ecc 12:3 In the day when the keepers of the house shall tremble, and the strong men shall bow themselves, and the grinders cease because they are few, and those that look out of the windows be darkened,
\fn{the grinders...: or, the grinders fail, because they grind little.}
\vs Ecc 12:4 And the doors shall be shut in the streets, when the sound of the grinding is low, and he shall rise up at the voice of the bird, and all the daughters of musick shall be brought low;
\vs Ecc 12:5 Also \bibemph{when} they shall be afraid of \bibemph{that which is} high, and fears \bibemph{shall be} in the way, and the almond tree shall flourish, and the grasshopper shall be a burden, and desire shall fail: because man goeth to his long home, and the mourners go about the streets:
\vs Ecc 12:6 Or ever the silver cord be loosed, or the golden bowl be broken, or the pitcher be broken at the fountain, or the wheel broken at the cistern.
\vs Ecc 12:7 Then shall the dust return to the earth as it was: and the spirit shall return unto God who gave it.
\vs Ecc 12:8 Vanity of vanities, saith the preacher; all \bibemph{is} vanity.
\vs Ecc 12:9 And moreover, because the preacher was wise, he still taught the people knowledge; yea, he gave good heed, and sought out, \bibemph{and} set in order many proverbs.
\fn{moreover...: or, the more wise the preacher was, etc.}
\vs Ecc 12:10 The preacher sought to find out acceptable words: and \bibemph{that which was} written \bibemph{was} upright, \bibemph{even} words of truth.
\fn{acceptable...: Heb. words of delight.}
\vs Ecc 12:11 The words of the wise \bibemph{are} as goads, and as nails fastened \bibemph{by} the masters of assemblies, \bibemph{which} are given from one shepherd.
\vs Ecc 12:12 And further, by these, my son, be admonished: of making many books \bibemph{there is} no end; and much study \bibemph{is} a weariness of the flesh.
\fn{study: or, reading.}
\vs Ecc 12:13 Let us hear the conclusion of the whole matter: Fear God, and keep his commandments: for this \bibemph{is} the whole \bibemph{duty} of man.
\fn{Let...: or, The end of the matter, even all that hath been heard, is.}
\vs Ecc 12:14 For God shall bring every work into judgment, with every secret thing, whether \bibemph{it be} good, or whether \bibemph{it be} evil.
