\bibbook{the book of the prophet\\\bibtitlemain{ezekiel}}{ezekiel}{Ezekiel}{Eze}
\vs Eze 1:1 Now it came to pass in the thirtieth year, in the fourth \bibemph{month}, in the fifth \bibemph{day} of the month, as I \bibemph{was} among the captives by the river of Chebar, \bibemph{that} the heavens were opened, and I saw visions of God.
\fn{captives: Heb. captivity.}
\vs Eze 1:2 In the fifth \bibemph{day} of the month, which \bibemph{was} the fifth year of king Jehoiachin's captivity,
\vs Eze 1:3 The word of the \textsc{Lord} came expressly unto Ezekiel the priest, the son of Buzi, in the land of the Chaldeans by the river Chebar; and the hand of the \textsc{Lord} was there upon him.
\fn{Ezekiel: Heb. Jehezkel.}
\vs Eze 1:4 And I looked, and, behold, a whirlwind came out of the north, a great cloud, and a fire infolding itself, and a brightness \bibemph{was} about it, and out of the midst thereof as the colour of amber, out of the midst of the fire.
\fn{infolding...: Heb. catching itself.}
\vs Eze 1:5 Also out of the midst thereof \bibemph{came} the likeness of four living creatures. And this \bibemph{was} their appearance; they had the likeness of a man.
\vs Eze 1:6 And every one had four faces, and every one had four wings.
\vs Eze 1:7 And their feet \bibemph{were} straight feet; and the sole of their feet \bibemph{was} like the sole of a calf's foot: and they sparkled like the colour of burnished brass.
\fn{straight...: Heb. a straight foot.}
\vs Eze 1:8 And \bibemph{they had} the hands of a man under their wings on their four sides; and they four had their faces and their wings.
\vs Eze 1:9 Their wings \bibemph{were} joined one to another; they turned not when they went; they went every one straight forward.
\vs Eze 1:10 As for the likeness of their faces, they four had the face of a man, and the face of a lion, on the right side: and they four had the face of an ox on the left side; they four also had the face of an eagle.
\vs Eze 1:11 Thus \bibemph{were} their faces: and their wings \bibemph{were} stretched upward; two \bibemph{wings} of every one \bibemph{were} joined one to another, and two covered their bodies.
\fn{stretched...: or, divided above.}
\vs Eze 1:12 And they went every one straight forward: whither the spirit was to go, they went; \bibemph{and} they turned not when they went.
\vs Eze 1:13 As for the likeness of the living creatures, their appearance \bibemph{was} like burning coals of fire, \bibemph{and} like the appearance of lamps: it went up and down among the living creatures; and the fire was bright, and out of the fire went forth lightning.
\vs Eze 1:14 And the living creatures ran and returned as the appearance of a flash of lightning.
\vs Eze 1:15 Now as I beheld the living creatures, behold one wheel upon the earth by the living creatures, with his four faces.
\vs Eze 1:16 The appearance of the wheels and their work \bibemph{was} like unto the colour of a beryl: and they four had one likeness: and their appearance and their work \bibemph{was} as it were a wheel in the middle of a wheel.
\vs Eze 1:17 When they went, they went upon their four sides: \bibemph{and} they turned not when they went.
\vs Eze 1:18 As for their rings, they were so high that they were dreadful; and their rings \bibemph{were} full of eyes round about them four.
\fn{and their rings: or, and their strakes.}
\vs Eze 1:19 And when the living creatures went, the wheels went by them: and when the living creatures were lifted up from the earth, the wheels were lifted up.
\vs Eze 1:20 Whithersoever the spirit was to go, they went, thither \bibemph{was their} spirit to go; and the wheels were lifted up over against them: for the spirit of the living creature \bibemph{was} in the wheels.
\fn{of...: or, of life.}
\vs Eze 1:21 When those went, \bibemph{these} went; and when those stood, \bibemph{these} stood; and when those were lifted up from the earth, the wheels were lifted up over against them: for the spirit of the living creature \bibemph{was} in the wheels.
\fn{of...: or, of life.}
\vs Eze 1:22 And the likeness of the firmament upon the heads of the living creature \bibemph{was} as the colour of the terrible crystal, stretched forth over their heads above.
\vs Eze 1:23 And under the firmament \bibemph{were} their wings straight, the one toward the other: every one had two, which covered on this side, and every one had two, which covered on that side, their bodies.
\vs Eze 1:24 And when they went, I heard the noise of their wings, like the noise of great waters, as the voice of the Almighty, the voice of speech, as the noise of an host: when they stood, they let down their wings.
\vs Eze 1:25 And there was a voice from the firmament that \bibemph{was} over their heads, when they stood, \bibemph{and} had let down their wings.
\vs Eze 1:26 And above the firmament that \bibemph{was} over their heads \bibemph{was} the likeness of a throne, as the appearance of a sapphire stone: and upon the likeness of the throne \bibemph{was} the likeness as the appearance of a man above upon it.
\vs Eze 1:27 And I saw as the colour of amber, as the appearance of fire round about within it, from the appearance of his loins even upward, and from the appearance of his loins even downward, I saw as it were the appearance of fire, and it had brightness round about.
\vs Eze 1:28 As the appearance of the bow that is in the cloud in the day of rain, so \bibemph{was} the appearance of the brightness round about. This \bibemph{was} the appearance of the likeness of the glory of the \textsc{Lord}. And when I saw \bibemph{it}, I fell upon my face, and I heard a voice of one that spake.
\vs Eze 2:1 And he said unto me, Son of man, stand upon thy feet, and I will speak unto thee.
\vs Eze 2:2 And the spirit entered into me when he spake unto me, and set me upon my feet, that I heard him that spake unto me.
\vs Eze 2:3 And he said unto me, Son of man, I send thee to the children of Israel, to a rebellious nation that hath rebelled against me: they and their fathers have transgressed against me, \bibemph{even} unto this very day.
\fn{nation: Heb. nations.}
\vs Eze 2:4 For \bibemph{they are} impudent children and stiffhearted. I do send thee unto them; and thou shalt say unto them, Thus saith the Lord GOD.
\fn{impudent: Heb. hard of face.}
\vs Eze 2:5 And they, whether they will hear, or whether they will forbear, (for they \bibemph{are} a rebellious house,) yet shall know that there hath been a prophet among them.
\vs Eze 2:6 And thou, son of man, be not afraid of them, neither be afraid of their words, though briers and thorns \bibemph{be} with thee, and thou dost dwell among scorpions: be not afraid of their words, nor be dismayed at their looks, though they \bibemph{be} a rebellious house.
\fn{briers...: or, rebels.}
\vs Eze 2:7 And thou shalt speak my words unto them, whether they will hear, or whether they will forbear: for they \bibemph{are} most rebellious.
\fn{most...: Heb. rebellion.}
\vs Eze 2:8 But thou, son of man, hear what I say unto thee; Be not thou rebellious like that rebellious house: open thy mouth, and eat that I give thee.
\vs Eze 2:9 And when I looked, behold, an hand \bibemph{was} sent unto me; and, lo, a roll of a book \bibemph{was} therein;
\vs Eze 2:10 And he spread it before me; and it \bibemph{was} written within and without: and \bibemph{there was} written therein lamentations, and mourning, and woe.
\vs Eze 3:1 Moreover he said unto me, Son of man, eat that thou findest; eat this roll, and go speak unto the house of Israel.
\vs Eze 3:2 So I opened my mouth, and he caused me to eat that roll.
\vs Eze 3:3 And he said unto me, Son of man, cause thy belly to eat, and fill thy bowels with this roll that I give thee. Then did I eat \bibemph{it}; and it was in my mouth as honey for sweetness.
\vs Eze 3:4 And he said unto me, Son of man, go, get thee unto the house of Israel, and speak with my words unto them.
\vs Eze 3:5 For thou \bibemph{art} not sent to a people of a strange speech and of an hard language, \bibemph{but} to the house of Israel;
\fn{of a...: Heb. deep of lip, and heavy of tongue.}
\vs Eze 3:6 Not to many people of a strange speech and of an hard language, whose words thou canst not understand. Surely, had I sent thee to them, they would have hearkened unto thee.
\fn{of a...: Heb. deep of lip, and heavy of language. Surely...: or, If I had sent thee, etc. would they not have hearkened unto thee?.}
\vs Eze 3:7 But the house of Israel will not hearken unto thee; for they will not hearken unto me: for all the house of Israel \bibemph{are} impudent and hardhearted.
\fn{impudent...: Heb. stiff of forehead, and hard of heart.}
\vs Eze 3:8 Behold, I have made thy face strong against their faces, and thy forehead strong against their foreheads.
\vs Eze 3:9 As an adamant harder than flint have I made thy forehead: fear them not, neither be dismayed at their looks, though they \bibemph{be} a rebellious house.
\vs Eze 3:10 Moreover he said unto me, Son of man, all my words that I shall speak unto thee receive in thine heart, and hear with thine ears.
\vs Eze 3:11 And go, get thee to them of the captivity, unto the children of thy people, and speak unto them, and tell them, Thus saith the Lord GOD; whether they will hear, or whether they will forbear.
\vs Eze 3:12 Then the spirit took me up, and I heard behind me a voice of a great rushing, \bibemph{saying}, Blessed \bibemph{be} the glory of the \textsc{Lord} from his place.
\vs Eze 3:13 \bibemph{I heard} also the noise of the wings of the living creatures that touched one another, and the noise of the wheels over against them, and a noise of a great rushing.
\fn{touched: Heb. kissed.}
\vs Eze 3:14 So the spirit lifted me up, and took me away, and I went in bitterness, in the heat of my spirit; but the hand of the \textsc{Lord} was strong upon me.
\fn{in bitterness: Heb. bitter. heat: Heb. hot anger.}
\vs Eze 3:15 Then I came to them of the captivity at Telabib, that dwelt by the river of Chebar, and I sat where they sat, and remained there astonished among them seven days.
\vs Eze 3:16 And it came to pass at the end of seven days, that the word of the \textsc{Lord} came unto me, saying,
\vs Eze 3:17 Son of man, I have made thee a watchman unto the house of Israel: therefore hear the word at my mouth, and give them warning from me.
\vs Eze 3:18 When I say unto the wicked, Thou shalt surely die; and thou givest him not warning, nor speakest to warn the wicked from his wicked way, to save his life; the same wicked \bibemph{man} shall die in his iniquity; but his blood will I require at thine hand.
\vs Eze 3:19 Yet if thou warn the wicked, and he turn not from his wickedness, nor from his wicked way, he shall die in his iniquity; but thou hast delivered thy soul.
\vs Eze 3:20 Again, When a righteous \bibemph{man} doth turn from his righteousness, and commit iniquity, and I lay a stumblingblock before him, he shall die: because thou hast not given him warning, he shall die in his sin, and his righteousness which he hath done shall not be remembered; but his blood will I require at thine hand.
\fn{righteousness which: Heb. righteousnesses.}
\vs Eze 3:21 Nevertheless if thou warn the righteous \bibemph{man}, that the righteous sin not, and he doth not sin, he shall surely live, because he is warned; also thou hast delivered thy soul.
\vs Eze 3:22 And the hand of the \textsc{Lord} was there upon me; and he said unto me, Arise, go forth into the plain, and I will there talk with thee.
\vs Eze 3:23 Then I arose, and went forth into the plain: and, behold, the glory of the \textsc{Lord} stood there, as the glory which I saw by the river of Chebar: and I fell on my face.
\vs Eze 3:24 Then the spirit entered into me, and set me upon my feet, and spake with me, and said unto me, Go, shut thyself within thine house.
\vs Eze 3:25 But thou, O son of man, behold, they shall put bands upon thee, and shall bind thee with them, and thou shalt not go out among them:
\vs Eze 3:26 And I will make thy tongue cleave to the roof of thy mouth, that thou shalt be dumb, and shalt not be to them a reprover: for they \bibemph{are} a rebellious house.
\fn{a reprover: Heb. a man reproving.}
\vs Eze 3:27 But when I speak with thee, I will open thy mouth, and thou shalt say unto them, Thus saith the Lord GOD; He that heareth, let him hear; and he that forbeareth, let him forbear: for they \bibemph{are} a rebellious house.
\vs Eze 4:1 Thou also, son of man, take thee a tile, and lay it before thee, and pourtray upon it the city, \bibemph{even} Jerusalem:
\vs Eze 4:2 And lay siege against it, and build a fort against it, and cast a mount against it; set the camp also against it, and set \bibemph{battering} rams against it round about.
\fn{battering...: chief leaders.}
\vs Eze 4:3 Moreover take thou unto thee an iron pan, and set it \bibemph{for} a wall of iron between thee and the city: and set thy face against it, and it shall be besieged, and thou shalt lay siege against it. This \bibemph{shall be} a sign to the house of Israel.
\fn{an...: or, a flat plate, or, slice.}
\vs Eze 4:4 Lie thou also upon thy left side, and lay the iniquity of the house of Israel upon it: \bibemph{according} to the number of the days that thou shalt lie upon it thou shalt bear their iniquity.
\vs Eze 4:5 For I have laid upon thee the years of their iniquity, according to the number of the days, three hundred and ninety days: so shalt thou bear the iniquity of the house of Israel.
\vs Eze 4:6 And when thou hast accomplished them, lie again on thy right side, and thou shalt bear the iniquity of the house of Judah forty days: I have appointed thee each day for a year.
\fn{each...: Heb. a day for a year, a day for a year.}
\vs Eze 4:7 Therefore thou shalt set thy face toward the siege of Jerusalem, and thine arm \bibemph{shall be} uncovered, and thou shalt prophesy against it.
\vs Eze 4:8 And, behold, I will lay bands upon thee, and thou shalt not turn thee from one side to another, till thou hast ended the days of thy siege.
\fn{from one...: Heb. from thy side to thy side.}
\vs Eze 4:9 Take thou also unto thee wheat, and barley, and beans, and lentiles, and millet, and fitches, and put them in one vessel, and make thee bread thereof, \bibemph{according} to the number of the days that thou shalt lie upon thy side, three hundred and ninety days shalt thou eat thereof.
\fn{fitches: or, spelt.}
\vs Eze 4:10 And thy meat which thou shalt eat \bibemph{shall be} by weight, twenty shekels a day: from time to time shalt thou eat it.
\vs Eze 4:11 Thou shalt drink also water by measure, the sixth part of an hin: from time to time shalt thou drink.
\vs Eze 4:12 And thou shalt eat it \bibemph{as} barley cakes, and thou shalt bake it with dung that cometh out of man, in their sight.
\vs Eze 4:13 And the \textsc{Lord} said, Even thus shall the children of Israel eat their defiled bread among the Gentiles, whither I will drive them.
\vs Eze 4:14 Then said I, Ah Lord GOD! behold, my soul hath not been polluted: for from my youth up even till now have I not eaten of that which dieth of itself, or is torn in pieces; neither came there abominable flesh into my mouth.
\vs Eze 4:15 Then he said unto me, Lo, I have given thee cow's dung for man's dung, and thou shalt prepare thy bread therewith.
\vs Eze 4:16 Moreover he said unto me, Son of man, behold, I will break the staff of bread in Jerusalem: and they shall eat bread by weight, and with care; and they shall drink water by measure, and with astonishment:
\vs Eze 4:17 That they may want bread and water, and be astonied one with another, and consume away for their iniquity.
\vs Eze 5:1 And thou, son of man, take thee a sharp knife, take thee a barber's razor, and cause \bibemph{it} to pass upon thine head and upon thy beard: then take thee balances to weigh, and divide the \bibemph{hair}.
\vs Eze 5:2 Thou shalt burn with fire a third part in the midst of the city, when the days of the siege are fulfilled: and thou shalt take a third part, \bibemph{and} smite about it with a knife: and a third part thou shalt scatter in the wind; and I will draw out a sword after them.
\vs Eze 5:3 Thou shalt also take thereof a few in number, and bind them in thy skirts.
\fn{skirts: Heb. wings.}
\vs Eze 5:4 Then take of them again, and cast them into the midst of the fire, and burn them in the fire; \bibemph{for} thereof shall a fire come forth into all the house of Israel.
\vs Eze 5:5 Thus saith the Lord GOD; This \bibemph{is} Jerusalem: I have set it in the midst of the nations and countries \bibemph{that are} round about her.
\vs Eze 5:6 And she hath changed my judgments into wickedness more than the nations, and my statutes more than the countries that \bibemph{are} round about her: for they have refused my judgments and my statutes, they have not walked in them.
\vs Eze 5:7 Therefore thus saith the Lord GOD; Because ye multiplied more than the nations that \bibemph{are} round about you, \bibemph{and} have not walked in my statutes, neither have kept my judgments, neither have done according to the judgments of the nations that \bibemph{are} round about you;
\vs Eze 5:8 Therefore thus saith the Lord GOD; Behold, I, even I, \bibemph{am} against thee, and will execute judgments in the midst of thee in the sight of the nations.
\vs Eze 5:9 And I will do in thee that which I have not done, and whereunto I will not do any more the like, because of all thine abominations.
\vs Eze 5:10 Therefore the fathers shall eat the sons in the midst of thee, and the sons shall eat their fathers; and I will execute judgments in thee, and the whole remnant of thee will I scatter into all the winds.
\vs Eze 5:11 Wherefore, \bibemph{as} I live, saith the Lord GOD; Surely, because thou hast defiled my sanctuary with all thy detestable things, and with all thine abominations, therefore will I also diminish \bibemph{thee}; neither shall mine eye spare, neither will I have any pity.
\vs Eze 5:12 A third part of thee shall die with the pestilence, and with famine shall they be consumed in the midst of thee: and a third part shall fall by the sword round about thee; and I will scatter a third part into all the winds, and I will draw out a sword after them.
\vs Eze 5:13 Thus shall mine anger be accomplished, and I will cause my fury to rest upon them, and I will be comforted: and they shall know that I the \textsc{Lord} have spoken \bibemph{it} in my zeal, when I have accomplished my fury in them.
\vs Eze 5:14 Moreover I will make thee waste, and a reproach among the nations that \bibemph{are} round about thee, in the sight of all that pass by.
\vs Eze 5:15 So it shall be a reproach and a taunt, an instruction and an astonishment unto the nations that \bibemph{are} round about thee, when I shall execute judgments in thee in anger and in fury and in furious rebukes. I the \textsc{Lord} have spoken \bibemph{it}.
\vs Eze 5:16 When I shall send upon them the evil arrows of famine, which shall be for \bibemph{their} destruction, \bibemph{and} which I will send to destroy you: and I will increase the famine upon you, and will break your staff of bread:
\vs Eze 5:17 So will I send upon you famine and evil beasts, and they shall bereave thee; and pestilence and blood shall pass through thee; and I will bring the sword upon thee. I the \textsc{Lord} have spoken \bibemph{it}.
\vs Eze 6:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 6:2 Son of man, set thy face toward the mountains of Israel, and prophesy against them,
\vs Eze 6:3 And say, Ye mountains of Israel, hear the word of the Lord GOD; Thus saith the Lord GOD to the mountains, and to the hills, to the rivers, and to the valleys; Behold, I, \bibemph{even} I, will bring a sword upon you, and I will destroy your high places.
\vs Eze 6:4 And your altars shall be desolate, and your images shall be broken: and I will cast down your slain \bibemph{men} before your idols.
\fn{images: or, sun images.}
\vs Eze 6:5 And I will lay the dead carcases of the children of Israel before their idols; and I will scatter your bones round about your altars.
\fn{lay: Heb. give.}
\vs Eze 6:6 In all your dwellingplaces the cities shall be laid waste, and the high places shall be desolate; that your altars may be laid waste and made desolate, and your idols may be broken and cease, and your images may be cut down, and your works may be abolished.
\vs Eze 6:7 And the slain shall fall in the midst of you, and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 6:8 Yet will I leave a remnant, that ye may have \bibemph{some} that shall escape the sword among the nations, when ye shall be scattered through the countries.
\vs Eze 6:9 And they that escape of you shall remember me among the nations whither they shall be carried captives, because I am broken with their whorish heart, which hath departed from me, and with their eyes, which go a whoring after their idols: and they shall lothe themselves for the evils which they have committed in all their abominations.
\vs Eze 6:10 And they shall know that I \bibemph{am} the \textsc{Lord}, \bibemph{and that} I have not said in vain that I would do this evil unto them.
\vs Eze 6:11 Thus saith the Lord GOD; Smite with thine hand, and stamp with thy foot, and say, Alas for all the evil abominations of the house of Israel! for they shall fall by the sword, by the famine, and by the pestilence.
\vs Eze 6:12 He that is far off shall die of the pestilence; and he that is near shall fall by the sword; and he that remaineth and is besieged shall die by the famine: thus will I accomplish my fury upon them.
\vs Eze 6:13 Then shall ye know that I \bibemph{am} the \textsc{Lord}, when their slain \bibemph{men} shall be among their idols round about their altars, upon every high hill, in all the tops of the mountains, and under every green tree, and under every thick oak, the place where they did offer sweet savour to all their idols.
\vs Eze 6:14 So will I stretch out my hand upon them, and make the land desolate, yea, more desolate than the wilderness toward Diblath, in all their habitations: and they shall know that I \bibemph{am} the \textsc{Lord}.
\fn{more...: or, desolate from the wilderness.}
\vs Eze 7:1 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 7:2 Also, thou son of man, thus saith the Lord GOD unto the land of Israel; An end, the end is come upon the four corners of the land.
\vs Eze 7:3 Now \bibemph{is} the end \bibemph{come} upon thee, and I will send mine anger upon thee, and will judge thee according to thy ways, and will recompense upon thee all thine abominations.
\fn{recompense: Heb. give.}
\vs Eze 7:4 And mine eye shall not spare thee, neither will I have pity: but I will recompense thy ways upon thee, and thine abominations shall be in the midst of thee: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 7:5 Thus saith the Lord GOD; An evil, an only evil, behold, is come.
\vs Eze 7:6 An end is come, the end is come: it watcheth for thee; behold, it is come.
\fn{watcheth for: Heb. awaketh against.}
\vs Eze 7:7 The morning is come unto thee, O thou that dwellest in the land: the time is come, the day of trouble \bibemph{is} near, and not the sounding again of the mountains.
\fn{sounding...: or, echo.}
\vs Eze 7:8 Now will I shortly pour out my fury upon thee, and accomplish mine anger upon thee: and I will judge thee according to thy ways, and will recompense thee for all thine abominations.
\vs Eze 7:9 And mine eye shall not spare, neither will I have pity: I will recompense thee according to thy ways and thine abominations \bibemph{that} are in the midst of thee; and ye shall know that I \bibemph{am} the \textsc{Lord} that smiteth.
\fn{thee according: Heb. upon thee, etc.}
\vs Eze 7:10 Behold the day, behold, it is come: the morning is gone forth; the rod hath blossomed, pride hath budded.
\vs Eze 7:11 Violence is risen up into a rod of wickedness: none of them \bibemph{shall remain}, nor of their multitude, nor of any of theirs: neither \bibemph{shall there be} wailing for them.
\fn{theirs: or, their tumultuous persons: Heb. tumult.}
\vs Eze 7:12 The time is come, the day draweth near: let not the buyer rejoice, nor the seller mourn: for wrath \bibemph{is} upon all the multitude thereof.
\vs Eze 7:13 For the seller shall not return to that which is sold, although they were yet alive: for the vision \bibemph{is} touching the whole multitude thereof, \bibemph{which} shall not return; neither shall any strengthen himself in the iniquity of his life.
\fn{although they...: Heb. though their life were yet among the living. in the...: or, whose life is in his iniquity. the iniquity: Heb. his iniquity.}
\vs Eze 7:14 They have blown the trumpet, even to make all ready; but none goeth to the battle: for my wrath \bibemph{is} upon all the multitude thereof.
\vs Eze 7:15 The sword \bibemph{is} without, and the pestilence and the famine within: he that \bibemph{is} in the field shall die with the sword; and he that \bibemph{is} in the city, famine and pestilence shall devour him.
\vs Eze 7:16 But they that escape of them shall escape, and shall be on the mountains like doves of the valleys, all of them mourning, every one for his iniquity.
\vs Eze 7:17 All hands shall be feeble, and all knees shall be weak \bibemph{as} water.
\fn{be weak...: Heb. go into water.}
\vs Eze 7:18 They shall also gird \bibemph{themselves} with sackcloth, and horror shall cover them; and shame \bibemph{shall be} upon all faces, and baldness upon all their heads.
\vs Eze 7:19 They shall cast their silver in the streets, and their gold shall be removed: their silver and their gold shall not be able to deliver them in the day of the wrath of the \textsc{Lord}: they shall not satisfy their souls, neither fill their bowels: because it is the stumblingblock of their iniquity.
\fn{removed: Heb. for a separation, or, uncleanness. it is...: or, their iniquity is their stumblingblock.}
\vs Eze 7:20 As for the beauty of his ornament, he set it in majesty: but they made the images of their abominations \bibemph{and} of their detestable things therein: therefore have I set it far from them.
\fn{set it far...: or, made it unto them an unclean thing.}
\vs Eze 7:21 And I will give it into the hands of the strangers for a prey, and to the wicked of the earth for a spoil; and they shall pollute it.
\vs Eze 7:22 My face will I turn also from them, and they shall pollute my secret \bibemph{place}: for the robbers shall enter into it, and defile it.
\fn{robbers: or, buglers.}
\vs Eze 7:23 Make a chain: for the land is full of bloody crimes, and the city is full of violence.
\vs Eze 7:24 Wherefore I will bring the worst of the heathen, and they shall possess their houses: I will also make the pomp of the strong to cease; and their holy places shall be defiled.
\fn{their holy...: or, they shall inherit their holy places.}
\vs Eze 7:25 Destruction cometh; and they shall seek peace, and \bibemph{there shall be} none.
\fn{Destruction: Heb. Cutting off.}
\vs Eze 7:26 Mischief shall come upon mischief, and rumour shall be upon rumour; then shall they seek a vision of the prophet; but the law shall perish from the priest, and counsel from the ancients.
\vs Eze 7:27 The king shall mourn, and the prince shall be clothed with desolation, and the hands of the people of the land shall be troubled: I will do unto them after their way, and according to their deserts will I judge them; and they shall know that I \bibemph{am} the \textsc{Lord}.
\fn{according...: Heb. with their judgments.}
\vs Eze 8:1 And it came to pass in the sixth year, in the sixth \bibemph{month}, in the fifth \bibemph{day} of the month, \bibemph{as} I sat in mine house, and the elders of Judah sat before me, that the hand of the Lord GOD fell there upon me.
\vs Eze 8:2 Then I beheld, and lo a likeness as the appearance of fire: from the appearance of his loins even downward, fire; and from his loins even upward, as the appearance of brightness, as the colour of amber.
\vs Eze 8:3 And he put forth the form of an hand, and took me by a lock of mine head; and the spirit lifted me up between the earth and the heaven, and brought me in the visions of God to Jerusalem, to the door of the inner gate that looketh toward the north; where \bibemph{was} the seat of the image of jealousy, which provoketh to jealousy.
\vs Eze 8:4 And, behold, the glory of the God of Israel \bibemph{was} there, according to the vision that I saw in the plain.
\vs Eze 8:5 Then said he unto me, Son of man, lift up thine eyes now the way toward the north. So I lifted up mine eyes the way toward the north, and behold northward at the gate of the altar this image of jealousy in the entry.
\vs Eze 8:6 He said furthermore unto me, Son of man, seest thou what they do? \bibemph{even} the great abominations that the house of Israel committeth here, that I should go far off from my sanctuary? but turn thee yet again, \bibemph{and} thou shalt see greater abominations.
\vs Eze 8:7 And he brought me to the door of the court; and when I looked, behold a hole in the wall.
\vs Eze 8:8 Then said he unto me, Son of man, dig now in the wall: and when I had digged in the wall, behold a door.
\vs Eze 8:9 And he said unto me, Go in, and behold the wicked abominations that they do here.
\vs Eze 8:10 So I went in and saw; and behold every form of creeping things, and abominable beasts, and all the idols of the house of Israel, pourtrayed upon the wall round about.
\vs Eze 8:11 And there stood before them seventy men of the ancients of the house of Israel, and in the midst of them stood Jaazaniah the son of Shaphan, with every man his censer in his hand; and a thick cloud of incense went up.
\vs Eze 8:12 Then said he unto me, Son of man, hast thou seen what the ancients of the house of Israel do in the dark, every man in the chambers of his imagery? for they say, The \textsc{Lord} seeth us not; the \textsc{Lord} hath forsaken the earth.
\vs Eze 8:13 He said also unto me, Turn thee yet again, \bibemph{and} thou shalt see greater abominations that they do.
\vs Eze 8:14 Then he brought me to the door of the gate of the \textsc{Lord's} house which \bibemph{was} toward the north; and, behold, there sat women weeping for Tammuz.
\vs Eze 8:15 Then said he unto me, Hast thou seen \bibemph{this}, O son of man? turn thee yet again, \bibemph{and} thou shalt see greater abominations than these.
\vs Eze 8:16 And he brought me into the inner court of the \textsc{Lord's} house, and, behold, at the door of the temple of the \textsc{Lord}, between the porch and the altar, \bibemph{were} about five and twenty men, with their backs toward the temple of the \textsc{Lord}, and their faces toward the east; and they worshipped the sun toward the east.
\vs Eze 8:17 Then he said unto me, Hast thou seen \bibemph{this}, O son of man? Is it a light thing to the house of Judah that they commit the abominations which they commit here? for they have filled the land with violence, and have returned to provoke me to anger: and, lo, they put the branch to their nose.
\fn{Is it...: or, Is there any thing lighter than to commit.}
\vs Eze 8:18 Therefore will I also deal in fury: mine eye shall not spare, neither will I have pity: and though they cry in mine ears with a loud voice, \bibemph{yet} will I not hear them.
\vs Eze 9:1 He cried also in mine ears with a loud voice, saying, Cause them that have charge over the city to draw near, even every man \bibemph{with} his destroying weapon in his hand.
\vs Eze 9:2 And, behold, six men came from the way of the higher gate, which lieth toward the north, and every man a slaughter weapon in his hand; and one man among them \bibemph{was} clothed with linen, with a writer's inkhorn by his side: and they went in, and stood beside the brasen altar.
\fn{which lieth: Heb. which is turned. by his...: Heb. upon his loins. a slaughter...: Heb. a weapon of his breaking in pieces.}
\vs Eze 9:3 And the glory of the God of Israel was gone up from the cherub, whereupon he was, to the threshold of the house. And he called to the man clothed with linen, which \bibemph{had} the writer's inkhorn by his side;
\vs Eze 9:4 And the \textsc{Lord} said unto him, Go through the midst of the city, through the midst of Jerusalem, and set a mark upon the foreheads of the men that sigh and that cry for all the abominations that be done in the midst thereof.
\fn{set a mark: Heb. mark a mark.}
\vs Eze 9:5 And to the others he said in mine hearing, Go ye after him through the city, and smite: let not your eye spare, neither have ye pity:
\fn{mine hearing: Heb. mine ears.}
\vs Eze 9:6 Slay utterly old \bibemph{and} young, both maids, and little children, and women: but come not near any man upon whom \bibemph{is} the mark; and begin at my sanctuary. Then they began at the ancient men which \bibemph{were} before the house.
\fn{utterly: Heb. to destruction.}
\vs Eze 9:7 And he said unto them, Defile the house, and fill the courts with the slain: go ye forth. And they went forth, and slew in the city.
\vs Eze 9:8 And it came to pass, while they were slaying them, and I was left, that I fell upon my face, and cried, and said, Ah Lord GOD! wilt thou destroy all the residue of Israel in thy pouring out of thy fury upon Jerusalem?
\vs Eze 9:9 Then said he unto me, The iniquity of the house of Israel and Judah \bibemph{is} exceeding great, and the land is full of blood, and the city full of perverseness: for they say, The \textsc{Lord} hath forsaken the earth, and the \textsc{Lord} seeth not.
\fn{full of blood: Heb. filled with, etc. perverseness: or, wresting of judgment.}
\vs Eze 9:10 And as for me also, mine eye shall not spare, neither will I have pity, \bibemph{but} I will recompense their way upon their head.
\vs Eze 9:11 And, behold, the man clothed with linen, which \bibemph{had} the inkhorn by his side, reported the matter, saying, I have done as thou hast commanded me.
\fn{reported...: Heb. returned the word.}
\vs Eze 10:1 Then I looked, and, behold, in the firmament that was above the head of the cherubims there appeared over them as it were a sapphire stone, as the appearance of the likeness of a throne.
\vs Eze 10:2 And he spake unto the man clothed with linen, and said, Go in between the wheels, \bibemph{even} under the cherub, and fill thine hand with coals of fire from between the cherubims, and scatter \bibemph{them} over the city. And he went in in my sight.
\fn{thine hand: Heb. the hollow of thine hand.}
\vs Eze 10:3 Now the cherubims stood on the right side of the house, when the man went in; and the cloud filled the inner court.
\vs Eze 10:4 Then the glory of the \textsc{Lord} went up from the cherub, \bibemph{and stood} over the threshold of the house; and the house was filled with the cloud, and the court was full of the brightness of the \textsc{Lord's} glory.
\fn{went up: Heb. was lifted up.}
\vs Eze 10:5 And the sound of the cherubims' wings was heard \bibemph{even} to the outer court, as the voice of the Almighty God when he speaketh.
\vs Eze 10:6 And it came to pass, \bibemph{that} when he had commanded the man clothed with linen, saying, Take fire from between the wheels, from between the cherubims; then he went in, and stood beside the wheels.
\vs Eze 10:7 And \bibemph{one} cherub stretched forth his hand from between the cherubims unto the fire that \bibemph{was} between the cherubims, and took \bibemph{thereof}, and put \bibemph{it} into the hands of \bibemph{him that was} clothed with linen: who took \bibemph{it}, and went out.
\fn{stretched forth: Heb. sent forth.}
\vs Eze 10:8 And there appeared in the cherubims the form of a man's hand under their wings.
\vs Eze 10:9 And when I looked, behold the four wheels by the cherubims, one wheel by one cherub, and another wheel by another cherub: and the appearance of the wheels \bibemph{was} as the colour of a beryl stone.
\vs Eze 10:10 And \bibemph{as for} their appearances, they four had one likeness, as if a wheel had been in the midst of a wheel.
\vs Eze 10:11 When they went, they went upon their four sides; they turned not as they went, but to the place whither the head looked they followed it; they turned not as they went.
\vs Eze 10:12 And their whole body, and their backs, and their hands, and their wings, and the wheels, \bibemph{were} full of eyes round about, \bibemph{even} the wheels that they four had.
\fn{body: Heb. flesh.}
\vs Eze 10:13 As for the wheels, it was cried unto them in my hearing, O wheel.
\fn{it was...: or, they were called in my hearing, wheel, or, galgal.}
\vs Eze 10:14 And every one had four faces: the first face \bibemph{was} the face of a cherub, and the second face \bibemph{was} the face of a man, and the third the face of a lion, and the fourth the face of an eagle.
\vs Eze 10:15 And the cherubims were lifted up. This \bibemph{is} the living creature that I saw by the river of Chebar.
\vs Eze 10:16 And when the cherubims went, the wheels went by them: and when the cherubims lifted up their wings to mount up from the earth, the same wheels also turned not from beside them.
\vs Eze 10:17 When they stood, \bibemph{these} stood; and when they were lifted up, \bibemph{these} lifted up themselves \bibemph{also}: for the spirit of the living creature \bibemph{was} in them.
\fn{of...: or, of life.}
\vs Eze 10:18 Then the glory of the \textsc{Lord} departed from off the threshold of the house, and stood over the cherubims.
\vs Eze 10:19 And the cherubims lifted up their wings, and mounted up from the earth in my sight: when they went out, the wheels also \bibemph{were} beside them, and \bibemph{every one} stood at the door of the east gate of the \textsc{Lord's} house; and the glory of the God of Israel \bibemph{was} over them above.
\vs Eze 10:20 This \bibemph{is} the living creature that I saw under the God of Israel by the river of Chebar; and I knew that they \bibemph{were} the cherubims.
\vs Eze 10:21 Every one had four faces apiece, and every one four wings; and the likeness of the hands of a man \bibemph{was} under their wings.
\vs Eze 10:22 And the likeness of their faces \bibemph{was} the same faces which I saw by the river of Chebar, their appearances and themselves: they went every one straight forward.
\vs Eze 11:1 Moreover the spirit lifted me up, and brought me unto the east gate of the \textsc{Lord's} house, which looketh eastward: and behold at the door of the gate five and twenty men; among whom I saw Jaazaniah the son of Azur, and Pelatiah the son of Benaiah, princes of the people.
\vs Eze 11:2 Then said he unto me, Son of man, these \bibemph{are} the men that devise mischief, and give wicked counsel in this city:
\vs Eze 11:3 Which say, \bibemph{It is} not near; let us build houses: this \bibemph{city is} the caldron, and we \bibemph{be} the flesh.
\fn{It is...: or, It is not for us to build houses near.}
\vs Eze 11:4 Therefore prophesy against them, prophesy, O son of man.
\vs Eze 11:5 And the Spirit of the \textsc{Lord} fell upon me, and said unto me, Speak; Thus saith the \textsc{Lord}; Thus have ye said, O house of Israel: for I know the things that come into your mind, \bibemph{every one of} them.
\vs Eze 11:6 Ye have multiplied your slain in this city, and ye have filled the streets thereof with the slain.
\vs Eze 11:7 Therefore thus saith the Lord GOD; Your slain whom ye have laid in the midst of it, they \bibemph{are} the flesh, and this \bibemph{city is} the caldron: but I will bring you forth out of the midst of it.
\vs Eze 11:8 Ye have feared the sword; and I will bring a sword upon you, saith the Lord GOD.
\vs Eze 11:9 And I will bring you out of the midst thereof, and deliver you into the hands of strangers, and will execute judgments among you.
\vs Eze 11:10 Ye shall fall by the sword; I will judge you in the border of Israel; and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 11:11 This \bibemph{city} shall not be your caldron, neither shall ye be the flesh in the midst thereof; \bibemph{but} I will judge you in the border of Israel:
\vs Eze 11:12 And ye shall know that I \bibemph{am} the \textsc{Lord}: for ye have not walked in my statutes, neither executed my judgments, but have done after the manners of the heathen that \bibemph{are} round about you.
\fn{for ye...: or, which have not walked.}
\vs Eze 11:13 And it came to pass, when I prophesied, that Pelatiah the son of Benaiah died. Then fell I down upon my face, and cried with a loud voice, and said, Ah Lord GOD! wilt thou make a full end of the remnant of Israel?
\vs Eze 11:14 Again the word of the \textsc{Lord} came unto me, saying,
\vs Eze 11:15 Son of man, thy brethren, \bibemph{even} thy brethren, the men of thy kindred, and all the house of Israel wholly, \bibemph{are} they unto whom the inhabitants of Jerusalem have said, Get you far from the \textsc{Lord}: unto us is this land given in possession.
\vs Eze 11:16 Therefore say, Thus saith the Lord GOD; Although I have cast them far off among the heathen, and although I have scattered them among the countries, yet will I be to them as a little sanctuary in the countries where they shall come.
\vs Eze 11:17 Therefore say, Thus saith the Lord GOD; I will even gather you from the people, and assemble you out of the countries where ye have been scattered, and I will give you the land of Israel.
\vs Eze 11:18 And they shall come thither, and they shall take away all the detestable things thereof and all the abominations thereof from thence.
\vs Eze 11:19 And I will give them one heart, and I will put a new spirit within you; and I will take the stony heart out of their flesh, and will give them an heart of flesh:
\vs Eze 11:20 That they may walk in my statutes, and keep mine ordinances, and do them: and they shall be my people, and I will be their God.
\vs Eze 11:21 But \bibemph{as for them} whose heart walketh after the heart of their detestable things and their abominations, I will recompense their way upon their own heads, saith the Lord GOD.
\vs Eze 11:22 Then did the cherubims lift up their wings, and the wheels beside them; and the glory of the God of Israel \bibemph{was} over them above.
\vs Eze 11:23 And the glory of the \textsc{Lord} went up from the midst of the city, and stood upon the mountain which \bibemph{is} on the east side of the city.
\vs Eze 11:24 Afterwards the spirit took me up, and brought me in a vision by the Spirit of God into Chaldea, to them of the captivity. So the vision that I had seen went up from me.
\vs Eze 11:25 Then I spake unto them of the captivity all the things that the \textsc{Lord} had shewed me.
\vs Eze 12:1 The word of the \textsc{Lord} also came unto me, saying,
\vs Eze 12:2 Son of man, thou dwellest in the midst of a rebellious house, which have eyes to see, and see not; they have ears to hear, and hear not: for they \bibemph{are} a rebellious house.
\vs Eze 12:3 Therefore, thou son of man, prepare thee stuff for removing, and remove by day in their sight; and thou shalt remove from thy place to another place in their sight: it may be they will consider, though they \bibemph{be} a rebellious house.
\fn{stuff: or, instruments.}
\vs Eze 12:4 Then shalt thou bring forth thy stuff by day in their sight, as stuff for removing: and thou shalt go forth at even in their sight, as they that go forth into captivity.
\fn{as they...: Heb. as the goings forth of captivity.}
\vs Eze 12:5 Dig thou through the wall in their sight, and carry out thereby.
\fn{Dig...: Heb. Dig for thee.}
\vs Eze 12:6 In their sight shalt thou bear \bibemph{it} upon \bibemph{thy} shoulders, \bibemph{and} carry \bibemph{it} forth in the twilight: thou shalt cover thy face, that thou see not the ground: for I have set thee \bibemph{for} a sign unto the house of Israel.
\vs Eze 12:7 And I did so as I was commanded: I brought forth my stuff by day, as stuff for captivity, and in the even I digged through the wall with mine hand; I brought \bibemph{it} forth in the twilight, \bibemph{and} I bare \bibemph{it} upon \bibemph{my} shoulder in their sight.
\fn{digged: Heb. digged for me.}
\vs Eze 12:8 And in the morning came the word of the \textsc{Lord} unto me, saying,
\vs Eze 12:9 Son of man, hath not the house of Israel, the rebellious house, said unto thee, What doest thou?
\vs Eze 12:10 Say thou unto them, Thus saith the Lord GOD; This burden \bibemph{concerneth} the prince in Jerusalem, and all the house of Israel that \bibemph{are} among them.
\vs Eze 12:11 Say, I \bibemph{am} your sign: like as I have done, so shall it be done unto them: they shall remove \bibemph{and} go into captivity.
\fn{they...: Heb. by removing go into captivity.}
\vs Eze 12:12 And the prince that \bibemph{is} among them shall bear upon \bibemph{his} shoulder in the twilight, and shall go forth: they shall dig through the wall to carry out thereby: he shall cover his face, that he see not the ground with \bibemph{his} eyes.
\vs Eze 12:13 My net also will I spread upon him, and he shall be taken in my snare: and I will bring him to Babylon \bibemph{to} the land of the Chaldeans; yet shall he not see it, though he shall die there.
\vs Eze 12:14 And I will scatter toward every wind all that \bibemph{are} about him to help him, and all his bands; and I will draw out the sword after them.
\vs Eze 12:15 And they shall know that I \bibemph{am} the \textsc{Lord}, when I shall scatter them among the nations, and disperse them in the countries.
\vs Eze 12:16 But I will leave a few men of them from the sword, from the famine, and from the pestilence; that they may declare all their abominations among the heathen whither they come; and they shall know that I \bibemph{am} the \textsc{Lord}.
\fn{a few...: Heb. men of number.}
\vs Eze 12:17 Moreover the word of the \textsc{Lord} came to me, saying,
\vs Eze 12:18 Son of man, eat thy bread with quaking, and drink thy water with trembling and with carefulness;
\vs Eze 12:19 And say unto the people of the land, Thus saith the Lord GOD of the inhabitants of Jerusalem, \bibemph{and} of the land of Israel; They shall eat their bread with carefulness, and drink their water with astonishment, that her land may be desolate from all that is therein, because of the violence of all them that dwell therein.
\fn{all that...: Heb. the fulness thereof.}
\vs Eze 12:20 And the cities that are inhabited shall be laid waste, and the land shall be desolate; and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 12:21 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 12:22 Son of man, what \bibemph{is} that proverb \bibemph{that} ye have in the land of Israel, saying, The days are prolonged, and every vision faileth?
\vs Eze 12:23 Tell them therefore, Thus saith the Lord GOD; I will make this proverb to cease, and they shall no more use it as a proverb in Israel; but say unto them, The days are at hand, and the effect of every vision.
\vs Eze 12:24 For there shall be no more any vain vision nor flattering divination within the house of Israel.
\vs Eze 12:25 For I \bibemph{am} the \textsc{Lord}: I will speak, and the word that I shall speak shall come to pass; it shall be no more prolonged: for in your days, O rebellious house, will I say the word, and will perform it, saith the Lord GOD.
\vs Eze 12:26 Again the word of the \textsc{Lord} came to me, saying,
\vs Eze 12:27 Son of man, behold, \bibemph{they of} the house of Israel say, The vision that he seeth \bibemph{is} for many days \bibemph{to come}, and he prophesieth of the times \bibemph{that are} far off.
\vs Eze 12:28 Therefore say unto them, Thus saith the Lord GOD; There shall none of my words be prolonged any more, but the word which I have spoken shall be done, saith the Lord GOD.
\vs Eze 13:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 13:2 Son of man, prophesy against the prophets of Israel that prophesy, and say thou unto them that prophesy out of their own hearts, Hear ye the word of the \textsc{Lord};
\fn{that prophesy out...: Heb. that are prophets out of their own hearts.}
\vs Eze 13:3 Thus saith the Lord GOD; Woe unto the foolish prophets, that follow their own spirit, and have seen nothing!
\fn{follow: Heb. walk after. and...: or, and things which they have not seen.}
\vs Eze 13:4 O Israel, thy prophets are like the foxes in the deserts.
\vs Eze 13:5 Ye have not gone up into the gaps, neither made up the hedge for the house of Israel to stand in the battle in the day of the \textsc{Lord}.
\fn{gaps: or, breaches. made...: Heb. hedged the hedge.}
\vs Eze 13:6 They have seen vanity and lying divination, saying, The \textsc{Lord} saith: and the \textsc{Lord} hath not sent them: and they have made \bibemph{others} to hope that they would confirm the word.
\vs Eze 13:7 Have ye not seen a vain vision, and have ye not spoken a lying divination, whereas ye say, The \textsc{Lord} saith \bibemph{it}; albeit I have not spoken?
\vs Eze 13:8 Therefore thus saith the Lord GOD; Because ye have spoken vanity, and seen lies, therefore, behold, I \bibemph{am} against you, saith the Lord GOD.
\vs Eze 13:9 And mine hand shall be upon the prophets that see vanity, and that divine lies: they shall not be in the assembly of my people, neither shall they be written in the writing of the house of Israel, neither shall they enter into the land of Israel; and ye shall know that I \bibemph{am} the Lord GOD.
\fn{assembly: or, secret, or, counsel.}
\vs Eze 13:10 Because, even because they have seduced my people, saying, Peace; and \bibemph{there was} no peace; and one built up a wall, and, lo, others daubed it with untempered \bibemph{morter}:
\fn{a wall: or, a slight wall.}
\vs Eze 13:11 Say unto them which daub \bibemph{it} with untempered \bibemph{morter}, that it shall fall: there shall be an overflowing shower; and ye, O great hailstones, shall fall; and a stormy wind shall rend \bibemph{it}.
\vs Eze 13:12 Lo, when the wall is fallen, shall it not be said unto you, Where \bibemph{is} the daubing wherewith ye have daubed \bibemph{it}?
\vs Eze 13:13 Therefore thus saith the Lord GOD; I will even rend \bibemph{it} with a stormy wind in my fury; and there shall be an overflowing shower in mine anger, and great hailstones in \bibemph{my} fury to consume \bibemph{it}.
\vs Eze 13:14 So will I break down the wall that ye have daubed with untempered \bibemph{morter}, and bring it down to the ground, so that the foundation thereof shall be discovered, and it shall fall, and ye shall be consumed in the midst thereof: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 13:15 Thus will I accomplish my wrath upon the wall, and upon them that have daubed it with untempered \bibemph{morter}, and will say unto you, The wall \bibemph{is} no \bibemph{more}, neither they that daubed it;
\vs Eze 13:16 \bibemph{To wit}, the prophets of Israel which prophesy concerning Jerusalem, and which see visions of peace for her, and \bibemph{there is} no peace, saith the Lord GOD.
\vs Eze 13:17 Likewise, thou son of man, set thy face against the daughters of thy people, which prophesy out of their own heart; and prophesy thou against them,
\vs Eze 13:18 And say, Thus saith the Lord GOD; Woe to the \bibemph{women} that sew pillows to all armholes, and make kerchiefs upon the head of every stature to hunt souls! Will ye hunt the souls of my people, and will ye save the souls alive \bibemph{that come} unto you?
\fn{armholes: or, elbows.}
\vs Eze 13:19 And will ye pollute me among my people for handfuls of barley and for pieces of bread, to slay the souls that should not die, and to save the souls alive that should not live, by your lying to my people that hear \bibemph{your} lies?
\vs Eze 13:20 Wherefore thus saith the Lord GOD; Behold, I \bibemph{am} against your pillows, wherewith ye there hunt the souls to make \bibemph{them} fly, and I will tear them from your arms, and will let the souls go, \bibemph{even} the souls that ye hunt to make \bibemph{them} fly.
\fn{to make...: or, into gardens.}
\vs Eze 13:21 Your kerchiefs also will I tear, and deliver my people out of your hand, and they shall be no more in your hand to be hunted; and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 13:22 Because with lies ye have made the heart of the righteous sad, whom I have not made sad; and strengthened the hands of the wicked, that he should not return from his wicked way, by promising him life:
\fn{by...: or, that I should save his life: Heb. by quickening him.}
\vs Eze 13:23 Therefore ye shall see no more vanity, nor divine divinations: for I will deliver my people out of your hand: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 14:1 Then came certain of the elders of Israel unto me, and sat before me.
\vs Eze 14:2 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 14:3 Son of man, these men have set up their idols in their heart, and put the stumblingblock of their iniquity before their face: should I be enquired of at all by them?
\vs Eze 14:4 Therefore speak unto them, and say unto them, Thus saith the Lord GOD; Every man of the house of Israel that setteth up his idols in his heart, and putteth the stumblingblock of his iniquity before his face, and cometh to the prophet; I the \textsc{Lord} will answer him that cometh according to the multitude of his idols;
\vs Eze 14:5 That I may take the house of Israel in their own heart, because they are all estranged from me through their idols.
\vs Eze 14:6 Therefore say unto the house of Israel, Thus saith the Lord GOD; Repent, and turn \bibemph{yourselves} from your idols; and turn away your faces from all your abominations.
\fn{yourselves: or, others.}
\vs Eze 14:7 For every one of the house of Israel, or of the stranger that sojourneth in Israel, which separateth himself from me, and setteth up his idols in his heart, and putteth the stumblingblock of his iniquity before his face, and cometh to a prophet to enquire of him concerning me; I the \textsc{Lord} will answer him by myself:
\vs Eze 14:8 And I will set my face against that man, and will make him a sign and a proverb, and I will cut him off from the midst of my people; and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 14:9 And if the prophet be deceived when he hath spoken a thing, I the \textsc{Lord} have deceived that prophet, and I will stretch out my hand upon him, and will destroy him from the midst of my people Israel.
\vs Eze 14:10 And they shall bear the punishment of their iniquity: the punishment of the prophet shall be even as the punishment of him that seeketh \bibemph{unto him};
\vs Eze 14:11 That the house of Israel may go no more astray from me, neither be polluted any more with all their transgressions; but that they may be my people, and I may be their God, saith the Lord GOD.
\vs Eze 14:12 The word of the \textsc{Lord} came again to me, saying,
\vs Eze 14:13 Son of man, when the land sinneth against me by trespassing grievously, then will I stretch out mine hand upon it, and will break the staff of the bread thereof, and will send famine upon it, and will cut off man and beast from it:
\vs Eze 14:14 Though these three men, Noah, Daniel, and Job, were in it, they should deliver \bibemph{but} their own souls by their righteousness, saith the Lord GOD.
\vs Eze 14:15 If I cause noisome beasts to pass through the land, and they spoil it, so that it be desolate, that no man may pass through because of the beasts:
\fn{spoil: or, bereave.}
\vs Eze 14:16 \bibemph{Though} these three men \bibemph{were} in it, \bibemph{as} I live, saith the Lord GOD, they shall deliver neither sons nor daughters; they only shall be delivered, but the land shall be desolate.
\fn{in it: Heb. in the midst of it.}
\vs Eze 14:17 Or \bibemph{if} I bring a sword upon that land, and say, Sword, go through the land; so that I cut off man and beast from it:
\vs Eze 14:18 Though these three men \bibemph{were} in it, \bibemph{as} I live, saith the Lord GOD, they shall deliver neither sons nor daughters, but they only shall be delivered themselves.
\vs Eze 14:19 Or \bibemph{if} I send a pestilence into that land, and pour out my fury upon it in blood, to cut off from it man and beast:
\vs Eze 14:20 Though Noah, Daniel, and Job, \bibemph{were} in it, \bibemph{as} I live, saith the Lord GOD, they shall deliver neither son nor daughter; they shall \bibemph{but} deliver their own souls by their righteousness.
\vs Eze 14:21 For thus saith the Lord GOD; How much more when I send my four sore judgments upon Jerusalem, the sword, and the famine, and the noisome beast, and the pestilence, to cut off from it man and beast?
\fn{How...: or, Also when.}
\vs Eze 14:22 Yet, behold, therein shall be left a remnant that shall be brought forth, \bibemph{both} sons and daughters: behold, they shall come forth unto you, and ye shall see their way and their doings: and ye shall be comforted concerning the evil that I have brought upon Jerusalem, \bibemph{even} concerning all that I have brought upon it.
\vs Eze 14:23 And they shall comfort you, when ye see their ways and their doings: and ye shall know that I have not done without cause all that I have done in it, saith the Lord GOD.
\vs Eze 15:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 15:2 Son of man, What is the vine tree more than any tree, \bibemph{or than} a branch which is among the trees of the forest?
\vs Eze 15:3 Shall wood be taken thereof to do any work? or will \bibemph{men} take a pin of it to hang any vessel thereon?
\vs Eze 15:4 Behold, it is cast into the fire for fuel; the fire devoureth both the ends of it, and the midst of it is burned. Is it meet for \bibemph{any} work?
\fn{Is it...: Heb. Will it prosper?.}
\vs Eze 15:5 Behold, when it was whole, it was meet for no work: how much less shall it be meet yet for \bibemph{any} work, when the fire hath devoured it, and it is burned?
\fn{meet: Heb. made fit.}
\vs Eze 15:6 Therefore thus saith the Lord GOD; As the vine tree among the trees of the forest, which I have given to the fire for fuel, so will I give the inhabitants of Jerusalem.
\vs Eze 15:7 And I will set my face against them; they shall go out from \bibemph{one} fire, and \bibemph{another} fire shall devour them; and ye shall know that I \bibemph{am} the \textsc{Lord}, when I set my face against them.
\vs Eze 15:8 And I will make the land desolate, because they have committed a trespass, saith the Lord GOD.
\fn{committed...: Heb. trespassed a trespass.}
\vs Eze 16:1 Again the word of the \textsc{Lord} came unto me, saying,
\vs Eze 16:2 Son of man, cause Jerusalem to know her abominations,
\vs Eze 16:3 And say, Thus saith the Lord GOD unto Jerusalem; Thy birth and thy nativity \bibemph{is} of the land of Canaan; thy father \bibemph{was} an Amorite, and thy mother an Hittite.
\fn{birth: Heb. cutting out, or, habitation.}
\vs Eze 16:4 And \bibemph{as for} thy nativity, in the day thou wast born thy navel was not cut, neither wast thou washed in water to supple \bibemph{thee}; thou wast not salted at all, nor swaddled at all.
\fn{to supple...: or, when I looked upon thee.}
\vs Eze 16:5 None eye pitied thee, to do any of these unto thee, to have compassion upon thee; but thou wast cast out in the open field, to the lothing of thy person, in the day that thou wast born.
\vs Eze 16:6 And when I passed by thee, and saw thee polluted in thine own blood, I said unto thee \bibemph{when thou wast} in thy blood, Live; yea, I said unto thee \bibemph{when thou wast} in thy blood, Live.
\fn{polluted: or, trodden under foot.}
\vs Eze 16:7 I have caused thee to multiply as the bud of the field, and thou hast increased and waxen great, and thou art come to excellent ornaments: \bibemph{thy} breasts are fashioned, and thine hair is grown, whereas thou \bibemph{wast} naked and bare.
\fn{caused...: Heb. made thee a million. excellent...: Heb. ornament of ornaments.}
\vs Eze 16:8 Now when I passed by thee, and looked upon thee, behold, thy time \bibemph{was} the time of love; and I spread my skirt over thee, and covered thy nakedness: yea, I sware unto thee, and entered into a covenant with thee, saith the Lord GOD, and thou becamest mine.
\vs Eze 16:9 Then washed I thee with water; yea, I throughly washed away thy blood from thee, and I anointed thee with oil.
\fn{blood: Heb. bloods.}
\vs Eze 16:10 I clothed thee also with broidered work, and shod thee with badgers' skin, and I girded thee about with fine linen, and I covered thee with silk.
\vs Eze 16:11 I decked thee also with ornaments, and I put bracelets upon thy hands, and a chain on thy neck.
\vs Eze 16:12 And I put a jewel on thy forehead, and earrings in thine ears, and a beautiful crown upon thine head.
\fn{forehead: Heb. nose.}
\vs Eze 16:13 Thus wast thou decked with gold and silver; and thy raiment \bibemph{was of} fine linen, and silk, and broidered work; thou didst eat fine flour, and honey, and oil: and thou wast exceeding beautiful, and thou didst prosper into a kingdom.
\vs Eze 16:14 And thy renown went forth among the heathen for thy beauty: for it \bibemph{was} perfect through my comeliness, which I had put upon thee, saith the Lord GOD.
\vs Eze 16:15 But thou didst trust in thine own beauty, and playedst the harlot because of thy renown, and pouredst out thy fornications on every one that passed by; his it was.
\vs Eze 16:16 And of thy garments thou didst take, and deckedst thy high places with divers colours, and playedst the harlot thereupon: \bibemph{the like things} shall not come, neither shall it be \bibemph{so}.
\vs Eze 16:17 Thou hast also taken thy fair jewels of my gold and of my silver, which I had given thee, and madest to thyself images of men, and didst commit whoredom with them,
\fn{of men: Heb. of a male.}
\vs Eze 16:18 And tookest thy broidered garments, and coveredst them: and thou hast set mine oil and mine incense before them.
\vs Eze 16:19 My meat also which I gave thee, fine flour, and oil, and honey, \bibemph{wherewith} I fed thee, thou hast even set it before them for a sweet savour: and \bibemph{thus} it was, saith the Lord GOD.
\fn{a sweet...: Heb. a savour of rest.}
\vs Eze 16:20 Moreover thou hast taken thy sons and thy daughters, whom thou hast borne unto me, and these hast thou sacrificed unto them to be devoured. \bibemph{Is this} of thy whoredoms a small matter,
\fn{to be...: Heb. to devour.}
\vs Eze 16:21 That thou hast slain my children, and delivered them to cause them to pass through \bibemph{the fire} for them?
\vs Eze 16:22 And in all thine abominations and thy whoredoms thou hast not remembered the days of thy youth, when thou wast naked and bare, \bibemph{and} wast polluted in thy blood.
\vs Eze 16:23 And it came to pass after all thy wickedness, (woe, woe unto thee! saith the Lord GOD;)
\vs Eze 16:24 \bibemph{That} thou hast also built unto thee an eminent place, and hast made thee an high place in every street.
\fn{eminent...: or, brothel house.}
\vs Eze 16:25 Thou hast built thy high place at every head of the way, and hast made thy beauty to be abhorred, and hast opened thy feet to every one that passed by, and multiplied thy whoredoms.
\vs Eze 16:26 Thou hast also committed fornication with the Egyptians thy neighbours, great of flesh; and hast increased thy whoredoms, to provoke me to anger.
\vs Eze 16:27 Behold, therefore I have stretched out my hand over thee, and have diminished thine ordinary \bibemph{food}, and delivered thee unto the will of them that hate thee, the daughters of the Philistines, which are ashamed of thy lewd way.
\fn{daughters: or, cities.}
\vs Eze 16:28 Thou hast played the whore also with the Assyrians, because thou wast unsatiable; yea, thou hast played the harlot with them, and yet couldest not be satisfied.
\vs Eze 16:29 Thou hast moreover multiplied thy fornication in the land of Canaan unto Chaldea; and yet thou wast not satisfied herewith.
\vs Eze 16:30 How weak is thine heart, saith the Lord GOD, seeing thou doest all these \bibemph{things}, the work of an imperious whorish woman;
\vs Eze 16:31 In that thou buildest thine eminent place in the head of every way, and makest thine high place in every street; and hast not been as an harlot, in that thou scornest hire;
\fn{In that thou buildest...: or, In thy daughters is thine, etc.}
\vs Eze 16:32 \bibemph{But as} a wife that committeth adultery, \bibemph{which} taketh strangers instead of her husband!
\vs Eze 16:33 They give gifts to all whores: but thou givest thy gifts to all thy lovers, and hirest them, that they may come unto thee on every side for thy whoredom.
\fn{hirest: Heb. bribest.}
\vs Eze 16:34 And the contrary is in thee from \bibemph{other} women in thy whoredoms, whereas none followeth thee to commit whoredoms: and in that thou givest a reward, and no reward is given unto thee, therefore thou art contrary.
\vs Eze 16:35 Wherefore, O harlot, hear the word of the \textsc{Lord}:
\vs Eze 16:36 Thus saith the Lord GOD; Because thy filthiness was poured out, and thy nakedness discovered through thy whoredoms with thy lovers, and with all the idols of thy abominations, and by the blood of thy children, which thou didst give unto them;
\vs Eze 16:37 Behold, therefore I will gather all thy lovers, with whom thou hast taken pleasure, and all \bibemph{them} that thou hast loved, with all \bibemph{them} that thou hast hated; I will even gather them round about against thee, and will discover thy nakedness unto them, that they may see all thy nakedness.
\vs Eze 16:38 And I will judge thee, as women that break wedlock and shed blood are judged; and I will give thee blood in fury and jealousy.
\fn{as: Heb. with judgments of.}
\vs Eze 16:39 And I will also give thee into their hand, and they shall throw down thine eminent place, and shall break down thy high places: they shall strip thee also of thy clothes, and shall take thy fair jewels, and leave thee naked and bare.
\fn{thy fair...: Heb. instruments of thine ornament.}
\vs Eze 16:40 They shall also bring up a company against thee, and they shall stone thee with stones, and thrust thee through with their swords.
\vs Eze 16:41 And they shall burn thine houses with fire, and execute judgments upon thee in the sight of many women: and I will cause thee to cease from playing the harlot, and thou also shalt give no hire any more.
\vs Eze 16:42 So will I make my fury toward thee to rest, and my jealousy shall depart from thee, and I will be quiet, and will be no more angry.
\vs Eze 16:43 Because thou hast not remembered the days of thy youth, but hast fretted me in all these \bibemph{things}; behold, therefore I also will recompense thy way upon \bibemph{thine} head, saith the Lord GOD: and thou shalt not commit this lewdness above all thine abominations.
\vs Eze 16:44 Behold, every one that useth proverbs shall use \bibemph{this} proverb against thee, saying, As \bibemph{is} the mother, \bibemph{so is} her daughter.
\vs Eze 16:45 Thou \bibemph{art} thy mother's daughter, that lotheth her husband and her children; and thou \bibemph{art} the sister of thy sisters, which lothed their husbands and their children: your mother \bibemph{was} an Hittite, and your father an Amorite.
\vs Eze 16:46 And thine elder sister \bibemph{is} Samaria, she and her daughters that dwell at thy left hand: and thy younger sister, that dwelleth at thy right hand, \bibemph{is} Sodom and her daughters.
\fn{thy younger: Heb. lesser than thou.}
\vs Eze 16:47 Yet hast thou not walked after their ways, nor done after their abominations: but, as \bibemph{if that were} a very little \bibemph{thing}, thou wast corrupted more than they in all thy ways.
\fn{as if...: or, that was lothed as a small thing.}
\vs Eze 16:48 \bibemph{As} I live, saith the Lord GOD, Sodom thy sister hath not done, she nor her daughters, as thou hast done, thou and thy daughters.
\vs Eze 16:49 Behold, this was the iniquity of thy sister Sodom, pride, fulness of bread, and abundance of idleness was in her and in her daughters, neither did she strengthen the hand of the poor and needy.
\vs Eze 16:50 And they were haughty, and committed abomination before me: therefore I took them away as I saw \bibemph{good}.
\vs Eze 16:51 Neither hath Samaria committed half of thy sins; but thou hast multiplied thine abominations more than they, and hast justified thy sisters in all thine abominations which thou hast done.
\vs Eze 16:52 Thou also, which hast judged thy sisters, bear thine own shame for thy sins that thou hast committed more abominable than they: they are more righteous than thou: yea, be thou confounded also, and bear thy shame, in that thou hast justified thy sisters.
\vs Eze 16:53 When I shall bring again their captivity, the captivity of Sodom and her daughters, and the captivity of Samaria and her daughters, then \bibemph{will I bring again} the captivity of thy captives in the midst of them:
\vs Eze 16:54 That thou mayest bear thine own shame, and mayest be confounded in all that thou hast done, in that thou art a comfort unto them.
\vs Eze 16:55 When thy sisters, Sodom and her daughters, shall return to their former estate, and Samaria and her daughters shall return to their former estate, then thou and thy daughters shall return to your former estate.
\vs Eze 16:56 For thy sister Sodom was not mentioned by thy mouth in the day of thy pride,
\fn{mentioned: Heb. for a report, or, hearing. pride: Heb. prides, or, excellencies.}
\vs Eze 16:57 Before thy wickedness was discovered, as at the time of \bibemph{thy} reproach of the daughters of Syria, and all \bibemph{that are} round about her, the daughters of the Philistines, which despise thee round about.
\fn{Syria: Heb. Aram. despise: or, spoil.}
\vs Eze 16:58 Thou hast borne thy lewdness and thine abominations, saith the \textsc{Lord}.
\fn{borne: Heb. borne them.}
\vs Eze 16:59 For thus saith the Lord GOD; I will even deal with thee as thou hast done, which hast despised the oath in breaking the covenant.
\vs Eze 16:60 Nevertheless I will remember my covenant with thee in the days of thy youth, and I will establish unto thee an everlasting covenant.
\vs Eze 16:61 Then thou shalt remember thy ways, and be ashamed, when thou shalt receive thy sisters, thine elder and thy younger: and I will give them unto thee for daughters, but not by thy covenant.
\vs Eze 16:62 And I will establish my covenant with thee; and thou shalt know that I \bibemph{am} the \textsc{Lord}:
\vs Eze 16:63 That thou mayest remember, and be confounded, and never open thy mouth any more because of thy shame, when I am pacified toward thee for all that thou hast done, saith the Lord GOD.
\vs Eze 17:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 17:2 Son of man, put forth a riddle, and speak a parable unto the house of Israel;
\vs Eze 17:3 And say, Thus saith the Lord GOD; A great eagle with great wings, longwinged, full of feathers, which had divers colours, came unto Lebanon, and took the highest branch of the cedar:
\fn{divers...: Heb. embroidering.}
\vs Eze 17:4 He cropped off the top of his young twigs, and carried it into a land of traffick; he set it in a city of merchants.
\vs Eze 17:5 He took also of the seed of the land, and planted it in a fruitful field; he placed \bibemph{it} by great waters, \bibemph{and} set it \bibemph{as} a willow tree.
\fn{planted...: Heb. put it in a field of seed.}
\vs Eze 17:6 And it grew, and became a spreading vine of low stature, whose branches turned toward him, and the roots thereof were under him: so it became a vine, and brought forth branches, and shot forth sprigs.
\vs Eze 17:7 There was also another great eagle with great wings and many feathers: and, behold, this vine did bend her roots toward him, and shot forth her branches toward him, that he might water it by the furrows of her plantation.
\vs Eze 17:8 It was planted in a good soil by great waters, that it might bring forth branches, and that it might bear fruit, that it might be a goodly vine.
\fn{soil: Heb. field.}
\vs Eze 17:9 Say thou, Thus saith the Lord GOD; Shall it prosper? shall he not pull up the roots thereof, and cut off the fruit thereof, that it wither? it shall wither in all the leaves of her spring, even without great power or many people to pluck it up by the roots thereof.
\vs Eze 17:10 Yea, behold, \bibemph{being} planted, shall it prosper? shall it not utterly wither, when the east wind toucheth it? it shall wither in the furrows where it grew.
\vs Eze 17:11 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 17:12 Say now to the rebellious house, Know ye not what these \bibemph{things mean}? tell \bibemph{them}, Behold, the king of Babylon is come to Jerusalem, and hath taken the king thereof, and the princes thereof, and led them with him to Babylon;
\vs Eze 17:13 And hath taken of the king's seed, and made a covenant with him, and hath taken an oath of him: he hath also taken the mighty of the land:
\fn{taken an...: Heb. brought him to an oath.}
\vs Eze 17:14 That the kingdom might be base, that it might not lift itself up, \bibemph{but} that by keeping of his covenant it might stand.
\fn{but...: Heb. to keep his covenant, to stand to it.}
\vs Eze 17:15 But he rebelled against him in sending his ambassadors into Egypt, that they might give him horses and much people. Shall he prosper? shall he escape that doeth such \bibemph{things}? or shall he break the covenant, and be delivered?
\vs Eze 17:16 \bibemph{As} I live, saith the Lord GOD, surely in the place \bibemph{where} the king \bibemph{dwelleth} that made him king, whose oath he despised, and whose covenant he brake, \bibemph{even} with him in the midst of Babylon he shall die.
\vs Eze 17:17 Neither shall Pharaoh with \bibemph{his} mighty army and great company make for him in the war, by casting up mounts, and building forts, to cut off many persons:
\vs Eze 17:18 Seeing he despised the oath by breaking the covenant, when, lo, he had given his hand, and hath done all these \bibemph{things}, he shall not escape.
\vs Eze 17:19 Therefore thus saith the Lord GOD; \bibemph{As} I live, surely mine oath that he hath despised, and my covenant that he hath broken, even it will I recompense upon his own head.
\vs Eze 17:20 And I will spread my net upon him, and he shall be taken in my snare, and I will bring him to Babylon, and will plead with him there for his trespass that he hath trespassed against me.
\vs Eze 17:21 And all his fugitives with all his bands shall fall by the sword, and they that remain shall be scattered toward all winds: and ye shall know that I the \textsc{Lord} have spoken \bibemph{it}.
\vs Eze 17:22 Thus saith the Lord GOD; I will also take of the highest branch of the high cedar, and will set \bibemph{it}; I will crop off from the top of his young twigs a tender one, and will plant \bibemph{it} upon an high mountain and eminent:
\vs Eze 17:23 In the mountain of the height of Israel will I plant it: and it shall bring forth boughs, and bear fruit, and be a goodly cedar: and under it shall dwell all fowl of every wing; in the shadow of the branches thereof shall they dwell.
\vs Eze 17:24 And all the trees of the field shall know that I the \textsc{Lord} have brought down the high tree, have exalted the low tree, have dried up the green tree, and have made the dry tree to flourish: I the \textsc{Lord} have spoken and have done \bibemph{it}.
\vs Eze 18:1 The word of the \textsc{Lord} came unto me again, saying,
\vs Eze 18:2 What mean ye, that ye use this proverb concerning the land of Israel, saying, The fathers have eaten sour grapes, and the children's teeth are set on edge?
\vs Eze 18:3 \bibemph{As} I live, saith the Lord GOD, ye shall not have \bibemph{occasion} any more to use this proverb in Israel.
\vs Eze 18:4 Behold, all souls are mine; as the soul of the father, so also the soul of the son is mine: the soul that sinneth, it shall die.
\vs Eze 18:5 But if a man be just, and do that which is lawful and right,
\fn{that...: Heb. judgment and justice.}
\vs Eze 18:6 \bibemph{And} hath not eaten upon the mountains, neither hath lifted up his eyes to the idols of the house of Israel, neither hath defiled his neighbour's wife, neither hath come near to a menstruous woman,
\vs Eze 18:7 And hath not oppressed any, \bibemph{but} hath restored to the debtor his pledge, hath spoiled none by violence, hath given his bread to the hungry, and hath covered the naked with a garment;
\vs Eze 18:8 He \bibemph{that} hath not given forth upon usury, neither hath taken any increase, \bibemph{that} hath withdrawn his hand from iniquity, hath executed true judgment between man and man,
\vs Eze 18:9 Hath walked in my statutes, and hath kept my judgments, to deal truly; he \bibemph{is} just, he shall surely live, saith the Lord GOD.
\vs Eze 18:10 If he beget a son \bibemph{that is} a robber, a shedder of blood, and \bibemph{that} doeth the like to \bibemph{any} one of these \bibemph{things},
\fn{robber: or, breaker up of an house. that doeth...: or, that doeth to his brother besides any of these.}
\vs Eze 18:11 And that doeth not any of those \bibemph{duties}, but even hath eaten upon the mountains, and defiled his neighbour's wife,
\vs Eze 18:12 Hath oppressed the poor and needy, hath spoiled by violence, hath not restored the pledge, and hath lifted up his eyes to the idols, hath committed abomination,
\vs Eze 18:13 Hath given forth upon usury, and hath taken increase: shall he then live? he shall not live: he hath done all these abominations; he shall surely die; his blood shall be upon him.
\fn{blood: Heb. bloods.}
\vs Eze 18:14 Now, lo, \bibemph{if} he beget a son, that seeth all his father's sins which he hath done, and considereth, and doeth not such like,
\vs Eze 18:15 \bibemph{That} hath not eaten upon the mountains, neither hath lifted up his eyes to the idols of the house of Israel, hath not defiled his neighbour's wife,
\vs Eze 18:16 Neither hath oppressed any, hath not withholden the pledge, neither hath spoiled by violence, \bibemph{but} hath given his bread to the hungry, and hath covered the naked with a garment,
\fn{hath not...: Heb. hath not pledged the pledge, or, taken to pledge.}
\vs Eze 18:17 \bibemph{That} hath taken off his hand from the poor, \bibemph{that} hath not received usury nor increase, hath executed my judgments, hath walked in my statutes; he shall not die for the iniquity of his father, he shall surely live.
\vs Eze 18:18 \bibemph{As for} his father, because he cruelly oppressed, spoiled his brother by violence, and did \bibemph{that} which \bibemph{is} not good among his people, lo, even he shall die in his iniquity.
\vs Eze 18:19 Yet say ye, Why? doth not the son bear the iniquity of the father? When the son hath done that which is lawful and right, \bibemph{and} hath kept all my statutes, and hath done them, he shall surely live.
\vs Eze 18:20 The soul that sinneth, it shall die. The son shall not bear the iniquity of the father, neither shall the father bear the iniquity of the son: the righteousness of the righteous shall be upon him, and the wickedness of the wicked shall be upon him.
\vs Eze 18:21 But if the wicked will turn from all his sins that he hath committed, and keep all my statutes, and do that which is lawful and right, he shall surely live, he shall not die.
\vs Eze 18:22 All his transgressions that he hath committed, they shall not be mentioned unto him: in his righteousness that he hath done he shall live.
\vs Eze 18:23 Have I any pleasure at all that the wicked should die? saith the Lord GOD: \bibemph{and} not that he should return from his ways, and live?
\vs Eze 18:24 But when the righteous turneth away from his righteousness, and committeth iniquity, \bibemph{and} doeth according to all the abominations that the wicked \bibemph{man} doeth, shall he live? All his righteousness that he hath done shall not be mentioned: in his trespass that he hath trespassed, and in his sin that he hath sinned, in them shall he die.
\vs Eze 18:25 Yet ye say, The way of the Lord is not equal. Hear now, O house of Israel; Is not my way equal? are not your ways unequal?
\vs Eze 18:26 When a righteous \bibemph{man} turneth away from his righteousness, and committeth iniquity, and dieth in them; for his iniquity that he hath done shall he die.
\vs Eze 18:27 Again, when the wicked \bibemph{man} turneth away from his wickedness that he hath committed, and doeth that which is lawful and right, he shall save his soul alive.
\vs Eze 18:28 Because he considereth, and turneth away from all his transgressions that he hath committed, he shall surely live, he shall not die.
\vs Eze 18:29 Yet saith the house of Israel, The way of the Lord is not equal. O house of Israel, are not my ways equal? are not your ways unequal?
\vs Eze 18:30 Therefore I will judge you, O house of Israel, every one according to his ways, saith the Lord GOD. Repent, and turn \bibemph{yourselves} from all your transgressions; so iniquity shall not be your ruin.
\fn{yourselves: or, others.}
\vs Eze 18:31 Cast away from you all your transgressions, whereby ye have transgressed; and make you a new heart and a new spirit: for why will ye die, O house of Israel?
\vs Eze 18:32 For I have no pleasure in the death of him that dieth, saith the Lord GOD: wherefore turn \bibemph{yourselves}, and live ye.
\fn{yourselves: or, others.}
\vs Eze 19:1 Moreover take thou up a lamentation for the princes of Israel,
\vs Eze 19:2 And say, What \bibemph{is} thy mother? A lioness: she lay down among lions, she nourished her whelps among young lions.
\vs Eze 19:3 And she brought up one of her whelps: it became a young lion, and it learned to catch the prey; it devoured men.
\vs Eze 19:4 The nations also heard of him; he was taken in their pit, and they brought him with chains unto the land of Egypt.
\vs Eze 19:5 Now when she saw that she had waited, \bibemph{and} her hope was lost, then she took another of her whelps, \bibemph{and} made him a young lion.
\vs Eze 19:6 And he went up and down among the lions, he became a young lion, and learned to catch the prey, \bibemph{and} devoured men.
\vs Eze 19:7 And he knew their desolate palaces, and he laid waste their cities; and the land was desolate, and the fulness thereof, by the noise of his roaring.
\fn{their desolate...: or, their widows. the fulness...: or, all it containeth.}
\vs Eze 19:8 Then the nations set against him on every side from the provinces, and spread their net over him: he was taken in their pit.
\vs Eze 19:9 And they put him in ward in chains, and brought him to the king of Babylon: they brought him into holds, that his voice should no more be heard upon the mountains of Israel.
\fn{in chains: or, in hooks.}
\vs Eze 19:10 Thy mother \bibemph{is} like a vine in thy blood, planted by the waters: she was fruitful and full of branches by reason of many waters.
\fn{in...: or, in thy quietness, or, in thy likeness.}
\vs Eze 19:11 And she had strong rods for the sceptres of them that bare rule, and her stature was exalted among the thick branches, and she appeared in her height with the multitude of her branches.
\vs Eze 19:12 But she was plucked up in fury, she was cast down to the ground, and the east wind dried up her fruit: her strong rods were broken and withered; the fire consumed them.
\vs Eze 19:13 And now she \bibemph{is} planted in the wilderness, in a dry and thirsty ground.
\vs Eze 19:14 And fire is gone out of a rod of her branches, \bibemph{which} hath devoured her fruit, so that she hath no strong rod \bibemph{to be} a sceptre to rule. This \bibemph{is} a lamentation, and shall be for a lamentation.
\vs Eze 20:1 And it came to pass in the seventh year, in the fifth \bibemph{month}, the tenth \bibemph{day} of the month, \bibemph{that} certain of the elders of Israel came to enquire of the \textsc{Lord}, and sat before me.
\vs Eze 20:2 Then came the word of the \textsc{Lord} unto me, saying,
\vs Eze 20:3 Son of man, speak unto the elders of Israel, and say unto them, Thus saith the Lord GOD; Are ye come to enquire of me? \bibemph{As} I live, saith the Lord GOD, I will not be enquired of by you.
\vs Eze 20:4 Wilt thou judge them, son of man, wilt thou judge \bibemph{them}? cause them to know the abominations of their fathers:
\fn{Wilt...: or, Wilt thou plead for them.}
\vs Eze 20:5 And say unto them, Thus saith the Lord GOD; In the day when I chose Israel, and lifted up mine hand unto the seed of the house of Jacob, and made myself known unto them in the land of Egypt, when I lifted up mine hand unto them, saying, I \bibemph{am} the \textsc{Lord} your God;
\fn{lifted...: or, sware.}
\vs Eze 20:6 In the day \bibemph{that} I lifted up mine hand unto them, to bring them forth of the land of Egypt into a land that I had espied for them, flowing with milk and honey, which \bibemph{is} the glory of all lands:
\vs Eze 20:7 Then said I unto them, Cast ye away every man the abominations of his eyes, and defile not yourselves with the idols of Egypt: I \bibemph{am} the \textsc{Lord} your God.
\vs Eze 20:8 But they rebelled against me, and would not hearken unto me: they did not every man cast away the abominations of their eyes, neither did they forsake the idols of Egypt: then I said, I will pour out my fury upon them, to accomplish my anger against them in the midst of the land of Egypt.
\vs Eze 20:9 But I wrought for my name's sake, that it should not be polluted before the heathen, among whom they \bibemph{were}, in whose sight I made myself known unto them, in bringing them forth out of the land of Egypt.
\vs Eze 20:10 Wherefore I caused them to go forth out of the land of Egypt, and brought them into the wilderness.
\vs Eze 20:11 And I gave them my statutes, and shewed them my judgments, which \bibemph{if} a man do, he shall even live in them.
\fn{shewed...: Heb. made them to know.}
\vs Eze 20:12 Moreover also I gave them my sabbaths, to be a sign between me and them, that they might know that I \bibemph{am} the \textsc{Lord} that sanctify them.
\vs Eze 20:13 But the house of Israel rebelled against me in the wilderness: they walked not in my statutes, and they despised my judgments, which \bibemph{if} a man do, he shall even live in them; and my sabbaths they greatly polluted: then I said, I would pour out my fury upon them in the wilderness, to consume them.
\vs Eze 20:14 But I wrought for my name's sake, that it should not be polluted before the heathen, in whose sight I brought them out.
\vs Eze 20:15 Yet also I lifted up my hand unto them in the wilderness, that I would not bring them into the land which I had given \bibemph{them}, flowing with milk and honey, which \bibemph{is} the glory of all lands;
\vs Eze 20:16 Because they despised my judgments, and walked not in my statutes, but polluted my sabbaths: for their heart went after their idols.
\vs Eze 20:17 Nevertheless mine eye spared them from destroying them, neither did I make an end of them in the wilderness.
\vs Eze 20:18 But I said unto their children in the wilderness, Walk ye not in the statutes of your fathers, neither observe their judgments, nor defile yourselves with their idols:
\vs Eze 20:19 I \bibemph{am} the \textsc{Lord} your God; walk in my statutes, and keep my judgments, and do them;
\vs Eze 20:20 And hallow my sabbaths; and they shall be a sign between me and you, that ye may know that I \bibemph{am} the \textsc{Lord} your God.
\vs Eze 20:21 Notwithstanding the children rebelled against me: they walked not in my statutes, neither kept my judgments to do them, which \bibemph{if} a man do, he shall even live in them; they polluted my sabbaths: then I said, I would pour out my fury upon them, to accomplish my anger against them in the wilderness.
\vs Eze 20:22 Nevertheless I withdrew mine hand, and wrought for my name's sake, that it should not be polluted in the sight of the heathen, in whose sight I brought them forth.
\vs Eze 20:23 I lifted up mine hand unto them also in the wilderness, that I would scatter them among the heathen, and disperse them through the countries;
\vs Eze 20:24 Because they had not executed my judgments, but had despised my statutes, and had polluted my sabbaths, and their eyes were after their fathers' idols.
\vs Eze 20:25 Wherefore I gave them also statutes \bibemph{that were} not good, and judgments whereby they should not live;
\vs Eze 20:26 And I polluted them in their own gifts, in that they caused to pass through \bibemph{the fire} all that openeth the womb, that I might make them desolate, to the end that they might know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 20:27 Therefore, son of man, speak unto the house of Israel, and say unto them, Thus saith the Lord GOD; Yet in this your fathers have blasphemed me, in that they have committed a trespass against me.
\fn{committed...: Heb. trespassed a trespass.}
\vs Eze 20:28 \bibemph{For} when I had brought them into the land, \bibemph{for} the which I lifted up mine hand to give it to them, then they saw every high hill, and all the thick trees, and they offered there their sacrifices, and there they presented the provocation of their offering: there also they made their sweet savour, and poured out there their drink offerings.
\vs Eze 20:29 Then I said unto them, What \bibemph{is} the high place whereunto ye go? And the name thereof is called Bamah unto this day.
\fn{I said...: or, I told them what the high place was, or, Bamah.}
\vs Eze 20:30 Wherefore say unto the house of Israel, Thus saith the Lord GOD; Are ye polluted after the manner of your fathers? and commit ye whoredom after their abominations?
\vs Eze 20:31 For when ye offer your gifts, when ye make your sons to pass through the fire, ye pollute yourselves with all your idols, even unto this day: and shall I be enquired of by you, O house of Israel? \bibemph{As} I live, saith the Lord GOD, I will not be enquired of by you.
\vs Eze 20:32 And that which cometh into your mind shall not be at all, that ye say, We will be as the heathen, as the families of the countries, to serve wood and stone.
\vs Eze 20:33 \bibemph{As} I live, saith the Lord GOD, surely with a mighty hand, and with a stretched out arm, and with fury poured out, will I rule over you:
\vs Eze 20:34 And I will bring you out from the people, and will gather you out of the countries wherein ye are scattered, with a mighty hand, and with a stretched out arm, and with fury poured out.
\vs Eze 20:35 And I will bring you into the wilderness of the people, and there will I plead with you face to face.
\vs Eze 20:36 Like as I pleaded with your fathers in the wilderness of the land of Egypt, so will I plead with you, saith the Lord GOD.
\vs Eze 20:37 And I will cause you to pass under the rod, and I will bring you into the bond of the covenant:
\fn{bond: or, delivering.}
\vs Eze 20:38 And I will purge out from among you the rebels, and them that transgress against me: I will bring them forth out of the country where they sojourn, and they shall not enter into the land of Israel: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 20:39 As for you, O house of Israel, thus saith the Lord GOD; Go ye, serve ye every one his idols, and hereafter \bibemph{also}, if ye will not hearken unto me: but pollute ye my holy name no more with your gifts, and with your idols.
\vs Eze 20:40 For in mine holy mountain, in the mountain of the height of Israel, saith the Lord GOD, there shall all the house of Israel, all of them in the land, serve me: there will I accept them, and there will I require your offerings, and the firstfruits of your oblations, with all your holy things.
\fn{firstfruits: or, chief.}
\vs Eze 20:41 I will accept you with your sweet savour, when I bring you out from the people, and gather you out of the countries wherein ye have been scattered; and I will be sanctified in you before the heathen.
\fn{sweet...: Heb. savour of rest.}
\vs Eze 20:42 And ye shall know that I \bibemph{am} the \textsc{Lord}, when I shall bring you into the land of Israel, into the country \bibemph{for} the which I lifted up mine hand to give it to your fathers.
\vs Eze 20:43 And there shall ye remember your ways, and all your doings, wherein ye have been defiled; and ye shall lothe yourselves in your own sight for all your evils that ye have committed.
\vs Eze 20:44 And ye shall know that I \bibemph{am} the \textsc{Lord}, when I have wrought with you for my name's sake, not according to your wicked ways, nor according to your corrupt doings, O ye house of Israel, saith the Lord GOD.
\vs Eze 20:45 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 20:46 Son of man, set thy face toward the south, and drop \bibemph{thy word} toward the south, and prophesy against the forest of the south field;
\vs Eze 20:47 And say to the forest of the south, Hear the word of the \textsc{Lord}; Thus saith the Lord GOD; Behold, I will kindle a fire in thee, and it shall devour every green tree in thee, and every dry tree: the flaming flame shall not be quenched, and all faces from the south to the north shall be burned therein.
\vs Eze 20:48 And all flesh shall see that I the \textsc{Lord} have kindled it: it shall not be quenched.
\vs Eze 20:49 Then said I, Ah Lord GOD! they say of me, Doth he not speak parables?
\vs Eze 21:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 21:2 Son of man, set thy face toward Jerusalem, and drop \bibemph{thy word} toward the holy places, and prophesy against the land of Israel,
\vs Eze 21:3 And say to the land of Israel, Thus saith the \textsc{Lord}; Behold, I \bibemph{am} against thee, and will draw forth my sword out of his sheath, and will cut off from thee the righteous and the wicked.
\vs Eze 21:4 Seeing then that I will cut off from thee the righteous and the wicked, therefore shall my sword go forth out of his sheath against all flesh from the south to the north:
\vs Eze 21:5 That all flesh may know that I the \textsc{Lord} have drawn forth my sword out of his sheath: it shall not return any more.
\vs Eze 21:6 Sigh therefore, thou son of man, with the breaking of \bibemph{thy} loins; and with bitterness sigh before their eyes.
\vs Eze 21:7 And it shall be, when they say unto thee, Wherefore sighest thou? that thou shalt answer, For the tidings; because it cometh: and every heart shall melt, and all hands shall be feeble, and every spirit shall faint, and all knees shall be weak \bibemph{as} water: behold, it cometh, and shall be brought to pass, saith the Lord GOD.
\fn{shall be weak...: Heb. shall go into water.}
\vs Eze 21:8 Again the word of the \textsc{Lord} came unto me, saying,
\vs Eze 21:9 Son of man, prophesy, and say, Thus saith the \textsc{Lord}; Say, A sword, a sword is sharpened, and also furbished:
\vs Eze 21:10 It is sharpened to make a sore slaughter; it is furbished that it may glitter: should we then make mirth? it contemneth the rod of my son, \bibemph{as} every tree.
\fn{it contemneth...: or, it is the rod of my son, it despiseth every tree.}
\vs Eze 21:11 And he hath given it to be furbished, that it may be handled: this sword is sharpened, and it is furbished, to give it into the hand of the slayer.
\vs Eze 21:12 Cry and howl, son of man: for it shall be upon my people, it \bibemph{shall be} upon all the princes of Israel: terrors by reason of the sword shall be upon my people: smite therefore upon \bibemph{thy} thigh.
\fn{terrors...: or, they are thrust down to the sword with my people.}
\vs Eze 21:13 Because \bibemph{it is} a trial, and what if \bibemph{the sword} contemn even the rod? it shall be no \bibemph{more}, saith the Lord GOD.
\fn{Because...: or, When the trial hath been, what then? shall they not also belong to the despising rod?.}
\vs Eze 21:14 Thou therefore, son of man, prophesy, and smite \bibemph{thine} hands together, and let the sword be doubled the third time, the sword of the slain: it \bibemph{is} the sword of the great \bibemph{men that are} slain, which entereth into their privy chambers.
\fn{hands...: Heb. hand to hand.}
\vs Eze 21:15 I have set the point of the sword against all their gates, that \bibemph{their} heart may faint, and \bibemph{their} ruins be multiplied: ah! \bibemph{it is} made bright, \bibemph{it is} wrapped up for the slaughter.
\fn{point: or, glittering, or, fear. wrapped up: or, sharpened.}
\vs Eze 21:16 Go thee one way or other, \bibemph{either} on the right hand, \bibemph{or} on the left, whithersoever thy face \bibemph{is} set.
\fn{or on...: Heb. set thyself, take the left hand.}
\vs Eze 21:17 I will also smite mine hands together, and I will cause my fury to rest: I the \textsc{Lord} have said \bibemph{it}.
\vs Eze 21:18 The word of the \textsc{Lord} came unto me again, saying,
\vs Eze 21:19 Also, thou son of man, appoint thee two ways, that the sword of the king of Babylon may come: both twain shall come forth out of one land: and choose thou a place, choose \bibemph{it} at the head of the way to the city.
\vs Eze 21:20 Appoint a way, that the sword may come to Rabbath of the Ammonites, and to Judah in Jerusalem the defenced.
\vs Eze 21:21 For the king of Babylon stood at the parting of the way, at the head of the two ways, to use divination: he made \bibemph{his} arrows bright, he consulted with images, he looked in the liver.
\fn{parting of: Heb. mother of. arrows: or, knives. images: Heb. teraphim.}
\vs Eze 21:22 At his right hand was the divination for Jerusalem, to appoint captains, to open the mouth in the slaughter, to lift up the voice with shouting, to appoint \bibemph{battering} rams against the gates, to cast a mount, \bibemph{and} to build a fort.
\fn{captains: or, battering rams: Heb. rams.}
\vs Eze 21:23 And it shall be unto them as a false divination in their sight, to them that have sworn oaths: but he will call to remembrance the iniquity, that they may be taken.
\fn{to them...: or, for the oaths made unto them.}
\vs Eze 21:24 Therefore thus saith the Lord GOD; Because ye have made your iniquity to be remembered, in that your transgressions are discovered, so that in all your doings your sins do appear; because, \bibemph{I say}, that ye are come to remembrance, ye shall be taken with the hand.
\vs Eze 21:25 And thou, profane wicked prince of Israel, whose day is come, when iniquity \bibemph{shall have} an end,
\vs Eze 21:26 Thus saith the Lord GOD; Remove the diadem, and take off the crown: this \bibemph{shall} not \bibemph{be} the same: exalt \bibemph{him that is} low, and abase \bibemph{him that is} high.
\vs Eze 21:27 I will overturn, overturn, overturn, it: and it shall be no \bibemph{more}, until he come whose right it is; and I will give it \bibemph{him}.
\fn{I will overturn...: Heb. Perverted, perverted, perverted, will I make it.}
\vs Eze 21:28 And thou, son of man, prophesy and say, Thus saith the Lord GOD concerning the Ammonites, and concerning their reproach; even say thou, The sword, the sword \bibemph{is} drawn: for the slaughter \bibemph{it is} furbished, to consume because of the glittering:
\vs Eze 21:29 Whiles they see vanity unto thee, whiles they divine a lie unto thee, to bring thee upon the necks of \bibemph{them that are} slain, of the wicked, whose day is come, when their iniquity \bibemph{shall have} an end.
\vs Eze 21:30 Shall I cause \bibemph{it} to return into his sheath? I will judge thee in the place where thou wast created, in the land of thy nativity.
\fn{Shall...: or, Cause it to return.}
\vs Eze 21:31 And I will pour out mine indignation upon thee, I will blow against thee in the fire of my wrath, and deliver thee into the hand of brutish men, \bibemph{and} skilful to destroy.
\fn{brutish: or, burning.}
\vs Eze 21:32 Thou shalt be for fuel to the fire; thy blood shall be in the midst of the land; thou shalt be no \bibemph{more} remembered: for I the \textsc{Lord} have spoken \bibemph{it}.
\vs Eze 22:1 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 22:2 Now, thou son of man, wilt thou judge, wilt thou judge the bloody city? yea, thou shalt shew her all her abominations.
\fn{judge: or, plead for. bloody...: Heb. city of bloods?. shew her: Heb. make her know.}
\vs Eze 22:3 Then say thou, Thus saith the Lord GOD, The city sheddeth blood in the midst of it, that her time may come, and maketh idols against herself to defile herself.
\vs Eze 22:4 Thou art become guilty in thy blood that thou hast shed; and hast defiled thyself in thine idols which thou hast made; and thou hast caused thy days to draw near, and art come \bibemph{even} unto thy years: therefore have I made thee a reproach unto the heathen, and a mocking to all countries.
\vs Eze 22:5 \bibemph{Those that be} near, and \bibemph{those that be} far from thee, shall mock thee, \bibemph{which art} infamous \bibemph{and} much vexed.
\fn{infamous...: Heb. polluted of name, much in vexation.}
\vs Eze 22:6 Behold, the princes of Israel, every one were in thee to their power to shed blood.
\fn{power: Heb. arm.}
\vs Eze 22:7 In thee have they set light by father and mother: in the midst of thee have they dealt by oppression with the stranger: in thee have they vexed the fatherless and the widow.
\fn{oppression: or, deceit.}
\vs Eze 22:8 Thou hast despised mine holy things, and hast profaned my sabbaths.
\vs Eze 22:9 In thee are men that carry tales to shed blood: and in thee they eat upon the mountains: in the midst of thee they commit lewdness.
\fn{that...: Heb. of slanders.}
\vs Eze 22:10 In thee have they discovered their fathers' nakedness: in thee have they humbled her that was set apart for pollution.
\vs Eze 22:11 And one hath committed abomination with his neighbour's wife; and another hath lewdly defiled his daughter in law; and another in thee hath humbled his sister, his father's daughter.
\fn{one: or, every one. another: or, every one. lewdly: or, by lewdness.}
\vs Eze 22:12 In thee have they taken gifts to shed blood; thou hast taken usury and increase, and thou hast greedily gained of thy neighbours by extortion, and hast forgotten me, saith the Lord GOD.
\vs Eze 22:13 Behold, therefore I have smitten mine hand at thy dishonest gain which thou hast made, and at thy blood which hath been in the midst of thee.
\vs Eze 22:14 Can thine heart endure, or can thine hands be strong, in the days that I shall deal with thee? I the \textsc{Lord} have spoken \bibemph{it}, and will do \bibemph{it}.
\vs Eze 22:15 And I will scatter thee among the heathen, and disperse thee in the countries, and will consume thy filthiness out of thee.
\vs Eze 22:16 And thou shalt take thine inheritance in thyself in the sight of the heathen, and thou shalt know that I \bibemph{am} the \textsc{Lord}.
\fn{shalt take...: or, shalt be profaned.}
\vs Eze 22:17 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 22:18 Son of man, the house of Israel is to me become dross: all they \bibemph{are} brass, and tin, and iron, and lead, in the midst of the furnace; they are \bibemph{even} the dross of silver.
\fn{dross of silver: Heb. drosses, etc.}
\vs Eze 22:19 Therefore thus saith the Lord GOD; Because ye are all become dross, behold, therefore I will gather you into the midst of Jerusalem.
\vs Eze 22:20 \bibemph{As} they gather silver, and brass, and iron, and lead, and tin, into the midst of the furnace, to blow the fire upon it, to melt \bibemph{it}; so will I gather \bibemph{you} in mine anger and in my fury, and I will leave \bibemph{you there}, and melt you.
\fn{As...: Heb. According to the gathering.}
\vs Eze 22:21 Yea, I will gather you, and blow upon you in the fire of my wrath, and ye shall be melted in the midst thereof.
\vs Eze 22:22 As silver is melted in the midst of the furnace, so shall ye be melted in the midst thereof; and ye shall know that I the \textsc{Lord} have poured out my fury upon you.
\vs Eze 22:23 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 22:24 Son of man, say unto her, Thou \bibemph{art} the land that is not cleansed, nor rained upon in the day of indignation.
\vs Eze 22:25 \bibemph{There is} a conspiracy of her prophets in the midst thereof, like a roaring lion ravening the prey; they have devoured souls; they have taken the treasure and precious things; they have made her many widows in the midst thereof.
\vs Eze 22:26 Her priests have violated my law, and have profaned mine holy things: they have put no difference between the holy and profane, neither have they shewed \bibemph{difference} between the unclean and the clean, and have hid their eyes from my sabbaths, and I am profaned among them.
\fn{violated: Heb. offered violence to.}
\vs Eze 22:27 Her princes in the midst thereof \bibemph{are} like wolves ravening the prey, to shed blood, \bibemph{and} to destroy souls, to get dishonest gain.
\vs Eze 22:28 And her prophets have daubed them with untempered \bibemph{morter}, seeing vanity, and divining lies unto them, saying, Thus saith the Lord GOD, when the \textsc{Lord} hath not spoken.
\vs Eze 22:29 The people of the land have used oppression, and exercised robbery, and have vexed the poor and needy: yea, they have oppressed the stranger wrongfully.
\fn{oppression: or, deceit. wrongfully: Heb. without right.}
\vs Eze 22:30 And I sought for a man among them, that should make up the hedge, and stand in the gap before me for the land, that I should not destroy it: but I found none.
\vs Eze 22:31 Therefore have I poured out mine indignation upon them; I have consumed them with the fire of my wrath: their own way have I recompensed upon their heads, saith the Lord GOD.
\vs Eze 23:1 The word of the \textsc{Lord} came again unto me, saying,
\vs Eze 23:2 Son of man, there were two women, the daughters of one mother:
\vs Eze 23:3 And they committed whoredoms in Egypt; they committed whoredoms in their youth: there were their breasts pressed, and there they bruised the teats of their virginity.
\vs Eze 23:4 And the names of them \bibemph{were} Aholah the elder, and Aholibah her sister: and they were mine, and they bare sons and daughters. Thus \bibemph{were} their names; Samaria \bibemph{is} Aholah, and Jerusalem Aholibah.
\fn{Aholah: that is, His tent, or, tabernacle. Aholibah: that is, My tabernacle in her.}
\vs Eze 23:5 And Aholah played the harlot when she was mine; and she doted on her lovers, on the Assyrians \bibemph{her} neighbours,
\vs Eze 23:6 \bibemph{Which were} clothed with blue, captains and rulers, all of them desirable young men, horsemen riding upon horses.
\vs Eze 23:7 Thus she committed her whoredoms with them, with all them \bibemph{that were} the chosen men of Assyria, and with all on whom she doted: with all their idols she defiled herself.
\fn{committed...: Heb. bestowed her whoredoms upon them. the chosen...: Heb. the choice of the children of Asshur.}
\vs Eze 23:8 Neither left she her whoredoms \bibemph{brought} from Egypt: for in her youth they lay with her, and they bruised the breasts of her virginity, and poured their whoredom upon her.
\vs Eze 23:9 Wherefore I have delivered her into the hand of her lovers, into the hand of the Assyrians, upon whom she doted.
\vs Eze 23:10 These discovered her nakedness: they took her sons and her daughters, and slew her with the sword: and she became famous among women; for they had executed judgment upon her.
\fn{famous: Heb. a name.}
\vs Eze 23:11 And when her sister Aholibah saw \bibemph{this}, she was more corrupt in her inordinate love than she, and in her whoredoms more than her sister in \bibemph{her} whoredoms.
\fn{she was...: Heb. she corrupted her inordinate love more than, etc. more than...: Heb. more than the whoredoms of her sister.}
\vs Eze 23:12 She doted upon the Assyrians \bibemph{her} neighbours, captains and rulers clothed most gorgeously, horsemen riding upon horses, all of them desirable young men.
\vs Eze 23:13 Then I saw that she was defiled, \bibemph{that} they \bibemph{took} both one way,
\vs Eze 23:14 And \bibemph{that} she increased her whoredoms: for when she saw men pourtrayed upon the wall, the images of the Chaldeans pourtrayed with vermilion,
\vs Eze 23:15 Girded with girdles upon their loins, exceeding in dyed attire upon their heads, all of them princes to look to, after the manner of the Babylonians of Chaldea, the land of their nativity:
\vs Eze 23:16 And as soon as she saw them with her eyes, she doted upon them, and sent messengers unto them into Chaldea.
\fn{as soon...: Heb. at the sight of her eyes.}
\vs Eze 23:17 And the Babylonians came to her into the bed of love, and they defiled her with their whoredom, and she was polluted with them, and her mind was alienated from them.
\fn{Babylonians: Heb. children of Babel. alienated: Heb. loosed, or, disjointed.}
\vs Eze 23:18 So she discovered her whoredoms, and discovered her nakedness: then my mind was alienated from her, like as my mind was alienated from her sister.
\vs Eze 23:19 Yet she multiplied her whoredoms, in calling to remembrance the days of her youth, wherein she had played the harlot in the land of Egypt.
\vs Eze 23:20 For she doted upon their paramours, whose flesh \bibemph{is as} the flesh of asses, and whose issue \bibemph{is like} the issue of horses.
\vs Eze 23:21 Thus thou calledst to remembrance the lewdness of thy youth, in bruising thy teats by the Egyptians for the paps of thy youth.
\vs Eze 23:22 Therefore, O Aholibah, thus saith the Lord GOD; Behold, I will raise up thy lovers against thee, from whom thy mind is alienated, and I will bring them against thee on every side;
\vs Eze 23:23 The Babylonians, and all the Chaldeans, Pekod, and Shoa, and Koa, \bibemph{and} all the Assyrians with them: all of them desirable young men, captains and rulers, great lords and renowned, all of them riding upon horses.
\vs Eze 23:24 And they shall come against thee with chariots, wagons, and wheels, and with an assembly of people, \bibemph{which} shall set against thee buckler and shield and helmet round about: and I will set judgment before them, and they shall judge thee according to their judgments.
\vs Eze 23:25 And I will set my jealousy against thee, and they shall deal furiously with thee: they shall take away thy nose and thine ears; and thy remnant shall fall by the sword: they shall take thy sons and thy daughters; and thy residue shall be devoured by the fire.
\vs Eze 23:26 They shall also strip thee out of thy clothes, and take away thy fair jewels.
\fn{fair...: Heb. instruments of thy decking.}
\vs Eze 23:27 Thus will I make thy lewdness to cease from thee, and thy whoredom \bibemph{brought} from the land of Egypt: so that thou shalt not lift up thine eyes unto them, nor remember Egypt any more.
\vs Eze 23:28 For thus saith the Lord GOD; Behold, I will deliver thee into the hand \bibemph{of them} whom thou hatest, into the hand \bibemph{of them} from whom thy mind is alienated:
\vs Eze 23:29 And they shall deal with thee hatefully, and shall take away all thy labour, and shall leave thee naked and bare: and the nakedness of thy whoredoms shall be discovered, both thy lewdness and thy whoredoms.
\vs Eze 23:30 I will do these \bibemph{things} unto thee, because thou hast gone a whoring after the heathen, \bibemph{and} because thou art polluted with their idols.
\vs Eze 23:31 Thou hast walked in the way of thy sister; therefore will I give her cup into thine hand.
\vs Eze 23:32 Thus saith the Lord GOD; Thou shalt drink of thy sister's cup deep and large: thou shalt be laughed to scorn and had in derision; it containeth much.
\vs Eze 23:33 Thou shalt be filled with drunkenness and sorrow, with the cup of astonishment and desolation, with the cup of thy sister Samaria.
\vs Eze 23:34 Thou shalt even drink it and suck \bibemph{it} out, and thou shalt break the sherds thereof, and pluck off thine own breasts: for I have spoken \bibemph{it}, saith the Lord GOD.
\vs Eze 23:35 Therefore thus saith the Lord GOD; Because thou hast forgotten me, and cast me behind thy back, therefore bear thou also thy lewdness and thy whoredoms.
\vs Eze 23:36 The \textsc{Lord} said moreover unto me; Son of man, wilt thou judge Aholah and Aholibah? yea, declare unto them their abominations;
\fn{judge: or, plead for.}
\vs Eze 23:37 That they have committed adultery, and blood \bibemph{is} in their hands, and with their idols have they committed adultery, and have also caused their sons, whom they bare unto me, to pass for them through \bibemph{the fire}, to devour \bibemph{them}.
\vs Eze 23:38 Moreover this they have done unto me: they have defiled my sanctuary in the same day, and have profaned my sabbaths.
\vs Eze 23:39 For when they had slain their children to their idols, then they came the same day into my sanctuary to profane it; and, lo, thus have they done in the midst of mine house.
\vs Eze 23:40 And furthermore, that ye have sent for men to come from far, unto whom a messenger \bibemph{was} sent; and, lo, they came: for whom thou didst wash thyself, paintedst thy eyes, and deckedst thyself with ornaments,
\fn{to come: Heb. coming.}
\vs Eze 23:41 And satest upon a stately bed, and a table prepared before it, whereupon thou hast set mine incense and mine oil.
\fn{stately: Heb. honourable.}
\vs Eze 23:42 And a voice of a multitude being at ease \bibemph{was} with her: and with the men of the common sort \bibemph{were} brought Sabeans from the wilderness, which put bracelets upon their hands, and beautiful crowns upon their heads.
\fn{of the...: Heb. of the multitude of men. Sabeans: or, drunkards.}
\vs Eze 23:43 Then said I unto \bibemph{her that was} old in adulteries, Will they now commit whoredoms with her, and she \bibemph{with them}?
\fn{whoredoms...: Heb. her whoredoms.}
\vs Eze 23:44 Yet they went in unto her, as they go in unto a woman that playeth the harlot: so went they in unto Aholah and unto Aholibah, the lewd women.
\vs Eze 23:45 And the righteous men, they shall judge them after the manner of adulteresses, and after the manner of women that shed blood; because they \bibemph{are} adulteresses, and blood \bibemph{is} in their hands.
\vs Eze 23:46 For thus saith the Lord GOD; I will bring up a company upon them, and will give them to be removed and spoiled.
\fn{to be...: Heb. for a removing and spoil.}
\vs Eze 23:47 And the company shall stone them with stones, and dispatch them with their swords; they shall slay their sons and their daughters, and burn up their houses with fire.
\fn{dispatch...: or, single them out.}
\vs Eze 23:48 Thus will I cause lewdness to cease out of the land, that all women may be taught not to do after your lewdness.
\vs Eze 23:49 And they shall recompense your lewdness upon you, and ye shall bear the sins of your idols: and ye shall know that I \bibemph{am} the Lord GOD.
\vs Eze 24:1 Again in the ninth year, in the tenth month, in the tenth \bibemph{day} of the month, the word of the \textsc{Lord} came unto me, saying,
\vs Eze 24:2 Son of man, write thee the name of the day, \bibemph{even} of this same day: the king of Babylon set himself against Jerusalem this same day.
\vs Eze 24:3 And utter a parable unto the rebellious house, and say unto them, Thus saith the Lord GOD; Set on a pot, set \bibemph{it} on, and also pour water into it:
\vs Eze 24:4 Gather the pieces thereof into it, \bibemph{even} every good piece, the thigh, and the shoulder; fill \bibemph{it} with the choice bones.
\vs Eze 24:5 Take the choice of the flock, and burn also the bones under it, \bibemph{and} make it boil well, and let them seethe the bones of it therein.
\fn{burn: or, heap.}
\vs Eze 24:6 Wherefore thus saith the Lord GOD; Woe to the bloody city, to the pot whose scum \bibemph{is} therein, and whose scum is not gone out of it! bring it out piece by piece; let no lot fall upon it.
\vs Eze 24:7 For her blood is in the midst of her; she set it upon the top of a rock; she poured it not upon the ground, to cover it with dust;
\vs Eze 24:8 That it might cause fury to come up to take vengeance; I have set her blood upon the top of a rock, that it should not be covered.
\vs Eze 24:9 Therefore thus saith the Lord GOD; Woe to the bloody city! I will even make the pile for fire great.
\vs Eze 24:10 Heap on wood, kindle the fire, consume the flesh, and spice it well, and let the bones be burned.
\vs Eze 24:11 Then set it empty upon the coals thereof, that the brass of it may be hot, and may burn, and \bibemph{that} the filthiness of it may be molten in it, \bibemph{that} the scum of it may be consumed.
\vs Eze 24:12 She hath wearied \bibemph{herself} with lies, and her great scum went not forth out of her: her scum \bibemph{shall be} in the fire.
\vs Eze 24:13 In thy filthiness \bibemph{is} lewdness: because I have purged thee, and thou wast not purged, thou shalt not be purged from thy filthiness any more, till I have caused my fury to rest upon thee.
\vs Eze 24:14 I the \textsc{Lord} have spoken \bibemph{it}: it shall come to pass, and I will do \bibemph{it}; I will not go back, neither will I spare, neither will I repent; according to thy ways, and according to thy doings, shall they judge thee, saith the Lord GOD.
\vs Eze 24:15 Also the word of the \textsc{Lord} came unto me, saying,
\vs Eze 24:16 Son of man, behold, I take away from thee the desire of thine eyes with a stroke: yet neither shalt thou mourn nor weep, neither shall thy tears run down.
\fn{run down: Heb. go.}
\vs Eze 24:17 Forbear to cry, make no mourning for the dead, bind the tire of thine head upon thee, and put on thy shoes upon thy feet, and cover not \bibemph{thy} lips, and eat not the bread of men.
\fn{Forbear: Heb. Be silent. lips: Heb. upper lip.}
\vs Eze 24:18 So I spake unto the people in the morning: and at even my wife died; and I did in the morning as I was commanded.
\vs Eze 24:19 And the people said unto me, Wilt thou not tell us what these \bibemph{things are} to us, that thou doest \bibemph{so}?
\vs Eze 24:20 Then I answered them, The word of the \textsc{Lord} came unto me, saying,
\vs Eze 24:21 Speak unto the house of Israel, Thus saith the Lord GOD; Behold, I will profane my sanctuary, the excellency of your strength, the desire of your eyes, and that which your soul pitieth; and your sons and your daughters whom ye have left shall fall by the sword.
\fn{that...: Heb. the pity of your soul.}
\vs Eze 24:22 And ye shall do as I have done: ye shall not cover \bibemph{your} lips, nor eat the bread of men.
\vs Eze 24:23 And your tires \bibemph{shall be} upon your heads, and your shoes upon your feet: ye shall not mourn nor weep; but ye shall pine away for your iniquities, and mourn one toward another.
\vs Eze 24:24 Thus Ezekiel is unto you a sign: according to all that he hath done shall ye do: and when this cometh, ye shall know that I \bibemph{am} the Lord GOD.
\vs Eze 24:25 Also, thou son of man, \bibemph{shall it} not \bibemph{be} in the day when I take from them their strength, the joy of their glory, the desire of their eyes, and that whereupon they set their minds, their sons and their daughters,
\fn{that...: Heb. the lifting up of their soul.}
\vs Eze 24:26 \bibemph{That} he that escapeth in that day shall come unto thee, to cause \bibemph{thee} to hear \bibemph{it} with \bibemph{thine} ears?
\vs Eze 24:27 In that day shall thy mouth be opened to him which is escaped, and thou shalt speak, and be no more dumb: and thou shalt be a sign unto them; and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 25:1 The word of the \textsc{Lord} came again unto me, saying,
\vs Eze 25:2 Son of man, set thy face against the Ammonites, and prophesy against them;
\vs Eze 25:3 And say unto the Ammonites, Hear the word of the Lord GOD; Thus saith the Lord GOD; Because thou saidst, Aha, against my sanctuary, when it was profaned; and against the land of Israel, when it was desolate; and against the house of Judah, when they went into captivity;
\vs Eze 25:4 Behold, therefore I will deliver thee to the men of the east for a possession, and they shall set their palaces in thee, and make their dwellings in thee: they shall eat thy fruit, and they shall drink thy milk.
\fn{men: Heb. children.}
\vs Eze 25:5 And I will make Rabbah a stable for camels, and the Ammonites a couchingplace for flocks: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 25:6 For thus saith the Lord GOD; Because thou hast clapped \bibemph{thine} hands, and stamped with the feet, and rejoiced in heart with all thy despite against the land of Israel;
\fn{hands: Heb. hand. feet: Heb. foot. heart: Heb. soul.}
\vs Eze 25:7 Behold, therefore I will stretch out mine hand upon thee, and will deliver thee for a spoil to the heathen; and I will cut thee off from the people, and I will cause thee to perish out of the countries: I will destroy thee; and thou shalt know that I \bibemph{am} the \textsc{Lord}.
\fn{a spoil: or, meat.}
\vs Eze 25:8 Thus saith the Lord GOD; Because that Moab and Seir do say, Behold, the house of Judah \bibemph{is} like unto all the heathen;
\vs Eze 25:9 Therefore, behold, I will open the side of Moab from the cities, from his cities \bibemph{which are} on his frontiers, the glory of the country, Bethjeshimoth, Baalmeon, and Kiriathaim,
\fn{side...: Heb. shoulder of Moab.}
\vs Eze 25:10 Unto the men of the east with the Ammonites, and will give them in possession, that the Ammonites may not be remembered among the nations.
\fn{with...: or, against the children of Ammon.}
\vs Eze 25:11 And I will execute judgments upon Moab; and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 25:12 Thus saith the Lord GOD; Because that Edom hath dealt against the house of Judah by taking vengeance, and hath greatly offended, and revenged himself upon them;
\fn{by...: Heb. by revenging revengement.}
\vs Eze 25:13 Therefore thus saith the Lord GOD; I will also stretch out mine hand upon Edom, and will cut off man and beast from it; and I will make it desolate from Teman; and they of Dedan shall fall by the sword.
\fn{they...: or, they shall fall by the sword unto Dedan.}
\vs Eze 25:14 And I will lay my vengeance upon Edom by the hand of my people Israel: and they shall do in Edom according to mine anger and according to my fury; and they shall know my vengeance, saith the Lord GOD.
\vs Eze 25:15 Thus saith the Lord GOD; Because the Philistines have dealt by revenge, and have taken vengeance with a despiteful heart, to destroy \bibemph{it} for the old hatred;
\fn{for the...: or, with perpetual hatred.}
\vs Eze 25:16 Therefore thus saith the Lord GOD; Behold, I will stretch out mine hand upon the Philistines, and I will cut off the Cherethims, and destroy the remnant of the sea coast.
\fn{sea coast: or, haven of the sea.}
\vs Eze 25:17 And I will execute great vengeance upon them with furious rebukes; and they shall know that I \bibemph{am} the \textsc{Lord}, when I shall lay my vengeance upon them.
\fn{great vengeance: Heb. great vengeances.}
\vs Eze 26:1 And it came to pass in the eleventh year, in the first \bibemph{day} of the month, \bibemph{that} the word of the \textsc{Lord} came unto me, saying,
\vs Eze 26:2 Son of man, because that Tyrus hath said against Jerusalem, Aha, she is broken \bibemph{that was} the gates of the people: she is turned unto me: I shall be replenished, \bibemph{now} she is laid waste:
\vs Eze 26:3 Therefore thus saith the Lord GOD; Behold, I \bibemph{am} against thee, O Tyrus, and will cause many nations to come up against thee, as the sea causeth his waves to come up.
\vs Eze 26:4 And they shall destroy the walls of Tyrus, and break down her towers: I will also scrape her dust from her, and make her like the top of a rock.
\vs Eze 26:5 It shall be \bibemph{a place for} the spreading of nets in the midst of the sea: for I have spoken \bibemph{it}, saith the Lord GOD: and it shall become a spoil to the nations.
\vs Eze 26:6 And her daughters which \bibemph{are} in the field shall be slain by the sword; and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 26:7 For thus saith the Lord GOD; Behold, I will bring upon Tyrus Nebuchadrezzar king of Babylon, a king of kings, from the north, with horses, and with chariots, and with horsemen, and companies, and much people.
\vs Eze 26:8 He shall slay with the sword thy daughters in the field: and he shall make a fort against thee, and cast a mount against thee, and lift up the buckler against thee.
\fn{cast...: or, pour out the engine of shot.}
\vs Eze 26:9 And he shall set engines of war against thy walls, and with his axes he shall break down thy towers.
\vs Eze 26:10 By reason of the abundance of his horses their dust shall cover thee: thy walls shall shake at the noise of the horsemen, and of the wheels, and of the chariots, when he shall enter into thy gates, as men enter into a city wherein is made a breach.
\fn{as men...: Heb. according to the enterings of a city broken up.}
\vs Eze 26:11 With the hoofs of his horses shall he tread down all thy streets: he shall slay thy people by the sword, and thy strong garrisons shall go down to the ground.
\vs Eze 26:12 And they shall make a spoil of thy riches, and make a prey of thy merchandise: and they shall break down thy walls, and destroy thy pleasant houses: and they shall lay thy stones and thy timber and thy dust in the midst of the water.
\fn{thy pleasant...: Heb. houses of thy desire.}
\vs Eze 26:13 And I will cause the noise of thy songs to cease; and the sound of thy harps shall be no more heard.
\vs Eze 26:14 And I will make thee like the top of a rock: thou shalt be \bibemph{a place} to spread nets upon; thou shalt be built no more: for I the \textsc{Lord} have spoken \bibemph{it}, saith the Lord GOD.
\vs Eze 26:15 Thus saith the Lord GOD to Tyrus; Shall not the isles shake at the sound of thy fall, when the wounded cry, when the slaughter is made in the midst of thee?
\vs Eze 26:16 Then all the princes of the sea shall come down from their thrones, and lay away their robes, and put off their broidered garments: they shall clothe themselves with trembling; they shall sit upon the ground, and shall tremble at \bibemph{every} moment, and be astonished at thee.
\fn{trembling: Heb. tremblings.}
\vs Eze 26:17 And they shall take up a lamentation for thee, and say to thee, How art thou destroyed, \bibemph{that wast} inhabited of seafaring men, the renowned city, which wast strong in the sea, she and her inhabitants, which cause their terror \bibemph{to be} on all that haunt it!
\fn{of...: Heb. of the seas.}
\vs Eze 26:18 Now shall the isles tremble in the day of thy fall; yea, the isles that \bibemph{are} in the sea shall be troubled at thy departure.
\vs Eze 26:19 For thus saith the Lord GOD; When I shall make thee a desolate city, like the cities that are not inhabited; when I shall bring up the deep upon thee, and great waters shall cover thee;
\vs Eze 26:20 When I shall bring thee down with them that descend into the pit, with the people of old time, and shall set thee in the low parts of the earth, in places desolate of old, with them that go down to the pit, that thou be not inhabited; and I shall set glory in the land of the living;
\vs Eze 26:21 I will make thee a terror, and thou \bibemph{shalt be} no \bibemph{more}: though thou be sought for, yet shalt thou never be found again, saith the Lord GOD.
\fn{a terror: Heb. terrors.}
\vs Eze 27:1 The word of the \textsc{Lord} came again unto me, saying,
\vs Eze 27:2 Now, thou son of man, take up a lamentation for Tyrus;
\vs Eze 27:3 And say unto Tyrus, O thou that art situate at the entry of the sea, \bibemph{which art} a merchant of the people for many isles, Thus saith the Lord GOD; O Tyrus, thou hast said, I \bibemph{am} of perfect beauty.
\fn{of perfect...: Heb. perfect of beauty.}
\vs Eze 27:4 Thy borders \bibemph{are} in the midst of the seas, thy builders have perfected thy beauty.
\fn{midst: Heb. heart.}
\vs Eze 27:5 They have made all thy \bibemph{ship} boards of fir trees of Senir: they have taken cedars from Lebanon to make masts for thee.
\fn{made: Heb. built.}
\vs Eze 27:6 \bibemph{Of} the oaks of Bashan have they made thine oars; the company of the Ashurites have made thy benches \bibemph{of} ivory, \bibemph{brought} out of the isles of Chittim.
\fn{the company...: or, they have made thy hatches of ivory well trodden. company: Heb. daughter.}
\vs Eze 27:7 Fine linen with broidered work from Egypt was that which thou spreadest forth to be thy sail; blue and purple from the isles of Elishah was that which covered thee.
\fn{blue...: or, purple and scarlet.}
\vs Eze 27:8 The inhabitants of Zidon and Arvad were thy mariners: thy wise \bibemph{men}, O Tyrus, \bibemph{that} were in thee, were thy pilots.
\vs Eze 27:9 The ancients of Gebal and the wise \bibemph{men} thereof were in thee thy calkers: all the ships of the sea with their mariners were in thee to occupy thy merchandise.
\fn{calkers: or, stoppers of chinks: Heb. strengtheners, etc.}
\vs Eze 27:10 They of Persia and of Lud and of Phut were in thine army, thy men of war: they hanged the shield and helmet in thee; they set forth thy comeliness.
\vs Eze 27:11 The men of Arvad with thine army \bibemph{were} upon thy walls round about, and the Gammadims were in thy towers: they hanged their shields upon thy walls round about; they have made thy beauty perfect.
\vs Eze 27:12 Tarshish \bibemph{was} thy merchant by reason of the multitude of all \bibemph{kind of} riches; with silver, iron, tin, and lead, they traded in thy fairs.
\vs Eze 27:13 Javan, Tubal, and Meshech, they \bibemph{were} thy merchants: they traded the persons of men and vessels of brass in thy market.
\fn{market: or, merchandise.}
\vs Eze 27:14 They of the house of Togarmah traded in thy fairs with horses and horsemen and mules.
\vs Eze 27:15 The men of Dedan \bibemph{were} thy merchants; many isles \bibemph{were} the merchandise of thine hand: they brought thee \bibemph{for} a present horns of ivory and ebony.
\vs Eze 27:16 Syria \bibemph{was} thy merchant by reason of the multitude of the wares of thy making: they occupied in thy fairs with emeralds, purple, and broidered work, and fine linen, and coral, and agate.
\fn{the wares...: Heb. thy works. agate: or, chrysoprase.}
\vs Eze 27:17 Judah, and the land of Israel, they \bibemph{were} thy merchants: they traded in thy market wheat of Minnith, and Pannag, and honey, and oil, and balm.
\fn{balm: or, rosin.}
\vs Eze 27:18 Damascus \bibemph{was} thy merchant in the multitude of the wares of thy making, for the multitude of all riches; in the wine of Helbon, and white wool.
\vs Eze 27:19 Dan also and Javan going to and fro occupied in thy fairs: bright iron, cassia, and calamus, were in thy market.
\fn{going...: or, Meuzal.}
\vs Eze 27:20 Dedan \bibemph{was} thy merchant in precious clothes for chariots.
\fn{precious...: Heb. clothes of freedom.}
\vs Eze 27:21 Arabia, and all the princes of Kedar, they occupied with thee in lambs, and rams, and goats: in these \bibemph{were they} thy merchants.
\fn{they occupied...: Heb. they were the merchants of thy hand.}
\vs Eze 27:22 The merchants of Sheba and Raamah, they \bibemph{were} thy merchants: they occupied in thy fairs with chief of all spices, and with all precious stones, and gold.
\vs Eze 27:23 Haran, and Canneh, and Eden, the merchants of Sheba, Asshur, \bibemph{and} Chilmad, \bibemph{were} thy merchants.
\vs Eze 27:24 These \bibemph{were} thy merchants in all sorts \bibemph{of things}, in blue clothes, and broidered work, and in chests of rich apparel, bound with cords, and made of cedar, among thy merchandise.
\fn{all...: or, excellent things. clothes: Heb. foldings.}
\vs Eze 27:25 The ships of Tarshish did sing of thee in thy market: and thou wast replenished, and made very glorious in the midst of the seas.
\vs Eze 27:26 Thy rowers have brought thee into great waters: the east wind hath broken thee in the midst of the seas.
\fn{midst: Heb. heart.}
\vs Eze 27:27 Thy riches, and thy fairs, thy merchandise, thy mariners, and thy pilots, thy calkers, and the occupiers of thy merchandise, and all thy men of war, that \bibemph{are} in thee, and in all thy company which \bibemph{is} in the midst of thee, shall fall into the midst of the seas in the day of thy ruin.
\fn{and in...: or, even with all. midst of the seas: Heb. heart, etc.}
\vs Eze 27:28 The suburbs shall shake at the sound of the cry of thy pilots.
\fn{suburbs: or, waves.}
\vs Eze 27:29 And all that handle the oar, the mariners, \bibemph{and} all the pilots of the sea, shall come down from their ships, they shall stand upon the land;
\vs Eze 27:30 And shall cause their voice to be heard against thee, and shall cry bitterly, and shall cast up dust upon their heads, they shall wallow themselves in the ashes:
\vs Eze 27:31 And they shall make themselves utterly bald for thee, and gird them with sackcloth, and they shall weep for thee with bitterness of heart \bibemph{and} bitter wailing.
\vs Eze 27:32 And in their wailing they shall take up a lamentation for thee, and lament over thee, \bibemph{saying}, What \bibemph{city is} like Tyrus, like the destroyed in the midst of the sea?
\vs Eze 27:33 When thy wares went forth out of the seas, thou filledst many people; thou didst enrich the kings of the earth with the multitude of thy riches and of thy merchandise.
\vs Eze 27:34 In the time \bibemph{when} thou shalt be broken by the seas in the depths of the waters thy merchandise and all thy company in the midst of thee shall fall.
\vs Eze 27:35 All the inhabitants of the isles shall be astonished at thee, and their kings shall be sore afraid, they shall be troubled in \bibemph{their} countenance.
\vs Eze 27:36 The merchants among the people shall hiss at thee; thou shalt be a terror, and never \bibemph{shalt be} any more.
\fn{a terror: Heb. terrors. never...: Heb. shalt not be for ever.}
\vs Eze 28:1 The word of the \textsc{Lord} came again unto me, saying,
\vs Eze 28:2 Son of man, say unto the prince of Tyrus, Thus saith the Lord GOD; Because thine heart \bibemph{is} lifted up, and thou hast said, I \bibemph{am} a God, I sit \bibemph{in} the seat of God, in the midst of the seas; yet thou \bibemph{art} a man, and not God, though thou set thine heart as the heart of God:
\fn{midst: Heb. heart.}
\vs Eze 28:3 Behold, thou \bibemph{art} wiser than Daniel; there is no secret that they can hide from thee:
\vs Eze 28:4 With thy wisdom and with thine understanding thou hast gotten thee riches, and hast gotten gold and silver into thy treasures:
\vs Eze 28:5 By thy great wisdom \bibemph{and} by thy traffick hast thou increased thy riches, and thine heart is lifted up because of thy riches:
\fn{thy great...: Heb. the greatness of thy wisdom.}
\vs Eze 28:6 Therefore thus saith the Lord GOD; Because thou hast set thine heart as the heart of God;
\vs Eze 28:7 Behold, therefore I will bring strangers upon thee, the terrible of the nations: and they shall draw their swords against the beauty of thy wisdom, and they shall defile thy brightness.
\vs Eze 28:8 They shall bring thee down to the pit, and thou shalt die the deaths of \bibemph{them that are} slain in the midst of the seas.
\vs Eze 28:9 Wilt thou yet say before him that slayeth thee, I \bibemph{am} God? but thou \bibemph{shalt be} a man, and no God, in the hand of him that slayeth thee.
\fn{of him that slayeth: or, of him that woundeth.}
\vs Eze 28:10 Thou shalt die the deaths of the uncircumcised by the hand of strangers: for I have spoken \bibemph{it}, saith the Lord GOD.
\vs Eze 28:11 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 28:12 Son of man, take up a lamentation upon the king of Tyrus, and say unto him, Thus saith the Lord GOD; Thou sealest up the sum, full of wisdom, and perfect in beauty.
\vs Eze 28:13 Thou hast been in Eden the garden of God; every precious stone \bibemph{was} thy covering, the sardius, topaz, and the diamond, the beryl, the onyx, and the jasper, the sapphire, the emerald, and the carbuncle, and gold: the workmanship of thy tabrets and of thy pipes was prepared in thee in the day that thou wast created.
\fn{sardius: or, ruby. beryl: or, chrysolite. emerald: or, chrysoprase.}
\vs Eze 28:14 Thou \bibemph{art} the anointed cherub that covereth; and I have set thee \bibemph{so}: thou wast upon the holy mountain of God; thou hast walked up and down in the midst of the stones of fire.
\vs Eze 28:15 Thou \bibemph{wast} perfect in thy ways from the day that thou wast created, till iniquity was found in thee.
\vs Eze 28:16 By the multitude of thy merchandise they have filled the midst of thee with violence, and thou hast sinned: therefore I will cast thee as profane out of the mountain of God: and I will destroy thee, O covering cherub, from the midst of the stones of fire.
\vs Eze 28:17 Thine heart was lifted up because of thy beauty, thou hast corrupted thy wisdom by reason of thy brightness: I will cast thee to the ground, I will lay thee before kings, that they may behold thee.
\vs Eze 28:18 Thou hast defiled thy sanctuaries by the multitude of thine iniquities, by the iniquity of thy traffick; therefore will I bring forth a fire from the midst of thee, it shall devour thee, and I will bring thee to ashes upon the earth in the sight of all them that behold thee.
\vs Eze 28:19 All they that know thee among the people shall be astonished at thee: thou shalt be a terror, and never \bibemph{shalt} thou \bibemph{be} any more.
\fn{a terror: Heb. terrors.}
\vs Eze 28:20 Again the word of the \textsc{Lord} came unto me, saying,
\vs Eze 28:21 Son of man, set thy face against Zidon, and prophesy against it,
\vs Eze 28:22 And say, Thus saith the Lord GOD; Behold, I \bibemph{am} against thee, O Zidon; and I will be glorified in the midst of thee: and they shall know that I \bibemph{am} the \textsc{Lord}, when I shall have executed judgments in her, and shall be sanctified in her.
\vs Eze 28:23 For I will send into her pestilence, and blood into her streets; and the wounded shall be judged in the midst of her by the sword upon her on every side; and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 28:24 And there shall be no more a pricking brier unto the house of Israel, nor \bibemph{any} grieving thorn of all \bibemph{that are} round about them, that despised them; and they shall know that I \bibemph{am} the Lord GOD.
\vs Eze 28:25 Thus saith the Lord GOD; When I shall have gathered the house of Israel from the people among whom they are scattered, and shall be sanctified in them in the sight of the heathen, then shall they dwell in their land that I have given to my servant Jacob.
\vs Eze 28:26 And they shall dwell safely therein, and shall build houses, and plant vineyards; yea, they shall dwell with confidence, when I have executed judgments upon all those that despise them round about them; and they shall know that I \bibemph{am} the \textsc{Lord} their God.
\fn{safely: or, with confidence. despise: or, spoil.}
\vs Eze 29:1 In the tenth year, in the tenth \bibemph{month}, in the twelfth \bibemph{day} of the month, the word of the \textsc{Lord} came unto me, saying,
\vs Eze 29:2 Son of man, set thy face against Pharaoh king of Egypt, and prophesy against him, and against all Egypt:
\vs Eze 29:3 Speak, and say, Thus saith the Lord GOD; Behold, I \bibemph{am} against thee, Pharaoh king of Egypt, the great dragon that lieth in the midst of his rivers, which hath said, My river \bibemph{is} mine own, and I have made \bibemph{it} for myself.
\vs Eze 29:4 But I will put hooks in thy jaws, and I will cause the fish of thy rivers to stick unto thy scales, and I will bring thee up out of the midst of thy rivers, and all the fish of thy rivers shall stick unto thy scales.
\vs Eze 29:5 And I will leave thee \bibemph{thrown} into the wilderness, thee and all the fish of thy rivers: thou shalt fall upon the open fields; thou shalt not be brought together, nor gathered: I have given thee for meat to the beasts of the field and to the fowls of the heaven.
\fn{open...: Heb. face of the field.}
\vs Eze 29:6 And all the inhabitants of Egypt shall know that I \bibemph{am} the \textsc{Lord}, because they have been a staff of reed to the house of Israel.
\vs Eze 29:7 When they took hold of thee by thy hand, thou didst break, and rend all their shoulder: and when they leaned upon thee, thou brakest, and madest all their loins to be at a stand.
\vs Eze 29:8 Therefore thus saith the Lord GOD; Behold, I will bring a sword upon thee, and cut off man and beast out of thee.
\vs Eze 29:9 And the land of Egypt shall be desolate and waste; and they shall know that I \bibemph{am} the \textsc{Lord}: because he hath said, The river \bibemph{is} mine, and I have made \bibemph{it}.
\vs Eze 29:10 Behold, therefore I \bibemph{am} against thee, and against thy rivers, and I will make the land of Egypt utterly waste \bibemph{and} desolate, from the tower of Syene even unto the border of Ethiopia.
\fn{utterly...: Heb. wastes of waste. from...: or, from Migdol to Syene. Syene: Heb. Seveneh.}
\vs Eze 29:11 No foot of man shall pass through it, nor foot of beast shall pass through it, neither shall it be inhabited forty years.
\vs Eze 29:12 And I will make the land of Egypt desolate in the midst of the countries \bibemph{that are} desolate, and her cities among the cities \bibemph{that are} laid waste shall be desolate forty years: and I will scatter the Egyptians among the nations, and will disperse them through the countries.
\vs Eze 29:13 Yet thus saith the Lord GOD; At the end of forty years will I gather the Egyptians from the people whither they were scattered:
\vs Eze 29:14 And I will bring again the captivity of Egypt, and will cause them to return \bibemph{into} the land of Pathros, into the land of their habitation; and they shall be there a base kingdom.
\fn{habitation: or, birth. base: Heb. low.}
\vs Eze 29:15 It shall be the basest of the kingdoms; neither shall it exalt itself any more above the nations: for I will diminish them, that they shall no more rule over the nations.
\vs Eze 29:16 And it shall be no more the confidence of the house of Israel, which bringeth \bibemph{their} iniquity to remembrance, when they shall look after them: but they shall know that I \bibemph{am} the Lord GOD.
\vs Eze 29:17 And it came to pass in the seven and twentieth year, in the first \bibemph{month}, in the first \bibemph{day} of the month, the word of the \textsc{Lord} came unto me, saying,
\vs Eze 29:18 Son of man, Nebuchadrezzar king of Babylon caused his army to serve a great service against Tyrus: every head \bibemph{was} made bald, and every shoulder \bibemph{was} peeled: yet had he no wages, nor his army, for Tyrus, for the service that he had served against it:
\vs Eze 29:19 Therefore thus saith the Lord GOD; Behold, I will give the land of Egypt unto Nebuchadrezzar king of Babylon; and he shall take her multitude, and take her spoil, and take her prey; and it shall be the wages for his army.
\fn{take her spoil...: Heb. spoil her spoil, and prey her prey.}
\vs Eze 29:20 I have given him the land of Egypt \bibemph{for} his labour wherewith he served against it, because they wrought for me, saith the Lord GOD.
\fn{for his...: or, for his hire.}
\vs Eze 29:21 In that day will I cause the horn of the house of Israel to bud forth, and I will give thee the opening of the mouth in the midst of them; and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 30:1 The word of the \textsc{Lord} came again unto me, saying,
\vs Eze 30:2 Son of man, prophesy and say, Thus saith the Lord GOD; Howl ye, Woe worth the day!
\vs Eze 30:3 For the day \bibemph{is} near, even the day of the \textsc{Lord} \bibemph{is} near, a cloudy day; it shall be the time of the heathen.
\vs Eze 30:4 And the sword shall come upon Egypt, and great pain shall be in Ethiopia, when the slain shall fall in Egypt, and they shall take away her multitude, and her foundations shall be broken down.
\fn{pain: or, fear.}
\vs Eze 30:5 Ethiopia, and Libya, and Lydia, and all the mingled people, and Chub, and the men of the land that is in league, shall fall with them by the sword.
\fn{Libya: Heb. Phut. men: Heb. children.}
\vs Eze 30:6 Thus saith the \textsc{Lord}; They also that uphold Egypt shall fall; and the pride of her power shall come down: from the tower of Syene shall they fall in it by the sword, saith the Lord GOD.
\fn{from...: or, from Migdol to Syene.}
\vs Eze 30:7 And they shall be desolate in the midst of the countries \bibemph{that are} desolate, and her cities shall be in the midst of the cities \bibemph{that are} wasted.
\vs Eze 30:8 And they shall know that I \bibemph{am} the \textsc{Lord}, when I have set a fire in Egypt, and \bibemph{when} all her helpers shall be destroyed.
\fn{destroyed: Heb. broken.}
\vs Eze 30:9 In that day shall messengers go forth from me in ships to make the careless Ethiopians afraid, and great pain shall come upon them, as in the day of Egypt: for, lo, it cometh.
\vs Eze 30:10 Thus saith the Lord GOD; I will also make the multitude of Egypt to cease by the hand of Nebuchadrezzar king of Babylon.
\vs Eze 30:11 He and his people with him, the terrible of the nations, shall be brought to destroy the land: and they shall draw their swords against Egypt, and fill the land with the slain.
\vs Eze 30:12 And I will make the rivers dry, and sell the land into the hand of the wicked: and I will make the land waste, and all that is therein, by the hand of strangers: I the \textsc{Lord} have spoken \bibemph{it}.
\fn{dry: Heb. drought. all...: Heb. the fulness thereof.}
\vs Eze 30:13 Thus saith the Lord GOD; I will also destroy the idols, and I will cause \bibemph{their} images to cease out of Noph; and there shall be no more a prince of the land of Egypt: and I will put a fear in the land of Egypt.
\vs Eze 30:14 And I will make Pathros desolate, and will set fire in Zoan, and will execute judgments in No.
\fn{Zoan: or, Tanis.}
\vs Eze 30:15 And I will pour my fury upon Sin, the strength of Egypt; and I will cut off the multitude of No.
\fn{Sin: or, Pelusium.}
\vs Eze 30:16 And I will set fire in Egypt: Sin shall have great pain, and No shall be rent asunder, and Noph \bibemph{shall have} distresses daily.
\vs Eze 30:17 The young men of Aven and of Pibeseth shall fall by the sword: and these \bibemph{cities} shall go into captivity.
\fn{Aven: or, Heliopolis. Pibeseth: or, Pubastum.}
\vs Eze 30:18 At Tehaphnehes also the day shall be darkened, when I shall break there the yokes of Egypt: and the pomp of her strength shall cease in her: as for her, a cloud shall cover her, and her daughters shall go into captivity.
\fn{darkened: or, restrained.}
\vs Eze 30:19 Thus will I execute judgments in Egypt: and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 30:20 And it came to pass in the eleventh year, in the first \bibemph{month}, in the seventh \bibemph{day} of the month, \bibemph{that} the word of the \textsc{Lord} came unto me, saying,
\vs Eze 30:21 Son of man, I have broken the arm of Pharaoh king of Egypt; and, lo, it shall not be bound up to be healed, to put a roller to bind it, to make it strong to hold the sword.
\vs Eze 30:22 Therefore thus saith the Lord GOD; Behold, I \bibemph{am} against Pharaoh king of Egypt, and will break his arms, the strong, and that which was broken; and I will cause the sword to fall out of his hand.
\vs Eze 30:23 And I will scatter the Egyptians among the nations, and will disperse them through the countries.
\vs Eze 30:24 And I will strengthen the arms of the king of Babylon, and put my sword in his hand: but I will break Pharaoh's arms, and he shall groan before him with the groanings of a deadly wounded \bibemph{man}.
\vs Eze 30:25 But I will strengthen the arms of the king of Babylon, and the arms of Pharaoh shall fall down; and they shall know that I \bibemph{am} the \textsc{Lord}, when I shall put my sword into the hand of the king of Babylon, and he shall stretch it out upon the land of Egypt.
\vs Eze 30:26 And I will scatter the Egyptians among the nations, and disperse them among the countries; and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 31:1 And it came to pass in the eleventh year, in the third \bibemph{month}, in the first \bibemph{day} of the month, \bibemph{that} the word of the \textsc{Lord} came unto me, saying,
\vs Eze 31:2 Son of man, speak unto Pharaoh king of Egypt, and to his multitude; Whom art thou like in thy greatness?
\vs Eze 31:3 Behold, the Assyrian \bibemph{was} a cedar in Lebanon with fair branches, and with a shadowing shroud, and of an high stature; and his top was among the thick boughs.
\fn{with fair branches: Heb. fair of branches.}
\vs Eze 31:4 The waters made him great, the deep set him up on high with her rivers running round about his plants, and sent out her little rivers unto all the trees of the field.
\fn{made...: or, nourished him. set...: or, brought him up. little rivers: or, conduits.}
\vs Eze 31:5 Therefore his height was exalted above all the trees of the field, and his boughs were multiplied, and his branches became long because of the multitude of waters, when he shot forth.
\fn{when...: or, when it sent them forth.}
\vs Eze 31:6 All the fowls of heaven made their nests in his boughs, and under his branches did all the beasts of the field bring forth their young, and under his shadow dwelt all great nations.
\vs Eze 31:7 Thus was he fair in his greatness, in the length of his branches: for his root was by great waters.
\vs Eze 31:8 The cedars in the garden of God could not hide him: the fir trees were not like his boughs, and the chesnut trees were not like his branches; nor any tree in the garden of God was like unto him in his beauty.
\vs Eze 31:9 I have made him fair by the multitude of his branches: so that all the trees of Eden, that \bibemph{were} in the garden of God, envied him.
\vs Eze 31:10 Therefore thus saith the Lord GOD; Because thou hast lifted up thyself in height, and he hath shot up his top among the thick boughs, and his heart is lifted up in his height;
\vs Eze 31:11 I have therefore delivered him into the hand of the mighty one of the heathen; he shall surely deal with him: I have driven him out for his wickedness.
\fn{he...: Heb. in doing he shall do unto him.}
\vs Eze 31:12 And strangers, the terrible of the nations, have cut him off, and have left him: upon the mountains and in all the valleys his branches are fallen, and his boughs are broken by all the rivers of the land; and all the people of the earth are gone down from his shadow, and have left him.
\vs Eze 31:13 Upon his ruin shall all the fowls of the heaven remain, and all the beasts of the field shall be upon his branches:
\vs Eze 31:14 To the end that none of all the trees by the waters exalt themselves for their height, neither shoot up their top among the thick boughs, neither their trees stand up in their height, all that drink water: for they are all delivered unto death, to the nether parts of the earth, in the midst of the children of men, with them that go down to the pit.
\fn{their trees...: or, stand upon themselves for their height.}
\vs Eze 31:15 Thus saith the Lord GOD; In the day when he went down to the grave I caused a mourning: I covered the deep for him, and I restrained the floods thereof, and the great waters were stayed: and I caused Lebanon to mourn for him, and all the trees of the field fainted for him.
\fn{to mourn: Heb. to be black.}
\vs Eze 31:16 I made the nations to shake at the sound of his fall, when I cast him down to hell with them that descend into the pit: and all the trees of Eden, the choice and best of Lebanon, all that drink water, shall be comforted in the nether parts of the earth.
\vs Eze 31:17 They also went down into hell with him unto \bibemph{them that be} slain with the sword; and \bibemph{they that were} his arm, \bibemph{that} dwelt under his shadow in the midst of the heathen.
\vs Eze 31:18 To whom art thou thus like in glory and in greatness among the trees of Eden? yet shalt thou be brought down with the trees of Eden unto the nether parts of the earth: thou shalt lie in the midst of the uncircumcised with \bibemph{them that be} slain by the sword. This \bibemph{is} Pharaoh and all his multitude, saith the Lord GOD.
\vs Eze 32:1 And it came to pass in the twelfth year, in the twelfth month, in the first \bibemph{day} of the month, \bibemph{that} the word of the \textsc{Lord} came unto me, saying,
\vs Eze 32:2 Son of man, take up a lamentation for Pharaoh king of Egypt, and say unto him, Thou art like a young lion of the nations, and thou \bibemph{art} as a whale in the seas: and thou camest forth with thy rivers, and troubledst the waters with thy feet, and fouledst their rivers.
\fn{whale: or, dragon.}
\vs Eze 32:3 Thus saith the Lord GOD; I will therefore spread out my net over thee with a company of many people; and they shall bring thee up in my net.
\vs Eze 32:4 Then will I leave thee upon the land, I will cast thee forth upon the open field, and will cause all the fowls of the heaven to remain upon thee, and I will fill the beasts of the whole earth with thee.
\vs Eze 32:5 And I will lay thy flesh upon the mountains, and fill the valleys with thy height.
\vs Eze 32:6 I will also water with thy blood the land wherein thou swimmest, \bibemph{even} to the mountains; and the rivers shall be full of thee.
\fn{the land wherein...: or, the land of thy swimming.}
\vs Eze 32:7 And when I shall put thee out, I will cover the heaven, and make the stars thereof dark; I will cover the sun with a cloud, and the moon shall not give her light.
\fn{put...: or, extinguish thee.}
\vs Eze 32:8 All the bright lights of heaven will I make dark over thee, and set darkness upon thy land, saith the Lord GOD.
\fn{bright...: Heb. lights of the light in heaven. dark: Heb. them dark.}
\vs Eze 32:9 I will also vex the hearts of many people, when I shall bring thy destruction among the nations, into the countries which thou hast not known.
\fn{vex: Heb. provoke to anger, or, grief.}
\vs Eze 32:10 Yea, I will make many people amazed at thee, and their kings shall be horribly afraid for thee, when I shall brandish my sword before them; and they shall tremble at \bibemph{every} moment, every man for his own life, in the day of thy fall.
\vs Eze 32:11 For thus saith the Lord GOD; The sword of the king of Babylon shall come upon thee.
\vs Eze 32:12 By the swords of the mighty will I cause thy multitude to fall, the terrible of the nations, all of them: and they shall spoil the pomp of Egypt, and all the multitude thereof shall be destroyed.
\vs Eze 32:13 I will destroy also all the beasts thereof from beside the great waters; neither shall the foot of man trouble them any more, nor the hoofs of beasts trouble them.
\vs Eze 32:14 Then will I make their waters deep, and cause their rivers to run like oil, saith the Lord GOD.
\vs Eze 32:15 When I shall make the land of Egypt desolate, and the country shall be destitute of that whereof it was full, when I shall smite all them that dwell therein, then shall they know that I \bibemph{am} the \textsc{Lord}.
\fn{destitute...: Heb. desolate from the fulness thereof.}
\vs Eze 32:16 This \bibemph{is} the lamentation wherewith they shall lament her: the daughters of the nations shall lament her: they shall lament for her, \bibemph{even} for Egypt, and for all her multitude, saith the Lord GOD.
\vs Eze 32:17 It came to pass also in the twelfth year, in the fifteenth \bibemph{day} of the month, \bibemph{that} the word of the \textsc{Lord} came unto me, saying,
\vs Eze 32:18 Son of man, wail for the multitude of Egypt, and cast them down, \bibemph{even} her, and the daughters of the famous nations, unto the nether parts of the earth, with them that go down into the pit.
\vs Eze 32:19 Whom dost thou pass in beauty? go down, and be thou laid with the uncircumcised.
\vs Eze 32:20 They shall fall in the midst of \bibemph{them that are} slain by the sword: she is delivered to the sword: draw her and all her multitudes.
\fn{she...: or, the sword is laid.}
\vs Eze 32:21 The strong among the mighty shall speak to him out of the midst of hell with them that help him: they are gone down, they lie uncircumcised, slain by the sword.
\vs Eze 32:22 Asshur \bibemph{is} there and all her company: his graves \bibemph{are} about him: all of them slain, fallen by the sword:
\vs Eze 32:23 Whose graves are set in the sides of the pit, and her company is round about her grave: all of them slain, fallen by the sword, which caused terror in the land of the living.
\fn{terror: or, dismaying.}
\vs Eze 32:24 There \bibemph{is} Elam and all her multitude round about her grave, all of them slain, fallen by the sword, which are gone down uncircumcised into the nether parts of the earth, which caused their terror in the land of the living; yet have they borne their shame with them that go down to the pit.
\vs Eze 32:25 They have set her a bed in the midst of the slain with all her multitude: her graves \bibemph{are} round about him: all of them uncircumcised, slain by the sword: though their terror was caused in the land of the living, yet have they borne their shame with them that go down to the pit: he is put in the midst of \bibemph{them that be} slain.
\vs Eze 32:26 There \bibemph{is} Meshech, Tubal, and all her multitude: her graves \bibemph{are} round about him: all of them uncircumcised, slain by the sword, though they caused their terror in the land of the living.
\vs Eze 32:27 And they shall not lie with the mighty \bibemph{that are} fallen of the uncircumcised, which are gone down to hell with their weapons of war: and they have laid their swords under their heads, but their iniquities shall be upon their bones, though \bibemph{they were} the terror of the mighty in the land of the living.
\fn{with their...: Heb. with weapons of their war.}
\vs Eze 32:28 Yea, thou shalt be broken in the midst of the uncircumcised, and shalt lie with \bibemph{them that are} slain with the sword.
\vs Eze 32:29 There \bibemph{is} Edom, her kings, and all her princes, which with their might are laid by \bibemph{them that were} slain by the sword: they shall lie with the uncircumcised, and with them that go down to the pit.
\fn{laid: Heb. given, or, put.}
\vs Eze 32:30 There \bibemph{be} the princes of the north, all of them, and all the Zidonians, which are gone down with the slain; with their terror they are ashamed of their might; and they lie uncircumcised with \bibemph{them that be} slain by the sword, and bear their shame with them that go down to the pit.
\vs Eze 32:31 Pharaoh shall see them, and shall be comforted over all his multitude, \bibemph{even} Pharaoh and all his army slain by the sword, saith the Lord GOD.
\vs Eze 32:32 For I have caused my terror in the land of the living: and he shall be laid in the midst of the uncircumcised with \bibemph{them that are} slain with the sword, \bibemph{even} Pharaoh and all his multitude, saith the Lord GOD.
\vs Eze 33:1 Again the word of the \textsc{Lord} came unto me, saying,
\vs Eze 33:2 Son of man, speak to the children of thy people, and say unto them, When I bring the sword upon a land, if the people of the land take a man of their coasts, and set him for their watchman:
\fn{When...: Heb. A land when I bring a sword upon her.}
\vs Eze 33:3 If when he seeth the sword come upon the land, he blow the trumpet, and warn the people;
\vs Eze 33:4 Then whosoever heareth the sound of the trumpet, and taketh not warning; if the sword come, and take him away, his blood shall be upon his own head.
\fn{whosoever...: Heb. he that hearing heareth.}
\vs Eze 33:5 He heard the sound of the trumpet, and took not warning; his blood shall be upon him. But he that taketh warning shall deliver his soul.
\vs Eze 33:6 But if the watchman see the sword come, and blow not the trumpet, and the people be not warned; if the sword come, and take \bibemph{any} person from among them, he is taken away in his iniquity; but his blood will I require at the watchman's hand.
\vs Eze 33:7 So thou, O son of man, I have set thee a watchman unto the house of Israel; therefore thou shalt hear the word at my mouth, and warn them from me.
\vs Eze 33:8 When I say unto the wicked, O wicked \bibemph{man}, thou shalt surely die; if thou dost not speak to warn the wicked from his way, that wicked \bibemph{man} shall die in his iniquity; but his blood will I require at thine hand.
\vs Eze 33:9 Nevertheless, if thou warn the wicked of his way to turn from it; if he do not turn from his way, he shall die in his iniquity; but thou hast delivered thy soul.
\vs Eze 33:10 Therefore, O thou son of man, speak unto the house of Israel; Thus ye speak, saying, If our transgressions and our sins \bibemph{be} upon us, and we pine away in them, how should we then live?
\vs Eze 33:11 Say unto them, \bibemph{As} I live, saith the Lord GOD, I have no pleasure in the death of the wicked; but that the wicked turn from his way and live: turn ye, turn ye from your evil ways; for why will ye die, O house of Israel?
\vs Eze 33:12 Therefore, thou son of man, say unto the children of thy people, The righteousness of the righteous shall not deliver him in the day of his transgression: as for the wickedness of the wicked, he shall not fall thereby in the day that he turneth from his wickedness; neither shall the righteous be able to live for his \bibemph{righteousness} in the day that he sinneth.
\vs Eze 33:13 When I shall say to the righteous, \bibemph{that} he shall surely live; if he trust to his own righteousness, and commit iniquity, all his righteousnesses shall not be remembered; but for his iniquity that he hath committed, he shall die for it.
\vs Eze 33:14 Again, when I say unto the wicked, Thou shalt surely die; if he turn from his sin, and do that which is lawful and right;
\fn{that...: Heb. judgment and justice.}
\vs Eze 33:15 \bibemph{If} the wicked restore the pledge, give again that he had robbed, walk in the statutes of life, without committing iniquity; he shall surely live, he shall not die.
\vs Eze 33:16 None of his sins that he hath committed shall be mentioned unto him: he hath done that which is lawful and right; he shall surely live.
\vs Eze 33:17 Yet the children of thy people say, The way of the Lord is not equal: but as for them, their way is not equal.
\vs Eze 33:18 When the righteous turneth from his righteousness, and committeth iniquity, he shall even die thereby.
\vs Eze 33:19 But if the wicked turn from his wickedness, and do that which is lawful and right, he shall live thereby.
\vs Eze 33:20 Yet ye say, The way of the Lord is not equal. O ye house of Israel, I will judge you every one after his ways.
\vs Eze 33:21 And it came to pass in the twelfth year of our captivity, in the tenth \bibemph{month}, in the fifth \bibemph{day} of the month, \bibemph{that} one that had escaped out of Jerusalem came unto me, saying, The city is smitten.
\vs Eze 33:22 Now the hand of the \textsc{Lord} was upon me in the evening, afore he that was escaped came; and had opened my mouth, until he came to me in the morning; and my mouth was opened, and I was no more dumb.
\vs Eze 33:23 Then the word of the \textsc{Lord} came unto me, saying,
\vs Eze 33:24 Son of man, they that inhabit those wastes of the land of Israel speak, saying, Abraham was one, and he inherited the land: but we \bibemph{are} many; the land is given us for inheritance.
\vs Eze 33:25 Wherefore say unto them, Thus saith the Lord GOD; Ye eat with the blood, and lift up your eyes toward your idols, and shed blood: and shall ye possess the land?
\vs Eze 33:26 Ye stand upon your sword, ye work abomination, and ye defile every one his neighbour's wife: and shall ye possess the land?
\vs Eze 33:27 Say thou thus unto them, Thus saith the Lord GOD; \bibemph{As} I live, surely they that \bibemph{are} in the wastes shall fall by the sword, and him that \bibemph{is} in the open field will I give to the beasts to be devoured, and they that \bibemph{be} in the forts and in the caves shall die of the pestilence.
\fn{to be...: Heb. to devour him.}
\vs Eze 33:28 For I will lay the land most desolate, and the pomp of her strength shall cease; and the mountains of Israel shall be desolate, that none shall pass through.
\fn{most...: Heb. desolation and desolation.}
\vs Eze 33:29 Then shall they know that I \bibemph{am} the \textsc{Lord}, when I have laid the land most desolate because of all their abominations which they have committed.
\vs Eze 33:30 Also, thou son of man, the children of thy people still are talking against thee by the walls and in the doors of the houses, and speak one to another, every one to his brother, saying, Come, I pray you, and hear what is the word that cometh forth from the \textsc{Lord}.
\fn{against thee: of, of thee.}
\vs Eze 33:31 And they come unto thee as the people cometh, and they sit before thee \bibemph{as} my people, and they hear thy words, but they will not do them: for with their mouth they shew much love, \bibemph{but} their heart goeth after their covetousness.
\fn{as the...: Heb. according to the coming of the people. they sit...: or, my people sit before thee. they shew...: Heb. they make loves, or, jests.}
\vs Eze 33:32 And, lo, thou \bibemph{art} unto them as a very lovely song of one that hath a pleasant voice, and can play well on an instrument: for they hear thy words, but they do them not.
\fn{a very...: Heb. a song of loves.}
\vs Eze 33:33 And when this cometh to pass, (lo, it will come,) then shall they know that a prophet hath been among them.
\vs Eze 34:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 34:2 Son of man, prophesy against the shepherds of Israel, prophesy, and say unto them, Thus saith the Lord GOD unto the shepherds; Woe \bibemph{be} to the shepherds of Israel that do feed themselves! should not the shepherds feed the flocks?
\vs Eze 34:3 Ye eat the fat, and ye clothe you with the wool, ye kill them that are fed: \bibemph{but} ye feed not the flock.
\vs Eze 34:4 The diseased have ye not strengthened, neither have ye healed that which was sick, neither have ye bound up \bibemph{that which was} broken, neither have ye brought again that which was driven away, neither have ye sought that which was lost; but with force and with cruelty have ye ruled them.
\vs Eze 34:5 And they were scattered, because \bibemph{there is} no shepherd: and they became meat to all the beasts of the field, when they were scattered.
\fn{because...: or, without a shepherd.}
\vs Eze 34:6 My sheep wandered through all the mountains, and upon every high hill: yea, my flock was scattered upon all the face of the earth, and none did search or seek \bibemph{after them}.
\vs Eze 34:7 Therefore, ye shepherds, hear the word of the \textsc{Lord};
\vs Eze 34:8 \bibemph{As} I live, saith the Lord GOD, surely because my flock became a prey, and my flock became meat to every beast of the field, because \bibemph{there was} no shepherd, neither did my shepherds search for my flock, but the shepherds fed themselves, and fed not my flock;
\vs Eze 34:9 Therefore, O ye shepherds, hear the word of the \textsc{Lord};
\vs Eze 34:10 Thus saith the Lord GOD; Behold, I \bibemph{am} against the shepherds; and I will require my flock at their hand, and cause them to cease from feeding the flock; neither shall the shepherds feed themselves any more; for I will deliver my flock from their mouth, that they may not be meat for them.
\vs Eze 34:11 For thus saith the Lord GOD; Behold, I, \bibemph{even} I, will both search my sheep, and seek them out.
\vs Eze 34:12 As a shepherd seeketh out his flock in the day that he is among his sheep \bibemph{that are} scattered; so will I seek out my sheep, and will deliver them out of all places where they have been scattered in the cloudy and dark day.
\fn{As...: Heb. According to the seeking.}
\vs Eze 34:13 And I will bring them out from the people, and gather them from the countries, and will bring them to their own land, and feed them upon the mountains of Israel by the rivers, and in all the inhabited places of the country.
\vs Eze 34:14 I will feed them in a good pasture, and upon the high mountains of Israel shall their fold be: there shall they lie in a good fold, and \bibemph{in} a fat pasture shall they feed upon the mountains of Israel.
\vs Eze 34:15 I will feed my flock, and I will cause them to lie down, saith the Lord GOD.
\vs Eze 34:16 I will seek that which was lost, and bring again that which was driven away, and will bind up \bibemph{that which was} broken, and will strengthen that which was sick: but I will destroy the fat and the strong; I will feed them with judgment.
\vs Eze 34:17 And \bibemph{as for} you, O my flock, thus saith the Lord GOD; Behold, I judge between cattle and cattle, between the rams and the he goats.
\fn{cattle: Heb. small cattle of lambs and kids. he goats: Heb. great he goats.}
\vs Eze 34:18 \bibemph{Seemeth it} a small thing unto you to have eaten up the good pasture, but ye must tread down with your feet the residue of your pastures? and to have drunk of the deep waters, but ye must foul the residue with your feet?
\vs Eze 34:19 And \bibemph{as for} my flock, they eat that which ye have trodden with your feet; and they drink that which ye have fouled with your feet.
\vs Eze 34:20 Therefore thus saith the Lord GOD unto them; Behold, I, \bibemph{even} I, will judge between the fat cattle and between the lean cattle.
\vs Eze 34:21 Because ye have thrust with side and with shoulder, and pushed all the diseased with your horns, till ye have scattered them abroad;
\vs Eze 34:22 Therefore will I save my flock, and they shall no more be a prey; and I will judge between cattle and cattle.
\vs Eze 34:23 And I will set up one shepherd over them, and he shall feed them, \bibemph{even} my servant David; he shall feed them, and he shall be their shepherd.
\vs Eze 34:24 And I the \textsc{Lord} will be their God, and my servant David a prince among them; I the \textsc{Lord} have spoken \bibemph{it}.
\vs Eze 34:25 And I will make with them a covenant of peace, and will cause the evil beasts to cease out of the land: and they shall dwell safely in the wilderness, and sleep in the woods.
\vs Eze 34:26 And I will make them and the places round about my hill a blessing; and I will cause the shower to come down in his season; there shall be showers of blessing.
\vs Eze 34:27 And the tree of the field shall yield her fruit, and the earth shall yield her increase, and they shall be safe in their land, and shall know that I \bibemph{am} the \textsc{Lord}, when I have broken the bands of their yoke, and delivered them out of the hand of those that served themselves of them.
\vs Eze 34:28 And they shall no more be a prey to the heathen, neither shall the beast of the land devour them; but they shall dwell safely, and none shall make \bibemph{them} afraid.
\vs Eze 34:29 And I will raise up for them a plant of renown, and they shall be no more consumed with hunger in the land, neither bear the shame of the heathen any more.
\fn{of renown: or, for renown. consumed: Heb. taken away.}
\vs Eze 34:30 Thus shall they know that I the \textsc{Lord} their God \bibemph{am} with them, and \bibemph{that} they, \bibemph{even} the house of Israel, \bibemph{are} my people, saith the Lord GOD.
\vs Eze 34:31 And ye my flock, the flock of my pasture, \bibemph{are} men, \bibemph{and} I \bibemph{am} your God, saith the Lord GOD.
\vs Eze 35:1 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 35:2 Son of man, set thy face against mount Seir, and prophesy against it,
\vs Eze 35:3 And say unto it, Thus saith the Lord GOD; Behold, O mount Seir, I \bibemph{am} against thee, and I will stretch out mine hand against thee, and I will make thee most desolate.
\fn{most...: Heb. desolation and desolation.}
\vs Eze 35:4 I will lay thy cities waste, and thou shalt be desolate, and thou shalt know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 35:5 Because thou hast had a perpetual hatred, and hast shed \bibemph{the blood of} the children of Israel by the force of the sword in the time of their calamity, in the time \bibemph{that their} iniquity \bibemph{had} an end:
\fn{perpetual...: or, hatred of old. shed...: Heb. poured out the children. force: Heb. hands.}
\vs Eze 35:6 Therefore, \bibemph{as} I live, saith the Lord GOD, I will prepare thee unto blood, and blood shall pursue thee: sith thou hast not hated blood, even blood shall pursue thee.
\vs Eze 35:7 Thus will I make mount Seir most desolate, and cut off from it him that passeth out and him that returneth.
\fn{most...: Heb. desolation and desolation.}
\vs Eze 35:8 And I will fill his mountains with his slain \bibemph{men}: in thy hills, and in thy valleys, and in all thy rivers, shall they fall that are slain with the sword.
\vs Eze 35:9 I will make thee perpetual desolations, and thy cities shall not return: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 35:10 Because thou hast said, These two nations and these two countries shall be mine, and we will possess it; whereas the \textsc{Lord} was there:
\fn{whereas...: or, though the \textsc{Lord} was there.}
\vs Eze 35:11 Therefore, \bibemph{as} I live, saith the Lord GOD, I will even do according to thine anger, and according to thine envy which thou hast used out of thy hatred against them; and I will make myself known among them, when I have judged thee.
\vs Eze 35:12 And thou shalt know that I \bibemph{am} the \textsc{Lord}, \bibemph{and that} I have heard all thy blasphemies which thou hast spoken against the mountains of Israel, saying, They are laid desolate, they are given us to consume.
\fn{to consume: Heb. to devour.}
\vs Eze 35:13 Thus with your mouth ye have boasted against me, and have multiplied your words against me: I have heard \bibemph{them}.
\fn{boasted: Heb. magnified.}
\vs Eze 35:14 Thus saith the Lord GOD; When the whole earth rejoiceth, I will make thee desolate.
\vs Eze 35:15 As thou didst rejoice at the inheritance of the house of Israel, because it was desolate, so will I do unto thee: thou shalt be desolate, O mount Seir, and all Idumea, \bibemph{even} all of it: and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 36:1 Also, thou son of man, prophesy unto the mountains of Israel, and say, Ye mountains of Israel, hear the word of the \textsc{Lord}:
\vs Eze 36:2 Thus saith the Lord GOD; Because the enemy hath said against you, Aha, even the ancient high places are ours in possession:
\vs Eze 36:3 Therefore prophesy and say, Thus saith the Lord GOD; Because they have made \bibemph{you} desolate, and swallowed you up on every side, that ye might be a possession unto the residue of the heathen, and ye are taken up in the lips of talkers, and \bibemph{are} an infamy of the people:
\fn{Because: Heb. Because for because. are...: or, ye are made to come up on the lip of the tongue.}
\vs Eze 36:4 Therefore, ye mountains of Israel, hear the word of the Lord GOD; Thus saith the Lord GOD to the mountains, and to the hills, to the rivers, and to the valleys, to the desolate wastes, and to the cities that are forsaken, which became a prey and derision to the residue of the heathen that \bibemph{are} round about;
\fn{rivers: or, bottoms, or, dales.}
\vs Eze 36:5 Therefore thus saith the Lord GOD; Surely in the fire of my jealousy have I spoken against the residue of the heathen, and against all Idumea, which have appointed my land into their possession with the joy of all \bibemph{their} heart, with despiteful minds, to cast it out for a prey.
\vs Eze 36:6 Prophesy therefore concerning the land of Israel, and say unto the mountains, and to the hills, to the rivers, and to the valleys, Thus saith the Lord GOD; Behold, I have spoken in my jealousy and in my fury, because ye have borne the shame of the heathen:
\vs Eze 36:7 Therefore thus saith the Lord GOD; I have lifted up mine hand, Surely the heathen that \bibemph{are} about you, they shall bear their shame.
\vs Eze 36:8 But ye, O mountains of Israel, ye shall shoot forth your branches, and yield your fruit to my people of Israel; for they are at hand to come.
\vs Eze 36:9 For, behold, I \bibemph{am} for you, and I will turn unto you, and ye shall be tilled and sown:
\vs Eze 36:10 And I will multiply men upon you, all the house of Israel, \bibemph{even} all of it: and the cities shall be inhabited, and the wastes shall be builded:
\vs Eze 36:11 And I will multiply upon you man and beast; and they shall increase and bring fruit: and I will settle you after your old estates, and will do better \bibemph{unto you} than at your beginnings: and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 36:12 Yea, I will cause men to walk upon you, \bibemph{even} my people Israel; and they shall possess thee, and thou shalt be their inheritance, and thou shalt no more henceforth bereave them \bibemph{of men}.
\vs Eze 36:13 Thus saith the Lord GOD; Because they say unto you, Thou \bibemph{land} devourest up men, and hast bereaved thy nations;
\vs Eze 36:14 Therefore thou shalt devour men no more, neither bereave thy nations any more, saith the Lord GOD.
\fn{bereave: or, cause to fall.}
\vs Eze 36:15 Neither will I cause \bibemph{men} to hear in thee the shame of the heathen any more, neither shalt thou bear the reproach of the people any more, neither shalt thou cause thy nations to fall any more, saith the Lord GOD.
\vs Eze 36:16 Moreover the word of the \textsc{Lord} came unto me, saying,
\vs Eze 36:17 Son of man, when the house of Israel dwelt in their own land, they defiled it by their own way and by their doings: their way was before me as the uncleanness of a removed woman.
\vs Eze 36:18 Wherefore I poured my fury upon them for the blood that they had shed upon the land, and for their idols \bibemph{wherewith} they had polluted it:
\vs Eze 36:19 And I scattered them among the heathen, and they were dispersed through the countries: according to their way and according to their doings I judged them.
\vs Eze 36:20 And when they entered unto the heathen, whither they went, they profaned my holy name, when they said to them, These \bibemph{are} the people of the \textsc{Lord}, and are gone forth out of his land.
\vs Eze 36:21 But I had pity for mine holy name, which the house of Israel had profaned among the heathen, whither they went.
\vs Eze 36:22 Therefore say unto the house of Israel, Thus saith the Lord GOD; I do not \bibemph{this} for your sakes, O house of Israel, but for mine holy name's sake, which ye have profaned among the heathen, whither ye went.
\vs Eze 36:23 And I will sanctify my great name, which was profaned among the heathen, which ye have profaned in the midst of them; and the heathen shall know that I \bibemph{am} the \textsc{Lord}, saith the Lord GOD, when I shall be sanctified in you before their eyes.
\fn{their: or, your.}
\vs Eze 36:24 For I will take you from among the heathen, and gather you out of all countries, and will bring you into your own land.
\vs Eze 36:25 Then will I sprinkle clean water upon you, and ye shall be clean: from all your filthiness, and from all your idols, will I cleanse you.
\vs Eze 36:26 A new heart also will I give you, and a new spirit will I put within you: and I will take away the stony heart out of your flesh, and I will give you an heart of flesh.
\vs Eze 36:27 And I will put my spirit within you, and cause you to walk in my statutes, and ye shall keep my judgments, and do \bibemph{them}.
\vs Eze 36:28 And ye shall dwell in the land that I gave to your fathers; and ye shall be my people, and I will be your God.
\vs Eze 36:29 I will also save you from all your uncleannesses: and I will call for the corn, and will increase it, and lay no famine upon you.
\vs Eze 36:30 And I will multiply the fruit of the tree, and the increase of the field, that ye shall receive no more reproach of famine among the heathen.
\vs Eze 36:31 Then shall ye remember your own evil ways, and your doings that \bibemph{were} not good, and shall lothe yourselves in your own sight for your iniquities and for your abominations.
\vs Eze 36:32 Not for your sakes do I \bibemph{this}, saith the Lord GOD, be it known unto you: be ashamed and confounded for your own ways, O house of Israel.
\vs Eze 36:33 Thus saith the Lord GOD; In the day that I shall have cleansed you from all your iniquities I will also cause \bibemph{you} to dwell in the cities, and the wastes shall be builded.
\vs Eze 36:34 And the desolate land shall be tilled, whereas it lay desolate in the sight of all that passed by.
\vs Eze 36:35 And they shall say, This land that was desolate is become like the garden of Eden; and the waste and desolate and ruined cities \bibemph{are become} fenced, \bibemph{and} are inhabited.
\vs Eze 36:36 Then the heathen that are left round about you shall know that I the \textsc{Lord} build the ruined \bibemph{places, and} plant that that was desolate: I the \textsc{Lord} have spoken \bibemph{it}, and I will do \bibemph{it}.
\vs Eze 36:37 Thus saith the Lord GOD; I will yet \bibemph{for} this be enquired of by the house of Israel, to do \bibemph{it} for them; I will increase them with men like a flock.
\vs Eze 36:38 As the holy flock, as the flock of Jerusalem in her solemn feasts; so shall the waste cities be filled with flocks of men: and they shall know that I \bibemph{am} the \textsc{Lord}.
\fn{holy...: Heb. flock of holy things.}
\vs Eze 37:1 The hand of the \textsc{Lord} was upon me, and carried me out in the spirit of the \textsc{Lord}, and set me down in the midst of the valley which \bibemph{was} full of bones,
\vs Eze 37:2 And caused me to pass by them round about: and, behold, \bibemph{there were} very many in the open valley; and, lo, \bibemph{they were} very dry.
\fn{valley: or, champaign.}
\vs Eze 37:3 And he said unto me, Son of man, can these bones live? And I answered, O Lord GOD, thou knowest.
\vs Eze 37:4 Again he said unto me, Prophesy upon these bones, and say unto them, O ye dry bones, hear the word of the \textsc{Lord}.
\vs Eze 37:5 Thus saith the Lord GOD unto these bones; Behold, I will cause breath to enter into you, and ye shall live:
\vs Eze 37:6 And I will lay sinews upon you, and will bring up flesh upon you, and cover you with skin, and put breath in you, and ye shall live; and ye shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 37:7 So I prophesied as I was commanded: and as I prophesied, there was a noise, and behold a shaking, and the bones came together, bone to his bone.
\vs Eze 37:8 And when I beheld, lo, the sinews and the flesh came up upon them, and the skin covered them above: but \bibemph{there was} no breath in them.
\vs Eze 37:9 Then said he unto me, Prophesy unto the wind, prophesy, son of man, and say to the wind, Thus saith the Lord GOD; Come from the four winds, O breath, and breathe upon these slain, that they may live.
\fn{wind: or, breath.}
\vs Eze 37:10 So I prophesied as he commanded me, and the breath came into them, and they lived, and stood up upon their feet, an exceeding great army.
\vs Eze 37:11 Then he said unto me, Son of man, these bones are the whole house of Israel: behold, they say, Our bones are dried, and our hope is lost: we are cut off for our parts.
\vs Eze 37:12 Therefore prophesy and say unto them, Thus saith the Lord GOD; Behold, O my people, I will open your graves, and cause you to come up out of your graves, and bring you into the land of Israel.
\vs Eze 37:13 And ye shall know that I \bibemph{am} the \textsc{Lord}, when I have opened your graves, O my people, and brought you up out of your graves,
\vs Eze 37:14 And shall put my spirit in you, and ye shall live, and I shall place you in your own land: then shall ye know that I the \textsc{Lord} have spoken \bibemph{it}, and performed \bibemph{it}, saith the \textsc{Lord}.
\vs Eze 37:15 The word of the \textsc{Lord} came again unto me, saying,
\vs Eze 37:16 Moreover, thou son of man, take thee one stick, and write upon it, For Judah, and for the children of Israel his companions: then take another stick, and write upon it, For Joseph, the stick of Ephraim, and \bibemph{for} all the house of Israel his companions:
\vs Eze 37:17 And join them one to another into one stick; and they shall become one in thine hand.
\vs Eze 37:18 And when the children of thy people shall speak unto thee, saying, Wilt thou not shew us what thou \bibemph{meanest} by these?
\vs Eze 37:19 Say unto them, Thus saith the Lord GOD; Behold, I will take the stick of Joseph, which \bibemph{is} in the hand of Ephraim, and the tribes of Israel his fellows, and will put them with him, \bibemph{even} with the stick of Judah, and make them one stick, and they shall be one in mine hand.
\vs Eze 37:20 And the sticks whereon thou writest shall be in thine hand before their eyes.
\vs Eze 37:21 And say unto them, Thus saith the Lord GOD; Behold, I will take the children of Israel from among the heathen, whither they be gone, and will gather them on every side, and bring them into their own land:
\vs Eze 37:22 And I will make them one nation in the land upon the mountains of Israel; and one king shall be king to them all: and they shall be no more two nations, neither shall they be divided into two kingdoms any more at all:
\vs Eze 37:23 Neither shall they defile themselves any more with their idols, nor with their detestable things, nor with any of their transgressions: but I will save them out of all their dwellingplaces, wherein they have sinned, and will cleanse them: so shall they be my people, and I will be their God.
\vs Eze 37:24 And David my servant \bibemph{shall be} king over them; and they all shall have one shepherd: they shall also walk in my judgments, and observe my statutes, and do them.
\vs Eze 37:25 And they shall dwell in the land that I have given unto Jacob my servant, wherein your fathers have dwelt; and they shall dwell therein, \bibemph{even} they, and their children, and their children's children for ever: and my servant David \bibemph{shall be} their prince for ever.
\vs Eze 37:26 Moreover I will make a covenant of peace with them; it shall be an everlasting covenant with them: and I will place them, and multiply them, and will set my sanctuary in the midst of them for evermore.
\vs Eze 37:27 My tabernacle also shall be with them: yea, I will be their God, and they shall be my people.
\vs Eze 37:28 And the heathen shall know that I the \textsc{Lord} do sanctify Israel, when my sanctuary shall be in the midst of them for evermore.
\vs Eze 38:1 And the word of the \textsc{Lord} came unto me, saying,
\vs Eze 38:2 Son of man, set thy face against Gog, the land of Magog, the chief prince of Meshech and Tubal, and prophesy against him,
\fn{the chief...: or, prince of the chief.}
\vs Eze 38:3 And say, Thus saith the Lord GOD; Behold, I \bibemph{am} against thee, O Gog, the chief prince of Meshech and Tubal:
\vs Eze 38:4 And I will turn thee back, and put hooks into thy jaws, and I will bring thee forth, and all thine army, horses and horsemen, all of them clothed with all sorts \bibemph{of armour, even} a great company \bibemph{with} bucklers and shields, all of them handling swords:
\vs Eze 38:5 Persia, Ethiopia, and Libya with them; all of them with shield and helmet:
\fn{Libya: or, Phut.}
\vs Eze 38:6 Gomer, and all his bands; the house of Togarmah of the north quarters, and all his bands: \bibemph{and} many people with thee.
\vs Eze 38:7 Be thou prepared, and prepare for thyself, thou, and all thy company that are assembled unto thee, and be thou a guard unto them.
\vs Eze 38:8 After many days thou shalt be visited: in the latter years thou shalt come into the land \bibemph{that is} brought back from the sword, \bibemph{and is} gathered out of many people, against the mountains of Israel, which have been always waste: but it is brought forth out of the nations, and they shall dwell safely all of them.
\vs Eze 38:9 Thou shalt ascend and come like a storm, thou shalt be like a cloud to cover the land, thou, and all thy bands, and many people with thee.
\vs Eze 38:10 Thus saith the Lord GOD; It shall also come to pass, \bibemph{that} at the same time shall things come into thy mind, and thou shalt think an evil thought:
\fn{think...: or, conceive a mischievous purpose.}
\vs Eze 38:11 And thou shalt say, I will go up to the land of unwalled villages; I will go to them that are at rest, that dwell safely, all of them dwelling without walls, and having neither bars nor gates,
\fn{safely: or, confidently.}
\vs Eze 38:12 To take a spoil, and to take a prey; to turn thine hand upon the desolate places \bibemph{that are now} inhabited, and upon the people \bibemph{that are} gathered out of the nations, which have gotten cattle and goods, that dwell in the midst of the land.
\fn{To take...: Heb. To spoil the spoil, and to prey the prey. midst: Heb. navel.}
\vs Eze 38:13 Sheba, and Dedan, and the merchants of Tarshish, with all the young lions thereof, shall say unto thee, Art thou come to take a spoil? hast thou gathered thy company to take a prey? to carry away silver and gold, to take away cattle and goods, to take a great spoil?
\vs Eze 38:14 Therefore, son of man, prophesy and say unto Gog, Thus saith the Lord GOD; In that day when my people of Israel dwelleth safely, shalt thou not know \bibemph{it}?
\vs Eze 38:15 And thou shalt come from thy place out of the north parts, thou, and many people with thee, all of them riding upon horses, a great company, and a mighty army:
\vs Eze 38:16 And thou shalt come up against my people of Israel, as a cloud to cover the land; it shall be in the latter days, and I will bring thee against my land, that the heathen may know me, when I shall be sanctified in thee, O Gog, before their eyes.
\vs Eze 38:17 Thus saith the Lord GOD; \bibemph{Art} thou he of whom I have spoken in old time by my servants the prophets of Israel, which prophesied in those days \bibemph{many} years that I would bring thee against them?
\fn{by: Heb. by the hand of.}
\vs Eze 38:18 And it shall come to pass at the same time when Gog shall come against the land of Israel, saith the Lord GOD, \bibemph{that} my fury shall come up in my face.
\vs Eze 38:19 For in my jealousy \bibemph{and} in the fire of my wrath have I spoken, Surely in that day there shall be a great shaking in the land of Israel;
\vs Eze 38:20 So that the fishes of the sea, and the fowls of the heaven, and the beasts of the field, and all creeping things that creep upon the earth, and all the men that \bibemph{are} upon the face of the earth, shall shake at my presence, and the mountains shall be thrown down, and the steep places shall fall, and every wall shall fall to the ground.
\fn{steep...: or, towers, or, stairs.}
\vs Eze 38:21 And I will call for a sword against him throughout all my mountains, saith the Lord GOD: every man's sword shall be against his brother.
\vs Eze 38:22 And I will plead against him with pestilence and with blood; and I will rain upon him, and upon his bands, and upon the many people that \bibemph{are} with him, an overflowing rain, and great hailstones, fire, and brimstone.
\vs Eze 38:23 Thus will I magnify myself, and sanctify myself; and I will be known in the eyes of many nations, and they shall know that I \bibemph{am} the \textsc{Lord}.
\vs Eze 39:1 Therefore, thou son of man, prophesy against Gog, and say, Thus saith the Lord GOD; Behold, I \bibemph{am} against thee, O Gog, the chief prince of Meshech and Tubal:
\vs Eze 39:2 And I will turn thee back, and leave but the sixth part of thee, and will cause thee to come up from the north parts, and will bring thee upon the mountains of Israel:
\fn{leave...: or, strike thee with six plagues: or, draw thee back with an hook of six teeth. the north...: Heb. the sides of the north.}
\vs Eze 39:3 And I will smite thy bow out of thy left hand, and will cause thine arrows to fall out of thy right hand.
\vs Eze 39:4 Thou shalt fall upon the mountains of Israel, thou, and all thy bands, and the people that \bibemph{is} with thee: I will give thee unto the ravenous birds of every sort, and \bibemph{to} the beasts of the field to be devoured.
\fn{sort: Heb. wing. to be...: Heb. to devour.}
\vs Eze 39:5 Thou shalt fall upon the open field: for I have spoken \bibemph{it}, saith the Lord GOD.
\fn{the open...: Heb. the face of the field.}
\vs Eze 39:6 And I will send a fire on Magog, and among them that dwell carelessly in the isles: and they shall know that I \bibemph{am} the \textsc{Lord}.
\fn{carelessly: or, confidently.}
\vs Eze 39:7 So will I make my holy name known in the midst of my people Israel; and I will not \bibemph{let them} pollute my holy name any more: and the heathen shall know that I \bibemph{am} the \textsc{Lord}, the Holy One in Israel.
\vs Eze 39:8 Behold, it is come, and it is done, saith the Lord GOD; this \bibemph{is} the day whereof I have spoken.
\vs Eze 39:9 And they that dwell in the cities of Israel shall go forth, and shall set on fire and burn the weapons, both the shields and the bucklers, the bows and the arrows, and the handstaves, and the spears, and they shall burn them with fire seven years:
\fn{handstaves: or, javelins. burn them...: or, make a fire of them.}
\vs Eze 39:10 So that they shall take no wood out of the field, neither cut down \bibemph{any} out of the forests; for they shall burn the weapons with fire: and they shall spoil those that spoiled them, and rob those that robbed them, saith the Lord GOD.
\vs Eze 39:11 And it shall come to pass in that day, \bibemph{that} I will give unto Gog a place there of graves in Israel, the valley of the passengers on the east of the sea: and it shall stop the \bibemph{noses} of the passengers: and there shall they bury Gog and all his multitude: and they shall call \bibemph{it} The valley of Hamongog.
\fn{noses: or, mouths. Hamongog: that is, The multitude of Gog.}
\vs Eze 39:12 And seven months shall the house of Israel be burying of them, that they may cleanse the land.
\vs Eze 39:13 Yea, all the people of the land shall bury \bibemph{them}; and it shall be to them a renown the day that I shall be glorified, saith the Lord GOD.
\vs Eze 39:14 And they shall sever out men of continual employment, passing through the land to bury with the passengers those that remain upon the face of the earth, to cleanse it: after the end of seven months shall they search.
\fn{men...: Heb. men of continuance.}
\vs Eze 39:15 And the passengers \bibemph{that} pass through the land, when \bibemph{any} seeth a man's bone, then shall he set up a sign by it, till the buriers have buried it in the valley of Hamongog.
\fn{set up: Heb. build.}
\vs Eze 39:16 And also the name of the city \bibemph{shall be} Hamonah. Thus shall they cleanse the land.
\fn{Hamonah: that is, The multitude.}
\vs Eze 39:17 And, thou son of man, thus saith the Lord GOD; Speak unto every feathered fowl, and to every beast of the field, Assemble yourselves, and come; gather yourselves on every side to my sacrifice that I do sacrifice for you, \bibemph{even} a great sacrifice upon the mountains of Israel, that ye may eat flesh, and drink blood.
\fn{unto...: Heb. to the fowl of every wing. my sacrifice: or, my slaughter.}
\vs Eze 39:18 Ye shall eat the flesh of the mighty, and drink the blood of the princes of the earth, of rams, of lambs, and of goats, of bullocks, all of them fatlings of Bashan.
\fn{goats: Heb. great goats.}
\vs Eze 39:19 And ye shall eat fat till ye be full, and drink blood till ye be drunken, of my sacrifice which I have sacrificed for you.
\vs Eze 39:20 Thus ye shall be filled at my table with horses and chariots, with mighty men, and with all men of war, saith the Lord GOD.
\vs Eze 39:21 And I will set my glory among the heathen, and all the heathen shall see my judgment that I have executed, and my hand that I have laid upon them.
\vs Eze 39:22 So the house of Israel shall know that I \bibemph{am} the \textsc{Lord} their God from that day and forward.
\vs Eze 39:23 And the heathen shall know that the house of Israel went into captivity for their iniquity: because they trespassed against me, therefore hid I my face from them, and gave them into the hand of their enemies: so fell they all by the sword.
\vs Eze 39:24 According to their uncleanness and according to their transgressions have I done unto them, and hid my face from them.
\vs Eze 39:25 Therefore thus saith the Lord GOD; Now will I bring again the captivity of Jacob, and have mercy upon the whole house of Israel, and will be jealous for my holy name;
\vs Eze 39:26 After that they have borne their shame, and all their trespasses whereby they have trespassed against me, when they dwelt safely in their land, and none made \bibemph{them} afraid.
\vs Eze 39:27 When I have brought them again from the people, and gathered them out of their enemies' lands, and am sanctified in them in the sight of many nations;
\vs Eze 39:28 Then shall they know that I \bibemph{am} the \textsc{Lord} their God, which caused them to be led into captivity among the heathen: but I have gathered them unto their own land, and have left none of them any more there.
\fn{which...: Heb. by my causing of them, etc.}
\vs Eze 39:29 Neither will I hide my face any more from them: for I have poured out my spirit upon the house of Israel, saith the Lord GOD.
\vs Eze 40:1 In the five and twentieth year of our captivity, in the beginning of the year, in the tenth \bibemph{day} of the month, in the fourteenth year after that the city was smitten, in the selfsame day the hand of the \textsc{Lord} was upon me, and brought me thither.
\vs Eze 40:2 In the visions of God brought he me into the land of Israel, and set me upon a very high mountain, by which \bibemph{was} as the frame of a city on the south.
\fn{by which: or, upon which.}
\vs Eze 40:3 And he brought me thither, and, behold, \bibemph{there was} a man, whose appearance \bibemph{was} like the appearance of brass, with a line of flax in his hand, and a measuring reed; and he stood in the gate.
\vs Eze 40:4 And the man said unto me, Son of man, behold with thine eyes, and hear with thine ears, and set thine heart upon all that I shall shew thee; for to the intent that I might shew \bibemph{them} unto thee \bibemph{art} thou brought hither: declare all that thou seest to the house of Israel.
\vs Eze 40:5 And behold a wall on the outside of the house round about, and in the man's hand a measuring reed of six cubits \bibemph{long} by the cubit and an hand breadth: so he measured the breadth of the building, one reed; and the height, one reed.
\vs Eze 40:6 Then came he unto the gate which looketh toward the east, and went up the stairs thereof, and measured the threshold of the gate, \bibemph{which was} one reed broad; and the other threshold \bibemph{of the gate, which was} one reed broad.
\fn{which looketh...: Heb. whose face was the way toward the east.}
\vs Eze 40:7 And \bibemph{every} little chamber \bibemph{was} one reed long, and one reed broad; and between the little chambers \bibemph{were} five cubits; and the threshold of the gate by the porch of the gate within \bibemph{was} one reed.
\vs Eze 40:8 He measured also the porch of the gate within, one reed.
\vs Eze 40:9 Then measured he the porch of the gate, eight cubits; and the posts thereof, two cubits; and the porch of the gate \bibemph{was} inward.
\vs Eze 40:10 And the little chambers of the gate eastward \bibemph{were} three on this side, and three on that side; they three \bibemph{were} of one measure: and the posts had one measure on this side and on that side.
\vs Eze 40:11 And he measured the breadth of the entry of the gate, ten cubits; \bibemph{and} the length of the gate, thirteen cubits.
\vs Eze 40:12 The space also before the little chambers \bibemph{was} one cubit \bibemph{on this side}, and the space \bibemph{was} one cubit on that side: and the little chambers \bibemph{were} six cubits on this side, and six cubits on that side.
\fn{space: Heb. limit, or, bound.}
\vs Eze 40:13 He measured then the gate from the roof of \bibemph{one} little chamber to the roof of another: the breadth \bibemph{was} five and twenty cubits, door against door.
\vs Eze 40:14 He made also posts of threescore cubits, even unto the post of the court round about the gate.
\vs Eze 40:15 And from the face of the gate of the entrance unto the face of the porch of the inner gate \bibemph{were} fifty cubits.
\vs Eze 40:16 And \bibemph{there were} narrow windows to the little chambers, and to their posts within the gate round about, and likewise to the arches: and windows \bibemph{were} round about inward: and upon \bibemph{each} post \bibemph{were} palm trees.
\fn{narrow: Heb. closed. arches: or, galleries, or, porches. inward: or, within.}
\vs Eze 40:17 Then brought he me into the outward court, and, lo, \bibemph{there were} chambers, and a pavement made for the court round about: thirty chambers \bibemph{were} upon the pavement.
\vs Eze 40:18 And the pavement by the side of the gates over against the length of the gates \bibemph{was} the lower pavement.
\vs Eze 40:19 Then he measured the breadth from the forefront of the lower gate unto the forefront of the inner court without, an hundred cubits eastward and northward.
\fn{without: or, from without.}
\vs Eze 40:20 And the gate of the outward court that looked toward the north, he measured the length thereof, and the breadth thereof.
\fn{that looked: Heb. whose face was.}
\vs Eze 40:21 And the little chambers thereof \bibemph{were} three on this side and three on that side; and the posts thereof and the arches thereof were after the measure of the first gate: the length thereof \bibemph{was} fifty cubits, and the breadth five and twenty cubits.
\fn{arches: or, galleries, or, porches.}
\vs Eze 40:22 And their windows, and their arches, and their palm trees, \bibemph{were} after the measure of the gate that looketh toward the east; and they went up unto it by seven steps; and the arches thereof \bibemph{were} before them.
\vs Eze 40:23 And the gate of the inner court \bibemph{was} over against the gate toward the north, and toward the east; and he measured from gate to gate an hundred cubits.
\vs Eze 40:24 After that he brought me toward the south, and behold a gate toward the south: and he measured the posts thereof and the arches thereof according to these measures.
\vs Eze 40:25 And \bibemph{there were} windows in it and in the arches thereof round about, like those windows: the length \bibemph{was} fifty cubits, and the breadth five and twenty cubits.
\vs Eze 40:26 And \bibemph{there were} seven steps to go up to it, and the arches thereof \bibemph{were} before them: and it had palm trees, one on this side, and another on that side, upon the posts thereof.
\vs Eze 40:27 And \bibemph{there was} a gate in the inner court toward the south: and he measured from gate to gate toward the south an hundred cubits.
\vs Eze 40:28 And he brought me to the inner court by the south gate: and he measured the south gate according to these measures;
\vs Eze 40:29 And the little chambers thereof, and the posts thereof, and the arches thereof, according to these measures: and \bibemph{there were} windows in it and in the arches thereof round about: \bibemph{it was} fifty cubits long, and five and twenty cubits broad.
\vs Eze 40:30 And the arches round about \bibemph{were} five and twenty cubits long, and five cubits broad.
\fn{broad: Heb. breadth.}
\vs Eze 40:31 And the arches thereof \bibemph{were} toward the utter court; and palm trees \bibemph{were} upon the posts thereof: and the going up to it \bibemph{had} eight steps.
\vs Eze 40:32 And he brought me into the inner court toward the east: and he measured the gate according to these measures.
\vs Eze 40:33 And the little chambers thereof, and the posts thereof, and the arches thereof, \bibemph{were} according to these measures: and \bibemph{there were} windows therein and in the arches thereof round about: \bibemph{it was} fifty cubits long, and five and twenty cubits broad.
\vs Eze 40:34 And the arches thereof \bibemph{were} toward the outward court; and palm trees \bibemph{were} upon the posts thereof, on this side, and on that side: and the going up to it \bibemph{had} eight steps.
\vs Eze 40:35 And he brought me to the north gate, and measured \bibemph{it} according to these measures;
\vs Eze 40:36 The little chambers thereof, the posts thereof, and the arches thereof, and the windows to it round about: the length \bibemph{was} fifty cubits, and the breadth five and twenty cubits.
\vs Eze 40:37 And the posts thereof \bibemph{were} toward the utter court; and palm trees \bibemph{were} upon the posts thereof, on this side, and on that side: and the going up to it \bibemph{had} eight steps.
\vs Eze 40:38 And the chambers and the entries thereof \bibemph{were} by the posts of the gates, where they washed the burnt offering.
\vs Eze 40:39 And in the porch of the gate \bibemph{were} two tables on this side, and two tables on that side, to slay thereon the burnt offering and the sin offering and the trespass offering.
\vs Eze 40:40 And at the side without, as one goeth up to the entry of the north gate, \bibemph{were} two tables; and on the other side, which \bibemph{was} at the porch of the gate, \bibemph{were} two tables.
\fn{as one...: or, at the step.}
\vs Eze 40:41 Four tables \bibemph{were} on this side, and four tables on that side, by the side of the gate; eight tables, whereupon they slew \bibemph{their sacrifices}.
\vs Eze 40:42 And the four tables \bibemph{were} of hewn stone for the burnt offering, of a cubit and an half long, and a cubit and an half broad, and one cubit high: whereupon also they laid the instruments wherewith they slew the burnt offering and the sacrifice.
\vs Eze 40:43 And within \bibemph{were} hooks, an hand broad, fastened round about: and upon the tables \bibemph{was} the flesh of the offering.
\fn{hooks: or, and irons, or, the two hearthstones.}
\vs Eze 40:44 And without the inner gate \bibemph{were} the chambers of the singers in the inner court, which \bibemph{was} at the side of the north gate; and their prospect \bibemph{was} toward the south: one at the side of the east gate \bibemph{having} the prospect toward the north.
\vs Eze 40:45 And he said unto me, This chamber, whose prospect \bibemph{is} toward the south, \bibemph{is} for the priests, the keepers of the charge of the house.
\fn{charge: or, ward, or, ordinance.}
\vs Eze 40:46 And the chamber whose prospect \bibemph{is} toward the north \bibemph{is} for the priests, the keepers of the charge of the altar: these \bibemph{are} the sons of Zadok among the sons of Levi, which come near to the \textsc{Lord} to minister unto him.
\vs Eze 40:47 So he measured the court, an hundred cubits long, and an hundred cubits broad, foursquare; and the altar \bibemph{that was} before the house.
\vs Eze 40:48 And he brought me to the porch of the house, and measured \bibemph{each} post of the porch, five cubits on this side, and five cubits on that side: and the breadth of the gate \bibemph{was} three cubits on this side, and three cubits on that side.
\vs Eze 40:49 The length of the porch \bibemph{was} twenty cubits, and the breadth eleven cubits; and \bibemph{he brought me} by the steps whereby they went up to it: and \bibemph{there were} pillars by the posts, one on this side, and another on that side.
\vs Eze 41:1 Afterward he brought me to the temple, and measured the posts, six cubits broad on the one side, and six cubits broad on the other side, \bibemph{which was} the breadth of the tabernacle.
\vs Eze 41:2 And the breadth of the door \bibemph{was} ten cubits; and the sides of the door \bibemph{were} five cubits on the one side, and five cubits on the other side: and he measured the length thereof, forty cubits: and the breadth, twenty cubits.
\fn{door: or, entrance.}
\vs Eze 41:3 Then went he inward, and measured the post of the door, two cubits; and the door, six cubits; and the breadth of the door, seven cubits.
\vs Eze 41:4 So he measured the length thereof, twenty cubits; and the breadth, twenty cubits, before the temple: and he said unto me, This \bibemph{is} the most holy \bibemph{place}.
\vs Eze 41:5 After he measured the wall of the house, six cubits; and the breadth of \bibemph{every} side chamber, four cubits, round about the house on every side.
\vs Eze 41:6 And the side chambers \bibemph{were} three, one over another, and thirty in order; and they entered into the wall which \bibemph{was} of the house for the side chambers round about, that they might have hold, but they had not hold in the wall of the house.
\fn{one...: Heb. side chamber over side chamber. have hold: Heb. be holden. thirty...: or, three and thirty times, or, foot.}
\vs Eze 41:7 And \bibemph{there was} an enlarging, and a winding about still upward to the side chambers: for the winding about of the house went still upward round about the house: therefore the breadth of the house \bibemph{was still} upward, and so increased \bibemph{from} the lowest \bibemph{chamber} to the highest by the midst.
\fn{there...: Heb. it was made broader, and went round.}
\vs Eze 41:8 I saw also the height of the house round about: the foundations of the side chambers \bibemph{were} a full reed of six great cubits.
\vs Eze 41:9 The thickness of the wall, which \bibemph{was} for the side chamber without, \bibemph{was} five cubits: and \bibemph{that} which \bibemph{was} left \bibemph{was} the place of the side chambers that \bibemph{were} within.
\vs Eze 41:10 And between the chambers \bibemph{was} the wideness of twenty cubits round about the house on every side.
\vs Eze 41:11 And the doors of the side chambers \bibemph{were} toward \bibemph{the place that was} left, one door toward the north, and another door toward the south: and the breadth of the place that was left \bibemph{was} five cubits round about.
\vs Eze 41:12 Now the building that \bibemph{was} before the separate place at the end toward the west \bibemph{was} seventy cubits broad; and the wall of the building \bibemph{was} five cubits thick round about, and the length thereof ninety cubits.
\vs Eze 41:13 So he measured the house, an hundred cubits long; and the separate place, and the building, with the walls thereof, an hundred cubits long;
\vs Eze 41:14 Also the breadth of the face of the house, and of the separate place toward the east, an hundred cubits.
\vs Eze 41:15 And he measured the length of the building over against the separate place which \bibemph{was} behind it, and the galleries thereof on the one side and on the other side, an hundred cubits, with the inner temple, and the porches of the court;
\fn{galleries: or, several walks, or, walks with pillars.}
\vs Eze 41:16 The door posts, and the narrow windows, and the galleries round about on their three stories, over against the door, cieled with wood round about, and from the ground up to the windows, and the windows \bibemph{were} covered;
\fn{cieled...: Heb. cieling of wood. and from...: or, and the ground unto the windows.}
\vs Eze 41:17 To that above the door, even unto the inner house, and without, and by all the wall round about within and without, by measure.
\fn{measure: Heb. measures.}
\vs Eze 41:18 And \bibemph{it was} made with cherubims and palm trees, so that a palm tree \bibemph{was} between a cherub and a cherub; and \bibemph{every} cherub had two faces;
\vs Eze 41:19 So that the face of a man \bibemph{was} toward the palm tree on the one side, and the face of a young lion toward the palm tree on the other side: \bibemph{it was} made through all the house round about.
\vs Eze 41:20 From the ground unto above the door \bibemph{were} cherubims and palm trees made, and \bibemph{on} the wall of the temple.
\vs Eze 41:21 The posts of the temple \bibemph{were} squared, \bibemph{and} the face of the sanctuary; the appearance \bibemph{of the one} as the appearance \bibemph{of the other}.
\fn{posts: Heb. post.}
\vs Eze 41:22 The altar of wood \bibemph{was} three cubits high, and the length thereof two cubits; and the corners thereof, and the length thereof, and the walls thereof, \bibemph{were} of wood: and he said unto me, This \bibemph{is} the table that \bibemph{is} before the \textsc{Lord}.
\vs Eze 41:23 And the temple and the sanctuary had two doors.
\vs Eze 41:24 And the doors had two leaves \bibemph{apiece}, two turning leaves; two \bibemph{leaves} for the one door, and two leaves for the other \bibemph{door}.
\vs Eze 41:25 And \bibemph{there were} made on them, on the doors of the temple, cherubims and palm trees, like as \bibemph{were} made upon the walls; and \bibemph{there were} thick planks upon the face of the porch without.
\vs Eze 41:26 And \bibemph{there were} narrow windows and palm trees on the one side and on the other side, on the sides of the porch, and \bibemph{upon} the side chambers of the house, and thick planks.
\vs Eze 42:1 Then he brought me forth into the utter court, the way toward the north: and he brought me into the chamber that \bibemph{was} over against the separate place, and which \bibemph{was} before the building toward the north.
\vs Eze 42:2 Before the length of an hundred cubits \bibemph{was} the north door, and the breadth \bibemph{was} fifty cubits.
\vs Eze 42:3 Over against the twenty \bibemph{cubits} which \bibemph{were} for the inner court, and over against the pavement which \bibemph{was} for the utter court, \bibemph{was} gallery against gallery in three \bibemph{stories}.
\vs Eze 42:4 And before the chambers \bibemph{was} a walk of ten cubits breadth inward, a way of one cubit; and their doors toward the north.
\vs Eze 42:5 Now the upper chambers \bibemph{were} shorter: for the galleries were higher than these, than the lower, and than the middlemost of the building.
\fn{were higher...: or, did eat of these. than the lower...: or, and the building consisted of the lower and the middlemost.}
\vs Eze 42:6 For they \bibemph{were} in three \bibemph{stories}, but had not pillars as the pillars of the courts: therefore \bibemph{the building} was straitened more than the lowest and the middlemost from the ground.
\vs Eze 42:7 And the wall that \bibemph{was} without over against the chambers, toward the utter court on the forepart of the chambers, the length thereof \bibemph{was} fifty cubits.
\vs Eze 42:8 For the length of the chambers that \bibemph{were} in the utter court \bibemph{was} fifty cubits: and, lo, before the temple \bibemph{were} an hundred cubits.
\vs Eze 42:9 And from under these chambers \bibemph{was} the entry on the east side, as one goeth into them from the utter court.
\fn{from under: or, from the place. the entry: or, he that brought me. as...: or, as he came.}
\vs Eze 42:10 The chambers \bibemph{were} in the thickness of the wall of the court toward the east, over against the separate place, and over against the building.
\vs Eze 42:11 And the way before them \bibemph{was} like the appearance of the chambers which \bibemph{were} toward the north, as long as they, \bibemph{and} as broad as they: and all their goings out \bibemph{were} both according to their fashions, and according to their doors.
\vs Eze 42:12 And according to the doors of the chambers that \bibemph{were} toward the south \bibemph{was} a door in the head of the way, \bibemph{even} the way directly before the wall toward the east, as one entereth into them.
\vs Eze 42:13 Then said he unto me, The north chambers \bibemph{and} the south chambers, which \bibemph{are} before the separate place, they \bibemph{be} holy chambers, where the priests that approach unto the \textsc{Lord} shall eat the most holy things: there shall they lay the most holy things, and the meat offering, and the sin offering, and the trespass offering; for the place \bibemph{is} holy.
\vs Eze 42:14 When the priests enter therein, then shall they not go out of the holy \bibemph{place} into the utter court, but there they shall lay their garments wherein they minister; for they \bibemph{are} holy; and shall put on other garments, and shall approach to \bibemph{those things} which \bibemph{are} for the people.
\vs Eze 42:15 Now when he had made an end of measuring the inner house, he brought me forth toward the gate whose prospect \bibemph{is} toward the east, and measured it round about.
\vs Eze 42:16 He measured the east side with the measuring reed, five hundred reeds, with the measuring reed round about.
\fn{side: Heb. wind.}
\vs Eze 42:17 He measured the north side, five hundred reeds, with the measuring reed round about.
\vs Eze 42:18 He measured the south side, five hundred reeds, with the measuring reed.
\vs Eze 42:19 He turned about to the west side, \bibemph{and} measured five hundred reeds with the measuring reed.
\vs Eze 42:20 He measured it by the four sides: it had a wall round about, five hundred \bibemph{reeds} long, and five hundred broad, to make a separation between the sanctuary and the profane place.
\vs Eze 43:1 Afterward he brought me to the gate, \bibemph{even} the gate that looketh toward the east:
\vs Eze 43:2 And, behold, the glory of the God of Israel came from the way of the east: and his voice \bibemph{was} like a noise of many waters: and the earth shined with his glory.
\vs Eze 43:3 And \bibemph{it was} according to the appearance of the vision which I saw, \bibemph{even} according to the vision that I saw when I came to destroy the city: and the visions \bibemph{were} like the vision that I saw by the river Chebar; and I fell upon my face.
\fn{when...: or, when I came to prophesy that the city should be destroyed.}
\vs Eze 43:4 And the glory of the \textsc{Lord} came into the house by the way of the gate whose prospect \bibemph{is} toward the east.
\vs Eze 43:5 So the spirit took me up, and brought me into the inner court; and, behold, the glory of the \textsc{Lord} filled the house.
\vs Eze 43:6 And I heard \bibemph{him} speaking unto me out of the house; and the man stood by me.
\vs Eze 43:7 And he said unto me, Son of man, the place of my throne, and the place of the soles of my feet, where I will dwell in the midst of the children of Israel for ever, and my holy name, shall the house of Israel no more defile, \bibemph{neither} they, nor their kings, by their whoredom, nor by the carcases of their kings in their high places.
\vs Eze 43:8 In their setting of their threshold by my thresholds, and their post by my posts, and the wall between me and them, they have even defiled my holy name by their abominations that they have committed: wherefore I have consumed them in mine anger.
\fn{and the...: or, for there was but a wall between me and them.}
\vs Eze 43:9 Now let them put away their whoredom, and the carcases of their kings, far from me, and I will dwell in the midst of them for ever.
\vs Eze 43:10 Thou son of man, shew the house to the house of Israel, that they may be ashamed of their iniquities: and let them measure the pattern.
\fn{pattern: or, sum, or, number.}
\vs Eze 43:11 And if they be ashamed of all that they have done, shew them the form of the house, and the fashion thereof, and the goings out thereof, and the comings in thereof, and all the forms thereof, and all the ordinances thereof, and all the forms thereof, and all the laws thereof: and write \bibemph{it} in their sight, that they may keep the whole form thereof, and all the ordinances thereof, and do them.
\vs Eze 43:12 This \bibemph{is} the law of the house; Upon the top of the mountain the whole limit thereof round about \bibemph{shall be} most holy. Behold, this \bibemph{is} the law of the house.
\vs Eze 43:13 And these \bibemph{are} the measures of the altar after the cubits: The cubit \bibemph{is} a cubit and an hand breadth; even the bottom \bibemph{shall be} a cubit, and the breadth a cubit, and the border thereof by the edge thereof round about \bibemph{shall be} a span: and this \bibemph{shall be} the higher place of the altar.
\fn{bottom: Heb. bosom. edge: Heb. lip.}
\vs Eze 43:14 And from the bottom \bibemph{upon} the ground \bibemph{even} to the lower settle \bibemph{shall be} two cubits, and the breadth one cubit; and from the lesser settle \bibemph{even} to the greater settle \bibemph{shall be} four cubits, and the breadth \bibemph{one} cubit.
\vs Eze 43:15 So the altar \bibemph{shall be} four cubits; and from the altar and upward \bibemph{shall be} four horns.
\fn{the altar (first): Heb. Harel, that is, the mountain of God. the altar (second): Heb. Ariel, that is, the lion of God.}
\vs Eze 43:16 And the altar \bibemph{shall be} twelve \bibemph{cubits} long, twelve broad, square in the four squares thereof.
\vs Eze 43:17 And the settle \bibemph{shall be} fourteen \bibemph{cubits} long and fourteen broad in the four squares thereof; and the border about it \bibemph{shall be} half a cubit; and the bottom thereof \bibemph{shall be} a cubit about; and his stairs shall look toward the east.
\vs Eze 43:18 And he said unto me, Son of man, thus saith the Lord GOD; These \bibemph{are} the ordinances of the altar in the day when they shall make it, to offer burnt offerings thereon, and to sprinkle blood thereon.
\vs Eze 43:19 And thou shalt give to the priests the Levites that be of the seed of Zadok, which approach unto me, to minister unto me, saith the Lord GOD, a young bullock for a sin offering.
\vs Eze 43:20 And thou shalt take of the blood thereof, and put \bibemph{it} on the four horns of it, and on the four corners of the settle, and upon the border round about: thus shalt thou cleanse and purge it.
\vs Eze 43:21 Thou shalt take the bullock also of the sin offering, and he shall burn it in the appointed place of the house, without the sanctuary.
\vs Eze 43:22 And on the second day thou shalt offer a kid of the goats without blemish for a sin offering; and they shall cleanse the altar, as they did cleanse \bibemph{it} with the bullock.
\vs Eze 43:23 When thou hast made an end of cleansing \bibemph{it}, thou shalt offer a young bullock without blemish, and a ram out of the flock without blemish.
\vs Eze 43:24 And thou shalt offer them before the \textsc{Lord}, and the priests shall cast salt upon them, and they shall offer them up \bibemph{for} a burnt offering unto the \textsc{Lord}.
\vs Eze 43:25 Seven days shalt thou prepare every day a goat \bibemph{for} a sin offering: they shall also prepare a young bullock, and a ram out of the flock, without blemish.
\vs Eze 43:26 Seven days shall they purge the altar and purify it; and they shall consecrate themselves.
\fn{consecrate...: Heb. fill their hands.}
\vs Eze 43:27 And when these days are expired, it shall be, \bibemph{that} upon the eighth day, and \bibemph{so} forward, the priests shall make your burnt offerings upon the altar, and your peace offerings; and I will accept you, saith the Lord GOD.
\fn{peace...: or, thank offerings.}
\vs Eze 44:1 Then he brought me back the way of the gate of the outward sanctuary which looketh toward the east; and it \bibemph{was} shut.
\vs Eze 44:2 Then said the \textsc{Lord} unto me; This gate shall be shut, it shall not be opened, and no man shall enter in by it; because the \textsc{Lord}, the God of Israel, hath entered in by it, therefore it shall be shut.
\vs Eze 44:3 \bibemph{It is} for the prince; the prince, he shall sit in it to eat bread before the \textsc{Lord}; he shall enter by the way of the porch of \bibemph{that} gate, and shall go out by the way of the same.
\vs Eze 44:4 Then brought he me the way of the north gate before the house: and I looked, and, behold, the glory of the \textsc{Lord} filled the house of the \textsc{Lord}: and I fell upon my face.
\vs Eze 44:5 And the \textsc{Lord} said unto me, Son of man, mark well, and behold with thine eyes, and hear with thine ears all that I say unto thee concerning all the ordinances of the house of the \textsc{Lord}, and all the laws thereof; and mark well the entering in of the house, with every going forth of the sanctuary.
\fn{mark well: Heb. set thine heart. mark well: Heb. set thine heart.}
\vs Eze 44:6 And thou shalt say to the rebellious, \bibemph{even} to the house of Israel, Thus saith the Lord GOD; O ye house of Israel, let it suffice you of all your abominations,
\vs Eze 44:7 In that ye have brought \bibemph{into my sanctuary} strangers, uncircumcised in heart, and uncircumcised in flesh, to be in my sanctuary, to pollute it, \bibemph{even} my house, when ye offer my bread, the fat and the blood, and they have broken my covenant because of all your abominations.
\fn{strangers: Heb. children of a stranger.}
\vs Eze 44:8 And ye have not kept the charge of mine holy things: but ye have set keepers of my charge in my sanctuary for yourselves.
\fn{my charge: or, my ward, or, ordinance.}
\vs Eze 44:9 Thus saith the Lord GOD; No stranger, uncircumcised in heart, nor uncircumcised in flesh, shall enter into my sanctuary, of any stranger that \bibemph{is} among the children of Israel.
\vs Eze 44:10 And the Levites that are gone away far from me, when Israel went astray, which went astray away from me after their idols; they shall even bear their iniquity.
\vs Eze 44:11 Yet they shall be ministers in my sanctuary, \bibemph{having} charge at the gates of the house, and ministering to the house: they shall slay the burnt offering and the sacrifice for the people, and they shall stand before them to minister unto them.
\vs Eze 44:12 Because they ministered unto them before their idols, and caused the house of Israel to fall into iniquity; therefore have I lifted up mine hand against them, saith the Lord GOD, and they shall bear their iniquity.
\fn{caused...: Heb. were for a stumblingblock of iniquity unto, etc.}
\vs Eze 44:13 And they shall not come near unto me, to do the office of a priest unto me, nor to come near to any of my holy things, in the most holy \bibemph{place}: but they shall bear their shame, and their abominations which they have committed.
\vs Eze 44:14 But I will make them keepers of the charge of the house, for all the service thereof, and for all that shall be done therein.
\vs Eze 44:15 But the priests the Levites, the sons of Zadok, that kept the charge of my sanctuary when the children of Israel went astray from me, they shall come near to me to minister unto me, and they shall stand before me to offer unto me the fat and the blood, saith the Lord GOD:
\vs Eze 44:16 They shall enter into my sanctuary, and they shall come near to my table, to minister unto me, and they shall keep my charge.
\vs Eze 44:17 And it shall come to pass, \bibemph{that} when they enter in at the gates of the inner court, they shall be clothed with linen garments; and no wool shall come upon them, whiles they minister in the gates of the inner court, and within.
\vs Eze 44:18 They shall have linen bonnets upon their heads, and shall have linen breeches upon their loins; they shall not gird \bibemph{themselves} with any thing that causeth sweat.
\fn{with...: or, in sweating places: Heb. in, or, with sweat.}
\vs Eze 44:19 And when they go forth into the utter court, \bibemph{even} into the utter court to the people, they shall put off their garments wherein they ministered, and lay them in the holy chambers, and they shall put on other garments; and they shall not sanctify the people with their garments.
\vs Eze 44:20 Neither shall they shave their heads, nor suffer their locks to grow long; they shall only poll their heads.
\vs Eze 44:21 Neither shall any priest drink wine, when they enter into the inner court.
\vs Eze 44:22 Neither shall they take for their wives a widow, nor her that is put away: but they shall take maidens of the seed of the house of Israel, or a widow that had a priest before.
\fn{put...: Heb. thrust forth. that had...: Heb. from a priest.}
\vs Eze 44:23 And they shall teach my people \bibemph{the difference} between the holy and profane, and cause them to discern between the unclean and the clean.
\vs Eze 44:24 And in controversy they shall stand in judgment; \bibemph{and} they shall judge it according to my judgments: and they shall keep my laws and my statutes in all mine assemblies; and they shall hallow my sabbaths.
\vs Eze 44:25 And they shall come at no dead person to defile themselves: but for father, or for mother, or for son, or for daughter, for brother, or for sister that hath had no husband, they may defile themselves.
\vs Eze 44:26 And after he is cleansed, they shall reckon unto him seven days.
\vs Eze 44:27 And in the day that he goeth into the sanctuary, unto the inner court, to minister in the sanctuary, he shall offer his sin offering, saith the Lord GOD.
\vs Eze 44:28 And it shall be unto them for an inheritance: I \bibemph{am} their inheritance: and ye shall give them no possession in Israel: I \bibemph{am} their possession.
\vs Eze 44:29 They shall eat the meat offering, and the sin offering, and the trespass offering; and every dedicated thing in Israel shall be theirs.
\fn{dedicated: or, devoted.}
\vs Eze 44:30 And the first of all the firstfruits of all \bibemph{things}, and every oblation of all, of every \bibemph{sort} of your oblations, shall be the priest's: ye shall also give unto the priest the first of your dough, that he may cause the blessing to rest in thine house.
\fn{And the first: or, And the chief.}
\vs Eze 44:31 The priests shall not eat of any thing that is dead of itself, or torn, whether it be fowl or beast.
\vs Eze 45:1 Moreover, when ye shall divide by lot the land for inheritance, ye shall offer an oblation unto the \textsc{Lord}, an holy portion of the land: the length \bibemph{shall be} the length of five and twenty thousand \bibemph{reeds}, and the breadth \bibemph{shall be} ten thousand. This \bibemph{shall be} holy in all the borders thereof round about.
\fn{when...: Heb. when ye cause the land to fall. an holy...: Heb. holiness.}
\vs Eze 45:2 Of this there shall be for the sanctuary five hundred \bibemph{in length}, with five hundred \bibemph{in breadth}, square round about; and fifty cubits round about for the suburbs thereof.
\fn{suburbs: or, void places.}
\vs Eze 45:3 And of this measure shalt thou measure the length of five and twenty thousand, and the breadth of ten thousand: and in it shall be the sanctuary \bibemph{and} the most holy \bibemph{place}.
\vs Eze 45:4 The holy \bibemph{portion} of the land shall be for the priests the ministers of the sanctuary, which shall come near to minister unto the \textsc{Lord}: and it shall be a place for their houses, and an holy place for the sanctuary.
\vs Eze 45:5 And the five and twenty thousand of length, and the ten thousand of breadth, shall also the Levites, the ministers of the house, have for themselves, for a possession for twenty chambers.
\vs Eze 45:6 And ye shall appoint the possession of the city five thousand broad, and five and twenty thousand long, over against the oblation of the holy \bibemph{portion}: it shall be for the whole house of Israel.
\vs Eze 45:7 And a \bibemph{portion shall be} for the prince on the one side and on the other side of the oblation of the holy \bibemph{portion}, and of the possession of the city, before the oblation of the holy \bibemph{portion}, and before the possession of the city, from the west side westward, and from the east side eastward: and the length \bibemph{shall be} over against one of the portions, from the west border unto the east border.
\vs Eze 45:8 In the land shall be his possession in Israel: and my princes shall no more oppress my people; and \bibemph{the rest of} the land shall they give to the house of Israel according to their tribes.
\vs Eze 45:9 Thus saith the Lord GOD; Let it suffice you, O princes of Israel: remove violence and spoil, and execute judgment and justice, take away your exactions from my people, saith the Lord GOD.
\fn{exactions: Heb. expulsions.}
\vs Eze 45:10 Ye shall have just balances, and a just ephah, and a just bath.
\vs Eze 45:11 The ephah and the bath shall be of one measure, that the bath may contain the tenth part of an homer, and the ephah the tenth part of an homer: the measure thereof shall be after the homer.
\vs Eze 45:12 And the shekel \bibemph{shall be} twenty gerahs: twenty shekels, five and twenty shekels, fifteen shekels, shall be your maneh.
\vs Eze 45:13 This \bibemph{is} the oblation that ye shall offer; the sixth part of an ephah of an homer of wheat, and ye shall give the sixth part of an ephah of an homer of barley:
\vs Eze 45:14 Concerning the ordinance of oil, the bath of oil, \bibemph{ye shall offer} the tenth part of a bath out of the cor, \bibemph{which is} an homer of ten baths; for ten baths \bibemph{are} an homer:
\vs Eze 45:15 And one lamb out of the flock, out of two hundred, out of the fat pastures of Israel; for a meat offering, and for a burnt offering, and for peace offerings, to make reconciliation for them, saith the Lord GOD.
\fn{lamb: or, kid. peace...: or, thank offerings.}
\vs Eze 45:16 All the people of the land shall give this oblation for the prince in Israel.
\fn{shall...: Heb. shall be for. for: or, with.}
\vs Eze 45:17 And it shall be the prince's part \bibemph{to give} burnt offerings, and meat offerings, and drink offerings, in the feasts, and in the new moons, and in the sabbaths, in all solemnities of the house of Israel: he shall prepare the sin offering, and the meat offering, and the burnt offering, and the peace offerings, to make reconciliation for the house of Israel.
\fn{peace...: or, thank offerings.}
\vs Eze 45:18 Thus saith the Lord GOD; In the first \bibemph{month}, in the first \bibemph{day} of the month, thou shalt take a young bullock without blemish, and cleanse the sanctuary:
\vs Eze 45:19 And the priest shall take of the blood of the sin offering, and put \bibemph{it} upon the posts of the house, and upon the four corners of the settle of the altar, and upon the posts of the gate of the inner court.
\vs Eze 45:20 And so thou shalt do the seventh \bibemph{day} of the month for every one that erreth, and for \bibemph{him that is} simple: so shall ye reconcile the house.
\vs Eze 45:21 In the first \bibemph{month}, in the fourteenth day of the month, ye shall have the passover, a feast of seven days; unleavened bread shall be eaten.
\vs Eze 45:22 And upon that day shall the prince prepare for himself and for all the people of the land a bullock \bibemph{for} a sin offering.
\vs Eze 45:23 And seven days of the feast he shall prepare a burnt offering to the \textsc{Lord}, seven bullocks and seven rams without blemish daily the seven days; and a kid of the goats daily \bibemph{for} a sin offering.
\vs Eze 45:24 And he shall prepare a meat offering of an ephah for a bullock, and an ephah for a ram, and an hin of oil for an ephah.
\vs Eze 45:25 In the seventh \bibemph{month}, in the fifteenth day of the month, shall he do the like in the feast of the seven days, according to the sin offering, according to the burnt offering, and according to the meat offering, and according to the oil.
\vs Eze 46:1 Thus saith the Lord GOD; The gate of the inner court that looketh toward the east shall be shut the six working days; but on the sabbath it shall be opened, and in the day of the new moon it shall be opened.
\vs Eze 46:2 And the prince shall enter by the way of the porch of \bibemph{that} gate without, and shall stand by the post of the gate, and the priests shall prepare his burnt offering and his peace offerings, and he shall worship at the threshold of the gate: then he shall go forth; but the gate shall not be shut until the evening.
\vs Eze 46:3 Likewise the people of the land shall worship at the door of this gate before the \textsc{Lord} in the sabbaths and in the new moons.
\vs Eze 46:4 And the burnt offering that the prince shall offer unto the \textsc{Lord} in the sabbath day \bibemph{shall be} six lambs without blemish, and a ram without blemish.
\vs Eze 46:5 And the meat offering \bibemph{shall be} an ephah for a ram, and the meat offering for the lambs as he shall be able to give, and an hin of oil to an ephah.
\fn{as he...: Heb. the gift of his hand.}
\vs Eze 46:6 And in the day of the new moon \bibemph{it shall be} a young bullock without blemish, and six lambs, and a ram: they shall be without blemish.
\vs Eze 46:7 And he shall prepare a meat offering, an ephah for a bullock, and an ephah for a ram, and for the lambs according as his hand shall attain unto, and an hin of oil to an ephah.
\vs Eze 46:8 And when the prince shall enter, he shall go in by the way of the porch of \bibemph{that} gate, and he shall go forth by the way thereof.
\vs Eze 46:9 But when the people of the land shall come before the \textsc{Lord} in the solemn feasts, he that entereth in by the way of the north gate to worship shall go out by the way of the south gate; and he that entereth by the way of the south gate shall go forth by the way of the north gate: he shall not return by the way of the gate whereby he came in, but shall go forth over against it.
\vs Eze 46:10 And the prince in the midst of them, when they go in, shall go in; and when they go forth, shall go forth.
\vs Eze 46:11 And in the feasts and in the solemnities the meat offering shall be an ephah to a bullock, and an ephah to a ram, and to the lambs as he is able to give, and an hin of oil to an ephah.
\vs Eze 46:12 Now when the prince shall prepare a voluntary burnt offering or peace offerings voluntarily unto the \textsc{Lord}, \bibemph{one} shall then open him the gate that looketh toward the east, and he shall prepare his burnt offering and his peace offerings, as he did on the sabbath day: then he shall go forth; and after his going forth \bibemph{one} shall shut the gate.
\vs Eze 46:13 Thou shalt daily prepare a burnt offering unto the \textsc{Lord} \bibemph{of} a lamb of the first year without blemish: thou shalt prepare it every morning.
\fn{of the...: Heb. a son of his year. every...: Heb. morning by morning.}
\vs Eze 46:14 And thou shalt prepare a meat offering for it every morning, the sixth part of an ephah, and the third part of an hin of oil, to temper with the fine flour; a meat offering continually by a perpetual ordinance unto the \textsc{Lord}.
\vs Eze 46:15 Thus shall they prepare the lamb, and the meat offering, and the oil, every morning \bibemph{for} a continual burnt offering.
\vs Eze 46:16 Thus saith the Lord GOD; If the prince give a gift unto any of his sons, the inheritance thereof shall be his sons'; it \bibemph{shall be} their possession by inheritance.
\vs Eze 46:17 But if he give a gift of his inheritance to one of his servants, then it shall be his to the year of liberty; after it shall return to the prince: but his inheritance shall be his sons' for them.
\vs Eze 46:18 Moreover the prince shall not take of the people's inheritance by oppression, to thrust them out of their possession; \bibemph{but} he shall give his sons inheritance out of his own possession: that my people be not scattered every man from his possession.
\vs Eze 46:19 After he brought me through the entry, which \bibemph{was} at the side of the gate, into the holy chambers of the priests, which looked toward the north: and, behold, there \bibemph{was} a place on the two sides westward.
\vs Eze 46:20 Then said he unto me, This \bibemph{is} the place where the priests shall boil the trespass offering and the sin offering, where they shall bake the meat offering; that they bear \bibemph{them} not out into the utter court, to sanctify the people.
\vs Eze 46:21 Then he brought me forth into the utter court, and caused me to pass by the four corners of the court; and, behold, in every corner of the court \bibemph{there was} a court.
\fn{in every...: Heb. a court in a corner of a court, and a court in a corner of a court.}
\vs Eze 46:22 In the four corners of the court \bibemph{there were} courts joined of forty \bibemph{cubits} long and thirty broad: these four corners \bibemph{were} of one measure.
\fn{joined: or, made with chimneys. corners were...: Heb. cornered, etc.}
\vs Eze 46:23 And \bibemph{there was} a row \bibemph{of building} round about in them, round about them four, and \bibemph{it was} made with boiling places under the rows round about.
\vs Eze 46:24 Then said he unto me, These \bibemph{are} the places of them that boil, where the ministers of the house shall boil the sacrifice of the people.
\vs Eze 47:1 Afterward he brought me again unto the door of the house; and, behold, waters issued out from under the threshold of the house eastward: for the forefront of the house \bibemph{stood toward} the east, and the waters came down from under from the right side of the house, at the south \bibemph{side} of the altar.
\vs Eze 47:2 Then brought he me out of the way of the gate northward, and led me about the way without unto the utter gate by the way that looketh eastward; and, behold, there ran out waters on the right side.
\vs Eze 47:3 And when the man that had the line in his hand went forth eastward, he measured a thousand cubits, and he brought me through the waters; the waters \bibemph{were} to the ankles.
\fn{the waters were...: Heb. waters of the ankles.}
\vs Eze 47:4 Again he measured a thousand, and brought me through the waters; the waters \bibemph{were} to the knees. Again he measured a thousand, and brought me through; the waters \bibemph{were} to the loins.
\vs Eze 47:5 Afterward he measured a thousand; \bibemph{and it was} a river that I could not pass over: for the waters were risen, waters to swim in, a river that could not be passed over.
\fn{waters to...: Heb. waters of swimming.}
\vs Eze 47:6 And he said unto me, Son of man, hast thou seen \bibemph{this}? Then he brought me, and caused me to return to the brink of the river.
\vs Eze 47:7 Now when I had returned, behold, at the bank of the river \bibemph{were} very many trees on the one side and on the other.
\fn{bank: Heb. lip.}
\vs Eze 47:8 Then said he unto me, These waters issue out toward the east country, and go down into the desert, and go into the sea: \bibemph{which being} brought forth into the sea, the waters shall be healed.
\fn{desert: or, plain.}
\vs Eze 47:9 And it shall come to pass, \bibemph{that} every thing that liveth, which moveth, whithersoever the rivers shall come, shall live: and there shall be a very great multitude of fish, because these waters shall come thither: for they shall be healed; and every thing shall live whither the river cometh.
\fn{rivers: Heb. two rivers.}
\vs Eze 47:10 And it shall come to pass, \bibemph{that} the fishers shall stand upon it from Engedi even unto Eneglaim; they shall be a \bibemph{place} to spread forth nets; their fish shall be according to their kinds, as the fish of the great sea, exceeding many.
\vs Eze 47:11 But the miry places thereof and the marishes thereof shall not be healed; they shall be given to salt.
\fn{shall not...: or, and that which shall not be healed.}
\vs Eze 47:12 And by the river upon the bank thereof, on this side and on that side, shall grow all trees for meat, whose leaf shall not fade, neither shall the fruit thereof be consumed: it shall bring forth new fruit according to his months, because their waters they issued out of the sanctuary: and the fruit thereof shall be for meat, and the leaf thereof for medicine.
\fn{shall grow: Heb. shall come up. new: or, principal. for medicine: or, for bruises and sores.}
\vs Eze 47:13 Thus saith the Lord GOD; This \bibemph{shall be} the border, whereby ye shall inherit the land according to the twelve tribes of Israel: Joseph \bibemph{shall have two} portions.
\vs Eze 47:14 And ye shall inherit it, one as well as another: \bibemph{concerning} the which I lifted up mine hand to give it unto your fathers: and this land shall fall unto you for inheritance.
\fn{lifted...: or, swore.}
\vs Eze 47:15 And this \bibemph{shall be} the border of the land toward the north side, from the great sea, the way of Hethlon, as men go to Zedad;
\vs Eze 47:16 Hamath, Berothah, Sibraim, which \bibemph{is} between the border of Damascus and the border of Hamath; Hazarhatticon, which \bibemph{is} by the coast of Hauran.
\fn{Hazarhatticon: or, the middle village.}
\vs Eze 47:17 And the border from the sea shall be Hazarenan, the border of Damascus, and the north northward, and the border of Hamath. And \bibemph{this is} the north side.
\vs Eze 47:18 And the east side ye shall measure from Hauran, and from Damascus, and from Gilead, and from the land of Israel \bibemph{by} Jordan, from the border unto the east sea. And \bibemph{this is} the east side.
\fn{from (Hauran, Damascus, Gilead, the land): Heb. from between.}
\vs Eze 47:19 And the south side southward, from Tamar \bibemph{even} to the waters of strife \bibemph{in} Kadesh, the river to the great sea. And \bibemph{this is} the south side southward.
\fn{strife: or, Meribah. river: or, valley. is the south side southward: or, is the south side toward Teman.}
\vs Eze 47:20 The west side also \bibemph{shall be} the great sea from the border, till a man come over against Hamath. This \bibemph{is} the west side.
\vs Eze 47:21 So shall ye divide this land unto you according to the tribes of Israel.
\vs Eze 47:22 And it shall come to pass, \bibemph{that} ye shall divide it by lot for an inheritance unto you, and to the strangers that sojourn among you, which shall beget children among you: and they shall be unto you as born in the country among the children of Israel; they shall have inheritance with you among the tribes of Israel.
\vs Eze 47:23 And it shall come to pass, \bibemph{that} in what tribe the stranger sojourneth, there shall ye give \bibemph{him} his inheritance, saith the Lord GOD.
\vs Eze 48:1 Now these \bibemph{are} the names of the tribes. From the north end to the coast of the way of Hethlon, as one goeth to Hamath, Hazarenan, the border of Damascus northward, to the coast of Hamath; for these are his sides east \bibemph{and} west; a \bibemph{portion for} Dan.
\fn{a portion: Heb. one portion.}
\vs Eze 48:2 And by the border of Dan, from the east side unto the west side, a \bibemph{portion for} Asher.
\vs Eze 48:3 And by the border of Asher, from the east side even unto the west side, a \bibemph{portion for} Naphtali.
\vs Eze 48:4 And by the border of Naphtali, from the east side unto the west side, a \bibemph{portion for} Manasseh.
\vs Eze 48:5 And by the border of Manasseh, from the east side unto the west side, a \bibemph{portion for} Ephraim.
\vs Eze 48:6 And by the border of Ephraim, from the east side even unto the west side, a \bibemph{portion for} Reuben.
\vs Eze 48:7 And by the border of Reuben, from the east side unto the west side, a \bibemph{portion for} Judah.
\vs Eze 48:8 And by the border of Judah, from the east side unto the west side, shall be the offering which ye shall offer of five and twenty thousand \bibemph{reeds in} breadth, and \bibemph{in} length as one of the \bibemph{other} parts, from the east side unto the west side: and the sanctuary shall be in the midst of it.
\vs Eze 48:9 The oblation that ye shall offer unto the \textsc{Lord} \bibemph{shall be} of five and twenty thousand in length, and of ten thousand in breadth.
\vs Eze 48:10 And for them, \bibemph{even} for the priests, shall be \bibemph{this} holy oblation; toward the north five and twenty thousand \bibemph{in length}, and toward the west ten thousand in breadth, and toward the east ten thousand in breadth, and toward the south five and twenty thousand in length: and the sanctuary of the \textsc{Lord} shall be in the midst thereof.
\vs Eze 48:11 \bibemph{It shall be} for the priests that are sanctified of the sons of Zadok; which have kept my charge, which went not astray when the children of Israel went astray, as the Levites went astray.
\fn{It...: or, The sanctified portion shall be for the priests. charge: or, ward, or, ordinance.}
\vs Eze 48:12 And \bibemph{this} oblation of the land that is offered shall be unto them a thing most holy by the border of the Levites.
\vs Eze 48:13 And over against the border of the priests the Levites \bibemph{shall have} five and twenty thousand in length, and ten thousand in breadth: all the length \bibemph{shall be} five and twenty thousand, and the breadth ten thousand.
\vs Eze 48:14 And they shall not sell of it, neither exchange, nor alienate the firstfruits of the land: for \bibemph{it is} holy unto the \textsc{Lord}.
\vs Eze 48:15 And the five thousand, that are left in the breadth over against the five and twenty thousand, shall be a profane \bibemph{place} for the city, for dwelling, and for suburbs: and the city shall be in the midst thereof.
\vs Eze 48:16 And these \bibemph{shall be} the measures thereof; the north side four thousand and five hundred, and the south side four thousand and five hundred, and on the east side four thousand and five hundred, and the west side four thousand and five hundred.
\vs Eze 48:17 And the suburbs of the city shall be toward the north two hundred and fifty, and toward the south two hundred and fifty, and toward the east two hundred and fifty, and toward the west two hundred and fifty.
\vs Eze 48:18 And the residue in length over against the oblation of the holy \bibemph{portion shall be} ten thousand eastward, and ten thousand westward: and it shall be over against the oblation of the holy \bibemph{portion}; and the increase thereof shall be for food unto them that serve the city.
\vs Eze 48:19 And they that serve the city shall serve it out of all the tribes of Israel.
\vs Eze 48:20 All the oblation \bibemph{shall be} five and twenty thousand by five and twenty thousand: ye shall offer the holy oblation foursquare, with the possession of the city.
\vs Eze 48:21 And the residue \bibemph{shall be} for the prince, on the one side and on the other of the holy oblation, and of the possession of the city, over against the five and twenty thousand of the oblation toward the east border, and westward over against the five and twenty thousand toward the west border, over against the portions for the prince: and it shall be the holy oblation; and the sanctuary of the house \bibemph{shall be} in the midst thereof.
\vs Eze 48:22 Moreover from the possession of the Levites, and from the possession of the city, \bibemph{being} in the midst \bibemph{of that} which is the prince's, between the border of Judah and the border of Benjamin, shall be for the prince.
\vs Eze 48:23 As for the rest of the tribes, from the east side unto the west side, Benjamin \bibemph{shall have} a \bibemph{portion}.
\fn{A portion: Heb. one portion.}
\vs Eze 48:24 And by the border of Benjamin, from the east side unto the west side, Simeon \bibemph{shall have} a \bibemph{portion}.
\vs Eze 48:25 And by the border of Simeon, from the east side unto the west side, Issachar a \bibemph{portion}.
\vs Eze 48:26 And by the border of Issachar, from the east side unto the west side, Zebulun a \bibemph{portion}.
\vs Eze 48:27 And by the border of Zebulun, from the east side unto the west side, Gad a \bibemph{portion}.
\vs Eze 48:28 And by the border of Gad, at the south side southward, the border shall be even from Tamar \bibemph{unto} the waters of strife \bibemph{in} Kadesh, \bibemph{and} to the river toward the great sea.
\fn{strife...: or, Meribahkadesh.}
\vs Eze 48:29 This \bibemph{is} the land which ye shall divide by lot unto the tribes of Israel for inheritance, and these \bibemph{are} their portions, saith the Lord GOD.
\vs Eze 48:30 And these \bibemph{are} the goings out of the city on the north side, four thousand and five hundred measures.
\vs Eze 48:31 And the gates of the city \bibemph{shall be} after the names of the tribes of Israel: three gates northward; one gate of Reuben, one gate of Judah, one gate of Levi.
\vs Eze 48:32 And at the east side four thousand and five hundred: and three gates; and one gate of Joseph, one gate of Benjamin, one gate of Dan.
\vs Eze 48:33 And at the south side four thousand and five hundred measures: and three gates; one gate of Simeon, one gate of Issachar, one gate of Zebulun.
\vs Eze 48:34 At the west side four thousand and five hundred, \bibemph{with} their three gates; one gate of Gad, one gate of Asher, one gate of Naphtali.
\vs Eze 48:35 \bibemph{It was} round about eighteen thousand \bibemph{measures}: and the name of the city from \bibemph{that} day \bibemph{shall be}, The \textsc{Lord} \bibemph{is} there.
\fn{The \textsc{Lord}...: Heb. Jehovahshammah.}
