\bibbook{the first book of moses,\\\LARGE called\\\bibtitlemain{genesis}}{genesis}{Genesis}{Gen}
\vs Gen 1:1 In the beginning God created the heaven and the earth.
\vs Gen 1:2 And the earth was without form, and void; and darkness \bibemph{was} upon the face of the deep. And the Spirit of God moved upon the face of the waters.
\vs Gen 1:3 And God said, Let there be light: and there was light.
\vs Gen 1:4 And God saw the light, that \bibemph{it was} good: and God divided the light from the darkness.
\fn{the light from...: Heb. between the light and between the darkness.}
\vs Gen 1:5 And God called the light Day, and the darkness he called Night. And the evening and the morning were the first day.
\fn{And the evening...: Heb. And the evening was, and the morning was.}
\vs Gen 1:6 And God said, Let there be a firmament in the midst of the waters, and let it divide the waters from the waters.
\fn{firmament: Heb. expansion.}
\vs Gen 1:7 And God made the firmament, and divided the waters which \bibemph{were} under the firmament from the waters which \bibemph{were} above the firmament: and it was so.
\vs Gen 1:8 And God called the firmament Heaven. And the evening and the morning were the second day.
\vs Gen 1:9 And God said, Let the waters under the heaven be gathered together unto one place, and let the dry \bibemph{land} appear: and it was so.
\vs Gen 1:10 And God called the dry \bibemph{land} Earth; and the gathering together of the waters called he Seas: and God saw that \bibemph{it was} good.
\vs Gen 1:11 And God said, Let the earth bring forth grass, the herb yielding seed, \bibemph{and} the fruit tree yielding fruit after his kind, whose seed \bibemph{is} in itself, upon the earth: and it was so.
\fn{grass: Heb. tender grass.}
\vs Gen 1:12 And the earth brought forth grass, \bibemph{and} herb yielding seed after his kind, and the tree yielding fruit, whose seed \bibemph{was} in itself, after his kind: and God saw that \bibemph{it was} good.
\vs Gen 1:13 And the evening and the morning were the third day.
\vs Gen 1:14 And God said, Let there be lights in the firmament of the heaven to divide the day from the night; and let them be for signs, and for seasons, and for days, and years:
\fn{the day...: Heb. between the day and between the night.}
\vs Gen 1:15 And let them be for lights in the firmament of the heaven to give light upon the earth: and it was so.
\vs Gen 1:16 And God made two great lights; the greater light to rule the day, and the lesser light to rule the night: \bibemph{he made} the stars also.
\fn{to rule the day...: Heb. for the rule of the day, etc..}
\vs Gen 1:17 And God set them in the firmament of the heaven to give light upon the earth,
\vs Gen 1:18 And to rule over the day and over the night, and to divide the light from the darkness: and God saw that \bibemph{it was} good.
\vs Gen 1:19 And the evening and the morning were the fourth day.
\vs Gen 1:20 And God said, Let the waters bring forth abundantly the moving creature that hath life, and fowl \bibemph{that} may fly above the earth in the open firmament of heaven.
\fn{moving: or, creeping. creature: Heb. soul. fowl...: Heb. let fowl fly. open...: Heb. face of the firmament of heaven.}
\vs Gen 1:21 And God created great whales, and every living creature that moveth, which the waters brought forth abundantly, after their kind, and every winged fowl after his kind: and God saw that \bibemph{it was} good.
\vs Gen 1:22 And God blessed them, saying, Be fruitful, and multiply, and fill the waters in the seas, and let fowl multiply in the earth.
\vs Gen 1:23 And the evening and the morning were the fifth day.
\vs Gen 1:24 And God said, Let the earth bring forth the living creature after his kind, cattle, and creeping thing, and beast of the earth after his kind: and it was so.
\vs Gen 1:25 And God made the beast of the earth after his kind, and cattle after their kind, and every thing that creepeth upon the earth after his kind: and God saw that \bibemph{it was} good.
\vs Gen 1:26 And God said, Let us make man in our image, after our likeness: and let them have dominion over the fish of the sea, and over the fowl of the air, and over the cattle, and over all the earth, and over every creeping thing that creepeth upon the earth.
\vs Gen 1:27 So God created man in his \bibemph{own} image, in the image of God created he him; male and female created he them.
\vs Gen 1:28 And God blessed them, and God said unto them, Be fruitful, and multiply, and replenish the earth, and subdue it: and have dominion over the fish of the sea, and over the fowl of the air, and over every living thing that moveth upon the earth.
\fn{moveth: Heb. creepeth.}
\vs Gen 1:29 And God said, Behold, I have given you every herb bearing seed, which \bibemph{is} upon the face of all the earth, and every tree, in the which \bibemph{is} the fruit of a tree yielding seed; to you it shall be for meat.
\fn{bearing...: Heb. seeding seed. yielding...: Heb. seeding seed.}
\vs Gen 1:30 And to every beast of the earth, and to every fowl of the air, and to every thing that creepeth upon the earth, wherein \bibemph{there is} life, \bibemph{I have given} every green herb for meat: and it was so.
\fn{life: Heb. a living soul.}
\vs Gen 1:31 And God saw every thing that he had made, and, behold, \bibemph{it was} very good. And the evening and the morning were the sixth day.
\vs Gen 2:1 Thus the heavens and the earth were finished, and all the host of them.
\vs Gen 2:2 And on the seventh day God ended his work which he had made; and he rested on the seventh day from all his work which he had made.
\vs Gen 2:3 And God blessed the seventh day, and sanctified it: because that in it he had rested from all his work which God created and made.
\fn{created...: Heb. created to make.}
\vs Gen 2:4 These \bibemph{are} the generations of the heavens and of the earth when they were created, in the day that the \textsc{Lord} God made the earth and the heavens,
\vs Gen 2:5 And every plant of the field before it was in the earth, and every herb of the field before it grew: for the \textsc{Lord} God had not caused it to rain upon the earth, and \bibemph{there was} not a man to till the ground.
\vs Gen 2:6 But there went up a mist from the earth, and watered the whole face of the ground.
\fn{there...: or, a mist which went up from, etc..}
\vs Gen 2:7 And the \textsc{Lord} God formed man \bibemph{of} the dust of the ground, and breathed into his nostrils the breath of life; and man became a living soul.
\fn{of the dust...: Heb. dust of the ground.}
\vs Gen 2:8 And the \textsc{Lord} God planted a garden eastward in Eden; and there he put the man whom he had formed.
\vs Gen 2:9 And out of the ground made the \textsc{Lord} God to grow every tree that is pleasant to the sight, and good for food; the tree of life also in the midst of the garden, and the tree of knowledge of good and evil.
\vs Gen 2:10 And a river went out of Eden to water the garden; and from thence it was parted, and became into four heads.
\vs Gen 2:11 The name of the first \bibemph{is} Pison: that \bibemph{is} it which compasseth the whole land of Havilah, where \bibemph{there is} gold;
\vs Gen 2:12 And the gold of that land \bibemph{is} good: there \bibemph{is} bdellium and the onyx stone.
\vs Gen 2:13 And the name of the second river \bibemph{is} Gihon: the same \bibemph{is} it that compasseth the whole land of Ethiopia.
\fn{Ethiopia: Heb. Cush.}
\vs Gen 2:14 And the name of the third river \bibemph{is} Hiddekel: that \bibemph{is} it which goeth toward the east of Assyria. And the fourth river \bibemph{is} Euphrates.
\fn{toward...: or, eastward to Assyria.}
\vs Gen 2:15 And the \textsc{Lord} God took the man, and put him into the garden of Eden to dress it and to keep it.
\fn{the man: or, Adam.}
\vs Gen 2:16 And the \textsc{Lord} God commanded the man, saying, Of every tree of the garden thou mayest freely eat:
\fn{thou...: Heb. eating thou shalt eat.}
\vs Gen 2:17 But of the tree of the knowledge of good and evil, thou shalt not eat of it: for in the day that thou eatest thereof thou shalt surely die.
\fn{thou shalt surely...: Heb. dying thou shalt die.}
\vs Gen 2:18 And the \textsc{Lord} God said, \bibemph{It is} not good that the man should be alone; I will make him an help meet for him.
\fn{meet...: Heb. as before him.}
\vs Gen 2:19 And out of the ground the \textsc{Lord} God formed every beast of the field, and every fowl of the air; and brought \bibemph{them} unto Adam to see what he would call them: and whatsoever Adam called every living creature, that \bibemph{was} the name thereof.
\fn{Adam: or, the man.}
\vs Gen 2:20 And Adam gave names to all cattle, and to the fowl of the air, and to every beast of the field; but for Adam there was not found an help meet for him.
\fn{gave: Heb. called.}
\vs Gen 2:21 And the \textsc{Lord} God caused a deep sleep to fall upon Adam, and he slept: and he took one of his ribs, and closed up the flesh instead thereof;
\vs Gen 2:22 And the rib, which the \textsc{Lord} God had taken from man, made he a woman, and brought her unto the man.
\fn{made: Heb. built.}
\vs Gen 2:23 And Adam said, This \bibemph{is} now bone of my bones, and flesh of my flesh: she shall be called Woman, because she was taken out of Man.
\fn{Woman: Heb. Isha. Man: Heb. Ish.}
\vs Gen 2:24 Therefore shall a man leave his father and his mother, and shall cleave unto his wife: and they shall be one flesh.
\vs Gen 2:25 And they were both naked, the man and his wife, and were not ashamed.
\vs Gen 3:1 Now the serpent was more subtil than any beast of the field which the \textsc{Lord} God had made. And he said unto the woman, Yea, hath God said, Ye shall not eat of every tree of the garden?
\fn{Yea...: Heb. Yea, because, etc..}
\vs Gen 3:2 And the woman said unto the serpent, We may eat of the fruit of the trees of the garden:
\vs Gen 3:3 But of the fruit of the tree which \bibemph{is} in the midst of the garden, God hath said, Ye shall not eat of it, neither shall ye touch it, lest ye die.
\vs Gen 3:4 And the serpent said unto the woman, Ye shall not surely die:
\vs Gen 3:5 For God doth know that in the day ye eat thereof, then your eyes shall be opened, and ye shall be as gods, knowing good and evil.
\vs Gen 3:6 And when the woman saw that the tree \bibemph{was} good for food, and that it \bibemph{was} pleasant to the eyes, and a tree to be desired to make \bibemph{one} wise, she took of the fruit thereof, and did eat, and gave also unto her husband with her; and he did eat.
\fn{pleasant: Heb. a desire.}
\vs Gen 3:7 And the eyes of them both were opened, and they knew that they \bibemph{were} naked; and they sewed fig leaves together, and made themselves aprons.
\fn{aprons: or, things to gird about.}
\vs Gen 3:8 And they heard the voice of the \textsc{Lord} God walking in the garden in the cool of the day: and Adam and his wife hid themselves from the presence of the \textsc{Lord} God amongst the trees of the garden.
\fn{cool: Heb. wind.}
\vs Gen 3:9 And the \textsc{Lord} God called unto Adam, and said unto him, Where \bibemph{art} thou?
\vs Gen 3:10 And he said, I heard thy voice in the garden, and I was afraid, because I \bibemph{was} naked; and I hid myself.
\vs Gen 3:11 And he said, Who told thee that thou \bibemph{wast} naked? Hast thou eaten of the tree, whereof I commanded thee that thou shouldest not eat?
\vs Gen 3:12 And the man said, The woman whom thou gavest \bibemph{to be} with me, she gave me of the tree, and I did eat.
\vs Gen 3:13 And the \textsc{Lord} God said unto the woman, What \bibemph{is} this \bibemph{that} thou hast done? And the woman said, The serpent beguiled me, and I did eat.
\vs Gen 3:14 And the \textsc{Lord} God said unto the serpent, Because thou hast done this, thou \bibemph{art} cursed above all cattle, and above every beast of the field; upon thy belly shalt thou go, and dust shalt thou eat all the days of thy life:
\vs Gen 3:15 And I will put enmity between thee and the woman, and between thy seed and her seed; it shall bruise thy head, and thou shalt bruise his heel.
\vs Gen 3:16 Unto the woman he said, I will greatly multiply thy sorrow and thy conception; in sorrow thou shalt bring forth children; and thy desire \bibemph{shall be} to thy husband, and he shall rule over thee.
\fn{to thy...: or, subject to thy husband.}
\vs Gen 3:17 And unto Adam he said, Because thou hast hearkened unto the voice of thy wife, and hast eaten of the tree, of which I commanded thee, saying, Thou shalt not eat of it: cursed \bibemph{is} the ground for thy sake; in sorrow shalt thou eat \bibemph{of} it all the days of thy life;
\vs Gen 3:18 Thorns also and thistles shall it bring forth to thee; and thou shalt eat the herb of the field;
\fn{bring...: Heb. cause to bud.}
\vs Gen 3:19 In the sweat of thy face shalt thou eat bread, till thou return unto the ground; for out of it wast thou taken: for dust thou \bibemph{art}, and unto dust shalt thou return.
\vs Gen 3:20 And Adam called his wife's name Eve; because she was the mother of all living.
\fn{Eve: Heb. Chavah: that is Living.}
\vs Gen 3:21 Unto Adam also and to his wife did the \textsc{Lord} God make coats of skins, and clothed them.
\vs Gen 3:22 And the \textsc{Lord} God said, Behold, the man is become as one of us, to know good and evil: and now, lest he put forth his hand, and take also of the tree of life, and eat, and live for ever:
\vs Gen 3:23 Therefore the \textsc{Lord} God sent him forth from the garden of Eden, to till the ground from whence he was taken.
\vs Gen 3:24 So he drove out the man; and he placed at the east of the garden of Eden Cherubims, and a flaming sword which turned every way, to keep the way of the tree of life.
\vs Gen 4:1 And Adam knew Eve his wife; and she conceived, and bare Cain, and said, I have gotten a man from the \textsc{Lord}.
\fn{Cain: that is Gotten, or, Acquired.}
\vs Gen 4:2 And she again bare his brother Abel. And Abel was a keeper of sheep, but Cain was a tiller of the ground.
\fn{Abel: Heb. Hebel. a keeper: Heb. a feeder.}
\vs Gen 4:3 And in process of time it came to pass, that Cain brought of the fruit of the ground an offering unto the \textsc{Lord}.
\fn{in process...: Heb. at the end of days.}
\vs Gen 4:4 And Abel, he also brought of the firstlings of his flock and of the fat thereof. And the \textsc{Lord} had respect unto Abel and to his offering:
\fn{flock: Heb. sheep, or, goats.}
\vs Gen 4:5 But unto Cain and to his offering he had not respect. And Cain was very wroth, and his countenance fell.
\vs Gen 4:6 And the \textsc{Lord} said unto Cain, Why art thou wroth? and why is thy countenance fallen?
\vs Gen 4:7 If thou doest well, shalt thou not be accepted? and if thou doest not well, sin lieth at the door. And unto thee \bibemph{shall be} his desire, and thou shalt rule over him.
\fn{be accepted: or, have the excellency. unto...: or, subject unto thee.}
\vs Gen 4:8 And Cain talked with Abel his brother: and it came to pass, when they were in the field, that Cain rose up against Abel his brother, and slew him.
\vs Gen 4:9 And the \textsc{Lord} said unto Cain, Where \bibemph{is} Abel thy brother? And he said, I know not: \bibemph{Am} I my brother's keeper?
\vs Gen 4:10 And he said, What hast thou done? the voice of thy brother's blood crieth unto me from the ground.
\fn{blood: Heb. bloods.}
\vs Gen 4:11 And now \bibemph{art} thou cursed from the earth, which hath opened her mouth to receive thy brother's blood from thy hand;
\vs Gen 4:12 When thou tillest the ground, it shall not henceforth yield unto thee her strength; a fugitive and a vagabond shalt thou be in the earth.
\vs Gen 4:13 And Cain said unto the \textsc{Lord}, My punishment \bibemph{is} greater than I can bear.
\fn{My...: or, Mine iniquity is greater than that it may be forgiven.}
\vs Gen 4:14 Behold, thou hast driven me out this day from the face of the earth; and from thy face shall I be hid; and I shall be a fugitive and a vagabond in the earth; and it shall come to pass, \bibemph{that} every one that findeth me shall slay me.
\vs Gen 4:15 And the \textsc{Lord} said unto him, Therefore whosoever slayeth Cain, vengeance shall be taken on him sevenfold. And the \textsc{Lord} set a mark upon Cain, lest any finding him should kill him.
\vs Gen 4:16 And Cain went out from the presence of the \textsc{Lord}, and dwelt in the land of Nod, on the east of Eden.
\vs Gen 4:17 And Cain knew his wife; and she conceived, and bare Enoch: and he builded a city, and called the name of the city, after the name of his son, Enoch.
\fn{Enoch: Heb. Chanoch.}
\vs Gen 4:18 And unto Enoch was born Irad: and Irad begat Mehujael: and Mehujael begat Methusael: and Methusael begat Lamech.
\fn{Lamech: Heb. Lemech.}
\vs Gen 4:19 And Lamech took unto him two wives: the name of the one \bibemph{was} Adah, and the name of the other Zillah.
\vs Gen 4:20 And Adah bare Jabal: he was the father of such as dwell in tents, and \bibemph{of such as have} cattle.
\vs Gen 4:21 And his brother's name \bibemph{was} Jubal: he was the father of all such as handle the harp and organ.
\vs Gen 4:22 And Zillah, she also bare Tubalcain, an instructer of every artificer in brass and iron: and the sister of Tubalcain \bibemph{was} Naamah.
\fn{instructor: Heb. whetter.}
\vs Gen 4:23 And Lamech said unto his wives, Adah and Zillah, Hear my voice; ye wives of Lamech, hearken unto my speech: for I have slain a man to my wounding, and a young man to my hurt.
\fn{I have...: or, I would slay a man in my wound, etc. to my hurt: or, in my hurt.}
\vs Gen 4:24 If Cain shall be avenged sevenfold, truly Lamech seventy and sevenfold.
\vs Gen 4:25 And Adam knew his wife again; and she bare a son, and called his name Seth: For God, \bibemph{said she}, hath appointed me another seed instead of Abel, whom Cain slew.
\fn{Seth: Heb. Sheth: that is Appointed, or, Put.}
\vs Gen 4:26 And to Seth, to him also there was born a son; and he called his name Enos: then began men to call upon the name of the \textsc{Lord}.
\fn{Enos: Heb. Enosh. to call...: or, to call themselves by the name of the Lord.}
\vs Gen 5:1 This \bibemph{is} the book of the generations of Adam. In the day that God created man, in the likeness of God made he him;
\vs Gen 5:2 Male and female created he them; and blessed them, and called their name Adam, in the day when they were created.
\vs Gen 5:3 And Adam lived an hundred and thirty years, and begat \bibemph{a son} in his own likeness, after his image; and called his name Seth:
\vs Gen 5:4 And the days of Adam after he had begotten Seth were eight hundred years: and he begat sons and daughters:
\vs Gen 5:5 And all the days that Adam lived were nine hundred and thirty years: and he died.
\vs Gen 5:6 And Seth lived an hundred and five years, and begat Enos:
\fn{Enos: Heb. Enosh.}
\vs Gen 5:7 And Seth lived after he begat Enos eight hundred and seven years, and begat sons and daughters:
\vs Gen 5:8 And all the days of Seth were nine hundred and twelve years: and he died.
\vs Gen 5:9 And Enos lived ninety years, and begat Cainan:
\fn{Cainan: Heb. Kenan.}
\vs Gen 5:10 And Enos lived after he begat Cainan eight hundred and fifteen years, and begat sons and daughters:
\vs Gen 5:11 And all the days of Enos were nine hundred and five years: and he died.
\vs Gen 5:12 And Cainan lived seventy years, and begat Mahalaleel:
\fn{Mahalaleel: Gr. Maleleel.}
\vs Gen 5:13 And Cainan lived after he begat Mahalaleel eight hundred and forty years, and begat sons and daughters:
\vs Gen 5:14 And all the days of Cainan were nine hundred and ten years: and he died.
\vs Gen 5:15 And Mahalaleel lived sixty and five years, and begat Jared:
\fn{Jared: Heb. Jered.}
\vs Gen 5:16 And Mahalaleel lived after he begat Jared eight hundred and thirty years, and begat sons and daughters:
\vs Gen 5:17 And all the days of Mahalaleel were eight hundred ninety and five years: and he died.
\vs Gen 5:18 And Jared lived an hundred sixty and two years, and he begat Enoch:
\vs Gen 5:19 And Jared lived after he begat Enoch eight hundred years, and begat sons and daughters:
\vs Gen 5:20 And all the days of Jared were nine hundred sixty and two years: and he died.
\vs Gen 5:21 And Enoch lived sixty and five years, and begat Methuselah:
\fn{Methuselah: Gr. Mathusala.}
\vs Gen 5:22 And Enoch walked with God after he begat Methuselah three hundred years, and begat sons and daughters:
\vs Gen 5:23 And all the days of Enoch were three hundred sixty and five years:
\vs Gen 5:24 And Enoch walked with God: and he \bibemph{was} not; for God took him.
\vs Gen 5:25 And Methuselah lived an hundred eighty and seven years, and begat Lamech:
\vs Gen 5:26 And Methuselah lived after he begat Lamech seven hundred eighty and two years, and begat sons and daughters:
\fn{Lamech: Heb. Lemech.}
\vs Gen 5:27 And all the days of Methuselah were nine hundred sixty and nine years: and he died.
\vs Gen 5:28 And Lamech lived an hundred eighty and two years, and begat a son:
\vs Gen 5:29 And he called his name Noah, saying, This \bibemph{same} shall comfort us concerning our work and toil of our hands, because of the ground which the \textsc{Lord} hath cursed.
\fn{Noah: Gr. Noe: that is Rest, or, Comfort.}
\vs Gen 5:30 And Lamech lived after he begat Noah five hundred ninety and five years, and begat sons and daughters:
\vs Gen 5:31 And all the days of Lamech were seven hundred seventy and seven years: and he died.
\vs Gen 5:32 And Noah was five hundred years old: and Noah begat Shem, Ham, and Japheth.
\vs Gen 6:1 And it came to pass, when men began to multiply on the face of the earth, and daughters were born unto them,
\vs Gen 6:2 That the sons of God saw the daughters of men that they \bibemph{were} fair; and they took them wives of all which they chose.
\vs Gen 6:3 And the \textsc{Lord} said, My spirit shall not always strive with man, for that he also \bibemph{is} flesh: yet his days shall be an hundred and twenty years.
\vs Gen 6:4 There were giants in the earth in those days; and also after that, when the sons of God came in unto the daughters of men, and they bare \bibemph{children} to them, the same \bibemph{became} mighty men which \bibemph{were} of old, men of renown.
\vs Gen 6:5 And GOD saw that the wickedness of man \bibemph{was} great in the earth, and \bibemph{that} every imagination of the thoughts of his heart \bibemph{was} only evil continually.
\fn{every...: or, the whole imagination: the Hebrew word signifieth not only the imagination, but also the purposes and desires. continually: Heb. every day.}
\vs Gen 6:6 And it repented the \textsc{Lord} that he had made man on the earth, and it grieved him at his heart.
\vs Gen 6:7 And the \textsc{Lord} said, I will destroy man whom I have created from the face of the earth; both man, and beast, and the creeping thing, and the fowls of the air; for it repenteth me that I have made them.
\fn{both...: Heb. from man unto beast.}
\vs Gen 6:8 But Noah found grace in the eyes of the \textsc{Lord}.
\vs Gen 6:9 These \bibemph{are} the generations of Noah: Noah was a just man \bibemph{and} perfect in his generations, \bibemph{and} Noah walked with God.
\fn{perfect: or, upright.}
\vs Gen 6:10 And Noah begat three sons, Shem, Ham, and Japheth.
\vs Gen 6:11 The earth also was corrupt before God, and the earth was filled with violence.
\vs Gen 6:12 And God looked upon the earth, and, behold, it was corrupt; for all flesh had corrupted his way upon the earth.
\vs Gen 6:13 And God said unto Noah, The end of all flesh is come before me; for the earth is filled with violence through them; and, behold, I will destroy them with the earth.
\fn{with the earth: or, from the earth.}
\vs Gen 6:14 Make thee an ark of gopher wood; rooms shalt thou make in the ark, and shalt pitch it within and without with pitch.
\fn{rooms: Heb. nests.}
\vs Gen 6:15 And this \bibemph{is the fashion} which thou shalt make it \bibemph{of}: The length of the ark \bibemph{shall be} three hundred cubits, the breadth of it fifty cubits, and the height of it thirty cubits.
\vs Gen 6:16 A window shalt thou make to the ark, and in a cubit shalt thou finish it above; and the door of the ark shalt thou set in the side thereof; \bibemph{with} lower, second, and third \bibemph{stories} shalt thou make it.
\vs Gen 6:17 And, behold, I, even I, do bring a flood of waters upon the earth, to destroy all flesh, wherein \bibemph{is} the breath of life, from under heaven; \bibemph{and} every thing that \bibemph{is} in the earth shall die.
\vs Gen 6:18 But with thee will I establish my covenant; and thou shalt come into the ark, thou, and thy sons, and thy wife, and thy sons' wives with thee.
\vs Gen 6:19 And of every living thing of all flesh, two of every \bibemph{sort} shalt thou bring into the ark, to keep \bibemph{them} alive with thee; they shall be male and female.
\vs Gen 6:20 Of fowls after their kind, and of cattle after their kind, of every creeping thing of the earth after his kind, two of every \bibemph{sort} shall come unto thee, to keep \bibemph{them} alive.
\vs Gen 6:21 And take thou unto thee of all food that is eaten, and thou shalt gather \bibemph{it} to thee; and it shall be for food for thee, and for them.
\vs Gen 6:22 Thus did Noah; according to all that God commanded him, so did he.
\vs Gen 7:1 And the \textsc{Lord} said unto Noah, Come thou and all thy house into the ark; for thee have I seen righteous before me in this generation.
\vs Gen 7:2 Of every clean beast thou shalt take to thee by sevens, the male and his female: and of beasts that \bibemph{are} not clean by two, the male and his female.
\fn{by sevens: Heb. seven seven.}
\vs Gen 7:3 Of fowls also of the air by sevens, the male and the female; to keep seed alive upon the face of all the earth.
\fn{by sevens: Heb. seven seven.}
\vs Gen 7:4 For yet seven days, and I will cause it to rain upon the earth forty days and forty nights; and every living substance that I have made will I destroy from off the face of the earth.
\fn{destroy: Heb. blot out.}
\vs Gen 7:5 And Noah did according unto all that the \textsc{Lord} commanded him.
\vs Gen 7:6 And Noah \bibemph{was} six hundred years old when the flood of waters was upon the earth.
\vs Gen 7:7 And Noah went in, and his sons, and his wife, and his sons' wives with him, into the ark, because of the waters of the flood.
\vs Gen 7:8 Of clean beasts, and of beasts that \bibemph{are} not clean, and of fowls, and of every thing that creepeth upon the earth,
\vs Gen 7:9 There went in two and two unto Noah into the ark, the male and the female, as God had commanded Noah.
\vs Gen 7:10 And it came to pass after seven days, that the waters of the flood were upon the earth.
\fn{after...: or, on the seventh day.}
\vs Gen 7:11 In the six hundredth year of Noah's life, in the second month, the seventeenth day of the month, the same day were all the fountains of the great deep broken up, and the windows of heaven were opened.
\fn{windows: or, floodgates.}
\vs Gen 7:12 And the rain was upon the earth forty days and forty nights.
\vs Gen 7:13 In the selfsame day entered Noah, and Shem, and Ham, and Japheth, the sons of Noah, and Noah's wife, and the three wives of his sons with them, into the ark;
\vs Gen 7:14 They, and every beast after his kind, and all the cattle after their kind, and every creeping thing that creepeth upon the earth after his kind, and every fowl after his kind, every bird of every sort.
\fn{sort: Heb. wing.}
\vs Gen 7:15 And they went in unto Noah into the ark, two and two of all flesh, wherein \bibemph{is} the breath of life.
\vs Gen 7:16 And they that went in, went in male and female of all flesh, as God had commanded him: and the \textsc{Lord} shut him in.
\vs Gen 7:17 And the flood was forty days upon the earth; and the waters increased, and bare up the ark, and it was lift up above the earth.
\vs Gen 7:18 And the waters prevailed, and were increased greatly upon the earth; and the ark went upon the face of the waters.
\vs Gen 7:19 And the waters prevailed exceedingly upon the earth; and all the high hills, that \bibemph{were} under the whole heaven, were covered.
\vs Gen 7:20 Fifteen cubits upward did the waters prevail; and the mountains were covered.
\vs Gen 7:21 And all flesh died that moved upon the earth, both of fowl, and of cattle, and of beast, and of every creeping thing that creepeth upon the earth, and every man:
\vs Gen 7:22 All in whose nostrils \bibemph{was} the breath of life, of all that \bibemph{was} in the dry \bibemph{land}, died.
\fn{the breath...: Heb. the breath of the spirit of life.}
\vs Gen 7:23 And every living substance was destroyed which was upon the face of the ground, both man, and cattle, and the creeping things, and the fowl of the heaven; and they were destroyed from the earth: and Noah only remained \bibemph{alive}, and they that \bibemph{were} with him in the ark.
\vs Gen 7:24 And the waters prevailed upon the earth an hundred and fifty days.
\vs Gen 8:1 And God remembered Noah, and every living thing, and all the cattle that \bibemph{was} with him in the ark: and God made a wind to pass over the earth, and the waters asswaged;
\vs Gen 8:2 The fountains also of the deep and the windows of heaven were stopped, and the rain from heaven was restrained;
\vs Gen 8:3 And the waters returned from off the earth continually: and after the end of the hundred and fifty days the waters were abated.
\fn{continually: Heb. in going and returning.}
\vs Gen 8:4 And the ark rested in the seventh month, on the seventeenth day of the month, upon the mountains of Ararat.
\vs Gen 8:5 And the waters decreased continually until the tenth month: in the tenth \bibemph{month}, on the first \bibemph{day} of the month, were the tops of the mountains seen.
\fn{decreased...: Heb. were in going and decreasing.}
\vs Gen 8:6 And it came to pass at the end of forty days, that Noah opened the window of the ark which he had made:
\vs Gen 8:7 And he sent forth a raven, which went forth to and fro, until the waters were dried up from off the earth.
\fn{to...: Heb. in going forth and returning.}
\vs Gen 8:8 Also he sent forth a dove from him, to see if the waters were abated from off the face of the ground;
\vs Gen 8:9 But the dove found no rest for the sole of her foot, and she returned unto him into the ark, for the waters \bibemph{were} on the face of the whole earth: then he put forth his hand, and took her, and pulled her in unto him into the ark.
\fn{pulled...: Heb. caused her to come.}
\vs Gen 8:10 And he stayed yet other seven days; and again he sent forth the dove out of the ark;
\vs Gen 8:11 And the dove came in to him in the evening; and, lo, in her mouth \bibemph{was} an olive leaf pluckt off: so Noah knew that the waters were abated from off the earth.
\vs Gen 8:12 And he stayed yet other seven days; and sent forth the dove; which returned not again unto him any more.
\vs Gen 8:13 And it came to pass in the six hundredth and first year, in the first \bibemph{month}, the first \bibemph{day} of the month, the waters were dried up from off the earth: and Noah removed the covering of the ark, and looked, and, behold, the face of the ground was dry.
\vs Gen 8:14 And in the second month, on the seven and twentieth day of the month, was the earth dried.
\vs Gen 8:15 And God spake unto Noah, saying,
\vs Gen 8:16 Go forth of the ark, thou, and thy wife, and thy sons, and thy sons' wives with thee.
\vs Gen 8:17 Bring forth with thee every living thing that \bibemph{is} with thee, of all flesh, \bibemph{both} of fowl, and of cattle, and of every creeping thing that creepeth upon the earth; that they may breed abundantly in the earth, and be fruitful, and multiply upon the earth.
\vs Gen 8:18 And Noah went forth, and his sons, and his wife, and his sons' wives with him:
\vs Gen 8:19 Every beast, every creeping thing, and every fowl, \bibemph{and} whatsoever creepeth upon the earth, after their kinds, went forth out of the ark.
\fn{kinds: Heb. families.}
\vs Gen 8:20 And Noah builded an altar unto the \textsc{Lord}; and took of every clean beast, and of every clean fowl, and offered burnt offerings on the altar.
\vs Gen 8:21 And the \textsc{Lord} smelled a sweet savour; and the \textsc{Lord} said in his heart, I will not again curse the ground any more for man's sake; for the imagination of man's heart \bibemph{is} evil from his youth; neither will I again smite any more every thing living, as I have done.
\fn{a sweet...: Heb. a savour of rest or, satisfaction. for the imagination: or, through the imagination.}
\vs Gen 8:22 While the earth remaineth, seedtime and harvest, and cold and heat, and summer and winter, and day and night shall not cease.
\fn{While...: Heb. As yet all the days of the earth.}
\vs Gen 9:1 And God blessed Noah and his sons, and said unto them, Be fruitful, and multiply, and replenish the earth.
\vs Gen 9:2 And the fear of you and the dread of you shall be upon every beast of the earth, and upon every fowl of the air, upon all that moveth \bibemph{upon} the earth, and upon all the fishes of the sea; into your hand are they delivered.
\vs Gen 9:3 Every moving thing that liveth shall be meat for you; even as the green herb have I given you all things.
\vs Gen 9:4 But flesh with the life thereof, \bibemph{which is} the blood thereof, shall ye not eat.
\vs Gen 9:5 And surely your blood of your lives will I require; at the hand of every beast will I require it, and at the hand of man; at the hand of every man's brother will I require the life of man.
\vs Gen 9:6 Whoso sheddeth man's blood, by man shall his blood be shed: for in the image of God made he man.
\vs Gen 9:7 And you, be ye fruitful, and multiply; bring forth abundantly in the earth, and multiply therein.
\vs Gen 9:8 And God spake unto Noah, and to his sons with him, saying,
\vs Gen 9:9 And I, behold, I establish my covenant with you, and with your seed after you;
\vs Gen 9:10 And with every living creature that \bibemph{is} with you, of the fowl, of the cattle, and of every beast of the earth with you; from all that go out of the ark, to every beast of the earth.
\vs Gen 9:11 And I will establish my covenant with you; neither shall all flesh be cut off any more by the waters of a flood; neither shall there any more be a flood to destroy the earth.
\vs Gen 9:12 And God said, This \bibemph{is} the token of the covenant which I make between me and you and every living creature that \bibemph{is} with you, for perpetual generations:
\vs Gen 9:13 I do set my bow in the cloud, and it shall be for a token of a covenant between me and the earth.
\vs Gen 9:14 And it shall come to pass, when I bring a cloud over the earth, that the bow shall be seen in the cloud:
\vs Gen 9:15 And I will remember my covenant, which \bibemph{is} between me and you and every living creature of all flesh; and the waters shall no more become a flood to destroy all flesh.
\vs Gen 9:16 And the bow shall be in the cloud; and I will look upon it, that I may remember the everlasting covenant between God and every living creature of all flesh that \bibemph{is} upon the earth.
\vs Gen 9:17 And God said unto Noah, This \bibemph{is} the token of the covenant, which I have established between me and all flesh that \bibemph{is} upon the earth.
\vs Gen 9:18 And the sons of Noah, that went forth of the ark, were Shem, and Ham, and Japheth: and Ham \bibemph{is} the father of Canaan.
\fn{Canaan: Heb. Chenaan.}
\vs Gen 9:19 These \bibemph{are} the three sons of Noah: and of them was the whole earth overspread.
\vs Gen 9:20 And Noah began \bibemph{to be} an husbandman, and he planted a vineyard:
\vs Gen 9:21 And he drank of the wine, and was drunken; and he was uncovered within his tent.
\vs Gen 9:22 And Ham, the father of Canaan, saw the nakedness of his father, and told his two brethren without.
\vs Gen 9:23 And Shem and Japheth took a garment, and laid \bibemph{it} upon both their shoulders, and went backward, and covered the nakedness of their father; and their faces \bibemph{were} backward, and they saw not their father's nakedness.
\vs Gen 9:24 And Noah awoke from his wine, and knew what his younger son had done unto him.
\vs Gen 9:25 And he said, Cursed \bibemph{be} Canaan; a servant of servants shall he be unto his brethren.
\vs Gen 9:26 And he said, Blessed \bibemph{be} the \textsc{Lord} God of Shem; and Canaan shall be his servant.
\fn{his servant: or, servant to them.}
\vs Gen 9:27 God shall enlarge Japheth, and he shall dwell in the tents of Shem; and Canaan shall be his servant.
\fn{enlarge: or, persuade.}
\vs Gen 9:28 And Noah lived after the flood three hundred and fifty years.
\vs Gen 9:29 And all the days of Noah were nine hundred and fifty years: and he died.
\vs Gen 10:1 Now these \bibemph{are} the generations of the sons of Noah, Shem, Ham, and Japheth: and unto them were sons born after the flood.
\vs Gen 10:2 The sons of Japheth; Gomer, and Magog, and Madai, and Javan, and Tubal, and Meshech, and Tiras.
\vs Gen 10:3 And the sons of Gomer; Ashkenaz, and Riphath, and Togarmah.
\vs Gen 10:4 And the sons of Javan; Elishah, and Tarshish, Kittim, and Dodanim.
\fn{Dodanim: or, as some read it, Rodanim.}
\vs Gen 10:5 By these were the isles of the Gentiles divided in their lands; every one after his tongue, after their families, in their nations.
\vs Gen 10:6 And the sons of Ham; Cush, and Mizraim, and Phut, and Canaan.
\vs Gen 10:7 And the sons of Cush; Seba, and Havilah, and Sabtah, and Raamah, and Sabtecha: and the sons of Raamah; Sheba, and Dedan.
\vs Gen 10:8 And Cush begat Nimrod: he began to be a mighty one in the earth.
\vs Gen 10:9 He was a mighty hunter before the \textsc{Lord}: wherefore it is said, Even as Nimrod the mighty hunter before the \textsc{Lord}.
\vs Gen 10:10 And the beginning of his kingdom was Babel, and Erech, and Accad, and Calneh, in the land of Shinar.
\fn{Babel: Gr. Babylon.}
\vs Gen 10:11 Out of that land went forth Asshur, and builded Nineveh, and the city Rehoboth, and Calah,
\fn{went...: or, he went out into Assyria. the city...: or, the streets of the city.}
\vs Gen 10:12 And Resen between Nineveh and Calah: the same \bibemph{is} a great city.
\vs Gen 10:13 And Mizraim begat Ludim, and Anamim, and Lehabim, and Naphtuhim,
\vs Gen 10:14 And Pathrusim, and Casluhim, (out of whom came Philistim,) and Caphtorim.
\vs Gen 10:15 And Canaan begat Sidon his firstborn, and Heth,
\fn{Sidon: Heb. Tzidon.}
\vs Gen 10:16 And the Jebusite, and the Amorite, and the Girgasite,
\vs Gen 10:17 And the Hivite, and the Arkite, and the Sinite,
\vs Gen 10:18 And the Arvadite, and the Zemarite, and the Hamathite: and afterward were the families of the Canaanites spread abroad.
\vs Gen 10:19 And the border of the Canaanites was from Sidon, as thou comest to Gerar, unto Gaza; as thou goest, unto Sodom, and Gomorrah, and Admah, and Zeboim, even unto Lasha.
\fn{Gaza: Heb. Azzah.}
\vs Gen 10:20 These \bibemph{are} the sons of Ham, after their families, after their tongues, in their countries, \bibemph{and} in their nations.
\vs Gen 10:21 Unto Shem also, the father of all the children of Eber, the brother of Japheth the elder, even to him were \bibemph{children} born.
\vs Gen 10:22 The children of Shem; Elam, and Asshur, and Arphaxad, and Lud, and Aram.
\fn{Arphaxad: Heb. Arpachshad.}
\vs Gen 10:23 And the children of Aram; Uz, and Hul, and Gether, and Mash.
\vs Gen 10:24 And Arphaxad begat Salah; and Salah begat Eber.
\fn{Salah: Heb. Shelah.}
\vs Gen 10:25 And unto Eber were born two sons: the name of one \bibemph{was} Peleg; for in his days was the earth divided; and his brother's name \bibemph{was} Joktan.
\fn{Peleg: that is Division.}
\vs Gen 10:26 And Joktan begat Almodad, and Sheleph, and Hazarmaveth, and Jerah,
\vs Gen 10:27 And Hadoram, and Uzal, and Diklah,
\vs Gen 10:28 And Obal, and Abimael, and Sheba,
\vs Gen 10:29 And Ophir, and Havilah, and Jobab: all these \bibemph{were} the sons of Joktan.
\vs Gen 10:30 And their dwelling was from Mesha, as thou goest unto Sephar a mount of the east.
\vs Gen 10:31 These \bibemph{are} the sons of Shem, after their families, after their tongues, in their lands, after their nations.
\vs Gen 10:32 These \bibemph{are} the families of the sons of Noah, after their generations, in their nations: and by these were the nations divided in the earth after the flood.
\vs Gen 11:1 And the whole earth was of one language, and of one speech.
\fn{language: Heb. lip.. speech: Heb. words.}
\vs Gen 11:2 And it came to pass, as they journeyed from the east, that they found a plain in the land of Shinar; and they dwelt there.
\fn{from...: or, eastward.}
\vs Gen 11:3 And they said one to another, Go to, let us make brick, and burn them throughly. And they had brick for stone, and slime had they for morter.
\fn{they said...: Heb. a man said to his neighbour. burn...: Heb. burn them to a burning.}
\vs Gen 11:4 And they said, Go to, let us build us a city and a tower, whose top \bibemph{may reach} unto heaven; and let us make us a name, lest we be scattered abroad upon the face of the whole earth.
\vs Gen 11:5 And the \textsc{Lord} came down to see the city and the tower, which the children of men builded.
\vs Gen 11:6 And the \textsc{Lord} said, Behold, the people \bibemph{is} one, and they have all one language; and this they begin to do: and now nothing will be restrained from them, which they have imagined to do.
\vs Gen 11:7 Go to, let us go down, and there confound their language, that they may not understand one another's speech.
\vs Gen 11:8 So the \textsc{Lord} scattered them abroad from thence upon the face of all the earth: and they left off to build the city.
\vs Gen 11:9 Therefore is the name of it called Babel; because the \textsc{Lord} did there confound the language of all the earth: and from thence did the \textsc{Lord} scatter them abroad upon the face of all the earth.
\fn{Babel: that is, Confusion.}
\vs Gen 11:10 These \bibemph{are} the generations of Shem: Shem \bibemph{was} an hundred years old, and begat Arphaxad two years after the flood:
\vs Gen 11:11 And Shem lived after he begat Arphaxad five hundred years, and begat sons and daughters.
\vs Gen 11:12 And Arphaxad lived five and thirty years, and begat Salah:
\vs Gen 11:13 And Arphaxad lived after he begat Salah four hundred and three years, and begat sons and daughters.
\vs Gen 11:14 And Salah lived thirty years, and begat Eber:
\vs Gen 11:15 And Salah lived after he begat Eber four hundred and three years, and begat sons and daughters.
\vs Gen 11:16 And Eber lived four and thirty years, and begat Peleg:
\fn{Peleg: Gr. Phalec.}
\vs Gen 11:17 And Eber lived after he begat Peleg four hundred and thirty years, and begat sons and daughters.
\vs Gen 11:18 And Peleg lived thirty years, and begat Reu:
\vs Gen 11:19 And Peleg lived after he begat Reu two hundred and nine years, and begat sons and daughters.
\vs Gen 11:20 And Reu lived two and thirty years, and begat Serug:
\fn{Serug: Gr. Saruch.}
\vs Gen 11:21 And Reu lived after he begat Serug two hundred and seven years, and begat sons and daughters.
\vs Gen 11:22 And Serug lived thirty years, and begat Nahor:
\vs Gen 11:23 And Serug lived after he begat Nahor two hundred years, and begat sons and daughters.
\vs Gen 11:24 And Nahor lived nine and twenty years, and begat Terah:
\fn{Terah: Gr. Thara.}
\vs Gen 11:25 And Nahor lived after he begat Terah an hundred and nineteen years, and begat sons and daughters.
\vs Gen 11:26 And Terah lived seventy years, and begat Abram, Nahor, and Haran.
\vs Gen 11:27 Now these \bibemph{are} the generations of Terah: Terah begat Abram, Nahor, and Haran; and Haran begat Lot.
\vs Gen 11:28 And Haran died before his father Terah in the land of his nativity, in Ur of the Chaldees.
\vs Gen 11:29 And Abram and Nahor took them wives: the name of Abram's wife \bibemph{was} Sarai; and the name of Nahor's wife, Milcah, the daughter of Haran, the father of Milcah, and the father of Iscah.
\vs Gen 11:30 But Sarai was barren; she \bibemph{had} no child.
\vs Gen 11:31 And Terah took Abram his son, and Lot the son of Haran his son's son, and Sarai his daughter in law, his son Abram's wife; and they went forth with them from Ur of the Chaldees, to go into the land of Canaan; and they came unto Haran, and dwelt there.
\vs Gen 11:32 And the days of Terah were two hundred and five years: and Terah died in Haran.
\vs Gen 12:1 Now the \textsc{Lord} had said unto Abram, Get thee out of thy country, and from thy kindred, and from thy father's house, unto a land that I will shew thee:
\vs Gen 12:2 And I will make of thee a great nation, and I will bless thee, and make thy name great; and thou shalt be a blessing:
\vs Gen 12:3 And I will bless them that bless thee, and curse him that curseth thee: and in thee shall all families of the earth be blessed.
\vs Gen 12:4 So Abram departed, as the \textsc{Lord} had spoken unto him; and Lot went with him: and Abram \bibemph{was} seventy and five years old when he departed out of Haran.
\vs Gen 12:5 And Abram took Sarai his wife, and Lot his brother's son, and all their substance that they had gathered, and the souls that they had gotten in Haran; and they went forth to go into the land of Canaan; and into the land of Canaan they came.
\vs Gen 12:6 And Abram passed through the land unto the place of Sichem, unto the plain of Moreh. And the Canaanite \bibemph{was} then in the land.
\vs Gen 12:7 And the \textsc{Lord} appeared unto Abram, and said, Unto thy seed will I give this land: and there builded he an altar unto the \textsc{Lord}, who appeared unto him.
\vs Gen 12:8 And he removed from thence unto a mountain on the east of Bethel, and pitched his tent, \bibemph{having} Bethel on the west, and Hai on the east: and there he builded an altar unto the \textsc{Lord}, and called upon the name of the \textsc{Lord}.
\vs Gen 12:9 And Abram journeyed, going on still toward the south.
\fn{going...: Heb. in going and journeying.}
\vs Gen 12:10 And there was a famine in the land: and Abram went down into Egypt to sojourn there; for the famine \bibemph{was} grievous in the land.
\vs Gen 12:11 And it came to pass, when he was come near to enter into Egypt, that he said unto Sarai his wife, Behold now, I know that thou \bibemph{art} a fair woman to look upon:
\vs Gen 12:12 Therefore it shall come to pass, when the Egyptians shall see thee, that they shall say, This \bibemph{is} his wife: and they will kill me, but they will save thee alive.
\vs Gen 12:13 Say, I pray thee, thou \bibemph{art} my sister: that it may be well with me for thy sake; and my soul shall live because of thee.
\vs Gen 12:14 And it came to pass, that, when Abram was come into Egypt, the Egyptians beheld the woman that she \bibemph{was} very fair.
\vs Gen 12:15 The princes also of Pharaoh saw her, and commended her before Pharaoh: and the woman was taken into Pharaoh's house.
\vs Gen 12:16 And he entreated Abram well for her sake: and he had sheep, and oxen, and he asses, and menservants, and maidservants, and she asses, and camels.
\vs Gen 12:17 And the \textsc{Lord} plagued Pharaoh and his house with great plagues because of Sarai Abram's wife.
\vs Gen 12:18 And Pharaoh called Abram, and said, What \bibemph{is} this \bibemph{that} thou hast done unto me? why didst thou not tell me that she \bibemph{was} thy wife?
\vs Gen 12:19 Why saidst thou, She \bibemph{is} my sister? so I might have taken her to me to wife: now therefore behold thy wife, take \bibemph{her}, and go thy way.
\vs Gen 12:20 And Pharaoh commanded \bibemph{his} men concerning him: and they sent him away, and his wife, and all that he had.
\vs Gen 13:1 And Abram went up out of Egypt, he, and his wife, and all that he had, and Lot with him, into the south.
\vs Gen 13:2 And Abram \bibemph{was} very rich in cattle, in silver, and in gold.
\vs Gen 13:3 And he went on his journeys from the south even to Bethel, unto the place where his tent had been at the beginning, between Bethel and Hai;
\vs Gen 13:4 Unto the place of the altar, which he had made there at the first: and there Abram called on the name of the \textsc{Lord}.
\vs Gen 13:5 And Lot also, which went with Abram, had flocks, and herds, and tents.
\vs Gen 13:6 And the land was not able to bear them, that they might dwell together: for their substance was great, so that they could not dwell together.
\vs Gen 13:7 And there was a strife between the herdmen of Abram's cattle and the herdmen of Lot's cattle: and the Canaanite and the Perizzite dwelled then in the land.
\vs Gen 13:8 And Abram said unto Lot, Let there be no strife, I pray thee, between me and thee, and between my herdmen and thy herdmen; for we \bibemph{be} brethren.
\fn{brethren: Heb. men brethren.}
\vs Gen 13:9 \bibemph{Is} not the whole land before thee? separate thyself, I pray thee, from me: if \bibemph{thou wilt take} the left hand, then I will go to the right; or if \bibemph{thou depart} to the right hand, then I will go to the left.
\vs Gen 13:10 And Lot lifted up his eyes, and beheld all the plain of Jordan, that it \bibemph{was} well watered every where, before the \textsc{Lord} destroyed Sodom and Gomorrah, \bibemph{even} as the garden of the \textsc{Lord}, like the land of Egypt, as thou comest unto Zoar.
\vs Gen 13:11 Then Lot chose him all the plain of Jordan; and Lot journeyed east: and they separated themselves the one from the other.
\vs Gen 13:12 Abram dwelled in the land of Canaan, and Lot dwelled in the cities of the plain, and pitched \bibemph{his} tent toward Sodom.
\vs Gen 13:13 But the men of Sodom \bibemph{were} wicked and sinners before the \textsc{Lord} exceedingly.
\vs Gen 13:14 And the \textsc{Lord} said unto Abram, after that Lot was separated from him, Lift up now thine eyes, and look from the place where thou art northward, and southward, and eastward, and westward:
\vs Gen 13:15 For all the land which thou seest, to thee will I give it, and to thy seed for ever.
\vs Gen 13:16 And I will make thy seed as the dust of the earth: so that if a man can number the dust of the earth, \bibemph{then} shall thy seed also be numbered.
\vs Gen 13:17 Arise, walk through the land in the length of it and in the breadth of it; for I will give it unto thee.
\vs Gen 13:18 Then Abram removed \bibemph{his} tent, and came and dwelt in the plain of Mamre, which \bibemph{is} in Hebron, and built there an altar unto the \textsc{Lord}.
\fn{plain: Heb. plains.}
\vs Gen 14:1 And it came to pass in the days of Amraphel king of Shinar, Arioch king of Ellasar, Chedorlaomer king of Elam, and Tidal king of nations;
\vs Gen 14:2 \bibemph{That these} made war with Bera king of Sodom, and with Birsha king of Gomorrah, Shinab king of Admah, and Shemeber king of Zeboiim, and the king of Bela, which is Zoar.
\vs Gen 14:3 All these were joined together in the vale of Siddim, which is the salt sea.
\vs Gen 14:4 Twelve years they served Chedorlaomer, and in the thirteenth year they rebelled.
\vs Gen 14:5 And in the fourteenth year came Chedorlaomer, and the kings that \bibemph{were} with him, and smote the Rephaims in Ashteroth Karnaim, and the Zuzims in Ham, and the Emims in Shaveh Kiriathaim,
\fn{Shaveh...: or, The plain of Kiriathaim.}
\vs Gen 14:6 And the Horites in their mount Seir, unto Elparan, which \bibemph{is} by the wilderness.
\fn{Elparan: or, The plain of Paran.}
\vs Gen 14:7 And they returned, and came to Enmishpat, which \bibemph{is} Kadesh, and smote all the country of the Amalekites, and also the Amorites, that dwelt in Hazezontamar.
\vs Gen 14:8 And there went out the king of Sodom, and the king of Gomorrah, and the king of Admah, and the king of Zeboiim, and the king of Bela (the same \bibemph{is} Zoar;) and they joined battle with them in the vale of Siddim;
\vs Gen 14:9 With Chedorlaomer the king of Elam, and with Tidal king of nations, and Amraphel king of Shinar, and Arioch king of Ellasar; four kings with five.
\vs Gen 14:10 And the vale of Siddim \bibemph{was full of} slimepits; and the kings of Sodom and Gomorrah fled, and fell there; and they that remained fled to the mountain.
\vs Gen 14:11 And they took all the goods of Sodom and Gomorrah, and all their victuals, and went their way.
\vs Gen 14:12 And they took Lot, Abram's brother's son, who dwelt in Sodom, and his goods, and departed.
\vs Gen 14:13 And there came one that had escaped, and told Abram the Hebrew; for he dwelt in the plain of Mamre the Amorite, brother of Eshcol, and brother of Aner: and these \bibemph{were} confederate with Abram.
\vs Gen 14:14 And when Abram heard that his brother was taken captive, he armed his trained \bibemph{servants}, born in his own house, three hundred and eighteen, and pursued \bibemph{them} unto Dan.
\fn{armed: or, led forth. trained: or, instructed.}
\vs Gen 14:15 And he divided himself against them, he and his servants, by night, and smote them, and pursued them unto Hobah, which \bibemph{is} on the left hand of Damascus.
\vs Gen 14:16 And he brought back all the goods, and also brought again his brother Lot, and his goods, and the women also, and the people.
\vs Gen 14:17 And the king of Sodom went out to meet him after his return from the slaughter of Chedorlaomer, and of the kings that \bibemph{were} with him, at the valley of Shaveh, which \bibemph{is} the king's dale.
\vs Gen 14:18 And Melchizedek king of Salem brought forth bread and wine: and he \bibemph{was} the priest of the most high God.
\vs Gen 14:19 And he blessed him, and said, Blessed \bibemph{be} Abram of the most high God, possessor of heaven and earth:
\vs Gen 14:20 And blessed be the most high God, which hath delivered thine enemies into thy hand. And he gave him tithes of all.
\vs Gen 14:21 And the king of Sodom said unto Abram, Give me the persons, and take the goods to thyself.
\fn{persons: Heb. souls.}
\vs Gen 14:22 And Abram said to the king of Sodom, I have lift up mine hand unto the \textsc{Lord}, the most high God, the possessor of heaven and earth,
\vs Gen 14:23 That I will not \bibemph{take} from a thread even to a shoelatchet, and that I will not take any thing that \bibemph{is} thine, lest thou shouldest say, I have made Abram rich:
\vs Gen 14:24 Save only that which the young men have eaten, and the portion of the men which went with me, Aner, Eshcol, and Mamre; let them take their portion.
\vs Gen 15:1 After these things the word of the \textsc{Lord} came unto Abram in a vision, saying, Fear not, Abram: I \bibemph{am} thy shield, \bibemph{and} thy exceeding great reward.
\vs Gen 15:2 And Abram said, Lord GOD, what wilt thou give me, seeing I go childless, and the steward of my house \bibemph{is} this Eliezer of Damascus?
\vs Gen 15:3 And Abram said, Behold, to me thou hast given no seed: and, lo, one born in my house is mine heir.
\vs Gen 15:4 And, behold, the word of the \textsc{Lord} \bibemph{came} unto him, saying, This shall not be thine heir; but he that shall come forth out of thine own bowels shall be thine heir.
\vs Gen 15:5 And he brought him forth abroad, and said, Look now toward heaven, and tell the stars, if thou be able to number them: and he said unto him, So shall thy seed be.
\vs Gen 15:6 And he believed in the \textsc{Lord}; and he counted it to him for righteousness.
\vs Gen 15:7 And he said unto him, I \bibemph{am} the \textsc{Lord} that brought thee out of Ur of the Chaldees, to give thee this land to inherit it.
\vs Gen 15:8 And he said, Lord GOD, whereby shall I know that I shall inherit it?
\vs Gen 15:9 And he said unto him, Take me an heifer of three years old, and a she goat of three years old, and a ram of three years old, and a turtledove, and a young pigeon.
\vs Gen 15:10 And he took unto him all these, and divided them in the midst, and laid each piece one against another: but the birds divided he not.
\vs Gen 15:11 And when the fowls came down upon the carcases, Abram drove them away.
\vs Gen 15:12 And when the sun was going down, a deep sleep fell upon Abram; and, lo, an horror of great darkness fell upon him.
\vs Gen 15:13 And he said unto Abram, Know of a surety that thy seed shall be a stranger in a land \bibemph{that is} not theirs, and shall serve them; and they shall afflict them four hundred years;
\vs Gen 15:14 And also that nation, whom they shall serve, will I judge: and afterward shall they come out with great substance.
\vs Gen 15:15 And thou shalt go to thy fathers in peace; thou shalt be buried in a good old age.
\vs Gen 15:16 But in the fourth generation they shall come hither again: for the iniquity of the Amorites \bibemph{is} not yet full.
\vs Gen 15:17 And it came to pass, that, when the sun went down, and it was dark, behold a smoking furnace, and a burning lamp that passed between those pieces.
\fn{a burning...: Heb. a lamp of fire.}
\vs Gen 15:18 In the same day the \textsc{Lord} made a covenant with Abram, saying, Unto thy seed have I given this land, from the river of Egypt unto the great river, the river Euphrates:
\vs Gen 15:19 The Kenites, and the Kenizzites, and the Kadmonites,
\vs Gen 15:20 And the Hittites, and the Perizzites, and the Rephaims,
\vs Gen 15:21 And the Amorites, and the Canaanites, and the Girgashites, and the Jebusites.
\vs Gen 16:1 Now Sarai Abram's wife bare him no children: and she had an handmaid, an Egyptian, whose name \bibemph{was} Hagar.
\vs Gen 16:2 And Sarai said unto Abram, Behold now, the \textsc{Lord} hath restrained me from bearing: I pray thee, go in unto my maid; it may be that I may obtain children by her. And Abram hearkened to the voice of Sarai.
\fn{obtain...: Heb. be built by her.}
\vs Gen 16:3 And Sarai Abram's wife took Hagar her maid the Egyptian, after Abram had dwelt ten years in the land of Canaan, and gave her to her husband Abram to be his wife.
\vs Gen 16:4 And he went in unto Hagar, and she conceived: and when she saw that she had conceived, her mistress was despised in her eyes.
\vs Gen 16:5 And Sarai said unto Abram, My wrong \bibemph{be} upon thee: I have given my maid into thy bosom; and when she saw that she had conceived, I was despised in her eyes: the \textsc{Lord} judge between me and thee.
\vs Gen 16:6 But Abram said unto Sarai, Behold, thy maid \bibemph{is} in thy hand; do to her as it pleaseth thee. And when Sarai dealt hardly with her, she fled from her face.
\fn{as...: Heb. that which is good in thine eyes. dealt...: Heb. afflicted her.}
\vs Gen 16:7 And the angel of the \textsc{Lord} found her by a fountain of water in the wilderness, by the fountain in the way to Shur.
\vs Gen 16:8 And he said, Hagar, Sarai's maid, whence camest thou? and whither wilt thou go? And she said, I flee from the face of my mistress Sarai.
\vs Gen 16:9 And the angel of the \textsc{Lord} said unto her, Return to thy mistress, and submit thyself under her hands.
\vs Gen 16:10 And the angel of the \textsc{Lord} said unto her, I will multiply thy seed exceedingly, that it shall not be numbered for multitude.
\vs Gen 16:11 And the angel of the \textsc{Lord} said unto her, Behold, thou \bibemph{art} with child, and shalt bear a son, and shalt call his name Ishmael; because the \textsc{Lord} hath heard thy affliction.
\fn{Ishmael: that is, God shall hear.}
\vs Gen 16:12 And he will be a wild man; his hand \bibemph{will be} against every man, and every man's hand against him; and he shall dwell in the presence of all his brethren.
\vs Gen 16:13 And she called the name of the \textsc{Lord} that spake unto her, Thou God seest me: for she said, Have I also here looked after him that seeth me?
\vs Gen 16:14 Wherefore the well was called Beerlahairoi; behold, \bibemph{it is} between Kadesh and Bered.
\fn{Beerlahairoi: that is, The well of him that liveth and seeth me.}
\vs Gen 16:15 And Hagar bare Abram a son: and Abram called his son's name, which Hagar bare, Ishmael.
\vs Gen 16:16 And Abram \bibemph{was} fourscore and six years old, when Hagar bare Ishmael to Abram.
\vs Gen 17:1 And when Abram was ninety years old and nine, the \textsc{Lord} appeared to Abram, and said unto him, I \bibemph{am} the Almighty God; walk before me, and be thou perfect.
\fn{perfect: or, upright, or, sincere.}
\vs Gen 17:2 And I will make my covenant between me and thee, and will multiply thee exceedingly.
\vs Gen 17:3 And Abram fell on his face: and God talked with him, saying,
\vs Gen 17:4 As for me, behold, my covenant \bibemph{is} with thee, and thou shalt be a father of many nations.
\fn{many...: Heb. multitude of nations.}
\vs Gen 17:5 Neither shall thy name any more be called Abram, but thy name shall be Abraham; for a father of many nations have I made thee.
\fn{Abraham: that is, Father of a great multitude.}
\vs Gen 17:6 And I will make thee exceeding fruitful, and I will make nations of thee, and kings shall come out of thee.
\vs Gen 17:7 And I will establish my covenant between me and thee and thy seed after thee in their generations for an everlasting covenant, to be a God unto thee, and to thy seed after thee.
\vs Gen 17:8 And I will give unto thee, and to thy seed after thee, the land wherein thou art a stranger, all the land of Canaan, for an everlasting possession; and I will be their God.
\fn{wherein...: Heb. of thy sojournings.}
\vs Gen 17:9 And God said unto Abraham, Thou shalt keep my covenant therefore, thou, and thy seed after thee in their generations.
\vs Gen 17:10 This \bibemph{is} my covenant, which ye shall keep, between me and you and thy seed after thee; Every man child among you shall be circumcised.
\vs Gen 17:11 And ye shall circumcise the flesh of your foreskin; and it shall be a token of the covenant betwixt me and you.
\vs Gen 17:12 And he that is eight days old shall be circumcised among you, every man child in your generations, he that is born in the house, or bought with money of any stranger, which \bibemph{is} not of thy seed.
\fn{he that is eight...: Heb. a son of eight days.}
\vs Gen 17:13 He that is born in thy house, and he that is bought with thy money, must needs be circumcised: and my covenant shall be in your flesh for an everlasting covenant.
\vs Gen 17:14 And the uncircumcised man child whose flesh of his foreskin is not circumcised, that soul shall be cut off from his people; he hath broken my covenant.
\vs Gen 17:15 And God said unto Abraham, As for Sarai thy wife, thou shalt not call her name Sarai, but Sarah \bibemph{shall} her name \bibemph{be}.
\fn{Sarah: that is Princess.}
\vs Gen 17:16 And I will bless her, and give thee a son also of her: yea, I will bless her, and she shall be \bibemph{a mother} of nations; kings of people shall be of her.
\fn{she...: Heb. she shall become nations.}
\vs Gen 17:17 Then Abraham fell upon his face, and laughed, and said in his heart, Shall \bibemph{a child} be born unto him that is an hundred years old? and shall Sarah, that is ninety years old, bear?
\vs Gen 17:18 And Abraham said unto God, O that Ishmael might live before thee!
\vs Gen 17:19 And God said, Sarah thy wife shall bear thee a son indeed; and thou shalt call his name Isaac: and I will establish my covenant with him for an everlasting covenant, \bibemph{and} with his seed after him.
\vs Gen 17:20 And as for Ishmael, I have heard thee: Behold, I have blessed him, and will make him fruitful, and will multiply him exceedingly; twelve princes shall he beget, and I will make him a great nation.
\vs Gen 17:21 But my covenant will I establish with Isaac, which Sarah shall bear unto thee at this set time in the next year.
\vs Gen 17:22 And he left off talking with him, and God went up from Abraham.
\vs Gen 17:23 And Abraham took Ishmael his son, and all that were born in his house, and all that were bought with his money, every male among the men of Abraham's house; and circumcised the flesh of their foreskin in the selfsame day, as God had said unto him.
\vs Gen 17:24 And Abraham \bibemph{was} ninety years old and nine, when he was circumcised in the flesh of his foreskin.
\vs Gen 17:25 And Ishmael his son \bibemph{was} thirteen years old, when he was circumcised in the flesh of his foreskin.
\vs Gen 17:26 In the selfsame day was Abraham circumcised, and Ishmael his son.
\vs Gen 17:27 And all the men of his house, born in the house, and bought with money of the stranger, were circumcised with him.
\vs Gen 18:1 And the \textsc{Lord} appeared unto him in the plains of Mamre: and he sat in the tent door in the heat of the day;
\vs Gen 18:2 And he lift up his eyes and looked, and, lo, three men stood by him: and when he saw \bibemph{them}, he ran to meet them from the tent door, and bowed himself toward the ground,
\vs Gen 18:3 And said, My Lord, if now I have found favour in thy sight, pass not away, I pray thee, from thy servant:
\vs Gen 18:4 Let a little water, I pray you, be fetched, and wash your feet, and rest yourselves under the tree:
\vs Gen 18:5 And I will fetch a morsel of bread, and comfort ye your hearts; after that ye shall pass on: for therefore are ye come to your servant. And they said, So do, as thou hast said.
\fn{comfort: Heb. stay. are...: Heb. you have passed.}
\vs Gen 18:6 And Abraham hastened into the tent unto Sarah, and said, Make ready quickly three measures of fine meal, knead \bibemph{it}, and make cakes upon the hearth.
\fn{Make ready...: Heb. Hasten.}
\vs Gen 18:7 And Abraham ran unto the herd, and fetcht a calf tender and good, and gave \bibemph{it} unto a young man; and he hasted to dress it.
\vs Gen 18:8 And he took butter, and milk, and the calf which he had dressed, and set \bibemph{it} before them; and he stood by them under the tree, and they did eat.
\vs Gen 18:9 And they said unto him, Where \bibemph{is} Sarah thy wife? And he said, Behold, in the tent.
\vs Gen 18:10 And he said, I will certainly return unto thee according to the time of life; and, lo, Sarah thy wife shall have a son. And Sarah heard \bibemph{it} in the tent door, which \bibemph{was} behind him.
\vs Gen 18:11 Now Abraham and Sarah \bibemph{were} old \bibemph{and} well stricken in age; \bibemph{and} it ceased to be with Sarah after the manner of women.
\vs Gen 18:12 Therefore Sarah laughed within herself, saying, After I am waxed old shall I have pleasure, my lord being old also?
\vs Gen 18:13 And the \textsc{Lord} said unto Abraham, Wherefore did Sarah laugh, saying, Shall I of a surety bear a child, which am old?
\vs Gen 18:14 Is any thing too hard for the \textsc{Lord}? At the time appointed I will return unto thee, according to the time of life, and Sarah shall have a son.
\vs Gen 18:15 Then Sarah denied, saying, I laughed not; for she was afraid. And he said, Nay; but thou didst laugh.
\vs Gen 18:16 And the men rose up from thence, and looked toward Sodom: and Abraham went with them to bring them on the way.
\vs Gen 18:17 And the \textsc{Lord} said, Shall I hide from Abraham that thing which I do;
\vs Gen 18:18 Seeing that Abraham shall surely become a great and mighty nation, and all the nations of the earth shall be blessed in him?
\vs Gen 18:19 For I know him, that he will command his children and his household after him, and they shall keep the way of the \textsc{Lord}, to do justice and judgment; that the \textsc{Lord} may bring upon Abraham that which he hath spoken of him.
\vs Gen 18:20 And the \textsc{Lord} said, Because the cry of Sodom and Gomorrah is great, and because their sin is very grievous;
\vs Gen 18:21 I will go down now, and see whether they have done altogether according to the cry of it, which is come unto me; and if not, I will know.
\vs Gen 18:22 And the men turned their faces from thence, and went toward Sodom: but Abraham stood yet before the \textsc{Lord}.
\vs Gen 18:23 And Abraham drew near, and said, Wilt thou also destroy the righteous with the wicked?
\vs Gen 18:24 Peradventure there be fifty righteous within the city: wilt thou also destroy and not spare the place for the fifty righteous that \bibemph{are} therein?
\vs Gen 18:25 That be far from thee to do after this manner, to slay the righteous with the wicked: and that the righteous should be as the wicked, that be far from thee: Shall not the Judge of all the earth do right?
\vs Gen 18:26 And the \textsc{Lord} said, If I find in Sodom fifty righteous within the city, then I will spare all the place for their sakes.
\vs Gen 18:27 And Abraham answered and said, Behold now, I have taken upon me to speak unto the Lord, which \bibemph{am but} dust and ashes:
\vs Gen 18:28 Peradventure there shall lack five of the fifty righteous: wilt thou destroy all the city for \bibemph{lack of} five? And he said, If I find there forty and five, I will not destroy \bibemph{it}.
\vs Gen 18:29 And he spake unto him yet again, and said, Peradventure there shall be forty found there. And he said, I will not do \bibemph{it} for forty's sake.
\vs Gen 18:30 And he said \bibemph{unto him}, Oh let not the Lord be angry, and I will speak: Peradventure there shall thirty be found there. And he said, I will not do \bibemph{it}, if I find thirty there.
\vs Gen 18:31 And he said, Behold now, I have taken upon me to speak unto the Lord: Peradventure there shall be twenty found there. And he said, I will not destroy \bibemph{it} for twenty's sake.
\vs Gen 18:32 And he said, Oh let not the Lord be angry, and I will speak yet but this once: Peradventure ten shall be found there. And he said, I will not destroy \bibemph{it} for ten's sake.
\vs Gen 18:33 And the \textsc{Lord} went his way, as soon as he had left communing with Abraham: and Abraham returned unto his place.
\vs Gen 19:1 And there came two angels to Sodom at even; and Lot sat in the gate of Sodom: and Lot seeing \bibemph{them} rose up to meet them; and he bowed himself with his face toward the ground;
\vs Gen 19:2 And he said, Behold now, my lords, turn in, I pray you, into your servant's house, and tarry all night, and wash your feet, and ye shall rise up early, and go on your ways. And they said, Nay; but we will abide in the street all night.
\vs Gen 19:3 And he pressed upon them greatly; and they turned in unto him, and entered into his house; and he made them a feast, and did bake unleavened bread, and they did eat.
\vs Gen 19:4 But before they lay down, the men of the city, \bibemph{even} the men of Sodom, compassed the house round, both old and young, all the people from every quarter:
\vs Gen 19:5 And they called unto Lot, and said unto him, Where \bibemph{are} the men which came in to thee this night? bring them out unto us, that we may know them.
\vs Gen 19:6 And Lot went out at the door unto them, and shut the door after him,
\vs Gen 19:7 And said, I pray you, brethren, do not so wickedly.
\vs Gen 19:8 Behold now, I have two daughters which have not known man; let me, I pray you, bring them out unto you, and do ye to them as \bibemph{is} good in your eyes: only unto these men do nothing; for therefore came they under the shadow of my roof.
\vs Gen 19:9 And they said, Stand back. And they said \bibemph{again}, This one \bibemph{fellow} came in to sojourn, and he will needs be a judge: now will we deal worse with thee, than with them. And they pressed sore upon the man, \bibemph{even} Lot, and came near to break the door.
\vs Gen 19:10 But the men put forth their hand, and pulled Lot into the house to them, and shut to the door.
\vs Gen 19:11 And they smote the men that \bibemph{were} at the door of the house with blindness, both small and great: so that they wearied themselves to find the door.
\vs Gen 19:12 And the men said unto Lot, Hast thou here any besides? son in law, and thy sons, and thy daughters, and whatsoever thou hast in the city, bring \bibemph{them} out of this place:
\vs Gen 19:13 For we will destroy this place, because the cry of them is waxen great before the face of the \textsc{Lord}; and the \textsc{Lord} hath sent us to destroy it.
\vs Gen 19:14 And Lot went out, and spake unto his sons in law, which married his daughters, and said, Up, get you out of this place; for the \textsc{Lord} will destroy this city. But he seemed as one that mocked unto his sons in law.
\vs Gen 19:15 And when the morning arose, then the angels hastened Lot, saying, Arise, take thy wife, and thy two daughters, which are here; lest thou be consumed in the iniquity of the city.
\fn{are here: Heb. are found. iniquity: or, punishment.}
\vs Gen 19:16 And while he lingered, the men laid hold upon his hand, and upon the hand of his wife, and upon the hand of his two daughters; the \textsc{Lord} being merciful unto him: and they brought him forth, and set him without the city.
\vs Gen 19:17 And it came to pass, when they had brought them forth abroad, that he said, Escape for thy life; look not behind thee, neither stay thou in all the plain; escape to the mountain, lest thou be consumed.
\vs Gen 19:18 And Lot said unto them, Oh, not so, my Lord:
\vs Gen 19:19 Behold now, thy servant hath found grace in thy sight, and thou hast magnified thy mercy, which thou hast shewed unto me in saving my life; and I cannot escape to the mountain, lest some evil take me, and I die:
\vs Gen 19:20 Behold now, this city \bibemph{is} near to flee unto, and it \bibemph{is} a little one: Oh, let me escape thither, (\bibemph{is} it not a little one?) and my soul shall live.
\vs Gen 19:21 And he said unto him, See, I have accepted thee concerning this thing also, that I will not overthrow this city, for the which thou hast spoken.
\fn{thee: Heb. thy face.}
\vs Gen 19:22 Haste thee, escape thither; for I cannot do any thing till thou be come thither. Therefore the name of the city was called Zoar.
\fn{Zoar: that is, Little.}
\vs Gen 19:23 The sun was risen upon the earth when Lot entered into Zoar.
\fn{risen: Heb. gone forth.}
\vs Gen 19:24 Then the \textsc{Lord} rained upon Sodom and upon Gomorrah brimstone and fire from the \textsc{Lord} out of heaven;
\vs Gen 19:25 And he overthrew those cities, and all the plain, and all the inhabitants of the cities, and that which grew upon the ground.
\vs Gen 19:26 But his wife looked back from behind him, and she became a pillar of salt.
\vs Gen 19:27 And Abraham gat up early in the morning to the place where he stood before the \textsc{Lord}:
\vs Gen 19:28 And he looked toward Sodom and Gomorrah, and toward all the land of the plain, and beheld, and, lo, the smoke of the country went up as the smoke of a furnace.
\vs Gen 19:29 And it came to pass, when God destroyed the cities of the plain, that God remembered Abraham, and sent Lot out of the midst of the overthrow, when he overthrew the cities in the which Lot dwelt.
\vs Gen 19:30 And Lot went up out of Zoar, and dwelt in the mountain, and his two daughters with him; for he feared to dwell in Zoar: and he dwelt in a cave, he and his two daughters.
\vs Gen 19:31 And the firstborn said unto the younger, Our father \bibemph{is} old, and \bibemph{there is} not a man in the earth to come in unto us after the manner of all the earth:
\vs Gen 19:32 Come, let us make our father drink wine, and we will lie with him, that we may preserve seed of our father.
\vs Gen 19:33 And they made their father drink wine that night: and the firstborn went in, and lay with her father; and he perceived not when she lay down, nor when she arose.
\vs Gen 19:34 And it came to pass on the morrow, that the firstborn said unto the younger, Behold, I lay yesternight with my father: let us make him drink wine this night also; and go thou in, \bibemph{and} lie with him, that we may preserve seed of our father.
\vs Gen 19:35 And they made their father drink wine that night also: and the younger arose, and lay with him; and he perceived not when she lay down, nor when she arose.
\vs Gen 19:36 Thus were both the daughters of Lot with child by their father.
\vs Gen 19:37 And the firstborn bare a son, and called his name Moab: the same \bibemph{is} the father of the Moabites unto this day.
\vs Gen 19:38 And the younger, she also bare a son, and called his name Benammi: the same \bibemph{is} the father of the children of Ammon unto this day.
\vs Gen 20:1 And Abraham journeyed from thence toward the south country, and dwelled between Kadesh and Shur, and sojourned in Gerar.
\vs Gen 20:2 And Abraham said of Sarah his wife, She \bibemph{is} my sister: and Abimelech king of Gerar sent, and took Sarah.
\vs Gen 20:3 But God came to Abimelech in a dream by night, and said to him, Behold, thou \bibemph{art but} a dead man, for the woman which thou hast taken; for she \bibemph{is} a man's wife.
\fn{a man's...: Heb. married to an husband.}
\vs Gen 20:4 But Abimelech had not come near her: and he said, Lord, wilt thou slay also a righteous nation?
\vs Gen 20:5 Said he not unto me, She \bibemph{is} my sister? and she, even she herself said, He \bibemph{is} my brother: in the integrity of my heart and innocency of my hands have I done this.
\fn{integrity: or, simplicity, or, sincerity.}
\vs Gen 20:6 And God said unto him in a dream, Yea, I know that thou didst this in the integrity of thy heart; for I also withheld thee from sinning against me: therefore suffered I thee not to touch her.
\vs Gen 20:7 Now therefore restore the man \bibemph{his} wife; for he \bibemph{is} a prophet, and he shall pray for thee, and thou shalt live: and if thou restore \bibemph{her} not, know thou that thou shalt surely die, thou, and all that \bibemph{are} thine.
\vs Gen 20:8 Therefore Abimelech rose early in the morning, and called all his servants, and told all these things in their ears: and the men were sore afraid.
\vs Gen 20:9 Then Abimelech called Abraham, and said unto him, What hast thou done unto us? and what have I offended thee, that thou hast brought on me and on my kingdom a great sin? thou hast done deeds unto me that ought not to be done.
\vs Gen 20:10 And Abimelech said unto Abraham, What sawest thou, that thou hast done this thing?
\vs Gen 20:11 And Abraham said, Because I thought, Surely the fear of God \bibemph{is} not in this place; and they will slay me for my wife's sake.
\vs Gen 20:12 And yet indeed \bibemph{she is} my sister; she \bibemph{is} the daughter of my father, but not the daughter of my mother; and she became my wife.
\vs Gen 20:13 And it came to pass, when God caused me to wander from my father's house, that I said unto her, This \bibemph{is} thy kindness which thou shalt shew unto me; at every place whither we shall come, say of me, He \bibemph{is} my brother.
\vs Gen 20:14 And Abimelech took sheep, and oxen, and menservants, and womenservants, and gave \bibemph{them} unto Abraham, and restored him Sarah his wife.
\vs Gen 20:15 And Abimelech said, Behold, my land \bibemph{is} before thee: dwell where it pleaseth thee.
\fn{where...: Heb. as is good in thine eyes.}
\vs Gen 20:16 And unto Sarah he said, Behold, I have given thy brother a thousand \bibemph{pieces} of silver: behold, he \bibemph{is} to thee a covering of the eyes, unto all that \bibemph{are} with thee, and with all \bibemph{other}: thus she was reproved.
\vs Gen 20:17 So Abraham prayed unto God: and God healed Abimelech, and his wife, and his maidservants; and they bare \bibemph{children}.
\vs Gen 20:18 For the \textsc{Lord} had fast closed up all the wombs of the house of Abimelech, because of Sarah Abraham's wife.
\vs Gen 21:1 And the \textsc{Lord} visited Sarah as he had said, and the \textsc{Lord} did unto Sarah as he had spoken.
\vs Gen 21:2 For Sarah conceived, and bare Abraham a son in his old age, at the set time of which God had spoken to him.
\vs Gen 21:3 And Abraham called the name of his son that was born unto him, whom Sarah bare to him, Isaac.
\vs Gen 21:4 And Abraham circumcised his son Isaac being eight days old, as God had commanded him.
\vs Gen 21:5 And Abraham was an hundred years old, when his son Isaac was born unto him.
\vs Gen 21:6 And Sarah said, God hath made me to laugh, \bibemph{so that} all that hear will laugh with me.
\vs Gen 21:7 And she said, Who would have said unto Abraham, that Sarah should have given children suck? for I have born \bibemph{him} a son in his old age.
\vs Gen 21:8 And the child grew, and was weaned: and Abraham made a great feast the \bibemph{same} day that Isaac was weaned.
\vs Gen 21:9 And Sarah saw the son of Hagar the Egyptian, which she had born unto Abraham, mocking.
\vs Gen 21:10 Wherefore she said unto Abraham, Cast out this bondwoman and her son: for the son of this bondwoman shall not be heir with my son, \bibemph{even} with Isaac.
\vs Gen 21:11 And the thing was very grievous in Abraham's sight because of his son.
\vs Gen 21:12 And God said unto Abraham, Let it not be grievous in thy sight because of the lad, and because of thy bondwoman; in all that Sarah hath said unto thee, hearken unto her voice; for in Isaac shall thy seed be called.
\vs Gen 21:13 And also of the son of the bondwoman will I make a nation, because he \bibemph{is} thy seed.
\vs Gen 21:14 And Abraham rose up early in the morning, and took bread, and a bottle of water, and gave \bibemph{it} unto Hagar, putting \bibemph{it} on her shoulder, and the child, and sent her away: and she departed, and wandered in the wilderness of Beersheba.
\vs Gen 21:15 And the water was spent in the bottle, and she cast the child under one of the shrubs.
\vs Gen 21:16 And she went, and sat her down over against \bibemph{him} a good way off, as it were a bowshot: for she said, Let me not see the death of the child. And she sat over against \bibemph{him}, and lift up her voice, and wept.
\vs Gen 21:17 And God heard the voice of the lad; and the angel of God called to Hagar out of heaven, and said unto her, What aileth thee, Hagar? fear not; for God hath heard the voice of the lad where he \bibemph{is}.
\vs Gen 21:18 Arise, lift up the lad, and hold him in thine hand; for I will make him a great nation.
\vs Gen 21:19 And God opened her eyes, and she saw a well of water; and she went, and filled the bottle with water, and gave the lad drink.
\vs Gen 21:20 And God was with the lad; and he grew, and dwelt in the wilderness, and became an archer.
\vs Gen 21:21 And he dwelt in the wilderness of Paran: and his mother took him a wife out of the land of Egypt.
\vs Gen 21:22 And it came to pass at that time, that Abimelech and Phichol the chief captain of his host spake unto Abraham, saying, God \bibemph{is} with thee in all that thou doest:
\vs Gen 21:23 Now therefore swear unto me here by God that thou wilt not deal falsely with me, nor with my son, nor with my son's son: \bibemph{but} according to the kindness that I have done unto thee, thou shalt do unto me, and to the land wherein thou hast sojourned.
\fn{that thou...: Heb. if thou shalt lie unto me.}
\vs Gen 21:24 And Abraham said, I will swear.
\vs Gen 21:25 And Abraham reproved Abimelech because of a well of water, which Abimelech's servants had violently taken away.
\vs Gen 21:26 And Abimelech said, I wot not who hath done this thing: neither didst thou tell me, neither yet heard I \bibemph{of it}, but to day.
\vs Gen 21:27 And Abraham took sheep and oxen, and gave them unto Abimelech; and both of them made a covenant.
\vs Gen 21:28 And Abraham set seven ewe lambs of the flock by themselves.
\vs Gen 21:29 And Abimelech said unto Abraham, What \bibemph{mean} these seven ewe lambs which thou hast set by themselves?
\vs Gen 21:30 And he said, For \bibemph{these} seven ewe lambs shalt thou take of my hand, that they may be a witness unto me, that I have digged this well.
\vs Gen 21:31 Wherefore he called that place Beersheba; because there they sware both of them.
\fn{Beersheba: that is, The well of the oath.}
\vs Gen 21:32 Thus they made a covenant at Beersheba: then Abimelech rose up, and Phichol the chief captain of his host, and they returned into the land of the Philistines.
\vs Gen 21:33 And \bibemph{Abraham} planted a grove in Beersheba, and called there on the name of the \textsc{Lord}, the everlasting God.
\fn{grove: or, tree.}
\vs Gen 21:34 And Abraham sojourned in the Philistines' land many days.
\vs Gen 22:1 And it came to pass after these things, that God did tempt Abraham, and said unto him, Abraham: and he said, Behold, \bibemph{here} I \bibemph{am}.
\fn{Behold...: Heb. Behold me.}
\vs Gen 22:2 And he said, Take now thy son, thine only \bibemph{son} Isaac, whom thou lovest, and get thee into the land of Moriah; and offer him there for a burnt offering upon one of the mountains which I will tell thee of.
\vs Gen 22:3 And Abraham rose up early in the morning, and saddled his ass, and took two of his young men with him, and Isaac his son, and clave the wood for the burnt offering, and rose up, and went unto the place of which God had told him.
\vs Gen 22:4 Then on the third day Abraham lifted up his eyes, and saw the place afar off.
\vs Gen 22:5 And Abraham said unto his young men, Abide ye here with the ass; and I and the lad will go yonder and worship, and come again to you.
\vs Gen 22:6 And Abraham took the wood of the burnt offering, and laid \bibemph{it} upon Isaac his son; and he took the fire in his hand, and a knife; and they went both of them together.
\vs Gen 22:7 And Isaac spake unto Abraham his father, and said, My father: and he said, Here \bibemph{am} I, my son. And he said, Behold the fire and the wood: but where \bibemph{is} the lamb for a burnt offering?
\fn{lamb: or, kid.}
\vs Gen 22:8 And Abraham said, My son, God will provide himself a lamb for a burnt offering: so they went both of them together.
\vs Gen 22:9 And they came to the place which God had told him of; and Abraham built an altar there, and laid the wood in order, and bound Isaac his son, and laid him on the altar upon the wood.
\vs Gen 22:10 And Abraham stretched forth his hand, and took the knife to slay his son.
\vs Gen 22:11 And the angel of the \textsc{Lord} called unto him out of heaven, and said, Abraham, Abraham: and he said, Here \bibemph{am} I.
\vs Gen 22:12 And he said, Lay not thine hand upon the lad, neither do thou any thing unto him: for now I know that thou fearest God, seeing thou hast not withheld thy son, thine only \bibemph{son} from me.
\vs Gen 22:13 And Abraham lifted up his eyes, and looked, and behold behind \bibemph{him} a ram caught in a thicket by his horns: and Abraham went and took the ram, and offered him up for a burnt offering in the stead of his son.
\vs Gen 22:14 And Abraham called the name of that place Jehovahjireh: as it is said \bibemph{to} this day, In the mount of the \textsc{Lord} it shall be seen.
\fn{Jehovahjireh: that is, The Lord will see, or, provide.}
\vs Gen 22:15 And the angel of the \textsc{Lord} called unto Abraham out of heaven the second time,
\vs Gen 22:16 And said, By myself have I sworn, saith the \textsc{Lord}, for because thou hast done this thing, and hast not withheld thy son, thine only \bibemph{son}:
\vs Gen 22:17 That in blessing I will bless thee, and in multiplying I will multiply thy seed as the stars of the heaven, and as the sand which \bibemph{is} upon the sea shore; and thy seed shall possess the gate of his enemies;
\fn{shore: Heb. lip.}
\vs Gen 22:18 And in thy seed shall all the nations of the earth be blessed; because thou hast obeyed my voice.
\vs Gen 22:19 So Abraham returned unto his young men, and they rose up and went together to Beersheba; and Abraham dwelt at Beersheba.
\vs Gen 22:20 And it came to pass after these things, that it was told Abraham, saying, Behold, Milcah, she hath also born children unto thy brother Nahor;
\vs Gen 22:21 Huz his firstborn, and Buz his brother, and Kemuel the father of Aram,
\vs Gen 22:22 And Chesed, and Hazo, and Pildash, and Jidlaph, and Bethuel.
\vs Gen 22:23 And Bethuel begat Rebekah: these eight Milcah did bear to Nahor, Abraham's brother.
\fn{Rebekah: Gr. Rebecca.}
\vs Gen 22:24 And his concubine, whose name \bibemph{was} Reumah, she bare also Tebah, and Gaham, and Thahash, and Maachah.
\vs Gen 23:1 And Sarah was an hundred and seven and twenty years old: \bibemph{these were} the years of the life of Sarah.
\vs Gen 23:2 And Sarah died in Kirjatharba; the same \bibemph{is} Hebron in the land of Canaan: and Abraham came to mourn for Sarah, and to weep for her.
\vs Gen 23:3 And Abraham stood up from before his dead, and spake unto the sons of Heth, saying,
\vs Gen 23:4 I \bibemph{am} a stranger and a sojourner with you: give me a possession of a buryingplace with you, that I may bury my dead out of my sight.
\vs Gen 23:5 And the children of Heth answered Abraham, saying unto him,
\vs Gen 23:6 Hear us, my lord: thou \bibemph{art} a mighty prince among us: in the choice of our sepulchres bury thy dead; none of us shall withhold from thee his sepulchre, but that thou mayest bury thy dead.
\fn{a mighty...: Heb. a prince of God.}
\vs Gen 23:7 And Abraham stood up, and bowed himself to the people of the land, \bibemph{even} to the children of Heth.
\vs Gen 23:8 And he communed with them, saying, If it be your mind that I should bury my dead out of my sight; hear me, and intreat for me to Ephron the son of Zohar,
\vs Gen 23:9 That he may give me the cave of Machpelah, which he hath, which \bibemph{is} in the end of his field; for as much money as it is worth he shall give it me for a possession of a buryingplace amongst you.
\fn{as much...: Heb. full money.}
\vs Gen 23:10 And Ephron dwelt among the children of Heth: and Ephron the Hittite answered Abraham in the audience of the children of Heth, \bibemph{even} of all that went in at the gate of his city, saying,
\fn{audience: Heb. ears.}
\vs Gen 23:11 Nay, my lord, hear me: the field give I thee, and the cave that \bibemph{is} therein, I give it thee; in the presence of the sons of my people give I it thee: bury thy dead.
\vs Gen 23:12 And Abraham bowed down himself before the people of the land.
\vs Gen 23:13 And he spake unto Ephron in the audience of the people of the land, saying, But if thou \bibemph{wilt give it}, I pray thee, hear me: I will give thee money for the field; take \bibemph{it} of me, and I will bury my dead there.
\vs Gen 23:14 And Ephron answered Abraham, saying unto him,
\vs Gen 23:15 My lord, hearken unto me: the land \bibemph{is worth} four hundred shekels of silver; what \bibemph{is} that betwixt me and thee? bury therefore thy dead.
\vs Gen 23:16 And Abraham hearkened unto Ephron; and Abraham weighed to Ephron the silver, which he had named in the audience of the sons of Heth, four hundred shekels of silver, current \bibemph{money} with the merchant.
\vs Gen 23:17 And the field of Ephron, which \bibemph{was} in Machpelah, which \bibemph{was} before Mamre, the field, and the cave which \bibemph{was} therein, and all the trees that \bibemph{were} in the field, that \bibemph{were} in all the borders round about, were made sure
\vs Gen 23:18 Unto Abraham for a possession in the presence of the children of Heth, before all that went in at the gate of his city.
\vs Gen 23:19 And after this, Abraham buried Sarah his wife in the cave of the field of Machpelah before Mamre: the same \bibemph{is} Hebron in the land of Canaan.
\vs Gen 23:20 And the field, and the cave that \bibemph{is} therein, were made sure unto Abraham for a possession of a buryingplace by the sons of Heth.
\vs Gen 24:1 And Abraham was old, \bibemph{and} well stricken in age: and the \textsc{Lord} had blessed Abraham in all things.
\fn{well...: Heb. gone into days.}
\vs Gen 24:2 And Abraham said unto his eldest servant of his house, that ruled over all that he had, Put, I pray thee, thy hand under my thigh:
\vs Gen 24:3 And I will make thee swear by the \textsc{Lord}, the God of heaven, and the God of the earth, that thou shalt not take a wife unto my son of the daughters of the Canaanites, among whom I dwell:
\vs Gen 24:4 But thou shalt go unto my country, and to my kindred, and take a wife unto my son Isaac.
\vs Gen 24:5 And the servant said unto him, Peradventure the woman will not be willing to follow me unto this land: must I needs bring thy son again unto the land from whence thou camest?
\vs Gen 24:6 And Abraham said unto him, Beware thou that thou bring not my son thither again.
\vs Gen 24:7 The \textsc{Lord} God of heaven, which took me from my father's house, and from the land of my kindred, and which spake unto me, and that sware unto me, saying, Unto thy seed will I give this land; he shall send his angel before thee, and thou shalt take a wife unto my son from thence.
\vs Gen 24:8 And if the woman will not be willing to follow thee, then thou shalt be clear from this my oath: only bring not my son thither again.
\vs Gen 24:9 And the servant put his hand under the thigh of Abraham his master, and sware to him concerning that matter.
\vs Gen 24:10 And the servant took ten camels of the camels of his master, and departed; for all the goods of his master \bibemph{were} in his hand: and he arose, and went to Mesopotamia, unto the city of Nahor.
\fn{for: or, and.}
\vs Gen 24:11 And he made his camels to kneel down without the city by a well of water at the time of the evening, \bibemph{even} the time that women go out to draw \bibemph{water}.
\fn{that...: Heb. that women who draw water go forth.}
\vs Gen 24:12 And he said, O \textsc{Lord} God of my master Abraham, I pray thee, send me good speed this day, and shew kindness unto my master Abraham.
\vs Gen 24:13 Behold, I stand \bibemph{here} by the well of water; and the daughters of the men of the city come out to draw water:
\vs Gen 24:14 And let it come to pass, that the damsel to whom I shall say, Let down thy pitcher, I pray thee, that I may drink; and she shall say, Drink, and I will give thy camels drink also: \bibemph{let the same be} she \bibemph{that} thou hast appointed for thy servant Isaac; and thereby shall I know that thou hast shewed kindness unto my master.
\vs Gen 24:15 And it came to pass, before he had done speaking, that, behold, Rebekah came out, who was born to Bethuel, son of Milcah, the wife of Nahor, Abraham's brother, with her pitcher upon her shoulder.
\vs Gen 24:16 And the damsel \bibemph{was} very fair to look upon, a virgin, neither had any man known her: and she went down to the well, and filled her pitcher, and came up.
\fn{very...: Heb. good of countenance.}
\vs Gen 24:17 And the servant ran to meet her, and said, Let me, I pray thee, drink a little water of thy pitcher.
\vs Gen 24:18 And she said, Drink, my lord: and she hasted, and let down her pitcher upon her hand, and gave him drink.
\vs Gen 24:19 And when she had done giving him drink, she said, I will draw \bibemph{water} for thy camels also, until they have done drinking.
\vs Gen 24:20 And she hasted, and emptied her pitcher into the trough, and ran again unto the well to draw \bibemph{water}, and drew for all his camels.
\vs Gen 24:21 And the man wondering at her held his peace, to wit whether the \textsc{Lord} had made his journey prosperous or not.
\vs Gen 24:22 And it came to pass, as the camels had done drinking, that the man took a golden earring of half a shekel weight, and two bracelets for her hands of ten \bibemph{shekels} weight of gold;
\fn{earring: or, jewel for the forehead.}
\vs Gen 24:23 And said, Whose daughter \bibemph{art} thou? tell me, I pray thee: is there room \bibemph{in} thy father's house for us to lodge in?
\vs Gen 24:24 And she said unto him, I \bibemph{am} the daughter of Bethuel the son of Milcah, which she bare unto Nahor.
\vs Gen 24:25 She said moreover unto him, We have both straw and provender enough, and room to lodge in.
\vs Gen 24:26 And the man bowed down his head, and worshipped the \textsc{Lord}.
\vs Gen 24:27 And he said, Blessed \bibemph{be} the \textsc{Lord} God of my master Abraham, who hath not left destitute my master of his mercy and his truth: I \bibemph{being} in the way, the \textsc{Lord} led me to the house of my master's brethren.
\vs Gen 24:28 And the damsel ran, and told \bibemph{them of} her mother's house these things.
\vs Gen 24:29 And Rebekah had a brother, and his name \bibemph{was} Laban: and Laban ran out unto the man, unto the well.
\vs Gen 24:30 And it came to pass, when he saw the earring and bracelets upon his sister's hands, and when he heard the words of Rebekah his sister, saying, Thus spake the man unto me; that he came unto the man; and, behold, he stood by the camels at the well.
\vs Gen 24:31 And he said, Come in, thou blessed of the \textsc{Lord}; wherefore standest thou without? for I have prepared the house, and room for the camels.
\vs Gen 24:32 And the man came into the house: and he ungirded his camels, and gave straw and provender for the camels, and water to wash his feet, and the men's feet that \bibemph{were} with him.
\vs Gen 24:33 And there was set \bibemph{meat} before him to eat: but he said, I will not eat, until I have told mine errand. And he said, Speak on.
\vs Gen 24:34 And he said, I \bibemph{am} Abraham's servant.
\vs Gen 24:35 And the \textsc{Lord} hath blessed my master greatly; and he is become great: and he hath given him flocks, and herds, and silver, and gold, and menservants, and maidservants, and camels, and asses.
\vs Gen 24:36 And Sarah my master's wife bare a son to my master when she was old: and unto him hath he given all that he hath.
\vs Gen 24:37 And my master made me swear, saying, Thou shalt not take a wife to my son of the daughters of the Canaanites, in whose land I dwell:
\vs Gen 24:38 But thou shalt go unto my father's house, and to my kindred, and take a wife unto my son.
\vs Gen 24:39 And I said unto my master, Peradventure the woman will not follow me.
\vs Gen 24:40 And he said unto me, The \textsc{Lord}, before whom I walk, will send his angel with thee, and prosper thy way; and thou shalt take a wife for my son of my kindred, and of my father's house:
\vs Gen 24:41 Then shalt thou be clear from \bibemph{this} my oath, when thou comest to my kindred; and if they give not thee \bibemph{one}, thou shalt be clear from my oath.
\vs Gen 24:42 And I came this day unto the well, and said, O \textsc{Lord} God of my master Abraham, if now thou do prosper my way which I go:
\vs Gen 24:43 Behold, I stand by the well of water; and it shall come to pass, that when the virgin cometh forth to draw \bibemph{water}, and I say to her, Give me, I pray thee, a little water of thy pitcher to drink;
\vs Gen 24:44 And she say to me, Both drink thou, and I will also draw for thy camels: \bibemph{let} the same \bibemph{be} the woman whom the \textsc{Lord} hath appointed out for my master's son.
\vs Gen 24:45 And before I had done speaking in mine heart, behold, Rebekah came forth with her pitcher on her shoulder; and she went down unto the well, and drew \bibemph{water}: and I said unto her, Let me drink, I pray thee.
\vs Gen 24:46 And she made haste, and let down her pitcher from her \bibemph{shoulder}, and said, Drink, and I will give thy camels drink also: so I drank, and she made the camels drink also.
\vs Gen 24:47 And I asked her, and said, Whose daughter \bibemph{art} thou? And she said, The daughter of Bethuel, Nahor's son, whom Milcah bare unto him: and I put the earring upon her face, and the bracelets upon her hands.
\vs Gen 24:48 And I bowed down my head, and worshipped the \textsc{Lord}, and blessed the \textsc{Lord} God of my master Abraham, which had led me in the right way to take my master's brother's daughter unto his son.
\vs Gen 24:49 And now if ye will deal kindly and truly with my master, tell me: and if not, tell me; that I may turn to the right hand, or to the left.
\vs Gen 24:50 Then Laban and Bethuel answered and said, The thing proceedeth from the \textsc{Lord}: we cannot speak unto thee bad or good.
\vs Gen 24:51 Behold, Rebekah \bibemph{is} before thee, take \bibemph{her}, and go, and let her be thy master's son's wife, as the \textsc{Lord} hath spoken.
\vs Gen 24:52 And it came to pass, that, when Abraham's servant heard their words, he worshipped the \textsc{Lord}, \bibemph{bowing himself} to the earth.
\vs Gen 24:53 And the servant brought forth jewels of silver, and jewels of gold, and raiment, and gave \bibemph{them} to Rebekah: he gave also to her brother and to her mother precious things.
\fn{jewels: Heb. vessels.}
\vs Gen 24:54 And they did eat and drink, he and the men that \bibemph{were} with him, and tarried all night; and they rose up in the morning, and he said, Send me away unto my master.
\vs Gen 24:55 And her brother and her mother said, Let the damsel abide with us \bibemph{a few} days, at the least ten; after that she shall go.
\fn{a few...: or, a full year, or ten months.}
\vs Gen 24:56 And he said unto them, Hinder me not, seeing the \textsc{Lord} hath prospered my way; send me away that I may go to my master.
\vs Gen 24:57 And they said, We will call the damsel, and enquire at her mouth.
\vs Gen 24:58 And they called Rebekah, and said unto her, Wilt thou go with this man? And she said, I will go.
\vs Gen 24:59 And they sent away Rebekah their sister, and her nurse, and Abraham's servant, and his men.
\vs Gen 24:60 And they blessed Rebekah, and said unto her, Thou \bibemph{art} our sister, be thou \bibemph{the mother} of thousands of millions, and let thy seed possess the gate of those which hate them.
\vs Gen 24:61 And Rebekah arose, and her damsels, and they rode upon the camels, and followed the man: and the servant took Rebekah, and went his way.
\vs Gen 24:62 And Isaac came from the way of the well Lahairoi; for he dwelt in the south country.
\vs Gen 24:63 And Isaac went out to meditate in the field at the eventide: and he lifted up his eyes, and saw, and, behold, the camels \bibemph{were} coming.
\fn{to meditate: or, to pray.}
\vs Gen 24:64 And Rebekah lifted up her eyes, and when she saw Isaac, she lighted off the camel.
\vs Gen 24:65 For she \bibemph{had} said unto the servant, What man \bibemph{is} this that walketh in the field to meet us? And the servant \bibemph{had} said, It \bibemph{is} my master: therefore she took a vail, and covered herself.
\vs Gen 24:66 And the servant told Isaac all things that he had done.
\vs Gen 24:67 And Isaac brought her into his mother Sarah's tent, and took Rebekah, and she became his wife; and he loved her: and Isaac was comforted after his mother's \bibemph{death}.
\vs Gen 25:1 Then again Abraham took a wife, and her name \bibemph{was} Keturah.
\vs Gen 25:2 And she bare him Zimran, and Jokshan, and Medan, and Midian, and Ishbak, and Shuah.
\vs Gen 25:3 And Jokshan begat Sheba, and Dedan. And the sons of Dedan were Asshurim, and Letushim, and Leummim.
\vs Gen 25:4 And the sons of Midian; Ephah, and Epher, and Hanoch, and Abida, and Eldaah. All these \bibemph{were} the children of Keturah.
\vs Gen 25:5 And Abraham gave all that he had unto Isaac.
\vs Gen 25:6 But unto the sons of the concubines, which Abraham had, Abraham gave gifts, and sent them away from Isaac his son, while he yet lived, eastward, unto the east country.
\vs Gen 25:7 And these \bibemph{are} the days of the years of Abraham's life which he lived, an hundred threescore and fifteen years.
\vs Gen 25:8 Then Abraham gave up the ghost, and died in a good old age, an old man, and full \bibemph{of years}; and was gathered to his people.
\vs Gen 25:9 And his sons Isaac and Ishmael buried him in the cave of Machpelah, in the field of Ephron the son of Zohar the Hittite, which \bibemph{is} before Mamre;
\vs Gen 25:10 The field which Abraham purchased of the sons of Heth: there was Abraham buried, and Sarah his wife.
\vs Gen 25:11 And it came to pass after the death of Abraham, that God blessed his son Isaac; and Isaac dwelt by the well Lahairoi.
\vs Gen 25:12 Now these \bibemph{are} the generations of Ishmael, Abraham's son, whom Hagar the Egyptian, Sarah's handmaid, bare unto Abraham:
\vs Gen 25:13 And these \bibemph{are} the names of the sons of Ishmael, by their names, according to their generations: the firstborn of Ishmael, Nebajoth; and Kedar, and Adbeel, and Mibsam,
\vs Gen 25:14 And Mishma, and Dumah, and Massa,
\vs Gen 25:15 Hadar, and Tema, Jetur, Naphish, and Kedemah:
\fn{Hadar: or, Hadad.}
\vs Gen 25:16 These \bibemph{are} the sons of Ishmael, and these \bibemph{are} their names, by their towns, and by their castles; twelve princes according to their nations.
\vs Gen 25:17 And these \bibemph{are} the years of the life of Ishmael, an hundred and thirty and seven years: and he gave up the ghost and died; and was gathered unto his people.
\vs Gen 25:18 And they dwelt from Havilah unto Shur, that \bibemph{is} before Egypt, as thou goest toward Assyria: \bibemph{and} he died in the presence of all his brethren.
\fn{died: Heb. fell.}
\vs Gen 25:19 And these \bibemph{are} the generations of Isaac, Abraham's son: Abraham begat Isaac:
\vs Gen 25:20 And Isaac was forty years old when he took Rebekah to wife, the daughter of Bethuel the Syrian of Padanaram, the sister to Laban the Syrian.
\vs Gen 25:21 And Isaac intreated the \textsc{Lord} for his wife, because she \bibemph{was} barren: and the \textsc{Lord} was intreated of him, and Rebekah his wife conceived.
\vs Gen 25:22 And the children struggled together within her; and she said, If \bibemph{it be} so, why \bibemph{am} I thus? And she went to enquire of the \textsc{Lord}.
\vs Gen 25:23 And the \textsc{Lord} said unto her, Two nations \bibemph{are} in thy womb, and two manner of people shall be separated from thy bowels; and \bibemph{the one} people shall be stronger than \bibemph{the other} people; and the elder shall serve the younger.
\vs Gen 25:24 And when her days to be delivered were fulfilled, behold, \bibemph{there were} twins in her womb.
\vs Gen 25:25 And the first came out red, all over like an hairy garment; and they called his name Esau.
\vs Gen 25:26 And after that came his brother out, and his hand took hold on Esau's heel; and his name was called Jacob: and Isaac \bibemph{was} threescore years old when she bare them.
\vs Gen 25:27 And the boys grew: and Esau was a cunning hunter, a man of the field; and Jacob \bibemph{was} a plain man, dwelling in tents.
\vs Gen 25:28 And Isaac loved Esau, because he did eat of \bibemph{his} venison: but Rebekah loved Jacob.
\fn{he...: Heb. venison was in his mouth.}
\vs Gen 25:29 And Jacob sod pottage: and Esau came from the field, and he \bibemph{was} faint:
\vs Gen 25:30 And Esau said to Jacob, Feed me, I pray thee, with that same red \bibemph{pottage}; for I \bibemph{am} faint: therefore was his name called Edom.
\fn{with...: Heb. with that red, with that red pottage. Edom: that is Red.}
\vs Gen 25:31 And Jacob said, Sell me this day thy birthright.
\vs Gen 25:32 And Esau said, Behold, I \bibemph{am} at the point to die: and what profit shall this birthright do to me?
\fn{at...: Heb. going to die.}
\vs Gen 25:33 And Jacob said, Swear to me this day; and he sware unto him: and he sold his birthright unto Jacob.
\vs Gen 25:34 Then Jacob gave Esau bread and pottage of lentiles; and he did eat and drink, and rose up, and went his way: thus Esau despised \bibemph{his} birthright.
\vs Gen 26:1 And there was a famine in the land, beside the first famine that was in the days of Abraham. And Isaac went unto Abimelech king of the Philistines unto Gerar.
\vs Gen 26:2 And the \textsc{Lord} appeared unto him, and said, Go not down into Egypt; dwell in the land which I shall tell thee of:
\vs Gen 26:3 Sojourn in this land, and I will be with thee, and will bless thee; for unto thee, and unto thy seed, I will give all these countries, and I will perform the oath which I sware unto Abraham thy father;
\vs Gen 26:4 And I will make thy seed to multiply as the stars of heaven, and will give unto thy seed all these countries; and in thy seed shall all the nations of the earth be blessed;
\vs Gen 26:5 Because that Abraham obeyed my voice, and kept my charge, my commandments, my statutes, and my laws.
\vs Gen 26:6 And Isaac dwelt in Gerar:
\vs Gen 26:7 And the men of the place asked \bibemph{him} of his wife; and he said, She \bibemph{is} my sister: for he feared to say, \bibemph{She is} my wife; lest, \bibemph{said he}, the men of the place should kill me for Rebekah; because she \bibemph{was} fair to look upon.
\vs Gen 26:8 And it came to pass, when he had been there a long time, that Abimelech king of the Philistines looked out at a window, and saw, and, behold, Isaac \bibemph{was} sporting with Rebekah his wife.
\vs Gen 26:9 And Abimelech called Isaac, and said, Behold, of a surety she \bibemph{is} thy wife: and how saidst thou, She \bibemph{is} my sister? And Isaac said unto him, Because I said, Lest I die for her.
\vs Gen 26:10 And Abimelech said, What \bibemph{is} this thou hast done unto us? one of the people might lightly have lien with thy wife, and thou shouldest have brought guiltiness upon us.
\vs Gen 26:11 And Abimelech charged all \bibemph{his} people, saying, He that toucheth this man or his wife shall surely be put to death.
\vs Gen 26:12 Then Isaac sowed in that land, and received in the same year an hundredfold: and the \textsc{Lord} blessed him.
\fn{received: Heb. found.}
\vs Gen 26:13 And the man waxed great, and went forward, and grew until he became very great:
\fn{went...: Heb. went going.}
\vs Gen 26:14 For he had possession of flocks, and possession of herds, and great store of servants: and the Philistines envied him.
\fn{servants: or, husbandry.}
\vs Gen 26:15 For all the wells which his father's servants had digged in the days of Abraham his father, the Philistines had stopped them, and filled them with earth.
\vs Gen 26:16 And Abimelech said unto Isaac, Go from us; for thou art much mightier than we.
\vs Gen 26:17 And Isaac departed thence, and pitched his tent in the valley of Gerar, and dwelt there.
\vs Gen 26:18 And Isaac digged again the wells of water, which they had digged in the days of Abraham his father; for the Philistines had stopped them after the death of Abraham: and he called their names after the names by which his father had called them.
\vs Gen 26:19 And Isaac's servants digged in the valley, and found there a well of springing water.
\fn{springing: Heb. living.}
\vs Gen 26:20 And the herdmen of Gerar did strive with Isaac's herdmen, saying, The water \bibemph{is} ours: and he called the name of the well Esek; because they strove with him.
\fn{Esek: that is, Contention.}
\vs Gen 26:21 And they digged another well, and strove for that also: and he called the name of it Sitnah.
\fn{Sitnah: that is, Hatred.}
\vs Gen 26:22 And he removed from thence, and digged another well; and for that they strove not: and he called the name of it Rehoboth; and he said, For now the \textsc{Lord} hath made room for us, and we shall be fruitful in the land.
\fn{Rehoboth: that is Room.}
\vs Gen 26:23 And he went up from thence to Beersheba.
\vs Gen 26:24 And the \textsc{Lord} appeared unto him the same night, and said, I \bibemph{am} the God of Abraham thy father: fear not, for I \bibemph{am} with thee, and will bless thee, and multiply thy seed for my servant Abraham's sake.
\vs Gen 26:25 And he builded an altar there, and called upon the name of the \textsc{Lord}, and pitched his tent there: and there Isaac's servants digged a well.
\vs Gen 26:26 Then Abimelech went to him from Gerar, and Ahuzzath one of his friends, and Phichol the chief captain of his army.
\vs Gen 26:27 And Isaac said unto them, Wherefore come ye to me, seeing ye hate me, and have sent me away from you?
\vs Gen 26:28 And they said, We saw certainly that the \textsc{Lord} was with thee: and we said, Let there be now an oath betwixt us, \bibemph{even} betwixt us and thee, and let us make a covenant with thee;
\fn{We saw...: Heb. Seeing we saw.}
\vs Gen 26:29 That thou wilt do us no hurt, as we have not touched thee, and as we have done unto thee nothing but good, and have sent thee away in peace: thou \bibemph{art} now the blessed of the \textsc{Lord}.
\fn{That...: Heb. If thou shalt.}
\vs Gen 26:30 And he made them a feast, and they did eat and drink.
\vs Gen 26:31 And they rose up betimes in the morning, and sware one to another: and Isaac sent them away, and they departed from him in peace.
\vs Gen 26:32 And it came to pass the same day, that Isaac's servants came, and told him concerning the well which they had digged, and said unto him, We have found water.
\vs Gen 26:33 And he called it Shebah: therefore the name of the city \bibemph{is} Beersheba unto this day.
\fn{Shebah: That is, an oath. Beersheba: that is, the well of the oath.}
\vs Gen 26:34 And Esau was forty years old when he took to wife Judith the daughter of Beeri the Hittite, and Bashemath the daughter of Elon the Hittite:
\vs Gen 26:35 Which were a grief of mind unto Isaac and to Rebekah.
\fn{a grief...: Heb. bitterness of spirit.}
\vs Gen 27:1 And it came to pass, that when Isaac was old, and his eyes were dim, so that he could not see, he called Esau his eldest son, and said unto him, My son: and he said unto him, Behold, \bibemph{here am} I.
\vs Gen 27:2 And he said, Behold now, I am old, I know not the day of my death:
\vs Gen 27:3 Now therefore take, I pray thee, thy weapons, thy quiver and thy bow, and go out to the field, and take me \bibemph{some} venison;
\fn{take: Heb. hunt.}
\vs Gen 27:4 And make me savoury meat, such as I love, and bring \bibemph{it} to me, that I may eat; that my soul may bless thee before I die.
\vs Gen 27:5 And Rebekah heard when Isaac spake to Esau his son. And Esau went to the field to hunt \bibemph{for} venison, \bibemph{and} to bring \bibemph{it}.
\vs Gen 27:6 And Rebekah spake unto Jacob her son, saying, Behold, I heard thy father speak unto Esau thy brother, saying,
\vs Gen 27:7 Bring me venison, and make me savoury meat, that I may eat, and bless thee before the \textsc{Lord} before my death.
\vs Gen 27:8 Now therefore, my son, obey my voice according to that which I command thee.
\vs Gen 27:9 Go now to the flock, and fetch me from thence two good kids of the goats; and I will make them savoury meat for thy father, such as he loveth:
\vs Gen 27:10 And thou shalt bring \bibemph{it} to thy father, that he may eat, and that he may bless thee before his death.
\vs Gen 27:11 And Jacob said to Rebekah his mother, Behold, Esau my brother \bibemph{is} a hairy man, and I \bibemph{am} a smooth man:
\vs Gen 27:12 My father peradventure will feel me, and I shall seem to him as a deceiver; and I shall bring a curse upon me, and not a blessing.
\vs Gen 27:13 And his mother said unto him, Upon me \bibemph{be} thy curse, my son: only obey my voice, and go fetch me \bibemph{them}.
\vs Gen 27:14 And he went, and fetched, and brought \bibemph{them} to his mother: and his mother made savoury meat, such as his father loved.
\vs Gen 27:15 And Rebekah took goodly raiment of her eldest son Esau, which \bibemph{were} with her in the house, and put them upon Jacob her younger son:
\fn{goodly: Heb. desirable.}
\vs Gen 27:16 And she put the skins of the kids of the goats upon his hands, and upon the smooth of his neck:
\vs Gen 27:17 And she gave the savoury meat and the bread, which she had prepared, into the hand of her son Jacob.
\vs Gen 27:18 And he came unto his father, and said, My father: and he said, Here \bibemph{am} I; who \bibemph{art} thou, my son?
\vs Gen 27:19 And Jacob said unto his father, I \bibemph{am} Esau thy firstborn; I have done according as thou badest me: arise, I pray thee, sit and eat of my venison, that thy soul may bless me.
\vs Gen 27:20 And Isaac said unto his son, How \bibemph{is it} that thou hast found \bibemph{it} so quickly, my son? And he said, Because the \textsc{Lord} thy God brought \bibemph{it} to me.
\fn{to me: Heb. before me.}
\vs Gen 27:21 And Isaac said unto Jacob, Come near, I pray thee, that I may feel thee, my son, whether thou \bibemph{be} my very son Esau or not.
\vs Gen 27:22 And Jacob went near unto Isaac his father; and he felt him, and said, The voice \bibemph{is} Jacob's voice, but the hands \bibemph{are} the hands of Esau.
\vs Gen 27:23 And he discerned him not, because his hands were hairy, as his brother Esau's hands: so he blessed him.
\vs Gen 27:24 And he said, \bibemph{Art} thou my very son Esau? And he said, I \bibemph{am}.
\vs Gen 27:25 And he said, Bring \bibemph{it} near to me, and I will eat of my son's venison, that my soul may bless thee. And he brought \bibemph{it} near to him, and he did eat: and he brought him wine, and he drank.
\vs Gen 27:26 And his father Isaac said unto him, Come near now, and kiss me, my son.
\vs Gen 27:27 And he came near, and kissed him: and he smelled the smell of his raiment, and blessed him, and said, See, the smell of my son \bibemph{is} as the smell of a field which the \textsc{Lord} hath blessed:
\vs Gen 27:28 Therefore God give thee of the dew of heaven, and the fatness of the earth, and plenty of corn and wine:
\vs Gen 27:29 Let people serve thee, and nations bow down to thee: be lord over thy brethren, and let thy mother's sons bow down to thee: cursed \bibemph{be} every one that curseth thee, and blessed \bibemph{be} he that blesseth thee.
\vs Gen 27:30 And it came to pass, as soon as Isaac had made an end of blessing Jacob, and Jacob was yet scarce gone out from the presence of Isaac his father, that Esau his brother came in from his hunting.
\vs Gen 27:31 And he also had made savoury meat, and brought it unto his father, and said unto his father, Let my father arise, and eat of his son's venison, that thy soul may bless me.
\vs Gen 27:32 And Isaac his father said unto him, Who \bibemph{art} thou? And he said, I \bibemph{am} thy son, thy firstborn Esau.
\vs Gen 27:33 And Isaac trembled very exceedingly, and said, Who? where \bibemph{is} he that hath taken venison, and brought \bibemph{it} me, and I have eaten of all before thou camest, and have blessed him? yea, \bibemph{and} he shall be blessed.
\fn{trembled...: Heb. trembled with a great trembling greatly. taken: Heb. hunted.}
\vs Gen 27:34 And when Esau heard the words of his father, he cried with a great and exceeding bitter cry, and said unto his father, Bless me, \bibemph{even} me also, O my father.
\vs Gen 27:35 And he said, Thy brother came with subtilty, and hath taken away thy blessing.
\vs Gen 27:36 And he said, Is not he rightly named Jacob? for he hath supplanted me these two times: he took away my birthright; and, behold, now he hath taken away my blessing. And he said, Hast thou not reserved a blessing for me?
\fn{Jacob: that is, A supplanter.}
\vs Gen 27:37 And Isaac answered and said unto Esau, Behold, I have made him thy lord, and all his brethren have I given to him for servants; and with corn and wine have I sustained him: and what shall I do now unto thee, my son?
\fn{sustained: or, supported.}
\vs Gen 27:38 And Esau said unto his father, Hast thou but one blessing, my father? bless me, \bibemph{even} me also, O my father. And Esau lifted up his voice, and wept.
\vs Gen 27:39 And Isaac his father answered and said unto him, Behold, thy dwelling shall be the fatness of the earth, and of the dew of heaven from above;
\fn{the fatness: or, of the fatness.}
\vs Gen 27:40 And by thy sword shalt thou live, and shalt serve thy brother; and it shall come to pass when thou shalt have the dominion, that thou shalt break his yoke from off thy neck.
\vs Gen 27:41 And Esau hated Jacob because of the blessing wherewith his father blessed him: and Esau said in his heart, The days of mourning for my father are at hand; then will I slay my brother Jacob.
\vs Gen 27:42 And these words of Esau her elder son were told to Rebekah: and she sent and called Jacob her younger son, and said unto him, Behold, thy brother Esau, as touching thee, doth comfort himself, \bibemph{purposing} to kill thee.
\vs Gen 27:43 Now therefore, my son, obey my voice; and arise, flee thou to Laban my brother to Haran;
\vs Gen 27:44 And tarry with him a few days, until thy brother's fury turn away;
\vs Gen 27:45 Until thy brother's anger turn away from thee, and he forget \bibemph{that} which thou hast done to him: then I will send, and fetch thee from thence: why should I be deprived also of you both in one day?
\vs Gen 27:46 And Rebekah said to Isaac, I am weary of my life because of the daughters of Heth: if Jacob take a wife of the daughters of Heth, such as these \bibemph{which are} of the daughters of the land, what good shall my life do me?
\vs Gen 28:1 And Isaac called Jacob, and blessed him, and charged him, and said unto him, Thou shalt not take a wife of the daughters of Canaan.
\vs Gen 28:2 Arise, go to Padanaram, to the house of Bethuel thy mother's father; and take thee a wife from thence of the daughters of Laban thy mother's brother.
\vs Gen 28:3 And God Almighty bless thee, and make thee fruitful, and multiply thee, that thou mayest be a multitude of people;
\fn{a multitude...: Heb. an assembly of people.}
\vs Gen 28:4 And give thee the blessing of Abraham, to thee, and to thy seed with thee; that thou mayest inherit the land wherein thou art a stranger, which God gave unto Abraham.
\fn{wherein...: Heb. of thy sojournings.}
\vs Gen 28:5 And Isaac sent away Jacob: and he went to Padanaram unto Laban, son of Bethuel the Syrian, the brother of Rebekah, Jacob's and Esau's mother.
\vs Gen 28:6 When Esau saw that Isaac had blessed Jacob, and sent him away to Padanaram, to take him a wife from thence; and that as he blessed him he gave him a charge, saying, Thou shalt not take a wife of the daughters of Canaan;
\vs Gen 28:7 And that Jacob obeyed his father and his mother, and was gone to Padanaram;
\vs Gen 28:8 And Esau seeing that the daughters of Canaan pleased not Isaac his father;
\fn{pleased...: Heb. were evil in the eyes, etc.}
\vs Gen 28:9 Then went Esau unto Ishmael, and took unto the wives which he had Mahalath the daughter of Ishmael Abraham's son, the sister of Nebajoth, to be his wife.
\fn{Mahalath: or, Bashemath.}
\vs Gen 28:10 And Jacob went out from Beersheba, and went toward Haran.
\fn{Haran: Gr. Charran.}
\vs Gen 28:11 And he lighted upon a certain place, and tarried there all night, because the sun was set; and he took of the stones of that place, and put \bibemph{them for} his pillows, and lay down in that place to sleep.
\vs Gen 28:12 And he dreamed, and behold a ladder set up on the earth, and the top of it reached to heaven: and behold the angels of God ascending and descending on it.
\vs Gen 28:13 And, behold, the \textsc{Lord} stood above it, and said, I \bibemph{am} the \textsc{Lord} God of Abraham thy father, and the God of Isaac: the land whereon thou liest, to thee will I give it, and to thy seed;
\vs Gen 28:14 And thy seed shall be as the dust of the earth, and thou shalt spread abroad to the west, and to the east, and to the north, and to the south: and in thee and in thy seed shall all the families of the earth be blessed.
\fn{spread...: Heb. break forth.}
\vs Gen 28:15 And, behold, I \bibemph{am} with thee, and will keep thee in all \bibemph{places} whither thou goest, and will bring thee again into this land; for I will not leave thee, until I have done \bibemph{that} which I have spoken to thee of.
\vs Gen 28:16 And Jacob awaked out of his sleep, and he said, Surely the \textsc{Lord} is in this place; and I knew \bibemph{it} not.
\vs Gen 28:17 And he was afraid, and said, How dreadful \bibemph{is} this place! this \bibemph{is} none other but the house of God, and this \bibemph{is} the gate of heaven.
\vs Gen 28:18 And Jacob rose up early in the morning, and took the stone that he had put \bibemph{for} his pillows, and set it up \bibemph{for} a pillar, and poured oil upon the top of it.
\vs Gen 28:19 And he called the name of that place Bethel: but the name of that city \bibemph{was called} Luz at the first.
\fn{Bethel: that is, The house of God.}
\vs Gen 28:20 And Jacob vowed a vow, saying, If God will be with me, and will keep me in this way that I go, and will give me bread to eat, and raiment to put on,
\vs Gen 28:21 So that I come again to my father's house in peace; then shall the \textsc{Lord} be my God:
\vs Gen 28:22 And this stone, which I have set \bibemph{for} a pillar, shall be God's house: and of all that thou shalt give me I will surely give the tenth unto thee.
\vs Gen 29:1 Then Jacob went on his journey, and came into the land of the people of the east.
\fn{went...: Heb. lift up his feet. people: Heb. children.}
\vs Gen 29:2 And he looked, and behold a well in the field, and, lo, there \bibemph{were} three flocks of sheep lying by it; for out of that well they watered the flocks: and a great stone \bibemph{was} upon the well's mouth.
\vs Gen 29:3 And thither were all the flocks gathered: and they rolled the stone from the well's mouth, and watered the sheep, and put the stone again upon the well's mouth in his place.
\vs Gen 29:4 And Jacob said unto them, My brethren, whence \bibemph{be} ye? And they said, Of Haran \bibemph{are} we.
\vs Gen 29:5 And he said unto them, Know ye Laban the son of Nahor? And they said, We know \bibemph{him}.
\vs Gen 29:6 And he said unto them, \bibemph{Is} he well? And they said, \bibemph{He is} well: and, behold, Rachel his daughter cometh with the sheep.
\fn{Is he...: Heb. Is there peace to him?.}
\vs Gen 29:7 And he said, Lo, \bibemph{it is} yet high day, neither \bibemph{is it} time that the cattle should be gathered together: water ye the sheep, and go \bibemph{and} feed \bibemph{them}.
\fn{it is...: Heb. yet the day is great.}
\vs Gen 29:8 And they said, We cannot, until all the flocks be gathered together, and \bibemph{till} they roll the stone from the well's mouth; then we water the sheep.
\vs Gen 29:9 And while he yet spake with them, Rachel came with her father's sheep: for she kept them.
\vs Gen 29:10 And it came to pass, when Jacob saw Rachel the daughter of Laban his mother's brother, and the sheep of Laban his mother's brother, that Jacob went near, and rolled the stone from the well's mouth, and watered the flock of Laban his mother's brother.
\vs Gen 29:11 And Jacob kissed Rachel, and lifted up his voice, and wept.
\vs Gen 29:12 And Jacob told Rachel that he \bibemph{was} her father's brother, and that he \bibemph{was} Rebekah's son: and she ran and told her father.
\vs Gen 29:13 And it came to pass, when Laban heard the tidings of Jacob his sister's son, that he ran to meet him, and embraced him, and kissed him, and brought him to his house. And he told Laban all these things.
\fn{tidings: Heb. hearing.}
\vs Gen 29:14 And Laban said to him, Surely thou \bibemph{art} my bone and my flesh. And he abode with him the space of a month.
\fn{the space...: Heb. a month of days.}
\vs Gen 29:15 And Laban said unto Jacob, Because thou \bibemph{art} my brother, shouldest thou therefore serve me for nought? tell me, what \bibemph{shall} thy wages \bibemph{be}?
\vs Gen 29:16 And Laban had two daughters: the name of the elder \bibemph{was} Leah, and the name of the younger \bibemph{was} Rachel.
\vs Gen 29:17 Leah \bibemph{was} tender eyed; but Rachel was beautiful and well favoured.
\vs Gen 29:18 And Jacob loved Rachel; and said, I will serve thee seven years for Rachel thy younger daughter.
\vs Gen 29:19 And Laban said, \bibemph{It is} better that I give her to thee, than that I should give her to another man: abide with me.
\vs Gen 29:20 And Jacob served seven years for Rachel; and they seemed unto him \bibemph{but} a few days, for the love he had to her.
\vs Gen 29:21 And Jacob said unto Laban, Give \bibemph{me} my wife, for my days are fulfilled, that I may go in unto her.
\vs Gen 29:22 And Laban gathered together all the men of the place, and made a feast.
\vs Gen 29:23 And it came to pass in the evening, that he took Leah his daughter, and brought her to him; and he went in unto her.
\vs Gen 29:24 And Laban gave unto his daughter Leah Zilpah his maid \bibemph{for} an handmaid.
\vs Gen 29:25 And it came to pass, that in the morning, behold, it \bibemph{was} Leah: and he said to Laban, What \bibemph{is} this thou hast done unto me? did not I serve with thee for Rachel? wherefore then hast thou beguiled me?
\vs Gen 29:26 And Laban said, It must not be so done in our country, to give the younger before the firstborn.
\fn{country: Heb. place.}
\vs Gen 29:27 Fulfil her week, and we will give thee this also for the service which thou shalt serve with me yet seven other years.
\vs Gen 29:28 And Jacob did so, and fulfilled her week: and he gave him Rachel his daughter to wife also.
\vs Gen 29:29 And Laban gave to Rachel his daughter Bilhah his handmaid to be her maid.
\vs Gen 29:30 And he went in also unto Rachel, and he loved also Rachel more than Leah, and served with him yet seven other years.
\vs Gen 29:31 And when the \textsc{Lord} saw that Leah \bibemph{was} hated, he opened her womb: but Rachel \bibemph{was} barren.
\vs Gen 29:32 And Leah conceived, and bare a son, and she called his name Reuben: for she said, Surely the \textsc{Lord} hath looked upon my affliction; now therefore my husband will love me.
\fn{Reuben: that is, See a son.}
\vs Gen 29:33 And she conceived again, and bare a son; and said, Because the \textsc{Lord} hath heard that I \bibemph{was} hated, he hath therefore given me this \bibemph{son} also: and she called his name Simeon.
\fn{Simeon: that is, Hearing.}
\vs Gen 29:34 And she conceived again, and bare a son; and said, Now this time will my husband be joined unto me, because I have born him three sons: therefore was his name called Levi.
\fn{Levi: that is, Joined.}
\vs Gen 29:35 And she conceived again, and bare a son: and she said, Now will I praise the \textsc{Lord}: therefore she called his name Judah; and left bearing.
\fn{Judah: that is, Praise. left...: Heb. stood from bearing.}
\vs Gen 30:1 And when Rachel saw that she bare Jacob no children, Rachel envied her sister; and said unto Jacob, Give me children, or else I die.
\vs Gen 30:2 And Jacob's anger was kindled against Rachel: and he said, \bibemph{Am} I in God's stead, who hath withheld from thee the fruit of the womb?
\vs Gen 30:3 And she said, Behold my maid Bilhah, go in unto her; and she shall bear upon my knees, that I may also have children by her.
\fn{have...: Heb. be built by her.}
\vs Gen 30:4 And she gave him Bilhah her handmaid to wife: and Jacob went in unto her.
\vs Gen 30:5 And Bilhah conceived, and bare Jacob a son.
\vs Gen 30:6 And Rachel said, God hath judged me, and hath also heard my voice, and hath given me a son: therefore called she his name Dan.
\fn{Dan: that is, Judging.}
\vs Gen 30:7 And Bilhah Rachel's maid conceived again, and bare Jacob a second son.
\vs Gen 30:8 And Rachel said, With great wrestlings have I wrestled with my sister, and I have prevailed: and she called his name Naphtali.
\fn{great...: Heb. wrestlings of God. Naphtali: that is, My wrestling: Gr. Nephthalim.}
\vs Gen 30:9 When Leah saw that she had left bearing, she took Zilpah her maid, and gave her Jacob to wife.
\vs Gen 30:10 And Zilpah Leah's maid bare Jacob a son.
\vs Gen 30:11 And Leah said, A troop cometh: and she called his name Gad.
\fn{Gad: that is, A troop, or, company.}
\vs Gen 30:12 And Zilpah Leah's maid bare Jacob a second son.
\vs Gen 30:13 And Leah said, Happy am I, for the daughters will call me blessed: and she called his name Asher.
\fn{Happy...: Heb. In my happiness. Asher: that is, Happy.}
\vs Gen 30:14 And Reuben went in the days of wheat harvest, and found mandrakes in the field, and brought them unto his mother Leah. Then Rachel said to Leah, Give me, I pray thee, of thy son's mandrakes.
\vs Gen 30:15 And she said unto her, \bibemph{Is it} a small matter that thou hast taken my husband? and wouldest thou take away my son's mandrakes also? And Rachel said, Therefore he shall lie with thee to night for thy son's mandrakes.
\vs Gen 30:16 And Jacob came out of the field in the evening, and Leah went out to meet him, and said, Thou must come in unto me; for surely I have hired thee with my son's mandrakes. And he lay with her that night.
\vs Gen 30:17 And God hearkened unto Leah, and she conceived, and bare Jacob the fifth son.
\vs Gen 30:18 And Leah said, God hath given me my hire, because I have given my maiden to my husband: and she called his name Issachar.
\fn{Issachar: that is, An hire.}
\vs Gen 30:19 And Leah conceived again, and bare Jacob the sixth son.
\vs Gen 30:20 And Leah said, God hath endued me \bibemph{with} a good dowry; now will my husband dwell with me, because I have born him six sons: and she called his name Zebulun.
\fn{Zebulun: that is, Dwelling: Gr. Zabulon.}
\vs Gen 30:21 And afterwards she bare a daughter, and called her name Dinah.
\fn{Dinah: that is Judgment.}
\vs Gen 30:22 And God remembered Rachel, and God hearkened to her, and opened her womb.
\vs Gen 30:23 And she conceived, and bare a son; and said, God hath taken away my reproach:
\vs Gen 30:24 And she called his name Joseph; and said, The \textsc{Lord} shall add to me another son.
\fn{Joseph: that is, Adding.}
\vs Gen 30:25 And it came to pass, when Rachel had born Joseph, that Jacob said unto Laban, Send me away, that I may go unto mine own place, and to my country.
\vs Gen 30:26 Give \bibemph{me} my wives and my children, for whom I have served thee, and let me go: for thou knowest my service which I have done thee.
\vs Gen 30:27 And Laban said unto him, I pray thee, if I have found favour in thine eyes, \bibemph{tarry: for} I have learned by experience that the \textsc{Lord} hath blessed me for thy sake.
\vs Gen 30:28 And he said, Appoint me thy wages, and I will give \bibemph{it}.
\vs Gen 30:29 And he said unto him, Thou knowest how I have served thee, and how thy cattle was with me.
\vs Gen 30:30 For \bibemph{it was} little which thou hadst before I \bibemph{came}, and it is \bibemph{now} increased unto a multitude; and the \textsc{Lord} hath blessed thee since my coming: and now when shall I provide for mine own house also?
\fn{increased: Heb. broken forth. since...: Heb. at my foot.}
\vs Gen 30:31 And he said, What shall I give thee? And Jacob said, Thou shalt not give me any thing: if thou wilt do this thing for me, I will again feed \bibemph{and} keep thy flock:
\vs Gen 30:32 I will pass through all thy flock to day, removing from thence all the speckled and spotted cattle, and all the brown cattle among the sheep, and the spotted and speckled among the goats: and \bibemph{of such} shall be my hire.
\vs Gen 30:33 So shall my righteousness answer for me in time to come, when it shall come for my hire before thy face: every one that \bibemph{is} not speckled and spotted among the goats, and brown among the sheep, that shall be counted stolen with me.
\fn{in time...: Heb. to morrow.}
\vs Gen 30:34 And Laban said, Behold, I would it might be according to thy word.
\vs Gen 30:35 And he removed that day the he goats that were ringstraked and spotted, and all the she goats that were speckled and spotted, \bibemph{and} every one that had \bibemph{some} white in it, and all the brown among the sheep, and gave \bibemph{them} into the hand of his sons.
\vs Gen 30:36 And he set three days' journey betwixt himself and Jacob: and Jacob fed the rest of Laban's flocks.
\vs Gen 30:37 And Jacob took him rods of green poplar, and of the hazel and chesnut tree; and pilled white strakes in them, and made the white appear which \bibemph{was} in the rods.
\vs Gen 30:38 And he set the rods which he had pilled before the flocks in the gutters in the watering troughs when the flocks came to drink, that they should conceive when they came to drink.
\vs Gen 30:39 And the flocks conceived before the rods, and brought forth cattle ringstraked, speckled, and spotted.
\vs Gen 30:40 And Jacob did separate the lambs, and set the faces of the flocks toward the ringstraked, and all the brown in the flock of Laban; and he put his own flocks by themselves, and put them not unto Laban's cattle.
\vs Gen 30:41 And it came to pass, whensoever the stronger cattle did conceive, that Jacob laid the rods before the eyes of the cattle in the gutters, that they might conceive among the rods.
\vs Gen 30:42 But when the cattle were feeble, he put \bibemph{them} not in: so the feebler were Laban's, and the stronger Jacob's.
\vs Gen 30:43 And the man increased exceedingly, and had much cattle, and maidservants, and menservants, and camels, and asses.
\vs Gen 31:1 And he heard the words of Laban's sons, saying, Jacob hath taken away all that \bibemph{was} our father's; and of \bibemph{that} which \bibemph{was} our father's hath he gotten all this glory.
\vs Gen 31:2 And Jacob beheld the countenance of Laban, and, behold, it \bibemph{was} not toward him as before.
\fn{as before: Heb. as yesterday and the day before.}
\vs Gen 31:3 And the \textsc{Lord} said unto Jacob, Return unto the land of thy fathers, and to thy kindred; and I will be with thee.
\vs Gen 31:4 And Jacob sent and called Rachel and Leah to the field unto his flock,
\vs Gen 31:5 And said unto them, I see your father's countenance, that it \bibemph{is} not toward me as before; but the God of my father hath been with me.
\vs Gen 31:6 And ye know that with all my power I have served your father.
\vs Gen 31:7 And your father hath deceived me, and changed my wages ten times; but God suffered him not to hurt me.
\vs Gen 31:8 If he said thus, The speckled shall be thy wages; then all the cattle bare speckled: and if he said thus, The ringstraked shall be thy hire; then bare all the cattle ringstraked.
\vs Gen 31:9 Thus God hath taken away the cattle of your father, and given \bibemph{them} to me.
\vs Gen 31:10 And it came to pass at the time that the cattle conceived, that I lifted up mine eyes, and saw in a dream, and, behold, the rams which leaped upon the cattle \bibemph{were} ringstraked, speckled, and grisled.
\fn{rams: or, he goats.}
\vs Gen 31:11 And the angel of God spake unto me in a dream, \bibemph{saying}, Jacob: And I said, Here \bibemph{am} I.
\vs Gen 31:12 And he said, Lift up now thine eyes, and see, all the rams which leap upon the cattle \bibemph{are} ringstraked, speckled, and grisled: for I have seen all that Laban doeth unto thee.
\vs Gen 31:13 I \bibemph{am} the God of Bethel, where thou anointedst the pillar, \bibemph{and} where thou vowedst a vow unto me: now arise, get thee out from this land, and return unto the land of thy kindred.
\vs Gen 31:14 And Rachel and Leah answered and said unto him, \bibemph{Is there} yet any portion or inheritance for us in our father's house?
\vs Gen 31:15 Are we not counted of him strangers? for he hath sold us, and hath quite devoured also our money.
\vs Gen 31:16 For all the riches which God hath taken from our father, that \bibemph{is} ours, and our children's: now then, whatsoever God hath said unto thee, do.
\vs Gen 31:17 Then Jacob rose up, and set his sons and his wives upon camels;
\vs Gen 31:18 And he carried away all his cattle, and all his goods which he had gotten, the cattle of his getting, which he had gotten in Padanaram, for to go to Isaac his father in the land of Canaan.
\vs Gen 31:19 And Laban went to shear his sheep: and Rachel had stolen the images that \bibemph{were} her father's.
\fn{images: Heb. teraphim.}
\vs Gen 31:20 And Jacob stole away unawares to Laban the Syrian, in that he told him not that he fled.
\fn{unawares...: Heb. the heart of Laban.}
\vs Gen 31:21 So he fled with all that he had; and he rose up, and passed over the river, and set his face \bibemph{toward} the mount Gilead.
\vs Gen 31:22 And it was told Laban on the third day that Jacob was fled.
\vs Gen 31:23 And he took his brethren with him, and pursued after him seven days' journey; and they overtook him in the mount Gilead.
\vs Gen 31:24 And God came to Laban the Syrian in a dream by night, and said unto him, Take heed that thou speak not to Jacob either good or bad.
\fn{either...: Heb. from good to bad.}
\vs Gen 31:25 Then Laban overtook Jacob. Now Jacob had pitched his tent in the mount: and Laban with his brethren pitched in the mount of Gilead.
\vs Gen 31:26 And Laban said to Jacob, What hast thou done, that thou hast stolen away unawares to me, and carried away my daughters, as captives \bibemph{taken} with the sword?
\vs Gen 31:27 Wherefore didst thou flee away secretly, and steal away from me; and didst not tell me, that I might have sent thee away with mirth, and with songs, with tabret, and with harp?
\fn{steal...: Heb. hast stolen me.}
\vs Gen 31:28 And hast not suffered me to kiss my sons and my daughters? thou hast now done foolishly in \bibemph{so} doing.
\vs Gen 31:29 It is in the power of my hand to do you hurt: but the God of your father spake unto me yesternight, saying, Take thou heed that thou speak not to Jacob either good or bad.
\vs Gen 31:30 And now, \bibemph{though} thou wouldest needs be gone, because thou sore longedst after thy father's house, \bibemph{yet} wherefore hast thou stolen my gods?
\vs Gen 31:31 And Jacob answered and said to Laban, Because I was afraid: for I said, Peradventure thou wouldest take by force thy daughters from me.
\vs Gen 31:32 With whomsoever thou findest thy gods, let him not live: before our brethren discern thou what \bibemph{is} thine with me, and take \bibemph{it} to thee. For Jacob knew not that Rachel had stolen them.
\vs Gen 31:33 And Laban went into Jacob's tent, and into Leah's tent, and into the two maidservants' tents; but he found \bibemph{them} not. Then went he out of Leah's tent, and entered into Rachel's tent.
\vs Gen 31:34 Now Rachel had taken the images, and put them in the camel's furniture, and sat upon them. And Laban searched all the tent, but found \bibemph{them} not.
\fn{searched: Heb. felt.}
\vs Gen 31:35 And she said to her father, Let it not displease my lord that I cannot rise up before thee; for the custom of women \bibemph{is} upon me. And he searched, but found not the images.
\vs Gen 31:36 And Jacob was wroth, and chode with Laban: and Jacob answered and said to Laban, What \bibemph{is} my trespass? what \bibemph{is} my sin, that thou hast so hotly pursued after me?
\vs Gen 31:37 Whereas thou hast searched all my stuff, what hast thou found of all thy household stuff? set \bibemph{it} here before my brethren and thy brethren, that they may judge betwixt us both.
\fn{searched: Heb. felt.}
\vs Gen 31:38 This twenty years \bibemph{have} I \bibemph{been} with thee; thy ewes and thy she goats have not cast their young, and the rams of thy flock have I not eaten.
\vs Gen 31:39 That which was torn \bibemph{of beasts} I brought not unto thee; I bare the loss of it; of my hand didst thou require it, \bibemph{whether} stolen by day, or stolen by night.
\vs Gen 31:40 \bibemph{Thus} I was; in the day the drought consumed me, and the frost by night; and my sleep departed from mine eyes.
\vs Gen 31:41 Thus have I been twenty years in thy house; I served thee fourteen years for thy two daughters, and six years for thy cattle: and thou hast changed my wages ten times.
\vs Gen 31:42 Except the God of my father, the God of Abraham, and the fear of Isaac, had been with me, surely thou hadst sent me away now empty. God hath seen mine affliction and the labour of my hands, and rebuked \bibemph{thee} yesternight.
\vs Gen 31:43 And Laban answered and said unto Jacob, \bibemph{These} daughters \bibemph{are} my daughters, and \bibemph{these} children \bibemph{are} my children, and \bibemph{these} cattle \bibemph{are} my cattle, and all that thou seest \bibemph{is} mine: and what can I do this day unto these my daughters, or unto their children which they have born?
\vs Gen 31:44 Now therefore come thou, let us make a covenant, I and thou; and let it be for a witness between me and thee.
\vs Gen 31:45 And Jacob took a stone, and set it up \bibemph{for} a pillar.
\vs Gen 31:46 And Jacob said unto his brethren, Gather stones; and they took stones, and made an heap: and they did eat there upon the heap.
\vs Gen 31:47 And Laban called it Jegarsahadutha: but Jacob called it Galeed.
\fn{Jegarsahadutha: that is, The heap of witness, Chaldee. Galeed: that is, The heap of witness, Heb..}
\vs Gen 31:48 And Laban said, This heap \bibemph{is} a witness between me and thee this day. Therefore was the name of it called Galeed;
\vs Gen 31:49 And Mizpah; for he said, The \textsc{Lord} watch between me and thee, when we are absent one from another.
\fn{Mizpah: that is, A beacon, or, watchtower.}
\vs Gen 31:50 If thou shalt afflict my daughters, or if thou shalt take \bibemph{other} wives beside my daughters, no man \bibemph{is} with us; see, God \bibemph{is} witness betwixt me and thee.
\vs Gen 31:51 And Laban said to Jacob, Behold this heap, and behold \bibemph{this} pillar, which I have cast betwixt me and thee;
\vs Gen 31:52 This heap \bibemph{be} witness, and \bibemph{this} pillar \bibemph{be} witness, that I will not pass over this heap to thee, and that thou shalt not pass over this heap and this pillar unto me, for harm.
\vs Gen 31:53 The God of Abraham, and the God of Nahor, the God of their father, judge betwixt us. And Jacob sware by the fear of his father Isaac.
\vs Gen 31:54 Then Jacob offered sacrifice upon the mount, and called his brethren to eat bread: and they did eat bread, and tarried all night in the mount.
\fn{offered...: or, killed beasts.}
\vs Gen 31:55 And early in the morning Laban rose up, and kissed his sons and his daughters, and blessed them: and Laban departed, and returned unto his place.
\vs Gen 32:1 And Jacob went on his way, and the angels of God met him.
\vs Gen 32:2 And when Jacob saw them, he said, This \bibemph{is} God's host: and he called the name of that place Mahanaim.
\fn{Mahanaim: that is, Two hosts, or, camps.}
\vs Gen 32:3 And Jacob sent messengers before him to Esau his brother unto the land of Seir, the country of Edom.
\fn{country: Heb. field.}
\vs Gen 32:4 And he commanded them, saying, Thus shall ye speak unto my lord Esau; Thy servant Jacob saith thus, I have sojourned with Laban, and stayed there until now:
\vs Gen 32:5 And I have oxen, and asses, flocks, and menservants, and womenservants: and I have sent to tell my lord, that I may find grace in thy sight.
\vs Gen 32:6 And the messengers returned to Jacob, saying, We came to thy brother Esau, and also he cometh to meet thee, and four hundred men with him.
\vs Gen 32:7 Then Jacob was greatly afraid and distressed: and he divided the people that \bibemph{was} with him, and the flocks, and herds, and the camels, into two bands;
\vs Gen 32:8 And said, If Esau come to the one company, and smite it, then the other company which is left shall escape.
\vs Gen 32:9 And Jacob said, O God of my father Abraham, and God of my father Isaac, the \textsc{Lord} which saidst unto me, Return unto thy country, and to thy kindred, and I will deal well with thee:
\vs Gen 32:10 I am not worthy of the least of all the mercies, and of all the truth, which thou hast shewed unto thy servant; for with my staff I passed over this Jordan; and now I am become two bands.
\fn{I am not...: Heb. I am less than all.}
\vs Gen 32:11 Deliver me, I pray thee, from the hand of my brother, from the hand of Esau: for I fear him, lest he will come and smite me, \bibemph{and} the mother with the children.
\fn{with: Heb. upon.}
\vs Gen 32:12 And thou saidst, I will surely do thee good, and make thy seed as the sand of the sea, which cannot be numbered for multitude.
\vs Gen 32:13 And he lodged there that same night; and took of that which came to his hand a present for Esau his brother;
\vs Gen 32:14 Two hundred she goats, and twenty he goats, two hundred ewes, and twenty rams,
\vs Gen 32:15 Thirty milch camels with their colts, forty kine, and ten bulls, twenty she asses, and ten foals.
\vs Gen 32:16 And he delivered \bibemph{them} into the hand of his servants, every drove by themselves; and said unto his servants, Pass over before me, and put a space betwixt drove and drove.
\vs Gen 32:17 And he commanded the foremost, saying, When Esau my brother meeteth thee, and asketh thee, saying, Whose \bibemph{art} thou? and whither goest thou? and whose \bibemph{are} these before thee?
\vs Gen 32:18 Then thou shalt say, \bibemph{They be} thy servant Jacob's; it \bibemph{is} a present sent unto my lord Esau: and, behold, also he \bibemph{is} behind us.
\vs Gen 32:19 And so commanded he the second, and the third, and all that followed the droves, saying, On this manner shall ye speak unto Esau, when ye find him.
\vs Gen 32:20 And say ye moreover, Behold, thy servant Jacob \bibemph{is} behind us. For he said, I will appease him with the present that goeth before me, and afterward I will see his face; peradventure he will accept of me.
\fn{of me: Heb. my face.}
\vs Gen 32:21 So went the present over before him: and himself lodged that night in the company.
\vs Gen 32:22 And he rose up that night, and took his two wives, and his two womenservants, and his eleven sons, and passed over the ford Jabbok.
\vs Gen 32:23 And he took them, and sent them over the brook, and sent over that he had.
\fn{sent them: Heb. caused to pass.}
\vs Gen 32:24 And Jacob was left alone; and there wrestled a man with him until the breaking of the day.
\fn{breaking...: Heb. ascending of the morning.}
\vs Gen 32:25 And when he saw that he prevailed not against him, he touched the hollow of his thigh; and the hollow of Jacob's thigh was out of joint, as he wrestled with him.
\vs Gen 32:26 And he said, Let me go, for the day breaketh. And he said, I will not let thee go, except thou bless me.
\vs Gen 32:27 And he said unto him, What \bibemph{is} thy name? And he said, Jacob.
\vs Gen 32:28 And he said, Thy name shall be called no more Jacob, but Israel: for as a prince hast thou power with God and with men, and hast prevailed.
\fn{Israel: that is, A prince of God.}
\vs Gen 32:29 And Jacob asked \bibemph{him}, and said, Tell \bibemph{me}, I pray thee, thy name. And he said, Wherefore \bibemph{is} it \bibemph{that} thou dost ask after my name? And he blessed him there.
\vs Gen 32:30 And Jacob called the name of the place Peniel: for I have seen God face to face, and my life is preserved.
\fn{Peniel: that is, The face of God.}
\vs Gen 32:31 And as he passed over Penuel the sun rose upon him, and he halted upon his thigh.
\vs Gen 32:32 Therefore the children of Israel eat not \bibemph{of} the sinew which shrank, which \bibemph{is} upon the hollow of the thigh, unto this day: because he touched the hollow of Jacob's thigh in the sinew that shrank.
\vs Gen 33:1 And Jacob lifted up his eyes, and looked, and, behold, Esau came, and with him four hundred men. And he divided the children unto Leah, and unto Rachel, and unto the two handmaids.
\vs Gen 33:2 And he put the handmaids and their children foremost, and Leah and her children after, and Rachel and Joseph hindermost.
\vs Gen 33:3 And he passed over before them, and bowed himself to the ground seven times, until he came near to his brother.
\vs Gen 33:4 And Esau ran to meet him, and embraced him, and fell on his neck, and kissed him: and they wept.
\vs Gen 33:5 And he lifted up his eyes, and saw the women and the children; and said, Who \bibemph{are} those with thee? And he said, The children which God hath graciously given thy servant.
\fn{with...: Heb. to thee.}
\vs Gen 33:6 Then the handmaidens came near, they and their children, and they bowed themselves.
\vs Gen 33:7 And Leah also with her children came near, and bowed themselves: and after came Joseph near and Rachel, and they bowed themselves.
\vs Gen 33:8 And he said, What \bibemph{meanest} thou by all this drove which I met? And he said, \bibemph{These are} to find grace in the sight of my lord.
\fn{What...: Heb. What is all this band to thee?.}
\vs Gen 33:9 And Esau said, I have enough, my brother; keep that thou hast unto thyself.
\fn{keep...: Heb. be that to thee that is thine.}
\vs Gen 33:10 And Jacob said, Nay, I pray thee, if now I have found grace in thy sight, then receive my present at my hand: for therefore I have seen thy face, as though I had seen the face of God, and thou wast pleased with me.
\vs Gen 33:11 Take, I pray thee, my blessing that is brought to thee; because God hath dealt graciously with me, and because I have enough. And he urged him, and he took \bibemph{it}.
\fn{enough: Heb. all things.}
\vs Gen 33:12 And he said, Let us take our journey, and let us go, and I will go before thee.
\vs Gen 33:13 And he said unto him, My lord knoweth that the children \bibemph{are} tender, and the flocks and herds with young \bibemph{are} with me: and if men should overdrive them one day, all the flock will die.
\vs Gen 33:14 Let my lord, I pray thee, pass over before his servant: and I will lead on softly, according as the cattle that goeth before me and the children be able to endure, until I come unto my lord unto Seir.
\fn{according...: Heb. according to the foot of the work, etc, and according to the foot of the children.}
\vs Gen 33:15 And Esau said, Let me now leave with thee \bibemph{some} of the folk that \bibemph{are} with me. And he said, What needeth it? let me find grace in the sight of my lord.
\fn{leave: Heb. set, or, place. What...: Heb. Wherefore is this?.}
\vs Gen 33:16 So Esau returned that day on his way unto Seir.
\vs Gen 33:17 And Jacob journeyed to Succoth, and built him an house, and made booths for his cattle: therefore the name of the place is called Succoth.
\fn{Succoth: that is, Booths.}
\vs Gen 33:18 And Jacob came to Shalem, a city of Shechem, which \bibemph{is} in the land of Canaan, when he came from Padanaram; and pitched his tent before the city.
\fn{Shechem: Gr. Sychem.}
\vs Gen 33:19 And he bought a parcel of a field, where he had spread his tent, at the hand of the children of Hamor, Shechem's father, for an hundred pieces of money.
\fn{Hamor: Gr. Emmor. pieces...: or, lambs.}
\vs Gen 33:20 And he erected there an altar, and called it Elelohe-Israel.
\fn{Elelohe-Israel: that is God the God of Israel.}
\vs Gen 34:1 And Dinah the daughter of Leah, which she bare unto Jacob, went out to see the daughters of the land.
\vs Gen 34:2 And when Shechem the son of Hamor the Hivite, prince of the country, saw her, he took her, and lay with her, and defiled her.
\fn{defiled...: Heb. humbled her.}
\vs Gen 34:3 And his soul clave unto Dinah the daughter of Jacob, and he loved the damsel, and spake kindly unto the damsel.
\fn{kindly...: Heb. to the heart of the damsel.}
\vs Gen 34:4 And Shechem spake unto his father Hamor, saying, Get me this damsel to wife.
\vs Gen 34:5 And Jacob heard that he had defiled Dinah his daughter: now his sons were with his cattle in the field: and Jacob held his peace until they were come.
\vs Gen 34:6 And Hamor the father of Shechem went out unto Jacob to commune with him.
\vs Gen 34:7 And the sons of Jacob came out of the field when they heard \bibemph{it}: and the men were grieved, and they were very wroth, because he had wrought folly in Israel in lying with Jacob's daughter; which thing ought not to be done.
\vs Gen 34:8 And Hamor communed with them, saying, The soul of my son Shechem longeth for your daughter: I pray you give her him to wife.
\vs Gen 34:9 And make ye marriages with us, \bibemph{and} give your daughters unto us, and take our daughters unto you.
\vs Gen 34:10 And ye shall dwell with us: and the land shall be before you; dwell and trade ye therein, and get you possessions therein.
\vs Gen 34:11 And Shechem said unto her father and unto her brethren, Let me find grace in your eyes, and what ye shall say unto me I will give.
\vs Gen 34:12 Ask me never so much dowry and gift, and I will give according as ye shall say unto me: but give me the damsel to wife.
\vs Gen 34:13 And the sons of Jacob answered Shechem and Hamor his father deceitfully, and said, because he had defiled Dinah their sister:
\vs Gen 34:14 And they said unto them, We cannot do this thing, to give our sister to one that is uncircumcised; for that \bibemph{were} a reproach unto us:
\vs Gen 34:15 But in this will we consent unto you: If ye will be as we \bibemph{be}, that every male of you be circumcised;
\vs Gen 34:16 Then will we give our daughters unto you, and we will take your daughters to us, and we will dwell with you, and we will become one people.
\vs Gen 34:17 But if ye will not hearken unto us, to be circumcised; then will we take our daughter, and we will be gone.
\vs Gen 34:18 And their words pleased Hamor, and Shechem Hamor's son.
\vs Gen 34:19 And the young man deferred not to do the thing, because he had delight in Jacob's daughter: and he \bibemph{was} more honourable than all the house of his father.
\vs Gen 34:20 And Hamor and Shechem his son came unto the gate of their city, and communed with the men of their city, saying,
\vs Gen 34:21 These men \bibemph{are} peaceable with us; therefore let them dwell in the land, and trade therein; for the land, behold, \bibemph{it is} large enough for them; let us take their daughters to us for wives, and let us give them our daughters.
\vs Gen 34:22 Only herein will the men consent unto us for to dwell with us, to be one people, if every male among us be circumcised, as they \bibemph{are} circumcised.
\vs Gen 34:23 \bibemph{Shall} not their cattle and their substance and every beast of theirs \bibemph{be} ours? only let us consent unto them, and they will dwell with us.
\vs Gen 34:24 And unto Hamor and unto Shechem his son hearkened all that went out of the gate of his city; and every male was circumcised, all that went out of the gate of his city.
\vs Gen 34:25 And it came to pass on the third day, when they were sore, that two of the sons of Jacob, Simeon and Levi, Dinah's brethren, took each man his sword, and came upon the city boldly, and slew all the males.
\vs Gen 34:26 And they slew Hamor and Shechem his son with the edge of the sword, and took Dinah out of Shechem's house, and went out.
\fn{edge: Heb. mouth.}
\vs Gen 34:27 The sons of Jacob came upon the slain, and spoiled the city, because they had defiled their sister.
\vs Gen 34:28 They took their sheep, and their oxen, and their asses, and that which \bibemph{was} in the city, and that which \bibemph{was} in the field,
\vs Gen 34:29 And all their wealth, and all their little ones, and their wives took they captive, and spoiled even all that \bibemph{was} in the house.
\vs Gen 34:30 And Jacob said to Simeon and Levi, Ye have troubled me to make me to stink among the inhabitants of the land, among the Canaanites and the Perizzites: and I \bibemph{being} few in number, they shall gather themselves together against me, and slay me; and I shall be destroyed, I and my house.
\vs Gen 34:31 And they said, Should he deal with our sister as with an harlot?
\vs Gen 35:1 And God said unto Jacob, Arise, go up to Bethel, and dwell there: and make there an altar unto God, that appeared unto thee when thou fleddest from the face of Esau thy brother.
\vs Gen 35:2 Then Jacob said unto his household, and to all that \bibemph{were} with him, Put away the strange gods that \bibemph{are} among you, and be clean, and change your garments:
\vs Gen 35:3 And let us arise, and go up to Bethel; and I will make there an altar unto God, who answered me in the day of my distress, and was with me in the way which I went.
\vs Gen 35:4 And they gave unto Jacob all the strange gods which \bibemph{were} in their hand, and \bibemph{all their} earrings which \bibemph{were} in their ears; and Jacob hid them under the oak which \bibemph{was} by Shechem.
\vs Gen 35:5 And they journeyed: and the terror of God was upon the cities that \bibemph{were} round about them, and they did not pursue after the sons of Jacob.
\vs Gen 35:6 So Jacob came to Luz, which \bibemph{is} in the land of Canaan, that \bibemph{is}, Bethel, he and all the people that \bibemph{were} with him.
\vs Gen 35:7 And he built there an altar, and called the place Elbethel: because there God appeared unto him, when he fled from the face of his brother.
\fn{Elbethel: that is, The God of Bethel.}
\vs Gen 35:8 But Deborah Rebekah's nurse died, and she was buried beneath Bethel under an oak: and the name of it was called Allonbachuth.
\fn{Allonbachuth: that is, The oak of weeping.}
\vs Gen 35:9 And God appeared unto Jacob again, when he came out of Padanaram, and blessed him.
\vs Gen 35:10 And God said unto him, Thy name \bibemph{is} Jacob: thy name shall not be called any more Jacob, but Israel shall be thy name: and he called his name Israel.
\vs Gen 35:11 And God said unto him, I \bibemph{am} God Almighty: be fruitful and multiply; a nation and a company of nations shall be of thee, and kings shall come out of thy loins;
\vs Gen 35:12 And the land which I gave Abraham and Isaac, to thee I will give it, and to thy seed after thee will I give the land.
\vs Gen 35:13 And God went up from him in the place where he talked with him.
\vs Gen 35:14 And Jacob set up a pillar in the place where he talked with him, \bibemph{even} a pillar of stone: and he poured a drink offering thereon, and he poured oil thereon.
\vs Gen 35:15 And Jacob called the name of the place where God spake with him, Bethel.
\vs Gen 35:16 And they journeyed from Bethel; and there was but a little way to come to Ephrath: and Rachel travailed, and she had hard labour.
\fn{a little...: Heb. a little piece of ground.}
\vs Gen 35:17 And it came to pass, when she was in hard labour, that the midwife said unto her, Fear not; thou shalt have this son also.
\vs Gen 35:18 And it came to pass, as her soul was in departing, (for she died) that she called his name Benoni: but his father called him Benjamin.
\fn{Benoni: that is, The son of my sorrow. Benjamin: that is, The son of the right hand.}
\vs Gen 35:19 And Rachel died, and was buried in the way to Ephrath, which \bibemph{is} Bethlehem.
\vs Gen 35:20 And Jacob set a pillar upon her grave: that \bibemph{is} the pillar of Rachel's grave unto this day.
\vs Gen 35:21 And Israel journeyed, and spread his tent beyond the tower of Edar.
\vs Gen 35:22 And it came to pass, when Israel dwelt in that land, that Reuben went and lay with Bilhah his father's concubine: and Israel heard \bibemph{it}. Now the sons of Jacob were twelve:
\vs Gen 35:23 The sons of Leah; Reuben, Jacob's firstborn, and Simeon, and Levi, and Judah, and Issachar, and Zebulun:
\vs Gen 35:24 The sons of Rachel; Joseph, and Benjamin:
\vs Gen 35:25 And the sons of Bilhah, Rachel's handmaid; Dan, and Naphtali:
\vs Gen 35:26 And the sons of Zilpah, Leah's handmaid; Gad, and Asher: these \bibemph{are} the sons of Jacob, which were born to him in Padanaram.
\vs Gen 35:27 And Jacob came unto Isaac his father unto Mamre, unto the city of Arbah, which \bibemph{is} Hebron, where Abraham and Isaac sojourned.
\vs Gen 35:28 And the days of Isaac were an hundred and fourscore years.
\vs Gen 35:29 And Isaac gave up the ghost, and died, and was gathered unto his people, \bibemph{being} old and full of days: and his sons Esau and Jacob buried him.
\vs Gen 36:1 Now these \bibemph{are} the generations of Esau, who \bibemph{is} Edom.
\vs Gen 36:2 Esau took his wives of the daughters of Canaan; Adah the daughter of Elon the Hittite, and Aholibamah the daughter of Anah the daughter of Zibeon the Hivite;
\vs Gen 36:3 And Bashemath Ishmael's daughter, sister of Nebajoth.
\vs Gen 36:4 And Adah bare to Esau Eliphaz; and Bashemath bare Reuel;
\vs Gen 36:5 And Aholibamah bare Jeush, and Jaalam, and Korah: these \bibemph{are} the sons of Esau, which were born unto him in the land of Canaan.
\vs Gen 36:6 And Esau took his wives, and his sons, and his daughters, and all the persons of his house, and his cattle, and all his beasts, and all his substance, which he had got in the land of Canaan; and went into the country from the face of his brother Jacob.
\fn{persons: Heb. souls.}
\vs Gen 36:7 For their riches were more than that they might dwell together; and the land wherein they were strangers could not bear them because of their cattle.
\vs Gen 36:8 Thus dwelt Esau in mount Seir: Esau \bibemph{is} Edom.
\vs Gen 36:9 And these \bibemph{are} the generations of Esau the father of the Edomites in mount Seir:
\fn{the Edomites: Heb. Edom.}
\vs Gen 36:10 These \bibemph{are} the names of Esau's sons; Eliphaz the son of Adah the wife of Esau, Reuel the son of Bashemath the wife of Esau.
\vs Gen 36:11 And the sons of Eliphaz were Teman, Omar, Zepho, and Gatam, and Kenaz.
\fn{Zepho: or, Zephi.}
\vs Gen 36:12 And Timna was concubine to Eliphaz Esau's son; and she bare to Eliphaz Amalek: these \bibemph{were} the sons of Adah Esau's wife.
\vs Gen 36:13 And these \bibemph{are} the sons of Reuel; Nahath, and Zerah, Shammah, and Mizzah: these were the sons of Bashemath Esau's wife.
\vs Gen 36:14 And these were the sons of Aholibamah, the daughter of Anah the daughter of Zibeon, Esau's wife: and she bare to Esau Jeush, and Jaalam, and Korah.
\vs Gen 36:15 These \bibemph{were} dukes of the sons of Esau: the sons of Eliphaz the firstborn \bibemph{son} of Esau; duke Teman, duke Omar, duke Zepho, duke Kenaz,
\vs Gen 36:16 Duke Korah, duke Gatam, \bibemph{and} duke Amalek: these \bibemph{are} the dukes \bibemph{that came} of Eliphaz in the land of Edom; these \bibemph{were} the sons of Adah.
\vs Gen 36:17 And these \bibemph{are} the sons of Reuel Esau's son; duke Nahath, duke Zerah, duke Shammah, duke Mizzah: these \bibemph{are} the dukes \bibemph{that came} of Reuel in the land of Edom; these \bibemph{are} the sons of Bashemath Esau's wife.
\vs Gen 36:18 And these \bibemph{are} the sons of Aholibamah Esau's wife; duke Jeush, duke Jaalam, duke Korah: these \bibemph{were} the dukes \bibemph{that came} of Aholibamah the daughter of Anah, Esau's wife.
\vs Gen 36:19 These \bibemph{are} the sons of Esau, who \bibemph{is} Edom, and these \bibemph{are} their dukes.
\vs Gen 36:20 These \bibemph{are} the sons of Seir the Horite, who inhabited the land; Lotan, and Shobal, and Zibeon, and Anah,
\vs Gen 36:21 And Dishon, and Ezer, and Dishan: these \bibemph{are} the dukes of the Horites, the children of Seir in the land of Edom.
\vs Gen 36:22 And the children of Lotan were Hori and Hemam; and Lotan's sister \bibemph{was} Timna.
\fn{Hemam: or, Homam.}
\vs Gen 36:23 And the children of Shobal \bibemph{were} these; Alvan, and Manahath, and Ebal, Shepho, and Onam.
\fn{Alvan: or, Alian. Shepho: or, Shephi.}
\vs Gen 36:24 And these \bibemph{are} the children of Zibeon; both Ajah, and Anah: this \bibemph{was that} Anah that found the mules in the wilderness, as he fed the asses of Zibeon his father.
\vs Gen 36:25 And the children of Anah \bibemph{were} these; Dishon, and Aholibamah the daughter of Anah.
\vs Gen 36:26 And these \bibemph{are} the children of Dishon; Hemdan, and Eshban, and Ithran, and Cheran.
\fn{Hemdan: or, Amram.}
\vs Gen 36:27 The children of Ezer \bibemph{are} these; Bilhan, and Zaavan, and Akan.
\fn{Akan: or, Jakan.}
\vs Gen 36:28 The children of Dishan \bibemph{are} these; Uz, and Aran.
\vs Gen 36:29 These \bibemph{are} the dukes \bibemph{that came} of the Horites; duke Lotan, duke Shobal, duke Zibeon, duke Anah,
\vs Gen 36:30 Duke Dishon, duke Ezer, duke Dishan: these \bibemph{are} the dukes \bibemph{that came} of Hori, among their dukes in the land of Seir.
\vs Gen 36:31 And these \bibemph{are} the kings that reigned in the land of Edom, before there reigned any king over the children of Israel.
\vs Gen 36:32 And Bela the son of Beor reigned in Edom: and the name of his city \bibemph{was} Dinhabah.
\vs Gen 36:33 And Bela died, and Jobab the son of Zerah of Bozrah reigned in his stead.
\vs Gen 36:34 And Jobab died, and Husham of the land of Temani reigned in his stead.
\vs Gen 36:35 And Husham died, and Hadad the son of Bedad, who smote Midian in the field of Moab, reigned in his stead: and the name of his city \bibemph{was} Avith.
\vs Gen 36:36 And Hadad died, and Samlah of Masrekah reigned in his stead.
\vs Gen 36:37 And Samlah died, and Saul of Rehoboth \bibemph{by} the river reigned in his stead.
\vs Gen 36:38 And Saul died, and Baalhanan the son of Achbor reigned in his stead.
\vs Gen 36:39 And Baalhanan the son of Achbor died, and Hadar reigned in his stead: and the name of his city \bibemph{was} Pau; and his wife's name \bibemph{was} Mehetabel, the daughter of Matred, the daughter of Mezahab.
\fn{Hadar, Pau: or, Hadad, Pai: after his death was an Aristocracy.}
\vs Gen 36:40 And these \bibemph{are} the names of the dukes \bibemph{that came} of Esau, according to their families, after their places, by their names; duke Timnah, duke Alvah, duke Jetheth,
\fn{Alvah: or, Aliah.}
\vs Gen 36:41 Duke Aholibamah, duke Elah, duke Pinon,
\vs Gen 36:42 Duke Kenaz, duke Teman, duke Mibzar,
\vs Gen 36:43 Duke Magdiel, duke Iram: these \bibemph{be} the dukes of Edom, according to their habitations in the land of their possession: he \bibemph{is} Esau the father of the Edomites.
\fn{the Edomites: Heb. Edom.}
\vs Gen 37:1 And Jacob dwelt in the land wherein his father was a stranger, in the land of Canaan.
\fn{wherein...: Heb. of his father's sojournings.}
\vs Gen 37:2 These \bibemph{are} the generations of Jacob. Joseph, \bibemph{being} seventeen years old, was feeding the flock with his brethren; and the lad \bibemph{was} with the sons of Bilhah, and with the sons of Zilpah, his father's wives: and Joseph brought unto his father their evil report.
\vs Gen 37:3 Now Israel loved Joseph more than all his children, because he \bibemph{was} the son of his old age: and he made him a coat of \bibemph{many} colours.
\fn{colours: or, pieces.}
\vs Gen 37:4 And when his brethren saw that their father loved him more than all his brethren, they hated him, and could not speak peaceably unto him.
\vs Gen 37:5 And Joseph dreamed a dream, and he told \bibemph{it} his brethren: and they hated him yet the more.
\vs Gen 37:6 And he said unto them, Hear, I pray you, this dream which I have dreamed:
\vs Gen 37:7 For, behold, we \bibemph{were} binding sheaves in the field, and, lo, my sheaf arose, and also stood upright; and, behold, your sheaves stood round about, and made obeisance to my sheaf.
\vs Gen 37:8 And his brethren said to him, Shalt thou indeed reign over us? or shalt thou indeed have dominion over us? And they hated him yet the more for his dreams, and for his words.
\vs Gen 37:9 And he dreamed yet another dream, and told it his brethren, and said, Behold, I have dreamed a dream more; and, behold, the sun and the moon and the eleven stars made obeisance to me.
\vs Gen 37:10 And he told \bibemph{it} to his father, and to his brethren: and his father rebuked him, and said unto him, What \bibemph{is} this dream that thou hast dreamed? Shall I and thy mother and thy brethren indeed come to bow down ourselves to thee to the earth?
\vs Gen 37:11 And his brethren envied him; but his father observed the saying.
\vs Gen 37:12 And his brethren went to feed their father's flock in Shechem.
\vs Gen 37:13 And Israel said unto Joseph, Do not thy brethren feed \bibemph{the flock} in Shechem? come, and I will send thee unto them. And he said to him, Here \bibemph{am I}.
\vs Gen 37:14 And he said to him, Go, I pray thee, see whether it be well with thy brethren, and well with the flocks; and bring me word again. So he sent him out of the vale of Hebron, and he came to Shechem.
\fn{see...: Heb. see the peace of thy brethren, etc.}
\vs Gen 37:15 And a certain man found him, and, behold, \bibemph{he was} wandering in the field: and the man asked him, saying, What seekest thou?
\vs Gen 37:16 And he said, I seek my brethren: tell me, I pray thee, where they feed \bibemph{their flocks}.
\vs Gen 37:17 And the man said, They are departed hence; for I heard them say, Let us go to Dothan. And Joseph went after his brethren, and found them in Dothan.
\vs Gen 37:18 And when they saw him afar off, even before he came near unto them, they conspired against him to slay him.
\vs Gen 37:19 And they said one to another, Behold, this dreamer cometh.
\fn{dreamer: Heb. master of dreams.}
\vs Gen 37:20 Come now therefore, and let us slay him, and cast him into some pit, and we will say, Some evil beast hath devoured him: and we shall see what will become of his dreams.
\vs Gen 37:21 And Reuben heard \bibemph{it}, and he delivered him out of their hands; and said, Let us not kill him.
\vs Gen 37:22 And Reuben said unto them, Shed no blood, \bibemph{but} cast him into this pit that \bibemph{is} in the wilderness, and lay no hand upon him; that he might rid him out of their hands, to deliver him to his father again.
\vs Gen 37:23 And it came to pass, when Joseph was come unto his brethren, that they stript Joseph out of his coat, \bibemph{his} coat of \bibemph{many} colours that \bibemph{was} on him;
\fn{colours: or, pieces.}
\vs Gen 37:24 And they took him, and cast him into a pit: and the pit \bibemph{was} empty, \bibemph{there was} no water in it.
\vs Gen 37:25 And they sat down to eat bread: and they lifted up their eyes and looked, and, behold, a company of Ishmeelites came from Gilead with their camels bearing spicery and balm and myrrh, going to carry \bibemph{it} down to Egypt.
\vs Gen 37:26 And Judah said unto his brethren, What profit \bibemph{is it} if we slay our brother, and conceal his blood?
\vs Gen 37:27 Come, and let us sell him to the Ishmeelites, and let not our hand be upon him; for he \bibemph{is} our brother \bibemph{and} our flesh. And his brethren were content.
\fn{were...: Heb. hearkened.}
\vs Gen 37:28 Then there passed by Midianites merchantmen; and they drew and lifted up Joseph out of the pit, and sold Joseph to the Ishmeelites for twenty \bibemph{pieces} of silver: and they brought Joseph into Egypt.
\vs Gen 37:29 And Reuben returned unto the pit; and, behold, Joseph \bibemph{was} not in the pit; and he rent his clothes.
\vs Gen 37:30 And he returned unto his brethren, and said, The child \bibemph{is} not; and I, whither shall I go?
\vs Gen 37:31 And they took Joseph's coat, and killed a kid of the goats, and dipped the coat in the blood;
\vs Gen 37:32 And they sent the coat of \bibemph{many} colours, and they brought \bibemph{it} to their father; and said, This have we found: know now whether it \bibemph{be} thy son's coat or no.
\vs Gen 37:33 And he knew it, and said, \bibemph{It is} my son's coat; an evil beast hath devoured him; Joseph is without doubt rent in pieces.
\vs Gen 37:34 And Jacob rent his clothes, and put sackcloth upon his loins, and mourned for his son many days.
\vs Gen 37:35 And all his sons and all his daughters rose up to comfort him; but he refused to be comforted; and he said, For I will go down into the grave unto my son mourning. Thus his father wept for him.
\vs Gen 37:36 And the Midianites sold him into Egypt unto Potiphar, an officer of Pharaoh's, \bibemph{and} captain of the guard.
\fn{officer: Heb. eunuch: but the word doth signify not only eunuchs, but also chamberlains, courtiers, and officers. captain...: or, chief marshal: Heb. chief of the slaughter men, or executioners.}
\vs Gen 38:1 And it came to pass at that time, that Judah went down from his brethren, and turned in to a certain Adullamite, whose name \bibemph{was} Hirah.
\vs Gen 38:2 And Judah saw there a daughter of a certain Canaanite, whose name \bibemph{was} Shuah; and he took her, and went in unto her.
\vs Gen 38:3 And she conceived, and bare a son; and he called his name Er.
\vs Gen 38:4 And she conceived again, and bare a son; and she called his name Onan.
\vs Gen 38:5 And she yet again conceived, and bare a son; and called his name Shelah: and he was at Chezib, when she bare him.
\vs Gen 38:6 And Judah took a wife for Er his firstborn, whose name \bibemph{was} Tamar.
\vs Gen 38:7 And Er, Judah's firstborn, was wicked in the sight of the \textsc{Lord}; and the \textsc{Lord} slew him.
\vs Gen 38:8 And Judah said unto Onan, Go in unto thy brother's wife, and marry her, and raise up seed to thy brother.
\vs Gen 38:9 And Onan knew that the seed should not be his; and it came to pass, when he went in unto his brother's wife, that he spilled \bibemph{it} on the ground, lest that he should give seed to his brother.
\vs Gen 38:10 And the thing which he did displeased the \textsc{Lord}: wherefore he slew him also.
\fn{displeased...: Heb. was evil in the eyes of the Lord.}
\vs Gen 38:11 Then said Judah to Tamar his daughter in law, Remain a widow at thy father's house, till Shelah my son be grown: for he said, Lest peradventure he die also, as his brethren \bibemph{did}. And Tamar went and dwelt in her father's house.
\vs Gen 38:12 And in process of time the daughter of Shuah Judah's wife died; and Judah was comforted, and went up unto his sheepshearers to Timnath, he and his friend Hirah the Adullamite.
\fn{in process...: Heb. the days were multiplied.}
\vs Gen 38:13 And it was told Tamar, saying, Behold thy father in law goeth up to Timnath to shear his sheep.
\vs Gen 38:14 And she put her widow's garments off from her, and covered her with a vail, and wrapped herself, and sat in an open place, which \bibemph{is} by the way to Timnath; for she saw that Shelah was grown, and she was not given unto him to wife.
\fn{an open...: Heb. the door of eyes, or, of Enajim.}
\vs Gen 38:15 When Judah saw her, he thought her \bibemph{to be} an harlot; because she had covered her face.
\vs Gen 38:16 And he turned unto her by the way, and said, Go to, I pray thee, let me come in unto thee; (for he knew not that she \bibemph{was} his daughter in law.) And she said, What wilt thou give me, that thou mayest come in unto me?
\vs Gen 38:17 And he said, I will send \bibemph{thee} a kid from the flock. And she said, Wilt thou give \bibemph{me} a pledge, till thou send \bibemph{it}?
\fn{a kid: Heb. a kid of the goats.}
\vs Gen 38:18 And he said, What pledge shall I give thee? And she said, Thy signet, and thy bracelets, and thy staff that \bibemph{is} in thine hand. And he gave \bibemph{it} her, and came in unto her, and she conceived by him.
\vs Gen 38:19 And she arose, and went away, and laid by her vail from her, and put on the garments of her widowhood.
\vs Gen 38:20 And Judah sent the kid by the hand of his friend the Adullamite, to receive \bibemph{his} pledge from the woman's hand: but he found her not.
\vs Gen 38:21 Then he asked the men of that place, saying, Where \bibemph{is} the harlot, that \bibemph{was} openly by the way side? And they said, There was no harlot in this \bibemph{place}.
\fn{openly: or, in Enajim.}
\vs Gen 38:22 And he returned to Judah, and said, I cannot find her; and also the men of the place said, \bibemph{that} there was no harlot in this \bibemph{place}.
\vs Gen 38:23 And Judah said, Let her take \bibemph{it} to her, lest we be shamed: behold, I sent this kid, and thou hast not found her.
\fn{be shamed: Heb. become a contempt.}
\vs Gen 38:24 And it came to pass about three months after, that it was told Judah, saying, Tamar thy daughter in law hath played the harlot; and also, behold, she \bibemph{is} with child by whoredom. And Judah said, Bring her forth, and let her be burnt.
\vs Gen 38:25 When she \bibemph{was} brought forth, she sent to her father in law, saying, By the man, whose these \bibemph{are, am} I with child: and she said, Discern, I pray thee, whose \bibemph{are} these, the signet, and bracelets, and staff.
\vs Gen 38:26 And Judah acknowledged \bibemph{them}, and said, She hath been more righteous than I; because that I gave her not to Shelah my son. And he knew her again no more.
\vs Gen 38:27 And it came to pass in the time of her travail, that, behold, twins \bibemph{were} in her womb.
\vs Gen 38:28 And it came to pass, when she travailed, that \bibemph{the one} put out \bibemph{his} hand: and the midwife took and bound upon his hand a scarlet thread, saying, This came out first.
\vs Gen 38:29 And it came to pass, as he drew back his hand, that, behold, his brother came out: and she said, How hast thou broken forth? \bibemph{this} breach \bibemph{be} upon thee: therefore his name was called Pharez.
\fn{How hast...: or, Wherefore hast thou made this breach against thee?. Pharez: that is A breach.}
\vs Gen 38:30 And afterward came out his brother, that had the scarlet thread upon his hand: and his name was called Zarah.
\vs Gen 39:1 And Joseph was brought down to Egypt; and Potiphar, an officer of Pharaoh, captain of the guard, an Egyptian, bought him of the hands of the Ishmeelites, which had brought him down thither.
\vs Gen 39:2 And the \textsc{Lord} was with Joseph, and he was a prosperous man; and he was in the house of his master the Egyptian.
\vs Gen 39:3 And his master saw that the \textsc{Lord} \bibemph{was} with him, and that the \textsc{Lord} made all that he did to prosper in his hand.
\vs Gen 39:4 And Joseph found grace in his sight, and he served him: and he made him overseer over his house, and all \bibemph{that} he had he put into his hand.
\vs Gen 39:5 And it came to pass from the time \bibemph{that} he had made him overseer in his house, and over all that he had, that the \textsc{Lord} blessed the Egyptian's house for Joseph's sake; and the blessing of the \textsc{Lord} was upon all that he had in the house, and in the field.
\vs Gen 39:6 And he left all that he had in Joseph's hand; and he knew not ought he had, save the bread which he did eat. And Joseph was \bibemph{a} goodly \bibemph{person}, and well favoured.
\vs Gen 39:7 And it came to pass after these things, that his master's wife cast her eyes upon Joseph; and she said, Lie with me.
\vs Gen 39:8 But he refused, and said unto his master's wife, Behold, my master wotteth not what \bibemph{is} with me in the house, and he hath committed all that he hath to my hand;
\vs Gen 39:9 \bibemph{There is} none greater in this house than I; neither hath he kept back any thing from me but thee, because thou \bibemph{art} his wife: how then can I do this great wickedness, and sin against God?
\vs Gen 39:10 And it came to pass, as she spake to Joseph day by day, that he hearkened not unto her, to lie by her, \bibemph{or} to be with her.
\vs Gen 39:11 And it came to pass about this time, that \bibemph{Joseph} went into the house to do his business; and \bibemph{there was} none of the men of the house there within.
\vs Gen 39:12 And she caught him by his garment, saying, Lie with me: and he left his garment in her hand, and fled, and got him out.
\vs Gen 39:13 And it came to pass, when she saw that he had left his garment in her hand, and was fled forth,
\vs Gen 39:14 That she called unto the men of her house, and spake unto them, saying, See, he hath brought in an Hebrew unto us to mock us; he came in unto me to lie with me, and I cried with a loud voice:
\fn{loud: Heb. great.}
\vs Gen 39:15 And it came to pass, when he heard that I lifted up my voice and cried, that he left his garment with me, and fled, and got him out.
\vs Gen 39:16 And she laid up his garment by her, until his lord came home.
\vs Gen 39:17 And she spake unto him according to these words, saying, The Hebrew servant, which thou hast brought unto us, came in unto me to mock me:
\vs Gen 39:18 And it came to pass, as I lifted up my voice and cried, that he left his garment with me, and fled out.
\vs Gen 39:19 And it came to pass, when his master heard the words of his wife, which she spake unto him, saying, After this manner did thy servant to me; that his wrath was kindled.
\vs Gen 39:20 And Joseph's master took him, and put him into the prison, a place where the king's prisoners \bibemph{were} bound: and he was there in the prison.
\vs Gen 39:21 But the \textsc{Lord} was with Joseph, and shewed him mercy, and gave him favour in the sight of the keeper of the prison.
\fn{shewed...: Heb. extended kindness unto him.}
\vs Gen 39:22 And the keeper of the prison committed to Joseph's hand all the prisoners that \bibemph{were} in the prison; and whatsoever they did there, he was the doer \bibemph{of it}.
\vs Gen 39:23 The keeper of the prison looked not to any thing \bibemph{that was} under his hand; because the \textsc{Lord} was with him, and \bibemph{that} which he did, the \textsc{Lord} made \bibemph{it} to prosper.
\vs Gen 40:1 And it came to pass after these things, \bibemph{that} the butler of the king of Egypt and \bibemph{his} baker had offended their lord the king of Egypt.
\vs Gen 40:2 And Pharaoh was wroth against two \bibemph{of} his officers, against the chief of the butlers, and against the chief of the bakers.
\vs Gen 40:3 And he put them in ward in the house of the captain of the guard, into the prison, the place where Joseph \bibemph{was} bound.
\vs Gen 40:4 And the captain of the guard charged Joseph with them, and he served them: and they continued a season in ward.
\vs Gen 40:5 And they dreamed a dream both of them, each man his dream in one night, each man according to the interpretation of his dream, the butler and the baker of the king of Egypt, which \bibemph{were} bound in the prison.
\vs Gen 40:6 And Joseph came in unto them in the morning, and looked upon them, and, behold, they \bibemph{were} sad.
\vs Gen 40:7 And he asked Pharaoh's officers that \bibemph{were} with him in the ward of his lord's house, saying, Wherefore look ye \bibemph{so} sadly to day?
\fn{look...: Heb. are your faces evil?.}
\vs Gen 40:8 And they said unto him, We have dreamed a dream, and \bibemph{there is} no interpreter of it. And Joseph said unto them, \bibemph{Do} not interpretations \bibemph{belong} to God? tell me \bibemph{them}, I pray you.
\vs Gen 40:9 And the chief butler told his dream to Joseph, and said to him, In my dream, behold, a vine \bibemph{was} before me;
\vs Gen 40:10 And in the vine \bibemph{were} three branches: and it \bibemph{was} as though it budded, \bibemph{and} her blossoms shot forth; and the clusters thereof brought forth ripe grapes:
\vs Gen 40:11 And Pharaoh's cup \bibemph{was} in my hand: and I took the grapes, and pressed them into Pharaoh's cup, and I gave the cup into Pharaoh's hand.
\vs Gen 40:12 And Joseph said unto him, This \bibemph{is} the interpretation of it: The three branches \bibemph{are} three days:
\vs Gen 40:13 Yet within three days shall Pharaoh lift up thine head, and restore thee unto thy place: and thou shalt deliver Pharaoh's cup into his hand, after the former manner when thou wast his butler.
\fn{lift...: or, reckon.}
\vs Gen 40:14 But think on me when it shall be well with thee, and shew kindness, I pray thee, unto me, and make mention of me unto Pharaoh, and bring me out of this house:
\fn{think...: Heb. remember me with thee.}
\vs Gen 40:15 For indeed I was stolen away out of the land of the Hebrews: and here also have I done nothing that they should put me into the dungeon.
\vs Gen 40:16 When the chief baker saw that the interpretation was good, he said unto Joseph, I also \bibemph{was} in my dream, and, behold, \bibemph{I had} three white baskets on my head:
\fn{white: or, full of holes.}
\vs Gen 40:17 And in the uppermost basket \bibemph{there was} of all manner of bakemeats for Pharaoh; and the birds did eat them out of the basket upon my head.
\fn{bakemeats...: Heb. meat of Pharaoh, the work of a baker, or, cook.}
\vs Gen 40:18 And Joseph answered and said, This \bibemph{is} the interpretation thereof: The three baskets \bibemph{are} three days:
\vs Gen 40:19 Yet within three days shall Pharaoh lift up thy head from off thee, and shall hang thee on a tree; and the birds shall eat thy flesh from off thee.
\fn{lift...: or, reckon thee, and take thy office from thee.}
\vs Gen 40:20 And it came to pass the third day, \bibemph{which was} Pharaoh's birthday, that he made a feast unto all his servants: and he lifted up the head of the chief butler and of the chief baker among his servants.
\fn{lifted...: or, reckoned.}
\vs Gen 40:21 And he restored the chief butler unto his butlership again; and he gave the cup into Pharaoh's hand:
\vs Gen 40:22 But he hanged the chief baker: as Joseph had interpreted to them.
\vs Gen 40:23 Yet did not the chief butler remember Joseph, but forgat him.
\vs Gen 41:1 And it came to pass at the end of two full years, that Pharaoh dreamed: and, behold, he stood by the river.
\vs Gen 41:2 And, behold, there came up out of the river seven well favoured kine and fatfleshed; and they fed in a meadow.
\vs Gen 41:3 And, behold, seven other kine came up after them out of the river, ill favoured and leanfleshed; and stood by the \bibemph{other} kine upon the brink of the river.
\vs Gen 41:4 And the ill favoured and leanfleshed kine did eat up the seven well favoured and fat kine. So Pharaoh awoke.
\vs Gen 41:5 And he slept and dreamed the second time: and, behold, seven ears of corn came up upon one stalk, rank and good.
\fn{rank: Heb. fat.}
\vs Gen 41:6 And, behold, seven thin ears and blasted with the east wind sprung up after them.
\vs Gen 41:7 And the seven thin ears devoured the seven rank and full ears. And Pharaoh awoke, and, behold, \bibemph{it was} a dream.
\vs Gen 41:8 And it came to pass in the morning that his spirit was troubled; and he sent and called for all the magicians of Egypt, and all the wise men thereof: and Pharaoh told them his dream; but \bibemph{there was} none that could interpret them unto Pharaoh.
\vs Gen 41:9 Then spake the chief butler unto Pharaoh, saying, I do remember my faults this day:
\vs Gen 41:10 Pharaoh was wroth with his servants, and put me in ward in the captain of the guard's house, \bibemph{both} me and the chief baker:
\vs Gen 41:11 And we dreamed a dream in one night, I and he; we dreamed each man according to the interpretation of his dream.
\vs Gen 41:12 And \bibemph{there was} there with us a young man, an Hebrew, servant to the captain of the guard; and we told him, and he interpreted to us our dreams; to each man according to his dream he did interpret.
\vs Gen 41:13 And it came to pass, as he interpreted to us, so it was; me he restored unto mine office, and him he hanged.
\vs Gen 41:14 Then Pharaoh sent and called Joseph, and they brought him hastily out of the dungeon: and he shaved \bibemph{himself}, and changed his raiment, and came in unto Pharaoh.
\fn{brought...: Heb. made him run.}
\vs Gen 41:15 And Pharaoh said unto Joseph, I have dreamed a dream, and \bibemph{there is} none that can interpret it: and I have heard say of thee, \bibemph{that} thou canst understand a dream to interpret it.
\fn{thou...: or, when thou hearest a dream thou canst interpret it.}
\vs Gen 41:16 And Joseph answered Pharaoh, saying, \bibemph{It is} not in me: God shall give Pharaoh an answer of peace.
\vs Gen 41:17 And Pharaoh said unto Joseph, In my dream, behold, I stood upon the bank of the river:
\vs Gen 41:18 And, behold, there came up out of the river seven kine, fatfleshed and well favoured; and they fed in a meadow:
\vs Gen 41:19 And, behold, seven other kine came up after them, poor and very ill favoured and leanfleshed, such as I never saw in all the land of Egypt for badness:
\vs Gen 41:20 And the lean and the ill favoured kine did eat up the first seven fat kine:
\vs Gen 41:21 And when they had eaten them up, it could not be known that they had eaten them; but they \bibemph{were} still ill favoured, as at the beginning. So I awoke.
\fn{eaten...: Heb. come to the inward parts of them.}
\vs Gen 41:22 And I saw in my dream, and, behold, seven ears came up in one stalk, full and good:
\vs Gen 41:23 And, behold, seven ears, withered, thin, \bibemph{and} blasted with the east wind, sprung up after them:
\fn{withered: or, small.}
\vs Gen 41:24 And the thin ears devoured the seven good ears: and I told \bibemph{this} unto the magicians; but \bibemph{there was} none that could declare \bibemph{it} to me.
\vs Gen 41:25 And Joseph said unto Pharaoh, The dream of Pharaoh \bibemph{is} one: God hath shewed Pharaoh what he \bibemph{is} about to do.
\vs Gen 41:26 The seven good kine \bibemph{are} seven years; and the seven good ears \bibemph{are} seven years: the dream \bibemph{is} one.
\vs Gen 41:27 And the seven thin and ill favoured kine that came up after them \bibemph{are} seven years; and the seven empty ears blasted with the east wind shall be seven years of famine.
\vs Gen 41:28 This \bibemph{is} the thing which I have spoken unto Pharaoh: What God \bibemph{is} about to do he sheweth unto Pharaoh.
\vs Gen 41:29 Behold, there come seven years of great plenty throughout all the land of Egypt:
\vs Gen 41:30 And there shall arise after them seven years of famine; and all the plenty shall be forgotten in the land of Egypt; and the famine shall consume the land;
\vs Gen 41:31 And the plenty shall not be known in the land by reason of that famine following; for it \bibemph{shall be} very grievous.
\fn{grievous: Heb. heavy.}
\vs Gen 41:32 And for that the dream was doubled unto Pharaoh twice; \bibemph{it is} because the thing \bibemph{is} established by God, and God will shortly bring it to pass.
\fn{established...: or, prepared of God.}
\vs Gen 41:33 Now therefore let Pharaoh look out a man discreet and wise, and set him over the land of Egypt.
\vs Gen 41:34 Let Pharaoh do \bibemph{this}, and let him appoint officers over the land, and take up the fifth part of the land of Egypt in the seven plenteous years.
\fn{officers: or, overseers.}
\vs Gen 41:35 And let them gather all the food of those good years that come, and lay up corn under the hand of Pharaoh, and let them keep food in the cities.
\vs Gen 41:36 And that food shall be for store to the land against the seven years of famine, which shall be in the land of Egypt; that the land perish not through the famine.
\fn{perish...: Heb. be not cut off.}
\vs Gen 41:37 And the thing was good in the eyes of Pharaoh, and in the eyes of all his servants.
\vs Gen 41:38 And Pharaoh said unto his servants, Can we find \bibemph{such a one} as this \bibemph{is}, a man in whom the Spirit of God \bibemph{is}?
\vs Gen 41:39 And Pharaoh said unto Joseph, Forasmuch as God hath shewed thee all this, \bibemph{there is} none so discreet and wise as thou \bibemph{art}:
\vs Gen 41:40 Thou shalt be over my house, and according unto thy word shall all my people be ruled: only in the throne will I be greater than thou.
\fn{be ruled: Heb. be armed, or, kiss.}
\vs Gen 41:41 And Pharaoh said unto Joseph, See, I have set thee over all the land of Egypt.
\vs Gen 41:42 And Pharaoh took off his ring from his hand, and put it upon Joseph's hand, and arrayed him in vestures of fine linen, and put a gold chain about his neck;
\fn{fine...: or, silk.}
\vs Gen 41:43 And he made him to ride in the second chariot which he had; and they cried before him, Bow the knee: and he made him \bibemph{ruler} over all the land of Egypt.
\fn{Bow...: or, Tender father: Heb. Abrech.}
\vs Gen 41:44 And Pharaoh said unto Joseph, I \bibemph{am} Pharaoh, and without thee shall no man lift up his hand or foot in all the land of Egypt.
\vs Gen 41:45 And Pharaoh called Joseph's name Zaphnathpaaneah; and he gave him to wife Asenath the daughter of Potipherah priest of On. And Joseph went out over \bibemph{all} the land of Egypt.
\fn{Zaphnathpaaneah: which in the Coptic signifies, A revealer of secrets, or, The man to whom secrets are revealed. priest: or, prince.}
\vs Gen 41:46 And Joseph \bibemph{was} thirty years old when he stood before Pharaoh king of Egypt. And Joseph went out from the presence of Pharaoh, and went throughout all the land of Egypt.
\vs Gen 41:47 And in the seven plenteous years the earth brought forth by handfuls.
\vs Gen 41:48 And he gathered up all the food of the seven years, which were in the land of Egypt, and laid up the food in the cities: the food of the field, which \bibemph{was} round about every city, laid he up in the same.
\vs Gen 41:49 And Joseph gathered corn as the sand of the sea, very much, until he left numbering; for \bibemph{it was} without number.
\vs Gen 41:50 And unto Joseph were born two sons before the years of famine came, which Asenath the daughter of Potipherah priest of On bare unto him.
\fn{priest: or, prince.}
\vs Gen 41:51 And Joseph called the name of the firstborn Manasseh: For God, \bibemph{said he}, hath made me forget all my toil, and all my father's house.
\fn{Manasseh: that is, Forgetting.}
\vs Gen 41:52 And the name of the second called he Ephraim: For God hath caused me to be fruitful in the land of my affliction.
\fn{Ephraim: that is, Fruitful.}
\vs Gen 41:53 And the seven years of plenteousness, that was in the land of Egypt, were ended.
\vs Gen 41:54 And the seven years of dearth began to come, according as Joseph had said: and the dearth was in all lands; but in all the land of Egypt there was bread.
\vs Gen 41:55 And when all the land of Egypt was famished, the people cried to Pharaoh for bread: and Pharaoh said unto all the Egyptians, Go unto Joseph; what he saith to you, do.
\vs Gen 41:56 And the famine was over all the face of the earth: And Joseph opened all the storehouses, and sold unto the Egyptians; and the famine waxed sore in the land of Egypt.
\fn{all the storehouses: Heb. all wherein was.}
\vs Gen 41:57 And all countries came into Egypt to Joseph for to buy \bibemph{corn}; because that the famine was \bibemph{so} sore in all lands.
\vs Gen 42:1 Now when Jacob saw that there was corn in Egypt, Jacob said unto his sons, Why do ye look one upon another?
\vs Gen 42:2 And he said, Behold, I have heard that there is corn in Egypt: get you down thither, and buy for us from thence; that we may live, and not die.
\vs Gen 42:3 And Joseph's ten brethren went down to buy corn in Egypt.
\vs Gen 42:4 But Benjamin, Joseph's brother, Jacob sent not with his brethren; for he said, Lest peradventure mischief befall him.
\vs Gen 42:5 And the sons of Israel came to buy \bibemph{corn} among those that came: for the famine was in the land of Canaan.
\vs Gen 42:6 And Joseph \bibemph{was} the governor over the land, \bibemph{and} he \bibemph{it was} that sold to all the people of the land: and Joseph's brethren came, and bowed down themselves before him \bibemph{with} their faces to the earth.
\vs Gen 42:7 And Joseph saw his brethren, and he knew them, but made himself strange unto them, and spake roughly unto them; and he said unto them, Whence come ye? And they said, From the land of Canaan to buy food.
\fn{roughly...: Heb. hard things with them.}
\vs Gen 42:8 And Joseph knew his brethren, but they knew not him.
\vs Gen 42:9 And Joseph remembered the dreams which he dreamed of them, and said unto them, Ye \bibemph{are} spies; to see the nakedness of the land ye are come.
\vs Gen 42:10 And they said unto him, Nay, my lord, but to buy food are thy servants come.
\vs Gen 42:11 We \bibemph{are} all one man's sons; we \bibemph{are} true \bibemph{men}, thy servants are no spies.
\vs Gen 42:12 And he said unto them, Nay, but to see the nakedness of the land ye are come.
\vs Gen 42:13 And they said, Thy servants \bibemph{are} twelve brethren, the sons of one man in the land of Canaan; and, behold, the youngest \bibemph{is} this day with our father, and one \bibemph{is} not.
\vs Gen 42:14 And Joseph said unto them, That \bibemph{is it} that I spake unto you, saying, Ye \bibemph{are} spies:
\vs Gen 42:15 Hereby ye shall be proved: By the life of Pharaoh ye shall not go forth hence, except your youngest brother come hither.
\vs Gen 42:16 Send one of you, and let him fetch your brother, and ye shall be kept in prison, that your words may be proved, whether \bibemph{there be any} truth in you: or else by the life of Pharaoh surely ye \bibemph{are} spies.
\fn{kept...: Heb. bound.}
\vs Gen 42:17 And he put them all together into ward three days.
\fn{put: Heb. gathered.}
\vs Gen 42:18 And Joseph said unto them the third day, This do, and live; \bibemph{for} I fear God:
\vs Gen 42:19 If ye \bibemph{be} true \bibemph{men}, let one of your brethren be bound in the house of your prison: go ye, carry corn for the famine of your houses:
\vs Gen 42:20 But bring your youngest brother unto me; so shall your words be verified, and ye shall not die. And they did so.
\vs Gen 42:21 And they said one to another, We \bibemph{are} verily guilty concerning our brother, in that we saw the anguish of his soul, when he besought us, and we would not hear; therefore is this distress come upon us.
\vs Gen 42:22 And Reuben answered them, saying, Spake I not unto you, saying, Do not sin against the child; and ye would not hear? therefore, behold, also his blood is required.
\vs Gen 42:23 And they knew not that Joseph understood \bibemph{them}; for he spake unto them by an interpreter.
\fn{he spake...: Heb. an interpreter was between them.}
\vs Gen 42:24 And he turned himself about from them, and wept; and returned to them again, and communed with them, and took from them Simeon, and bound him before their eyes.
\vs Gen 42:25 Then Joseph commanded to fill their sacks with corn, and to restore every man's money into his sack, and to give them provision for the way: and thus did he unto them.
\vs Gen 42:26 And they laded their asses with the corn, and departed thence.
\vs Gen 42:27 And as one of them opened his sack to give his ass provender in the inn, he espied his money; for, behold, it \bibemph{was} in his sack's mouth.
\vs Gen 42:28 And he said unto his brethren, My money is restored; and, lo, \bibemph{it is} even in my sack: and their heart failed \bibemph{them}, and they were afraid, saying one to another, What \bibemph{is} this \bibemph{that} God hath done unto us?
\fn{failed...: Heb. went forth.}
\vs Gen 42:29 And they came unto Jacob their father unto the land of Canaan, and told him all that befell unto them; saying,
\vs Gen 42:30 The man, \bibemph{who is} the lord of the land, spake roughly to us, and took us for spies of the country.
\fn{roughly...: Heb. with us hard things.}
\vs Gen 42:31 And we said unto him, We \bibemph{are} true \bibemph{men}; we are no spies:
\vs Gen 42:32 We \bibemph{be} twelve brethren, sons of our father; one \bibemph{is} not, and the youngest \bibemph{is} this day with our father in the land of Canaan.
\vs Gen 42:33 And the man, the lord of the country, said unto us, Hereby shall I know that ye \bibemph{are} true \bibemph{men}; leave one of your brethren \bibemph{here} with me, and take \bibemph{food for} the famine of your households, and be gone:
\vs Gen 42:34 And bring your youngest brother unto me: then shall I know that ye \bibemph{are} no spies, but \bibemph{that} ye \bibemph{are} true \bibemph{men: so} will I deliver you your brother, and ye shall traffick in the land.
\vs Gen 42:35 And it came to pass as they emptied their sacks, that, behold, every man's bundle of money \bibemph{was} in his sack: and when \bibemph{both} they and their father saw the bundles of money, they were afraid.
\vs Gen 42:36 And Jacob their father said unto them, Me have ye bereaved \bibemph{of my children}: Joseph \bibemph{is} not, and Simeon \bibemph{is} not, and ye will take Benjamin \bibemph{away}: all these things are against me.
\vs Gen 42:37 And Reuben spake unto his father, saying, Slay my two sons, if I bring him not to thee: deliver him into my hand, and I will bring him to thee again.
\vs Gen 42:38 And he said, My son shall not go down with you; for his brother is dead, and he is left alone: if mischief befall him by the way in the which ye go, then shall ye bring down my gray hairs with sorrow to the grave.
\vs Gen 43:1 And the famine \bibemph{was} sore in the land.
\vs Gen 43:2 And it came to pass, when they had eaten up the corn which they had brought out of Egypt, their father said unto them, Go again, buy us a little food.
\vs Gen 43:3 And Judah spake unto him, saying, The man did solemnly protest unto us, saying, Ye shall not see my face, except your brother \bibemph{be} with you.
\fn{did...: Heb. protesting protested.}
\vs Gen 43:4 If thou wilt send our brother with us, we will go down and buy thee food:
\vs Gen 43:5 But if thou wilt not send \bibemph{him}, we will not go down: for the man said unto us, Ye shall not see my face, except your brother \bibemph{be} with you.
\vs Gen 43:6 And Israel said, Wherefore dealt ye \bibemph{so} ill with me, \bibemph{as} to tell the man whether ye had yet a brother?
\vs Gen 43:7 And they said, The man asked us straitly of our state, and of our kindred, saying, \bibemph{Is} your father yet alive? have ye \bibemph{another} brother? and we told him according to the tenor of these words: could we certainly know that he would say, Bring your brother down?
\fn{asked...: Heb. asking asked us. tenor: Heb. mouth. could...: Heb. knowing could we know.}
\vs Gen 43:8 And Judah said unto Israel his father, Send the lad with me, and we will arise and go; that we may live, and not die, both we, and thou, \bibemph{and} also our little ones.
\vs Gen 43:9 I will be surety for him; of my hand shalt thou require him: if I bring him not unto thee, and set him before thee, then let me bear the blame for ever:
\vs Gen 43:10 For except we had lingered, surely now we had returned this second time.
\fn{this: or, twice by this.}
\vs Gen 43:11 And their father Israel said unto them, If \bibemph{it must be} so now, do this; take of the best fruits in the land in your vessels, and carry down the man a present, a little balm, and a little honey, spices, and myrrh, nuts, and almonds:
\vs Gen 43:12 And take double money in your hand; and the money that was brought again in the mouth of your sacks, carry \bibemph{it} again in your hand; peradventure it \bibemph{was} an oversight:
\vs Gen 43:13 Take also your brother, and arise, go again unto the man:
\vs Gen 43:14 And God Almighty give you mercy before the man, that he may send away your other brother, and Benjamin. If I be bereaved \bibemph{of my children}, I am bereaved.
\fn{If...: or, And I, as I have been, etc.}
\vs Gen 43:15 And the men took that present, and they took double money in their hand, and Benjamin; and rose up, and went down to Egypt, and stood before Joseph.
\vs Gen 43:16 And when Joseph saw Benjamin with them, he said to the ruler of his house, Bring \bibemph{these} men home, and slay, and make ready; for \bibemph{these} men shall dine with me at noon.
\fn{slay: Heb. kill a killing. dine: Heb. eat.}
\vs Gen 43:17 And the man did as Joseph bade; and the man brought the men into Joseph's house.
\vs Gen 43:18 And the men were afraid, because they were brought into Joseph's house; and they said, Because of the money that was returned in our sacks at the first time are we brought in; that he may seek occasion against us, and fall upon us, and take us for bondmen, and our asses.
\fn{seek...: Heb. roll himself upon us.}
\vs Gen 43:19 And they came near to the steward of Joseph's house, and they communed with him at the door of the house,
\vs Gen 43:20 And said, O sir, we came indeed down at the first time to buy food:
\fn{we...: Heb. coming down we came down.}
\vs Gen 43:21 And it came to pass, when we came to the inn, that we opened our sacks, and, behold, \bibemph{every} man's money \bibemph{was} in the mouth of his sack, our money in full weight: and we have brought it again in our hand.
\vs Gen 43:22 And other money have we brought down in our hands to buy food: we cannot tell who put our money in our sacks.
\vs Gen 43:23 And he said, Peace \bibemph{be} to you, fear not: your God, and the God of your father, hath given you treasure in your sacks: I had your money. And he brought Simeon out unto them.
\fn{I had...: Heb. your money came to me.}
\vs Gen 43:24 And the man brought the men into Joseph's house, and gave \bibemph{them} water, and they washed their feet; and he gave their asses provender.
\vs Gen 43:25 And they made ready the present against Joseph came at noon: for they heard that they should eat bread there.
\vs Gen 43:26 And when Joseph came home, they brought him the present which \bibemph{was} in their hand into the house, and bowed themselves to him to the earth.
\vs Gen 43:27 And he asked them of \bibemph{their} welfare, and said, \bibemph{Is} your father well, the old man of whom ye spake? \bibemph{Is} he yet alive?
\fn{welfare: Heb. peace. Is your...: Heb. Is there peace to your father?.}
\vs Gen 43:28 And they answered, Thy servant our father \bibemph{is} in good health, he \bibemph{is} yet alive. And they bowed down their heads, and made obeisance.
\vs Gen 43:29 And he lifted up his eyes, and saw his brother Benjamin, his mother's son, and said, \bibemph{Is} this your younger brother, of whom ye spake unto me? And he said, God be gracious unto thee, my son.
\vs Gen 43:30 And Joseph made haste; for his bowels did yearn upon his brother: and he sought \bibemph{where} to weep; and he entered into \bibemph{his} chamber, and wept there.
\vs Gen 43:31 And he washed his face, and went out, and refrained himself, and said, Set on bread.
\vs Gen 43:32 And they set on for him by himself, and for them by themselves, and for the Egyptians, which did eat with him, by themselves: because the Egyptians might not eat bread with the Hebrews; for that \bibemph{is} an abomination unto the Egyptians.
\vs Gen 43:33 And they sat before him, the firstborn according to his birthright, and the youngest according to his youth: and the men marvelled one at another.
\vs Gen 43:34 And he took \bibemph{and sent} messes unto them from before him: but Benjamin's mess was five times so much as any of theirs. And they drank, and were merry with him.
\fn{were...: Heb. drank largely.}
\vs Gen 44:1 And he commanded the steward of his house, saying, Fill the men's sacks \bibemph{with} food, as much as they can carry, and put every man's money in his sack's mouth.
\fn{the steward...: Heb. him that was over his house.}
\vs Gen 44:2 And put my cup, the silver cup, in the sack's mouth of the youngest, and his corn money. And he did according to the word that Joseph had spoken.
\vs Gen 44:3 As soon as the morning was light, the men were sent away, they and their asses.
\vs Gen 44:4 \bibemph{And} when they were gone out of the city, \bibemph{and} not \bibemph{yet} far off, Joseph said unto his steward, Up, follow after the men; and when thou dost overtake them, say unto them, Wherefore have ye rewarded evil for good?
\vs Gen 44:5 \bibemph{Is} not this \bibemph{it} in which my lord drinketh, and whereby indeed he divineth? ye have done evil in so doing.
\fn{divineth: or, maketh trial?.}
\vs Gen 44:6 And he overtook them, and he spake unto them these same words.
\vs Gen 44:7 And they said unto him, Wherefore saith my lord these words? God forbid that thy servants should do according to this thing:
\vs Gen 44:8 Behold, the money, which we found in our sacks' mouths, we brought again unto thee out of the land of Canaan: how then should we steal out of thy lord's house silver or gold?
\vs Gen 44:9 With whomsoever of thy servants it be found, both let him die, and we also will be my lord's bondmen.
\vs Gen 44:10 And he said, Now also \bibemph{let} it \bibemph{be} according unto your words: he with whom it is found shall be my servant; and ye shall be blameless.
\vs Gen 44:11 Then they speedily took down every man his sack to the ground, and opened every man his sack.
\vs Gen 44:12 And he searched, \bibemph{and} began at the eldest, and left at the youngest: and the cup was found in Benjamin's sack.
\vs Gen 44:13 Then they rent their clothes, and laded every man his ass, and returned to the city.
\vs Gen 44:14 And Judah and his brethren came to Joseph's house; for he \bibemph{was} yet there: and they fell before him on the ground.
\vs Gen 44:15 And Joseph said unto them, What deed \bibemph{is} this that ye have done? wot ye not that such a man as I can certainly divine?
\fn{divine: or, make trial?.}
\vs Gen 44:16 And Judah said, What shall we say unto my lord? what shall we speak? or how shall we clear ourselves? God hath found out the iniquity of thy servants: behold, we \bibemph{are} my lord's servants, both we, and \bibemph{he} also with whom the cup is found.
\vs Gen 44:17 And he said, God forbid that I should do so: \bibemph{but} the man in whose hand the cup is found, he shall be my servant; and as for you, get you up in peace unto your father.
\vs Gen 44:18 Then Judah came near unto him, and said, Oh my lord, let thy servant, I pray thee, speak a word in my lord's ears, and let not thine anger burn against thy servant: for thou \bibemph{art} even as Pharaoh.
\vs Gen 44:19 My lord asked his servants, saying, Have ye a father, or a brother?
\vs Gen 44:20 And we said unto my lord, We have a father, an old man, and a child of his old age, a little one; and his brother is dead, and he alone is left of his mother, and his father loveth him.
\vs Gen 44:21 And thou saidst unto thy servants, Bring him down unto me, that I may set mine eyes upon him.
\vs Gen 44:22 And we said unto my lord, The lad cannot leave his father: for \bibemph{if} he should leave his father, \bibemph{his father} would die.
\vs Gen 44:23 And thou saidst unto thy servants, Except your youngest brother come down with you, ye shall see my face no more.
\vs Gen 44:24 And it came to pass when we came up unto thy servant my father, we told him the words of my lord.
\vs Gen 44:25 And our father said, Go again, \bibemph{and} buy us a little food.
\vs Gen 44:26 And we said, We cannot go down: if our youngest brother be with us, then will we go down: for we may not see the man's face, except our youngest brother \bibemph{be} with us.
\vs Gen 44:27 And thy servant my father said unto us, Ye know that my wife bare me two \bibemph{sons}:
\vs Gen 44:28 And the one went out from me, and I said, Surely he is torn in pieces; and I saw him not since:
\vs Gen 44:29 And if ye take this also from me, and mischief befall him, ye shall bring down my gray hairs with sorrow to the grave.
\vs Gen 44:30 Now therefore when I come to thy servant my father, and the lad \bibemph{be} not with us; seeing that his life is bound up in the lad's life;
\vs Gen 44:31 It shall come to pass, when he seeth that the lad \bibemph{is} not \bibemph{with us}, that he will die: and thy servants shall bring down the gray hairs of thy servant our father with sorrow to the grave.
\vs Gen 44:32 For thy servant became surety for the lad unto my father, saying, If I bring him not unto thee, then I shall bear the blame to my father for ever.
\vs Gen 44:33 Now therefore, I pray thee, let thy servant abide instead of the lad a bondman to my lord; and let the lad go up with his brethren.
\vs Gen 44:34 For how shall I go up to my father, and the lad \bibemph{be} not with me? lest peradventure I see the evil that shall come on my father.
\fn{come...: Heb. find my father.}
\vs Gen 45:1 Then Joseph could not refrain himself before all them that stood by him; and he cried, Cause every man to go out from me. And there stood no man with him, while Joseph made himself known unto his brethren.
\vs Gen 45:2 And he wept aloud: and the Egyptians and the house of Pharaoh heard.
\fn{wept...: Heb. gave forth his voice in weeping.}
\vs Gen 45:3 And Joseph said unto his brethren, I \bibemph{am} Joseph; doth my father yet live? And his brethren could not answer him; for they were troubled at his presence.
\fn{troubled: or, terrified.}
\vs Gen 45:4 And Joseph said unto his brethren, Come near to me, I pray you. And they came near. And he said, I \bibemph{am} Joseph your brother, whom ye sold into Egypt.
\vs Gen 45:5 Now therefore be not grieved, nor angry with yourselves, that ye sold me hither: for God did send me before you to preserve life.
\fn{nor...: Heb. neither let there be anger in your eyes.}
\vs Gen 45:6 For these two years \bibemph{hath} the famine \bibemph{been} in the land: and yet \bibemph{there are} five years, in the which \bibemph{there shall} neither \bibemph{be} earing nor harvest.
\vs Gen 45:7 And God sent me before you to preserve you a posterity in the earth, and to save your lives by a great deliverance.
\fn{to preserve...: Heb. to put for you a remnant.}
\vs Gen 45:8 So now \bibemph{it was} not you \bibemph{that} sent me hither, but God: and he hath made me a father to Pharaoh, and lord of all his house, and a ruler throughout all the land of Egypt.
\vs Gen 45:9 Haste ye, and go up to my father, and say unto him, Thus saith thy son Joseph, God hath made me lord of all Egypt: come down unto me, tarry not:
\vs Gen 45:10 And thou shalt dwell in the land of Goshen, and thou shalt be near unto me, thou, and thy children, and thy children's children, and thy flocks, and thy herds, and all that thou hast:
\vs Gen 45:11 And there will I nourish thee; for yet \bibemph{there are} five years of famine; lest thou, and thy household, and all that thou hast, come to poverty.
\vs Gen 45:12 And, behold, your eyes see, and the eyes of my brother Benjamin, that \bibemph{it is} my mouth that speaketh unto you.
\vs Gen 45:13 And ye shall tell my father of all my glory in Egypt, and of all that ye have seen; and ye shall haste and bring down my father hither.
\vs Gen 45:14 And he fell upon his brother Benjamin's neck, and wept; and Benjamin wept upon his neck.
\vs Gen 45:15 Moreover he kissed all his brethren, and wept upon them: and after that his brethren talked with him.
\vs Gen 45:16 And the fame thereof was heard in Pharaoh's house, saying, Joseph's brethren are come: and it pleased Pharaoh well, and his servants.
\fn{pleased...: Heb. was good in the eyes of Pharaoh.}
\vs Gen 45:17 And Pharaoh said unto Joseph, Say unto thy brethren, This do ye; lade your beasts, and go, get you unto the land of Canaan;
\vs Gen 45:18 And take your father and your households, and come unto me: and I will give you the good of the land of Egypt, and ye shall eat the fat of the land.
\vs Gen 45:19 Now thou art commanded, this do ye; take you wagons out of the land of Egypt for your little ones, and for your wives, and bring your father, and come.
\vs Gen 45:20 Also regard not your stuff; for the good of all the land of Egypt \bibemph{is} yours.
\fn{regard...: Heb. let not your eye spare, etc.}
\vs Gen 45:21 And the children of Israel did so: and Joseph gave them wagons, according to the commandment of Pharaoh, and gave them provision for the way.
\fn{commandment: Heb. mouth.}
\vs Gen 45:22 To all of them he gave each man changes of raiment; but to Benjamin he gave three hundred \bibemph{pieces} of silver, and five changes of raiment.
\vs Gen 45:23 And to his father he sent after this \bibemph{manner}; ten asses laden with the good things of Egypt, and ten she asses laden with corn and bread and meat for his father by the way.
\fn{laden...: Heb. carrying.}
\vs Gen 45:24 So he sent his brethren away, and they departed: and he said unto them, See that ye fall not out by the way.
\vs Gen 45:25 And they went up out of Egypt, and came into the land of Canaan unto Jacob their father,
\vs Gen 45:26 And told him, saying, Joseph \bibemph{is} yet alive, and he \bibemph{is} governor over all the land of Egypt. And Jacob's heart fainted, for he believed them not.
\fn{Jacob's: Heb. his.}
\vs Gen 45:27 And they told him all the words of Joseph, which he had said unto them: and when he saw the wagons which Joseph had sent to carry him, the spirit of Jacob their father revived:
\vs Gen 45:28 And Israel said, \bibemph{It is} enough; Joseph my son \bibemph{is} yet alive: I will go and see him before I die.
\vs Gen 46:1 And Israel took his journey with all that he had, and came to Beersheba, and offered sacrifices unto the God of his father Isaac.
\vs Gen 46:2 And God spake unto Israel in the visions of the night, and said, Jacob, Jacob. And he said, Here \bibemph{am} I.
\vs Gen 46:3 And he said, I \bibemph{am} God, the God of thy father: fear not to go down into Egypt; for I will there make of thee a great nation:
\vs Gen 46:4 I will go down with thee into Egypt; and I will also surely bring thee up \bibemph{again}: and Joseph shall put his hand upon thine eyes.
\vs Gen 46:5 And Jacob rose up from Beersheba: and the sons of Israel carried Jacob their father, and their little ones, and their wives, in the wagons which Pharaoh had sent to carry him.
\vs Gen 46:6 And they took their cattle, and their goods, which they had gotten in the land of Canaan, and came into Egypt, Jacob, and all his seed with him:
\vs Gen 46:7 His sons, and his sons' sons with him, his daughters, and his sons' daughters, and all his seed brought he with him into Egypt.
\vs Gen 46:8 And these \bibemph{are} the names of the children of Israel, which came into Egypt, Jacob and his sons: Reuben, Jacob's firstborn.
\vs Gen 46:9 And the sons of Reuben; Hanoch, and Phallu, and Hezron, and Carmi.
\vs Gen 46:10 And the sons of Simeon; Jemuel, and Jamin, and Ohad, and Jachin, and Zohar, and Shaul the son of a Canaanitish woman.
\fn{Jemuel: or, Nemuel. Jachin: or, Jarib. Zohar: or, Zerah.}
\vs Gen 46:11 And the sons of Levi; Gershon, Kohath, and Merari.
\fn{Gershon: or, Gershom.}
\vs Gen 46:12 And the sons of Judah; Er, and Onan, and Shelah, and Pharez, and Zerah: but Er and Onan died in the land of Canaan. And the sons of Pharez were Hezron and Hamul.
\vs Gen 46:13 And the sons of Issachar; Tola, and Phuvah, and Job, and Shimron.
\fn{Phuvah, and Job: or, Puah, and Jashub.}
\vs Gen 46:14 And the sons of Zebulun; Sered, and Elon, and Jahleel.
\vs Gen 46:15 These \bibemph{be} the sons of Leah, which she bare unto Jacob in Padanaram, with his daughter Dinah: all the souls of his sons and his daughters \bibemph{were} thirty and three.
\vs Gen 46:16 And the sons of Gad; Ziphion, and Haggi, Shuni, and Ezbon, Eri, and Arodi, and Areli.
\fn{Ziphion: or, Zephon. Ezbon: or, Ozni. Arodi: or, Arod.}
\vs Gen 46:17 And the sons of Asher; Jimnah, and Ishuah, and Isui, and Beriah, and Serah their sister: and the sons of Beriah; Heber, and Malchiel.
\vs Gen 46:18 These \bibemph{are} the sons of Zilpah, whom Laban gave to Leah his daughter, and these she bare unto Jacob, \bibemph{even} sixteen souls.
\vs Gen 46:19 The sons of Rachel Jacob's wife; Joseph, and Benjamin.
\vs Gen 46:20 And unto Joseph in the land of Egypt were born Manasseh and Ephraim, which Asenath the daughter of Potipherah priest of On bare unto him.
\fn{priest: or, prince.}
\vs Gen 46:21 And the sons of Benjamin \bibemph{were} Belah, and Becher, and Ashbel, Gera, and Naaman, Ehi, and Rosh, Muppim, and Huppim, and Ard.
\fn{Ehi: or, Ahiram. Muppim: or, Shupham or, Shuppim. Huppim: or, Hupham.}
\vs Gen 46:22 These \bibemph{are} the sons of Rachel, which were born to Jacob: all the souls \bibemph{were} fourteen.
\vs Gen 46:23 And the sons of Dan; Hushim.
\fn{Hushim: or, Shuham.}
\vs Gen 46:24 And the sons of Naphtali; Jahzeel, and Guni, and Jezer, and Shillem.
\vs Gen 46:25 These \bibemph{are} the sons of Bilhah, which Laban gave unto Rachel his daughter, and she bare these unto Jacob: all the souls \bibemph{were} seven.
\vs Gen 46:26 All the souls that came with Jacob into Egypt, which came out of his loins, besides Jacob's sons' wives, all the souls \bibemph{were} threescore and six;
\fn{loins: Heb. thigh.}
\vs Gen 46:27 And the sons of Joseph, which were born him in Egypt, \bibemph{were} two souls: all the souls of the house of Jacob, which came into Egypt, \bibemph{were} threescore and ten.
\vs Gen 46:28 And he sent Judah before him unto Joseph, to direct his face unto Goshen; and they came into the land of Goshen.
\vs Gen 46:29 And Joseph made ready his chariot, and went up to meet Israel his father, to Goshen, and presented himself unto him; and he fell on his neck, and wept on his neck a good while.
\vs Gen 46:30 And Israel said unto Joseph, Now let me die, since I have seen thy face, because thou \bibemph{art} yet alive.
\vs Gen 46:31 And Joseph said unto his brethren, and unto his father's house, I will go up, and shew Pharaoh, and say unto him, My brethren, and my father's house, which \bibemph{were} in the land of Canaan, are come unto me;
\vs Gen 46:32 And the men \bibemph{are} shepherds, for their trade hath been to feed cattle; and they have brought their flocks, and their herds, and all that they have.
\fn{their trade...: Heb. they are men of cattle.}
\vs Gen 46:33 And it shall come to pass, when Pharaoh shall call you, and shall say, What \bibemph{is} your occupation?
\vs Gen 46:34 That ye shall say, Thy servants' trade hath been about cattle from our youth even until now, both we, \bibemph{and} also our fathers: that ye may dwell in the land of Goshen; for every shepherd \bibemph{is} an abomination unto the Egyptians.
\vs Gen 47:1 Then Joseph came and told Pharaoh, and said, My father and my brethren, and their flocks, and their herds, and all that they have, are come out of the land of Canaan; and, behold, they \bibemph{are} in the land of Goshen.
\vs Gen 47:2 And he took some of his brethren, \bibemph{even} five men, and presented them unto Pharaoh.
\vs Gen 47:3 And Pharaoh said unto his brethren, What \bibemph{is} your occupation? And they said unto Pharaoh, Thy servants \bibemph{are} shepherds, both we, \bibemph{and} also our fathers.
\vs Gen 47:4 They said moreover unto Pharaoh, For to sojourn in the land are we come; for thy servants have no pasture for their flocks; for the famine \bibemph{is} sore in the land of Canaan: now therefore, we pray thee, let thy servants dwell in the land of Goshen.
\vs Gen 47:5 And Pharaoh spake unto Joseph, saying, Thy father and thy brethren are come unto thee:
\vs Gen 47:6 The land of Egypt \bibemph{is} before thee; in the best of the land make thy father and brethren to dwell; in the land of Goshen let them dwell: and if thou knowest \bibemph{any} men of activity among them, then make them rulers over my cattle.
\vs Gen 47:7 And Joseph brought in Jacob his father, and set him before Pharaoh: and Jacob blessed Pharaoh.
\vs Gen 47:8 And Pharaoh said unto Jacob, How old \bibemph{art} thou?
\fn{How...: Heb. How many are the days of the years of thy life?.}
\vs Gen 47:9 And Jacob said unto Pharaoh, The days of the years of my pilgrimage \bibemph{are} an hundred and thirty years: few and evil have the days of the years of my life been, and have not attained unto the days of the years of the life of my fathers in the days of their pilgrimage.
\vs Gen 47:10 And Jacob blessed Pharaoh, and went out from before Pharaoh.
\vs Gen 47:11 And Joseph placed his father and his brethren, and gave them a possession in the land of Egypt, in the best of the land, in the land of Rameses, as Pharaoh had commanded.
\vs Gen 47:12 And Joseph nourished his father, and his brethren, and all his father's household, with bread, according to \bibemph{their} families.
\fn{according...: or, as a little child is nourished: Heb. according to the little ones.}
\vs Gen 47:13 And \bibemph{there was} no bread in all the land; for the famine \bibemph{was} very sore, so that the land of Egypt and \bibemph{all} the land of Canaan fainted by reason of the famine.
\vs Gen 47:14 And Joseph gathered up all the money that was found in the land of Egypt, and in the land of Canaan, for the corn which they bought: and Joseph brought the money into Pharaoh's house.
\vs Gen 47:15 And when money failed in the land of Egypt, and in the land of Canaan, all the Egyptians came unto Joseph, and said, Give us bread: for why should we die in thy presence? for the money faileth.
\vs Gen 47:16 And Joseph said, Give your cattle; and I will give you for your cattle, if money fail.
\vs Gen 47:17 And they brought their cattle unto Joseph: and Joseph gave them bread \bibemph{in exchange} for horses, and for the flocks, and for the cattle of the herds, and for the asses: and he fed them with bread for all their cattle for that year.
\fn{fed...: Heb. led them.}
\vs Gen 47:18 When that year was ended, they came unto him the second year, and said unto him, We will not hide \bibemph{it} from my lord, how that our money is spent; my lord also hath our herds of cattle; there is not ought left in the sight of my lord, but our bodies, and our lands:
\vs Gen 47:19 Wherefore shall we die before thine eyes, both we and our land? buy us and our land for bread, and we and our land will be servants unto Pharaoh: and give \bibemph{us} seed, that we may live, and not die, that the land be not desolate.
\vs Gen 47:20 And Joseph bought all the land of Egypt for Pharaoh; for the Egyptians sold every man his field, because the famine prevailed over them: so the land became Pharaoh's.
\vs Gen 47:21 And as for the people, he removed them to cities from \bibemph{one} end of the borders of Egypt even to the \bibemph{other} end thereof.
\vs Gen 47:22 Only the land of the priests bought he not; for the priests had a portion \bibemph{assigned them} of Pharaoh, and did eat their portion which Pharaoh gave them: wherefore they sold not their lands.
\fn{priests: or, princes.}
\vs Gen 47:23 Then Joseph said unto the people, Behold, I have bought you this day and your land for Pharaoh: lo, \bibemph{here is} seed for you, and ye shall sow the land.
\vs Gen 47:24 And it shall come to pass in the increase, that ye shall give the fifth \bibemph{part} unto Pharaoh, and four parts shall be your own, for seed of the field, and for your food, and for them of your households, and for food for your little ones.
\vs Gen 47:25 And they said, Thou hast saved our lives: let us find grace in the sight of my lord, and we will be Pharaoh's servants.
\vs Gen 47:26 And Joseph made it a law over the land of Egypt unto this day, \bibemph{that} Pharaoh should have the fifth \bibemph{part}; except the land of the priests only, \bibemph{which} became not Pharaoh's.
\fn{priests: or, princes.}
\vs Gen 47:27 And Israel dwelt in the land of Egypt, in the country of Goshen; and they had possessions therein, and grew, and multiplied exceedingly.
\vs Gen 47:28 And Jacob lived in the land of Egypt seventeen years: so the whole age of Jacob was an hundred forty and seven years.
\fn{the whole...: Heb. the days of the years of his life.}
\vs Gen 47:29 And the time drew nigh that Israel must die: and he called his son Joseph, and said unto him, If now I have found grace in thy sight, put, I pray thee, thy hand under my thigh, and deal kindly and truly with me; bury me not, I pray thee, in Egypt:
\vs Gen 47:30 But I will lie with my fathers, and thou shalt carry me out of Egypt, and bury me in their buryingplace. And he said, I will do as thou hast said.
\vs Gen 47:31 And he said, Swear unto me. And he sware unto him. And Israel bowed himself upon the bed's head.
\vs Gen 48:1 And it came to pass after these things, that \bibemph{one} told Joseph, Behold, thy father \bibemph{is} sick: and he took with him his two sons, Manasseh and Ephraim.
\vs Gen 48:2 And \bibemph{one} told Jacob, and said, Behold, thy son Joseph cometh unto thee: and Israel strengthened himself, and sat upon the bed.
\vs Gen 48:3 And Jacob said unto Joseph, God Almighty appeared unto me at Luz in the land of Canaan, and blessed me,
\vs Gen 48:4 And said unto me, Behold, I will make thee fruitful, and multiply thee, and I will make of thee a multitude of people; and will give this land to thy seed after thee \bibemph{for} an everlasting possession.
\vs Gen 48:5 And now thy two sons, Ephraim and Manasseh, which were born unto thee in the land of Egypt before I came unto thee into Egypt, \bibemph{are} mine; as Reuben and Simeon, they shall be mine.
\vs Gen 48:6 And thy issue, which thou begettest after them, shall be thine, \bibemph{and} shall be called after the name of their brethren in their inheritance.
\vs Gen 48:7 And as for me, when I came from Padan, Rachel died by me in the land of Canaan in the way, when yet \bibemph{there was} but a little way to come unto Ephrath: and I buried her there in the way of Ephrath; the same \bibemph{is} Bethlehem.
\vs Gen 48:8 And Israel beheld Joseph's sons, and said, Who \bibemph{are} these?
\vs Gen 48:9 And Joseph said unto his father, They \bibemph{are} my sons, whom God hath given me in this \bibemph{place}. And he said, Bring them, I pray thee, unto me, and I will bless them.
\vs Gen 48:10 Now the eyes of Israel were dim for age, \bibemph{so that} he could not see. And he brought them near unto him; and he kissed them, and embraced them.
\fn{dim: Heb. heavy.}
\vs Gen 48:11 And Israel said unto Joseph, I had not thought to see thy face: and, lo, God hath shewed me also thy seed.
\vs Gen 48:12 And Joseph brought them out from between his knees, and he bowed himself with his face to the earth.
\vs Gen 48:13 And Joseph took them both, Ephraim in his right hand toward Israel's left hand, and Manasseh in his left hand toward Israel's right hand, and brought \bibemph{them} near unto him.
\vs Gen 48:14 And Israel stretched out his right hand, and laid \bibemph{it} upon Ephraim's head, who \bibemph{was} the younger, and his left hand upon Manasseh's head, guiding his hands wittingly; for Manasseh \bibemph{was} the firstborn.
\vs Gen 48:15 And he blessed Joseph, and said, God, before whom my fathers Abraham and Isaac did walk, the God which fed me all my life long unto this day,
\vs Gen 48:16 The Angel which redeemed me from all evil, bless the lads; and let my name be named on them, and the name of my fathers Abraham and Isaac; and let them grow into a multitude in the midst of the earth.
\fn{grow: Heb. as fishes do increase.}
\vs Gen 48:17 And when Joseph saw that his father laid his right hand upon the head of Ephraim, it displeased him: and he held up his father's hand, to remove it from Ephraim's head unto Manasseh's head.
\fn{displeased...: was evil in his eyes.}
\vs Gen 48:18 And Joseph said unto his father, Not so, my father: for this \bibemph{is} the firstborn; put thy right hand upon his head.
\vs Gen 48:19 And his father refused, and said, I know \bibemph{it}, my son, I know \bibemph{it}: he also shall become a people, and he also shall be great: but truly his younger brother shall be greater than he, and his seed shall become a multitude of nations.
\fn{multitude: Heb. fulness.}
\vs Gen 48:20 And he blessed them that day, saying, In thee shall Israel bless, saying, God make thee as Ephraim and as Manasseh: and he set Ephraim before Manasseh.
\vs Gen 48:21 And Israel said unto Joseph, Behold, I die: but God shall be with you, and bring you again unto the land of your fathers.
\vs Gen 48:22 Moreover I have given to thee one portion above thy brethren, which I took out of the hand of the Amorite with my sword and with my bow.
\vs Gen 49:1 And Jacob called unto his sons, and said, Gather yourselves together, that I may tell you \bibemph{that} which shall befall you in the last days.
\vs Gen 49:2 Gather yourselves together, and hear, ye sons of Jacob; and hearken unto Israel your father.
\vs Gen 49:3 Reuben, thou \bibemph{art} my firstborn, my might, and the beginning of my strength, the excellency of dignity, and the excellency of power:
\vs Gen 49:4 Unstable as water, thou shalt not excel; because thou wentest up to thy father's bed; then defiledst thou \bibemph{it}: he went up to my couch.
\fn{thou shalt...: Heb. do not thou excel. he went...: or, my couch is gone.}
\vs Gen 49:5 Simeon and Levi \bibemph{are} brethren; instruments of cruelty \bibemph{are in} their habitations.
\fn{instruments...: or, their swords are weapons of violence.}
\vs Gen 49:6 O my soul, come not thou into their secret; unto their assembly, mine honour, be not thou united: for in their anger they slew a man, and in their selfwill they digged down a wall.
\fn{digged...: or, lamed oxen.}
\vs Gen 49:7 Cursed \bibemph{be} their anger, for \bibemph{it was} fierce; and their wrath, for it was cruel: I will divide them in Jacob, and scatter them in Israel.
\vs Gen 49:8 Judah, thou \bibemph{art he} whom thy brethren shall praise: thy hand \bibemph{shall be} in the neck of thine enemies; thy father's children shall bow down before thee.
\vs Gen 49:9 Judah \bibemph{is} a lion's whelp: from the prey, my son, thou art gone up: he stooped down, he couched as a lion, and as an old lion; who shall rouse him up?
\vs Gen 49:10 The sceptre shall not depart from Judah, nor a lawgiver from between his feet, until Shiloh come; and unto him \bibemph{shall} the gathering of the people \bibemph{be}.
\vs Gen 49:11 Binding his foal unto the vine, and his ass's colt unto the choice vine; he washed his garments in wine, and his clothes in the blood of grapes:
\vs Gen 49:12 His eyes \bibemph{shall be} red with wine, and his teeth white with milk.
\vs Gen 49:13 Zebulun shall dwell at the haven of the sea; and he \bibemph{shall be} for an haven of ships; and his border \bibemph{shall be} unto Zidon.
\vs Gen 49:14 Issachar \bibemph{is} a strong ass couching down between two burdens:
\vs Gen 49:15 And he saw that rest \bibemph{was} good, and the land that \bibemph{it was} pleasant; and bowed his shoulder to bear, and became a servant unto tribute.
\vs Gen 49:16 Dan shall judge his people, as one of the tribes of Israel.
\vs Gen 49:17 Dan shall be a serpent by the way, an adder in the path, that biteth the horse heels, so that his rider shall fall backward.
\fn{an adder: Heb. an arrow snake.}
\vs Gen 49:18 I have waited for thy salvation, O \textsc{Lord}.
\vs Gen 49:19 Gad, a troop shall overcome him: but he shall overcome at the last.
\vs Gen 49:20 Out of Asher his bread \bibemph{shall be} fat, and he shall yield royal dainties.
\vs Gen 49:21 Naphtali \bibemph{is} a hind let loose: he giveth goodly words.
\vs Gen 49:22 Joseph \bibemph{is} a fruitful bough, \bibemph{even} a fruitful bough by a well; \bibemph{whose} branches run over the wall:
\fn{branches: Heb. daughters.}
\vs Gen 49:23 The archers have sorely grieved him, and shot \bibemph{at him}, and hated him:
\vs Gen 49:24 But his bow abode in strength, and the arms of his hands were made strong by the hands of the mighty \bibemph{God} of Jacob; (from thence \bibemph{is} the shepherd, the stone of Israel:)
\vs Gen 49:25 \bibemph{Even} by the God of thy father, who shall help thee; and by the Almighty, who shall bless thee with blessings of heaven above, blessings of the deep that lieth under, blessings of the breasts, and of the womb:
\vs Gen 49:26 The blessings of thy father have prevailed above the blessings of my progenitors unto the utmost bound of the everlasting hills: they shall be on the head of Joseph, and on the crown of the head of him that was separate from his brethren.
\vs Gen 49:27 Benjamin shall ravin \bibemph{as} a wolf: in the morning he shall devour the prey, and at night he shall divide the spoil.
\vs Gen 49:28 All these \bibemph{are} the twelve tribes of Israel: and this \bibemph{is it} that their father spake unto them, and blessed them; every one according to his blessing he blessed them.
\vs Gen 49:29 And he charged them, and said unto them, I am to be gathered unto my people: bury me with my fathers in the cave that \bibemph{is} in the field of Ephron the Hittite,
\vs Gen 49:30 In the cave that \bibemph{is} in the field of Machpelah, which \bibemph{is} before Mamre, in the land of Canaan, which Abraham bought with the field of Ephron the Hittite for a possession of a buryingplace.
\vs Gen 49:31 There they buried Abraham and Sarah his wife; there they buried Isaac and Rebekah his wife; and there I buried Leah.
\vs Gen 49:32 The purchase of the field and of the cave that \bibemph{is} therein \bibemph{was} from the children of Heth.
\vs Gen 49:33 And when Jacob had made an end of commanding his sons, he gathered up his feet into the bed, and yielded up the ghost, and was gathered unto his people.
\vs Gen 50:1 And Joseph fell upon his father's face, and wept upon him, and kissed him.
\vs Gen 50:2 And Joseph commanded his servants the physicians to embalm his father: and the physicians embalmed Israel.
\vs Gen 50:3 And forty days were fulfilled for him; for so are fulfilled the days of those which are embalmed: and the Egyptians mourned for him threescore and ten days.
\fn{mourned: Heb. wept.}
\vs Gen 50:4 And when the days of his mourning were past, Joseph spake unto the house of Pharaoh, saying, If now I have found grace in your eyes, speak, I pray you, in the ears of Pharaoh, saying,
\vs Gen 50:5 My father made me swear, saying, Lo, I die: in my grave which I have digged for me in the land of Canaan, there shalt thou bury me. Now therefore let me go up, I pray thee, and bury my father, and I will come again.
\vs Gen 50:6 And Pharaoh said, Go up, and bury thy father, according as he made thee swear.
\vs Gen 50:7 And Joseph went up to bury his father: and with him went up all the servants of Pharaoh, the elders of his house, and all the elders of the land of Egypt,
\vs Gen 50:8 And all the house of Joseph, and his brethren, and his father's house: only their little ones, and their flocks, and their herds, they left in the land of Goshen.
\vs Gen 50:9 And there went up with him both chariots and horsemen: and it was a very great company.
\vs Gen 50:10 And they came to the threshingfloor of Atad, which \bibemph{is} beyond Jordan, and there they mourned with a great and very sore lamentation: and he made a mourning for his father seven days.
\vs Gen 50:11 And when the inhabitants of the land, the Canaanites, saw the mourning in the floor of Atad, they said, This \bibemph{is} a grievous mourning to the Egyptians: wherefore the name of it was called Abelmizraim, which \bibemph{is} beyond Jordan.
\fn{Abelmizraim: that is, The mourning of the Egyptians.}
\vs Gen 50:12 And his sons did unto him according as he commanded them:
\vs Gen 50:13 For his sons carried him into the land of Canaan, and buried him in the cave of the field of Machpelah, which Abraham bought with the field for a possession of a buryingplace of Ephron the Hittite, before Mamre.
\vs Gen 50:14 And Joseph returned into Egypt, he, and his brethren, and all that went up with him to bury his father, after he had buried his father.
\vs Gen 50:15 And when Joseph's brethren saw that their father was dead, they said, Joseph will peradventure hate us, and will certainly requite us all the evil which we did unto him.
\vs Gen 50:16 And they sent a messenger unto Joseph, saying, Thy father did command before he died, saying,
\fn{sent: Heb. charged.}
\vs Gen 50:17 So shall ye say unto Joseph, Forgive, I pray thee now, the trespass of thy brethren, and their sin; for they did unto thee evil: and now, we pray thee, forgive the trespass of the servants of the God of thy father. And Joseph wept when they spake unto him.
\vs Gen 50:18 And his brethren also went and fell down before his face; and they said, Behold, we \bibemph{be} thy servants.
\vs Gen 50:19 And Joseph said unto them, Fear not: for \bibemph{am} I in the place of God?
\vs Gen 50:20 But as for you, ye thought evil against me; \bibemph{but} God meant it unto good, to bring to pass, as \bibemph{it is} this day, to save much people alive.
\vs Gen 50:21 Now therefore fear ye not: I will nourish you, and your little ones. And he comforted them, and spake kindly unto them.
\fn{kindly...: Heb. to their hearts.}
\vs Gen 50:22 And Joseph dwelt in Egypt, he, and his father's house: and Joseph lived an hundred and ten years.
\vs Gen 50:23 And Joseph saw Ephraim's children of the third \bibemph{generation}: the children also of Machir the son of Manasseh were brought up upon Joseph's knees.
\fn{brought...: Heb. born.}
\vs Gen 50:24 And Joseph said unto his brethren, I die: and God will surely visit you, and bring you out of this land unto the land which he sware to Abraham, to Isaac, and to Jacob.
\vs Gen 50:25 And Joseph took an oath of the children of Israel, saying, God will surely visit you, and ye shall carry up my bones from hence.
\vs Gen 50:26 So Joseph died, \bibemph{being} an hundred and ten years old: and they embalmed him, and he was put in a coffin in Egypt.
