\bibbook{\bibtitlemain{habakkuk}}{habakkuk}{Habakkuk}{Hab}
\vs Hab 1:1 The burden which Habakkuk the prophet did see.
\vs Hab 1:2 O \textsc{Lord}, how long shall I cry, and thou wilt not hear! \bibemph{even} cry out unto thee \bibemph{of} violence, and thou wilt not save!
\vs Hab 1:3 Why dost thou shew me iniquity, and cause \bibemph{me} to behold grievance? for spoiling and violence \bibemph{are} before me: and there are \bibemph{that} raise up strife and contention.
\vs Hab 1:4 Therefore the law is slacked, and judgment doth never go forth: for the wicked doth compass about the righteous; therefore wrong judgment proceedeth.
\fn{wrong: or, wrested.}
\vs Hab 1:5 Behold ye among the heathen, and regard, and wonder marvellously: for \bibemph{I} will work a work in your days, \bibemph{which} ye will not believe, though it be told \bibemph{you}.
\vs Hab 1:6 For, lo, I raise up the Chaldeans, \bibemph{that} bitter and hasty nation, which shall march through the breadth of the land, to possess the dwellingplaces \bibemph{that are} not theirs.
\fn{breadth: Heb. breadths.}
\vs Hab 1:7 They \bibemph{are} terrible and dreadful: their judgment and their dignity shall proceed of themselves.
\fn{their judgment...: or, from them shall proceed the judgment of these, and the captivity of these.}
\vs Hab 1:8 Their horses also are swifter than the leopards, and are more fierce than the evening wolves: and their horsemen shall spread themselves, and their horsemen shall come from far; they shall fly as the eagle \bibemph{that} hasteth to eat.
\fn{fierce: Heb. sharp.}
\vs Hab 1:9 They shall come all for violence: their faces shall sup up \bibemph{as} the east wind, and they shall gather the captivity as the sand.
\fn{their...: or, the supping up of their faces, etc: or, their faces shall look (Heb. the opposition of their faces) toward the east.}
\vs Hab 1:10 And they shall scoff at the kings, and the princes shall be a scorn unto them: they shall deride every strong hold; for they shall heap dust, and take it.
\vs Hab 1:11 Then shall \bibemph{his} mind change, and he shall pass over, and offend, \bibemph{imputing} this his power unto his god.
\vs Hab 1:12 \bibemph{Art} thou not from everlasting, O \textsc{Lord} my God, mine Holy One? we shall not die. O \textsc{Lord}, thou hast ordained them for judgment; and, O mighty God, thou hast established them for correction.
\fn{mighty...: Heb. rock established: Heb. founded.}
\vs Hab 1:13 \bibemph{Thou art} of purer eyes than to behold evil, and canst not look on iniquity: wherefore lookest thou upon them that deal treacherously, \bibemph{and} holdest thy tongue when the wicked devoureth \bibemph{the man that is} more righteous than he?
\fn{iniquity: or, grievance.}
\vs Hab 1:14 And makest men as the fishes of the sea, as the creeping things, \bibemph{that have} no ruler over them?
\fn{creeping: or, moving.}
\vs Hab 1:15 They take up all of them with the angle, they catch them in their net, and gather them in their drag: therefore they rejoice and are glad.
\fn{drag: or, flue net.}
\vs Hab 1:16 Therefore they sacrifice unto their net, and burn incense unto their drag; because by them their portion \bibemph{is} fat, and their meat plenteous.
\fn{plenteous: or, dainty: Heb. fat.}
\vs Hab 1:17 Shall they therefore empty their net, and not spare continually to slay the nations?
\vs Hab 2:1 I will stand upon my watch, and set me upon the tower, and will watch to see what he will say unto me, and what I shall answer when I am reproved.
\fn{tower: Heb. fenced place. unto me: or, in me. when...: or, when I am argued with: Heb. upon my reproof, or, arguing.}
\vs Hab 2:2 And the \textsc{Lord} answered me, and said, Write the vision, and make \bibemph{it} plain upon tables, that he may run that readeth it.
\vs Hab 2:3 For the vision \bibemph{is} yet for an appointed time, but at the end it shall speak, and not lie: though it tarry, wait for it; because it will surely come, it will not tarry.
\vs Hab 2:4 Behold, his soul \bibemph{which} is lifted up is not upright in him: but the just shall live by his faith.
\vs Hab 2:5 Yea also, because he transgresseth by wine, \bibemph{he is} a proud man, neither keepeth at home, who enlargeth his desire as hell, and \bibemph{is} as death, and cannot be satisfied, but gathereth unto him all nations, and heapeth unto him all people:
\fn{Yea...: or, How much more.}
\vs Hab 2:6 Shall not all these take up a parable against him, and a taunting proverb against him, and say, Woe to him that increaseth \bibemph{that which is} not his! how long? and to him that ladeth himself with thick clay!
\fn{Woe...: or, Ho, he.}
\vs Hab 2:7 Shall they not rise up suddenly that shall bite thee, and awake that shall vex thee, and thou shalt be for booties unto them?
\vs Hab 2:8 Because thou hast spoiled many nations, all the remnant of the people shall spoil thee; because of men's blood, and \bibemph{for} the violence of the land, of the city, and of all that dwell therein.
\fn{blood: Heb. bloods.}
\vs Hab 2:9 Woe to him that coveteth an evil covetousness to his house, that he may set his nest on high, that he may be delivered from the power of evil!
\fn{coveteth...: or, gaineth and evil gain. power...: Heb. palm of the hand.}
\vs Hab 2:10 Thou hast consulted shame to thy house by cutting off many people, and hast sinned \bibemph{against} thy soul.
\vs Hab 2:11 For the stone shall cry out of the wall, and the beam out of the timber shall answer it.
\fn{beam: or, piece, or, fastening. answer it: or, witness against it.}
\vs Hab 2:12 Woe to him that buildeth a town with blood, and stablisheth a city by iniquity!
\fn{blood: Heb. bloods.}
\vs Hab 2:13 Behold, \bibemph{is it} not of the \textsc{Lord} of hosts that the people shall labour in the very fire, and the people shall weary themselves for very vanity?
\fn{for...: or, in vain?.}
\vs Hab 2:14 For the earth shall be filled with the knowledge of the glory of the \textsc{Lord}, as the waters cover the sea.
\fn{with...: or, by knowing the glory.}
\vs Hab 2:15 Woe unto him that giveth his neighbour drink, that puttest thy bottle to \bibemph{him}, and makest \bibemph{him} drunken also, that thou mayest look on their nakedness!
\vs Hab 2:16 Thou art filled with shame for glory: drink thou also, and let thy foreskin be uncovered: the cup of the \textsc{Lord's} right hand shall be turned unto thee, and shameful spewing \bibemph{shall be} on thy glory.
\fn{with...: or, more with shame than with glory.}
\vs Hab 2:17 For the violence of Lebanon shall cover thee, and the spoil of beasts, \bibemph{which} made them afraid, because of men's blood, and for the violence of the land, of the city, and of all that dwell therein.
\vs Hab 2:18 What profiteth the graven image that the maker thereof hath graven it; the molten image, and a teacher of lies, that the maker of his work trusteth therein, to make dumb idols?
\fn{maker of...: Heb. fashioner of his fashion.}
\vs Hab 2:19 Woe unto him that saith to the wood, Awake; to the dumb stone, Arise, it shall teach! Behold, it \bibemph{is} laid over with gold and silver, and \bibemph{there is} no breath at all in the midst of it.
\vs Hab 2:20 But the \textsc{Lord} \bibemph{is} in his holy temple: let all the earth keep silence before him.
\fn{let...: Heb. be silent all the earth before him.}
\vs Hab 3:1 A prayer of Habakkuk the prophet upon Shigionoth.
\fn{upon...: or, according to variable songs, or, tunes, called in Hebrew, Shigionoth.}
\vs Hab 3:2 O \textsc{Lord}, I have heard thy speech, \bibemph{and} was afraid: O \textsc{Lord}, revive thy work in the midst of the years, in the midst of the years make known; in wrath remember mercy.
\fn{speech: Heb. report, or, hearing. revive: or, preserve alive.}
\vs Hab 3:3 God came from Teman, and the Holy One from mount Paran. Selah. His glory covered the heavens, and the earth was full of his praise.
\fn{Teman: or, the south.}
\vs Hab 3:4 And \bibemph{his} brightness was as the light; he had horns \bibemph{coming} out of his hand: and there \bibemph{was} the hiding of his power.
\fn{horns...: or, bright beams out of his side.}
\vs Hab 3:5 Before him went the pestilence, and burning coals went forth at his feet.
\fn{coals: or, diseases.}
\vs Hab 3:6 He stood, and measured the earth: he beheld, and drove asunder the nations; and the everlasting mountains were scattered, the perpetual hills did bow: his ways \bibemph{are} everlasting.
\vs Hab 3:7 I saw the tents of Cushan in affliction: \bibemph{and} the curtains of the land of Midian did tremble.
\fn{Cushan: or, Ethiopia. in...: or, under affliction, or, vanity.}
\vs Hab 3:8 Was the \textsc{Lord} displeased against the rivers? \bibemph{was} thine anger against the rivers? \bibemph{was} thy wrath against the sea, that thou didst ride upon thine horses \bibemph{and} thy chariots of salvation?
\fn{of...: or, were salvation?.}
\vs Hab 3:9 Thy bow was made quite naked, \bibemph{according} to the oaths of the tribes, \bibemph{even thy} word. Selah. Thou didst cleave the earth with rivers.
\fn{earth...: or, rivers of the earth.}
\vs Hab 3:10 The mountains saw thee, \bibemph{and} they trembled: the overflowing of the water passed by: the deep uttered his voice, \bibemph{and} lifted up his hands on high.
\vs Hab 3:11 The sun \bibemph{and} moon stood still in their habitation: at the light of thine arrows they went, \bibemph{and} at the shining of thy glittering spear.
\fn{at the light...: or, thine arrows walked in the light.}
\vs Hab 3:12 Thou didst march through the land in indignation, thou didst thresh the heathen in anger.
\vs Hab 3:13 Thou wentest forth for the salvation of thy people, \bibemph{even} for salvation with thine anointed; thou woundedst the head out of the house of the wicked, by discovering the foundation unto the neck. Selah.
\fn{by...: Heb. making naked.}
\vs Hab 3:14 Thou didst strike through with his staves the head of his villages: they came out as a whirlwind to scatter me: their rejoicing \bibemph{was} as to devour the poor secretly.
\fn{came...: Heb. were tempestuous.}
\vs Hab 3:15 Thou didst walk through the sea with thine horses, \bibemph{through} the heap of great waters.
\fn{heap: or, mud.}
\vs Hab 3:16 When I heard, my belly trembled; my lips quivered at the voice: rottenness entered into my bones, and I trembled in myself, that I might rest in the day of trouble: when he cometh up unto the people, he will invade them with his troops.
\fn{invade...: or, cut them in pieces.}
\vs Hab 3:17 Although the fig tree shall not blossom, neither \bibemph{shall} fruit \bibemph{be} in the vines; the labour of the olive shall fail, and the fields shall yield no meat; the flock shall be cut off from the fold, and \bibemph{there shall be} no herd in the stalls:
\fn{fall: Heb. lie.}
\vs Hab 3:18 Yet I will rejoice in the \textsc{Lord}, I will joy in the God of my salvation.
\vs Hab 3:19 The \textsc{Lord} God \bibemph{is} my strength, and he will make my feet like hinds' \bibemph{feet}, and he will make me to walk upon mine high places. To the chief singer on my stringed instruments.
\fn{stringed...: Heb. Neginoth.}
