\bibbook{the book of the prophet\\\bibtitlemain{jeremiah}}{jeremiah}{Jeremiah}{Jer}
\vs Jer 1:1 The words of Jeremiah the son of Hilkiah, of the priests that \bibemph{were} in Anathoth in the land of Benjamin:
\vs Jer 1:2 To whom the word of the \textsc{Lord} came in the days of Josiah the son of Amon king of Judah, in the thirteenth year of his reign.
\vs Jer 1:3 It came also in the days of Jehoiakim the son of Josiah king of Judah, unto the end of the eleventh year of Zedekiah the son of Josiah king of Judah, unto the carrying away of Jerusalem captive in the fifth month.
\vs Jer 1:4 Then the word of the \textsc{Lord} came unto me, saying,
\vs Jer 1:5 Before I formed thee in the belly I knew thee; and before thou camest forth out of the womb I sanctified thee, \bibemph{and} I ordained thee a prophet unto the nations.
\fn{ordained: Heb. gave.}
\vs Jer 1:6 Then said I, Ah, Lord GOD! behold, I cannot speak: for I \bibemph{am} a child.
\vs Jer 1:7 But the \textsc{Lord} said unto me, Say not, I \bibemph{am} a child: for thou shalt go to all that I shall send thee, and whatsoever I command thee thou shalt speak.
\vs Jer 1:8 Be not afraid of their faces: for I \bibemph{am} with thee to deliver thee, saith the \textsc{Lord}.
\vs Jer 1:9 Then the \textsc{Lord} put forth his hand, and touched my mouth. And the \textsc{Lord} said unto me, Behold, I have put my words in thy mouth.
\vs Jer 1:10 See, I have this day set thee over the nations and over the kingdoms, to root out, and to pull down, and to destroy, and to throw down, to build, and to plant.
\vs Jer 1:11 Moreover the word of the \textsc{Lord} came unto me, saying, Jeremiah, what seest thou? And I said, I see a rod of an almond tree.
\vs Jer 1:12 Then said the \textsc{Lord} unto me, Thou hast well seen: for I will hasten my word to perform it.
\vs Jer 1:13 And the word of the \textsc{Lord} came unto me the second time, saying, What seest thou? And I said, I see a seething pot; and the face thereof \bibemph{is} toward the north.
\fn{toward...: Heb. from the face of the north.}
\vs Jer 1:14 Then the \textsc{Lord} said unto me, Out of the north an evil shall break forth upon all the inhabitants of the land.
\fn{shall...: Heb. shall be opened.}
\vs Jer 1:15 For, lo, I will call all the families of the kingdoms of the north, saith the \textsc{Lord}; and they shall come, and they shall set every one his throne at the entering of the gates of Jerusalem, and against all the walls thereof round about, and against all the cities of Judah.
\vs Jer 1:16 And I will utter my judgments against them touching all their wickedness, who have forsaken me, and have burned incense unto other gods, and worshipped the works of their own hands.
\vs Jer 1:17 Thou therefore gird up thy loins, and arise, and speak unto them all that I command thee: be not dismayed at their faces, lest I confound thee before them.
\fn{confound: or, break to pieces.}
\vs Jer 1:18 For, behold, I have made thee this day a defenced city, and an iron pillar, and brasen walls against the whole land, against the kings of Judah, against the princes thereof, against the priests thereof, and against the people of the land.
\vs Jer 1:19 And they shall fight against thee; but they shall not prevail against thee; for I \bibemph{am} with thee, saith the \textsc{Lord}, to deliver thee.
\vs Jer 2:1 Moreover the word of the \textsc{Lord} came to me, saying,
\vs Jer 2:2 Go and cry in the ears of Jerusalem, saying, Thus saith the \textsc{Lord}; I remember thee, the kindness of thy youth, the love of thine espousals, when thou wentest after me in the wilderness, in a land \bibemph{that was} not sown.
\fn{thee: or, for thy sake.}
\vs Jer 2:3 Israel \bibemph{was} holiness unto the \textsc{Lord}, \bibemph{and} the firstfruits of his increase: all that devour him shall offend; evil shall come upon them, saith the \textsc{Lord}.
\vs Jer 2:4 Hear ye the word of the \textsc{Lord}, O house of Jacob, and all the families of the house of Israel:
\vs Jer 2:5 Thus saith the \textsc{Lord}, What iniquity have your fathers found in me, that they are gone far from me, and have walked after vanity, and are become vain?
\vs Jer 2:6 Neither said they, Where \bibemph{is} the \textsc{Lord} that brought us up out of the land of Egypt, that led us through the wilderness, through a land of deserts and of pits, through a land of drought, and of the shadow of death, through a land that no man passed through, and where no man dwelt?
\vs Jer 2:7 And I brought you into a plentiful country, to eat the fruit thereof and the goodness thereof; but when ye entered, ye defiled my land, and made mine heritage an abomination.
\fn{a plentiful...: or, the land of Carmel.}
\vs Jer 2:8 The priests said not, Where \bibemph{is} the \textsc{Lord}? and they that handle the law knew me not: the pastors also transgressed against me, and the prophets prophesied by Baal, and walked after \bibemph{things that} do not profit.
\vs Jer 2:9 Wherefore I will yet plead with you, saith the \textsc{Lord}, and with your children's children will I plead.
\vs Jer 2:10 For pass over the isles of Chittim, and see; and send unto Kedar, and consider diligently, and see if there be such a thing.
\fn{over: or, over to.}
\vs Jer 2:11 Hath a nation changed \bibemph{their} gods, which \bibemph{are} yet no gods? but my people have changed their glory for \bibemph{that which} doth not profit.
\vs Jer 2:12 Be astonished, O ye heavens, at this, and be horribly afraid, be ye very desolate, saith the \textsc{Lord}.
\vs Jer 2:13 For my people have committed two evils; they have forsaken me the fountain of living waters, \bibemph{and} hewed them out cisterns, broken cisterns, that can hold no water.
\vs Jer 2:14 \bibemph{Is} Israel a servant? \bibemph{is} he a homeborn \bibemph{slave}? why is he spoiled?
\fn{spoiled: Heb. become a spoil?.}
\vs Jer 2:15 The young lions roared upon him, \bibemph{and} yelled, and they made his land waste: his cities are burned without inhabitant.
\fn{yelled: Heb. gave out their voice.}
\vs Jer 2:16 Also the children of Noph and Tahapanes have broken the crown of thy head.
\fn{have...: or, feed on thy crown.}
\vs Jer 2:17 Hast thou not procured this unto thyself, in that thou hast forsaken the \textsc{Lord} thy God, when he led thee by the way?
\vs Jer 2:18 And now what hast thou to do in the way of Egypt, to drink the waters of Sihor? or what hast thou to do in the way of Assyria, to drink the waters of the river?
\vs Jer 2:19 Thine own wickedness shall correct thee, and thy backslidings shall reprove thee: know therefore and see that \bibemph{it is} an evil \bibemph{thing} and bitter, that thou hast forsaken the \textsc{Lord} thy God, and that my fear \bibemph{is} not in thee, saith the Lord GOD of hosts.
\vs Jer 2:20 For of old time I have broken thy yoke, \bibemph{and} burst thy bands; and thou saidst, I will not transgress; when upon every high hill and under every green tree thou wanderest, playing the harlot.
\fn{transgress: or, serve.}
\vs Jer 2:21 Yet I had planted thee a noble vine, wholly a right seed: how then art thou turned into the degenerate plant of a strange vine unto me?
\vs Jer 2:22 For though thou wash thee with nitre, and take thee much soap, \bibemph{yet} thine iniquity is marked before me, saith the Lord GOD.
\vs Jer 2:23 How canst thou say, I am not polluted, I have not gone after Baalim? see thy way in the valley, know what thou hast done: \bibemph{thou art} a swift dromedary traversing her ways;
\fn{thou art...: or, O swift dromedary.}
\vs Jer 2:24 A wild ass used to the wilderness, \bibemph{that} snuffeth up the wind at her pleasure; in her occasion who can turn her away? all they that seek her will not weary themselves; in her month they shall find her.
\fn{A wild...: or, O wild ass, etc. turn...: or, reverse it?. used: Heb. taught. her pleasure: Heb. the desire of her heart.}
\vs Jer 2:25 Withhold thy foot from being unshod, and thy throat from thirst: but thou saidst, There is no hope: no; for I have loved strangers, and after them will I go.
\fn{There...: or, Is the case desperate?.}
\vs Jer 2:26 As the thief is ashamed when he is found, so is the house of Israel ashamed; they, their kings, their princes, and their priests, and their prophets,
\vs Jer 2:27 Saying to a stock, Thou \bibemph{art} my father; and to a stone, Thou hast brought me forth: for they have turned \bibemph{their} back unto me, and not \bibemph{their} face: but in the time of their trouble they will say, Arise, and save us.
\fn{brought...: or, begotten me. their back: Heb. the hinder part of the neck.}
\vs Jer 2:28 But where \bibemph{are} thy gods that thou hast made thee? let them arise, if they can save thee in the time of thy trouble: for \bibemph{according to} the number of thy cities are thy gods, O Judah.
\fn{trouble: Heb. evil.}
\vs Jer 2:29 Wherefore will ye plead with me? ye all have transgressed against me, saith the \textsc{Lord}.
\vs Jer 2:30 In vain have I smitten your children; they received no correction: your own sword hath devoured your prophets, like a destroying lion.
\vs Jer 2:31 O generation, see ye the word of the \textsc{Lord}. Have I been a wilderness unto Israel? a land of darkness? wherefore say my people, We are lords; we will come no more unto thee?
\fn{We are...: Heb. We have dominion.}
\vs Jer 2:32 Can a maid forget her ornaments, \bibemph{or} a bride her attire? yet my people have forgotten me days without number.
\vs Jer 2:33 Why trimmest thou thy way to seek love? therefore hast thou also taught the wicked ones thy ways.
\vs Jer 2:34 Also in thy skirts is found the blood of the souls of the poor innocents: I have not found it by secret search, but upon all these.
\fn{secret...: Heb. digging.}
\vs Jer 2:35 Yet thou sayest, Because I am innocent, surely his anger shall turn from me. Behold, I will plead with thee, because thou sayest, I have not sinned.
\vs Jer 2:36 Why gaddest thou about so much to change thy way? thou also shalt be ashamed of Egypt, as thou wast ashamed of Assyria.
\vs Jer 2:37 Yea, thou shalt go forth from him, and thine hands upon thine head: for the \textsc{Lord} hath rejected thy confidences, and thou shalt not prosper in them.
\vs Jer 3:1 They say, If a man put away his wife, and she go from him, and become another man's, shall he return unto her again? shall not that land be greatly polluted? but thou hast played the harlot with many lovers; yet return again to me, saith the \textsc{Lord}.
\fn{They say: Heb. Saying.}
\vs Jer 3:2 Lift up thine eyes unto the high places, and see where thou hast not been lien with. In the ways hast thou sat for them, as the Arabian in the wilderness; and thou hast polluted the land with thy whoredoms and with thy wickedness.
\vs Jer 3:3 Therefore the showers have been withholden, and there hath been no latter rain; and thou hadst a whore's forehead, thou refusedst to be ashamed.
\vs Jer 3:4 Wilt thou not from this time cry unto me, My father, thou \bibemph{art} the guide of my youth?
\vs Jer 3:5 Will he reserve \bibemph{his anger} for ever? will he keep \bibemph{it} to the end? Behold, thou hast spoken and done evil things as thou couldest.
\vs Jer 3:6 The \textsc{Lord} said also unto me in the days of Josiah the king, Hast thou seen \bibemph{that} which backsliding Israel hath done? she is gone up upon every high mountain and under every green tree, and there hath played the harlot.
\vs Jer 3:7 And I said after she had done all these \bibemph{things}, Turn thou unto me. But she returned not. And her treacherous sister Judah saw \bibemph{it}.
\vs Jer 3:8 And I saw, when for all the causes whereby backsliding Israel committed adultery I had put her away, and given her a bill of divorce; yet her treacherous sister Judah feared not, but went and played the harlot also.
\vs Jer 3:9 And it came to pass through the lightness of her whoredom, that she defiled the land, and committed adultery with stones and with stocks.
\fn{lightness: or, fame.}
\vs Jer 3:10 And yet for all this her treacherous sister Judah hath not turned unto me with her whole heart, but feignedly, saith the \textsc{Lord}.
\fn{feignedly: Heb. in falsehood.}
\vs Jer 3:11 And the \textsc{Lord} said unto me, The backsliding Israel hath justified herself more than treacherous Judah.
\vs Jer 3:12 Go and proclaim these words toward the north, and say, Return, thou backsliding Israel, saith the \textsc{Lord}; \bibemph{and} I will not cause mine anger to fall upon you: for I \bibemph{am} merciful, saith the \textsc{Lord}, \bibemph{and} I will not keep \bibemph{anger} for ever.
\vs Jer 3:13 Only acknowledge thine iniquity, that thou hast transgressed against the \textsc{Lord} thy God, and hast scattered thy ways to the strangers under every green tree, and ye have not obeyed my voice, saith the \textsc{Lord}.
\vs Jer 3:14 Turn, O backsliding children, saith the \textsc{Lord}; for I am married unto you: and I will take you one of a city, and two of a family, and I will bring you to Zion:
\vs Jer 3:15 And I will give you pastors according to mine heart, which shall feed you with knowledge and understanding.
\vs Jer 3:16 And it shall come to pass, when ye be multiplied and increased in the land, in those days, saith the \textsc{Lord}, they shall say no more, The ark of the covenant of the \textsc{Lord}: neither shall it come to mind: neither shall they remember it; neither shall they visit \bibemph{it}; neither shall \bibemph{that} be done any more.
\fn{come to mind: Heb. come upon the heart. that...: or, it be magnified.}
\vs Jer 3:17 At that time they shall call Jerusalem the throne of the \textsc{Lord}; and all the nations shall be gathered unto it, to the name of the \textsc{Lord}, to Jerusalem: neither shall they walk any more after the imagination of their evil heart.
\fn{imagination: or, stubbornness.}
\vs Jer 3:18 In those days the house of Judah shall walk with the house of Israel, and they shall come together out of the land of the north to the land that I have given for an inheritance unto your fathers.
\fn{with: or, to. given...: or, caused your fathers to possess.}
\vs Jer 3:19 But I said, How shall I put thee among the children, and give thee a pleasant land, a goodly heritage of the hosts of nations? and I said, Thou shalt call me, My father; and shalt not turn away from me.
\fn{pleasant...: Heb. land of desire. from me: Heb. from after me. a goodly...: Heb. an heritage of glory, or, beauty.}
\vs Jer 3:20 Surely \bibemph{as} a wife treacherously departeth from her husband, so have ye dealt treacherously with me, O house of Israel, saith the \textsc{Lord}.
\fn{husband: Heb. friend.}
\vs Jer 3:21 A voice was heard upon the high places, weeping \bibemph{and} supplications of the children of Israel: for they have perverted their way, \bibemph{and} they have forgotten the \textsc{Lord} their God.
\vs Jer 3:22 Return, ye backsliding children, \bibemph{and} I will heal your backslidings. Behold, we come unto thee; for thou \bibemph{art} the \textsc{Lord} our God.
\vs Jer 3:23 Truly in vain \bibemph{is salvation hoped for} from the hills, \bibemph{and from} the multitude of mountains: truly in the \textsc{Lord} our God \bibemph{is} the salvation of Israel.
\vs Jer 3:24 For shame hath devoured the labour of our fathers from our youth; their flocks and their herds, their sons and their daughters.
\vs Jer 3:25 We lie down in our shame, and our confusion covereth us: for we have sinned against the \textsc{Lord} our God, we and our fathers, from our youth even unto this day, and have not obeyed the voice of the \textsc{Lord} our God.
\vs Jer 4:1 If thou wilt return, O Israel, saith the \textsc{Lord}, return unto me: and if thou wilt put away thine abominations out of my sight, then shalt thou not remove.
\vs Jer 4:2 And thou shalt swear, The \textsc{Lord} liveth, in truth, in judgment, and in righteousness; and the nations shall bless themselves in him, and in him shall they glory.
\vs Jer 4:3 For thus saith the \textsc{Lord} to the men of Judah and Jerusalem, Break up your fallow ground, and sow not among thorns.
\vs Jer 4:4 Circumcise yourselves to the \textsc{Lord}, and take away the foreskins of your heart, ye men of Judah and inhabitants of Jerusalem: lest my fury come forth like fire, and burn that none can quench \bibemph{it}, because of the evil of your doings.
\vs Jer 4:5 Declare ye in Judah, and publish in Jerusalem; and say, Blow ye the trumpet in the land: cry, gather together, and say, Assemble yourselves, and let us go into the defenced cities.
\vs Jer 4:6 Set up the standard toward Zion: retire, stay not: for I will bring evil from the north, and a great destruction.
\fn{retire: or, strengthen. destruction: Heb. breaking.}
\vs Jer 4:7 The lion is come up from his thicket, and the destroyer of the Gentiles is on his way; he is gone forth from his place to make thy land desolate; \bibemph{and} thy cities shall be laid waste, without an inhabitant.
\vs Jer 4:8 For this gird you with sackcloth, lament and howl: for the fierce anger of the \textsc{Lord} is not turned back from us.
\vs Jer 4:9 And it shall come to pass at that day, saith the \textsc{Lord}, \bibemph{that} the heart of the king shall perish, and the heart of the princes; and the priests shall be astonished, and the prophets shall wonder.
\vs Jer 4:10 Then said I, Ah, Lord GOD! surely thou hast greatly deceived this people and Jerusalem, saying, Ye shall have peace; whereas the sword reacheth unto the soul.
\vs Jer 4:11 At that time shall it be said to this people and to Jerusalem, A dry wind of the high places in the wilderness toward the daughter of my people, not to fan, nor to cleanse,
\vs Jer 4:12 \bibemph{Even} a full wind from those \bibemph{places} shall come unto me: now also will I give sentence against them.
\fn{a full...: or, a fuller wind than those. give...: Heb. utter judgments.}
\vs Jer 4:13 Behold, he shall come up as clouds, and his chariots \bibemph{shall be} as a whirlwind: his horses are swifter than eagles. Woe unto us! for we are spoiled.
\vs Jer 4:14 O Jerusalem, wash thine heart from wickedness, that thou mayest be saved. How long shall thy vain thoughts lodge within thee?
\vs Jer 4:15 For a voice declareth from Dan, and publisheth affliction from mount Ephraim.
\vs Jer 4:16 Make ye mention to the nations; behold, publish against Jerusalem, \bibemph{that} watchers come from a far country, and give out their voice against the cities of Judah.
\vs Jer 4:17 As keepers of a field, are they against her round about; because she hath been rebellious against me, saith the \textsc{Lord}.
\vs Jer 4:18 Thy way and thy doings have procured these \bibemph{things} unto thee; this \bibemph{is} thy wickedness, because it is bitter, because it reacheth unto thine heart.
\vs Jer 4:19 My bowels, my bowels! I am pained at my very heart; my heart maketh a noise in me; I cannot hold my peace, because thou hast heard, O my soul, the sound of the trumpet, the alarm of war.
\fn{my very...: Heb. the walls of my heart.}
\vs Jer 4:20 Destruction upon destruction is cried; for the whole land is spoiled: suddenly are my tents spoiled, \bibemph{and} my curtains in a moment.
\vs Jer 4:21 How long shall I see the standard, \bibemph{and} hear the sound of the trumpet?
\vs Jer 4:22 For my people \bibemph{is} foolish, they have not known me; they \bibemph{are} sottish children, and they have none understanding: they \bibemph{are} wise to do evil, but to do good they have no knowledge.
\vs Jer 4:23 I beheld the earth, and, lo, \bibemph{it was} without form, and void; and the heavens, and they \bibemph{had} no light.
\vs Jer 4:24 I beheld the mountains, and, lo, they trembled, and all the hills moved lightly.
\vs Jer 4:25 I beheld, and, lo, \bibemph{there was} no man, and all the birds of the heavens were fled.
\vs Jer 4:26 I beheld, and, lo, the fruitful place \bibemph{was} a wilderness, and all the cities thereof were broken down at the presence of the \textsc{Lord}, \bibemph{and} by his fierce anger.
\vs Jer 4:27 For thus hath the \textsc{Lord} said, The whole land shall be desolate; yet will I not make a full end.
\vs Jer 4:28 For this shall the earth mourn, and the heavens above be black: because I have spoken \bibemph{it}, I have purposed \bibemph{it}, and will not repent, neither will I turn back from it.
\vs Jer 4:29 The whole city shall flee for the noise of the horsemen and bowmen; they shall go into thickets, and climb up upon the rocks: every city \bibemph{shall be} forsaken, and not a man dwell therein.
\vs Jer 4:30 And \bibemph{when} thou \bibemph{art} spoiled, what wilt thou do? Though thou clothest thyself with crimson, though thou deckest thee with ornaments of gold, though thou rentest thy face with painting, in vain shalt thou make thyself fair; \bibemph{thy} lovers will despise thee, they will seek thy life.
\fn{face: Heb. eyes.}
\vs Jer 4:31 For I have heard a voice as of a woman in travail, \bibemph{and} the anguish as of her that bringeth forth her first child, the voice of the daughter of Zion, \bibemph{that} bewaileth herself, \bibemph{that} spreadeth her hands, \bibemph{saying}, Woe \bibemph{is} me now! for my soul is wearied because of murderers.
\vs Jer 5:1 Run ye to and fro through the streets of Jerusalem, and see now, and know, and seek in the broad places thereof, if ye can find a man, if there be \bibemph{any} that executeth judgment, that seeketh the truth; and I will pardon it.
\vs Jer 5:2 And though they say, The \textsc{Lord} liveth; surely they swear falsely.
\vs Jer 5:3 O \textsc{Lord}, \bibemph{are} not thine eyes upon the truth? thou hast stricken them, but they have not grieved; thou hast consumed them, \bibemph{but} they have refused to receive correction: they have made their faces harder than a rock; they have refused to return.
\vs Jer 5:4 Therefore I said, Surely these \bibemph{are} poor; they are foolish: for they know not the way of the \textsc{Lord}, \bibemph{nor} the judgment of their God.
\vs Jer 5:5 I will get me unto the great men, and will speak unto them; for they have known the way of the \textsc{Lord}, \bibemph{and} the judgment of their God: but these have altogether broken the yoke, \bibemph{and} burst the bonds.
\vs Jer 5:6 Wherefore a lion out of the forest shall slay them, \bibemph{and} a wolf of the evenings shall spoil them, a leopard shall watch over their cities: every one that goeth out thence shall be torn in pieces: because their transgressions are many, \bibemph{and} their backslidings are increased.
\fn{evenings: or, deserts. are increased: Heb. are strong.}
\vs Jer 5:7 How shall I pardon thee for this? thy children have forsaken me, and sworn by \bibemph{them that are} no gods: when I had fed them to the full, they then committed adultery, and assembled themselves by troops in the harlots' houses.
\vs Jer 5:8 They were \bibemph{as} fed horses in the morning: every one neighed after his neighbour's wife.
\vs Jer 5:9 Shall I not visit for these \bibemph{things}? saith the \textsc{Lord}: and shall not my soul be avenged on such a nation as this?
\vs Jer 5:10 Go ye up upon her walls, and destroy; but make not a full end: take away her battlements; for they \bibemph{are} not the \textsc{Lord's}.
\vs Jer 5:11 For the house of Israel and the house of Judah have dealt very treacherously against me, saith the \textsc{Lord}.
\vs Jer 5:12 They have belied the \textsc{Lord}, and said, \bibemph{It is} not he; neither shall evil come upon us; neither shall we see sword nor famine:
\vs Jer 5:13 And the prophets shall become wind, and the word \bibemph{is} not in them: thus shall it be done unto them.
\vs Jer 5:14 Wherefore thus saith the \textsc{Lord} God of hosts, Because ye speak this word, behold, I will make my words in thy mouth fire, and this people wood, and it shall devour them.
\vs Jer 5:15 Lo, I will bring a nation upon you from far, O house of Israel, saith the \textsc{Lord}: it \bibemph{is} a mighty nation, it \bibemph{is} an ancient nation, a nation whose language thou knowest not, neither understandest what they say.
\vs Jer 5:16 Their quiver \bibemph{is} as an open sepulchre, they \bibemph{are} all mighty men.
\vs Jer 5:17 And they shall eat up thine harvest, and thy bread, \bibemph{which} thy sons and thy daughters should eat: they shall eat up thy flocks and thine herds: they shall eat up thy vines and thy fig trees: they shall impoverish thy fenced cities, wherein thou trustedst, with the sword.
\vs Jer 5:18 Nevertheless in those days, saith the \textsc{Lord}, I will not make a full end with you.
\vs Jer 5:19 And it shall come to pass, when ye shall say, Wherefore doeth the \textsc{Lord} our God all these \bibemph{things} unto us? then shalt thou answer them, Like as ye have forsaken me, and served strange gods in your land, so shall ye serve strangers in a land \bibemph{that is} not yours.
\vs Jer 5:20 Declare this in the house of Jacob, and publish it in Judah, saying,
\vs Jer 5:21 Hear now this, O foolish people, and without understanding; which have eyes, and see not; which have ears, and hear not:
\fn{understanding: Heb. heart.}
\vs Jer 5:22 Fear ye not me? saith the \textsc{Lord}: will ye not tremble at my presence, which have placed the sand \bibemph{for} the bound of the sea by a perpetual decree, that it cannot pass it: and though the waves thereof toss themselves, yet can they not prevail; though they roar, yet can they not pass over it?
\vs Jer 5:23 But this people hath a revolting and a rebellious heart; they are revolted and gone.
\vs Jer 5:24 Neither say they in their heart, Let us now fear the \textsc{Lord} our God, that giveth rain, both the former and the latter, in his season: he reserveth unto us the appointed weeks of the harvest.
\vs Jer 5:25 Your iniquities have turned away these \bibemph{things}, and your sins have withholden good \bibemph{things} from you.
\vs Jer 5:26 For among my people are found wicked \bibemph{men}: they lay wait, as he that setteth snares; they set a trap, they catch men.
\fn{they lay...: or, they pry as fowlers lie in wait.}
\vs Jer 5:27 As a cage is full of birds, so \bibemph{are} their houses full of deceit: therefore they are become great, and waxen rich.
\fn{cage: or, coop.}
\vs Jer 5:28 They are waxen fat, they shine: yea, they overpass the deeds of the wicked: they judge not the cause, the cause of the fatherless, yet they prosper; and the right of the needy do they not judge.
\vs Jer 5:29 Shall I not visit for these \bibemph{things}? saith the \textsc{Lord}: shall not my soul be avenged on such a nation as this?
\vs Jer 5:30 A wonderful and horrible thing is committed in the land;
\fn{A wonderful...: or, Astonishment and filthiness.}
\vs Jer 5:31 The prophets prophesy falsely, and the priests bear rule by their means; and my people love \bibemph{to have it} so: and what will ye do in the end thereof?
\fn{bear...: or, take into their hands.}
\vs Jer 6:1 O ye children of Benjamin, gather yourselves to flee out of the midst of Jerusalem, and blow the trumpet in Tekoa, and set up a sign of fire in Bethhaccerem: for evil appeareth out of the north, and great destruction.
\vs Jer 6:2 I have likened the daughter of Zion to a comely and delicate \bibemph{woman}.
\fn{comely: or, dwelling at home.}
\vs Jer 6:3 The shepherds with their flocks shall come unto her; they shall pitch \bibemph{their} tents against her round about; they shall feed every one in his place.
\vs Jer 6:4 Prepare ye war against her; arise, and let us go up at noon. Woe unto us! for the day goeth away, for the shadows of the evening are stretched out.
\vs Jer 6:5 Arise, and let us go by night, and let us destroy her palaces.
\vs Jer 6:6 For thus hath the \textsc{Lord} of hosts said, Hew ye down trees, and cast a mount against Jerusalem: this \bibemph{is} the city to be visited; she \bibemph{is} wholly oppression in the midst of her.
\fn{cast...: or, pour out the engine of shot.}
\vs Jer 6:7 As a fountain casteth out her waters, so she casteth out her wickedness: violence and spoil is heard in her; before me continually \bibemph{is} grief and wounds.
\vs Jer 6:8 Be thou instructed, O Jerusalem, lest my soul depart from thee; lest I make thee desolate, a land not inhabited.
\fn{depart: Heb. be loosed, or, disjointed.}
\vs Jer 6:9 Thus saith the \textsc{Lord} of hosts, They shall throughly glean the remnant of Israel as a vine: turn back thine hand as a grapegatherer into the baskets.
\vs Jer 6:10 To whom shall I speak, and give warning, that they may hear? behold, their ear \bibemph{is} uncircumcised, and they cannot hearken: behold, the word of the \textsc{Lord} is unto them a reproach; they have no delight in it.
\vs Jer 6:11 Therefore I am full of the fury of the \textsc{Lord}; I am weary with holding in: I will pour it out upon the children abroad, and upon the assembly of young men together: for even the husband with the wife shall be taken, the aged with \bibemph{him that is} full of days.
\vs Jer 6:12 And their houses shall be turned unto others, \bibemph{with their} fields and wives together: for I will stretch out my hand upon the inhabitants of the land, saith the \textsc{Lord}.
\vs Jer 6:13 For from the least of them even unto the greatest of them every one \bibemph{is} given to covetousness; and from the prophet even unto the priest every one dealeth falsely.
\vs Jer 6:14 They have healed also the hurt \bibemph{of the daughter} of my people slightly, saying, Peace, peace; when \bibemph{there is} no peace.
\fn{hurt: Heb. bruise, or, breach.}
\vs Jer 6:15 Were they ashamed when they had committed abomination? nay, they were not at all ashamed, neither could they blush: therefore they shall fall among them that fall: at the time \bibemph{that} I visit them they shall be cast down, saith the \textsc{Lord}.
\vs Jer 6:16 Thus saith the \textsc{Lord}, Stand ye in the ways, and see, and ask for the old paths, where \bibemph{is} the good way, and walk therein, and ye shall find rest for your souls. But they said, We will not walk \bibemph{therein}.
\vs Jer 6:17 Also I set watchmen over you, \bibemph{saying}, Hearken to the sound of the trumpet. But they said, We will not hearken.
\vs Jer 6:18 Therefore hear, ye nations, and know, O congregation, what \bibemph{is} among them.
\vs Jer 6:19 Hear, O earth: behold, I will bring evil upon this people, \bibemph{even} the fruit of their thoughts, because they have not hearkened unto my words, nor to my law, but rejected it.
\vs Jer 6:20 To what purpose cometh there to me incense from Sheba, and the sweet cane from a far country? your burnt offerings \bibemph{are} not acceptable, nor your sacrifices sweet unto me.
\vs Jer 6:21 Therefore thus saith the \textsc{Lord}, Behold, I will lay stumblingblocks before this people, and the fathers and the sons together shall fall upon them; the neighbour and his friend shall perish.
\vs Jer 6:22 Thus saith the \textsc{Lord}, Behold, a people cometh from the north country, and a great nation shall be raised from the sides of the earth.
\vs Jer 6:23 They shall lay hold on bow and spear; they \bibemph{are} cruel, and have no mercy; their voice roareth like the sea; and they ride upon horses, set in array as men for war against thee, O daughter of Zion.
\vs Jer 6:24 We have heard the fame thereof: our hands wax feeble: anguish hath taken hold of us, \bibemph{and} pain, as of a woman in travail.
\vs Jer 6:25 Go not forth into the field, nor walk by the way; for the sword of the enemy \bibemph{and} fear \bibemph{is} on every side.
\vs Jer 6:26 O daughter of my people, gird \bibemph{thee} with sackcloth, and wallow thyself in ashes: make thee mourning, \bibemph{as for} an only son, most bitter lamentation: for the spoiler shall suddenly come upon us.
\vs Jer 6:27 I have set thee \bibemph{for} a tower \bibemph{and} a fortress among my people, that thou mayest know and try their way.
\vs Jer 6:28 They \bibemph{are} all grievous revolters, walking with slanders: \bibemph{they are} brass and iron; they \bibemph{are} all corrupters.
\vs Jer 6:29 The bellows are burned, the lead is consumed of the fire; the founder melteth in vain: for the wicked are not plucked away.
\vs Jer 6:30 Reprobate silver shall \bibemph{men} call them, because the \textsc{Lord} hath rejected them.
\fn{Reprobate...: or, Refuse silver.}
\vs Jer 7:1 The word that came to Jeremiah from the \textsc{Lord}, saying,
\vs Jer 7:2 Stand in the gate of the \textsc{Lord's} house, and proclaim there this word, and say, Hear the word of the \textsc{Lord}, all \bibemph{ye of} Judah, that enter in at these gates to worship the \textsc{Lord}.
\vs Jer 7:3 Thus saith the \textsc{Lord} of hosts, the God of Israel, Amend your ways and your doings, and I will cause you to dwell in this place.
\vs Jer 7:4 Trust ye not in lying words, saying, The temple of the \textsc{Lord}, The temple of the \textsc{Lord}, The temple of the \textsc{Lord}, \bibemph{are} these.
\vs Jer 7:5 For if ye throughly amend your ways and your doings; if ye throughly execute judgment between a man and his neighbour;
\vs Jer 7:6 \bibemph{If} ye oppress not the stranger, the fatherless, and the widow, and shed not innocent blood in this place, neither walk after other gods to your hurt:
\vs Jer 7:7 Then will I cause you to dwell in this place, in the land that I gave to your fathers, for ever and ever.
\vs Jer 7:8 Behold, ye trust in lying words, that cannot profit.
\vs Jer 7:9 Will ye steal, murder, and commit adultery, and swear falsely, and burn incense unto Baal, and walk after other gods whom ye know not;
\vs Jer 7:10 And come and stand before me in this house, which is called by my name, and say, We are delivered to do all these abominations?
\fn{which...: Heb. whereupon my name is called.}
\vs Jer 7:11 Is this house, which is called by my name, become a den of robbers in your eyes? Behold, even I have seen \bibemph{it}, saith the \textsc{Lord}.
\vs Jer 7:12 But go ye now unto my place which \bibemph{was} in Shiloh, where I set my name at the first, and see what I did to it for the wickedness of my people Israel.
\vs Jer 7:13 And now, because ye have done all these works, saith the \textsc{Lord}, and I spake unto you, rising up early and speaking, but ye heard not; and I called you, but ye answered not;
\vs Jer 7:14 Therefore will I do unto \bibemph{this} house, which is called by my name, wherein ye trust, and unto the place which I gave to you and to your fathers, as I have done to Shiloh.
\vs Jer 7:15 And I will cast you out of my sight, as I have cast out all your brethren, \bibemph{even} the whole seed of Ephraim.
\vs Jer 7:16 Therefore pray not thou for this people, neither lift up cry nor prayer for them, neither make intercession to me: for I will not hear thee.
\vs Jer 7:17 Seest thou not what they do in the cities of Judah and in the streets of Jerusalem?
\vs Jer 7:18 The children gather wood, and the fathers kindle the fire, and the women knead \bibemph{their} dough, to make cakes to the queen of heaven, and to pour out drink offerings unto other gods, that they may provoke me to anger.
\fn{queen...: or, frame, or, workmanship of heaven.}
\vs Jer 7:19 Do they provoke me to anger? saith the \textsc{Lord}: \bibemph{do they} not \bibemph{provoke} themselves to the confusion of their own faces?
\vs Jer 7:20 Therefore thus saith the Lord GOD; Behold, mine anger and my fury shall be poured out upon this place, upon man, and upon beast, and upon the trees of the field, and upon the fruit of the ground; and it shall burn, and shall not be quenched.
\vs Jer 7:21 Thus saith the \textsc{Lord} of hosts, the God of Israel; Put your burnt offerings unto your sacrifices, and eat flesh.
\vs Jer 7:22 For I spake not unto your fathers, nor commanded them in the day that I brought them out of the land of Egypt, concerning burnt offerings or sacrifices:
\fn{concerning: Heb. concerning the matter of.}
\vs Jer 7:23 But this thing commanded I them, saying, Obey my voice, and I will be your God, and ye shall be my people: and walk ye in all the ways that I have commanded you, that it may be well unto you.
\vs Jer 7:24 But they hearkened not, nor inclined their ear, but walked in the counsels \bibemph{and} in the imagination of their evil heart, and went backward, and not forward.
\fn{imagination: or, stubbornness. went: Heb. were.}
\vs Jer 7:25 Since the day that your fathers came forth out of the land of Egypt unto this day I have even sent unto you all my servants the prophets, daily rising up early and sending \bibemph{them}:
\vs Jer 7:26 Yet they hearkened not unto me, nor inclined their ear, but hardened their neck: they did worse than their fathers.
\vs Jer 7:27 Therefore thou shalt speak all these words unto them; but they will not hearken to thee: thou shalt also call unto them; but they will not answer thee.
\vs Jer 7:28 But thou shalt say unto them, This \bibemph{is} a nation that obeyeth not the voice of the \textsc{Lord} their God, nor receiveth correction: truth is perished, and is cut off from their mouth.
\fn{correction: or, instruction.}
\vs Jer 7:29 Cut off thine hair, \bibemph{O Jerusalem}, and cast \bibemph{it} away, and take up a lamentation on high places; for the \textsc{Lord} hath rejected and forsaken the generation of his wrath.
\vs Jer 7:30 For the children of Judah have done evil in my sight, saith the \textsc{Lord}: they have set their abominations in the house which is called by my name, to pollute it.
\vs Jer 7:31 And they have built the high places of Tophet, which \bibemph{is} in the valley of the son of Hinnom, to burn their sons and their daughters in the fire; which I commanded \bibemph{them} not, neither came it into my heart.
\fn{came...: Heb. came it upon my heart.}
\vs Jer 7:32 Therefore, behold, the days come, saith the \textsc{Lord}, that it shall no more be called Tophet, nor the valley of the son of Hinnom, but the valley of slaughter: for they shall bury in Tophet, till there be no place.
\vs Jer 7:33 And the carcases of this people shall be meat for the fowls of the heaven, and for the beasts of the earth; and none shall fray \bibemph{them} away.
\vs Jer 7:34 Then will I cause to cease from the cities of Judah, and from the streets of Jerusalem, the voice of mirth, and the voice of gladness, the voice of the bridegroom, and the voice of the bride: for the land shall be desolate.
\vs Jer 8:1 At that time, saith the \textsc{Lord}, they shall bring out the bones of the kings of Judah, and the bones of his princes, and the bones of the priests, and the bones of the prophets, and the bones of the inhabitants of Jerusalem, out of their graves:
\vs Jer 8:2 And they shall spread them before the sun, and the moon, and all the host of heaven, whom they have loved, and whom they have served, and after whom they have walked, and whom they have sought, and whom they have worshipped: they shall not be gathered, nor be buried; they shall be for dung upon the face of the earth.
\vs Jer 8:3 And death shall be chosen rather than life by all the residue of them that remain of this evil family, which remain in all the places whither I have driven them, saith the \textsc{Lord} of hosts.
\vs Jer 8:4 Moreover thou shalt say unto them, Thus saith the \textsc{Lord}; Shall they fall, and not arise? shall he turn away, and not return?
\vs Jer 8:5 Why \bibemph{then} is this people of Jerusalem slidden back by a perpetual backsliding? they hold fast deceit, they refuse to return.
\vs Jer 8:6 I hearkened and heard, \bibemph{but} they spake not aright: no man repented him of his wickedness, saying, What have I done? every one turned to his course, as the horse rusheth into the battle.
\vs Jer 8:7 Yea, the stork in the heaven knoweth her appointed times; and the turtle and the crane and the swallow observe the time of their coming; but my people know not the judgment of the \textsc{Lord}.
\vs Jer 8:8 How do ye say, We \bibemph{are} wise, and the law of the \textsc{Lord} \bibemph{is} with us? Lo, certainly in vain made he \bibemph{it}; the pen of the scribes \bibemph{is} in vain.
\fn{in vain made...: or, the false pen of the scribes worketh for falsehood.}
\vs Jer 8:9 The wise \bibemph{men} are ashamed, they are dismayed and taken: lo, they have rejected the word of the \textsc{Lord}; and what wisdom \bibemph{is} in them?
\fn{The wise...: or, Have they been ashamed, etc. what...: Heb. the wisdom of what thing.}
\vs Jer 8:10 Therefore will I give their wives unto others, \bibemph{and} their fields to them that shall inherit \bibemph{them}: for every one from the least even unto the greatest is given to covetousness, from the prophet even unto the priest every one dealeth falsely.
\vs Jer 8:11 For they have healed the hurt of the daughter of my people slightly, saying, Peace, peace; when \bibemph{there is} no peace.
\vs Jer 8:12 Were they ashamed when they had committed abomination? nay, they were not at all ashamed, neither could they blush: therefore shall they fall among them that fall: in the time of their visitation they shall be cast down, saith the \textsc{Lord}.
\fn{I will...: or, In gathering I will consume.}
\vs Jer 8:13 I will surely consume them, saith the \textsc{Lord}: \bibemph{there shall be} no grapes on the vine, nor figs on the fig tree, and the leaf shall fade; and \bibemph{the things that} I have given them shall pass away from them.
\vs Jer 8:14 Why do we sit still? assemble yourselves, and let us enter into the defenced cities, and let us be silent there: for the \textsc{Lord} our God hath put us to silence, and given us water of gall to drink, because we have sinned against the \textsc{Lord}.
\fn{gall: or, poison.}
\vs Jer 8:15 We looked for peace, but no good \bibemph{came; and} for a time of health, and behold trouble!
\vs Jer 8:16 The snorting of his horses was heard from Dan: the whole land trembled at the sound of the neighing of his strong ones; for they are come, and have devoured the land, and all that is in it; the city, and those that dwell therein.
\fn{all...: Heb. the fulness thereof.}
\vs Jer 8:17 For, behold, I will send serpents, cockatrices, among you, which \bibemph{will} not \bibemph{be} charmed, and they shall bite you, saith the \textsc{Lord}.
\vs Jer 8:18 \bibemph{When} I would comfort myself against sorrow, my heart \bibemph{is} faint in me.
\fn{in: Heb. upon.}
\vs Jer 8:19 Behold the voice of the cry of the daughter of my people because of them that dwell in a far country: \bibemph{Is} not the \textsc{Lord} in Zion? \bibemph{is} not her king in her? Why have they provoked me to anger with their graven images, \bibemph{and} with strange vanities?
\fn{them...: Heb. the country of them that are far off.}
\vs Jer 8:20 The harvest is past, the summer is ended, and we are not saved.
\vs Jer 8:21 For the hurt of the daughter of my people am I hurt; I am black; astonishment hath taken hold on me.
\vs Jer 8:22 \bibemph{Is there} no balm in Gilead; \bibemph{is there} no physician there? why then is not the health of the daughter of my people recovered?
\fn{recovered: Heb. gone up?.}
\vs Jer 9:1 Oh that my head were waters, and mine eyes a fountain of tears, that I might weep day and night for the slain of the daughter of my people!
\fn{Oh...: Heb. Who will give my head, etc.}
\vs Jer 9:2 Oh that I had in the wilderness a lodging place of wayfaring men; that I might leave my people, and go from them! for they \bibemph{be} all adulterers, an assembly of treacherous men.
\vs Jer 9:3 And they bend their tongues \bibemph{like} their bow \bibemph{for} lies: but they are not valiant for the truth upon the earth; for they proceed from evil to evil, and they know not me, saith the \textsc{Lord}.
\vs Jer 9:4 Take ye heed every one of his neighbour, and trust ye not in any brother: for every brother will utterly supplant, and every neighbour will walk with slanders.
\fn{neighbour: or, friend.}
\vs Jer 9:5 And they will deceive every one his neighbour, and will not speak the truth: they have taught their tongue to speak lies, \bibemph{and} weary themselves to commit iniquity.
\fn{deceive: or, mock.}
\vs Jer 9:6 Thine habitation \bibemph{is} in the midst of deceit; through deceit they refuse to know me, saith the \textsc{Lord}.
\vs Jer 9:7 Therefore thus saith the \textsc{Lord} of hosts, Behold, I will melt them, and try them; for how shall I do for the daughter of my people?
\vs Jer 9:8 Their tongue \bibemph{is as} an arrow shot out; it speaketh deceit: \bibemph{one} speaketh peaceably to his neighbour with his mouth, but in heart he layeth his wait.
\fn{in heart: Heb. in the midst of him. his wait: or, wait for him.}
\vs Jer 9:9 Shall I not visit them for these \bibemph{things}? saith the \textsc{Lord}: shall not my soul be avenged on such a nation as this?
\vs Jer 9:10 For the mountains will I take up a weeping and wailing, and for the habitations of the wilderness a lamentation, because they are burned up, so that none can pass through \bibemph{them}; neither can \bibemph{men} hear the voice of the cattle; both the fowl of the heavens and the beast are fled; they are gone.
\fn{habitations: or, pastures. burned up: or, desolate. both...: Heb. from the fowl even to, etc.}
\vs Jer 9:11 And I will make Jerusalem heaps, \bibemph{and} a den of dragons; and I will make the cities of Judah desolate, without an inhabitant.
\fn{desolate: Heb. desolation.}
\vs Jer 9:12 Who \bibemph{is} the wise man, that may understand this? and \bibemph{who is he} to whom the mouth of the \textsc{Lord} hath spoken, that he may declare it, for what the land perisheth \bibemph{and} is burned up like a wilderness, that none passeth through?
\vs Jer 9:13 And the \textsc{Lord} saith, Because they have forsaken my law which I set before them, and have not obeyed my voice, neither walked therein;
\vs Jer 9:14 But have walked after the imagination of their own heart, and after Baalim, which their fathers taught them:
\fn{imagination: or, stubbornness.}
\vs Jer 9:15 Therefore thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will feed them, \bibemph{even} this people, with wormwood, and give them water of gall to drink.
\vs Jer 9:16 I will scatter them also among the heathen, whom neither they nor their fathers have known: and I will send a sword after them, till I have consumed them.
\vs Jer 9:17 Thus saith the \textsc{Lord} of hosts, Consider ye, and call for the mourning women, that they may come; and send for cunning \bibemph{women}, that they may come:
\vs Jer 9:18 And let them make haste, and take up a wailing for us, that our eyes may run down with tears, and our eyelids gush out with waters.
\vs Jer 9:19 For a voice of wailing is heard out of Zion, How are we spoiled! we are greatly confounded, because we have forsaken the land, because our dwellings have cast \bibemph{us} out.
\vs Jer 9:20 Yet hear the word of the \textsc{Lord}, O ye women, and let your ear receive the word of his mouth, and teach your daughters wailing, and every one her neighbour lamentation.
\vs Jer 9:21 For death is come up into our windows, \bibemph{and} is entered into our palaces, to cut off the children from without, \bibemph{and} the young men from the streets.
\vs Jer 9:22 Speak, Thus saith the \textsc{Lord}, Even the carcases of men shall fall as dung upon the open field, and as the handful after the harvestman, and none shall gather \bibemph{them}.
\vs Jer 9:23 Thus saith the \textsc{Lord}, Let not the wise \bibemph{man} glory in his wisdom, neither let the mighty \bibemph{man} glory in his might, let not the rich \bibemph{man} glory in his riches:
\vs Jer 9:24 But let him that glorieth glory in this, that he understandeth and knoweth me, that I \bibemph{am} the \textsc{Lord} which exercise lovingkindness, judgment, and righteousness, in the earth: for in these \bibemph{things} I delight, saith the \textsc{Lord}.
\vs Jer 9:25 Behold, the days come, saith the \textsc{Lord}, that I will punish all \bibemph{them which are} circumcised with the uncircumcised;
\fn{punish: Heb. visit upon.}
\vs Jer 9:26 Egypt, and Judah, and Edom, and the children of Ammon, and Moab, and all \bibemph{that are} in the utmost corners, that dwell in the wilderness: for all \bibemph{these} nations \bibemph{are} uncircumcised, and all the house of Israel \bibemph{are} uncircumcised in the heart.
\fn{in the utmost...: Heb. cut off into corners, or, having the corners of their hair polled.}
\vs Jer 10:1 Hear ye the word which the \textsc{Lord} speaketh unto you, O house of Israel:
\vs Jer 10:2 Thus saith the \textsc{Lord}, Learn not the way of the heathen, and be not dismayed at the signs of heaven; for the heathen are dismayed at them.
\vs Jer 10:3 For the customs of the people \bibemph{are} vain: for \bibemph{one} cutteth a tree out of the forest, the work of the hands of the workman, with the axe.
\fn{customs...: Heb. statutes, or, ordinances are vanity.}
\vs Jer 10:4 They deck it with silver and with gold; they fasten it with nails and with hammers, that it move not.
\vs Jer 10:5 They \bibemph{are} upright as the palm tree, but speak not: they must needs be borne, because they cannot go. Be not afraid of them; for they cannot do evil, neither also \bibemph{is it} in them to do good.
\vs Jer 10:6 Forasmuch as \bibemph{there is} none like unto thee, O \textsc{Lord}; thou \bibemph{art} great, and thy name \bibemph{is} great in might.
\vs Jer 10:7 Who would not fear thee, O King of nations? for to thee doth it appertain: forasmuch as among all the wise \bibemph{men} of the nations, and in all their kingdoms, \bibemph{there is} none like unto thee.
\fn{to...: or, it liketh thee.}
\vs Jer 10:8 But they are altogether brutish and foolish: the stock \bibemph{is} a doctrine of vanities.
\fn{altogether: Heb. in one, or, at once.}
\vs Jer 10:9 Silver spread into plates is brought from Tarshish, and gold from Uphaz, the work of the workman, and of the hands of the founder: blue and purple \bibemph{is} their clothing: they \bibemph{are} all the work of cunning \bibemph{men}.
\vs Jer 10:10 But the \textsc{Lord} \bibemph{is} the true God, he \bibemph{is} the living God, and an everlasting king: at his wrath the earth shall tremble, and the nations shall not be able to abide his indignation.
\fn{true...: Heb. God of truth. everlasting...: Heb. king of eternity.}
\vs Jer 10:11 Thus shall ye say unto them, The gods that have not made the heavens and the earth, \bibemph{even} they shall perish from the earth, and from under these heavens.
\vs Jer 10:12 He hath made the earth by his power, he hath established the world by his wisdom, and hath stretched out the heavens by his discretion.
\vs Jer 10:13 When he uttereth his voice, \bibemph{there is} a multitude of waters in the heavens, and he causeth the vapours to ascend from the ends of the earth; he maketh lightnings with rain, and bringeth forth the wind out of his treasures.
\fn{multitude: or, noise. with: or, for.}
\vs Jer 10:14 Every man is brutish in \bibemph{his} knowledge: every founder is confounded by the graven image: for his molten image \bibemph{is} falsehood, and \bibemph{there is} no breath in them.
\fn{brutish in his knowledge: or, more brutish than to know.}
\vs Jer 10:15 They \bibemph{are} vanity, \bibemph{and} the work of errors: in the time of their visitation they shall perish.
\vs Jer 10:16 The portion of Jacob \bibemph{is} not like them: for he \bibemph{is} the former of all \bibemph{things}; and Israel \bibemph{is} the rod of his inheritance: The \textsc{Lord} of hosts \bibemph{is} his name.
\vs Jer 10:17 Gather up thy wares out of the land, O inhabitant of the fortress.
\fn{inhabitant: Heb. inhabitress.}
\vs Jer 10:18 For thus saith the \textsc{Lord}, Behold, I will sling out the inhabitants of the land at this once, and will distress them, that they may find \bibemph{it so}.
\vs Jer 10:19 Woe is me for my hurt! my wound is grievous: but I said, Truly this \bibemph{is} a grief, and I must bear it.
\vs Jer 10:20 My tabernacle is spoiled, and all my cords are broken: my children are gone forth of me, and they \bibemph{are} not: \bibemph{there is} none to stretch forth my tent any more, and to set up my curtains.
\vs Jer 10:21 For the pastors are become brutish, and have not sought the \textsc{Lord}: therefore they shall not prosper, and all their flocks shall be scattered.
\vs Jer 10:22 Behold, the noise of the bruit is come, and a great commotion out of the north country, to make the cities of Judah desolate, \bibemph{and} a den of dragons.
\vs Jer 10:23 O \textsc{Lord}, I know that the way of man \bibemph{is} not in himself: \bibemph{it is} not in man that walketh to direct his steps.
\vs Jer 10:24 O \textsc{Lord}, correct me, but with judgment; not in thine anger, lest thou bring me to nothing.
\fn{bring...: Heb. diminish me.}
\vs Jer 10:25 Pour out thy fury upon the heathen that know thee not, and upon the families that call not on thy name: for they have eaten up Jacob, and devoured him, and consumed him, and have made his habitation desolate.
\vs Jer 11:1 The word that came to Jeremiah from the \textsc{Lord}, saying,
\vs Jer 11:2 Hear ye the words of this covenant, and speak unto the men of Judah, and to the inhabitants of Jerusalem;
\vs Jer 11:3 And say thou unto them, Thus saith the \textsc{Lord} God of Israel; Cursed \bibemph{be} the man that obeyeth not the words of this covenant,
\vs Jer 11:4 Which I commanded your fathers in the day \bibemph{that} I brought them forth out of the land of Egypt, from the iron furnace, saying, Obey my voice, and do them, according to all which I command you: so shall ye be my people, and I will be your God:
\vs Jer 11:5 That I may perform the oath which I have sworn unto your fathers, to give them a land flowing with milk and honey, as \bibemph{it is} this day. Then answered I, and said, So be it, O \textsc{Lord}.
\fn{So...: Heb. Amen.}
\vs Jer 11:6 Then the \textsc{Lord} said unto me, Proclaim all these words in the cities of Judah, and in the streets of Jerusalem, saying, Hear ye the words of this covenant, and do them.
\vs Jer 11:7 For I earnestly protested unto your fathers in the day \bibemph{that} I brought them up out of the land of Egypt, \bibemph{even} unto this day, rising early and protesting, saying, Obey my voice.
\vs Jer 11:8 Yet they obeyed not, nor inclined their ear, but walked every one in the imagination of their evil heart: therefore I will bring upon them all the words of this covenant, which I commanded \bibemph{them} to do; but they did \bibemph{them} not.
\fn{imagination: or, stubbornness.}
\vs Jer 11:9 And the \textsc{Lord} said unto me, A conspiracy is found among the men of Judah, and among the inhabitants of Jerusalem.
\vs Jer 11:10 They are turned back to the iniquities of their forefathers, which refused to hear my words; and they went after other gods to serve them: the house of Israel and the house of Judah have broken my covenant which I made with their fathers.
\vs Jer 11:11 Therefore thus saith the \textsc{Lord}, Behold, I will bring evil upon them, which they shall not be able to escape; and though they shall cry unto me, I will not hearken unto them.
\fn{to escape: Heb. to go forth of.}
\vs Jer 11:12 Then shall the cities of Judah and inhabitants of Jerusalem go, and cry unto the gods unto whom they offer incense: but they shall not save them at all in the time of their trouble.
\fn{trouble: Heb. evil.}
\vs Jer 11:13 For \bibemph{according to} the number of thy cities were thy gods, O Judah; and \bibemph{according to} the number of the streets of Jerusalem have ye set up altars to \bibemph{that} shameful thing, \bibemph{even} altars to burn incense unto Baal.
\fn{shameful...: Heb. shame.}
\vs Jer 11:14 Therefore pray not thou for this people, neither lift up a cry or prayer for them: for I will not hear \bibemph{them} in the time that they cry unto me for their trouble.
\fn{trouble: Heb. evil.}
\vs Jer 11:15 What hath my beloved to do in mine house, \bibemph{seeing} she hath wrought lewdness with many, and the holy flesh is passed from thee? when thou doest evil, then thou rejoicest.
\fn{What...: Heb. What is to my beloved in my house. when...: or, when thy evil is.}
\vs Jer 11:16 The \textsc{Lord} called thy name, A green olive tree, fair, \bibemph{and} of goodly fruit: with the noise of a great tumult he hath kindled fire upon it, and the branches of it are broken.
\vs Jer 11:17 For the \textsc{Lord} of hosts, that planted thee, hath pronounced evil against thee, for the evil of the house of Israel and of the house of Judah, which they have done against themselves to provoke me to anger in offering incense unto Baal.
\vs Jer 11:18 And the \textsc{Lord} hath given me knowledge \bibemph{of it}, and I know \bibemph{it}: then thou shewedst me their doings.
\vs Jer 11:19 But I \bibemph{was} like a lamb \bibemph{or} an ox \bibemph{that} is brought to the slaughter; and I knew not that they had devised devices against me, \bibemph{saying}, Let us destroy the tree with the fruit thereof, and let us cut him off from the land of the living, that his name may be no more remembered.
\fn{the tree...: Heb. the stalk with his bread.}
\vs Jer 11:20 But, O \textsc{Lord} of hosts, that judgest righteously, that triest the reins and the heart, let me see thy vengeance on them: for unto thee have I revealed my cause.
\vs Jer 11:21 Therefore thus saith the \textsc{Lord} of the men of Anathoth, that seek thy life, saying, Prophesy not in the name of the \textsc{Lord}, that thou die not by our hand:
\vs Jer 11:22 Therefore thus saith the \textsc{Lord} of hosts, Behold, I will punish them: the young men shall die by the sword; their sons and their daughters shall die by famine:
\fn{punish: Heb. visit upon.}
\vs Jer 11:23 And there shall be no remnant of them: for I will bring evil upon the men of Anathoth, \bibemph{even} the year of their visitation.
\vs Jer 12:1 Righteous \bibemph{art} thou, O \textsc{Lord}, when I plead with thee: yet let me talk with thee of \bibemph{thy} judgments: Wherefore doth the way of the wicked prosper? \bibemph{wherefore} are all they happy that deal very treacherously?
\fn{talk...: or, reason the case with thee.}
\vs Jer 12:2 Thou hast planted them, yea, they have taken root: they grow, yea, they bring forth fruit: thou \bibemph{art} near in their mouth, and far from their reins.
\fn{they grow: Heb. they go on.}
\vs Jer 12:3 But thou, O \textsc{Lord}, knowest me: thou hast seen me, and tried mine heart toward thee: pull them out like sheep for the slaughter, and prepare them for the day of slaughter.
\fn{toward...: Heb. with thee.}
\vs Jer 12:4 How long shall the land mourn, and the herbs of every field wither, for the wickedness of them that dwell therein? the beasts are consumed, and the birds; because they said, He shall not see our last end.
\vs Jer 12:5 If thou hast run with the footmen, and they have wearied thee, then how canst thou contend with horses? and \bibemph{if} in the land of peace, \bibemph{wherein} thou trustedst, \bibemph{they wearied thee}, then how wilt thou do in the swelling of Jordan?
\vs Jer 12:6 For even thy brethren, and the house of thy father, even they have dealt treacherously with thee; yea, they have called a multitude after thee: believe them not, though they speak fair words unto thee.
\fn{they have called...: or, they cried after thee fully. fair...: Heb. good things.}
\vs Jer 12:7 I have forsaken mine house, I have left mine heritage; I have given the dearly beloved of my soul into the hand of her enemies.
\fn{the dearly...: Heb. the love.}
\vs Jer 12:8 Mine heritage is unto me as a lion in the forest; it crieth out against me: therefore have I hated it.
\fn{crieth...: or, yelleth: Heb. giveth out his voice.}
\vs Jer 12:9 Mine heritage \bibemph{is} unto me \bibemph{as} a speckled bird, the birds round about \bibemph{are} against her; come ye, assemble all the beasts of the field, come to devour.
\fn{speckled: or, taloned. come to: or, cause them to come to.}
\vs Jer 12:10 Many pastors have destroyed my vineyard, they have trodden my portion under foot, they have made my pleasant portion a desolate wilderness.
\fn{pleasant...: Heb. portion of desire.}
\vs Jer 12:11 They have made it desolate, \bibemph{and being} desolate it mourneth unto me; the whole land is made desolate, because no man layeth \bibemph{it} to heart.
\vs Jer 12:12 The spoilers are come upon all high places through the wilderness: for the sword of the \textsc{Lord} shall devour from the \bibemph{one} end of the land even to the \bibemph{other} end of the land: no flesh shall have peace.
\vs Jer 12:13 They have sown wheat, but shall reap thorns: they have put themselves to pain, \bibemph{but} shall not profit: and they shall be ashamed of your revenues because of the fierce anger of the \textsc{Lord}.
\fn{they shall: or, ye shall.}
\vs Jer 12:14 Thus saith the \textsc{Lord} against all mine evil neighbours, that touch the inheritance which I have caused my people Israel to inherit; Behold, I will pluck them out of their land, and pluck out the house of Judah from among them.
\vs Jer 12:15 And it shall come to pass, after that I have plucked them out I will return, and have compassion on them, and will bring them again, every man to his heritage, and every man to his land.
\vs Jer 12:16 And it shall come to pass, if they will diligently learn the ways of my people, to swear by my name, The \textsc{Lord} liveth; as they taught my people to swear by Baal; then shall they be built in the midst of my people.
\vs Jer 12:17 But if they will not obey, I will utterly pluck up and destroy that nation, saith the \textsc{Lord}.
\vs Jer 13:1 Thus saith the \textsc{Lord} unto me, Go and get thee a linen girdle, and put it upon thy loins, and put it not in water.
\vs Jer 13:2 So I got a girdle according to the word of the \textsc{Lord}, and put \bibemph{it} on my loins.
\vs Jer 13:3 And the word of the \textsc{Lord} came unto me the second time, saying,
\vs Jer 13:4 Take the girdle that thou hast got, which \bibemph{is} upon thy loins, and arise, go to Euphrates, and hide it there in a hole of the rock.
\vs Jer 13:5 So I went, and hid it by Euphrates, as the \textsc{Lord} commanded me.
\vs Jer 13:6 And it came to pass after many days, that the \textsc{Lord} said unto me, Arise, go to Euphrates, and take the girdle from thence, which I commanded thee to hide there.
\vs Jer 13:7 Then I went to Euphrates, and digged, and took the girdle from the place where I had hid it: and, behold, the girdle was marred, it was profitable for nothing.
\vs Jer 13:8 Then the word of the \textsc{Lord} came unto me, saying,
\vs Jer 13:9 Thus saith the \textsc{Lord}, After this manner will I mar the pride of Judah, and the great pride of Jerusalem.
\vs Jer 13:10 This evil people, which refuse to hear my words, which walk in the imagination of their heart, and walk after other gods, to serve them, and to worship them, shall even be as this girdle, which is good for nothing.
\fn{imagination: or, stubbornness.}
\vs Jer 13:11 For as the girdle cleaveth to the loins of a man, so have I caused to cleave unto me the whole house of Israel and the whole house of Judah, saith the \textsc{Lord}; that they might be unto me for a people, and for a name, and for a praise, and for a glory: but they would not hear.
\vs Jer 13:12 Therefore thou shalt speak unto them this word; Thus saith the \textsc{Lord} God of Israel, Every bottle shall be filled with wine: and they shall say unto thee, Do we not certainly know that every bottle shall be filled with wine?
\vs Jer 13:13 Then shalt thou say unto them, Thus saith the \textsc{Lord}, Behold, I will fill all the inhabitants of this land, even the kings that sit upon David's throne, and the priests, and the prophets, and all the inhabitants of Jerusalem, with drunkenness.
\vs Jer 13:14 And I will dash them one against another, even the fathers and the sons together, saith the \textsc{Lord}: I will not pity, nor spare, nor have mercy, but destroy them.
\fn{one...: Heb. a man against his brother. but...: Heb. from destroying them.}
\vs Jer 13:15 Hear ye, and give ear; be not proud: for the \textsc{Lord} hath spoken.
\vs Jer 13:16 Give glory to the \textsc{Lord} your God, before he cause darkness, and before your feet stumble upon the dark mountains, and, while ye look for light, he turn it into the shadow of death, \bibemph{and} make \bibemph{it} gross darkness.
\vs Jer 13:17 But if ye will not hear it, my soul shall weep in secret places for \bibemph{your} pride; and mine eye shall weep sore, and run down with tears, because the \textsc{Lord's} flock is carried away captive.
\vs Jer 13:18 Say unto the king and to the queen, Humble yourselves, sit down: for your principalities shall come down, \bibemph{even} the crown of your glory.
\fn{principalities: or, head tires.}
\vs Jer 13:19 The cities of the south shall be shut up, and none shall open \bibemph{them}: Judah shall be carried away captive all of it, it shall be wholly carried away captive.
\vs Jer 13:20 Lift up your eyes, and behold them that come from the north: where \bibemph{is} the flock \bibemph{that} was given thee, thy beautiful flock?
\vs Jer 13:21 What wilt thou say when he shall punish thee? for thou hast taught them \bibemph{to be} captains, \bibemph{and} as chief over thee: shall not sorrows take thee, as a woman in travail?
\fn{punish: Heb visit upon.}
\vs Jer 13:22 And if thou say in thine heart, Wherefore come these things upon me? For the greatness of thine iniquity are thy skirts discovered, \bibemph{and} thy heels made bare.
\fn{made...: or, shall be violently taken away.}
\vs Jer 13:23 Can the Ethiopian change his skin, or the leopard his spots? \bibemph{then} may ye also do good, that are accustomed to do evil.
\fn{accustomed: Heb. taught.}
\vs Jer 13:24 Therefore will I scatter them as the stubble that passeth away by the wind of the wilderness.
\vs Jer 13:25 This \bibemph{is} thy lot, the portion of thy measures from me, saith the \textsc{Lord}; because thou hast forgotten me, and trusted in falsehood.
\vs Jer 13:26 Therefore will I discover thy skirts upon thy face, that thy shame may appear.
\vs Jer 13:27 I have seen thine adulteries, and thy neighings, the lewdness of thy whoredom, \bibemph{and} thine abominations on the hills in the fields. Woe unto thee, O Jerusalem! wilt thou not be made clean? when \bibemph{shall it} once \bibemph{be}?
\fn{when...: Heb. after when yet?.}
\vs Jer 14:1 The word of the \textsc{Lord} that came to Jeremiah concerning the dearth.
\fn{the dearth: Heb. the words of the dearths, or, restraints.}
\vs Jer 14:2 Judah mourneth, and the gates thereof languish; they are black unto the ground; and the cry of Jerusalem is gone up.
\vs Jer 14:3 And their nobles have sent their little ones to the waters: they came to the pits, \bibemph{and} found no water; they returned with their vessels empty; they were ashamed and confounded, and covered their heads.
\vs Jer 14:4 Because the ground is chapt, for there was no rain in the earth, the plowmen were ashamed, they covered their heads.
\vs Jer 14:5 Yea, the hind also calved in the field, and forsook \bibemph{it}, because there was no grass.
\vs Jer 14:6 And the wild asses did stand in the high places, they snuffed up the wind like dragons; their eyes did fail, because \bibemph{there was} no grass.
\vs Jer 14:7 O \textsc{Lord}, though our iniquities testify against us, do thou \bibemph{it} for thy name's sake: for our backslidings are many; we have sinned against thee.
\vs Jer 14:8 O the hope of Israel, the saviour thereof in time of trouble, why shouldest thou be as a stranger in the land, and as a wayfaring man \bibemph{that} turneth aside to tarry for a night?
\vs Jer 14:9 Why shouldest thou be as a man astonied, as a mighty man \bibemph{that} cannot save? yet thou, O \textsc{Lord}, \bibemph{art} in the midst of us, and we are called by thy name; leave us not.
\fn{we...: Heb. thy name is called upon us.}
\vs Jer 14:10 Thus saith the \textsc{Lord} unto this people, Thus have they loved to wander, they have not refrained their feet, therefore the \textsc{Lord} doth not accept them; he will now remember their iniquity, and visit their sins.
\vs Jer 14:11 Then said the \textsc{Lord} unto me, Pray not for this people for \bibemph{their} good.
\vs Jer 14:12 When they fast, I will not hear their cry; and when they offer burnt offering and an oblation, I will not accept them: but I will consume them by the sword, and by the famine, and by the pestilence.
\vs Jer 14:13 Then said I, Ah, Lord GOD! behold, the prophets say unto them, Ye shall not see the sword, neither shall ye have famine; but I will give you assured peace in this place.
\fn{assured...: Heb. peace of truth.}
\vs Jer 14:14 Then the \textsc{Lord} said unto me, The prophets prophesy lies in my name: I sent them not, neither have I commanded them, neither spake unto them: they prophesy unto you a false vision and divination, and a thing of nought, and the deceit of their heart.
\vs Jer 14:15 Therefore thus saith the \textsc{Lord} concerning the prophets that prophesy in my name, and I sent them not, yet they say, Sword and famine shall not be in this land; By sword and famine shall those prophets be consumed.
\vs Jer 14:16 And the people to whom they prophesy shall be cast out in the streets of Jerusalem because of the famine and the sword; and they shall have none to bury them, them, their wives, nor their sons, nor their daughters: for I will pour their wickedness upon them.
\vs Jer 14:17 Therefore thou shalt say this word unto them; Let mine eyes run down with tears night and day, and let them not cease: for the virgin daughter of my people is broken with a great breach, with a very grievous blow.
\vs Jer 14:18 If I go forth into the field, then behold the slain with the sword! and if I enter into the city, then behold them that are sick with famine! yea, both the prophet and the priest go about into a land that they know not.
\fn{go about...: or, make merchandise against a land, and men acknowledge it not.}
\vs Jer 14:19 Hast thou utterly rejected Judah? hath thy soul lothed Zion? why hast thou smitten us, and \bibemph{there is} no healing for us? we looked for peace, and \bibemph{there is} no good; and for the time of healing, and behold trouble!
\vs Jer 14:20 We acknowledge, O \textsc{Lord}, our wickedness, \bibemph{and} the iniquity of our fathers: for we have sinned against thee.
\vs Jer 14:21 Do not abhor \bibemph{us}, for thy name's sake, do not disgrace the throne of thy glory: remember, break not thy covenant with us.
\vs Jer 14:22 Are there \bibemph{any} among the vanities of the Gentiles that can cause rain? or can the heavens give showers? \bibemph{art} not thou he, O \textsc{Lord} our God? therefore we will wait upon thee: for thou hast made all these \bibemph{things}.
\vs Jer 15:1 Then said the \textsc{Lord} unto me, Though Moses and Samuel stood before me, \bibemph{yet} my mind \bibemph{could} not \bibemph{be} toward this people: cast \bibemph{them} out of my sight, and let them go forth.
\vs Jer 15:2 And it shall come to pass, if they say unto thee, Whither shall we go forth? then thou shalt tell them, Thus saith the \textsc{Lord}; Such as \bibemph{are} for death, to death; and such as \bibemph{are} for the sword, to the sword; and such as \bibemph{are} for the famine, to the famine; and such as \bibemph{are} for the captivity, to the captivity.
\vs Jer 15:3 And I will appoint over them four kinds, saith the \textsc{Lord}: the sword to slay, and the dogs to tear, and the fowls of the heaven, and the beasts of the earth, to devour and destroy.
\fn{kinds: Heb. families.}
\vs Jer 15:4 And I will cause them to be removed into all kingdoms of the earth, because of Manasseh the son of Hezekiah king of Judah, for \bibemph{that} which he did in Jerusalem.
\fn{cause...: Heb. give them for a removing.}
\vs Jer 15:5 For who shall have pity upon thee, O Jerusalem? or who shall bemoan thee? or who shall go aside to ask how thou doest?
\fn{how...: Heb. of thy peace?.}
\vs Jer 15:6 Thou hast forsaken me, saith the \textsc{Lord}, thou art gone backward: therefore will I stretch out my hand against thee, and destroy thee; I am weary with repenting.
\vs Jer 15:7 And I will fan them with a fan in the gates of the land; I will bereave \bibemph{them} of children, I will destroy my people, \bibemph{since} they return not from their ways.
\fn{children: or, whatsoever is dear.}
\vs Jer 15:8 Their widows are increased to me above the sand of the seas: I have brought upon them against the mother of the young men a spoiler at noonday: I have caused \bibemph{him} to fall upon it suddenly, and terrors upon the city.
\fn{the mother...: or, the mother city a young man spoiling, etc, or, the mother and the young men.}
\vs Jer 15:9 She that hath borne seven languisheth: she hath given up the ghost; her sun is gone down while \bibemph{it was} yet day: she hath been ashamed and confounded: and the residue of them will I deliver to the sword before their enemies, saith the \textsc{Lord}.
\vs Jer 15:10 Woe is me, my mother, that thou hast borne me a man of strife and a man of contention to the whole earth! I have neither lent on usury, nor men have lent to me on usury; \bibemph{yet} every one of them doth curse me.
\vs Jer 15:11 The \textsc{Lord} said, Verily it shall be well with thy remnant; verily I will cause the enemy to entreat thee \bibemph{well} in the time of evil and in the time of affliction.
\fn{cause...: or, intreat the enemy for thee.}
\vs Jer 15:12 Shall iron break the northern iron and the steel?
\vs Jer 15:13 Thy substance and thy treasures will I give to the spoil without price, and \bibemph{that} for all thy sins, even in all thy borders.
\vs Jer 15:14 And I will make \bibemph{thee} to pass with thine enemies into a land \bibemph{which} thou knowest not: for a fire is kindled in mine anger, \bibemph{which} shall burn upon you.
\vs Jer 15:15 O \textsc{Lord}, thou knowest: remember me, and visit me, and revenge me of my persecutors; take me not away in thy longsuffering: know that for thy sake I have suffered rebuke.
\vs Jer 15:16 Thy words were found, and I did eat them; and thy word was unto me the joy and rejoicing of mine heart: for I am called by thy name, O \textsc{Lord} God of hosts.
\fn{I am...: Heb. thy name is called upon me.}
\vs Jer 15:17 I sat not in the assembly of the mockers, nor rejoiced; I sat alone because of thy hand: for thou hast filled me with indignation.
\vs Jer 15:18 Why is my pain perpetual, and my wound incurable, \bibemph{which} refuseth to be healed? wilt thou be altogether unto me as a liar, \bibemph{and as} waters \bibemph{that} fail?
\fn{fail: Heb. be not sure?.}
\vs Jer 15:19 Therefore thus saith the \textsc{Lord}, If thou return, then will I bring thee again, \bibemph{and} thou shalt stand before me: and if thou take forth the precious from the vile, thou shalt be as my mouth: let them return unto thee; but return not thou unto them.
\vs Jer 15:20 And I will make thee unto this people a fenced brasen wall: and they shall fight against thee, but they shall not prevail against thee: for I \bibemph{am} with thee to save thee and to deliver thee, saith the \textsc{Lord}.
\vs Jer 15:21 And I will deliver thee out of the hand of the wicked, and I will redeem thee out of the hand of the terrible.
\vs Jer 16:1 The word of the \textsc{Lord} came also unto me, saying,
\vs Jer 16:2 Thou shalt not take thee a wife, neither shalt thou have sons or daughters in this place.
\vs Jer 16:3 For thus saith the \textsc{Lord} concerning the sons and concerning the daughters that are born in this place, and concerning their mothers that bare them, and concerning their fathers that begat them in this land;
\vs Jer 16:4 They shall die of grievous deaths; they shall not be lamented; neither shall they be buried; \bibemph{but} they shall be as dung upon the face of the earth: and they shall be consumed by the sword, and by famine; and their carcases shall be meat for the fowls of heaven, and for the beasts of the earth.
\vs Jer 16:5 For thus saith the \textsc{Lord}, Enter not into the house of mourning, neither go to lament nor bemoan them: for I have taken away my peace from this people, saith the \textsc{Lord}, \bibemph{even} lovingkindness and mercies.
\fn{mourning: or, mourning feast.}
\vs Jer 16:6 Both the great and the small shall die in this land: they shall not be buried, neither shall \bibemph{men} lament for them, nor cut themselves, nor make themselves bald for them:
\vs Jer 16:7 Neither shall \bibemph{men} tear \bibemph{themselves} for them in mourning, to comfort them for the dead; neither shall \bibemph{men} give them the cup of consolation to drink for their father or for their mother.
\fn{tear...: or, break bread for them.}
\vs Jer 16:8 Thou shalt not also go into the house of feasting, to sit with them to eat and to drink.
\vs Jer 16:9 For thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will cause to cease out of this place in your eyes, and in your days, the voice of mirth, and the voice of gladness, the voice of the bridegroom, and the voice of the bride.
\vs Jer 16:10 And it shall come to pass, when thou shalt shew this people all these words, and they shall say unto thee, Wherefore hath the \textsc{Lord} pronounced all this great evil against us? or what \bibemph{is} our iniquity? or what \bibemph{is} our sin that we have committed against the \textsc{Lord} our God?
\vs Jer 16:11 Then shalt thou say unto them, Because your fathers have forsaken me, saith the \textsc{Lord}, and have walked after other gods, and have served them, and have worshipped them, and have forsaken me, and have not kept my law;
\vs Jer 16:12 And ye have done worse than your fathers; for, behold, ye walk every one after the imagination of his evil heart, that they may not hearken unto me:
\fn{imagination: or, stubbornness.}
\vs Jer 16:13 Therefore will I cast you out of this land into a land that ye know not, \bibemph{neither} ye nor your fathers; and there shall ye serve other gods day and night; where I will not shew you favour.
\vs Jer 16:14 Therefore, behold, the days come, saith the \textsc{Lord}, that it shall no more be said, The \textsc{Lord} liveth, that brought up the children of Israel out of the land of Egypt;
\vs Jer 16:15 But, The \textsc{Lord} liveth, that brought up the children of Israel from the land of the north, and from all the lands whither he had driven them: and I will bring them again into their land that I gave unto their fathers.
\vs Jer 16:16 Behold, I will send for many fishers, saith the \textsc{Lord}, and they shall fish them; and after will I send for many hunters, and they shall hunt them from every mountain, and from every hill, and out of the holes of the rocks.
\vs Jer 16:17 For mine eyes \bibemph{are} upon all their ways: they are not hid from my face, neither is their iniquity hid from mine eyes.
\vs Jer 16:18 And first I will recompense their iniquity and their sin double; because they have defiled my land, they have filled mine inheritance with the carcases of their detestable and abominable things.
\vs Jer 16:19 O \textsc{Lord}, my strength, and my fortress, and my refuge in the day of affliction, the Gentiles shall come unto thee from the ends of the earth, and shall say, Surely our fathers have inherited lies, vanity, and \bibemph{things} wherein \bibemph{there is} no profit.
\vs Jer 16:20 Shall a man make gods unto himself, and they \bibemph{are} no gods?
\vs Jer 16:21 Therefore, behold, I will this once cause them to know, I will cause them to know mine hand and my might; and they shall know that my name \bibemph{is} The \textsc{Lord}.
\fn{The \textsc{Lord}: or, JEHOVAH.}
\vs Jer 17:1 The sin of Judah \bibemph{is} written with a pen of iron, \bibemph{and} with the point of a diamond: \bibemph{it is} graven upon the table of their heart, and upon the horns of your altars;
\fn{point: Heb. nail.}
\vs Jer 17:2 Whilst their children remember their altars and their groves by the green trees upon the high hills.
\vs Jer 17:3 O my mountain in the field, I will give thy substance \bibemph{and} all thy treasures to the spoil, \bibemph{and} thy high places for sin, throughout all thy borders.
\vs Jer 17:4 And thou, even thyself, shalt discontinue from thine heritage that I gave thee; and I will cause thee to serve thine enemies in the land which thou knowest not: for ye have kindled a fire in mine anger, \bibemph{which} shall burn for ever.
\fn{thyself: Heb. in thyself.}
\vs Jer 17:5 Thus saith the \textsc{Lord}; Cursed \bibemph{be} the man that trusteth in man, and maketh flesh his arm, and whose heart departeth from the \textsc{Lord}.
\vs Jer 17:6 For he shall be like the heath in the desert, and shall not see when good cometh; but shall inhabit the parched places in the wilderness, \bibemph{in} a salt land and not inhabited.
\vs Jer 17:7 Blessed \bibemph{is} the man that trusteth in the \textsc{Lord}, and whose hope the \textsc{Lord} is.
\vs Jer 17:8 For he shall be as a tree planted by the waters, and \bibemph{that} spreadeth out her roots by the river, and shall not see when heat cometh, but her leaf shall be green; and shall not be careful in the year of drought, neither shall cease from yielding fruit.
\fn{drought: or, restraint.}
\vs Jer 17:9 The heart \bibemph{is} deceitful above all \bibemph{things}, and desperately wicked: who can know it?
\vs Jer 17:10 I the \textsc{Lord} search the heart, \bibemph{I} try the reins, even to give every man according to his ways, \bibemph{and} according to the fruit of his doings.
\vs Jer 17:11 \bibemph{As} the partridge sitteth \bibemph{on eggs}, and hatcheth \bibemph{them} not; \bibemph{so} he that getteth riches, and not by right, shall leave them in the midst of his days, and at his end shall be a fool.
\fn{sitteth...: or, gathereth young which she hath not brought forth.}
\vs Jer 17:12 A glorious high throne from the beginning \bibemph{is} the place of our sanctuary.
\vs Jer 17:13 O \textsc{Lord}, the hope of Israel, all that forsake thee shall be ashamed, \bibemph{and} they that depart from me shall be written in the earth, because they have forsaken the \textsc{Lord}, the fountain of living waters.
\vs Jer 17:14 Heal me, O \textsc{Lord}, and I shall be healed; save me, and I shall be saved: for thou \bibemph{art} my praise.
\vs Jer 17:15 Behold, they say unto me, Where \bibemph{is} the word of the \textsc{Lord}? let it come now.
\vs Jer 17:16 As for me, I have not hastened from \bibemph{being} a pastor to follow thee: neither have I desired the woeful day; thou knowest: that which came out of my lips was \bibemph{right} before thee.
\fn{to...: Heb. after thee.}
\vs Jer 17:17 Be not a terror unto me: thou \bibemph{art} my hope in the day of evil.
\vs Jer 17:18 Let them be confounded that persecute me, but let not me be confounded: let them be dismayed, but let not me be dismayed: bring upon them the day of evil, and destroy them with double destruction.
\fn{destroy...: Heb. break them with a double breach.}
\vs Jer 17:19 Thus said the \textsc{Lord} unto me; Go and stand in the gate of the children of the people, whereby the kings of Judah come in, and by the which they go out, and in all the gates of Jerusalem;
\vs Jer 17:20 And say unto them, Hear ye the word of the \textsc{Lord}, ye kings of Judah, and all Judah, and all the inhabitants of Jerusalem, that enter in by these gates:
\vs Jer 17:21 Thus saith the \textsc{Lord}; Take heed to yourselves, and bear no burden on the sabbath day, nor bring \bibemph{it} in by the gates of Jerusalem;
\vs Jer 17:22 Neither carry forth a burden out of your houses on the sabbath day, neither do ye any work, but hallow ye the sabbath day, as I commanded your fathers.
\vs Jer 17:23 But they obeyed not, neither inclined their ear, but made their neck stiff, that they might not hear, nor receive instruction.
\vs Jer 17:24 And it shall come to pass, if ye diligently hearken unto me, saith the \textsc{Lord}, to bring in no burden through the gates of this city on the sabbath day, but hallow the sabbath day, to do no work therein;
\vs Jer 17:25 Then shall there enter into the gates of this city kings and princes sitting upon the throne of David, riding in chariots and on horses, they, and their princes, the men of Judah, and the inhabitants of Jerusalem: and this city shall remain for ever.
\vs Jer 17:26 And they shall come from the cities of Judah, and from the places about Jerusalem, and from the land of Benjamin, and from the plain, and from the mountains, and from the south, bringing burnt offerings, and sacrifices, and meat offerings, and incense, and bringing sacrifices of praise, unto the house of the \textsc{Lord}.
\vs Jer 17:27 But if ye will not hearken unto me to hallow the sabbath day, and not to bear a burden, even entering in at the gates of Jerusalem on the sabbath day; then will I kindle a fire in the gates thereof, and it shall devour the palaces of Jerusalem, and it shall not be quenched.
\vs Jer 18:1 The word which came to Jeremiah from the \textsc{Lord}, saying,
\vs Jer 18:2 Arise, and go down to the potter's house, and there I will cause thee to hear my words.
\vs Jer 18:3 Then I went down to the potter's house, and, behold, he wrought a work on the wheels.
\fn{wheels: or, frames, or, seats.}
\vs Jer 18:4 And the vessel that he made of clay was marred in the hand of the potter: so he made it again another vessel, as seemed good to the potter to make \bibemph{it}.
\fn{of clay...: or, was marred, as clay in the hand of the potter. made it...: Heb. returned and made, etc.}
\vs Jer 18:5 Then the word of the \textsc{Lord} came to me, saying,
\vs Jer 18:6 O house of Israel, cannot I do with you as this potter? saith the \textsc{Lord}. Behold, as the clay \bibemph{is} in the potter's hand, so \bibemph{are} ye in mine hand, O house of Israel.
\vs Jer 18:7 \bibemph{At what} instant I shall speak concerning a nation, and concerning a kingdom, to pluck up, and to pull down, and to destroy \bibemph{it};
\vs Jer 18:8 If that nation, against whom I have pronounced, turn from their evil, I will repent of the evil that I thought to do unto them.
\vs Jer 18:9 And \bibemph{at what} instant I shall speak concerning a nation, and concerning a kingdom, to build and to plant \bibemph{it};
\vs Jer 18:10 If it do evil in my sight, that it obey not my voice, then I will repent of the good, wherewith I said I would benefit them.
\vs Jer 18:11 Now therefore go to, speak to the men of Judah, and to the inhabitants of Jerusalem, saying, Thus saith the \textsc{Lord}; Behold, I frame evil against you, and devise a device against you: return ye now every one from his evil way, and make your ways and your doings good.
\vs Jer 18:12 And they said, There is no hope: but we will walk after our own devices, and we will every one do the imagination of his evil heart.
\vs Jer 18:13 Therefore thus saith the \textsc{Lord}; Ask ye now among the heathen, who hath heard such things: the virgin of Israel hath done a very horrible thing.
\vs Jer 18:14 Will \bibemph{a man} leave the snow of Lebanon \bibemph{which cometh} from the rock of the field? \bibemph{or} shall the cold flowing waters that come from another place be forsaken?
\fn{the snow...: or, my fields for a rock, or for the snow of Lebanon? shall the running waters be forsaken for the strange cold waters?.}
\vs Jer 18:15 Because my people hath forgotten me, they have burned incense to vanity, and they have caused them to stumble in their ways \bibemph{from} the ancient paths, to walk in paths, \bibemph{in} a way not cast up;
\vs Jer 18:16 To make their land desolate, \bibemph{and} a perpetual hissing; every one that passeth thereby shall be astonished, and wag his head.
\vs Jer 18:17 I will scatter them as with an east wind before the enemy; I will shew them the back, and not the face, in the day of their calamity.
\vs Jer 18:18 Then said they, Come, and let us devise devices against Jeremiah; for the law shall not perish from the priest, nor counsel from the wise, nor the word from the prophet. Come, and let us smite him with the tongue, and let us not give heed to any of his words.
\fn{with...: or, for the tongue.}
\vs Jer 18:19 Give heed to me, O \textsc{Lord}, and hearken to the voice of them that contend with me.
\vs Jer 18:20 Shall evil be recompensed for good? for they have digged a pit for my soul. Remember that I stood before thee to speak good for them, \bibemph{and} to turn away thy wrath from them.
\vs Jer 18:21 Therefore deliver up their children to the famine, and pour out their \bibemph{blood} by the force of the sword; and let their wives be bereaved of their children, and \bibemph{be} widows; and let their men be put to death; \bibemph{let} their young men \bibemph{be} slain by the sword in battle.
\fn{pour...: Heb. pour them out.}
\vs Jer 18:22 Let a cry be heard from their houses, when thou shalt bring a troop suddenly upon them: for they have digged a pit to take me, and hid snares for my feet.
\vs Jer 18:23 Yet, \textsc{Lord}, thou knowest all their counsel against me to slay \bibemph{me}: forgive not their iniquity, neither blot out their sin from thy sight, but let them be overthrown before thee; deal \bibemph{thus} with them in the time of thine anger.
\fn{to slay...: Heb. for death.}
\vs Jer 19:1 Thus saith the \textsc{Lord}, Go and get a potter's earthen bottle, and \bibemph{take} of the ancients of the people, and of the ancients of the priests;
\vs Jer 19:2 And go forth unto the valley of the son of Hinnom, which \bibemph{is} by the entry of the east gate, and proclaim there the words that I shall tell thee,
\fn{the east...: Heb. the sun gate.}
\vs Jer 19:3 And say, Hear ye the word of the \textsc{Lord}, O kings of Judah, and inhabitants of Jerusalem; Thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will bring evil upon this place, the which whosoever heareth, his ears shall tingle.
\vs Jer 19:4 Because they have forsaken me, and have estranged this place, and have burned incense in it unto other gods, whom neither they nor their fathers have known, nor the kings of Judah, and have filled this place with the blood of innocents;
\vs Jer 19:5 They have built also the high places of Baal, to burn their sons with fire \bibemph{for} burnt offerings unto Baal, which I commanded not, nor spake \bibemph{it}, neither came \bibemph{it} into my mind:
\vs Jer 19:6 Therefore, behold, the days come, saith the \textsc{Lord}, that this place shall no more be called Tophet, nor The valley of the son of Hinnom, but The valley of slaughter.
\vs Jer 19:7 And I will make void the counsel of Judah and Jerusalem in this place; and I will cause them to fall by the sword before their enemies, and by the hands of them that seek their lives: and their carcases will I give to be meat for the fowls of the heaven, and for the beasts of the earth.
\vs Jer 19:8 And I will make this city desolate, and an hissing; every one that passeth thereby shall be astonished and hiss because of all the plagues thereof.
\vs Jer 19:9 And I will cause them to eat the flesh of their sons and the flesh of their daughters, and they shall eat every one the flesh of his friend in the siege and straitness, wherewith their enemies, and they that seek their lives, shall straiten them.
\vs Jer 19:10 Then shalt thou break the bottle in the sight of the men that go with thee,
\vs Jer 19:11 And shalt say unto them, Thus saith the \textsc{Lord} of hosts; Even so will I break this people and this city, as \bibemph{one} breaketh a potter's vessel, that cannot be made whole again: and they shall bury \bibemph{them} in Tophet, till \bibemph{there be} no place to bury.
\fn{be made...: Heb. be healed.}
\vs Jer 19:12 Thus will I do unto this place, saith the \textsc{Lord}, and to the inhabitants thereof, and \bibemph{even} make this city as Tophet:
\vs Jer 19:13 And the houses of Jerusalem, and the houses of the kings of Judah, shall be defiled as the place of Tophet, because of all the houses upon whose roofs they have burned incense unto all the host of heaven, and have poured out drink offerings unto other gods.
\vs Jer 19:14 Then came Jeremiah from Tophet, whither the \textsc{Lord} had sent him to prophesy; and he stood in the court of the \textsc{Lord's} house; and said to all the people,
\vs Jer 19:15 Thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will bring upon this city and upon all her towns all the evil that I have pronounced against it, because they have hardened their necks, that they might not hear my words.
\vs Jer 20:1 Now Pashur the son of Immer the priest, who \bibemph{was} also chief governor in the house of the \textsc{Lord}, heard that Jeremiah prophesied these things.
\vs Jer 20:2 Then Pashur smote Jeremiah the prophet, and put him in the stocks that \bibemph{were} in the high gate of Benjamin, which \bibemph{was} by the house of the \textsc{Lord}.
\vs Jer 20:3 And it came to pass on the morrow, that Pashur brought forth Jeremiah out of the stocks. Then said Jeremiah unto him, The \textsc{Lord} hath not called thy name Pashur, but Magormissabib.
\fn{Magormissabib: that is, Fear round about.}
\vs Jer 20:4 For thus saith the \textsc{Lord}, Behold, I will make thee a terror to thyself, and to all thy friends: and they shall fall by the sword of their enemies, and thine eyes shall behold \bibemph{it}: and I will give all Judah into the hand of the king of Babylon, and he shall carry them captive into Babylon, and shall slay them with the sword.
\vs Jer 20:5 Moreover I will deliver all the strength of this city, and all the labours thereof, and all the precious things thereof, and all the treasures of the kings of Judah will I give into the hand of their enemies, which shall spoil them, and take them, and carry them to Babylon.
\vs Jer 20:6 And thou, Pashur, and all that dwell in thine house shall go into captivity: and thou shalt come to Babylon, and there thou shalt die, and shalt be buried there, thou, and all thy friends, to whom thou hast prophesied lies.
\vs Jer 20:7 O \textsc{Lord}, thou hast deceived me, and I was deceived: thou art stronger than I, and hast prevailed: I am in derision daily, every one mocketh me.
\fn{was deceived: or, was enticed.}
\vs Jer 20:8 For since I spake, I cried out, I cried violence and spoil; because the word of the \textsc{Lord} was made a reproach unto me, and a derision, daily.
\vs Jer 20:9 Then I said, I will not make mention of him, nor speak any more in his name. But \bibemph{his word} was in mine heart as a burning fire shut up in my bones, and I was weary with forbearing, and I could not \bibemph{stay}.
\vs Jer 20:10 For I heard the defaming of many, fear on every side. Report, \bibemph{say they}, and we will report it. All my familiars watched for my halting, \bibemph{saying}, Peradventure he will be enticed, and we shall prevail against him, and we shall take our revenge on him.
\fn{All...: Heb. Every man of my peace.}
\vs Jer 20:11 But the \textsc{Lord} \bibemph{is} with me as a mighty terrible one: therefore my persecutors shall stumble, and they shall not prevail: they shall be greatly ashamed; for they shall not prosper: \bibemph{their} everlasting confusion shall never be forgotten.
\vs Jer 20:12 But, O \textsc{Lord} of hosts, that triest the righteous, \bibemph{and} seest the reins and the heart, let me see thy vengeance on them: for unto thee have I opened my cause.
\vs Jer 20:13 Sing unto the \textsc{Lord}, praise ye the \textsc{Lord}: for he hath delivered the soul of the poor from the hand of evildoers.
\vs Jer 20:14 Cursed \bibemph{be} the day wherein I was born: let not the day wherein my mother bare me be blessed.
\vs Jer 20:15 Cursed \bibemph{be} the man who brought tidings to my father, saying, A man child is born unto thee; making him very glad.
\vs Jer 20:16 And let that man be as the cities which the \textsc{Lord} overthrew, and repented not: and let him hear the cry in the morning, and the shouting at noontide;
\vs Jer 20:17 Because he slew me not from the womb; or that my mother might have been my grave, and her womb \bibemph{to be} always great \bibemph{with me}.
\vs Jer 20:18 Wherefore came I forth out of the womb to see labour and sorrow, that my days should be consumed with shame?
\vs Jer 21:1 The word which came unto Jeremiah from the \textsc{Lord}, when king Zedekiah sent unto him Pashur the son of Melchiah, and Zephaniah the son of Maaseiah the priest, saying,
\vs Jer 21:2 Enquire, I pray thee, of the \textsc{Lord} for us; for Nebuchadrezzar king of Babylon maketh war against us; if so be that the \textsc{Lord} will deal with us according to all his wondrous works, that he may go up from us.
\vs Jer 21:3 Then said Jeremiah unto them, Thus shall ye say to Zedekiah:
\vs Jer 21:4 Thus saith the \textsc{Lord} God of Israel; Behold, I will turn back the weapons of war that \bibemph{are} in your hands, wherewith ye fight against the king of Babylon, and \bibemph{against} the Chaldeans, which besiege you without the walls, and I will assemble them into the midst of this city.
\vs Jer 21:5 And I myself will fight against you with an outstretched hand and with a strong arm, even in anger, and in fury, and in great wrath.
\vs Jer 21:6 And I will smite the inhabitants of this city, both man and beast: they shall die of a great pestilence.
\vs Jer 21:7 And afterward, saith the \textsc{Lord}, I will deliver Zedekiah king of Judah, and his servants, and the people, and such as are left in this city from the pestilence, from the sword, and from the famine, into the hand of Nebuchadrezzar king of Babylon, and into the hand of their enemies, and into the hand of those that seek their life: and he shall smite them with the edge of the sword; he shall not spare them, neither have pity, nor have mercy.
\vs Jer 21:8 And unto this people thou shalt say, Thus saith the \textsc{Lord}; Behold, I set before you the way of life, and the way of death.
\vs Jer 21:9 He that abideth in this city shall die by the sword, and by the famine, and by the pestilence: but he that goeth out, and falleth to the Chaldeans that besiege you, he shall live, and his life shall be unto him for a prey.
\vs Jer 21:10 For I have set my face against this city for evil, and not for good, saith the \textsc{Lord}: it shall be given into the hand of the king of Babylon, and he shall burn it with fire.
\vs Jer 21:11 And touching the house of the king of Judah, \bibemph{say}, Hear ye the word of the \textsc{Lord};
\vs Jer 21:12 O house of David, thus saith the \textsc{Lord}; Execute judgment in the morning, and deliver \bibemph{him that is} spoiled out of the hand of the oppressor, lest my fury go out like fire, and burn that none can quench \bibemph{it}, because of the evil of your doings.
\fn{Execute: Heb. Judge.}
\vs Jer 21:13 Behold, I \bibemph{am} against thee, O inhabitant of the valley, \bibemph{and} rock of the plain, saith the \textsc{Lord}; which say, Who shall come down against us? or who shall enter into our habitations?
\fn{inhabitant: Heb. inhabitress.}
\vs Jer 21:14 But I will punish you according to the fruit of your doings, saith the \textsc{Lord}: and I will kindle a fire in the forest thereof, and it shall devour all things round about it.
\fn{punish: Heb. visit upon.}
\vs Jer 22:1 Thus saith the \textsc{Lord}; Go down to the house of the king of Judah, and speak there this word,
\vs Jer 22:2 And say, Hear the word of the \textsc{Lord}, O king of Judah, that sittest upon the throne of David, thou, and thy servants, and thy people that enter in by these gates:
\vs Jer 22:3 Thus saith the \textsc{Lord}; Execute ye judgment and righteousness, and deliver the spoiled out of the hand of the oppressor: and do no wrong, do no violence to the stranger, the fatherless, nor the widow, neither shed innocent blood in this place.
\vs Jer 22:4 For if ye do this thing indeed, then shall there enter in by the gates of this house kings sitting upon the throne of David, riding in chariots and on horses, he, and his servants, and his people.
\fn{upon...: Heb. for David upon his throne.}
\vs Jer 22:5 But if ye will not hear these words, I swear by myself, saith the \textsc{Lord}, that this house shall become a desolation.
\vs Jer 22:6 For thus saith the \textsc{Lord} unto the king's house of Judah; Thou \bibemph{art} Gilead unto me, \bibemph{and} the head of Lebanon: \bibemph{yet} surely I will make thee a wilderness, \bibemph{and} cities \bibemph{which} are not inhabited.
\vs Jer 22:7 And I will prepare destroyers against thee, every one with his weapons: and they shall cut down thy choice cedars, and cast \bibemph{them} into the fire.
\vs Jer 22:8 And many nations shall pass by this city, and they shall say every man to his neighbour, Wherefore hath the \textsc{Lord} done thus unto this great city?
\vs Jer 22:9 Then they shall answer, Because they have forsaken the covenant of the \textsc{Lord} their God, and worshipped other gods, and served them.
\vs Jer 22:10 Weep ye not for the dead, neither bemoan him: \bibemph{but} weep sore for him that goeth away: for he shall return no more, nor see his native country.
\vs Jer 22:11 For thus saith the \textsc{Lord} touching Shallum the son of Josiah king of Judah, which reigned instead of Josiah his father, which went forth out of this place; He shall not return thither any more:
\vs Jer 22:12 But he shall die in the place whither they have led him captive, and shall see this land no more.
\vs Jer 22:13 Woe unto him that buildeth his house by unrighteousness, and his chambers by wrong; \bibemph{that} useth his neighbour's service without wages, and giveth him not for his work;
\vs Jer 22:14 That saith, I will build me a wide house and large chambers, and cutteth him out windows; and \bibemph{it is} cieled with cedar, and painted with vermilion.
\fn{large: Heb. through-aired. windows: or, my windows.}
\vs Jer 22:15 Shalt thou reign, because thou closest \bibemph{thyself} in cedar? did not thy father eat and drink, and do judgment and justice, \bibemph{and} then \bibemph{it was} well with him?
\vs Jer 22:16 He judged the cause of the poor and needy; then \bibemph{it was} well \bibemph{with him: was} not this to know me? saith the \textsc{Lord}.
\vs Jer 22:17 But thine eyes and thine heart \bibemph{are} not but for thy covetousness, and for to shed innocent blood, and for oppression, and for violence, to do \bibemph{it}.
\fn{violence: or, incursion.}
\vs Jer 22:18 Therefore thus saith the \textsc{Lord} concerning Jehoiakim the son of Josiah king of Judah; They shall not lament for him, \bibemph{saying}, Ah my brother! or, Ah sister! they shall not lament for him, \bibemph{saying}, Ah lord! or, Ah his glory!
\vs Jer 22:19 He shall be buried with the burial of an ass, drawn and cast forth beyond the gates of Jerusalem.
\vs Jer 22:20 Go up to Lebanon, and cry; and lift up thy voice in Bashan, and cry from the passages: for all thy lovers are destroyed.
\vs Jer 22:21 I spake unto thee in thy prosperity; \bibemph{but} thou saidst, I will not hear. This \bibemph{hath been} thy manner from thy youth, that thou obeyedst not my voice.
\fn{prosperity: Heb. prosperities.}
\vs Jer 22:22 The wind shall eat up all thy pastors, and thy lovers shall go into captivity: surely then shalt thou be ashamed and confounded for all thy wickedness.
\vs Jer 22:23 O inhabitant of Lebanon, that makest thy nest in the cedars, how gracious shalt thou be when pangs come upon thee, the pain as of a woman in travail!
\fn{inhabitant: Heb. inhabitress.}
\vs Jer 22:24 \bibemph{As} I live, saith the \textsc{Lord}, though Coniah the son of Jehoiakim king of Judah were the signet upon my right hand, yet would I pluck thee thence;
\vs Jer 22:25 And I will give thee into the hand of them that seek thy life, and into the hand \bibemph{of them} whose face thou fearest, even into the hand of Nebuchadrezzar king of Babylon, and into the hand of the Chaldeans.
\vs Jer 22:26 And I will cast thee out, and thy mother that bare thee, into another country, where ye were not born; and there shall ye die.
\vs Jer 22:27 But to the land whereunto they desire to return, thither shall they not return.
\fn{desire: Heb. lift up their mind.}
\vs Jer 22:28 \bibemph{Is} this man Coniah a despised broken idol? \bibemph{is he} a vessel wherein \bibemph{is} no pleasure? wherefore are they cast out, he and his seed, and are cast into a land which they know not?
\vs Jer 22:29 O earth, earth, earth, hear the word of the \textsc{Lord}.
\vs Jer 22:30 Thus saith the \textsc{Lord}, Write ye this man childless, a man \bibemph{that} shall not prosper in his days: for no man of his seed shall prosper, sitting upon the throne of David, and ruling any more in Judah.
\vs Jer 23:1 Woe be unto the pastors that destroy and scatter the sheep of my pasture! saith the \textsc{Lord}.
\vs Jer 23:2 Therefore thus saith the \textsc{Lord} God of Israel against the pastors that feed my people; Ye have scattered my flock, and driven them away, and have not visited them: behold, I will visit upon you the evil of your doings, saith the \textsc{Lord}.
\vs Jer 23:3 And I will gather the remnant of my flock out of all countries whither I have driven them, and will bring them again to their folds; and they shall be fruitful and increase.
\vs Jer 23:4 And I will set up shepherds over them which shall feed them: and they shall fear no more, nor be dismayed, neither shall they be lacking, saith the \textsc{Lord}.
\vs Jer 23:5 Behold, the days come, saith the \textsc{Lord}, that I will raise unto David a righteous Branch, and a King shall reign and prosper, and shall execute judgment and justice in the earth.
\vs Jer 23:6 In his days Judah shall be saved, and Israel shall dwell safely: and this \bibemph{is} his name whereby he shall be called, THE \textsc{Lord} OUR RIGHTEOUSNESS.
\fn{THE \textsc{Lord}...: Heb. Jehovahtsidkenu.}
\vs Jer 23:7 Therefore, behold, the days come, saith the \textsc{Lord}, that they shall no more say, The \textsc{Lord} liveth, which brought up the children of Israel out of the land of Egypt;
\vs Jer 23:8 But, The \textsc{Lord} liveth, which brought up and which led the seed of the house of Israel out of the north country, and from all countries whither I had driven them; and they shall dwell in their own land.
\vs Jer 23:9 Mine heart within me is broken because of the prophets; all my bones shake; I am like a drunken man, and like a man whom wine hath overcome, because of the \textsc{Lord}, and because of the words of his holiness.
\vs Jer 23:10 For the land is full of adulterers; for because of swearing the land mourneth; the pleasant places of the wilderness are dried up, and their course is evil, and their force \bibemph{is} not right.
\fn{swearing: or, cursing. course: or, violence.}
\vs Jer 23:11 For both prophet and priest are profane; yea, in my house have I found their wickedness, saith the \textsc{Lord}.
\vs Jer 23:12 Wherefore their way shall be unto them as slippery \bibemph{ways} in the darkness: they shall be driven on, and fall therein: for I will bring evil upon them, \bibemph{even} the year of their visitation, saith the \textsc{Lord}.
\vs Jer 23:13 And I have seen folly in the prophets of Samaria; they prophesied in Baal, and caused my people Israel to err.
\fn{folly: or, an absurd thing: Heb. unsavoury.}
\vs Jer 23:14 I have seen also in the prophets of Jerusalem an horrible thing: they commit adultery, and walk in lies: they strengthen also the hands of evildoers, that none doth return from his wickedness: they are all of them unto me as Sodom, and the inhabitants thereof as Gomorrah.
\fn{an...: or, filthiness.}
\vs Jer 23:15 Therefore thus saith the \textsc{Lord} of hosts concerning the prophets; Behold, I will feed them with wormwood, and make them drink the water of gall: for from the prophets of Jerusalem is profaneness gone forth into all the land.
\fn{profaneness: or, hypocrisy.}
\vs Jer 23:16 Thus saith the \textsc{Lord} of hosts, Hearken not unto the words of the prophets that prophesy unto you: they make you vain: they speak a vision of their own heart, \bibemph{and} not out of the mouth of the \textsc{Lord}.
\vs Jer 23:17 They say still unto them that despise me, The \textsc{Lord} hath said, Ye shall have peace; and they say unto every one that walketh after the imagination of his own heart, No evil shall come upon you.
\fn{imagination: or, stubbornness.}
\vs Jer 23:18 For who hath stood in the counsel of the \textsc{Lord}, and hath perceived and heard his word? who hath marked his word, and heard \bibemph{it}?
\fn{counsel: or, secret.}
\vs Jer 23:19 Behold, a whirlwind of the \textsc{Lord} is gone forth in fury, even a grievous whirlwind: it shall fall grievously upon the head of the wicked.
\vs Jer 23:20 The anger of the \textsc{Lord} shall not return, until he have executed, and till he have performed the thoughts of his heart: in the latter days ye shall consider it perfectly.
\vs Jer 23:21 I have not sent these prophets, yet they ran: I have not spoken to them, yet they prophesied.
\vs Jer 23:22 But if they had stood in my counsel, and had caused my people to hear my words, then they should have turned them from their evil way, and from the evil of their doings.
\vs Jer 23:23 \bibemph{Am} I a God at hand, saith the \textsc{Lord}, and not a God afar off?
\vs Jer 23:24 Can any hide himself in secret places that I shall not see him? saith the \textsc{Lord}. Do not I fill heaven and earth? saith the \textsc{Lord}.
\vs Jer 23:25 I have heard what the prophets said, that prophesy lies in my name, saying, I have dreamed, I have dreamed.
\vs Jer 23:26 How long shall \bibemph{this} be in the heart of the prophets that prophesy lies? yea, \bibemph{they are} prophets of the deceit of their own heart;
\vs Jer 23:27 Which think to cause my people to forget my name by their dreams which they tell every man to his neighbour, as their fathers have forgotten my name for Baal.
\vs Jer 23:28 The prophet that hath a dream, let him tell a dream; and he that hath my word, let him speak my word faithfully. What \bibemph{is} the chaff to the wheat? saith the \textsc{Lord}.
\fn{that hath a dream: Heb. with whom is, etc.}
\vs Jer 23:29 \bibemph{Is} not my word like as a fire? saith the \textsc{Lord}; and like a hammer \bibemph{that} breaketh the rock in pieces?
\vs Jer 23:30 Therefore, behold, I \bibemph{am} against the prophets, saith the \textsc{Lord}, that steal my words every one from his neighbour.
\vs Jer 23:31 Behold, I \bibemph{am} against the prophets, saith the \textsc{Lord}, that use their tongues, and say, He saith.
\fn{that...: or, that smooth their tongues.}
\vs Jer 23:32 Behold, I \bibemph{am} against them that prophesy false dreams, saith the \textsc{Lord}, and do tell them, and cause my people to err by their lies, and by their lightness; yet I sent them not, nor commanded them: therefore they shall not profit this people at all, saith the \textsc{Lord}.
\vs Jer 23:33 And when this people, or the prophet, or a priest, shall ask thee, saying, What \bibemph{is} the burden of the \textsc{Lord}? thou shalt then say unto them, What burden? I will even forsake you, saith the \textsc{Lord}.
\vs Jer 23:34 And \bibemph{as for} the prophet, and the priest, and the people, that shall say, The burden of the \textsc{Lord}, I will even punish that man and his house.
\fn{punish: Heb. visit upon.}
\vs Jer 23:35 Thus shall ye say every one to his neighbour, and every one to his brother, What hath the \textsc{Lord} answered? and, What hath the \textsc{Lord} spoken?
\vs Jer 23:36 And the burden of the \textsc{Lord} shall ye mention no more: for every man's word shall be his burden; for ye have perverted the words of the living God, of the \textsc{Lord} of hosts our God.
\vs Jer 23:37 Thus shalt thou say to the prophet, What hath the \textsc{Lord} answered thee? and, What hath the \textsc{Lord} spoken?
\vs Jer 23:38 But since ye say, The burden of the \textsc{Lord}; therefore thus saith the \textsc{Lord}; Because ye say this word, The burden of the \textsc{Lord}, and I have sent unto you, saying, Ye shall not say, The burden of the \textsc{Lord};
\vs Jer 23:39 Therefore, behold, I, even I, will utterly forget you, and I will forsake you, and the city that I gave you and your fathers, \bibemph{and cast you} out of my presence:
\vs Jer 23:40 And I will bring an everlasting reproach upon you, and a perpetual shame, which shall not be forgotten.
\vs Jer 24:1 The \textsc{Lord} shewed me, and, behold, two baskets of figs \bibemph{were} set before the temple of the \textsc{Lord}, after that Nebuchadrezzar king of Babylon had carried away captive Jeconiah the son of Jehoiakim king of Judah, and the princes of Judah, with the carpenters and smiths, from Jerusalem, and had brought them to Babylon.
\vs Jer 24:2 One basket \bibemph{had} very good figs, \bibemph{even} like the figs \bibemph{that are} first ripe: and the other basket \bibemph{had} very naughty figs, which could not be eaten, they were so bad.
\fn{they...: Heb. for badness.}
\vs Jer 24:3 Then said the \textsc{Lord} unto me, What seest thou, Jeremiah? And I said, Figs; the good figs, very good; and the evil, very evil, that cannot be eaten, they are so evil.
\vs Jer 24:4 Again the word of the \textsc{Lord} came unto me, saying,
\vs Jer 24:5 Thus saith the \textsc{Lord}, the God of Israel; Like these good figs, so will I acknowledge them that are carried away captive of Judah, whom I have sent out of this place into the land of the Chaldeans for \bibemph{their} good.
\fn{them...: Heb. the captivity.}
\vs Jer 24:6 For I will set mine eyes upon them for good, and I will bring them again to this land: and I will build them, and not pull \bibemph{them} down; and I will plant them, and not pluck \bibemph{them} up.
\vs Jer 24:7 And I will give them an heart to know me, that I \bibemph{am} the \textsc{Lord}: and they shall be my people, and I will be their God: for they shall return unto me with their whole heart.
\vs Jer 24:8 And as the evil figs, which cannot be eaten, they are so evil; surely thus saith the \textsc{Lord}, So will I give Zedekiah the king of Judah, and his princes, and the residue of Jerusalem, that remain in this land, and them that dwell in the land of Egypt:
\vs Jer 24:9 And I will deliver them to be removed into all the kingdoms of the earth for \bibemph{their} hurt, \bibemph{to be} a reproach and a proverb, a taunt and a curse, in all places whither I shall drive them.
\fn{to be removed: Heb. for removing, or, vexation.}
\vs Jer 24:10 And I will send the sword, the famine, and the pestilence, among them, till they be consumed from off the land that I gave unto them and to their fathers.
\vs Jer 25:1 The word that came to Jeremiah concerning all the people of Judah in the fourth year of Jehoiakim the son of Josiah king of Judah, that \bibemph{was} the first year of Nebuchadrezzar king of Babylon;
\vs Jer 25:2 The which Jeremiah the prophet spake unto all the people of Judah, and to all the inhabitants of Jerusalem, saying,
\vs Jer 25:3 From the thirteenth year of Josiah the son of Amon king of Judah, even unto this day, that \bibemph{is} the three and twentieth year, the word of the \textsc{Lord} hath come unto me, and I have spoken unto you, rising early and speaking; but ye have not hearkened.
\vs Jer 25:4 And the \textsc{Lord} hath sent unto you all his servants the prophets, rising early and sending \bibemph{them}; but ye have not hearkened, nor inclined your ear to hear.
\vs Jer 25:5 They said, Turn ye again now every one from his evil way, and from the evil of your doings, and dwell in the land that the \textsc{Lord} hath given unto you and to your fathers for ever and ever:
\vs Jer 25:6 And go not after other gods to serve them, and to worship them, and provoke me not to anger with the works of your hands; and I will do you no hurt.
\vs Jer 25:7 Yet ye have not hearkened unto me, saith the \textsc{Lord}; that ye might provoke me to anger with the works of your hands to your own hurt.
\vs Jer 25:8 Therefore thus saith the \textsc{Lord} of hosts; Because ye have not heard my words,
\vs Jer 25:9 Behold, I will send and take all the families of the north, saith the \textsc{Lord}, and Nebuchadrezzar the king of Babylon, my servant, and will bring them against this land, and against the inhabitants thereof, and against all these nations round about, and will utterly destroy them, and make them an astonishment, and an hissing, and perpetual desolations.
\vs Jer 25:10 Moreover I will take from them the voice of mirth, and the voice of gladness, the voice of the bridegroom, and the voice of the bride, the sound of the millstones, and the light of the candle.
\fn{I will...: Heb. I will cause to perish from them.}
\vs Jer 25:11 And this whole land shall be a desolation, \bibemph{and} an astonishment; and these nations shall serve the king of Babylon seventy years.
\vs Jer 25:12 And it shall come to pass, when seventy years are accomplished, \bibemph{that} I will punish the king of Babylon, and that nation, saith the \textsc{Lord}, for their iniquity, and the land of the Chaldeans, and will make it perpetual desolations.
\fn{punish: Heb. visit upon.}
\vs Jer 25:13 And I will bring upon that land all my words which I have pronounced against it, \bibemph{even} all that is written in this book, which Jeremiah hath prophesied against all the nations.
\vs Jer 25:14 For many nations and great kings shall serve themselves of them also: and I will recompense them according to their deeds, and according to the works of their own hands.
\vs Jer 25:15 For thus saith the \textsc{Lord} God of Israel unto me; Take the wine cup of this fury at my hand, and cause all the nations, to whom I send thee, to drink it.
\vs Jer 25:16 And they shall drink, and be moved, and be mad, because of the sword that I will send among them.
\vs Jer 25:17 Then took I the cup at the \textsc{Lord's} hand, and made all the nations to drink, unto whom the \textsc{Lord} had sent me:
\vs Jer 25:18 \bibemph{To wit}, Jerusalem, and the cities of Judah, and the kings thereof, and the princes thereof, to make them a desolation, an astonishment, an hissing, and a curse; as \bibemph{it is} this day;
\vs Jer 25:19 Pharaoh king of Egypt, and his servants, and his princes, and all his people;
\vs Jer 25:20 And all the mingled people, and all the kings of the land of Uz, and all the kings of the land of the Philistines, and Ashkelon, and Azzah, and Ekron, and the remnant of Ashdod,
\vs Jer 25:21 Edom, and Moab, and the children of Ammon,
\vs Jer 25:22 And all the kings of Tyrus, and all the kings of Zidon, and the kings of the isles which \bibemph{are} beyond the sea,
\fn{isles: or, region by the sea side.}
\vs Jer 25:23 Dedan, and Tema, and Buz, and all \bibemph{that are} in the utmost corners,
\fn{that...: Heb. cut off into corners, or, having the corners of the hair polled.}
\vs Jer 25:24 And all the kings of Arabia, and all the kings of the mingled people that dwell in the desert,
\vs Jer 25:25 And all the kings of Zimri, and all the kings of Elam, and all the kings of the Medes,
\vs Jer 25:26 And all the kings of the north, far and near, one with another, and all the kingdoms of the world, which \bibemph{are} upon the face of the earth: and the king of Sheshach shall drink after them.
\vs Jer 25:27 Therefore thou shalt say unto them, Thus saith the \textsc{Lord} of hosts, the God of Israel; Drink ye, and be drunken, and spue, and fall, and rise no more, because of the sword which I will send among you.
\vs Jer 25:28 And it shall be, if they refuse to take the cup at thine hand to drink, then shalt thou say unto them, Thus saith the \textsc{Lord} of hosts; Ye shall certainly drink.
\vs Jer 25:29 For, lo, I begin to bring evil on the city which is called by my name, and should ye be utterly unpunished? Ye shall not be unpunished: for I will call for a sword upon all the inhabitants of the earth, saith the \textsc{Lord} of hosts.
\fn{which...: Heb. upon which my name is called.}
\vs Jer 25:30 Therefore prophesy thou against them all these words, and say unto them, The \textsc{Lord} shall roar from on high, and utter his voice from his holy habitation; he shall mightily roar upon his habitation; he shall give a shout, as they that tread \bibemph{the grapes}, against all the inhabitants of the earth.
\vs Jer 25:31 A noise shall come \bibemph{even} to the ends of the earth; for the \textsc{Lord} hath a controversy with the nations, he will plead with all flesh; he will give them \bibemph{that are} wicked to the sword, saith the \textsc{Lord}.
\vs Jer 25:32 Thus saith the \textsc{Lord} of hosts, Behold, evil shall go forth from nation to nation, and a great whirlwind shall be raised up from the coasts of the earth.
\vs Jer 25:33 And the slain of the \textsc{Lord} shall be at that day from \bibemph{one} end of the earth even unto the \bibemph{other} end of the earth: they shall not be lamented, neither gathered, nor buried; they shall be dung upon the ground.
\vs Jer 25:34 Howl, ye shepherds, and cry; and wallow yourselves \bibemph{in the ashes}, ye principal of the flock: for the days of your slaughter and of your dispersions are accomplished; and ye shall fall like a pleasant vessel.
\fn{the days...: Heb. your days for slaughter. a pleasant...: Heb. a vessel of desire.}
\vs Jer 25:35 And the shepherds shall have no way to flee, nor the principal of the flock to escape.
\fn{the shepherds...: Heb. flight shall perish from the shepherds, and escaping from, etc.}
\vs Jer 25:36 A voice of the cry of the shepherds, and an howling of the principal of the flock, \bibemph{shall be heard}: for the \textsc{Lord} hath spoiled their pasture.
\vs Jer 25:37 And the peaceable habitations are cut down because of the fierce anger of the \textsc{Lord}.
\vs Jer 25:38 He hath forsaken his covert, as the lion: for their land is desolate because of the fierceness of the oppressor, and because of his fierce anger.
\fn{desolate: Heb. a desolation.}
\vs Jer 26:1 In the beginning of the reign of Jehoiakim the son of Josiah king of Judah came this word from the \textsc{Lord}, saying,
\vs Jer 26:2 Thus saith the \textsc{Lord}; Stand in the court of the \textsc{Lord's} house, and speak unto all the cities of Judah, which come to worship in the \textsc{Lord's} house, all the words that I command thee to speak unto them; diminish not a word:
\vs Jer 26:3 If so be they will hearken, and turn every man from his evil way, that I may repent me of the evil, which I purpose to do unto them because of the evil of their doings.
\vs Jer 26:4 And thou shalt say unto them, Thus saith the \textsc{Lord}; If ye will not hearken to me, to walk in my law, which I have set before you,
\vs Jer 26:5 To hearken to the words of my servants the prophets, whom I sent unto you, both rising up early, and sending \bibemph{them}, but ye have not hearkened;
\vs Jer 26:6 Then will I make this house like Shiloh, and will make this city a curse to all the nations of the earth.
\vs Jer 26:7 So the priests and the prophets and all the people heard Jeremiah speaking these words in the house of the \textsc{Lord}.
\vs Jer 26:8 Now it came to pass, when Jeremiah had made an end of speaking all that the \textsc{Lord} had commanded \bibemph{him} to speak unto all the people, that the priests and the prophets and all the people took him, saying, Thou shalt surely die.
\vs Jer 26:9 Why hast thou prophesied in the name of the \textsc{Lord}, saying, This house shall be like Shiloh, and this city shall be desolate without an inhabitant? And all the people were gathered against Jeremiah in the house of the \textsc{Lord}.
\vs Jer 26:10 When the princes of Judah heard these things, then they came up from the king's house unto the house of the \textsc{Lord}, and sat down in the entry of the new gate of the \textsc{Lord's} \bibemph{house}.
\fn{in the...: or, at the door.}
\vs Jer 26:11 Then spake the priests and the prophets unto the princes and to all the people, saying, This man \bibemph{is} worthy to die; for he hath prophesied against this city, as ye have heard with your ears.
\fn{This man...: Heb. The judgment of death is for this man.}
\vs Jer 26:12 Then spake Jeremiah unto all the princes and to all the people, saying, The \textsc{Lord} sent me to prophesy against this house and against this city all the words that ye have heard.
\vs Jer 26:13 Therefore now amend your ways and your doings, and obey the voice of the \textsc{Lord} your God; and the \textsc{Lord} will repent him of the evil that he hath pronounced against you.
\vs Jer 26:14 As for me, behold, I \bibemph{am} in your hand: do with me as seemeth good and meet unto you.
\fn{as...: Heb. as it is good and right in your eyes.}
\vs Jer 26:15 But know ye for certain, that if ye put me to death, ye shall surely bring innocent blood upon yourselves, and upon this city, and upon the inhabitants thereof: for of a truth the \textsc{Lord} hath sent me unto you to speak all these words in your ears.
\vs Jer 26:16 Then said the princes and all the people unto the priests and to the prophets; This man \bibemph{is} not worthy to die: for he hath spoken to us in the name of the \textsc{Lord} our God.
\vs Jer 26:17 Then rose up certain of the elders of the land, and spake to all the assembly of the people, saying,
\vs Jer 26:18 Micah the Morasthite prophesied in the days of Hezekiah king of Judah, and spake to all the people of Judah, saying, Thus saith the \textsc{Lord} of hosts; Zion shall be plowed \bibemph{like} a field, and Jerusalem shall become heaps, and the mountain of the house as the high places of a forest.
\vs Jer 26:19 Did Hezekiah king of Judah and all Judah put him at all to death? did he not fear the \textsc{Lord}, and besought the \textsc{Lord}, and the \textsc{Lord} repented him of the evil which he had pronounced against them? Thus might we procure great evil against our souls.
\fn{the \textsc{Lord}, and the: Heb. the face of the \textsc{Lord}, etc.}
\vs Jer 26:20 And there was also a man that prophesied in the name of the \textsc{Lord}, Urijah the son of Shemaiah of Kirjathjearim, who prophesied against this city and against this land according to all the words of Jeremiah:
\vs Jer 26:21 And when Jehoiakim the king, with all his mighty men, and all the princes, heard his words, the king sought to put him to death: but when Urijah heard it, he was afraid, and fled, and went into Egypt;
\vs Jer 26:22 And Jehoiakim the king sent men into Egypt, \bibemph{namely}, Elnathan the son of Achbor, and \bibemph{certain} men with him into Egypt.
\vs Jer 26:23 And they fetched forth Urijah out of Egypt, and brought him unto Jehoiakim the king; who slew him with the sword, and cast his dead body into the graves of the common people.
\fn{common...: Heb. sons of the people.}
\vs Jer 26:24 Nevertheless the hand of Ahikam the son of Shaphan was with Jeremiah, that they should not give him into the hand of the people to put him to death.
\vs Jer 27:1 In the beginning of the reign of Jehoiakim the son of Josiah king of Judah came this word unto Jeremiah from the \textsc{Lord}, saying,
\vs Jer 27:2 Thus saith the \textsc{Lord} to me; Make thee bonds and yokes, and put them upon thy neck,
\fn{saith...: or, hath the \textsc{Lord} said.}
\vs Jer 27:3 And send them to the king of Edom, and to the king of Moab, and to the king of the Ammonites, and to the king of Tyrus, and to the king of Zidon, by the hand of the messengers which come to Jerusalem unto Zedekiah king of Judah;
\vs Jer 27:4 And command them to say unto their masters, Thus saith the \textsc{Lord} of hosts, the God of Israel; Thus shall ye say unto your masters;
\fn{to say...: or, concerning their masters, saying.}
\vs Jer 27:5 I have made the earth, the man and the beast that \bibemph{are} upon the ground, by my great power and by my outstretched arm, and have given it unto whom it seemed meet unto me.
\vs Jer 27:6 And now have I given all these lands into the hand of Nebuchadnezzar the king of Babylon, my servant; and the beasts of the field have I given him also to serve him.
\vs Jer 27:7 And all nations shall serve him, and his son, and his son's son, until the very time of his land come: and then many nations and great kings shall serve themselves of him.
\vs Jer 27:8 And it shall come to pass, \bibemph{that} the nation and kingdom which will not serve the same Nebuchadnezzar the king of Babylon, and that will not put their neck under the yoke of the king of Babylon, that nation will I punish, saith the \textsc{Lord}, with the sword, and with the famine, and with the pestilence, until I have consumed them by his hand.
\vs Jer 27:9 Therefore hearken not ye to your prophets, nor to your diviners, nor to your dreamers, nor to your enchanters, nor to your sorcerers, which speak unto you, saying, Ye shall not serve the king of Babylon:
\fn{dreamers: Heb. dreams.}
\vs Jer 27:10 For they prophesy a lie unto you, to remove you far from your land; and that I should drive you out, and ye should perish.
\vs Jer 27:11 But the nations that bring their neck under the yoke of the king of Babylon, and serve him, those will I let remain still in their own land, saith the \textsc{Lord}; and they shall till it, and dwell therein.
\vs Jer 27:12 I spake also to Zedekiah king of Judah according to all these words, saying, Bring your necks under the yoke of the king of Babylon, and serve him and his people, and live.
\vs Jer 27:13 Why will ye die, thou and thy people, by the sword, by the famine, and by the pestilence, as the \textsc{Lord} hath spoken against the nation that will not serve the king of Babylon?
\vs Jer 27:14 Therefore hearken not unto the words of the prophets that speak unto you, saying, Ye shall not serve the king of Babylon: for they prophesy a lie unto you.
\vs Jer 27:15 For I have not sent them, saith the \textsc{Lord}, yet they prophesy a lie in my name; that I might drive you out, and that ye might perish, ye, and the prophets that prophesy unto you.
\fn{a lie: Heb. in a lie, or, lyingly.}
\vs Jer 27:16 Also I spake to the priests and to all this people, saying, Thus saith the \textsc{Lord}; Hearken not to the words of your prophets that prophesy unto you, saying, Behold, the vessels of the \textsc{Lord's} house shall now shortly be brought again from Babylon: for they prophesy a lie unto you.
\vs Jer 27:17 Hearken not unto them; serve the king of Babylon, and live: wherefore should this city be laid waste?
\vs Jer 27:18 But if they \bibemph{be} prophets, and if the word of the \textsc{Lord} be with them, let them now make intercession to the \textsc{Lord} of hosts, that the vessels which are left in the house of the \textsc{Lord}, and \bibemph{in} the house of the king of Judah, and at Jerusalem, go not to Babylon.
\vs Jer 27:19 For thus saith the \textsc{Lord} of hosts concerning the pillars, and concerning the sea, and concerning the bases, and concerning the residue of the vessels that remain in this city,
\vs Jer 27:20 Which Nebuchadnezzar king of Babylon took not, when he carried away captive Jeconiah the son of Jehoiakim king of Judah from Jerusalem to Babylon, and all the nobles of Judah and Jerusalem;
\vs Jer 27:21 Yea, thus saith the \textsc{Lord} of hosts, the God of Israel, concerning the vessels that remain \bibemph{in} the house of the \textsc{Lord}, and \bibemph{in} the house of the king of Judah and of Jerusalem;
\vs Jer 27:22 They shall be carried to Babylon, and there shall they be until the day that I visit them, saith the \textsc{Lord}; then will I bring them up, and restore them to this place.
\vs Jer 28:1 And it came to pass the same year, in the beginning of the reign of Zedekiah king of Judah, in the fourth year, \bibemph{and} in the fifth month, \bibemph{that} Hananiah the son of Azur the prophet, which \bibemph{was} of Gibeon, spake unto me in the house of the \textsc{Lord}, in the presence of the priests and of all the people, saying,
\vs Jer 28:2 Thus speaketh the \textsc{Lord} of hosts, the God of Israel, saying, I have broken the yoke of the king of Babylon.
\vs Jer 28:3 Within two full years will I bring again into this place all the vessels of the \textsc{Lord's} house, that Nebuchadnezzar king of Babylon took away from this place, and carried them to Babylon:
\fn{two...: Heb. two years of days.}
\vs Jer 28:4 And I will bring again to this place Jeconiah the son of Jehoiakim king of Judah, with all the captives of Judah, that went into Babylon, saith the \textsc{Lord}: for I will break the yoke of the king of Babylon.
\fn{captives: Heb. captivity.}
\vs Jer 28:5 Then the prophet Jeremiah said unto the prophet Hananiah in the presence of the priests, and in the presence of all the people that stood in the house of the \textsc{Lord},
\vs Jer 28:6 Even the prophet Jeremiah said, Amen: the \textsc{Lord} do so: the \textsc{Lord} perform thy words which thou hast prophesied, to bring again the vessels of the \textsc{Lord's} house, and all that is carried away captive, from Babylon into this place.
\vs Jer 28:7 Nevertheless hear thou now this word that I speak in thine ears, and in the ears of all the people;
\vs Jer 28:8 The prophets that have been before me and before thee of old prophesied both against many countries, and against great kingdoms, of war, and of evil, and of pestilence.
\vs Jer 28:9 The prophet which prophesieth of peace, when the word of the prophet shall come to pass, \bibemph{then} shall the prophet be known, that the \textsc{Lord} hath truly sent him.
\vs Jer 28:10 Then Hananiah the prophet took the yoke from off the prophet Jeremiah's neck, and brake it.
\vs Jer 28:11 And Hananiah spake in the presence of all the people, saying, Thus saith the \textsc{Lord}; Even so will I break the yoke of Nebuchadnezzar king of Babylon from the neck of all nations within the space of two full years. And the prophet Jeremiah went his way.
\vs Jer 28:12 Then the word of the \textsc{Lord} came unto Jeremiah \bibemph{the prophet}, after that Hananiah the prophet had broken the yoke from off the neck of the prophet Jeremiah, saying,
\vs Jer 28:13 Go and tell Hananiah, saying, Thus saith the \textsc{Lord}; Thou hast broken the yokes of wood; but thou shalt make for them yokes of iron.
\vs Jer 28:14 For thus saith the \textsc{Lord} of hosts, the God of Israel; I have put a yoke of iron upon the neck of all these nations, that they may serve Nebuchadnezzar king of Babylon; and they shall serve him: and I have given him the beasts of the field also.
\vs Jer 28:15 Then said the prophet Jeremiah unto Hananiah the prophet, Hear now, Hananiah; The \textsc{Lord} hath not sent thee; but thou makest this people to trust in a lie.
\vs Jer 28:16 Therefore thus saith the \textsc{Lord}; Behold, I will cast thee from off the face of the earth: this year thou shalt die, because thou hast taught rebellion against the \textsc{Lord}.
\fn{rebellion: Heb. revolt.}
\vs Jer 28:17 So Hananiah the prophet died the same year in the seventh month.
\vs Jer 29:1 Now these \bibemph{are} the words of the letter that Jeremiah the prophet sent from Jerusalem unto the residue of the elders which were carried away captives, and to the priests, and to the prophets, and to all the people whom Nebuchadnezzar had carried away captive from Jerusalem to Babylon;
\vs Jer 29:2 (After that Jeconiah the king, and the queen, and the eunuchs, the princes of Judah and Jerusalem, and the carpenters, and the smiths, were departed from Jerusalem;)
\fn{eunuchs: or, chamberlains.}
\vs Jer 29:3 By the hand of Elasah the son of Shaphan, and Gemariah the son of Hilkiah, (whom Zedekiah king of Judah sent unto Babylon to Nebuchadnezzar king of Babylon) saying,
\vs Jer 29:4 Thus saith the \textsc{Lord} of hosts, the God of Israel, unto all that are carried away captives, whom I have caused to be carried away from Jerusalem unto Babylon;
\vs Jer 29:5 Build ye houses, and dwell \bibemph{in them}; and plant gardens, and eat the fruit of them;
\vs Jer 29:6 Take ye wives, and beget sons and daughters; and take wives for your sons, and give your daughters to husbands, that they may bear sons and daughters; that ye may be increased there, and not diminished.
\vs Jer 29:7 And seek the peace of the city whither I have caused you to be carried away captives, and pray unto the \textsc{Lord} for it: for in the peace thereof shall ye have peace.
\vs Jer 29:8 For thus saith the \textsc{Lord} of hosts, the God of Israel; Let not your prophets and your diviners, that \bibemph{be} in the midst of you, deceive you, neither hearken to your dreams which ye cause to be dreamed.
\vs Jer 29:9 For they prophesy falsely unto you in my name: I have not sent them, saith the \textsc{Lord}.
\fn{falsely: Heb. in a lie.}
\vs Jer 29:10 For thus saith the \textsc{Lord}, That after seventy years be accomplished at Babylon I will visit you, and perform my good word toward you, in causing you to return to this place.
\vs Jer 29:11 For I know the thoughts that I think toward you, saith the \textsc{Lord}, thoughts of peace, and not of evil, to give you an expected end.
\fn{expected...: Heb. end and expectation.}
\vs Jer 29:12 Then shall ye call upon me, and ye shall go and pray unto me, and I will hearken unto you.
\vs Jer 29:13 And ye shall seek me, and find \bibemph{me}, when ye shall search for me with all your heart.
\vs Jer 29:14 And I will be found of you, saith the \textsc{Lord}: and I will turn away your captivity, and I will gather you from all the nations, and from all the places whither I have driven you, saith the \textsc{Lord}; and I will bring you again into the place whence I caused you to be carried away captive.
\vs Jer 29:15 Because ye have said, The \textsc{Lord} hath raised us up prophets in Babylon;
\vs Jer 29:16 \bibemph{Know} that thus saith the \textsc{Lord} of the king that sitteth upon the throne of David, and of all the people that dwelleth in this city, \bibemph{and} of your brethren that are not gone forth with you into captivity;
\vs Jer 29:17 Thus saith the \textsc{Lord} of hosts; Behold, I will send upon them the sword, the famine, and the pestilence, and will make them like vile figs, that cannot be eaten, they are so evil.
\vs Jer 29:18 And I will persecute them with the sword, with the famine, and with the pestilence, and will deliver them to be removed to all the kingdoms of the earth, to be a curse, and an astonishment, and an hissing, and a reproach, among all the nations whither I have driven them:
\fn{to be a curse: Heb. for a curse.}
\vs Jer 29:19 Because they have not hearkened to my words, saith the \textsc{Lord}, which I sent unto them by my servants the prophets, rising up early and sending \bibemph{them}; but ye would not hear, saith the \textsc{Lord}.
\vs Jer 29:20 Hear ye therefore the word of the \textsc{Lord}, all ye of the captivity, whom I have sent from Jerusalem to Babylon:
\vs Jer 29:21 Thus saith the \textsc{Lord} of hosts, the God of Israel, of Ahab the son of Kolaiah, and of Zedekiah the son of Maaseiah, which prophesy a lie unto you in my name; Behold, I will deliver them into the hand of Nebuchadrezzar king of Babylon; and he shall slay them before your eyes;
\vs Jer 29:22 And of them shall be taken up a curse by all the captivity of Judah which \bibemph{are} in Babylon, saying, The \textsc{Lord} make thee like Zedekiah and like Ahab, whom the king of Babylon roasted in the fire;
\vs Jer 29:23 Because they have committed villany in Israel, and have committed adultery with their neighbours' wives, and have spoken lying words in my name, which I have not commanded them; even I know, and \bibemph{am} a witness, saith the \textsc{Lord}.
\vs Jer 29:24 \bibemph{Thus} shalt thou also speak to Shemaiah the Nehelamite, saying,
\fn{Nehelamite: or, dreamer.}
\vs Jer 29:25 Thus speaketh the \textsc{Lord} of hosts, the God of Israel, saying, Because thou hast sent letters in thy name unto all the people that \bibemph{are} at Jerusalem, and to Zephaniah the son of Maaseiah the priest, and to all the priests, saying,
\vs Jer 29:26 The \textsc{Lord} hath made thee priest in the stead of Jehoiada the priest, that ye should be officers in the house of the \textsc{Lord}, for every man \bibemph{that is} mad, and maketh himself a prophet, that thou shouldest put him in prison, and in the stocks.
\vs Jer 29:27 Now therefore why hast thou not reproved Jeremiah of Anathoth, which maketh himself a prophet to you?
\vs Jer 29:28 For therefore he sent unto us \bibemph{in} Babylon, saying, This \bibemph{captivity is} long: build ye houses, and dwell \bibemph{in them}; and plant gardens, and eat the fruit of them.
\vs Jer 29:29 And Zephaniah the priest read this letter in the ears of Jeremiah the prophet.
\vs Jer 29:30 Then came the word of the \textsc{Lord} unto Jeremiah, saying,
\vs Jer 29:31 Send to all them of the captivity, saying, Thus saith the \textsc{Lord} concerning Shemaiah the Nehelamite; Because that Shemaiah hath prophesied unto you, and I sent him not, and he caused you to trust in a lie:
\vs Jer 29:32 Therefore thus saith the \textsc{Lord}; Behold, I will punish Shemaiah the Nehelamite, and his seed: he shall not have a man to dwell among this people; neither shall he behold the good that I will do for my people, saith the \textsc{Lord}; because he hath taught rebellion against the \textsc{Lord}.
\fn{rebellion: Heb. revolt.}
\vs Jer 30:1 The word that came to Jeremiah from the \textsc{Lord}, saying,
\vs Jer 30:2 Thus speaketh the \textsc{Lord} God of Israel, saying, Write thee all the words that I have spoken unto thee in a book.
\vs Jer 30:3 For, lo, the days come, saith the \textsc{Lord}, that I will bring again the captivity of my people Israel and Judah, saith the \textsc{Lord}: and I will cause them to return to the land that I gave to their fathers, and they shall possess it.
\vs Jer 30:4 And these \bibemph{are} the words that the \textsc{Lord} spake concerning Israel and concerning Judah.
\vs Jer 30:5 For thus saith the \textsc{Lord}; We have heard a voice of trembling, of fear, and not of peace.
\fn{of fear...: or, there is fear, and not peace.}
\vs Jer 30:6 Ask ye now, and see whether a man doth travail with child? wherefore do I see every man with his hands on his loins, as a woman in travail, and all faces are turned into paleness?
\fn{a man: Heb. a male.}
\vs Jer 30:7 Alas! for that day \bibemph{is} great, so that none \bibemph{is} like it: it \bibemph{is} even the time of Jacob's trouble; but he shall be saved out of it.
\vs Jer 30:8 For it shall come to pass in that day, saith the \textsc{Lord} of hosts, \bibemph{that} I will break his yoke from off thy neck, and will burst thy bonds, and strangers shall no more serve themselves of him:
\vs Jer 30:9 But they shall serve the \textsc{Lord} their God, and David their king, whom I will raise up unto them.
\vs Jer 30:10 Therefore fear thou not, O my servant Jacob, saith the \textsc{Lord}; neither be dismayed, O Israel: for, lo, I will save thee from afar, and thy seed from the land of their captivity; and Jacob shall return, and shall be in rest, and be quiet, and none shall make \bibemph{him} afraid.
\vs Jer 30:11 For I \bibemph{am} with thee, saith the \textsc{Lord}, to save thee: though I make a full end of all nations whither I have scattered thee, yet will I not make a full end of thee: but I will correct thee in measure, and will not leave thee altogether unpunished.
\vs Jer 30:12 For thus saith the \textsc{Lord}, Thy bruise \bibemph{is} incurable, \bibemph{and} thy wound \bibemph{is} grievous.
\vs Jer 30:13 \bibemph{There is} none to plead thy cause, that thou mayest be bound up: thou hast no healing medicines.
\fn{that...: Heb. for binding up, or, pressing.}
\vs Jer 30:14 All thy lovers have forgotten thee; they seek thee not; for I have wounded thee with the wound of an enemy, with the chastisement of a cruel one, for the multitude of thine iniquity; \bibemph{because} thy sins were increased.
\vs Jer 30:15 Why criest thou for thine affliction? thy sorrow \bibemph{is} incurable for the multitude of thine iniquity: \bibemph{because} thy sins were increased, I have done these things unto thee.
\vs Jer 30:16 Therefore all they that devour thee shall be devoured; and all thine adversaries, every one of them, shall go into captivity; and they that spoil thee shall be a spoil, and all that prey upon thee will I give for a prey.
\vs Jer 30:17 For I will restore health unto thee, and I will heal thee of thy wounds, saith the \textsc{Lord}; because they called thee an Outcast, \bibemph{saying}, This \bibemph{is} Zion, whom no man seeketh after.
\vs Jer 30:18 Thus saith the \textsc{Lord}; Behold, I will bring again the captivity of Jacob's tents, and have mercy on his dwellingplaces; and the city shall be builded upon her own heap, and the palace shall remain after the manner thereof.
\fn{heap: or, little hill.}
\vs Jer 30:19 And out of them shall proceed thanksgiving and the voice of them that make merry: and I will multiply them, and they shall not be few; I will also glorify them, and they shall not be small.
\vs Jer 30:20 Their children also shall be as aforetime, and their congregation shall be established before me, and I will punish all that oppress them.
\vs Jer 30:21 And their nobles shall be of themselves, and their governor shall proceed from the midst of them; and I will cause him to draw near, and he shall approach unto me: for who \bibemph{is} this that engaged his heart to approach unto me? saith the \textsc{Lord}.
\vs Jer 30:22 And ye shall be my people, and I will be your God.
\vs Jer 30:23 Behold, the whirlwind of the \textsc{Lord} goeth forth with fury, a continuing whirlwind: it shall fall with pain upon the head of the wicked.
\fn{continuing: Heb. cutting. fall...: or, remain.}
\vs Jer 30:24 The fierce anger of the \textsc{Lord} shall not return, until he have done \bibemph{it}, and until he have performed the intents of his heart: in the latter days ye shall consider it.
\vs Jer 31:1 At the same time, saith the \textsc{Lord}, will I be the God of all the families of Israel, and they shall be my people.
\vs Jer 31:2 Thus saith the \textsc{Lord}, The people \bibemph{which were} left of the sword found grace in the wilderness; \bibemph{even} Israel, when I went to cause him to rest.
\vs Jer 31:3 The \textsc{Lord} hath appeared of old unto me, \bibemph{saying}, Yea, I have loved thee with an everlasting love: therefore with lovingkindness have I drawn thee.
\fn{of...: Heb. from afar. with lovingkindness...: have I extended lovingkindness unto thee.}
\vs Jer 31:4 Again I will build thee, and thou shalt be built, O virgin of Israel: thou shalt again be adorned with thy tabrets, and shalt go forth in the dances of them that make merry.
\fn{tabrets: or, timbrels.}
\vs Jer 31:5 Thou shalt yet plant vines upon the mountains of Samaria: the planters shall plant, and shall eat \bibemph{them} as common things.
\fn{eat...: Heb. profane them.}
\vs Jer 31:6 For there shall be a day, \bibemph{that} the watchmen upon the mount Ephraim shall cry, Arise ye, and let us go up to Zion unto the \textsc{Lord} our God.
\vs Jer 31:7 For thus saith the \textsc{Lord}; Sing with gladness for Jacob, and shout among the chief of the nations: publish ye, praise ye, and say, O \textsc{Lord}, save thy people, the remnant of Israel.
\vs Jer 31:8 Behold, I will bring them from the north country, and gather them from the coasts of the earth, \bibemph{and} with them the blind and the lame, the woman with child and her that travaileth with child together: a great company shall return thither.
\vs Jer 31:9 They shall come with weeping, and with supplications will I lead them: I will cause them to walk by the rivers of waters in a straight way, wherein they shall not stumble: for I am a father to Israel, and Ephraim \bibemph{is} my firstborn.
\fn{supplications: or, favours.}
\vs Jer 31:10 Hear the word of the \textsc{Lord}, O ye nations, and declare \bibemph{it} in the isles afar off, and say, He that scattered Israel will gather him, and keep him, as a shepherd \bibemph{doth} his flock.
\vs Jer 31:11 For the \textsc{Lord} hath redeemed Jacob, and ransomed him from the hand of \bibemph{him that was} stronger than he.
\vs Jer 31:12 Therefore they shall come and sing in the height of Zion, and shall flow together to the goodness of the \textsc{Lord}, for wheat, and for wine, and for oil, and for the young of the flock and of the herd: and their soul shall be as a watered garden; and they shall not sorrow any more at all.
\vs Jer 31:13 Then shall the virgin rejoice in the dance, both young men and old together: for I will turn their mourning into joy, and will comfort them, and make them rejoice from their sorrow.
\vs Jer 31:14 And I will satiate the soul of the priests with fatness, and my people shall be satisfied with my goodness, saith the \textsc{Lord}.
\vs Jer 31:15 Thus saith the \textsc{Lord}; A voice was heard in Ramah, lamentation, \bibemph{and} bitter weeping; Rahel weeping for her children refused to be comforted for her children, because they \bibemph{were} not.
\vs Jer 31:16 Thus saith the \textsc{Lord}; Refrain thy voice from weeping, and thine eyes from tears: for thy work shall be rewarded, saith the \textsc{Lord}; and they shall come again from the land of the enemy.
\vs Jer 31:17 And there is hope in thine end, saith the \textsc{Lord}, that thy children shall come again to their own border.
\vs Jer 31:18 I have surely heard Ephraim bemoaning himself \bibemph{thus}; Thou hast chastised me, and I was chastised, as a bullock unaccustomed \bibemph{to the yoke}: turn thou me, and I shall be turned; for thou \bibemph{art} the \textsc{Lord} my God.
\vs Jer 31:19 Surely after that I was turned, I repented; and after that I was instructed, I smote upon \bibemph{my} thigh: I was ashamed, yea, even confounded, because I did bear the reproach of my youth.
\vs Jer 31:20 \bibemph{Is} Ephraim my dear son? \bibemph{is he} a pleasant child? for since I spake against him, I do earnestly remember him still: therefore my bowels are troubled for him; I will surely have mercy upon him, saith the \textsc{Lord}.
\fn{are...: Heb. sound.}
\vs Jer 31:21 Set thee up waymarks, make thee high heaps: set thine heart toward the highway, \bibemph{even} the way \bibemph{which} thou wentest: turn again, O virgin of Israel, turn again to these thy cities.
\vs Jer 31:22 How long wilt thou go about, O thou backsliding daughter? for the \textsc{Lord} hath created a new thing in the earth, A woman shall compass a man.
\vs Jer 31:23 Thus saith the \textsc{Lord} of hosts, the God of Israel; As yet they shall use this speech in the land of Judah and in the cities thereof, when I shall bring again their captivity; The \textsc{Lord} bless thee, O habitation of justice, \bibemph{and} mountain of holiness.
\vs Jer 31:24 And there shall dwell in Judah itself, and in all the cities thereof together, husbandmen, and they \bibemph{that} go forth with flocks.
\vs Jer 31:25 For I have satiated the weary soul, and I have replenished every sorrowful soul.
\vs Jer 31:26 Upon this I awaked, and beheld; and my sleep was sweet unto me.
\vs Jer 31:27 Behold, the days come, saith the \textsc{Lord}, that I will sow the house of Israel and the house of Judah with the seed of man, and with the seed of beast.
\vs Jer 31:28 And it shall come to pass, \bibemph{that} like as I have watched over them, to pluck up, and to break down, and to throw down, and to destroy, and to afflict; so will I watch over them, to build, and to plant, saith the \textsc{Lord}.
\vs Jer 31:29 In those days they shall say no more, The fathers have eaten a sour grape, and the children's teeth are set on edge.
\vs Jer 31:30 But every one shall die for his own iniquity: every man that eateth the sour grape, his teeth shall be set on edge.
\vs Jer 31:31 Behold, the days come, saith the \textsc{Lord}, that I will make a new covenant with the house of Israel, and with the house of Judah:
\vs Jer 31:32 Not according to the covenant that I made with their fathers in the day \bibemph{that} I took them by the hand to bring them out of the land of Egypt; which my covenant they brake, although I was an husband unto them, saith the \textsc{Lord}:
\fn{although...: or, should I have continued an husband unto them?.}
\vs Jer 31:33 But this \bibemph{shall be} the covenant that I will make with the house of Israel; After those days, saith the \textsc{Lord}, I will put my law in their inward parts, and write it in their hearts; and will be their God, and they shall be my people.
\vs Jer 31:34 And they shall teach no more every man his neighbour, and every man his brother, saying, Know the \textsc{Lord}: for they shall all know me, from the least of them unto the greatest of them, saith the \textsc{Lord}: for I will forgive their iniquity, and I will remember their sin no more.
\vs Jer 31:35 Thus saith the \textsc{Lord}, which giveth the sun for a light by day, \bibemph{and} the ordinances of the moon and of the stars for a light by night, which divideth the sea when the waves thereof roar; The \textsc{Lord} of hosts \bibemph{is} his name:
\vs Jer 31:36 If those ordinances depart from before me, saith the \textsc{Lord}, \bibemph{then} the seed of Israel also shall cease from being a nation before me for ever.
\vs Jer 31:37 Thus saith the \textsc{Lord}; If heaven above can be measured, and the foundations of the earth searched out beneath, I will also cast off all the seed of Israel for all that they have done, saith the \textsc{Lord}.
\vs Jer 31:38 Behold, the days come, saith the \textsc{Lord}, that the city shall be built to the \textsc{Lord} from the tower of Hananeel unto the gate of the corner.
\vs Jer 31:39 And the measuring line shall yet go forth over against it upon the hill Gareb, and shall compass about to Goath.
\vs Jer 31:40 And the whole valley of the dead bodies, and of the ashes, and all the fields unto the brook of Kidron, unto the corner of the horse gate toward the east, \bibemph{shall be} holy unto the \textsc{Lord}; it shall not be plucked up, nor thrown down any more for ever.
\vs Jer 32:1 The word that came to Jeremiah from the \textsc{Lord} in the tenth year of Zedekiah king of Judah, which \bibemph{was} the eighteenth year of Nebuchadrezzar.
\vs Jer 32:2 For then the king of Babylon's army besieged Jerusalem: and Jeremiah the prophet was shut up in the court of the prison, which \bibemph{was} in the king of Judah's house.
\vs Jer 32:3 For Zedekiah king of Judah had shut him up, saying, Wherefore dost thou prophesy, and say, Thus saith the \textsc{Lord}, Behold, I will give this city into the hand of the king of Babylon, and he shall take it;
\vs Jer 32:4 And Zedekiah king of Judah shall not escape out of the hand of the Chaldeans, but shall surely be delivered into the hand of the king of Babylon, and shall speak with him mouth to mouth, and his eyes shall behold his eyes;
\vs Jer 32:5 And he shall lead Zedekiah to Babylon, and there shall he be until I visit him, saith the \textsc{Lord}: though ye fight with the Chaldeans, ye shall not prosper.
\vs Jer 32:6 And Jeremiah said, The word of the \textsc{Lord} came unto me, saying,
\vs Jer 32:7 Behold, Hanameel the son of Shallum thine uncle shall come unto thee, saying, Buy thee my field that \bibemph{is} in Anathoth: for the right of redemption \bibemph{is} thine to buy \bibemph{it}.
\vs Jer 32:8 So Hanameel mine uncle's son came to me in the court of the prison according to the word of the \textsc{Lord}, and said unto me, Buy my field, I pray thee, that \bibemph{is} in Anathoth, which \bibemph{is} in the country of Benjamin: for the right of inheritance \bibemph{is} thine, and the redemption \bibemph{is} thine; buy \bibemph{it} for thyself. Then I knew that this \bibemph{was} the word of the \textsc{Lord}.
\vs Jer 32:9 And I bought the field of Hanameel my uncle's son, that \bibemph{was} in Anathoth, and weighed him the money, \bibemph{even} seventeen shekels of silver.
\fn{seventeen...: or, seven shekels and ten pieces of silver.}
\vs Jer 32:10 And I subscribed the evidence, and sealed \bibemph{it}, and took witnesses, and weighed \bibemph{him} the money in the balances.
\fn{subscribed...: Heb. wrote in the book.}
\vs Jer 32:11 So I took the evidence of the purchase, \bibemph{both} that which was sealed \bibemph{according} to the law and custom, and that which was open:
\vs Jer 32:12 And I gave the evidence of the purchase unto Baruch the son of Neriah, the son of Maaseiah, in the sight of Hanameel mine uncle's \bibemph{son}, and in the presence of the witnesses that subscribed the book of the purchase, before all the Jews that sat in the court of the prison.
\vs Jer 32:13 And I charged Baruch before them, saying,
\vs Jer 32:14 Thus saith the \textsc{Lord} of hosts, the God of Israel; Take these evidences, this evidence of the purchase, both which is sealed, and this evidence which is open; and put them in an earthen vessel, that they may continue many days.
\vs Jer 32:15 For thus saith the \textsc{Lord} of hosts, the God of Israel; Houses and fields and vineyards shall be possessed again in this land.
\vs Jer 32:16 Now when I had delivered the evidence of the purchase unto Baruch the son of Neriah, I prayed unto the \textsc{Lord}, saying,
\vs Jer 32:17 Ah Lord GOD! behold, thou hast made the heaven and the earth by thy great power and stretched out arm, \bibemph{and} there is nothing too hard for thee:
\fn{too...: or, hid from thee.}
\vs Jer 32:18 Thou shewest lovingkindness unto thousands, and recompensest the iniquity of the fathers into the bosom of their children after them: the Great, the Mighty God, the \textsc{Lord} of hosts, \bibemph{is} his name,
\vs Jer 32:19 Great in counsel, and mighty in work: for thine eyes \bibemph{are} open upon all the ways of the sons of men: to give every one according to his ways, and according to the fruit of his doings:
\fn{work: Heb. doing.}
\vs Jer 32:20 Which hast set signs and wonders in the land of Egypt, \bibemph{even} unto this day, and in Israel, and among \bibemph{other} men; and hast made thee a name, as at this day;
\vs Jer 32:21 And hast brought forth thy people Israel out of the land of Egypt with signs, and with wonders, and with a strong hand, and with a stretched out arm, and with great terror;
\vs Jer 32:22 And hast given them this land, which thou didst swear to their fathers to give them, a land flowing with milk and honey;
\vs Jer 32:23 And they came in, and possessed it; but they obeyed not thy voice, neither walked in thy law; they have done nothing of all that thou commandedst them to do: therefore thou hast caused all this evil to come upon them:
\vs Jer 32:24 Behold the mounts, they are come unto the city to take it; and the city is given into the hand of the Chaldeans, that fight against it, because of the sword, and of the famine, and of the pestilence: and what thou hast spoken is come to pass; and, behold, thou seest \bibemph{it}.
\fn{mounts: or, engines of shot.}
\vs Jer 32:25 And thou hast said unto me, O Lord GOD, Buy thee the field for money, and take witnesses; for the city is given into the hand of the Chaldeans.
\fn{for the: or, though the.}
\vs Jer 32:26 Then came the word of the \textsc{Lord} unto Jeremiah, saying,
\vs Jer 32:27 Behold, I \bibemph{am} the \textsc{Lord}, the God of all flesh: is there any thing too hard for me?
\vs Jer 32:28 Therefore thus saith the \textsc{Lord}; Behold, I will give this city into the hand of the Chaldeans, and into the hand of Nebuchadrezzar king of Babylon, and he shall take it:
\vs Jer 32:29 And the Chaldeans, that fight against this city, shall come and set fire on this city, and burn it with the houses, upon whose roofs they have offered incense unto Baal, and poured out drink offerings unto other gods, to provoke me to anger.
\vs Jer 32:30 For the children of Israel and the children of Judah have only done evil before me from their youth: for the children of Israel have only provoked me to anger with the work of their hands, saith the \textsc{Lord}.
\vs Jer 32:31 For this city hath been to me \bibemph{as} a provocation of mine anger and of my fury from the day that they built it even unto this day; that I should remove it from before my face,
\fn{a provocation...: Heb. for my anger.}
\vs Jer 32:32 Because of all the evil of the children of Israel and of the children of Judah, which they have done to provoke me to anger, they, their kings, their princes, their priests, and their prophets, and the men of Judah, and the inhabitants of Jerusalem.
\vs Jer 32:33 And they have turned unto me the back, and not the face: though I taught them, rising up early and teaching \bibemph{them}, yet they have not hearkened to receive instruction.
\fn{back: Heb. neck.}
\vs Jer 32:34 But they set their abominations in the house, which is called by my name, to defile it.
\vs Jer 32:35 And they built the high places of Baal, which \bibemph{are} in the valley of the son of Hinnom, to cause their sons and their daughters to pass through \bibemph{the fire} unto Molech; which I commanded them not, neither came it into my mind, that they should do this abomination, to cause Judah to sin.
\vs Jer 32:36 And now therefore thus saith the \textsc{Lord}, the God of Israel, concerning this city, whereof ye say, It shall be delivered into the hand of the king of Babylon by the sword, and by the famine, and by the pestilence;
\vs Jer 32:37 Behold, I will gather them out of all countries, whither I have driven them in mine anger, and in my fury, and in great wrath; and I will bring them again unto this place, and I will cause them to dwell safely:
\vs Jer 32:38 And they shall be my people, and I will be their God:
\vs Jer 32:39 And I will give them one heart, and one way, that they may fear me for ever, for the good of them, and of their children after them:
\fn{for ever: Heb. all days.}
\vs Jer 32:40 And I will make an everlasting covenant with them, that I will not turn away from them, to do them good; but I will put my fear in their hearts, that they shall not depart from me.
\fn{from them: Heb. from after them.}
\vs Jer 32:41 Yea, I will rejoice over them to do them good, and I will plant them in this land assuredly with my whole heart and with my whole soul.
\fn{assuredly: Heb. in truth, or, stability.}
\vs Jer 32:42 For thus saith the \textsc{Lord}; Like as I have brought all this great evil upon this people, so will I bring upon them all the good that I have promised them.
\vs Jer 32:43 And fields shall be bought in this land, whereof ye say, \bibemph{It is} desolate without man or beast; it is given into the hand of the Chaldeans.
\vs Jer 32:44 Men shall buy fields for money, and subscribe evidences, and seal \bibemph{them}, and take witnesses in the land of Benjamin, and in the places about Jerusalem, and in the cities of Judah, and in the cities of the mountains, and in the cities of the valley, and in the cities of the south: for I will cause their captivity to return, saith the \textsc{Lord}.
\vs Jer 33:1 Moreover the word of the \textsc{Lord} came unto Jeremiah the second time, while he was yet shut up in the court of the prison, saying,
\vs Jer 33:2 Thus saith the \textsc{Lord} the maker thereof, the \textsc{Lord} that formed it, to establish it; the \textsc{Lord} \bibemph{is} his name;
\fn{the \textsc{Lord} is: or, JEHOVAH, etc.}
\vs Jer 33:3 Call unto me, and I will answer thee, and shew thee great and mighty things, which thou knowest not.
\fn{mighty: or, hidden.}
\vs Jer 33:4 For thus saith the \textsc{Lord}, the God of Israel, concerning the houses of this city, and concerning the houses of the kings of Judah, which are thrown down by the mounts, and by the sword;
\vs Jer 33:5 They come to fight with the Chaldeans, but \bibemph{it is} to fill them with the dead bodies of men, whom I have slain in mine anger and in my fury, and for all whose wickedness I have hid my face from this city.
\vs Jer 33:6 Behold, I will bring it health and cure, and I will cure them, and will reveal unto them the abundance of peace and truth.
\vs Jer 33:7 And I will cause the captivity of Judah and the captivity of Israel to return, and will build them, as at the first.
\vs Jer 33:8 And I will cleanse them from all their iniquity, whereby they have sinned against me; and I will pardon all their iniquities, whereby they have sinned, and whereby they have transgressed against me.
\vs Jer 33:9 And it shall be to me a name of joy, a praise and an honour before all the nations of the earth, which shall hear all the good that I do unto them: and they shall fear and tremble for all the goodness and for all the prosperity that I procure unto it.
\vs Jer 33:10 Thus saith the \textsc{Lord}; Again there shall be heard in this place, which ye say \bibemph{shall be} desolate without man and without beast, \bibemph{even} in the cities of Judah, and in the streets of Jerusalem, that are desolate, without man, and without inhabitant, and without beast,
\vs Jer 33:11 The voice of joy, and the voice of gladness, the voice of the bridegroom, and the voice of the bride, the voice of them that shall say, Praise the \textsc{Lord} of hosts: for the \textsc{Lord} \bibemph{is} good; for his mercy \bibemph{endureth} for ever: \bibemph{and} of them that shall bring the sacrifice of praise into the house of the \textsc{Lord}. For I will cause to return the captivity of the land, as at the first, saith the \textsc{Lord}.
\vs Jer 33:12 Thus saith the \textsc{Lord} of hosts; Again in this place, which is desolate without man and without beast, and in all the cities thereof, shall be an habitation of shepherds causing \bibemph{their} flocks to lie down.
\vs Jer 33:13 In the cities of the mountains, in the cities of the vale, and in the cities of the south, and in the land of Benjamin, and in the places about Jerusalem, and in the cities of Judah, shall the flocks pass again under the hands of him that telleth \bibemph{them}, saith the \textsc{Lord}.
\vs Jer 33:14 Behold, the days come, saith the \textsc{Lord}, that I will perform that good thing which I have promised unto the house of Israel and to the house of Judah.
\vs Jer 33:15 In those days, and at that time, will I cause the Branch of righteousness to grow up unto David; and he shall execute judgment and righteousness in the land.
\vs Jer 33:16 In those days shall Judah be saved, and Jerusalem shall dwell safely: and this \bibemph{is the name} wherewith she shall be called, The \textsc{Lord} our righteousness.
\fn{The \textsc{Lord}...: Heb. Jehovahtsidkenu.}
\vs Jer 33:17 For thus saith the \textsc{Lord}; David shall never want a man to sit upon the throne of the house of Israel;
\fn{David...: Heb. There shall not be cut off from David.}
\vs Jer 33:18 Neither shall the priests the Levites want a man before me to offer burnt offerings, and to kindle meat offerings, and to do sacrifice continually.
\vs Jer 33:19 And the word of the \textsc{Lord} came unto Jeremiah, saying,
\vs Jer 33:20 Thus saith the \textsc{Lord}; If ye can break my covenant of the day, and my covenant of the night, and that there should not be day and night in their season;
\vs Jer 33:21 \bibemph{Then} may also my covenant be broken with David my servant, that he should not have a son to reign upon his throne; and with the Levites the priests, my ministers.
\vs Jer 33:22 As the host of heaven cannot be numbered, neither the sand of the sea measured: so will I multiply the seed of David my servant, and the Levites that minister unto me.
\vs Jer 33:23 Moreover the word of the \textsc{Lord} came to Jeremiah, saying,
\vs Jer 33:24 Considerest thou not what this people have spoken, saying, The two families which the \textsc{Lord} hath chosen, he hath even cast them off? thus they have despised my people, that they should be no more a nation before them.
\vs Jer 33:25 Thus saith the \textsc{Lord}; If my covenant \bibemph{be} not with day and night, \bibemph{and if} I have not appointed the ordinances of heaven and earth;
\vs Jer 33:26 Then will I cast away the seed of Jacob, and David my servant, \bibemph{so} that I will not take \bibemph{any} of his seed \bibemph{to be} rulers over the seed of Abraham, Isaac, and Jacob: for I will cause their captivity to return, and have mercy on them.
\vs Jer 34:1 The word which came unto Jeremiah from the \textsc{Lord}, when Nebuchadnezzar king of Babylon, and all his army, and all the kingdoms of the earth of his dominion, and all the people, fought against Jerusalem, and against all the cities thereof, saying,
\fn{of his...: Heb. the dominion of his hand.}
\vs Jer 34:2 Thus saith the \textsc{Lord}, the God of Israel; Go and speak to Zedekiah king of Judah, and tell him, Thus saith the \textsc{Lord}; Behold, I will give this city into the hand of the king of Babylon, and he shall burn it with fire:
\vs Jer 34:3 And thou shalt not escape out of his hand, but shalt surely be taken, and delivered into his hand; and thine eyes shall behold the eyes of the king of Babylon, and he shall speak with thee mouth to mouth, and thou shalt go to Babylon.
\fn{he shall...: Heb. his mouth shall speak to thy mouth.}
\vs Jer 34:4 Yet hear the word of the \textsc{Lord}, O Zedekiah king of Judah; Thus saith the \textsc{Lord} of thee, Thou shalt not die by the sword:
\vs Jer 34:5 \bibemph{But} thou shalt die in peace: and with the burnings of thy fathers, the former kings which were before thee, so shall they burn \bibemph{odours} for thee; and they will lament thee, \bibemph{saying}, Ah lord! for I have pronounced the word, saith the \textsc{Lord}.
\vs Jer 34:6 Then Jeremiah the prophet spake all these words unto Zedekiah king of Judah in Jerusalem,
\vs Jer 34:7 When the king of Babylon's army fought against Jerusalem, and against all the cities of Judah that were left, against Lachish, and against Azekah: for these defenced cities remained of the cities of Judah.
\vs Jer 34:8 \bibemph{This is} the word that came unto Jeremiah from the \textsc{Lord}, after that the king Zedekiah had made a covenant with all the people which \bibemph{were} at Jerusalem, to proclaim liberty unto them;
\vs Jer 34:9 That every man should let his manservant, and every man his maidservant, \bibemph{being} an Hebrew or an Hebrewess, go free; that none should serve himself of them, \bibemph{to wit}, of a Jew his brother.
\vs Jer 34:10 Now when all the princes, and all the people, which had entered into the covenant, heard that every one should let his manservant, and every one his maidservant, go free, that none should serve themselves of them any more, then they obeyed, and let \bibemph{them} go.
\vs Jer 34:11 But afterward they turned, and caused the servants and the handmaids, whom they had let go free, to return, and brought them into subjection for servants and for handmaids.
\vs Jer 34:12 Therefore the word of the \textsc{Lord} came to Jeremiah from the \textsc{Lord}, saying,
\vs Jer 34:13 Thus saith the \textsc{Lord}, the God of Israel; I made a covenant with your fathers in the day that I brought them forth out of the land of Egypt, out of the house of bondmen, saying,
\vs Jer 34:14 At the end of seven years let ye go every man his brother an Hebrew, which hath been sold unto thee; and when he hath served thee six years, thou shalt let him go free from thee: but your fathers hearkened not unto me, neither inclined their ear.
\fn{hath been...: or, hath sold himself.}
\vs Jer 34:15 And ye were now turned, and had done right in my sight, in proclaiming liberty every man to his neighbour; and ye had made a covenant before me in the house which is called by my name:
\fn{now: Heb. to day. which...: Heb. whereupon my name is called.}
\vs Jer 34:16 But ye turned and polluted my name, and caused every man his servant, and every man his handmaid, whom ye had set at liberty at their pleasure, to return, and brought them into subjection, to be unto you for servants and for handmaids.
\vs Jer 34:17 Therefore thus saith the \textsc{Lord}; Ye have not hearkened unto me, in proclaiming liberty, every one to his brother, and every man to his neighbour: behold, I proclaim a liberty for you, saith the \textsc{Lord}, to the sword, to the pestilence, and to the famine; and I will make you to be removed into all the kingdoms of the earth.
\fn{to be...: Heb. for a removing.}
\vs Jer 34:18 And I will give the men that have transgressed my covenant, which have not performed the words of the covenant which they had made before me, when they cut the calf in twain, and passed between the parts thereof,
\vs Jer 34:19 The princes of Judah, and the princes of Jerusalem, the eunuchs, and the priests, and all the people of the land, which passed between the parts of the calf;
\vs Jer 34:20 I will even give them into the hand of their enemies, and into the hand of them that seek their life: and their dead bodies shall be for meat unto the fowls of the heaven, and to the beasts of the earth.
\vs Jer 34:21 And Zedekiah king of Judah and his princes will I give into the hand of their enemies, and into the hand of them that seek their life, and into the hand of the king of Babylon's army, which are gone up from you.
\vs Jer 34:22 Behold, I will command, saith the \textsc{Lord}, and cause them to return to this city; and they shall fight against it, and take it, and burn it with fire: and I will make the cities of Judah a desolation without an inhabitant.
\vs Jer 35:1 The word which came unto Jeremiah from the \textsc{Lord} in the days of Jehoiakim the son of Josiah king of Judah, saying,
\vs Jer 35:2 Go unto the house of the Rechabites, and speak unto them, and bring them into the house of the \textsc{Lord}, into one of the chambers, and give them wine to drink.
\vs Jer 35:3 Then I took Jaazaniah the son of Jeremiah, the son of Habaziniah, and his brethren, and all his sons, and the whole house of the Rechabites;
\vs Jer 35:4 And I brought them into the house of the \textsc{Lord}, into the chamber of the sons of Hanan, the son of Igdaliah, a man of God, which \bibemph{was} by the chamber of the princes, which \bibemph{was} above the chamber of Maaseiah the son of Shallum, the keeper of the door:
\fn{door: Heb. threshold, or, vessel.}
\vs Jer 35:5 And I set before the sons of the house of the Rechabites pots full of wine, and cups, and I said unto them, Drink ye wine.
\vs Jer 35:6 But they said, We will drink no wine: for Jonadab the son of Rechab our father commanded us, saying, Ye shall drink no wine, \bibemph{neither ye}, nor your sons for ever:
\vs Jer 35:7 Neither shall ye build house, nor sow seed, nor plant vineyard, nor have \bibemph{any}: but all your days ye shall dwell in tents; that ye may live many days in the land where ye \bibemph{be} strangers.
\vs Jer 35:8 Thus have we obeyed the voice of Jonadab the son of Rechab our father in all that he hath charged us, to drink no wine all our days, we, our wives, our sons, nor our daughters;
\vs Jer 35:9 Nor to build houses for us to dwell in: neither have we vineyard, nor field, nor seed:
\vs Jer 35:10 But we have dwelt in tents, and have obeyed, and done according to all that Jonadab our father commanded us.
\vs Jer 35:11 But it came to pass, when Nebuchadrezzar king of Babylon came up into the land, that we said, Come, and let us go to Jerusalem for fear of the army of the Chaldeans, and for fear of the army of the Syrians: so we dwell at Jerusalem.
\vs Jer 35:12 Then came the word of the \textsc{Lord} unto Jeremiah, saying,
\vs Jer 35:13 Thus saith the \textsc{Lord} of hosts, the God of Israel; Go and tell the men of Judah and the inhabitants of Jerusalem, Will ye not receive instruction to hearken to my words? saith the \textsc{Lord}.
\vs Jer 35:14 The words of Jonadab the son of Rechab, that he commanded his sons not to drink wine, are performed; for unto this day they drink none, but obey their father's commandment: notwithstanding I have spoken unto you, rising early and speaking; but ye hearkened not unto me.
\vs Jer 35:15 I have sent also unto you all my servants the prophets, rising up early and sending \bibemph{them}, saying, Return ye now every man from his evil way, and amend your doings, and go not after other gods to serve them, and ye shall dwell in the land which I have given to you and to your fathers: but ye have not inclined your ear, nor hearkened unto me.
\vs Jer 35:16 Because the sons of Jonadab the son of Rechab have performed the commandment of their father, which he commanded them; but this people hath not hearkened unto me:
\vs Jer 35:17 Therefore thus saith the \textsc{Lord} God of hosts, the God of Israel; Behold, I will bring upon Judah and upon all the inhabitants of Jerusalem all the evil that I have pronounced against them: because I have spoken unto them, but they have not heard; and I have called unto them, but they have not answered.
\vs Jer 35:18 And Jeremiah said unto the house of the Rechabites, Thus saith the \textsc{Lord} of hosts, the God of Israel; Because ye have obeyed the commandment of Jonadab your father, and kept all his precepts, and done according unto all that he hath commanded you:
\vs Jer 35:19 Therefore thus saith the \textsc{Lord} of hosts, the God of Israel; Jonadab the son of Rechab shall not want a man to stand before me for ever.
\fn{Jonadab...: Heb. There shall not a man be cut off from Jonadab the son of Rechab to stand, etc.}
\vs Jer 36:1 And it came to pass in the fourth year of Jehoiakim the son of Josiah king of Judah, \bibemph{that} this word came unto Jeremiah from the \textsc{Lord}, saying,
\vs Jer 36:2 Take thee a roll of a book, and write therein all the words that I have spoken unto thee against Israel, and against Judah, and against all the nations, from the day I spake unto thee, from the days of Josiah, even unto this day.
\vs Jer 36:3 It may be that the house of Judah will hear all the evil which I purpose to do unto them; that they may return every man from his evil way; that I may forgive their iniquity and their sin.
\vs Jer 36:4 Then Jeremiah called Baruch the son of Neriah: and Baruch wrote from the mouth of Jeremiah all the words of the \textsc{Lord}, which he had spoken unto him, upon a roll of a book.
\vs Jer 36:5 And Jeremiah commanded Baruch, saying, I \bibemph{am} shut up; I cannot go into the house of the \textsc{Lord}:
\vs Jer 36:6 Therefore go thou, and read in the roll, which thou hast written from my mouth, the words of the \textsc{Lord} in the ears of the people in the \textsc{Lord's} house upon the fasting day: and also thou shalt read them in the ears of all Judah that come out of their cities.
\vs Jer 36:7 It may be they will present their supplication before the \textsc{Lord}, and will return every one from his evil way: for great \bibemph{is} the anger and the fury that the \textsc{Lord} hath pronounced against this people.
\fn{they...: Heb. their supplication shall fail.}
\vs Jer 36:8 And Baruch the son of Neriah did according to all that Jeremiah the prophet commanded him, reading in the book the words of the \textsc{Lord} in the \textsc{Lord's} house.
\vs Jer 36:9 And it came to pass in the fifth year of Jehoiakim the son of Josiah king of Judah, in the ninth month, \bibemph{that} they proclaimed a fast before the \textsc{Lord} to all the people in Jerusalem, and to all the people that came from the cities of Judah unto Jerusalem.
\vs Jer 36:10 Then read Baruch in the book the words of Jeremiah in the house of the \textsc{Lord}, in the chamber of Gemariah the son of Shaphan the scribe, in the higher court, at the entry of the new gate of the \textsc{Lord's} house, in the ears of all the people.
\fn{entry: or, door.}
\vs Jer 36:11 When Michaiah the son of Gemariah, the son of Shaphan, had heard out of the book all the words of the \textsc{Lord},
\vs Jer 36:12 Then he went down into the king's house, into the scribe's chamber: and, lo, all the princes sat there, \bibemph{even} Elishama the scribe, and Delaiah the son of Shemaiah, and Elnathan the son of Achbor, and Gemariah the son of Shaphan, and Zedekiah the son of Hananiah, and all the princes.
\vs Jer 36:13 Then Michaiah declared unto them all the words that he had heard, when Baruch read the book in the ears of the people.
\vs Jer 36:14 Therefore all the princes sent Jehudi the son of Nethaniah, the son of Shelemiah, the son of Cushi, unto Baruch, saying, Take in thine hand the roll wherein thou hast read in the ears of the people, and come. So Baruch the son of Neriah took the roll in his hand, and came unto them.
\vs Jer 36:15 And they said unto him, Sit down now, and read it in our ears. So Baruch read \bibemph{it} in their ears.
\vs Jer 36:16 Now it came to pass, when they had heard all the words, they were afraid both one and other, and said unto Baruch, We will surely tell the king of all these words.
\vs Jer 36:17 And they asked Baruch, saying, Tell us now, How didst thou write all these words at his mouth?
\vs Jer 36:18 Then Baruch answered them, He pronounced all these words unto me with his mouth, and I wrote \bibemph{them} with ink in the book.
\vs Jer 36:19 Then said the princes unto Baruch, Go, hide thee, thou and Jeremiah; and let no man know where ye be.
\vs Jer 36:20 And they went in to the king into the court, but they laid up the roll in the chamber of Elishama the scribe, and told all the words in the ears of the king.
\vs Jer 36:21 So the king sent Jehudi to fetch the roll: and he took it out of Elishama the scribe's chamber. And Jehudi read it in the ears of the king, and in the ears of all the princes which stood beside the king.
\vs Jer 36:22 Now the king sat in the winterhouse in the ninth month: and \bibemph{there was a fire} on the hearth burning before him.
\vs Jer 36:23 And it came to pass, \bibemph{that} when Jehudi had read three or four leaves, he cut it with the penknife, and cast \bibemph{it} into the fire that \bibemph{was} on the hearth, until all the roll was consumed in the fire that \bibemph{was} on the hearth.
\vs Jer 36:24 Yet they were not afraid, nor rent their garments, \bibemph{neither} the king, nor any of his servants that heard all these words.
\vs Jer 36:25 Nevertheless Elnathan and Delaiah and Gemariah had made intercession to the king that he would not burn the roll: but he would not hear them.
\vs Jer 36:26 But the king commanded Jerahmeel the son of Hammelech, and Seraiah the son of Azriel, and Shelemiah the son of Abdeel, to take Baruch the scribe and Jeremiah the prophet: but the \textsc{Lord} hid them.
\fn{of Hammelech: or, of the king.}
\vs Jer 36:27 Then the word of the \textsc{Lord} came to Jeremiah, after that the king had burned the roll, and the words which Baruch wrote at the mouth of Jeremiah, saying,
\vs Jer 36:28 Take thee again another roll, and write in it all the former words that were in the first roll, which Jehoiakim the king of Judah hath burned.
\vs Jer 36:29 And thou shalt say to Jehoiakim king of Judah, Thus saith the \textsc{Lord}; Thou hast burned this roll, saying, Why hast thou written therein, saying, The king of Babylon shall certainly come and destroy this land, and shall cause to cease from thence man and beast?
\vs Jer 36:30 Therefore thus saith the \textsc{Lord} of Jehoiakim king of Judah; He shall have none to sit upon the throne of David: and his dead body shall be cast out in the day to the heat, and in the night to the frost.
\vs Jer 36:31 And I will punish him and his seed and his servants for their iniquity; and I will bring upon them, and upon the inhabitants of Jerusalem, and upon the men of Judah, all the evil that I have pronounced against them; but they hearkened not.
\fn{punish: Heb. visit upon.}
\vs Jer 36:32 Then took Jeremiah another roll, and gave it to Baruch the scribe, the son of Neriah; who wrote therein from the mouth of Jeremiah all the words of the book which Jehoiakim king of Judah had burned in the fire: and there were added besides unto them many like words.
\fn{like: Heb. as they.}
\vs Jer 37:1 And king Zedekiah the son of Josiah reigned instead of Coniah the son of Jehoiakim, whom Nebuchadrezzar king of Babylon made king in the land of Judah.
\vs Jer 37:2 But neither he, nor his servants, nor the people of the land, did hearken unto the words of the \textsc{Lord}, which he spake by the prophet Jeremiah.
\fn{by...: Heb. by the hand of the prophet.}
\vs Jer 37:3 And Zedekiah the king sent Jehucal the son of Shelemiah and Zephaniah the son of Maaseiah the priest to the prophet Jeremiah, saying, Pray now unto the \textsc{Lord} our God for us.
\vs Jer 37:4 Now Jeremiah came in and went out among the people: for they had not put him into prison.
\vs Jer 37:5 Then Pharaoh's army was come forth out of Egypt: and when the Chaldeans that besieged Jerusalem heard tidings of them, they departed from Jerusalem.
\vs Jer 37:6 Then came the word of the \textsc{Lord} unto the prophet Jeremiah, saying,
\vs Jer 37:7 Thus saith the \textsc{Lord}, the God of Israel; Thus shall ye say to the king of Judah, that sent you unto me to enquire of me; Behold, Pharaoh's army, which is come forth to help you, shall return to Egypt into their own land.
\vs Jer 37:8 And the Chaldeans shall come again, and fight against this city, and take it, and burn it with fire.
\vs Jer 37:9 Thus saith the \textsc{Lord}; Deceive not yourselves, saying, The Chaldeans shall surely depart from us: for they shall not depart.
\fn{yourselves: Heb. your souls.}
\vs Jer 37:10 For though ye had smitten the whole army of the Chaldeans that fight against you, and there remained \bibemph{but} wounded men among them, \bibemph{yet} should they rise up every man in his tent, and burn this city with fire.
\fn{wounded: Heb. thrust through.}
\vs Jer 37:11 And it came to pass, that when the army of the Chaldeans was broken up from Jerusalem for fear of Pharaoh's army,
\fn{broken...: Heb. made to ascend.}
\vs Jer 37:12 Then Jeremiah went forth out of Jerusalem to go into the land of Benjamin, to separate himself thence in the midst of the people.
\fn{separate...: or, to slip away from thence in the midst of the people.}
\vs Jer 37:13 And when he was in the gate of Benjamin, a captain of the ward \bibemph{was} there, whose name \bibemph{was} Irijah, the son of Shelemiah, the son of Hananiah; and he took Jeremiah the prophet, saying, Thou fallest away to the Chaldeans.
\vs Jer 37:14 Then said Jeremiah, \bibemph{It is} false; I fall not away to the Chaldeans. But he hearkened not to him: so Irijah took Jeremiah, and brought him to the princes.
\fn{false: Heb. falsehood, or, a lie.}
\vs Jer 37:15 Wherefore the princes were wroth with Jeremiah, and smote him, and put him in prison in the house of Jonathan the scribe: for they had made that the prison.
\vs Jer 37:16 When Jeremiah was entered into the dungeon, and into the cabins, and Jeremiah had remained there many days;
\fn{cabins: or, cells.}
\vs Jer 37:17 Then Zedekiah the king sent, and took him out: and the king asked him secretly in his house, and said, Is there \bibemph{any} word from the \textsc{Lord}? And Jeremiah said, There is: for, said he, thou shalt be delivered into the hand of the king of Babylon.
\vs Jer 37:18 Moreover Jeremiah said unto king Zedekiah, What have I offended against thee, or against thy servants, or against this people, that ye have put me in prison?
\vs Jer 37:19 Where \bibemph{are} now your prophets which prophesied unto you, saying, The king of Babylon shall not come against you, nor against this land?
\vs Jer 37:20 Therefore hear now, I pray thee, O my lord the king: let my supplication, I pray thee, be accepted before thee; that thou cause me not to return to the house of Jonathan the scribe, lest I die there.
\fn{let...: Heb. let my supplication fall.}
\vs Jer 37:21 Then Zedekiah the king commanded that they should commit Jeremiah into the court of the prison, and that they should give him daily a piece of bread out of the bakers' street, until all the bread in the city were spent. Thus Jeremiah remained in the court of the prison.
\vs Jer 38:1 Then Shephatiah the son of Mattan, and Gedaliah the son of Pashur, and Jucal the son of Shelemiah, and Pashur the son of Malchiah, heard the words that Jeremiah had spoken unto all the people, saying,
\vs Jer 38:2 Thus saith the \textsc{Lord}, He that remaineth in this city shall die by the sword, by the famine, and by the pestilence: but he that goeth forth to the Chaldeans shall live; for he shall have his life for a prey, and shall live.
\vs Jer 38:3 Thus saith the \textsc{Lord}, This city shall surely be given into the hand of the king of Babylon's army, which shall take it.
\vs Jer 38:4 Therefore the princes said unto the king, We beseech thee, let this man be put to death: for thus he weakeneth the hands of the men of war that remain in this city, and the hands of all the people, in speaking such words unto them: for this man seeketh not the welfare of this people, but the hurt.
\fn{welfare: Heb. peace.}
\vs Jer 38:5 Then Zedekiah the king said, Behold, he \bibemph{is} in your hand: for the king \bibemph{is} not \bibemph{he that} can do \bibemph{any} thing against you.
\vs Jer 38:6 Then took they Jeremiah, and cast him into the dungeon of Malchiah the son of Hammelech, that \bibemph{was} in the court of the prison: and they let down Jeremiah with cords. And in the dungeon \bibemph{there was} no water, but mire: so Jeremiah sunk in the mire.
\fn{of Hammelech: or, of the king.}
\vs Jer 38:7 Now when Ebedmelech the Ethiopian, one of the eunuchs which was in the king's house, heard that they had put Jeremiah in the dungeon; the king then sitting in the gate of Benjamin;
\vs Jer 38:8 Ebedmelech went forth out of the king's house, and spake to the king, saying,
\vs Jer 38:9 My lord the king, these men have done evil in all that they have done to Jeremiah the prophet, whom they have cast into the dungeon; and he is like to die for hunger in the place where he is: for \bibemph{there is} no more bread in the city.
\fn{he is like...: Heb. he will die.}
\vs Jer 38:10 Then the king commanded Ebedmelech the Ethiopian, saying, Take from hence thirty men with thee, and take up Jeremiah the prophet out of the dungeon, before he die.
\fn{with thee: Heb. in thine hand.}
\vs Jer 38:11 So Ebedmelech took the men with him, and went into the house of the king under the treasury, and took thence old cast clouts and old rotten rags, and let them down by cords into the dungeon to Jeremiah.
\vs Jer 38:12 And Ebedmelech the Ethiopian said unto Jeremiah, Put now \bibemph{these} old cast clouts and rotten rags under thine armholes under the cords. And Jeremiah did so.
\vs Jer 38:13 So they drew up Jeremiah with cords, and took him up out of the dungeon: and Jeremiah remained in the court of the prison.
\vs Jer 38:14 Then Zedekiah the king sent, and took Jeremiah the prophet unto him into the third entry that \bibemph{is} in the house of the \textsc{Lord}: and the king said unto Jeremiah, I will ask thee a thing; hide nothing from me.
\fn{third: or, principal.}
\vs Jer 38:15 Then Jeremiah said unto Zedekiah, If I declare \bibemph{it} unto thee, wilt thou not surely put me to death? and if I give thee counsel, wilt thou not hearken unto me?
\vs Jer 38:16 So Zedekiah the king sware secretly unto Jeremiah, saying, \bibemph{As} the \textsc{Lord} liveth, that made us this soul, I will not put thee to death, neither will I give thee into the hand of these men that seek thy life.
\vs Jer 38:17 Then said Jeremiah unto Zedekiah, Thus saith the \textsc{Lord}, the God of hosts, the God of Israel; If thou wilt assuredly go forth unto the king of Babylon's princes, then thy soul shall live, and this city shall not be burned with fire; and thou shalt live, and thine house:
\vs Jer 38:18 But if thou wilt not go forth to the king of Babylon's princes, then shall this city be given into the hand of the Chaldeans, and they shall burn it with fire, and thou shalt not escape out of their hand.
\vs Jer 38:19 And Zedekiah the king said unto Jeremiah, I am afraid of the Jews that are fallen to the Chaldeans, lest they deliver me into their hand, and they mock me.
\vs Jer 38:20 But Jeremiah said, They shall not deliver \bibemph{thee}. Obey, I beseech thee, the voice of the \textsc{Lord}, which I speak unto thee: so it shall be well unto thee, and thy soul shall live.
\vs Jer 38:21 But if thou refuse to go forth, this \bibemph{is} the word that the \textsc{Lord} hath shewed me:
\vs Jer 38:22 And, behold, all the women that are left in the king of Judah's house \bibemph{shall be} brought forth to the king of Babylon's princes, and those \bibemph{women} shall say, Thy friends have set thee on, and have prevailed against thee: thy feet are sunk in the mire, \bibemph{and} they are turned away back.
\fn{Thy friends: Heb. Men of thy peace.}
\vs Jer 38:23 So they shall bring out all thy wives and thy children to the Chaldeans: and thou shalt not escape out of their hand, but shalt be taken by the hand of the king of Babylon: and thou shalt cause this city to be burned with fire.
\fn{thou shalt cause...: Heb. thou shalt burn, etc.}
\vs Jer 38:24 Then said Zedekiah unto Jeremiah, Let no man know of these words, and thou shalt not die.
\vs Jer 38:25 But if the princes hear that I have talked with thee, and they come unto thee, and say unto thee, Declare unto us now what thou hast said unto the king, hide it not from us, and we will not put thee to death; also what the king said unto thee:
\vs Jer 38:26 Then thou shalt say unto them, I presented my supplication before the king, that he would not cause me to return to Jonathan's house, to die there.
\vs Jer 38:27 Then came all the princes unto Jeremiah, and asked him: and he told them according to all these words that the king had commanded. So they left off speaking with him; for the matter was not perceived.
\fn{they...: Heb. they were silent from him.}
\vs Jer 38:28 So Jeremiah abode in the court of the prison until the day that Jerusalem was taken: and he was \bibemph{there} when Jerusalem was taken.
\vs Jer 39:1 In the ninth year of Zedekiah king of Judah, in the tenth month, came Nebuchadrezzar king of Babylon and all his army against Jerusalem, and they besieged it.
\vs Jer 39:2 \bibemph{And} in the eleventh year of Zedekiah, in the fourth month, the ninth \bibemph{day} of the month, the city was broken up.
\vs Jer 39:3 And all the princes of the king of Babylon came in, and sat in the middle gate, \bibemph{even} Nergalsharezer, Samgarnebo, Sarsechim, Rabsaris, Nergalsharezer, Rabmag, with all the residue of the princes of the king of Babylon.
\vs Jer 39:4 And it came to pass, \bibemph{that} when Zedekiah the king of Judah saw them, and all the men of war, then they fled, and went forth out of the city by night, by the way of the king's garden, by the gate betwixt the two walls: and he went out the way of the plain.
\vs Jer 39:5 But the Chaldeans' army pursued after them, and overtook Zedekiah in the plains of Jericho: and when they had taken him, they brought him up to Nebuchadnezzar king of Babylon to Riblah in the land of Hamath, where he gave judgment upon him.
\fn{gave...: Heb. spake with him judgments.}
\vs Jer 39:6 Then the king of Babylon slew the sons of Zedekiah in Riblah before his eyes: also the king of Babylon slew all the nobles of Judah.
\vs Jer 39:7 Moreover he put out Zedekiah's eyes, and bound him with chains, to carry him to Babylon.
\fn{with...: Heb. with two brasen chains, or, fetters.}
\vs Jer 39:8 And the Chaldeans burned the king's house, and the houses of the people, with fire, and brake down the walls of Jerusalem.
\vs Jer 39:9 Then Nebuzaradan the captain of the guard carried away captive into Babylon the remnant of the people that remained in the city, and those that fell away, that fell to him, with the rest of the people that remained.
\fn{captain...: or, chief marshal: Heb. chief of the executioners, or, slaughtermen.}
\vs Jer 39:10 But Nebuzaradan the captain of the guard left of the poor of the people, which had nothing, in the land of Judah, and gave them vineyards and fields at the same time.
\fn{at the...: Heb. in that day.}
\vs Jer 39:11 Now Nebuchadrezzar king of Babylon gave charge concerning Jeremiah to Nebuzaradan the captain of the guard, saying,
\fn{to: Heb. by the hand of.}
\vs Jer 39:12 Take him, and look well to him, and do him no harm; but do unto him even as he shall say unto thee.
\fn{look...: Heb. set thine eyes upon him.}
\vs Jer 39:13 So Nebuzaradan the captain of the guard sent, and Nebushasban, Rabsaris, and Nergalsharezer, Rabmag, and all the king of Babylon's princes;
\vs Jer 39:14 Even they sent, and took Jeremiah out of the court of the prison, and committed him unto Gedaliah the son of Ahikam the son of Shaphan, that he should carry him home: so he dwelt among the people.
\vs Jer 39:15 Now the word of the \textsc{Lord} came unto Jeremiah, while he was shut up in the court of the prison, saying,
\vs Jer 39:16 Go and speak to Ebedmelech the Ethiopian, saying, Thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will bring my words upon this city for evil, and not for good; and they shall be \bibemph{accomplished} in that day before thee.
\vs Jer 39:17 But I will deliver thee in that day, saith the \textsc{Lord}: and thou shalt not be given into the hand of the men of whom thou \bibemph{art} afraid.
\vs Jer 39:18 For I will surely deliver thee, and thou shalt not fall by the sword, but thy life shall be for a prey unto thee: because thou hast put thy trust in me, saith the \textsc{Lord}.
\vs Jer 40:1 The word that came to Jeremiah from the \textsc{Lord}, after that Nebuzaradan the captain of the guard had let him go from Ramah, when he had taken him being bound in chains among all that were carried away captive of Jerusalem and Judah, which were carried away captive unto Babylon.
\fn{chains: or, manicles.}
\vs Jer 40:2 And the captain of the guard took Jeremiah, and said unto him, The \textsc{Lord} thy God hath pronounced this evil upon this place.
\vs Jer 40:3 Now the \textsc{Lord} hath brought \bibemph{it}, and done according as he hath said: because ye have sinned against the \textsc{Lord}, and have not obeyed his voice, therefore this thing is come upon you.
\vs Jer 40:4 And now, behold, I loose thee this day from the chains which \bibemph{were} upon thine hand. If it seem good unto thee to come with me into Babylon, come; and I will look well unto thee: but if it seem ill unto thee to come with me into Babylon, forbear: behold, all the land \bibemph{is} before thee: whither it seemeth good and convenient for thee to go, thither go.
\fn{were...: or, are upon thine hand. I will...: Heb. I will set mine eye upon thee.}
\vs Jer 40:5 Now while he was not yet gone back, \bibemph{he said}, Go back also to Gedaliah the son of Ahikam the son of Shaphan, whom the king of Babylon hath made governor over the cities of Judah, and dwell with him among the people: or go wheresoever it seemeth convenient unto thee to go. So the captain of the guard gave him victuals and a reward, and let him go.
\vs Jer 40:6 Then went Jeremiah unto Gedaliah the son of Ahikam to Mizpah; and dwelt with him among the people that were left in the land.
\vs Jer 40:7 Now when all the captains of the forces which \bibemph{were} in the fields, \bibemph{even} they and their men, heard that the king of Babylon had made Gedaliah the son of Ahikam governor in the land, and had committed unto him men, and women, and children, and of the poor of the land, of them that were not carried away captive to Babylon;
\vs Jer 40:8 Then they came to Gedaliah to Mizpah, even Ishmael the son of Nethaniah, and Johanan and Jonathan the sons of Kareah, and Seraiah the son of Tanhumeth, and the sons of Ephai the Netophathite, and Jezaniah the son of a Maachathite, they and their men.
\vs Jer 40:9 And Gedaliah the son of Ahikam the son of Shaphan sware unto them and to their men, saying, Fear not to serve the Chaldeans: dwell in the land, and serve the king of Babylon, and it shall be well with you.
\vs Jer 40:10 As for me, behold, I will dwell at Mizpah to serve the Chaldeans, which will come unto us: but ye, gather ye wine, and summer fruits, and oil, and put \bibemph{them} in your vessels, and dwell in your cities that ye have taken.
\fn{to serve: Heb. to stand before.}
\vs Jer 40:11 Likewise when all the Jews that \bibemph{were} in Moab, and among the Ammonites, and in Edom, and that \bibemph{were} in all the countries, heard that the king of Babylon had left a remnant of Judah, and that he had set over them Gedaliah the son of Ahikam the son of Shaphan;
\vs Jer 40:12 Even all the Jews returned out of all places whither they were driven, and came to the land of Judah, to Gedaliah, unto Mizpah, and gathered wine and summer fruits very much.
\vs Jer 40:13 Moreover Johanan the son of Kareah, and all the captains of the forces that \bibemph{were} in the fields, came to Gedaliah to Mizpah,
\vs Jer 40:14 And said unto him, Dost thou certainly know that Baalis the king of the Ammonites hath sent Ishmael the son of Nethaniah to slay thee? But Gedaliah the son of Ahikam believed them not.
\fn{to slay...: Heb. to strike thee in soul?.}
\vs Jer 40:15 Then Johanan the son of Kareah spake to Gedaliah in Mizpah secretly, saying, Let me go, I pray thee, and I will slay Ishmael the son of Nethaniah, and no man shall know \bibemph{it}: wherefore should he slay thee, that all the Jews which are gathered unto thee should be scattered, and the remnant in Judah perish?
\vs Jer 40:16 But Gedaliah the son of Ahikam said unto Johanan the son of Kareah, Thou shalt not do this thing: for thou speakest falsely of Ishmael.
\vs Jer 41:1 Now it came to pass in the seventh month, \bibemph{that} Ishmael the son of Nethaniah the son of Elishama, of the seed royal, and the princes of the king, even ten men with him, came unto Gedaliah the son of Ahikam to Mizpah; and there they did eat bread together in Mizpah.
\vs Jer 41:2 Then arose Ishmael the son of Nethaniah, and the ten men that were with him, and smote Gedaliah the son of Ahikam the son of Shaphan with the sword, and slew him, whom the king of Babylon had made governor over the land.
\vs Jer 41:3 Ishmael also slew all the Jews that were with him, \bibemph{even} with Gedaliah, at Mizpah, and the Chaldeans that were found there, \bibemph{and} the men of war.
\vs Jer 41:4 And it came to pass the second day after he had slain Gedaliah, and no man knew \bibemph{it},
\vs Jer 41:5 That there came certain from Shechem, from Shiloh, and from Samaria, \bibemph{even} fourscore men, having their beards shaven, and their clothes rent, and having cut themselves, with offerings and incense in their hand, to bring \bibemph{them} to the house of the \textsc{Lord}.
\vs Jer 41:6 And Ishmael the son of Nethaniah went forth from Mizpah to meet them, weeping all along as he went: and it came to pass, as he met them, he said unto them, Come to Gedaliah the son of Ahikam.
\fn{weeping...: Heb. in going and weeping.}
\vs Jer 41:7 And it was \bibemph{so}, when they came into the midst of the city, that Ishmael the son of Nethaniah slew them, \bibemph{and cast them} into the midst of the pit, he, and the men that \bibemph{were} with him.
\vs Jer 41:8 But ten men were found among them that said unto Ishmael, Slay us not: for we have treasures in the field, of wheat, and of barley, and of oil, and of honey. So he forbare, and slew them not among their brethren.
\vs Jer 41:9 Now the pit wherein Ishmael had cast all the dead bodies of the men, whom he had slain because of Gedaliah, \bibemph{was} it which Asa the king had made for fear of Baasha king of Israel: \bibemph{and} Ishmael the son of Nethaniah filled it with \bibemph{them that were} slain.
\fn{because...: or, near Gedaliah: Heb. by the hand, or, by the side of Gedaliah.}
\vs Jer 41:10 Then Ishmael carried away captive all the residue of the people that \bibemph{were} in Mizpah, \bibemph{even} the king's daughters, and all the people that remained in Mizpah, whom Nebuzaradan the captain of the guard had committed to Gedaliah the son of Ahikam: and Ishmael the son of Nethaniah carried them away captive, and departed to go over to the Ammonites.
\vs Jer 41:11 But when Johanan the son of Kareah, and all the captains of the forces that \bibemph{were} with him, heard of all the evil that Ishmael the son of Nethaniah had done,
\vs Jer 41:12 Then they took all the men, and went to fight with Ishmael the son of Nethaniah, and found him by the great waters that \bibemph{are} in Gibeon.
\vs Jer 41:13 Now it came to pass, \bibemph{that} when all the people which \bibemph{were} with Ishmael saw Johanan the son of Kareah, and all the captains of the forces that \bibemph{were} with him, then they were glad.
\vs Jer 41:14 So all the people that Ishmael had carried away captive from Mizpah cast about and returned, and went unto Johanan the son of Kareah.
\vs Jer 41:15 But Ishmael the son of Nethaniah escaped from Johanan with eight men, and went to the Ammonites.
\vs Jer 41:16 Then took Johanan the son of Kareah, and all the captains of the forces that \bibemph{were} with him, all the remnant of the people whom he had recovered from Ishmael the son of Nethaniah, from Mizpah, after \bibemph{that} he had slain Gedaliah the son of Ahikam, \bibemph{even} mighty men of war, and the women, and the children, and the eunuchs, whom he had brought again from Gibeon:
\vs Jer 41:17 And they departed, and dwelt in the habitation of Chimham, which is by Bethlehem, to go to enter into Egypt,
\vs Jer 41:18 Because of the Chaldeans: for they were afraid of them, because Ishmael the son of Nethaniah had slain Gedaliah the son of Ahikam, whom the king of Babylon made governor in the land.
\vs Jer 42:1 Then all the captains of the forces, and Johanan the son of Kareah, and Jezaniah the son of Hoshaiah, and all the people from the least even unto the greatest, came near,
\vs Jer 42:2 And said unto Jeremiah the prophet, Let, we beseech thee, our supplication be accepted before thee, and pray for us unto the \textsc{Lord} thy God, \bibemph{even} for all this remnant; (for we are left \bibemph{but} a few of many, as thine eyes do behold us:)
\fn{Let...: or, Let our supplication fall before thee.}
\vs Jer 42:3 That the \textsc{Lord} thy God may shew us the way wherein we may walk, and the thing that we may do.
\vs Jer 42:4 Then Jeremiah the prophet said unto them, I have heard \bibemph{you}; behold, I will pray unto the \textsc{Lord} your God according to your words; and it shall come to pass, \bibemph{that} whatsoever thing the \textsc{Lord} shall answer you, I will declare \bibemph{it} unto you; I will keep nothing back from you.
\vs Jer 42:5 Then they said to Jeremiah, The \textsc{Lord} be a true and faithful witness between us, if we do not even according to all things for the which the \textsc{Lord} thy God shall send thee to us.
\vs Jer 42:6 Whether \bibemph{it be} good, or whether \bibemph{it be} evil, we will obey the voice of the \textsc{Lord} our God, to whom we send thee; that it may be well with us, when we obey the voice of the \textsc{Lord} our God.
\vs Jer 42:7 And it came to pass after ten days, that the word of the \textsc{Lord} came unto Jeremiah.
\vs Jer 42:8 Then called he Johanan the son of Kareah, and all the captains of the forces which \bibemph{were} with him, and all the people from the least even to the greatest,
\vs Jer 42:9 And said unto them, Thus saith the \textsc{Lord}, the God of Israel, unto whom ye sent me to present your supplication before him;
\vs Jer 42:10 If ye will still abide in this land, then will I build you, and not pull \bibemph{you} down, and I will plant you, and not pluck \bibemph{you} up: for I repent me of the evil that I have done unto you.
\vs Jer 42:11 Be not afraid of the king of Babylon, of whom ye are afraid; be not afraid of him, saith the \textsc{Lord}: for I \bibemph{am} with you to save you, and to deliver you from his hand.
\vs Jer 42:12 And I will shew mercies unto you, that he may have mercy upon you, and cause you to return to your own land.
\vs Jer 42:13 But if ye say, We will not dwell in this land, neither obey the voice of the \textsc{Lord} your God,
\vs Jer 42:14 Saying, No; but we will go into the land of Egypt, where we shall see no war, nor hear the sound of the trumpet, nor have hunger of bread; and there will we dwell:
\vs Jer 42:15 And now therefore hear the word of the \textsc{Lord}, ye remnant of Judah; Thus saith the \textsc{Lord} of hosts, the God of Israel; If ye wholly set your faces to enter into Egypt, and go to sojourn there;
\vs Jer 42:16 Then it shall come to pass, \bibemph{that} the sword, which ye feared, shall overtake you there in the land of Egypt, and the famine, whereof ye were afraid, shall follow close after you there in Egypt; and there ye shall die.
\fn{shall follow...: Heb. shall cleave after you.}
\vs Jer 42:17 So shall it be with all the men that set their faces to go into Egypt to sojourn there; they shall die by the sword, by the famine, and by the pestilence: and none of them shall remain or escape from the evil that I will bring upon them.
\fn{So...: Heb. So shall all the men be.}
\vs Jer 42:18 For thus saith the \textsc{Lord} of hosts, the God of Israel; As mine anger and my fury hath been poured forth upon the inhabitants of Jerusalem; so shall my fury be poured forth upon you, when ye shall enter into Egypt: and ye shall be an execration, and an astonishment, and a curse, and a reproach; and ye shall see this place no more.
\vs Jer 42:19 The \textsc{Lord} hath said concerning you, O ye remnant of Judah; Go ye not into Egypt: know certainly that I have admonished you this day.
\fn{admonished...: Heb. testified against you.}
\vs Jer 42:20 For ye dissembled in your hearts, when ye sent me unto the \textsc{Lord} your God, saying, Pray for us unto the \textsc{Lord} our God; and according unto all that the \textsc{Lord} our God shall say, so declare unto us, and we will do \bibemph{it}.
\fn{ye dissembled...: or, ye have used deceit against your souls.}
\vs Jer 42:21 And \bibemph{now} I have this day declared \bibemph{it} to you; but ye have not obeyed the voice of the \textsc{Lord} your God, nor any \bibemph{thing} for the which he hath sent me unto you.
\vs Jer 42:22 Now therefore know certainly that ye shall die by the sword, by the famine, and by the pestilence, in the place whither ye desire to go \bibemph{and} to sojourn.
\fn{to go...: or, to go to sojourn.}
\vs Jer 43:1 And it came to pass, \bibemph{that} when Jeremiah had made an end of speaking unto all the people all the words of the \textsc{Lord} their God, for which the \textsc{Lord} their God had sent him to them, \bibemph{even} all these words,
\vs Jer 43:2 Then spake Azariah the son of Hoshaiah, and Johanan the son of Kareah, and all the proud men, saying unto Jeremiah, Thou speakest falsely: the \textsc{Lord} our God hath not sent thee to say, Go not into Egypt to sojourn there:
\vs Jer 43:3 But Baruch the son of Neriah setteth thee on against us, for to deliver us into the hand of the Chaldeans, that they might put us to death, and carry us away captives into Babylon.
\vs Jer 43:4 So Johanan the son of Kareah, and all the captains of the forces, and all the people, obeyed not the voice of the \textsc{Lord}, to dwell in the land of Judah.
\vs Jer 43:5 But Johanan the son of Kareah, and all the captains of the forces, took all the remnant of Judah, that were returned from all nations, whither they had been driven, to dwell in the land of Judah;
\vs Jer 43:6 \bibemph{Even} men, and women, and children, and the king's daughters, and every person that Nebuzaradan the captain of the guard had left with Gedaliah the son of Ahikam the son of Shaphan, and Jeremiah the prophet, and Baruch the son of Neriah.
\vs Jer 43:7 So they came into the land of Egypt: for they obeyed not the voice of the \textsc{Lord}: thus came they \bibemph{even} to Tahpanhes.
\vs Jer 43:8 Then came the word of the \textsc{Lord} unto Jeremiah in Tahpanhes, saying,
\vs Jer 43:9 Take great stones in thine hand, and hide them in the clay in the brickkiln, which \bibemph{is} at the entry of Pharaoh's house in Tahpanhes, in the sight of the men of Judah;
\vs Jer 43:10 And say unto them, Thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will send and take Nebuchadrezzar the king of Babylon, my servant, and will set his throne upon these stones that I have hid; and he shall spread his royal pavilion over them.
\vs Jer 43:11 And when he cometh, he shall smite the land of Egypt, \bibemph{and deliver} such \bibemph{as are} for death to death; and such \bibemph{as are} for captivity to captivity; and such \bibemph{as are} for the sword to the sword.
\vs Jer 43:12 And I will kindle a fire in the houses of the gods of Egypt; and he shall burn them, and carry them away captives: and he shall array himself with the land of Egypt, as a shepherd putteth on his garment; and he shall go forth from thence in peace.
\vs Jer 43:13 He shall break also the images of Bethshemesh, that \bibemph{is} in the land of Egypt; and the houses of the gods of the Egyptians shall he burn with fire.
\fn{images: Heb. statues, or, standing images. Bethshemesh: or, The house of the sun.}
\vs Jer 44:1 The word that came to Jeremiah concerning all the Jews which dwell in the land of Egypt, which dwell at Migdol, and at Tahpanhes, and at Noph, and in the country of Pathros, saying,
\vs Jer 44:2 Thus saith the \textsc{Lord} of hosts, the God of Israel; Ye have seen all the evil that I have brought upon Jerusalem, and upon all the cities of Judah; and, behold, this day they \bibemph{are} a desolation, and no man dwelleth therein,
\vs Jer 44:3 Because of their wickedness which they have committed to provoke me to anger, in that they went to burn incense, \bibemph{and} to serve other gods, whom they knew not, \bibemph{neither} they, ye, nor your fathers.
\vs Jer 44:4 Howbeit I sent unto you all my servants the prophets, rising early and sending \bibemph{them}, saying, Oh, do not this abominable thing that I hate.
\vs Jer 44:5 But they hearkened not, nor inclined their ear to turn from their wickedness, to burn no incense unto other gods.
\vs Jer 44:6 Wherefore my fury and mine anger was poured forth, and was kindled in the cities of Judah and in the streets of Jerusalem; and they are wasted \bibemph{and} desolate, as at this day.
\vs Jer 44:7 Therefore now thus saith the \textsc{Lord}, the God of hosts, the God of Israel; Wherefore commit ye \bibemph{this} great evil against your souls, to cut off from you man and woman, child and suckling, out of Judah, to leave you none to remain;
\fn{out...: Heb. out of the midst of Judah.}
\vs Jer 44:8 In that ye provoke me unto wrath with the works of your hands, burning incense unto other gods in the land of Egypt, whither ye be gone to dwell, that ye might cut yourselves off, and that ye might be a curse and a reproach among all the nations of the earth?
\vs Jer 44:9 Have ye forgotten the wickedness of your fathers, and the wickedness of the kings of Judah, and the wickedness of their wives, and your own wickedness, and the wickedness of your wives, which they have committed in the land of Judah, and in the streets of Jerusalem?
\fn{wickedness...: Heb. wickednesses, or, punishments, etc.}
\vs Jer 44:10 They are not humbled \bibemph{even} unto this day, neither have they feared, nor walked in my law, nor in my statutes, that I set before you and before your fathers.
\fn{humbled: Heb. contrite.}
\vs Jer 44:11 Therefore thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will set my face against you for evil, and to cut off all Judah.
\vs Jer 44:12 And I will take the remnant of Judah, that have set their faces to go into the land of Egypt to sojourn there, and they shall all be consumed, \bibemph{and} fall in the land of Egypt; they shall \bibemph{even} be consumed by the sword \bibemph{and} by the famine: they shall die, from the least even unto the greatest, by the sword and by the famine: and they shall be an execration, \bibemph{and} an astonishment, and a curse, and a reproach.
\vs Jer 44:13 For I will punish them that dwell in the land of Egypt, as I have punished Jerusalem, by the sword, by the famine, and by the pestilence:
\vs Jer 44:14 So that none of the remnant of Judah, which are gone into the land of Egypt to sojourn there, shall escape or remain, that they should return into the land of Judah, to the which they have a desire to return to dwell there: for none shall return but such as shall escape.
\fn{have...: Heb. lift up their soul.}
\vs Jer 44:15 Then all the men which knew that their wives had burned incense unto other gods, and all the women that stood by, a great multitude, even all the people that dwelt in the land of Egypt, in Pathros, answered Jeremiah, saying,
\vs Jer 44:16 \bibemph{As for} the word that thou hast spoken unto us in the name of the \textsc{Lord}, we will not hearken unto thee.
\vs Jer 44:17 But we will certainly do whatsoever thing goeth forth out of our own mouth, to burn incense unto the queen of heaven, and to pour out drink offerings unto her, as we have done, we, and our fathers, our kings, and our princes, in the cities of Judah, and in the streets of Jerusalem: for \bibemph{then} had we plenty of victuals, and were well, and saw no evil.
\fn{queen...: or, frame of heaven. victuals: Heb. bread.}
\vs Jer 44:18 But since we left off to burn incense to the queen of heaven, and to pour out drink offerings unto her, we have wanted all \bibemph{things}, and have been consumed by the sword and by the famine.
\vs Jer 44:19 And when we burned incense to the queen of heaven, and poured out drink offerings unto her, did we make her cakes to worship her, and pour out drink offerings unto her, without our men?
\fn{men: or, husbands?.}
\vs Jer 44:20 Then Jeremiah said unto all the people, to the men, and to the women, and to all the people which had given him \bibemph{that} answer, saying,
\vs Jer 44:21 The incense that ye burned in the cities of Judah, and in the streets of Jerusalem, ye, and your fathers, your kings, and your princes, and the people of the land, did not the \textsc{Lord} remember them, and came it \bibemph{not} into his mind?
\vs Jer 44:22 So that the \textsc{Lord} could no longer bear, because of the evil of your doings, \bibemph{and} because of the abominations which ye have committed; therefore is your land a desolation, and an astonishment, and a curse, without an inhabitant, as at this day.
\vs Jer 44:23 Because ye have burned incense, and because ye have sinned against the \textsc{Lord}, and have not obeyed the voice of the \textsc{Lord}, nor walked in his law, nor in his statutes, nor in his testimonies; therefore this evil is happened unto you, as at this day.
\vs Jer 44:24 Moreover Jeremiah said unto all the people, and to all the women, Hear the word of the \textsc{Lord}, all Judah that \bibemph{are} in the land of Egypt:
\vs Jer 44:25 Thus saith the \textsc{Lord} of hosts, the God of Israel, saying; Ye and your wives have both spoken with your mouths, and fulfilled with your hand, saying, We will surely perform our vows that we have vowed, to burn incense to the queen of heaven, and to pour out drink offerings unto her: ye will surely accomplish your vows, and surely perform your vows.
\vs Jer 44:26 Therefore hear ye the word of the \textsc{Lord}, all Judah that dwell in the land of Egypt; Behold, I have sworn by my great name, saith the \textsc{Lord}, that my name shall no more be named in the mouth of any man of Judah in all the land of Egypt, saying, The Lord GOD liveth.
\vs Jer 44:27 Behold, I will watch over them for evil, and not for good: and all the men of Judah that \bibemph{are} in the land of Egypt shall be consumed by the sword and by the famine, until there be an end of them.
\vs Jer 44:28 Yet a small number that escape the sword shall return out of the land of Egypt into the land of Judah, and all the remnant of Judah, that are gone into the land of Egypt to sojourn there, shall know whose words shall stand, mine, or theirs.
\fn{mine...: Heb. from me, or, from them.}
\vs Jer 44:29 And this \bibemph{shall be} a sign unto you, saith the \textsc{Lord}, that I will punish you in this place, that ye may know that my words shall surely stand against you for evil:
\vs Jer 44:30 Thus saith the \textsc{Lord}; Behold, I will give Pharaohhophra king of Egypt into the hand of his enemies, and into the hand of them that seek his life; as I gave Zedekiah king of Judah into the hand of Nebuchadrezzar king of Babylon, his enemy, and that sought his life.
\vs Jer 45:1 The word that Jeremiah the prophet spake unto Baruch the son of Neriah, when he had written these words in a book at the mouth of Jeremiah, in the fourth year of Jehoiakim the son of Josiah king of Judah, saying,
\vs Jer 45:2 Thus saith the \textsc{Lord}, the God of Israel, unto thee, O Baruch;
\vs Jer 45:3 Thou didst say, Woe is me now! for the \textsc{Lord} hath added grief to my sorrow; I fainted in my sighing, and I find no rest.
\vs Jer 45:4 Thus shalt thou say unto him, The \textsc{Lord} saith thus; Behold, \bibemph{that} which I have built will I break down, and that which I have planted I will pluck up, even this whole land.
\vs Jer 45:5 And seekest thou great things for thyself? seek \bibemph{them} not: for, behold, I will bring evil upon all flesh, saith the \textsc{Lord}: but thy life will I give unto thee for a prey in all places whither thou goest.
\vs Jer 46:1 The word of the \textsc{Lord} which came to Jeremiah the prophet against the Gentiles;
\vs Jer 46:2 Against Egypt, against the army of Pharaohnecho king of Egypt, which was by the river Euphrates in Carchemish, which Nebuchadrezzar king of Babylon smote in the fourth year of Jehoiakim the son of Josiah king of Judah.
\vs Jer 46:3 Order ye the buckler and shield, and draw near to battle.
\vs Jer 46:4 Harness the horses; and get up, ye horsemen, and stand forth with \bibemph{your} helmets; furbish the spears, \bibemph{and} put on the brigandines.
\vs Jer 46:5 Wherefore have I seen them dismayed \bibemph{and} turned away back? and their mighty ones are beaten down, and are fled apace, and look not back: \bibemph{for} fear \bibemph{was} round about, saith the \textsc{Lord}.
\fn{beaten...: Heb. broken in pieces. fled...: Heb. fled a flight.}
\vs Jer 46:6 Let not the swift flee away, nor the mighty man escape; they shall stumble, and fall toward the north by the river Euphrates.
\vs Jer 46:7 Who \bibemph{is} this \bibemph{that} cometh up as a flood, whose waters are moved as the rivers?
\vs Jer 46:8 Egypt riseth up like a flood, and \bibemph{his} waters are moved like the rivers; and he saith, I will go up, \bibemph{and} will cover the earth; I will destroy the city and the inhabitants thereof.
\vs Jer 46:9 Come up, ye horses; and rage, ye chariots; and let the mighty men come forth; the Ethiopians and the Libyans, that handle the shield; and the Lydians, that handle \bibemph{and} bend the bow.
\fn{the Ethiopians: Heb. Cush. the Libyans: Heb. Put.}
\vs Jer 46:10 For this \bibemph{is} the day of the Lord GOD of hosts, a day of vengeance, that he may avenge him of his adversaries: and the sword shall devour, and it shall be satiate and made drunk with their blood: for the Lord GOD of hosts hath a sacrifice in the north country by the river Euphrates.
\vs Jer 46:11 Go up into Gilead, and take balm, O virgin, the daughter of Egypt: in vain shalt thou use many medicines; \bibemph{for} thou shalt not be cured.
\fn{thou shalt...: Heb. no cure shall be unto thee.}
\vs Jer 46:12 The nations have heard of thy shame, and thy cry hath filled the land: for the mighty man hath stumbled against the mighty, \bibemph{and} they are fallen both together.
\vs Jer 46:13 The word that the \textsc{Lord} spake to Jeremiah the prophet, how Nebuchadrezzar king of Babylon should come \bibemph{and} smite the land of Egypt.
\vs Jer 46:14 Declare ye in Egypt, and publish in Migdol, and publish in Noph and in Tahpanhes: say ye, Stand fast, and prepare thee; for the sword shall devour round about thee.
\vs Jer 46:15 Why are thy valiant \bibemph{men} swept away? they stood not, because the \textsc{Lord} did drive them.
\vs Jer 46:16 He made many to fall, yea, one fell upon another: and they said, Arise, and let us go again to our own people, and to the land of our nativity, from the oppressing sword.
\fn{made...: Heb. multiplied the faller.}
\vs Jer 46:17 They did cry there, Pharaoh king of Egypt \bibemph{is but} a noise; he hath passed the time appointed.
\vs Jer 46:18 \bibemph{As} I live, saith the King, whose name \bibemph{is} the \textsc{Lord} of hosts, Surely as Tabor \bibemph{is} among the mountains, and as Carmel by the sea, \bibemph{so} shall he come.
\vs Jer 46:19 O thou daughter dwelling in Egypt, furnish thyself to go into captivity: for Noph shall be waste and desolate without an inhabitant.
\fn{furnish...: Heb. make thee instruments of captivity.}
\vs Jer 46:20 Egypt \bibemph{is like} a very fair heifer, \bibemph{but} destruction cometh; it cometh out of the north.
\vs Jer 46:21 Also her hired men \bibemph{are} in the midst of her like fatted bullocks; for they also are turned back, \bibemph{and} are fled away together: they did not stand, because the day of their calamity was come upon them, \bibemph{and} the time of their visitation.
\fn{fatted...: Heb. bullocks of the stall.}
\vs Jer 46:22 The voice thereof shall go like a serpent; for they shall march with an army, and come against her with axes, as hewers of wood.
\vs Jer 46:23 They shall cut down her forest, saith the \textsc{Lord}, though it cannot be searched; because they are more than the grasshoppers, and \bibemph{are} innumerable.
\vs Jer 46:24 The daughter of Egypt shall be confounded; she shall be delivered into the hand of the people of the north.
\vs Jer 46:25 The \textsc{Lord} of hosts, the God of Israel, saith; Behold, I will punish the multitude of No, and Pharaoh, and Egypt, with their gods, and their kings; even Pharaoh, and \bibemph{all} them that trust in him:
\fn{multitude: or, nourisher: Heb. Amon.}
\vs Jer 46:26 And I will deliver them into the hand of those that seek their lives, and into the hand of Nebuchadrezzar king of Babylon, and into the hand of his servants: and afterward it shall be inhabited, as in the days of old, saith the \textsc{Lord}.
\vs Jer 46:27 But fear not thou, O my servant Jacob, and be not dismayed, O Israel: for, behold, I will save thee from afar off, and thy seed from the land of their captivity; and Jacob shall return, and be in rest and at ease, and none shall make \bibemph{him} afraid.
\vs Jer 46:28 Fear thou not, O Jacob my servant, saith the \textsc{Lord}: for I \bibemph{am} with thee; for I will make a full end of all the nations whither I have driven thee: but I will not make a full end of thee, but correct thee in measure; yet will I not leave thee wholly unpunished.
\fn{not leave...: or, not utterly cut thee off.}
\vs Jer 47:1 The word of the \textsc{Lord} that came to Jeremiah the prophet against the Philistines, before that Pharaoh smote Gaza.
\fn{Gaza: Heb. Azzah.}
\vs Jer 47:2 Thus saith the \textsc{Lord}; Behold, waters rise up out of the north, and shall be an overflowing flood, and shall overflow the land, and all that is therein; the city, and them that dwell therein: then the men shall cry, and all the inhabitants of the land shall howl.
\fn{all that...: Heb. the fulness thereof.}
\vs Jer 47:3 At the noise of the stamping of the hoofs of his strong \bibemph{horses}, at the rushing of his chariots, \bibemph{and at} the rumbling of his wheels, the fathers shall not look back to \bibemph{their} children for feebleness of hands;
\vs Jer 47:4 Because of the day that cometh to spoil all the Philistines, \bibemph{and} to cut off from Tyrus and Zidon every helper that remaineth: for the \textsc{Lord} will spoil the Philistines, the remnant of the country of Caphtor.
\fn{the country: Heb. the isle.}
\vs Jer 47:5 Baldness is come upon Gaza; Ashkelon is cut off \bibemph{with} the remnant of their valley: how long wilt thou cut thyself?
\vs Jer 47:6 O thou sword of the \textsc{Lord}, how long \bibemph{will it be} ere thou be quiet? put up thyself into thy scabbard, rest, and be still.
\fn{put...: Heb. gather thyself.}
\vs Jer 47:7 How can it be quiet, seeing the \textsc{Lord} hath given it a charge against Ashkelon, and against the sea shore? there hath he appointed it.
\fn{How...: Heb. How canst thou.}
\vs Jer 48:1 Against Moab thus saith the \textsc{Lord} of hosts, the God of Israel; Woe unto Nebo! for it is spoiled: Kiriathaim is confounded \bibemph{and} taken: Misgab is confounded and dismayed.
\fn{Misgab: or, the high place.}
\vs Jer 48:2 \bibemph{There shall be} no more praise of Moab: in Heshbon they have devised evil against it; come, and let us cut it off from \bibemph{being} a nation. Also thou shalt be cut down, O Madmen; the sword shall pursue thee.
\fn{be cut...: or, be brought to silence. pursue...: Heb. go after thee.}
\vs Jer 48:3 A voice of crying \bibemph{shall be} from Horonaim, spoiling and great destruction.
\vs Jer 48:4 Moab is destroyed; her little ones have caused a cry to be heard.
\vs Jer 48:5 For in the going up of Luhith continual weeping shall go up; for in the going down of Horonaim the enemies have heard a cry of destruction.
\fn{continual...: Heb. weeping with weeping.}
\vs Jer 48:6 Flee, save your lives, and be like the heath in the wilderness.
\fn{the heath: or, a naked tree.}
\vs Jer 48:7 For because thou hast trusted in thy works and in thy treasures, thou shalt also be taken: and Chemosh shall go forth into captivity \bibemph{with} his priests and his princes together.
\vs Jer 48:8 And the spoiler shall come upon every city, and no city shall escape: the valley also shall perish, and the plain shall be destroyed, as the \textsc{Lord} hath spoken.
\vs Jer 48:9 Give wings unto Moab, that it may flee and get away: for the cities thereof shall be desolate, without any to dwell therein.
\vs Jer 48:10 Cursed \bibemph{be} he that doeth the work of the \textsc{Lord} deceitfully, and cursed \bibemph{be} he that keepeth back his sword from blood.
\fn{deceitfully: or, negligently.}
\vs Jer 48:11 Moab hath been at ease from his youth, and he hath settled on his lees, and hath not been emptied from vessel to vessel, neither hath he gone into captivity: therefore his taste remained in him, and his scent is not changed.
\fn{remained: Heb. stood.}
\vs Jer 48:12 Therefore, behold, the days come, saith the \textsc{Lord}, that I will send unto him wanderers, that shall cause him to wander, and shall empty his vessels, and break their bottles.
\vs Jer 48:13 And Moab shall be ashamed of Chemosh, as the house of Israel was ashamed of Bethel their confidence.
\vs Jer 48:14 How say ye, We \bibemph{are} mighty and strong men for the war?
\vs Jer 48:15 Moab is spoiled, and gone up \bibemph{out of} her cities, and his chosen young men are gone down to the slaughter, saith the King, whose name \bibemph{is} the \textsc{Lord} of hosts.
\fn{his...: Heb. the choice of.}
\vs Jer 48:16 The calamity of Moab \bibemph{is} near to come, and his affliction hasteth fast.
\vs Jer 48:17 All ye that are about him, bemoan him; and all ye that know his name, say, How is the strong staff broken, \bibemph{and} the beautiful rod!
\vs Jer 48:18 Thou daughter that dost inhabit Dibon, come down from \bibemph{thy} glory, and sit in thirst; for the spoiler of Moab shall come upon thee, \bibemph{and} he shall destroy thy strong holds.
\vs Jer 48:19 O inhabitant of Aroer, stand by the way, and espy; ask him that fleeth, and her that escapeth, \bibemph{and} say, What is done?
\fn{inhabitant: Heb. inhabitress.}
\vs Jer 48:20 Moab is confounded; for it is broken down: howl and cry; tell ye it in Arnon, that Moab is spoiled,
\vs Jer 48:21 And judgment is come upon the plain country; upon Holon, and upon Jahazah, and upon Mephaath,
\vs Jer 48:22 And upon Dibon, and upon Nebo, and upon Bethdiblathaim,
\vs Jer 48:23 And upon Kiriathaim, and upon Bethgamul, and upon Bethmeon,
\vs Jer 48:24 And upon Kerioth, and upon Bozrah, and upon all the cities of the land of Moab, far or near.
\vs Jer 48:25 The horn of Moab is cut off, and his arm is broken, saith the \textsc{Lord}.
\vs Jer 48:26 Make ye him drunken: for he magnified \bibemph{himself} against the \textsc{Lord}: Moab also shall wallow in his vomit, and he also shall be in derision.
\vs Jer 48:27 For was not Israel a derision unto thee? was he found among thieves? for since thou spakest of him, thou skippedst for joy.
\fn{skippedst...: or, movedst thyself.}
\vs Jer 48:28 O ye that dwell in Moab, leave the cities, and dwell in the rock, and be like the dove \bibemph{that} maketh her nest in the sides of the hole's mouth.
\vs Jer 48:29 We have heard the pride of Moab, (he is exceeding proud) his loftiness, and his arrogancy, and his pride, and the haughtiness of his heart.
\vs Jer 48:30 I know his wrath, saith the \textsc{Lord}; but \bibemph{it shall} not \bibemph{be} so; his lies shall not so effect \bibemph{it}.
\fn{his lies...: or, those on whom he stayeth (Heb. his bars) do not right.}
\vs Jer 48:31 Therefore will I howl for Moab, and I will cry out for all Moab; \bibemph{mine heart} shall mourn for the men of Kirheres.
\vs Jer 48:32 O vine of Sibmah, I will weep for thee with the weeping of Jazer: thy plants are gone over the sea, they reach \bibemph{even} to the sea of Jazer: the spoiler is fallen upon thy summer fruits and upon thy vintage.
\vs Jer 48:33 And joy and gladness is taken from the plentiful field, and from the land of Moab; and I have caused wine to fail from the winepresses: none shall tread with shouting; \bibemph{their} shouting \bibemph{shall be} no shouting.
\vs Jer 48:34 From the cry of Heshbon \bibemph{even} unto Elealeh, \bibemph{and even} unto Jahaz, have they uttered their voice, from Zoar \bibemph{even} unto Horonaim, \bibemph{as} an heifer of three years old: for the waters also of Nimrim shall be desolate.
\fn{desolate: Heb. desolations.}
\vs Jer 48:35 Moreover I will cause to cease in Moab, saith the \textsc{Lord}, him that offereth in the high places, and him that burneth incense to his gods.
\vs Jer 48:36 Therefore mine heart shall sound for Moab like pipes, and mine heart shall sound like pipes for the men of Kirheres: because the riches \bibemph{that} he hath gotten are perished.
\vs Jer 48:37 For every head \bibemph{shall be} bald, and every beard clipped: upon all the hands \bibemph{shall be} cuttings, and upon the loins sackcloth.
\fn{clipped: Heb. diminished.}
\vs Jer 48:38 \bibemph{There shall be} lamentation generally upon all the housetops of Moab, and in the streets thereof: for I have broken Moab like a vessel wherein \bibemph{is} no pleasure, saith the \textsc{Lord}.
\vs Jer 48:39 They shall howl, \bibemph{saying}, How is it broken down! how hath Moab turned the back with shame! so shall Moab be a derision and a dismaying to all them about him.
\fn{back: Heb. neck.}
\vs Jer 48:40 For thus saith the \textsc{Lord}; Behold, he shall fly as an eagle, and shall spread his wings over Moab.
\vs Jer 48:41 Kerioth is taken, and the strong holds are surprised, and the mighty men's hearts in Moab at that day shall be as the heart of a woman in her pangs.
\fn{Kerioth: or, The cities.}
\vs Jer 48:42 And Moab shall be destroyed from \bibemph{being} a people, because he hath magnified \bibemph{himself} against the \textsc{Lord}.
\vs Jer 48:43 Fear, and the pit, and the snare, \bibemph{shall be} upon thee, O inhabitant of Moab, saith the \textsc{Lord}.
\vs Jer 48:44 He that fleeth from the fear shall fall into the pit; and he that getteth up out of the pit shall be taken in the snare: for I will bring upon it, \bibemph{even} upon Moab, the year of their visitation, saith the \textsc{Lord}.
\vs Jer 48:45 They that fled stood under the shadow of Heshbon because of the force: but a fire shall come forth out of Heshbon, and a flame from the midst of Sihon, and shall devour the corner of Moab, and the crown of the head of the tumultuous ones.
\fn{tumultuous...: Heb. children of noise.}
\vs Jer 48:46 Woe be unto thee, O Moab! the people of Chemosh perisheth: for thy sons are taken captives, and thy daughters captives.
\fn{captives, and: Heb. in captivity, etc.}
\vs Jer 48:47 Yet will I bring again the captivity of Moab in the latter days, saith the \textsc{Lord}. Thus far \bibemph{is} the judgment of Moab.
\vs Jer 49:1 Concerning the Ammonites, thus saith the \textsc{Lord}; Hath Israel no sons? hath he no heir? why \bibemph{then} doth their king inherit Gad, and his people dwell in his cities?
\fn{Concerning: or, Against. their king: or, Melcom.}
\vs Jer 49:2 Therefore, behold, the days come, saith the \textsc{Lord}, that I will cause an alarm of war to be heard in Rabbah of the Ammonites; and it shall be a desolate heap, and her daughters shall be burned with fire: then shall Israel be heir unto them that were his heirs, saith the \textsc{Lord}.
\vs Jer 49:3 Howl, O Heshbon, for Ai is spoiled: cry, ye daughters of Rabbah, gird you with sackcloth; lament, and run to and fro by the hedges; for their king shall go into captivity, \bibemph{and} his priests and his princes together.
\fn{their king: or, Melcom.}
\vs Jer 49:4 Wherefore gloriest thou in the valleys, thy flowing valley, O backsliding daughter? that trusted in her treasures, \bibemph{saying}, Who shall come unto me?
\fn{thy...: or, thy valley floweth away.}
\vs Jer 49:5 Behold, I will bring a fear upon thee, saith the Lord GOD of hosts, from all those that be about thee; and ye shall be driven out every man right forth; and none shall gather up him that wandereth.
\vs Jer 49:6 And afterward I will bring again the captivity of the children of Ammon, saith the \textsc{Lord}.
\vs Jer 49:7 Concerning Edom, thus saith the \textsc{Lord} of hosts; \bibemph{Is} wisdom no more in Teman? is counsel perished from the prudent? is their wisdom vanished?
\vs Jer 49:8 Flee ye, turn back, dwell deep, O inhabitants of Dedan; for I will bring the calamity of Esau upon him, the time \bibemph{that} I will visit him.
\fn{turn...: or, they are turned back.}
\vs Jer 49:9 If grapegatherers come to thee, would they not leave \bibemph{some} gleaning grapes? if thieves by night, they will destroy till they have enough.
\fn{till...: Heb. their sufficiency.}
\vs Jer 49:10 But I have made Esau bare, I have uncovered his secret places, and he shall not be able to hide himself: his seed is spoiled, and his brethren, and his neighbours, and he \bibemph{is} not.
\vs Jer 49:11 Leave thy fatherless children, I will preserve \bibemph{them} alive; and let thy widows trust in me.
\vs Jer 49:12 For thus saith the \textsc{Lord}; Behold, they whose judgment \bibemph{was} not to drink of the cup have assuredly drunken; and \bibemph{art} thou he \bibemph{that} shall altogether go unpunished? thou shalt not go unpunished, but thou shalt surely drink \bibemph{of it}.
\vs Jer 49:13 For I have sworn by myself, saith the \textsc{Lord}, that Bozrah shall become a desolation, a reproach, a waste, and a curse; and all the cities thereof shall be perpetual wastes.
\vs Jer 49:14 I have heard a rumour from the \textsc{Lord}, and an ambassador is sent unto the heathen, \bibemph{saying}, Gather ye together, and come against her, and rise up to the battle.
\vs Jer 49:15 For, lo, I will make thee small among the heathen, \bibemph{and} despised among men.
\vs Jer 49:16 Thy terribleness hath deceived thee, \bibemph{and} the pride of thine heart, O thou that dwellest in the clefts of the rock, that holdest the height of the hill: though thou shouldest make thy nest as high as the eagle, I will bring thee down from thence, saith the \textsc{Lord}.
\vs Jer 49:17 Also Edom shall be a desolation: every one that goeth by it shall be astonished, and shall hiss at all the plagues thereof.
\vs Jer 49:18 As in the overthrow of Sodom and Gomorrah and the neighbour \bibemph{cities} thereof, saith the \textsc{Lord}, no man shall abide there, neither shall a son of man dwell in it.
\vs Jer 49:19 Behold, he shall come up like a lion from the swelling of Jordan against the habitation of the strong: but I will suddenly make him run away from her: and who \bibemph{is} a chosen \bibemph{man, that} I may appoint over her? for who \bibemph{is} like me? and who will appoint me the time? and who \bibemph{is} that shepherd that will stand before me?
\fn{appoint me...: or, convent me in judgment?.}
\vs Jer 49:20 Therefore hear the counsel of the \textsc{Lord}, that he hath taken against Edom; and his purposes, that he hath purposed against the inhabitants of Teman: Surely the least of the flock shall draw them out: surely he shall make their habitations desolate with them.
\vs Jer 49:21 The earth is moved at the noise of their fall, at the cry the noise thereof was heard in the Red sea.
\fn{Red sea: Heb. Weedy sea.}
\vs Jer 49:22 Behold, he shall come up and fly as the eagle, and spread his wings over Bozrah: and at that day shall the heart of the mighty men of Edom be as the heart of a woman in her pangs.
\vs Jer 49:23 Concerning Damascus. Hamath is confounded, and Arpad: for they have heard evil tidings: they are fainthearted; \bibemph{there is} sorrow on the sea; it cannot be quiet.
\fn{fainthearted: Heb. melted. on...: or, as on the sea.}
\vs Jer 49:24 Damascus is waxed feeble, \bibemph{and} turneth herself to flee, and fear hath seized on \bibemph{her}: anguish and sorrows have taken her, as a woman in travail.
\vs Jer 49:25 How is the city of praise not left, the city of my joy!
\vs Jer 49:26 Therefore her young men shall fall in her streets, and all the men of war shall be cut off in that day, saith the \textsc{Lord} of hosts.
\vs Jer 49:27 And I will kindle a fire in the wall of Damascus, and it shall consume the palaces of Benhadad.
\vs Jer 49:28 Concerning Kedar, and concerning the kingdoms of Hazor, which Nebuchadrezzar king of Babylon shall smite, thus saith the \textsc{Lord}; Arise ye, go up to Kedar, and spoil the men of the east.
\vs Jer 49:29 Their tents and their flocks shall they take away: they shall take to themselves their curtains, and all their vessels, and their camels; and they shall cry unto them, Fear \bibemph{is} on every side.
\vs Jer 49:30 Flee, get you far off, dwell deep, O ye inhabitants of Hazor, saith the \textsc{Lord}; for Nebuchadrezzar king of Babylon hath taken counsel against you, and hath conceived a purpose against you.
\fn{get...: Heb. flit greatly.}
\vs Jer 49:31 Arise, get you up unto the wealthy nation, that dwelleth without care, saith the \textsc{Lord}, which have neither gates nor bars, \bibemph{which} dwell alone.
\fn{wealthy: or, that is at ease.}
\vs Jer 49:32 And their camels shall be a booty, and the multitude of their cattle a spoil: and I will scatter into all winds them \bibemph{that are} in the utmost corners; and I will bring their calamity from all sides thereof, saith the \textsc{Lord}.
\fn{in...: Heb. cut off into corners, or, that have the corners of their hair polled.}
\vs Jer 49:33 And Hazor shall be a dwelling for dragons, \bibemph{and} a desolation for ever: there shall no man abide there, nor \bibemph{any} son of man dwell in it.
\vs Jer 49:34 The word of the \textsc{Lord} that came to Jeremiah the prophet against Elam in the beginning of the reign of Zedekiah king of Judah, saying,
\vs Jer 49:35 Thus saith the \textsc{Lord} of hosts; Behold, I will break the bow of Elam, the chief of their might.
\vs Jer 49:36 And upon Elam will I bring the four winds from the four quarters of heaven, and will scatter them toward all those winds; and there shall be no nation whither the outcasts of Elam shall not come.
\vs Jer 49:37 For I will cause Elam to be dismayed before their enemies, and before them that seek their life: and I will bring evil upon them, \bibemph{even} my fierce anger, saith the \textsc{Lord}; and I will send the sword after them, till I have consumed them:
\vs Jer 49:38 And I will set my throne in Elam, and will destroy from thence the king and the princes, saith the \textsc{Lord}.
\vs Jer 49:39 But it shall come to pass in the latter days, \bibemph{that} I will bring again the captivity of Elam, saith the \textsc{Lord}.
\vs Jer 50:1 The word that the \textsc{Lord} spake against Babylon \bibemph{and} against the land of the Chaldeans by Jeremiah the prophet.
\fn{by...: Heb. by the hand of Jeremiah.}
\vs Jer 50:2 Declare ye among the nations, and publish, and set up a standard; publish, \bibemph{and} conceal not: say, Babylon is taken, Bel is confounded, Merodach is broken in pieces; her idols are confounded, her images are broken in pieces.
\fn{set up: Heb. lift up.}
\vs Jer 50:3 For out of the north there cometh up a nation against her, which shall make her land desolate, and none shall dwell therein: they shall remove, they shall depart, both man and beast.
\vs Jer 50:4 In those days, and in that time, saith the \textsc{Lord}, the children of Israel shall come, they and the children of Judah together, going and weeping: they shall go, and seek the \textsc{Lord} their God.
\vs Jer 50:5 They shall ask the way to Zion with their faces thitherward, \bibemph{saying}, Come, and let us join ourselves to the \textsc{Lord} in a perpetual covenant \bibemph{that} shall not be forgotten.
\vs Jer 50:6 My people hath been lost sheep: their shepherds have caused them to go astray, they have turned them away \bibemph{on} the mountains: they have gone from mountain to hill, they have forgotten their restingplace.
\fn{restingplace: Heb. place to lie down in.}
\vs Jer 50:7 All that found them have devoured them: and their adversaries said, We offend not, because they have sinned against the \textsc{Lord}, the habitation of justice, even the \textsc{Lord}, the hope of their fathers.
\vs Jer 50:8 Remove out of the midst of Babylon, and go forth out of the land of the Chaldeans, and be as the he goats before the flocks.
\vs Jer 50:9 For, lo, I will raise and cause to come up against Babylon an assembly of great nations from the north country: and they shall set themselves in array against her; from thence she shall be taken: their arrows \bibemph{shall be} as of a mighty expert man; none shall return in vain.
\fn{expert...: or, destroyer.}
\vs Jer 50:10 And Chaldea shall be a spoil: all that spoil her shall be satisfied, saith the \textsc{Lord}.
\vs Jer 50:11 Because ye were glad, because ye rejoiced, O ye destroyers of mine heritage, because ye are grown fat as the heifer at grass, and bellow as bulls;
\fn{fat: Heb. big, or, corpulent. bellow...: or, neigh as steeds.}
\vs Jer 50:12 Your mother shall be sore confounded; she that bare you shall be ashamed: behold, the hindermost of the nations \bibemph{shall be} a wilderness, a dry land, and a desert.
\vs Jer 50:13 Because of the wrath of the \textsc{Lord} it shall not be inhabited, but it shall be wholly desolate: every one that goeth by Babylon shall be astonished, and hiss at all her plagues.
\vs Jer 50:14 Put yourselves in array against Babylon round about: all ye that bend the bow, shoot at her, spare no arrows: for she hath sinned against the \textsc{Lord}.
\vs Jer 50:15 Shout against her round about: she hath given her hand: her foundations are fallen, her walls are thrown down: for it \bibemph{is} the vengeance of the \textsc{Lord}: take vengeance upon her; as she hath done, do unto her.
\vs Jer 50:16 Cut off the sower from Babylon, and him that handleth the sickle in the time of harvest: for fear of the oppressing sword they shall turn every one to his people, and they shall flee every one to his own land.
\fn{sickle; or, scythe.}
\vs Jer 50:17 Israel \bibemph{is} a scattered sheep; the lions have driven \bibemph{him} away: first the king of Assyria hath devoured him; and last this Nebuchadrezzar king of Babylon hath broken his bones.
\vs Jer 50:18 Therefore thus saith the \textsc{Lord} of hosts, the God of Israel; Behold, I will punish the king of Babylon and his land, as I have punished the king of Assyria.
\vs Jer 50:19 And I will bring Israel again to his habitation, and he shall feed on Carmel and Bashan, and his soul shall be satisfied upon mount Ephraim and Gilead.
\vs Jer 50:20 In those days, and in that time, saith the \textsc{Lord}, the iniquity of Israel shall be sought for, and \bibemph{there shall be} none; and the sins of Judah, and they shall not be found: for I will pardon them whom I reserve.
\vs Jer 50:21 Go up against the land of Merathaim, \bibemph{even} against it, and against the inhabitants of Pekod: waste and utterly destroy after them, saith the \textsc{Lord}, and do according to all that I have commanded thee.
\fn{of Merathaim: or, of the rebels. Pekod: or, Visitation.}
\vs Jer 50:22 A sound of battle \bibemph{is} in the land, and of great destruction.
\vs Jer 50:23 How is the hammer of the whole earth cut asunder and broken! how is Babylon become a desolation among the nations!
\vs Jer 50:24 I have laid a snare for thee, and thou art also taken, O Babylon, and thou wast not aware: thou art found, and also caught, because thou hast striven against the \textsc{Lord}.
\vs Jer 50:25 The \textsc{Lord} hath opened his armoury, and hath brought forth the weapons of his indignation: for this \bibemph{is} the work of the Lord GOD of hosts in the land of the Chaldeans.
\vs Jer 50:26 Come against her from the utmost border, open her storehouses: cast her up as heaps, and destroy her utterly: let nothing of her be left.
\fn{from...: Heb. from the end. cast...: or, tread her.}
\vs Jer 50:27 Slay all her bullocks; let them go down to the slaughter: woe unto them! for their day is come, the time of their visitation.
\vs Jer 50:28 The voice of them that flee and escape out of the land of Babylon, to declare in Zion the vengeance of the \textsc{Lord} our God, the vengeance of his temple.
\vs Jer 50:29 Call together the archers against Babylon: all ye that bend the bow, camp against it round about; let none thereof escape: recompense her according to her work; according to all that she hath done, do unto her: for she hath been proud against the \textsc{Lord}, against the Holy One of Israel.
\vs Jer 50:30 Therefore shall her young men fall in the streets, and all her men of war shall be cut off in that day, saith the \textsc{Lord}.
\vs Jer 50:31 Behold, I \bibemph{am} against thee, \bibemph{O thou} most proud, saith the Lord GOD of hosts: for thy day is come, the time \bibemph{that} I will visit thee.
\fn{most...: Heb. pride.}
\vs Jer 50:32 And the most proud shall stumble and fall, and none shall raise him up: and I will kindle a fire in his cities, and it shall devour all round about him.
\fn{the...: Heb. pride.}
\vs Jer 50:33 Thus saith the \textsc{Lord} of hosts; The children of Israel and the children of Judah \bibemph{were} oppressed together: and all that took them captives held them fast; they refused to let them go.
\vs Jer 50:34 Their Redeemer \bibemph{is} strong; the \textsc{Lord} of hosts \bibemph{is} his name: he shall throughly plead their cause, that he may give rest to the land, and disquiet the inhabitants of Babylon.
\vs Jer 50:35 A sword \bibemph{is} upon the Chaldeans, saith the \textsc{Lord}, and upon the inhabitants of Babylon, and upon her princes, and upon her wise \bibemph{men}.
\vs Jer 50:36 A sword \bibemph{is} upon the liars; and they shall dote: a sword \bibemph{is} upon her mighty men; and they shall be dismayed.
\fn{liars: or, chief stays: Heb. bars.}
\vs Jer 50:37 A sword \bibemph{is} upon their horses, and upon their chariots, and upon all the mingled people that \bibemph{are} in the midst of her; and they shall become as women: a sword \bibemph{is} upon her treasures; and they shall be robbed.
\vs Jer 50:38 A drought \bibemph{is} upon her waters; and they shall be dried up: for it \bibemph{is} the land of graven images, and they are mad upon \bibemph{their} idols.
\vs Jer 50:39 Therefore the wild beasts of the desert with the wild beasts of the islands shall dwell \bibemph{there}, and the owls shall dwell therein: and it shall be no more inhabited for ever; neither shall it be dwelt in from generation to generation.
\vs Jer 50:40 As God overthrew Sodom and Gomorrah and the neighbour \bibemph{cities} thereof, saith the \textsc{Lord}; \bibemph{so} shall no man abide there, neither shall any son of man dwell therein.
\vs Jer 50:41 Behold, a people shall come from the north, and a great nation, and many kings shall be raised up from the coasts of the earth.
\vs Jer 50:42 They shall hold the bow and the lance: they \bibemph{are} cruel, and will not shew mercy: their voice shall roar like the sea, and they shall ride upon horses, \bibemph{every one} put in array, like a man to the battle, against thee, O daughter of Babylon.
\vs Jer 50:43 The king of Babylon hath heard the report of them, and his hands waxed feeble: anguish took hold of him, \bibemph{and} pangs as of a woman in travail.
\vs Jer 50:44 Behold, he shall come up like a lion from the swelling of Jordan unto the habitation of the strong: but I will make them suddenly run away from her: and who \bibemph{is} a chosen \bibemph{man, that} I may appoint over her? for who \bibemph{is} like me? and who will appoint me the time? and who \bibemph{is} that shepherd that will stand before me?
\fn{appoint me...: or, convent me to plead?.}
\vs Jer 50:45 Therefore hear ye the counsel of the \textsc{Lord}, that he hath taken against Babylon; and his purposes, that he hath purposed against the land of the Chaldeans: Surely the least of the flock shall draw them out: surely he shall make \bibemph{their} habitation desolate with them.
\vs Jer 50:46 At the noise of the taking of Babylon the earth is moved, and the cry is heard among the nations.
\vs Jer 51:1 Thus saith the \textsc{Lord}; Behold, I will raise up against Babylon, and against them that dwell in the midst of them that rise up against me, a destroying wind;
\fn{midst: Heb. heart.}
\vs Jer 51:2 And will send unto Babylon fanners, that shall fan her, and shall empty her land: for in the day of trouble they shall be against her round about.
\vs Jer 51:3 Against \bibemph{him that} bendeth let the archer bend his bow, and against \bibemph{him that} lifteth himself up in his brigandine: and spare ye not her young men; destroy ye utterly all her host.
\vs Jer 51:4 Thus the slain shall fall in the land of the Chaldeans, and \bibemph{they that are} thrust through in her streets.
\vs Jer 51:5 For Israel \bibemph{hath} not \bibemph{been} forsaken, nor Judah of his God, of the \textsc{Lord} of hosts; though their land was filled with sin against the Holy One of Israel.
\vs Jer 51:6 Flee out of the midst of Babylon, and deliver every man his soul: be not cut off in her iniquity; for this \bibemph{is} the time of the \textsc{Lord's} vengeance; he will render unto her a recompence.
\vs Jer 51:7 Babylon \bibemph{hath been} a golden cup in the \textsc{Lord's} hand, that made all the earth drunken: the nations have drunken of her wine; therefore the nations are mad.
\vs Jer 51:8 Babylon is suddenly fallen and destroyed: howl for her; take balm for her pain, if so be she may be healed.
\vs Jer 51:9 We would have healed Babylon, but she is not healed: forsake her, and let us go every one into his own country: for her judgment reacheth unto heaven, and is lifted up \bibemph{even} to the skies.
\vs Jer 51:10 The \textsc{Lord} hath brought forth our righteousness: come, and let us declare in Zion the work of the \textsc{Lord} our God.
\vs Jer 51:11 Make bright the arrows; gather the shields: the \textsc{Lord} hath raised up the spirit of the kings of the Medes: for his device \bibemph{is} against Babylon, to destroy it; because it \bibemph{is} the vengeance of the \textsc{Lord}, the vengeance of his temple.
\fn{bright: Heb. pure.}
\vs Jer 51:12 Set up the standard upon the walls of Babylon, make the watch strong, set up the watchmen, prepare the ambushes: for the \textsc{Lord} hath both devised and done that which he spake against the inhabitants of Babylon.
\fn{ambushes: Heb. liers in wait.}
\vs Jer 51:13 O thou that dwellest upon many waters, abundant in treasures, thine end is come, \bibemph{and} the measure of thy covetousness.
\vs Jer 51:14 The \textsc{Lord} of hosts hath sworn by himself, \bibemph{saying}, Surely I will fill thee with men, as with caterpillers; and they shall lift up a shout against thee.
\fn{by himself: Heb. by his soul. lift up: Heb. utter.}
\vs Jer 51:15 He hath made the earth by his power, he hath established the world by his wisdom, and hath stretched out the heaven by his understanding.
\vs Jer 51:16 When he uttereth \bibemph{his} voice, \bibemph{there is} a multitude of waters in the heavens; and he causeth the vapours to ascend from the ends of the earth: he maketh lightnings with rain, and bringeth forth the wind out of his treasures.
\fn{multitude: or, noise.}
\vs Jer 51:17 Every man is brutish by \bibemph{his} knowledge; every founder is confounded by the graven image: for his molten image \bibemph{is} falsehood, and \bibemph{there is} no breath in them.
\fn{is brutish...: or, is more brutish than to know.}
\vs Jer 51:18 They \bibemph{are} vanity, the work of errors: in the time of their visitation they shall perish.
\vs Jer 51:19 The portion of Jacob \bibemph{is} not like them; for he \bibemph{is} the former of all things: and \bibemph{Israel is} the rod of his inheritance: the \textsc{Lord} of hosts \bibemph{is} his name.
\vs Jer 51:20 Thou \bibemph{art} my battle axe \bibemph{and} weapons of war: for with thee will I break in pieces the nations, and with thee will I destroy kingdoms;
\fn{with thee: or, in thee, or, by thee.}
\vs Jer 51:21 And with thee will I break in pieces the horse and his rider; and with thee will I break in pieces the chariot and his rider;
\vs Jer 51:22 With thee also will I break in pieces man and woman; and with thee will I break in pieces old and young; and with thee will I break in pieces the young man and the maid;
\vs Jer 51:23 I will also break in pieces with thee the shepherd and his flock; and with thee will I break in pieces the husbandman and his yoke of oxen; and with thee will I break in pieces captains and rulers.
\vs Jer 51:24 And I will render unto Babylon and to all the inhabitants of Chaldea all their evil that they have done in Zion in your sight, saith the \textsc{Lord}.
\vs Jer 51:25 Behold, I \bibemph{am} against thee, O destroying mountain, saith the \textsc{Lord}, which destroyest all the earth: and I will stretch out mine hand upon thee, and roll thee down from the rocks, and will make thee a burnt mountain.
\vs Jer 51:26 And they shall not take of thee a stone for a corner, nor a stone for foundations; but thou shalt be desolate for ever, saith the \textsc{Lord}.
\fn{desolate...: Heb. everlasting desolations.}
\vs Jer 51:27 Set ye up a standard in the land, blow the trumpet among the nations, prepare the nations against her, call together against her the kingdoms of Ararat, Minni, and Ashchenaz; appoint a captain against her; cause the horses to come up as the rough caterpillers.
\vs Jer 51:28 Prepare against her the nations with the kings of the Medes, the captains thereof, and all the rulers thereof, and all the land of his dominion.
\vs Jer 51:29 And the land shall tremble and sorrow: for every purpose of the \textsc{Lord} shall be performed against Babylon, to make the land of Babylon a desolation without an inhabitant.
\vs Jer 51:30 The mighty men of Babylon have forborn to fight, they have remained in \bibemph{their} holds: their might hath failed; they became as women: they have burned her dwellingplaces; her bars are broken.
\vs Jer 51:31 One post shall run to meet another, and one messenger to meet another, to shew the king of Babylon that his city is taken at \bibemph{one} end,
\vs Jer 51:32 And that the passages are stopped, and the reeds they have burned with fire, and the men of war are affrighted.
\vs Jer 51:33 For thus saith the \textsc{Lord} of hosts, the God of Israel; The daughter of Babylon \bibemph{is} like a threshingfloor, \bibemph{it is} time to thresh her: yet a little while, and the time of her harvest shall come.
\fn{it is...: or, in the time that he thresheth her.}
\vs Jer 51:34 Nebuchadrezzar the king of Babylon hath devoured me, he hath crushed me, he hath made me an empty vessel, he hath swallowed me up like a dragon, he hath filled his belly with my delicates, he hath cast me out.
\vs Jer 51:35 The violence done to me and to my flesh \bibemph{be} upon Babylon, shall the inhabitant of Zion say; and my blood upon the inhabitants of Chaldea, shall Jerusalem say.
\fn{The violence...: Heb. My violence. flesh: or, remainder. inhabitant: Heb. inhabitress.}
\vs Jer 51:36 Therefore thus saith the \textsc{Lord}; Behold, I will plead thy cause, and take vengeance for thee; and I will dry up her sea, and make her springs dry.
\vs Jer 51:37 And Babylon shall become heaps, a dwellingplace for dragons, an astonishment, and an hissing, without an inhabitant.
\vs Jer 51:38 They shall roar together like lions: they shall yell as lions' whelps.
\fn{yell: or, shake themselves.}
\vs Jer 51:39 In their heat I will make their feasts, and I will make them drunken, that they may rejoice, and sleep a perpetual sleep, and not wake, saith the \textsc{Lord}.
\vs Jer 51:40 I will bring them down like lambs to the slaughter, like rams with he goats.
\vs Jer 51:41 How is Sheshach taken! and how is the praise of the whole earth surprised! how is Babylon become an astonishment among the nations!
\vs Jer 51:42 The sea is come up upon Babylon: she is covered with the multitude of the waves thereof.
\vs Jer 51:43 Her cities are a desolation, a dry land, and a wilderness, a land wherein no man dwelleth, neither doth \bibemph{any} son of man pass thereby.
\vs Jer 51:44 And I will punish Bel in Babylon, and I will bring forth out of his mouth that which he hath swallowed up: and the nations shall not flow together any more unto him: yea, the wall of Babylon shall fall.
\vs Jer 51:45 My people, go ye out of the midst of her, and deliver ye every man his soul from the fierce anger of the \textsc{Lord}.
\vs Jer 51:46 And lest your heart faint, and ye fear for the rumour that shall be heard in the land; a rumour shall both come \bibemph{one} year, and after that in \bibemph{another} year \bibemph{shall come} a rumour, and violence in the land, ruler against ruler.
\fn{lest: or, let not.}
\vs Jer 51:47 Therefore, behold, the days come, that I will do judgment upon the graven images of Babylon: and her whole land shall be confounded, and all her slain shall fall in the midst of her.
\fn{do...: Heb. visit upon.}
\vs Jer 51:48 Then the heaven and the earth, and all that \bibemph{is} therein, shall sing for Babylon: for the spoilers shall come unto her from the north, saith the \textsc{Lord}.
\vs Jer 51:49 As Babylon \bibemph{hath caused} the slain of Israel to fall, so at Babylon shall fall the slain of all the earth.
\fn{As...: or, Both Babylon is to fall, O ye slain of Israel, and with Babylon, etc. the earth: or, the country.}
\vs Jer 51:50 Ye that have escaped the sword, go away, stand not still: remember the \textsc{Lord} afar off, and let Jerusalem come into your mind.
\vs Jer 51:51 We are confounded, because we have heard reproach: shame hath covered our faces: for strangers are come into the sanctuaries of the \textsc{Lord's} house.
\vs Jer 51:52 Wherefore, behold, the days come, saith the \textsc{Lord}, that I will do judgment upon her graven images: and through all her land the wounded shall groan.
\vs Jer 51:53 Though Babylon should mount up to heaven, and though she should fortify the height of her strength, \bibemph{yet} from me shall spoilers come unto her, saith the \textsc{Lord}.
\vs Jer 51:54 A sound of a cry \bibemph{cometh} from Babylon, and great destruction from the land of the Chaldeans:
\vs Jer 51:55 Because the \textsc{Lord} hath spoiled Babylon, and destroyed out of her the great voice; when her waves do roar like great waters, a noise of their voice is uttered:
\vs Jer 51:56 Because the spoiler is come upon her, \bibemph{even} upon Babylon, and her mighty men are taken, every one of their bows is broken: for the \textsc{Lord} God of recompences shall surely requite.
\vs Jer 51:57 And I will make drunk her princes, and her wise \bibemph{men}, her captains, and her rulers, and her mighty men: and they shall sleep a perpetual sleep, and not wake, saith the King, whose name \bibemph{is} the \textsc{Lord} of hosts.
\vs Jer 51:58 Thus saith the \textsc{Lord} of hosts; The broad walls of Babylon shall be utterly broken, and her high gates shall be burned with fire; and the people shall labour in vain, and the folk in the fire, and they shall be weary.
\fn{The broad...: or, The walls of broad Babylon. broken: or, made naked.}
\vs Jer 51:59 The word which Jeremiah the prophet commanded Seraiah the son of Neriah, the son of Maaseiah, when he went with Zedekiah the king of Judah into Babylon in the fourth year of his reign. And \bibemph{this} Seraiah \bibemph{was} a quiet prince.
\fn{with: or, on the behalf of. quiet...: or, prince of Menucha, or, chief chamberlain.}
\vs Jer 51:60 So Jeremiah wrote in a book all the evil that should come upon Babylon, \bibemph{even} all these words that are written against Babylon.
\vs Jer 51:61 And Jeremiah said to Seraiah, When thou comest to Babylon, and shalt see, and shalt read all these words;
\vs Jer 51:62 Then shalt thou say, O \textsc{Lord}, thou hast spoken against this place, to cut it off, that none shall remain in it, neither man nor beast, but that it shall be desolate for ever.
\fn{desolate: Heb. desolations.}
\vs Jer 51:63 And it shall be, when thou hast made an end of reading this book, \bibemph{that} thou shalt bind a stone to it, and cast it into the midst of Euphrates:
\vs Jer 51:64 And thou shalt say, Thus shall Babylon sink, and shall not rise from the evil that I will bring upon her: and they shall be weary. Thus far \bibemph{are} the words of Jeremiah.
\vs Jer 52:1 Zedekiah \bibemph{was} one and twenty years old when he began to reign, and he reigned eleven years in Jerusalem. And his mother's name \bibemph{was} Hamutal the daughter of Jeremiah of Libnah.
\fn{began...: Heb. reigned.}
\vs Jer 52:2 And he did \bibemph{that which was} evil in the eyes of the \textsc{Lord}, according to all that Jehoiakim had done.
\vs Jer 52:3 For through the anger of the \textsc{Lord} it came to pass in Jerusalem and Judah, till he had cast them out from his presence, that Zedekiah rebelled against the king of Babylon.
\vs Jer 52:4 And it came to pass in the ninth year of his reign, in the tenth month, in the tenth \bibemph{day} of the month, \bibemph{that} Nebuchadrezzar king of Babylon came, he and all his army, against Jerusalem, and pitched against it, and built forts against it round about.
\vs Jer 52:5 So the city was besieged unto the eleventh year of king Zedekiah.
\vs Jer 52:6 And in the fourth month, in the ninth \bibemph{day} of the month, the famine was sore in the city, so that there was no bread for the people of the land.
\vs Jer 52:7 Then the city was broken up, and all the men of war fled, and went forth out of the city by night by the way of the gate between the two walls, which \bibemph{was} by the king's garden; (now the Chaldeans \bibemph{were} by the city round about:) and they went by the way of the plain.
\vs Jer 52:8 But the army of the Chaldeans pursued after the king, and overtook Zedekiah in the plains of Jericho; and all his army was scattered from him.
\vs Jer 52:9 Then they took the king, and carried him up unto the king of Babylon to Riblah in the land of Hamath; where he gave judgment upon him.
\vs Jer 52:10 And the king of Babylon slew the sons of Zedekiah before his eyes: he slew also all the princes of Judah in Riblah.
\vs Jer 52:11 Then he put out the eyes of Zedekiah; and the king of Babylon bound him in chains, and carried him to Babylon, and put him in prison till the day of his death.
\fn{put out: Heb. blinded. chains: or, fetters. prison: Heb. house of the wards.}
\vs Jer 52:12 Now in the fifth month, in the tenth \bibemph{day} of the month, which \bibemph{was} the nineteenth year of Nebuchadrezzar king of Babylon, came Nebuzaradan, captain of the guard, \bibemph{which} served the king of Babylon, into Jerusalem,
\fn{captain...: or, chief marshal: Heb. chief of the executioners, or, slaughtermen. served: Heb. stood before.}
\vs Jer 52:13 And burned the house of the \textsc{Lord}, and the king's house; and all the houses of Jerusalem, and all the houses of the great \bibemph{men}, burned he with fire:
\vs Jer 52:14 And all the army of the Chaldeans, that \bibemph{were} with the captain of the guard, brake down all the walls of Jerusalem round about.
\vs Jer 52:15 Then Nebuzaradan the captain of the guard carried away captive \bibemph{certain} of the poor of the people, and the residue of the people that remained in the city, and those that fell away, that fell to the king of Babylon, and the rest of the multitude.
\vs Jer 52:16 But Nebuzaradan the captain of the guard left \bibemph{certain} of the poor of the land for vinedressers and for husbandmen.
\vs Jer 52:17 Also the pillars of brass that \bibemph{were} in the house of the \textsc{Lord}, and the bases, and the brasen sea that \bibemph{was} in the house of the \textsc{Lord}, the Chaldeans brake, and carried all the brass of them to Babylon.
\vs Jer 52:18 The caldrons also, and the shovels, and the snuffers, and the bowls, and the spoons, and all the vessels of brass wherewith they ministered, took they away.
\fn{shovels: or, instruments to remove the ashes. bowls: or, basons.}
\vs Jer 52:19 And the basons, and the firepans, and the bowls, and the caldrons, and the candlesticks, and the spoons, and the cups; \bibemph{that} which \bibemph{was} of gold \bibemph{in} gold, and \bibemph{that} which \bibemph{was} of silver \bibemph{in} silver, took the captain of the guard away.
\fn{firepans: or, censers.}
\vs Jer 52:20 The two pillars, one sea, and twelve brasen bulls that \bibemph{were} under the bases, which king Solomon had made in the house of the \textsc{Lord}: the brass of all these vessels was without weight.
\fn{the brass: Heb. their brass.}
\vs Jer 52:21 And \bibemph{concerning} the pillars, the height of one pillar \bibemph{was} eighteen cubits; and a fillet of twelve cubits did compass it; and the thickness thereof \bibemph{was} four fingers: \bibemph{it was} hollow.
\fn{fillet: Heb. thread.}
\vs Jer 52:22 And a chapiter of brass \bibemph{was} upon it; and the height of one chapiter \bibemph{was} five cubits, with network and pomegranates upon the chapiters round about, all \bibemph{of} brass. The second pillar also and the pomegranates \bibemph{were} like unto these.
\vs Jer 52:23 And there were ninety and six pomegranates on a side; \bibemph{and} all the pomegranates upon the network \bibemph{were} an hundred round about.
\vs Jer 52:24 And the captain of the guard took Seraiah the chief priest, and Zephaniah the second priest, and the three keepers of the door:
\fn{door: Heb. threshold.}
\vs Jer 52:25 He took also out of the city an eunuch, which had the charge of the men of war; and seven men of them that were near the king's person, which were found in the city; and the principal scribe of the host, who mustered the people of the land; and threescore men of the people of the land, that were found in the midst of the city.
\fn{were near...: Heb. saw the face of the king. principal...: or, scribe of the captain of the host.}
\vs Jer 52:26 So Nebuzaradan the captain of the guard took them, and brought them to the king of Babylon to Riblah.
\vs Jer 52:27 And the king of Babylon smote them, and put them to death in Riblah in the land of Hamath. Thus Judah was carried away captive out of his own land.
\vs Jer 52:28 This \bibemph{is} the people whom Nebuchadrezzar carried away captive: in the seventh year three thousand Jews and three and twenty:
\vs Jer 52:29 In the eighteenth year of Nebuchadrezzar he carried away captive from Jerusalem eight hundred thirty and two persons:
\fn{persons: Heb. souls.}
\vs Jer 52:30 In the three and twentieth year of Nebuchadrezzar Nebuzaradan the captain of the guard carried away captive of the Jews seven hundred forty and five persons: all the persons \bibemph{were} four thousand and six hundred.
\vs Jer 52:31 And it came to pass in the seven and thirtieth year of the captivity of Jehoiachin king of Judah, in the twelfth month, in the five and twentieth \bibemph{day} of the month, \bibemph{that} Evilmerodach king of Babylon in the \bibemph{first} year of his reign lifted up the head of Jehoiachin king of Judah, and brought him forth out of prison,
\vs Jer 52:32 And spake kindly unto him, and set his throne above the throne of the kings that \bibemph{were} with him in Babylon,
\fn{kindly...: Heb. good things with him.}
\vs Jer 52:33 And changed his prison garments: and he did continually eat bread before him all the days of his life.
\vs Jer 52:34 And \bibemph{for} his diet, there was a continual diet given him of the king of Babylon, every day a portion until the day of his death, all the days of his life.
\fn{every...: Heb. the matter of the day in his day.}
