\bibbook{\bibtitlemain{jonah}}{jonah}{Jonah}{Jon}
\vs Jon 1:1 Now the word of the \textsc{Lord} came unto Jonah the son of Amittai, saying,
\fn{Jonah: Gr. Jonas.}
\vs Jon 1:2 Arise, go to Nineveh, that great city, and cry against it; for their wickedness is come up before me.
\vs Jon 1:3 But Jonah rose up to flee unto Tarshish from the presence of the \textsc{Lord}, and went down to Joppa; and he found a ship going to Tarshish: so he paid the fare thereof, and went down into it, to go with them unto Tarshish from the presence of the \textsc{Lord}.
\vs Jon 1:4 But the \textsc{Lord} sent out a great wind into the sea, and there was a mighty tempest in the sea, so that the ship was like to be broken.
\fn{sent out: Heb. cast forth. was like...: Heb. thought to be broken.}
\vs Jon 1:5 Then the mariners were afraid, and cried every man unto his god, and cast forth the wares that \bibemph{were} in the ship into the sea, to lighten \bibemph{it} of them. But Jonah was gone down into the sides of the ship; and he lay, and was fast asleep.
\vs Jon 1:6 So the shipmaster came to him, and said unto him, What meanest thou, O sleeper? arise, call upon thy God, if so be that God will think upon us, that we perish not.
\vs Jon 1:7 And they said every one to his fellow, Come, and let us cast lots, that we may know for whose cause this evil \bibemph{is} upon us. So they cast lots, and the lot fell upon Jonah.
\vs Jon 1:8 Then said they unto him, Tell us, we pray thee, for whose cause this evil \bibemph{is} upon us; What \bibemph{is} thine occupation? and whence comest thou? what \bibemph{is} thy country? and of what people \bibemph{art} thou?
\vs Jon 1:9 And he said unto them, I \bibemph{am} an Hebrew; and I fear the \textsc{Lord}, the God of heaven, which hath made the sea and the dry \bibemph{land}.
\fn{the \textsc{Lord}: or, JEHOVAH.}
\vs Jon 1:10 Then were the men exceedingly afraid, and said unto him, Why hast thou done this? For the men knew that he fled from the presence of the \textsc{Lord}, because he had told them.
\fn{exceedingly...: Heb. with great fear.}
\vs Jon 1:11 Then said they unto him, What shall we do unto thee, that the sea may be calm unto us? for the sea wrought, and was tempestuous.
\fn{may...: Heb. may be silent from us. wrought...: or, grew more and more tempestuous: Heb. went.}
\vs Jon 1:12 And he said unto them, Take me up, and cast me forth into the sea; so shall the sea be calm unto you: for I know that for my sake this great tempest \bibemph{is} upon you.
\vs Jon 1:13 Nevertheless the men rowed hard to bring \bibemph{it} to the land; but they could not: for the sea wrought, and was tempestuous against them.
\fn{rowed: Heb. digged. wrought...: or, grew more and more tempestuous: Heb. went.}
\vs Jon 1:14 Wherefore they cried unto the \textsc{Lord}, and said, We beseech thee, O \textsc{Lord}, we beseech thee, let us not perish for this man's life, and lay not upon us innocent blood: for thou, O \textsc{Lord}, hast done as it pleased thee.
\vs Jon 1:15 So they took up Jonah, and cast him forth into the sea: and the sea ceased from her raging.
\fn{ceased: Heb. stood.}
\vs Jon 1:16 Then the men feared the \textsc{Lord} exceedingly, and offered a sacrifice unto the \textsc{Lord}, and made vows.
\fn{offered...: Heb. sacrifice unto the \textsc{Lord}, and vowed vows.}
\vs Jon 1:17 Now the \textsc{Lord} had prepared a great fish to swallow up Jonah. And Jonah was in the belly of the fish three days and three nights.
\fn{belly: Heb. bowels.}
\vs Jon 2:1 Then Jonah prayed unto the \textsc{Lord} his God out of the fish's belly,
\vs Jon 2:2 And said, I cried by reason of mine affliction unto the \textsc{Lord}, and he heard me; out of the belly of hell cried I, \bibemph{and} thou heardest my voice.
\fn{by...: or, out of mine affliction. hell: or, the grave.}
\vs Jon 2:3 For thou hadst cast me into the deep, in the midst of the seas; and the floods compassed me about: all thy billows and thy waves passed over me.
\fn{midst: Heb. heart.}
\vs Jon 2:4 Then I said, I am cast out of thy sight; yet I will look again toward thy holy temple.
\vs Jon 2:5 The waters compassed me about, \bibemph{even} to the soul: the depth closed me round about, the weeds were wrapped about my head.
\vs Jon 2:6 I went down to the bottoms of the mountains; the earth with her bars \bibemph{was} about me for ever: yet hast thou brought up my life from corruption, O \textsc{Lord} my God.
\fn{bottoms: Heb. cuttings off. corruption: or, the pit.}
\vs Jon 2:7 When my soul fainted within me I remembered the \textsc{Lord}: and my prayer came in unto thee, into thine holy temple.
\vs Jon 2:8 They that observe lying vanities forsake their own mercy.
\vs Jon 2:9 But I will sacrifice unto thee with the voice of thanksgiving; I will pay \bibemph{that} that I have vowed. Salvation \bibemph{is} of the \textsc{Lord}.
\vs Jon 2:10 And the \textsc{Lord} spake unto the fish, and it vomited out Jonah upon the dry \bibemph{land}.
\vs Jon 3:1 And the word of the \textsc{Lord} came unto Jonah the second time, saying,
\vs Jon 3:2 Arise, go unto Nineveh, that great city, and preach unto it the preaching that I bid thee.
\vs Jon 3:3 So Jonah arose, and went unto Nineveh, according to the word of the \textsc{Lord}. Now Nineveh was an exceeding great city of three days' journey.
\fn{exceeding: Heb. of God.}
\vs Jon 3:4 And Jonah began to enter into the city a day's journey, and he cried, and said, Yet forty days, and Nineveh shall be overthrown.
\vs Jon 3:5 So the people of Nineveh believed God, and proclaimed a fast, and put on sackcloth, from the greatest of them even to the least of them.
\vs Jon 3:6 For word came unto the king of Nineveh, and he arose from his throne, and he laid his robe from him, and covered \bibemph{him} with sackcloth, and sat in ashes.
\vs Jon 3:7 And he caused \bibemph{it} to be proclaimed and published through Nineveh by the decree of the king and his nobles, saying, Let neither man nor beast, herd nor flock, taste any thing: let them not feed, nor drink water:
\fn{published: Heb. said. nobles: Heb. great men.}
\vs Jon 3:8 But let man and beast be covered with sackcloth, and cry mightily unto God: yea, let them turn every one from his evil way, and from the violence that \bibemph{is} in their hands.
\vs Jon 3:9 Who can tell \bibemph{if} God will turn and repent, and turn away from his fierce anger, that we perish not?
\vs Jon 3:10 And God saw their works, that they turned from their evil way; and God repented of the evil, that he had said that he would do unto them; and he did \bibemph{it} not.
\vs Jon 4:1 But it displeased Jonah exceedingly, and he was very angry.
\vs Jon 4:2 And he prayed unto the \textsc{Lord}, and said, I pray thee, O \textsc{Lord}, \bibemph{was} not this my saying, when I was yet in my country? Therefore I fled before unto Tarshish: for I knew that thou \bibemph{art} a gracious God, and merciful, slow to anger, and of great kindness, and repentest thee of the evil.
\vs Jon 4:3 Therefore now, O \textsc{Lord}, take, I beseech thee, my life from me; for \bibemph{it is} better for me to die than to live.
\vs Jon 4:4 Then said the \textsc{Lord}, Doest thou well to be angry?
\fn{Doest...: or, Art thou greatly angry?.}
\vs Jon 4:5 So Jonah went out of the city, and sat on the east side of the city, and there made him a booth, and sat under it in the shadow, till he might see what would become of the city.
\vs Jon 4:6 And the \textsc{Lord} God prepared a gourd, and made \bibemph{it} to come up over Jonah, that it might be a shadow over his head, to deliver him from his grief. So Jonah was exceeding glad of the gourd.
\fn{gourd: or, palmcrist: Heb. Kikajon. was...: Heb. rejoiced with great joy.}
\vs Jon 4:7 But God prepared a worm when the morning rose the next day, and it smote the gourd that it withered.
\vs Jon 4:8 And it came to pass, when the sun did arise, that God prepared a vehement east wind; and the sun beat upon the head of Jonah, that he fainted, and wished in himself to die, and said, \bibemph{It is} better for me to die than to live.
\fn{vehement: or, silent.}
\vs Jon 4:9 And God said to Jonah, Doest thou well to be angry for the gourd? And he said, I do well to be angry, \bibemph{even} unto death.
\fn{Doest...: or, Art thou greatly angry?. I do well...: or, I am greatly angry.}
\vs Jon 4:10 Then said the \textsc{Lord}, Thou hast had pity on the gourd, for the which thou hast not laboured, neither madest it grow; which came up in a night, and perished in a night:
\fn{had pity: or, spared. came...: Heb. was the son of the night.}
\vs Jon 4:11 And should not I spare Nineveh, that great city, wherein are more than sixscore thousand persons that cannot discern between their right hand and their left hand; and \bibemph{also} much cattle?
