\bibbook{\bibtitlemain{malachi}}{malachi}{Malachi}{Mal}
\vs Mal 1:1 The burden of the word of the \textsc{Lord} to Israel by Malachi.
\fn{by...: Heb. by the hand of.}
\vs Mal 1:2 I have loved you, saith the \textsc{Lord}. Yet ye say, Wherein hast thou loved us? \bibemph{Was} not Esau Jacob's brother? saith the \textsc{Lord}: yet I loved Jacob,
\vs Mal 1:3 And I hated Esau, and laid his mountains and his heritage waste for the dragons of the wilderness.
\vs Mal 1:4 Whereas Edom saith, We are impoverished, but we will return and build the desolate places; thus saith the \textsc{Lord} of hosts, They shall build, but I will throw down; and they shall call them, The border of wickedness, and, The people against whom the \textsc{Lord} hath indignation for ever.
\vs Mal 1:5 And your eyes shall see, and ye shall say, The \textsc{Lord} will be magnified from the border of Israel.
\fn{from: or, upon: Heb. from upon.}
\vs Mal 1:6 A son honoureth \bibemph{his} father, and a servant his master: if then I \bibemph{be} a father, where \bibemph{is} mine honour? and if I \bibemph{be} a master, where \bibemph{is} my fear? saith the \textsc{Lord} of hosts unto you, O priests, that despise my name. And ye say, Wherein have we despised thy name?
\vs Mal 1:7 Ye offer polluted bread upon mine altar; and ye say, Wherein have we polluted thee? In that ye say, The table of the \textsc{Lord} \bibemph{is} contemptible.
\fn{offer...: or, bring unto, etc.}
\vs Mal 1:8 And if ye offer the blind for sacrifice, \bibemph{is it} not evil? and if ye offer the lame and sick, \bibemph{is it} not evil? offer it now unto thy governor; will he be pleased with thee, or accept thy person? saith the \textsc{Lord} of hosts.
\fn{for: Heb. to.}
\vs Mal 1:9 And now, I pray you, beseech God that he will be gracious unto us: this hath been by your means: will he regard your persons? saith the \textsc{Lord} of hosts.
\fn{God: Heb. the face of God. by...: Heb. from your hand.}
\vs Mal 1:10 Who \bibemph{is there} even among you that would shut the doors \bibemph{for nought}? neither do ye kindle \bibemph{fire} on mine altar for nought. I have no pleasure in you, saith the \textsc{Lord} of hosts, neither will I accept an offering at your hand.
\vs Mal 1:11 For from the rising of the sun even unto the going down of the same my name \bibemph{shall be} great among the Gentiles; and in every place incense \bibemph{shall be} offered unto my name, and a pure offering: for my name \bibemph{shall be} great among the heathen, saith the \textsc{Lord} of hosts.
\vs Mal 1:12 But ye have profaned it, in that ye say, The table of the \textsc{Lord} \bibemph{is} polluted; and the fruit thereof, \bibemph{even} his meat, \bibemph{is} contemptible.
\vs Mal 1:13 Ye said also, Behold, what a weariness \bibemph{is it}! and ye have snuffed at it, saith the \textsc{Lord} of hosts; and ye brought \bibemph{that which was} torn, and the lame, and the sick; thus ye brought an offering: should I accept this of your hand? saith the \textsc{Lord}.
\fn{and ye have...: or, whereas ye might have blown it away.}
\vs Mal 1:14 But cursed \bibemph{be} the deceiver, which hath in his flock a male, and voweth, and sacrificeth unto the Lord a corrupt thing: for I \bibemph{am} a great King, saith the \textsc{Lord} of hosts, and my name \bibemph{is} dreadful among the heathen.
\fn{which...: Heb. in whose flock is.}
\vs Mal 2:1 And now, O ye priests, this commandment \bibemph{is} for you.
\vs Mal 2:2 If ye will not hear, and if ye will not lay \bibemph{it} to heart, to give glory unto my name, saith the \textsc{Lord} of hosts, I will even send a curse upon you, and I will curse your blessings: yea, I have cursed them already, because ye do not lay \bibemph{it} to heart.
\vs Mal 2:3 Behold, I will corrupt your seed, and spread dung upon your faces, \bibemph{even} the dung of your solemn feasts; and \bibemph{one} shall take you away with it.
\fn{corrupt: or, reprove. spread: Heb. scatter. one...: or, it shall take you away to it.}
\vs Mal 2:4 And ye shall know that I have sent this commandment unto you, that my covenant might be with Levi, saith the \textsc{Lord} of hosts.
\vs Mal 2:5 My covenant was with him of life and peace; and I gave them to him \bibemph{for} the fear wherewith he feared me, and was afraid before my name.
\vs Mal 2:6 The law of truth was in his mouth, and iniquity was not found in his lips: he walked with me in peace and equity, and did turn many away from iniquity.
\vs Mal 2:7 For the priest's lips should keep knowledge, and they should seek the law at his mouth: for he \bibemph{is} the messenger of the \textsc{Lord} of hosts.
\vs Mal 2:8 But ye are departed out of the way; ye have caused many to stumble at the law; ye have corrupted the covenant of Levi, saith the \textsc{Lord} of hosts.
\fn{stumble at: or, fall in.}
\vs Mal 2:9 Therefore have I also made you contemptible and base before all the people, according as ye have not kept my ways, but have been partial in the law.
\fn{have been...: or, lifted up the face against: Heb. accepted faces.}
\vs Mal 2:10 Have we not all one father? hath not one God created us? why do we deal treacherously every man against his brother, by profaning the covenant of our fathers?
\vs Mal 2:11 Judah hath dealt treacherously, and an abomination is committed in Israel and in Jerusalem; for Judah hath profaned the holiness of the \textsc{Lord} which he loved, and hath married the daughter of a strange god.
\fn{loved: or, ought to love.}
\vs Mal 2:12 The \textsc{Lord} will cut off the man that doeth this, the master and the scholar, out of the tabernacles of Jacob, and him that offereth an offering unto the \textsc{Lord} of hosts.
\fn{the master...: or, him that waketh, and him that answereth.}
\vs Mal 2:13 And this have ye done again, covering the altar of the \textsc{Lord} with tears, with weeping, and with crying out, insomuch that he regardeth not the offering any more, or receiveth \bibemph{it} with good will at your hand.
\vs Mal 2:14 Yet ye say, Wherefore? Because the \textsc{Lord} hath been witness between thee and the wife of thy youth, against whom thou hast dealt treacherously: yet \bibemph{is} she thy companion, and the wife of thy covenant.
\vs Mal 2:15 And did not he make one? Yet had he the residue of the spirit. And wherefore one? That he might seek a godly seed. Therefore take heed to your spirit, and let none deal treacherously against the wife of his youth.
\fn{residue: or, excellency. godly...: Heb. seed of God. treacherously: or, unfaithfully.}
\vs Mal 2:16 For the \textsc{Lord}, the God of Israel, saith that he hateth putting away: for \bibemph{one} covereth violence with his garment, saith the \textsc{Lord} of hosts: therefore take heed to your spirit, that ye deal not treacherously.
\fn{that he...: or, if he hate her, put her away. putting...: Heb. to put away.}
\vs Mal 2:17 Ye have wearied the \textsc{Lord} with your words. Yet ye say, Wherein have we wearied \bibemph{him}? When ye say, Every one that doeth evil \bibemph{is} good in the sight of the \textsc{Lord}, and he delighteth in them; or, Where \bibemph{is} the God of judgment?
\vs Mal 3:1 Behold, I will send my messenger, and he shall prepare the way before me: and the Lord, whom ye seek, shall suddenly come to his temple, even the messenger of the covenant, whom ye delight in: behold, he shall come, saith the \textsc{Lord} of hosts.
\vs Mal 3:2 But who may abide the day of his coming? and who shall stand when he appeareth? for he \bibemph{is} like a refiner's fire, and like fullers' soap:
\vs Mal 3:3 And he shall sit \bibemph{as} a refiner and purifier of silver: and he shall purify the sons of Levi, and purge them as gold and silver, that they may offer unto the \textsc{Lord} an offering in righteousness.
\vs Mal 3:4 Then shall the offering of Judah and Jerusalem be pleasant unto the \textsc{Lord}, as in the days of old, and as in former years.
\fn{former: or, ancient.}
\vs Mal 3:5 And I will come near to you to judgment; and I will be a swift witness against the sorcerers, and against the adulterers, and against false swearers, and against those that oppress the hireling in \bibemph{his} wages, the widow, and the fatherless, and that turn aside the stranger \bibemph{from his right}, and fear not me, saith the \textsc{Lord} of hosts.
\fn{oppress: or, defraud.}
\vs Mal 3:6 For I \bibemph{am} the \textsc{Lord}, I change not; therefore ye sons of Jacob are not consumed.
\vs Mal 3:7 Even from the days of your fathers ye are gone away from mine ordinances, and have not kept \bibemph{them}. Return unto me, and I will return unto you, saith the \textsc{Lord} of hosts. But ye said, Wherein shall we return?
\vs Mal 3:8 Will a man rob God? Yet ye have robbed me. But ye say, Wherein have we robbed thee? In tithes and offerings.
\vs Mal 3:9 Ye \bibemph{are} cursed with a curse: for ye have robbed me, \bibemph{even} this whole nation.
\vs Mal 3:10 Bring ye all the tithes into the storehouse, that there may be meat in mine house, and prove me now herewith, saith the \textsc{Lord} of hosts, if I will not open you the windows of heaven, and pour you out a blessing, that \bibemph{there shall} not \bibemph{be room} enough \bibemph{to receive it}.
\fn{pour...: Heb. empty out.}
\vs Mal 3:11 And I will rebuke the devourer for your sakes, and he shall not destroy the fruits of your ground; neither shall your vine cast her fruit before the time in the field, saith the \textsc{Lord} of hosts.
\fn{destroy: Heb. corrupt.}
\vs Mal 3:12 And all nations shall call you blessed: for ye shall be a delightsome land, saith the \textsc{Lord} of hosts.
\vs Mal 3:13 Your words have been stout against me, saith the \textsc{Lord}. Yet ye say, What have we spoken \bibemph{so much} against thee?
\vs Mal 3:14 Ye have said, It \bibemph{is} vain to serve God: and what profit \bibemph{is it} that we have kept his ordinance, and that we have walked mournfully before the \textsc{Lord} of hosts?
\fn{ordinance: Heb. observation. mournfully: Heb. in black.}
\vs Mal 3:15 And now we call the proud happy; yea, they that work wickedness are set up; yea, \bibemph{they that} tempt God are even delivered.
\fn{are set up: Heb. are built.}
\vs Mal 3:16 Then they that feared the \textsc{Lord} spake often one to another: and the \textsc{Lord} hearkened, and heard \bibemph{it}, and a book of remembrance was written before him for them that feared the \textsc{Lord}, and that thought upon his name.
\vs Mal 3:17 And they shall be mine, saith the \textsc{Lord} of hosts, in that day when I make up my jewels; and I will spare them, as a man spareth his own son that serveth him.
\fn{jewels: or, special treasure.}
\vs Mal 3:18 Then shall ye return, and discern between the righteous and the wicked, between him that serveth God and him that serveth him not.
\vs Mal 4:1 For, behold, the day cometh, that shall burn as an oven; and all the proud, yea, and all that do wickedly, shall be stubble: and the day that cometh shall burn them up, saith the \textsc{Lord} of hosts, that it shall leave them neither root nor branch.
\vs Mal 4:2 But unto you that fear my name shall the Sun of righteousness arise with healing in his wings; and ye shall go forth, and grow up as calves of the stall.
\vs Mal 4:3 And ye shall tread down the wicked; for they shall be ashes under the soles of your feet in the day that I shall do \bibemph{this}, saith the \textsc{Lord} of hosts.
\vs Mal 4:4 Remember ye the law of Moses my servant, which I commanded unto him in Horeb for all Israel, \bibemph{with} the statutes and judgments.
\vs Mal 4:5 Behold, I will send you Elijah the prophet before the coming of the great and dreadful day of the \textsc{Lord}:
\vs Mal 4:6 And he shall turn the heart of the fathers to the children, and the heart of the children to their fathers, lest I come and smite the earth with a curse.
\bibtheend{THE END OF THE PROPHETS.}
