\upaper{16}{The Seven Master Spirits}
\uminitoc{Relation to Triune Deity}
\uminitoc{Relation to the Infinite Spirit}
\uminitoc{Identity and Diversity of the Master Spirits}
\uminitoc{Attributes and Functions of the Master Spirits}
\uminitoc{Relation to Creatures}
\uminitoc{The Cosmic Mind}
\uminitoc{Morals, Virtue, and Personality}
\uminitoc{Urantia Personality}
\uminitoc{Reality of Human Consciousness}
\author{Universal Censor}
\vs p016 0:1 The Seven Master Spirits of Paradise are the primary personalities of the Infinite Spirit. In this sevenfold creative act of self\hyp{}duplication the Infinite Spirit exhausted the associative possibilities mathematically inherent in the factual existence of the three persons of Deity. Had it been possible to produce a larger number of Master Spirits, they would have been created, but there are just seven associative possibilities, and only seven, inherent in three Deities. And this explains why the universe is operated in seven grand divisions, and why the number seven is basically fundamental in its organization and administration.
\vs p016 0:2 The Seven Master Spirits thus have their origin in, and derive their individual characteristics from, the following seven likenesses:
\vs p016 0:3 \ublistelem{1.}\bibnobreakspace The Universal Father.
\vs p016 0:4 \ublistelem{2.}\bibnobreakspace The Eternal Son.
\vs p016 0:5 \ublistelem{3.}\bibnobreakspace The Infinite Spirit.
\vs p016 0:6 \ublistelem{4.}\bibnobreakspace The Father and the Son.
\vs p016 0:7 \ublistelem{5.}\bibnobreakspace The Father and the Spirit.
\vs p016 0:8 \ublistelem{6.}\bibnobreakspace The Son and the Spirit.
\vs p016 0:9 \ublistelem{7.}\bibnobreakspace The Father, Son, and Spirit.
\vs p016 0:10 \pc We know very little about the action of the Father and the Son in the creation of the Master Spirits. Apparently they were brought into existence by the personal acts of the Infinite Spirit, but we have been definitely instructed that both the Father and the Son participated in their origin.
\vs p016 0:11 In spirit character and nature these Seven Spirits of Paradise are as one, but in all other aspects of identity they are very unlike, and the results of their functioning in the superuniverses are such that the individual differences of each are unmistakably discernible. All the afterplans of the seven segments of the grand universe --- and even the correlative segments of outer space --- have been conditioned by the other\hyp{}than\hyp{}spiritual diversity of these Seven Master Spirits of supreme and ultimate supervision.
\vs p016 0:12 The Master Spirits have many functions, but at the present time their particular domain is the central supervision of the seven superuniverses. Each Master Spirit maintains an enormous force\hyp{}focal headquarters, which slowly circulates around the periphery of Paradise, always maintaining a position opposite the superuniverse of immediate supervision and at the Paradise focal point of its specialized power control and segmental energy distribution. The radial boundary lines of any one of the superuniverses do actually converge at the Paradise headquarters of the supervising Master Spirit.
\usection{Relation to Triune Deity}
\vs p016 1:1 The Conjoint Creator, the Infinite Spirit, is necessary to the completion of the triune personalization of undivided Deity. This threefold Deity personalization is inherently sevenfold in possibility of individual and associative expression; hence the subsequent plan to create universes inhabited by intelligent and potentially spiritual beings, duly expressive of the Father, Son, and Spirit, made the personalization of the Seven Master Spirits inescapable. We have come to speak of the threefold personalization of Deity as the \bibemph{absolute inevitability,} while we have come to look upon the appearance of the Seven Master Spirits as the \bibemph{subabsolute inevitability.}
\vs p016 1:2 While the Seven Master Spirits are hardly expressive of \bibemph{threefold} Deity, they are the eternal portrayal of \bibemph{sevenfold} Deity, the active and associative functions of the three ever\hyp{}existent persons of Deity. By and in and through these Seven Spirits, the Universal Father, the Eternal Son, or the Infinite Spirit, or any dual association, is able to function as such. When the Father, the Son, and the Spirit act together, they can and do function through Master Spirit Number Seven, but not as the Trinity. The Master Spirits singly and collectively represent any and all possible Deity functions, single and several, but not collective, not the Trinity. Master Spirit Number Seven is personally nonfunctional with regard to the Paradise Trinity, and that is just why he can function \bibemph{personally} for the Supreme Being.
\vs p016 1:3 But when the Seven Master Spirits vacate their individual seats of personal power and superuniverse authority and assemble about the Conjoint Actor in the triune presence of Paradise Deity, then and there are they collectively representative of the functional power, wisdom, and authority of undivided Deity --- the Trinity --- to and in the evolving universes. Such a Paradise union of the primal sevenfold expression of Deity does actually embrace, literally encompass, all of every attribute and attitude of the three eternal Deities in Supremacy and in Ultimacy. To all practical intents and purposes the Seven Master Spirits do, then and there, encompass the functional domain of the Supreme\hyp{}Ultimate to and in the master universe.
\vs p016 1:4 As far as we can discern, these Seven Spirits are associated with the divine activities of the three eternal persons of Deity; we detect no evidence of direct association with the functioning presences of the three eternal phases of the Absolute. When associated, the Master Spirits represent the Paradise Deities in what may be roughly conceived as the finite domain of action. It might embrace much that is ultimate but \bibemph{not} absolute.
\usection{Relation to the Infinite Spirit}
\vs p016 2:1 Just as the Eternal and Original Son is revealed through the persons of the constantly increasing number of divine Sons, so is the Infinite and Divine Spirit revealed through the channels of the Seven Master Spirits and their associated spirit groups. At the centre of centres the Infinite Spirit is approachable, but not all who attain Paradise are immediately able to discern his personality and differentiated presence; but all who attain the central universe can and do immediately commune with one of the Seven Master Spirits, the one presiding over the superuniverse from which the newly arrived space pilgrim hails.
\vs p016 2:2 To the universe of universes the Paradise Father speaks only through his Son, while he and the Son conjointly act only through the Infinite Spirit. Outside of Paradise and Havona the Infinite Spirit \bibemph{speaks} only by the voices of the Seven Master Spirits.
\vs p016 2:3 \pc The Infinite Spirit exerts an influence of \bibemph{personal presence} within the confines of the Paradise\hyp{}Havona system; elsewhere his personal spirit presence is exerted by and through one of the Seven Master Spirits. Therefore is the superuniverse spirit presence of the Third Source and Centre on any world or in any individual conditioned by the unique nature of the supervisory Master Spirit of that segment of creation. Conversely, the combined lines of spirit force and intelligence pass inward to the Third Person of Deity by way of the Seven Master Spirits.
\vs p016 2:4 \pc The Seven Master Spirits are collectively endowed with the supreme\hyp{}ultimate attributes of the Third Source and Centre. While each one individually partakes of this endowment, only collectively do they disclose the attributes of omnipotence, omniscience, and omnipresence. No one of them can so function universally; as individuals and in the exercise of these powers of supremacy and ultimacy each is personally limited to the superuniverse of immediate supervision.
\vs p016 2:5 All of everything which has been told you concerning the divinity and personality of the Conjoint Actor applies equally and fully to the Seven Master Spirits, who so effectively distribute the Infinite Spirit to the seven segments of the grand universe in accordance with their divine endowment and in the manner of their differing and individually unique natures. It would therefore be proper to apply to the collective group of seven any or all of the names of the Infinite Spirit. Collectively they are one with the Conjoint Creator on all subabsolute levels.
\usection{Identity and Diversity of the Master Spirits}
\vs p016 3:1 The Seven Master Spirits are indescribable beings, but they are distinctly and definitely personal. They have names, but we elect to introduce them by number. As primary personalizations of the Infinite Spirit, they are akin, but as primary expressions of the seven possible associations of triune Deity, they are essentially diverse in nature, and this diversity of nature determines their differential of superuniverse conduct. These Seven Master Spirits may be described as follows:
\vs p016 3:2 \bibemph{Master Spirit Number One.} In a special manner this Spirit is the direct representation of the Paradise Father. He is a peculiar and efficient manifestation of the power, love, and wisdom of the Universal Father. He is the close associate and supernal adviser of the chief of Mystery Monitors, that being who presides over the College of Personalized Adjusters on Divinington. In all associations of the Seven Master Spirits, it is always Master Spirit Number One who speaks for the Universal Father.
\vs p016 3:3 This Spirit presides over the first superuniverse and, while unfailingly exhibiting the divine nature of a primary personalization of the Infinite Spirit, seems more especially to resemble the Universal Father in character. He is always in personal liaison with the seven Reflective Spirits at the headquarters of the first superuniverse.
\vs p016 3:4 \pc \bibemph{Master Spirit Number Two.} This Spirit adequately portrays the matchless nature and charming character of the Eternal Son, the first\hyp{}born of all creation. He is always in close association with all orders of the Sons of God whenever they may happen to be in the residential universe as individuals or in joyous conclave. In all the assemblies of the Seven Master Spirits he always speaks for, and in behalf of, the Eternal Son.
\vs p016 3:5 This Spirit directs the destinies of superuniverse number two and rules this vast domain much as would the Eternal Son. He is always in liaison with the seven Reflective Spirits situated at the capital of the second superuniverse.
\vs p016 3:6 \pc \bibemph{Master Spirit Number Three.} This Spirit personality especially resembles the Infinite Spirit, and he directs the movements and work of many of the high personalities of the Infinite Spirit. He presides over their assemblies and is closely associated with all personalities who take exclusive origin in the Third Source and Centre. When the Seven Master Spirits are in council, it is Master Spirit Number Three who always speaks for the Infinite Spirit.
\vs p016 3:7 This Spirit is in charge of superuniverse number three, and he administers the affairs of this segment much as would the Infinite Spirit. He is always in liaison with the Reflective Spirits at the headquarters of the third superuniverse.
\vs p016 3:8 \pc \bibemph{Master Spirit Number Four.} Partaking of the combined natures of the Father and the Son, this Master Spirit is the determining influence regarding Father\hyp{}Son policies and procedures in the councils of the Seven Master Spirits. This Spirit is the chief director and adviser of those ascendant beings who have attained the Infinite Spirit and thus have become candidates for seeing the Son and the Father. He fosters that enormous group of personalities taking origin in the Father and the Son. When it becomes necessary to represent the Father and the Son in the association of the Seven Master Spirits, it is always Master Spirit Number Four who speaks.
\vs p016 3:9 This Spirit fosters the fourth segment of the grand universe in accordance with his peculiar association of the attributes of the Universal Father and the Eternal Son. He is always in personal liaison with the Reflective Spirits of the headquarters of the fourth superuniverse.
\vs p016 3:10 \pc \bibemph{Master Spirit Number Five.} This divine personality who exquisitely blends the character of the Universal Father and the Infinite Spirit is the adviser of that enormous group of beings known as the power directors, power centres, and physical controllers. This Spirit also fosters all personalities taking origin in the Father and the Conjoint Actor. In the councils of the Seven Master Spirits, when the Father\hyp{}Spirit attitude is in question, it is always Master Spirit Number Five who speaks.
\vs p016 3:11 This Spirit directs the welfare of the fifth superuniverse in such a way as to suggest the combined action of the Universal Father and the Infinite Spirit. He is always in liaison with the Reflective Spirits at the headquarters of the fifth superuniverse.
\vs p016 3:12 \pc \bibemph{Master Spirit Number Six.} This divine being seems to portray the combined character of the Eternal Son and the Infinite Spirit. Whenever the creatures jointly created by the Son and the Spirit forgather in the central universe, it is this Master Spirit who is their adviser; and whenever, in the councils of the Seven Master Spirits, it becomes necessary to speak conjointly for the Eternal Son and the Infinite Spirit, it is Master Spirit Number Six who responds.
\vs p016 3:13 This Spirit directs the affairs of the sixth superuniverse much as would the Eternal Son and the Infinite Spirit. He is always in liaison with the Reflective Spirits at the headquarters of the sixth superuniverse.
\vs p016 3:14 \pc \bibemph{Master Spirit Number Seven.} The presiding Spirit of the seventh superuniverse is a uniquely equal portrayal of the Universal Father, the Eternal Son, and the Infinite Spirit. The Seventh Spirit, the fostering adviser of all triune\hyp{}origin beings, is also the adviser and director of all the ascending pilgrims of Havona, those lowly beings who have attained the courts of glory through the combined ministry of the Father, the Son, and the Spirit.
\vs p016 3:15 The Seventh Master Spirit is not organically representative of the Paradise Trinity; but it is a known fact that his personal and spiritual nature \bibemph{is} the Conjoint Actor’s portraiture in equal proportions of the three infinite persons whose Deity union \bibemph{is} the Paradise Trinity, and whose function as such \bibemph{is} the source of the personal and spiritual nature of God the Supreme. Hence the Seventh Master Spirit discloses a personal and organic relationship to the spirit person of the evolving Supreme. Therefore in the Master Spirit councils on high, when it becomes necessary to cast the ballot for the combined personal attitude of the Father, Son, and Spirit or to depict the spiritual attitude of the Supreme Being, it is Master Spirit Number Seven who functions. He thus inherently becomes the presiding head of the Paradise council of the Seven Master Spirits.
\vs p016 3:16 No one of the Seven Spirits is organically representative of the Paradise Trinity, but when they unite as sevenfold Deity, this union in a deity sense --- not in a personal sense --- equivalates to a functional level associable with Trinity functions. In this sense the “Sevenfold Spirit” is functionally associable with the Paradise Trinity. It is also in this sense that Master Spirit Number Seven sometimes speaks in confirmation of Trinity attitudes or, rather, acts as spokesman for the attitude of the Sevenfold\hyp{}Spirit\hyp{}union regarding the attitude of the Threefold\hyp{}Deity\hyp{}union, the attitude of the Paradise Trinity.
\vs p016 3:17 The multiple functions of the Seventh Master Spirit thus range from a combined portraiture of the \bibemph{personal natures} of the Father, Son, and Spirit, through a representation of the \bibemph{personal attitude} of God the Supreme, to a disclosure of the \bibemph{deity attitude} of the Paradise Trinity. And in certain respects this presiding Spirit is similarly expressive of the \bibemph{attitudes} of the Ultimate and of the Supreme\hyp{}Ultimate.
\vs p016 3:18 It is Master Spirit Number Seven who, in his multiple capacities, personally sponsors the progress of the ascension candidates from the worlds of time in their attempts to achieve comprehension of the undivided Deity of Supremacy. Such comprehension involves a grasp of the existential sovereignty of the Trinity of Supremacy so co\hyp{}ordinated with a concept of the growing experiential sovereignty of the Supreme Being as to constitute the creature grasp of the unity of Supremacy. Creature realization of these three factors equals Havona comprehension of Trinity reality and endows the pilgrims of time with the ability eventually to penetrate the Trinity, to discover the three infinite persons of Deity.
\vs p016 3:19 The inability of the Havona pilgrims fully to find God the Supreme is compensated by the Seventh Master Spirit, whose triune nature in such a peculiar manner is revelatory of the spirit person of the Supreme. During the present universe age of the noncontactability of the person of the Supreme, Master Spirit Number Seven functions in the place of the God of ascendant creatures in the matter of personal relationships. He is the one high spirit being that all ascenders are certain to recognize and somewhat comprehend when they reach the centres of glory.
\vs p016 3:20 This Master Spirit is always in liaison with the Reflective Spirits of Uversa, the headquarters of the seventh superuniverse, our own segment of creation. His administration of Orvonton discloses the marvellous symmetry of the co\hyp{}ordinate blending of the divine natures of Father, Son, and Spirit.
\usection{Attributes and Functions of the Master Spirits}
\vs p016 4:1 The Seven Master Spirits are the full representation of the Infinite Spirit to the evolutionary universes. They represent the Third Source and Centre in the relationships of energy, mind, and spirit. While they function as the co\hyp{}ordinating heads of the universal administrative control of the Conjoint Actor, do not forget that they have their origin in the creative acts of the Paradise Deities. It is literally true that these Seven Spirits are the personalized physical power, cosmic mind, and spiritual presence of the triune Deity, “the Seven Spirits of God sent forth to all the universe.”
\vs p016 4:2 The Master Spirits are unique in that they function on all universe levels of reality excepting the absolute. They are, therefore, efficient and perfect supervisors of all phases of administrative affairs on all levels of superuniverse activities. It is difficult for the mortal mind to understand very much about the Master Spirits because their work is so highly specialized yet all\hyp{}embracing, so exceptionally material and at the same time so exquisitely spiritual. These versatile creators of the cosmic mind are the ancestors of the Universe Power Directors and are, themselves, supreme directors of the vast and far\hyp{}flung spirit\hyp{}creature creation.
\vs p016 4:3 The Seven Master Spirits are the creators of the Universe Power Directors and their associates, entities who are indispensable to the organization, control, and regulation of the physical energies of the grand universe. And these same Master Spirits very materially assist the Creator Sons in the work of shaping and organizing the local universes.
\vs p016 4:4 We are unable to trace any personal connection between the cosmic\hyp{}energy work of the Master Spirits and the force functions of the Unqualified Absolute. The energy manifestations under the jurisdiction of the Master Spirits are all directed from the periphery of Paradise; they do not appear to be in any direct manner associated with the force phenomena identified with the nether surface of Paradise.
\vs p016 4:5 Unquestionably, when we encounter the functional activities of the various Morontia Power Supervisors, we are face to face with certain of the unrevealed activities of the Master Spirits. Who, aside from these ancestors of both physical controllers and spirit ministers, could have contrived so to combine and associate material and spiritual energies as to produce a hitherto nonexistent phase of universe reality --- morontia substance and morontia mind?
\vs p016 4:6 Much of the reality of the spiritual worlds is of the morontia order, a phase of universe reality wholly unknown on Urantia. The goal of personality existence is spiritual, but the morontia creations always intervene, bridging the gulf between the material realms of mortal origin and the superuniverse spheres of advancing spiritual status. It is in this realm that the Master Spirits make their great contribution to the plan of man’s Paradise ascension.
\vs p016 4:7 The Seven Master Spirits have personal representatives who function throughout the grand universe; but since a large majority of these subordinate beings are not directly concerned with the ascendant scheme of mortal progression in the path of Paradise perfection, little or nothing has been revealed about them. Much, very much, of the activity of the Seven Master Spirits remains hidden from human understanding because in no way does it directly pertain to your problem of Paradise ascent.
\vs p016 4:8 \pc It is highly probable, though we cannot offer definite proof, that the Master Spirit of Orvonton exerts a decided influence in the following spheres of activity:
\vs p016 4:9 \ublistelem{1.}\bibnobreakspace The life\hyp{}initiation procedures of the local universe Life Carriers.
\vs p016 4:10 \ublistelem{2.}\bibnobreakspace The life activations of the adjutant mind\hyp{}spirits bestowed upon the worlds by a local universe Creative Spirit.
\vs p016 4:11 \ublistelem{3.}\bibnobreakspace The fluctuations in energy manifestations exhibited by the linear\hyp{}gravity\hyp{}responding units of organized matter.
\vs p016 4:12 \ublistelem{4.}\bibnobreakspace The behaviour of emergent energy when fully liberated from the grasp of the Unqualified Absolute, thus becoming responsive to the direct influence of linear gravity and to the manipulations of the Universe Power Directors and their associates.
\vs p016 4:13 \ublistelem{5.}\bibnobreakspace The bestowal of the ministry spirit of a local universe Creative Spirit, known on Urantia as the Holy Spirit.
\vs p016 4:14 \ublistelem{6.}\bibnobreakspace The subsequent bestowal of the spirit of the bestowal Sons, on Urantia called the Comforter or the Spirit of Truth.
\vs p016 4:15 \ublistelem{7.}\bibnobreakspace The reflectivity mechanism of the local universes and the superuniverse. Many features connected with this extraordinary phenomenon can hardly be reasonably explained or rationally understood without postulating the activity of the Master Spirits in association with the Conjoint Actor and the Supreme Being.
\vs p016 4:16 \pc Notwithstanding our failure adequately to comprehend the manifold workings of the Seven Master Spirits, we are confident there are two realms in the vast range of universe activities with which they have nothing whatever to do: the bestowal and ministry of the Thought Adjusters and the inscrutable functions of the Unqualified Absolute.
\usection{Relation to Creatures}
\vs p016 5:1 Each segment of the grand universe, each individual universe and world, enjoys the benefits of the united counsel and wisdom of all Seven Master Spirits but receives the personal touch and tinge of only one. And the personal nature of each Master Spirit entirely pervades and uniquely conditions his superuniverse.
\vs p016 5:2 Through this personal influence of the Seven Master Spirits every creature of every order of intelligent beings, outside of Paradise and Havona, must bear the characteristic stamp of individuality indicative of the ancestral nature of some one of these Seven Paradise Spirits. As concerns the seven superuniverses, each native creature, man or angel, will forever bear this badge of natal identification.
\vs p016 5:3 The Seven Master Spirits do not directly invade the material minds of the individual creatures on the evolutionary worlds of space. The mortals of Urantia do not experience the personal presence of the mind\hyp{}spirit influence of the Master Spirit of Orvonton. If this Master Spirit does attain any sort of contact with the individual mortal mind during the earlier evolutionary ages of an inhabited world, it must occur through the ministry of the local universe Creative Spirit, the consort and associate of the Creator Son of God who presides over the destinies of each local creation. But this very Creative Mother Spirit is, in nature and character, quite like the Master Spirit of Orvonton.
\vs p016 5:4 The physical stamp of a Master Spirit is a part of man’s material origin. The entire morontia career is lived under the continuing influence of this same Master Spirit. It is hardly strange that the subsequent spirit career of such an ascending mortal never fully eradicates the characteristic stamp of this same supervising Spirit. The impress of a Master Spirit is basic to the very existence of every pre\hyp{}Havona stage of mortal ascension.
\vs p016 5:5 The distinctive personality trends exhibited in the life experience of evolutionary mortals, which are characteristic in each superuniverse, and which are directly expressive of the nature of the dominating Master Spirit, are never fully effaced, not even after such ascenders are subjected to the long training and unifying discipline encountered on the one billion educational spheres of Havona. Even the subsequent intense Paradise culture does not suffice to eradicate the earmarks of superuniverse origin. Throughout all eternity an ascendant mortal will exhibit traits indicative of the presiding Spirit of his superuniverse of nativity. Even in the Corps of the Finality, when it is desired to arrive at or to portray a \bibemph{complete} Trinity relationship to the evolutionary creation, always a group of seven finaliters is assembled, one from each superuniverse.
\usection{The Cosmic Mind}
\vs p016 6:1 The Master Spirits are the sevenfold source of the cosmic mind, the intellectual potential of the grand universe. This cosmic mind is a subabsolute manifestation of the mind of the Third Source and Centre and, in certain ways, is functionally related to the mind of the evolving Supreme Being.
\vs p016 6:2 On a world like Urantia we do not encounter the direct influence of the Seven Master Spirits in the affairs of the human races. You live under the immediate influence of the Creative Spirit of Nebadon. Nevertheless these same Master Spirits dominate the basic reactions of all creature mind because they are the actual sources of the intellectual and spiritual potentials which have been specialized in the local universes for function in the lives of those individuals who inhabit the evolutionary worlds of time and space.
\vs p016 6:3 The fact of the cosmic mind explains the kinship of various types of human and superhuman minds. Not only are kindred spirits attracted to each other, but kindred minds are also very fraternal and inclined towards co\hyp{}operation the one with the other. Human minds are sometimes observed to be running in channels of astonishing similarity and inexplicable agreement.
\vs p016 6:4 \pc There exists in all personality associations of the cosmic mind a quality which might be denominated the “reality response.” It is this universal cosmic endowment of will creatures which saves them from becoming helpless victims of the implied a priori assumptions of science, philosophy, and religion. This reality sensitivity of the cosmic mind responds to certain phases of reality just as energy\hyp{}material responds to gravity. It would be still more correct to say that these supermaterial realities so respond to the mind of the cosmos.
\vs p016 6:5 The cosmic mind unfailingly responds (recognizes response) on three levels of universe reality. These responses are self\hyp{}evident to clear\hyp{}reasoning and deep\hyp{}thinking minds. These levels of reality are:
\vs p016 6:6 \ublistelem{1.}\bibnobreakspace \bibemph{Causation ---} the reality domain of the physical senses, the scientific realms of logical uniformity, the differentiation of the factual and the nonfactual, reflective conclusions based on cosmic response. This is the mathematical form of the cosmic discrimination.
\vs p016 6:7 \ublistelem{2.}\bibnobreakspace \bibemph{Duty ---} the reality domain of morals in the philosophic realm, the arena of reason, the recognition of relative right and wrong. This is the judicial form of the cosmic discrimination.
\vs p016 6:8 \ublistelem{3.}\bibnobreakspace \bibemph{Worship ---} the spiritual domain of the reality of religious experience, the personal realization of divine fellowship, the recognition of spirit values, the assurance of eternal survival, the ascent from the status of servants of God to the joy and liberty of the sons of God. This is the highest insight of the cosmic mind, the reverential and worshipful form of the cosmic discrimination.
\vs p016 6:9 \pc These scientific, moral, and spiritual insights, these cosmic responses, are innate in the cosmic mind, which endows all will creatures. The experience of living never fails to develop these three cosmic intuitions; they are constitutive in the self\hyp{}consciousness of reflective thinking. But it is sad to record that so few persons on Urantia take delight in cultivating these qualities of courageous and independent cosmic thinking.
\vs p016 6:10 \pc In the local universe mind bestowals, these three insights of the cosmic mind constitute the a priori assumptions which make it possible for man to function as a rational and self\hyp{}conscious personality in the realms of science, philosophy, and religion. Stated otherwise, the recognition of the \bibemph{reality} of these three manifestations of the Infinite is by a cosmic technique of self\hyp{}revelation. Matter\hyp{}energy is recognized by the mathematical logic of the senses; mind\hyp{}reason intuitively knows its moral duty; spirit\hyp{}faith (worship) is the religion of the reality of spiritual experience. These three basic factors in reflective thinking may be unified and co\hyp{}ordinated in personality development, or they may become disproportionate and virtually unrelated in their respective functions. But when they become unified, they produce a strong character consisting in the correlation of a factual science, a moral philosophy, and a genuine religious experience. And it is these three cosmic intuitions that give objective validity, reality, to man’s experience in and with things, meanings, and values.
\vs p016 6:11 It is the purpose of education to develop and sharpen these innate endowments of the human mind; of civilization to express them; of life experience to realize them; of religion to ennoble them; and of personality to unify them.
\usection{Morals, Virtue, and Personality}
\vs p016 7:1 Intelligence alone cannot explain the moral nature. Morality, virtue, is indigenous to human personality. Moral intuition, the realization of duty, is a component of human mind endowment and is associated with the other inalienables of human nature: scientific curiosity and spiritual insight. Man’s mentality far transcends that of his animal cousins, but it is his moral and religious natures that especially distinguish him from the animal world.
\vs p016 7:2 The selective response of an animal is limited to the motor level of behaviour. The supposed insight of the higher animals is on a motor level and usually appears only after the experience of motor trial and error. Man is able to exercise scientific, moral, and spiritual insight prior to all exploration or experimentation.
\vs p016 7:3 Only a personality can know what it is doing before it does it; only personalities possess insight in advance of experience. A personality can look before it leaps and can therefore learn from looking as well as from leaping. A nonpersonal animal ordinarily learns only by leaping.
\vs p016 7:4 As a result of experience an animal becomes able to examine the different ways of attaining a goal and to select an approach based on accumulated experience. But a personality can also examine the goal itself and pass judgment on its worth\hyp{}whileness, its value. Intelligence alone can discriminate as to the best means of attaining indiscriminate ends, but a moral being possesses an insight which enables him to discriminate between ends as well as between means. And a moral being in choosing virtue is nonetheless intelligent. He knows what he is doing, why he is doing it, where he is going, and how he will get there.
\vs p016 7:5 When man fails to discriminate the ends of his mortal striving, he finds himself functioning on the animal level of existence. He has failed to avail himself of the superior advantages of that material acumen, moral discrimination, and spiritual insight which are an integral part of his cosmic\hyp{}mind endowment as a personal being.
\vs p016 7:6 \pc Virtue is righteousness --- conformity with the cosmos. To name virtues is not to define them, but to live them is to know them. Virtue is not mere knowledge nor yet wisdom but rather the reality of progressive experience in the attainment of ascending levels of cosmic achievement. In the day\hyp{}by\hyp{}day life of mortal man, virtue is realized by the consistent choosing of good rather than evil, and such choosing ability is evidence of the possession of a moral nature.
\vs p016 7:7 Man’s choosing between good and evil is influenced, not only by the keenness of his moral nature, but also by such influences as ignorance, immaturity, and delusion. A sense of proportion is also concerned in the exercise of virtue because evil may be perpetrated when the lesser is chosen in the place of the greater as a result of distortion or deception. The art of relative estimation or comparative measurement enters into the practice of the virtues of the moral realm.
\vs p016 7:8 \pc Man’s moral nature would be impotent without the art of measurement, the discrimination embodied in his ability to scrutinize meanings. Likewise would moral choosing be futile without that cosmic insight which yields the consciousness of spiritual values. From the standpoint of intelligence, man ascends to the level of a moral being because he is endowed with personality.
\vs p016 7:9 \pc Morality can never be advanced by law or by force. It is a personal and freewill matter and must be disseminated by the contagion of the contact of morally fragrant persons with those who are less morally responsive, but who are also in some measure desirous of doing the Father’s will.
\vs p016 7:10 Moral acts are those human performances which are characterized by the highest intelligence, directed by selective discrimination in the choice of superior ends as well as in the selection of moral means to attain these ends. Such conduct is virtuous. Supreme virtue, then, is wholeheartedly to choose to do the will of the Father in heaven.
\usection{Urantia Personality}
\vs p016 8:1 The Universal Father bestows personality upon numerous orders of beings as they function on diverse levels of universe actuality. Urantia human beings are endowed with personality of the finite\hyp{}mortal type, functioning on the level of the ascending sons of God.
\vs p016 8:2 Though we can hardly undertake to define personality, we may attempt to narrate our understanding of the known factors which go to make up the ensemble of material, mental, and spiritual energies whose interassociation constitutes the mechanism wherein and whereon and wherewith the Universal Father causes his bestowed personality to function.
\vs p016 8:3 Personality is a unique endowment of original nature whose existence is independent of, and antecedent to, the bestowal of the Thought Adjuster. Nevertheless, the presence of the Adjuster does augment the qualitative manifestation of personality. Thought Adjusters, when they come forth from the Father, are identical in nature, but personality is diverse, original, and exclusive; and the manifestation of personality is further conditioned and qualified by the nature and qualities of the associated energies of a material, mindal, and spiritual nature which constitute the organismal vehicle for personality manifestation.
\vs p016 8:4 Personalities may be similar, but they are never the same. Persons of a given series, type, order, or pattern may and do resemble one another, but they are never identical. Personality is that feature of an individual which we \bibemph{know,} and which enables us to identify such a being at some future time regardless of the nature and extent of changes in form, mind, or spirit status. Personality is that part of any individual which enables us to recognize and positively identify that person as the one we have previously known, no matter how much he may have changed because of the modification of the vehicle of expression and manifestation of his personality.
\vs p016 8:5 \pc Creature personality is distinguished by two self\hyp{}manifesting and characteristic phenomena of mortal reactive behaviour: self\hyp{}consciousness and associated relative free will.
\vs p016 8:6 Self\hyp{}consciousness consists in intellectual\tunemarkup{pgkoboaurahd}{\linebreak} awareness of personality actuality; it includes the ability to recognize the reality of other personalities. It indicates capacity for individualized experience in and with cosmic realities, equivalating to the attainment of identity status in the personality relationships of the universe. Self\hyp{}consciousness connotes recognition of the actuality of mind ministration and the realization of relative independence of creative and determinative free will.
\vs p016 8:7 \pc The relative free will which characterizes the self\hyp{}consciousness of human personality is involved in:
\vs p016 8:8 \ublistelem{1.}\bibnobreakspace Moral decision, highest wisdom.
\vs p016 8:9 \ublistelem{2.}\bibnobreakspace Spiritual choice, truth discernment.
\vs p016 8:10 \ublistelem{3.}\bibnobreakspace Unselfish love, brotherhood service.
\vs p016 8:11 \ublistelem{4.}\bibnobreakspace Purposeful co\hyp{}operation, group loyalty.
\vs p016 8:12 \ublistelem{5.}\bibnobreakspace Cosmic insight, the grasp of universe meanings.
\vs p016 8:13 \ublistelem{6.}\bibnobreakspace Personality dedication, wholehearted devotion to doing the Father’s will.
\vs p016 8:14 \ublistelem{7.}\bibnobreakspace Worship, the sincere pursuit of divine values and the wholehearted love of the divine Value\hyp{}Giver.
\vs p016 8:15 \pc The Urantia type of human personality may be viewed as functioning in a physical mechanism consisting of the planetary modification of the Nebadon type of organism belonging to the electrochemical order of life activation and endowed with the Nebadon order of the Orvonton series of the cosmic mind of parental reproductive pattern. The bestowal of the divine gift of personality upon such a mind\hyp{}endowed mortal mechanism confers the dignity of cosmic citizenship and enables such a mortal creature forthwith to become reactive to the constitutive recognition of the three basic mind realities of the cosmos:
\vs p016 8:16 \ublistelem{1.}\bibnobreakspace The mathematical or logical recognition of the uniformity of physical causation.
\vs p016 8:17 \ublistelem{2.}\bibnobreakspace The reasoned recognition of the obligation of moral conduct.
\vs p016 8:18 \ublistelem{3.}\bibnobreakspace The faith\hyp{}grasp of the fellowship worship of Deity, associated with the loving service of humanity.
\vs p016 8:19 \pc The full function of such a personality endowment is the beginning realization of Deity kinship. Such a selfhood, indwelt by a prepersonal fragment of God the Father, is in truth and in fact a spiritual son of God. Such a creature not only discloses capacity for the reception of the gift of the divine presence but also exhibits reactive response to the personality\hyp{}gravity circuit of the Paradise Father of all personalities.
\usection{Reality of Human Consciousness}
\vs p016 9:1 The cosmic\hyp{}mind\hyp{}endowed, Adjuster\hyp{}indwelt, personal\tunemarkup{pgnexus10}{\linebreak} creature possesses innate recognition\hyp{}realization of energy reality, mind reality, and spirit reality. The will creature is thus equipped to discern the fact, the law, and the love of God. Aside from these three inalienables of human consciousness, all human experience is really subjective except that intuitive realization of validity attaches to the \bibemph{unification} of these three universe reality responses of cosmic recognition.
\vs p016 9:2 The God\hyp{}discerning mortal is able to sense the unification value of these three cosmic qualities in the evolution of the surviving soul, man’s supreme undertaking in the physical tabernacle where the moral mind collaborates with the indwelling divine spirit to dualize the immortal soul. From its earliest inception the soul is \bibemph{real;} it has cosmic survival qualities.
\vs p016 9:3 If mortal man fails to survive natural death, the real spiritual values of his human experience survive as a part of the continuing experience of the Thought Adjuster. The personality values of such a nonsurvivor persist as a factor in the personality of the actualizing Supreme Being. Such persisting qualities of personality are deprived of identity but not of experiential values accumulated during the mortal life in the flesh. The survival of identity is dependent on the survival of the immortal soul of morontia status and increasingly divine value. Personality identity survives in and by the survival of the soul.
\vs p016 9:4 \pc Human self\hyp{}consciousness implies the recognition of the reality of selves other than the conscious self and further implies that such awareness is mutual; that the self is known as it knows. This is shown in a purely human manner in man’s social life. But you cannot become so absolutely certain of a fellow being’s reality as you can of the reality of the presence of God that lives within you. The social consciousness is not inalienable like the God\hyp{}consciousness; it is a cultural development and is dependent on knowledge, symbols, and the contributions of the constitutive endowments of man --- science, morality, and religion. And these cosmic gifts, socialized, constitute civilization.
\vs p016 9:5 Civilizations are unstable because they are not cosmic; they are not innate in the individuals of the races. They must be nurtured by the combined contributions of the constitutive factors of man --- science, morality, and religion. Civilizations come and go, but science, morality, and religion always survive the crash.
\vs p016 9:6 Jesus not only revealed God to man, but he also made a new revelation of man to himself and to other men. In the life of Jesus you see man at his best. Man thus becomes so beautifully real because Jesus had so much of God in his life, and the realization (recognition) of God is inalienable and constitutive in all men.
\vs p016 9:7 \pc Unselfishness, aside from parental instinct, is not altogether natural; other persons are not naturally loved or socially served. It requires the enlightenment of reason, morality, and the urge of religion, God\hyp{}knowingness, to generate an unselfish and altruistic social order. Man’s own personality awareness, self\hyp{}consciousness, is also directly dependent on this very fact of innate other\hyp{}awareness, this innate ability to recognize and grasp the reality of other personality, ranging from the human to the divine.
\vs p016 9:8 Unselfish social consciousness must be, at bottom, a religious consciousness; that is, if it is objective; otherwise it is a purely subjective philosophic abstraction and therefore devoid of love. Only a God\hyp{}knowing individual can love another person as he loves himself.
\vs p016 9:9 Self\hyp{}consciousness is in essence a communal consciousness: God and man, Father and son, Creator and creature. In human self\hyp{}consciousness four universe\hyp{}reality realizations are latent and inherent:
\vs p016 9:10 \ublistelem{1.}\bibnobreakspace The quest for knowledge, the logic of science.
\vs p016 9:11 \ublistelem{2.}\bibnobreakspace The quest for moral values, the sense of duty.
\vs p016 9:12 \ublistelem{3.}\bibnobreakspace The quest for spiritual values, the religious experience.
\vs p016 9:13 \ublistelem{4.}\bibnobreakspace The quest for personality values, the ability to recognize the reality of God as a personality and the concurrent realization of our fraternal relationship with fellow personalities.
\vs p016 9:14 \pc You become conscious of man as your creature brother because you are already conscious of God as your Creator Father. Fatherhood is the relationship out of which we reason ourselves into the recognition of brotherhood. And Fatherhood becomes, or may become, a universe reality to all moral creatures because the Father has himself bestowed personality upon all such beings and has encircuited them within the grasp of the universal personality circuit. We worship God, first, because \bibemph{he is,} then, because \bibemph{he is in us,} and last, because \bibemph{we are in him.}
\vs p016 9:15 Is it strange that the cosmic mind should be self\hyp{}consciously aware of its own source, the infinite mind of the Infinite Spirit, and at the same time conscious of the physical reality of the far\hyp{}flung universes, the spiritual reality of the Eternal Son, and the personality reality of the Universal Father?
\vsetoff
\vs p016 9:16 [Sponsored by a Universal Censor from Uversa.]
\quizlink
\begin{thebibliography}{100}
\bibitem{Griffiths1}
Rees Griffiths.
{<<God in Idea and Experience, Or, The A Priori Elements of the Religious Consciousness: An Epistemological Study>>.}
{\em Edinburgh: T. \&\ T. Clark}, 1931.
\bibitem{Schoen1}
Max Schoen.
{<<A Scientific Basis for Moral Action>>.}
{\em The Scientific Monthly, Vol.\,48, No.\,3}, March 1939.
\end{thebibliography}
