\upaper{22}{The Trinitized Sons of God}
\author{Mighty Messenger}
\vs p022 0:1 There are three groups of beings who are called Sons of God. In addition to descending and ascending orders of sonship there is a third group known as the Trinitized Sons of God. The trinitized order of sonship is subdivided into three primary divisions in accordance with the origins of its many types of personalities, revealed and unrevealed. These primary divisions are:
\vs p022 0:2 \ublistelem{1.}\bibnobreakspace Deity\hyp{}trinitized Sons.
\vs p022 0:3 \ublistelem{2.}\bibnobreakspace Trinity\hyp{}embraced Sons.
\vs p022 0:4 \ublistelem{3.}\bibnobreakspace Creature\hyp{}trinitized Sons.
\vs p022 0:5 Irrespective of origin all Trinitized Sons of God have in common the experience of trinitization, either as a part of their origin or as an experience of Trinity embrace subsequently attained. The Deity\hyp{}trinitized Sons are unrevealed in these narratives; therefore will this presentation be confined to a portrayal of the remaining two groups, more particularly the Trinity\hyp{}embraced sons of God.
\usection{1.\bibnobreakspace The Trinity\hyp{}Embraced Sons}
\vs p022 1:1 All Trinity\hyp{}embraced sons are originally of dual or single origin, but subsequent to the Trinity embrace they are forever devoted to Trinity service and assignment. This corps, as revealed and as organized for superuniverse service, embraces seven orders of personalities:
\vs p022 1:2 \ublistelem{1.}\bibnobreakspace Mighty Messengers.
\vs p022 1:3 \ublistelem{2.}\bibnobreakspace Those High in Authority.
\vs p022 1:4 \ublistelem{3.}\bibnobreakspace Those without Name and Number.
\vs p022 1:5 \ublistelem{4.}\bibnobreakspace Trinitized Custodians.
\vs p022 1:6 \ublistelem{5.}\bibnobreakspace Trinitized Ambassadors.
\vs p022 1:7 \ublistelem{6.}\bibnobreakspace Celestial Guardians.
\vs p022 1:8 \ublistelem{7.}\bibnobreakspace High Son Assistants.
\vs p022 1:9 \pc These seven groups of personalities are further classified, according to origin, nature, and function, into three major divisions: the Trinitized Sons of Attainment, the Trinitized Sons of Selection, and the Trinitized Sons of Perfection.
\vs p022 1:10 \pc \bibemph{The Trinitized Sons of Attainment ---} the Mighty Messengers, Those High in Authority, and Those without Name and Number --- are all Adjuster\hyp{}fused ascendant mortals who have attained Paradise and the Corps of the Finality. But they are not finaliters; when they have been Trinity embraced, their names are removed from the finaliter roll call. The new sons of this order pass through specific courses of training, for comparatively short periods, on the circuit headquarters planets of the Havona circuits under the direction of the Eternals of Days. Thereafter they are assigned to the services of the Ancients of Days in the seven superuniverses.
\vs p022 1:11 \pc \bibemph{The Trinitized Sons of Selection} embrace the Trinitized Custodians and the Trinitized Ambassadors. They are recruited from certain of the evolutionary seraphim and translated midway creatures who have traversed Havona and have attained Paradise, as well as from certain of the Spirit\hyp{}fused and the Son\hyp{}fused mortals who have likewise ascended to the central Isle of Light and Life. Subsequent to their embrace by the Paradise Trinity and after a brief training in Havona, the Trinitized Sons of Selection are assigned to the courts of the Ancients of Days.
\vs p022 1:12 \pc \bibemph{The Trinitized Sons of Perfection.} The Celestial Guardians and their co\hyp{}ordinates, the High Son Assistants, comprise a unique group of twice\hyp{}trinitized personalities. They are the creature\hyp{}trinitized sons of Paradise\hyp{}Havona personalities or of perfected ascendant mortals who have long distinguished themselves in the Corps of the Finality. Some of these creature\hyp{}trinitized sons, after service with the Supreme Executives of the Seven Master Spirits and after serving under the Trinity Teacher Sons, are retrinitized (embraced) by the Paradise Trinity and then commissioned to the courts of the Ancients of Days as Celestial Guardians and as High Son Assistants. Trinitized Sons of Perfection are assigned directly to the superuniverse service without further training.
\vs p022 1:13 \pc Our Trinity\hyp{}origin associates --- Perfectors of Wisdom, Divine Counsellors, and Universal Censors --- are of stationary numbers, but the Trinity\hyp{}embraced sons are constantly increasing. All seven orders of Trinity\hyp{}embraced sons are commissioned as members of one of the seven superuniverse governments, and the number in the service of each superuniverse is exactly the same; not one has ever been lost. Trinity\hyp{}embraced beings have never gone astray; they may stumble temporarily, but not one has ever been adjudged in contempt of the superuniverse governments. The Sons of Attainment and the Sons of Selection have never faltered in the service of Orvonton, but the Trinitized Sons of Perfection have sometimes erred in judgment and thereby caused transient confusion.
\vs p022 1:14 Under the direction of the Ancients of Days all seven orders function very much as self\hyp{}governing groups. Their scope of service is far\hyp{}flung; Trinitized Sons of Perfection do not leave the superuniverse of assignment, but their trinitized associates range the grand universe, journeying from the evolutionary worlds of time and space to the eternal Isle of Paradise. They may function in any of the superuniverses, but they do so always as members of the supergovernment of original designation.
\vs p022 1:15 Apparently the Trinity\hyp{}embraced sons have been permanently assigned to the service of the seven superuniverses; certainly this assignment is for the duration of the present universe age, but we have never been informed that it is to be eternal.
\usection{2.\bibnobreakspace The Mighty Messengers}
\vs p022 2:1 Mighty Messengers belong to the ascendant group of the Trinitized Sons. They are a class of perfected mortals who have been rebellion tested or otherwise equally proved as to their personal loyalty; all have passed through some definite test of universe allegiance. At some time in their Paradise ascent they stood firm and loyal in the face of the disloyalty of their superiors, and some did actively and loyally function in the places of such unfaithful leaders.
\vs p022 2:2 With such personal records of fidelity and devotion, these ascending mortals pass on through Havona with the stream of the pilgrims of time, attain Paradise, graduate therefrom, and are mustered into the Corps of the Finality. Thereupon they are trinitized in the secret embrace of the Paradise Trinity and subsequently are commissioned to become associated with the Ancients of Days in the administration of the governments of the seven superuniverses.
\vs p022 2:3 Every ascendant mortal of insurrectionary experience who functions loyally in the face of rebellion is eventually destined to become a Mighty Messenger of the superuniverse service. Likewise is any ascendant creature who effectively prevents such upheavals of error, evil, or sin; for action designed to prevent rebellion or to effect higher types of loyalty in a universe crisis is regarded as of even greater value than loyalty in the face of actual rebellion.
\vs p022 2:4 The senior Mighty Messengers were chosen from those ascendant mortals of time and space who were among the earlier Paradise arrivals, many having traversed Havona in the times of Grandfanda. But the first trinitizing of Mighty Messengers was not effected until the candidate corps contained representatives from each of the seven superuniverses. And the last group of this order to qualify on Paradise embraced ascendant pilgrims from the local universe of Nebadon.
\vs p022 2:5 Mighty Messengers are embraced by the Paradise Trinity in classes of 700,000, 100,000 for assignment to each superuniverse. Almost one trillion Mighty Messengers are commissioned on Uversa, and there is every reason to believe that the number serving in each of the seven superuniverses is exactly the same.
\vs p022 2:6 \pc I am a Mighty Messenger, and it may interest Urantians to know that the companion and associate of my mortal experience was also triumphant in the great test, and that, though we were many times and for long periods separated in the agelong inward ascent to Havona, we were embraced in the same 700,000 group, and that we spent our time passing through Vicegerington in close and loving association. We were finally commissioned and together assigned to Uversa of Orvonton, and we are often dispatched in company for the execution of assignments requiring the services of two Messengers.
\vs p022 2:7 \pc Mighty Messengers, in common with all Trinity\hyp{}embraced sons, are assigned to all phases of superuniverse activities. They maintain constant connection with their headquarters through the superuniverse reflectivity service. Mighty Messengers serve in all sectors of a superuniverse and frequently execute missions to the local universes and even to the individual worlds, as I do on this occasion.
\vs p022 2:8 In the superuniverse courts, Mighty Messengers act as defenders of both individuals and planets when they come up for adjudication; they also assist the Perfections of Days in the direction of the affairs of the major sectors. As a group, their chief assignment is that of superuniverse observers. They are stationed on the various headquarters worlds and on individual planets of importance as the official observers of the Ancients of Days. When so assigned, they also serve as advisers to the authorities directing the affairs of the spheres of their sojourn. The Messengers take active part in all phases of the ascendant scheme of mortal progression. With their associates of mortal origin they keep the supergovernments in close and personal touch with the status and progression of the plans of the descending Sons of God.
\vs p022 2:9 Mighty Messengers are fully conscious of their entire ascendant careers, and that is why they are such useful and sympathetic ministers, understanding messengers, for service on any world of space and to any creature of time. As soon as you are delivered from the flesh, you will communicate freely and understandingly with us since we spring from all the races on all the evolutionary worlds of space, that is, from those mortal races that are indwelt by, and subsequently fused with, Thought Adjusters.
\usection{3.\bibnobreakspace Those High in Authority}
\vs p022 3:1 Those High in Authority, the second group of the Trinitized Sons of Attainment, are all Adjuster\hyp{}fused beings of mortal origin. These are the perfected mortals who have exhibited superior administrative ability and have shown extraordinary executive genius throughout their long ascending careers. They are the cream of governing ability derived from the surviving mortals of space.
\vs p022 3:2 70,000 of Those High in Authority are trinitized at each Trinity liaison. Though the local universe of Nebadon is a comparatively young creation, it has representatives among a recently trinitized class of this order. There are now commissioned in Orvonton more than ten billion of these skillful administrators. Like all separate orders of celestial beings, they maintain their own headquarters on Uversa, and like the other Trinity\hyp{}embraced sons, their reserves on Uversa act as the central directing body of their order in Orvonton.
\vs p022 3:3 Those High in Authority are administrators without limitation. They are the everywhere\hyp{}present and always\hyp{}efficient executives of the Ancients of Days. They serve on any sphere, on any inhabited world, and in any phase of activity in any of the seven superuniverses.
\vs p022 3:4 Having superb administrative wisdom and unusual executive skill, these brilliant beings assume to present the cause of justice in behalf of the superuniverse tribunals; they foster the execution of justice and the rectification of misadaptations in the evolutionary universes. Therefore, if you should ever be cited for errors of judgment while you are ascending the worlds and spheres of your ordained cosmic progression, it is hardly likely that you would suffer injustice since your prosecutors would be onetime ascendant creatures who are personally familiar with every step of the career you have traversed and are traversing.
\usection{4.\bibnobreakspace Those Without Name and Number}
\vs p022 4:1 Those without Name and Number constitute the third and last group of the Trinitized Sons of Attainment; they are the ascendant souls who have developed the ability to worship beyond the skill of all the sons and daughters of the evolutionary races from the worlds of time and space. They have acquired a spiritual concept of the eternal purpose of the Universal Father which comparatively transcends the comprehension of the evolutionary creatures of name or number; therefore are they denominated Those without Name and Number. More strictly translated, their name would be “Those \bibemph{above} Name and Number.”
\vs p022 4:2 This order of sons is embraced by the Paradise Trinity in groups of 7,000. There are of record on Uversa over 100,000,000 of these sons commissioned in Orvonton.
\vs p022 4:3 Since Those without Name and Number are the superior spiritual minds of the survival races, they are especially qualified to sit in judgment and to render opinions when a spiritual viewpoint is desirable, and when experience in the ascendant career is essential to an adequate comprehension of the questions involved in the problem to be adjudicated. They are the supreme jurors of Orvonton. A maladministered jury system may be more or less of a travesty of justice on some worlds, but on Uversa and its extension tribunals we employ the highest type of evolved spiritual mentality as juror\hyp{}judges. Adjudication is the highest function of any government, and those who are entrusted with verdict rendering should be chosen from the highest and most noble types of the most experienced and understanding individuals.
\vs p022 4:4 \pc The selection of candidates for the trinitization classes of Mighty Messengers, Those High in Authority, and Those without Name and Number is inherent and automatic. The selective techniques of Paradise are not in any sense arbitrary. Personal experience and spiritual values determine the personnel of the Trinitized Sons of Attainment. Such beings are equal in authority and uniform in administrative status, but they all possess individuality and diverse characters; they are not standardized beings. All are characteristically different, depending on the differentials of their ascendant careers.
\vs p022 4:5 In addition to these experiential qualifications, the Trinitized Sons of Attainment have been trinitized in the divine embrace of the Paradise Deities. Consequently they function as the co\hyp{}ordinate associates of the Stationary Sons of the Trinity, for the Trinity embrace does seem to precipitate out of the stream of future time many of the unrealized potentials of creature beings. But this is true concerning only that which pertains to the present universe age.
\vs p022 4:6 This group of sons is chiefly, but not wholly, concerned with the services of the ascendant career of the time\hyp{}space mortals. If the viewpoint of a mortal creature is ever in doubt, the question is settled by appeal to an ascendant commission consisting of a Mighty Messenger, One High in Authority, and One without Name and Number.
\vs p022 4:7 You mortals who read this message may yourselves ascend to Paradise, attain the Trinity embrace, and in remote future ages be attached to the service of the Ancients of Days in one of the seven superuniverses, and sometime be assigned to enlarge the revelation of truth to some evolving inhabited planet, even as I am now functioning on Urantia.
\usection{5.\bibnobreakspace The Trinitized Custodians}
\vs p022 5:1 The Trinitized Custodians are Trinitized Sons of Selection. Not only do your races and other mortals of survival value traverse Havona, attain Paradise, and sometimes find themselves destined to superuniverse service with the Stationary Sons of the Trinity, but your faithful seraphic guardians and your equally faithful midway associates may also become candidates for the same Trinity recognition and superb personality destiny.
\vs p022 5:2 Trinitized Custodians are ascendant seraphim and translated midway creatures who have passed through Havona and have attained Paradise and the Corps of the Finality. Subsequently they were embraced by the Paradise Trinity and were assigned to the service of the Ancients of Days.
\vs p022 5:3 The candidates for the Trinity embrace from among the ascendant seraphim are accorded this recognition because of their valiant co\hyp{}operation with some ascendant mortal who attained the Corps of the Finality and was subsequently trinitized. My own seraphic guardian of the mortal career went through with me, was later trinitized, and now is attached to the Uversa government as a Trinitized Custodian.
\vs p022 5:4 And so with the midway creatures; many are translated and achieve Paradise and, along with the seraphim and for the same reasons, are Trinity embraced and commissioned as Custodians in the superuniverses.
\vs p022 5:5 The Trinitized Custodians are embraced by the Paradise Trinity in groups of 70,000, and \bibfrac{1}{7} of each group is assigned to a superuniverse. There are now in the service of Orvonton slightly over 10,000,000 of these trusted and high Custodians. They serve on Uversa and on the major and minor headquarters spheres. In their labours they are assisted by a corps of several billion seconaphim and other able superuniverse personalities.
\vs p022 5:6 The Trinitized Custodians start out their careers as custodians, and they continue as such in the affairs of the supergovernments. In a way, they are officers of their superuniverse governments, but they do not deal with individuals, as do the Celestial Guardians. The Trinitized Custodians administer group affairs and foster collective projects. They are the custodians of records, plans, and institutions; they act as the trustees of undertakings, personality groups, ascendant projects, morontia plans, universe projections, and innumerable other enterprises.
\usection{6.\bibnobreakspace The Trinitized Ambassadors}
\vs p022 6:1 Trinitized Ambassadors are the second order of the Trinitized Sons of Selection and like their associates, the Custodians, are recruited from two types of ascendant creatures. Not all ascending mortals are Adjuster or Father fused; some are Spirit fused, some are Son fused. Certain of these Spirit\hyp{} and Son\hyp{}fused mortals reach Havona and attain Paradise. From among these Paradise ascenders, candidates are selected for the Trinity embrace, and from time to time they are trinitized in classes of 7,000. They are then commissioned in the superuniverses as Trinitized Ambassadors of the Ancients of Days. Almost 500,000,000 are registered on Uversa.
\vs p022 6:2 Trinitized Ambassadors are selected for the Trinity embrace upon the advices of their Havona teachers. They represent the superior minds of their respective groups and are, therefore, best qualified to assist the superuniverse rulers in understanding and in administering the interests of those worlds from which the Spirit\hyp{}fused mortals hail. The Son\hyp{}fused Ambassadors are of great assistance in our dealings with problems involving the Son\hyp{}fused order of personality.
\vs p022 6:3 Trinitized Ambassadors are the emissaries of the Ancients of Days for any and all purposes, to any and all worlds or universes within the superuniverse of their assignment. They render particular and important services on the headquarters of the minor sectors, and they perform the numberless miscellaneous assignments of a superuniverse. They are the emergency or reserve corps of the Trinitized Sons of the supergovernments, and they are therefore available for a great range of duties. They engage in thousands upon thousands of undertakings in superuniverse affairs which it is impossible to portray to human minds since there is nothing transpiring on Urantia that is in any way analogous to these activities.
\usection{7.\bibnobreakspace Technique of Trinitization}
\vs p022 7:1 I cannot fully unfold to the material mind the experience of the supreme creative performance of perfect and perfected spiritual beings --- the act of trinitization. The techniques of trinitization are among the secrets of Vicegerington and Solitarington and are revealable to, and understandable by, none save those who have passed through these unique experiences. Therefore is it beyond the possibility of any being successfully to portray to the human mind the nature and purport of this extraordinary transaction.
\vs p022 7:2 Aside from the Deities, only Paradise\hyp{}Havona personalities and certain members of each of the finaliter corps engage in trinitization. Under specialized conditions of Paradise perfection, these superb beings may embark upon the unique adventure of concept\hyp{}identity, and they are many times successful in the production of a new being, a creature\hyp{}trinitized son.
\vs p022 7:3 The glorified creatures who engage in such adventures of trinitization may participate in only one such experience, whereas with the Paradise Deities there seems to be no limit to the continued enactment of trinitization episodes. Deity seems to be limited in just one respect: There can be only one Original and Infinite Spirit, only one infinite executive of the united will of the Father\hyp{}Son.
\vs p022 7:4 The ascendant Adjuster\hyp{}fused mortal finaliters who have attained certain levels of Paradise culture and spiritual development are among those who can essay to trinitize a creature being. Mortal\hyp{}finaliter companies, when stationed on Paradise, are granted a recess every millennium of Havona time. There are seven different ways such finaliters may elect to spend this duty\hyp{}free period, and one of these is, in association with some fellow finaliter or some Paradise\hyp{}Havona personality, to attempt the enactment of creature trinitization.
\vs p022 7:5 If two mortal finaliters, on going before the Architects of the Master Universe, demonstrate that they have independently chosen an identical concept for trinitization, the Architects are empowered, on their own discretion, to promulgate mandates permitting these glorified mortal ascenders to extend their recess and to remove themselves for a time to the trinitizing sector of the Paradise Citizens. At the end of this assigned retreat, if they report that they have singly and jointly elected to make the paradisiacal effort to spiritualize, idealize, and actualize a selected and original concept which has not theretofore been trinitized, then does Master Spirit Number Seven issue orders authorizing such an extraordinary undertaking.
\vs p022 7:6 Unbelievably long periods of time are sometimes consumed in these adventures; an age seems to pass before these faithful and determined onetime mortals --- and sometimes Paradise\hyp{}Havona personalities --- finally achieve their goal, really succeed in bringing their chosen concept of universal truth into actual being. And not always do these devoted couples meet with success; many times they fail, and that through no discoverable error on their part. Candidates for trinitization who thus fail are admitted to a special group of finaliters who are designated as beings who have made the supreme effort and sustained the supreme disappointment. When the Paradise Deities unite to trinitize, they always succeed, but not so with a homogeneous pair of creatures, the attempted union of two members of the same order of being.
\vs p022 7:7 \pc When a new and original being is trinitized by the Gods, the divine parents are in deity potential unchanged; but when exalted creature beings enact such a creative episode, one of the contracting and participating individuals undergoes a unique personality modification. The two ancestors of a creature\hyp{}trinitized son become in a certain sense spiritually as one. We believe that this status of bi\hyp{}unification of certain spiritual phases of personality will probably prevail until such time as the Supreme Being shall have attained full and completed manifestation of personality in the grand universe.
\vs p022 7:8 Simultaneously with the appearance of a new creature\hyp{}trinitized son, there occurs this functional spiritual union of the two ancestors; the two trinitizing parents become one on the ultimate functional level. No created being in the universe can fully explain this amazing phenomenon; it is a near\hyp{}divine experience. When the Father and the Son united to eternalize the Infinite Spirit, upon the accomplishment of their purpose they immediately became as one and ever since have been one. And while the trinitization union of two creatures is on the order of the infinite scope of the perfect Deity union of the Universal Father and the Eternal Son, the repercussions of creature trinitization are not eternal in nature; they will terminate upon the completed factualization of the experiential Deities.
\vs p022 7:9 While these parents of creature\hyp{}trinitized sons become as one in their universe assignments, they continue to be reckoned as two personalities in the make\hyp{}up and roll calls of the Corps of the Finality and of the Architects of the Master Universe. During the current universe age, all trinitization\hyp{}united parents are inseparable in assignment and function; where one goes the other goes, what one does the other does. If parental bi\hyp{}unification involves a mortal (or other) finaliter and a Paradise\hyp{}Havona personality, the united parental beings function neither with the Paradisers, Havoners, nor finaliters. Such mixed unions forgather in a special corps made up of similar beings. And in all trinitization unions, mixed or otherwise, the parental beings are conscious of, and can communicate with, each other, and they can perform duties that neither could have previously discharged.
\vs p022 7:10 \pc The Seven Master Spirits have authority to sanction the trinitizing union of finaliters and Paradise\hyp{}Havona personalities, and such mixed liaisons are always successful. The resultant magnificent creature\hyp{}trinitized sons are representative of concepts unsuited to the comprehension of either the eternal creatures of Paradise or the time creatures of space; hence they become the wards of the Architects of the Master Universe. These trinitized sons of destiny embody ideas, ideals, and \bibemph{experience} which apparently pertain to a future universe age and are therefore of no immediate practical value to either the super\hyp{} or central universe administrations. These unique sons of the children of time and the citizens of eternity are all held in reserve on Vicegerington, where they are engaged in the study of the concepts of time and the realities of eternity in a special sector of the sphere occupied by the secret colleges of the corps of the Creator Sons.
\vs p022 7:11 \pc The Supreme Being is the unification of three phases of Deity reality: God the Supreme, the spiritual unification of certain finite aspects of the Paradise Trinity; the Almighty Supreme, the power unification of the grand universe Creators; and the Supreme Mind, the individual contribution of the Third Source and Centre and his co\hyp{}ordinates to the reality of the Supreme Being. In their trinitization adventures the superb creatures of the central universe and Paradise are engaged in a threefold exploration of the Deity of the Supreme which results in the production of three orders of creature\hyp{}trinitized sons:
\vs p022 7:12 \ublistelem{1.}\bibnobreakspace \bibemph{Ascender\hyp{}trinitized Sons.} In their creative efforts the finaliters are attempting to trinitize certain conceptual realities of the Almighty Supreme which they have experientially acquired in their ascension through time and space to Paradise.
\vs p022 7:13 \ublistelem{2.}\bibnobreakspace \bibemph{Paradise\hyp{}Havona\hyp{}trinitized Sons.} The creative efforts of the Paradise Citizens and the Havoners result in the trinitization of certain high spiritual aspects of the Supreme Being which they have experientially acquired on a supersupreme background bordering on the Ultimate and the Eternal.
\vs p022 7:14 \ublistelem{3.}\bibnobreakspace \bibemph{Trinitized Sons of Destiny.} But when a finaliter and a Paradise\hyp{}Havoner together trinitize a new creature, this conjoint effort repercusses in certain phases of the Supreme\hyp{}Ultimate Mind. The resulting creature\hyp{}trinitized sons are supercreational; they represent actualities of Supreme\hyp{}Ultimate Deity which have not been otherwise experientially attained, and which, therefore, automatically fall within the province of the Architects of the Master Universe, custodians of those things which transcend the creational limits of the present universe age. The trinitized sons of destiny embody certain aspects of the unrevealed master universe function of the Supreme\hyp{}Ultimate. We do not know a great deal about these conjoint children of time and eternity, but we know much more than we are permitted to reveal.
\usection{8.\bibnobreakspace The Creature\hyp{}Trinitized Sons}
\vs p022 8:1 In addition to the creature\hyp{}trinitized sons considered in this narrative, there are numerous unrevealed orders of creature\hyp{}trinitized beings --- the diverse progeny of the multiple liaisons of seven finaliter corps and Paradise\hyp{}Havona personalities. But all these creature\hyp{}trinitized beings, revealed and unrevealed, are endowed with personality by the Universal Father.
\vs p022 8:2 \pc When new ascender\hyp{}trinitized and Paradise\hyp{}Havona\hyp{}trinitized sons are young and untrained, they are usually dispatched for long periods of service on the seven Paradise spheres of the Infinite Spirit, where they serve under the tutelage of the Seven Supreme Executives. Subsequently they may be adopted for further training in the local universes by the Trinity Teacher Sons.
\vs p022 8:3 These adopted sons of high and glorified creature origin are the apprentices, student helpers, of the Teacher Sons, and as regards classification they are often temporarily numbered with these Sons. They may and do execute many noble assignments in self\hyp{}denial in behalf of their chosen realms of service.
\vs p022 8:4 The Teacher Sons in the local universes may nominate their creature\hyp{}trinitized wards for embrace by the Paradise Trinity. Emerging from this embrace as Trinitized Sons of Perfection, they enter the service of the Ancients of Days in the seven superuniverses, that being the present known destiny of this unique group of twice\hyp{}trinitized beings.
\vs p022 8:5 Not all creature\hyp{}trinitized sons are Trinity embraced; many become the associates and ambassadors of the Seven Master Spirits of Paradise, of the Reflective Spirits of the superuniverses, and of the Mother Spirits of the local creations. Others may accept special assignments on the eternal Isle. Still others may enter the special services on the secret worlds of the Father and on the Paradise spheres of the Spirit. Eventually many find their way into the conjoint corps of the Trinitized Sons on the inner circuit of Havona.
\vs p022 8:6 Excepting the Trinitized Sons of Perfection and those who are forgathering on Vicegerington, the supreme destiny of all creature\hyp{}trinitized sons appears to be entrance into the Corps of Trinitized Finaliters, one of the seven Paradise Corps of the Finality.
\usection{9.\bibnobreakspace The Celestial Guardians}
\vs p022 9:1 Creature\hyp{}trinitized sons are embraced by the Paradise Trinity in classes of 7,000. These trinitized offspring of perfected humans and of Paradise\hyp{}Havona personalities are all equally embraced by the Deities, but they are assigned to the superuniverses in accordance with the advice of their former instructors, the Trinity Teacher Sons. Those of more acceptable service are commissioned High Son Assistants; those of less distinguished performance are designated Celestial Guardians.
\vs p022 9:2 When these unique beings have been Trinity embraced, they become valuable adjuncts to the superuniverse governments. They are versed in the affairs of the ascendant career, not by personal ascension, but as a result of their service with the Trinity Teacher Sons on the worlds of space.
\vs p022 9:3 \pc Almost one billion Celestial Guardians have been commissioned in Orvonton. They are chiefly assigned to the administrations of the Perfections of Days on the headquarters of the major sectors and are ably assisted by a corps of ascendant Son\hyp{}fused mortals.
\vs p022 9:4 The Celestial Guardians are the officers of the courts of the Ancients of Days, functioning as court messengers and as bearers of the summonses and decisions of the various tribunals of the superuniverse governments. They are the apprehending agents of the Ancients of Days; they go forth from Uversa to bring back beings whose presence is required before the superuniverse judges; they execute the mandates for the detention of any personality in the superuniverse. They also accompany Spirit\hyp{}fused mortals of the local universes when, for any reason, their presence is required on Uversa.
\vs p022 9:5 \pc The Celestial Guardians and their associates, the High Son Assistants, have never been indwelt by Adjusters. Neither are they Spirit nor Son fused. The embrace of the Paradise Trinity does, however, compensate for the nonfused status of the Trinitized Sons of Perfection. The Trinity embrace may act solely upon the idea which is personified in a creature\hyp{}trinitized son, leaving the embraced son otherwise unchanged, but such a limitation occurs only when so planned.
\vs p022 9:6 These twice\hyp{}trinitized sons are marvellous beings, but they are neither as versatile nor dependable as their ascendant associates; they lack that tremendous and profound personal experience which the rest of the sons belonging to this group have acquired by actually climbing up to glory from the dark domains of space. We of the ascendant career love them and do all in our power to compensate their deficiencies, but they make us ever grateful for our lowly origin and our capacity for experience. Their willingness to recognize and acknowledge their deficiencies in the experiencible realities of universe ascension is transcendently beautiful and sometimes most touchingly pathetic.
\vs p022 9:7 Trinitized Sons of Perfection are limited in contrast to other Trinity\hyp{}embraced sons because their experiential capacity is time\hyp{}space inhibited. They are experience\hyp{}deficient, despite long training with the Supreme Executives and the Teacher Sons, and if this were not the case, experiential saturation would preclude their being left in reserve for acquiring experience in a future universe age. There is simply nothing in all universal existence which can take the place of actual personal experience, and these creature\hyp{}trinitized sons are held in reserve for experiential function in some future universe epoch.
\vs p022 9:8 On the mansion worlds I have often seen these dignified officers of the high courts of the superuniverse look so longingly and appealingly at even the recent arrivals from the evolutionary worlds of space that one could not help realizing that these possessors of nonexperiential trinitization really envied their supposedly less fortunate brethren who ascend the universal path by steps of bona fide experience and actual living. Notwithstanding their handicaps and limitations they are a wonderfully useful and ever\hyp{}willing corps of workers when it comes to the execution of the complex administrative plans of the superuniverse governments.
\usection{10.\bibnobreakspace High Son Assistants}
\vs p022 10:1 The High Son Assistants are the superior group of the retrinitized trinitized sons of glorified ascendant beings of the Mortal Corps of the Finality and of their eternal associates, the Paradise\hyp{}Havona personalities. They are assigned to the superuniverse service and function as personal aids to the high sons of the governments of the Ancients of Days. They might fittingly be denominated private secretaries. They act, from time to time, as clerks for special commissions and other group associations of the high sons. They serve Perfectors of Wisdom, Divine Counsellors, Universal Censors, Mighty Messengers, Those High in Authority, and Those without Name and Number.
\vs p022 10:2 \pc If, in discussing the Celestial Guardians, I have seemed to call attention to the limitations and handicaps of these twice\hyp{}trinitized sons, let me now, in all fairness, call attention to their one point of great strength, the attribute which makes them almost invaluable to us. These beings owe their very existence to the fact that they are the personification of a single and supreme concept. They are the personality embodiment of some divine idea, some universal ideal, as it has never before been conceived, expressed, or trinitized. And they have subsequently been Trinity embraced; thus they show forth and actually embody the very wisdom of the divine Trinity as concerns the idea\hyp{}ideal of their personality existence. As far as that particular concept is revealable to the universes, these personalities embody all of everything that any creature or Creator intelligence could possibly conceive, express, or exemplify. \bibemph{They are that idea personified.}
\vs p022 10:3 Can you not see that such living concentrations of a single supreme concept of universe reality would be of untold service to those who are entrusted with the administration of the superuniverses?
\vs p022 10:4 \pc Not long since I was directed to head a commission of six --- one of each of the high sons --- assigned to the study of three problems pertaining to a group of new universes in the south parts of Orvonton. I was made acutely aware of the value of the High Son Assistants when I made requisition on the chief of their order on Uversa for temporary assignment of such secretaries to my commission. The first of our ideas was represented by a High Son Assistant on Uversa, who was forthwith attached to our group. Our second problem was embodied in a High Son Assistant assigned to superuniverse number three. We secured much help from this source through the central universe clearinghouse for the co\hyp{}ordination and dissemination of essential knowledge, but nothing comparable to the assistance afforded by the actual presence of a personality who \bibemph{is} a concept creature\hyp{}trinitized in supremacy and Deity\hyp{}trinitized in finality. Concerning our third problem, the records of Paradise disclosed that such an idea had never been creature trinitized.
\vs p022 10:5 \pc High Son Assistants are unique and original personalizations of tremendous concepts and stupendous ideals. And as such they are able to impart inexpressible illumination to our deliberations from time to time. When I am acting on some remote assignment out in the universes of space, think what it means, by way of assistance, if I am so fortunate as to have attached to my mission a High Son Assistant who is the fullness of divine concept regarding the very problem I have been sent to attack and solve; and I have repeatedly had this very experience. The only difficulty with this plan is that no superuniverse can have a complete edition of these trinitized ideas; we only get \bibfrac{1}{7} of these beings; so it is only about one time in seven that we enjoy the personal association of these beings even when the records indicate that the idea has been trinitized.
\vs p022 10:6 We could use to great advantage much larger numbers of these beings on Uversa. Because of their value to the superuniverse administrations, we, in every way possible, encourage the pilgrims of space and also the residents of Paradise to attempt trinitization after they have contributed to one another those experiential realities which are essential to the enactment of such creative adventures.
\vs p022 10:7 \pc We now have in our superuniverse about $1.25 \times 10^6$ High Son Assistants, and they serve on both the major and minor sectors, even as they function on Uversa. They very often accompany us on our assignments to the remote universes. High Son Assistants are not permanently assigned to any Son or to any commission. They are in constant circulation, serving where the idea or ideal which they \bibemph{are} can best further the eternal purposes of the Paradise Trinity, whose sons they have become.
\vs p022 10:8 They are touchingly affectionate, superbly loyal, exquisitely intelligent, supremely wise --- regarding a single idea --- and transcendently humble. While they can impart to you the lore of the universe concerning their one idea or ideal, it is well\hyp{}nigh pathetic to observe them seeking knowledge and information on hosts of other subjects, even from the ascending mortals.
\vs p022 10:9 \pc And this is the narrative of the origin, nature, and functioning of certain of those who are called the Trinitized Sons of God, more particularly of those who have passed through the divine embrace of the Paradise Trinity, and who have then been assigned to the services of the superuniverses, there to give wise and understanding co\hyp{}operation with the administrators of the Ancients of Days in their untiring efforts to facilitate the inward progress of the ascending mortals of time toward their immediate Havona destination and their eventual Paradise goal.
\vsetoff
\vs p022 10:10 [Narrated by a Mighty Messenger of the revelatory corps of Orvonton.]
\quizlink
