\upaper{29}{The Universe Power Directors}
\uminitoc{The Seven Supreme Power Directors}
\uminitoc{The Supreme Power Centres}
\uminitoc{The Domain of Power Centres}
\uminitoc{The Master Physical Controllers}
\uminitoc{The Master Force Organizers}
\author{Universal Censor}
\vs p029 0:1 Of all the universe personalities concerned in the regulation of interplanetary and interuniverse affairs, the power directors and their associates have been the least understood on Urantia. While your races have long known of the existence of angels and similar orders of celestial beings, little information concerning the controllers and regulators of the physical domain has ever been imparted. Even now I am permitted fully to disclose only the last of the following three groups of living beings having to do with force control and energy regulation in the master universe:
\vs p029 0:2 \ublistelem{1.}\bibnobreakspace Primary Eventuated Master Force Organizers.
\vs p029 0:3 \ublistelem{2.}\bibnobreakspace Associate Transcendental Master Force Organizers.
\vs p029 0:4 \ublistelem{3.}\bibnobreakspace Universe Power Directors.
\vs p029 0:5 \pc Though I deem it impossible to portray the individuality of the various groups of directors, centres, and controllers of universe power, I hope to be able to explain something about the domain of their activities. They are a unique group of living beings having to do with the intelligent regulation of energy throughout the grand universe. Including the supreme directors, they embrace the following major divisions:
\vs p029 0:6 \ublistelem{1.}\bibnobreakspace The Seven Supreme Power Directors.
\vs p029 0:7 \ublistelem{2.}\bibnobreakspace The Supreme Power Centres.
\vs p029 0:8 \ublistelem{3.}\bibnobreakspace The Master Physical Controllers.
\vs p029 0:9 \ublistelem{4.}\bibnobreakspace The Morontia Power Supervisors.
\vs p029 0:10 \pc The Supreme Power Directors and Centres have existed from the near times of eternity, and as far as we know, no more beings of these orders have been created. The Seven Supreme Directors were personalized by the Seven Master Spirits, and then they collaborated with their parents in the production of more than ten billion associates. Before the days of the power directors the energy circuits of space outside of the central universe were under the intelligent supervision of the Master Force Organizers of Paradise.
\vs p029 0:11 Having knowledge about material creatures, you have at least a contrastive conception of spiritual beings; but it is very difficult for the mortal mind to envisage the power directors. In the scheme of ascendant progression to higher levels of existence you have nothing directly to do with either the supreme directors or the power centres. On certain rare occasions you will have dealings with the physical controllers, and you will work freely with the supervisors of morontia power upon reaching the mansion worlds. These Morontia Power Supervisors function so exclusively in the morontia regime of the local creations that it is deemed best to narrate their activities in the section dealing with the local universe.
\usection{The Seven Supreme Power Directors}
\vs p029 1:1 The Seven Supreme Power Directors are the physical\hyp{}energy regulators of the grand universe. Their creation by the Seven Master Spirits is the first recorded instance of the derivation of semimaterial progeny from true spirit ancestry. When the Seven Master Spirits create individually, they bring forth highly spiritual personalities on the angelic order; when they create collectively, they sometimes produce these high types of semimaterial beings. But even these quasi\hyp{}physical beings would be invisible to the short\hyp{}range vision of Urantia mortals.
\vs p029 1:2 The Supreme Power Directors are seven in number, and they are identical in appearance and function. One cannot be distinguished from another except by that Master Spirit with whom each is in immediate association, and to whom each is in complete functional subservience. Each of the Master Spirits is thus in eternal union with one of their collective offspring. The same director is always in association with the same Spirit, and their working partnership results in a unique association of physical and spiritual energies, of a semiphysical being and a spirit personality.
\vs p029 1:3 The Seven Supreme Power Directors are stationed on peripheral Paradise, where their slowly circulating presences indicate the whereabouts of the force\hyp{}focal headquarters of the Master Spirits. These power directors function singly in the power\hyp{}energy regulation of the superuniverses but collectively in the administration of the central creation. They operate from Paradise but maintain themselves as effective power centres in all divisions of the grand universe.
\vs p029 1:4 These mighty beings are the physical ancestors of the vast host of the power centres and, through them, of the physical controllers scattered throughout the seven superuniverses. Such subordinate physical\hyp{}control organisms are basically uniform, identical except for the differential toning of each superuniverse corps. In order to change in superuniverse service, they would merely have to return to Paradise for retoning. The physical creation is fundamentally uniform in administration.
\usection{The Supreme Power Centres}
\vs p029 2:1 The Seven Supreme Power Directors are not able, individually, to reproduce themselves, but collectively, and in association with the Seven Master Spirits, they can and do reproduce --- create --- other beings like themselves. Such is the origin of the Supreme Power Centres of the grand universe, who function in the following seven groups:
\vs p029 2:2 \ublistelem{1.}\bibnobreakspace Supreme Centre Supervisors.
\vs p029 2:3 \ublistelem{2.}\bibnobreakspace Havona Centres.
\vs p029 2:4 \ublistelem{3.}\bibnobreakspace Superuniverse Centres.
\vs p029 2:5 \ublistelem{4.}\bibnobreakspace Local Universe Centres.
\vs p029 2:6 \ublistelem{5.}\bibnobreakspace Constellation Centres.
\vs p029 2:7 \ublistelem{6.}\bibnobreakspace System Centres.
\vs p029 2:8 \ublistelem{7.}\bibnobreakspace Unclassified Centres.
\vs p029 2:9 \pc These power centres together with the Supreme Power Directors are beings of high will freedom and action. They are all endowed with Third\hyp{}Source personality and disclose unquestioned volitional capacity of a high order. These directing centres of the universe power system are the possessors of exquisite intelligence endowment; they are the intellect of the power system of the grand universe and the secret of the technique of the mind control of all the vast network of the far\hyp{}flung functions of the Master Physical Controllers and the Morontia Power Supervisors.
\vs p029 2:10 \ublistelem{1.}\bibnobreakspace \bibemph{Supreme Centre Supervisors.} These seven co\hyp{}ordinates and associates of the Supreme Power Directors are the regulators of the master energy circuits of the grand universe. Each centre supervisor is headquartered on one of the special worlds of the Seven Supreme Executives, and they work in close association with these co\hyp{}ordinators of general universe affairs.
\vs p029 2:11 The Supreme Power Directors and the Supreme Centre Supervisors function both as individuals and conjointly with regard to all cosmic phenomena below the levels of “gravity energy.” When acting in liaison, these 14 beings are to universe power what the Seven Supreme Executives are to general universe affairs, and what the Seven Master Spirits are to the cosmic mind.
\vs p029 2:12 \ublistelem{2.}\bibnobreakspace \bibemph{Havona Centres.} Before the creation of the universes of time and space, power centres were not required in Havona, but ever since these far\hyp{}distant times, 1,000,000 have functioned in the central creation, each centre having the supervision of 1,000 Havona worlds. Here in the divine universe there is perfection of energy control, a condition not existing elsewhere. Perfection of energy regulation is the ultimate goal of all the power centres and the physical controllers of space.
\vs p029 2:13 \ublistelem{3.}\bibnobreakspace \bibemph{Superuniverse Centres.} Occupying an enormous area on the capital sphere of each of the seven superuniverses are 1,000 power centres of the 3\ts{rd} order. Three currents of primary energy of 10 segregations each come in to these power centres, but seven specialized and well\hyp{}directed, though imperfectly controlled, circuits of power go forth from their seat of united action. This is the electronic organization of universe power.
\vs p029 2:14 All energy is circuited in the Paradise cycle, but the Universe Power Directors \bibemph{direct} the force\hyp{}energies of nether Paradise as they find them modified in the space functions of the central and superuniverses, converting and directing these energies into channels of useful and constructive application. There is a difference between Havona energy and the energies of the superuniverses. The power charge of a superuniverse consists of 3 phases of energy of 10 segregations each. This threefold energy charge spreads throughout the space of the grand universe; it is like a vast moving ocean of energy which engulfs and bathes the whole of each of the seven supercreations.
\vs p029 2:15 The electronic organization of universe power functions in seven phases and discloses varying response to local or linear gravity. This sevenfold circuit proceeds from the superuniverse power centres and pervades each supercreation. Such specialized currents of time and space are definite and localized energy movements initiated and directed for specific purposes, much as the Gulf Stream functions as a circumscribed phenomenon in the midst of the Atlantic Ocean.
\vs p029 2:16 \ublistelem{4.}\bibnobreakspace \bibemph{Local Universe Centres.} On the headquarters of each local universe are stationed 100 power centres of the 4\ts{th} order. They function to downstep and otherwise to modify the seven power circuits emanating from superuniverse headquarters, thus making them applicable to the services of the constellations and systems. The local astronomical catastrophes of space are of passing concern to these power centres; they are engaged in the orderly dispatch of effective energy to the subsidiary constellations and systems. They are of great assistance to the Creator Sons during the later times of universe organization and energy mobilization. These centres are able to provide intensified lanes of energy useful for interplanetary communication between important inhabited points. Such a \bibemph{lane} or \bibemph{line} of energy, sometimes also called an energy path, is a direct circuit of energy from one power centre to another power centre or from one physical controller to another controller. It is an individualized stream of power and stands in contrast to the free space movements of undifferentiated energy.
\vs p029 2:17 \ublistelem{5.}\bibnobreakspace \bibemph{Constellation Centres.} Ten of these living power centres are stationed in each constellation, functioning as energy projectors to the 100 tributary local systems. From these beings there go forth the power lines for communication and transport and for the energizing of those living creatures who are dependent upon certain forms of physical energy for the maintenance of life. But neither the power centres nor the subordinate physical controllers are otherwise concerned with life as a functional organization.
\vs p029 2:18 \ublistelem{6.}\bibnobreakspace \bibemph{System Centres.} One Supreme Power Centre is permanently assigned to each local system. These system centres dispatch the power circuits to the inhabited worlds of time and space. They co\hyp{}ordinate the activities of the subordinate physical controllers and otherwise function to ensure the satisfactory distribution of power in the local system. The circuit relay between the planets depends upon the perfect co\hyp{}ordination of certain material energies and upon the efficient regulation of physical power.
\vs p029 2:19 \ublistelem{7.}\bibnobreakspace \bibemph{Unclassified Centres.} These are the centres who function in special local situations but not on the inhabited planets. The individual worlds are in the charge of Master Physical Controllers and receive the encircuited lines of power dispatched by the power centre of their system. Only those spheres of the most extraordinary energy relationships have power centres of the 7\ts{th} order acting as universe balance wheels or energy governors. In every phase of activity these power centres are the full equals of those who function on the higher units of control, but not 1 space body in 1,000,000 harbours such a living power organization.
\usection{The Domain of Power Centres}
\vs p029 3:1 The Supreme Power Centres distributed throughout the superuniverses number, with their associates and subordinates, upward of ten billion. And they are all in perfect synchrony and complete liaison with their Paradise progenitors, the Seven Supreme Power Directors. The power control of the grand universe is thus entrusted to the keeping and direction of the Seven Master Spirits, the creators of the Seven Supreme Power Directors.
\vs p029 3:2 The Supreme Power Directors and all their associates, assistants, and subordinates are forever exempt from apprehension or interference by all the tribunals of all space; neither are they subject to the administrative direction either of the superuniverse government of the Ancients of Days or of the local universe administration of the Creator Sons.
\vs p029 3:3 These power centres and directors are brought into being by the children of the Infinite Spirit. They are not germane to the administration of the Sons of God, though they affiliate with the Creator Sons during the later epochs of universe material organization. But the power centres are in some way closely associated with the cosmic overcontrol of the Supreme Being.
\vs p029 3:4 \pc Power centres and physical controllers undergo no training; they are all created in perfection and are inherently perfect in action. Never do they pass from one function to another; always do they serve as originally assigned. There is no evolution in their ranks, and this is true of all seven divisions of both orders.
\vs p029 3:5 Having no ascendant past to revert to in memory, power centres and physical controllers never play; they are thoroughly businesslike in all their actions. They are always on duty; there is no provision in the universal scheme for the interruption of the physical lines of energy; never for a fraction of a second can these beings relinquish their direct supervision of the energy circuits of time and space.
\vs p029 3:6 \pc The directors, centres, and controllers of power have nothing to do with anything in all creation except power, material or semiphysical energy; they do not originate it, but they do modify, manipulate, and directionize it. Neither do they have anything whatever to do with physical gravity except to resist its drawing power. Their relation to gravity is wholly negative.
\vs p029 3:7 The power centres utilize vast mechanisms and co\hyp{}ordinations of a material order in liaison with the living mechanisms of the various segregated energy concentrations. Each individual power centre is constituted in exactly 1,000,000 units of functional control, and these energy\hyp{}modifying units are not stationary as are the vital organs of man’s physical body; these “vital organs” of power regulation are mobile and truly kaleidoscopic in associative possibilities.
\vs p029 3:8 It is utterly beyond my ability to explain the manner in which these living beings encompass the manipulation and regulation of the master circuits of universe energy. To undertake to inform you further concerning the size and function of these gigantic and almost perfectly efficient power centres, would only add to your confusion and consternation. They are both living and “personal,” but they are beyond your comprehension.
\vs p029 3:9 \pc Outside of Havona the Supreme Power Centres function only on especially constructed (architectural) spheres or on otherwise suitably constituted space bodies. The architectural worlds are so constructed that the living power centres can act as selective switches to directionize, modify, and concentrate the energies of space as they pour over these spheres. They could not so function on an ordinary evolutionary sun or planet. Certain groups are also concerned in the heating and other material necessities of these special headquarters worlds. And though it is beyond the scope of Urantia knowledge, I may state that these orders of living power personalities have much to do with the distribution of the light that shines without heat. They do not produce this phenomenon, but they are concerned with its dissemination and directionization.
\vs p029 3:10 \pc The power centres and their subordinate controllers are assigned to the working of all of the physical energies of organized space. They work with the three basic currents of ten energies each. That is the energy charge of organized space; and organized space is their domain. The Universe Power Directors have nothing whatever to do with those tremendous actions of force which are now taking place outside the present boundaries of the seven superuniverses.
\vs p029 3:11 The power centres and controllers exert perfect control over only seven of the ten forms of energy contained in each basic universe current; those forms which are partly or wholly exempt from their control must represent the unpredictable realms of energy manifestation dominated by the Unqualified Absolute. If they exert an influence upon the primordial forces of this Absolute, we are not cognizant of such functions, though there is some slight evidence which would warrant the opinion that certain of the physical controllers are sometimes automatically reactive to certain impulses of the Universal Absolute.
\vs p029 3:12 These living power mechanisms are not consciously related to the master universe energy overcontrol of the Unqualified Absolute, but we surmise that their entire and almost perfect scheme of power direction is in some unknown manner subordinated to this supergravity presence. In any local energy situation the centres and controllers exert near\hyp{}supremacy, but they are always conscious of the superenergy presence and the unrecognizable performance of the Unqualified Absolute.
\usection{The Master Physical Controllers}
\vs p029 4:1 These beings are the mobile subordinates of the Supreme Power Centres. The physical controllers are endowed with capabilities of individuality metamorphosis of such a nature that they can engage in a remarkable variety of autotransport, being able to traverse local space at velocities approaching the flight of Solitary Messengers. But like all other space traversers they require the assistance of both their fellows and certain other types of beings in overcoming the action of gravity and the resistance of inertia in departing from a material sphere.
\vs p029 4:2 The Master Physical Controllers serve\tunemarkup{pgauraone}{\linebreak}\tunemarkup{pgkoboaurahd}{\linebreak} throughout the grand universe. They are directly governed from Paradise by the Seven Supreme Power Directors as far as the headquarters of the superuniverses; from here they are directed and distributed by the Council of Equilibrium, the high commissioners of power dispatched by the Seven Master Spirits from the personnel of the Associate Master Force Organizers. These high commissioners are empowered to interpret the readings and registrations of the master frandalanks, those living instruments which indicate the power pressure and the energy charge of an entire superuniverse.
\vs p029 4:3 While the presence of the Paradise Deities encircles the grand universe and sweeps around the circle of eternity, the influence of any one of the Seven Master Spirits is limited to a single superuniverse. There is a distinct segregation of energy and a separation of the circuits of power between each of the seven supercreations; hence individualized control methods must and do prevail.
\vs p029 4:4 \pc The Master Physical Controllers are the direct offspring of the Supreme Power Centres, and their numbers include the following:
\vs p029 4:5 \ublistelem{1.}\bibnobreakspace Associate Power Directors.
\vs p029 4:6 \ublistelem{2.}\bibnobreakspace Mechanical Controllers.
\vs p029 4:7 \ublistelem{3.}\bibnobreakspace Energy Transformers.
\vs p029 4:8 \ublistelem{4.}\bibnobreakspace Energy Transmitters.
\vs p029 4:9 \ublistelem{5.}\bibnobreakspace Primary Associators.
\vs p029 4:10 \ublistelem{6.}\bibnobreakspace Secondary Dissociators.
\vs p029 4:11 \ublistelem{7.}\bibnobreakspace The Frandalanks and Chronoldeks.
\vs p029 4:12 \pc Not all of these orders are persons in the sense of possessing individual powers of choice. Especially do the last four seem to be wholly automatic and mechanical in response to the impulses of their superiors and in reaction to existing energy conditions. But though such response appears wholly mechanistic, it is not; they may seem to be automatons, but all of them disclose the differential function of intelligence.
\vs p029 4:13 Personality is not necessarily a concomitant of mind. Mind can think even when deprived of all power of choice, as in numerous of the lower types of animals and in certain of these subordinate physical controllers. Many of these more automatic regulators of physical power are not persons in any sense of the term. They are not endowed with will and independence of decision, being wholly subservient to the mechanical perfection of design for the tasks of their allotment. Nonetheless all of them are highly intelligent beings.
\vs p029 4:14 The physical controllers are chiefly occupied in the adjustment of basic energies undiscovered on Urantia. These unknown energies are very essential to the interplanetary system of transport and to certain techniques of communication. When we lay lines of energy for the purpose of conveying sound equivalents or of extending vision, these undiscovered forms of energy are utilized by the living physical controllers and their associates. These same energies are also, on occasion, used by the midway creatures in their routine work.
\vs p029 4:15 \ublistelem{1.}\bibnobreakspace \bibemph{Associate Power Directors.} These marvellously efficient beings are entrusted with the assignment and dispatch of all orders of the Master Physical Controllers in accordance with the ever\hyp{}shifting needs of the constantly changing energy status of the realms. The vast reserves of the physical controllers are maintained on the headquarters worlds of the minor sectors, and from these concentration points they are periodically dispatched by the associate power directors to the headquarters of the universes, constellations, and systems, and to the individual planets. When thus assigned, the physical controllers are provisionally subject to the orders of the divine executioners of the conciliating commissions but are otherwise solely amenable to their associate directors and to the Supreme Power Centres.
\vs p029 4:16 3,000,000 associate power directors are assigned to each of the Orvonton minor sectors, making a total of 3,000,000,000 as the superuniverse quota of these amazingly versatile beings. Their own reserves are maintained on these same minor sector worlds, where they also serve as instructors of all who study the sciences of the techniques of intelligent energy control and transmutation.
\vs p029 4:17 These directors alternate periods of executive service in the minor sectors with equal periods of inspection service to the realms of space. At least one acting inspector is always present in each local system, maintaining headquarters on its capital sphere. They keep the whole vast living energy aggregation in harmonious synchrony.
\vs p029 4:18 \ublistelem{2.}\bibnobreakspace \bibemph{Mechanical Controllers.} These are the exceedingly versatile and mobile assistants of the associate power directors. Trillions upon trillions of them are commissioned in Ensa, your minor sector. These beings are called mechanical controllers because they are so completely dominated by their superiors, so fully subservient to the will of the associate power directors. Nevertheless they are, themselves, very intelligent, and their work, though mechanical and matter\hyp{}of\hyp{}fact in nature, is skillfully performed.
\vs p029 4:19 Of all the Master Physical Controllers assigned to the inhabited worlds, the mechanical controllers are by far the most powerful. Possessing the living endowment of antigravity in excess of all other beings, each controller has a gravity resistance equalled only by enormous spheres revolving at tremendous velocity. 10 of these controllers are now stationed on Urantia, and one of their most important planetary activities is to facilitate the departure of seraphic transports. In so functioning, all 10 of the mechanical controllers act in unison while a battery of 1,000 energy transmitters provides the initial momentum for the seraphic departure.
\vs p029 4:20 The mechanical controllers are competent to directionize the flow of energy and to facilitate its concentration into the specialized currents or circuits. These mighty beings have much to do with the segregation, directionization, and intensification of the physical energies and with the equalization of the pressures of the interplanetary circuits. They are expert in the manipulation of 21 of the 30 physical energies of space, constituting the power charge of a superuniverse. They are also able to accomplish much towards the management and control of six of the nine more subtle forms of physical energy. By placing these controllers in proper technical relationship to each other and to certain of the power centres, the associate power directors are enabled to effect unbelievable changes in power adjustment and energy control.
\vs p029 4:21 The Master Physical Controllers often function in batteries of hundreds, thousands, and even millions and by varying their positions and formations are able to effect energy control in a collective as well as an individual capacity. As requirements vary, they can upstep and accelerate the energy volume and movement or detain, condense, and retard the energy currents. They influence energy and power transformations somewhat as so\hyp{}called catalytic agents augment chemical reactions. They function by inherent ability and in co\hyp{}operation with the Supreme Power Centres.
\vs p029 4:22 \ublistelem{3.}\bibnobreakspace \bibemph{Energy Transformers.} The number of these beings in a superuniverse is unbelievable. There are almost 1,000,000 in Satania alone, and the usual quota is 100 for each inhabited world.
\vs p029 4:23 The energy transformers are the conjoint creation of the Seven Supreme Power Directors and the Seven Centre\fnst{In 1955 text ``Central'', but there is no mention of the Seven \bibemph{Central} Supervisors anywhere in the text and there are multiple references to the Seven \bibemph{Centre} Supervisors; the primary description of them is given at \bibref[29:2.10]{p029 2:10}.} Supervisors. They are among the more personal orders of physical controllers, and except when an associate power director is present on an inhabited world, the transformers are in command. They are the planetary inspectors of all departing seraphic transports. All classes of celestial life can utilize the less personal orders of the physical controllers only by liaison with the more personal orders of the associate directors and the energy transformers.
\vs p029 4:24 These transformers are powerful and effective living\tunemarkup{pgnexus10}{\linebreak} switches, being able to dispose themselves for or against a given power disposition or directionization. They are also skillful in their efforts to insulate the planets against the powerful energy streams passing between gigantic planetary and starry neighbours. Their energy\hyp{}transmutive attributes render them most serviceable in the important task of maintaining universal energy balance, or power equilibrium. At one time they seem to consume or store energy; at other times they appear to exude or liberate energy. The transformers are able to increase or to diminish the “storage\hyp{}battery” potential of the living and dead energies of their respective realms. But they deal only with physical and semimaterial energies, they do not directly function in the domain of life, neither do they change the forms of living beings.
\vs p029 4:25 In some respects the energy transformers are the most remarkable and mysterious of all semimaterial living creatures. They are in some unknown manner physically differentiated, and by varying their liaison relationships, they are able to exert a profound influence upon the energy which passes through their associated presences. The status of the physical realms seems to undergo a transformation under their skillful manipulation. \bibemph{They can and do change the physical form of the energies of space.} With the aid of their fellow controllers they are actually able to change the form and potential of 27 of the 30 physical energies of the superuniverse power charge. That three of these energies are beyond their control proves that they are not instrumentalities of the Unqualified Absolute.
\vs p029 4:26 \pc The remaining four groups of the Master Physical Controllers are hardly persons within any acceptable definition of that word. These transmitters, associators, dissociators, and frandalanks are wholly automatic in their reactions; nevertheless they are in every sense intelligent. We are greatly limited in our knowledge of these wonderful entities because we cannot communicate with them. They appear to understand the language of the realm, but they cannot communicate with us. They seem fully able to receive our communications but quite powerless to make response.
\vs p029 4:27 \ublistelem{4.}\bibnobreakspace \bibemph{Energy Transmitters.} These beings function chiefly, but not wholly, in an intraplanetary capacity. They are marvellous dispatchers of energy as it is manifested on the individual worlds.
\vs p029 4:28 When energy is to be diverted to a new circuit, the transmitters deploy themselves in a line along the desired energy path, and by virtue of their unique attributes of energy\hyp{}attraction, they can actually induce an increased energy flow in the desired direction. This they do just as literally as certain metallic circuits directionize the flow of certain forms of electric energy; and they are living superconductors for more than half of the 30 forms of physical energy.
\vs p029 4:29 Transmitters form skillful liaisons which are effective in rehabilitating the weakening currents of specialized energy passing from planet to planet and from station to station on an individual planet. They can detect currents which are much too feeble to be recognized by any other type of living being, and they can so augment these energies that the accompanying message becomes perfectly intelligible. Their services are invaluable to the broadcast receivers.
\vs p029 4:30 Energy transmitters can function with regard to all forms of communicable perception; they can render a distant scene “visible” as well as a distant sound “audible.” They provide the emergency lines of communication in the local systems and on the individual planets. These services must be used by practically all creatures for purposes of communication outside of the regularly established circuits.
\vs p029 4:31 These beings, together with the energy transformers, are indispensable to the maintenance of mortal existence on those worlds having an impoverished atmosphere, and they are an integral part of the technique of life on the nonbreathing planets.
\vs p029 4:32 \ublistelem{5.}\bibnobreakspace \bibemph{Primary Associators.} These interesting and invaluable entities are masterly energy conservators and custodians. Somewhat as a plant stores solar light, so do these living organisms store energy during times of plus manifestations. They work on a gigantic scale, converting the energies of space into a physical state not known on Urantia. They are also able to carry forward these transformations to the point of producing some of the primitive units of material existence. These beings simply act by their presence. They are in no way exhausted or depleted by this function; they act like living catalytic agents.
\vs p029 4:33 During seasons of minus manifestations they are empowered to release these accumulated energies. But your knowledge of energy and matter is not sufficiently advanced to make it possible to explain the technique of this phase of their work. They always labour in compliance with universal law, handling and manipulating atoms, electrons, and ultimatons much as you manoeuvre adjustable type to make the same alphabetical symbols tell vastly different stories.
\vs p029 4:34 The associators are the first group of life to appear on an organizing material sphere, and they can function at physical temperatures which you would regard as utterly incompatible with the existence of living beings. They represent an order of life which is simply beyond the range of human imagination. Together with their co\hyp{}workers, the dissociators, they are the most slavish of all intelligent creatures.
\vs p029 4:35 \ublistelem{6.}\bibnobreakspace \bibemph{Secondary Dissociators.} Compared with the primary associators, these beings of enormous antigravity endowment are the reverse workers. There is never any danger that the special or modified forms of physical energy on the local worlds or in the local systems will be exhausted, for these living organizations are endowed with the unique power of evolving limitless supplies of energy. They are chiefly concerned with the evolution of a form of energy which is hardly known on Urantia from a form of matter which is recognized still less. They are truly the alchemists of space and the wonder\hyp{}workers of time. But in all the wonders they work, they never transgress the mandates of Cosmic Supremacy.
\vs p029 4:36 \ublistelem{7.}\bibnobreakspace \bibemph{The Frandalanks.} These beings are the joint creation of all three orders of energy\hyp{}control beings: the primary and secondary force organizers and the power directors. Frandalanks are the most numerous of all the Master Physical Controllers; the number functioning in Satania alone is beyond your numerical concept. They are stationed on all inhabited worlds and are always attached to the higher orders of physical controllers. They function interchangeably in the central and superuniverses and in the domains of outer space.
\vs p029 4:37 The frandalanks are created in 30 divisions, one for each form of basic universe force, and they function exclusively as living and automatic presence, pressure, and velocity gauges. These living barometers are solely concerned with the automatic and unerring registration of the status of all forms of force\hyp{}energy. They are to the physical universe what the vast reflectivity mechanism is to the minded universe. The frandalanks that register time in addition to quantitative and qualitative energy presence are called \bibemph{chronoldeks.}
\vs p029 4:38 I recognize that the frandalanks are intelligent, but I cannot classify them as other than living machines. About the only way I can help you to understand these living mechanisms is to compare them to your own mechanical contrivances which perform with almost intelligentlike precision and accuracy. Then if you would conceive of these beings, draw upon your imagination to the extent of recognizing that in the grand universe we actually have intelligent and \bibemph{living} mechanisms (entities) that can perform more intricate tasks involving more stupendous computations with even greater delicacy of accuracy, even with ultimacy of precision.
\usection{The Master Force Organizers}
\vs p029 5:1 The force organizers are resident on Paradise, but they function throughout the master universe, more particularly in the domains of unorganized space. These extraordinary beings are neither creators nor creatures, and they comprise two grand divisions of service:
\vs p029 5:2 \ublistelem{1.}\bibnobreakspace Primary Eventuated Master Force Organizers.
\vs p029 5:3 \ublistelem{2.}\bibnobreakspace Associate Transcendental Master Force Organizers.
\vs p029 5:4 \pc These two mighty orders of primordial\hyp{}force manipulators work exclusively under the supervision of the Architects of the Master Universe, and at the present time they do not function extensively within the boundaries of the grand universe.
\vs p029 5:5 \pc Primary Master Force Organizers are the manipulators of the primordial or basic space\hyp{}forces of the Unqualified Absolute; they are nebulae creators. They are the living instigators of the energy cyclones of space and the early organizers and directionizers of these gigantic manifestations. These force organizers transmute \bibemph{primordial force} (pre\hyp{}energy not responsive to direct Paradise gravity) into primary or \bibemph{puissant energy,} energy transmuting from the exclusive grasp of the Unqualified Absolute to the gravity grasp of the Isle of Paradise. They are thereupon succeeded by the associate force organizers, who continue the process of energy transmutation from the primary through the secondary or \bibemph{gravity\hyp{}energy} stage.
\vs p029 5:6 Upon the completion of the plans for the creation of a local universe, signalized by the arrival of a Creator Son, the Associate Master Force Organizers give way to the orders of power directors acting in the superuniverse of astronomic jurisdiction. But in the absence of such plans the associate force organizers continue on indefinitely in charge of these material creations, even as they now operate in outer space.
\vs p029 5:7 The Master Force Organizers withstand temperatures and function under physical conditions which would be intolerable even to the versatile power centres and physical controllers of Orvonton. The only other types of revealed beings capable of functioning in these realms of outer space are the Solitary Messengers and the Inspired Trinity Spirits.
\vsetoff
\vs p029 5:8 [Sponsored by a Universal Censor acting by authority of the Ancients of Days on Uversa.]
\quizlink
