\upaper{40}{The Ascending Sons of God}
\uminitoc{Evolutionary Seraphim}
\uminitoc{Ascending Material Sons}
\uminitoc{Translated Midwayers}
\uminitoc{Personalized Adjusters}
\uminitoc{Mortals of Time and Space}
\uminitoc{The Faith Sons of God}
\uminitoc{Father-Fused Mortals}
\uminitoc{Son-Fused Mortals}
\uminitoc{Spirit-Fused Mortals}
\uminitoc{Ascendant Destinies}
\author{Mighty Messenger}
\vs p040 0:1 As in many of the major groups of universe beings, seven general classes of the Ascending Sons of God have been revealed:
\vs p040 0:2 \ublistelem{1.}\bibnobreakspace Father\hyp{}fused Mortals.
\vs p040 0:3 \ublistelem{2.}\bibnobreakspace Son\hyp{}fused Mortals.
\vs p040 0:4 \ublistelem{3.}\bibnobreakspace Spirit\hyp{}fused Mortals.
\vs p040 0:5 \ublistelem{4.}\bibnobreakspace Evolutionary Seraphim.
\vs p040 0:6 \ublistelem{5.}\bibnobreakspace Ascending Material Sons.
\vs p040 0:7 \ublistelem{6.}\bibnobreakspace Translated Midwayers.
\vs p040 0:8 \ublistelem{7.}\bibnobreakspace Personalized Adjusters.
\vs p040 0:9 \pc The story of these beings, from the lowly animal\hyp{}origin mortals of the evolutionary worlds to the Personalized Adjusters of the Universal Father, presents a glorious recital of the unstinted bestowal of divine love and gracious condescension throughout all time and in all universes of the far\hyp{}flung creation of the Paradise Deities.
\vs p040 0:10 These presentations began with a description of the Deities, and group by group, the narrative has descended the universal scale of living beings until it has reached the lowest order of life endowed with the potential of immortality; and now am I dispatched from Salvington --- onetime a mortal of origin on an evolutionary world of space --- to elaborate and continue the recital of the eternal purpose of the Gods respecting the ascending orders of sonship, more particularly with regard to the mortal creatures of time and space.
\vs p040 0:11 Since the greater part of this narrative will be devoted to a discussion of the three basic orders of ascending mortals, consideration will first be given to the nonmortal ascending orders of sonship --- seraphic, Adamic, midwayer, and Adjuster.
\usection{Evolutionary Seraphim}
\vs p040 1:1 Mortal creatures of animal origin are not the only beings privileged to enjoy sonship; the angelic hosts also share the supernal opportunity to attain Paradise. Guardian seraphim, through experience and service with the ascending mortals of time, also achieve the status of ascendant sonship. Such angels attain Paradise through Seraphington, and many are even mustered into the Corps of Mortal Finality.
\vs p040 1:2 To climb to the supernal heights of finaliter sonship with God is a masterly achievement for an angel, an accomplishment far transcending your attainment of eternal survival\tunemarkup{pgnexus10}{\linebreak} through the plan of the Eternal Son and the ever\hyp{}present help of the indwelling Adjuster; but the guardian seraphim, and occasionally others, do actually effect such ascensions.
\usection{Ascending Material Sons}
\vs p040 2:1 The Material Sons of God are created in the local universe along with the Melchizedeks and their associates, who are all classified as descending Sons. And indeed, the Planetary Adams --- the Material Sons and Daughters of the evolutionary worlds --- are descending Sons, coming down to the inhabited worlds from their spheres of origin, the capitals of the local systems.
\vs p040 2:2 When such an Adam and Eve are wholly successful in their joint planetary mission as biologic uplifters, they share the destiny of the inhabitants of their world. When such a world is settled in the advanced stages of light and life, this faithful Material Son and Daughter are permitted to resign all planetary administrative duties, and after being thus liberated from the descending adventure, they are permitted to register themselves as perfected Material Sons on the records of the local universe. Likewise, when planetary assignment is long delayed, may the Material Sons of stationary status --- the citizens of the local systems --- withdraw from the activities of their status spheres and similarly register as perfected Material Sons. After these formalities such liberated Adams and Eves are accredited as ascending Sons of God and may immediately begin the long journey to Havona and Paradise, starting at the exact point of their then present status and spiritual attainment. And they make this journey in company with the mortal and other ascending Sons, continuing until they have found God and have achieved the Corps of Mortal Finality in the eternal service of the Paradise Deities.
\usection{Translated Midwayers}
\vs p040 3:1 Although deprived of the immediate benefits of the planetary bestowals of the descending Sons of God, though the Paradise ascent is long deferred, nevertheless, soon after an evolutionary planet has attained the intermediate epochs of light and life (if not before), both groups of midway creatures are released from planetary duty. Sometimes the majority of them are translated, along with their human cousins, on the day of the descent of the temple of light and the elevation of the Planetary Prince to the dignity of Planetary Sovereign. Upon being relieved of planetary service, both orders are registered in the local universe as ascending Sons of God and immediately begin the long Paradise ascent by the very routes ordained for the progression of the mortal races of the material worlds. The primary group are destined to various finaliter corps, but the secondary or Adamic midwayers are all routed for enrolment in the Mortal Corps of Finality.
\usection{Personalized Adjusters}
\vs p040 4:1 When the mortals of time fail to achieve the eternal survival of their souls in planetary association with the spirit gifts of the Universal Father, such failure is never in any way due to neglect of duty, ministry, service, or devotion on the part of the Adjuster. At mortal death, such deserted Monitors return to Divinington, and subsequently, following the adjudication of the nonsurvivor, they may be reassigned to the worlds of time and space. Sometimes, after repeated services of this sort or following some unusual experience, such as functioning as the indwelling Adjuster of an incarnated bestowal Son, these efficient Adjusters are personalized by the Universal Father.
\vs p040 4:2 Personalized Adjusters are beings of a unique and unfathomable order. Originally of existential prepersonal status, they have experientialized by participation in the lives and careers of the lowly mortals of the material worlds. And since the personality bestowed upon these experienced Thought Adjusters takes origin, and has its wellspring, in the Universal Father’s personal and continuing ministry of the bestowals of experiential personality upon his creature creation, these Personalized Adjusters are classified as ascending Sons of God, the highest of all such orders of sonship.
\usection{Mortals of Time and Space}
\vs p040 5:1 Mortals represent the last link in the chain of those beings who are called sons of God. The personal touch of the Original and Eternal Son passes on down through a series of decreasingly divine and increasingly human personalizations until there arrives a being much like yourselves, one you can see, hear, and touch. And then you are made spiritually aware of the great truth which your faith may grasp --- sonship with the eternal God!
\vs p040 5:2 Likewise does the Original and Infinite Spirit, by a long series of decreasingly divine and increasingly human orders, draw nearer and nearer to the struggling creatures of the realms, reaching the limit of expression in the angels --- than whom you were created but a little lower --- who personally guard and guide you in the life journey of the mortal career of time.
\vs p040 5:3 God the Father does not, cannot, thus downstep himself to make such near personal contact with the almost limitless number of ascending creatures throughout the universe of universes. But the Father is not deprived of personal contact with his lowly creatures; you are not without the divine presence. Although God the Father cannot be with you by direct personality manifestation, he is in you and of you in the identity of the indwelling Thought Adjusters, the divine Monitors. Thus does the Father, who is the farthest from you in personality and in spirit, draw the nearest to you in the personality circuit and in the spirit touch of inner communion with the very souls of his mortal sons and daughters.
\vs p040 5:4 \pc Spirit identification constitutes the secret of personal survival and determines the destiny of spiritual ascension. And since the Thought Adjusters are the only spirits of fusion potential to be identified with man during the life in the flesh, the mortals of time and space are primarily classified in accordance with their relation to these divine gifts, the indwelling Mystery Monitors. This classification is as follows:
\vs p040 5:5 \ublistelem{1.}\bibnobreakspace Mortals of the transient or experiential Adjuster sojourn.
\vs p040 5:6 \ublistelem{2.}\bibnobreakspace Mortals of the non\hyp{}Adjuster\hyp{}fusion types.
\vs p040 5:7 \ublistelem{3.}\bibnobreakspace Mortals of Adjuster\hyp{}fusion potential.
\vs p040 5:8 \pc \bibemph{Series one --- mortals of the transient or experiential Adjuster sojourn.} This series designation is temporary for any evolving planet, being used during the early stages of all inhabited worlds except those of the second series.
\vs p040 5:9 Mortals of series one inhabit the worlds of space during the earlier epochs of the evolution of mankind and embrace the most primitive types of human minds. On many worlds like pre\hyp{}Adamic Urantia great numbers of the higher and more advanced types of primitive men acquire survival capacity but fail to attain Adjuster fusion. For ages upon ages, before man’s ascent to the level of higher spiritual volition, the Adjusters occupy the minds of these struggling creatures during their short lives in the flesh, and the moment such will creatures are indwelt by Adjusters, the group guardian angels begin to function. While these mortals of the first series do not have personal guardians, they do have group custodians.
\vs p040 5:10 An experiential Adjuster remains with a primitive human being throughout his entire lifetime in the flesh. The Adjusters contribute much to the advancement of primitive men but are unable to form eternal unions with such mortals. This transient ministry of the Adjusters accomplishes two things: First, they gain valuable and actual experience in the nature and working of the evolutionary intellect, an experience which will be invaluable in connection with later contacts on other worlds with beings of higher development. Second, the transient sojourn of the Adjusters contributes much towards preparing their mortal subjects for possible subsequent Spirit fusion. All God\hyp{}seeking souls of this type achieve eternal life through the spiritual embrace of the Mother Spirit of the local universe, thus becoming ascending mortals of the local universe regime. Many persons from pre\hyp{}Adamic Urantia were thus advanced to the mansion worlds of Satania.
\vs p040 5:11 \pc The Gods who ordained that mortal man should climb to higher levels of spiritual intelligence through long ages of evolutionary trials and tribulations, take note of his status and needs at every stage of the ascent; and always are they divinely fair and just, even charmingly merciful, in the final judgments of these struggling mortals of the early days of the evolving races.
\vs p040 5:12 \pc \bibemph{Series two --- mortals of the non\hyp{}Adjuster\hyp{}fusion types.} These are specialized types of human beings who are not able to effect eternal union with their indwelling Adjusters. Type classification among the one-, two-, and three\hyp{}brained races is not a factor in Adjuster fusion; all such mortals are akin, but these non\hyp{}Adjuster\hyp{}fusion types are a wholly different and markedly modified order of will creatures. Many of the nonbreathers belong to this series, and there are numerous other groups who do not ordinarily fuse with Adjusters.
\vs p040 5:13 Like series number one, each member of this group enjoys the ministry of a single Adjuster during lifetime in the flesh. During temporal life these Adjusters do everything for their subjects of temporary indwelling that is done on other worlds where the mortals are of fusion potential. The mortals of this second series are often indwelt by virgin Adjusters, but the higher human types are often in liaison with masterful and experienced Monitors.
\vs p040 5:14 In the ascendant plan for upstepping the animal\hyp{}origin creatures, these beings enjoy the same devoted service of the Sons of God as is extended to the Urantia type of mortals. Seraphic co\hyp{}operation with Adjusters on the nonfusion planets is just as fully provided as on the worlds of fusion potential; the guardians of destiny minister on such spheres just as on Urantia and similarly function at the time of mortal survival, at which time the surviving soul becomes Spirit fused.
\vs p040 5:15 When you encounter these modified mortal types on the mansion worlds, you will find no difficulty in communicating with them. There they speak the same system language but by a modified technique. These beings are identical with your order of creature life in spirit and personality manifestations, differing only in certain physical features and in the fact that they are nonfusible with Thought Adjusters.
\vs p040 5:16 As to just why this type of creature is never able to fuse with the Adjusters of the Universal Father, I am unable to say. Some of us incline to the belief that the Life Carriers, in their efforts to formulate beings capable of maintaining existence in an unusual planetary environment, are confronted with the necessity of making such radical modifications in the universe plan of intelligent will creatures that it becomes inherently impossible to bring about permanent union with the Adjusters. Often have we asked: Is this an intended or an unintended part of the ascension plan? but we have not found the answer.
\vs p040 5:17 \pc \bibemph{Series three --- mortals of Adjuster\hyp{}fusion potential.} All Father\hyp{}fused mortals are of animal origin, just like the Urantia races. They embrace mortals of the one\hyp{}brained, two\hyp{}brained, and three\hyp{}brained types of Adjuster\hyp{}fusion potential. Urantians are of the intermediate or two\hyp{}brained type, being in many ways humanly superior to the one\hyp{}brained groups but definitely limited in comparison with the three\hyp{}brained orders. These three types of physical\hyp{}brain endowment are not factors in Adjuster bestowal, in seraphic service, or in any other phase of spirit ministry. The intellectual and spiritual differential between the three brain types characterizes individuals who are otherwise quite alike in mind endowment and spiritual potential, being greatest in the temporal life and tending to diminish as the mansion worlds are traversed one by one. From the system headquarters on, the progression of these three types is the same, and their eventual Paradise destiny is identical.
\vs p040 5:18 \pc \bibemph{The unnumbered series.} These narratives cannot possibly embrace all of the fascinating variations in the evolutionary worlds. You know that every tenth world is a decimal or experimental planet, but you know nothing of the other variables that punctuate the processional of the evolutionary spheres. There are differences too numerous to narrate even between the revealed orders of living creatures as between planets of the same group, but this presentation makes clear the essential differences in relation to the ascension career. And the ascension career is the most important factor in any consideration of the mortals of time and space.
\vs p040 5:19 \pc As to the chances of mortal survival, let it be made forever clear: All souls of every possible phase of mortal existence will survive provided they manifest willingness to co\hyp{}operate with their indwelling Adjusters and exhibit a desire to find God and to attain divine perfection, even though these desires be but the first faint flickers of the primitive comprehension of that “true light which lights every man who comes into the world.”
\usection{The Faith Sons of God}
\vs p040 6:1 The mortal races stand as the representatives of the lowest order of intelligent and personal creation. You mortals are divinely beloved, and every one of you may choose to accept the certain destiny of a glorious experience, but you are not yet by nature of the divine order; you are wholly mortal. You will be reckoned as ascending sons the instant fusion takes place, but the status of the mortals of time and space is that of faith sons prior to the event of the final amalgamation of the surviving mortal soul with some type of eternal and immortal spirit.
\vs p040 6:2 It is a solemn and supernal fact that such lowly and material creatures as Urantia human beings are the sons of God, faith children of the Highest. “Behold, what manner of love the Father has bestowed upon us that we should be called the sons of God.” “As many as received him, to them gave he the power to recognize that they are the sons of God.” While “it does not yet appear what you shall be,” even now “you are the faith sons of God”; “for you have not received the spirit of bondage again to fear, but you have received the spirit of sonship, whereby you cry, ‘our Father.’” Spoke the prophet of old in the name of the eternal God: “Even to them will I give in my house a place and a name better than sons; I will give them an everlasting name, one that shall not be cut off.” “And because you are sons, God has sent forth the spirit of his Son into your hearts.”
\vs p040 6:3 All evolutionary worlds of mortal habitation harbour these faith sons of God, sons of grace and mercy, mortal beings belonging to the divine family and accordingly called the sons of God. Urantia mortals are entitled to regard themselves as being the sons of God because:
\vs p040 6:4 \ublistelem{1.}\bibnobreakspace You are sons of spiritual promise, faith sons; you have accepted the status of sonship. You believe in the reality of your sonship, and thus does your sonship with God become eternally real.
\vs p040 6:5 \ublistelem{2.}\bibnobreakspace A Creator Son of God became one of you; he is your elder brother in fact; and if in spirit you become truly related brothers of Christ, the victorious Michael, then in spirit must you also be sons of that Father which you have in common --- even the Universal Father of all.
\vs p040 6:6 \ublistelem{3.}\bibnobreakspace You are sons because the spirit of a Son has been poured out upon you, has been freely and certainly bestowed upon all Urantia races. This spirit ever draws you toward the divine Son, who is its source, and toward the Paradise Father, who is the source of that divine Son.
\vs p040 6:7 \ublistelem{4.}\bibnobreakspace Of his divine free\hyp{}willness, the Universal Father has given you your creature personalities. You have been endowed with a measure of that divine spontaneity of freewill action which God shares with all who may become his sons.
\vs p040 6:8 \ublistelem{5.}\bibnobreakspace There dwells within you a fragment of the Universal Father, and you are thus directly related to the divine Father of all the Sons of God.
\usection{Father\hyp{}Fused Mortals}
\vs p040 7:1 The sending of Adjusters, their indwelling, is indeed one of the unfathomable mysteries of God the Father. These fragments of the divine nature of the Universal Father carry with them the potential of creature immortality. Adjusters are immortal spirits, and union with them confers eternal life upon the soul of the fused mortal.
\vs p040 7:2 Your own races of surviving mortals belong to this group of the ascending Sons of God. You are now planetary sons, evolutionary creatures derived from the Life Carrier implantations and modified by the Adamic\hyp{}life infusion, hardly yet ascending sons; but you are indeed sons of ascension potential --- even to the highest heights of glory and divinity attainment --- and this spiritual status of ascending sonship you may attain by faith and by freewill co\hyp{}operation with the spiritualizing activities of the indwelling Adjuster. When you and your Adjusters are finally and forever fused, when you two are made one, even as in Christ Michael the Son of God and the Son of Man are one, then in fact have you become the ascending sons of God.
\vs p040 7:3 The details of the Adjuster career of indwelling ministry on a probationary and evolutionary planet are not a part of my assignment; the elaboration of this great truth embraces your whole career. I include the mention of certain Adjuster functions in order to make a replete statement regarding Adjuster\hyp{}fused mortals. These indwelling fragments of God are with your order of being from the early days of physical existence through all of the ascending career in Nebadon and Orvonton and on through Havona to Paradise itself. Thereafter, in the eternal adventure, this same Adjuster is one with you and of you.
\vs p040 7:4 These are the mortals who have been commanded by the Universal Father, “Be you perfect, even as I am perfect.” The Father has bestowed himself upon you, placed his own spirit within you; \bibemph{therefore} does he demand ultimate perfection of you. The narrative of human ascent from the mortal spheres of time to the divine realms of eternity constitutes an intriguing recital not included in my assignment, but this supernal adventure should be the supreme study of mortal man.
\vs p040 7:5 Fusion with a fragment of the Universal Father is equivalent to a divine validation of eventual Paradise attainment, and such Adjuster\hyp{}fused mortals are the only class of human beings who all traverse the Havona circuits and find God on Paradise. To the Adjuster\hyp{}fused mortal the career of universal service is wide open. What dignity of destiny and glory of attainment await every one of you! Do you fully appreciate what has been done for you? Do you comprehend the grandeur of the heights of eternal achievement which are spread out before you? --- even you who now trudge on in the lowly path of life through your so\hyp{}called “vale of tears”?
\usection{Son\hyp{}Fused Mortals}
\vs p040 8:1 While practically all surviving mortals are fused with their Adjusters on one of the mansion worlds or immediately upon their arrival on the higher morontia spheres, there are certain cases of delayed fusion, some not experiencing this final surety of survival until they reach the last educational worlds of the universe headquarters; and a few of these mortal candidates for never\hyp{}ending life utterly fail to attain identity fusion with their faithful Adjusters.
\vs p040 8:2 Such mortals have been deemed worthy of survival by the adjudicational authorities, and even their Adjusters, by returning from Divinington, have concurred in their ascension to the mansion worlds. Such beings have ascended through a system, a constellation, and through the educational worlds of the Salvington circuit; they have enjoyed the “seventy times seven” opportunities for fusion and still have been unable to attain oneness with their Adjusters.
\vs p040 8:3 When it becomes apparent that some synchronizing difficulty is inhibiting Father fusion, the survival referees of the Creator Son are convened. And when this court of inquiry, sanctioned by a personal representative of the Ancients of Days, finally determines that the ascending mortal is not guilty of any discoverable cause for failure to attain fusion, they so certify on the records of the local universe and duly transmit this finding to the Ancients of Days. Thereupon does the indwelling Adjuster return forthwith to Divinington for confirmation by the Personalized Monitors, and upon this leave\hyp{}taking the morontia mortal is immediately fused with an individualized gift of the spirit of the Creator Son.
\vs p040 8:4 \pc Much as the morontia spheres of Nebadon are shared with the Spirit\hyp{}fused mortals, so do these Son\hyp{}fused creatures share the services of Orvonton with their Adjuster\hyp{}fused brethren who are journeying inward towards the far\hyp{}distant Isle of Paradise. They are truly your brethren, and you will greatly enjoy their association as you pass through the training worlds of the superuniverse.
\vs p040 8:5 Son\hyp{}fused mortals are not a numerous group, there being less than 1,000,000 of them in the superuniverse of Orvonton. Aside from residential destiny on Paradise they are in every way the equals of their Adjuster\hyp{}fused associates. They frequently journey to Paradise on superuniverse assignment but seldom permanently reside there, being, as a class, confined to the superuniverse of their nativity.
\usection{Spirit\hyp{}Fused Mortals}
\vs p040 9:1 Ascending Spirit\hyp{}fused mortals are not Third Source personalities; they are included in the Father’s personality circuit, but they have fused with individualizations of the premind spirit of the Third Source and Centre. Such Spirit fusion never occurs during the span of natural life; it takes place only at the time of mortal reawakening in the morontia existence on the mansion worlds. In the fusion experience there is no overlapping; the will creature is either Spirit fused, Son fused, or Father fused. Those who are Adjuster or Father fused are never Spirit or Son fused.
\vs p040 9:2 \pc The fact that these types of mortal creatures are not Adjuster\hyp{}fusion candidates does not prevent the Adjusters from indwelling them during the life in the flesh. Adjusters do work in the minds of such beings during the span of material life but never become everlastingly one with their pupil souls. During this temporary sojourn the Adjusters effectively build up the same spirit counterpart of mortal nature --- the soul --- that they do in the candidates for Adjuster fusion. Up to the time of mortal death the work of the Adjusters is wholly akin to their function in your own races, but upon mortal dissolution the Adjusters take eternal leave of these Spirit\hyp{}fusion candidates and, proceeding directly to Divinington, the headquarters of all divine Monitors, there await the new assignments of their order.
\vs p040 9:3 When such sleeping survivors are repersonalized on the mansion worlds, the place of the departed Adjuster is filled by an individualization of the spirit of the Divine Minister, the representative of the Infinite Spirit in the local universe concerned. This spirit infusion constitutes these surviving creatures Spirit\hyp{}fused mortals. Such beings are in every way your equals in mind and spirit; and they are indeed your contemporaries, sharing the mansion and morontia spheres in common with your order of fusion candidates and with those who are to be Son fused.
\vs p040 9:4 \pc There is, however, one particular in which Spirit\hyp{}fused mortals differ from their ascendant brethren: Mortal memory of human experience on the material worlds of origin survives death in the flesh because the indwelling Adjuster has acquired a spirit counterpart, or transcript, of those events of human life which were of spiritual significance. But with Spirit\hyp{}fused mortals there exists no such mechanism whereby human memory may persist. The Adjuster transcripts of memory are full and intact, but these acquisitions are experiential possessions of the departed Adjusters and are not available to the creatures of their former indwelling, who therefore awaken in the resurrection halls of the morontia spheres of Nebadon as if they were newly created beings, creatures without consciousness of former existence.
\vs p040 9:5 Such children of the local universe are enabled to repossess themselves of much of their former human memory experience through having it retold by the associated seraphim and cherubim and by consulting the records of the mortal career filed by the recording angels. This they can do with undoubted assurance because the surviving soul, of experiential origin in the material and mortal life, while having no memory of mortal events, does have a residual experiential\hyp{}recognition\hyp{}response to these unremembered events of past experience.
\vs p040 9:6 When a Spirit\hyp{}fused mortal is told about the events of the unremembered past experience, there is an immediate response of experiential recognition within the soul (identity) of such a survivor which instantly invests the narrated event with the emotional tinge of reality and with the intellectual quality of fact; and this dual response constitutes the reconstruction, recognition, and validation of an unremembered facet of mortal experience.
\vs p040 9:7 Even with Adjuster\hyp{}fusion candidates, only those human experiences which were of spiritual value are common possessions of the surviving mortal and the returning Adjuster and hence are immediately remembered subsequent to mortal survival. Concerning those happenings which were not of spiritual significance, even these Adjuster\hyp{}fusers must depend upon the attribute of recognition\hyp{}response in the surviving soul. And since any one event may have a spiritual connotation to one mortal but not to another, it becomes possible for a group of contemporary ascenders from the same planet to pool their store of Adjuster\hyp{}remembered events and thus to reconstruct any experience which they had in common, and which was of spiritual value in the life of any one of them.
\vs p040 9:8 \pc While we understand such techniques of memory reconstruction fairly well, we do not grasp the technique of personality recognition. Personalities of onetime association mutually respond quite independently of the operation of memory, albeit, memory itself and the techniques of its reconstruction are necessary to invest such mutual personality response with the fullness of recognition.
\vs p040 9:9 \pc A Spirit\hyp{}fused survivor is also able to learn much about the life he lived in the flesh by revisiting his nativity world subsequent to the planetary dispensation in which he lived. Such children of Spirit fusion are enabled to enjoy these opportunities for investigating their human careers since they are in general confined to the service of the local universe. They do not share your high and exalted destiny in the Paradise Corps of the Finality; only Adjuster\hyp{}fused mortals or other especially embraced ascendant beings are mustered into the ranks of those who await the eternal Deity adventure. Spirit\hyp{}fused mortals are the permanent citizens of the local universes; they may aspire to Paradise destiny, but they cannot be sure of it. In Nebadon their universe home is the eighth group of worlds encircling Salvington, a destiny\hyp{}heaven of nature and location much like the one envisioned by the planetary traditions of Urantia.
\usection{Ascendant Destinies}
\vs p040 10:1 Spirit\hyp{}fused mortals are, generally speaking, confined to a local universe; Son\hyp{}fused survivors are restricted to a superuniverse; Adjuster\hyp{}fused mortals are destined to penetrate the universe of universes. The spirits of mortal fusion always ascend to the level of origin; such spirit entities unfailingly return to the sphere of primal source.
\vs p040 10:2 Spirit\hyp{}fused mortals are of the local universe; they do not, ordinarily, ascend beyond the confines of their native realm, beyond the boundaries of the space range of the spirit that pervades them. Son\hyp{}fused ascenders likewise rise to the source of spirit endowment, for much as the Truth Spirit of a Creator Son focalizes in the associated Divine Minister, so is his “fusion spirit” implemented in the Reflective Spirits of the higher universes. Such spirit relationship between the local and the superuniverse levels of God the Sevenfold may be difficult of explanation but not of discernment, being unmistakably revealed in those children of the Reflective Spirits --- the secoraphic Voices of the Creator Sons. The Thought Adjuster, hailing from the Father on Paradise, never stops until the mortal son stands face to face with the eternal God.
\vs p040 10:3 \pc The mysterious variable in associative technique whereby a mortal being does not or cannot become eternally fused with the indwelling Thought Adjuster may seem to disclose a flaw in the ascension scheme; Son and Spirit fusion do, superficially, resemble compensations of unexplained failures in some detail of the Paradise\hyp{}attainment plan; but all such conclusions stand in error; we are taught that all these happenings unfold in obedience to the established laws of the Supreme Universe Rulers.
\vs p040 10:4 We have analysed this problem and have reached the undoubted conclusion that the consignment of all mortals to an ultimate Paradise destiny would be unfair to the time\hyp{}space universes inasmuch as the courts of the Creator Sons and of the Ancients of Days would then be wholly dependent on the services of those who were in transit to higher realms. And it does seem to be no more than fitting that the local and the superuniverse governments should each be provided with a permanent group of ascendant citizenship; that the functions of these administrations should be enriched by the efforts of certain groups of glorified mortals who are of permanent status, evolutionary complements of the abandonters and of the susatia. Now it is quite obvious that the present ascension scheme effectively provides the time\hyp{}space administrations with just such groups of ascendant creatures; and we have many times wondered: Does all this represent an intended part of the all\hyp{}wise plans of the Architects of the Master Universe designed to provide the Creator Sons and the Ancients of Days with a permanent ascendant population? with evolved orders of citizenship that will become increasingly competent to carry forward the affairs of these realms in the universe ages to come?
\vs p040 10:5 \pc That mortal destinies do thus vary in no wise proves that one is necessarily greater or lesser than another, merely that they differ. Adjuster\hyp{}fused ascenders do indeed have a grand and glorious career as finaliters spread out before them in the eternal future, but this does not mean that they are preferred above their ascendant brethren. There is no favouritism, nothing arbitrary, in the selective operation of the divine plan of mortal survival.
\vs p040 10:6 \pc While the Adjuster\hyp{}fused finaliters obviously enjoy the widest service opportunity of all, the attainment of this goal automatically shuts them off from the chance to participate in the agelong struggle of some one universe or superuniverse, from the earlier and less settled epochs to the later and established eras of relative perfection attainment. Finaliters acquire a marvellous and far\hyp{}flung experience of transient service in all seven segments of the grand universe, but they do not ordinarily acquire that intimate knowledge of any one universe which even now characterizes the Spirit\hyp{}fused veterans of the Nebadon Corps of Completion. These individuals enjoy an opportunity to witness the ascending processional of the planetary ages as they unfold one by one on 10,000,000 inhabited worlds. And in the faithful service of such local universe citizens, experience superimposes upon experience until the fullness of time ripens that high quality of wisdom which is engendered by focalized experience --- \bibemph{authoritative} wisdom --- and this in itself is a vital factor in the settling of any local universe.
\vs p040 10:7 As it is with the Spirit fusers, so is it with those Son\hyp{}fused mortals who have achieved residential status on Uversa. Some of these beings hail from the earliest epochs of Orvonton, and they represent a slowly accumulating body of insight\hyp{}deepening wisdom which is making ever\hyp{}augmenting service contributions to the welfare and eventual settlement of the seventh superuniverse.
\vs p040 10:8 \pc What the ultimate destiny of these stationary orders of local and of superuniverse citizenship will be we do not know, but it is quite possible that, when the Paradise finaliters are pioneering the expanding frontiers of divinity in the planetary systems of the first outer space level, their Son\hyp{} and Spirit\hyp{}fused brethren of the ascendant evolutionary struggle will be acceptably contributing to the maintenance of the experiential equilibrium of the perfected superuniverses while they stand ready to welcome the incoming stream of Paradise pilgrims who may, at that distant day, pour in through Orvonton and its sister creations as a vast spirit\hyp{}questing torrent from these now uncharted and uninhabited galaxies of outer space.
\vs p040 10:9 \pc While the majority of Spirit fusers serve permanently as citizens of the local universes, all do not. If some phase of their universe ministry should require their personal presence in the superuniverse, then would such transformations of being be wrought in these citizens as would enable them to ascend to the higher universe; and upon the arrival of the Celestial Guardians with orders to present such Spirit\hyp{}fused mortals at the courts of the Ancients of Days, they would so ascend, never to return. They become wards of the superuniverse, serving as assistants to the Celestial Guardians and permanently, save for those few who are in turn summoned to the service of Paradise and Havona.
\vs p040 10:10 Like their Spirit\hyp{}fused brethren, the Son fusers neither traverse Havona nor attain Paradise unless they have undergone certain modifying transformations. For good and sufficient reasons, such changes have been wrought in certain Son\hyp{}fused survivors, and these beings are to be encountered ever and anon on the seven circuits of the central universe. Thus it is that certain numbers of both the Son\hyp{} and the Spirit\hyp{}fused mortals do actually ascend to Paradise, do attain a goal in many ways equal to that which awaits the Father\hyp{}fused mortals.
\vs p040 10:11 Father\hyp{}fused mortals are potential finaliters; their destination is the Universal Father, and him they do attain, but within the purview of the present universe age, finaliters, as such, are not destiny attainers. They remain unfinished creatures --- sixth\hyp{}stage spirits --- and hence nonactive in the evolutionary domains of prelight\hyp{}and\hyp{}life status.
\vs p040 10:12 When a mortal finaliter is Trinity embraced --- becomes a Trinitized Son, such as a Mighty Messenger --- then has that finaliter attained destiny, at least for the present universe age. Mighty Messengers and their fellows may not in the exact sense be seventh\hyp{}stage spirits, but in addition to other things the Trinity embrace endows them with everything which a finaliter will sometime achieve as a seventh\hyp{}stage spirit. After Spirit\hyp{}fused or Son\hyp{}fused mortals are trinitized, they pass through the Paradise experience with the Adjuster\hyp{}fused ascenders, with whom they are then identical in all matters pertaining to superuniverse administration. These Trinitized Sons of Selection or of Attainment at least for now are finished creatures, in contrast to the finaliters, who are at present unfinished creatures.
\vs p040 10:13 \pc Thus, in the final analysis, it would be hardly proper to use the words “greater” or “lesser” in contrasting the destinies of the ascending orders of sonship. Every such son of God shares the fatherhood of God, and God loves each of his creature sons alike; he is no more a respecter of ascendant destinies than is he of the creatures who may attain such destinies. The Father loves \bibemph{each} of his sons, and that affection is not less than true, holy, divine, unlimited, eternal, and unique --- a love bestowed upon \bibemph{this} son and upon \bibemph{that} son, individually, personally, and exclusively. And such a love utterly eclipses all other facts. Sonship is the supreme relationship of the creature to the Creator.
\vs p040 10:14 As mortals you can now recognize your place in the family of divine sonship and begin to sense the obligation to avail yourselves of the advantages so freely provided in and by the Paradise plan for mortal survival, which plan has been so enhanced and illuminated by the life experience of a bestowal Son. Every facility and all power have been provided for ensuring your ultimate attainment of the Paradise goal of divine perfection.
\vsetoff
\vs p040 10:15 [Presented by a Mighty Messenger temporarily attached to the staff of Gabriel of Salvington.]
\quizlink
