\upaper{46}{The Local System Headquarters}
\uminitoc{Physical Aspects of Jerusem}
\uminitoc{Physical Features of Jerusem}
\uminitoc{The Jerusem Broadcasts}
\uminitoc{Residential and Administrative Areas}
\uminitoc{The Jerusem Circles}
\uminitoc{The Executive-Administrative Squares}
\uminitoc{The Rectangles --- The Spornagia}
\uminitoc{The Jerusem Triangles}
\author{Archangel}
\vs p046 0:1 Jerusem, the headquarters of Satania, is an average capital of a local system, and aside from numerous irregularities occasioned by the Lucifer rebellion and the bestowal of Michael on Urantia, it is typical of similar spheres. Your local system has passed through some stormy experiences, but it is at present being administered most efficiently, and as the ages pass, the results of disharmony are being slowly but surely eradicated. Order and good will are being restored, and the conditions on Jerusem are more and more approaching the heavenly status of your traditions, for the system headquarters is truly the heaven visualized by the majority of XX century religious believers.
\usection{Physical Aspects of Jerusem}
\vs p046 1:1 Jerusem is divided into 1,000 latitudinal sectors and 10,000 longitudinal zones. The sphere has 7 major capitals and 70 minor administrative centres. The 7 sectional capitals are concerned with diverse activities, and the System Sovereign is present in each at least once a year.
\vs p046 1:2 \pc The standard mile of Jerusem is equivalent to about 11 Urantia km. The standard weight, the “gradant,” is built up through the decimal system from the mature ultimaton\fnst{This suggests that there is no need for the concept of \bibemph{amount of substance} or the equivalent of our Avogadro's number on Jerusem. It should be recalled that on Urantia the concept of the ``amount of substance'' exists because the unit of mass is not defined as a multiple of some elementary such unit, unlike the metre and second. And the accuracy with which the mass can be measured differs from that of the number of structural elements (electrons, ions, molecules, atoms etc). But on Jerusem, the mass is defined in the same way as our metre and second, i.e. as multiple of an elementary unit.} and represents almost exactly 283 grams\fnst{Thus the mass of the mature ultimaton can be written as $2.8349523125 \times 10^{-N}$\,kg, where $N$ is a natural number. I think that $N=8$, i.e. ultimaton has the mass close to the value of \bibemph{Planck mass}.} of your weight. The Satania day equals 3 days of Urantia time, less 1 hour, 4 minutes, and 15 seconds, that being the time of the axial revolution of Jerusem. The system year consists of 100 Jerusem days. The time of the system is broadcast by the master chronoldeks.
\vs p046 1:3 \pc The energy of Jerusem is superbly controlled and circulates about the sphere in the zone channels, which are directly fed from the energy charges of space and expertly administered by the Master Physical Controllers. The natural resistance to the passage of these energies through the physical channels of conduction yields the heat required for the production of the equable temperature of Jerusem. The full\hyp{}light temperature is maintained at about 21\,°C, while during the period of light recession it falls to a little lower than 10\,°C.
\vs p046 1:4 \pc The lighting system of Jerusem should not be so difficult for you to comprehend. There are no days and nights, no seasons of heat and cold. The power transformers maintain 100,000 centres from which rarefied energies are projected upward through the planetary atmosphere, undergoing certain changes, until they reach the electric air\hyp{}ceiling of the sphere; and then these energies are reflected back and down as a gentle, sifting, and even light of about the intensity of Urantia sunlight when the sun is shining overhead at 10:00.
\vs p046 1:5 Under such conditions of lighting, the light rays do not seem to come from one place; they just sift out of the sky, emanating equally from all space directions. This light is very similar to natural sunlight except that it contains very much less heat. Thus it will be recognized that such headquarters worlds are not luminous in space; if Jerusem were very near Urantia, it would not be visible.
\vs p046 1:6 The gases which reflect this light\hyp{}energy from the Jerusem upper ionosphere back to the ground are very similar to those in the Urantia upper air belts which are concerned with the auroral phenomena of your so\hyp{}called northern lights, although these are produced by different causes. On Urantia it is this same gas shield which prevents the escape of the terrestrial broadcast waves, reflecting them earthward when they strike this gas belt in their direct outward flight. In this way broadcasts are held near the surface as they journey through the air around your world.
\vs p046 1:7 This lighting of the sphere is uniformly maintained for 75\% of the Jerusem day, and then there is a gradual recession until, at the time of minimum illumination, the light is about that of your full moon on a clear night. This is the quiet hour for all Jerusem. Only the broadcast\hyp{}receiving stations are in operation during this period of rest and rehabilitation.
\vs p046 1:8 \pc Jerusem receives faint light from several near\hyp{}by suns --- a sort of brilliant starlight --- but it is not dependent on them; worlds like Jerusem are not subject to the vicissitudes of sun disturbances, neither are they confronted with the problem of a cooling or dying sun.
\vs p046 1:9 The 7 transitional study worlds and their 49 satellites are heated, lighted, energized, and watered by the Jerusem technique.
\usection{Physical Features of Jerusem}
\vs p046 2:1 On Jerusem you will miss the rugged mountain ranges of Urantia and other evolved worlds since there are neither earthquakes nor rainfalls, but you will enjoy the beauteous highlands and other unique variations of topography and landscape. Enormous areas of Jerusem are preserved in a “natural state,” and the grandeur of such districts is quite beyond the powers of human imagination.
\vs p046 2:2 There are thousands upon thousands of small lakes but no raging rivers nor expansive oceans. There is no rainfall, neither storms nor blizzards, on any of the architectural worlds, but there is the daily precipitation of the condensation of moisture during the time of lowest temperature attending the light recession. (The dew point is higher on a 3\hyp{}gas world than on a 2\hyp{}gas planet like Urantia.) The physical plant life and the morontia world of living things both require moisture, but this is largely supplied by the subsoil system of circulation which extends all over the sphere, even up to the very tops of the highlands. This water system is not entirely subsurface, for there are many canals interconnecting the sparkling lakes of Jerusem.
\vs p046 2:3 The atmosphere of Jerusem is a 3\hyp{}gas mixture. This air is very similar to that of Urantia with the addition of a gas adapted to the respiration of the morontia order of life. This third gas in no way unfits the air for the respiration of animals or plants of the material orders.
\vs p046 2:4 The transportation system is allied with the circulatory streams of energy movement, these main energy currents being located at 10\hyp{}mile intervals. By adjustment of physical mechanisms the material beings of the planet can proceed at a pace varying from 2 to 500 miles\fnst{Here the ``mile'' could be referring to a ``standard Jerusem mile'' and so was left unconverted.} per hour. The transport birds fly at about 100 miles an hour. The air mechanisms of the Material Sons travel around 500 miles per hour. Material and early morontia beings must utilize these mechanical means of transport, but spirit personalities proceed by liaison with the superior forces and spirit sources of energy.
\vs p046 2:5 \pc Jerusem and its associated worlds are endowed with the ten standard divisions of physical life characteristic of the architectural spheres of Nebadon. And since there is no organic evolution on Jerusem, there are no conflicting forms of life, no struggle for existence, no survival of the fittest. Rather is there a creative adaptation which foreshadows the beauty, the harmony, and the perfection of the eternal worlds of the central and divine universe. And in all this creative perfection there is the most amazing intermingling of physical and of morontia life, artistically contrasted by the celestial artisans and their fellows.
\vs p046 2:6 Jerusem is indeed a foretaste of paradisiacal glory and grandeur. But you can never hope to gain an adequate idea of these glorious architectural worlds by any attempted description. There is so little that can be compared with aught on your world, and even then the things of Jerusem so transcend the things of Urantia that the comparison is almost grotesque. Until you actually arrive on Jerusem, you can hardly entertain anything like a true concept of the heavenly worlds, but that is not so long a time in the future when your coming experience on the system capital is compared with your sometime arrival on the more remote training spheres of the universe, the superuniverse, and of Havona.
\vs p046 2:7 \pc The manufacturing or laboratory sector of Jerusem is an extensive domain, one which Urantians would hardly recognize since it has no smoking chimneys; nevertheless, there is an intricate material economy associated with these special worlds, and there is a perfection of mechanical technique and physical achievement which would astonish and even awe your most experienced chemists and inventors. Pause to consider that this first world of detention in the Paradise journey is far more material than spiritual. Throughout your stay on Jerusem and its transition worlds you are far nearer your earth life of material things than your later life of advancing spirit existence.
\vs p046 2:8 \pc Mount Seraph is the highest elevation on Jerusem, almost 4.5\,km, and is the point of departure for all transport seraphim. Numerous mechanical developments are used in providing initial energy for escaping the planetary gravity and overcoming the air resistance. A seraphic transport departs every 3 seconds of Urantia time throughout the light period and, sometimes, far into the recession. The transporters take off at about 25 standard miles per second of Urantia time and do not attain standard velocity until they are over 2,000 miles away from Jerusem.
\vs p046 2:9 Transports arrive on the crystal field, the so\hyp{}called sea of glass. Around this area are the receiving stations for the various orders of beings who traverse space by seraphic transport. Near the polar crystal receiving station for student visitors you may ascend the pearly observatory and view the immense relief map of the entire headquarters planet.
\usection{The Jerusem Broadcasts}
\vs p046 3:1 The superuniverse and Paradise\hyp{}Havona broadcasts are received on Jerusem in liaison with Salvington and by a technique involving the polar crystal, the sea of glass. In addition to provisions for the reception of these extra\hyp{}Nebadon communications, there are 3 distinct groups of receiving stations. These separate but tricircular groups of stations are adjusted to the reception of broadcasts from the local worlds, from the constellation headquarters, and from the capital of the local universe. All these broadcasts are automatically displayed so as to be discernible by all types of beings present in the central broadcast amphitheatre; of all preoccupations for an ascendant mortal on Jerusem, none is more engaging and engrossing than that of listening in on the never\hyp{}ending stream of universe space reports.
\vs p046 3:2 This Jerusem broadcast\hyp{}receiving station is encircled by an enormous amphitheatre, constructed of scintillating materials largely unknown on Urantia and seating over five billion beings --- material and morontia --- besides accommodating innumerable spirit personalities. It is the favourite diversion for all Jerusem to spend their leisure at the broadcast station, there to learn of the welfare and state of the universe. And this is the only planetary activity which is not slowed down during the recession of light.
\vs p046 3:3 At this broadcast\hyp{}receiving amphitheatre the Salvington messages are coming in continuously. Near by, the Edentia word of the Most High Constellation Fathers is received at least once a day. Periodically the regular and special broadcasts of Uversa are relayed through Salvington, and when Paradise messages are in reception, the entire population is assembled around the sea of glass, and the Uversa friends add the reflectivity phenomena to the technique of the Paradise broadcast so that everything heard becomes visible. And it is in this manner that continual foretastes of advancing beauty and grandeur are afforded the mortal survivors as they journey inward on the eternal adventure.
\vs p046 3:4 \pc The Jerusem sending station is located at the opposite pole of the sphere. All broadcasts to the individual worlds are relayed from the system capitals except the Michael messages, which sometimes go direct to their destinations over the archangels’ circuit.
\usection{Residential and Administrative Areas}
\vs p046 4:1 Considerable portions of Jerusem are assigned as residential areas, while other portions of the system capital are given over to the necessary administrative functions involving the supervision of the affairs of 619 inhabited spheres, 56 transitional\hyp{}culture worlds, and the system capital itself. On Jerusem and in Nebadon these arrangements are designed as follows:
\vs p046 4:2 \ublistelem{1.}\bibnobreakspace \bibemph{The circles ---} the nonnative residential areas.
\vs p046 4:3 \ublistelem{2.}\bibnobreakspace \bibemph{The squares ---} the system executive\hyp{}administrative areas.
\vs p046 4:4 \ublistelem{3.}\bibnobreakspace \bibemph{The rectangles ---} the rendezvous of the lower native life.
\vs p046 4:5 \ublistelem{4.}\bibnobreakspace \bibemph{The triangles ---} the local or Jerusem administrative areas.
\vs p046 4:6 \pc This arrangement of the system activities into circles, squares, rectangles, and triangles is common to all the system capitals of Nebadon. In another universe an entirely different arrangement might prevail. These are matters determined by the diverse plans of the Creator Sons.
\vs p046 4:7 \pc Our narrative of these residential and administrative areas takes no account of the vast and beautiful estates of the Material Sons of God, the permanent citizens of Jerusem, neither do we mention numerous other fascinating orders of spirit and near\hyp{}spirit creatures. For example: Jerusem enjoys the efficient services of the spironga of design for system function. These beings are devoted to spiritual ministry in behalf of the supermaterial residents and visitors. They are a wonderful group of intelligent and beautiful beings who are the transition servants of the higher morontia creatures and of the morontia helpers who labour for the upkeep and embellishment of all morontia creations. They are on Jerusem what the midway creatures are on Urantia, midway helpers functioning between the material and the spiritual.
\vs p046 4:8 The system capitals are unique in that they are the only worlds which exhibit well\hyp{}nigh perfectly all 3 phases of universe existence: the material, the morontial, and the spiritual. Whether you are a material, morontia, or spirit personality, you will feel at home on Jerusem; so also do the combined beings, such as the midway creatures and the Material Sons.
\vs p046 4:9 Jerusem has great buildings of both material and morontia types, while the embellishment of the purely spiritual zones is no less exquisite and replete. If I only had words to tell you of the morontia counterparts of the marvellous physical equipment of Jerusem! If I could only go on to portray the sublime grandeur and exquisite perfection of the spiritual appointments of this headquarters world! Your most imaginative concept of perfection of beauty and repleteness of appointment would hardly approach these grandeurs. And Jerusem is but the first step on the way to the supernal perfection of Paradise beauty.
\usection{The Jerusem Circles}
\vs p046 5:1 The residential reservations assigned to the major groups of universe life are designated the Jerusem circles. Those circle groups which find mention in these narratives are the following:
\vs p046 5:2 \ublistelem{1.}\bibnobreakspace The circles of the Sons of God.
\vs p046 5:3 \ublistelem{2.}\bibnobreakspace The circles of the angels and higher spirits.
\vs p046 5:4 \ublistelem{3.}\bibnobreakspace The circles of the Universe Aids, including the creature\hyp{}trinitized sons not assigned to the Trinity Teacher Sons.
\vs p046 5:5 \ublistelem{4.}\bibnobreakspace The circles of the Master Physical Controllers.
\vs p046 5:6 \ublistelem{5.}\bibnobreakspace The circles of the assigned ascending mortals, including the midway creatures.
\vs p046 5:7 \ublistelem{6.}\bibnobreakspace The circles of the courtesy colonies.
\vs p046 5:8 \ublistelem{7.}\bibnobreakspace The circles of the Corps of the Finality.
\vs p046 5:9 \pc Each of these residential groupings consists of 7 concentric and successively elevated circles. They are all constructed along the same lines but are of different sizes and are fashioned of differing materials. They are all surrounded by far\hyp{}reaching enclosures, which mount up to form extensive promenades entirely encompassing every group of 7 concentric circles.
\vs p046 5:10 \ublistelem{1.}\bibnobreakspace \bibemph{Circles of the Sons of God.} Though the Sons of God possess a social planet of their own, one of the transitional\hyp{}culture worlds, they also occupy these extensive domains on Jerusem. On their transitional\hyp{}culture world the ascending mortals freely mingle with all orders of divine sonship. There you will personally know and love these Sons, but their social life is largely confined to this special world and its satellites. In the Jerusem circles, however, these various groups of sonship may be observed at work. And since morontia vision is of enormous range, you can walk about on the Sons’ promenades and overlook the intriguing activities of their numerous orders.
\vs p046 5:11 These 7 circles of the Sons are concentric and successively elevated so that each of the outer and larger circles overlooks the inner and smaller ones, each being surrounded by a public promenade wall. These walls are constructed of crystal gems of gleaming brightness and are so elevated as to overlook all of their respective residential circles. The many gates --- from 50 to 150,000 --- which penetrate each of these walls consist of single pearly crystals.
\vs p046 5:12 The 1\ts{st} circle of the domain of the Sons is occupied by the Magisterial Sons and their personal staffs. Here centre all of the plans and immediate activities of the bestowal and adjudicational services of these juridical Sons. It is also through this centre that the Avonals of the system maintain contact with the universe.
\vs p046 5:13 The 2\ts{nd} circle is occupied by the Trinity Teacher Sons. In this sacred domain the Daynals and their associates carry forward the training of the newly arrived primary Teacher Sons. And in all of this work they are ably assisted by a division of certain co\hyp{}ordinates of the Brilliant Evening Stars. The creature\hyp{}trinitized sons occupy a sector of the Daynal circle. The Trinity Teacher Sons come the nearest to being the personal representatives of the Universal Father in a local system; they are at least Trinity\hyp{}origin beings. This 2\ts{nd} circle is a domain of extraordinary interest to all the peoples of Jerusem.
\vs p046 5:14 The 3\ts{rd} circle is devoted to the Melchizedeks. Here the system chiefs reside and supervise the almost endless activities of these versatile Sons. From the first of the mansion worlds on through all the Jerusem career of ascending mortals, the Melchizedeks are foster fathers and ever\hyp{}present advisers. It would not be amiss to say that they are the dominant influence on Jerusem aside from the ever\hyp{}present activities of the Material Sons and Daughters.
\vs p046 5:15 The 4\ts{th} circle is the home of the Vorondadeks and all other orders of the visiting and observer Sons who are not otherwise provided for. The Most High Constellation Fathers take up their abode in this circle when on visits of inspection to the local system. Perfectors of Wisdom, Divine Counsellors, and Universal Censors all reside in this circle when on duty in the system.
\vs p046 5:16 The 5\ts{th} circle is the abode of the Lanonandeks, the sonship order of the System Sovereigns and the Planetary Princes. The 3 groups mingle as one when at home in this domain. The system reserves are held in this circle, while the System Sovereign has a temple situated at the centre of the governing group of structures on administration hill.
\vs p046 5:17 The 6\ts{th} circle is the tarrying place of the system Life Carriers. All orders of these Sons are here assembled, and from here they go forth on their world assignments.
\vs p046 5:18 The 7\ts{th} circle is the rendezvous of the ascending sons, those assigned mortals who may be temporarily functioning on the system headquarters, together with their seraphic consorts. All ex\hyp{}mortals above the status of Jerusem citizens and below that of finaliters are reckoned as belonging to the group having its headquarters in this circle.
\vs p046 5:19 These circular reservations of the Sons occupy an enormous area, and until 1900 years ago there existed a great open space at its centre. This central region is now occupied by the Michael memorial, completed some 500 years ago. 495 years ago, when this temple was dedicated, Michael was present in person, and all Jerusem heard the touching story of the Master Son’s bestowal on Urantia, the least of Satania. The Michael memorial is now the centre of all activities embraced in the modified management of the system occasioned by Michael’s bestowal, including most of the more recently transplanted Salvington activities. The memorial staff consists of over 1,000,000 personalities.
\vs p046 5:20 \ublistelem{2.}\bibnobreakspace \bibemph{The circles of the angels.} Like the residential area of the Sons, these circles of the angels consist of 7 concentric and successively elevated circles, each overlooking the inner areas.
\vs p046 5:21 \pc The 1\ts{st} circle of the angels is occupied by the Higher Personalities of the Infinite Spirit who may be stationed on the headquarters world --- Solitary Messengers and their associates. The 2\ts{nd} circle is dedicated to the messenger hosts, Technical Advisers, companions, inspectors, and recorders as they may chance to function on Jerusem from time to time. The 3\ts{rd} circle is held by the ministering spirits of the higher orders and groupings.
\vs p046 5:22 The 4\ts{th} circle is held by the administrator seraphim, and the seraphim serving in a local system like Satania are an “innumerable host of angels.” The 5\ts{th} circle is occupied by the planetary seraphim, while the 6\ts{th} is the home of the transition ministers. The 7\ts{th} circle is the tarrying sphere of certain unrevealed orders of seraphim. The recorders of all these groups of angels do not sojourn with their fellows, being domiciled in the Jerusem temple of records. All records are preserved in triplicate in this threefold hall of archives. On a system headquarters, records are always preserved in material, in morontia, and in spirit form.
\vs p046 5:23 These 7 circles are surrounded by the exhibit panorama of Jerusem, 5,000 standard miles in circumference, which is devoted to the presentation of the advancing status of the peopled worlds of Satania and is constantly revised so as to truly represent up\hyp{}to\hyp{}date conditions on the individual planets. I doubt not that this vast promenade overlooking the circles of the angels will be the first sight of Jerusem to claim your attention when you are permitted extended leisure on your earlier visits.
\vs p046 5:24 These exhibits are in charge of the native life of Jerusem, but they are assisted by the ascenders from the various Satania worlds who are tarrying on Jerusem en route to Edentia. The portrayal of planetary conditions and world progress is effected by many methods, some known to you, but mostly by techniques unknown on Urantia. These exhibits occupy the outer edge of this vast wall. The remainder of the promenade is almost entirely open, being highly and magnificently embellished.
\vs p046 5:25 \ublistelem{3.}\bibnobreakspace \bibemph{The circles of the Universe Aids} have the headquarters of the Evening Stars situated in the enormous central space. Here is located the system headquarters of Galantia, the associate head of this powerful group of superangels, being the first commissioned of all the ascendant Evening Stars. This is one of the most magnificent of all the administrative sectors of Jerusem, even though it is among the more recent constructions. This centre is 50 miles in diameter. The Galantia headquarters is a monolithic cast crystal, wholly transparent. These material\hyp{}morontia crystals are greatly appreciated by both morontia and material beings. The created Evening Stars exert their influence all over Jerusem, being possessed of such extrapersonality attributes. The entire world has been rendered spiritually fragrant since so many of their activities were transferred here from Salvington.
\vs p046 5:26 \ublistelem{4.}\bibnobreakspace \bibemph{The circles of the Master Physical Controllers.} The various orders of the Master Physical Controllers are concentrically arranged around the vast temple of power, wherein presides the power chief of the system in association with the chief of the Morontia Power Supervisors. This temple of power is one of two sectors on Jerusem where ascending mortals and midway creatures are not permitted. The other one is the dematerializing sector in the area of the Material Sons, a series of laboratories wherein the transport seraphim transform material beings into a state quite like that of the morontia order of existence.
\vs p046 5:27 \ublistelem{5.}\bibnobreakspace \bibemph{The circles of the ascending mortals.} The central area of the circles of the ascending mortals is occupied by a group of 619 planetary memorials representative of the inhabited worlds of the system, and these structures periodically undergo extensive changes. It is the privilege of the mortals from each world to agree, from time to time, upon certain of the alterations or additions to their planetary memorials. Many changes are even now being made in the Urantia structures. The centre of these 619 temples is occupied by a working model of Edentia and its many worlds of ascendant culture. This model is 40 miles in diameter and is an actual reproduction of the Edentia system, true to the original in every detail.
\vs p046 5:28 Ascenders enjoy their Jerusem services and take pleasure in observing the techniques of other groups. Everything done in these various circles is open to the full observation of all Jerusem.
\vs p046 5:29 The activities of such a world are of 3 distinct varieties: work, progress, and play. Stated otherwise, they are: service, study, and relaxation. The composite activities consist of social intercourse, group entertainment, and divine worship. There is great educational value in mingling with diverse groups of personalities, orders very different from one’s own fellows.
\vs p046 5:30 \ublistelem{6.}\bibnobreakspace \bibemph{The circles of the courtesy colonies.} The 7 circles of the courtesy colonies are graced by 3 enormous structures: the vast astronomic observatory of Jerusem, the gigantic art gallery of Satania, and the immense assembly hall of the reversion directors, the theatre of morontia activities devoted to rest and recreation.
\vs p046 5:31 The celestial artisans direct the spornagia and provide the host of creative decorations and monumental memorials which abound in every place of public assembly. The studios of these artisans are among the largest and most beautiful of all the matchless structures of this wonderful world. The other courtesy colonies maintain extensive and beautiful headquarters. Many of these buildings are constructed wholly of crystal gems. All the architectural worlds abound in crystals and the so\hyp{}called precious metals.
\vs p046 5:32 \ublistelem{7.}\bibnobreakspace \bibemph{The circles of the finaliters} have a unique structure at the centre. And this same vacant temple is found on every system headquarters world throughout Nebadon. This edifice on Jerusem is sealed with the insignia of Michael, and it bears this inscription: “Undedicated to the seventh stage of spirit --- to the eternal assignment.” Gabriel placed the seal on this temple of mystery, and none but Michael can or may break the seal of sovereignty affixed by the Bright and Morning Star. Some day you shall look upon this silent temple, even though you may not penetrate its mystery.
\vs p046 5:33 \pc \bibemph{Other Jerusem circles:} In addition to these residential circles there are on Jerusem numerous additional designated abodes.
\usection{The Executive\hyp{}Administrative Squares}
\vs p046 6:1 The executive\hyp{}administrative divisions of the system are located in the immense departmental squares, 1,000 in number. Each administrative unit is divided into 100 subdivisions of 10 subgroups each. These 1,000 squares are clustered in 10 grand divisions, thus constituting the following 10 administrative departments:
\vs p046 6:2 \ublistelem{1.}\bibnobreakspace Physical maintenance and material improvement, the domains of physical power and energy.
\vs p046 6:3 \ublistelem{2.}\bibnobreakspace Arbitration, ethics, and administrative adjudication.
\vs p046 6:4 \ublistelem{3.}\bibnobreakspace Planetary and local affairs.
\vs p046 6:5 \ublistelem{4.}\bibnobreakspace Constellation and universe affairs.
\vs p046 6:6 \ublistelem{5.}\bibnobreakspace Education and other Melchizedek activities.
\vs p046 6:7 \ublistelem{6.}\bibnobreakspace Planetary and system physical progress, the scientific domains of Satania activities.
\vs p046 6:8 \ublistelem{7.}\bibnobreakspace Morontia affairs.
\vs p046 6:9 \ublistelem{8.}\bibnobreakspace Pure spirit activities and ethics.
\vs p046 6:10 \ublistelem{9.}\bibnobreakspace Ascendant ministry.
\vs p046 6:11 \ublistelem{10.}\bibnobreakspace Grand universe philosophy.
\vs p046 6:12 \pc These structures are transparent; hence all system activities can be viewed even by student visitors.
\usection{The Rectangles --- The Spornagia}
\vs p046 7:1 The 1,000 \bibemph{rectangles} of Jerusem are occupied by the lower native life of the headquarters planet, and at their centre is situated the vast circular headquarters of the spornagia.
\vs p046 7:2 On Jerusem you will be amazed by the agricultural achievements of the wonderful spornagia. There the land is cultivated largely for aesthetic and ornamental effects. The spornagia are the landscape gardeners of the headquarters worlds, and they are both original and artistic in their treatment of the open spaces of Jerusem. They utilize both animals and numerous mechanical contrivances in the culture of the soil. They are intelligently expert in the employment of the power agencies of their realms as well as in the utilization of numerous orders of their lesser brethren of the lower animal creations, many of which are provided them on these special worlds. This order of animal life is now largely directed by the ascending midway creatures from the evolutionary spheres.
\vs p046 7:3 Spornagia are not Adjuster indwelt. They do not possess survival souls, but they do enjoy long lives, sometimes to the extent of 40,000--50,000 standard years. Their number is legion, and they afford physical ministry to all orders of universe personalities requiring material service.
\vs p046 7:4 \pc Although spornagia neither possess nor evolve survival souls, though they do not have personality, nevertheless, they do evolve an individuality which can experience reincarnation. When, with the passing of time, the physical bodies of these unique creatures deteriorate from usage and age, their creators, in collaboration with the Life Carriers, fabricate new bodies in which the old spornagia re\hyp{}establish their residences.
\vs p046 7:5 Spornagia are the only creatures in all the universe of Nebadon who experience this or any other sort of reincarnation. They are only reactive to the first five of the adjutant mind\hyp{}spirits; they are not responsive to the spirits of worship and wisdom. But the five\hyp{}adjutant mind equivalates to a totality or sixth reality level, and it is this factor which persists as an experiential identity.
\vs p046 7:6 \pc I am quite without comparisons in undertaking to describe these useful and unusual creatures as there are no animals on the evolutionary worlds comparable to them. They are not evolutionary beings, having been projected by the Life Carriers in their present form and status. They are bisexual and procreate as they are required to meet the needs of a growing population.
\vs p046 7:7 Perhaps I can best suggest to Urantia minds something of the nature of these beautiful and serviceable creatures by saying that they embrace the combined traits of a faithful horse and an affectionate dog and manifest an intelligence exceeding that of the highest type of chimpanzee. And they are very beautiful, as judged by the physical standards of Urantia. They are most appreciative of the attentions shown them by the material and semimaterial sojourners on these architectural worlds. They have a vision which permits them to recognize --- in addition to material beings --- the morontia creations, the lower angelic orders, midway creatures, and some of the lower orders of spirit personalities. They do not comprehend worship of the Infinite, nor do they grasp the import of the Eternal, but they do, through affection for their masters, join in the outward spiritual devotions of their realms.
\vs p046 7:8 \pc There are those who believe that, in a future universe age, these faithful spornagia will escape from their animal level of existence and attain a worthy evolutional destiny of progressive intellectual growth and even spiritual achievement.
\usection{The Jerusem Triangles}
\vs p046 8:1 The purely local and routine affairs of Jerusem are directed from the 100 \bibemph{triangles.} These units are clustered around the ten marvellous structures domiciling the local administration of Jerusem. The triangles are surrounded by the panoramic depiction of the system headquarters history. At present there is an erasure of over 2 standard miles in this circular story. This sector will be restored upon the readmission of Satania into the constellation family. Every provision for this event has been made by the decrees of Michael, but the tribunal of the Ancients of Days has not yet finished the adjudication of the affairs of the Lucifer rebellion. Satania may not come back into the full fellowship of Norlatiadek so long as it harbours archrebels, high created beings who have fallen from light into darkness.
\vs p046 8:2 When Satania can return to the constellation fold, then will come up for consideration the readmission of the isolated worlds into the system family of inhabited planets, accompanied by their restoration to the spiritual communion of the realms. But even if Urantia were restored to the system circuits, you would still be embarrassed by the fact that your whole system rests under a Norlatiadek quarantine partially segregating it from all other systems.
\vs p046 8:3 \pc But ere long, the adjudication of Lucifer and his associates will restore the Satania system to the Norlatiadek constellation, and subsequently, Urantia and the other isolated spheres will be restored to the Satania circuits, and again will such worlds enjoy the privileges of interplanetary communication and intersystem communion.
\vs p046 8:4 \pc There will come an end for rebels and rebellion. The Supreme Rulers are merciful and patient, but the law of deliberately nourished evil is universally and unerringly executed. “The wages of sin is death” --- eternal obliteration.
\vsetoff
\vs p046 8:5 [Presented by an Archangel of Nebadon.]
\quizlink
