\upaper{50}{The Planetary Princes}
\uminitoc{Mission of the Princes}
\uminitoc{Planetary Administration}
\uminitoc{The Prince’s Corporeal Staff}
\uminitoc{The Planetary Headquarters and Schools}
\uminitoc{Progressive Civilization}
\uminitoc{Planetary Culture}
\uminitoc{The Rewards of Isolation}
\author{Secondary Lanonandek}
\vs p050 0:1 While belonging to the order of Lanonandek Sons, the Planetary Princes are so specialized in service that they are commonly regarded as a distinct group. After their Melchizedek certification as secondary Lanonandeks, these local universe Sons are assigned to the reserves of their order on the constellation headquarters. From here they are assigned to various duties by the System Sovereign and eventually commissioned as Planetary Princes and sent forth to rule the evolving inhabited worlds.
\vs p050 0:2 The signal for a System Sovereign to act in the matter of assigning a ruler to a given planet is the reception of a request from the Life Carriers for the dispatch of an administrative head to function on this planet whereon they have established life and developed intelligent evolutionary beings. All planets which are inhabited by evolutionary mortal creatures have assigned to them a planetary ruler of this order of sonship.
\usection{Mission of the Princes}
\vs p050 1:1 The Planetary Prince and his assistant brethren represent the nearest personalized approach (aside from incarnation) that the Eternal Son of Paradise can make to the lowly creatures of time and space. True, the Creator Son touches the creatures of the realms through his spirit, but the Planetary Prince is the last of the orders of personal Sons extending out from Paradise to the children of men. The Infinite Spirit comes very near in the persons of the guardians of destiny and other angelic beings; the Universal Father lives in man by the prepersonal presence of the Mystery Monitors; but the Planetary Prince represents the last effort of the Eternal Son and his Sons to draw near you. On a newly inhabited world the Planetary Prince is the sole representative of complete divinity, springing from the Creator Son (the offspring of the Universal Father and the Eternal Son) and the Divine Minister (the universe Daughter of the Infinite Spirit).
\vs p050 1:2 The prince of a newly inhabited world is surrounded by a loyal corps of helpers and assistants and by large numbers of the ministering spirits. But the directing corps of such new worlds must be of the lower orders of the administrators of a system in order to be innately sympathetic with, and understanding of, the planetary problems and difficulties. And all of this effort to provide sympathetic rulership for the evolutionary worlds entails the increased liability that these near\hyp{}human personalities may be led astray by the exaltation of their own minds over and above the will of the Supreme Rulers.
\vs p050 1:3 Being quite alone as representatives of divinity on the individual planets, these Sons are tested severely, and Nebadon has suffered the misfortune of several rebellions. In the creation of the System Sovereigns and the Planetary Princes there occurs the personalization of a concept that has been getting farther and farther away from the Universal Father and the Eternal Son, and there is an increasing danger of losing the sense of proportion as to one’s self\hyp{}importance and a greater likelihood of failure to keep a proper grasp of the values and relationships of the numerous orders of divine beings and their gradations of authority. That the Father is not personally present in the local universe also imposes a certain test of faith and loyalty on all these Sons.
\vs p050 1:4 But not often do these world princes fail in their missions of organizing and administering the inhabited spheres, and their success greatly facilitates the subsequent missions of the Material Sons, who come to engraft the higher forms of creature life on the primitive men of the worlds. Their rule also does much to prepare the planets for the Paradise Sons of God, who subsequently come to judge the worlds and to inaugurate successive dispensations.
\usection{Planetary Administration}
\vs p050 2:1 All Planetary Princes are under the universe administrative jurisdiction of Gabriel, the chief executive of Michael, while in immediate authority they are subject to the executive mandates of the System Sovereigns.
\vs p050 2:2 The Planetary Princes may at any time seek the counsel of the Melchizedeks, their former instructors and sponsors, but they are not arbitrarily required to ask for such assistance, and if such aid is not voluntarily requested, the Melchizedeks do not interfere with the planetary administration. These world rulers may also avail themselves of the advice of the four and twenty counsellors, assembled from the bestowal worlds of the system. In Satania these counsellors are at present all natives of Urantia. And there is an analogous council of 70 at the constellation headquarters also selected from the evolutionary beings of the realms.
\vs p050 2:3 The rule of the evolutionary planets in their early and unsettled careers is largely autocratic. The Planetary Princes organize their specialized groups of assistants from among their corps of planetary aids. They usually surround themselves with a supreme council of twelve, but this is variously chosen and diversely constituted on the different worlds. A Planetary Prince may also have as assistants one or more of the third order of his own group of sonship and sometimes, on certain worlds, one of his own order, a secondary Lanonandek associate.
\vs p050 2:4 The entire staff of a world ruler consists of personalities of the Infinite Spirit and certain types of higher evolved beings and ascending mortals from other worlds. Such a staff averages about 1,000, and as the planet progresses, this corps of helpers may be increased up to 100,000 or more. At any time need is felt for more helpers, the Planetary Princes have only to make request of their brothers, the System Sovereigns, and the petition is granted forthwith.
\vs p050 2:5 Planets vary greatly in nature and organization and in administration, but all provide for tribunals of justice. The judicial system of the local universe has its beginnings in the tribunals of a Planetary Prince, which are presided over by a member of his personal staff; the decrees of such courts reflect a highly fatherly and discretionary attitude. All problems involving more than the regulation of the planetary inhabitants are subject to appeal to the higher tribunals, but the affairs of his world domain are largely adjusted in accordance with the personal discretion of the prince.
\vs p050 2:6 The roving commissions of conciliators serve and supplement the planetary tribunals, and both spirit and physical controllers are subject to the findings of these conciliators. But no arbitrary execution is ever carried out without the consent of the Constellation Father, for the “Most Highs rule in the kingdoms of men.”
\vs p050 2:7 The controllers and transformers of planetary assignment are also able to collaborate with angels and other orders of celestial beings in rendering these latter personalities visible to mortal creatures. On special occasions the seraphic helpers and even the Melchizedeks can and do make themselves visible to the inhabitants of the evolutionary worlds. The principal reason for bringing mortal ascenders from the system capital as a part of the staff of the Planetary Prince is to facilitate communication with the inhabitants of the realm.
\usection{The Prince’s Corporeal Staff}
\vs p050 3:1 On going to a young world, a Planetary Prince usually takes with him a group of volunteer ascending beings from the local system headquarters. These ascenders accompany the prince as advisers and helpers in the work of early race improvement. This corps of material helpers constitutes the connecting link between the prince and the world races. The Urantia Prince, Caligastia, had a corps of 100 such helpers.
\vs p050 3:2 \pc Such volunteer assistants are citizens of a system capital, and none of them have fused with their indwelling Adjusters. The status of the Adjusters of such volunteer servers remains as of the residential standing on the system headquarters while these morontia progressors temporarily revert to a former material state.
\vs p050 3:3 The Life Carriers, the architects of form, provide such volunteers with new physical bodies, which they occupy for the periods of their planetary sojourn. These personality forms, while exempt from the ordinary diseases of the realms, are, like the early morontia bodies, subject to certain accidents of a mechanical nature.
\vs p050 3:4 \pc The prince’s corporeal staff are usually removed from the planet in connection with the next adjudication at the time of the second Son’s arrival on the sphere. Before leaving, they customarily assign their various duties to their mutual offspring and to certain superior native volunteers. On those worlds where these helpers of the prince have been permitted to mate with the superior groups of the native races, such offspring usually succeed them.
\vs p050 3:5 These assistants to the Planetary Prince seldom mate with the world races, but they do always mate among themselves. Two classes of beings result from these unions: the primary type of midway creatures and certain high types of material beings who remain attached to the prince’s staff after their parents have been removed from the planet at the time of the arrival of Adam and Eve. These children do not mate with the mortal races except in certain emergencies and then only by direction of the Planetary Prince. In such an event, their children --- the grandchildren of the corporeal staff --- are in status as of the superior races of their day and generation. All the offspring of these semimaterial assistants of the Planetary Prince are Adjuster indwelt.
\vs p050 3:6 At the end of the prince’s dispensation, when the time comes for this “reversion staff” to be returned to the system headquarters for the resumption of the Paradise career, these ascenders present themselves to the Life Carriers for the purpose of yielding up their material bodies. They enter the transition slumber and awaken delivered from their mortal investment and clothed with morontia forms, ready for seraphic transportation back to the system capital, where their detached Adjusters await them. They are a whole dispensation behind their Jerusem class, but they have gained a unique and extraordinary experience, a rare chapter in the career of an ascending mortal.
\usection{The Planetary Headquarters and Schools}
\vs p050 4:1 The prince’s corporeal staff early organize the planetary schools of training and culture, wherein the cream of the evolutionary races are instructed and then sent forth to teach these better ways to their people. These schools of the prince are located at the material headquarters of the planet.
\vs p050 4:2 Much of the physical work connected with the establishment of this headquarters city is performed by the corporeal staff. Such headquarters cities, or settlements, of the early times of the Planetary Prince are very different from what a Urantia mortal might imagine. They are, in comparison with later ages, simple, being characterized by mineral embellishment and by relatively advanced material construction. And all of this stands in contrast with the Adamic regime centring around a garden headquarters, from which their work in behalf of the races is prosecuted during the second dispensation of the universe Sons.
\vs p050 4:3 \pc In the headquarters settlement on your world every human habitation was provided with abundance of land. Although the remote tribes continued in hunting and food foraging, the students and teachers in the Prince’s schools were all agriculturists and horticulturists. The time was about equally divided between the following pursuits:
\vs p050 4:4 \ublistelem{1.}\bibnobreakspace \bibemph{Physical labour.} Cultivation of the soil, associated with home building and embellishment.
\vs p050 4:5 \ublistelem{2.}\bibnobreakspace \bibemph{Social activities.} Play performances and cultural social groupings.
\vs p050 4:6 \ublistelem{3.}\bibnobreakspace \bibemph{Educational application.} Individual instruction in connection with family\hyp{}group teaching, supplemented by specialized class training.
\vs p050 4:7 \ublistelem{4.}\bibnobreakspace \bibemph{Vocational training.} Schools of marriage and homemaking, the schools of art and craft training, and the classes for the training of teachers --- secular, cultural, and religious.
\vs p050 4:8 \ublistelem{5.}\bibnobreakspace \bibemph{Spiritual culture.} The teacher brotherhood, the enlightenment of childhood and youth groups, and the training of adopted native children as missionaries to their people.
\vs p050 4:9 \pc A Planetary Prince is not visible to mortal beings; it is a test of faith to believe the representations of the semimaterial beings of his staff. But these schools of culture and training are well adapted to the needs of each planet, and there soon develops a keen and laudatory rivalry among the races of men in their efforts to gain entrance to these various institutions of learning.
\vs p050 4:10 From such a world centre of culture and achievement there gradually radiates to all peoples an uplifting and civilizing influence which slowly and certainly transforms the evolutionary races. Meantime the educated and spiritualized children of the surrounding peoples who have been adopted and trained in the prince’s schools are returning to their native groups and, to the best of their ability, are there establishing new and potent centres of learning and culture which they carry on according to the plan of the prince’s schools.
\vs p050 4:11 \pc On Urantia these plans for planetary progress and cultural advancement were well under way, proceeding most satisfactorily, when the whole enterprise was brought to a rather sudden and most inglorious end by Caligastia’s adherence to the Lucifer rebellion.
\vs p050 4:12 It was one of the most profoundly shocking episodes of this rebellion for me to learn of the callous perfidy of one of my own order of sonship, Caligastia, who, in deliberation and with malice aforethought, systematically perverted the instruction and poisoned the teaching provided in all the Urantia planetary schools in operation at that time. The wreck of these schools was speedy and complete.
\vs p050 4:13 Many of the offspring of the ascenders of the Prince’s materialized staff remained loyal, deserting the ranks of Caligastia. These loyalists were encouraged by the Melchizedek receivers of Urantia, and in later times their descendants did much to uphold the planetary concepts of truth and righteousness. The work of these loyal evangels helped to prevent the total obliteration of spiritual truth on Urantia. These courageous souls and their descendants kept alive some knowledge of the Father’s rule and preserved for the world races the concept of the successive planetary dispensations of the various orders of divine Sons.
\usection{Progressive Civilization}
\vs p050 5:1 The loyal princes of the inhabited worlds are permanently attached to the planets of their original assignment. Paradise Sons and their dispensations may come and go, but a successful Planetary Prince continues on as the ruler of his realm. His work is quite independent of the missions of the higher Sons, being designed to foster the development of planetary civilization.
\vs p050 5:2 The progress of civilization is hardly alike on any two planets. The details of the unfoldment of mortal evolution are very different on numerous dissimilar worlds. Notwithstanding these many diversifications of planetary development along physical, intellectual, and social lines, all evolutionary spheres progress in certain well\hyp{}defined directions.
\vs p050 5:3 Under the benign rule of a Planetary Prince, augmented by the Material Sons and punctuated by the periodic missions of the Paradise Sons, the mortal races on an average world of time and space will successively pass through the following seven developmental epochs:
\vs p050 5:4 \ublistelem{1.}\bibnobreakspace \bibemph{The nutrition epoch.} The prehuman creatures and the dawn races of primitive man are chiefly concerned with food problems. These evolving beings spend their waking hours either in seeking food or in fighting, offensively or defensively. The food quest is paramount in the minds of these early ancestors of subsequent civilization.
\vs p050 5:5 \ublistelem{2.}\bibnobreakspace \bibemph{The security age.} Just as soon as the primitive hunter can spare any time from the search for food, he turns this leisure to augmenting his security. More and more attention is devoted to the technique of war. Homes are fortified, and the clans are solidified by mutual fear and by the inculcation of hate for foreign groups. Self\hyp{}preservation is a pursuit which always follows self\hyp{}maintenance.
\vs p050 5:6 \ublistelem{3.}\bibnobreakspace \bibemph{The material\hyp{}comfort era.} After food problems have been partially solved and some degree of security has been attained, the additional leisure is utilized to promote personal comfort. Luxury vies with necessity in occupying the centre of the stage of human activities. Such an age is all too often characterized by tyranny, intolerance, gluttony, and drunkenness. The weaker elements of the races incline towards excesses and brutality. Gradually these pleasure\hyp{}seeking weaklings are subjugated by the more strong and truth\hyp{}loving elements of the advancing civilization.
\vs p050 5:7 \ublistelem{4.}\bibnobreakspace \bibemph{The quest for knowledge and wisdom.} Food, security, pleasure, and leisure provide the foundation for the development of culture and the spread of knowledge. The effort to execute knowledge results in wisdom, and when a culture has learned how to profit and improve by experience, civilization has really arrived. Food, security, and material comfort still dominate society, but many forward\hyp{}looking individuals are hungering for knowledge and thirsting for wisdom. Every child is provided an opportunity to learn by doing; education is the watchword of these ages.
\vs p050 5:8 \ublistelem{5.}\bibnobreakspace \bibemph{The epoch of philosophy and brotherhood.} When mortals learn to think and begin to profit by experience, they become philosophical --- they start out to reason within themselves and to exercise discriminative judgment. The society of this age becomes ethical, and the mortals of such an era are truly becoming moral beings. Wise moral beings are capable of establishing human brotherhood on such a progressing world. Ethical and moral beings can learn how to live in accordance with the golden rule.
\vs p050 5:9 \ublistelem{6.}\bibnobreakspace \bibemph{The age of spiritual striving.} When evolving mortals have passed through the physical, intellectual, and social stages of development, sooner or later they attain those levels of personal insight which impel them to seek for spiritual satisfactions and cosmic understandings. Religion is completing the ascent from the emotional domains of fear and superstition to the high levels of cosmic wisdom and personal spiritual experience. Education aspires to the attainment of meanings, and culture grasps at cosmic relationships and true values. Such evolving mortals are genuinely cultured, truly educated, and exquisitely God\hyp{}knowing.
\vs p050 5:10 \ublistelem{7.}\bibnobreakspace \bibemph{The era of light and life.} This is the flowering of the successive ages of physical security, intellectual expansion, social culture, and spiritual achievement. These human accomplishments are now blended, associated, and co\hyp{}ordinated in cosmic unity and unselfish service. Within the limitations of finite nature and material endowments there are no bounds set upon the possibilities of evolutionary attainment by the advancing generations who successively live upon these supernal and settled worlds of time and space.
\vs p050 5:11 \pc After serving their spheres through successive dispensations of world history and the progressing epochs of planetary progress, the Planetary Princes are elevated to the position of Planetary Sovereigns upon the inauguration of the era of light and life.
\usection{Planetary Culture}
\vs p050 6:1 The isolation of Urantia renders it impossible to undertake the presentation of many details of the life and environment of your Satania neighbours. In these presentations we are limited by the planetary quarantine and by the system isolation. We must be guided by these restrictions in all our efforts to enlighten Urantia mortals, but in so far as is permissible, you have been instructed in the progress of an average evolutionary world, and you are able to compare such a world’s career with the present state of Urantia.
\vs p050 6:2 The development of civilization on Urantia has not differed so greatly from that of other worlds which have sustained the misfortune of spiritual isolation. But when compared with the loyal worlds of the universe, your planet seems most confused and greatly retarded in all phases of intellectual progress and spiritual attainment.
\vs p050 6:3 Because of your planetary misfortunes, Urantians are prevented from understanding very much about the culture of normal worlds. But you should not envisage the evolutionary worlds, even the most ideal, as spheres whereon life is a flowery bed of ease. The initial life of the mortal races is always attended by struggle. Effort and decision are an essential part of the acquirement of survival values.
\vs p050 6:4 Culture presupposes quality of mind; culture cannot be enhanced unless mind is elevated. Superior intellect will seek a noble culture and find some way to attain such a goal. Inferior minds will spurn the highest culture even when presented to them ready\hyp{}made. Much depends, also, upon the successive missions of the divine Sons and upon the extent to which enlightenment is received by the ages of their respective dispensations.
\vs p050 6:5 \pc You should not forget that for 200,000 years all the worlds of Satania have rested under the spiritual ban of Norlatiadek in consequence of the Lucifer rebellion. And it will require age upon age to retrieve the resultant handicaps of sin and secession. Your world still continues to pursue an irregular and checkered career as a result of the double tragedy of a rebellious Planetary Prince and a defaulting Material Son. Even the bestowal of Christ Michael on Urantia did not immediately set aside the temporal consequences of these serious blunders in the earlier administration of the world.
\usection{The Rewards of Isolation}
\vs p050 7:1 On first thought it might appear that Urantia and its associated isolated worlds are most unfortunate in being deprived of the beneficent presence and influence of such superhuman personalities as a Planetary Prince and a Material Son and Daughter. But isolation of these spheres affords their races a unique opportunity for the exercise of faith and for the development of a peculiar quality of confidence in cosmic reliability which is not dependent on sight or any other material consideration. It may turn out, eventually, that mortal creatures hailing from the worlds quarantined in consequence of rebellion are extremely fortunate. We have discovered that such ascenders are very early entrusted with numerous special assignments to cosmic undertakings where unquestioned faith and sublime confidence are essential to achievement.
\vs p050 7:2 On Jerusem the ascenders from these isolated worlds occupy a residential sector by themselves and are known as the \bibemph{agondonters,} meaning evolutionary will creatures who can believe without seeing, persevere when isolated, and triumph over insuperable difficulties even when alone. This functional grouping of the agondonters persists throughout the ascension of the local universe and the traversal of the superuniverse; it disappears during the sojourn in Havona but promptly reappears upon the attainment of Paradise and definitely persists in the Corps of the Mortal Finality. Tabamantia is an \bibemph{agondonter} of finaliter status, having survived from one of the quarantined spheres involved in the first rebellion ever to take place in the universes of time and space.
\vs p050 7:3 All through the Paradise career, reward follows effort as the result of causes. Such rewards set off the individual from the average, provide a differential of creature experience, and contribute to the versatility of ultimate performances in the collective body of the finaliters.
\vsetoff
\vs p050 7:4 [Presented by a Secondary Lanonandek Son of the Reserve Corps.]
\quizlink
