\upaper{52}{Planetary Mortal Epochs}
\author{Mighty Messenger}
\vs p052 0:1 From the inception of life on an evolutionary planet to the time of its final flowering in the era of light and life, there appear upon the stage of world action at least seven epochs of human life. These successive ages are determined by the planetary missions of the divine Sons, and on an average inhabited world these epochs appear in the following order:
\vs p052 0:2 \ublistelem{1.}\bibnobreakspace Pre\hyp{}Planetary Prince Man.
\vs p052 0:3 \ublistelem{2.}\bibnobreakspace Post\hyp{}Planetary Prince Man.
\vs p052 0:4 \ublistelem{3.}\bibnobreakspace Post\hyp{}Adamic Man.
\vs p052 0:5 \ublistelem{4.}\bibnobreakspace Post\hyp{}Magisterial Son Man.
\vs p052 0:6 \ublistelem{5.}\bibnobreakspace Post\hyp{}Bestowal Son Man.
\vs p052 0:7 \ublistelem{6.}\bibnobreakspace Post\hyp{}Teacher Son Man.
\vs p052 0:8 \ublistelem{7.}\bibnobreakspace The Era of Light and Life.
\vs p052 0:9 \pc The worlds of space, as soon as they are physically suitable for life, are placed on the registry of the Life Carriers, and in due time these Sons are dispatched to such planets for the purpose of initiating life. The entire period from life initiation to the appearance of man is designated the prehuman era and precedes the successive mortal epochs considered in this narrative.
\usection{1.\bibnobreakspace Primitive Man}
\vs p052 1:1 From the time of man’s emergence from the animal level --- when he can choose to worship the Creator --- to the arrival of the Planetary Prince, mortal will creatures are called \bibemph{primitive men.} There are six basic types or races of primitive men, and these early peoples successively appear in the order of the spectrum colours, beginning with the red. The length of time consumed in this early life evolution varies greatly on the different worlds, ranging from 150,000 years to over 1,000,000 years of Urantia time.
\vs p052 1:2 The evolutionary races of colour --- red, orange, yellow, green, blue, and indigo --- begin to appear about the time that primitive man is developing a simple language and is beginning to exercise the creative imagination. By this time man is well accustomed to standing erect.
\vs p052 1:3 \pc Primitive men are mighty hunters and fierce fighters. The law of this age is the physical survival of the fittest; the government of these times is wholly tribal. During the early racial struggles on many worlds some of the evolutionary races are obliterated, as occurred on Urantia. Those who survive are usually subsequently blended with the later imported violet race, the Adamic peoples.
\vs p052 1:4 In the light of subsequent civilization, this era of primitive man is a long, dark, and bloody chapter. The ethics of the jungle and the morals of the primeval forests are not in keeping with the standards of later dispensations of revealed religion and higher spiritual development. On normal and nonexperimental worlds this epoch is very different from the prolonged and extraordinarily brutal struggles which characterized this age on Urantia. When you have emerged from your first world experience, you will begin to see why this long and painful struggle on the evolutionary worlds occurs, and as you go forward in the Paradise path, you will increasingly understand the wisdom of these apparently strange doings. But notwithstanding all the vicissitudes of the early ages of human emergence, the performances of primitive man represent a splendid, even a heroic, chapter in the annals of an evolutionary world of time and space.
\vs p052 1:5 \pc Early evolutionary man is not a colourful creature. In general, these primitive mortals are cave dwellers or cliff residents. They also build crude huts in the large trees. Before they acquire a high order of intelligence, the planets are sometimes overrun with the larger types of animals. But early in this era mortals learn to kindle and maintain fire, and with the increase of inventive imagination and the improvement in tools, evolving man soon vanquishes the larger and more unwieldy animals. The early races also make extensive use of the larger flying animals. These enormous birds are able to carry one or two average\hyp{}sized men for a nonstop flight of over 800\,km. On some planets these birds are of great service since they possess a high order of intelligence, often being able to speak many words of the languages of the realm. These birds are most intelligent, very obedient, and unbelievably affectionate. Such passenger birds have been long extinct on Urantia, but your early ancestors enjoyed their services.
\vs p052 1:6 \pc Man’s acquirement of ethical judgment, moral will, is usually coincident with the appearance of early language. Upon attaining the human level, after this emergence of mortal will, these beings become receptive to the temporary indwelling of the divine Adjusters, and upon death many are duly elected as survivors and sealed by the archangels for subsequent resurrection and Spirit fusion. The archangels always accompany the Planetary Princes, and a dispensational adjudication of the realm is simultaneous with the prince’s arrival.
\vs p052 1:7 All mortals who are indwelt by Thought Adjusters are potential worshippers; they have been “lighted by the true light,” and they possess capacity for seeking reciprocal contact with divinity. Nevertheless, the early or biologic religion of primitive man is largely a persistence of animal fear coupled with ignorant awe and tribal superstition. The survival of superstition in the Urantia races is hardly complimentary to your evolutionary development nor compatible with your otherwise splendid achievements in material progress. But this early fear religion serves a very valuable purpose in subduing the fiery tempers of these primitive creatures. It is the forerunner of civilization and the soil for the subsequent planting of the seeds of revealed religion by the Planetary Prince and his ministers.
\vs p052 1:8 \pc Within 100,000 years from the time man acquires erect posture, the Planetary Prince usually arrives, having been dispatched by the System Sovereign upon the report of the Life Carriers that will is functioning, even though comparatively few individuals have thus developed. Primitive mortals usually welcome the Planetary Prince and his visible staff; in fact, they often look upon them with awe and reverence, almost with worshipfulness, if they are not restrained.
\usection{2.\bibnobreakspace Post\hyp{}Planetary Prince Man}
\vs p052 2:1 With the arrival of the Planetary Prince a new dispensation begins. Government appears on earth, and the advanced tribal epoch is attained. Great social strides are made during a few thousand years of this regime. Under normal conditions mortals attain a high state of civilization during this age. They do not struggle so long in barbarism as did the Urantia races. But life on an inhabited world is so changed by rebellion that you can have little or no idea of such a regime on a normal planet.
\vs p052 2:2 The average length of this dispensation is around 500,000 years, some longer, some shorter. During this era the planet is established in the circuits of the system, and a full quota of seraphic and other celestial helpers is assigned to its administration. The Thought Adjusters come in increasing numbers, and the seraphic guardians amplify their regime of mortal supervision.
\vs p052 2:3 When the Planetary Prince arrives on a primitive world, the evolved religion of fear and ignorance prevails. The prince and his staff make the first revelations of higher truth and universe organization. These initial presentations of revealed religion are very simple, and they usually pertain to the affairs of the local system. Religion is wholly an evolutionary process prior to the arrival of the Planetary Prince. Subsequently, religion progresses by graduated revelation as well as by evolutionary growth. Each dispensation, each mortal epoch, receives an enlarged presentation of spiritual truth and religious ethics. The evolution of the religious capacity of receptivity in the inhabitants of a world largely determines their rate of spiritual advancement and the extent of religious revelation.
\vs p052 2:4 This dispensation witnesses a spiritual dawn, and the different races and their various tribes tend to develop specialized systems of religious and philosophic thought. There uniformly run through all of these racial religions two strains: the early fears of primitive men and the later revelations of the Planetary Prince. In some respects Urantians do not seem to have wholly emerged from this stage of planetary evolution. As you pursue this study, you will the more clearly discern how far your world departs from the average course of evolutionary progress and development.
\vs p052 2:5 \pc But the Planetary Prince is not “the Prince of Peace.” Racial struggles and tribal wars continue over into this dispensation but with diminishing frequency and severity. This is the great age of racial dispersion, and it culminates in a period of intense nationalism. Colour is the basis of tribal and national groupings, and the different races often develop separate languages. Each expanding group of mortals tends to seek isolation. This segregation is favoured by the existence of many languages. Before the unification of the several races their relentless warfare sometimes results in the obliteration of whole peoples; the orange and green men are particularly subject to such extinction.
\vs p052 2:6 On average worlds, during the latter part of the prince’s rule, national life begins to replace tribal organization or rather to be superimposed upon the existing tribal groupings. But the great social achievement of the prince’s epoch is the emergence of family life. Heretofore, human relationships have been chiefly tribal; now, the home begins to materialize.
\vs p052 2:7 This is the dispensation of the realization of sex equality. On some planets the male may rule the female; on others the reverse prevails. During this age normal worlds establish full equality of the sexes, this being preliminary to the fuller realization of the ideals of home life. This is the dawn of the golden age of the home. The idea of tribal rule gradually gives way to the dual concept of national life and family life.
\vs p052 2:8 During this age agriculture makes its appearance. The growth of the family idea is incompatible with the roving and unsettled life of the hunter. Gradually the practices of settled habitations and the cultivation of the soil become established. The domestication of animals and the development of home arts proceed apace. Upon reaching the apex of biologic evolution, a high level of civilization has been attained, but there is little development of a mechanical order; invention is the characteristic of the succeeding age.
\vs p052 2:9 \pc The races are purified and brought up to a high state of physical perfection and intellectual strength before the end of this era. The early development of a normal world is greatly helped by the plan of promoting the increase of the higher types of mortals with proportionate curtailment of the lower. And it is the failure of your early peoples to thus discriminate between these types that accounts for the presence of so many defective and degenerate individuals among the present\hyp{}day Urantia races.
\vs p052 2:10 One of the great achievements of the age of the prince is this restriction of the multiplication of mentally defective and socially unfit individuals. Long before the times of the arrival of the second Sons, the Adams, most worlds seriously address themselves to the tasks of race purification, something which the Urantia peoples have not even yet seriously undertaken.
\vs p052 2:11 This problem of race improvement is not such an extensive undertaking when it is attacked at this early date in human evolution. The preceding period of tribal struggles and rugged competition in race survival has weeded out most of the abnormal and defective strains. An idiot does not have much chance of survival in a primitive and warring tribal social organization. It is the false sentiment of your partially perfected civilizations that fosters, protects, and perpetuates the hopelessly defective strains of evolutionary human stocks.
\vs p052 2:12 It is neither tenderness nor altruism to bestow futile sympathy upon degenerated human beings, unsalvable abnormal and inferior mortals. There exist on even the most normal of the evolutionary worlds sufficient differences between individuals and between numerous social groups to provide for the full exercise of all those noble traits of altruistic sentiment and unselfish mortal ministry without perpetuating the socially unfit and the morally degenerate strains of evolving humanity. There is abundant opportunity for the exercise of tolerance and the function of altruism in behalf of those unfortunate and needy individuals who have not irretrievably lost their moral heritage and forever destroyed their spiritual birthright.
\usection{3.\bibnobreakspace Post\hyp{}Adamic Man}
\vs p052 3:1 When the original impetus of evolutionary life has run its biologic course, when man has reached the apex of animal development, there arrives the second order of sonship, and the second dispensation of grace and ministry is inaugurated. This is true on all evolutionary worlds. When the highest possible level of evolutionary life has been attained, when primitive man has ascended as far as possible in the biologic scale, a Material Son and Daughter always appear on the planet, having been dispatched by the System Sovereign.
\vs p052 3:2 Thought Adjusters are increasingly bestowed upon the post\hyp{}Adamic men, and in constantly augmented numbers these mortals attain capacity for subsequent Adjuster fusion. While functioning as descending Sons, the Adams do not possess Adjusters, but their planetary offspring --- direct and mixed --- become legitimate candidates for the reception, in due time, of the Mystery Monitors. By the termination of the post\hyp{}Adamic age the planet is in possession of its full quota of celestial ministers; only the fusion Adjusters are not yet universally bestowed.
\vs p052 3:3 \pc It is the prime purpose of the Adamic regime to influence evolving man to complete the transit from the hunter and herder stage of civilization to that of the agriculturist and horticulturist, to be later supplemented by the appearance of the urban and industrial adjuncts to civilization. 10,000 years of this dispensation of the biologic uplifters is sufficient to effect a marvellous transformation. 25,000 years of such an administration of the conjoint wisdom of the Planetary Prince and the Material Sons usually ripens the sphere for the advent of a Magisterial Son.
\vs p052 3:4 \pc This age usually witnesses the completion of the elimination of the unfit and the still further purification of the racial strains; on normal worlds the defective bestial tendencies are very nearly eliminated from the reproducing stocks of the realm.
\vs p052 3:5 The Adamic progeny never amalgamate with the inferior strains of the evolutionary races. Neither is it the divine plan for the Planetary Adam or Eve to mate, personally, with the evolutionary peoples. This race\hyp{}improvement project is the task of their progeny. But the offspring of the Material Son and Daughter are mobilized for generations before the racial\hyp{}amalgamation ministry is inaugurated.
\vs p052 3:6 The result of the gift of the Adamic life plasm to the mortal races is an immediate upstepping of intellectual capacity and an acceleration of spiritual progress. There is usually some physical improvement also. On an average world the post\hyp{}Adamic dispensation is an age of great invention, energy control, and mechanical development. This is the era of the appearance of multiform manufacture and the control of natural forces; it is the golden age of exploration and the final subduing of the planet. Much of the material progress of a world occurs during this time of the inauguration of the development of the physical sciences, just such an epoch as Urantia is now experiencing. Your world is a full dispensation and more behind the average planetary schedule.
\vs p052 3:7 By the end of the Adamic dispensation on a normal planet the races are practically blended, so that it can be truly proclaimed that “God has made of one blood all the nations,” and that his Son “has made of one colour all peoples.” The colour of such an amalgamated race is somewhat of an olive shade of the violet hue, the racial “white” of the spheres.
\vs p052 3:8 \pc Primitive man is for the most part carnivorous; the Material Sons and Daughters do not eat meat, but their offspring within a few generations usually gravitate to the omnivorous level, although whole groups of their descendants sometimes remain nonflesh eaters. This double origin of the post\hyp{}Adamic races explains how such blended human stocks exhibit anatomic vestiges belonging to both the herbivorous and carnivorous animal groups.
\vs p052 3:9 Within 10,000 years of racial amalgamation the resultant stocks show varying degrees of anatomic blend, some strains carrying more of the marks of the nonflesh\hyp{}eating ancestry, others exhibiting more of the distinguishing traits and physical characteristics of their carnivorous evolutionary progenitors. The majority of these world races soon become omnivorous, subsisting upon a wide range of viands from both the animal and vegetable kingdoms.
\vs p052 3:10 \pc The post\hyp{}Adamic epoch is the dispensation of internationalism. With the near completion of the task of race blending, nationalism wanes, and the brotherhood of man really begins to materialize. Representative government begins to take the place of the monarchial or paternal form of rulership. The educational system becomes world\hyp{}wide, and gradually the languages of the races give way to the tongue of the violet people. Universal peace and co\hyp{}operation are seldom attained until the races are fairly well blended, and until they speak a common language.
\vs p052 3:11 During the closing centuries of the post\hyp{}Adamic age there develops new interest in art, music, and literature, and this world\hyp{}wide awakening is the signal for the appearance of a Magisterial Son. The crowning development of this era is the universal interest in intellectual realities, true philosophy. Religion becomes less nationalistic, becomes more and more a planetary affair. New revelations of truth characterize these ages, and the Most Highs of the constellations begin to rule in the affairs of men. Truth is revealed up to the administration of the constellations.
\vs p052 3:12 Great ethical advancement characterizes this era; the brotherhood of man is the goal of its society. World\hyp{}wide peace --- the cessation of race conflict and national animosity --- is the indicator of planetary ripeness for the advent of the third order of sonship, the Magisterial Son.
\usection{4.\bibnobreakspace Post\hyp{}Magisterial Son Man}
\vs p052 4:1 On normal and loyal planets this age opens with the mortal races blended and biologically fit. There are no race or colour problems; literally all nations and races are of one blood. The brotherhood of man flourishes, and the nations are learning to live on earth in peace and tranquillity. Such a world stands on the eve of a great and culminating intellectual development.
\vs p052 4:2 \pc When an evolutionary world becomes thus ripe for the magisterial age, one of the high order of Avonal Sons makes his appearance on a magisterial mission. The Planetary Prince and the Material Sons are of local universe origin; the Magisterial Son hails from Paradise.
\vs p052 4:3 When the Paradise Avonals come to the mortal spheres on judicial actions, solely as dispensation adjudicators, they are never incarnated. But when they come on magisterial missions, at least the initial one, they are always incarnated, though they do not experience birth, neither do they die the death of the realm. They may live on for generations in those cases where they remain as rulers on certain planets. When their missions are concluded, they yield up their planetary lives and return to their former status of divine sonship.
\vs p052 4:4 Each new dispensation extends the horizon of revealed religion, and the Magisterial Sons extend the revelation of truth to portray the affairs of the local universe and all its tributaries.
\vs p052 4:5 \pc After the initial visitation of a Magisterial Son the races soon effect their economic liberation. The daily work required to sustain one’s independence would be represented by 2½ hours of your time. It is perfectly safe to liberate such ethical and intelligent mortals. Such refined peoples well know how to utilize leisure for self\hyp{}improvement and planetary advancement. This age witnesses the further purification of the racial stocks by the restriction of reproduction among the less fit and poorly endowed individuals.
\vs p052 4:6 The political government and social administration of the races continue to improve, self\hyp{}government being fairly well established by the end of this age. By self\hyp{}government we refer to the highest type of representative government. Such worlds advance and honour only those leaders and rulers who are most fit to bear social and political responsibilities.
\vs p052 4:7 During this epoch the majority of the world mortals are Adjuster indwelt. But even yet the bestowal of divine Monitors is not always universal. The Adjusters of fusion destiny are not yet bestowed upon all planetary mortals; it is still necessary for the will creatures to choose the Mystery Monitors.
\vs p052 4:8 During the closing ages of this dispensation, society begins to return to more simplified forms of living. The complex nature of an advancing civilization is running its course, and mortals are learning to live more naturally and effectively. And this trend increases with each succeeding epoch. This is the age of the flowering of art, music, and higher learning. The physical sciences have already reached their height of development. The termination of this age, on an ideal world, witnesses the fullness of a great religious awakening, a world\hyp{}wide spiritual enlightenment. And this extensive arousal of the spiritual natures of the races is the signal for the arrival of the bestowal Son and for the inauguration of the fifth mortal epoch.
\vs p052 4:9 \pc On many worlds it develops that the planet is not made ready for a bestowal Son by one magisterial mission; in that event there will be a second, even a succession of Magisterial Sons, each of whom will advance the races from one dispensation to another until the planet is made ready for the gift of the bestowal Son. On the second and subsequent missions the Magisterial Sons may or may not be incarnated. But no matter how many Magisterial Sons may appear --- and they may also come as such after the bestowal Son --- the advent of each one marks the end of one dispensation and the beginning of another.
\vs p052 4:10 \pc These dispensations of the Magisterial Sons cover anywhere from 25,000 to 50,000 years of Urantia time. Sometimes such an epoch is much shorter and in rare instances even longer. But in the fullness of time one of these same Magisterial Sons will be born as the Paradise bestowal Son.
\usection{5.\bibnobreakspace Post\hyp{}Bestowal Son Man}
\vs p052 5:1 When a certain standard of intellectual and spiritual development is attained on an inhabited world, a Paradise bestowal Son always arrives. On normal worlds he does not appear in the flesh until the races have ascended to the highest levels of intellectual development and ethical attainment. But on Urantia the bestowal Son, even your own Creator Son, appeared at the close of the Adamic dispensation, but that is not the usual order of events on the worlds of space.
\vs p052 5:2 When the worlds have become ripe for spiritualization, the bestowal Son arrives. These Sons always belong to the Magisterial or Avonal order except in that case, once in each local universe, when the Creator Son prepares for his terminal bestowal on some evolutionary world, as occurred when Michael of Nebadon appeared on Urantia to bestow himself upon your mortal races. Only 1 world in near 10,000,000 can enjoy such a gift; all other worlds are spiritually advanced by the bestowal of a Paradise Son of the Avonal order.
\vs p052 5:3 \pc The bestowal Son arrives on a world of high educational culture and encounters a race spiritually trained and prepared to assimilate advanced teachings and to appreciate the bestowal mission. This is an age characterized by the world\hyp{}wide pursuit of moral culture and spiritual truth. The mortal passion of this dispensation is the penetration of cosmic reality and communion with spiritual reality. The revelations of truth are extended to include the superuniverse. Entirely new systems of education and government grow up to supplant the crude regimes of former times. The joy of living takes on new colour, and the reactions of life are exalted to heavenly heights of tone and timbre.
\vs p052 5:4 The bestowal Son lives and dies for the spiritual uplift of the mortal races of a world. He establishes the “new and living way”; his life is an incarnation of Paradise truth in mortal flesh, that very truth --- even the Spirit of Truth --- in the knowledge of which men shall be free.
\vs p052 5:5 On Urantia the establishment of this “new and living way” was a matter of fact as well as of truth. The isolation of Urantia in the Lucifer rebellion had suspended the procedure whereby mortals can pass, upon death, directly to the shores of the mansion worlds. Before the days of Christ Michael on Urantia all souls slept on until the dispensational or special millennial resurrections. Even Moses was not permitted to go over to the other side until the occasion of a special resurrection, the fallen Planetary Prince, Caligastia, contesting such a deliverance. But ever since the day of Pentecost, Urantia mortals again may proceed directly to the morontia spheres.
\vs p052 5:6 \pc Upon the resurrection of a bestowal Son, on the third day after yielding up his incarnated life, he ascends to the right hand of the Universal Father, receives the assurance of the acceptance of the bestowal mission, and returns to the Creator Son at the headquarters of the local universe. Thereupon the bestowal Avonal and the Creator Michael send their joint spirit, the Spirit of Truth, into the bestowal world. This is the occasion when the “spirit of the triumphant Son is poured out upon all flesh.” The Universe Mother Spirit also participates in this bestowal of the Spirit of Truth, and concomitant therewith there issues the bestowal edict of the Thought Adjusters. Thereafter all normal\hyp{}minded will creatures of that world will receive Adjusters as soon as they attain the age of moral responsibility, of spiritual choice.
\vs p052 5:7 If such a bestowal Avonal should return to a world after the bestowal mission, he would not incarnate but would come “in glory with the seraphic hosts.”
\vs p052 5:8 \pc The postbestowal Son age may extend from 10,000 to 100,000 years. There is no arbitrary time allotted to any of these dispensational eras. This is a time of great ethical and spiritual progress. Under the spiritual influence of these ages, human character undergoes tremendous transformations and experiences phenomenal development. It becomes possible to put the golden rule into practical operation. The teachings of Jesus are really applicable to a mortal world which has had the preliminary training of the prebestowal Sons with their dispensations of character ennoblement and culture augmentation.
\vs p052 5:9 During this era the problems of disease and delinquency are virtually solved. Degeneracy has already been largely eliminated by selective reproduction. Disease has been practically mastered through the high resistant qualities of the Adamic strains and by the intelligent and world\hyp{}wide application of the discoveries of the physical sciences of preceding ages. The average length of life, during this period, climbs well above the equivalent of 300 years of Urantia time.
\vs p052 5:10 Throughout this epoch there is a gradual lessening of governmental supervision. True self\hyp{}government is beginning to function; fewer and fewer restrictive laws are necessary. The military branches of national resistance are passing away; the era of international harmony is really arriving. There are many nations, mostly determined by land distribution, but only one race, one language, and one religion. Mortal affairs are almost, but not quite, utopian. This truly is a great and glorious age!
\usection{6.\bibnobreakspace Urantia’s Post\hyp{}Bestowal Age}
\vs p052 6:1 The bestowal Son is the Prince of Peace. He arrives with the message, “Peace on earth and good will among men.” On normal worlds this is a dispensation of world\hyp{}wide peace; the nations no more learn war. But such salutary influences did not attend the coming of your bestowal Son, Christ Michael. Urantia is not proceeding in the normal order. Your world is out of step in the planetary procession. Your Master, when on earth, warned his disciples that his advent would not bring the usual reign of peace on Urantia. He distinctly told them that there would be “wars and rumours of wars,” and that nation would rise against nation. At another time he said, “Think not that I have come to bring peace upon earth.”
\vs p052 6:2 Even on normal evolutionary worlds the realization of the world\hyp{}wide brotherhood of man is not an easy accomplishment. On a confused and disordered planet like Urantia such an achievement requires a much longer time and necessitates far greater effort. Unaided social evolution can hardly achieve such happy results on a spiritually isolated sphere. Religious revelation is essential to the realization of brotherhood on Urantia. While Jesus has shown the way to the immediate attainment of spiritual brotherhood, the realization of social brotherhood on your world depends much on the achievement of the following personal transformations and planetary adjustments:
\vs p052 6:3 \ublistelem{1.}\bibnobreakspace \bibemph{Social fraternity.} Multiplication of international and interracial social contacts and fraternal associations through travel, commerce, and competitive play. Development of a common language and the multiplication of multilinguists. The racial and national interchange of students, teachers, industrialists, and religious philosophers.
\vs p052 6:4 \ublistelem{2.}\bibnobreakspace \bibemph{Intellectual cross\hyp{}fertilization.} Brotherhood is impossible on a world whose inhabitants are so primitive that they fail to recognize the folly of unmitigated selfishness. There must occur an exchange of national and racial literature. Each race must become familiar with the thought of all races; each nation must know the feelings of all nations. Ignorance breeds suspicion, and suspicion is incompatible with the essential attitude of sympathy and love.
\vs p052 6:5 \ublistelem{3.}\bibnobreakspace \bibemph{Ethical awakening.} Only ethical consciousness can unmask the immorality of human intolerance and the sinfulness of fratricidal strife. Only a moral conscience can condemn the evils of national envy and racial jealousy. Only moral beings will ever seek for that spiritual insight which is essential to living the golden rule.
\vs p052 6:6 \ublistelem{4.}\bibnobreakspace \bibemph{Political wisdom.} Emotional maturity is essential to self\hyp{}control. Only emotional maturity will ensure the substitution of international techniques of civilized adjudication for the barbarous arbitrament of war. Wise statesmen will sometime work for the welfare of humanity even while they strive to promote the interest of their national or racial groups. Selfish political sagacity is ultimately suicidal --- destructive of all those enduring qualities which ensure planetary group survival.
\vs p052 6:7 \ublistelem{5.}\bibnobreakspace \bibemph{Spiritual insight.} The brotherhood of man is, after all, predicated on the recognition of the fatherhood of God. The quickest way to realize the brotherhood of man on Urantia is to effect the spiritual transformation of present\hyp{}day humanity. The only technique for accelerating the natural trend of social evolution is that of applying spiritual pressure from above, thus augmenting moral insight while enhancing the soul capacity of every mortal to understand and love every other mortal. Mutual understanding and fraternal love are transcendent civilizers and mighty factors in the world\hyp{}wide realization of the brotherhood of man.
\vs p052 6:8 \pc If you could be transplanted from your backward and confused world to some normal planet now in the postbestowal Son age, you would think you had been translated to the heaven of your traditions. You would hardly believe that you were observing the normal evolutionary workings of a mortal sphere of human habitation. These worlds are in the spiritual circuits of their realm, and they enjoy all the advantages of the universe broadcasts and the reflectivity services of the superuniverse.
\usection{7.\bibnobreakspace Post\hyp{}Teacher Son Man}
\vs p052 7:1 The Sons of the next order to arrive on the average evolutionary world are the Trinity Teacher Sons, the divine Sons of the Paradise Trinity. Again we find Urantia out of step with its sister spheres in that your Jesus has promised to return. That promise he will certainly fulfil, but no one knows whether his second coming will precede or follow the appearances of Magisterial or Teacher Sons on Urantia.
\vs p052 7:2 The Teacher Sons come in groups to the spiritualizing worlds. A planetary Teacher Son is assisted and supported by 70 primary Sons, 12 secondary Sons, and 3 of the highest and most experienced of the supreme order of Daynals. This corps will remain for some time on the world, long enough to effect the transition from the evolutionary ages to the era of light and life --- not less than 1,000 years of planetary time and often considerably longer. This mission is a Trinity contribution to the antecedent efforts of all the divine personalities who have ministered to an inhabited world.
\vs p052 7:3 \pc The revelation of truth is now extended to the central universe and to Paradise. The races are becoming highly spiritual. A great people has evolved and a great age is approaching. The educational, economic, and administrative systems of the planet are undergoing radical transformations. New values and relationships are being established. The kingdom of heaven is appearing on earth, and the glory of God is being shed abroad in the world.
\vs p052 7:4 This is the dispensation when many mortals are translated from among the living. As the era of Trinity Teacher Sons progresses, the spiritual allegiance of the mortals of time becomes more and more universal. Natural death becomes less frequent as the Adjusters increasingly fuse with their subjects during the lifetime in the flesh. The planet eventually is classed as of the primary modified order of mortal ascension.
\vs p052 7:5 \pc Life during this era is pleasant and profitable. Degeneracy and the antisocial end products of the long evolutionary struggle have been virtually obliterated. The length of life approaches 500 Urantia years, and the reproductive rate of racial increase is intelligently controlled. An entirely new order of society has arrived. There are still great differences among mortals, but the state of society more nearly approaches the ideals of social brotherhood and spiritual equality. Representative government is vanishing, and the world is passing under the rule of individual self\hyp{}control. The function of government is chiefly directed to collective tasks of social administration and economic co\hyp{}ordination. The golden age is coming on apace; the temporal goal of the long and intense planetary evolutionary struggle is in sight. The reward of the ages is soon to be realized; the wisdom of the Gods is about to be manifested.
\vs p052 7:6 The physical administration of a world during this age requires about one hour each day on the part of every adult individual; that is, the equivalent of one Urantia hour. The planet is in close touch with universe affairs, and its people scan the latest broadcasts with the same keen interest you now manifest in the latest editions of your daily newspapers. These races are occupied with a thousand things of interest unknown on your world.
\vs p052 7:7 \pc Increasingly, true planetary allegiance to the Supreme Being grows. Generation after generation, more and more of the race step into line with those who practise justice and live mercy. Slowly but surely the world is being won to the joyous service of the Sons of God. The physical difficulties and material problems have been largely solved; the planet is ripening for advanced life and a more settled existence.
\vs p052 7:8 \pc From time to time throughout their dispensation, Teacher Sons continue to come to these peaceful worlds. They do not leave a world until they observe that the evolutionary plan, as it concerns that planet, is working smoothly. A Magisterial Son of judgment usually accompanies the Teacher Sons on their successive missions, while another such Son functions at the time of their departure, and these judicial actions continue from age to age throughout the duration of the mortal regime of time and space.
\vs p052 7:9 Each recurring mission of the Trinity Teacher Sons successively exalts such a supernal world to ever\hyp{}ascending heights of wisdom, spirituality, and cosmic illumination. But the noble natives of such a sphere are still finite and mortal. Nothing is perfect; nevertheless, there is evolving a quality of near perfection in the operation of an imperfect world and in the lives of its human inhabitants.
\vs p052 7:10 \pc The Trinity Teacher Sons may return many times to the same world. But sooner or later, in connection with the termination of one of their missions, the Planetary Prince is elevated to the position of Planetary Sovereign, and the System Sovereign appears to proclaim the entrance of such a world upon the era of light and life.
\vs p052 7:11 It was of the conclusion of the terminal mission of the Teacher Sons (at least that would be the chronology on a normal world) that John wrote: “I saw a new heaven and a new earth and the new Jerusalem coming down from God out of heaven, prepared as a princess adorned for the prince.”
\vs p052 7:12 This is the same renovated earth, the advanced planetary stage, that the olden seer envisioned when he wrote: “‘For, as the new heavens and the new earth, which I will make, shall remain before me, so shall you and your children survive; and it shall come to pass that from one new moon to another and from one Sabbath to another all flesh shall come to worship before me,’ says the Lord.”
\vs p052 7:13 It is the mortals of such an age who are described as “a chosen generation, a royal priesthood, a holy nation, an exalted people; and you shall show forth the praises of Him who has called you out of darkness into this marvellous light.”
\vs p052 7:14 \pc No matter what the special natural history of an individual planet may be, no difference whether a realm has been wholly loyal, tainted with evil, or cursed by sin --- no matter what the antecedents may be --- sooner or later the grace of God and the ministry of angels will usher in the day of the advent of the Trinity Teacher Sons; and their departure, following their final mission, will inaugurate this superb era of light and life.
\vs p052 7:15 All the worlds of Satania can join in the hope of the one who wrote: “Nevertheless we, according to His promise, look for a new heaven and a new earth, wherein dwells righteousness. Wherefore, beloved, seeing that you look for such things, be diligent that you may be found by Him in peace, without spot and blameless.”
\vs p052 7:16 \pc The departure of the Teacher Son corps, at the end of their first or some subsequent reign, ushers in the dawn of the era of light and life --- the threshold of the transition from time to the vestibule of eternity. The planetary realization of this era of light and life far more than equals the fondest expectations of Urantia mortals who have entertained no more farseeing concepts of the future life than those embraced within religious beliefs which depict heaven as the immediate destiny and final dwelling place of surviving mortals.
\vsetoff
\vs p052 7:17 [Sponsored by a Mighty Messenger temporarily attached to the staff of Gabriel.]
