\upaper{53}{The Lucifer Rebellion}
\uminitoc{The Leaders of Rebellion}
\uminitoc{The Causes of Rebellion}
\uminitoc{The Lucifer Manifesto}
\uminitoc{Outbreak of the Rebellion}
\uminitoc{Nature of the Conflict}
\uminitoc{A Loyal Seraphic Commander}
\uminitoc{History of the Rebellion}
\uminitoc{The Son of Man on Urantia}
\uminitoc{Present Status of the Rebellion}
\author{Manovandet Melchizedek}
\vs p053 0:1 Lucifer was a brilliant primary Lanonandek Son of Nebadon. He had experienced service in many systems, had been a high counsellor of his group, and was distinguished for wisdom, sagacity, and efficiency. Lucifer was number 37 of his order, and when commissioned by the Melchizedeks, he was designated as one of the 100 most able and brilliant personalities in more than 700,000 of his kind. From such a magnificent beginning, through evil and error, he embraced sin and now is numbered as one of three System Sovereigns in Nebadon who have succumbed to the urge of self and surrendered to the sophistry of spurious personal liberty --- rejection of universe allegiance and disregard of fraternal obligations, blindness to cosmic relationships.
\vs p053 0:2 In the universe of Nebadon, the domain of Christ Michael, there are 10,000 systems of inhabited worlds. In all the history of Lanonandek Sons, in all their work throughout these thousands of systems and at the universe headquarters, only three System Sovereigns have ever been found in contempt of the government of the Creator Son.
\usection{The Leaders of Rebellion}
\vs p053 1:1 Lucifer was not an ascendant being; he was a created Son of the local universe, and of him it was said: “You were perfect in all your ways from the day you were created till unrighteousness was found in you.” Many times had he been in counsel with the Most Highs of Edentia. And Lucifer reigned “upon the holy mountain of God,” the administrative mount of Jerusem, for he was the chief executive of a great system of 607 inhabited worlds.
\vs p053 1:2 Lucifer was a magnificent being, a brilliant personality; he stood next to the Most High Fathers of the constellations in the direct line of universe authority. Notwithstanding Lucifer’s transgression, subordinate intelligences refrained from showing him disrespect and disdain prior to Michael’s bestowal on Urantia. Even the archangel of Michael, at the time of Moses’ resurrection, “did not bring against him an accusing judgment but simply said, ‘the Judge rebuke you.’” Judgment in such matters belongs to the Ancients of Days, the rulers of the superuniverse.
\vs p053 1:3 Lucifer is now the fallen and deposed Sovereign of Satania. Self\hyp{}contemplation is most disastrous, even to the exalted personalities of the celestial world. Of Lucifer it was said: “Your heart was lifted up because of your beauty; you corrupted your wisdom because of your brightness.” Your olden prophet saw his sad estate when he wrote: “How are you fallen from heaven, O Lucifer, son of the morning! How are you cast down, you who dared to confuse the worlds!”
\vs p053 1:4 Very little was heard of Lucifer on Urantia owing to the fact that he assigned his first lieutenant, Satan, to advocate his cause on your planet. Satan was a member of the same primary group of Lanonandeks but had never functioned as a System Sovereign; he entered fully into the Lucifer insurrection. The “devil” is none other than Caligastia, the deposed Planetary Prince of Urantia and a Son of the secondary order of Lanonandeks. At the time Michael was on Urantia in the flesh, Lucifer, Satan, and Caligastia were leagued together to effect the miscarriage of his bestowal mission. But they signally failed.
\vs p053 1:5 Abaddon was the chief of the staff of Caligastia. He followed his master into rebellion and has ever since acted as chief executive of the Urantia rebels. Beelzebub was the leader of the disloyal midway creatures who allied themselves with the forces of the traitorous Caligastia.
\vs p053 1:6 \pc The dragon eventually became the symbolic representation of all these evil personages. Upon the triumph of Michael, “Gabriel came down from Salvington and bound the dragon (all the rebel leaders) for an age.”\fnst{We must not forget, however, the following verse, (Rev~20:3), which reads “\ldots\ and after that he must be loosed a little season.”} Of the Jerusem seraphic rebels it is written: “And the angels who kept not their first estate but left their own habitation, he has reserved in sure chains of darkness to the judgment of the great day.”
\usection{The Causes of Rebellion}
\vs p053 2:1 Lucifer and his first assistant, Satan, had reigned on Jerusem for more than 500,000 years when in their hearts they began to array themselves against the Universal Father and his then vicegerent Son, Michael.
\vs p053 2:2 There were no peculiar or special conditions in the system of Satania which suggested or favoured rebellion. It is our belief that the idea took origin and form in Lucifer’s mind, and that he might have instigated such a rebellion no matter where he might have been stationed. Lucifer first announced his plans to Satan, but it required several months to corrupt the mind of his able and brilliant associate. However, when once converted to the rebel theories, he became a bold and earnest advocate of “self\hyp{}assertion and liberty.”
\vs p053 2:3 \pc No one ever suggested rebellion to Lucifer. The idea of self\hyp{}assertion in opposition to the will of Michael and to the plans of the Universal Father, as they are represented in Michael, had its origin in his own mind. His relations with the Creator Son had been intimate and always cordial. At no time prior to the exaltation of his own mind did Lucifer openly express dissatisfaction about the universe administration. Notwithstanding his silence, for more than 100 years of standard time the Union of Days on Salvington had been reflectivating to Uversa that all was not at peace in Lucifer’s mind. This information was also communicated to the Creator Son and the Constellation Fathers of Norlatiadek.
\vs p053 2:4 Throughout this period Lucifer became increasingly critical of the entire plan of universe administration but always professed wholehearted loyalty to the Supreme Rulers. His first outspoken disloyalty was manifested on the occasion of a visit of Gabriel to Jerusem just a few days before the open proclamation of the Lucifer Declaration of Liberty. Gabriel was so profoundly impressed with the certainty of the impending outbreak that he went direct to Edentia to confer with the Constellation Fathers regarding the measures to be employed in case of open rebellion.
\vs p053 2:5 It is very difficult to point out the exact cause or causes which finally culminated in the Lucifer rebellion. We are certain of only one thing, and that is: Whatever these first beginnings were, they had their origin in Lucifer’s mind. There must have been a pride of self that nourished itself to the point of self\hyp{}deception, so that Lucifer for a time really persuaded himself that his contemplation of rebellion was actually for the good of the system, if not of the universe. By the time his plans had developed to the point of disillusionment, no doubt he had gone too far for his original and mischief\hyp{}making pride to permit him to stop. At some point in this experience he became insincere, and evil evolved into deliberate and wilful sin. That this happened is proved by the subsequent conduct of this brilliant executive. He was long offered opportunity for repentance, but only some of his subordinates ever accepted the proffered mercy. The Faithful of Days of Edentia, on the request of the Constellation Fathers, in person presented the plan of Michael for the saving of these flagrant rebels, but always was the mercy of the Creator Son rejected and rejected with increasing contempt and disdain.
\usection{The Lucifer Manifesto}
\vs p053 3:1 Whatever the early origins of trouble in the hearts of Lucifer and Satan, the final outbreak took form as the Lucifer Declaration of Liberty. The cause of the rebels was stated under three heads:
\vs p053 3:2 \ublistelem{1.}\bibnobreakspace \bibemph{The reality of the Universal Father.} Lucifer charged that the Universal Father did not really exist, that physical gravity and space\hyp{}energy were inherent in the universe, and that the Father was a myth invented by the Paradise Sons to enable them to maintain the rule of the universes in the Father’s name. He denied that personality was a gift of the Universal Father. He even intimated that the finaliters were in collusion with the Paradise Sons to foist fraud upon all creation since they never brought back a very clear\hyp{}cut idea of the Father’s actual personality as it is discernible on Paradise. He traded on reverence as ignorance. The charge was sweeping, terrible, and blasphemous. It was this veiled attack upon the finaliters that no doubt influenced the ascendant citizens then on Jerusem to stand firm and remain steadfast in resistance to all the rebel’s proposals.
\vs p053 3:3 \ublistelem{2.}\bibnobreakspace \bibemph{The universe government of the Creator Son --- Michael.} Lucifer contended that the local systems should be autonomous. He protested against the right of Michael, the Creator Son, to assume sovereignty of Nebadon in the name of a hypothetical Paradise Father and require all personalities to acknowledge allegiance to this unseen Father. He asserted that the whole plan of worship was a clever scheme to aggrandize the Paradise Sons. He was willing to acknowledge Michael as his Creator\hyp{}father but not as his God and rightful ruler.
\vs p053 3:4 Most bitterly did he attack the right of the Ancients of Days --- “foreign potentates” --- to interfere in the affairs of the local systems and universes. These rulers he denounced as tyrants and usurpers. He exhorted his followers to believe that none of these rulers could do aught to interfere with the operation of complete home rule if men and angels only had the courage to assert themselves and boldly claim their rights.
\vs p053 3:5 He contended that the executioners of the Ancients of Days could be debarred from functioning in the local systems if the native beings would only assert their independence. He maintained that immortality was inherent in the system personalities, that resurrection was natural and automatic, and that all beings would live eternally except for the arbitrary and unjust acts of the executioners of the Ancients of Days.
\vs p053 3:6 \ublistelem{3.}\bibnobreakspace \bibemph{The attack upon the universal plan of ascendant mortal training.} Lucifer maintained that far too much time and energy were expended upon the scheme of so thoroughly training ascending mortals in the principles of universe administration, principles which he alleged were unethical and unsound. He protested against the agelong program for preparing the mortals of space for some unknown destiny and pointed to the presence of the finaliter corps on Jerusem as proof that these mortals had spent ages of preparation for some destiny of pure fiction. With derision he pointed out that the finaliters had encountered a destiny no more glorious than to be returned to humble spheres similar to those of their origin. He intimated that they had been debauched by overmuch discipline and prolonged training, and that they were in reality traitors to their mortal fellows since they were now co\hyp{}operating with the scheme of enslaving all creation to the fictions of a mythical eternal destiny for ascending mortals. He advocated that ascenders should enjoy the liberty of individual self\hyp{}determination. He challenged and condemned the entire plan of mortal ascension as sponsored by the Paradise Sons of God and supported by the Infinite Spirit.
\vs p053 3:7 \pc And it was with such a Declaration of Liberty that Lucifer launched his orgy of darkness and death.
\usection{Outbreak of the Rebellion}
\vs p053 4:1 The Lucifer manifesto was issued at the annual conclave of Satania on the sea of glass, in the presence of the assembled hosts of Jerusem, on the last day of the year, about 200,000 years ago, Urantia time. Satan proclaimed that worship could be accorded the universal forces --- physical, intellectual, and spiritual --- but that allegiance could be acknowledged only to the actual and present ruler, Lucifer, the “friend of men and angels” and the “God of liberty.”
\vs p053 4:2 Self\hyp{}assertion was the battle cry of the Lucifer rebellion. One of his chief arguments was that, if self\hyp{}government was good and right for the Melchizedeks and other groups, it was equally good for all orders of intelligence. He was bold and persistent in the advocacy of the “equality of mind” and “the brotherhood of intelligence.” He maintained that all government should be limited to the local planets and their voluntary confederation into the local systems. All other supervision he disallowed. He promised the Planetary Princes that they should rule the worlds as supreme executives. He denounced the location of legislative activities on the constellation headquarters and the conduct of judicial affairs on the universe capital. He contended that all these functions of government should be concentrated on the system capitals and proceeded to set up his own legislative assembly and organized his own tribunals under the jurisdiction of Satan. And he directed that the princes on the apostate worlds do the same.
\vs p053 4:3 The entire administrative cabinet of Lucifer went over in a body and were sworn in publicly as the officers of the administration of the new head of “the liberated worlds and systems.”
\vs p053 4:4 \pc While there had been two previous rebellions in Nebadon, they were in distant constellations. Lucifer held that these insurrections were unsuccessful because the majority of the intelligences failed to follow their leaders. He contended that “majorities rule,”\fnst{Hence we learn that Lucifer was the first protagonist of democracy in our Universe.} that “mind is infallible.” The freedom allowed him by the universe rulers apparently sustained many of his nefarious contentions. He defied all his superiors; yet they apparently took no note of his doings. He was given a free hand to prosecute his seductive plan without let or hindrance.
\vs p053 4:5 \pc All the merciful delays of justice Lucifer pointed to as evidence of the inability of the government of the Paradise Sons to stop the rebellion. He would openly defy and arrogantly challenge Michael, Immanuel, and the Ancients of Days and then point to the fact that no action ensued as positive evidence of the impotency of the universe and the superuniverse governments.
\vs p053 4:6 Gabriel was personally present throughout all these disloyal proceedings and only announced that he would, in due time, speak for Michael, and that all beings would be left free and unmolested in their choice; that the “government of the Sons for the Father desired only that loyalty and devotion which was voluntary, wholehearted, and sophistry\hyp{}proof.”
\vs p053 4:7 \pc Lucifer was permitted fully to establish and thoroughly to organize his rebel government before Gabriel made any effort to contest the right of secession or to counterwork the rebel propaganda. But the Constellation Fathers immediately confined the action of these disloyal personalities to the system of Satania. Nevertheless, this period of delay was a time of great trial and testing to the loyal beings of all Satania. All was chaotic for a few years, and there was great confusion on the mansion worlds.
\usection{Nature of the Conflict}
\vs p053 5:1 Upon the outbreak of the Satania rebellion, Michael took counsel of his Paradise brother, Immanuel. Following this momentous conference, Michael announced that he would pursue the same policy which had characterized his dealings with similar upheavals in the past, an attitude of noninterference.
\vs p053 5:2 \pc At the time of this rebellion and the two which preceded it there was no absolute and personal sovereign authority in the universe of Nebadon. Michael ruled by divine right, as vicegerent of the Universal Father, but not yet in his own personal right. He had not completed his bestowal career; he had not yet been vested with “all power in heaven and on earth.”
\vs p053 5:3 From the outbreak of rebellion to the day of his enthronement as sovereign ruler of Nebadon, Michael never interfered with the rebel forces of Lucifer; they were allowed to run a free course for almost 200,000 years of Urantia time. Christ Michael now has ample power and authority to deal promptly, even summarily, with such outbreaks of disloyalty, but we doubt that this sovereign authority would lead him to act differently if another such upheaval should occur.
\vs p053 5:4 \pc Since Michael elected to remain aloof from the actual warfare of the Lucifer rebellion, Gabriel called his personal staff together on Edentia and, in counsel with the Most Highs, elected to assume command of the loyal hosts of Satania. Michael remained on Salvington while Gabriel proceeded to Jerusem, and establishing himself on the sphere dedicated to the Father --- the same Universal Father whose personality Lucifer and Satan had questioned --- in the presence of the forgathered hosts of loyal personalities, he displayed the banner of Michael, the material emblem of the Trinity government of all creation, the three azure blue concentric circles\fnst{Many prophets (including the present editor) have visualised the evolving morontia soul as a blue sphere of light, see for example the presentation ``Glorious Ascent'' by Troy R.~Bishop.} on a white background.
\vs p053 5:5 The Lucifer emblem was a banner of white with one red circle, in the centre of which a black solid circle appeared.
\vs p053 5:6 “There was war in heaven; Michael’s commander and his angels fought against the dragon (Lucifer, Satan, and the apostate princes); and the dragon and his rebellious angels fought but prevailed not.” This “war in heaven” was not a physical battle as such a conflict might be conceived on Urantia. In the early days of the struggle Lucifer held forth continuously in the planetary amphitheatre. Gabriel conducted an unceasing exposure of the rebel sophistries from his headquarters taken up near at hand. The various personalities present on the sphere who were in doubt as to their attitude would journey back and forth between these discussions until they arrived at a final decision.
\vs p053 5:7 But this war in heaven was very terrible and very real. While displaying none of the barbarities so characteristic of physical warfare on the immature worlds, this conflict was far more deadly; material life is in jeopardy in material combat, but the war in heaven was fought in terms of life eternal.
\usection{A Loyal Seraphic Commander}
\vs p053 6:1 There were many noble and inspiring acts of devotion and loyalty which were performed by numerous personalities during the interim between the outbreak of hostilities and the arrival of the new system ruler and his staff. But the most thrilling of all these daring feats of devotion was the courageous conduct of Manotia, the second in command of the Satania headquarters’ seraphim.
\vs p053 6:2 At the outbreak of rebellion on Jerusem the head of the seraphic hosts joined the Lucifer cause. This no doubt explains why such a large number of the fourth order, the system administrator seraphim, went astray. The seraphic leader was spiritually blinded by the brilliant personality of Lucifer; his charming ways fascinated the lower orders of celestial beings. They simply could not comprehend that it was possible for such a dazzling personality to go wrong.
\vs p053 6:3 \pc Not long since, in describing the experiences associated with the onset of the Lucifer rebellion, Manotia said: “But my most exhilarating moment was the thrilling adventure connected with the Lucifer rebellion when, as second seraphic commander, I refused to participate in the projected insult to Michael; and the powerful rebels sought my destruction by means of the liaison forces they had arranged. There was a tremendous upheaval on Jerusem, but not a single loyal seraphim was harmed.
\vs p053 6:4 “Upon the default of my immediate superior it devolved upon me to assume command of the angelic hosts of Jerusem as the titular director of the confused seraphic affairs of the system. I was morally upheld by the Melchizedeks, ably assisted by a majority of the Material Sons, deserted by a tremendous group of my own order, but magnificently supported by the ascendant mortals on Jerusem.
\vs p053 6:5 “Having been automatically thrown out of the constellation circuits by the secession of Lucifer, we were dependent on the loyalty of our intelligence corps, who forwarded calls for help to Edentia from the near\hyp{}by system of Rantulia; and we found that the kingdom of order, the intellect of loyalty, and the spirit of truth were inherently triumphant over rebellion, self\hyp{}assertion, and so\hyp{}called personal liberty; we were able to carry on until the arrival of the new System Sovereign, the worthy successor of Lucifer. And immediately thereafter I was assigned to the corps of the Melchizedek receivership of Urantia, assuming jurisdiction over the loyal seraphic orders on the world of the traitorous Caligastia, who had proclaimed his sphere a member of the newly projected system of ‘liberated worlds and emancipated personalities’ proposed in the infamous Declaration of Liberty issued by Lucifer in his call to the ‘liberty\hyp{}loving, free\hyp{}thinking, and forward\hyp{}looking intelligences of the misruled and maladministered worlds of Satania.’”
\vs p053 6:6 \pc This angel is still in service on Urantia, functioning as associate chief of seraphim.
\usection{History of the Rebellion}
\vs p053 7:1 The Lucifer rebellion was system wide. 37 seceding Planetary Princes swung their world administrations largely to the side of the archrebel. Only on Panoptia did the Planetary Prince fail to carry his people with him. On this world, under the guidance of the Melchizedeks, the people rallied to the support of Michael. Ellanora\fnst{Possibly derived from the Greek \textit{eleos} ``mercy'' and Hebrew \textit{nur} ``shining'', combined to give the meaning ``the one shining with mercy''.}, a young woman of that mortal realm, grasped the leadership of the human races, and not a single soul on that strife\hyp{}torn world enlisted under the Lucifer banner. And ever since have these loyal Panoptians served on the seventh Jerusem transition world as the caretakers and builders on the Father’s sphere and its surrounding seven detention worlds. The Panoptians not only act as the literal custodians of these worlds, but they also execute the personal orders of Michael for the embellishment of these spheres for some future and unknown use. They do this work as they tarry en route to Edentia.
\vs p053 7:2 Throughout this period Caligastia was advocating the cause of Lucifer on Urantia. The Melchizedeks ably opposed the apostate Planetary Prince, but the sophistries of unbridled liberty and the delusions of self\hyp{}assertion had every opportunity for deceiving the primitive peoples of a young and undeveloped world.
\vs p053 7:3 All secession propaganda had to be carried on by personal effort because the broadcast service and all other avenues of interplanetary communication were suspended by the action of the system circuit supervisors. Upon the actual outbreak of the insurrection the entire system of Satania was isolated in both the constellation and the universe circuits. During this time all incoming and outgoing messages were dispatched by seraphic agents and Solitary Messengers. The circuits to the fallen worlds were also cut off, so that Lucifer could not utilize this avenue for the furtherance of his nefarious scheme. And these circuits will not be restored so long as the archrebel lives within the confines of Satania.
\vs p053 7:4 This was a Lanonandek rebellion. The higher orders of local universe sonship did not join the Lucifer secession, although a few of the Life Carriers stationed on the rebel planets were somewhat influenced by the rebellion of the disloyal princes. None of the Trinitized Sons went astray. The Melchizedeks, archangels, and the Brilliant Evening Stars were all loyal to Michael and, with Gabriel, valiantly contended for the Father’s will and the Son’s rule.
\vs p053 7:5 No beings of Paradise origin were involved in disloyalty. Together with the Solitary Messengers they took up headquarters on the world of the Spirit and remained under the leadership of the Faithful of Days of Edentia. None of the conciliators apostatized, nor did a single one of the Celestial Recorders go astray. But a heavy toll was taken of the Morontia Companions and the Mansion World Teachers.
\vs p053 7:6 Of the supreme order of seraphim, not an angel was lost, but a considerable group of the next order, the superior, were deceived and ensnared. Likewise a few of the third or supervisor order of angels were misled. But the terrible breakdown came in the fourth group, the administrator angels, those seraphim who are normally assigned to the duties of the system capitals. Manotia saved almost \bibfrac{2}{3}\ts{rds} of them, but slightly over \bibfrac{1}{3}\ts{rd} followed their chief into the rebel ranks. \bibfrac{1}{3}\ts{rd} of all the Jerusem cherubim attached to the administrator angels were lost with their disloyal seraphim.
\vs p053 7:7 Of the planetary angelic helpers, those assigned to the Material Sons, about \bibfrac{1}{3}\ts{rd} were deceived, and almost 10\% of the transition ministers were ensnared. In symbol John saw this when he wrote\fnst{Note that the quotation from Revelation~12:4 is not exact. The Greek (and almost all other) sources say ``cast them into the earth'' and not ``cast them down in darkness''. However, it is worth mentioning that some ancient Armenian MSS omit the phrase ``into the earth'' and instead say ``from heaven''.} of the great red dragon, saying: “And his tail drew a third part of the stars of heaven and cast them down in darkness.”
\vs p053 7:8 The greatest loss occurred in the angelic ranks, but most of the lower orders of intelligence were involved in disloyalty. Of the 681,217\fnc{In 1955 text ``681,227''. The change was made because in \bibref[51:1.5]{p051 1:5} it is stated that the total number of Material Sons lost is $13+681,204=681,217$.}} Material Sons lost in Satania, 95\% were casualties of the Lucifer rebellion. Large numbers of midway creatures were lost on those individual planets whose Planetary Princes joined the Lucifer cause.
\vs p053 7:9 \pc In many respects this rebellion was the most widespread and disastrous of all such occurrences in Nebadon. More personalities were involved in this insurrection than in both of the others. And it is to their everlasting dishonour that the emissaries of Lucifer and Satan spared not the infant\hyp{}training schools on the finaliter cultural planet but rather sought to corrupt these developing minds in mercy salvaged from the evolutionary worlds.
\vs p053 7:10 \pc The ascending mortals were vulnerable, but they withstood the sophistries of rebellion better than the lower spirits. While many on the lower mansion worlds, those who had not attained final fusion with their Adjusters, fell, it is recorded to the glory of the wisdom of the ascension scheme that not a single member of the Satania ascendant citizenship resident on Jerusem participated in the Lucifer rebellion.
\vs p053 7:11 Hour by hour and day by day the broadcast stations of all Nebadon were thronged by the anxious watchers of every imaginable class of celestial intelligence, who intently perused the bulletins of the Satania rebellion and rejoiced as the reports continuously narrated the unswerving loyalty of the ascending mortals who, under their Melchizedek leadership, successfully withstood the combined and protracted efforts of all the subtle evil forces which so swiftly gathered around the banners of secession and sin.
\vs p053 7:12 It was over two years of system time from the beginning of the “war in heaven” until the installation of Lucifer’s successor. But at last the new Sovereign came, landing on the sea of glass with his staff. I was among the reserves mobilized on Edentia by Gabriel, and I well remember the first message of Lanaforge to the Constellation Father of Norlatiadek. It read: “Not a single Jerusem citizen was lost. Every ascendant mortal survived the fiery trial and emerged from the crucial test triumphant and altogether victorious.” And on to Salvington, Uversa, and Paradise went this message of assurance that the survival experience of mortal ascension is the greatest security against rebellion and the surest safeguard against sin. This noble Jerusem band of faithful mortals numbered just 187,432,811.
\vs p053 7:13 \pc With the arrival of Lanaforge the archrebels were dethroned and shorn of all governing powers, though they were permitted freely to go about Jerusem, the morontia spheres, and even to the individual inhabited worlds. They continued their deceptive and seductive efforts to confuse and mislead the minds of men and angels. But as concerned their work on the administrative mount of Jerusem, “their place was found no more.”
\vs p053 7:14 \pc While Lucifer was deprived of all administrative authority in Satania, there then existed no local universe power nor tribunal which could detain or destroy this wicked rebel; at that time Michael was not a sovereign ruler. The Ancients of Days sustained the Constellation Fathers in their seizure of the system government, but they have never handed down any subsequent decisions in the many appeals still pending with regard to the present status and future disposition of Lucifer, Satan, and their associates.
\vs p053 7:15 Thus were these archrebels allowed to roam the entire system to seek further penetration for their doctrines of discontent and self\hyp{}assertion. But in almost 200,000 Urantia years they have been unable to deceive another world. No Satania worlds have been lost since the fall of the 37, not even those younger worlds peopled since that day of rebellion.
\usection{The Son of Man on Urantia}
\vs p053 8:1 Lucifer and Satan freely roamed the Satania system until the completion of the bestowal mission of Michael on Urantia. They were last on your world together during the time of their combined assault upon the Son of Man.
\vs p053 8:2 Formerly, when the Planetary Princes, the “Sons of God,” were periodically assembled, “Satan came also,” claiming that he represented all of the isolated worlds of the fallen Planetary Princes. But he has not been accorded such liberty on Jerusem since Michael’s terminal bestowal. Subsequent to their effort to corrupt Michael when in the bestowal flesh, all sympathy for Lucifer and Satan has perished throughout all Satania, that is, outside the isolated worlds of sin.
\vs p053 8:3 \pc The bestowal of Michael terminated the Lucifer rebellion in all Satania aside from the planets of the apostate Planetary Princes. And this was the significance of Jesus’ personal experience, just before his death in the flesh, when he one day exclaimed to his disciples, \textcolour{ubdarkred}{“And I beheld Satan fall as lightning from heaven.”} He had come with Lucifer to Urantia for the last crucial struggle.
\vs p053 8:4 The Son of Man was confident of success, and he knew that his triumph on your world would forever settle the status of his agelong enemies, not only in Satania but also in the other two systems where sin had entered. There was survival for mortals and security for angels when your Master, in reply to the Lucifer proposals, calmly and with divine assurance replied, \textcolour{ubdarkred}{“Get you behind me, Satan.”} That was, in principle, the real end of the Lucifer rebellion. True, the Uversa tribunals have not yet rendered the executive decision regarding the appeal of Gabriel praying for the destruction of the rebels, but such a decree will, no doubt, be forthcoming in the fullness of time since the first step in the hearing of this case has already been taken.
\vs p053 8:5 Caligastia was recognized by the Son of Man as the technical Prince of Urantia up to near the time of his death. Said Jesus: \textcolour{ubdarkred}{“Now is the judgment of this world; now shall the prince of this world be cast down.”} And then still nearer the completion of his lifework he announced, \textcolour{ubdarkred}{“The prince of this world is judged.”} And it is this same dethroned and discredited Prince who was once termed “God of Urantia.”
\vs p053 8:6 \pc The last act of Michael before leaving Urantia was to offer mercy to Caligastia and Daligastia\fnst{Since the English letter `D' follows `C' in alphabetical order, the probable meaning of this name is ``the one who is next to Caligastia''. On the etymology of the name `Caligastia' see the note on \bibref[4:3.3]{p004 3:3}.}, but they spurned his tender proffer. Caligastia, your apostate Planetary Prince, is still free on Urantia to prosecute his nefarious designs, but he has absolutely no power to enter the minds of men, neither can he draw near to their souls to tempt or corrupt them unless they really desire to be cursed with his wicked presence.
\vs p053 8:7 \pc Before the bestowal of Michael these rulers of darkness sought to maintain their authority on Urantia, and they persistently withstood the minor and subordinate celestial personalities. But since the day of Pentecost this traitorous Caligastia and his equally contemptible associate, Daligastia, are servile before the divine majesty of the Paradise Thought Adjusters and the protective Spirit of Truth, the spirit of Michael, which has been poured out upon all flesh.
\vs p053 8:8 But even so, no fallen spirit ever did have the power to invade the minds or to harass the souls of the children of God. Neither Satan nor Caligastia could ever touch or approach the faith sons of God; faith is an effective armour against sin and iniquity. It is true: “He who is born of God keeps himself, and the wicked one touches him not.”
\vs p053 8:9 In general, when weak and dissolute mortals are supposed to be under the influence of devils and demons, they are merely being dominated by their own inherent and debased tendencies, being led away by their own natural propensities. The devil has been given a great deal of credit for evil which does not belong to him. Caligastia has been comparatively impotent since the cross of Christ.
\usection{Present Status of the Rebellion}
\vs p053 9:1 Early in the days of the Lucifer rebellion, salvation was offered all rebels by Michael. To all who would show proof of sincere repentance, he offered, upon his attainment of complete universe sovereignty, forgiveness and reinstatement in some form of universe service. None of the leaders accepted this merciful proffer. But thousands of the angels and the lower orders of celestial beings, including hundreds of the Material Sons and Daughters, accepted the mercy proclaimed by the Panoptians and were given rehabilitation at the time of Jesus’ resurrection 1900 years ago. These beings have since been transferred to the Father’s world of Jerusem, where they must be held, technically, until the Uversa courts hand down a decision in the matter of Gabriel \bibemph{vs.} Lucifer. But no one doubts that, when the annihilation verdict is issued, these repentant and salvaged personalities will be exempted from the decree of extinction. These probationary souls now labour with the Panoptians in the work of caring for the Father’s world.
\vs p053 9:2 \pc The archdeceiver has never been on Urantia since the days when he sought to turn back Michael from the purpose to complete the bestowal and to establish himself finally and securely as the unqualified ruler of Nebadon. Upon Michael’s becoming the settled head of the universe of Nebadon, Lucifer was taken into custody by the agents of the Uversa Ancients of Days and has since been a prisoner on satellite number one of the Father’s group of the transition spheres of Jerusem. And here the rulers of other worlds and systems behold the end of the unfaithful Sovereign of Satania. Paul knew of the status of these rebellious leaders following Michael’s bestowal, for he wrote of Caligastia’s chiefs as “spiritual hosts of wickedness in the heavenly places.”
\vs p053 9:3 \pc Michael, upon assuming the supreme sovereignty of Nebadon, petitioned the Ancients of Days for authority to intern all personalities concerned in the Lucifer rebellion pending the rulings of the superuniverse tribunals in the case of Gabriel \bibemph{vs.} Lucifer, placed on the records of the Uversa supreme court almost 200,000 years ago, as you reckon time. Concerning the system capital group, the Ancients of Days granted the Michael petition with but a single exception: Satan was allowed to make periodic visits to the apostate princes on the fallen worlds until another Son of God should be accepted by such apostate worlds, or until such time as the courts of Uversa should begin the adjudication of the case of Gabriel \bibemph{vs.} Lucifer.
\vs p053 9:4 Satan could come to Urantia because you had no Son of standing in residence --- neither Planetary Prince nor Material Son. Machiventa Melchizedek has since been proclaimed vicegerent Planetary Prince of Urantia, and the opening of the case of Gabriel \bibemph{vs.} Lucifer has signalized the inauguration of temporary planetary regimes on all the isolated worlds. It is true that Satan did periodically visit Caligastia and others of the fallen princes right up to the time of the presentation of these revelations, when there occurred the first hearing of Gabriel’s plea for the annihilation of the archrebels. Satan is now unqualifiedly detained on the Jerusem prison worlds.
\vs p053 9:5 \pc Since Michael’s final bestowal no one in all Satania has desired to go to the prison worlds to minister to the interned rebels. And no more beings have been won to the deceiver’s cause. For 1900 years the status has been unchanged.
\vs p053 9:6 We do not look for a removal of the present Satania restrictions until the Ancients of Days make final disposition of the archrebels. The system circuits will not be reinstated so long as Lucifer lives. Meantime, he is wholly inactive.
\vs p053 9:7 The rebellion has ended on Jerusem. It ends on the fallen worlds as fast as divine Sons arrive. We believe that all rebels who will ever accept mercy have done so. We await the flashing broadcast that will deprive these traitors of personality existence. We anticipate the verdict of Uversa will be announced by the executionary broadcast which will effect the annihilation of these interned rebels. Then will you look for their places, but they shall not be found. “And they who know you among the worlds will be astonished at you; you have been a terror, but never shall you be any more.” And thus shall all of these unworthy traitors “become as though they had not been.” All await the Uversa decree.
\vs p053 9:8 But for ages the seven prison worlds of spiritual darkness in Satania have constituted a solemn warning to all Nebadon, eloquently and effectively proclaiming the great truth “that the way of the transgressor is hard”; “that within every sin is concealed the seed of its own destruction”; that “the wages of sin is death.”
\vsetoff
\vs p053 9:9 [Presented by Manovandet Melchizedek, onetime attached to the receivership of Urantia.]
\quizlink
