\upaper{56}{Universal Unity}
\uminitoc{Physical Co-ordination}
\uminitoc{Intellectual Unity}
\uminitoc{Spiritual Unification}
\uminitoc{Personality Unification}
\uminitoc{Deity Unity}
\uminitoc{Unification of Evolutionary Deity}
\uminitoc{Universal Evolutionary Repercussions}
\uminitoc{The Supreme Unifier}
\uminitoc{Universal Absolute Unity}
\uminitoc{Truth, Beauty, and Goodness}
\author{Mighty Messenger and Machiventa Melchizedek}
\vs p056 0:1 God is unity. Deity is universally co\hyp{}ordinated. The universe of universes is one vast integrated mechanism which is absolutely controlled by one infinite mind. The physical, intellectual, and spiritual domains of universal creation are divinely correlated. The perfect and imperfect are truly interrelated, and therefore may the finite evolutionary creature ascend to Paradise in obedience to the Universal Father’s mandate: “Be you perfect, even as I am perfect.”
\vs p056 0:2 The diverse levels of creation are all unified in the plans and administration of the Architects of the Master Universe. To the circumscribed minds of time\hyp{}space mortals the universe may present many problems and situations which apparently portray disharmony and indicate absence of effective co\hyp{}ordination; but those of us who are able to observe wider stretches of universal phenomena, and who are more experienced in this art of detecting the basic unity which underlies creative diversity and of discovering the divine oneness which overspreads all this functioning of plurality, better perceive the divine and single purpose exhibited in all these manifold manifestations of universal creative energy.
\usection{Physical Co\hyp{}ordination}
\vs p056 1:1 The physical or material creation is not infinite, but it is perfectly co\hyp{}ordinated. There are force, energy, and power, but they are all one in origin. The seven superuniverses are seemingly dual; the central universe, triune; but Paradise is of single constitution. And Paradise is the actual source of all material universes --- past, present, and future. But this cosmic derivation is an \bibemph{eternity} event; at no \bibemph{time ---} past, present, or future --- does either space or the material cosmos come forth from the nuclear Isle of Light. As the cosmic source, Paradise functions prior to space and before time; hence would its derivations seem to be orphaned in time and space did they not emerge through the Unqualified Absolute, their ultimate repository in space and their revealer and regulator in time.
\vs p056 1:2 \pc The Unqualified Absolute upholds the physical universe, while the Deity Absolute motivates the exquisite overcontrol of all material reality; and both Absolutes are functionally unified by the Universal Absolute. This cohesive correlation of the material universe is best understood by all personalities --- material, morontia, absonite, or spiritual --- by the observation of the gravity response of all bona fide material reality to the gravity centring on nether Paradise.
\vs p056 1:3 Gravity unification is universal and unvarying; pure\hyp{}energy response is likewise universal and inescapable. Pure energy (primordial force) and pure spirit are wholly preresponsive to gravity. These primal forces, inhering in the Absolutes, are personally controlled by the Universal Father; hence does all gravity centre in the personal presence of the Paradise Father of pure energy and pure spirit and in his supermaterial abode.
\vs p056 1:4 \pc Pure energy is the ancestor of all relative, nonspirit functional realities, while pure spirit is the potential of the divine and directive overcontrol of all basic energy systems. And these realities, so diverse as manifested throughout space and as observed in the motions of time, are both centred in the person of the Paradise Father. In him they are one --- must be unified --- because God is one. The Father’s personality is absolutely unified.
\vs p056 1:5 In the infinite nature of God the Father there could not possibly exist duality of reality, such as physical and spiritual; but the instant we look aside from the infinite levels and absolute reality of the personal values of the Paradise Father, we observe the existence of these two realities and recognize that they are fully responsive to his personal presence; in him all things consist.
\vs p056 1:6 The moment you depart from the unqualified concept of the infinite personality of the Paradise Father, you must postulate MIND as the inevitable technique of unifying the ever\hyp{}widening divergence of these dual universe manifestations of the original monothetic Creator personality, the First Source and Centre --- the I AM.
\usection{Intellectual Unity}
\vs p056 2:1 The Thought\hyp{}Father realizes spirit expression in the Word\hyp{}Son and attains reality expansion through Paradise in the far\hyp{}flung material universes. The spiritual expressions of the Eternal Son are correlated with the material levels of creation by the functions of the Infinite Spirit, by whose spirit\hyp{}responsive ministry of mind, and in whose physical\hyp{}directive acts of mind, the spiritual realities of Deity and the material repercussions of Deity are correlated the one with the other.
\vs p056 2:2 Mind is the functional endowment of the Infinite Spirit, therefore infinite in potential and universal in bestowal. The primal thought of the Universal Father eternalizes in dual expression: the Isle of Paradise and his Deity equal, the spiritual and Eternal Son. Such duality of eternal reality renders the mind God, the Infinite Spirit, inevitable. Mind is the indispensable channel of communication between spiritual and material realities. The material evolutionary creature can conceive and comprehend the indwelling spirit only by the ministry of mind.
\vs p056 2:3 This infinite and universal mind is ministered in the universes of time and space as the cosmic mind; and though extending from the primitive ministry of the adjutant spirits up to the magnificent mind of the chief executive of a universe, even this cosmic mind is adequately unified in the supervision of the Seven Master Spirits, who are in turn co\hyp{}ordinated with the Supreme Mind of time and space and perfectly correlated with the all\hyp{}embracing mind of the Infinite Spirit.
\usection{Spiritual Unification}
\vs p056 3:1 As the universal mind gravity is centred in the Paradise personal presence of the Infinite Spirit, so does the universal spirit gravity centre in the Paradise personal presence of the Eternal Son. The Universal Father is one, but to time\hyp{}space he is revealed in the dual phenomena of pure energy and pure spirit.
\vs p056 3:2 Paradise spirit realities are likewise one, but in all time\hyp{}space situations and relations this single spirit is revealed in the dual phenomena of the spirit personalities and emanations of the Eternal Son and the spirit personalities and influences of the Infinite Spirit and associated creations; and there is yet a third --- pure\hyp{}spirit fragmentations --- the Father’s bestowal of the Thought Adjusters and other spirit entities which are prepersonal.
\vs p056 3:3 \pc No matter on what level of universe activities you may encounter spiritual phenomena or contact with spirit beings, you may know that they are all derived from the God who is spirit by the ministry of the Spirit Son and the Infinite Mind Spirit. And this far\hyp{}flung spirit functions as a phenomenon on the evolutionary worlds of time as it is directed from the headquarters of the local universes. From these capitals of the Creator Sons come the Holy Spirit and the Spirit of Truth, together with the ministry of the adjutant mind\hyp{}spirits, to the lower and evolving levels of material minds.
\vs p056 3:4 While mind is more unified on the level of the Master Spirits in association with the Supreme Being and as the cosmic mind in subordination to the Absolute Mind, the spirit ministry to the evolving worlds is more directly unified in the personalities resident on the headquarters of the local universes and in the persons of the presiding Divine Ministers, who are in turn well\hyp{}nigh perfectly correlated with the Paradise gravity circuit of the Eternal Son, wherein occurs final unification of all time\hyp{}space spirit manifestations.
\vs p056 3:5 \pc Perfected creature existence can be attained, sustained, and eternalized by the fusion of self\hyp{}conscious mind with a fragment of the pre\hyp{}Trinity spirit endowment of some one of the persons of the Paradise Trinity. The mortal mind is the creation of the Sons and Daughters of the Eternal Son and the Infinite Spirit and, when fused with the Thought Adjuster from the Father, partakes of the threefold spirit endowment of the evolutionary realms. But these three spirit expressions become perfectly unified in the finaliters, even as they were in eternity so unified in the Universal I AM ere he ever became the Universal Father of the Eternal Son and the Infinite Spirit.
\vs p056 3:6 Spirit must always and ultimately become threefold in expression and Trinity\hyp{}unified in final realization. Spirit originates from one source through a threefold expression; and in finality it must and does attain its full realization in that divine unification which is experienced in finding God --- oneness with divinity --- in eternity, and by means of the ministry of the cosmic mind of the infinite expression of the eternal word of the Father’s universal thought.
\usection{Personality Unification}
\vs p056 4:1 The Universal Father is a divinely unified personality; hence will all his ascendant children who are carried to Paradise by the rebound momentum of the Thought Adjusters, who went forth from Paradise to indwell material mortals in obedience to the Father’s mandate, likewise be fully unified personalities ere they reach Havona.
\vs p056 4:2 Personality inherently reaches out to unify all constituent realities. The infinite personality of the First Source and Centre, the Universal Father, unifies all seven constituent Absolutes of Infinity; and the personality of mortal man, being an exclusive and direct bestowal of the Universal Father, likewise possesses the potential of unifying the constituent factors of the mortal creature. Such unifying creativity of all creature personality is a birthmark of its high and exclusive source and is further evidential of its unbroken contact with this same source through the personality circuit, by means of which the personality of the creature maintains direct and sustaining contact with the Father of all personality on Paradise.
\vs p056 4:3 Notwithstanding that God is manifest from the domains of the Sevenfold up through supremacy and ultimacy to God the Absolute, the personality circuit, centring on Paradise and in the person of God the Father, provides for the complete and perfect unification of all these diverse expressions of divine personality so far as concerns all creature personalities on all levels of intelligent existence and in all the realms of the perfect, perfected, and perfecting universes.
\vs p056 4:4 \pc While God is to and in the universes all that we have portrayed, nevertheless, to you and to all other God\hyp{}knowing creatures he is one, your Father and their Father. To personality God cannot be plural. God is Father to each of his creatures, and it is literally impossible for any child to have more than one father.
\vs p056 4:5 Philosophically, cosmically, and with reference to differential levels and locations of manifestation, you may and perforce must conceive of the functioning of plural Deities and postulate the existence of plural Trinities; but in the worshipful experience of the personal contact of every worshipping personality throughout the master universe, God is one; and that unified and personal Deity is our Paradise parent, God the Father, the bestower, conservator, and Father of all personalities from mortal man on the inhabited worlds to the Eternal Son on the central Isle of Light.
\usection{Deity Unity}
\vs p056 5:1 The oneness, the indivisibility, of Paradise Deity is existential and absolute. There are three eternal personalizations of Deity --- the Universal Father, the Eternal Son, and the Infinite Spirit --- but in the Paradise Trinity they are \bibemph{actually} one Deity, undivided and indivisible.
\vs p056 5:2 \pc From the original Paradise\hyp{}Havona level of existential reality, two subabsolute levels have differentiated, and thereon have the Father, Son, and Spirit engaged in the creation of numerous personal associates and subordinates. And while it is inappropriate in this connection to undertake the consideration of absonite deity unification on transcendental levels of ultimacy, it is feasible to look at some features of the unifying function of the various Deity personalizations in whom divinity is functionally manifest to the diverse sectors of creation and to the different orders of intelligent beings.
\vs p056 5:3 The present functioning of divinity in the superuniverses is actively manifest in the operations of the Supreme Creators --- the local universe Creator Sons and Spirits, the superuniverse Ancients of Days, and the Seven Master Spirits of Paradise. These beings constitute the first three levels of God the Sevenfold leading inward to the Universal Father, and this entire domain of God the Sevenfold is co\hyp{}ordinating on the first level of experiential deity in the evolving Supreme Being.
\vs p056 5:4 \pc On Paradise and in the central universe, Deity unity is a fact of existence. Throughout the evolving universes of time and space, Deity unity is an achievement.
\usection{Unification of Evolutionary Deity}
\vs p056 6:1 When the three eternal persons of Deity function as undivided Deity in the Paradise Trinity, they achieve perfect unity; likewise, when they create, either associatively or severally, their Paradise progeny exhibit the characteristic unity of divinity. And this divinity of purpose manifested by the Supreme Creators and Rulers of the time\hyp{}space domains eventuates in the unifying power potential of the sovereignty of experiential supremacy which, in the presence of the impersonal energy unity of the universe, constitutes a reality tension that can be resolved only through adequate unification with the experiential personality realities of experiential Deity.
\vs p056 6:2 The personality realities of the Supreme Being come forth from the Paradise Deities and on the pilot world of the outer Havona circuit unify with the power prerogatives of the Almighty Supreme coming up from the Creator divinities of the grand universe. God the Supreme as a person existed in Havona before the creation of the seven superuniverses, but he functioned only on spiritual levels. The evolution of the Almighty power of Supremacy by diverse divinity synthesis in the evolving universes eventuated in a new power presence of Deity which co\hyp{}ordinated with the spiritual person of the Supreme in Havona by means of the Supreme Mind, which concomitantly translated from the potential resident in the infinite mind of the Infinite Spirit to the active functional mind of the Supreme Being.
\vs p056 6:3 \pc The material\hyp{}minded creatures of the evolutionary worlds of the seven superuniverses can comprehend Deity unity only as it is evolving in this power\hyp{}personality synthesis of the Supreme Being. On any level of existence God cannot exceed the conceptual capacity of the beings who live on such a level. Mortal man must, through the recognition of truth, the appreciation of beauty, and the worship of goodness, evolve the recognition of a God of love and then progress through ascending deity levels to the comprehension of the Supreme. Deity, having been thus grasped as unified in power, can then be personalized in spirit to creature understanding and attainment.
\vs p056 6:4 While ascending mortals achieve power comprehension of the Almighty on the capitals of the superuniverses and personality comprehension of the Supreme on the outer circuits of Havona, they do not actually find the Supreme Being as they are destined to find the Paradise Deities. Even the finaliters, sixth\hyp{}stage spirits, have not found the Supreme Being, nor are they likely to until they have achieved seventh\hyp{}stage\hyp{}spirit status, and until the Supreme has become actually functional in the activities of the future outer universes.
\vs p056 6:5 But when ascenders find the Universal Father as the 7\ts{th} level of God the Sevenfold, they have attained the personality of the First Person of \bibemph{all} deity levels of personal relationships with universe creatures.
\usection{Universal Evolutionary Repercussions}
\vs p056 7:1 The steady progress of evolution in the time\hyp{}space universes is accompanied by ever\hyp{}enlarging revelations of Deity to all intelligent creatures. The attainment of the height of evolutionary progress on a world, in a system, constellation, universe, superuniverse, or in the grand universe signalizes corresponding enlargements of deity function to and in these progressive units of creation. And every such local enhancement of divinity realization is accompanied by certain well\hyp{}defined repercussions of enlarged deity manifestation to all other sectors of creation. Extending outward from Paradise, each new domain of realized and attained evolution constitutes a new and enlarged revelation of experiential Deity to the universe of universes.
\vs p056 7:2 As the components of a local universe are progressively settled in light and life, God the Sevenfold is increasingly made manifest. Time\hyp{}space evolution begins on a planet with the first expression of God the Sevenfold --- the Creator Son\hyp{}Creative Spirit association --- in control. With the settling of a system in light, this Son\hyp{}Spirit liaison attains the fullness of function; and when an entire constellation is thus settled, the second phase of God the Sevenfold becomes more active throughout such a realm. The completed administrative evolution of a local universe is attended by new and more direct ministrations of the superuniverse Master Spirits; and at this point there also begins that ever\hyp{}expanding revelation and realization of God the Supreme which culminates in the ascender’s comprehension of the Supreme Being while passing through the worlds of the sixth Havona circuit.
\vs p056 7:3 \pc The Universal Father, the Eternal Son, and the Infinite Spirit are existential deity manifestations to intelligent creatures and are not, therefore, similarly expanded in personality relations with the mind and spirit creatures of all creation.
\vs p056 7:4 \pc It should be noted that ascending mortals may experience the impersonal presence of successive levels of Deity long before they become sufficiently spiritual and adequately educated to attain experiential personal recognition of, and contact with, these Deities as personal beings.
\vs p056 7:5 Each new evolutionary attainment within a sector of creation, as well as every new invasion of space by divinity manifestations, is attended by simultaneous expansions of Deity functional\hyp{}revelation within the then existing and previously organized units of all creation. This new invasion of the administrative work of the universes and their component units may not always appear to be executed exactly in accordance with the technique herewith outlined because it is the practice to send forth advance groups of administrators to prepare the way for the subsequent and successive eras of new administrative overcontrol. Even God the Ultimate foreshadows his transcendental overcontrol of the universes during the later stages of a local universe settled in light and life.
\vs p056 7:6 It is a fact that, as the creations of time and space are progressively settled in evolutionary status, there is observed a new and fuller functioning of God the Supreme concomitant with a corresponding withdrawing of the first three manifestations of God the Sevenfold. If and when the grand universe becomes settled in light and life, what then will be the future function of the Creator\hyp{}Creative manifestations of God the Sevenfold if God the Supreme assumes direct control of these creations of time and space? Are these organizers and pioneers of the time\hyp{}space universes to be liberated for similar activities in outer space? We do not know, but we speculate much concerning these and related matters.
\vs p056 7:7 \pc As the frontiers of experiential Deity are extended out into the domains of the Unqualified Absolute, we envision the activity of God the Sevenfold during the earlier evolutionary epochs of these creations of the future. We are not all in agreement respecting the future status of the Ancients of Days and the superuniverse Master Spirits. Neither do we know whether or not the Supreme Being will therein function as in the seven superuniverses. But we all conjecture that the Michaels, the Creator Sons, are destined to function in these outer universes. Some hold that the future ages will witness some closer form of union between the associated Creator Sons and Divine Ministers; it is even possible that such a creator union might eventuate in some new expression of associate\hyp{}creator identity of an ultimate nature. But we really know nothing about these possibilities of the unrevealed future.
\vs p056 7:8 We do know, however, that in the universes of time and space, God the Sevenfold provides a progressive approach to the Universal Father, and that this evolutionary approach is experientially unified in God the Supreme. We might conjecture that such a plan must prevail in the outer universes; on the other hand,\fnst{In 1955 text no comma after ``hand''.} the new orders of beings that may sometime inhabit these universes may be able to approach Deity on ultimate levels and by absonite techniques. In short, we have not the slightest concept of what technique of deity approach may become operative in the future universes of outer space.
\vs p056 7:9 Nevertheless, we deem that the perfected superuniverses will in some way become a part of the Paradise\hyp{}ascension careers of those beings who may inhabit these outer creations. It is quite possible that in that future age we may witness outer\hyp{}spacers approaching Havona through the seven superuniverses, administered by God the Supreme with or without the collaboration of the Seven Master Spirits.
\usection{The Supreme Unifier}
\vs p056 8:1 The Supreme Being has a threefold function in the experience of mortal man: First, he is the unifier of time\hyp{}space divinity, God the Sevenfold; second, he is the maximum of Deity which finite creatures can actually comprehend; third, he is mortal man’s only avenue of approach to the transcendental experience of consorting with absonite mind, eternal spirit, and Paradise personality.
\vs p056 8:2 Ascendant finaliters, having been born in the local universes, nurtured in the superuniverses, and trained in the central universe, embrace in their personal experiences the full potential of the comprehension of the time\hyp{}space divinity of God the Sevenfold unifying in the Supreme. Finaliters serve successively in superuniverses other than those of nativity, thereby superimposing experience upon experience until the fullness of the sevenfold diversity of possible creature experience has been encompassed. Through the ministry of the indwelling Adjusters the finaliters are enabled to \bibemph{find} the Universal Father, but it is by these techniques of experience that such finaliters come really to \bibemph{know} the Supreme Being, and they are destined to the service and the \bibemph{revelation} of this Supreme Deity in and to the future universes of outer space.
\vs p056 8:3 Bear in mind, all that God the Father and his Paradise Sons do for us, we in turn and in spirit have the opportunity to do for and in the emerging Supreme Being. The experience of love, joy, and service in the universe is mutual. God the Father does not need that his sons should return to him all that he bestows upon them, but they do (or may) in turn bestow all of this upon their fellows and upon the evolving Supreme Being.
\vs p056 8:4 All creational phenomena are reflective of antecedent creator\hyp{}spirit activities. Said Jesus, and it is literally true, “The Son does only those things which he sees the Father do.” In time you mortals may begin the revelation of the Supreme to your fellows, and increasingly may you augment this revelation as you ascend Paradiseward. In eternity you may be permitted to make increasing revelations of this God of evolutionary creatures on supreme levels --- even ultimate --- as seventh\hyp{}stage finaliters.
\usection{Universal Absolute Unity}
\vs p056 9:1 The Unqualified Absolute and the Deity Absolute are unified in the Universal Absolute. The Absolutes are co\hyp{}ordinated in the Ultimate, conditioned in the Supreme, and time\hyp{}space modified in God the Sevenfold. On subinfinite levels there are \bibemph{three} Absolutes, but in infinity they appear to be \bibemph{one.} On Paradise there are three personalizations of Deity, but in the Trinity they \bibemph{are} one.
\vs p056 9:2 \pc The major philosophic proposition of the master universe is this: Did the Absolute (the three Absolutes as one in infinity) exist before the Trinity? and is the Absolute ancestral to the Trinity? or is the Trinity antecedent to the Absolute?
\vs p056 9:3 Is the Unqualified Absolute a force presence independent of the Trinity? Does the presence of the Deity Absolute connote the unlimited function of the Trinity? and is the Universal Absolute the final function of the Trinity, even a Trinity of Trinities?
\vs p056 9:4 On first thought, a concept of the Absolute as ancestor to all things --- even the Trinity --- seems to afford transitory satisfaction of consistency gratification and philosophic unification, but any such conclusion is invalidated by the actuality of the eternity of the Paradise Trinity. We are taught, and we believe, that the Universal Father and his Trinity associates are eternal in nature and existence. There is, then, but one consistent philosophic conclusion, and that is: The Absolute is, to all universe intelligences, the impersonal and co\hyp{}ordinate reaction of the Trinity (of Trinities) to all basic and primary space situations, intrauniversal and extrauniversal. To all personality intelligences of the grand universe the Paradise Trinity forever stands in finality, eternity, supremacy, and ultimacy and, for all practical purposes of personal comprehension and creature realization, as absolute.
\vs p056 9:5 As creature minds may view this problem, they are led to the final postulate of the Universal I AM as the primal cause and the unqualified source of both the Trinity and the Absolute. When, therefore, we crave to entertain a personal concept of the Absolute, we revert to our ideas and ideals of the Paradise Father. When we desire to facilitate comprehension or to augment consciousness of this otherwise impersonal Absolute, we revert to the fact that the Universal Father is the existential Father of absolute personality; the Eternal Son is the Absolute Person, though not, in the experiential sense, the personalization of the Absolute. And then we go on to envisage the experiential Trinities as culminating in the experiential personalization of the Deity Absolute, while conceiving the Universal Absolute as constituting the universe and the extrauniverse phenomena of the manifest presence of the impersonal activities of the unified and co\hyp{}ordinated Deity associations of supremacy, ultimacy, and infinity --- the Trinity of Trinities.
\vs p056 9:6 \pc God the Father is discernible on all levels from the finite to the infinite, and though his creatures from Paradise to the evolutionary worlds have variously perceived him, only the Eternal Son and the Infinite Spirit know him as an infinity.
\vs p056 9:7 Spiritual personality is absolute only on Paradise, and the concept of the Absolute is unqualified only in infinity. Deity presence is absolute only on Paradise, and the revelation of God must always be partial, relative, and progressive until his power becomes experientially infinite in the space potency of the Unqualified Absolute, while his personality manifestation becomes experientially infinite in the manifest presence of the Deity Absolute, and while these two potentials of infinity become reality\hyp{}unified in the Universal Absolute.
\vs p056 9:8 But beyond subinfinite levels the three Absolutes \bibemph{are} one, and thereby is infinity Deity\hyp{}realized regardless of whether any other order of existence ever self\hyp{}realizes consciousness of infinity.
\vs p056 9:9 Existential status in eternity implies existential self\hyp{}consciousness of infinity, even though another eternity may be required to experience self\hyp{}realization of the experiential potentialities inherent in an infinity eternity --- an eternal infinity.
\vs p056 9:10 \pc And God the Father is the personal source of all manifestations of Deity and reality to all intelligent creatures and spirit beings throughout all the universe of universes. As personalities, now or in the successive universe experiences of the eternal future, no matter if you achieve the attainment of God the Sevenfold, comprehend God the Supreme, find God the Ultimate, or attempt to grasp the concept of God the Absolute, you will discover to your eternal satisfaction that in the consummation of each adventure you have, on new experiential levels, rediscovered the eternal God --- the Paradise Father of all universe personalities.
\vs p056 9:11 The Universal Father is the explanation of universal unity as it must be supremely, even ultimately, realized in the postultimate unity of absolute values and meanings --- unqualified Reality.
\vs p056 9:12 The Master Force Organizers go out into space and mobilize its energies to become gravity responsive to the Paradise pull of the Universal Father; and subsequently there come the Creator Sons, who organize these gravity\hyp{}responding forces into inhabited universes and therein evolve intelligent creatures who receive unto themselves the spirit of the Paradise Father and subsequently ascend to the Father to become like him in all possible divinity attributes.
\vs p056 9:13 The ceaseless and expanding march of the Paradise creative forces through space seems to presage the ever\hyp{}extending domain of the gravity grasp of the Universal Father and the never\hyp{}ending multiplication of varied types of intelligent creatures who are able to love God and be loved by him, and who, by thus becoming God\hyp{}knowing, may choose to be like him, may elect to attain Paradise and find God.
\vs p056 9:14 The universe of universes is altogether unified. God is one in power and personality. There is co\hyp{}ordination of all levels of energy and all phases of personality. Philosophically and experientially, in concept and in reality, all things and beings centre in the Paradise Father. God is all and in all, and no things or beings exist without him.
\usection{Truth, Beauty, and Goodness}
\vs p056 10:1 As the worlds settled in life and light progress from the initial stage to the 7\ts{th} epoch, they successively grasp for the realization of the reality of God the Sevenfold, ranging from the adoration of the Creator Son to the worship of his Paradise Father. Throughout the continuing 7\ts{th} stage of such a world’s history the ever\hyp{}progressing mortals grow in the knowledge of God the Supreme, while they vaguely discern the reality of the overshadowing ministry of God the Ultimate.
\vs p056 10:2 Throughout this glorious age the chief pursuit of the ever\hyp{}advancing mortals is the quest for a better understanding and a fuller realization of the comprehensible elements of Deity --- truth, beauty, and goodness. This represents man’s effort to discern God in mind, matter, and spirit. And as the mortal pursues this quest, he finds himself increasingly absorbed in the experiential study of philosophy, cosmology, and divinity.
\vs p056 10:3 \pc Philosophy you somewhat grasp, and divinity you comprehend in worship, social service, and personal spiritual experience, but the pursuit of beauty --- cosmology --- you all too often limit to the study of man’s crude artistic endeavours. Beauty, art, is largely a matter of the unification of contrasts. Variety is essential to the concept of beauty. The supreme beauty, the height of finite art, is the drama of the unification of the vastness of the cosmic extremes of Creator and creature. Man finding God and God finding man --- the creature becoming perfect as is the Creator --- that is the supernal achievement of the supremely beautiful, the attainment of the apex of cosmic art.
\vs p056 10:4 Hence materialism, atheism, is the maximation of ugliness, the climax of the finite antithesis of the beautiful. Highest beauty consists in the panorama of the unification of the variations which have been born of pre\hyp{}existent harmonious reality.
\vs p056 10:5 The attainment of cosmologic levels of thought includes:
\vs p056 10:6 \ublistelem{1.}\bibnobreakspace \bibemph{Curiosity.} Hunger for harmony and thirst for beauty. Persistent attempts to discover new levels of harmonious cosmic relationships.
\vs p056 10:7 \ublistelem{2.}\bibnobreakspace \bibemph{Aesthetic appreciation.} Love of the beautiful and ever\hyp{}advancing appreciation of the artistic touch of all creative manifestations on all levels of reality.
\vs p056 10:8 \ublistelem{3.}\bibnobreakspace \bibemph{Ethic sensitivity.} Through the realization of truth the appreciation of beauty leads to the sense of the eternal fitness of those things which impinge upon the recognition of divine goodness in Deity relations with all beings; and thus even cosmology leads to the pursuit of divine reality values --- to God\hyp{}consciousness.
\vs p056 10:9 \pc The worlds settled in light and life are so fully concerned with the comprehension of truth, beauty, and goodness because these quality values embrace the revelation of Deity to the realms of time and space. The meanings of eternal truth make a combined appeal to the intellectual and spiritual natures of mortal man. Universal beauty embraces the harmonious relations and rhythms of the cosmic creation; this is more distinctly the intellectual appeal and leads towards unified and synchronous comprehension of the material universe. Divine goodness represents the revelation of infinite values to the finite mind, therein to be perceived and elevated to the very threshold of the spiritual level of human comprehension.
\vs p056 10:10 Truth is the basis of science and philosophy, presenting the intellectual foundation of religion. Beauty sponsors art, music, and the meaningful rhythms of all human experience. Goodness embraces the sense of ethics, morality, and religion --- experiential perfection\hyp{}hunger.
\vs p056 10:11 The existence of beauty implies the presence of appreciative creature mind just as certainly as the fact of progressive evolution indicates the dominance of the Supreme Mind. Beauty is the intellectual recognition of the harmonious time\hyp{}space synthesis of the far\hyp{}flung diversification of phenomenal reality, all of which stems from pre\hyp{}existent and eternal oneness.
\vs p056 10:12 Goodness is the mental recognition of the relative values of the diverse levels of divine perfection. The recognition of goodness implies a mind of moral status, a personal mind with ability to discriminate between good and evil. But the possession of goodness, greatness, is the measure of real divinity attainment.
\vs p056 10:13 \pc The recognition of \bibemph{true relations} implies a mind competent to discriminate between truth and error. The bestowal Spirit of Truth which invests the human minds of Urantia is unerringly responsive to truth --- the living spirit relationship of all things and all beings as they are co\hyp{}ordinated in the eternal ascent Godward.
\vs p056 10:14 Every impulse of every electron, thought, or spirit is an acting unit in the whole universe. Only sin is isolated and evil gravity resisting on the mental and spiritual levels. The universe is a whole; no thing or being exists or lives in isolation. Self\hyp{}realization is potentially evil if it is antisocial. It is literally true: “No man lives by himself.” Cosmic socialization constitutes the highest form of personality unification. Said Jesus: \textcolour{ubdarkred}{“He who would be greatest among you, let him become server of all.”}
\vs p056 10:15 Even truth, beauty, and goodness --- man’s intellectual approach to the universe of mind, matter, and spirit --- must be combined into one unified concept of a divine and supreme \bibemph{ideal.} As mortal personality unifies the human experience with matter, mind, and spirit, so does this divine and supreme ideal become power\hyp{}unified in Supremacy and then personalized as a God of fatherly love.
\vs p056 10:16 All insight into the relations of the parts to any given whole requires an understanding grasp of the relation of all parts to that whole; and in the universe this means the relation of created parts to the Creative Whole. Deity thus becomes the transcendental, even the infinite, goal of universal and eternal attainment.
\vs p056 10:17 \pc Universal beauty is the recognition of the reflection of the Isle of Paradise in the material creation, while eternal truth is the special ministry of the Paradise Sons who not only bestow themselves upon the mortal races but even pour out their Spirit of Truth upon all peoples. Divine goodness is more fully shown forth in the loving ministry of the manifold personalities of the Infinite Spirit. But love, the sum total of these three qualities, is man’s perception of God as his spirit Father.
\vs p056 10:18 Physical matter is the time\hyp{}space shadow of the Paradise energy\hyp{}shining of the absolute Deities. Truth meanings are the mortal\hyp{}intellect repercussions of the eternal word of Deity --- the time\hyp{}space comprehension of supreme concepts. The goodness values of divinity are the merciful ministries of the spirit personalities of the Universal, the Eternal, and the Infinite to the time\hyp{}space finite creatures of the evolutionary spheres.
\vs p056 10:19 These meaningful reality values of divinity are blended in the Father’s relation with each personal creature as divine love. They are co\hyp{}ordinated in the Son and his Sons as divine mercy. They manifest their qualities through the Spirit and his spirit children as divine ministry, the portrayal of loving mercy to the children of time. These three divinities are primarily manifested by the Supreme Being as power\hyp{}personality synthesis. They are variously shown forth by God the Sevenfold in seven differing associations of divine meanings and values on seven ascending levels.
\vs p056 10:20 \pc To finite man truth, beauty, and goodness embrace the full revelation of divinity reality. As this love\hyp{}comprehension of Deity finds spiritual expression in the lives of God\hyp{}knowing mortals, there are yielded the fruits of divinity: intellectual peace, social progress, moral satisfaction, spiritual joy, and cosmic wisdom. The advanced mortals on a world in the 7\ts{th} stage of light and life have learned that love is the greatest thing in the universe --- and they know that God is love.
\vs p056 10:21 \pc Love is the desire to do good to others.
\vsetoff
\vs p056 10:22 [Presented by a Mighty Messenger visiting on Urantia, by request of the Nebadon Revelatory Corps and in collaboration with a certain Melchizedek, the vicegerent Planetary Prince of Urantia.]
\vs p056 10:23 \separatorline
\vs p056 10:24 This paper on Universal Unity is the 25\ts{th} of a series of presentations by various authors, having been sponsored as a group by a commission of Nebadon personalities numbering 12 and acting under the direction of Mantutia Melchizedek. We indited these narratives and put them in the English language, by a technique authorized by our superiors, in the year 1934\fnst{Many of the human sources used in these papers were written and published in 1940s, so the date is spurious and was probably added later.} of Urantia time.
\quizlink
