\upaper{65}{The Overcontrol of Evolution}
\uminitoc{Life Carrier Functions}
\uminitoc{The Evolutionary Panorama}
\uminitoc{The Fostering of Evolution}
\uminitoc{The Urantia Adventure}
\uminitoc{Life-Evolution Vicissitudes}
\uminitoc{Evolutionary Techniques of Life}
\uminitoc{Evolutionary Mind Levels}
\uminitoc{Evolution in Time and Space}
\author{Life Carrier}
\vs p065 0:1 Basic evolutionary material life --- premind life --- is the formulation of the Master Physical Controllers and the life\hyp{}impartation ministry of the Seven Master Spirits in conjunction with the active ministration of the ordained Life Carriers. As a result of the co\hyp{}ordinate function of this threefold creativity there develops organismal physical capacity for mind --- material mechanisms for intelligent reaction to external environmental stimuli and, later on, to internal stimuli, influences taking origin in the organismal mind itself.
\vs p065 0:2 \pc There are, then, three distinct levels of life production and evolution:
\vs p065 0:3 \ublistelem{1.}\bibnobreakspace The physical\hyp{}energy domain --- mind\hyp{}capacity production.
\vs p065 0:4 \ublistelem{2.}\bibnobreakspace The mind ministry of the adjutant spirits --- impinging upon spirit capacity.
\vs p065 0:5 \ublistelem{3.}\bibnobreakspace The spirit endowment of mortal mind --- culminating in Thought Adjuster bestowal.
\vs p065 0:6 \pc The mechanical\hyp{}nonteachable levels of organismal environmental response are the domains of the physical controllers. The adjutant mind\hyp{}spirits activate and regulate the adaptative or nonmechanical\hyp{}teachable types of mind --- those response mechanisms of organisms capable of learning from experience. And as the spirit adjutants thus manipulate mind potentials, so do the Life Carriers exercise considerable discretionary control over the environmental aspects of evolutionary processes right up to the time of the appearance of human will --- the ability to know God and the power of choosing to worship him.
\vs p065 0:7 It is the integrated functioning of the Life Carriers, the physical controllers, and the spirit adjutants that conditions the course of organic evolution on the inhabited worlds. And this is why evolution --- on Urantia or elsewhere --- is always purposeful and never accidental.
\usection{Life Carrier Functions}
\vs p065 1:1 The Life Carriers are endowed with potentials of personality metamorphosis which but few orders of creatures possess. These Sons of the local universe are capable of functioning in three diverse phases of being. They ordinarily perform their duties as mid\hyp{}phase Sons, that being the state of their origin. But a Life Carrier in such a stage of existence could not possibly function in the electrochemical domains as a fabricator of physical energies and material particles into units of living existence.
\vs p065 1:2 Life Carriers are able to function and do function on the following three levels:
\vs p065 1:3 \ublistelem{1.}\bibnobreakspace The physical level of electrochemistry.
\vs p065 1:4 \ublistelem{2.}\bibnobreakspace The usual mid\hyp{}phase of quasi\hyp{}morontial existence.
\vs p065 1:5 \ublistelem{3.}\bibnobreakspace The advanced semispiritual level.
\vs p065 1:6 \pc When the Life Carriers make ready to engage in life implantation, and after they have selected the sites for such an undertaking, they summon the archangel commission of Life Carrier transmutation. This group consists of ten orders of diverse personalities, including the physical controllers and their associates, and is presided over by the chief of archangels, who acts in this capacity by the mandate of Gabriel and with the permission of the Ancients of Days. When these beings are properly encircuited, they can effect such modifications in the Life Carriers as will enable them immediately to function on the physical levels of electrochemistry.
\vs p065 1:7 After the life patterns have been formulated and the material organizations have been duly completed, the supermaterial forces concerned in life propagation become forthwith active, and life is existent. Whereupon the Life Carriers are immediately returned to their normal mid\hyp{}phase of personality existence, in which estate they can manipulate the living units and manoeuvre the evolving organisms, even though they are shorn of all ability to organize --- create --- new patterns of living matter.
\vs p065 1:8 After organic evolution has run a certain course and free will of the human type has appeared in the highest evolving organisms, the Life Carriers must either leave the planet or take renunciation vows; that is, they must pledge themselves to refrain from all attempts further to influence the course of organic evolution. And when such vows are voluntarily taken by those Life Carriers who choose to remain on the planet as future advisers to those who shall be entrusted with the fostering of the newly evolved will creatures, there is summoned a commission of 12, presided over by the chief of the Evening Stars, acting by authority of the System Sovereign and with permission of Gabriel; and forthwith these Life Carriers are transmuted to the third phase of personality existence --- the semispiritual level of being. And I have functioned on Urantia in this third phase of existence ever since the times of Andon and Fonta.
\vs p065 1:9 We look forward to a time when the universe may be settled in light and life, to a possible fourth stage of being wherein we shall be wholly spiritual, but it has never been revealed to us by what technique we may attain this desirable and advanced estate.
\usection{The Evolutionary Panorama}
\vs p065 2:1 The story of man’s ascent from seaweed to the lordship of earthly creation is indeed a romance of biologic struggle and mind survival. Man’s primordial ancestors were literally the slime and ooze of the ocean bed in the sluggish and warm\hyp{}water bays and lagoons of the vast shore lines of the ancient inland seas, those very waters in which the Life Carriers established the three independent life implantations on Urantia.
\vs p065 2:2 Very few species of the early types of marine vegetation that participated in those epochal changes which resulted in the animallike borderland organisms are in existence today. The sponges are the survivors of one of these early midway types, those organisms through which the \bibemph{gradual} transition from the vegetable to the animal took place. These early transition forms, while not identical with modern sponges, were much like them; they were true borderline organisms --- neither vegetable nor animal --- but they eventually led to the development of the true animal forms of life.
\vs p065 2:3 The bacteria, simple vegetable organisms of a very primitive nature, are very little changed from the early dawn of life; they even exhibit a degree of retrogression in their parasitic behaviour. Many of the fungi also represent a retrograde movement in evolution, being plants which have lost their chlorophyll\hyp{}making ability and have become more or less parasitic. The majority of disease\hyp{}causing bacteria and their auxiliary virus bodies really belong to this group of renegade parasitic fungi. During the intervening ages all of the vast kingdom of plant life has evolved from ancestors from which the bacteria have also descended.
\vs p065 2:4 The higher protozoan type of animal life soon appeared, and appeared \bibemph{suddenly.} And from these far\hyp{}distant times the amoeba, the typical single\hyp{}celled animal organism, has come on down but little modified. He disports himself today much as he did when he was the last and greatest achievement in life evolution. This minute creature and his protozoan cousins are to the animal creation what bacteria are to the plant kingdom; they represent the survival of the first early evolutionary steps in life differentiation together with \bibemph{failure of subsequent development.}
\vs p065 2:5 Before long the early single\hyp{}celled animal types associated themselves in communities, first on the plan of the Volvox and presently along the lines of the Hydra and jellyfish. Still later there evolved the starfish, stone lilies, sea urchins, sea cucumbers, centipedes, insects, spiders, crustaceans, and the closely related groups of earthworms and leeches, soon followed by the molluscs --- the oyster, octopus, and snail. Hundreds upon hundreds of species intervened and perished; mention is made only of those which survived the long, long struggle. Such nonprogressive specimens, together with the later appearing fish family, today represent the stationary types of early and lower animals, branches of the tree of life which failed to progress.
\vs p065 2:6 \pc The stage was thus set for the appearance of the first backboned animals, the fishes. From this fish family there sprang two unique modifications, the frog and the salamander. And it was the frog which began that series of progressive differentiations in animal life that finally culminated in man himself.
\vs p065 2:7 The frog is one of the earliest of surviving human\hyp{}race ancestors, but it also failed to progress, persisting today much as in those remote times. The frog is the only species ancestor of the early dawn races now living on the face of the earth. The human race has no surviving ancestry between the frog and the Eskimo.
\vs p065 2:8 \pc The frogs gave rise to the Reptilia, a great animal family which is virtually extinct, but which, before passing out of existence, gave origin to the whole bird family and the numerous orders of mammals.
\vs p065 2:9 Probably the greatest single leap of all prehuman evolution was executed when the reptile became a bird. The bird types of today --- eagles, ducks, pigeons, and ostriches --- all descended from the enormous reptiles of long, long ago.
\vs p065 2:10 The kingdom of reptiles, descended from the frog family, is today represented by four surviving divisions: two nonprogressive, snakes and lizards, together with their cousins, alligators and turtles; one partially progressive, the bird family, and the fourth, the ancestors of mammals and the direct line of descent of the human species. But though long departed, the massiveness of the passing Reptilia found echo in the elephant and mastodon, while their peculiar forms were perpetuated in the leaping kangaroos.
\vs p065 2:11 \pc Only 14 phyla have appeared on Urantia, the fishes being the last, and no new classes have developed since birds and mammals.
\vs p065 2:12 \pc It was from an agile little reptilian dinosaur of carnivorous habits but having a comparatively large brain that the placental mammals \bibemph{suddenly} sprang. These mammals developed rapidly and in many different ways, not only giving rise to the common modern varieties but also evolving into marine types, such as whales and seals, and into air navigators like the bat family.
\vs p065 2:13 Man thus evolved from the higher mammals derived principally from the \bibemph{western implantation} of life in the ancient east\hyp{}west sheltered seas. The \bibemph{eastern} and \bibemph{central groups} of living organisms were early progressing favourably toward the attainment of prehuman levels of animal existence. But as the ages passed, the eastern focus of life emplacement failed to attain a satisfactory level of intelligent prehuman status, having suffered such repeated and irretrievable losses of its highest types of germ plasm that it was forever shorn of the power to rehabilitate human potentialities.
\vs p065 2:14 Since the quality of the mind capacity for development in this eastern group was so definitely inferior to that of the other two groups, the Life Carriers, with the consent of their superiors, so manipulated the environment as further to circumscribe these inferior prehuman strains of evolving life. To all outward appearances the elimination of these inferior groups of creatures was accidental, but in reality it was altogether purposeful.
\vs p065 2:15 Later in the evolutionary unfolding of intelligence, the lemur ancestors of the human species were far more advanced in North America than in other regions; and they were therefore led to migrate from the arena of western life implantation over the Bering land bridge and down the coast to south\hyp{}western Asia, where they continued to evolve and to benefit by the addition of certain strains of the central life group. Man thus evolved out of certain western and central life strains but in the central to near\hyp{}eastern regions.
\vs p065 2:16 In this way the life that was planted on Urantia evolved until the ice age, when man himself first appeared and began his eventful planetary career. And this appearance of primitive man on earth during the ice age was not just an accident; it was by design. The rigors and climatic severity of the glacial era were in every way adapted to the purpose of fostering the production of a hardy type of human being with tremendous survival endowment.
\usection{The Fostering of Evolution}
\vs p065 3:1 It will hardly be possible to explain to the present\hyp{}day human mind many of the queer and apparently grotesque occurrences of early evolutionary progress. A purposeful plan was functioning throughout all of these seemingly strange evolutions of living things, but we are not allowed arbitrarily to interfere with the development of the life patterns after they have once been set in operation.
\vs p065 3:2 \pc Life Carriers may employ every possible natural resource and may utilize any and all fortuitous circumstances which will enhance the developmental progress of the life experiment, but we are not permitted mechanically to intervene in, or arbitrarily to manipulate the conduct and course of, either plant or animal evolution.
\vs p065 3:3 You have been informed that Urantia mortals evolved by way of primitive frog development, and that this ascending strain, carried in potential in a single frog, narrowly escaped extinction on a certain occasion. But it should not be inferred that the evolution of mankind would have been terminated by an accident at this juncture. At that very moment we were observing and fostering no less than 1,000 different and remotely situated mutating strains of life which could have been directed into various different patterns of prehuman development. This particular ancestral frog represented our third selection, the two prior life strains having perished in spite of all our efforts toward their conservation.
\vs p065 3:4 \pc Even the loss of Andon and Fonta before they had offspring, though delaying human evolution, would not have prevented it. Subsequent to the appearance of Andon and Fonta and before the mutating human potentials of animal life were exhausted, there evolved no less than 7,000 favourable strains which could have achieved some sort of human type of development. And many of these better stocks were subsequently assimilated by the various branches of the expanding human species.
\vs p065 3:5 Long before the Material Son and Daughter, the biologic uplifters, arrive on a planet, the human potentials of the evolving animal species have been exhausted. This biologic status of animal life is disclosed to the Life Carriers by the phenomenon of the third phase of adjutant spirit mobilization, which automatically occurs concomitantly with the exhaustion of the capacity of all animal life to give origin to the mutant potentials of prehuman individuals.
\vs p065 3:6 \pc Mankind on Urantia must solve its problems of mortal development with the human stocks it has --- no more races will evolve from prehuman sources throughout all future time. But this fact does not preclude the possibility of the attainment of vastly higher levels of human development through the intelligent fostering of the evolutionary potentials still resident in the mortal races. That which we, the Life Carriers, do toward fostering and conserving the life strains before the appearance of human will, man must do for himself after such an event and subsequent to our retirement from active participation in evolution. In a general way, man’s evolutionary destiny is in his own hands, and scientific intelligence must sooner or later supersede the random functioning of uncontrolled natural selection and chance survival.
\vs p065 3:7 And in discussing the fostering of evolution, it would not be amiss to point out that, in the long future ahead, when you may sometime be attached to a corps of Life Carriers, you will have abundant and ample opportunity to offer suggestions and make any possible improvements in the plans and technique of life management and transplantation. Be patient! If you have good ideas, if your minds are fertile with better methods of administration for any part of the universal domains, you are certainly going to have an opportunity to present them to your associates and fellow administrators in the ages to come.
\usection{The Urantia Adventure}
\vs p065 4:1 Do not overlook the fact that Urantia was assigned to us as a life\hyp{}experiment world. On this planet we made our 60\ts{th} attempt to modify and, if possible, improve the Satania adaptation of the Nebadon life designs, and it is of record that we achieved numerous beneficial modifications of the standard life patterns. To be specific, on Urantia we worked out and have satisfactorily demonstrated not less than 28 features of life modification which will be of service to all Nebadon throughout all future time.
\vs p065 4:2 But the establishment of life on no world is ever experimental in the sense that something untried and unknown is attempted. The evolution of life is a technique ever progressive, differential, and variable, but never haphazard, uncontrolled, nor wholly experimental, in the accidental sense.
\vs p065 4:3 \pc Many features of human life afford abundant evidence that the phenomenon of mortal existence was intelligently planned, that organic evolution is not a mere cosmic accident. When a living cell is injured, it possesses the ability to elaborate certain chemical substances which are empowered so to stimulate and activate the neighbouring normal cells that they immediately begin the secretion of certain substances which facilitate healing processes in the wound; and at the same time these normal and uninjured cells begin to proliferate --- they actually start to work creating new cells to replace any fellow cells which may have been destroyed by the accident.
\vs p065 4:4 This chemical action and reaction concerned in wound healing and cell reproduction represents the choice of the Life Carriers of a formula embracing over 100,000 phases and features of possible chemical reactions and biologic repercussions. More than 500,000 specific experiments were made by the Life Carriers in their laboratories before they finally settled upon this formula for the Urantia life experiment.
\vs p065 4:5 When Urantia scientists know more of these healing chemicals, they will become more efficient in the treatment of injuries, and indirectly they will know more about controlling certain serious diseases.
\vs p065 4:6 Since life was established on Urantia, the Life Carriers have improved this healing technique as it has been introduced on another Satania world, in that it affords more pain relief and exercises better control over the proliferation capacity of the associated normal cells.
\vs p065 4:7 \pc There were many unique features of the Urantia life experiment, but the two outstanding episodes were the appearance of the Andonic race prior to the evolution of the 6 coloured peoples and the later simultaneous appearance of the Sangik mutants in a single family. Urantia is the first world in Satania where the 6 coloured races sprang from the same human family. They ordinarily arise in diversified strains from independent mutations within the prehuman animal stock and usually appear on earth one at a time and successively over long periods of time, beginning with the red man and passing on down through the colours to indigo.
\vs p065 4:8 Another outstanding variation of procedure was the late arrival of the Planetary Prince. As a rule, the prince appears on a planet about the time of will development; and if such a plan had been followed, Caligastia might have come to Urantia even during the lifetimes of Andon and Fonta instead of almost 500,000 years later, simultaneously with the appearance of the 6 Sangik races.
\vs p065 4:9 On an ordinary inhabited world a Planetary Prince would have been granted on the request of the Life Carriers at, or sometime after, the appearance of Andon and Fonta. But Urantia having been designated a life\hyp{}modification planet, it was by preagreement that the Melchizedek observers, 12 in number, were sent as advisers to the Life Carriers and as overseers of the planet until the subsequent arrival of the Planetary Prince. These Melchizedeks came at the time Andon and Fonta made the decisions which enabled Thought Adjusters to indwell their mortal minds.
\vs p065 4:10 \pc On Urantia the endeavours of the Life Carriers to improve the Satania life patterns necessarily resulted in the production of many apparently useless forms of transition life. But the gains already accrued are sufficient to justify the Urantia modifications of the standard life designs.
\vs p065 4:11 It was our intention to produce an early manifestation of will in the evolutionary life of Urantia, and we succeeded. Ordinarily, will does not emerge until the coloured races have long been in existence, usually first appearing among the superior types of the red man. Your world is the only planet in Satania where the human type of will has appeared in a precoloured race.
\vs p065 4:12 But in our effort to provide for that combination and association of inheritance factors which finally gave rise to the mammalian ancestors of the human race, we were confronted with the necessity of permitting hundreds and thousands of other and comparatively useless combinations and associations of inheritance factors to take place. Many of these seemingly strange by\hyp{}products of our efforts are certain to meet your gaze as you dig back into the planetary past, and I can well understand how puzzling some of these things must be to the limited human viewpoint.
\usection{Life\hyp{}Evolution Vicissitudes}
\vs p065 5:1 It was a source of regret to the Life Carriers that our special efforts to modify intelligent life on Urantia should have been so handicapped by tragic perversions beyond our control: the Caligastia betrayal and the Adamic default.
\vs p065 5:2 But throughout all of this biologic adventure our greatest disappointment grew out of the reversion of certain primitive plant life to the prechlorophyll levels of parasitic bacteria on such an extensive and unexpected scale. This eventuality in plant\hyp{}life evolution caused many distressful diseases in the higher mammals, particularly in the more vulnerable human species. When we were confronted with this perplexing situation, we somewhat discounted the difficulties involved because we knew that the subsequent admixture of the Adamic life plasm would so reinforce the resisting powers of the resulting blended race as to make it practically immune to all diseases produced by the vegetable type of organism. But our hopes were doomed to disappointment owing to the misfortune of the Adamic default.
\vs p065 5:3 The universe of universes, including this small world called Urantia, is not being managed merely to meet our approval nor just to suit our convenience, much less to gratify our whims and satisfy our curiosity. The wise and all\hyp{}powerful beings who are responsible for universe management undoubtedly know exactly what they are about; and so it becomes Life Carriers and behoves mortal minds to enlist in patient waiting and hearty co\hyp{}operation with the rule of wisdom, the reign of power, and the march of progress.
\vs p065 5:4 There are, of course, certain compensations for tribulation, such as Michael’s bestowal on Urantia. But irrespective of all such considerations, the later celestial supervisors of this planet express complete confidence in the ultimate evolutionary triumph of the human race and in the eventual vindication of our original plans and life patterns.
\usection{Evolutionary Techniques of Life}
\vs p065 6:1 It is impossible accurately to determine, simultaneously, the exact location and the velocity of a moving object\fnst{\textbf{the exact location and the velocity}, This is known in quantum mechanics as the \bibemph{Heisenberg's Uncertainty Principle}: $\Delta x\Delta p\geqslant\hbar/2$. It is interesting to notice that this uncertainty is related to the \bibemph{measurement} and is not a fundamental property of the so-called ``quantum systems''.}; any attempt at measurement of either inevitably involves change in the other. The same sort of a paradox confronts mortal man when he undertakes the chemical analysis of protoplasm. The chemist can elucidate the chemistry of \bibemph{dead} protoplasm, but he cannot discern either the physical organization or the dynamic performance of \bibemph{living} protoplasm. Ever will the scientist come nearer and nearer the secrets of life, but never will he find them and for no other reason than that he must kill protoplasm in order to analyse it\fnst{\textbf{kill protoplasm in order to analyse it}, The idea that our knowledge of the living cell is complementary to the complete knowledge of its molecular structure goes back to Niels Bohr. He suggested that the complete knowledge of this structure can be achieved only by operations that destroy the life of the cell and so it is logically possible that life precludes the complete determinations of its underlying physicochemical structure. See W.~Heisenberg, ``Physics and Philosophy'', 1958, p.\,104.}. Dead protoplasm weighs the same as living protoplasm, but it is not the same.
\vs p065 6:2 \pc There is original endowment of adaptation in living things and beings. In every \bibemph{living} plant or animal cell, in every \bibemph{living} organism --- material or spiritual --- there is an insatiable craving for the attainment of ever\hyp{}increasing perfection of environmental adjustment, organismal adaptation, and augmented life realization. These interminable efforts of all living things evidence the existence within them of an innate striving for perfection.
\vs p065 6:3 The most important step in plant evolution was the development of chlorophyll\hyp{}making ability, and the second greatest advance was the evolution of the spore into the complex seed. The spore is most efficient as a reproductive agent, but it lacks the potentials of variety and versatility inherent in the seed.
\vs p065 6:4 One of the most serviceable and complex episodes in the evolution of the higher types of animals consisted in the development of the ability of the iron in the circulating blood cells to perform in the double role of oxygen carrier and carbon dioxide remover. And this performance of the red blood cells illustrates how evolving organisms are able to adapt their functions to varying or changing environment. The higher animals, including man, oxygenate their tissues by the action of the iron of the red blood cells, which carries oxygen to the living cells and just as efficiently removes the carbon dioxide. But other metals can be made to serve the same purpose. The cuttlefish employs copper for this function, and the sea squirt utilizes vanadium.
\vs p065 6:5 The continuation of such biologic adjustments is illustrated by the evolution of teeth in the higher Urantia mammals; these attained to 36 in man’s remote ancestors, and then began an adaptative readjustment toward 32 in the dawn man and his near relatives. Now the human species is slowly gravitating toward 28. The process of evolution is still actively and adaptatively in progress on this planet.
\vs p065 6:6 But many seemingly mysterious adjustments of living organisms are purely chemical, wholly physical. At any moment of time, in the blood stream of any human being there exists the possibility of upward of 15,000,000 chemical reactions between the hormone output of a dozen ductless glands.
\vs p065 6:7 \pc The lower forms of plant life are wholly responsive to physical, chemical, and electrical environment. But as the scale of life ascends, one by one the mind ministries of the 7 adjutant spirits become operative, and the mind becomes increasingly adjustive, creative, co\hyp{}ordinative, and dominative. The ability of animals to adapt themselves to air, water, and land is not a supernatural endowment, but it is a superphysical adjustment.
\vs p065 6:8 Physics and chemistry alone cannot explain how a human being evolved out of the primeval protoplasm of the early seas. The ability to learn, memory and differential response to environment, is the endowment of mind. The laws of physics are not responsive to training; they are immutable and unchanging. The reactions of chemistry are not modified by education; they are uniform and dependable. Aside from the presence of the Unqualified Absolute, electrical and chemical reactions are predictable. But mind can profit from experience, can learn from reactive habits of behaviour in response to repetition of stimuli.
\vs p065 6:9 Preintelligent organisms react to environmental stimuli, but those organisms which are reactive to mind ministry can adjust and manipulate the environment itself.
\vs p065 6:10 The physical brain with its associated nervous system possesses innate capacity for response to mind ministry just as the developing mind of a personality possesses a certain innate capacity for spirit receptivity and therefore contains the potentials of spiritual progress and attainment. Intellectual, social, moral, and spiritual evolution are dependent on the mind ministry of the 7 adjutant spirits and their superphysical associates.
\usection{Evolutionary Mind Levels}
\vs p065 7:1 The 7 adjutant mind\hyp{}spirits are the versatile mind ministers to the lower intelligent existences of a local universe. This order of mind is ministered from the local universe headquarters or from some world connected therewith, but there is influential direction of lower\hyp{}mind function from the system capitals.
\vs p065 7:2 On an evolutionary world much, very much, depends on the work of these 7 adjutants. But they are mind ministers; they are not concerned in physical evolution, the domain of the Life Carriers. Nevertheless, the perfect integration of these spirit endowments with the ordained and natural procedure of the unfolding and inherent regime of the Life Carriers is responsible for the mortal inability to discern, in the phenomenon of mind, aught but the hand of nature and the outworking of natural processes, albeit you are occasionally somewhat perplexed in explaining all of everything connected with the natural reactions of mind as it is associated with matter. And if Urantia were operating more in accordance with the original plans, you would observe even less to arrest your attention in the phenomenon of mind.
\vs p065 7:3 The 7 adjutant spirits are more circuitlike than entitylike, and on ordinary worlds they are encircuited with other adjutant functionings throughout the local universe. On life\hyp{}experiment planets, however, they are relatively isolated. And on Urantia, owing to the unique nature of the life patterns, the lower adjutants experienced far more difficulty in contacting with the evolutionary organisms than would have been the case in a more standardized type of life endowment.
\vs p065 7:4 Again, on an average evolutionary world the 7 adjutant spirits are far better synchronized with the advancing stages of animal development than they were on Urantia. With but a single exception, the adjutants experienced the greatest difficulty in contacting with the evolving minds of Urantia organisms that they had ever had in all their functioning throughout the universe of Nebadon. On this world there developed many forms of border phenomena --- confusional combinations of the mechanical\hyp{}nonteachable and the nonmechanical\hyp{}teachable types of organismal response.
\vs p065 7:5 The 7 adjutant spirits do not make contact with the purely mechanical orders of organismal environmental response. Such preintelligent responses of living organisms pertain purely to the energy domains of the power centres, the physical controllers, and their associates.
\vs p065 7:6 The acquisition of the potential of the ability to \bibemph{learn} from experience marks the beginning of the functioning of the adjutant spirits, and they function from the lowliest minds of primitive and invisible existences up to the highest types in the evolutionary scale of human beings. They are the source and pattern for the otherwise more or less mysterious behaviour and incompletely understood quick reactions of mind to the material environment. Long must these faithful and always dependable influences carry forward their preliminary ministry before the animal mind attains the human levels of spirit receptivity.
\vs p065 7:7 The adjutants function exclusively in the evolution of experiencing mind up to the level of the 6\ts{th} phase, the spirit of worship. At this level there occurs that inevitable overlapping of ministry --- the phenomenon of the higher reaching down to co\hyp{}ordinate with the lower in anticipation of subsequent attainment of advanced levels of development. And still additional spirit ministry accompanies the action of the 7\ts{th} and last adjutant, the spirit of wisdom. Throughout the ministry of the spirit world the individual never experiences abrupt transitions of spirit co\hyp{}operation; always are these changes gradual and reciprocal.
\vs p065 7:8 Always should the domains of the physical (electrochemical) and the mental response to environmental stimuli be differentiated, and in turn must they all be recognized as phenomena apart from spiritual activities. The domains of physical, mental, and spiritual gravity are distinct realms of cosmic reality, notwithstanding their intimate interrelations.
\usection{Evolution in Time and Space}
\vs p065 8:1 Time and space are indissolubly linked; there is an innate association. The delays of time are inevitable in the presence of certain space conditions.
\vs p065 8:2 If spending so much time in effecting the evolutionary changes of life development occasions perplexity, I would say that we cannot time the life processes to unfold any faster than the physical metamorphoses of a planet will permit. We must wait upon the natural, physical development of a planet; we have absolutely no control over geologic evolution. If the physical conditions would allow, we could arrange for the completed evolution of life in considerably less than 1,000,000 years. But we are all under the jurisdiction of the Supreme Rulers of Paradise, and time is nonexistent on Paradise.
\vs p065 8:3 The individual’s yardstick for time measurement is the length of his life. All creatures are thus time conditioned, and therefore do they regard evolution as being a long\hyp{}drawn\hyp{}out process. To those of us whose life span is not limited by a temporal existence, evolution does not seem to be such a protracted transaction. On Paradise, where time is nonexistent, these things are all \bibemph{present} in the mind of Infinity and the acts of Eternity.
\vs p065 8:4 As mind evolution is dependent on, and delayed by, the slow development of physical conditions, so is spiritual progress dependent on mental expansion and unfailingly delayed by intellectual retardation. But this does not mean that spiritual evolution is dependent on education, culture, or wisdom. The soul may evolve regardless of mental culture but not in the absence of mental capacity and desire --- the choice of survival and the decision to achieve ever\hyp{}increasing perfection --- to do the will of the Father in heaven. Although survival may not depend on the possession of knowledge and wisdom, progression most certainly does.
\vs p065 8:5 \pc In the cosmic evolutionary laboratories mind is always dominant over matter, and spirit is ever correlated with mind. Failure of these diverse endowments to synchronize and co\hyp{}ordinate may cause time delays, but if the individual really knows God and desires to find him and become like him, then survival is assured regardless of the handicaps of time. Physical status may handicap mind, and mental perversity may delay spiritual attainment, but none of these obstacles can defeat the whole\hyp{}souled choice of will.
\vs p065 8:6 When physical conditions are ripe, \bibemph{sudden} mental evolutions may take place; when mind status is propitious, \bibemph{sudden} spiritual transformations may occur; when spiritual values receive proper recognition, then cosmic meanings become discernible, and increasingly the personality is released from the handicaps of time and delivered from the limitations of space.
\vsetoff
\vs p065 8:7 [Sponsored by a Life Carrier of Nebadon resident on Urantia.]
\quizlink
