\upaper{106}{Universe Levels of Reality}
\author{Melchizedek}
\vs p106 0:1 It is not enough that the ascending mortal should know something of the relations of Deity to the genesis and manifestations of cosmic reality; he should also comprehend something of the relationships existing between himself and the numerous levels of existential and experiential realities, of potential and actual realities. Man’s terrestrial orientation, his cosmic insight, and his spiritual directionization are all enhanced by a better comprehension of universe realities and their techniques of interassociation, integration, and unification.
\vs p106 0:2 The present grand universe and the emerging master universe are made up of many forms and phases of reality which, in turn, are existent on several levels of functional activity. These manifold existents and latents have been previously suggested in these papers, and they are now grouped for conceptual convenience in the following categories:
\vs p106 0:3 \ublistelem{1.}\bibnobreakspace \bibemph{Incomplete finites.} This is the present status of the ascending creatures of the grand universe, the present status of Urantia mortals. This level embraces creature existence from the planetary human up to, but not including, destiny attainers. It pertains to universes from early physical beginnings up to, but not including, settlement in light and life. This level constitutes the present periphery of creative activity in time and space. It appears to be moving outward from Paradise, for the closing of the present universe age, which will witness the grand universe attainment of light and life, will also and surely witness the appearance of some new order of developmental growth in the first outer space level.
\vs p106 0:4 \ublistelem{2.}\bibnobreakspace \bibemph{Maximum finites.} This is the present status of all experiential creatures who have attained destiny --- destiny as revealed within the scope of the present universe age. Even universes can attain to the maximum of status, both spiritually and physically. But the term “maximum” is itself a relative term --- maximum in relation to what? And that which is maximum, seemingly final, in the present universe age may be no more than a real beginning in terms of the ages to come. Some phases of Havona appear to be on the maximum order.
\vs p106 0:5 \ublistelem{3.}\bibnobreakspace \bibemph{Transcendentals.} This superfinite level (antecedently) follows finite progression. It implies the prefinite genesis of finite beginnings and the postfinite significance of all apparent finite endings or destinies. Much of Paradise\hyp{}Havona appears to be on the transcendental order.
\vs p106 0:6 \ublistelem{4.}\bibnobreakspace \bibemph{Ultimates.} This level encompasses that which is of master universe significance and impinges on the destiny level of the completed master universe. Paradise\hyp{}Havona (especially the circuit of the Father’s worlds) is in many respects of ultimate significance.
\vs p106 0:7 \ublistelem{5.}\bibnobreakspace \bibemph{Coabsolutes.} This level implies the projection of experientials upon a supermaster universe field of creative expression.
\vs p106 0:8 \ublistelem{6.}\bibnobreakspace \bibemph{Absolutes.} This level connotes the eternity presence of the seven existential Absolutes. It may also involve some degree of associative experiential attainment, but if so, we do not understand how, perhaps through the contact potential of personality.
\vs p106 0:9 \ublistelem{7.}\bibnobreakspace \bibemph{Infinity.} This level is pre\hyp{}existential and postexperiential. Unqualified unity of infinity is a hypothetical reality before all beginnings and after all destinies.
\vs p106 0:10 \pc These levels of reality are convenient compromise symbolizations of the present universe age and for the mortal perspective. There are a number of other ways of looking at reality from other\hyp{}than\hyp{}mortal perspective and from the standpoint of other universe ages. Thus it should be recognized that the concepts herewith presented are entirely relative, relative in the sense of being conditioned and limited by:
\vs p106 0:11 \ublistelem{1.}\bibnobreakspace The limitations of mortal language.
\vs p106 0:12 \ublistelem{2.}\bibnobreakspace The limitations of the mortal mind.
\vs p106 0:13 \ublistelem{3.}\bibnobreakspace The limited development of the seven superuniverses.
\vs p106 0:14 \ublistelem{4.}\bibnobreakspace Your ignorance of the six prime purposes of superuniverse development which do not pertain to the mortal ascent to Paradise.
\vs p106 0:15 \ublistelem{5.}\bibnobreakspace Your inability to grasp even a partial eternity viewpoint.
\vs p106 0:16 \ublistelem{6.}\bibnobreakspace The impossibility of depicting cosmic evolution and destiny in relation to all universe ages, not just in regard to the present age of the evolutionary unfolding of the seven superuniverses.
\vs p106 0:17 \ublistelem{7.}\bibnobreakspace The inability of any creature to grasp what is really meant by pre\hyp{}existentials or by postexperientials --- that which lies before beginnings and after destinies.
\vs p106 0:18 \pc Reality growth is conditioned by the circumstances of the successive universe ages. The central universe underwent no evolutionary change in the Havona age, but in the present epochs of the superuniverse age it is undergoing certain progressive changes induced by co\hyp{}ordination with the evolutionary superuniverses. The seven superuniverses, now evolving, will sometime attain the settled status of light and life, will attain the growth limit for the present universe age. But beyond doubt, the next age, the age of the first outer space level, will release the superuniverses from the destiny limitations of the present age. Repletion is continually being superimposed upon completion.
\vs p106 0:19 \pc These are some of the limitations which we encounter in attempting to present a unified concept of the cosmic growth of things, meanings, and values and of their synthesis on ever\hyp{}ascending levels of reality.
\usection{1.\bibnobreakspace Primary Association of Finite Functionals}
\vs p106 1:1 The primary or spirit\hyp{}origin phases of finite reality find immediate expression on creature levels as perfect personalities and on universe levels as the perfect Havona creation. Even experiential Deity is thus expressed in the spirit person of God the Supreme in Havona. But the secondary, evolutionary, time\hyp{}and\hyp{}matter\hyp{}conditioned phases of the finite become cosmically integrated only as a result of growth and attainment. Eventually all secondary or perfecting finites are to attain a level equal to that of primary perfection, but such destiny is subject to a time delay, a constitutive superuniverse qualification which is not genetically found in the central creation. (We know of the existence of tertiary finites, but the technique of their integration is as yet unrevealed.)
\vs p106 1:2 This superuniverse time lag, this obstacle to perfection attainment, provides for creature participation in evolutionary growth. It thus makes it possible for the creature to enter into partnership with the Creator in the evolution of that selfsame creature. And during these times of expanding growth the incomplete is correlated with the perfect through the ministry of God the Sevenfold.
\vs p106 1:3 God the Sevenfold signifies the recognition by Paradise Deity of the barriers of time in the evolutionary universes of space. No matter how remote from Paradise, how deep in space, a material survival personality may take origin, God the Sevenfold will be found there present and engaged in the loving and merciful ministry of truth, beauty, and goodness to such an incomplete, struggling, and evolutionary creature. The divinity ministry of the Sevenfold reaches inward through the Eternal Son to the Paradise Father and outward through the Ancients of Days to the universe Fathers --- the Creator Sons.
\vs p106 1:4 Man, being personal and ascending by spiritual progression, finds the personal and spiritual divinity of the Sevenfold Deity; but there are other phases of the Sevenfold which are not concerned with the progression of personality. The divinity aspects of this Deity grouping are at present integrated in the liaison between the Seven Master Spirits and the Conjoint Actor, but they are destined to be eternally unified in the emerging personality of the Supreme Being. The other phases of the Sevenfold Deity are variously integrated in the present universe age, but all are likewise destined to be unified in the Supreme. The Sevenfold, in all phases, is the source of the relative unity of the functional reality of the present grand universe.
\usection{2.\bibnobreakspace Secondary Supreme Finite Integration}
\vs p106 2:1 As God the Sevenfold functionally co\hyp{}ordinates finite evolution, so does the Supreme Being eventually synthesize destiny attainment. The Supreme Being is the deity culmination of grand universe evolution --- physical evolution around a spirit nucleus and eventual dominance of the spirit nucleus over the encircling and whirling domains of physical evolution. And all of this takes place in accordance with the mandates of personality: Paradise personality in the highest sense, Creator personality in the universe sense, mortal personality in the human sense, Supreme personality in the culminating or experiential totalling sense.
\vs p106 2:2 \pc The concept of the Supreme must provide for the differential recognition of spirit person, evolutionary power, and power\hyp{}personality synthesis --- the unification of evolutionary power with, and its dominance by, spirit personality.
\vs p106 2:3 Spirit, in the last analysis, comes from Paradise through Havona. Energy\hyp{}matter seemingly evolves in the depths of space and is organized as power by the children of the Infinite Spirit in conjunction with the Creator Sons of God. And all of this is experiential; it is a transaction in time and space involving a wide range of living beings including even Creator divinities and evolutionary creatures. The power mastery of the Creator divinities in the grand universe slowly expands to encompass the evolutionary settling and stabilizing of the time\hyp{}space creations, and this is the flowering of the experiential power of God the Sevenfold. It encompasses the whole gamut of divinity attainment in time and space from the Adjuster bestowals of the Universal Father to the life bestowals of the Paradise Sons. This is earned power, demonstrated power, experiential power; it stands in contrast to the eternity power, the unfathomable power, the existential power of the Paradise Deities.
\vs p106 2:4 This experiential power arising out of the divinity achievements of God the Sevenfold itself manifests the cohesive qualities of divinity by synthesizing --- totalizing --- as the almighty power of the attained experiential mastery of the evolving creations. And this almighty power in turn finds spirit\hyp{}personality cohesion on the pilot sphere of the outer belt of Havona worlds in union with the spirit personality of the Havona presence of God the Supreme. Thus does experiential Deity culminate the long evolutionary struggle by investing the power product of time and space with the spirit presence and divine personality resident in the central creation.
\vs p106 2:5 Thus does the Supreme Being eventually attain to the embrace of all of everything evolving in time and space while investing these qualities with spirit personality. Since creatures, even mortals, are personality participants in this majestic transaction, so do they certainly attain the capacity to know the Supreme and to perceive the Supreme as true children of such an evolutionary Deity.
\vs p106 2:6 \pc Michael of Nebadon is like the Paradise Father because he shares his Paradise perfection; so will evolutionary mortals sometime attain to kinship with the experiential Supreme, for they will truly share his evolutionary perfection.
\vs p106 2:7 \pc God the Supreme is experiential; therefore is he completely experiencible. The existential realities of the seven Absolutes are not perceivable by the technique of experience; only the \bibemph{personality realities} of the Father, Son, and Spirit can be grasped by the personality of the finite creature in the prayer\hyp{}worship attitude.
\vs p106 2:8 Within the completed power\hyp{}personality synthesis of the Supreme Being there will be associated all of the absoluteness of the several triodities which could be so associated, and this majestic personality of evolution will be experientially attainable and understandable by all finite personalities. When ascenders attain the postulated seventh stage of spirit existence, they will therein experience the realization of a new meaning\hyp{}value of the absoluteness and infinity of the triodities as such is revealed on subabsolute levels in the Supreme Being, who is experiencible. But the attainment of these stages of maximum development will probably await the co\hyp{}ordinate settling of the entire grand universe in light and life.
\usection{3.\bibnobreakspace Transcendental Tertiary Reality Association}
\vs p106 3:1 The absonite architects eventuate the plan; the Supreme Creators bring it into existence; the Supreme Being will consummate its fullness as it was time created by the Supreme Creators, and as it was space forecast by the Master Architects.
\vs p106 3:2 During the present universe age the administrative co\hyp{}ordination of the master universe is the function of the Architects of the Master Universe. But the appearance of the Almighty Supreme at the termination of the present universe age will signify that the evolutionary finite has attained the first stage of experiential destiny. This happening will certainly lead to the completed function of the first experiential Trinity --- the union of the Supreme Creators, the Supreme Being, and the Architects of the Master Universe. This Trinity is destined to effect the further evolutionary integration of the master creation.
\vs p106 3:3 The Paradise Trinity is truly one of infinity, and no Trinity can possibly be infinite that does not include this original Trinity. But the original Trinity is an eventuality of the exclusive association of absolute Deities; subabsolute beings had nothing to do with this primal association. The subsequently appearing and experiential Trinities embrace the contributions of even creature personalities. Certainly this is true of the Trinity Ultimate, wherein the very presence of the Master Creator Sons among the Supreme Creator members thereof betokens the concomitant presence of actual and bona fide creature experience \bibemph{within} this Trinity association.
\vs p106 3:4 The first experiential Trinity provides for group attainment of ultimate eventualities. Group associations are enabled to anticipate, even to transcend, individual capacities; and this is true even beyond the finite level. In the ages to come, after the seven superuniverses have been settled in light and life, the Corps of the Finality will doubtless be promulgating the purposes of the Paradise Deities as they are dictated by the Trinity Ultimate, and as they are power\hyp{}personality unified in the Supreme Being.
\vs p106 3:5 \pc Throughout all the gigantic universe developments of past and future eternity, we detect the expansion of the comprehensible elements of the Universal Father. As the I AM, we philosophically postulate his permeation of total infinity, but no creature is able experientially to encompass such a postulate. As the universes expand, and as gravity and love reach out into time\hyp{}organizing space, we are able to understand more and more of the First Source and Centre. We observe gravity action penetrating the space presence of the Unqualified Absolute, and we detect spirit creatures evolving and expanding within the divinity presence of the Deity Absolute while both cosmic and spirit evolution are by mind and experience unifying on finite deity levels as the Supreme Being and are co\hyp{}ordinating on transcendental levels as the Trinity Ultimate.
\usection{4.\bibnobreakspace Ultimate Quartan Integration}
\vs p106 4:1 The Paradise Trinity certainly co\hyp{}ordinates in the ultimate sense but functions in this respect as a self\hyp{}qualified absolute; the experiential Trinity Ultimate co\hyp{}ordinates the transcendental as a transcendental. In the eternal future this experiential Trinity will, through augmenting unity, further activate the eventuating presence of Ultimate Deity.
\vs p106 4:2 While the Trinity Ultimate is destined to co\hyp{}ordinate the master creation, God the Ultimate is the transcendental power\hyp{}personalization of the directionization of the entire master universe. The completed eventuation of the Ultimate implies the completion of the master creation and connotes the full emergence of this transcendental Deity.
\vs p106 4:3 What changes will be inaugurated by the full emergence of the Ultimate we do not know. But as the Supreme is now spiritually and personally present in Havona, so also is the Ultimate there present but in the absonite and superpersonal sense. And you have been informed of the existence of the Qualified Vicegerents of the Ultimate, though you have not been informed of their present whereabouts or function.
\vs p106 4:4 But irrespective of the administrative repercussions attendant upon the emergence of Ultimate Deity, the personal values of his transcendental divinity will be experiencible by all personalities who have been participants in the actualization of this Deity level. Transcendence of the finite can lead only to ultimate attainment. God the Ultimate exists in transcendence of time and space but is nonetheless subabsolute notwithstanding inherent capacity for functional association with absolutes.
\usection{5.\bibnobreakspace Coabsolute or Fifth\hyp{}Phase Association}
\vs p106 5:1 The Ultimate is the apex of transcendental reality even as the Supreme is the capstone of evolutionary\hyp{}experiential reality. And the actual emergence of these two experiential Deities lays the foundation for the second experiential Trinity. This is the Trinity Absolute, the union of God the Supreme, God the Ultimate, and the unrevealed Consummator of Universe Destiny. And this Trinity has theoretical capacity to activate the Absolutes of potentiality --- Deity, Universal, and Unqualified. But the completed formation of this Trinity Absolute could take place only after the completed evolution of the entire master universe, from Havona to the fourth and outermost space level.\fnc{\ldots{}the union of God the Supreme, God the Ultimate, and the \bibtextul{Unrevealed} Consummator of Universe Destiny. \bibexpl{The lowercase version appears to be correct because “unrevealed” does not seem to be part of the name but is solely descriptive (the title being found in several places without unrevealed preceding it). In the one other case in which “unrevealed” is found in conjunction with Consummator of Universe Destiny, it is not capitalized \bibref[0:12.7]{p00 12:7} in the text. [“Unrevealed” is found in one other location as a capitalized component of a title --- The “Unrevealed Creative Agencies of the Ancients of Days” \bibref[30:1.108]{p030 1:108} --- so such a format is possible.]}}
\vs p106 5:2 It should be made clear that these experiential Trinities are correlative, not only of the personality qualities of experiential Divinity, but also of all the other\hyp{}than\hyp{}personal qualities which characterize their attained Deity unity. While this presentation deals primarily with the personal phases of the unification of the cosmos, it is nonetheless true that the impersonal aspects of the universe of universes are likewise destined to undergo unification as is illustrated by the power\hyp{}personality synthesis now going on in connection with the evolution of the Supreme Being. The spirit\hyp{}personal qualities of the Supreme are inseparable from the power prerogatives of the Almighty, and both are complemented by the unknown potential of Supreme mind. Neither can God the Ultimate as a person be considered apart from the other\hyp{}than\hyp{}personal aspects of Ultimate Deity. And on the absolute level the Deity and the Unqualified Absolutes are inseparable and indistinguishable in the presence of the Universal Absolute.
\vs p106 5:3 Trinities are, in and of themselves, not personal, but neither do they contravene personality. Rather do they encompass it and correlate it, in a collective sense, with impersonal functions. Trinities are, then, always \bibemph{deity} reality but never \bibemph{personality} reality. The personality aspects of a trinity are inherent in its individual members, and as individual persons they are \bibemph{not} that trinity. Only as a collective are they trinity; that \bibemph{is} trinity. But always is trinity inclusive of all encompassed deity; trinity is deity unity.
\vs p106 5:4 The three Absolutes --- Deity, Universal, and Unqualified --- are not trinity, for all are not deity. Only the deified can become trinity; all other associations are triunities or triodities.
\usection{6.\bibnobreakspace Absolute or Sixth\hyp{}Phase Integration}
\vs p106 6:1 The present potential of the master universe is hardly absolute, though it may well be near\hyp{}ultimate, and we deem it impossible to achieve the full revelation of absolute meaning\hyp{}values within the scope of a subabsolute cosmos. We therefore encounter considerable difficulty in attempting to conceive of a total expression of the limitless possibilities of the three Absolutes or even in attempting to visualize the experiential personalization of God the Absolute on the now impersonal level of the Deity Absolute.
\vs p106 6:2 The space\hyp{}stage of the master universe seems to be adequate for the actualization of the Supreme Being, for the formation and full function of the Trinity Ultimate, for the eventuation of God the Ultimate, and even for the inception of the Trinity Absolute. But our concepts regarding the full function of this second experiential Trinity seem to imply something beyond even the wide\hyp{}spreading master universe.
\vs p106 6:3 If we assume a cosmos\hyp{}infinite --- some illimitable cosmos on beyond the master universe --- and if we conceive that the final developments of the Absolute Trinity will take place out on such a superultimate stage of action, then it becomes possible to conjecture that the completed function of the Trinity Absolute will achieve final expression in the creations of infinity and will consummate the absolute actualization of \bibemph{all} potentials. The integration and association of ever\hyp{}enlarging segments of reality will approach absoluteness of status proportional to the inclusion of all reality within the segments thus associated.
\vs p106 6:4 Stated otherwise: The Trinity Absolute, as its name implies, is really absolute in total function. We do not know how an absolute function can achieve total expression on a qualified, limited, or otherwise restricted basis. Hence we must assume that any such totality function will be unconditioned (in potential). And it would also appear that the unconditioned would also be unlimited, at least from a qualitative standpoint, though we are not so sure regarding quantitative relationships.
\vs p106 6:5 Of this, however, we are certain: While the existential Paradise Trinity is infinite, and while the experiential Trinity Ultimate is subinfinite, the Trinity Absolute is not so easy to classify. Though experiential in genesis and constitution, it definitely impinges upon the existential Absolutes of potentiality.
\vs p106 6:6 While it is hardly profitable for the human mind to seek to grasp such faraway and superhuman concepts, we would suggest that the eternity action of the Trinity Absolute may be thought of as culminating in some kind of experientialization of the Absolutes of potentiality. This would appear to be a reasonable conclusion with respect to the Universal Absolute, if not the Unqualified Absolute; at least we know that the Universal Absolute is not only static and potential but also associative in the total Deity sense of those words. But in regard to the conceivable values of divinity and personality, these conjectured happenings imply the personalization of the Deity Absolute and the appearance of those superpersonal values and those ultrapersonal meanings inherent in the personality completion of God the Absolute --- the third and last of the experiential Deities.
\usection{7.\bibnobreakspace Finality of Destiny}
\vs p106 7:1 Some of the difficulties in forming concepts of infinite reality integration are inherent in the fact that all such ideas embrace something of the finality of universal development, some kind of an experiential realization of all that could ever be. And it is inconceivable that quantitative infinity could ever be completely realized in finality. Always there must remain unexplored possibilities in the three potential Absolutes which no quantity of experiential development could ever exhaust. Eternity itself, though absolute, is not more than absolute.
\vs p106 7:2 Even a tentative concept of final integration is inseparable from the fruitions of unqualified eternity and is, therefore, practically nonrealizable at any conceivable future time.
\vs p106 7:3 \pc Destiny is established by the volitional act of the Deities who constitute the Paradise Trinity; destiny is established in the vastness of the three great potentials whose absoluteness encompasses the possibilities of all future development; destiny is probably consummated by the act of the Consummator of Universe Destiny, and this act is probably involved with the Supreme and the Ultimate in the Trinity Absolute. Any experiential destiny can be at least partially comprehended by experiencing creatures; but a destiny which impinges on infinite existentials is hardly comprehensible. Finality destiny is an existential\hyp{}experiential attainment which appears to involve the Deity Absolute. But the Deity Absolute stands in eternity relationship with the Unqualified Absolute by virtue of the Universal Absolute. And these three Absolutes, experiential in possibility, are actually existential and more, being limitless, timeless, spaceless, boundless, and measureless --- truly infinite.
\vs p106 7:4 The improbability of goal attainment does not, however, prevent philosophical theorizing about such hypothetical destinies. The actualization of the Deity Absolute as an attainable absolute God may be practically impossible of realization; nevertheless, such a finality fruition remains a theoretical possibility. The involvement of the Unqualified Absolute in some inconceivable cosmos\hyp{}infinite may be measurelessly remote in the futurity of endless eternity, but such a hypothesis is nonetheless valid. Mortals, morontians, spirits, finaliters, Transcendentalers, and others, together with the universes themselves and all other phases of reality, certainly do have a \bibemph{potentially final destiny that is absolute in value;} but we doubt that any being or universe will ever completely attain all of the aspects of such a destiny.
\vs p106 7:5 \pc No matter how much you may grow in Father comprehension, your mind will always be staggered by the unrevealed infinity of the Father\hyp{}I AM, the unexplored vastness of which will always remain unfathomable and incomprehensible throughout all the cycles of eternity. No matter how much of God you may attain, there will always remain much more of him, the existence of which you will not even suspect. And we believe that this is just as true on transcendental levels as it is in the domains of finite existence. The quest for God is endless!
\vs p106 7:6 Such inability to attain God in a final sense should in no manner discourage universe creatures; indeed, you can and do attain Deity levels of the Sevenfold, the Supreme, and the Ultimate, which mean to you what the infinite realization of God the Father means to the Eternal Son and to the Conjoint Actor in their absolute status of eternity existence. Far from harassing the creature, the infinity of God should be the supreme assurance that throughout all endless futurity an ascending personality will have before him the possibilities of personality development and Deity association which even eternity will neither exhaust nor terminate.
\vs p106 7:7 \pc To finite creatures of the grand universe the concept of the master universe seems to be well\hyp{}nigh infinite, but doubtless the absonite architects thereof perceive its relatedness to future and unimagined developments within the unending I AM. Even space itself is but an ultimate condition, a condition of qualification \bibemph{within} the relative absoluteness of the quiet zones of midspace.
\vs p106 7:8 At the inconceivably distant future eternity moment of the final completion of the entire master universe, no doubt we will all look back upon its entire history as only the beginning, simply the creation of certain finite and transcendental foundations for even greater and more enthralling metamorphoses in uncharted infinity. At such a future eternity moment the master universe will still seem youthful; indeed, it will be always young in the face of the limitless possibilities of never\hyp{}ending eternity.
\vs p106 7:9 \pc The improbability of infinite destiny attainment does not in the least prevent the entertainment of ideas about such destiny, and we do not hesitate to say that, if the three absolute potentials could ever become completely actualized, it would be possible to conceive of the final integration of total reality. This developmental realization is predicated on the completed actualization of the Unqualified, Universal, and Deity Absolutes, the three potentialities whose union constitutes the latency of the I AM, the suspended realities of eternity, the abeyant possibilities of all futurity, and more.
\vs p106 7:10 Such eventualities are rather remote to say the least; nevertheless, in the mechanisms, personalities, and associations of the three Trinities we believe we detect the theoretical possibility of the reuniting of the seven absolute phases of the Father\hyp{}I AM. And this brings us face to face with the concept of the threefold Trinity encompassing the Paradise Trinity of existential status and the two subsequently appearing Trinities of experiential nature and origin.
\usection{8.\bibnobreakspace The Trinity of Trinities}
\vs p106 8:1 The nature of the Trinity of Trinities is difficult to portray to the human mind; it is the actual summation of the entirety of experiential infinity as such is manifested in a theoretical infinity of eternity realization. In the Trinity of Trinities the experiential infinite attains to identity with the existential infinite, and both are as one in the pre\hyp{}experiential, pre\hyp{}existential I AM. The Trinity of Trinities is the final expression of all that is implied in the 15 triunities and associated triodities. Finalities are difficult for relative beings to comprehend, be they existential or experiential; therefore must they always be presented as relativities.
\vs p106 8:2 The Trinity of Trinities exists in several phases. It contains possibilities, probabilities, and inevitabilities that stagger the imaginations of beings far above the human level. It has implications that are probably unsuspected by the celestial philosophers, for its implications are in the triunities, and the triunities are, in the last analysis, unfathomable.
\vs p106 8:3 There are a number of ways in which the Trinity of Trinities can be portrayed. We elect to present the three\hyp{}level concept, which is as follows:
\vs p106 8:4 \ublistelem{1.}\bibnobreakspace The level of the three Trinities.
\vs p106 8:5 \ublistelem{2.}\bibnobreakspace The level of experiential Deity.
\vs p106 8:6 \ublistelem{3.}\bibnobreakspace The level of the I AM.
\vs p106 8:7 \pc These are levels of increasing unification. Actually the Trinity of Trinities is the first level, while the second and third levels are unification\hyp{}derivatives of the first.
\vs p106 8:8 \pc THE FIRST LEVEL: On this initial level of association it is believed that the three Trinities function as perfectly synchronized, though distinct, groupings of Deity personalities.
\vs p106 8:9 \ublistelem{1.}\bibnobreakspace \bibemph{The Paradise Trinity,} the association of the three Paradise Deities --- Father, Son, and Spirit. It should be remembered that the Paradise Trinity implies a threefold function --- an absolute function, a transcendental function (Trinity of Ultimacy), and a finite function (Trinity of Supremacy). The Paradise Trinity is any and all of these at any and all times.
\vs p106 8:10 \ublistelem{2.}\bibnobreakspace \bibemph{The Ultimate Trinity.} This is the deity association of the Supreme Creators, God the Supreme, and the Architects of the Master Universe. While this is an adequate presentation of the divinity aspects of this Trinity, it should be recorded that there are other phases of this Trinity, which, however, appear to be perfectly co\hyp{}ordinating with the divinity aspects.
\vs p106 8:11 \ublistelem{3.}\bibnobreakspace \bibemph{The Absolute Trinity.} This is the grouping of God the Supreme, God the Ultimate, and the Consummator of Universe Destiny in regard to all divinity values. Certain other phases of this triune grouping have to do with other\hyp{}than\hyp{}divinity values in the expanding cosmos. But these are unifying with the divinity phases just as the power and the personality aspects of the experiential Deities are now in process of experiential synthesis.
\vs p106 8:12 \pc The association of these three Trinities in the Trinity of Trinities provides for a possible unlimited integration of reality. This grouping contains causes, intermediates, and finals; inceptors, realizers, and consummators; beginnings, existences, and destinies. The Father\hyp{}Son partnership has become Son\hyp{}Spirit and then Spirit\hyp{}Supreme and on to Supreme\hyp{}Ultimate and Ultimate\hyp{}Absolute, even to Absolute and Father\hyp{}Infinite --- the completion of the cycle of reality. Likewise, in other phases not so immediately concerned with divinity and personality, does the First Great Source and Centre self\hyp{}realize the limitlessness of reality around the circle of eternity, from the absoluteness of self\hyp{}existence, through the endlessness of self\hyp{}revelation, to the finality of self\hyp{}realization --- from the absolute of existentials to the finality of experientials.
\vs p106 8:13 \pc THE SECOND LEVEL: The co\hyp{}ordination of the three Trinities inevitably involves the associative union of the experiential Deities, who are genetically associated with these Trinities. The nature of this second level has been sometimes presented as:
\vs p106 8:14 \ublistelem{1.}\bibnobreakspace \bibemph{The Supreme.} This is the deity consequence of the unity of the Paradise Trinity in experiential liaison with the Creator\hyp{}Creative children of the Paradise Deities. The Supreme is the deity embodiment of the completion of the first stage of finite evolution.
\vs p106 8:15 \ublistelem{2.}\bibnobreakspace \bibemph{The Ultimate.} This is the deity consequence of the eventuated unity of the second Trinity, the transcendental and absonite personification of divinity. The Ultimate consists in a variably regarded unity of many qualities, and the human conception thereof would do well to include at least those phases of ultimacy which are control directing, personally experiencible, and tensionally unifying, but there are many other unrevealed aspects of the eventuated Deity. While the Ultimate and the Supreme are comparable, they are not identical, neither is the Ultimate merely an amplification of the Supreme.
\vs p106 8:16 \ublistelem{3.}\bibnobreakspace \bibemph{The Absolute.} There are many theories held as to the character of the third member of the second level of the Trinity of Trinities. God the Absolute is undoubtedly involved in this association as the personality consequence of the final function of the Trinity Absolute, yet the Deity Absolute is an existential reality of eternity status.
\vs p106 8:17 The concept difficulty regarding this third member is inherent in the fact that the presupposition of such a membership really implies just one Absolute. Theoretically, if such an event could take place, we should witness the \bibemph{experiential} unification of the three Absolutes as one. And we are taught that, in infinity and \bibemph{existentially,} there is one Absolute. While it is least clear as to who this third member can be, it is often postulated that such may consist of the Deity, Universal, and Unqualified Absolutes in some form of unimagined liaison and cosmic manifestation. Certainly, the Trinity of Trinities could hardly attain to complete function short of the full unification of the three Absolutes, and the three Absolutes can hardly be unified short of the complete realization of all infinite potentials.
\vs p106 8:18 It will probably represent a minimum distortion of truth if the third member of the Trinity of Trinities is conceived as the Universal Absolute, provided this conception envisions the Universal not only as static and potential but also as associative. But we still do not perceive the relationship to the creative and evolutional aspects of the function of total Deity.
\vs p106 8:19 Though a completed concept of the Trinity of Trinities is difficult to form, a qualified concept is not so difficult. If the second level of the Trinity of Trinities is conceived as essentially personal, it becomes quite possible to postulate the union of God the Supreme, God the Ultimate, and God the Absolute as the personal repercussion of the union of the personal Trinities who are ancestral to these experiential Deities. We venture the opinion that these three experiential Deities will certainly unify on the second level as the direct consequence of the growing unity of their ancestral and causative Trinities who constitute the first level.
\vs p106 8:20 The first level consists of three Trinities; the second level exists as the personality association of experiential\hyp{}evolved, experiential\hyp{}eventuated, and experiential\hyp{}existential Deity personalities. And regardless of any conceptual difficulty in understanding the complete Trinity of Trinities, the personal association of these three Deities on the second level has become manifest to our own universe age in the phenomenon of the deitization of Majeston, who was actualized on this second level by the Deity Absolute, acting through the Ultimate and in response to the initial creative mandate of the Supreme Being.
\vs p106 8:21 THE THIRD LEVEL: In an unqualified hypothesis of the second level of the Trinity of Trinities, there is embraced the correlation of every phase of every kind of reality that is, or was, or could be in the entirety of infinity. The Supreme Being is not only spirit but also mind and power and experience. The Ultimate is all this and much more, while, in the conjoined concept of the oneness of the Deity, Universal, and Unqualified Absolutes, there is included the absolute finality of all reality realization.
\vs p106 8:22 In the union of the Supreme, Ultimate, and the complete Absolute, there could occur the functional reassembly of those aspects of infinity which were originally segmentalized by the I AM, and which resulted in the appearance of the Seven Absolutes of Infinity. Though the universe philosophers deem this to be a most remote probability, still, we often ask this question: If the second level of the Trinity of Trinities could ever achieve trinity unity, what then would transpire as a consequence of such deity unity? We do not know, but we are confident that it would lead directly to the realization of the I AM as an experiential attainable. From the standpoint of personal beings it could mean that the unknowable I AM had become experiencible as the Father\hyp{}Infinite. What these absolute destinies might mean from a nonpersonal standpoint is another matter and one which only eternity could possibly clarify. But as we view these remote eventualities as personal creatures, we deduce that the final destiny of all personalities is the final knowing of the Universal Father of these selfsame personalities.
\vs p106 8:23 As we philosophically conceive of the I AM in past eternity, he is alone, there is none beside him. Looking forward into future eternity, we do not see that the I AM could possibly change as an existential, but we are inclined to forecast a vast experiential difference. Such a concept of the I AM implies full self\hyp{}realization --- it embraces that limitless galaxy of personalities who have become volitional participants in the self\hyp{}revelation of the I AM, and who will remain eternally as absolute volitional parts of the totality of infinity, final sons of the absolute Father.
\usection{9.\bibnobreakspace Existential Infinite Unification}
\vs p106 9:1 In the concept of the Trinity of Trinities we postulate the possible experiential unification of limitless reality, and we sometimes theorize that all this may happen in the utter remoteness of far\hyp{}distant eternity. But there is nonetheless an actual and present unification of infinity in this very age as in all past and future universe ages; such unification is existential in the Paradise Trinity. Infinity unification as an experiential reality is unthinkably remote, but an unqualified unity of infinity now dominates the present moment of universe existence and unites the divergencies of all reality with an existential majesty that is \bibemph{absolute.}
\vs p106 9:2 When finite creatures attempt to conceive of infinite unification on the finality levels of consummated eternity, they are face to face with intellect limitations inherent in their finite existences. Time, space, and experience constitute barriers to creature concept; and yet, without time, apart from space, and except for experience, no creature could achieve even a limited comprehension of universe reality. Without time sensitivity, no evolutionary creature could possibly perceive the relations of sequence. Without space perception, no creature could fathom the relations of simultaneity. Without experience, no evolutionary creature could even exist; only the Seven Absolutes of Infinity really transcend experience, and even these may be experiential in certain phases.
\vs p106 9:3 Time, space, and experience are man’s greatest aids to relative reality perception and yet his most formidable obstacles to complete reality perception. Mortals and many other universe creatures find it necessary to think of potentials as being actualized in space and evolving to fruition in time, but this entire process is a time\hyp{}space phenomenon which does not actually take place on Paradise and in eternity. On the absolute level there is neither time nor space; all potentials may be there perceived as actuals.
\vs p106 9:4 The concept of the unification of all reality, be it in this or any other universe age, is basically twofold: existential and experiential. Such a unity is in process of experiential realization in the Trinity of Trinities, but the degree of the apparent actualization of this threefold Trinity is directly proportional to the disappearance of the qualifications and imperfections of reality in the cosmos. But total integration of reality is unqualifiedly and eternally and existentially present in the Paradise Trinity, within which, at this very universe moment, infinite reality is absolutely unified.
\vs p106 9:5 \pc The paradox created by the experiential and the existential viewpoints is inevitable and is predicated in part on the fact that the Paradise Trinity and the Trinity of Trinities are each an eternity relationship which mortals can only perceive as a time\hyp{}space relativity. The human concept of the gradual experiential actualization of the Trinity of Trinities --- the time viewpoint --- must be supplemented by the additional postulate that this \bibemph{is} already a factualization --- the eternity viewpoint. But how can these two viewpoints be reconciled? To finite mortals we suggest the acceptance of the truth that the Paradise Trinity is the existential unification of infinity, and that the inability to detect the actual presence and completed manifestation of the experiential Trinity of Trinities is in part due to reciprocal distortion because of:
\vs p106 9:6 \ublistelem{1.}\bibnobreakspace The limited human viewpoint, the inability to grasp the concept of unqualified eternity.
\vs p106 9:7 \ublistelem{2.}\bibnobreakspace The imperfect human status, the remoteness from the absolute level of experientials.
\vs p106 9:8 \ublistelem{3.}\bibnobreakspace The purpose of human existence, the fact that mankind is designed to evolve by the technique of experience and, therefore, must be inherently and constitutively dependent on experience. Only an Absolute can be both existential and experiential.
\vs p106 9:9 \pc The Universal Father in the Paradise Trinity is the I AM of the Trinity of Trinities, and the failure to experience the Father as infinite is due to finite limitations. The concept of the \bibemph{existential,} solitary, pre\hyp{}Trinity nonattainable I AM and the postulate of the \bibemph{experiential} post\hyp{}Trinity of Trinities and attainable I AM are one and the same hypothesis; no actual change has taken place in the Infinite; all apparent developments are due to increased capacities for reality reception and cosmic appreciation.
\vs p106 9:10 The I AM, in the final analysis, must exist \bibemph{before} all existentials and \bibemph{after} all experientials. While these ideas may not clarify the paradoxes of eternity and infinity in the human mind, they should at least stimulate such finite intellects to grapple anew with these never\hyp{}ending problems, problems which will continue to intrigue you on Salvington and later as finaliters and on throughout the unending future of your eternal careers in the wide\hyp{}spreading universes.
\vs p106 9:11 \pc Sooner or later all universe personalities begin to realize that the final quest of eternity is the endless exploration of infinity, the never\hyp{}ending voyage of discovery into the absoluteness of the First Source and Centre. Sooner or later we all become aware that all creature growth is proportional to Father identification. We arrive at the understanding that living the will of God is the eternal passport to the endless possibility of infinity itself. Mortals will sometime realize that success in the quest of the Infinite is directly proportional to the achievement of Fatherlikeness, and that in this universe age the realities of the Father are revealed within the qualities of divinity. And these qualities of divinity are personally appropriated by universe creatures in the experience of living divinely, and to live divinely means actually to live the will of God.
\vs p106 9:12 To material, evolutionary, finite creatures, a life predicated on the living of the Father’s will leads directly to the attainment of spirit supremacy in the personality arena and brings such creatures one step nearer the comprehension of the Father\hyp{}Infinite. Such a Father life is one predicated on truth, sensitive to beauty, and dominated by goodness. Such a God\hyp{}knowing person is inwardly illuminated by worship and outwardly devoted to the wholehearted service of the universal brotherhood of all personalities, a service ministry which is filled with mercy and motivated by love, while all these life qualities are unified in the evolving personality on ever\hyp{}ascending levels of cosmic wisdom, self\hyp{}realization, God\hyp{}finding, and Father worship.
\vsetoff
\vs p106 9:13 [Presented by a Melchizedek of Nebadon.]
