\upaper{115}{The Supreme Being}
\author{Mighty Messenger}
\vs p115 0:1 With God the Father, sonship is the great relationship. With God the Supreme, achievement is the prerequisite to status --- one must do something as well as be something.
\usection{1.\bibnobreakspace Relativity of Concept Frames}
\vs p115 1:1 Partial, incomplete, and evolving intellects would be helpless in the master universe, would be unable to form the first rational thought pattern, were it not for the innate ability of all mind, high or low, to form a \bibemph{universe frame} in which to think. If mind cannot fathom conclusions, if it cannot penetrate to true origins, then will such mind unfailingly postulate conclusions and invent origins that it may have a means of logical thought within the frame of these mind\hyp{}created postulates. And while such universe frames for creature thought are indispensable to rational intellectual operations, they are, without exception, erroneous to a greater or lesser degree.
\vs p115 1:2 Conceptual frames of the universe are only relatively true; they are serviceable scaffolding which must eventually give way before the expansions of enlarging cosmic comprehension. The understandings of truth, beauty, and goodness, morality, ethics, duty, love, divinity, origin, existence, purpose, destiny, time, space, even Deity, are only relatively true. God is much, much more than a Father, but the Father is man’s highest concept of God; nonetheless, the Father\hyp{}Son portrayal of Creator\hyp{}creature relationship will be augmented by those supermortal conceptions of Deity which will be attained in Orvonton, in Havona, and on Paradise. Man must think in a mortal universe frame, but that does not mean that he cannot envision other and higher frames within which thought can take place.
\vs p115 1:3 In order to facilitate mortal comprehension of the universe of universes, the diverse levels of cosmic reality have been designated as finite, absonite, and absolute. Of these only the absolute is unqualifiedly eternal, truly existential. Absonites and finites are derivatives, modifications, qualifications, and attenuations of the original and primordial absolute reality of infinity.
\vs p115 1:4 The realms of the finite exist by virtue of the eternal purpose of God. Finite creatures, high and low, may propound theories, and have done so, as to the necessity of the finite in the cosmic economy, but in the last analysis it exists because God so willed. The universe cannot be explained, neither can a finite creature offer a rational reason for his own individual existence without appealing to the prior acts and pre\hyp{}existent volition of ancestral beings, Creators or procreators.
\usection{2.\bibnobreakspace The Absolute Basis for Supremacy}
\vs p115 2:1 From the existential standpoint, nothing new can happen throughout the galaxies, for the completion of infinity inherent in the I AM is eternally present in the seven Absolutes, is functionally associated in the triunities, and is transmitively associated in the triodities. But the fact that infinity is thus existentially present in these absolute associations in no way makes it impossible to realize new cosmic experientials. From a finite creature’s viewpoint, infinity contains much that is potential, much that is on the order of a future possibility rather than a present actuality.
\vs p115 2:2 Value is a unique element in universe reality. We do not comprehend how the value of anything infinite and divine could possibly be increased. But we discover that \bibemph{meanings} can be modified if not augmented even in the relations of infinite Deity. To the experiential universes even divine values are increased as actualities by enlarged comprehension of reality meanings.
\vs p115 2:3 The entire scheme of universal creation and evolution on all experiencing levels is apparently a matter of the conversion of potentialities into actualities; and this transmutation has to do equally with the realms of space potency, mind potency, and spirit potency.
\vs p115 2:4 The apparent method whereby the possibilities of the cosmos are brought into actual existence varies from level to level, being experiential evolution in the finite and experiential eventuation in the absonite. Existential infinity is indeed unqualified in all\hyp{}inclusiveness, and this very all\hyp{}inclusiveness must, perforce, encompass even the possibility for evolutionary finite experiencing. And the possibility for such experiential growth becomes a universe actuality through triodity relationships impinging upon and in the Supreme.
\usection{3.\bibnobreakspace Original, Actual, and Potential}
\vs p115 3:1 The absolute cosmos is conceptually without limit; to define the extent and nature of this primal reality is to place qualifications upon infinity and to attenuate the pure concept of eternity. The idea of the infinite\hyp{}eternal, the eternal\hyp{}infinite, is unqualified in extent and absolute in fact. There is no language in the past, present, or future of Urantia adequate to express the reality of infinity or the infinity of reality. Man, a finite creature in an infinite cosmos, must content himself with distorted reflections and attenuated conceptions of that limitless, boundless, never\hyp{}beginning, never\hyp{}ending existence the comprehension of which is really beyond his ability.
\vs p115 3:2 Mind can never hope to grasp the concept of an Absolute without attempting first to break the unity of such a reality. Mind is unifying of all divergencies, but in the very absence of such divergencies, mind finds no basis upon which to attempt to formulate understanding concepts.
\vs p115 3:3 The primordial stasis of infinity requires segmentation prior to human attempts at comprehension. There is a unity in infinity which has been expressed in these papers as the I AM --- the premier postulate of the creature mind. But never can a creature understand how it is that this unity becomes duality, triunity, and diversity while yet remaining an unqualified unity. Man encounters a similar problem when he pauses to contemplate the undivided Deity of Trinity alongside the plural personalization of God.
\vs p115 3:4 It is only man’s distance from infinity that causes this concept to be expressed as one word. While infinity is on the one hand UNITY, on the other it is DIVERSITY without end or limit. Infinity, as it is observed by finite intelligences, is the maximum paradox of creature philosophy and finite metaphysics. Though man’s spiritual nature reaches up in the worship experience to the Father who is infinite, man’s intellectual comprehension capacity is exhausted by the maximum conception of the Supreme Being. Beyond the Supreme, concepts are increasingly names; less and less are they true designations of reality; more and more do they become the creature’s projection of finite understanding toward the superfinite.
\vs p115 3:5 \pc One basic conception of the absolute level involves a postulate of three phases:
\vs p115 3:6 \ublistelem{1.}\bibnobreakspace \bibemph{The Original.} The unqualified concept of the First Source and Centre, that source manifestation of the I AM from which all reality takes origin.
\vs p115 3:7 \ublistelem{2.}\bibnobreakspace \bibemph{The Actual.} The union of the three Absolutes of actuality, the Second, Third, and Paradise Sources and Centres. This triodity of the Eternal Son, the Infinite Spirit, and the Paradise Isle constitutes the actual revelation of the originality of the First Source and Centre.
\vs p115 3:8 \ublistelem{3.}\bibnobreakspace \bibemph{The Potential.} The union of the three Absolutes of potentiality, the Deity, Unqualified, and Universal Absolutes. This triodity of existential potentiality constitutes the potential revelation of the originality of the First Source and Centre.
\vs p115 3:9 \pc The interassociation of the Original, the Actual, and the Potential yields the tensions within infinity which result in the possibility for all universe growth; and growth is the nature of the Sevenfold, the Supreme, and the Ultimate.
\vs p115 3:10 In the association of the Deity, Universal, and Unqualified Absolutes, potentiality is absolute while actuality is emergent; in the association of the Second, Third, and Paradise Sources and Centres, actuality is absolute while potentiality is emergent; in the originality of the First Source and Centre, we cannot say that either actuality or potentiality is either existent or emergent --- \bibemph{the Father is.}
\vs p115 3:11 From the time viewpoint, the Actual is that which was and is; the Potential is that which is becoming and will be; the Original is that which is. From the eternity viewpoint, the differences between the Original, the Actual, and the Potential are not thus apparent. These triune qualities are not so distinguished on Paradise\hyp{}eternity levels. In eternity all is --- only has all not yet been revealed in time and space.
\vs p115 3:12 From a creature’s viewpoint, actuality is substance, potentiality is capacity. Actuality exists centremost and expands therefrom into peripheral infinity; potentiality comes inward from the infinity periphery and converges at the centre of all things. Originality is that which first causes and then balances the dual motions of the cycle of reality metamorphosis from potentials to actuals and the potentializing of existing actuals.
\vs p115 3:13 The three Absolutes of potentiality are operative on the purely eternal level of the cosmos, hence never function as such on subabsolute levels. On the descending levels of reality the triodity of potentiality is manifest with the Ultimate and upon the Supreme. The potential may fail to time\hyp{}actualize with respect to a part on some subabsolute level, but never in the aggregate. The will of God does ultimately prevail, not always concerning the individual but invariably concerning the total.
\vs p115 3:14 It is in the triodity of actuality that the existents of the cosmos have their centre; be it spirit, mind, or energy, all centre in this association of the Son, the Spirit, and Paradise. The personality of the spirit Son is the master pattern for all personality throughout all universes. The substance of the Paradise Isle is the master pattern of which Havona is a perfect, and the superuniverses are a perfecting, revelation. The Conjoint Actor is at one and the same time the mind activation of cosmic energy, the conceptualization of spirit purpose, and the integration of the mathematical causes and effects of the material levels with the volitional purposes and motives of the spiritual level. In and to a finite universe the Son, Spirit, and Paradise function in and upon the Ultimate as he is conditioned and qualified in the Supreme.
\vs p115 3:15 Actuality (of Deity) is what man seeks in the Paradise ascent. Potentiality (of human divinity) is what man evolves in that search. The Original is what makes possible the coexistence and integration of man the actual, man the potential, and man the eternal.
\vs p115 3:16 \pc The final dynamics of the cosmos have to do with the continual transfer of reality from potentiality to actuality. In theory, there may be an end to this metamorphosis, but in fact, such is impossible since the Potential and the Actual are both encircuited in the Original (the I AM), and this identification makes it forever impossible to place a limit on the developmental progression of the universe. Whatsoever is identified with the I AM can never find an end to progression since the actuality of the potentials of the I AM is absolute, and the potentiality of the actuals of the I AM is also absolute. Always will actuals be opening up new avenues of the realization of hitherto impossible potentials --- every human decision not only actualizes a new reality in human experience but also opens up a new capacity for human growth. The man lives in every child, and the morontia progressor is resident in the mature God\hyp{}knowing man.
\vs p115 3:17 Statics in growth can never appear in the total cosmos since the basis for growth --- the absolute actuals --- is unqualified, and since the possibilities for growth --- the absolute potentials --- are unlimited. From a practical viewpoint the philosophers of the universe have come to the conclusion that there is no such thing as an \bibemph{end.}
\vs p115 3:18 From a circumscribed view there are, indeed, many ends, many terminations of activities, but from a larger viewpoint on a higher universe level, there are no endings, merely transitions from one phase of development to another. The major chronicity of the master universe is concerned with the several universe ages, the Havona, the superuniverse, and the outer universe ages. But even these basic divisions of sequence relationships cannot be more than relative landmarks on the unending highway of eternity.
\vs p115 3:19 The final penetration of the truth, beauty, and goodness of the Supreme Being could only open up to the progressing creature those absonite qualities of ultimate divinity which lie beyond the concept levels of truth, beauty, and goodness.
\usection{4.\bibnobreakspace Sources of Supreme Reality}
\vs p115 4:1 Any consideration of the \bibemph{origins} of God the Supreme must begin with the Paradise Trinity, for the Trinity is original Deity while the Supreme is derived Deity. Any consideration of the \bibemph{growth} of the Supreme must give consideration to the existential triodities, for they encompass all absolute actuality and all infinite potentiality (in conjunction with the First Source and Centre). And the evolutionary Supreme is the culminating and personally volitional focus of the transmutation --- the transformation --- of potentials to actuals in and on the finite level of existence. The two triodities, actual and potential, encompass the totality of the interrelationships of growth in the universes.
\vs p115 4:2 The source of the Supreme is in the Paradise Trinity --- eternal, actual, and undivided Deity. The Supreme is first of all a spirit person, and this spirit person stems from the Trinity. But the Supreme is secondly a Deity of growth --- evolutionary growth --- and this growth derives from the two triodities, actual and potential.
\vs p115 4:3 If it is difficult to comprehend that the infinite triodities can function on the finite level, pause to consider that their very infinity must in itself contain the potentiality of the finite; infinity encompasses all things ranging from the lowest and most qualified finite existence to the highest and unqualifiedly absolute realities.
\vs p115 4:4 It is not so difficult to comprehend that the infinite does contain the finite as it is to understand just how this infinite actually is manifest to the finite. But the Thought Adjusters indwelling mortal man are one of the eternal proofs that even the absolute God (as absolute) can and does actually make direct contact with even the lowest and least of all universe will creatures.
\vs p115 4:5 The triodities which collectively encompass the actual and the potential are manifest on the finite level in conjunction with the Supreme Being. The technique of such manifestation is both direct and indirect: direct in so far as triodity relations repercuss directly in the Supreme and indirect in so far as they are derived through the eventuated level of the absonite.
\vs p115 4:6 Supreme reality, which is total finite reality, is in process of dynamic growth between the unqualified potentials of outer space and the unqualified actuals at the centre of all things. The finite domain thus factualizes through the co\hyp{}operation of the absonite agencies of Paradise and the Supreme Creator Personalities of time. The act of maturing the qualified possibilities of the three great potential Absolutes is the absonite function of the Architects of the Master Universe and their transcendental associates. And when these eventualities have attained to a certain point of maturation, the Supreme Creator Personalities emerge from Paradise to engage in the agelong task of bringing the evolving universes into factual being.
\vs p115 4:7 The growth of Supremacy derives from the triodities; the spirit person of the Supreme, from the Trinity; but the power prerogatives of the Almighty are predicated on the divinity successes of God the Sevenfold, while the conjoining of the power prerogatives of the Almighty Supreme with the spirit person of God the Supreme takes place by virtue of the ministry of the Conjoint Actor, who bestowed the mind of the Supreme as the conjoining factor in this evolutionary Deity.
\usection{5.\bibnobreakspace Relation of the Supreme to the Paradise Trinity}
\vs p115 5:1 The Supreme Being is absolutely dependent on the existence and action of the Paradise Trinity for the reality of his personal and spirit nature. While the growth of the Supreme is a matter of triodity relationship, the spirit personality of God the Supreme is dependent upon, and is derived from, the Paradise Trinity, which ever remains as the absolute centre\hyp{}source of perfect and infinite stability around which the evolutionary growth of the Supreme progressively unfolds.
\vs p115 5:2 The function of the Trinity is related to the function of the Supreme, for the Trinity is functional on all (total) levels, including the level of the function of Supremacy. But as the age of Havona gives way to the age of the superuniverses, so does the discernible action of the Trinity as immediate creator give way to the creative acts of the children of the Paradise Deities.
\usection{6.\bibnobreakspace Relation of the Supreme to the Triodities}
\vs p115 6:1 The triodity of actuality continues to function directly in the post\hyp{}Havona epochs; Paradise gravity grasps the basic units of material existence, the spirit gravity of the Eternal Son operates directly upon the fundamental values of spirit existence, and the mind gravity of the Conjoint Actor unerringly clutches all vital meanings of intellectual existence.
\vs p115 6:2 But as each stage of creative activity proceeds out through uncharted space, it functions and exists farther and farther removed from direct action by the creative forces and divine personalities of central emplacement --- the absolute Isle of Paradise and the infinite Deities resident thereon. These successive levels of cosmic existence become, therefore, increasingly dependent upon developments within the three Absolute potentialities of infinity.
\vs p115 6:3 The Supreme Being embraces possibilities for cosmic ministry that are not apparently manifested in the Eternal Son, the Infinite Spirit, or the nonpersonal realities of the Isle of Paradise. This statement is made with due regard for the absoluteness of these three basic actualities, but the growth of the Supreme is not only predicated on these actualities of Deity and Paradise but is also involved in developments within the Deity, Universal, and Unqualified Absolutes.
\vs p115 6:4 \pc The Supreme not only grows as the Creators and creatures of the evolving universes attain to Godlikeness, but this finite Deity also experiences growth as a result of the creature and Creator mastery of the finite possibilities of the grand universe. The motion of the Supreme is twofold: intensively toward Paradise and Deity and extensively toward the limitlessness of the Absolutes of potential.
\vs p115 6:5 In the present universe age this dual motion is revealed in the descending and ascending personalities of the grand universe. The Supreme Creator Personalities and all their divine associates are reflective of the outward, diverging motion of the Supreme, while the ascending pilgrims from the seven superuniverses are indicative of the inward, converging trend of Supremacy.
\vs p115 6:6 Always is the finite Deity seeking for dual correlation, inward toward Paradise and the Deities thereof and outward toward infinity and the Absolutes therein. The mighty eruption of the Paradise\hyp{}creative divinity personalizing in the Creator Sons and powerizing in the power controllers, signifies the vast outsurge of Supremacy into the domains of potentiality, while the endless procession of the ascending creatures of the grand universe witnesses the mighty insurge of Supremacy toward unity with Paradise Deity.
\vs p115 6:7 Human beings have learned that the motion of the invisible may sometimes be discerned by observing its effects on the visible; and we in the universes have long since learned to detect the movements and trends of Supremacy by observing the repercussions of such evolutions in the personalities and patterns of the grand universe.
\vs p115 6:8 Though we are not sure, we believe that, as a finite reflection of Paradise Deity, the Supreme is engaged in an eternal progression into outer space; but as a qualification of the three Absolute potentials of outer space, this Supreme Being is forever seeking for Paradise coherence. And these dual motions seem to account for most of the basic activities in the presently organized universes.
\usection{7.\bibnobreakspace The Nature of the Supreme}
\vs p115 7:1 In the Deity of the Supreme the Father\hyp{}I AM has achieved relatively complete liberation from the limitations inherent in infinity of status, eternity of being, and absoluteness of nature. But God the Supreme has been freed from all existential limitations only by having become subject to experiential qualifications of universal function. In attaining capacity for experience, the finite God also becomes subject to the necessity therefor; in achieving liberation from eternity, the Almighty encounters the barriers of time; and the Supreme could only know growth and development as a consequence of partiality of existence and incompleteness of nature, nonabsoluteness of being.
\vs p115 7:2 All this must be according to the Father’s plan, which has predicated finite progress upon effort, creature achievement upon perseverance, and personality development upon faith. By thus ordaining the experience\hyp{}evolution of the Supreme, the Father has made it possible for finite creatures to exist in the universes and, by experiential progression, sometime to attain the divinity of Supremacy.
\vs p115 7:3 \pc Including the Supreme and even the Ultimate, all reality, excepting the unqualified values of the seven Absolutes, is relative. The fact of Supremacy is predicated on Paradise power, Son personality, and Conjoint action, but the growth of the Supreme is involved in the Deity Absolute, the Unqualified Absolute, and the Universal Absolute. And this synthesizing and unifying Deity --- God the Supreme --- is the personification of the finite shadow cast athwart the grand universe by the infinite unity of the unsearchable nature of the Paradise Father, the First Source and Centre.
\vs p115 7:4 To the extent that the triodities are directly operative on the finite level, they impinge upon the Supreme, who is the Deity focalization and cosmic summation of the finite qualifications of the natures of the Absolute Actual and the Absolute Potential.
\vs p115 7:5 \pc The Paradise Trinity is considered to be the absolute inevitability; the Seven Master Spirits are apparently Trinity inevitabilities; the power\hyp{}mind\hyp{}spirit\hyp{}personality actualization of the Supreme must be the evolutionary inevitability.
\vs p115 7:6 God the Supreme does not appear to have been inevitable in unqualified infinity, but he seems to be on all relativity levels. He is the indispensable focalizer, summarizer, and encompasser of evolutionary experience, effectively unifying the results of this mode of reality perception in his Deity nature. And all this he appears to do for the purpose of contributing to the appearance of the \bibemph{inevitable eventuation,} the superexperience and superfinite manifestation of God the Ultimate.
\vs p115 7:7 The Supreme Being cannot be fully appreciated without taking into consideration source, function, and destiny: relationship to the originating Trinity, the universe of activity, and the Trinity Ultimate of immediate destiny.
\vs p115 7:8 By the process of summating evolutionary experience the Supreme connects the finite with the absonite, even as the mind of the Conjoint Actor integrates the divine spirituality of the personal Son with the immutable energies of the Paradise pattern, and as the presence of the Universal Absolute unifies Deity activation with the Unqualified reactivity. And this unity must be a revelation of the undetected working of the original unity of the First Father\hyp{}Cause and Source\hyp{}Pattern of all things and all beings.
\vsetoff
\vs p115 7:9 [Sponsored by a Mighty Messenger temporarily sojourning on Urantia.]
\quizlink
