\bibbook{\bibtitlemain{the proverbs}}{proverbs}{Proverbs}{Pro}
\vs Pro 1:1 The proverbs of Solomon the son of David, king of Israel;
\vs Pro 1:2 To know wisdom and instruction; to perceive the words of understanding;
\vs Pro 1:3 To receive the instruction of wisdom, justice, and judgment, and equity;
\fn{equity: Heb. equities.}
\vs Pro 1:4 To give subtilty to the simple, to the young man knowledge and discretion.
\fn{discretion: or, advisement.}
\vs Pro 1:5 A wise \bibemph{man} will hear, and will increase learning; and a man of understanding shall attain unto wise counsels:
\vs Pro 1:6 To understand a proverb, and the interpretation; the words of the wise, and their dark sayings.
\fn{the interpretation: or, an eloquent speech.}
\vs Pro 1:7 The fear of the \textsc{Lord} \bibemph{is} the beginning of knowledge: \bibemph{but} fools despise wisdom and instruction.
\fn{the beginning: or, the principal part.}
\vs Pro 1:8 My son, hear the instruction of thy father, and forsake not the law of thy mother:
\vs Pro 1:9 For they \bibemph{shall be} an ornament of grace unto thy head, and chains about thy neck.
\fn{an...: Heb. an adding.}
\vs Pro 1:10 My son, if sinners entice thee, consent thou not.
\vs Pro 1:11 If they say, Come with us, let us lay wait for blood, let us lurk privily for the innocent without cause:
\vs Pro 1:12 Let us swallow them up alive as the grave; and whole, as those that go down into the pit:
\vs Pro 1:13 We shall find all precious substance, we shall fill our houses with spoil:
\vs Pro 1:14 Cast in thy lot among us; let us all have one purse:
\vs Pro 1:15 My son, walk not thou in the way with them; refrain thy foot from their path:
\vs Pro 1:16 For their feet run to evil, and make haste to shed blood.
\vs Pro 1:17 Surely in vain the net is spread in the sight of any bird.
\fn{in the...: Heb. in the eyes of every thing that hath a wing.}
\vs Pro 1:18 And they lay wait for their \bibemph{own} blood; they lurk privily for their \bibemph{own} lives.
\vs Pro 1:19 So \bibemph{are} the ways of every one that is greedy of gain; \bibemph{which} taketh away the life of the owners thereof.
\vs Pro 1:20 Wisdom crieth without; she uttereth her voice in the streets:
\fn{Wisdom: Heb. Wisdoms, that is, Excellent wisdom.}
\vs Pro 1:21 She crieth in the chief place of concourse, in the openings of the gates: in the city she uttereth her words, \bibemph{saying},
\vs Pro 1:22 How long, ye simple ones, will ye love simplicity? and the scorners delight in their scorning, and fools hate knowledge?
\vs Pro 1:23 Turn you at my reproof: behold, I will pour out my spirit unto you, I will make known my words unto you.
\vs Pro 1:24 Because I have called, and ye refused; I have stretched out my hand, and no man regarded;
\vs Pro 1:25 But ye have set at nought all my counsel, and would none of my reproof:
\vs Pro 1:26 I also will laugh at your calamity; I will mock when your fear cometh;
\vs Pro 1:27 When your fear cometh as desolation, and your destruction cometh as a whirlwind; when distress and anguish cometh upon you.
\vs Pro 1:28 Then shall they call upon me, but I will not answer; they shall seek me early, but they shall not find me:
\vs Pro 1:29 For that they hated knowledge, and did not choose the fear of the \textsc{Lord}:
\vs Pro 1:30 They would none of my counsel: they despised all my reproof.
\vs Pro 1:31 Therefore shall they eat of the fruit of their own way, and be filled with their own devices.
\vs Pro 1:32 For the turning away of the simple shall slay them, and the prosperity of fools shall destroy them.
\fn{turning...: or, ease of the simple.}
\vs Pro 1:33 But whoso hearkeneth unto me shall dwell safely, and shall be quiet from fear of evil.
\vs Pro 2:1 My son, if thou wilt receive my words, and hide my commandments with thee;
\vs Pro 2:2 So that thou incline thine ear unto wisdom, \bibemph{and} apply thine heart to understanding;
\vs Pro 2:3 Yea, if thou criest after knowledge, \bibemph{and} liftest up thy voice for understanding;
\fn{liftest...: Heb. givest thy voice.}
\vs Pro 2:4 If thou seekest her as silver, and searchest for her as \bibemph{for} hid treasures;
\vs Pro 2:5 Then shalt thou understand the fear of the \textsc{Lord}, and find the knowledge of God.
\vs Pro 2:6 For the \textsc{Lord} giveth wisdom: out of his mouth \bibemph{cometh} knowledge and understanding.
\vs Pro 2:7 He layeth up sound wisdom for the righteous: \bibemph{he is} a buckler to them that walk uprightly.
\vs Pro 2:8 He keepeth the paths of judgment, and preserveth the way of his saints.
\vs Pro 2:9 Then shalt thou understand righteousness, and judgment, and equity; \bibemph{yea}, every good path.
\vs Pro 2:10 When wisdom entereth into thine heart, and knowledge is pleasant unto thy soul;
\vs Pro 2:11 Discretion shall preserve thee, understanding shall keep thee:
\vs Pro 2:12 To deliver thee from the way of the evil \bibemph{man}, from the man that speaketh froward things;
\vs Pro 2:13 Who leave the paths of uprightness, to walk in the ways of darkness;
\vs Pro 2:14 Who rejoice to do evil, \bibemph{and} delight in the frowardness of the wicked;
\vs Pro 2:15 Whose ways \bibemph{are} crooked, and \bibemph{they} froward in their paths:
\vs Pro 2:16 To deliver thee from the strange woman, \bibemph{even} from the stranger \bibemph{which} flattereth with her words;
\vs Pro 2:17 Which forsaketh the guide of her youth, and forgetteth the covenant of her God.
\vs Pro 2:18 For her house inclineth unto death, and her paths unto the dead.
\vs Pro 2:19 None that go unto her return again, neither take they hold of the paths of life.
\vs Pro 2:20 That thou mayest walk in the way of good \bibemph{men}, and keep the paths of the righteous.
\vs Pro 2:21 For the upright shall dwell in the land, and the perfect shall remain in it.
\vs Pro 2:22 But the wicked shall be cut off from the earth, and the transgressors shall be rooted out of it.
\fn{rooted: or, plucked up.}
\vs Pro 3:1 My son, forget not my law; but let thine heart keep my commandments:
\vs Pro 3:2 For length of days, and long life, and peace, shall they add to thee.
\fn{long...: Heb. years of life.}
\vs Pro 3:3 Let not mercy and truth forsake thee: bind them about thy neck; write them upon the table of thine heart:
\vs Pro 3:4 So shalt thou find favour and good understanding in the sight of God and man.
\fn{good...: or, good success.}
\vs Pro 3:5 Trust in the \textsc{Lord} with all thine heart; and lean not unto thine own understanding.
\vs Pro 3:6 In all thy ways acknowledge him, and he shall direct thy paths.
\vs Pro 3:7 Be not wise in thine own eyes: fear the \textsc{Lord}, and depart from evil.
\vs Pro 3:8 It shall be health to thy navel, and marrow to thy bones.
\fn{health: Heb. medicine. marrow: Heb. watering, or, moistening.}
\vs Pro 3:9 Honour the \textsc{Lord} with thy substance, and with the firstfruits of all thine increase:
\vs Pro 3:10 So shall thy barns be filled with plenty, and thy presses shall burst out with new wine.
\vs Pro 3:11 My son, despise not the chastening of the \textsc{Lord}; neither be weary of his correction:
\vs Pro 3:12 For whom the \textsc{Lord} loveth he correcteth; even as a father the son \bibemph{in whom} he delighteth.
\vs Pro 3:13 Happy \bibemph{is} the man \bibemph{that} findeth wisdom, and the man \bibemph{that} getteth understanding.
\fn{that getteth...: Heb. that draweth out understanding.}
\vs Pro 3:14 For the merchandise of it \bibemph{is} better than the merchandise of silver, and the gain thereof than fine gold.
\vs Pro 3:15 She \bibemph{is} more precious than rubies: and all the things thou canst desire are not to be compared unto her.
\vs Pro 3:16 Length of days \bibemph{is} in her right hand; \bibemph{and} in her left hand riches and honour.
\vs Pro 3:17 Her ways \bibemph{are} ways of pleasantness, and all her paths \bibemph{are} peace.
\vs Pro 3:18 She \bibemph{is} a tree of life to them that lay hold upon her: and happy \bibemph{is every one} that retaineth her.
\vs Pro 3:19 The \textsc{Lord} by wisdom hath founded the earth; by understanding hath he established the heavens.
\fn{established: or, prepared.}
\vs Pro 3:20 By his knowledge the depths are broken up, and the clouds drop down the dew.
\vs Pro 3:21 My son, let not them depart from thine eyes: keep sound wisdom and discretion:
\vs Pro 3:22 So shall they be life unto thy soul, and grace to thy neck.
\vs Pro 3:23 Then shalt thou walk in thy way safely, and thy foot shall not stumble.
\vs Pro 3:24 When thou liest down, thou shalt not be afraid: yea, thou shalt lie down, and thy sleep shall be sweet.
\vs Pro 3:25 Be not afraid of sudden fear, neither of the desolation of the wicked, when it cometh.
\vs Pro 3:26 For the \textsc{Lord} shall be thy confidence, and shall keep thy foot from being taken.
\vs Pro 3:27 Withhold not good from them to whom it is due, when it is in the power of thine hand to do \bibemph{it}.
\fn{them...: Heb. the owners thereof.}
\vs Pro 3:28 Say not unto thy neighbour, Go, and come again, and to morrow I will give; when thou hast it by thee.
\vs Pro 3:29 Devise not evil against thy neighbour, seeing he dwelleth securely by thee.
\fn{Devise...: or, Practise no evil.}
\vs Pro 3:30 Strive not with a man without cause, if he have done thee no harm.
\vs Pro 3:31 Envy thou not the oppressor, and choose none of his ways.
\fn{the oppressor: Heb. a man of violence.}
\vs Pro 3:32 For the froward \bibemph{is} abomination to the \textsc{Lord}: but his secret \bibemph{is} with the righteous.
\vs Pro 3:33 The curse of the \textsc{Lord} \bibemph{is} in the house of the wicked: but he blesseth the habitation of the just.
\vs Pro 3:34 Surely he scorneth the scorners: but he giveth grace unto the lowly.
\vs Pro 3:35 The wise shall inherit glory: but shame shall be the promotion of fools.
\fn{shall be...: Heb. exalteth the fools.}
\vs Pro 4:1 Hear, ye children, the instruction of a father, and attend to know understanding.
\vs Pro 4:2 For I give you good doctrine, forsake ye not my law.
\vs Pro 4:3 For I was my father's son, tender and only \bibemph{beloved} in the sight of my mother.
\vs Pro 4:4 He taught me also, and said unto me, Let thine heart retain my words: keep my commandments, and live.
\vs Pro 4:5 Get wisdom, get understanding: forget \bibemph{it} not; neither decline from the words of my mouth.
\vs Pro 4:6 Forsake her not, and she shall preserve thee: love her, and she shall keep thee.
\vs Pro 4:7 Wisdom \bibemph{is} the principal thing; \bibemph{therefore} get wisdom: and with all thy getting get understanding.
\vs Pro 4:8 Exalt her, and she shall promote thee: she shall bring thee to honour, when thou dost embrace her.
\vs Pro 4:9 She shall give to thine head an ornament of grace: a crown of glory shall she deliver to thee.
\fn{a crown...: or, she shall compass thee with a crown of glory.}
\vs Pro 4:10 Hear, O my son, and receive my sayings; and the years of thy life shall be many.
\vs Pro 4:11 I have taught thee in the way of wisdom; I have led thee in right paths.
\vs Pro 4:12 When thou goest, thy steps shall not be straitened; and when thou runnest, thou shalt not stumble.
\vs Pro 4:13 Take fast hold of instruction; let \bibemph{her} not go: keep her; for she \bibemph{is} thy life.
\vs Pro 4:14 Enter not into the path of the wicked, and go not in the way of evil \bibemph{men}.
\vs Pro 4:15 Avoid it, pass not by it, turn from it, and pass away.
\vs Pro 4:16 For they sleep not, except they have done mischief; and their sleep is taken away, unless they cause \bibemph{some} to fall.
\vs Pro 4:17 For they eat the bread of wickedness, and drink the wine of violence.
\vs Pro 4:18 But the path of the just \bibemph{is} as the shining light, that shineth more and more unto the perfect day.
\vs Pro 4:19 The way of the wicked \bibemph{is} as darkness: they know not at what they stumble.
\vs Pro 4:20 My son, attend to my words; incline thine ear unto my sayings.
\vs Pro 4:21 Let them not depart from thine eyes; keep them in the midst of thine heart.
\vs Pro 4:22 For they \bibemph{are} life unto those that find them, and health to all their flesh.
\fn{health: Heb. medicine.}
\vs Pro 4:23 Keep thy heart with all diligence; for out of it \bibemph{are} the issues of life.
\fn{with...: Heb. above all keeping.}
\vs Pro 4:24 Put away from thee a froward mouth, and perverse lips put far from thee.
\fn{a froward...: Heb. frowardness of mouth and perverseness of lips.}
\vs Pro 4:25 Let thine eyes look right on, and let thine eyelids look straight before thee.
\vs Pro 4:26 Ponder the path of thy feet, and let all thy ways be established.
\fn{let...: or, all thy ways shall be ordered aright.}
\vs Pro 4:27 Turn not to the right hand nor to the left: remove thy foot from evil.
\vs Pro 5:1 My son, attend unto my wisdom, \bibemph{and} bow thine ear to my understanding:
\vs Pro 5:2 That thou mayest regard discretion, and \bibemph{that} thy lips may keep knowledge.
\vs Pro 5:3 For the lips of a strange woman drop \bibemph{as} an honeycomb, and her mouth \bibemph{is} smoother than oil:
\fn{mouth: Heb. palate.}
\vs Pro 5:4 But her end is bitter as wormwood, sharp as a twoedged sword.
\vs Pro 5:5 Her feet go down to death; her steps take hold on hell.
\vs Pro 5:6 Lest thou shouldest ponder the path of life, her ways are moveable, \bibemph{that} thou canst not know \bibemph{them}.
\vs Pro 5:7 Hear me now therefore, O ye children, and depart not from the words of my mouth.
\vs Pro 5:8 Remove thy way far from her, and come not nigh the door of her house:
\vs Pro 5:9 Lest thou give thine honour unto others, and thy years unto the cruel:
\vs Pro 5:10 Lest strangers be filled with thy wealth; and thy labours \bibemph{be} in the house of a stranger;
\fn{thy wealth: Heb. thy strength.}
\vs Pro 5:11 And thou mourn at the last, when thy flesh and thy body are consumed,
\vs Pro 5:12 And say, How have I hated instruction, and my heart despised reproof;
\vs Pro 5:13 And have not obeyed the voice of my teachers, nor inclined mine ear to them that instructed me!
\vs Pro 5:14 I was almost in all evil in the midst of the congregation and assembly.
\vs Pro 5:15 Drink waters out of thine own cistern, and running waters out of thine own well.
\vs Pro 5:16 Let thy fountains be dispersed abroad, \bibemph{and} rivers of waters in the streets.
\vs Pro 5:17 Let them be only thine own, and not strangers' with thee.
\vs Pro 5:18 Let thy fountain be blessed: and rejoice with the wife of thy youth.
\vs Pro 5:19 \bibemph{Let her be as} the loving hind and pleasant roe; let her breasts satisfy thee at all times; and be thou ravished always with her love.
\fn{satisfy...: Heb. water thee. be thou...: Heb. err thou always in her love.}
\vs Pro 5:20 And why wilt thou, my son, be ravished with a strange woman, and embrace the bosom of a stranger?
\vs Pro 5:21 For the ways of man \bibemph{are} before the eyes of the \textsc{Lord}, and he pondereth all his goings.
\vs Pro 5:22 His own iniquities shall take the wicked himself, and he shall be holden with the cords of his sins.
\fn{sins: Heb. sin.}
\vs Pro 5:23 He shall die without instruction; and in the greatness of his folly he shall go astray.
\vs Pro 6:1 My son, if thou be surety for thy friend, \bibemph{if} thou hast stricken thy hand with a stranger,
\vs Pro 6:2 Thou art snared with the words of thy mouth, thou art taken with the words of thy mouth.
\vs Pro 6:3 Do this now, my son, and deliver thyself, when thou art come into the hand of thy friend; go, humble thyself, and make sure thy friend.
\fn{and make...: or, so shalt thou prevail with thy friend.}
\vs Pro 6:4 Give not sleep to thine eyes, nor slumber to thine eyelids.
\vs Pro 6:5 Deliver thyself as a roe from the hand \bibemph{of the hunter}, and as a bird from the hand of the fowler.
\vs Pro 6:6 Go to the ant, thou sluggard; consider her ways, and be wise:
\vs Pro 6:7 Which having no guide, overseer, or ruler,
\vs Pro 6:8 Provideth her meat in the summer, \bibemph{and} gathereth her food in the harvest.
\vs Pro 6:9 How long wilt thou sleep, O sluggard? when wilt thou arise out of thy sleep?
\vs Pro 6:10 \bibemph{Yet} a little sleep, a little slumber, a little folding of the hands to sleep:
\vs Pro 6:11 So shall thy poverty come as one that travelleth, and thy want as an armed man.
\vs Pro 6:12 A naughty person, a wicked man, walketh with a froward mouth.
\vs Pro 6:13 He winketh with his eyes, he speaketh with his feet, he teacheth with his fingers;
\vs Pro 6:14 Frowardness \bibemph{is} in his heart, he deviseth mischief continually; he soweth discord.
\fn{soweth: Heb. casteth forth.}
\vs Pro 6:15 Therefore shall his calamity come suddenly; suddenly shall he be broken without remedy.
\vs Pro 6:16 These six \bibemph{things} doth the \textsc{Lord} hate: yea, seven \bibemph{are} an abomination unto him:
\fn{unto...: Heb. of his soul.}
\vs Pro 6:17 A proud look, a lying tongue, and hands that shed innocent blood,
\fn{A proud...: Heb. Haughty eyes.}
\vs Pro 6:18 An heart that deviseth wicked imaginations, feet that be swift in running to mischief,
\vs Pro 6:19 A false witness \bibemph{that} speaketh lies, and he that soweth discord among brethren.
\vs Pro 6:20 My son, keep thy father's commandment, and forsake not the law of thy mother:
\vs Pro 6:21 Bind them continually upon thine heart, \bibemph{and} tie them about thy neck.
\vs Pro 6:22 When thou goest, it shall lead thee; when thou sleepest, it shall keep thee; and \bibemph{when} thou awakest, it shall talk with thee.
\vs Pro 6:23 For the commandment \bibemph{is} a lamp; and the law \bibemph{is} light; and reproofs of instruction \bibemph{are} the way of life:
\fn{lamp: or, candle.}
\vs Pro 6:24 To keep thee from the evil woman, from the flattery of the tongue of a strange woman.
\fn{of the...: or, of the strange tongue.}
\vs Pro 6:25 Lust not after her beauty in thine heart; neither let her take thee with her eyelids.
\vs Pro 6:26 For by means of a whorish woman \bibemph{a man is brought} to a piece of bread: and the adulteress will hunt for the precious life.
\fn{the adulteress: Heb. the woman of a man, or, a man's wife.}
\vs Pro 6:27 Can a man take fire in his bosom, and his clothes not be burned?
\vs Pro 6:28 Can one go upon hot coals, and his feet not be burned?
\vs Pro 6:29 So he that goeth in to his neighbour's wife; whosoever toucheth her shall not be innocent.
\vs Pro 6:30 \bibemph{Men} do not despise a thief, if he steal to satisfy his soul when he is hungry;
\vs Pro 6:31 But \bibemph{if} he be found, he shall restore sevenfold; he shall give all the substance of his house.
\vs Pro 6:32 \bibemph{But} whoso committeth adultery with a woman lacketh understanding: he \bibemph{that} doeth it destroyeth his own soul.
\fn{understanding: Heb. heart.}
\vs Pro 6:33 A wound and dishonour shall he get; and his reproach shall not be wiped away.
\vs Pro 6:34 For jealousy \bibemph{is} the rage of a man: therefore he will not spare in the day of vengeance.
\vs Pro 6:35 He will not regard any ransom; neither will he rest content, though thou givest many gifts.
\fn{He will...: Heb. He will not accept the face of any ransom.}
\vs Pro 7:1 My son, keep my words, and lay up my commandments with thee.
\vs Pro 7:2 Keep my commandments, and live; and my law as the apple of thine eye.
\vs Pro 7:3 Bind them upon thy fingers, write them upon the table of thine heart.
\vs Pro 7:4 Say unto wisdom, Thou \bibemph{art} my sister; and call understanding \bibemph{thy} kinswoman:
\vs Pro 7:5 That they may keep thee from the strange woman, from the stranger \bibemph{which} flattereth with her words.
\vs Pro 7:6 For at the window of my house I looked through my casement,
\vs Pro 7:7 And beheld among the simple ones, I discerned among the youths, a young man void of understanding,
\fn{the youths: Heb. the sons.}
\vs Pro 7:8 Passing through the street near her corner; and he went the way to her house,
\vs Pro 7:9 In the twilight, in the evening, in the black and dark night:
\fn{in the evening of days.}
\vs Pro 7:10 And, behold, there met him a woman \bibemph{with} the attire of an harlot, and subtil of heart.
\vs Pro 7:11 (She \bibemph{is} loud and stubborn; her feet abide not in her house:
\vs Pro 7:12 Now \bibemph{is she} without, now in the streets, and lieth in wait at every corner.)
\vs Pro 7:13 So she caught him, and kissed him, \bibemph{and} with an impudent face said unto him,
\fn{with...: Heb. she strengthened her face and said.}
\vs Pro 7:14 \bibemph{I have} peace offerings with me; this day have I payed my vows.
\fn{I have...: Heb. Peace offerings are upon me.}
\vs Pro 7:15 Therefore came I forth to meet thee, diligently to seek thy face, and I have found thee.
\vs Pro 7:16 I have decked my bed with coverings of tapestry, with carved \bibemph{works}, with fine linen of Egypt.
\vs Pro 7:17 I have perfumed my bed with myrrh, aloes, and cinnamon.
\vs Pro 7:18 Come, let us take our fill of love until the morning: let us solace ourselves with loves.
\vs Pro 7:19 For the goodman \bibemph{is} not at home, he is gone a long journey:
\vs Pro 7:20 He hath taken a bag of money with him, \bibemph{and} will come home at the day appointed.
\fn{with...: Heb. in his hand. the day...: or, the new moon.}
\vs Pro 7:21 With her much fair speech she caused him to yield, with the flattering of her lips she forced him.
\vs Pro 7:22 He goeth after her straightway, as an ox goeth to the slaughter, or as a fool to the correction of the stocks;
\fn{straightway: Heb. suddenly.}
\vs Pro 7:23 Till a dart strike through his liver; as a bird hasteth to the snare, and knoweth not that it \bibemph{is} for his life.
\vs Pro 7:24 Hearken unto me now therefore, O ye children, and attend to the words of my mouth.
\vs Pro 7:25 Let not thine heart decline to her ways, go not astray in her paths.
\vs Pro 7:26 For she hath cast down many wounded: yea, many strong \bibemph{men} have been slain by her.
\vs Pro 7:27 Her house \bibemph{is} the way to hell, going down to the chambers of death.
\vs Pro 8:1 Doth not wisdom cry? and understanding put forth her voice?
\vs Pro 8:2 She standeth in the top of high places, by the way in the places of the paths.
\vs Pro 8:3 She crieth at the gates, at the entry of the city, at the coming in at the doors.
\vs Pro 8:4 Unto you, O men, I call; and my voice \bibemph{is} to the sons of man.
\vs Pro 8:5 O ye simple, understand wisdom: and, ye fools, be ye of an understanding heart.
\vs Pro 8:6 Hear; for I will speak of excellent things; and the opening of my lips \bibemph{shall be} right things.
\vs Pro 8:7 For my mouth shall speak truth; and wickedness \bibemph{is} an abomination to my lips.
\fn{an...: Heb. the abomination of my lips.}
\vs Pro 8:8 All the words of my mouth \bibemph{are} in righteousness; \bibemph{there is} nothing froward or perverse in them.
\fn{froward: Heb. wreathed.}
\vs Pro 8:9 They \bibemph{are} all plain to him that understandeth, and right to them that find knowledge.
\vs Pro 8:10 Receive my instruction, and not silver; and knowledge rather than choice gold.
\vs Pro 8:11 For wisdom \bibemph{is} better than rubies; and all the things that may be desired are not to be compared to it.
\vs Pro 8:12 I wisdom dwell with prudence, and find out knowledge of witty inventions.
\fn{prudence: or, subtilty.}
\vs Pro 8:13 The fear of the \textsc{Lord} \bibemph{is} to hate evil: pride, and arrogancy, and the evil way, and the froward mouth, do I hate.
\vs Pro 8:14 Counsel \bibemph{is} mine, and sound wisdom: I \bibemph{am} understanding; I have strength.
\vs Pro 8:15 By me kings reign, and princes decree justice.
\vs Pro 8:16 By me princes rule, and nobles, \bibemph{even} all the judges of the earth.
\vs Pro 8:17 I love them that love me; and those that seek me early shall find me.
\vs Pro 8:18 Riches and honour \bibemph{are} with me; \bibemph{yea}, durable riches and righteousness.
\vs Pro 8:19 My fruit \bibemph{is} better than gold, yea, than fine gold; and my revenue than choice silver.
\vs Pro 8:20 I lead in the way of righteousness, in the midst of the paths of judgment:
\fn{lead: or, walk.}
\vs Pro 8:21 That I may cause those that love me to inherit substance; and I will fill their treasures.
\vs Pro 8:22 The \textsc{Lord} possessed me in the beginning of his way, before his works of old.
\vs Pro 8:23 I was set up from everlasting, from the beginning, or ever the earth was.
\vs Pro 8:24 When \bibemph{there were} no depths, I was brought forth; when \bibemph{there were} no fountains abounding with water.
\vs Pro 8:25 Before the mountains were settled, before the hills was I brought forth:
\vs Pro 8:26 While as yet he had not made the earth, nor the fields, nor the highest part of the dust of the world.
\fn{fields: or, open places. the highest...: or, the chief part.}
\vs Pro 8:27 When he prepared the heavens, I \bibemph{was} there: when he set a compass upon the face of the depth:
\fn{a compass: or, a circle.}
\vs Pro 8:28 When he established the clouds above: when he strengthened the fountains of the deep:
\vs Pro 8:29 When he gave to the sea his decree, that the waters should not pass his commandment: when he appointed the foundations of the earth:
\vs Pro 8:30 Then I was by him, \bibemph{as} one brought up \bibemph{with him}: and I was daily \bibemph{his} delight, rejoicing always before him;
\vs Pro 8:31 Rejoicing in the habitable part of his earth; and my delights \bibemph{were} with the sons of men.
\vs Pro 8:32 Now therefore hearken unto me, O ye children: for blessed \bibemph{are they that} keep my ways.
\vs Pro 8:33 Hear instruction, and be wise, and refuse it not.
\vs Pro 8:34 Blessed \bibemph{is} the man that heareth me, watching daily at my gates, waiting at the posts of my doors.
\vs Pro 8:35 For whoso findeth me findeth life, and shall obtain favour of the \textsc{Lord}.
\fn{obtain: Heb. bring forth.}
\vs Pro 8:36 But he that sinneth against me wrongeth his own soul: all they that hate me love death.
\vs Pro 9:1 Wisdom hath builded her house, she hath hewn out her seven pillars:
\vs Pro 9:2 She hath killed her beasts; she hath mingled her wine; she hath also furnished her table.
\fn{her beasts: Heb. her killing.}
\vs Pro 9:3 She hath sent forth her maidens: she crieth upon the highest places of the city,
\vs Pro 9:4 Whoso \bibemph{is} simple, let him turn in hither: \bibemph{as for} him that wanteth understanding, she saith to him,
\vs Pro 9:5 Come, eat of my bread, and drink of the wine \bibemph{which} I have mingled.
\vs Pro 9:6 Forsake the foolish, and live; and go in the way of understanding.
\vs Pro 9:7 He that reproveth a scorner getteth to himself shame: and he that rebuketh a wicked \bibemph{man getteth} himself a blot.
\vs Pro 9:8 Reprove not a scorner, lest he hate thee: rebuke a wise man, and he will love thee.
\vs Pro 9:9 Give \bibemph{instruction} to a wise \bibemph{man}, and he will be yet wiser: teach a just \bibemph{man}, and he will increase in learning.
\vs Pro 9:10 The fear of the \textsc{Lord} \bibemph{is} the beginning of wisdom: and the knowledge of the holy \bibemph{is} understanding.
\vs Pro 9:11 For by me thy days shall be multiplied, and the years of thy life shall be increased.
\vs Pro 9:12 If thou be wise, thou shalt be wise for thyself: but \bibemph{if} thou scornest, thou alone shalt bear \bibemph{it}.
\vs Pro 9:13 A foolish woman \bibemph{is} clamorous: \bibemph{she is} simple, and knoweth nothing.
\vs Pro 9:14 For she sitteth at the door of her house, on a seat in the high places of the city,
\vs Pro 9:15 To call passengers who go right on their ways:
\vs Pro 9:16 Whoso \bibemph{is} simple, let him turn in hither: and \bibemph{as for} him that wanteth understanding, she saith to him,
\vs Pro 9:17 Stolen waters are sweet, and bread \bibemph{eaten} in secret is pleasant.
\fn{eaten...: Heb. of secrecies.}
\vs Pro 9:18 But he knoweth not that the dead \bibemph{are} there; \bibemph{and that} her guests \bibemph{are} in the depths of hell.
\vs Pro 10:1 The proverbs of Solomon. A wise son maketh a glad father: but a foolish son \bibemph{is} the heaviness of his mother.
\vs Pro 10:2 Treasures of wickedness profit nothing: but righteousness delivereth from death.
\vs Pro 10:3 The \textsc{Lord} will not suffer the soul of the righteous to famish: but he casteth away the substance of the wicked.
\fn{the substance...: or, the wicked for their wickedness.}
\vs Pro 10:4 He becometh poor that dealeth \bibemph{with} a slack hand: but the hand of the diligent maketh rich.
\vs Pro 10:5 He that gathereth in summer \bibemph{is} a wise son: \bibemph{but} he that sleepeth in harvest \bibemph{is} a son that causeth shame.
\vs Pro 10:6 Blessings \bibemph{are} upon the head of the just: but violence covereth the mouth of the wicked.
\vs Pro 10:7 The memory of the just \bibemph{is} blessed: but the name of the wicked shall rot.
\vs Pro 10:8 The wise in heart will receive commandments: but a prating fool shall fall.
\fn{a prating...: Heb. a fool of lips. fall: Heb. be beaten.}
\vs Pro 10:9 He that walketh uprightly walketh surely: but he that perverteth his ways shall be known.
\vs Pro 10:10 He that winketh with the eye causeth sorrow: but a prating fool shall fall.
\fn{fall: Heb. be beaten.}
\vs Pro 10:11 The mouth of a righteous \bibemph{man is} a well of life: but violence covereth the mouth of the wicked.
\vs Pro 10:12 Hatred stirreth up strifes: but love covereth all sins.
\vs Pro 10:13 In the lips of him that hath understanding wisdom is found: but a rod \bibemph{is} for the back of him that is void of understanding.
\fn{of understanding: Heb. of heart.}
\vs Pro 10:14 Wise \bibemph{men} lay up knowledge: but the mouth of the foolish \bibemph{is} near destruction.
\vs Pro 10:15 The rich man's wealth \bibemph{is} his strong city: the destruction of the poor \bibemph{is} their poverty.
\vs Pro 10:16 The labour of the righteous \bibemph{tendeth} to life: the fruit of the wicked to sin.
\vs Pro 10:17 He \bibemph{is in} the way of life that keepeth instruction: but he that refuseth reproof erreth.
\fn{erreth: or, causeth to err.}
\vs Pro 10:18 He that hideth hatred \bibemph{with} lying lips, and he that uttereth a slander, \bibemph{is} a fool.
\vs Pro 10:19 In the multitude of words there wanteth not sin: but he that refraineth his lips \bibemph{is} wise.
\vs Pro 10:20 The tongue of the just \bibemph{is as} choice silver: the heart of the wicked \bibemph{is} little worth.
\vs Pro 10:21 The lips of the righteous feed many: but fools die for want of wisdom.
\fn{of wisdom: Heb. of heart.}
\vs Pro 10:22 The blessing of the \textsc{Lord}, it maketh rich, and he addeth no sorrow with it.
\vs Pro 10:23 \bibemph{It is} as sport to a fool to do mischief: but a man of understanding hath wisdom.
\vs Pro 10:24 The fear of the wicked, it shall come upon him: but the desire of the righteous shall be granted.
\vs Pro 10:25 As the whirlwind passeth, so \bibemph{is} the wicked no \bibemph{more}: but the righteous \bibemph{is} an everlasting foundation.
\vs Pro 10:26 As vinegar to the teeth, and as smoke to the eyes, so \bibemph{is} the sluggard to them that send him.
\vs Pro 10:27 The fear of the \textsc{Lord} prolongeth days: but the years of the wicked shall be shortened.
\fn{prolongeth: Heb. addeth.}
\vs Pro 10:28 The hope of the righteous \bibemph{shall be} gladness: but the expectation of the wicked shall perish.
\vs Pro 10:29 The way of the \textsc{Lord} \bibemph{is} strength to the upright: but destruction \bibemph{shall be} to the workers of iniquity.
\vs Pro 10:30 The righteous shall never be removed: but the wicked shall not inhabit the earth.
\vs Pro 10:31 The mouth of the just bringeth forth wisdom: but the froward tongue shall be cut out.
\vs Pro 10:32 The lips of the righteous know what is acceptable: but the mouth of the wicked \bibemph{speaketh} frowardness.
\fn{frowardness: Heb. frowardnesses.}
\vs Pro 11:1 A false balance \bibemph{is} abomination to the \textsc{Lord}: but a just weight \bibemph{is} his delight.
\fn{A false...: Heb. Balances of deceit. just...: Heb. perfect stone.}
\vs Pro 11:2 \bibemph{When} pride cometh, then cometh shame: but with the lowly \bibemph{is} wisdom.
\vs Pro 11:3 The integrity of the upright shall guide them: but the perverseness of transgressors shall destroy them.
\vs Pro 11:4 Riches profit not in the day of wrath: but righteousness delivereth from death.
\vs Pro 11:5 The righteousness of the perfect shall direct his way: but the wicked shall fall by his own wickedness.
\fn{direct: Heb. rectify.}
\vs Pro 11:6 The righteousness of the upright shall deliver them: but transgressors shall be taken in \bibemph{their own} naughtiness.
\vs Pro 11:7 When a wicked man dieth, \bibemph{his} expectation shall perish: and the hope of unjust \bibemph{men} perisheth.
\vs Pro 11:8 The righteous is delivered out of trouble, and the wicked cometh in his stead.
\vs Pro 11:9 An hypocrite with \bibemph{his} mouth destroyeth his neighbour: but through knowledge shall the just be delivered.
\vs Pro 11:10 When it goeth well with the righteous, the city rejoiceth: and when the wicked perish, \bibemph{there is} shouting.
\vs Pro 11:11 By the blessing of the upright the city is exalted: but it is overthrown by the mouth of the wicked.
\vs Pro 11:12 He that is void of wisdom despiseth his neighbour: but a man of understanding holdeth his peace.
\fn{void...: Heb. destitute of heart.}
\vs Pro 11:13 A talebearer revealeth secrets: but he that is of a faithful spirit concealeth the matter.
\fn{A talebearer: Heb. He that walketh, being a talebearer.}
\vs Pro 11:14 Where no counsel \bibemph{is}, the people fall: but in the multitude of counsellors \bibemph{there is} safety.
\vs Pro 11:15 He that is surety for a stranger shall smart \bibemph{for it}: and he that hateth suretiship is sure.
\fn{smart: Heb. be sore broken. suretiship: Heb. those that strike hands.}
\vs Pro 11:16 A gracious woman retaineth honour: and strong \bibemph{men} retain riches.
\vs Pro 11:17 The merciful man doeth good to his own soul: but \bibemph{he that is} cruel troubleth his own flesh.
\vs Pro 11:18 The wicked worketh a deceitful work: but to him that soweth righteousness \bibemph{shall be} a sure reward.
\vs Pro 11:19 As righteousness \bibemph{tendeth} to life: so he that pursueth evil \bibemph{pursueth it} to his own death.
\vs Pro 11:20 They that are of a froward heart \bibemph{are} abomination to the \textsc{Lord}: but \bibemph{such as are} upright in \bibemph{their} way \bibemph{are} his delight.
\vs Pro 11:21 \bibemph{Though} hand \bibemph{join} in hand, the wicked shall not be unpunished: but the seed of the righteous shall be delivered.
\vs Pro 11:22 \bibemph{As} a jewel of gold in a swine's snout, \bibemph{so is} a fair woman which is without discretion.
\fn{is without: Heb. departeth from.}
\vs Pro 11:23 The desire of the righteous \bibemph{is} only good: \bibemph{but} the expectation of the wicked \bibemph{is} wrath.
\vs Pro 11:24 There is that scattereth, and yet increaseth; and \bibemph{there is} that withholdeth more than is meet, but \bibemph{it tendeth} to poverty.
\vs Pro 11:25 The liberal soul shall be made fat: and he that watereth shall be watered also himself.
\fn{liberal...: Heb. soul of blessing.}
\vs Pro 11:26 He that withholdeth corn, the people shall curse him: but blessing \bibemph{shall be} upon the head of him that selleth \bibemph{it}.
\vs Pro 11:27 He that diligently seeketh good procureth favour: but he that seeketh mischief, it shall come unto him.
\vs Pro 11:28 He that trusteth in his riches shall fall: but the righteous shall flourish as a branch.
\vs Pro 11:29 He that troubleth his own house shall inherit the wind: and the fool \bibemph{shall be} servant to the wise of heart.
\vs Pro 11:30 The fruit of the righteous \bibemph{is} a tree of life; and he that winneth souls \bibemph{is} wise.
\fn{winneth: Heb. taketh.}
\vs Pro 11:31 Behold, the righteous shall be recompensed in the earth: much more the wicked and the sinner.
\vs Pro 12:1 Whoso loveth instruction loveth knowledge: but he that hateth reproof \bibemph{is} brutish.
\vs Pro 12:2 A good \bibemph{man} obtaineth favour of the \textsc{Lord}: but a man of wicked devices will he condemn.
\vs Pro 12:3 A man shall not be established by wickedness: but the root of the righteous shall not be moved.
\vs Pro 12:4 A virtuous woman \bibemph{is} a crown to her husband: but she that maketh ashamed \bibemph{is} as rottenness in his bones.
\vs Pro 12:5 The thoughts of the righteous \bibemph{are} right: \bibemph{but} the counsels of the wicked \bibemph{are} deceit.
\vs Pro 12:6 The words of the wicked \bibemph{are} to lie in wait for blood: but the mouth of the upright shall deliver them.
\vs Pro 12:7 The wicked are overthrown, and \bibemph{are} not: but the house of the righteous shall stand.
\vs Pro 12:8 A man shall be commended according to his wisdom: but he that is of a perverse heart shall be despised.
\fn{of a...: Heb. perverse of heart.}
\vs Pro 12:9 \bibemph{He that is} despised, and hath a servant, \bibemph{is} better than he that honoureth himself, and lacketh bread.
\vs Pro 12:10 A righteous \bibemph{man} regardeth the life of his beast: but the tender mercies of the wicked \bibemph{are} cruel.
\fn{tender...: or, bowels.}
\vs Pro 12:11 He that tilleth his land shall be satisfied with bread: but he that followeth vain \bibemph{persons is} void of understanding.
\vs Pro 12:12 The wicked desireth the net of evil \bibemph{men}: but the root of the righteous yieldeth \bibemph{fruit}.
\fn{the net: or, the fortress.}
\vs Pro 12:13 The wicked is snared by the transgression of \bibemph{his} lips: but the just shall come out of trouble.
\fn{The wicked...: Heb. The snare of the wicked is in the transgression of lips.}
\vs Pro 12:14 A man shall be satisfied with good by the fruit of \bibemph{his} mouth: and the recompence of a man's hands shall be rendered unto him.
\vs Pro 12:15 The way of a fool \bibemph{is} right in his own eyes: but he that hearkeneth unto counsel \bibemph{is} wise.
\vs Pro 12:16 A fool's wrath is presently known: but a prudent \bibemph{man} covereth shame.
\fn{presently: Heb. in that day.}
\vs Pro 12:17 \bibemph{He that} speaketh truth sheweth forth righteousness: but a false witness deceit.
\vs Pro 12:18 There is that speaketh like the piercings of a sword: but the tongue of the wise \bibemph{is} health.
\vs Pro 12:19 The lip of truth shall be established for ever: but a lying tongue \bibemph{is} but for a moment.
\vs Pro 12:20 Deceit \bibemph{is} in the heart of them that imagine evil: but to the counsellors of peace \bibemph{is} joy.
\vs Pro 12:21 There shall no evil happen to the just: but the wicked shall be filled with mischief.
\vs Pro 12:22 Lying lips \bibemph{are} abomination to the \textsc{Lord}: but they that deal truly \bibemph{are} his delight.
\vs Pro 12:23 A prudent man concealeth knowledge: but the heart of fools proclaimeth foolishness.
\vs Pro 12:24 The hand of the diligent shall bear rule: but the slothful shall be under tribute.
\fn{slothful: or, deceitful.}
\vs Pro 12:25 Heaviness in the heart of man maketh it stoop: but a good word maketh it glad.
\vs Pro 12:26 The righteous \bibemph{is} more excellent than his neighbour: but the way of the wicked seduceth them.
\fn{excellent: or, abundant.}
\vs Pro 12:27 The slothful \bibemph{man} roasteth not that which he took in hunting: but the substance of a diligent man \bibemph{is} precious.
\vs Pro 12:28 In the way of righteousness \bibemph{is} life; and \bibemph{in} the pathway \bibemph{thereof there is} no death.
\vs Pro 13:1 A wise son \bibemph{heareth} his father's instruction: but a scorner heareth not rebuke.
\vs Pro 13:2 A man shall eat good by the fruit of \bibemph{his} mouth: but the soul of the transgressors \bibemph{shall eat} violence.
\vs Pro 13:3 He that keepeth his mouth keepeth his life: \bibemph{but} he that openeth wide his lips shall have destruction.
\vs Pro 13:4 The soul of the sluggard desireth, and \bibemph{hath} nothing: but the soul of the diligent shall be made fat.
\vs Pro 13:5 A righteous \bibemph{man} hateth lying: but a wicked \bibemph{man} is loathsome, and cometh to shame.
\vs Pro 13:6 Righteousness keepeth \bibemph{him that is} upright in the way: but wickedness overthroweth the sinner.
\fn{the sinner: Heb. sin.}
\vs Pro 13:7 There is that maketh himself rich, yet \bibemph{hath} nothing: \bibemph{there is} that maketh himself poor, yet \bibemph{hath} great riches.
\vs Pro 13:8 The ransom of a man's life \bibemph{are} his riches: but the poor heareth not rebuke.
\vs Pro 13:9 The light of the righteous rejoiceth: but the lamp of the wicked shall be put out.
\fn{lamp: or, candle.}
\vs Pro 13:10 Only by pride cometh contention: but with the well advised \bibemph{is} wisdom.
\vs Pro 13:11 Wealth \bibemph{gotten} by vanity shall be diminished: but he that gathereth by labour shall increase.
\fn{by labour: Heb. with the hand.}
\vs Pro 13:12 Hope deferred maketh the heart sick: but \bibemph{when} the desire cometh, \bibemph{it is} a tree of life.
\vs Pro 13:13 Whoso despiseth the word shall be destroyed: but he that feareth the commandment shall be rewarded.
\fn{shall be rewarded: or, shall be in peace.}
\vs Pro 13:14 The law of the wise \bibemph{is} a fountain of life, to depart from the snares of death.
\vs Pro 13:15 Good understanding giveth favour: but the way of transgressors \bibemph{is} hard.
\vs Pro 13:16 Every prudent \bibemph{man} dealeth with knowledge: but a fool layeth open \bibemph{his} folly.
\fn{layeth...: Heb. spreadeth.}
\vs Pro 13:17 A wicked messenger falleth into mischief: but a faithful ambassador \bibemph{is} health.
\fn{a faithful...: Heb. an ambassador of faithfulness.}
\vs Pro 13:18 Poverty and shame \bibemph{shall be to} him that refuseth instruction: but he that regardeth reproof shall be honoured.
\vs Pro 13:19 The desire accomplished is sweet to the soul: but \bibemph{it is} abomination to fools to depart from evil.
\vs Pro 13:20 He that walketh with wise \bibemph{men} shall be wise: but a companion of fools shall be destroyed.
\fn{destroyed: Heb. broken.}
\vs Pro 13:21 Evil pursueth sinners: but to the righteous good shall be repayed.
\vs Pro 13:22 A good \bibemph{man} leaveth an inheritance to his children's children: and the wealth of the sinner \bibemph{is} laid up for the just.
\vs Pro 13:23 Much food \bibemph{is in} the tillage of the poor: but there is \bibemph{that is} destroyed for want of judgment.
\vs Pro 13:24 He that spareth his rod hateth his son: but he that loveth him chasteneth him betimes.
\vs Pro 13:25 The righteous eateth to the satisfying of his soul: but the belly of the wicked shall want.
\vs Pro 14:1 Every wise woman buildeth her house: but the foolish plucketh it down with her hands.
\vs Pro 14:2 He that walketh in his uprightness feareth the \textsc{Lord}: but \bibemph{he that is} perverse in his ways despiseth him.
\vs Pro 14:3 In the mouth of the foolish \bibemph{is} a rod of pride: but the lips of the wise shall preserve them.
\vs Pro 14:4 Where no oxen \bibemph{are}, the crib \bibemph{is} clean: but much increase \bibemph{is} by the strength of the ox.
\vs Pro 14:5 A faithful witness will not lie: but a false witness will utter lies.
\vs Pro 14:6 A scorner seeketh wisdom, and \bibemph{findeth it} not: but knowledge \bibemph{is} easy unto him that understandeth.
\vs Pro 14:7 Go from the presence of a foolish man, when thou perceivest not \bibemph{in him} the lips of knowledge.
\vs Pro 14:8 The wisdom of the prudent \bibemph{is} to understand his way: but the folly of fools \bibemph{is} deceit.
\vs Pro 14:9 Fools make a mock at sin: but among the righteous \bibemph{there is} favour.
\vs Pro 14:10 The heart knoweth his own bitterness; and a stranger doth not intermeddle with his joy.
\fn{his own...: Heb. the bitterness of his soul.}
\vs Pro 14:11 The house of the wicked shall be overthrown: but the tabernacle of the upright shall flourish.
\vs Pro 14:12 There is a way which seemeth right unto a man, but the end thereof \bibemph{are} the ways of death.
\vs Pro 14:13 Even in laughter the heart is sorrowful; and the end of that mirth \bibemph{is} heaviness.
\vs Pro 14:14 The backslider in heart shall be filled with his own ways: and a good man \bibemph{shall be satisfied} from himself.
\vs Pro 14:15 The simple believeth every word: but the prudent \bibemph{man} looketh well to his going.
\vs Pro 14:16 A wise \bibemph{man} feareth, and departeth from evil: but the fool rageth, and is confident.
\vs Pro 14:17 \bibemph{He that is} soon angry dealeth foolishly: and a man of wicked devices is hated.
\vs Pro 14:18 The simple inherit folly: but the prudent are crowned with knowledge.
\vs Pro 14:19 The evil bow before the good; and the wicked at the gates of the righteous.
\vs Pro 14:20 The poor is hated even of his own neighbour: but the rich \bibemph{hath} many friends.
\fn{the rich...: Heb. many are the lovers of the rich.}
\vs Pro 14:21 He that despiseth his neighbour sinneth: but he that hath mercy on the poor, happy \bibemph{is} he.
\vs Pro 14:22 Do they not err that devise evil? but mercy and truth \bibemph{shall be} to them that devise good.
\vs Pro 14:23 In all labour there is profit: but the talk of the lips \bibemph{tendeth} only to penury.
\vs Pro 14:24 The crown of the wise \bibemph{is} their riches: \bibemph{but} the foolishness of fools \bibemph{is} folly.
\vs Pro 14:25 A true witness delivereth souls: but a deceitful \bibemph{witness} speaketh lies.
\vs Pro 14:26 In the fear of the \textsc{Lord} \bibemph{is} strong confidence: and his children shall have a place of refuge.
\vs Pro 14:27 The fear of the \textsc{Lord} \bibemph{is} a fountain of life, to depart from the snares of death.
\vs Pro 14:28 In the multitude of people \bibemph{is} the king's honour: but in the want of people \bibemph{is} the destruction of the prince.
\vs Pro 14:29 \bibemph{He that is} slow to wrath \bibemph{is} of great understanding: but \bibemph{he that is} hasty of spirit exalteth folly.
\fn{hasty...: Heb. short of spirit.}
\vs Pro 14:30 A sound heart \bibemph{is} the life of the flesh: but envy the rottenness of the bones.
\vs Pro 14:31 He that oppresseth the poor reproacheth his Maker: but he that honoureth him hath mercy on the poor.
\vs Pro 14:32 The wicked is driven away in his wickedness: but the righteous hath hope in his death.
\vs Pro 14:33 Wisdom resteth in the heart of him that hath understanding: but \bibemph{that which is} in the midst of fools is made known.
\vs Pro 14:34 Righteousness exalteth a nation: but sin \bibemph{is} a reproach to any people.
\fn{to any...: Heb. to nations.}
\vs Pro 14:35 The king's favour \bibemph{is} toward a wise servant: but his wrath is \bibemph{against} him that causeth shame.
\vs Pro 15:1 A soft answer turneth away wrath: but grievous words stir up anger.
\vs Pro 15:2 The tongue of the wise useth knowledge aright: but the mouth of fools poureth out foolishness.
\fn{poureth: Heb. belcheth, or, bubbleth.}
\vs Pro 15:3 The eyes of the \textsc{Lord} \bibemph{are} in every place, beholding the evil and the good.
\vs Pro 15:4 A wholesome tongue \bibemph{is} a tree of life: but perverseness therein \bibemph{is} a breach in the spirit.
\fn{A wholesome...: Heb. The healing of the tongue.}
\vs Pro 15:5 A fool despiseth his father's instruction: but he that regardeth reproof is prudent.
\vs Pro 15:6 In the house of the righteous \bibemph{is} much treasure: but in the revenues of the wicked is trouble.
\vs Pro 15:7 The lips of the wise disperse knowledge: but the heart of the foolish \bibemph{doeth} not so.
\vs Pro 15:8 The sacrifice of the wicked \bibemph{is} an abomination to the \textsc{Lord}: but the prayer of the upright \bibemph{is} his delight.
\vs Pro 15:9 The way of the wicked \bibemph{is} an abomination unto the \textsc{Lord}: but he loveth him that followeth after righteousness.
\vs Pro 15:10 Correction \bibemph{is} grievous unto him that forsaketh the way: \bibemph{and} he that hateth reproof shall die.
\fn{Correction: or, Instruction.}
\vs Pro 15:11 Hell and destruction \bibemph{are} before the \textsc{Lord}: how much more then the hearts of the children of men?
\vs Pro 15:12 A scorner loveth not one that reproveth him: neither will he go unto the wise.
\vs Pro 15:13 A merry heart maketh a cheerful countenance: but by sorrow of the heart the spirit is broken.
\vs Pro 15:14 The heart of him that hath understanding seeketh knowledge: but the mouth of fools feedeth on foolishness.
\vs Pro 15:15 All the days of the afflicted \bibemph{are} evil: but he that is of a merry heart \bibemph{hath} a continual feast.
\vs Pro 15:16 Better \bibemph{is} little with the fear of the \textsc{Lord} than great treasure and trouble therewith.
\vs Pro 15:17 Better \bibemph{is} a dinner of herbs where love is, than a stalled ox and hatred therewith.
\vs Pro 15:18 A wrathful man stirreth up strife: but \bibemph{he that is} slow to anger appeaseth strife.
\vs Pro 15:19 The way of the slothful \bibemph{man is} as an hedge of thorns: but the way of the righteous \bibemph{is} made plain.
\fn{is made...: Heb. is raised up as a causey.}
\vs Pro 15:20 A wise son maketh a glad father: but a foolish man despiseth his mother.
\vs Pro 15:21 Folly \bibemph{is} joy to \bibemph{him that is} destitute of wisdom: but a man of understanding walketh uprightly.
\fn{destitute...: Heb. void of heart.}
\vs Pro 15:22 Without counsel purposes are disappointed: but in the multitude of counsellors they are established.
\vs Pro 15:23 A man hath joy by the answer of his mouth: and a word \bibemph{spoken} in due season, how good \bibemph{is it}!
\fn{due...: Heb. his season.}
\vs Pro 15:24 The way of life \bibemph{is} above to the wise, that he may depart from hell beneath.
\vs Pro 15:25 The \textsc{Lord} will destroy the house of the proud: but he will establish the border of the widow.
\vs Pro 15:26 The thoughts of the wicked \bibemph{are} an abomination to the \textsc{Lord}: but \bibemph{the words} of the pure \bibemph{are} pleasant words.
\fn{pleasant...: Heb. words of pleasantness.}
\vs Pro 15:27 He that is greedy of gain troubleth his own house; but he that hateth gifts shall live.
\vs Pro 15:28 The heart of the righteous studieth to answer: but the mouth of the wicked poureth out evil things.
\vs Pro 15:29 The \textsc{Lord} \bibemph{is} far from the wicked: but he heareth the prayer of the righteous.
\vs Pro 15:30 The light of the eyes rejoiceth the heart: \bibemph{and} a good report maketh the bones fat.
\vs Pro 15:31 The ear that heareth the reproof of life abideth among the wise.
\vs Pro 15:32 He that refuseth instruction despiseth his own soul: but he that heareth reproof getteth understanding.
\fn{instruction: or, correction. heareth: or, obeyeth. getteth...: Heb. possesseth an heart.}
\vs Pro 15:33 The fear of the \textsc{Lord} \bibemph{is} the instruction of wisdom; and before honour \bibemph{is} humility.
\vs Pro 16:1 The preparations of the heart in man, and the answer of the tongue, \bibemph{is} from the \textsc{Lord}.
\fn{preparations: or, disposings.}
\vs Pro 16:2 All the ways of a man \bibemph{are} clean in his own eyes; but the \textsc{Lord} weigheth the spirits.
\vs Pro 16:3 Commit thy works unto the \textsc{Lord}, and thy thoughts shall be established.
\fn{Commit: Heb. Roll.}
\vs Pro 16:4 The \textsc{Lord} hath made all \bibemph{things} for himself: yea, even the wicked for the day of evil.
\vs Pro 16:5 Every one \bibemph{that is} proud in heart \bibemph{is} an abomination to the \textsc{Lord}: \bibemph{though} hand \bibemph{join} in hand, he shall not be unpunished.
\fn{unpunished: Heb. held innocent.}
\vs Pro 16:6 By mercy and truth iniquity is purged: and by the fear of the \textsc{Lord} \bibemph{men} depart from evil.
\vs Pro 16:7 When a man's ways please the \textsc{Lord}, he maketh even his enemies to be at peace with him.
\vs Pro 16:8 Better \bibemph{is} a little with righteousness than great revenues without right.
\vs Pro 16:9 A man's heart deviseth his way: but the \textsc{Lord} directeth his steps.
\vs Pro 16:10 A divine sentence \bibemph{is} in the lips of the king: his mouth transgresseth not in judgment.
\fn{A divine...: Heb. Divination.}
\vs Pro 16:11 A just weight and balance \bibemph{are} the \textsc{Lord's}: all the weights of the bag \bibemph{are} his work.
\fn{the weights: Heb. the stones.}
\vs Pro 16:12 \bibemph{It is} an abomination to kings to commit wickedness: for the throne is established by righteousness.
\vs Pro 16:13 Righteous lips \bibemph{are} the delight of kings; and they love him that speaketh right.
\vs Pro 16:14 The wrath of a king \bibemph{is as} messengers of death: but a wise man will pacify it.
\vs Pro 16:15 In the light of the king's countenance \bibemph{is} life; and his favour \bibemph{is} as a cloud of the latter rain.
\vs Pro 16:16 How much better \bibemph{is it} to get wisdom than gold! and to get understanding rather to be chosen than silver!
\vs Pro 16:17 The highway of the upright \bibemph{is} to depart from evil: he that keepeth his way preserveth his soul.
\vs Pro 16:18 Pride \bibemph{goeth} before destruction, and an haughty spirit before a fall.
\vs Pro 16:19 Better \bibemph{it is to be} of an humble spirit with the lowly, than to divide the spoil with the proud.
\vs Pro 16:20 He that handleth a matter wisely shall find good: and whoso trusteth in the \textsc{Lord}, happy \bibemph{is} he.
\fn{handleth...: or, understandeth a matter.}
\vs Pro 16:21 The wise in heart shall be called prudent: and the sweetness of the lips increaseth learning.
\vs Pro 16:22 Understanding \bibemph{is} a wellspring of life unto him that hath it: but the instruction of fools \bibemph{is} folly.
\vs Pro 16:23 The heart of the wise teacheth his mouth, and addeth learning to his lips.
\fn{teacheth: Heb. maketh wise.}
\vs Pro 16:24 Pleasant words \bibemph{are as} an honeycomb, sweet to the soul, and health to the bones.
\vs Pro 16:25 There is a way that seemeth right unto a man, but the end thereof \bibemph{are} the ways of death.
\vs Pro 16:26 He that laboureth laboureth for himself; for his mouth craveth it of him.
\fn{He...: Heb. The soul of him that. craveth...: Heb. boweth unto him.}
\vs Pro 16:27 An ungodly man diggeth up evil: and in his lips \bibemph{there is} as a burning fire.
\fn{An...: Heb. A man of Belial.}
\vs Pro 16:28 A froward man soweth strife: and a whisperer separateth chief friends.
\fn{soweth: Heb. sendeth forth.}
\vs Pro 16:29 A violent man enticeth his neighbour, and leadeth him into the way \bibemph{that is} not good.
\vs Pro 16:30 He shutteth his eyes to devise froward things: moving his lips he bringeth evil to pass.
\vs Pro 16:31 The hoary head \bibemph{is} a crown of glory, \bibemph{if} it be found in the way of righteousness.
\vs Pro 16:32 \bibemph{He that is} slow to anger \bibemph{is} better than the mighty; and he that ruleth his spirit than he that taketh a city.
\vs Pro 16:33 The lot is cast into the lap; but the whole disposing thereof \bibemph{is} of the \textsc{Lord}.
\vs Pro 17:1 Better \bibemph{is} a dry morsel, and quietness therewith, than an house full of sacrifices \bibemph{with} strife.
\fn{sacrifices: or, good cheer.}
\vs Pro 17:2 A wise servant shall have rule over a son that causeth shame, and shall have part of the inheritance among the brethren.
\vs Pro 17:3 The fining pot \bibemph{is} for silver, and the furnace for gold: but the \textsc{Lord} trieth the hearts.
\vs Pro 17:4 A wicked doer giveth heed to false lips; \bibemph{and} a liar giveth ear to a naughty tongue.
\vs Pro 17:5 Whoso mocketh the poor reproacheth his Maker: \bibemph{and} he that is glad at calamities shall not be unpunished.
\fn{unpunished: Heb. held innocent.}
\vs Pro 17:6 Children's children \bibemph{are} the crown of old men; and the glory of children \bibemph{are} their fathers.
\vs Pro 17:7 Excellent speech becometh not a fool: much less do lying lips a prince.
\fn{Excellent...: Heb. A lip of excellency. lying...: Heb. a lip of lying.}
\vs Pro 17:8 A gift \bibemph{is as} a precious stone in the eyes of him that hath it: whithersoever it turneth, it prospereth.
\fn{a precious...: Heb. a stone of grace.}
\vs Pro 17:9 He that covereth a transgression seeketh love; but he that repeateth a matter separateth \bibemph{very} friends.
\fn{seeketh: or, procureth.}
\vs Pro 17:10 A reproof entereth more into a wise man than an hundred stripes into a fool.
\fn{entereth...: or, aweth more a wise man, than to strike a fool an hundred times.}
\vs Pro 17:11 An evil \bibemph{man} seeketh only rebellion: therefore a cruel messenger shall be sent against him.
\vs Pro 17:12 Let a bear robbed of her whelps meet a man, rather than a fool in his folly.
\vs Pro 17:13 Whoso rewardeth evil for good, evil shall not depart from his house.
\vs Pro 17:14 The beginning of strife \bibemph{is as} when one letteth out water: therefore leave off contention, before it be meddled with.
\vs Pro 17:15 He that justifieth the wicked, and he that condemneth the just, even they both \bibemph{are} abomination to the \textsc{Lord}.
\vs Pro 17:16 Wherefore \bibemph{is there} a price in the hand of a fool to get wisdom, seeing \bibemph{he hath} no heart \bibemph{to it}?
\vs Pro 17:17 A friend loveth at all times, and a brother is born for adversity.
\vs Pro 17:18 A man void of understanding striketh hands, \bibemph{and} becometh surety in the presence of his friend.
\fn{understanding: Heb. heart.}
\vs Pro 17:19 He loveth transgression that loveth strife: \bibemph{and} he that exalteth his gate seeketh destruction.
\vs Pro 17:20 He that hath a froward heart findeth no good: and he that hath a perverse tongue falleth into mischief.
\fn{He that hath a froward...: Heb. The froward of heart.}
\vs Pro 17:21 He that begetteth a fool \bibemph{doeth it} to his sorrow: and the father of a fool hath no joy.
\vs Pro 17:22 A merry heart doeth good \bibemph{like} a medicine: but a broken spirit drieth the bones.
\fn{like: or, to.}
\vs Pro 17:23 A wicked \bibemph{man} taketh a gift out of the bosom to pervert the ways of judgment.
\vs Pro 17:24 Wisdom \bibemph{is} before him that hath understanding; but the eyes of a fool \bibemph{are} in the ends of the earth.
\vs Pro 17:25 A foolish son \bibemph{is} a grief to his father, and bitterness to her that bare him.
\vs Pro 17:26 Also to punish the just \bibemph{is} not good, \bibemph{nor} to strike princes for equity.
\vs Pro 17:27 He that hath knowledge spareth his words: \bibemph{and} a man of understanding is of an excellent spirit.
\fn{an...: or, a cool.}
\vs Pro 17:28 Even a fool, when he holdeth his peace, is counted wise: \bibemph{and} he that shutteth his lips \bibemph{is esteemed} a man of understanding.
\vs Pro 18:1 Through desire a man, having separated himself, seeketh \bibemph{and} intermeddleth with all wisdom.
\fn{Through...: or, He that separateth himself seeketh according to his desire, and intermeddleth in every business.}
\vs Pro 18:2 A fool hath no delight in understanding, but that his heart may discover itself.
\vs Pro 18:3 When the wicked cometh, \bibemph{then} cometh also contempt, and with ignominy reproach.
\vs Pro 18:4 The words of a man's mouth \bibemph{are as} deep waters, \bibemph{and} the wellspring of wisdom \bibemph{as} a flowing brook.
\vs Pro 18:5 \bibemph{It is} not good to accept the person of the wicked, to overthrow the righteous in judgment.
\vs Pro 18:6 A fool's lips enter into contention, and his mouth calleth for strokes.
\vs Pro 18:7 A fool's mouth \bibemph{is} his destruction, and his lips \bibemph{are} the snare of his soul.
\vs Pro 18:8 The words of a talebearer \bibemph{are} as wounds, and they go down into the innermost parts of the belly.
\fn{talebearer: or, whisperer. most...: Heb. chambers. as wounds: or, like as when men are wounded.}
\vs Pro 18:9 He also that is slothful in his work is brother to him that is a great waster.
\vs Pro 18:10 The name of the \textsc{Lord} \bibemph{is} a strong tower: the righteous runneth into it, and is safe.
\fn{safe: Heb. set aloft.}
\vs Pro 18:11 The rich man's wealth \bibemph{is} his strong city, and as an high wall in his own conceit.
\vs Pro 18:12 Before destruction the heart of man is haughty, and before honour \bibemph{is} humility.
\vs Pro 18:13 He that answereth a matter before he heareth \bibemph{it}, it \bibemph{is} folly and shame unto him.
\fn{answereth...: Heb. returneth a word.}
\vs Pro 18:14 The spirit of a man will sustain his infirmity; but a wounded spirit who can bear?
\vs Pro 18:15 The heart of the prudent getteth knowledge; and the ear of the wise seeketh knowledge.
\vs Pro 18:16 A man's gift maketh room for him, and bringeth him before great men.
\vs Pro 18:17 \bibemph{He that is} first in his own cause \bibemph{seemeth} just; but his neighbour cometh and searcheth him.
\vs Pro 18:18 The lot causeth contentions to cease, and parteth between the mighty.
\vs Pro 18:19 A brother offended \bibemph{is harder to be won} than a strong city: and \bibemph{their} contentions \bibemph{are} like the bars of a castle.
\vs Pro 18:20 A man's belly shall be satisfied with the fruit of his mouth; \bibemph{and} with the increase of his lips shall he be filled.
\vs Pro 18:21 Death and life \bibemph{are} in the power of the tongue: and they that love it shall eat the fruit thereof.
\vs Pro 18:22 \bibemph{Whoso} findeth a wife findeth a good \bibemph{thing}, and obtaineth favour of the \textsc{Lord}.
\vs Pro 18:23 The poor useth intreaties; but the rich answereth roughly.
\vs Pro 18:24 A man \bibemph{that hath} friends must shew himself friendly: and there is a friend \bibemph{that} sticketh closer than a brother.
\vs Pro 19:1 Better \bibemph{is} the poor that walketh in his integrity, than \bibemph{he that is} perverse in his lips, and is a fool.
\vs Pro 19:2 Also, \bibemph{that} the soul \bibemph{be} without knowledge, \bibemph{it is} not good; and he that hasteth with \bibemph{his} feet sinneth.
\vs Pro 19:3 The foolishness of man perverteth his way: and his heart fretteth against the \textsc{Lord}.
\vs Pro 19:4 Wealth maketh many friends; but the poor is separated from his neighbour.
\vs Pro 19:5 A false witness shall not be unpunished, and \bibemph{he that} speaketh lies shall not escape.
\fn{unpunished: Heb. held innocent.}
\vs Pro 19:6 Many will intreat the favour of the prince: and every man \bibemph{is} a friend to him that giveth gifts.
\fn{him...: Heb. a man of gifts.}
\vs Pro 19:7 All the brethren of the poor do hate him: how much more do his friends go far from him? he pursueth \bibemph{them with} words, \bibemph{yet} they \bibemph{are} wanting \bibemph{to him}.
\vs Pro 19:8 He that getteth wisdom loveth his own soul: he that keepeth understanding shall find good.
\fn{wisdom: Heb. an heart.}
\vs Pro 19:9 A false witness shall not be unpunished, and \bibemph{he that} speaketh lies shall perish.
\vs Pro 19:10 Delight is not seemly for a fool; much less for a servant to have rule over princes.
\vs Pro 19:11 The discretion of a man deferreth his anger; and \bibemph{it is} his glory to pass over a transgression.
\fn{discretion: or, prudence.}
\vs Pro 19:12 The king's wrath \bibemph{is} as the roaring of a lion; but his favour \bibemph{is} as dew upon the grass.
\vs Pro 19:13 A foolish son \bibemph{is} the calamity of his father: and the contentions of a wife \bibemph{are} a continual dropping.
\vs Pro 19:14 House and riches \bibemph{are} the inheritance of fathers: and a prudent wife \bibemph{is} from the \textsc{Lord}.
\vs Pro 19:15 Slothfulness casteth into a deep sleep; and an idle soul shall suffer hunger.
\vs Pro 19:16 He that keepeth the commandment keepeth his own soul; \bibemph{but} he that despiseth his ways shall die.
\vs Pro 19:17 He that hath pity upon the poor lendeth unto the \textsc{Lord}; and that which he hath given will he pay him again.
\fn{that which...: or, his deed.}
\vs Pro 19:18 Chasten thy son while there is hope, and let not thy soul spare for his crying.
\fn{for...: or, to his destruction: or, to cause him to die.}
\vs Pro 19:19 A man of great wrath shall suffer punishment: for if thou deliver \bibemph{him}, yet thou must do it again.
\fn{do: Heb. add.}
\vs Pro 19:20 Hear counsel, and receive instruction, that thou mayest be wise in thy latter end.
\vs Pro 19:21 \bibemph{There are} many devices in a man's heart; nevertheless the counsel of the \textsc{Lord}, that shall stand.
\vs Pro 19:22 The desire of a man \bibemph{is} his kindness: and a poor man \bibemph{is} better than a liar.
\vs Pro 19:23 The fear of the \textsc{Lord} \bibemph{tendeth} to life: and \bibemph{he that hath it} shall abide satisfied; he shall not be visited with evil.
\vs Pro 19:24 A slothful \bibemph{man} hideth his hand in \bibemph{his} bosom, and will not so much as bring it to his mouth again.
\vs Pro 19:25 Smite a scorner, and the simple will beware: and reprove one that hath understanding, \bibemph{and} he will understand knowledge.
\fn{will beware: Heb. will be cunning.}
\vs Pro 19:26 He that wasteth \bibemph{his} father, \bibemph{and} chaseth away \bibemph{his} mother, \bibemph{is} a son that causeth shame, and bringeth reproach.
\vs Pro 19:27 Cease, my son, to hear the instruction \bibemph{that causeth} to err from the words of knowledge.
\vs Pro 19:28 An ungodly witness scorneth judgment: and the mouth of the wicked devoureth iniquity.
\fn{An...: Heb. A witness of Belial.}
\vs Pro 19:29 Judgments are prepared for scorners, and stripes for the back of fools.
\vs Pro 20:1 Wine \bibemph{is} a mocker, strong drink \bibemph{is} raging: and whosoever is deceived thereby is not wise.
\vs Pro 20:2 The fear of a king \bibemph{is} as the roaring of a lion: \bibemph{whoso} provoketh him to anger sinneth \bibemph{against} his own soul.
\vs Pro 20:3 \bibemph{It is} an honour for a man to cease from strife: but every fool will be meddling.
\vs Pro 20:4 The sluggard will not plow by reason of the cold; \bibemph{therefore} shall he beg in harvest, and \bibemph{have} nothing.
\fn{cold: or, winter.}
\vs Pro 20:5 Counsel in the heart of man \bibemph{is like} deep water; but a man of understanding will draw it out.
\vs Pro 20:6 Most men will proclaim every one his own goodness: but a faithful man who can find?
\fn{goodness: or, bounty.}
\vs Pro 20:7 The just \bibemph{man} walketh in his integrity: his children \bibemph{are} blessed after him.
\vs Pro 20:8 A king that sitteth in the throne of judgment scattereth away all evil with his eyes.
\vs Pro 20:9 Who can say, I have made my heart clean, I am pure from my sin?
\vs Pro 20:10 Divers weights, \bibemph{and} divers measures, both of them \bibemph{are} alike abomination to the \textsc{Lord}.
\fn{Divers weights: Heb. A stone and a stone. divers measures: Heb. an ephah and an ephah.}
\vs Pro 20:11 Even a child is known by his doings, whether his work \bibemph{be} pure, and whether \bibemph{it be} right.
\vs Pro 20:12 The hearing ear, and the seeing eye, the \textsc{Lord} hath made even both of them.
\vs Pro 20:13 Love not sleep, lest thou come to poverty; open thine eyes, \bibemph{and} thou shalt be satisfied with bread.
\vs Pro 20:14 \bibemph{It is} naught, \bibemph{it is} naught, saith the buyer: but when he is gone his way, then he boasteth.
\vs Pro 20:15 There is gold, and a multitude of rubies: but the lips of knowledge \bibemph{are} a precious jewel.
\vs Pro 20:16 Take his garment that is surety \bibemph{for} a stranger: and take a pledge of him for a strange woman.
\vs Pro 20:17 Bread of deceit \bibemph{is} sweet to a man; but afterwards his mouth shall be filled with gravel.
\fn{deceit: Heb. lying, or, falsehood.}
\vs Pro 20:18 \bibemph{Every} purpose is established by counsel: and with good advice make war.
\vs Pro 20:19 He that goeth about \bibemph{as} a talebearer revealeth secrets: therefore meddle not with him that flattereth with his lips.
\fn{flattereth: or, enticeth.}
\vs Pro 20:20 Whoso curseth his father or his mother, his lamp shall be put out in obscure darkness.
\fn{lamp: or, candle.}
\vs Pro 20:21 An inheritance \bibemph{may be} gotten hastily at the beginning; but the end thereof shall not be blessed.
\vs Pro 20:22 Say not thou, I will recompense evil; \bibemph{but} wait on the \textsc{Lord}, and he shall save thee.
\vs Pro 20:23 Divers weights \bibemph{are} an abomination unto the \textsc{Lord}; and a false balance \bibemph{is} not good.
\fn{a false...: Heb. balance of deceit.}
\vs Pro 20:24 Man's goings \bibemph{are} of the \textsc{Lord}; how can a man then understand his own way?
\vs Pro 20:25 \bibemph{It is} a snare to the man \bibemph{who} devoureth \bibemph{that which is} holy, and after vows to make enquiry.
\vs Pro 20:26 A wise king scattereth the wicked, and bringeth the wheel over them.
\vs Pro 20:27 The spirit of man \bibemph{is} the candle of the \textsc{Lord}, searching all the inward parts of the belly.
\fn{candle: or, lamp.}
\vs Pro 20:28 Mercy and truth preserve the king: and his throne is upholden by mercy.
\vs Pro 20:29 The glory of young men \bibemph{is} their strength: and the beauty of old men \bibemph{is} the gray head.
\vs Pro 20:30 The blueness of a wound cleanseth away evil: so \bibemph{do} stripes the inward parts of the belly.
\fn{cleanseth...: Heb. is a purging medicine against.}
\vs Pro 21:1 The king's heart \bibemph{is} in the hand of the \textsc{Lord}, \bibemph{as} the rivers of water: he turneth it whithersoever he will.
\vs Pro 21:2 Every way of a man \bibemph{is} right in his own eyes: but the \textsc{Lord} pondereth the hearts.
\vs Pro 21:3 To do justice and judgment \bibemph{is} more acceptable to the \textsc{Lord} than sacrifice.
\vs Pro 21:4 An high look, and a proud heart, \bibemph{and} the plowing of the wicked, \bibemph{is} sin.
\fn{An...: Heb. Haughtiness of eyes. the plowing: or, the light.}
\vs Pro 21:5 The thoughts of the diligent \bibemph{tend} only to plenteousness; but of every one \bibemph{that is} hasty only to want.
\vs Pro 21:6 The getting of treasures by a lying tongue \bibemph{is} a vanity tossed to and fro of them that seek death.
\vs Pro 21:7 The robbery of the wicked shall destroy them; because they refuse to do judgment.
\fn{destroy...: Heb. saw them, or, dwell with them.}
\vs Pro 21:8 The way of man \bibemph{is} froward and strange: but \bibemph{as for} the pure, his work \bibemph{is} right.
\vs Pro 21:9 \bibemph{It is} better to dwell in a corner of the housetop, than with a brawling woman in a wide house.
\fn{a brawling...: Heb. a woman of contentions. a wide...: Heb. an house of society.}
\vs Pro 21:10 The soul of the wicked desireth evil: his neighbour findeth no favour in his eyes.
\fn{findeth...: Heb. is not favoured.}
\vs Pro 21:11 When the scorner is punished, the simple is made wise: and when the wise is instructed, he receiveth knowledge.
\vs Pro 21:12 The righteous \bibemph{man} wisely considereth the house of the wicked: \bibemph{but God} overthroweth the wicked for \bibemph{their} wickedness.
\vs Pro 21:13 Whoso stoppeth his ears at the cry of the poor, he also shall cry himself, but shall not be heard.
\vs Pro 21:14 A gift in secret pacifieth anger: and a reward in the bosom strong wrath.
\vs Pro 21:15 \bibemph{It is} joy to the just to do judgment: but destruction \bibemph{shall be} to the workers of iniquity.
\vs Pro 21:16 The man that wandereth out of the way of understanding shall remain in the congregation of the dead.
\vs Pro 21:17 He that loveth pleasure \bibemph{shall be} a poor man: he that loveth wine and oil shall not be rich.
\fn{pleasure: or, sport.}
\vs Pro 21:18 The wicked \bibemph{shall be} a ransom for the righteous, and the transgressor for the upright.
\vs Pro 21:19 \bibemph{It is} better to dwell in the wilderness, than with a contentious and an angry woman.
\fn{in...: Heb. in the land of the desert.}
\vs Pro 21:20 \bibemph{There is} treasure to be desired and oil in the dwelling of the wise; but a foolish man spendeth it up.
\vs Pro 21:21 He that followeth after righteousness and mercy findeth life, righteousness, and honour.
\vs Pro 21:22 A wise \bibemph{man} scaleth the city of the mighty, and casteth down the strength of the confidence thereof.
\vs Pro 21:23 Whoso keepeth his mouth and his tongue keepeth his soul from troubles.
\vs Pro 21:24 Proud \bibemph{and} haughty scorner \bibemph{is} his name, who dealeth in proud wrath.
\fn{in proud...: Heb. in the wrath of pride.}
\vs Pro 21:25 The desire of the slothful killeth him; for his hands refuse to labour.
\vs Pro 21:26 He coveteth greedily all the day long: but the righteous giveth and spareth not.
\vs Pro 21:27 The sacrifice of the wicked \bibemph{is} abomination: how much more, \bibemph{when} he bringeth it with a wicked mind?
\fn{with...: Heb. in wickedness?.}
\vs Pro 21:28 A false witness shall perish: but the man that heareth speaketh constantly.
\fn{A...: Heb. A witness of lies.}
\vs Pro 21:29 A wicked man hardeneth his face: but \bibemph{as for} the upright, he directeth his way.
\fn{directeth: or, considereth.}
\vs Pro 21:30 \bibemph{There is} no wisdom nor understanding nor counsel against the \textsc{Lord}.
\vs Pro 21:31 The horse \bibemph{is} prepared against the day of battle: but safety \bibemph{is} of the \textsc{Lord}.
\fn{safety: or, victory.}
\vs Pro 22:1 A \bibemph{good} name \bibemph{is} rather to be chosen than great riches, \bibemph{and} loving favour rather than silver and gold.
\fn{loving...: or, favour is better than, etc.}
\vs Pro 22:2 The rich and poor meet together: the \textsc{Lord} \bibemph{is} the maker of them all.
\vs Pro 22:3 A prudent \bibemph{man} foreseeth the evil, and hideth himself: but the simple pass on, and are punished.
\vs Pro 22:4 By humility \bibemph{and} the fear of the \textsc{Lord} \bibemph{are} riches, and honour, and life.
\fn{By...: or, The reward of humility, etc.}
\vs Pro 22:5 Thorns \bibemph{and} snares \bibemph{are} in the way of the froward: he that doth keep his soul shall be far from them.
\vs Pro 22:6 Train up a child in the way he should go: and when he is old, he will not depart from it.
\fn{Train...: or, Catechise. in...: Heb. in his way.}
\vs Pro 22:7 The rich ruleth over the poor, and the borrower \bibemph{is} servant to the lender.
\fn{the lender: Heb. the man that lendeth.}
\vs Pro 22:8 He that soweth iniquity shall reap vanity: and the rod of his anger shall fail.
\fn{the rod...: or, with the rod of his anger he shall be consumed.}
\vs Pro 22:9 He that hath a bountiful eye shall be blessed; for he giveth of his bread to the poor.
\fn{He that...: Heb. Good of eye.}
\vs Pro 22:10 Cast out the scorner, and contention shall go out; yea, strife and reproach shall cease.
\vs Pro 22:11 He that loveth pureness of heart, \bibemph{for} the grace of his lips the king \bibemph{shall be} his friend.
\fn{for...: or, and hath grace in his lips.}
\vs Pro 22:12 The eyes of the \textsc{Lord} preserve knowledge, and he overthroweth the words of the transgressor.
\fn{the words: or, the matters.}
\vs Pro 22:13 The slothful \bibemph{man} saith, \bibemph{There is} a lion without, I shall be slain in the streets.
\vs Pro 22:14 The mouth of strange women \bibemph{is} a deep pit: he that is abhorred of the \textsc{Lord} shall fall therein.
\vs Pro 22:15 Foolishness \bibemph{is} bound in the heart of a child; \bibemph{but} the rod of correction shall drive it far from him.
\vs Pro 22:16 He that oppresseth the poor to increase his \bibemph{riches, and} he that giveth to the rich, \bibemph{shall} surely \bibemph{come} to want.
\vs Pro 22:17 Bow down thine ear, and hear the words of the wise, and apply thine heart unto my knowledge.
\vs Pro 22:18 For \bibemph{it is} a pleasant thing if thou keep them within thee; they shall withal be fitted in thy lips.
\fn{within...: Heb. in thy belly.}
\vs Pro 22:19 That thy trust may be in the \textsc{Lord}, I have made known to thee this day, even to thee.
\fn{even...: or, trust thou also.}
\vs Pro 22:20 Have not I written to thee excellent things in counsels and knowledge,
\vs Pro 22:21 That I might make thee know the certainty of the words of truth; that thou mightest answer the words of truth to them that send unto thee?
\fn{to them...: or, to those that send thee?.}
\vs Pro 22:22 Rob not the poor, because he \bibemph{is} poor: neither oppress the afflicted in the gate:
\vs Pro 22:23 For the \textsc{Lord} will plead their cause, and spoil the soul of those that spoiled them.
\vs Pro 22:24 Make no friendship with an angry man; and with a furious man thou shalt not go:
\vs Pro 22:25 Lest thou learn his ways, and get a snare to thy soul.
\vs Pro 22:26 Be not thou \bibemph{one} of them that strike hands, \bibemph{or} of them that are sureties for debts.
\vs Pro 22:27 If thou hast nothing to pay, why should he take away thy bed from under thee?
\vs Pro 22:28 Remove not the ancient landmark, which thy fathers have set.
\fn{landmark: or, bound.}
\vs Pro 22:29 Seest thou a man diligent in his business? he shall stand before kings; he shall not stand before mean \bibemph{men}.
\fn{mean...: Heb. obscure men.}
\vs Pro 23:1 When thou sittest to eat with a ruler, consider diligently what \bibemph{is} before thee:
\vs Pro 23:2 And put a knife to thy throat, if thou \bibemph{be} a man given to appetite.
\vs Pro 23:3 Be not desirous of his dainties: for they \bibemph{are} deceitful meat.
\vs Pro 23:4 Labour not to be rich: cease from thine own wisdom.
\vs Pro 23:5 Wilt thou set thine eyes upon that which is not? for \bibemph{riches} certainly make themselves wings; they fly away as an eagle toward heaven.
\fn{set...: Heb. cause thine eyes to fly upon.}
\vs Pro 23:6 Eat thou not the bread of \bibemph{him that hath} an evil eye, neither desire thou his dainty meats:
\vs Pro 23:7 For as he thinketh in his heart, so \bibemph{is} he: Eat and drink, saith he to thee; but his heart \bibemph{is} not with thee.
\vs Pro 23:8 The morsel \bibemph{which} thou hast eaten shalt thou vomit up, and lose thy sweet words.
\vs Pro 23:9 Speak not in the ears of a fool: for he will despise the wisdom of thy words.
\vs Pro 23:10 Remove not the old landmark; and enter not into the fields of the fatherless:
\fn{landmark: or, bound.}
\vs Pro 23:11 For their redeemer \bibemph{is} mighty; he shall plead their cause with thee.
\vs Pro 23:12 Apply thine heart unto instruction, and thine ears to the words of knowledge.
\vs Pro 23:13 Withhold not correction from the child: for \bibemph{if} thou beatest him with the rod, he shall not die.
\vs Pro 23:14 Thou shalt beat him with the rod, and shalt deliver his soul from hell.
\vs Pro 23:15 My son, if thine heart be wise, my heart shall rejoice, even mine.
\fn{even...: or, even I will rejoice.}
\vs Pro 23:16 Yea, my reins shall rejoice, when thy lips speak right things.
\vs Pro 23:17 Let not thine heart envy sinners: but \bibemph{be thou} in the fear of the \textsc{Lord} all the day long.
\vs Pro 23:18 For surely there is an end; and thine expectation shall not be cut off.
\fn{end: or, reward.}
\vs Pro 23:19 Hear thou, my son, and be wise, and guide thine heart in the way.
\vs Pro 23:20 Be not among winebibbers; among riotous eaters of flesh:
\fn{of flesh: Heb. of their flesh.}
\vs Pro 23:21 For the drunkard and the glutton shall come to poverty: and drowsiness shall clothe \bibemph{a man} with rags.
\vs Pro 23:22 Hearken unto thy father that begat thee, and despise not thy mother when she is old.
\vs Pro 23:23 Buy the truth, and sell \bibemph{it} not; \bibemph{also} wisdom, and instruction, and understanding.
\vs Pro 23:24 The father of the righteous shall greatly rejoice: and he that begetteth a wise \bibemph{child} shall have joy of him.
\vs Pro 23:25 Thy father and thy mother shall be glad, and she that bare thee shall rejoice.
\vs Pro 23:26 My son, give me thine heart, and let thine eyes observe my ways.
\vs Pro 23:27 For a whore \bibemph{is} a deep ditch; and a strange woman \bibemph{is} a narrow pit.
\vs Pro 23:28 She also lieth in wait as \bibemph{for} a prey, and increaseth the transgressors among men.
\fn{as for...: or, as a robber.}
\vs Pro 23:29 Who hath woe? who hath sorrow? who hath contentions? who hath babbling? who hath wounds without cause? who hath redness of eyes?
\vs Pro 23:30 They that tarry long at the wine; they that go to seek mixed wine.
\vs Pro 23:31 Look not thou upon the wine when it is red, when it giveth his colour in the cup, \bibemph{when} it moveth itself aright.
\vs Pro 23:32 At the last it biteth like a serpent, and stingeth like an adder.
\fn{an...: or, a cockatrice.}
\vs Pro 23:33 Thine eyes shall behold strange women, and thine heart shall utter perverse things.
\vs Pro 23:34 Yea, thou shalt be as he that lieth down in the midst of the sea, or as he that lieth upon the top of a mast.
\fn{the midst...: Heb. the heart of the sea.}
\vs Pro 23:35 They have stricken me, \bibemph{shalt thou say, and} I was not sick; they have beaten me, \bibemph{and} I felt \bibemph{it} not: when shall I awake? I will seek it yet again.
\fn{I felt...: Heb. I knew it not.}
\vs Pro 24:1 Be not thou envious against evil men, neither desire to be with them.
\vs Pro 24:2 For their heart studieth destruction, and their lips talk of mischief.
\vs Pro 24:3 Through wisdom is an house builded; and by understanding it is established:
\vs Pro 24:4 And by knowledge shall the chambers be filled with all precious and pleasant riches.
\vs Pro 24:5 A wise man \bibemph{is} strong; yea, a man of knowledge increaseth strength.
\fn{is strong: Heb. is in strength. increaseth...: Heb. strengtheneth might.}
\vs Pro 24:6 For by wise counsel thou shalt make thy war: and in multitude of counsellors \bibemph{there is} safety.
\vs Pro 24:7 Wisdom \bibemph{is} too high for a fool: he openeth not his mouth in the gate.
\vs Pro 24:8 He that deviseth to do evil shall be called a mischievous person.
\vs Pro 24:9 The thought of foolishness \bibemph{is} sin: and the scorner \bibemph{is} an abomination to men.
\vs Pro 24:10 \bibemph{If} thou faint in the day of adversity, thy strength \bibemph{is} small.
\fn{small: Heb. narrow.}
\vs Pro 24:11 If thou forbear to deliver \bibemph{them that are} drawn unto death, and \bibemph{those that are} ready to be slain;
\vs Pro 24:12 If thou sayest, Behold, we knew it not; doth not he that pondereth the heart consider \bibemph{it}? and he that keepeth thy soul, doth \bibemph{not} he know \bibemph{it}? and shall \bibemph{not} he render to \bibemph{every} man according to his works?
\vs Pro 24:13 My son, eat thou honey, because \bibemph{it is} good; and the honeycomb, \bibemph{which is} sweet to thy taste:
\fn{to...: Heb. upon thy palate.}
\vs Pro 24:14 So \bibemph{shall} the knowledge of wisdom \bibemph{be} unto thy soul: when thou hast found \bibemph{it}, then there shall be a reward, and thy expectation shall not be cut off.
\vs Pro 24:15 Lay not wait, O wicked \bibemph{man}, against the dwelling of the righteous; spoil not his resting place:
\vs Pro 24:16 For a just \bibemph{man} falleth seven times, and riseth up again: but the wicked shall fall into mischief.
\vs Pro 24:17 Rejoice not when thine enemy falleth, and let not thine heart be glad when he stumbleth:
\vs Pro 24:18 Lest the \textsc{Lord} see \bibemph{it}, and it displease him, and he turn away his wrath from him.
\fn{it displeaseth...: Heb. it be evil in his eyes.}
\vs Pro 24:19 Fret not thyself because of evil \bibemph{men}, neither be thou envious at the wicked;
\fn{Fret...: or, Keep not company with the wicked.}
\vs Pro 24:20 For there shall be no reward to the evil \bibemph{man}; the candle of the wicked shall be put out.
\fn{candle: or, lamp.}
\vs Pro 24:21 My son, fear thou the \textsc{Lord} and the king: \bibemph{and} meddle not with them that are given to change:
\fn{them...: Heb. changers.}
\vs Pro 24:22 For their calamity shall rise suddenly; and who knoweth the ruin of them both?
\vs Pro 24:23 These \bibemph{things} also \bibemph{belong} to the wise. \bibemph{It is} not good to have respect of persons in judgment.
\vs Pro 24:24 He that saith unto the wicked, Thou \bibemph{art} righteous; him shall the people curse, nations shall abhor him:
\vs Pro 24:25 But to them that rebuke \bibemph{him} shall be delight, and a good blessing shall come upon them.
\fn{a good...: Heb. a blessing of good.}
\vs Pro 24:26 \bibemph{Every man} shall kiss \bibemph{his} lips that giveth a right answer.
\fn{that...: Heb. that answereth right words.}
\vs Pro 24:27 Prepare thy work without, and make it fit for thyself in the field; and afterwards build thine house.
\vs Pro 24:28 Be not a witness against thy neighbour without cause; and deceive \bibemph{not} with thy lips.
\vs Pro 24:29 Say not, I will do so to him as he hath done to me: I will render to the man according to his work.
\vs Pro 24:30 I went by the field of the slothful, and by the vineyard of the man void of understanding;
\vs Pro 24:31 And, lo, it was all grown over with thorns, \bibemph{and} nettles had covered the face thereof, and the stone wall thereof was broken down.
\vs Pro 24:32 Then I saw, \bibemph{and} considered \bibemph{it} well: I looked upon \bibemph{it, and} received instruction.
\fn{considered...: Heb. set my heart.}
\vs Pro 24:33 \bibemph{Yet} a little sleep, a little slumber, a little folding of the hands to sleep:
\vs Pro 24:34 So shall thy poverty come \bibemph{as} one that travelleth; and thy want as an armed man.
\fn{an...: Heb. a man of shield.}
\vs Pro 25:1 These \bibemph{are} also proverbs of Solomon, which the men of Hezekiah king of Judah copied out.
\vs Pro 25:2 \bibemph{It is} the glory of God to conceal a thing: but the honour of kings \bibemph{is} to search out a matter.
\vs Pro 25:3 The heaven for height, and the earth for depth, and the heart of kings \bibemph{is} unsearchable.
\fn{is...: Heb. there is no searching.}
\vs Pro 25:4 Take away the dross from the silver, and there shall come forth a vessel for the finer.
\vs Pro 25:5 Take away the wicked \bibemph{from} before the king, and his throne shall be established in righteousness.
\vs Pro 25:6 Put not forth thyself in the presence of the king, and stand not in the place of great \bibemph{men}:
\fn{Put...: Heb. Set not out thy glory.}
\vs Pro 25:7 For better \bibemph{it is} that it be said unto thee, Come up hither; than that thou shouldest be put lower in the presence of the prince whom thine eyes have seen.
\vs Pro 25:8 Go not forth hastily to strive, lest \bibemph{thou know not} what to do in the end thereof, when thy neighbour hath put thee to shame.
\vs Pro 25:9 Debate thy cause with thy neighbour \bibemph{himself}; and discover not a secret to another:
\fn{a secret...: or, the secret of.}
\vs Pro 25:10 Lest he that heareth \bibemph{it} put thee to shame, and thine infamy turn not away.
\vs Pro 25:11 A word fitly spoken \bibemph{is like} apples of gold in pictures of silver.
\fn{fitly...: Heb. spoken upon his wheels.}
\vs Pro 25:12 \bibemph{As} an earring of gold, and an ornament of fine gold, \bibemph{so is} a wise reprover upon an obedient ear.
\vs Pro 25:13 As the cold of snow in the time of harvest, \bibemph{so is} a faithful messenger to them that send him: for he refresheth the soul of his masters.
\vs Pro 25:14 Whoso boasteth himself of a false gift \bibemph{is like} clouds and wind without rain.
\fn{of a...: Heb. in a gift of falsehood.}
\vs Pro 25:15 By long forbearing is a prince persuaded, and a soft tongue breaketh the bone.
\vs Pro 25:16 Hast thou found honey? eat so much as is sufficient for thee, lest thou be filled therewith, and vomit it.
\vs Pro 25:17 Withdraw thy foot from thy neighbour's house; lest he be weary of thee, and \bibemph{so} hate thee.
\fn{Withdraw...: or, Let thy foot be seldom in. weary...: Heb. full of thee.}
\vs Pro 25:18 A man that beareth false witness against his neighbour \bibemph{is} a maul, and a sword, and a sharp arrow.
\vs Pro 25:19 Confidence in an unfaithful man in time of trouble \bibemph{is like} a broken tooth, and a foot out of joint.
\vs Pro 25:20 \bibemph{As} he that taketh away a garment in cold weather, \bibemph{and as} vinegar upon nitre, so \bibemph{is} he that singeth songs to an heavy heart.
\vs Pro 25:21 If thine enemy be hungry, give him bread to eat; and if he be thirsty, give him water to drink:
\vs Pro 25:22 For thou shalt heap coals of fire upon his head, and the \textsc{Lord} shall reward thee.
\vs Pro 25:23 The north wind driveth away rain: so \bibemph{doth} an angry countenance a backbiting tongue.
\fn{driveth...: or, bringeth forth rain: so doth a backbiting tongue an angry countenance.}
\vs Pro 25:24 \bibemph{It is} better to dwell in the corner of the housetop, than with a brawling woman and in a wide house.
\vs Pro 25:25 \bibemph{As} cold waters to a thirsty soul, so \bibemph{is} good news from a far country.
\vs Pro 25:26 A righteous man falling down before the wicked \bibemph{is as} a troubled fountain, and a corrupt spring.
\vs Pro 25:27 \bibemph{It is} not good to eat much honey: so \bibemph{for men} to search their own glory \bibemph{is not} glory.
\vs Pro 25:28 He that \bibemph{hath} no rule over his own spirit \bibemph{is like} a city \bibemph{that is} broken down, \bibemph{and} without walls.
\vs Pro 26:1 As snow in summer, and as rain in harvest, so honour is not seemly for a fool.
\vs Pro 26:2 As the bird by wandering, as the swallow by flying, so the curse causeless shall not come.
\vs Pro 26:3 A whip for the horse, a bridle for the ass, and a rod for the fool's back.
\vs Pro 26:4 Answer not a fool according to his folly, lest thou also be like unto him.
\vs Pro 26:5 Answer a fool according to his folly, lest he be wise in his own conceit.
\fn{conceit: Heb. eyes.}
\vs Pro 26:6 He that sendeth a message by the hand of a fool cutteth off the feet, \bibemph{and} drinketh damage.
\fn{damage: or, violence.}
\vs Pro 26:7 The legs of the lame are not equal: so \bibemph{is} a parable in the mouth of fools.
\fn{are...: Heb. are lifted up.}
\vs Pro 26:8 As he that bindeth a stone in a sling, so \bibemph{is} he that giveth honour to a fool.
\fn{bindeth...: or, putteth a precious stone in an heap of stones.}
\vs Pro 26:9 \bibemph{As} a thorn goeth up into the hand of a drunkard, so \bibemph{is} a parable in the mouth of fools.
\vs Pro 26:10 The great \bibemph{God} that formed all \bibemph{things} both rewardeth the fool, and rewardeth transgressors.
\fn{The great...: or, A great man grieveth all, and he hireth the fool, he hireth also transgressors.}
\vs Pro 26:11 As a dog returneth to his vomit, \bibemph{so} a fool returneth to his folly.
\fn{returneth to his folly: Heb. iterateth his folly.}
\vs Pro 26:12 Seest thou a man wise in his own conceit? \bibemph{there is} more hope of a fool than of him.
\vs Pro 26:13 The slothful \bibemph{man} saith, \bibemph{There is} a lion in the way; a lion \bibemph{is} in the streets.
\vs Pro 26:14 \bibemph{As} the door turneth upon his hinges, so \bibemph{doth} the slothful upon his bed.
\vs Pro 26:15 The slothful hideth his hand in \bibemph{his} bosom; it grieveth him to bring it again to his mouth.
\fn{it grieveth...: or, he is weary.}
\vs Pro 26:16 The sluggard \bibemph{is} wiser in his own conceit than seven men that can render a reason.
\vs Pro 26:17 He that passeth by, \bibemph{and} meddleth with strife \bibemph{belonging} not to him, \bibemph{is like} one that taketh a dog by the ears.
\fn{meddleth: or, is enraged.}
\vs Pro 26:18 As a mad \bibemph{man} who casteth firebrands, arrows, and death,
\fn{firebrands: Heb. flames, or, sparks.}
\vs Pro 26:19 So \bibemph{is} the man \bibemph{that} deceiveth his neighbour, and saith, Am not I in sport?
\vs Pro 26:20 Where no wood is, \bibemph{there} the fire goeth out: so where \bibemph{there is} no talebearer, the strife ceaseth.
\fn{Where no...: Heb. Without wood. talebearer: or, whisperer. ceaseth: Heb. is silent.}
\vs Pro 26:21 \bibemph{As} coals \bibemph{are} to burning coals, and wood to fire; so \bibemph{is} a contentious man to kindle strife.
\vs Pro 26:22 The words of a talebearer \bibemph{are} as wounds, and they go down into the innermost parts of the belly.
\fn{innermost...: Heb. chambers.}
\vs Pro 26:23 Burning lips and a wicked heart \bibemph{are like} a potsherd covered with silver dross.
\vs Pro 26:24 He that hateth dissembleth with his lips, and layeth up deceit within him;
\fn{dissembleth: or, is known.}
\vs Pro 26:25 When he speaketh fair, believe him not: for \bibemph{there are} seven abominations in his heart.
\fn{speaketh...: Heb. maketh his voice gracious.}
\vs Pro 26:26 \bibemph{Whose} hatred is covered by deceit, his wickedness shall be shewed before the \bibemph{whole} congregation.
\fn{by...: or, in secret.}
\vs Pro 26:27 Whoso diggeth a pit shall fall therein: and he that rolleth a stone, it will return upon him.
\vs Pro 26:28 A lying tongue hateth \bibemph{those that are} afflicted by it; and a flattering mouth worketh ruin.
\vs Pro 27:1 Boast not thyself of to morrow; for thou knowest not what a day may bring forth.
\fn{to...: Heb. to morrow day.}
\vs Pro 27:2 Let another man praise thee, and not thine own mouth; a stranger, and not thine own lips.
\vs Pro 27:3 A stone \bibemph{is} heavy, and the sand weighty; but a fool's wrath \bibemph{is} heavier than them both.
\fn{heavy: Heb. heaviness.}
\vs Pro 27:4 Wrath \bibemph{is} cruel, and anger \bibemph{is} outrageous; but who \bibemph{is} able to stand before envy?
\fn{Wrath...: Heb. Wrath is cruelty, and anger an overflowing. envy: or, jealousy?.}
\vs Pro 27:5 Open rebuke \bibemph{is} better than secret love.
\vs Pro 27:6 Faithful \bibemph{are} the wounds of a friend; but the kisses of an enemy \bibemph{are} deceitful.
\fn{deceitful: or, earnest, or, frequent.}
\vs Pro 27:7 The full soul loatheth an honeycomb; but to the hungry soul every bitter thing is sweet.
\fn{loatheth: Heb. treadeth under foot.}
\vs Pro 27:8 As a bird that wandereth from her nest, so \bibemph{is} a man that wandereth from his place.
\vs Pro 27:9 Ointment and perfume rejoice the heart: so \bibemph{doth} the sweetness of a man's friend by hearty counsel.
\fn{by...: Heb. from the counsel of the soul.}
\vs Pro 27:10 Thine own friend, and thy father's friend, forsake not; neither go into thy brother's house in the day of thy calamity: \bibemph{for} better \bibemph{is} a neighbour \bibemph{that is} near than a brother far off.
\vs Pro 27:11 My son, be wise, and make my heart glad, that I may answer him that reproacheth me.
\vs Pro 27:12 A prudent \bibemph{man} foreseeth the evil, \bibemph{and} hideth himself; \bibemph{but} the simple pass on, \bibemph{and} are punished.
\vs Pro 27:13 Take his garment that is surety for a stranger, and take a pledge of him for a strange woman.
\vs Pro 27:14 He that blesseth his friend with a loud voice, rising early in the morning, it shall be counted a curse to him.
\vs Pro 27:15 A continual dropping in a very rainy day and a contentious woman are alike.
\vs Pro 27:16 Whosoever hideth her hideth the wind, and the ointment of his right hand, \bibemph{which} bewrayeth \bibemph{itself}.
\vs Pro 27:17 Iron sharpeneth iron; so a man sharpeneth the countenance of his friend.
\vs Pro 27:18 Whoso keepeth the fig tree shall eat the fruit thereof: so he that waiteth on his master shall be honoured.
\vs Pro 27:19 As in water face \bibemph{answereth} to face, so the heart of man to man.
\vs Pro 27:20 Hell and destruction are never full; so the eyes of man are never satisfied.
\fn{never: Heb. not.}
\vs Pro 27:21 \bibemph{As} the fining pot for silver, and the furnace for gold; so \bibemph{is} a man to his praise.
\vs Pro 27:22 Though thou shouldest bray a fool in a mortar among wheat with a pestle, \bibemph{yet} will not his foolishness depart from him.
\vs Pro 27:23 Be thou diligent to know the state of thy flocks, \bibemph{and} look well to thy herds.
\fn{look...: Heb. set thy heart.}
\vs Pro 27:24 For riches \bibemph{are} not for ever: and doth the crown \bibemph{endure} to every generation?
\fn{riches: Heb. strength. to...: Heb. to generation and generation?.}
\vs Pro 27:25 The hay appeareth, and the tender grass sheweth itself, and herbs of the mountains are gathered.
\vs Pro 27:26 The lambs \bibemph{are} for thy clothing, and the goats \bibemph{are} the price of the field.
\vs Pro 27:27 And \bibemph{thou shalt have} goats' milk enough for thy food, for the food of thy household, and \bibemph{for} the maintenance for thy maidens.
\fn{maintenance: Heb. life.}
\vs Pro 28:1 The wicked flee when no man pursueth: but the righteous are bold as a lion.
\vs Pro 28:2 For the transgression of a land many \bibemph{are} the princes thereof: but by a man of understanding \bibemph{and} knowledge the state \bibemph{thereof} shall be prolonged.
\fn{by...: or, by men of understanding and wisdom shall they likewise be prolonged.}
\vs Pro 28:3 A poor man that oppresseth the poor \bibemph{is like} a sweeping rain which leaveth no food.
\fn{which...: Heb. without food.}
\vs Pro 28:4 They that forsake the law praise the wicked: but such as keep the law contend with them.
\vs Pro 28:5 Evil men understand not judgment: but they that seek the \textsc{Lord} understand all \bibemph{things}.
\vs Pro 28:6 Better \bibemph{is} the poor that walketh in his uprightness, than \bibemph{he that is} perverse \bibemph{in his} ways, though he \bibemph{be} rich.
\vs Pro 28:7 Whoso keepeth the law \bibemph{is} a wise son: but he that is a companion of riotous \bibemph{men} shameth his father.
\fn{is a companion...: or, feedeth gluttons.}
\vs Pro 28:8 He that by usury and unjust gain increaseth his substance, he shall gather it for him that will pity the poor.
\fn{unjust...: Heb. by increase.}
\vs Pro 28:9 He that turneth away his ear from hearing the law, even his prayer \bibemph{shall be} abomination.
\vs Pro 28:10 Whoso causeth the righteous to go astray in an evil way, he shall fall himself into his own pit: but the upright shall have good \bibemph{things} in possession.
\vs Pro 28:11 The rich man \bibemph{is} wise in his own conceit; but the poor that hath understanding searcheth him out.
\fn{own...: Heb. eyes.}
\vs Pro 28:12 When righteous \bibemph{men} do rejoice, \bibemph{there is} great glory: but when the wicked rise, a man is hidden.
\fn{hidden: or, sought for.}
\vs Pro 28:13 He that covereth his sins shall not prosper: but whoso confesseth and forsaketh \bibemph{them} shall have mercy.
\vs Pro 28:14 Happy \bibemph{is} the man that feareth alway: but he that hardeneth his heart shall fall into mischief.
\vs Pro 28:15 \bibemph{As} a roaring lion, and a ranging bear; \bibemph{so is} a wicked ruler over the poor people.
\vs Pro 28:16 The prince that wanteth understanding \bibemph{is} also a great oppressor: \bibemph{but} he that hateth covetousness shall prolong \bibemph{his} days.
\vs Pro 28:17 A man that doeth violence to the blood of \bibemph{any} person shall flee to the pit; let no man stay him.
\vs Pro 28:18 Whoso walketh uprightly shall be saved: but \bibemph{he that is} perverse \bibemph{in his} ways shall fall at once.
\vs Pro 28:19 He that tilleth his land shall have plenty of bread: but he that followeth after vain \bibemph{persons} shall have poverty enough.
\vs Pro 28:20 A faithful man shall abound with blessings: but he that maketh haste to be rich shall not be innocent.
\fn{innocent: or, unpunished.}
\vs Pro 28:21 To have respect of persons \bibemph{is} not good: for for a piece of bread \bibemph{that} man will transgress.
\vs Pro 28:22 He that hasteth to be rich \bibemph{hath} an evil eye, and considereth not that poverty shall come upon him.
\fn{hasteth...: or, hath and evil eye hasteth to be rich.}
\vs Pro 28:23 He that rebuketh a man afterwards shall find more favour than he that flattereth with the tongue.
\vs Pro 28:24 Whoso robbeth his father or his mother, and saith, \bibemph{It is} no transgression; the same \bibemph{is} the companion of a destroyer.
\fn{a destroyer: Heb. a man destroying.}
\vs Pro 28:25 He that is of a proud heart stirreth up strife: but he that putteth his trust in the \textsc{Lord} shall be made fat.
\vs Pro 28:26 He that trusteth in his own heart is a fool: but whoso walketh wisely, he shall be delivered.
\vs Pro 28:27 He that giveth unto the poor shall not lack: but he that hideth his eyes shall have many a curse.
\vs Pro 28:28 When the wicked rise, men hide themselves: but when they perish, the righteous increase.
\vs Pro 29:1 He, that being often reproved hardeneth \bibemph{his} neck, shall suddenly be destroyed, and that without remedy.
\fn{He...: Heb. A man of reproofs.}
\vs Pro 29:2 When the righteous are in authority, the people rejoice: but when the wicked beareth rule, the people mourn.
\fn{in...: or, increased.}
\vs Pro 29:3 Whoso loveth wisdom rejoiceth his father: but he that keepeth company with harlots spendeth \bibemph{his} substance.
\vs Pro 29:4 The king by judgment establisheth the land: but he that receiveth gifts overthroweth it.
\fn{he...: Heb. a man of oblations.}
\vs Pro 29:5 A man that flattereth his neighbour spreadeth a net for his feet.
\vs Pro 29:6 In the transgression of an evil man \bibemph{there is} a snare: but the righteous doth sing and rejoice.
\vs Pro 29:7 The righteous considereth the cause of the poor: \bibemph{but} the wicked regardeth not to know \bibemph{it}.
\vs Pro 29:8 Scornful men bring a city into a snare: but wise \bibemph{men} turn away wrath.
\fn{bring...: or, set a city on fire.}
\vs Pro 29:9 \bibemph{If} a wise man contendeth with a foolish man, whether he rage or laugh, \bibemph{there is} no rest.
\vs Pro 29:10 The bloodthirsty hate the upright: but the just seek his soul.
\fn{The bloodthirsty: Heb. Men of blood.}
\vs Pro 29:11 A fool uttereth all his mind: but a wise \bibemph{man} keepeth it in till afterwards.
\vs Pro 29:12 If a ruler hearken to lies, all his servants \bibemph{are} wicked.
\vs Pro 29:13 The poor and the deceitful man meet together: the \textsc{Lord} lighteneth both their eyes.
\fn{the deceitful...: or, the usurer.}
\vs Pro 29:14 The king that faithfully judgeth the poor, his throne shall be established for ever.
\vs Pro 29:15 The rod and reproof give wisdom: but a child left \bibemph{to himself} bringeth his mother to shame.
\vs Pro 29:16 When the wicked are multiplied, transgression increaseth: but the righteous shall see their fall.
\vs Pro 29:17 Correct thy son, and he shall give thee rest; yea, he shall give delight unto thy soul.
\vs Pro 29:18 Where \bibemph{there is} no vision, the people perish: but he that keepeth the law, happy \bibemph{is} he.
\fn{perish: or, is made naked.}
\vs Pro 29:19 A servant will not be corrected by words: for though he understand he will not answer.
\vs Pro 29:20 Seest thou a man \bibemph{that is} hasty in his words? \bibemph{there is} more hope of a fool than of him.
\fn{words: or, matters?.}
\vs Pro 29:21 He that delicately bringeth up his servant from a child shall have him become \bibemph{his} son at the length.
\vs Pro 29:22 An angry man stirreth up strife, and a furious man aboundeth in transgression.
\vs Pro 29:23 A man's pride shall bring him low: but honour shall uphold the humble in spirit.
\vs Pro 29:24 Whoso is partner with a thief hateth his own soul: he heareth cursing, and bewrayeth \bibemph{it} not.
\vs Pro 29:25 The fear of man bringeth a snare: but whoso putteth his trust in the \textsc{Lord} shall be safe.
\fn{safe: Heb. set on high.}
\vs Pro 29:26 Many seek the ruler's favour; but \bibemph{every} man's judgment \bibemph{cometh} from the \textsc{Lord}.
\fn{the ruler's...: Heb. the face of a ruler.}
\vs Pro 29:27 An unjust man \bibemph{is} an abomination to the just: and \bibemph{he that is} upright in the way \bibemph{is} abomination to the wicked.
\vs Pro 30:1 The words of Agur the son of Jakeh, \bibemph{even} the prophecy: the man spake unto Ithiel, even unto Ithiel and Ucal,
\vs Pro 30:2 Surely I \bibemph{am} more brutish than \bibemph{any} man, and have not the understanding of a man.
\vs Pro 30:3 I neither learned wisdom, nor have the knowledge of the holy.
\fn{have: Heb. know.}
\vs Pro 30:4 Who hath ascended up into heaven, or descended? who hath gathered the wind in his fists? who hath bound the waters in a garment? who hath established all the ends of the earth? what \bibemph{is} his name, and what \bibemph{is} his son's name, if thou canst tell?
\vs Pro 30:5 Every word of God \bibemph{is} pure: he \bibemph{is} a shield unto them that put their trust in him.
\fn{pure: Heb. purified.}
\vs Pro 30:6 Add thou not unto his words, lest he reprove thee, and thou be found a liar.
\vs Pro 30:7 Two \bibemph{things} have I required of thee; deny me \bibemph{them} not before I die:
\fn{deny...: Heb. withhold not from me.}
\vs Pro 30:8 Remove far from me vanity and lies: give me neither poverty nor riches; feed me with food convenient for me:
\fn{convenient...: Heb. of my allowance.}
\vs Pro 30:9 Lest I be full, and deny \bibemph{thee}, and say, Who \bibemph{is} the \textsc{Lord}? or lest I be poor, and steal, and take the name of my God \bibemph{in vain}.
\fn{deny...: Heb. belie thee.}
\vs Pro 30:10 Accuse not a servant unto his master, lest he curse thee, and thou be found guilty.
\fn{Accuse...: Heb. Hurt not with thy tongue.}
\vs Pro 30:11 \bibemph{There is} a generation \bibemph{that} curseth their father, and doth not bless their mother.
\vs Pro 30:12 \bibemph{There is} a generation \bibemph{that are} pure in their own eyes, and \bibemph{yet} is not washed from their filthiness.
\vs Pro 30:13 \bibemph{There is} a generation, O how lofty are their eyes! and their eyelids are lifted up.
\vs Pro 30:14 \bibemph{There is} a generation, whose teeth \bibemph{are as} swords, and their jaw teeth \bibemph{as} knives, to devour the poor from off the earth, and the needy from \bibemph{among} men.
\vs Pro 30:15 The horseleach hath two daughters, \bibemph{crying}, Give, give. There are three \bibemph{things that} are never satisfied, \bibemph{yea}, four \bibemph{things} say not, \bibemph{It is} enough:
\fn{It is...: Heb. Wealth.}
\vs Pro 30:16 The grave; and the barren womb; the earth \bibemph{that} is not filled with water; and the fire \bibemph{that} saith not, \bibemph{It is} enough.
\vs Pro 30:17 The eye \bibemph{that} mocketh at \bibemph{his} father, and despiseth to obey \bibemph{his} mother, the ravens of the valley shall pick it out, and the young eagles shall eat it.
\fn{the valley: or, the brook.}
\vs Pro 30:18 There be three \bibemph{things which} are too wonderful for me, yea, four which I know not:
\vs Pro 30:19 The way of an eagle in the air; the way of a serpent upon a rock; the way of a ship in the midst of the sea; and the way of a man with a maid.
\fn{midst: Heb. heart.}
\vs Pro 30:20 Such \bibemph{is} the way of an adulterous woman; she eateth, and wipeth her mouth, and saith, I have done no wickedness.
\vs Pro 30:21 For three \bibemph{things} the earth is disquieted, and for four \bibemph{which} it cannot bear:
\vs Pro 30:22 For a servant when he reigneth; and a fool when he is filled with meat;
\vs Pro 30:23 For an odious \bibemph{woman} when she is married; and an handmaid that is heir to her mistress.
\vs Pro 30:24 There be four \bibemph{things which are} little upon the earth, but they \bibemph{are} exceeding wise:
\fn{exceeding...: Heb. wise, made wise.}
\vs Pro 30:25 The ants \bibemph{are} a people not strong, yet they prepare their meat in the summer;
\vs Pro 30:26 The conies \bibemph{are but} a feeble folk, yet make they their houses in the rocks;
\vs Pro 30:27 The locusts have no king, yet go they forth all of them by bands;
\fn{by...: Heb. gathered together.}
\vs Pro 30:28 The spider taketh hold with her hands, and is in kings' palaces.
\vs Pro 30:29 There be three \bibemph{things} which go well, yea, four are comely in going:
\vs Pro 30:30 A lion \bibemph{which is} strongest among beasts, and turneth not away for any;
\vs Pro 30:31 A greyhound; an he goat also; and a king, against whom \bibemph{there is} no rising up.
\fn{greyhound: or, horse: Heb. girt in the loins.}
\vs Pro 30:32 If thou hast done foolishly in lifting up thyself, or if thou hast thought evil, \bibemph{lay} thine hand upon thy mouth.
\vs Pro 30:33 Surely the churning of milk bringeth forth butter, and the wringing of the nose bringeth forth blood: so the forcing of wrath bringeth forth strife.
\vs Pro 31:1 The words of king Lemuel, the prophecy that his mother taught him.
\vs Pro 31:2 What, my son? and what, the son of my womb? and what, the son of my vows?
\vs Pro 31:3 Give not thy strength unto women, nor thy ways to that which destroyeth kings.
\vs Pro 31:4 \bibemph{It is} not for kings, O Lemuel, \bibemph{it is} not for kings to drink wine; nor for princes strong drink:
\vs Pro 31:5 Lest they drink, and forget the law, and pervert the judgment of any of the afflicted.
\fn{prevert: Heb. alter. of any...: Heb. of all the sons of affliction.}
\vs Pro 31:6 Give strong drink unto him that is ready to perish, and wine unto those that be of heavy hearts.
\fn{of...: Heb. bitter of soul.}
\vs Pro 31:7 Let him drink, and forget his poverty, and remember his misery no more.
\vs Pro 31:8 Open thy mouth for the dumb in the cause of all such as are appointed to destruction.
\fn{such...: Heb. the sons of destruction.}
\vs Pro 31:9 Open thy mouth, judge righteously, and plead the cause of the poor and needy.
\vs Pro 31:10 Who can find a virtuous woman? for her price \bibemph{is} far above rubies.
\vs Pro 31:11 The heart of her husband doth safely trust in her, so that he shall have no need of spoil.
\vs Pro 31:12 She will do him good and not evil all the days of her life.
\vs Pro 31:13 She seeketh wool, and flax, and worketh willingly with her hands.
\vs Pro 31:14 She is like the merchants' ships; she bringeth her food from afar.
\vs Pro 31:15 She riseth also while it is yet night, and giveth meat to her household, and a portion to her maidens.
\vs Pro 31:16 She considereth a field, and buyeth it: with the fruit of her hands she planteth a vineyard.
\fn{buyeth: Heb. taketh.}
\vs Pro 31:17 She girdeth her loins with strength, and strengtheneth her arms.
\vs Pro 31:18 She perceiveth that her merchandise \bibemph{is} good: her candle goeth not out by night.
\fn{She...: Heb. She tasteth.}
\vs Pro 31:19 She layeth her hands to the spindle, and her hands hold the distaff.
\vs Pro 31:20 She stretcheth out her hand to the poor; yea, she reacheth forth her hands to the needy.
\fn{She...: Heb. She spreadeth.}
\vs Pro 31:21 She is not afraid of the snow for her household: for all her household \bibemph{are} clothed with scarlet.
\fn{scarlet: or, double garments.}
\vs Pro 31:22 She maketh herself coverings of tapestry; her clothing \bibemph{is} silk and purple.
\vs Pro 31:23 Her husband is known in the gates, when he sitteth among the elders of the land.
\vs Pro 31:24 She maketh fine linen, and selleth \bibemph{it}; and delivereth girdles unto the merchant.
\vs Pro 31:25 Strength and honour \bibemph{are} her clothing; and she shall rejoice in time to come.
\vs Pro 31:26 She openeth her mouth with wisdom; and in her tongue \bibemph{is} the law of kindness.
\vs Pro 31:27 She looketh well to the ways of her household, and eateth not the bread of idleness.
\vs Pro 31:28 Her children arise up, and call her blessed; her husband \bibemph{also}, and he praiseth her.
\vs Pro 31:29 Many daughters have done virtuously, but thou excellest them all.
\fn{have...: or, have gotten riches.}
\vs Pro 31:30 Favour \bibemph{is} deceitful, and beauty \bibemph{is} vain: \bibemph{but} a woman \bibemph{that} feareth the \textsc{Lord}, she shall be praised.
\vs Pro 31:31 Give her of the fruit of her hands; and let her own works praise her in the gates.
