\bibbook{the\\\bibtitlemain{book of psalms}}{psalms}{Psalms}{Psa}
\vs Psa 1:1 Blessed \bibemph{is} the man that walketh not in the counsel of the ungodly, nor standeth in the way of sinners, nor sitteth in the seat of the scornful.
\fn{ungodly: or, wicked.}
\vs Psa 1:2 But his delight \bibemph{is} in the law of the \textsc{Lord}; and in his law doth he meditate day and night.
\vs Psa 1:3 And he shall be like a tree planted by the rivers of water, that bringeth forth his fruit in his season; his leaf also shall not wither; and whatsoever he doeth shall prosper.
\fn{wither: Heb. fade.}
\vs Psa 1:4 The ungodly \bibemph{are} not so: but \bibemph{are} like the chaff which the wind driveth away.
\vs Psa 1:5 Therefore the ungodly shall not stand in the judgment, nor sinners in the congregation of the righteous.
\vs Psa 1:6 For the \textsc{Lord} knoweth the way of the righteous: but the way of the ungodly shall perish.
\vs Psa 2:1 Why do the heathen rage, and the people imagine a vain thing?
\fn{rage: or, tumultuously assemble. imagine: Heb. meditate.}
\vs Psa 2:2 The kings of the earth set themselves, and the rulers take counsel together, against the \textsc{Lord}, and against his anointed, \bibemph{saying},
\vs Psa 2:3 Let us break their bands asunder, and cast away their cords from us.
\vs Psa 2:4 He that sitteth in the heavens shall laugh: the Lord shall have them in derision.
\vs Psa 2:5 Then shall he speak unto them in his wrath, and vex them in his sore displeasure.
\fn{vex: or, trouble.}
\vs Psa 2:6 Yet have I set my king upon my holy hill of Zion.
\fn{set: Heb. anointed. upon...: Heb. upon Zion, the hill of my holiness.}
\vs Psa 2:7 I will declare the decree: the \textsc{Lord} hath said unto me, Thou \bibemph{art} my Son; this day have I begotten thee.
\fn{the decree: or, for a decree.}
\vs Psa 2:8 Ask of me, and I shall give \bibemph{thee} the heathen \bibemph{for} thine inheritance, and the uttermost parts of the earth \bibemph{for} thy possession.
\vs Psa 2:9 Thou shalt break them with a rod of iron; thou shalt dash them in pieces like a potter's vessel.
\vs Psa 2:10 Be wise now therefore, O ye kings: be instructed, ye judges of the earth.
\vs Psa 2:11 Serve the \textsc{Lord} with fear, and rejoice with trembling.
\vs Psa 2:12 Kiss the Son, lest he be angry, and ye perish \bibemph{from} the way, when his wrath is kindled but a little. Blessed \bibemph{are} all they that put their trust in him.
\vs Psa 3:1 A Psalm of David, when he fled from Absalom his son. \textsc{Lord}, how are they increased that trouble me! many \bibemph{are} they that rise up against me.
\vs Psa 3:2 Many \bibemph{there be} which say of my soul, \bibemph{There is} no help for him in God. Selah.
\vs Psa 3:3 But thou, O \textsc{Lord}, \bibemph{art} a shield for me; my glory, and the lifter up of mine head.
\fn{for: or, about.}
\vs Psa 3:4 I cried unto the \textsc{Lord} with my voice, and he heard me out of his holy hill. Selah.
\vs Psa 3:5 I laid me down and slept; I awaked; for the \textsc{Lord} sustained me.
\vs Psa 3:6 I will not be afraid of ten thousands of people, that have set \bibemph{themselves} against me round about.
\vs Psa 3:7 Arise, O \textsc{Lord}; save me, O my God: for thou hast smitten all mine enemies \bibemph{upon} the cheek bone; thou hast broken the teeth of the ungodly.
\vs Psa 3:8 Salvation \bibemph{belongeth} unto the \textsc{Lord}: thy blessing \bibemph{is} upon thy people. Selah.
\vs Psa 4:1 To the chief Musician on Neginoth, A Psalm of David. Hear me when I call, O God of my righteousness: thou hast enlarged me \bibemph{when I was} in distress; have mercy upon me, and hear my prayer.
\fn{chief...: or, overseer. have...: or, be gracious unto me.}
\vs Psa 4:2 O ye sons of men, how long \bibemph{will ye turn} my glory into shame? \bibemph{how long} will ye love vanity, \bibemph{and} seek after leasing? Selah.
\vs Psa 4:3 But know that the \textsc{Lord} hath set apart him that is godly for himself: the \textsc{Lord} will hear when I call unto him.
\vs Psa 4:4 Stand in awe, and sin not: commune with your own heart upon your bed, and be still. Selah.
\vs Psa 4:5 Offer the sacrifices of righteousness, and put your trust in the \textsc{Lord}.
\vs Psa 4:6 \bibemph{There be} many that say, Who will shew us \bibemph{any} good? \textsc{Lord}, lift thou up the light of thy countenance upon us.
\vs Psa 4:7 Thou hast put gladness in my heart, more than in the time \bibemph{that} their corn and their wine increased.
\vs Psa 4:8 I will both lay me down in peace, and sleep: for thou, \textsc{Lord}, only makest me dwell in safety.
\vs Psa 5:1 To the chief Musician upon Nehiloth, A Psalm of David. Give ear to my words, O \textsc{Lord}, consider my meditation.
\vs Psa 5:2 Hearken unto the voice of my cry, my King, and my God: for unto thee will I pray.
\vs Psa 5:3 My voice shalt thou hear in the morning, O \textsc{Lord}; in the morning will I direct \bibemph{my prayer} unto thee, and will look up.
\vs Psa 5:4 For thou \bibemph{art} not a God that hath pleasure in wickedness: neither shall evil dwell with thee.
\vs Psa 5:5 The foolish shall not stand in thy sight: thou hatest all workers of iniquity.
\fn{in...: Heb. before thine eyes.}
\vs Psa 5:6 Thou shalt destroy them that speak leasing: the \textsc{Lord} will abhor the bloody and deceitful man.
\fn{bloody...: Heb. man of bloods and deceit.}
\vs Psa 5:7 But as for me, I will come \bibemph{into} thy house in the multitude of thy mercy: \bibemph{and} in thy fear will I worship toward thy holy temple.
\fn{thy holy...: Heb. the temple of thy holiness.}
\vs Psa 5:8 Lead me, O \textsc{Lord}, in thy righteousness because of mine enemies; make thy way straight before my face.
\fn{mine...: Heb. those which observe me.}
\vs Psa 5:9 For \bibemph{there is} no faithfulness in their mouth; their inward part \bibemph{is} very wickedness; their throat \bibemph{is} an open sepulchre; they flatter with their tongue.
\fn{faithfulness: or, stedfastness. very...: Heb. wickednesses. their mouth: Heb. his mouth, that is, the mouth of any of them.}
\vs Psa 5:10 Destroy thou them, O God; let them fall by their own counsels; cast them out in the multitude of their transgressions; for they have rebelled against thee.
\fn{Destroy...: or, Make them guilty. by...: or, from their counsels.}
\vs Psa 5:11 But let all those that put their trust in thee rejoice: let them ever shout for joy, because thou defendest them: let them also that love thy name be joyful in thee.
\fn{defendest...: Heb. coverest over, or, protectest them.}
\vs Psa 5:12 For thou, \textsc{Lord}, wilt bless the righteous; with favour wilt thou compass him as \bibemph{with} a shield.
\fn{compass: Heb. crown.}
\vs Psa 6:1 To the chief Musician on Neginoth upon Sheminith, A Psalm of David. O \textsc{Lord}, rebuke me not in thine anger, neither chasten me in thy hot displeasure.
\fn{Sheminith: or, the eighth.}
\vs Psa 6:2 Have mercy upon me, O \textsc{Lord}; for I \bibemph{am} weak: O \textsc{Lord}, heal me; for my bones are vexed.
\vs Psa 6:3 My soul is also sore vexed: but thou, O \textsc{Lord}, how long?
\vs Psa 6:4 Return, O \textsc{Lord}, deliver my soul: oh save me for thy mercies' sake.
\vs Psa 6:5 For in death \bibemph{there is} no remembrance of thee: in the grave who shall give thee thanks?
\vs Psa 6:6 I am weary with my groaning; all the night make I my bed to swim; I water my couch with my tears.
\fn{all...: or, every night.}
\vs Psa 6:7 Mine eye is consumed because of grief; it waxeth old because of all mine enemies.
\vs Psa 6:8 Depart from me, all ye workers of iniquity; for the \textsc{Lord} hath heard the voice of my weeping.
\vs Psa 6:9 The \textsc{Lord} hath heard my supplication; the \textsc{Lord} will receive my prayer.
\vs Psa 6:10 Let all mine enemies be ashamed and sore vexed: let them return \bibemph{and} be ashamed suddenly.
\vs Psa 7:1 Shiggaion of David, which he sang unto the \textsc{Lord}, concerning the words of Cush the Benjamite. O \textsc{Lord} my God, in thee do I put my trust: save me from all them that persecute me, and deliver me:
\fn{words: or, business.}
\vs Psa 7:2 Lest he tear my soul like a lion, rending \bibemph{it} in pieces, while \bibemph{there is} none to deliver.
\fn{none...: Heb. not a deliverer.}
\vs Psa 7:3 O \textsc{Lord} my God, if I have done this; if there be iniquity in my hands;
\vs Psa 7:4 If I have rewarded evil unto him that was at peace with me; (yea, I have delivered him that without cause is mine enemy:)
\vs Psa 7:5 Let the enemy persecute my soul, and take \bibemph{it}; yea, let him tread down my life upon the earth, and lay mine honour in the dust. Selah.
\vs Psa 7:6 Arise, O \textsc{Lord}, in thine anger, lift up thyself because of the rage of mine enemies: and awake for me \bibemph{to} the judgment \bibemph{that} thou hast commanded.
\vs Psa 7:7 So shall the congregation of the people compass thee about: for their sakes therefore return thou on high.
\vs Psa 7:8 The \textsc{Lord} shall judge the people: judge me, O \textsc{Lord}, according to my righteousness, and according to mine integrity \bibemph{that is} in me.
\vs Psa 7:9 Oh let the wickedness of the wicked come to an end; but establish the just: for the righteous God trieth the hearts and reins.
\vs Psa 7:10 My defence \bibemph{is} of God, which saveth the upright in heart.
\fn{My...: Heb. My buckler is upon God.}
\vs Psa 7:11 God judgeth the righteous, and God is angry \bibemph{with the wicked} every day.
\fn{judgeth...: or, is a righteous judge.}
\vs Psa 7:12 If he turn not, he will whet his sword; he hath bent his bow, and made it ready.
\vs Psa 7:13 He hath also prepared for him the instruments of death; he ordaineth his arrows against the persecutors.
\vs Psa 7:14 Behold, he travaileth with iniquity, and hath conceived mischief, and brought forth falsehood.
\vs Psa 7:15 He made a pit, and digged it, and is fallen into the ditch \bibemph{which} he made.
\fn{He made a pit: Heb. He hath digged a pit.}
\vs Psa 7:16 His mischief shall return upon his own head, and his violent dealing shall come down upon his own pate.
\vs Psa 7:17 I will praise the \textsc{Lord} according to his righteousness: and will sing praise to the name of the \textsc{Lord} most high.
\vs Psa 8:1 To the chief Musician upon Gittith, A Psalm of David. O \textsc{Lord} our Lord, how excellent \bibemph{is} thy name in all the earth! who hast set thy glory above the heavens.
\vs Psa 8:2 Out of the mouth of babes and sucklings hast thou ordained strength because of thine enemies, that thou mightest still the enemy and the avenger.
\fn{ordained: Heb. founded.}
\vs Psa 8:3 When I consider thy heavens, the work of thy fingers, the moon and the stars, which thou hast ordained;
\vs Psa 8:4 What is man, that thou art mindful of him? and the son of man, that thou visitest him?
\vs Psa 8:5 For thou hast made him a little lower than the angels, and hast crowned him with glory and honour.
\vs Psa 8:6 Thou madest him to have dominion over the works of thy hands; thou hast put all \bibemph{things} under his feet:
\vs Psa 8:7 All sheep and oxen, yea, and the beasts of the field;
\fn{All...: Heb. Flocks and oxen all of them.}
\vs Psa 8:8 The fowl of the air, and the fish of the sea, \bibemph{and whatsoever} passeth through the paths of the seas.
\vs Psa 8:9 O \textsc{Lord} our Lord, how excellent \bibemph{is} thy name in all the earth!
\vs Psa 9:1 To the chief Musician upon Muthlabben, A Psalm of David. I will praise \bibemph{thee}, O \textsc{Lord}, with my whole heart; I will shew forth all thy marvellous works.
\vs Psa 9:2 I will be glad and rejoice in thee: I will sing praise to thy name, O thou most High.
\vs Psa 9:3 When mine enemies are turned back, they shall fall and perish at thy presence.
\vs Psa 9:4 For thou hast maintained my right and my cause; thou satest in the throne judging right.
\fn{maintained...: Heb. made my judgment. judging right: Heb. judging in righteousness.}
\vs Psa 9:5 Thou hast rebuked the heathen, thou hast destroyed the wicked, thou hast put out their name for ever and ever.
\vs Psa 9:6 O thou enemy, destructions are come to a perpetual end: and thou hast destroyed cities; their memorial is perished with them.
\fn{O thou...: or, The destructions of the enemy are come to a perpetual end: and their cities hast thou destroyed, etc.}
\vs Psa 9:7 But the \textsc{Lord} shall endure for ever: he hath prepared his throne for judgment.
\vs Psa 9:8 And he shall judge the world in righteousness, he shall minister judgment to the people in uprightness.
\vs Psa 9:9 The \textsc{Lord} also will be a refuge for the oppressed, a refuge in times of trouble.
\fn{a refuge: Heb. an high place.}
\vs Psa 9:10 And they that know thy name will put their trust in thee: for thou, \textsc{Lord}, hast not forsaken them that seek thee.
\vs Psa 9:11 Sing praises to the \textsc{Lord}, which dwelleth in Zion: declare among the people his doings.
\vs Psa 9:12 When he maketh inquisition for blood, he remembereth them: he forgetteth not the cry of the humble.
\fn{humble: or, afflicted.}
\vs Psa 9:13 Have mercy upon me, O \textsc{Lord}; consider my trouble \bibemph{which I suffer} of them that hate me, thou that liftest me up from the gates of death:
\vs Psa 9:14 That I may shew forth all thy praise in the gates of the daughter of Zion: I will rejoice in thy salvation.
\vs Psa 9:15 The heathen are sunk down in the pit \bibemph{that} they made: in the net which they hid is their own foot taken.
\vs Psa 9:16 The \textsc{Lord} is known \bibemph{by} the judgment \bibemph{which} he executeth: the wicked is snared in the work of his own hands. Higgaion. Selah.
\fn{Higgaion: that is, Meditation.}
\vs Psa 9:17 The wicked shall be turned into hell, \bibemph{and} all the nations that forget God.
\vs Psa 9:18 For the needy shall not alway be forgotten: the expectation of the poor shall \bibemph{not} perish for ever.
\vs Psa 9:19 Arise, O \textsc{Lord}; let not man prevail: let the heathen be judged in thy sight.
\vs Psa 9:20 Put them in fear, O \textsc{Lord}: \bibemph{that} the nations may know themselves \bibemph{to be but} men. Selah.
\vs Psa 10:1 Why standest thou afar off, O \textsc{Lord}? \bibemph{why} hidest thou \bibemph{thyself} in times of trouble?
\vs Psa 10:2 The wicked in \bibemph{his} pride doth persecute the poor: let them be taken in the devices that they have imagined.
\fn{The wicked...: Heb. In the pride of the wicked he doth persecute.}
\vs Psa 10:3 For the wicked boasteth of his heart's desire, and blesseth the covetous, \bibemph{whom} the \textsc{Lord} abhorreth.
\fn{heart's: Heb. soul's. blesseth...: or, the covetous blesseth himself, he abhorreth the \textsc{Lord}.}
\vs Psa 10:4 The wicked, through the pride of his countenance, will not seek \bibemph{after God}: God \bibemph{is} not in all his thoughts.
\fn{God is...: or, all his thoughts are, There is no God.}
\vs Psa 10:5 His ways are always grievous; thy judgments \bibemph{are} far above out of his sight: \bibemph{as for} all his enemies, he puffeth at them.
\vs Psa 10:6 He hath said in his heart, I shall not be moved: for \bibemph{I shall} never \bibemph{be} in adversity.
\fn{never: Heb. unto generation and generation.}
\vs Psa 10:7 His mouth is full of cursing and deceit and fraud: under his tongue \bibemph{is} mischief and vanity.
\fn{deceit: Heb. deceits. vanity: or, iniquity.}
\vs Psa 10:8 He sitteth in the lurking places of the villages: in the secret places doth he murder the innocent: his eyes are privily set against the poor.
\fn{are...: Heb. hide themselves.}
\vs Psa 10:9 He lieth in wait secretly as a lion in his den: he lieth in wait to catch the poor: he doth catch the poor, when he draweth him into his net.
\fn{secretly: Heb. in the secret places.}
\vs Psa 10:10 He croucheth, \bibemph{and} humbleth himself, that the poor may fall by his strong ones.
\fn{He...: Heb. He breaketh himself. by...: or, into his strong parts.}
\vs Psa 10:11 He hath said in his heart, God hath forgotten: he hideth his face; he will never see \bibemph{it}.
\vs Psa 10:12 Arise, O \textsc{Lord}; O God, lift up thine hand: forget not the humble.
\fn{humble: or, afflicted.}
\vs Psa 10:13 Wherefore doth the wicked contemn God? he hath said in his heart, Thou wilt not require \bibemph{it}.
\vs Psa 10:14 Thou hast seen \bibemph{it}; for thou beholdest mischief and spite, to requite \bibemph{it} with thy hand: the poor committeth himself unto thee; thou art the helper of the fatherless.
\fn{committeth: Heb. leaveth.}
\vs Psa 10:15 Break thou the arm of the wicked and the evil \bibemph{man}: seek out his wickedness \bibemph{till} thou find none.
\vs Psa 10:16 The \textsc{Lord} \bibemph{is} King for ever and ever: the heathen are perished out of his land.
\vs Psa 10:17 \textsc{Lord}, thou hast heard the desire of the humble: thou wilt prepare their heart, thou wilt cause thine ear to hear:
\fn{prepare: or, establish.}
\vs Psa 10:18 To judge the fatherless and the oppressed, that the man of the earth may no more oppress.
\fn{oppress: or, terrify.}
\vs Psa 11:1 To the chief Musician, \bibemph{A Psalm} of David. In the \textsc{Lord} put I my trust: how say ye to my soul, Flee \bibemph{as} a bird to your mountain?
\vs Psa 11:2 For, lo, the wicked bend \bibemph{their} bow, they make ready their arrow upon the string, that they may privily shoot at the upright in heart.
\fn{privily: Heb. in darkness.}
\vs Psa 11:3 If the foundations be destroyed, what can the righteous do?
\vs Psa 11:4 The \textsc{Lord} \bibemph{is} in his holy temple, the \textsc{Lord's} throne \bibemph{is} in heaven: his eyes behold, his eyelids try, the children of men.
\vs Psa 11:5 The \textsc{Lord} trieth the righteous: but the wicked and him that loveth violence his soul hateth.
\vs Psa 11:6 Upon the wicked he shall rain snares, fire and brimstone, and an horrible tempest: \bibemph{this shall be} the portion of their cup.
\fn{an horrible...: or, a burning tempest.}
\vs Psa 11:7 For the righteous \textsc{Lord} loveth righteousness; his countenance doth behold the upright.
\vs Psa 12:1 To the chief Musician upon Sheminith, A Psalm of David. Help, \textsc{Lord}; for the godly man ceaseth; for the faithful fail from among the children of men.
\fn{upon...: or, upon the eighth. Help: or, Save.}
\vs Psa 12:2 They speak vanity every one with his neighbour: \bibemph{with} flattering lips \bibemph{and} with a double heart do they speak.
\fn{a double...: Heb. an heart and an heart.}
\vs Psa 12:3 The \textsc{Lord} shall cut off all flattering lips, \bibemph{and} the tongue that speaketh proud things:
\fn{proud: Heb. great.}
\vs Psa 12:4 Who have said, With our tongue will we prevail; our lips \bibemph{are} our own: who \bibemph{is} lord over us?
\fn{are...: Heb. are with us.}
\vs Psa 12:5 For the oppression of the poor, for the sighing of the needy, now will I arise, saith the \textsc{Lord}; I will set \bibemph{him} in safety \bibemph{from him that} puffeth at him.
\fn{puffeth...: or, would ensnare him.}
\vs Psa 12:6 The words of the \textsc{Lord} \bibemph{are} pure words: \bibemph{as} silver tried in a furnace of earth, purified seven times.
\vs Psa 12:7 Thou shalt keep them, O \textsc{Lord}, thou shalt preserve them from this generation for ever.
\fn{them from: Heb. him, etc: that is, every one of them, etc.}
\vs Psa 12:8 The wicked walk on every side, when the vilest men are exalted.
\fn{men: Heb. of the sons of the men.}
\vs Psa 13:1 To the chief Musician, A Psalm of David. How long wilt thou forget me, O \textsc{Lord}? for ever? how long wilt thou hide thy face from me?
\fn{chief...: or, overseer.}
\vs Psa 13:2 How long shall I take counsel in my soul, \bibemph{having} sorrow in my heart daily? how long shall mine enemy be exalted over me?
\vs Psa 13:3 Consider \bibemph{and} hear me, O \textsc{Lord} my God: lighten mine eyes, lest I sleep the \bibemph{sleep of} death;
\vs Psa 13:4 Lest mine enemy say, I have prevailed against him; \bibemph{and} those that trouble me rejoice when I am moved.
\vs Psa 13:5 But I have trusted in thy mercy; my heart shall rejoice in thy salvation.
\vs Psa 13:6 I will sing unto the \textsc{Lord}, because he hath dealt bountifully with me.
\vs Psa 14:1 To the chief Musician, \bibemph{A Psalm} of David. The fool hath said in his heart, \bibemph{There is} no God. They are corrupt, they have done abominable works, \bibemph{there is} none that doeth good.
\vs Psa 14:2 The \textsc{Lord} looked down from heaven upon the children of men, to see if there were any that did understand, \bibemph{and} seek God.
\vs Psa 14:3 They are all gone aside, they are \bibemph{all} together become filthy: \bibemph{there is} none that doeth good, no, not one.
\fn{filthy: Heb. stinking.}
\vs Psa 14:4 Have all the workers of iniquity no knowledge? who eat up my people \bibemph{as} they eat bread, and call not upon the \textsc{Lord}.
\vs Psa 14:5 There were they in great fear: for God \bibemph{is} in the generation of the righteous.
\fn{were...: Heb. they feared a fear.}
\vs Psa 14:6 Ye have shamed the counsel of the poor, because the \textsc{Lord} \bibemph{is} his refuge.
\vs Psa 14:7 Oh that the salvation of Israel \bibemph{were come} out of Zion! when the \textsc{Lord} bringeth back the captivity of his people, Jacob shall rejoice, \bibemph{and} Israel shall be glad.
\fn{Oh...: Heb. Who will give.}
\vs Psa 15:1 A Psalm of David. \textsc{Lord}, who shall abide in thy tabernacle? who shall dwell in thy holy hill?
\fn{abide: Heb. sojourn.}
\vs Psa 15:2 He that walketh uprightly, and worketh righteousness, and speaketh the truth in his heart.
\vs Psa 15:3 \bibemph{He that} backbiteth not with his tongue, nor doeth evil to his neighbour, nor taketh up a reproach against his neighbour.
\fn{taketh...: or, receiveth, or, endureth.}
\vs Psa 15:4 In whose eyes a vile person is contemned; but he honoureth them that fear the \textsc{Lord}. \bibemph{He that} sweareth to \bibemph{his own} hurt, and changeth not.
\vs Psa 15:5 \bibemph{He that} putteth not out his money to usury, nor taketh reward against the innocent. He that doeth these \bibemph{things} shall never be moved.
\vs Psa 16:1 Michtam of David. Preserve me, O God: for in thee do I put my trust.
\fn{Michtam: or, A golden Psalm.}
\vs Psa 16:2 \bibemph{O my soul}, thou hast said unto the \textsc{Lord}, Thou \bibemph{art} my Lord: my goodness \bibemph{extendeth} not to thee;
\vs Psa 16:3 \bibemph{But} to the saints that \bibemph{are} in the earth, and \bibemph{to} the excellent, in whom \bibemph{is} all my delight.
\vs Psa 16:4 Their sorrows shall be multiplied \bibemph{that} hasten \bibemph{after} another \bibemph{god}: their drink offerings of blood will I not offer, nor take up their names into my lips.
\fn{hasten...: or, give gifts to another.}
\vs Psa 16:5 The \textsc{Lord} \bibemph{is} the portion of mine inheritance and of my cup: thou maintainest my lot.
\fn{of mine...: Heb. of my part.}
\vs Psa 16:6 The lines are fallen unto me in pleasant \bibemph{places}; yea, I have a goodly heritage.
\vs Psa 16:7 I will bless the \textsc{Lord}, who hath given me counsel: my reins also instruct me in the night seasons.
\vs Psa 16:8 I have set the \textsc{Lord} always before me: because \bibemph{he is} at my right hand, I shall not be moved.
\vs Psa 16:9 Therefore my heart is glad, and my glory rejoiceth: my flesh also shall rest in hope.
\fn{rest...: Heb. dwell confidently.}
\vs Psa 16:10 For thou wilt not leave my soul in hell; neither wilt thou suffer thine Holy One to see corruption.
\vs Psa 16:11 Thou wilt shew me the path of life: in thy presence \bibemph{is} fulness of joy; at thy right hand \bibemph{there are} pleasures for evermore.
\vs Psa 17:1 A Prayer of David. Hear the right, O \textsc{Lord}, attend unto my cry, give ear unto my prayer, \bibemph{that goeth} not out of feigned lips.
\fn{the right: Heb. justice. not...: Heb. without lips of deceit.}
\vs Psa 17:2 Let my sentence come forth from thy presence; let thine eyes behold the things that are equal.
\vs Psa 17:3 Thou hast proved mine heart; thou hast visited \bibemph{me} in the night; thou hast tried me, \bibemph{and} shalt find nothing; I am purposed \bibemph{that} my mouth shall not transgress.
\vs Psa 17:4 Concerning the works of men, by the word of thy lips I have kept \bibemph{me from} the paths of the destroyer.
\vs Psa 17:5 Hold up my goings in thy paths, \bibemph{that} my footsteps slip not.
\fn{slip...: Heb. be not moved.}
\vs Psa 17:6 I have called upon thee, for thou wilt hear me, O God: incline thine ear unto me, \bibemph{and hear} my speech.
\vs Psa 17:7 Shew thy marvellous lovingkindness, O thou that savest by thy right hand them which put their trust \bibemph{in thee} from those that rise up \bibemph{against them}.
\fn{by thy...: or, them which trust in thee from those that rise up against thy right hand.}
\vs Psa 17:8 Keep me as the apple of the eye, hide me under the shadow of thy wings,
\vs Psa 17:9 From the wicked that oppress me, \bibemph{from} my deadly enemies, \bibemph{who} compass me about.
\fn{oppress: Heb. waste. my...: Heb. my enemies against the soul.}
\vs Psa 17:10 They are inclosed in their own fat: with their mouth they speak proudly.
\vs Psa 17:11 They have now compassed us in our steps: they have set their eyes bowing down to the earth;
\vs Psa 17:12 Like as a lion \bibemph{that} is greedy of his prey, and as it were a young lion lurking in secret places.
\fn{Like...: Heb. The likeness of him (that is, of every one of them) is as a lion that desireth to ravin. lurking: Heb. sitting.}
\vs Psa 17:13 Arise, O \textsc{Lord}, disappoint him, cast him down: deliver my soul from the wicked, \bibemph{which is} thy sword:
\fn{disappoint...: Heb. prevent his face. which is: or, by.}
\vs Psa 17:14 From men \bibemph{which are} thy hand, O \textsc{Lord}, from men of the world, \bibemph{which have} their portion in \bibemph{this} life, and whose belly thou fillest with thy hid \bibemph{treasure}: they are full of children, and leave the rest of their \bibemph{substance} to their babes.
\fn{which are: or, by. they...: or, their children are full.}
\vs Psa 17:15 As for me, I will behold thy face in righteousness: I shall be satisfied, when I awake, with thy likeness.
\vs Psa 18:1 To the chief Musician, \bibemph{A Psalm} of David, the servant of the \textsc{Lord}, who spake unto the \textsc{Lord} the words of this song in the day \bibemph{that} the \textsc{Lord} delivered him from the hand of all his enemies, and from the hand of Saul: And he said, I will love thee, O \textsc{Lord}, my strength.
\vs Psa 18:2 The \textsc{Lord} \bibemph{is} my rock, and my fortress, and my deliverer; my God, my strength, in whom I will trust; my buckler, and the horn of my salvation, \bibemph{and} my high tower.
\fn{my strength: Heb. my rock.}
\vs Psa 18:3 I will call upon the \textsc{Lord}, \bibemph{who is worthy} to be praised: so shall I be saved from mine enemies.
\vs Psa 18:4 The sorrows of death compassed me, and the floods of ungodly men made me afraid.
\fn{ungodly men: Heb. Belial.}
\vs Psa 18:5 The sorrows of hell compassed me about: the snares of death prevented me.
\fn{sorrows: or, cords.}
\vs Psa 18:6 In my distress I called upon the \textsc{Lord}, and cried unto my God: he heard my voice out of his temple, and my cry came before him, \bibemph{even} into his ears.
\vs Psa 18:7 Then the earth shook and trembled; the foundations also of the hills moved and were shaken, because he was wroth.
\vs Psa 18:8 There went up a smoke out of his nostrils, and fire out of his mouth devoured: coals were kindled by it.
\fn{out of his nostrils: Heb. by his, etc.}
\vs Psa 18:9 He bowed the heavens also, and came down: and darkness \bibemph{was} under his feet.
\vs Psa 18:10 And he rode upon a cherub, and did fly: yea, he did fly upon the wings of the wind.
\vs Psa 18:11 He made darkness his secret place; his pavilion round about him \bibemph{were} dark waters \bibemph{and} thick clouds of the skies.
\vs Psa 18:12 At the brightness \bibemph{that was} before him his thick clouds passed, hail \bibemph{stones} and coals of fire.
\vs Psa 18:13 The \textsc{Lord} also thundered in the heavens, and the Highest gave his voice; hail \bibemph{stones} and coals of fire.
\vs Psa 18:14 Yea, he sent out his arrows, and scattered them; and he shot out lightnings, and discomfited them.
\vs Psa 18:15 Then the channels of waters were seen, and the foundations of the world were discovered at thy rebuke, O \textsc{Lord}, at the blast of the breath of thy nostrils.
\vs Psa 18:16 He sent from above, he took me, he drew me out of many waters.
\fn{many: or, great.}
\vs Psa 18:17 He delivered me from my strong enemy, and from them which hated me: for they were too strong for me.
\vs Psa 18:18 They prevented me in the day of my calamity: but the \textsc{Lord} was my stay.
\vs Psa 18:19 He brought me forth also into a large place; he delivered me, because he delighted in me.
\vs Psa 18:20 The \textsc{Lord} rewarded me according to my righteousness; according to the cleanness of my hands hath he recompensed me.
\vs Psa 18:21 For I have kept the ways of the \textsc{Lord}, and have not wickedly departed from my God.
\vs Psa 18:22 For all his judgments \bibemph{were} before me, and I did not put away his statutes from me.
\vs Psa 18:23 I was also upright before him, and I kept myself from mine iniquity.
\fn{before: Heb. with.}
\vs Psa 18:24 Therefore hath the \textsc{Lord} recompensed me according to my righteousness, according to the cleanness of my hands in his eyesight.
\fn{in...: Heb. before his eyes.}
\vs Psa 18:25 With the merciful thou wilt shew thyself merciful; with an upright man thou wilt shew thyself upright;
\vs Psa 18:26 With the pure thou wilt shew thyself pure; and with the froward thou wilt shew thyself froward.
\fn{shew thyself froward: or, wrestle.}
\vs Psa 18:27 For thou wilt save the afflicted people; but wilt bring down high looks.
\vs Psa 18:28 For thou wilt light my candle: the \textsc{Lord} my God will enlighten my darkness.
\fn{candle: or, lamp.}
\vs Psa 18:29 For by thee I have run through a troop; and by my God have I leaped over a wall.
\fn{run...: or, broken.}
\vs Psa 18:30 \bibemph{As for} God, his way \bibemph{is} perfect: the word of the \textsc{Lord} is tried: he \bibemph{is} a buckler to all those that trust in him.
\fn{tried: or, refined.}
\vs Psa 18:31 For who \bibemph{is} God save the \textsc{Lord}? or who \bibemph{is} a rock save our God?
\vs Psa 18:32 \bibemph{It is} God that girdeth me with strength, and maketh my way perfect.
\vs Psa 18:33 He maketh my feet like hinds' \bibemph{feet}, and setteth me upon my high places.
\vs Psa 18:34 He teacheth my hands to war, so that a bow of steel is broken by mine arms.
\vs Psa 18:35 Thou hast also given me the shield of thy salvation: and thy right hand hath holden me up, and thy gentleness hath made me great.
\fn{thy gentleness...: or, with thy meekness thou hast multiplied me.}
\vs Psa 18:36 Thou hast enlarged my steps under me, that my feet did not slip.
\fn{my feet: Heb. mine ankles.}
\vs Psa 18:37 I have pursued mine enemies, and overtaken them: neither did I turn again till they were consumed.
\vs Psa 18:38 I have wounded them that they were not able to rise: they are fallen under my feet.
\vs Psa 18:39 For thou hast girded me with strength unto the battle: thou hast subdued under me those that rose up against me.
\fn{subdued: Heb. caused to bow.}
\vs Psa 18:40 Thou hast also given me the necks of mine enemies; that I might destroy them that hate me.
\vs Psa 18:41 They cried, but \bibemph{there was} none to save \bibemph{them: even} unto the \textsc{Lord}, but he answered them not.
\vs Psa 18:42 Then did I beat them small as the dust before the wind: I did cast them out as the dirt in the streets.
\vs Psa 18:43 Thou hast delivered me from the strivings of the people; \bibemph{and} thou hast made me the head of the heathen: a people \bibemph{whom} I have not known shall serve me.
\vs Psa 18:44 As soon as they hear of me, they shall obey me: the strangers shall submit themselves unto me.
\fn{As soon...: Heb. At the hearing of the ear. submit...: or, yield feigned obedience. strangers: Heb. sons of the stranger.}
\vs Psa 18:45 The strangers shall fade away, and be afraid out of their close places.
\fn{strangers: Heb. sons of the stranger.}
\vs Psa 18:46 The \textsc{Lord} liveth; and blessed \bibemph{be} my rock; and let the God of my salvation be exalted.
\vs Psa 18:47 \bibemph{It is} God that avengeth me, and subdueth the people under me.
\fn{avengeth: Heb. giveth avengements for. subdueth: or, destroyeth.}
\vs Psa 18:48 He delivereth me from mine enemies: yea, thou liftest me up above those that rise up against me: thou hast delivered me from the violent man.
\fn{violent...: Heb. man of violence.}
\vs Psa 18:49 Therefore will I give thanks unto thee, O \textsc{Lord}, among the heathen, and sing praises unto thy name.
\fn{give...: or, confess.}
\vs Psa 18:50 Great deliverance giveth he to his king; and sheweth mercy to his anointed, to David, and to his seed for evermore.
\vs Psa 19:1 To the chief Musician, A Psalm of David. The heavens declare the glory of God; and the firmament sheweth his handywork.
\vs Psa 19:2 Day unto day uttereth speech, and night unto night sheweth knowledge.
\vs Psa 19:3 \bibemph{There is} no speech nor language, \bibemph{where} their voice is not heard.
\fn{where...: or, without these their voice is heard: Heb. without their voice heard.}
\vs Psa 19:4 Their line is gone out through all the earth, and their words to the end of the world. In them hath he set a tabernacle for the sun,
\fn{line: or, rule, or, direction.}
\vs Psa 19:5 Which \bibemph{is} as a bridegroom coming out of his chamber, \bibemph{and} rejoiceth as a strong man to run a race.
\vs Psa 19:6 His going forth \bibemph{is} from the end of the heaven, and his circuit unto the ends of it: and there is nothing hid from the heat thereof.
\vs Psa 19:7 The law of the \textsc{Lord} \bibemph{is} perfect, converting the soul: the testimony of the \textsc{Lord} \bibemph{is} sure, making wise the simple.
\fn{law: or, doctrine. converting: or, restoring.}
\vs Psa 19:8 The statutes of the \textsc{Lord} \bibemph{are} right, rejoicing the heart: the commandment of the \textsc{Lord} \bibemph{is} pure, enlightening the eyes.
\vs Psa 19:9 The fear of the \textsc{Lord} \bibemph{is} clean, enduring for ever: the judgments of the \textsc{Lord} \bibemph{are} true \bibemph{and} righteous altogether.
\fn{true: Heb. truth.}
\vs Psa 19:10 More to be desired \bibemph{are they} than gold, yea, than much fine gold: sweeter also than honey and the honeycomb.
\fn{the honeycomb: Heb. the dropping of honeycomb.}
\vs Psa 19:11 Moreover by them is thy servant warned: \bibemph{and} in keeping of them \bibemph{there is} great reward.
\vs Psa 19:12 Who can understand \bibemph{his} errors? cleanse thou me from secret \bibemph{faults}.
\vs Psa 19:13 Keep back thy servant also from presumptuous \bibemph{sins}; let them not have dominion over me: then shall I be upright, and I shall be innocent from the great transgression.
\fn{the great: or, much.}
\vs Psa 19:14 Let the words of my mouth, and the meditation of my heart, be acceptable in thy sight, O \textsc{Lord}, my strength, and my redeemer.
\fn{strength: Heb. rock.}
\vs Psa 20:1 To the chief Musician, A Psalm of David. The \textsc{Lord} hear thee in the day of trouble; the name of the God of Jacob defend thee;
\fn{defend...: Heb. set thee on an high place.}
\vs Psa 20:2 Send thee help from the sanctuary, and strengthen thee out of Zion;
\fn{thee help: Heb. thy help. strengthen: Heb. support.}
\vs Psa 20:3 Remember all thy offerings, and accept thy burnt sacrifice; Selah.
\fn{accept: Heb. turn to ashes: or, make fat.}
\vs Psa 20:4 Grant thee according to thine own heart, and fulfil all thy counsel.
\vs Psa 20:5 We will rejoice in thy salvation, and in the name of our God we will set up \bibemph{our} banners: the \textsc{Lord} fulfil all thy petitions.
\vs Psa 20:6 Now know I that the \textsc{Lord} saveth his anointed; he will hear him from his holy heaven with the saving strength of his right hand.
\fn{his holy...: Heb. the heaven of his holiness. with...: Heb. by the strength of the salvation of.}
\vs Psa 20:7 Some \bibemph{trust} in chariots, and some in horses: but we will remember the name of the \textsc{Lord} our God.
\vs Psa 20:8 They are brought down and fallen: but we are risen, and stand upright.
\vs Psa 20:9 Save, \textsc{Lord}: let the king hear us when we call.
\vs Psa 21:1 To the chief Musician, A Psalm of David. The king shall joy in thy strength, O \textsc{Lord}; and in thy salvation how greatly shall he rejoice!
\vs Psa 21:2 Thou hast given him his heart's desire, and hast not withholden the request of his lips. Selah.
\vs Psa 21:3 For thou preventest him with the blessings of goodness: thou settest a crown of pure gold on his head.
\vs Psa 21:4 He asked life of thee, \bibemph{and} thou gavest \bibemph{it} him, \bibemph{even} length of days for ever and ever.
\vs Psa 21:5 His glory \bibemph{is} great in thy salvation: honour and majesty hast thou laid upon him.
\vs Psa 21:6 For thou hast made him most blessed for ever: thou hast made him exceeding glad with thy countenance.
\fn{made him most...: Heb. set him to be blessings. made him exceeding...: Heb. make him glad with joy.}
\vs Psa 21:7 For the king trusteth in the \textsc{Lord}, and through the mercy of the most High he shall not be moved.
\vs Psa 21:8 Thine hand shall find out all thine enemies: thy right hand shall find out those that hate thee.
\vs Psa 21:9 Thou shalt make them as a fiery oven in the time of thine anger: the \textsc{Lord} shall swallow them up in his wrath, and the fire shall devour them.
\vs Psa 21:10 Their fruit shalt thou destroy from the earth, and their seed from among the children of men.
\vs Psa 21:11 For they intended evil against thee: they imagined a mischievous device, \bibemph{which} they are not able \bibemph{to perform}.
\vs Psa 21:12 Therefore shalt thou make them turn their back, \bibemph{when} thou shalt make ready \bibemph{thine arrows} upon thy strings against the face of them.
\fn{shalt thou...: or, thou shalt set them as a butt. back: Heb. shoulder.}
\vs Psa 21:13 Be thou exalted, \textsc{Lord}, in thine own strength: \bibemph{so} will we sing and praise thy power.
\vs Psa 22:1 To the chief Musician upon Aijeleth Shahar, A Psalm of David. My God, my God, why hast thou forsaken me? \bibemph{why art thou so} far from helping me, \bibemph{and from} the words of my roaring?
\fn{Aijeleth...: or, the hind of the morning. helping...: Heb. my salvation.}
\vs Psa 22:2 O my God, I cry in the daytime, but thou hearest not; and in the night season, and am not silent.
\fn{am...: Heb. there is no silence to me.}
\vs Psa 22:3 But thou \bibemph{art} holy, \bibemph{O thou} that inhabitest the praises of Israel.
\vs Psa 22:4 Our fathers trusted in thee: they trusted, and thou didst deliver them.
\vs Psa 22:5 They cried unto thee, and were delivered: they trusted in thee, and were not confounded.
\vs Psa 22:6 But I \bibemph{am} a worm, and no man; a reproach of men, and despised of the people.
\vs Psa 22:7 All they that see me laugh me to scorn: they shoot out the lip, they shake the head, \bibemph{saying},
\fn{shoot...: Heb. open.}
\vs Psa 22:8 He trusted on the \textsc{Lord} \bibemph{that} he would deliver him: let him deliver him, seeing he delighted in him.
\fn{He trusted...: Heb. He rolled himself on. seeing...: or, if he delight in.}
\vs Psa 22:9 But thou \bibemph{art} he that took me out of the womb: thou didst make me hope \bibemph{when I was} upon my mother's breasts.
\fn{didst...: or, kept me in safety.}
\vs Psa 22:10 I was cast upon thee from the womb: thou \bibemph{art} my God from my mother's belly.
\vs Psa 22:11 Be not far from me; for trouble \bibemph{is} near; for \bibemph{there is} none to help.
\fn{none...: Heb. not a helper.}
\vs Psa 22:12 Many bulls have compassed me: strong \bibemph{bulls} of Bashan have beset me round.
\vs Psa 22:13 They gaped upon me \bibemph{with} their mouths, \bibemph{as} a ravening and a roaring lion.
\fn{gaped...: Heb. opened their mouths against me.}
\vs Psa 22:14 I am poured out like water, and all my bones are out of joint: my heart is like wax; it is melted in the midst of my bowels.
\fn{out of...: or, sundered.}
\vs Psa 22:15 My strength is dried up like a potsherd; and my tongue cleaveth to my jaws; and thou hast brought me into the dust of death.
\vs Psa 22:16 For dogs have compassed me: the assembly of the wicked have inclosed me: they pierced my hands and my feet.
\vs Psa 22:17 I may tell all my bones: they look \bibemph{and} stare upon me.
\vs Psa 22:18 They part my garments among them, and cast lots upon my vesture.
\vs Psa 22:19 But be not thou far from me, O \textsc{Lord}: O my strength, haste thee to help me.
\vs Psa 22:20 Deliver my soul from the sword; my darling from the power of the dog.
\fn{my darling: Heb. my only one. power: Heb. hand.}
\vs Psa 22:21 Save me from the lion's mouth: for thou hast heard me from the horns of the unicorns.
\vs Psa 22:22 I will declare thy name unto my brethren: in the midst of the congregation will I praise thee.
\vs Psa 22:23 Ye that fear the \textsc{Lord}, praise him; all ye the seed of Jacob, glorify him; and fear him, all ye the seed of Israel.
\vs Psa 22:24 For he hath not despised nor abhorred the affliction of the afflicted; neither hath he hid his face from him; but when he cried unto him, he heard.
\vs Psa 22:25 My praise \bibemph{shall be} of thee in the great congregation: I will pay my vows before them that fear him.
\vs Psa 22:26 The meek shall eat and be satisfied: they shall praise the \textsc{Lord} that seek him: your heart shall live for ever.
\vs Psa 22:27 All the ends of the world shall remember and turn unto the \textsc{Lord}: and all the kindreds of the nations shall worship before thee.
\vs Psa 22:28 For the kingdom \bibemph{is} the \textsc{Lord's}: and he \bibemph{is} the governor among the nations.
\vs Psa 22:29 All \bibemph{they that be} fat upon earth shall eat and worship: all they that go down to the dust shall bow before him: and none can keep alive his own soul.
\vs Psa 22:30 A seed shall serve him; it shall be accounted to the Lord for a generation.
\vs Psa 22:31 They shall come, and shall declare his righteousness unto a people that shall be born, that he hath done \bibemph{this}.
\vs Psa 23:1 A Psalm of David. The \textsc{Lord} \bibemph{is} my shepherd; I shall not want.
\vs Psa 23:2 He maketh me to lie down in green pastures: he leadeth me beside the still waters.
\fn{green...: Heb. pastures of tender grass. still...: Heb. waters of quietness.}
\vs Psa 23:3 He restoreth my soul: he leadeth me in the paths of righteousness for his name's sake.
\vs Psa 23:4 Yea, though I walk through the valley of the shadow of death, I will fear no evil: for thou \bibemph{art} with me; thy rod and thy staff they comfort me.
\vs Psa 23:5 Thou preparest a table before me in the presence of mine enemies: thou anointest my head with oil; my cup runneth over.
\fn{anointest: Heb. makest fat.}
\vs Psa 23:6 Surely goodness and mercy shall follow me all the days of my life: and I will dwell in the house of the \textsc{Lord} for ever.
\fn{for ever: Heb. to length of days.}
\vs Psa 24:1 A Psalm of David. The earth \bibemph{is} the \textsc{Lord's}, and the fulness thereof; the world, and they that dwell therein.
\vs Psa 24:2 For he hath founded it upon the seas, and established it upon the floods.
\vs Psa 24:3 Who shall ascend into the hill of the \textsc{Lord}? or who shall stand in his holy place?
\vs Psa 24:4 He that hath clean hands, and a pure heart; who hath not lifted up his soul unto vanity, nor sworn deceitfully.
\fn{He...: Heb. The clean of hands.}
\vs Psa 24:5 He shall receive the blessing from the \textsc{Lord}, and righteousness from the God of his salvation.
\vs Psa 24:6 This \bibemph{is} the generation of them that seek him, that seek thy face, O Jacob. Selah.
\fn{O Jacob: or, O God of Jacob.}
\vs Psa 24:7 Lift up your heads, O ye gates; and be ye lift up, ye everlasting doors; and the King of glory shall come in.
\vs Psa 24:8 Who \bibemph{is} this King of glory? The \textsc{Lord} strong and mighty, the \textsc{Lord} mighty in battle.
\vs Psa 24:9 Lift up your heads, O ye gates; even lift \bibemph{them} up, ye everlasting doors; and the King of glory shall come in.
\vs Psa 24:10 Who is this King of glory? The \textsc{Lord} of hosts, he \bibemph{is} the King of glory. Selah.
\vs Psa 25:1 \bibemph{A Psalm} of David. Unto thee, O \textsc{Lord}, do I lift up my soul.
\vs Psa 25:2 O my God, I trust in thee: let me not be ashamed, let not mine enemies triumph over me.
\vs Psa 25:3 Yea, let none that wait on thee be ashamed: let them be ashamed which transgress without cause.
\vs Psa 25:4 Shew me thy ways, O \textsc{Lord}; teach me thy paths.
\vs Psa 25:5 Lead me in thy truth, and teach me: for thou \bibemph{art} the God of my salvation; on thee do I wait all the day.
\vs Psa 25:6 Remember, O \textsc{Lord}, thy tender mercies and thy lovingkindnesses; for they \bibemph{have been} ever of old.
\fn{tender...: Heb. bowels.}
\vs Psa 25:7 Remember not the sins of my youth, nor my transgressions: according to thy mercy remember thou me for thy goodness' sake, O \textsc{Lord}.
\vs Psa 25:8 Good and upright \bibemph{is} the \textsc{Lord}: therefore will he teach sinners in the way.
\vs Psa 25:9 The meek will he guide in judgment: and the meek will he teach his way.
\vs Psa 25:10 All the paths of the \textsc{Lord} \bibemph{are} mercy and truth unto such as keep his covenant and his testimonies.
\vs Psa 25:11 For thy name's sake, O \textsc{Lord}, pardon mine iniquity; for it \bibemph{is} great.
\vs Psa 25:12 What man \bibemph{is} he that feareth the \textsc{Lord}? him shall he teach in the way \bibemph{that} he shall choose.
\vs Psa 25:13 His soul shall dwell at ease; and his seed shall inherit the earth.
\fn{dwell...: Heb. lodge in goodness.}
\vs Psa 25:14 The secret of the \textsc{Lord} \bibemph{is} with them that fear him; and he will shew them his covenant.
\fn{and...: or, and his covenant to make them know it.}
\vs Psa 25:15 Mine eyes \bibemph{are} ever toward the \textsc{Lord}; for he shall pluck my feet out of the net.
\fn{pluck: Heb. bring forth.}
\vs Psa 25:16 Turn thee unto me, and have mercy upon me; for I \bibemph{am} desolate and afflicted.
\vs Psa 25:17 The troubles of my heart are enlarged: \bibemph{O} bring thou me out of my distresses.
\vs Psa 25:18 Look upon mine affliction and my pain; and forgive all my sins.
\vs Psa 25:19 Consider mine enemies; for they are many; and they hate me with cruel hatred.
\fn{cruel...: Heb. hatred of violence.}
\vs Psa 25:20 O keep my soul, and deliver me: let me not be ashamed; for I put my trust in thee.
\vs Psa 25:21 Let integrity and uprightness preserve me; for I wait on thee.
\vs Psa 25:22 Redeem Israel, O God, out of all his troubles.
\vs Psa 26:1 \bibemph{A Psalm} of David. Judge me, O \textsc{Lord}; for I have walked in mine integrity: I have trusted also in the \textsc{Lord}; \bibemph{therefore} I shall not slide.
\vs Psa 26:2 Examine me, O \textsc{Lord}, and prove me; try my reins and my heart.
\vs Psa 26:3 For thy lovingkindness \bibemph{is} before mine eyes: and I have walked in thy truth.
\vs Psa 26:4 I have not sat with vain persons, neither will I go in with dissemblers.
\vs Psa 26:5 I have hated the congregation of evil doers; and will not sit with the wicked.
\vs Psa 26:6 I will wash mine hands in innocency: so will I compass thine altar, O \textsc{Lord}:
\vs Psa 26:7 That I may publish with the voice of thanksgiving, and tell of all thy wondrous works.
\vs Psa 26:8 \textsc{Lord}, I have loved the habitation of thy house, and the place where thine honour dwelleth.
\fn{where...: Heb. of the tabernacle of thy honour.}
\vs Psa 26:9 Gather not my soul with sinners, nor my life with bloody men:
\fn{Gather...: or, Take not away. bloody...: Heb. men of blood.}
\vs Psa 26:10 In whose hands \bibemph{is} mischief, and their right hand is full of bribes.
\fn{full...: Heb. filled with.}
\vs Psa 26:11 But as for me, I will walk in mine integrity: redeem me, and be merciful unto me.
\vs Psa 26:12 My foot standeth in an even place: in the congregations will I bless the \textsc{Lord}.
\vs Psa 27:1 \bibemph{A Psalm} of David. The \textsc{Lord} \bibemph{is} my light and my salvation; whom shall I fear? the \textsc{Lord} \bibemph{is} the strength of my life; of whom shall I be afraid?
\vs Psa 27:2 When the wicked, \bibemph{even} mine enemies and my foes, came upon me to eat up my flesh, they stumbled and fell.
\fn{came...: Heb. approached against me.}
\vs Psa 27:3 Though an host should encamp against me, my heart shall not fear: though war should rise against me, in this \bibemph{will} I \bibemph{be} confident.
\vs Psa 27:4 One \bibemph{thing} have I desired of the \textsc{Lord}, that will I seek after; that I may dwell in the house of the \textsc{Lord} all the days of my life, to behold the beauty of the \textsc{Lord}, and to enquire in his temple.
\fn{the beauty: or, the delight.}
\vs Psa 27:5 For in the time of trouble he shall hide me in his pavilion: in the secret of his tabernacle shall he hide me; he shall set me up upon a rock.
\vs Psa 27:6 And now shall mine head be lifted up above mine enemies round about me: therefore will I offer in his tabernacle sacrifices of joy; I will sing, yea, I will sing praises unto the \textsc{Lord}.
\fn{joy: Heb. shouting.}
\vs Psa 27:7 Hear, O \textsc{Lord}, \bibemph{when} I cry with my voice: have mercy also upon me, and answer me.
\vs Psa 27:8 \bibemph{When thou saidst}, Seek ye my face; my heart said unto thee, Thy face, \textsc{Lord}, will I seek.
\fn{When...: or, My heart said unto thee, Let my face seek thy face, etc.}
\vs Psa 27:9 Hide not thy face \bibemph{far} from me; put not thy servant away in anger: thou hast been my help; leave me not, neither forsake me, O God of my salvation.
\vs Psa 27:10 When my father and my mother forsake me, then the \textsc{Lord} will take me up.
\fn{take...: Heb. gather me.}
\vs Psa 27:11 Teach me thy way, O \textsc{Lord}, and lead me in a plain path, because of mine enemies.
\fn{a plain...: Heb. a way of plainness. mine...: Heb. those which observe me.}
\vs Psa 27:12 Deliver me not over unto the will of mine enemies: for false witnesses are risen up against me, and such as breathe out cruelty.
\vs Psa 27:13 \bibemph{I had fainted}, unless I had believed to see the goodness of the \textsc{Lord} in the land of the living.
\vs Psa 27:14 Wait on the \textsc{Lord}: be of good courage, and he shall strengthen thine heart: wait, I say, on the \textsc{Lord}.
\vs Psa 28:1 \bibemph{A Psalm} of David. Unto thee will I cry, O \textsc{Lord} my rock; be not silent to me: lest, \bibemph{if} thou be silent to me, I become like them that go down into the pit.
\fn{to me: Heb. from me.}
\vs Psa 28:2 Hear the voice of my supplications, when I cry unto thee, when I lift up my hands toward thy holy oracle.
\fn{thy...: or, the oracle of thy sanctuary.}
\vs Psa 28:3 Draw me not away with the wicked, and with the workers of iniquity, which speak peace to their neighbours, but mischief \bibemph{is} in their hearts.
\vs Psa 28:4 Give them according to their deeds, and according to the wickedness of their endeavours: give them after the work of their hands; render to them their desert.
\vs Psa 28:5 Because they regard not the works of the \textsc{Lord}, nor the operation of his hands, he shall destroy them, and not build them up.
\vs Psa 28:6 Blessed \bibemph{be} the \textsc{Lord}, because he hath heard the voice of my supplications.
\vs Psa 28:7 The \textsc{Lord} \bibemph{is} my strength and my shield; my heart trusted in him, and I am helped: therefore my heart greatly rejoiceth; and with my song will I praise him.
\vs Psa 28:8 The \textsc{Lord} \bibemph{is} their strength, and he \bibemph{is} the saving strength of his anointed.
\fn{their: or, his. saving...: Heb. strength of salvations.}
\vs Psa 28:9 Save thy people, and bless thine inheritance: feed them also, and lift them up for ever.
\fn{feed: or, rule.}
\vs Psa 29:1 A Psalm of David. Give unto the \textsc{Lord}, O ye mighty, give unto the \textsc{Lord} glory and strength.
\fn{ye mighty: Heb. ye sons of the mighty.}
\vs Psa 29:2 Give unto the \textsc{Lord} the glory due unto his name; worship the \textsc{Lord} in the beauty of holiness.
\fn{the glory...: Heb. the honour of his name. in...: or, in his glorious sanctuary.}
\vs Psa 29:3 The voice of the \textsc{Lord} \bibemph{is} upon the waters: the God of glory thundereth: the \textsc{Lord} \bibemph{is} upon many waters.
\fn{many: or, great.}
\vs Psa 29:4 The voice of the \textsc{Lord} \bibemph{is} powerful; the voice of the \textsc{Lord} \bibemph{is} full of majesty.
\fn{powerful: Heb. in power. full...: Heb. in majesty.}
\vs Psa 29:5 The voice of the \textsc{Lord} breaketh the cedars; yea, the \textsc{Lord} breaketh the cedars of Lebanon.
\vs Psa 29:6 He maketh them also to skip like a calf; Lebanon and Sirion like a young unicorn.
\vs Psa 29:7 The voice of the \textsc{Lord} divideth the flames of fire.
\fn{divideth: Heb. cutteth out.}
\vs Psa 29:8 The voice of the \textsc{Lord} shaketh the wilderness; the \textsc{Lord} shaketh the wilderness of Kadesh.
\vs Psa 29:9 The voice of the \textsc{Lord} maketh the hinds to calve, and discovereth the forests: and in his temple doth every one speak of \bibemph{his} glory.
\fn{to calve: or, to be in pain. doth...: or, every whit of it uttereth, etc.}
\vs Psa 29:10 The \textsc{Lord} sitteth upon the flood; yea, the \textsc{Lord} sitteth King for ever.
\vs Psa 29:11 The \textsc{Lord} will give strength unto his people; the \textsc{Lord} will bless his people with peace.
\vs Psa 30:1 A Psalm \bibemph{and} Song \bibemph{at} the dedication of the house of David. I will extol thee, O \textsc{Lord}; for thou hast lifted me up, and hast not made my foes to rejoice over me.
\vs Psa 30:2 O \textsc{Lord} my God, I cried unto thee, and thou hast healed me.
\vs Psa 30:3 O \textsc{Lord}, thou hast brought up my soul from the grave: thou hast kept me alive, that I should not go down to the pit.
\vs Psa 30:4 Sing unto the \textsc{Lord}, O ye saints of his, and give thanks at the remembrance of his holiness.
\fn{at...: or, to the memorial.}
\vs Psa 30:5 For his anger \bibemph{endureth but} a moment; in his favour \bibemph{is} life: weeping may endure for a night, but joy \bibemph{cometh} in the morning.
\fn{his anger...: Heb. there is but a moment in his anger. for a night: Heb. in the evening. joy: Heb. singing.}
\vs Psa 30:6 And in my prosperity I said, I shall never be moved.
\vs Psa 30:7 \textsc{Lord}, by thy favour thou hast made my mountain to stand strong: thou didst hide thy face, \bibemph{and} I was troubled.
\fn{made...: Heb. settled strength for my mountain.}
\vs Psa 30:8 I cried to thee, O \textsc{Lord}; and unto the \textsc{Lord} I made supplication.
\vs Psa 30:9 What profit \bibemph{is there} in my blood, when I go down to the pit? Shall the dust praise thee? shall it declare thy truth?
\vs Psa 30:10 Hear, O \textsc{Lord}, and have mercy upon me: \textsc{Lord}, be thou my helper.
\vs Psa 30:11 Thou hast turned for me my mourning into dancing: thou hast put off my sackcloth, and girded me with gladness;
\vs Psa 30:12 To the end that \bibemph{my} glory may sing praise to thee, and not be silent. O \textsc{Lord} my God, I will give thanks unto thee for ever.
\fn{my glory: that is, my tongue, or, my soul.}
\vs Psa 31:1 To the chief Musician, A Psalm of David. In thee, O \textsc{Lord}, do I put my trust; let me never be ashamed: deliver me in thy righteousness.
\vs Psa 31:2 Bow down thine ear to me; deliver me speedily: be thou my strong rock, for an house of defence to save me.
\fn{my...: Heb. to me for rock of strength.}
\vs Psa 31:3 For thou \bibemph{art} my rock and my fortress; therefore for thy name's sake lead me, and guide me.
\vs Psa 31:4 Pull me out of the net that they have laid privily for me: for thou \bibemph{art} my strength.
\vs Psa 31:5 Into thine hand I commit my spirit: thou hast redeemed me, O \textsc{Lord} God of truth.
\vs Psa 31:6 I have hated them that regard lying vanities: but I trust in the \textsc{Lord}.
\vs Psa 31:7 I will be glad and rejoice in thy mercy: for thou hast considered my trouble; thou hast known my soul in adversities;
\vs Psa 31:8 And hast not shut me up into the hand of the enemy: thou hast set my feet in a large room.
\vs Psa 31:9 Have mercy upon me, O \textsc{Lord}, for I am in trouble: mine eye is consumed with grief, \bibemph{yea}, my soul and my belly.
\vs Psa 31:10 For my life is spent with grief, and my years with sighing: my strength faileth because of mine iniquity, and my bones are consumed.
\vs Psa 31:11 I was a reproach among all mine enemies, but especially among my neighbours, and a fear to mine acquaintance: they that did see me without fled from me.
\vs Psa 31:12 I am forgotten as a dead man out of mind: I am like a broken vessel.
\fn{a broken...: Heb. a vessel that perisheth.}
\vs Psa 31:13 For I have heard the slander of many: fear \bibemph{was} on every side: while they took counsel together against me, they devised to take away my life.
\vs Psa 31:14 But I trusted in thee, O \textsc{Lord}: I said, Thou \bibemph{art} my God.
\vs Psa 31:15 My times \bibemph{are} in thy hand: deliver me from the hand of mine enemies, and from them that persecute me.
\vs Psa 31:16 Make thy face to shine upon thy servant: save me for thy mercies' sake.
\vs Psa 31:17 Let me not be ashamed, O \textsc{Lord}; for I have called upon thee: let the wicked be ashamed, \bibemph{and} let them be silent in the grave.
\fn{silent...: or, cut off for.}
\vs Psa 31:18 Let the lying lips be put to silence; which speak grievous things proudly and contemptuously against the righteous.
\fn{grievous...: Heb. a hard thing.}
\vs Psa 31:19 \bibemph{Oh} how great \bibemph{is} thy goodness, which thou hast laid up for them that fear thee; \bibemph{which} thou hast wrought for them that trust in thee before the sons of men!
\vs Psa 31:20 Thou shalt hide them in the secret of thy presence from the pride of man: thou shalt keep them secretly in a pavilion from the strife of tongues.
\vs Psa 31:21 Blessed \bibemph{be} the \textsc{Lord}: for he hath shewed me his marvellous kindness in a strong city.
\fn{strong: or, fenced.}
\vs Psa 31:22 For I said in my haste, I am cut off from before thine eyes: nevertheless thou heardest the voice of my supplications when I cried unto thee.
\vs Psa 31:23 O love the \textsc{Lord}, all ye his saints: \bibemph{for} the \textsc{Lord} preserveth the faithful, and plentifully rewardeth the proud doer.
\vs Psa 31:24 Be of good courage, and he shall strengthen your heart, all ye that hope in the \textsc{Lord}.
\vs Psa 32:1 \bibemph{A Psalm} of David, Maschil. Blessed \bibemph{is he whose} transgression \bibemph{is} forgiven, \bibemph{whose} sin \bibemph{is} covered.
\fn{A Psalm...: or, A Psalm of David giving instruction.}
\vs Psa 32:2 Blessed \bibemph{is} the man unto whom the \textsc{Lord} imputeth not iniquity, and in whose spirit \bibemph{there is} no guile.
\vs Psa 32:3 When I kept silence, my bones waxed old through my roaring all the day long.
\vs Psa 32:4 For day and night thy hand was heavy upon me: my moisture is turned into the drought of summer. Selah.
\vs Psa 32:5 I acknowledged my sin unto thee, and mine iniquity have I not hid. I said, I will confess my transgressions unto the \textsc{Lord}; and thou forgavest the iniquity of my sin. Selah.
\vs Psa 32:6 For this shall every one that is godly pray unto thee in a time when thou mayest be found: surely in the floods of great waters they shall not come nigh unto him.
\fn{when...: Heb. of finding.}
\vs Psa 32:7 Thou \bibemph{art} my hiding place; thou shalt preserve me from trouble; thou shalt compass me about with songs of deliverance. Selah.
\vs Psa 32:8 I will instruct thee and teach thee in the way which thou shalt go: I will guide thee with mine eye.
\fn{guide...: Heb. counsel thee, mine eye shall be upon thee.}
\vs Psa 32:9 Be ye not as the horse, \bibemph{or} as the mule, \bibemph{which} have no understanding: whose mouth must be held in with bit and bridle, lest they come near unto thee.
\vs Psa 32:10 Many sorrows \bibemph{shall be} to the wicked: but he that trusteth in the \textsc{Lord}, mercy shall compass him about.
\vs Psa 32:11 Be glad in the \textsc{Lord}, and rejoice, ye righteous: and shout for joy, all \bibemph{ye that are} upright in heart.
\vs Psa 33:1 Rejoice in the \textsc{Lord}, O ye righteous: \bibemph{for} praise is comely for the upright.
\vs Psa 33:2 Praise the \textsc{Lord} with harp: sing unto him with the psaltery \bibemph{and} an instrument of ten strings.
\vs Psa 33:3 Sing unto him a new song; play skilfully with a loud noise.
\vs Psa 33:4 For the word of the \textsc{Lord} \bibemph{is} right; and all his works \bibemph{are done} in truth.
\vs Psa 33:5 He loveth righteousness and judgment: the earth is full of the goodness of the \textsc{Lord}.
\fn{goodness: or, mercy.}
\vs Psa 33:6 By the word of the \textsc{Lord} were the heavens made; and all the host of them by the breath of his mouth.
\vs Psa 33:7 He gathereth the waters of the sea together as an heap: he layeth up the depth in storehouses.
\vs Psa 33:8 Let all the earth fear the \textsc{Lord}: let all the inhabitants of the world stand in awe of him.
\vs Psa 33:9 For he spake, and it was \bibemph{done}; he commanded, and it stood fast.
\vs Psa 33:10 The \textsc{Lord} bringeth the counsel of the heathen to nought: he maketh the devices of the people of none effect.
\fn{bringeth: Heb. maketh frustrate.}
\vs Psa 33:11 The counsel of the \textsc{Lord} standeth for ever, the thoughts of his heart to all generations.
\fn{to all...: Heb. to generation and generation.}
\vs Psa 33:12 Blessed \bibemph{is} the nation whose God \bibemph{is} the \textsc{Lord}; \bibemph{and} the people \bibemph{whom} he hath chosen for his own inheritance.
\vs Psa 33:13 The \textsc{Lord} looketh from heaven; he beholdeth all the sons of men.
\vs Psa 33:14 From the place of his habitation he looketh upon all the inhabitants of the earth.
\vs Psa 33:15 He fashioneth their hearts alike; he considereth all their works.
\vs Psa 33:16 There is no king saved by the multitude of an host: a mighty man is not delivered by much strength.
\vs Psa 33:17 An horse \bibemph{is} a vain thing for safety: neither shall he deliver \bibemph{any} by his great strength.
\vs Psa 33:18 Behold, the eye of the \textsc{Lord} \bibemph{is} upon them that fear him, upon them that hope in his mercy;
\vs Psa 33:19 To deliver their soul from death, and to keep them alive in famine.
\vs Psa 33:20 Our soul waiteth for the \textsc{Lord}: he \bibemph{is} our help and our shield.
\vs Psa 33:21 For our heart shall rejoice in him, because we have trusted in his holy name.
\vs Psa 33:22 Let thy mercy, O \textsc{Lord}, be upon us, according as we hope in thee.
\vs Psa 34:1 \bibemph{A Psalm} of David, when he changed his behaviour before Abimelech; who drove him away, and he departed. I will bless the \textsc{Lord} at all times: his praise \bibemph{shall} continually \bibemph{be} in my mouth.
\fn{Abimelech: or, Achish.}
\vs Psa 34:2 My soul shall make her boast in the \textsc{Lord}: the humble shall hear \bibemph{thereof}, and be glad.
\vs Psa 34:3 O magnify the \textsc{Lord} with me, and let us exalt his name together.
\vs Psa 34:4 I sought the \textsc{Lord}, and he heard me, and delivered me from all my fears.
\vs Psa 34:5 They looked unto him, and were lightened: and their faces were not ashamed.
\fn{were lightened: or, they flowed unto him.}
\vs Psa 34:6 This poor man cried, and the \textsc{Lord} heard \bibemph{him}, and saved him out of all his troubles.
\vs Psa 34:7 The angel of the \textsc{Lord} encampeth round about them that fear him, and delivereth them.
\vs Psa 34:8 O taste and see that the \textsc{Lord} \bibemph{is} good: blessed \bibemph{is} the man \bibemph{that} trusteth in him.
\vs Psa 34:9 O fear the \textsc{Lord}, ye his saints: for \bibemph{there is} no want to them that fear him.
\vs Psa 34:10 The young lions do lack, and suffer hunger: but they that seek the \textsc{Lord} shall not want any good \bibemph{thing}.
\vs Psa 34:11 Come, ye children, hearken unto me: I will teach you the fear of the \textsc{Lord}.
\vs Psa 34:12 What man \bibemph{is he that} desireth life, \bibemph{and} loveth \bibemph{many} days, that he may see good?
\vs Psa 34:13 Keep thy tongue from evil, and thy lips from speaking guile.
\vs Psa 34:14 Depart from evil, and do good; seek peace, and pursue it.
\vs Psa 34:15 The eyes of the \textsc{Lord} \bibemph{are} upon the righteous, and his ears \bibemph{are open} unto their cry.
\vs Psa 34:16 The face of the \textsc{Lord} \bibemph{is} against them that do evil, to cut off the remembrance of them from the earth.
\vs Psa 34:17 \bibemph{The righteous} cry, and the \textsc{Lord} heareth, and delivereth them out of all their troubles.
\vs Psa 34:18 The \textsc{Lord} \bibemph{is} nigh unto them that are of a broken heart; and saveth such as be of a contrite spirit.
\fn{unto...: Heb. to the broken of heart. of a contrite...: Heb. contrite of spirit.}
\vs Psa 34:19 Many \bibemph{are} the afflictions of the righteous: but the \textsc{Lord} delivereth him out of them all.
\vs Psa 34:20 He keepeth all his bones: not one of them is broken.
\vs Psa 34:21 Evil shall slay the wicked: and they that hate the righteous shall be desolate.
\fn{desolate: or, guilty.}
\vs Psa 34:22 The \textsc{Lord} redeemeth the soul of his servants: and none of them that trust in him shall be desolate.
\fn{desolate: or, guilty.}
\vs Psa 35:1 \bibemph{A Psalm} of David. Plead \bibemph{my cause}, O \textsc{Lord}, with them that strive with me: fight against them that fight against me.
\vs Psa 35:2 Take hold of shield and buckler, and stand up for mine help.
\vs Psa 35:3 Draw out also the spear, and stop \bibemph{the way} against them that persecute me: say unto my soul, I \bibemph{am} thy salvation.
\vs Psa 35:4 Let them be confounded and put to shame that seek after my soul: let them be turned back and brought to confusion that devise my hurt.
\vs Psa 35:5 Let them be as chaff before the wind: and let the angel of the \textsc{Lord} chase \bibemph{them}.
\vs Psa 35:6 Let their way be dark and slippery: and let the angel of the \textsc{Lord} persecute them.
\fn{dark...: Heb. darkness and slipperiness.}
\vs Psa 35:7 For without cause have they hid for me their net \bibemph{in} a pit, \bibemph{which} without cause they have digged for my soul.
\vs Psa 35:8 Let destruction come upon him at unawares; and let his net that he hath hid catch himself: into that very destruction let him fall.
\fn{at...: Heb. which he knoweth not of.}
\vs Psa 35:9 And my soul shall be joyful in the \textsc{Lord}: it shall rejoice in his salvation.
\vs Psa 35:10 All my bones shall say, \textsc{Lord}, who \bibemph{is} like unto thee, which deliverest the poor from him that is too strong for him, yea, the poor and the needy from him that spoileth him?
\vs Psa 35:11 False witnesses did rise up; they laid to my charge \bibemph{things} that I knew not.
\fn{False...: Heb. Witnesses of wrong. they...: Heb. they asked me.}
\vs Psa 35:12 They rewarded me evil for good \bibemph{to} the spoiling of my soul.
\fn{spoiling: Heb. depriving.}
\vs Psa 35:13 But as for me, when they were sick, my clothing \bibemph{was} sackcloth: I humbled my soul with fasting; and my prayer returned into mine own bosom.
\fn{humbled: or, afflicted.}
\vs Psa 35:14 I behaved myself as though \bibemph{he had been} my friend \bibemph{or} brother: I bowed down heavily, as one that mourneth \bibemph{for his} mother.
\fn{behaved myself: Heb. walked. as though...: Heb. as a friend, as a brother to me.}
\vs Psa 35:15 But in mine adversity they rejoiced, and gathered themselves together: \bibemph{yea}, the abjects gathered themselves together against me, and I knew \bibemph{it} not; they did tear \bibemph{me}, and ceased not:
\fn{adversity: Heb. halting.}
\vs Psa 35:16 With hypocritical mockers in feasts, they gnashed upon me with their teeth.
\vs Psa 35:17 Lord, how long wilt thou look on? rescue my soul from their destructions, my darling from the lions.
\fn{darling: Heb. only one.}
\vs Psa 35:18 I will give thee thanks in the great congregation: I will praise thee among much people.
\fn{much: Heb. strong.}
\vs Psa 35:19 Let not them that are mine enemies wrongfully rejoice over me: \bibemph{neither} let them wink with the eye that hate me without a cause.
\fn{wrongfully: Heb. falsely.}
\vs Psa 35:20 For they speak not peace: but they devise deceitful matters against \bibemph{them that are} quiet in the land.
\vs Psa 35:21 Yea, they opened their mouth wide against me, \bibemph{and} said, Aha, aha, our eye hath seen \bibemph{it}.
\vs Psa 35:22 \bibemph{This} thou hast seen, O \textsc{Lord}: keep not silence: O Lord, be not far from me.
\vs Psa 35:23 Stir up thyself, and awake to my judgment, \bibemph{even} unto my cause, my God and my Lord.
\vs Psa 35:24 Judge me, O \textsc{Lord} my God, according to thy righteousness; and let them not rejoice over me.
\vs Psa 35:25 Let them not say in their hearts, Ah, so would we have it: let them not say, We have swallowed him up.
\fn{Ah...: Heb. Ah, ah, our soul.}
\vs Psa 35:26 Let them be ashamed and brought to confusion together that rejoice at mine hurt: let them be clothed with shame and dishonour that magnify \bibemph{themselves} against me.
\vs Psa 35:27 Let them shout for joy, and be glad, that favour my righteous cause: yea, let them say continually, Let the \textsc{Lord} be magnified, which hath pleasure in the prosperity of his servant.
\fn{righteous...: Heb. righteousness.}
\vs Psa 35:28 And my tongue shall speak of thy righteousness \bibemph{and} of thy praise all the day long.
\vs Psa 36:1 To the chief Musician, \bibemph{A Psalm} of David the servant of the \textsc{Lord}. The transgression of the wicked saith within my heart, \bibemph{that there is} no fear of God before his eyes.
\vs Psa 36:2 For he flattereth himself in his own eyes, until his iniquity be found to be hateful.
\fn{until...: Heb. to find his iniquity to hate.}
\vs Psa 36:3 The words of his mouth \bibemph{are} iniquity and deceit: he hath left off to be wise, \bibemph{and} to do good.
\vs Psa 36:4 He deviseth mischief upon his bed; he setteth himself in a way \bibemph{that is} not good; he abhorreth not evil.
\fn{mischief: or, vanity.}
\vs Psa 36:5 Thy mercy, O \textsc{Lord}, \bibemph{is} in the heavens; \bibemph{and} thy faithfulness \bibemph{reacheth} unto the clouds.
\vs Psa 36:6 Thy righteousness \bibemph{is} like the great mountains; thy judgments \bibemph{are} a great deep: O \textsc{Lord}, thou preservest man and beast.
\fn{the...: Heb. the mountains of God.}
\vs Psa 36:7 How excellent \bibemph{is} thy lovingkindness, O God! therefore the children of men put their trust under the shadow of thy wings.
\fn{excellent: Heb. precious.}
\vs Psa 36:8 They shall be abundantly satisfied with the fatness of thy house; and thou shalt make them drink of the river of thy pleasures.
\fn{abundantly...: Heb. watered.}
\vs Psa 36:9 For with thee \bibemph{is} the fountain of life: in thy light shall we see light.
\vs Psa 36:10 O continue thy lovingkindness unto them that know thee; and thy righteousness to the upright in heart.
\fn{continue: Heb. draw out at length.}
\vs Psa 36:11 Let not the foot of pride come against me, and let not the hand of the wicked remove me.
\vs Psa 36:12 There are the workers of iniquity fallen: they are cast down, and shall not be able to rise.
\vs Psa 37:1 \bibemph{A Psalm} of David. Fret not thyself because of evildoers, neither be thou envious against the workers of iniquity.
\vs Psa 37:2 For they shall soon be cut down like the grass, and wither as the green herb.
\vs Psa 37:3 Trust in the \textsc{Lord}, and do good; \bibemph{so} shalt thou dwell in the land, and verily thou shalt be fed.
\fn{verily: Heb. in truth, or, stableness.}
\vs Psa 37:4 Delight thyself also in the \textsc{Lord}; and he shall give thee the desires of thine heart.
\vs Psa 37:5 Commit thy way unto the \textsc{Lord}; trust also in him; and he shall bring \bibemph{it} to pass.
\fn{Commit...: Heb. Roll thy way upon.}
\vs Psa 37:6 And he shall bring forth thy righteousness as the light, and thy judgment as the noonday.
\vs Psa 37:7 Rest in the \textsc{Lord}, and wait patiently for him: fret not thyself because of him who prospereth in his way, because of the man who bringeth wicked devices to pass.
\fn{Rest in: Heb. Be silent to.}
\vs Psa 37:8 Cease from anger, and forsake wrath: fret not thyself in any wise to do evil.
\vs Psa 37:9 For evildoers shall be cut off: but those that wait upon the \textsc{Lord}, they shall inherit the earth.
\vs Psa 37:10 For yet a little while, and the wicked \bibemph{shall} not \bibemph{be}: yea, thou shalt diligently consider his place, and it \bibemph{shall} not \bibemph{be}.
\vs Psa 37:11 But the meek shall inherit the earth; and shall delight themselves in the abundance of peace.
\vs Psa 37:12 The wicked plotteth against the just, and gnasheth upon him with his teeth.
\fn{plotteth: or, practiseth.}
\vs Psa 37:13 The Lord shall laugh at him: for he seeth that his day is coming.
\vs Psa 37:14 The wicked have drawn out the sword, and have bent their bow, to cast down the poor and needy, \bibemph{and} to slay such as be of upright conversation.
\fn{such...: Heb. the upright of way.}
\vs Psa 37:15 Their sword shall enter into their own heart, and their bows shall be broken.
\vs Psa 37:16 A little that a righteous man hath \bibemph{is} better than the riches of many wicked.
\vs Psa 37:17 For the arms of the wicked shall be broken: but the \textsc{Lord} upholdeth the righteous.
\vs Psa 37:18 The \textsc{Lord} knoweth the days of the upright: and their inheritance shall be for ever.
\vs Psa 37:19 They shall not be ashamed in the evil time: and in the days of famine they shall be satisfied.
\vs Psa 37:20 But the wicked shall perish, and the enemies of the \textsc{Lord} \bibemph{shall be} as the fat of lambs: they shall consume; into smoke shall they consume away.
\fn{the fat: Heb. the preciousness.}
\vs Psa 37:21 The wicked borroweth, and payeth not again: but the righteous sheweth mercy, and giveth.
\vs Psa 37:22 For \bibemph{such as be} blessed of him shall inherit the earth; and \bibemph{they that be} cursed of him shall be cut off.
\vs Psa 37:23 The steps of a \bibemph{good} man are ordered by the \textsc{Lord}: and he delighteth in his way.
\fn{ordered: or, established.}
\vs Psa 37:24 Though he fall, he shall not be utterly cast down: for the \textsc{Lord} upholdeth \bibemph{him with} his hand.
\vs Psa 37:25 I have been young, and \bibemph{now} am old; yet have I not seen the righteous forsaken, nor his seed begging bread.
\vs Psa 37:26 \bibemph{He is} ever merciful, and lendeth; and his seed \bibemph{is} blessed.
\fn{ever: Heb. all the day.}
\vs Psa 37:27 Depart from evil, and do good; and dwell for evermore.
\vs Psa 37:28 For the \textsc{Lord} loveth judgment, and forsaketh not his saints; they are preserved for ever: but the seed of the wicked shall be cut off.
\vs Psa 37:29 The righteous shall inherit the land, and dwell therein for ever.
\vs Psa 37:30 The mouth of the righteous speaketh wisdom, and his tongue talketh of judgment.
\vs Psa 37:31 The law of his God \bibemph{is} in his heart; none of his steps shall slide.
\fn{steps: or, goings.}
\vs Psa 37:32 The wicked watcheth the righteous, and seeketh to slay him.
\vs Psa 37:33 The \textsc{Lord} will not leave him in his hand, nor condemn him when he is judged.
\vs Psa 37:34 Wait on the \textsc{Lord}, and keep his way, and he shall exalt thee to inherit the land: when the wicked are cut off, thou shalt see \bibemph{it}.
\vs Psa 37:35 I have seen the wicked in great power, and spreading himself like a green bay tree.
\fn{a green...: or, a green tree that groweth in his own soil.}
\vs Psa 37:36 Yet he passed away, and, lo, he \bibemph{was} not: yea, I sought him, but he could not be found.
\vs Psa 37:37 Mark the perfect \bibemph{man}, and behold the upright: for the end of \bibemph{that} man \bibemph{is} peace.
\vs Psa 37:38 But the transgressors shall be destroyed together: the end of the wicked shall be cut off.
\vs Psa 37:39 But the salvation of the righteous \bibemph{is} of the \textsc{Lord}: \bibemph{he is} their strength in the time of trouble.
\vs Psa 37:40 And the \textsc{Lord} shall help them, and deliver them: he shall deliver them from the wicked, and save them, because they trust in him.
\vs Psa 38:1 A Psalm of David, to bring to remembrance. O \textsc{Lord}, rebuke me not in thy wrath: neither chasten me in thy hot displeasure.
\vs Psa 38:2 For thine arrows stick fast in me, and thy hand presseth me sore.
\vs Psa 38:3 \bibemph{There is} no soundness in my flesh because of thine anger; neither \bibemph{is there any} rest in my bones because of my sin.
\fn{rest: Heb. peace, or, health.}
\vs Psa 38:4 For mine iniquities are gone over mine head: as an heavy burden they are too heavy for me.
\vs Psa 38:5 My wounds stink \bibemph{and} are corrupt because of my foolishness.
\vs Psa 38:6 I am troubled; I am bowed down greatly; I go mourning all the day long.
\fn{troubled: Heb. wried.}
\vs Psa 38:7 For my loins are filled with a loathsome \bibemph{disease}: and \bibemph{there is} no soundness in my flesh.
\vs Psa 38:8 I am feeble and sore broken: I have roared by reason of the disquietness of my heart.
\vs Psa 38:9 Lord, all my desire \bibemph{is} before thee; and my groaning is not hid from thee.
\vs Psa 38:10 My heart panteth, my strength faileth me: as for the light of mine eyes, it also is gone from me.
\fn{is gone...: Heb. is not with me.}
\vs Psa 38:11 My lovers and my friends stand aloof from my sore; and my kinsmen stand afar off.
\fn{sore: Heb. stroke. my kinsmen: or, my neighbours.}
\vs Psa 38:12 They also that seek after my life lay snares \bibemph{for me}: and they that seek my hurt speak mischievous things, and imagine deceits all the day long.
\vs Psa 38:13 But I, as a deaf \bibemph{man}, heard not; and \bibemph{I was} as a dumb man \bibemph{that} openeth not his mouth.
\vs Psa 38:14 Thus I was as a man that heareth not, and in whose mouth \bibemph{are} no reproofs.
\vs Psa 38:15 For in thee, O \textsc{Lord}, do I hope: thou wilt hear, O Lord my God.
\fn{in...: or, thee do I wait for. hear: or, answer.}
\vs Psa 38:16 For I said, \bibemph{Hear me}, lest \bibemph{otherwise} they should rejoice over me: when my foot slippeth, they magnify \bibemph{themselves} against me.
\vs Psa 38:17 For I \bibemph{am} ready to halt, and my sorrow \bibemph{is} continually before me.
\fn{to halt: Heb. for halting.}
\vs Psa 38:18 For I will declare mine iniquity; I will be sorry for my sin.
\vs Psa 38:19 But mine enemies \bibemph{are} lively, \bibemph{and} they are strong: and they that hate me wrongfully are multiplied.
\fn{are lively...: Heb. being living, are strong.}
\vs Psa 38:20 They also that render evil for good are mine adversaries; because I follow \bibemph{the thing that} good \bibemph{is}.
\vs Psa 38:21 Forsake me not, O \textsc{Lord}: O my God, be not far from me.
\vs Psa 38:22 Make haste to help me, O Lord my salvation.
\fn{to...: Heb. for my help.}
\vs Psa 39:1 To the chief Musician, \bibemph{even} to Jeduthun, A Psalm of David. I said, I will take heed to my ways, that I sin not with my tongue: I will keep my mouth with a bridle, while the wicked is before me.
\fn{my mouth...: Heb. a bridle, or, muzzle for my mouth.}
\vs Psa 39:2 I was dumb with silence, I held my peace, \bibemph{even} from good; and my sorrow was stirred.
\fn{stirred: Heb. troubled.}
\vs Psa 39:3 My heart was hot within me, while I was musing the fire burned: \bibemph{then} spake I with my tongue,
\vs Psa 39:4 \textsc{Lord}, make me to know mine end, and the measure of my days, what it \bibemph{is; that} I may know how frail I \bibemph{am}.
\fn{how...: or, what time I have here.}
\vs Psa 39:5 Behold, thou hast made my days \bibemph{as} an handbreadth; and mine age \bibemph{is} as nothing before thee: verily every man at his best state \bibemph{is} altogether vanity. Selah.
\fn{at...: Heb. settled.}
\vs Psa 39:6 Surely every man walketh in a vain shew: surely they are disquieted in vain: he heapeth up \bibemph{riches}, and knoweth not who shall gather them.
\fn{a vain...: Heb. an image.}
\vs Psa 39:7 And now, Lord, what wait I for? my hope \bibemph{is} in thee.
\vs Psa 39:8 Deliver me from all my transgressions: make me not the reproach of the foolish.
\vs Psa 39:9 I was dumb, I opened not my mouth; because thou didst \bibemph{it}.
\vs Psa 39:10 Remove thy stroke away from me: I am consumed by the blow of thine hand.
\fn{blow: Heb. conflict.}
\vs Psa 39:11 When thou with rebukes dost correct man for iniquity, thou makest his beauty to consume away like a moth: surely every man \bibemph{is} vanity. Selah.
\fn{his...: Heb. that which is to be desired in him to melt away.}
\vs Psa 39:12 Hear my prayer, O \textsc{Lord}, and give ear unto my cry; hold not thy peace at my tears: for I \bibemph{am} a stranger with thee, \bibemph{and} a sojourner, as all my fathers \bibemph{were}.
\vs Psa 39:13 O spare me, that I may recover strength, before I go hence, and be no more.
\vs Psa 40:1 To the chief Musician, A Psalm of David. I waited patiently for the \textsc{Lord}; and he inclined unto me, and heard my cry.
\fn{I waited...: Heb. In waiting I waited.}
\vs Psa 40:2 He brought me up also out of an horrible pit, out of the miry clay, and set my feet upon a rock, \bibemph{and} established my goings.
\fn{an...: Heb. a pit of noise.}
\vs Psa 40:3 And he hath put a new song in my mouth, \bibemph{even} praise unto our God: many shall see \bibemph{it}, and fear, and shall trust in the \textsc{Lord}.
\vs Psa 40:4 Blessed \bibemph{is} that man that maketh the \textsc{Lord} his trust, and respecteth not the proud, nor such as turn aside to lies.
\vs Psa 40:5 Many, O \textsc{Lord} my God, \bibemph{are} thy wonderful works \bibemph{which} thou hast done, and thy thoughts \bibemph{which are} to us-ward: they cannot be reckoned up in order unto thee: \bibemph{if} I would declare and speak \bibemph{of them}, they are more than can be numbered.
\fn{they can not...: or, none can order them unto thee.}
\vs Psa 40:6 Sacrifice and offering thou didst not desire; mine ears hast thou opened: burnt offering and sin offering hast thou not required.
\fn{opened: Heb. digged.}
\vs Psa 40:7 Then said I, Lo, I come: in the volume of the book \bibemph{it is} written of me,
\vs Psa 40:8 I delight to do thy will, O my God: yea, thy law \bibemph{is} within my heart.
\fn{within...: Heb. in the midst of my bowels.}
\vs Psa 40:9 I have preached righteousness in the great congregation: lo, I have not refrained my lips, O \textsc{Lord}, thou knowest.
\vs Psa 40:10 I have not hid thy righteousness within my heart; I have declared thy faithfulness and thy salvation: I have not concealed thy lovingkindness and thy truth from the great congregation.
\vs Psa 40:11 Withhold not thou thy tender mercies from me, O \textsc{Lord}: let thy lovingkindness and thy truth continually preserve me.
\vs Psa 40:12 For innumerable evils have compassed me about: mine iniquities have taken hold upon me, so that I am not able to look up; they are more than the hairs of mine head: therefore my heart faileth me.
\fn{falleth: Heb. forsaketh.}
\vs Psa 40:13 Be pleased, O \textsc{Lord}, to deliver me: O \textsc{Lord}, make haste to help me.
\vs Psa 40:14 Let them be ashamed and confounded together that seek after my soul to destroy it; let them be driven backward and put to shame that wish me evil.
\vs Psa 40:15 Let them be desolate for a reward of their shame that say unto me, Aha, aha.
\vs Psa 40:16 Let all those that seek thee rejoice and be glad in thee: let such as love thy salvation say continually, The \textsc{Lord} be magnified.
\vs Psa 40:17 But I \bibemph{am} poor and needy; \bibemph{yet} the Lord thinketh upon me: thou \bibemph{art} my help and my deliverer; make no tarrying, O my God.
\vs Psa 41:1 To the chief Musician, A Psalm of David. Blessed \bibemph{is} he that considereth the poor: the \textsc{Lord} will deliver him in time of trouble.
\fn{the poor: or, the weak, or, sick. in time...: Heb. in the day of evil.}
\vs Psa 41:2 The \textsc{Lord} will preserve him, and keep him alive; \bibemph{and} he shall be blessed upon the earth: and thou wilt not deliver him unto the will of his enemies.
\fn{thou...: or, do not thou deliver.}
\vs Psa 41:3 The \textsc{Lord} will strengthen him upon the bed of languishing: thou wilt make all his bed in his sickness.
\fn{make: Heb. turn.}
\vs Psa 41:4 I said, \textsc{Lord}, be merciful unto me: heal my soul; for I have sinned against thee.
\vs Psa 41:5 Mine enemies speak evil of me, When shall he die, and his name perish?
\vs Psa 41:6 And if he come to see \bibemph{me}, he speaketh vanity: his heart gathereth iniquity to itself; \bibemph{when} he goeth abroad, he telleth \bibemph{it}.
\vs Psa 41:7 All that hate me whisper together against me: against me do they devise my hurt.
\fn{my...: Heb. evil to me.}
\vs Psa 41:8 An evil disease, \bibemph{say they}, cleaveth fast unto him: and \bibemph{now} that he lieth he shall rise up no more.
\fn{An...: Heb. A thing of Belial.}
\vs Psa 41:9 Yea, mine own familiar friend, in whom I trusted, which did eat of my bread, hath lifted up \bibemph{his} heel against me.
\fn{mine...: Heb. the man of my peace. lifted...: Heb. magnified.}
\vs Psa 41:10 But thou, O \textsc{Lord}, be merciful unto me, and raise me up, that I may requite them.
\vs Psa 41:11 By this I know that thou favourest me, because mine enemy doth not triumph over me.
\vs Psa 41:12 And as for me, thou upholdest me in mine integrity, and settest me before thy face for ever.
\vs Psa 41:13 Blessed \bibemph{be} the \textsc{Lord} God of Israel from everlasting, and to everlasting. Amen, and Amen.
\vs Psa 42:1 To the chief Musician, Maschil, for the sons of Korah. As the hart panteth after the water brooks, so panteth my soul after thee, O God.
\fn{Maschil...: or, A Psalm giving instruction of the sons, etc. panteth: Heb. brayeth.}
\vs Psa 42:2 My soul thirsteth for God, for the living God: when shall I come and appear before God?
\vs Psa 42:3 My tears have been my meat day and night, while they continually say unto me, Where \bibemph{is} thy God?
\vs Psa 42:4 When I remember these \bibemph{things}, I pour out my soul in me: for I had gone with the multitude, I went with them to the house of God, with the voice of joy and praise, with a multitude that kept holyday.
\vs Psa 42:5 Why art thou cast down, O my soul? and \bibemph{why} art thou disquieted in me? hope thou in God: for I shall yet praise him \bibemph{for} the help of his countenance.
\fn{cast: Heb. bowed. praise: or, give thanks. for the...: or, his presence is salvation.}
\vs Psa 42:6 O my God, my soul is cast down within me: therefore will I remember thee from the land of Jordan, and of the Hermonites, from the hill Mizar.
\fn{the hill...: or, the little hill.}
\vs Psa 42:7 Deep calleth unto deep at the noise of thy waterspouts: all thy waves and thy billows are gone over me.
\vs Psa 42:8 \bibemph{Yet} the \textsc{Lord} will command his lovingkindness in the daytime, and in the night his song \bibemph{shall be} with me, \bibemph{and} my prayer unto the God of my life.
\vs Psa 42:9 I will say unto God my rock, Why hast thou forgotten me? why go I mourning because of the oppression of the enemy?
\vs Psa 42:10 \bibemph{As} with a sword in my bones, mine enemies reproach me; while they say daily unto me, Where \bibemph{is} thy God?
\fn{sword: or, killing.}
\vs Psa 42:11 Why art thou cast down, O my soul? and why art thou disquieted within me? hope thou in God: for I shall yet praise him, \bibemph{who is} the health of my countenance, and my God.
\vs Psa 43:1 Judge me, O God, and plead my cause against an ungodly nation: O deliver me from the deceitful and unjust man.
\fn{ungodly: or, unmerciful. the deceitful...: Heb. a man of deceit and iniquity.}
\vs Psa 43:2 For thou \bibemph{art} the God of my strength: why dost thou cast me off? why go I mourning because of the oppression of the enemy?
\vs Psa 43:3 O send out thy light and thy truth: let them lead me; let them bring me unto thy holy hill, and to thy tabernacles.
\vs Psa 43:4 Then will I go unto the altar of God, unto God my exceeding joy: yea, upon the harp will I praise thee, O God my God.
\fn{my exceeding...: Heb. the gladness of my joy.}
\vs Psa 43:5 Why art thou cast down, O my soul? and why art thou disquieted within me? hope in God: for I shall yet praise him, \bibemph{who is} the health of my countenance, and my God.
\vs Psa 44:1 To the chief Musician for the sons of Korah, Maschil. We have heard with our ears, O God, our fathers have told us, \bibemph{what} work thou didst in their days, in the times of old.
\vs Psa 44:2 \bibemph{How} thou didst drive out the heathen with thy hand, and plantedst them; \bibemph{how} thou didst afflict the people, and cast them out.
\vs Psa 44:3 For they got not the land in possession by their own sword, neither did their own arm save them: but thy right hand, and thine arm, and the light of thy countenance, because thou hadst a favour unto them.
\vs Psa 44:4 Thou art my King, O God: command deliverances for Jacob.
\vs Psa 44:5 Through thee will we push down our enemies: through thy name will we tread them under that rise up against us.
\vs Psa 44:6 For I will not trust in my bow, neither shall my sword save me.
\vs Psa 44:7 But thou hast saved us from our enemies, and hast put them to shame that hated us.
\vs Psa 44:8 In God we boast all the day long, and praise thy name for ever. Selah.
\vs Psa 44:9 But thou hast cast off, and put us to shame; and goest not forth with our armies.
\vs Psa 44:10 Thou makest us to turn back from the enemy: and they which hate us spoil for themselves.
\vs Psa 44:11 Thou hast given us like sheep \bibemph{appointed} for meat; and hast scattered us among the heathen.
\fn{like...: Heb. as sheep of meat.}
\vs Psa 44:12 Thou sellest thy people for nought, and dost not increase \bibemph{thy wealth} by their price.
\fn{for...: Heb. without riches.}
\vs Psa 44:13 Thou makest us a reproach to our neighbours, a scorn and a derision to them that are round about us.
\vs Psa 44:14 Thou makest us a byword among the heathen, a shaking of the head among the people.
\vs Psa 44:15 My confusion \bibemph{is} continually before me, and the shame of my face hath covered me,
\vs Psa 44:16 For the voice of him that reproacheth and blasphemeth; by reason of the enemy and avenger.
\vs Psa 44:17 All this is come upon us; yet have we not forgotten thee, neither have we dealt falsely in thy covenant.
\vs Psa 44:18 Our heart is not turned back, neither have our steps declined from thy way;
\fn{steps: or, goings.}
\vs Psa 44:19 Though thou hast sore broken us in the place of dragons, and covered us with the shadow of death.
\vs Psa 44:20 If we have forgotten the name of our God, or stretched out our hands to a strange god;
\vs Psa 44:21 Shall not God search this out? for he knoweth the secrets of the heart.
\vs Psa 44:22 Yea, for thy sake are we killed all the day long; we are counted as sheep for the slaughter.
\vs Psa 44:23 Awake, why sleepest thou, O Lord? arise, cast \bibemph{us} not off for ever.
\vs Psa 44:24 Wherefore hidest thou thy face, \bibemph{and} forgettest our affliction and our oppression?
\vs Psa 44:25 For our soul is bowed down to the dust: our belly cleaveth unto the earth.
\vs Psa 44:26 Arise for our help, and redeem us for thy mercies' sake.
\fn{for our...: Heb. a help for us.}
\vs Psa 45:1 To the chief Musician upon Shoshannim, for the sons of Korah, Maschil, A Song of loves. My heart is inditing a good matter: I speak of the things which I have made touching the king: my tongue \bibemph{is} the pen of a ready writer.
\fn{Maschil: or, of instruction. is inditing: Heb. boileth, or, bubbleth up.}
\vs Psa 45:2 Thou art fairer than the children of men: grace is poured into thy lips: therefore God hath blessed thee for ever.
\vs Psa 45:3 Gird thy sword upon \bibemph{thy} thigh, O \bibemph{most} mighty, with thy glory and thy majesty.
\vs Psa 45:4 And in thy majesty ride prosperously because of truth and meekness \bibemph{and} righteousness; and thy right hand shall teach thee terrible things.
\fn{ride...: Heb. prosper thou, ride thou.}
\vs Psa 45:5 Thine arrows \bibemph{are} sharp in the heart of the king's enemies; \bibemph{whereby} the people fall under thee.
\vs Psa 45:6 Thy throne, O God, \bibemph{is} for ever and ever: the sceptre of thy kingdom \bibemph{is} a right sceptre.
\vs Psa 45:7 Thou lovest righteousness, and hatest wickedness: therefore God, thy God, hath anointed thee with the oil of gladness above thy fellows.
\vs Psa 45:8 All thy garments \bibemph{smell} of myrrh, and aloes, \bibemph{and} cassia, out of the ivory palaces, whereby they have made thee glad.
\vs Psa 45:9 Kings' daughters \bibemph{were} among thy honourable women: upon thy right hand did stand the queen in gold of Ophir.
\vs Psa 45:10 Hearken, O daughter, and consider, and incline thine ear; forget also thine own people, and thy father's house;
\vs Psa 45:11 So shall the king greatly desire thy beauty: for he \bibemph{is} thy Lord; and worship thou him.
\vs Psa 45:12 And the daughter of Tyre \bibemph{shall be there} with a gift; \bibemph{even} the rich among the people shall intreat thy favour.
\fn{favour: Heb. face.}
\vs Psa 45:13 The king's daughter \bibemph{is} all glorious within: her clothing \bibemph{is} of wrought gold.
\vs Psa 45:14 She shall be brought unto the king in raiment of needlework: the virgins her companions that follow her shall be brought unto thee.
\vs Psa 45:15 With gladness and rejoicing shall they be brought: they shall enter into the king's palace.
\vs Psa 45:16 Instead of thy fathers shall be thy children, whom thou mayest make princes in all the earth.
\vs Psa 45:17 I will make thy name to be remembered in all generations: therefore shall the people praise thee for ever and ever.
\vs Psa 46:1 To the chief Musician for the sons of Korah, A Song upon Alamoth. God \bibemph{is} our refuge and strength, a very present help in trouble.
\fn{for: or, of.}
\vs Psa 46:2 Therefore will not we fear, though the earth be removed, and though the mountains be carried into the midst of the sea;
\fn{midst...: Heb. heart of the seas.}
\vs Psa 46:3 \bibemph{Though} the waters thereof roar \bibemph{and} be troubled, \bibemph{though} the mountains shake with the swelling thereof. Selah.
\vs Psa 46:4 \bibemph{There is} a river, the streams whereof shall make glad the city of God, the holy \bibemph{place} of the tabernacles of the most High.
\vs Psa 46:5 God \bibemph{is} in the midst of her; she shall not be moved: God shall help her, \bibemph{and that} right early.
\fn{and...: Heb. when the morning appeareth.}
\vs Psa 46:6 The heathen raged, the kingdoms were moved: he uttered his voice, the earth melted.
\vs Psa 46:7 The \textsc{Lord} of hosts \bibemph{is} with us; the God of Jacob \bibemph{is} our refuge. Selah.
\fn{our...: Heb. an high place for us.}
\vs Psa 46:8 Come, behold the works of the \textsc{Lord}, what desolations he hath made in the earth.
\vs Psa 46:9 He maketh wars to cease unto the end of the earth; he breaketh the bow, and cutteth the spear in sunder; he burneth the chariot in the fire.
\vs Psa 46:10 Be still, and know that I \bibemph{am} God: I will be exalted among the heathen, I will be exalted in the earth.
\vs Psa 46:11 The \textsc{Lord} of hosts \bibemph{is} with us; the God of Jacob \bibemph{is} our refuge. Selah.
\vs Psa 47:1 To the chief Musician, A Psalm for the sons of Korah. O clap your hands, all ye people; shout unto God with the voice of triumph.
\fn{for: or, of.}
\vs Psa 47:2 For the \textsc{Lord} most high \bibemph{is} terrible; \bibemph{he is} a great King over all the earth.
\vs Psa 47:3 He shall subdue the people under us, and the nations under our feet.
\vs Psa 47:4 He shall choose our inheritance for us, the excellency of Jacob whom he loved. Selah.
\vs Psa 47:5 God is gone up with a shout, the \textsc{Lord} with the sound of a trumpet.
\vs Psa 47:6 Sing praises to God, sing praises: sing praises unto our King, sing praises.
\vs Psa 47:7 For God \bibemph{is} the King of all the earth: sing ye praises with understanding.
\fn{with: or, every one that hath.}
\vs Psa 47:8 God reigneth over the heathen: God sitteth upon the throne of his holiness.
\vs Psa 47:9 The princes of the people are gathered together, \bibemph{even} the people of the God of Abraham: for the shields of the earth \bibemph{belong} unto God: he is greatly exalted.
\fn{princes...: or, voluntary of the people are gathered unto the people of the God of Abraham.}
\vs Psa 48:1 A Song \bibemph{and} Psalm for the sons of Korah. Great \bibemph{is} the \textsc{Lord}, and greatly to be praised in the city of our God, \bibemph{in} the mountain of his holiness.
\fn{for: or, of.}
\vs Psa 48:2 Beautiful for situation, the joy of the whole earth, \bibemph{is} mount Zion, \bibemph{on} the sides of the north, the city of the great King.
\vs Psa 48:3 God is known in her palaces for a refuge.
\vs Psa 48:4 For, lo, the kings were assembled, they passed by together.
\vs Psa 48:5 They saw \bibemph{it, and} so they marvelled; they were troubled, \bibemph{and} hasted away.
\vs Psa 48:6 Fear took hold upon them there, \bibemph{and} pain, as of a woman in travail.
\vs Psa 48:7 Thou breakest the ships of Tarshish with an east wind.
\vs Psa 48:8 As we have heard, so have we seen in the city of the \textsc{Lord} of hosts, in the city of our God: God will establish it for ever. Selah.
\vs Psa 48:9 We have thought of thy lovingkindness, O God, in the midst of thy temple.
\vs Psa 48:10 According to thy name, O God, so \bibemph{is} thy praise unto the ends of the earth: thy right hand is full of righteousness.
\vs Psa 48:11 Let mount Zion rejoice, let the daughters of Judah be glad, because of thy judgments.
\vs Psa 48:12 Walk about Zion, and go round about her: tell the towers thereof.
\vs Psa 48:13 Mark ye well her bulwarks, consider her palaces; that ye may tell \bibemph{it} to the generation following.
\fn{Mark...: Heb. Set your heart to. consider: or, raise up.}
\vs Psa 48:14 For this God \bibemph{is} our God for ever and ever: he will be our guide \bibemph{even} unto death.
\vs Psa 49:1 To the chief Musician, A Psalm for the sons of Korah. Hear this, all \bibemph{ye} people; give ear, all \bibemph{ye} inhabitants of the world:
\fn{for: or, of.}
\vs Psa 49:2 Both low and high, rich and poor, together.
\vs Psa 49:3 My mouth shall speak of wisdom; and the meditation of my heart \bibemph{shall be} of understanding.
\vs Psa 49:4 I will incline mine ear to a parable: I will open my dark saying upon the harp.
\vs Psa 49:5 Wherefore should I fear in the days of evil, \bibemph{when} the iniquity of my heels shall compass me about?
\vs Psa 49:6 They that trust in their wealth, and boast themselves in the multitude of their riches;
\vs Psa 49:7 None \bibemph{of them} can by any means redeem his brother, nor give to God a ransom for him:
\vs Psa 49:8 (For the redemption of their soul \bibemph{is} precious, and it ceaseth for ever:)
\vs Psa 49:9 That he should still live for ever, \bibemph{and} not see corruption.
\vs Psa 49:10 For he seeth \bibemph{that} wise men die, likewise the fool and the brutish person perish, and leave their wealth to others.
\vs Psa 49:11 Their inward thought \bibemph{is, that} their houses \bibemph{shall continue} for ever, \bibemph{and} their dwelling places to all generations; they call \bibemph{their} lands after their own names.
\fn{all...: Heb. generation and generation.}
\vs Psa 49:12 Nevertheless man \bibemph{being} in honour abideth not: he is like the beasts \bibemph{that} perish.
\vs Psa 49:13 This their way \bibemph{is} their folly: yet their posterity approve their sayings. Selah.
\fn{approve...: Heb. delight in their mouth.}
\vs Psa 49:14 Like sheep they are laid in the grave; death shall feed on them; and the upright shall have dominion over them in the morning; and their beauty shall consume in the grave from their dwelling.
\fn{beauty: or, strength. in the grave from...: or, the grave being an habitation to every one of them.}
\vs Psa 49:15 But God will redeem my soul from the power of the grave: for he shall receive me. Selah.
\fn{power: Heb. hand. the grave: or, hell.}
\vs Psa 49:16 Be not thou afraid when one is made rich, when the glory of his house is increased;
\vs Psa 49:17 For when he dieth he shall carry nothing away: his glory shall not descend after him.
\vs Psa 49:18 Though while he lived he blessed his soul: and \bibemph{men} will praise thee, when thou doest well to thyself.
\fn{while...: Heb. in his life.}
\vs Psa 49:19 He shall go to the generation of his fathers; they shall never see light.
\fn{He...: Heb. The soul shall.}
\vs Psa 49:20 Man \bibemph{that is} in honour, and understandeth not, is like the beasts \bibemph{that} perish.
\vs Psa 50:1 A Psalm of Asaph. The mighty God, \bibemph{even} the \textsc{Lord}, hath spoken, and called the earth from the rising of the sun unto the going down thereof.
\fn{of...: or, for Asaph.}
\vs Psa 50:2 Out of Zion, the perfection of beauty, God hath shined.
\vs Psa 50:3 Our God shall come, and shall not keep silence: a fire shall devour before him, and it shall be very tempestuous round about him.
\vs Psa 50:4 He shall call to the heavens from above, and to the earth, that he may judge his people.
\vs Psa 50:5 Gather my saints together unto me; those that have made a covenant with me by sacrifice.
\vs Psa 50:6 And the heavens shall declare his righteousness: for God \bibemph{is} judge himself. Selah.
\vs Psa 50:7 Hear, O my people, and I will speak; O Israel, and I will testify against thee: I \bibemph{am} God, \bibemph{even} thy God.
\vs Psa 50:8 I will not reprove thee for thy sacrifices or thy burnt offerings, \bibemph{to have been} continually before me.
\vs Psa 50:9 I will take no bullock out of thy house, \bibemph{nor} he goats out of thy folds.
\vs Psa 50:10 For every beast of the forest \bibemph{is} mine, \bibemph{and} the cattle upon a thousand hills.
\vs Psa 50:11 I know all the fowls of the mountains: and the wild beasts of the field \bibemph{are} mine.
\fn{mine: Heb. with me.}
\vs Psa 50:12 If I were hungry, I would not tell thee: for the world \bibemph{is} mine, and the fulness thereof.
\vs Psa 50:13 Will I eat the flesh of bulls, or drink the blood of goats?
\vs Psa 50:14 Offer unto God thanksgiving; and pay thy vows unto the most High:
\vs Psa 50:15 And call upon me in the day of trouble: I will deliver thee, and thou shalt glorify me.
\vs Psa 50:16 But unto the wicked God saith, What hast thou to do to declare my statutes, or \bibemph{that} thou shouldest take my covenant in thy mouth?
\vs Psa 50:17 Seeing thou hatest instruction, and castest my words behind thee.
\vs Psa 50:18 When thou sawest a thief, then thou consentedst with him, and hast been partaker with adulterers.
\fn{hast...: Heb. thy portion was with.}
\vs Psa 50:19 Thou givest thy mouth to evil, and thy tongue frameth deceit.
\fn{givest: Heb. sendest.}
\vs Psa 50:20 Thou sittest \bibemph{and} speakest against thy brother; thou slanderest thine own mother's son.
\vs Psa 50:21 These \bibemph{things} hast thou done, and I kept silence; thou thoughtest that I was altogether \bibemph{such an one} as thyself: \bibemph{but} I will reprove thee, and set \bibemph{them} in order before thine eyes.
\vs Psa 50:22 Now consider this, ye that forget God, lest I tear \bibemph{you} in pieces, and \bibemph{there be} none to deliver.
\vs Psa 50:23 Whoso offereth praise glorifieth me: and to him that ordereth \bibemph{his} conversation \bibemph{aright} will I shew the salvation of God.
\fn{that...: Heb. that disposeth his way.}
\vs Psa 51:1 To the chief Musician, A Psalm of David, when Nathan the prophet came unto him, after he had gone in to Bathsheba. Have mercy upon me, O God, according to thy lovingkindness: according unto the multitude of thy tender mercies blot out my transgressions.
\vs Psa 51:2 Wash me throughly from mine iniquity, and cleanse me from my sin.
\vs Psa 51:3 For I acknowledge my transgressions: and my sin \bibemph{is} ever before me.
\vs Psa 51:4 Against thee, thee only, have I sinned, and done \bibemph{this} evil in thy sight: that thou mightest be justified when thou speakest, \bibemph{and} be clear when thou judgest.
\vs Psa 51:5 Behold, I was shapen in iniquity; and in sin did my mother conceive me.
\fn{conceive...: Heb. warm me.}
\vs Psa 51:6 Behold, thou desirest truth in the inward parts: and in the hidden \bibemph{part} thou shalt make me to know wisdom.
\vs Psa 51:7 Purge me with hyssop, and I shall be clean: wash me, and I shall be whiter than snow.
\vs Psa 51:8 Make me to hear joy and gladness; \bibemph{that} the bones \bibemph{which} thou hast broken may rejoice.
\vs Psa 51:9 Hide thy face from my sins, and blot out all mine iniquities.
\vs Psa 51:10 Create in me a clean heart, O God; and renew a right spirit within me.
\fn{right: or, constant.}
\vs Psa 51:11 Cast me not away from thy presence; and take not thy holy spirit from me.
\vs Psa 51:12 Restore unto me the joy of thy salvation; and uphold me \bibemph{with thy} free spirit.
\vs Psa 51:13 \bibemph{Then} will I teach transgressors thy ways; and sinners shall be converted unto thee.
\vs Psa 51:14 Deliver me from bloodguiltiness, O God, thou God of my salvation: \bibemph{and} my tongue shall sing aloud of thy righteousness.
\fn{bloodguiltiness: Heb. bloods.}
\vs Psa 51:15 O Lord, open thou my lips; and my mouth shall shew forth thy praise.
\vs Psa 51:16 For thou desirest not sacrifice; else would I give \bibemph{it}: thou delightest not in burnt offering.
\fn{else...: or, that I should.}
\vs Psa 51:17 The sacrifices of God \bibemph{are} a broken spirit: a broken and a contrite heart, O God, thou wilt not despise.
\vs Psa 51:18 Do good in thy good pleasure unto Zion: build thou the walls of Jerusalem.
\vs Psa 51:19 Then shalt thou be pleased with the sacrifices of righteousness, with burnt offering and whole burnt offering: then shall they offer bullocks upon thine altar.
\vs Psa 52:1 To the chief Musician, Maschil, \bibemph{A Psalm} of David, when Doeg the Edomite came and told Saul, and said unto him, David is come to the house of Ahimelech. Why boastest thou thyself in mischief, O mighty man? the goodness of God \bibemph{endureth} continually.
\vs Psa 52:2 Thy tongue deviseth mischiefs; like a sharp razor, working deceitfully.
\vs Psa 52:3 Thou lovest evil more than good; \bibemph{and} lying rather than to speak righteousness. Selah.
\vs Psa 52:4 Thou lovest all devouring words, O \bibemph{thou} deceitful tongue.
\fn{O thou...: or, and the deceitful tongue.}
\vs Psa 52:5 God shall likewise destroy thee for ever, he shall take thee away, and pluck thee out of \bibemph{thy} dwelling place, and root thee out of the land of the living. Selah.
\fn{destroy...: Heb. beat thee down.}
\vs Psa 52:6 The righteous also shall see, and fear, and shall laugh at him:
\vs Psa 52:7 Lo, \bibemph{this is} the man \bibemph{that} made not God his strength; but trusted in the abundance of his riches, \bibemph{and} strengthened himself in his wickedness.
\fn{wickedness: or, substance.}
\vs Psa 52:8 But I \bibemph{am} like a green olive tree in the house of God: I trust in the mercy of God for ever and ever.
\vs Psa 52:9 I will praise thee for ever, because thou hast done \bibemph{it}: and I will wait on thy name; for \bibemph{it is} good before thy saints.
\vs Psa 53:1 To the chief Musician upon Mahalath, Maschil, \bibemph{A Psalm} of David. The fool hath said in his heart, \bibemph{There is} no God. Corrupt are they, and have done abominable iniquity: \bibemph{there is} none that doeth good.
\vs Psa 53:2 God looked down from heaven upon the children of men, to see if there were \bibemph{any} that did understand, that did seek God.
\vs Psa 53:3 Every one of them is gone back: they are altogether become filthy; \bibemph{there is} none that doeth good, no, not one.
\vs Psa 53:4 Have the workers of iniquity no knowledge? who eat up my people \bibemph{as} they eat bread: they have not called upon God.
\vs Psa 53:5 There were they in great fear, \bibemph{where} no fear was: for God hath scattered the bones of him that encampeth \bibemph{against} thee: thou hast put \bibemph{them} to shame, because God hath despised them.
\fn{were...: Heb. they feared a fear.}
\vs Psa 53:6 Oh that the salvation of Israel \bibemph{were come} out of Zion! When God bringeth back the captivity of his people, Jacob shall rejoice, \bibemph{and} Israel shall be glad.
\fn{Oh that...: Heb. Who will give salvation, etc.}
\vs Psa 54:1 To the chief Musician on Neginoth, Maschil, \bibemph{A Psalm} of David, when the Ziphims came and said to Saul, Doth not David hide himself with us? Save me, O God, by thy name, and judge me by thy strength.
\vs Psa 54:2 Hear my prayer, O God; give ear to the words of my mouth.
\vs Psa 54:3 For strangers are risen up against me, and oppressors seek after my soul: they have not set God before them. Selah.
\vs Psa 54:4 Behold, God \bibemph{is} mine helper: the Lord \bibemph{is} with them that uphold my soul.
\vs Psa 54:5 He shall reward evil unto mine enemies: cut them off in thy truth.
\fn{mine...: Heb. those that observe me.}
\vs Psa 54:6 I will freely sacrifice unto thee: I will praise thy name, O \textsc{Lord}; for \bibemph{it is} good.
\vs Psa 54:7 For he hath delivered me out of all trouble: and mine eye hath seen \bibemph{his desire} upon mine enemies.
\vs Psa 55:1 To the chief Musician on Neginoth, Maschil, \bibemph{A Psalm} of David. Give ear to my prayer, O God; and hide not thyself from my supplication.
\vs Psa 55:2 Attend unto me, and hear me: I mourn in my complaint, and make a noise;
\vs Psa 55:3 Because of the voice of the enemy, because of the oppression of the wicked: for they cast iniquity upon me, and in wrath they hate me.
\vs Psa 55:4 My heart is sore pained within me: and the terrors of death are fallen upon me.
\vs Psa 55:5 Fearfulness and trembling are come upon me, and horror hath overwhelmed me.
\fn{overwhelmed: Heb. covered.}
\vs Psa 55:6 And I said, Oh that I had wings like a dove! \bibemph{for then} would I fly away, and be at rest.
\vs Psa 55:7 Lo, \bibemph{then} would I wander far off, \bibemph{and} remain in the wilderness. Selah.
\vs Psa 55:8 I would hasten my escape from the windy storm \bibemph{and} tempest.
\vs Psa 55:9 Destroy, O Lord, \bibemph{and} divide their tongues: for I have seen violence and strife in the city.
\vs Psa 55:10 Day and night they go about it upon the walls thereof: mischief also and sorrow \bibemph{are} in the midst of it.
\vs Psa 55:11 Wickedness \bibemph{is} in the midst thereof: deceit and guile depart not from her streets.
\vs Psa 55:12 For \bibemph{it was} not an enemy \bibemph{that} reproached me; then I could have borne \bibemph{it}: neither \bibemph{was it} he that hated me \bibemph{that} did magnify \bibemph{himself} against me; then I would have hid myself from him:
\vs Psa 55:13 But \bibemph{it was} thou, a man mine equal, my guide, and mine acquaintance.
\fn{mine equal: Heb. according to my rank.}
\vs Psa 55:14 We took sweet counsel together, \bibemph{and} walked unto the house of God in company.
\fn{We...: Heb. Who sweetened counsel.}
\vs Psa 55:15 Let death seize upon them, \bibemph{and} let them go down quick into hell: for wickedness \bibemph{is} in their dwellings, \bibemph{and} among them.
\fn{hell: or, the grave.}
\vs Psa 55:16 As for me, I will call upon God; and the \textsc{Lord} shall save me.
\vs Psa 55:17 Evening, and morning, and at noon, will I pray, and cry aloud: and he shall hear my voice.
\vs Psa 55:18 He hath delivered my soul in peace from the battle \bibemph{that was} against me: for there were many with me.
\vs Psa 55:19 God shall hear, and afflict them, even he that abideth of old. Selah. Because they have no changes, therefore they fear not God.
\fn{Because...: or, With whom also there be no changes, yet they.}
\vs Psa 55:20 He hath put forth his hands against such as be at peace with him: he hath broken his covenant.
\fn{broken: Heb. profaned.}
\vs Psa 55:21 \bibemph{The words} of his mouth were smoother than butter, but war \bibemph{was} in his heart: his words were softer than oil, yet \bibemph{were} they drawn swords.
\vs Psa 55:22 Cast thy burden upon the \textsc{Lord}, and he shall sustain thee: he shall never suffer the righteous to be moved.
\fn{burden: or, gift.}
\vs Psa 55:23 But thou, O God, shalt bring them down into the pit of destruction: bloody and deceitful men shall not live out half their days; but I will trust in thee.
\fn{bloody...: Heb. men of bloods and deceit. shall...: Heb. shall not half their days.}
\vs Psa 56:1 To the chief Musician upon Jonathelemrechokim, Michtam of David, when the Philistines took him in Gath. Be merciful unto me, O God: for man would swallow me up; he fighting daily oppresseth me.
\fn{Michtam...: or, A golden Psalm of David.}
\vs Psa 56:2 Mine enemies would daily swallow \bibemph{me} up: for \bibemph{they be} many that fight against me, O thou most High.
\fn{enemies: Heb. observers.}
\vs Psa 56:3 What time I am afraid, I will trust in thee.
\vs Psa 56:4 In God I will praise his word, in God I have put my trust; I will not fear what flesh can do unto me.
\vs Psa 56:5 Every day they wrest my words: all their thoughts \bibemph{are} against me for evil.
\vs Psa 56:6 They gather themselves together, they hide themselves, they mark my steps, when they wait for my soul.
\vs Psa 56:7 Shall they escape by iniquity? in \bibemph{thine} anger cast down the people, O God.
\vs Psa 56:8 Thou tellest my wanderings: put thou my tears into thy bottle: \bibemph{are they} not in thy book?
\vs Psa 56:9 When I cry \bibemph{unto thee}, then shall mine enemies turn back: this I know; for God \bibemph{is} for me.
\vs Psa 56:10 In God will I praise \bibemph{his} word: in the \textsc{Lord} will I praise \bibemph{his} word.
\vs Psa 56:11 In God have I put my trust: I will not be afraid what man can do unto me.
\vs Psa 56:12 Thy vows \bibemph{are} upon me, O God: I will render praises unto thee.
\vs Psa 56:13 For thou hast delivered my soul from death: \bibemph{wilt} not \bibemph{thou deliver} my feet from falling, that I may walk before God in the light of the living?
\vs Psa 57:1 To the chief Musician, Altaschith, Michtam of David, when he fled from Saul in the cave. Be merciful unto me, O God, be merciful unto me: for my soul trusteth in thee: yea, in the shadow of thy wings will I make my refuge, until \bibemph{these} calamities be overpast.
\fn{Altaschith: or, Destroy not.}
\vs Psa 57:2 I will cry unto God most high; unto God that performeth \bibemph{all things} for me.
\vs Psa 57:3 He shall send from heaven, and save me \bibemph{from} the reproach of him that would swallow me up. Selah. God shall send forth his mercy and his truth.
\fn{from the...: or, he reproacheth him that.}
\vs Psa 57:4 My soul \bibemph{is} among lions: \bibemph{and} I lie \bibemph{even among} them that are set on fire, \bibemph{even} the sons of men, whose teeth \bibemph{are} spears and arrows, and their tongue a sharp sword.
\vs Psa 57:5 Be thou exalted, O God, above the heavens; \bibemph{let} thy glory \bibemph{be} above all the earth.
\vs Psa 57:6 They have prepared a net for my steps; my soul is bowed down: they have digged a pit before me, into the midst whereof they are fallen \bibemph{themselves}. Selah.
\vs Psa 57:7 My heart is fixed, O God, my heart is fixed: I will sing and give praise.
\fn{fixed: or, prepared.}
\vs Psa 57:8 Awake up, my glory; awake, psaltery and harp: I \bibemph{myself} will awake early.
\vs Psa 57:9 I will praise thee, O Lord, among the people: I will sing unto thee among the nations.
\vs Psa 57:10 For thy mercy \bibemph{is} great unto the heavens, and thy truth unto the clouds.
\vs Psa 57:11 Be thou exalted, O God, above the heavens: \bibemph{let} thy glory \bibemph{be} above all the earth.
\vs Psa 58:1 To the chief Musician, Altaschith, Michtam of David. Do ye indeed speak righteousness, O congregation? do ye judge uprightly, O ye sons of men?
\fn{Altaschith...: or, Destroy not, A golden Psalm of David.}
\vs Psa 58:2 Yea, in heart ye work wickedness; ye weigh the violence of your hands in the earth.
\vs Psa 58:3 The wicked are estranged from the womb: they go astray as soon as they be born, speaking lies.
\fn{as soon...: Heb. from the belly.}
\vs Psa 58:4 Their poison \bibemph{is} like the poison of a serpent: \bibemph{they are} like the deaf adder \bibemph{that} stoppeth her ear;
\fn{like the poison: Heb. according to the likeness, etc. adder: or, asp.}
\vs Psa 58:5 Which will not hearken to the voice of charmers, charming never so wisely.
\fn{charming...: or, be the charmer never so cunning.}
\vs Psa 58:6 Break their teeth, O God, in their mouth: break out the great teeth of the young lions, O \textsc{Lord}.
\vs Psa 58:7 Let them melt away as waters \bibemph{which} run continually: \bibemph{when} he bendeth \bibemph{his bow to shoot} his arrows, let them be as cut in pieces.
\vs Psa 58:8 As a snail \bibemph{which} melteth, let \bibemph{every one of them} pass away: \bibemph{like} the untimely birth of a woman, \bibemph{that} they may not see the sun.
\vs Psa 58:9 Before your pots can feel the thorns, he shall take them away as with a whirlwind, both living, and in \bibemph{his} wrath.
\fn{both...: Heb. as living as wrath.}
\vs Psa 58:10 The righteous shall rejoice when he seeth the vengeance: he shall wash his feet in the blood of the wicked.
\vs Psa 58:11 So that a man shall say, Verily \bibemph{there is} a reward for the righteous: verily he is a God that judgeth in the earth.
\fn{a reward...: Heb. fruit of the, etc.}
\vs Psa 59:1 To the chief Musician, Altaschith, Michtam of David; when Saul sent, and they watched the house to kill him. Deliver me from mine enemies, O my God: defend me from them that rise up against me.
\fn{Altaschith...: or, Destroy not, A golden Psalm of David. defend...: Heb. set me on high.}
\vs Psa 59:2 Deliver me from the workers of iniquity, and save me from bloody men.
\vs Psa 59:3 For, lo, they lie in wait for my soul: the mighty are gathered against me; not \bibemph{for} my transgression, nor \bibemph{for} my sin, O \textsc{Lord}.
\vs Psa 59:4 They run and prepare themselves without \bibemph{my} fault: awake to help me, and behold.
\fn{help: Heb. meet.}
\vs Psa 59:5 Thou therefore, O \textsc{Lord} God of hosts, the God of Israel, awake to visit all the heathen: be not merciful to any wicked transgressors. Selah.
\vs Psa 59:6 They return at evening: they make a noise like a dog, and go round about the city.
\vs Psa 59:7 Behold, they belch out with their mouth: swords \bibemph{are} in their lips: for who, \bibemph{say they}, doth hear?
\vs Psa 59:8 But thou, O \textsc{Lord}, shalt laugh at them; thou shalt have all the heathen in derision.
\vs Psa 59:9 \bibemph{Because of} his strength will I wait upon thee: for God \bibemph{is} my defence.
\fn{defence: Heb. high place.}
\vs Psa 59:10 The God of my mercy shall prevent me: God shall let me see \bibemph{my desire} upon mine enemies.
\fn{enemies: Heb. observers.}
\vs Psa 59:11 Slay them not, lest my people forget: scatter them by thy power; and bring them down, O Lord our shield.
\vs Psa 59:12 \bibemph{For} the sin of their mouth \bibemph{and} the words of their lips let them even be taken in their pride: and for cursing and lying \bibemph{which} they speak.
\vs Psa 59:13 Consume \bibemph{them} in wrath, consume \bibemph{them}, that they \bibemph{may} not \bibemph{be}: and let them know that God ruleth in Jacob unto the ends of the earth. Selah.
\vs Psa 59:14 And at evening let them return; \bibemph{and} let them make a noise like a dog, and go round about the city.
\vs Psa 59:15 Let them wander up and down for meat, and grudge if they be not satisfied.
\fn{for meat: Heb. to eat. and grudge...: or, if they be not satisfied, then they will stay all night.}
\vs Psa 59:16 But I will sing of thy power; yea, I will sing aloud of thy mercy in the morning: for thou hast been my defence and refuge in the day of my trouble.
\vs Psa 59:17 Unto thee, O my strength, will I sing: for God \bibemph{is} my defence, \bibemph{and} the God of my mercy.
\vs Psa 60:1 To the chief Musician upon Shushaneduth, Michtam of David, to teach; when he strove with Aramnaharaim and with Aramzobah, when Joab returned, and smote of Edom in the valley of salt twelve thousand. O God, thou hast cast us off, thou hast scattered us, thou hast been displeased; O turn thyself to us again.
\fn{Michtam: or, A golden Psalm. scattered: Heb. broken.}
\vs Psa 60:2 Thou hast made the earth to tremble; thou hast broken it: heal the breaches thereof; for it shaketh.
\vs Psa 60:3 Thou hast shewed thy people hard things: thou hast made us to drink the wine of astonishment.
\vs Psa 60:4 Thou hast given a banner to them that fear thee, that it may be displayed because of the truth. Selah.
\vs Psa 60:5 That thy beloved may be delivered; save \bibemph{with} thy right hand, and hear me.
\vs Psa 60:6 God hath spoken in his holiness; I will rejoice, I will divide Shechem, and mete out the valley of Succoth.
\vs Psa 60:7 Gilead \bibemph{is} mine, and Manasseh \bibemph{is} mine; Ephraim also \bibemph{is} the strength of mine head; Judah \bibemph{is} my lawgiver;
\vs Psa 60:8 Moab \bibemph{is} my washpot; over Edom will I cast out my shoe: Philistia, triumph thou because of me.
\fn{triumph...: or, triumph thou over me: (by an irony).}
\vs Psa 60:9 Who will bring me \bibemph{into} the strong city? who will lead me into Edom?
\fn{strong...: Heb. city of strength?.}
\vs Psa 60:10 \bibemph{Wilt} not thou, O God, \bibemph{which} hadst cast us off? and \bibemph{thou}, O God, \bibemph{which} didst not go out with our armies?
\vs Psa 60:11 Give us help from trouble: for vain \bibemph{is} the help of man.
\fn{help of man: Heb. salvation, etc.}
\vs Psa 60:12 Through God we shall do valiantly: for he \bibemph{it is that} shall tread down our enemies.
\vs Psa 61:1 To the chief Musician upon Neginah, \bibemph{A Psalm} of David. Hear my cry, O God; attend unto my prayer.
\vs Psa 61:2 From the end of the earth will I cry unto thee, when my heart is overwhelmed: lead me to the rock \bibemph{that} is higher than I.
\vs Psa 61:3 For thou hast been a shelter for me, \bibemph{and} a strong tower from the enemy.
\vs Psa 61:4 I will abide in thy tabernacle for ever: I will trust in the covert of thy wings. Selah.
\fn{trust: or, make my refuge.}
\vs Psa 61:5 For thou, O God, hast heard my vows: thou hast given \bibemph{me} the heritage of those that fear thy name.
\vs Psa 61:6 Thou wilt prolong the king's life: \bibemph{and} his years as many generations.
\fn{Thou...: Heb. Thou shalt add days to the days of the king. as...: Heb. as generation and generation.}
\vs Psa 61:7 He shall abide before God for ever: O prepare mercy and truth, \bibemph{which} may preserve him.
\vs Psa 61:8 So will I sing praise unto thy name for ever, that I may daily perform my vows.
\vs Psa 62:1 To the chief Musician, to Jeduthun, A Psalm of David. Truly my soul waiteth upon God: from him \bibemph{cometh} my salvation.
\fn{Truly: or, Only. waiteth: Heb. is silent.}
\vs Psa 62:2 He only \bibemph{is} my rock and my salvation; \bibemph{he is} my defence; I shall not be greatly moved.
\fn{defence: Heb. high place.}
\vs Psa 62:3 How long will ye imagine mischief against a man? ye shall be slain all of you: as a bowing wall \bibemph{shall ye be, and as} a tottering fence.
\vs Psa 62:4 They only consult to cast \bibemph{him} down from his excellency: they delight in lies: they bless with their mouth, but they curse inwardly. Selah.
\fn{inwardly: Heb. in their inward parts.}
\vs Psa 62:5 My soul, wait thou only upon God; for my expectation \bibemph{is} from him.
\vs Psa 62:6 He only \bibemph{is} my rock and my salvation: \bibemph{he is} my defence; I shall not be moved.
\vs Psa 62:7 In God \bibemph{is} my salvation and my glory: the rock of my strength, \bibemph{and} my refuge, \bibemph{is} in God.
\vs Psa 62:8 Trust in him at all times; \bibemph{ye} people, pour out your heart before him: God \bibemph{is} a refuge for us. Selah.
\vs Psa 62:9 Surely men of low degree \bibemph{are} vanity, \bibemph{and} men of high degree \bibemph{are} a lie: to be laid in the balance, they \bibemph{are} altogether \bibemph{lighter} than vanity.
\fn{altogether: or, alike.}
\vs Psa 62:10 Trust not in oppression, and become not vain in robbery: if riches increase, set not your heart \bibemph{upon them}.
\vs Psa 62:11 God hath spoken once; twice have I heard this; that power \bibemph{belongeth} unto God.
\fn{power: or, strength.}
\vs Psa 62:12 Also unto thee, O Lord, \bibemph{belongeth} mercy: for thou renderest to every man according to his work.
\vs Psa 63:1 A Psalm of David, when he was in the wilderness of Judah. O God, thou \bibemph{art} my God; early will I seek thee: my soul thirsteth for thee, my flesh longeth for thee in a dry and thirsty land, where no water is;
\fn{thirsty: Heb. weary. where...: without water.}
\vs Psa 63:2 To see thy power and thy glory, so \bibemph{as} I have seen thee in the sanctuary.
\vs Psa 63:3 Because thy lovingkindness \bibemph{is} better than life, my lips shall praise thee.
\vs Psa 63:4 Thus will I bless thee while I live: I will lift up my hands in thy name.
\vs Psa 63:5 My soul shall be satisfied as \bibemph{with} marrow and fatness; and my mouth shall praise \bibemph{thee} with joyful lips:
\fn{marrow: Heb. fatness.}
\vs Psa 63:6 When I remember thee upon my bed, \bibemph{and} meditate on thee in the \bibemph{night} watches.
\vs Psa 63:7 Because thou hast been my help, therefore in the shadow of thy wings will I rejoice.
\vs Psa 63:8 My soul followeth hard after thee: thy right hand upholdeth me.
\vs Psa 63:9 But those \bibemph{that} seek my soul, to destroy \bibemph{it}, shall go into the lower parts of the earth.
\vs Psa 63:10 They shall fall by the sword: they shall be a portion for foxes.
\fn{They shall fall...: Heb. They shall make him run out like water by the hands of.}
\vs Psa 63:11 But the king shall rejoice in God; every one that sweareth by him shall glory: but the mouth of them that speak lies shall be stopped.
\vs Psa 64:1 To the chief Musician, A Psalm of David. Hear my voice, O God, in my prayer: preserve my life from fear of the enemy.
\vs Psa 64:2 Hide me from the secret counsel of the wicked; from the insurrection of the workers of iniquity:
\vs Psa 64:3 Who whet their tongue like a sword, \bibemph{and} bend \bibemph{their bows to shoot} their arrows, \bibemph{even} bitter words:
\vs Psa 64:4 That they may shoot in secret at the perfect: suddenly do they shoot at him, and fear not.
\vs Psa 64:5 They encourage themselves \bibemph{in} an evil matter: they commune of laying snares privily; they say, Who shall see them?
\fn{matter: or, speech. of...: Heb. to hide his snares.}
\vs Psa 64:6 They search out iniquities; they accomplish a diligent search: both the inward \bibemph{thought} of every one \bibemph{of them}, and the heart, \bibemph{is} deep.
\fn{they...: or, we are consumed by that which they have throughly searched. a diligent...: Heb. a search searched.}
\vs Psa 64:7 But God shall shoot at them \bibemph{with} an arrow; suddenly shall they be wounded.
\fn{shall they...: Heb. their wound shall be.}
\vs Psa 64:8 So they shall make their own tongue to fall upon themselves: all that see them shall flee away.
\vs Psa 64:9 And all men shall fear, and shall declare the work of God; for they shall wisely consider of his doing.
\vs Psa 64:10 The righteous shall be glad in the \textsc{Lord}, and shall trust in him; and all the upright in heart shall glory.
\vs Psa 65:1 To the chief Musician, A Psalm \bibemph{and} Song of David. Praise waiteth for thee, O God, in Sion: and unto thee shall the vow be performed.
\fn{waiteth: Heb. is silent.}
\vs Psa 65:2 O thou that hearest prayer, unto thee shall all flesh come.
\vs Psa 65:3 Iniquities prevail against me: \bibemph{as for} our transgressions, thou shalt purge them away.
\fn{iniquities: Heb. Words, or, Matters of iniquities.}
\vs Psa 65:4 Blessed \bibemph{is the man whom} thou choosest, and causest to approach \bibemph{unto thee, that} he may dwell in thy courts: we shall be satisfied with the goodness of thy house, \bibemph{even} of thy holy temple.
\vs Psa 65:5 \bibemph{By} terrible things in righteousness wilt thou answer us, O God of our salvation; \bibemph{who art} the confidence of all the ends of the earth, and of them that are afar off \bibemph{upon} the sea:
\vs Psa 65:6 Which by his strength setteth fast the mountains; \bibemph{being} girded with power:
\vs Psa 65:7 Which stilleth the noise of the seas, the noise of their waves, and the tumult of the people.
\vs Psa 65:8 They also that dwell in the uttermost parts are afraid at thy tokens: thou makest the outgoings of the morning and evening to rejoice.
\fn{rejoice: or, sing.}
\vs Psa 65:9 Thou visitest the earth, and waterest it: thou greatly enrichest it with the river of God, \bibemph{which} is full of water: thou preparest them corn, when thou hast so provided for it.
\fn{and...: or, after thou hadst made it to desire rain.}
\vs Psa 65:10 Thou waterest the ridges thereof abundantly: thou settlest the furrows thereof: thou makest it soft with showers: thou blessest the springing thereof.
\fn{settlest: or, causest rain to descend into. makest...: Heb. dissolvest it.}
\vs Psa 65:11 Thou crownest the year with thy goodness; and thy paths drop fatness.
\fn{with: Heb. of.}
\vs Psa 65:12 They drop \bibemph{upon} the pastures of the wilderness: and the little hills rejoice on every side.
\fn{rejoice...: Heb. are girded with joy.}
\vs Psa 65:13 The pastures are clothed with flocks; the valleys also are covered over with corn; they shout for joy, they also sing.
\vs Psa 66:1 To the chief Musician, A Song \bibemph{or} Psalm. Make a joyful noise unto God, all ye lands:
\fn{all...: Heb. all the earth.}
\vs Psa 66:2 Sing forth the honour of his name: make his praise glorious.
\vs Psa 66:3 Say unto God, How terrible \bibemph{art thou in} thy works! through the greatness of thy power shall thine enemies submit themselves unto thee.
\fn{submit...: or, yield feigned obedience: Heb. lie.}
\vs Psa 66:4 All the earth shall worship thee, and shall sing unto thee; they shall sing \bibemph{to} thy name. Selah.
\vs Psa 66:5 Come and see the works of God: \bibemph{he is} terrible \bibemph{in his} doing toward the children of men.
\vs Psa 66:6 He turned the sea into dry \bibemph{land}: they went through the flood on foot: there did we rejoice in him.
\vs Psa 66:7 He ruleth by his power for ever; his eyes behold the nations: let not the rebellious exalt themselves. Selah.
\vs Psa 66:8 O bless our God, ye people, and make the voice of his praise to be heard:
\vs Psa 66:9 Which holdeth our soul in life, and suffereth not our feet to be moved.
\fn{holdeth: Heb. putteth.}
\vs Psa 66:10 For thou, O God, hast proved us: thou hast tried us, as silver is tried.
\vs Psa 66:11 Thou broughtest us into the net; thou laidst affliction upon our loins.
\vs Psa 66:12 Thou hast caused men to ride over our heads; we went through fire and through water: but thou broughtest us out into a wealthy \bibemph{place}.
\fn{wealthy: Heb. moist.}
\vs Psa 66:13 I will go into thy house with burnt offerings: I will pay thee my vows,
\vs Psa 66:14 Which my lips have uttered, and my mouth hath spoken, when I was in trouble.
\fn{uttered: Heb. opened.}
\vs Psa 66:15 I will offer unto thee burnt sacrifices of fatlings, with the incense of rams; I will offer bullocks with goats. Selah.
\fn{fatlings: Heb. marrow.}
\vs Psa 66:16 Come \bibemph{and} hear, all ye that fear God, and I will declare what he hath done for my soul.
\vs Psa 66:17 I cried unto him with my mouth, and he was extolled with my tongue.
\vs Psa 66:18 If I regard iniquity in my heart, the Lord will not hear \bibemph{me}:
\vs Psa 66:19 \bibemph{But} verily God hath heard \bibemph{me}; he hath attended to the voice of my prayer.
\vs Psa 66:20 Blessed \bibemph{be} God, which hath not turned away my prayer, nor his mercy from me.
\vs Psa 67:1 To the chief Musician on Neginoth, A Psalm \bibemph{or} Song. God be merciful unto us, and bless us; \bibemph{and} cause his face to shine upon us; Selah.
\fn{upon: Heb. with.}
\vs Psa 67:2 That thy way may be known upon earth, thy saving health among all nations.
\vs Psa 67:3 Let the people praise thee, O God; let all the people praise thee.
\vs Psa 67:4 O let the nations be glad and sing for joy: for thou shalt judge the people righteously, and govern the nations upon earth. Selah.
\fn{govern: Heb. lead.}
\vs Psa 67:5 Let the people praise thee, O God; let all the people praise thee.
\vs Psa 67:6 \bibemph{Then} shall the earth yield her increase; \bibemph{and} God, \bibemph{even} our own God, shall bless us.
\vs Psa 67:7 God shall bless us; and all the ends of the earth shall fear him.
\vs Psa 68:1 To the chief Musician, A Psalm \bibemph{or} Song of David. Let God arise, let his enemies be scattered: let them also that hate him flee before him.
\fn{before...: Heb. from his face.}
\vs Psa 68:2 As smoke is driven away, \bibemph{so} drive \bibemph{them} away: as wax melteth before the fire, \bibemph{so} let the wicked perish at the presence of God.
\vs Psa 68:3 But let the righteous be glad; let them rejoice before God: yea, let them exceedingly rejoice.
\fn{exceedingly...: Heb. rejoice with gladness.}
\vs Psa 68:4 Sing unto God, sing praises to his name: extol him that rideth upon the heavens by his name JAH, and rejoice before him.
\vs Psa 68:5 A father of the fatherless, and a judge of the widows, \bibemph{is} God in his holy habitation.
\vs Psa 68:6 God setteth the solitary in families: he bringeth out those which are bound with chains: but the rebellious dwell in a dry \bibemph{land}.
\fn{in families: Heb. in a house.}
\vs Psa 68:7 O God, when thou wentest forth before thy people, when thou didst march through the wilderness; Selah:
\vs Psa 68:8 The earth shook, the heavens also dropped at the presence of God: \bibemph{even} Sinai itself \bibemph{was moved} at the presence of God, the God of Israel.
\vs Psa 68:9 Thou, O God, didst send a plentiful rain, whereby thou didst confirm thine inheritance, when it was weary.
\fn{send: Heb. shake out. confirm: Heb. confirm it.}
\vs Psa 68:10 Thy congregation hath dwelt therein: thou, O God, hast prepared of thy goodness for the poor.
\vs Psa 68:11 The Lord gave the word: great \bibemph{was} the company of those that published \bibemph{it}.
\fn{company: Heb. army.}
\vs Psa 68:12 Kings of armies did flee apace: and she that tarried at home divided the spoil.
\fn{did...: Heb. did flee, did flee.}
\vs Psa 68:13 Though ye have lien among the pots, \bibemph{yet shall ye be as} the wings of a dove covered with silver, and her feathers with yellow gold.
\vs Psa 68:14 When the Almighty scattered kings in it, it was \bibemph{white} as snow in Salmon.
\fn{in it...: or, for her, she.}
\vs Psa 68:15 The hill of God \bibemph{is as} the hill of Bashan; an high hill \bibemph{as} the hill of Bashan.
\vs Psa 68:16 Why leap ye, ye high hills? \bibemph{this is} the hill \bibemph{which} God desireth to dwell in; yea, the \textsc{Lord} will dwell \bibemph{in it} for ever.
\vs Psa 68:17 The chariots of God \bibemph{are} twenty thousand, \bibemph{even} thousands of angels: the Lord \bibemph{is} among them, \bibemph{as in} Sinai, in the holy \bibemph{place}.
\fn{even...: or, even many thousands.}
\vs Psa 68:18 Thou hast ascended on high, thou hast led captivity captive: thou hast received gifts for men; yea, \bibemph{for} the rebellious also, that the \textsc{Lord} God might dwell \bibemph{among them}.
\fn{for men: Heb. in the man.}
\vs Psa 68:19 Blessed \bibemph{be} the Lord, \bibemph{who} daily loadeth us \bibemph{with benefits, even} the God of our salvation. Selah.
\vs Psa 68:20 \bibemph{He that is} our God \bibemph{is} the God of salvation; and unto GOD the Lord \bibemph{belong} the issues from death.
\vs Psa 68:21 But God shall wound the head of his enemies, \bibemph{and} the hairy scalp of such an one as goeth on still in his trespasses.
\vs Psa 68:22 The Lord said, I will bring again from Bashan, I will bring \bibemph{my people} again from the depths of the sea:
\vs Psa 68:23 That thy foot may be dipped in the blood of \bibemph{thine} enemies, \bibemph{and} the tongue of thy dogs in the same.
\fn{dipped: or, red.}
\vs Psa 68:24 They have seen thy goings, O God; \bibemph{even} the goings of my God, my King, in the sanctuary.
\vs Psa 68:25 The singers went before, the players on instruments \bibemph{followed} after; among \bibemph{them were} the damsels playing with timbrels.
\vs Psa 68:26 Bless ye God in the congregations, \bibemph{even} the Lord, from the fountain of Israel.
\fn{from...: or, ye that are of the fountain of Israel.}
\vs Psa 68:27 There \bibemph{is} little Benjamin \bibemph{with} their ruler, the princes of Judah \bibemph{and} their council, the princes of Zebulun, \bibemph{and} the princes of Naphtali.
\fn{and their...: or, with their company.}
\vs Psa 68:28 Thy God hath commanded thy strength: strengthen, O God, that which thou hast wrought for us.
\vs Psa 68:29 Because of thy temple at Jerusalem shall kings bring presents unto thee.
\vs Psa 68:30 Rebuke the company of spearmen, the multitude of the bulls, with the calves of the people, \bibemph{till every one} submit himself with pieces of silver: scatter thou the people \bibemph{that} delight in war.
\fn{the company...: or, the beasts of the reeds. scatter...: or, he scattereth.}
\vs Psa 68:31 Princes shall come out of Egypt; Ethiopia shall soon stretch out her hands unto God.
\vs Psa 68:32 Sing unto God, ye kingdoms of the earth; O sing praises unto the Lord; Selah:
\vs Psa 68:33 To him that rideth upon the heavens of heavens, \bibemph{which were} of old; lo, he doth send out his voice, \bibemph{and that} a mighty voice.
\fn{send...: Heb. give.}
\vs Psa 68:34 Ascribe ye strength unto God: his excellency \bibemph{is} over Israel, and his strength \bibemph{is} in the clouds.
\fn{clouds: or, heavens.}
\vs Psa 68:35 O God, \bibemph{thou art} terrible out of thy holy places: the God of Israel \bibemph{is} he that giveth strength and power unto \bibemph{his} people. Blessed \bibemph{be} God.
\vs Psa 69:1 To the chief Musician upon Shoshannim, \bibemph{A Psalm} of David. Save me, O God; for the waters are come in unto \bibemph{my} soul.
\vs Psa 69:2 I sink in deep mire, where \bibemph{there is} no standing: I am come into deep waters, where the floods overflow me.
\fn{deep mire: Heb. the mire of depth. deep waters: Heb. depth of waters.}
\vs Psa 69:3 I am weary of my crying: my throat is dried: mine eyes fail while I wait for my God.
\vs Psa 69:4 They that hate me without a cause are more than the hairs of mine head: they that would destroy me, \bibemph{being} mine enemies wrongfully, are mighty: then I restored \bibemph{that} which I took not away.
\vs Psa 69:5 O God, thou knowest my foolishness; and my sins are not hid from thee.
\fn{sins: Heb. guiltiness.}
\vs Psa 69:6 Let not them that wait on thee, O Lord GOD of hosts, be ashamed for my sake: let not those that seek thee be confounded for my sake, O God of Israel.
\vs Psa 69:7 Because for thy sake I have borne reproach; shame hath covered my face.
\vs Psa 69:8 I am become a stranger unto my brethren, and an alien unto my mother's children.
\vs Psa 69:9 For the zeal of thine house hath eaten me up; and the reproaches of them that reproached thee are fallen upon me.
\vs Psa 69:10 When I wept, \bibemph{and chastened} my soul with fasting, that was to my reproach.
\vs Psa 69:11 I made sackcloth also my garment; and I became a proverb to them.
\vs Psa 69:12 They that sit in the gate speak against me; and I \bibemph{was} the song of the drunkards.
\fn{drunkards: Heb. drinkers of strong drink.}
\vs Psa 69:13 But as for me, my prayer \bibemph{is} unto thee, O \textsc{Lord}, \bibemph{in} an acceptable time: O God, in the multitude of thy mercy hear me, in the truth of thy salvation.
\vs Psa 69:14 Deliver me out of the mire, and let me not sink: let me be delivered from them that hate me, and out of the deep waters.
\vs Psa 69:15 Let not the waterflood overflow me, neither let the deep swallow me up, and let not the pit shut her mouth upon me.
\vs Psa 69:16 Hear me, O \textsc{Lord}; for thy lovingkindness \bibemph{is} good: turn unto me according to the multitude of thy tender mercies.
\vs Psa 69:17 And hide not thy face from thy servant; for I am in trouble: hear me speedily.
\fn{hear...: Heb. make haste to hear me.}
\vs Psa 69:18 Draw nigh unto my soul, \bibemph{and} redeem it: deliver me because of mine enemies.
\vs Psa 69:19 Thou hast known my reproach, and my shame, and my dishonour: mine adversaries \bibemph{are} all before thee.
\vs Psa 69:20 Reproach hath broken my heart; and I am full of heaviness: and I looked \bibemph{for some} to take pity, but \bibemph{there was} none; and for comforters, but I found none.
\fn{to take...: Heb. to lament with me.}
\vs Psa 69:21 They gave me also gall for my meat; and in my thirst they gave me vinegar to drink.
\vs Psa 69:22 Let their table become a snare before them: and \bibemph{that which should have been} for \bibemph{their} welfare, \bibemph{let it become} a trap.
\vs Psa 69:23 Let their eyes be darkened, that they see not; and make their loins continually to shake.
\vs Psa 69:24 Pour out thine indignation upon them, and let thy wrathful anger take hold of them.
\vs Psa 69:25 Let their habitation be desolate; \bibemph{and} let none dwell in their tents.
\fn{their habitation: Heb. their palace. let none...: Heb. let there not be a dweller.}
\vs Psa 69:26 For they persecute \bibemph{him} whom thou hast smitten; and they talk to the grief of those whom thou hast wounded.
\fn{those...: Heb. thy wounded.}
\vs Psa 69:27 Add iniquity unto their iniquity: and let them not come into thy righteousness.
\fn{iniquity unto...: or, punishment of iniquity, etc.}
\vs Psa 69:28 Let them be blotted out of the book of the living, and not be written with the righteous.
\vs Psa 69:29 But I \bibemph{am} poor and sorrowful: let thy salvation, O God, set me up on high.
\vs Psa 69:30 I will praise the name of God with a song, and will magnify him with thanksgiving.
\vs Psa 69:31 \bibemph{This} also shall please the \textsc{Lord} better than an ox \bibemph{or} bullock that hath horns and hoofs.
\vs Psa 69:32 The humble shall see \bibemph{this, and} be glad: and your heart shall live that seek God.
\fn{humble: or, meek.}
\vs Psa 69:33 For the \textsc{Lord} heareth the poor, and despiseth not his prisoners.
\vs Psa 69:34 Let the heaven and earth praise him, the seas, and every thing that moveth therein.
\fn{moveth: Heb. creepeth.}
\vs Psa 69:35 For God will save Zion, and will build the cities of Judah: that they may dwell there, and have it in possession.
\vs Psa 69:36 The seed also of his servants shall inherit it: and they that love his name shall dwell therein.
\vs Psa 70:1 To the chief Musician, \bibemph{A Psalm} of David, to bring to remembrance. \bibemph{Make haste}, O God, to deliver me; make haste to help me, O \textsc{Lord}.
\fn{to help...: Heb. to my help.}
\vs Psa 70:2 Let them be ashamed and confounded that seek after my soul: let them be turned backward, and put to confusion, that desire my hurt.
\vs Psa 70:3 Let them be turned back for a reward of their shame that say, Aha, aha.
\vs Psa 70:4 Let all those that seek thee rejoice and be glad in thee: and let such as love thy salvation say continually, Let God be magnified.
\vs Psa 70:5 But I \bibemph{am} poor and needy: make haste unto me, O God: thou \bibemph{art} my help and my deliverer; O \textsc{Lord}, make no tarrying.
\vs Psa 71:1 In thee, O \textsc{Lord}, do I put my trust: let me never be put to confusion.
\vs Psa 71:2 Deliver me in thy righteousness, and cause me to escape: incline thine ear unto me, and save me.
\vs Psa 71:3 Be thou my strong habitation, whereunto I may continually resort: thou hast given commandment to save me; for thou \bibemph{art} my rock and my fortress.
\fn{Be...: Heb. Be thou to me for a rock of habitation.}
\vs Psa 71:4 Deliver me, O my God, out of the hand of the wicked, out of the hand of the unrighteous and cruel man.
\vs Psa 71:5 For thou \bibemph{art} my hope, O Lord GOD: \bibemph{thou art} my trust from my youth.
\vs Psa 71:6 By thee have I been holden up from the womb: thou art he that took me out of my mother's bowels: my praise \bibemph{shall be} continually of thee.
\vs Psa 71:7 I am as a wonder unto many; but thou \bibemph{art} my strong refuge.
\vs Psa 71:8 Let my mouth be filled \bibemph{with} thy praise \bibemph{and with} thy honour all the day.
\vs Psa 71:9 Cast me not off in the time of old age; forsake me not when my strength faileth.
\vs Psa 71:10 For mine enemies speak against me; and they that lay wait for my soul take counsel together,
\fn{lay...: Heb. watch, or, observe.}
\vs Psa 71:11 Saying, God hath forsaken him: persecute and take him; for \bibemph{there is} none to deliver \bibemph{him}.
\vs Psa 71:12 O God, be not far from me: O my God, make haste for my help.
\vs Psa 71:13 Let them be confounded \bibemph{and} consumed that are adversaries to my soul; let them be covered \bibemph{with} reproach and dishonour that seek my hurt.
\vs Psa 71:14 But I will hope continually, and will yet praise thee more and more.
\vs Psa 71:15 My mouth shall shew forth thy righteousness \bibemph{and} thy salvation all the day; for I know not the numbers \bibemph{thereof}.
\vs Psa 71:16 I will go in the strength of the Lord GOD: I will make mention of thy righteousness, \bibemph{even} of thine only.
\vs Psa 71:17 O God, thou hast taught me from my youth: and hitherto have I declared thy wondrous works.
\vs Psa 71:18 Now also when I am old and grayheaded, O God, forsake me not; until I have shewed thy strength unto \bibemph{this} generation, \bibemph{and} thy power to every one \bibemph{that} is to come.
\fn{when...: Heb. unto old age and gray hairs. thy strength: Heb. thine arm.}
\vs Psa 71:19 Thy righteousness also, O God, \bibemph{is} very high, who hast done great things: O God, who \bibemph{is} like unto thee!
\vs Psa 71:20 \bibemph{Thou}, which hast shewed me great and sore troubles, shalt quicken me again, and shalt bring me up again from the depths of the earth.
\vs Psa 71:21 Thou shalt increase my greatness, and comfort me on every side.
\vs Psa 71:22 I will also praise thee with the psaltery, \bibemph{even} thy truth, O my God: unto thee will I sing with the harp, O thou Holy One of Israel.
\fn{the psaltery: Heb. the instrument of psaltery.}
\vs Psa 71:23 My lips shall greatly rejoice when I sing unto thee; and my soul, which thou hast redeemed.
\vs Psa 71:24 My tongue also shall talk of thy righteousness all the day long: for they are confounded, for they are brought unto shame, that seek my hurt.
\vs Psa 72:1 \bibemph{A Psalm} for Solomon. Give the king thy judgments, O God, and thy righteousness unto the king's son.
\fn{for: or, of.}
\vs Psa 72:2 He shall judge thy people with righteousness, and thy poor with judgment.
\vs Psa 72:3 The mountains shall bring peace to the people, and the little hills, by righteousness.
\vs Psa 72:4 He shall judge the poor of the people, he shall save the children of the needy, and shall break in pieces the oppressor.
\vs Psa 72:5 They shall fear thee as long as the sun and moon endure, throughout all generations.
\vs Psa 72:6 He shall come down like rain upon the mown grass: as showers \bibemph{that} water the earth.
\vs Psa 72:7 In his days shall the righteous flourish; and abundance of peace so long as the moon endureth.
\fn{so long...: Heb. till there be no moon.}
\vs Psa 72:8 He shall have dominion also from sea to sea, and from the river unto the ends of the earth.
\vs Psa 72:9 They that dwell in the wilderness shall bow before him; and his enemies shall lick the dust.
\vs Psa 72:10 The kings of Tarshish and of the isles shall bring presents: the kings of Sheba and Seba shall offer gifts.
\vs Psa 72:11 Yea, all kings shall fall down before him: all nations shall serve him.
\vs Psa 72:12 For he shall deliver the needy when he crieth; the poor also, and \bibemph{him} that hath no helper.
\vs Psa 72:13 He shall spare the poor and needy, and shall save the souls of the needy.
\vs Psa 72:14 He shall redeem their soul from deceit and violence: and precious shall their blood be in his sight.
\vs Psa 72:15 And he shall live, and to him shall be given of the gold of Sheba: prayer also shall be made for him continually; \bibemph{and} daily shall he be praised.
\fn{shall be given: Heb. one shall give.}
\vs Psa 72:16 There shall be an handful of corn in the earth upon the top of the mountains; the fruit thereof shall shake like Lebanon: and \bibemph{they} of the city shall flourish like grass of the earth.
\vs Psa 72:17 His name shall endure for ever: his name shall be continued as long as the sun: and \bibemph{men} shall be blessed in him: all nations shall call him blessed.
\fn{endure: Heb. be. his name shall be...: Heb. shall be as a son to continue his father's name for ever.}
\vs Psa 72:18 Blessed \bibemph{be} the \textsc{Lord} God, the God of Israel, who only doeth wondrous things.
\vs Psa 72:19 And blessed \bibemph{be} his glorious name for ever: and let the whole earth be filled \bibemph{with} his glory; Amen, and Amen.
\vs Psa 72:20 The prayers of David the son of Jesse are ended.
\vs Psa 73:1 A Psalm of Asaph. Truly God \bibemph{is} good to Israel, \bibemph{even} to such as are of a clean heart.
\fn{Truly: or, Yet. of...: Heb. clean of heart. of: or, for.}
\vs Psa 73:2 But as for me, my feet were almost gone; my steps had well nigh slipped.
\vs Psa 73:3 For I was envious at the foolish, \bibemph{when} I saw the prosperity of the wicked.
\vs Psa 73:4 For \bibemph{there are} no bands in their death: but their strength \bibemph{is} firm.
\fn{firm: Heb. fat.}
\vs Psa 73:5 They \bibemph{are} not in trouble \bibemph{as other} men; neither are they plagued like \bibemph{other} men.
\fn{in...: Heb. in the trouble of other men. like: Heb. with.}
\vs Psa 73:6 Therefore pride compasseth them about as a chain; violence covereth them \bibemph{as} a garment.
\vs Psa 73:7 Their eyes stand out with fatness: they have more than heart could wish.
\fn{have...: Heb. pass the thoughts of the heart.}
\vs Psa 73:8 They are corrupt, and speak wickedly \bibemph{concerning} oppression: they speak loftily.
\vs Psa 73:9 They set their mouth against the heavens, and their tongue walketh through the earth.
\vs Psa 73:10 Therefore his people return hither: and waters of a full \bibemph{cup} are wrung out to them.
\vs Psa 73:11 And they say, How doth God know? and is there knowledge in the most High?
\vs Psa 73:12 Behold, these \bibemph{are} the ungodly, who prosper in the world; they increase \bibemph{in} riches.
\vs Psa 73:13 Verily I have cleansed my heart \bibemph{in} vain, and washed my hands in innocency.
\vs Psa 73:14 For all the day long have I been plagued, and chastened every morning.
\fn{chastened: Heb. my chastisement was.}
\vs Psa 73:15 If I say, I will speak thus; behold, I should offend \bibemph{against} the generation of thy children.
\vs Psa 73:16 When I thought to know this, it \bibemph{was} too painful for me;
\fn{too...: Heb. labour in mine eyes.}
\vs Psa 73:17 Until I went into the sanctuary of God; \bibemph{then} understood I their end.
\vs Psa 73:18 Surely thou didst set them in slippery places: thou castedst them down into destruction.
\vs Psa 73:19 How are they \bibemph{brought} into desolation, as in a moment! they are utterly consumed with terrors.
\vs Psa 73:20 As a dream when \bibemph{one} awaketh; \bibemph{so}, O Lord, when thou awakest, thou shalt despise their image.
\vs Psa 73:21 Thus my heart was grieved, and I was pricked in my reins.
\vs Psa 73:22 So foolish \bibemph{was} I, and ignorant: I was \bibemph{as} a beast before thee.
\fn{ignorant: Heb. I knew not. before Heb. with.}
\vs Psa 73:23 Nevertheless I \bibemph{am} continually with thee: thou hast holden \bibemph{me} by my right hand.
\vs Psa 73:24 Thou shalt guide me with thy counsel, and afterward receive me \bibemph{to} glory.
\vs Psa 73:25 Whom have I in heaven \bibemph{but thee}? and \bibemph{there is} none upon earth \bibemph{that} I desire beside thee.
\vs Psa 73:26 My flesh and my heart faileth: \bibemph{but} God \bibemph{is} the strength of my heart, and my portion for ever.
\fn{strength: Heb. rock.}
\vs Psa 73:27 For, lo, they that are far from thee shall perish: thou hast destroyed all them that go a whoring from thee.
\vs Psa 73:28 But \bibemph{it is} good for me to draw near to God: I have put my trust in the Lord GOD, that I may declare all thy works.
\vs Psa 74:1 Maschil of Asaph. O God, why hast thou cast \bibemph{us} off for ever? \bibemph{why} doth thine anger smoke against the sheep of thy pasture?
\fn{Maschil...: or, A Psalm for Asaph to give instruction.}
\vs Psa 74:2 Remember thy congregation, \bibemph{which} thou hast purchased of old; the rod of thine inheritance, \bibemph{which} thou hast redeemed; this mount Zion, wherein thou hast dwelt.
\fn{rod: or, tribe.}
\vs Psa 74:3 Lift up thy feet unto the perpetual desolations; \bibemph{even} all \bibemph{that} the enemy hath done wickedly in the sanctuary.
\vs Psa 74:4 Thine enemies roar in the midst of thy congregations; they set up their ensigns \bibemph{for} signs.
\vs Psa 74:5 \bibemph{A man} was famous according as he had lifted up axes upon the thick trees.
\vs Psa 74:6 But now they break down the carved work thereof at once with axes and hammers.
\vs Psa 74:7 They have cast fire into thy sanctuary, they have defiled \bibemph{by casting down} the dwelling place of thy name to the ground.
\fn{They have cast...: Heb. They have sent thy sanctuary unto the fire.}
\vs Psa 74:8 They said in their hearts, Let us destroy them together: they have burned up all the synagogues of God in the land.
\fn{destroy: Heb. break.}
\vs Psa 74:9 We see not our signs: \bibemph{there is} no more any prophet: neither \bibemph{is there} among us any that knoweth how long.
\vs Psa 74:10 O God, how long shall the adversary reproach? shall the enemy blaspheme thy name for ever?
\vs Psa 74:11 Why withdrawest thou thy hand, even thy right hand? pluck \bibemph{it} out of thy bosom.
\vs Psa 74:12 For God \bibemph{is} my King of old, working salvation in the midst of the earth.
\vs Psa 74:13 Thou didst divide the sea by thy strength: thou brakest the heads of the dragons in the waters.
\fn{divide: Heb. break. dragons: or, whales.}
\vs Psa 74:14 Thou brakest the heads of leviathan in pieces, \bibemph{and} gavest him \bibemph{to be} meat to the people inhabiting the wilderness.
\vs Psa 74:15 Thou didst cleave the fountain and the flood: thou driedst up mighty rivers.
\fn{mighty...: Heb. rivers of strength.}
\vs Psa 74:16 The day \bibemph{is} thine, the night also \bibemph{is} thine: thou hast prepared the light and the sun.
\vs Psa 74:17 Thou hast set all the borders of the earth: thou hast made summer and winter.
\fn{made: Heb. made them.}
\vs Psa 74:18 Remember this, \bibemph{that} the enemy hath reproached, O \textsc{Lord}, and \bibemph{that} the foolish people have blasphemed thy name.
\vs Psa 74:19 O deliver not the soul of thy turtledove unto the multitude \bibemph{of the wicked}: forget not the congregation of thy poor for ever.
\vs Psa 74:20 Have respect unto the covenant: for the dark places of the earth are full of the habitations of cruelty.
\vs Psa 74:21 O let not the oppressed return ashamed: let the poor and needy praise thy name.
\vs Psa 74:22 Arise, O God, plead thine own cause: remember how the foolish man reproacheth thee daily.
\vs Psa 74:23 Forget not the voice of thine enemies: the tumult of those that rise up against thee increaseth continually.
\fn{increaseth: Heb. ascendeth.}
\vs Psa 75:1 To the chief Musician, Altaschith, A Psalm \bibemph{or} Song of Asaph. Unto thee, O God, do we give thanks, \bibemph{unto thee} do we give thanks: for \bibemph{that} thy name is near thy wondrous works declare.
\fn{Altaschith: or, Destroy not. of: or, for.}
\vs Psa 75:2 When I shall receive the congregation I will judge uprightly.
\fn{receive...: or, take a set time.}
\vs Psa 75:3 The earth and all the inhabitants thereof are dissolved: I bear up the pillars of it. Selah.
\vs Psa 75:4 I said unto the fools, Deal not foolishly: and to the wicked, Lift not up the horn:
\vs Psa 75:5 Lift not up your horn on high: speak \bibemph{not with} a stiff neck.
\vs Psa 75:6 For promotion \bibemph{cometh} neither from the east, nor from the west, nor from the south.
\fn{south: Heb. desert.}
\vs Psa 75:7 But God \bibemph{is} the judge: he putteth down one, and setteth up another.
\vs Psa 75:8 For in the hand of the \textsc{Lord} \bibemph{there is} a cup, and the wine is red; it is full of mixture; and he poureth out of the same: but the dregs thereof, all the wicked of the earth shall wring \bibemph{them} out, \bibemph{and} drink \bibemph{them}.
\vs Psa 75:9 But I will declare for ever; I will sing praises to the God of Jacob.
\vs Psa 75:10 All the horns of the wicked also will I cut off; \bibemph{but} the horns of the righteous shall be exalted.
\vs Psa 76:1 To the chief Musician on Neginoth, A Psalm \bibemph{or} Song of Asaph. In Judah \bibemph{is} God known: his name \bibemph{is} great in Israel.
\fn{of: or, for.}
\vs Psa 76:2 In Salem also is his tabernacle, and his dwelling place in Zion.
\vs Psa 76:3 There brake he the arrows of the bow, the shield, and the sword, and the battle. Selah.
\vs Psa 76:4 Thou \bibemph{art} more glorious \bibemph{and} excellent than the mountains of prey.
\vs Psa 76:5 The stouthearted are spoiled, they have slept their sleep: and none of the men of might have found their hands.
\vs Psa 76:6 At thy rebuke, O God of Jacob, both the chariot and horse are cast into a dead sleep.
\vs Psa 76:7 Thou, \bibemph{even} thou, \bibemph{art} to be feared: and who may stand in thy sight when once thou art angry?
\vs Psa 76:8 Thou didst cause judgment to be heard from heaven; the earth feared, and was still,
\vs Psa 76:9 When God arose to judgment, to save all the meek of the earth. Selah.
\vs Psa 76:10 Surely the wrath of man shall praise thee: the remainder of wrath shalt thou restrain.
\vs Psa 76:11 Vow, and pay unto the \textsc{Lord} your God: let all that be round about him bring presents unto him that ought to be feared.
\fn{unto him...: Heb. to fear.}
\vs Psa 76:12 He shall cut off the spirit of princes: \bibemph{he is} terrible to the kings of the earth.
\vs Psa 77:1 To the chief Musician, to Jeduthun, A Psalm of Asaph. I cried unto God with my voice, \bibemph{even} unto God with my voice; and he gave ear unto me.
\fn{of: or, for.}
\vs Psa 77:2 In the day of my trouble I sought the Lord: my sore ran in the night, and ceased not: my soul refused to be comforted.
\fn{sore: Heb. hand.}
\vs Psa 77:3 I remembered God, and was troubled: I complained, and my spirit was overwhelmed. Selah.
\vs Psa 77:4 Thou holdest mine eyes waking: I am so troubled that I cannot speak.
\vs Psa 77:5 I have considered the days of old, the years of ancient times.
\vs Psa 77:6 I call to remembrance my song in the night: I commune with mine own heart: and my spirit made diligent search.
\vs Psa 77:7 Will the Lord cast off for ever? and will he be favourable no more?
\vs Psa 77:8 Is his mercy clean gone for ever? doth \bibemph{his} promise fail for evermore?
\fn{for evermore: Heb. to generation and generation?.}
\vs Psa 77:9 Hath God forgotten to be gracious? hath he in anger shut up his tender mercies? Selah.
\vs Psa 77:10 And I said, This \bibemph{is} my infirmity: \bibemph{but I will remember} the years of the right hand of the most High.
\vs Psa 77:11 I will remember the works of the \textsc{Lord}: surely I will remember thy wonders of old.
\vs Psa 77:12 I will meditate also of all thy work, and talk of thy doings.
\vs Psa 77:13 Thy way, O God, \bibemph{is} in the sanctuary: who \bibemph{is so} great a God as \bibemph{our} God?
\vs Psa 77:14 Thou \bibemph{art} the God that doest wonders: thou hast declared thy strength among the people.
\vs Psa 77:15 Thou hast with \bibemph{thine} arm redeemed thy people, the sons of Jacob and Joseph. Selah.
\vs Psa 77:16 The waters saw thee, O God, the waters saw thee; they were afraid: the depths also were troubled.
\vs Psa 77:17 The clouds poured out water: the skies sent out a sound: thine arrows also went abroad.
\fn{poured...: Heb. were poured forth with water.}
\vs Psa 77:18 The voice of thy thunder \bibemph{was} in the heaven: the lightnings lightened the world: the earth trembled and shook.
\vs Psa 77:19 Thy way \bibemph{is} in the sea, and thy path in the great waters, and thy footsteps are not known.
\vs Psa 77:20 Thou leddest thy people like a flock by the hand of Moses and Aaron.
\vs Psa 78:1 Maschil of Asaph. Give ear, O my people, \bibemph{to} my law: incline your ears to the words of my mouth.
\fn{Maschil...: or, A Psalm for Asaph to give instruction.}
\vs Psa 78:2 I will open my mouth in a parable: I will utter dark sayings of old:
\vs Psa 78:3 Which we have heard and known, and our fathers have told us.
\vs Psa 78:4 We will not hide \bibemph{them} from their children, shewing to the generation to come the praises of the \textsc{Lord}, and his strength, and his wonderful works that he hath done.
\vs Psa 78:5 For he established a testimony in Jacob, and appointed a law in Israel, which he commanded our fathers, that they should make them known to their children:
\vs Psa 78:6 That the generation to come might know \bibemph{them, even} the children \bibemph{which} should be born; \bibemph{who} should arise and declare \bibemph{them} to their children:
\vs Psa 78:7 That they might set their hope in God, and not forget the works of God, but keep his commandments:
\vs Psa 78:8 And might not be as their fathers, a stubborn and rebellious generation; a generation \bibemph{that} set not their heart aright, and whose spirit was not stedfast with God.
\fn{that...: Heb. that prepared not their heart.}
\vs Psa 78:9 The children of Ephraim, \bibemph{being} armed, \bibemph{and} carrying bows, turned back in the day of battle.
\fn{carrying: Heb. throwing forth.}
\vs Psa 78:10 They kept not the covenant of God, and refused to walk in his law;
\vs Psa 78:11 And forgat his works, and his wonders that he had shewed them.
\vs Psa 78:12 Marvellous things did he in the sight of their fathers, in the land of Egypt, \bibemph{in} the field of Zoan.
\vs Psa 78:13 He divided the sea, and caused them to pass through; and he made the waters to stand as an heap.
\vs Psa 78:14 In the daytime also he led them with a cloud, and all the night with a light of fire.
\vs Psa 78:15 He clave the rocks in the wilderness, and gave \bibemph{them} drink as \bibemph{out of} the great depths.
\vs Psa 78:16 He brought streams also out of the rock, and caused waters to run down like rivers.
\vs Psa 78:17 And they sinned yet more against him by provoking the most High in the wilderness.
\vs Psa 78:18 And they tempted God in their heart by asking meat for their lust.
\vs Psa 78:19 Yea, they spake against God; they said, Can God furnish a table in the wilderness?
\fn{furnish: Heb. order.}
\vs Psa 78:20 Behold, he smote the rock, that the waters gushed out, and the streams overflowed; can he give bread also? can he provide flesh for his people?
\vs Psa 78:21 Therefore the \textsc{Lord} heard \bibemph{this}, and was wroth: so a fire was kindled against Jacob, and anger also came up against Israel;
\vs Psa 78:22 Because they believed not in God, and trusted not in his salvation:
\vs Psa 78:23 Though he had commanded the clouds from above, and opened the doors of heaven,
\vs Psa 78:24 And had rained down manna upon them to eat, and had given them of the corn of heaven.
\vs Psa 78:25 Man did eat angels' food: he sent them meat to the full.
\fn{Man...: or, Every one did eat the bread of the mighty.}
\vs Psa 78:26 He caused an east wind to blow in the heaven: and by his power he brought in the south wind.
\fn{to blow: Heb. to go.}
\vs Psa 78:27 He rained flesh also upon them as dust, and feathered fowls like as the sand of the sea:
\fn{feathered...: Heb. fowl of wing.}
\vs Psa 78:28 And he let \bibemph{it} fall in the midst of their camp, round about their habitations.
\vs Psa 78:29 So they did eat, and were well filled: for he gave them their own desire;
\vs Psa 78:30 They were not estranged from their lust. But while their meat \bibemph{was} yet in their mouths,
\vs Psa 78:31 The wrath of God came upon them, and slew the fattest of them, and smote down the chosen \bibemph{men} of Israel.
\fn{smote...: Heb. made to bow. Chosen...: or, young men.}
\vs Psa 78:32 For all this they sinned still, and believed not for his wondrous works.
\vs Psa 78:33 Therefore their days did he consume in vanity, and their years in trouble.
\vs Psa 78:34 When he slew them, then they sought him: and they returned and enquired early after God.
\vs Psa 78:35 And they remembered that God \bibemph{was} their rock, and the high God their redeemer.
\vs Psa 78:36 Nevertheless they did flatter him with their mouth, and they lied unto him with their tongues.
\vs Psa 78:37 For their heart was not right with him, neither were they stedfast in his covenant.
\vs Psa 78:38 But he, \bibemph{being} full of compassion, forgave \bibemph{their} iniquity, and destroyed \bibemph{them} not: yea, many a time turned he his anger away, and did not stir up all his wrath.
\vs Psa 78:39 For he remembered that they \bibemph{were but} flesh; a wind that passeth away, and cometh not again.
\vs Psa 78:40 How oft did they provoke him in the wilderness, \bibemph{and} grieve him in the desert!
\fn{provoke: or, rebel against.}
\vs Psa 78:41 Yea, they turned back and tempted God, and limited the Holy One of Israel.
\vs Psa 78:42 They remembered not his hand, \bibemph{nor} the day when he delivered them from the enemy.
\fn{from...: or, from affliction.}
\vs Psa 78:43 How he had wrought his signs in Egypt, and his wonders in the field of Zoan:
\fn{wrought: Heb. set.}
\vs Psa 78:44 And had turned their rivers into blood; and their floods, that they could not drink.
\vs Psa 78:45 He sent divers sorts of flies among them, which devoured them; and frogs, which destroyed them.
\vs Psa 78:46 He gave also their increase unto the caterpiller, and their labour unto the locust.
\vs Psa 78:47 He destroyed their vines with hail, and their sycomore trees with frost.
\fn{destroyed: Heb. killed. frost: or, great hailstones.}
\vs Psa 78:48 He gave up their cattle also to the hail, and their flocks to hot thunderbolts.
\fn{gave...: Heb. shut up. hot...: or, lightnings.}
\vs Psa 78:49 He cast upon them the fierceness of his anger, wrath, and indignation, and trouble, by sending evil angels \bibemph{among them}.
\vs Psa 78:50 He made a way to his anger; he spared not their soul from death, but gave their life over to the pestilence;
\fn{He made...: Heb. He weighed a path. life...: or, beasts to the murrain.}
\vs Psa 78:51 And smote all the firstborn in Egypt; the chief of \bibemph{their} strength in the tabernacles of Ham:
\vs Psa 78:52 But made his own people to go forth like sheep, and guided them in the wilderness like a flock.
\vs Psa 78:53 And he led them on safely, so that they feared not: but the sea overwhelmed their enemies.
\fn{overwhelmed: Heb. covered.}
\vs Psa 78:54 And he brought them to the border of his sanctuary, \bibemph{even to} this mountain, \bibemph{which} his right hand had purchased.
\vs Psa 78:55 He cast out the heathen also before them, and divided them an inheritance by line, and made the tribes of Israel to dwell in their tents.
\vs Psa 78:56 Yet they tempted and provoked the most high God, and kept not his testimonies:
\vs Psa 78:57 But turned back, and dealt unfaithfully like their fathers: they were turned aside like a deceitful bow.
\vs Psa 78:58 For they provoked him to anger with their high places, and moved him to jealousy with their graven images.
\vs Psa 78:59 When God heard \bibemph{this}, he was wroth, and greatly abhorred Israel:
\vs Psa 78:60 So that he forsook the tabernacle of Shiloh, the tent \bibemph{which} he placed among men;
\vs Psa 78:61 And delivered his strength into captivity, and his glory into the enemy's hand.
\vs Psa 78:62 He gave his people over also unto the sword; and was wroth with his inheritance.
\vs Psa 78:63 The fire consumed their young men; and their maidens were not given to marriage.
\fn{given...: Heb. praised.}
\vs Psa 78:64 Their priests fell by the sword; and their widows made no lamentation.
\vs Psa 78:65 Then the Lord awaked as one out of sleep, \bibemph{and} like a mighty man that shouteth by reason of wine.
\vs Psa 78:66 And he smote his enemies in the hinder parts: he put them to a perpetual reproach.
\vs Psa 78:67 Moreover he refused the tabernacle of Joseph, and chose not the tribe of Ephraim:
\vs Psa 78:68 But chose the tribe of Judah, the mount Zion which he loved.
\vs Psa 78:69 And he built his sanctuary like high \bibemph{palaces}, like the earth which he hath established for ever.
\fn{established: Heb. founded.}
\vs Psa 78:70 He chose David also his servant, and took him from the sheepfolds:
\vs Psa 78:71 From following the ewes great with young he brought him to feed Jacob his people, and Israel his inheritance.
\fn{following: Heb. after.}
\vs Psa 78:72 So he fed them according to the integrity of his heart; and guided them by the skilfulness of his hands.
\vs Psa 79:1 A Psalm of Asaph. O God, the heathen are come into thine inheritance; thy holy temple have they defiled; they have laid Jerusalem on heaps.
\fn{of: or, for.}
\vs Psa 79:2 The dead bodies of thy servants have they given \bibemph{to be} meat unto the fowls of the heaven, the flesh of thy saints unto the beasts of the earth.
\vs Psa 79:3 Their blood have they shed like water round about Jerusalem; and \bibemph{there was} none to bury \bibemph{them}.
\vs Psa 79:4 We are become a reproach to our neighbours, a scorn and derision to them that are round about us.
\vs Psa 79:5 How long, \textsc{Lord}? wilt thou be angry for ever? shall thy jealousy burn like fire?
\vs Psa 79:6 Pour out thy wrath upon the heathen that have not known thee, and upon the kingdoms that have not called upon thy name.
\vs Psa 79:7 For they have devoured Jacob, and laid waste his dwelling place.
\vs Psa 79:8 O remember not against us former iniquities: let thy tender mercies speedily prevent us: for we are brought very low.
\fn{former...: or, the iniquities of them that were before us.}
\vs Psa 79:9 Help us, O God of our salvation, for the glory of thy name: and deliver us, and purge away our sins, for thy name's sake.
\vs Psa 79:10 Wherefore should the heathen say, Where \bibemph{is} their God? let him be known among the heathen in our sight \bibemph{by} the revenging of the blood of thy servants \bibemph{which is} shed.
\fn{revenging: Heb. vengeance.}
\vs Psa 79:11 Let the sighing of the prisoner come before thee; according to the greatness of thy power preserve thou those that are appointed to die;
\fn{thy...: Heb. thine arm. preserve...: Heb. reserve the children of death.}
\vs Psa 79:12 And render unto our neighbours sevenfold into their bosom their reproach, wherewith they have reproached thee, O Lord.
\vs Psa 79:13 So we thy people and sheep of thy pasture will give thee thanks for ever: we will shew forth thy praise to all generations.
\fn{to all...: Heb. to generation and generation.}
\vs Psa 80:1 To the chief Musician upon Shoshannimeduth, A Psalm of Asaph. Give ear, O Shepherd of Israel, thou that leadest Joseph like a flock; thou that dwellest \bibemph{between} the cherubims, shine forth.
\fn{of: or, for.}
\vs Psa 80:2 Before Ephraim and Benjamin and Manasseh stir up thy strength, and come \bibemph{and} save us.
\fn{come...: Heb. come for salvation to us.}
\vs Psa 80:3 Turn us again, O God, and cause thy face to shine; and we shall be saved.
\vs Psa 80:4 O \textsc{Lord} God of hosts, how long wilt thou be angry against the prayer of thy people?
\fn{be...: Heb. smoke.}
\vs Psa 80:5 Thou feedest them with the bread of tears; and givest them tears to drink in great measure.
\vs Psa 80:6 Thou makest us a strife unto our neighbours: and our enemies laugh among themselves.
\vs Psa 80:7 Turn us again, O God of hosts, and cause thy face to shine; and we shall be saved.
\vs Psa 80:8 Thou hast brought a vine out of Egypt: thou hast cast out the heathen, and planted it.
\vs Psa 80:9 Thou preparedst \bibemph{room} before it, and didst cause it to take deep root, and it filled the land.
\vs Psa 80:10 The hills were covered with the shadow of it, and the boughs thereof \bibemph{were like} the goodly cedars.
\fn{goodly...: Heb. cedars of God.}
\vs Psa 80:11 She sent out her boughs unto the sea, and her branches unto the river.
\vs Psa 80:12 Why hast thou \bibemph{then} broken down her hedges, so that all they which pass by the way do pluck her?
\vs Psa 80:13 The boar out of the wood doth waste it, and the wild beast of the field doth devour it.
\vs Psa 80:14 Return, we beseech thee, O God of hosts: look down from heaven, and behold, and visit this vine;
\vs Psa 80:15 And the vineyard which thy right hand hath planted, and the branch \bibemph{that} thou madest strong for thyself.
\vs Psa 80:16 \bibemph{It is} burned with fire, \bibemph{it is} cut down: they perish at the rebuke of thy countenance.
\vs Psa 80:17 Let thy hand be upon the man of thy right hand, upon the son of man \bibemph{whom} thou madest strong for thyself.
\vs Psa 80:18 So will not we go back from thee: quicken us, and we will call upon thy name.
\vs Psa 80:19 Turn us again, O \textsc{Lord} God of hosts, cause thy face to shine; and we shall be saved.
\vs Psa 81:1 To the chief Musician upon Gittith, \bibemph{A Psalm} of Asaph. Sing aloud unto God our strength: make a joyful noise unto the God of Jacob.
\fn{of Asaph: or, for Asaph.}
\vs Psa 81:2 Take a psalm, and bring hither the timbrel, the pleasant harp with the psaltery.
\vs Psa 81:3 Blow up the trumpet in the new moon, in the time appointed, on our solemn feast day.
\vs Psa 81:4 For this \bibemph{was} a statute for Israel, \bibemph{and} a law of the God of Jacob.
\vs Psa 81:5 This he ordained in Joseph \bibemph{for} a testimony, when he went out through the land of Egypt: \bibemph{where} I heard a language \bibemph{that} I understood not.
\fn{through: or, against.}
\vs Psa 81:6 I removed his shoulder from the burden: his hands were delivered from the pots.
\fn{were...: Heb. passed away.}
\vs Psa 81:7 Thou calledst in trouble, and I delivered thee; I answered thee in the secret place of thunder: I proved thee at the waters of Meribah. Selah.
\fn{Meribah: or, Strife.}
\vs Psa 81:8 Hear, O my people, and I will testify unto thee: O Israel, if thou wilt hearken unto me;
\vs Psa 81:9 There shall no strange god be in thee; neither shalt thou worship any strange god.
\vs Psa 81:10 I \bibemph{am} the \textsc{Lord} thy God, which brought thee out of the land of Egypt: open thy mouth wide, and I will fill it.
\vs Psa 81:11 But my people would not hearken to my voice; and Israel would none of me.
\vs Psa 81:12 So I gave them up unto their own hearts' lust: \bibemph{and} they walked in their own counsels.
\fn{unto...: or, to the hardness of their hearts, or, imagination.}
\vs Psa 81:13 Oh that my people had hearkened unto me, \bibemph{and} Israel had walked in my ways!
\vs Psa 81:14 I should soon have subdued their enemies, and turned my hand against their adversaries.
\vs Psa 81:15 The haters of the \textsc{Lord} should have submitted themselves unto him: but their time should have endured for ever.
\fn{submitted...: or, yielded feigned obedience: Heb. lied.}
\vs Psa 81:16 He should have fed them also with the finest of the wheat: and with honey out of the rock should I have satisfied thee.
\fn{finest...: Heb. fat of wheat.}
\vs Psa 82:1 A Psalm of Asaph. God standeth in the congregation of the mighty; he judgeth among the gods.
\fn{of Asaph: or, for Asaph.}
\vs Psa 82:2 How long will ye judge unjustly, and accept the persons of the wicked? Selah.
\vs Psa 82:3 Defend the poor and fatherless: do justice to the afflicted and needy.
\fn{Defend: Heb. Judge.}
\vs Psa 82:4 Deliver the poor and needy: rid \bibemph{them} out of the hand of the wicked.
\vs Psa 82:5 They know not, neither will they understand; they walk on in darkness: all the foundations of the earth are out of course.
\fn{out...: Heb. moved.}
\vs Psa 82:6 I have said, Ye \bibemph{are} gods; and all of you \bibemph{are} children of the most High.
\vs Psa 82:7 But ye shall die like men, and fall like one of the princes.
\vs Psa 82:8 Arise, O God, judge the earth: for thou shalt inherit all nations.
\vs Psa 83:1 A Song \bibemph{or} Psalm of Asaph. Keep not thou silence, O God: hold not thy peace, and be not still, O God.
\fn{of Asaph: or, for Asaph.}
\vs Psa 83:2 For, lo, thine enemies make a tumult: and they that hate thee have lifted up the head.
\vs Psa 83:3 They have taken crafty counsel against thy people, and consulted against thy hidden ones.
\vs Psa 83:4 They have said, Come, and let us cut them off from \bibemph{being} a nation; that the name of Israel may be no more in remembrance.
\vs Psa 83:5 For they have consulted together with one consent: they are confederate against thee:
\fn{consent: Heb. heart.}
\vs Psa 83:6 The tabernacles of Edom, and the Ishmaelites; of Moab, and the Hagarenes;
\vs Psa 83:7 Gebal, and Ammon, and Amalek; the Philistines with the inhabitants of Tyre;
\vs Psa 83:8 Assur also is joined with them: they have holpen the children of Lot. Selah.
\fn{holpen: Heb. been an arm to.}
\vs Psa 83:9 Do unto them as \bibemph{unto} the Midianites; as \bibemph{to} Sisera, as \bibemph{to} Jabin, at the brook of Kison:
\vs Psa 83:10 \bibemph{Which} perished at Endor: they became \bibemph{as} dung for the earth.
\vs Psa 83:11 Make their nobles like Oreb, and like Zeeb: yea, all their princes as Zebah, and as Zalmunna:
\vs Psa 83:12 Who said, Let us take to ourselves the houses of God in possession.
\vs Psa 83:13 O my God, make them like a wheel; as the stubble before the wind.
\vs Psa 83:14 As the fire burneth a wood, and as the flame setteth the mountains on fire;
\vs Psa 83:15 So persecute them with thy tempest, and make them afraid with thy storm.
\vs Psa 83:16 Fill their faces with shame; that they may seek thy name, O \textsc{Lord}.
\vs Psa 83:17 Let them be confounded and troubled for ever; yea, let them be put to shame, and perish:
\vs Psa 83:18 That \bibemph{men} may know that thou, whose name alone \bibemph{is} JEHOVAH, \bibemph{art} the most high over all the earth.
\vs Psa 84:1 To the chief Musician upon Gittith, A Psalm for the sons of Korah. How amiable \bibemph{are} thy tabernacles, O \textsc{Lord} of hosts!
\fn{for the sons: or, of the sons.}
\vs Psa 84:2 My soul longeth, yea, even fainteth for the courts of the \textsc{Lord}: my heart and my flesh crieth out for the living God.
\vs Psa 84:3 Yea, the sparrow hath found an house, and the swallow a nest for herself, where she may lay her young, \bibemph{even} thine altars, O \textsc{Lord} of hosts, my King, and my God.
\vs Psa 84:4 Blessed \bibemph{are} they that dwell in thy house: they will be still praising thee. Selah.
\vs Psa 84:5 Blessed \bibemph{is} the man whose strength \bibemph{is} in thee; in whose heart \bibemph{are} the ways \bibemph{of them}.
\vs Psa 84:6 \bibemph{Who} passing through the valley of Baca make it a well; the rain also filleth the pools.
\fn{Baca...: or, mulberry trees make him a well, etc. filleth: Heb. covereth.}
\vs Psa 84:7 They go from strength to strength, \bibemph{every one of them} in Zion appeareth before God.
\fn{strength to...: or, company to company.}
\vs Psa 84:8 O \textsc{Lord} God of hosts, hear my prayer: give ear, O God of Jacob. Selah.
\vs Psa 84:9 Behold, O God our shield, and look upon the face of thine anointed.
\vs Psa 84:10 For a day in thy courts \bibemph{is} better than a thousand. I had rather be a doorkeeper in the house of my God, than to dwell in the tents of wickedness.
\fn{I had...: Heb. I would choose rather to sit at the threshold.}
\vs Psa 84:11 For the \textsc{Lord} God \bibemph{is} a sun and shield: the \textsc{Lord} will give grace and glory: no good \bibemph{thing} will he withhold from them that walk uprightly.
\vs Psa 84:12 O \textsc{Lord} of hosts, blessed \bibemph{is} the man that trusteth in thee.
\vs Psa 85:1 To the chief Musician, A Psalm for the sons of Korah. \textsc{Lord}, thou hast been favourable unto thy land: thou hast brought back the captivity of Jacob.
\fn{for the sons: or, of the sons. favourable: or, well pleased.}
\vs Psa 85:2 Thou hast forgiven the iniquity of thy people, thou hast covered all their sin. Selah.
\vs Psa 85:3 Thou hast taken away all thy wrath: thou hast turned \bibemph{thyself} from the fierceness of thine anger.
\fn{thyself...: or, thine anger from waxing hot.}
\vs Psa 85:4 Turn us, O God of our salvation, and cause thine anger toward us to cease.
\vs Psa 85:5 Wilt thou be angry with us for ever? wilt thou draw out thine anger to all generations?
\vs Psa 85:6 Wilt thou not revive us again: that thy people may rejoice in thee?
\vs Psa 85:7 Shew us thy mercy, O \textsc{Lord}, and grant us thy salvation.
\vs Psa 85:8 I will hear what God the \textsc{Lord} will speak: for he will speak peace unto his people, and to his saints: but let them not turn again to folly.
\vs Psa 85:9 Surely his salvation \bibemph{is} nigh them that fear him; that glory may dwell in our land.
\vs Psa 85:10 Mercy and truth are met together; righteousness and peace have kissed \bibemph{each other}.
\vs Psa 85:11 Truth shall spring out of the earth; and righteousness shall look down from heaven.
\vs Psa 85:12 Yea, the \textsc{Lord} shall give \bibemph{that which is} good; and our land shall yield her increase.
\vs Psa 85:13 Righteousness shall go before him; and shall set \bibemph{us} in the way of his steps.
\vs Psa 86:1 A Prayer of David. Bow down thine ear, O \textsc{Lord}, hear me: for I \bibemph{am} poor and needy.
\fn{A Prayer...: or, A Prayer, being a psalm of David.}
\vs Psa 86:2 Preserve my soul; for I \bibemph{am} holy: O thou my God, save thy servant that trusteth in thee.
\fn{holy: or, one whom thou favourest.}
\vs Psa 86:3 Be merciful unto me, O Lord: for I cry unto thee daily.
\fn{daily: or, all the day.}
\vs Psa 86:4 Rejoice the soul of thy servant: for unto thee, O Lord, do I lift up my soul.
\vs Psa 86:5 For thou, Lord, \bibemph{art} good, and ready to forgive; and plenteous in mercy unto all them that call upon thee.
\vs Psa 86:6 Give ear, O \textsc{Lord}, unto my prayer; and attend to the voice of my supplications.
\vs Psa 86:7 In the day of my trouble I will call upon thee: for thou wilt answer me.
\vs Psa 86:8 Among the gods \bibemph{there is} none like unto thee, O Lord; neither \bibemph{are there any works} like unto thy works.
\vs Psa 86:9 All nations whom thou hast made shall come and worship before thee, O Lord; and shall glorify thy name.
\vs Psa 86:10 For thou \bibemph{art} great, and doest wondrous things: thou \bibemph{art} God alone.
\vs Psa 86:11 Teach me thy way, O \textsc{Lord}; I will walk in thy truth: unite my heart to fear thy name.
\vs Psa 86:12 I will praise thee, O Lord my God, with all my heart: and I will glorify thy name for evermore.
\vs Psa 86:13 For great \bibemph{is} thy mercy toward me: and thou hast delivered my soul from the lowest hell.
\fn{hell: or, grave.}
\vs Psa 86:14 O God, the proud are risen against me, and the assemblies of violent \bibemph{men} have sought after my soul; and have not set thee before them.
\fn{violent: Heb. terrible.}
\vs Psa 86:15 But thou, O Lord, \bibemph{art} a God full of compassion, and gracious, longsuffering, and plenteous in mercy and truth.
\vs Psa 86:16 O turn unto me, and have mercy upon me; give thy strength unto thy servant, and save the son of thine handmaid.
\vs Psa 86:17 Shew me a token for good; that they which hate me may see \bibemph{it}, and be ashamed: because thou, \textsc{Lord}, hast holpen me, and comforted me.
\vs Psa 87:1 A Psalm \bibemph{or} Song for the sons of Korah. His foundation \bibemph{is} in the holy mountains.
\fn{for the sons: or, of the sons.}
\vs Psa 87:2 The \textsc{Lord} loveth the gates of Zion more than all the dwellings of Jacob.
\vs Psa 87:3 Glorious things are spoken of thee, O city of God. Selah.
\vs Psa 87:4 I will make mention of Rahab and Babylon to them that know me: behold Philistia, and Tyre, with Ethiopia; this \bibemph{man} was born there.
\vs Psa 87:5 And of Zion it shall be said, This and that man was born in her: and the highest himself shall establish her.
\vs Psa 87:6 The \textsc{Lord} shall count, when he writeth up the people, \bibemph{that} this \bibemph{man} was born there. Selah.
\vs Psa 87:7 As well the singers as the players on instruments \bibemph{shall be there}: all my springs \bibemph{are} in thee.
\vs Psa 88:1 A Song \bibemph{or} Psalm for the sons of Korah, to the chief Musician upon Mahalath Leannoth, Maschil of Heman the Ezrahite. O \textsc{Lord} God of my salvation, I have cried day \bibemph{and} night before thee:
\fn{for the sons: or, of the sons. Maschil...: or, A Psalm of Heman the Ezrahite, giving instruction.}
\vs Psa 88:2 Let my prayer come before thee: incline thine ear unto my cry;
\vs Psa 88:3 For my soul is full of troubles: and my life draweth nigh unto the grave.
\vs Psa 88:4 I am counted with them that go down into the pit: I am as a man \bibemph{that hath} no strength:
\vs Psa 88:5 Free among the dead, like the slain that lie in the grave, whom thou rememberest no more: and they are cut off from thy hand.
\fn{from: or, by.}
\vs Psa 88:6 Thou hast laid me in the lowest pit, in darkness, in the deeps.
\vs Psa 88:7 Thy wrath lieth hard upon me, and thou hast afflicted \bibemph{me} with all thy waves. Selah.
\vs Psa 88:8 Thou hast put away mine acquaintance far from me; thou hast made me an abomination unto them: \bibemph{I am} shut up, and I cannot come forth.
\vs Psa 88:9 Mine eye mourneth by reason of affliction: \textsc{Lord}, I have called daily upon thee, I have stretched out my hands unto thee.
\vs Psa 88:10 Wilt thou shew wonders to the dead? shall the dead arise \bibemph{and} praise thee? Selah.
\vs Psa 88:11 Shall thy lovingkindness be declared in the grave? \bibemph{or} thy faithfulness in destruction?
\vs Psa 88:12 Shall thy wonders be known in the dark? and thy righteousness in the land of forgetfulness?
\vs Psa 88:13 But unto thee have I cried, O \textsc{Lord}; and in the morning shall my prayer prevent thee.
\vs Psa 88:14 \textsc{Lord}, why castest thou off my soul? \bibemph{why} hidest thou thy face from me?
\vs Psa 88:15 I \bibemph{am} afflicted and ready to die from \bibemph{my} youth up: \bibemph{while} I suffer thy terrors I am distracted.
\vs Psa 88:16 Thy fierce wrath goeth over me; thy terrors have cut me off.
\vs Psa 88:17 They came round about me daily like water; they compassed me about together.
\fn{daily: or, all the day.}
\vs Psa 88:18 Lover and friend hast thou put far from me, \bibemph{and} mine acquaintance into darkness.
\vs Psa 89:1 Maschil of Ethan the Ezrahite. I will sing of the mercies of the \textsc{Lord} for ever: with my mouth will I make known thy faithfulness to all generations.
\fn{Maschil...: or, A Psalm for Ethan the Ezrahite, to give instruction. to all...: Heb. to generation and generation.}
\vs Psa 89:2 For I have said, Mercy shall be built up for ever: thy faithfulness shalt thou establish in the very heavens.
\vs Psa 89:3 I have made a covenant with my chosen, I have sworn unto David my servant,
\vs Psa 89:4 Thy seed will I establish for ever, and build up thy throne to all generations. Selah.
\vs Psa 89:5 And the heavens shall praise thy wonders, O \textsc{Lord}: thy faithfulness also in the congregation of the saints.
\vs Psa 89:6 For who in the heaven can be compared unto the \textsc{Lord}? \bibemph{who} among the sons of the mighty can be likened unto the \textsc{Lord}?
\vs Psa 89:7 God is greatly to be feared in the assembly of the saints, and to be had in reverence of all \bibemph{them that are} about him.
\vs Psa 89:8 O \textsc{Lord} God of hosts, who \bibemph{is} a strong \textsc{Lord} like unto thee? or to thy faithfulness round about thee?
\vs Psa 89:9 Thou rulest the raging of the sea: when the waves thereof arise, thou stillest them.
\vs Psa 89:10 Thou hast broken Rahab in pieces, as one that is slain; thou hast scattered thine enemies with thy strong arm.
\fn{Rahab: or, Egypt. thy...: Heb. the arm of thy strength.}
\vs Psa 89:11 The heavens \bibemph{are} thine, the earth also \bibemph{is} thine: \bibemph{as for} the world and the fulness thereof, thou hast founded them.
\fn{the fulness...: or, all it containeth.}
\vs Psa 89:12 The north and the south thou hast created them: Tabor and Hermon shall rejoice in thy name.
\vs Psa 89:13 Thou hast a mighty arm: strong is thy hand, \bibemph{and} high is thy right hand.
\fn{a...: Heb. an arm with might.}
\vs Psa 89:14 Justice and judgment \bibemph{are} the habitation of thy throne: mercy and truth shall go before thy face.
\fn{habitation: or, establishment.}
\vs Psa 89:15 Blessed \bibemph{is} the people that know the joyful sound: they shall walk, O \textsc{Lord}, in the light of thy countenance.
\vs Psa 89:16 In thy name shall they rejoice all the day: and in thy righteousness shall they be exalted.
\vs Psa 89:17 For thou \bibemph{art} the glory of their strength: and in thy favour our horn shall be exalted.
\vs Psa 89:18 For the \textsc{Lord} \bibemph{is} our defence; and the Holy One of Israel \bibemph{is} our king.
\fn{the \textsc{Lord}...: or, our shield is of the \textsc{Lord}, and our king is of the Holy One of Israel.}
\vs Psa 89:19 Then thou spakest in vision to thy holy one, and saidst, I have laid help upon \bibemph{one that is} mighty; I have exalted \bibemph{one} chosen out of the people.
\vs Psa 89:20 I have found David my servant; with my holy oil have I anointed him:
\vs Psa 89:21 With whom my hand shall be established: mine arm also shall strengthen him.
\vs Psa 89:22 The enemy shall not exact upon him; nor the son of wickedness afflict him.
\vs Psa 89:23 And I will beat down his foes before his face, and plague them that hate him.
\vs Psa 89:24 But my faithfulness and my mercy \bibemph{shall be} with him: and in my name shall his horn be exalted.
\vs Psa 89:25 I will set his hand also in the sea, and his right hand in the rivers.
\vs Psa 89:26 He shall cry unto me, Thou \bibemph{art} my father, my God, and the rock of my salvation.
\vs Psa 89:27 Also I will make him \bibemph{my} firstborn, higher than the kings of the earth.
\vs Psa 89:28 My mercy will I keep for him for evermore, and my covenant shall stand fast with him.
\vs Psa 89:29 His seed also will I make \bibemph{to endure} for ever, and his throne as the days of heaven.
\vs Psa 89:30 If his children forsake my law, and walk not in my judgments;
\vs Psa 89:31 If they break my statutes, and keep not my commandments;
\fn{break: Heb. profane.}
\vs Psa 89:32 Then will I visit their transgression with the rod, and their iniquity with stripes.
\vs Psa 89:33 Nevertheless my lovingkindness will I not utterly take from him, nor suffer my faithfulness to fail.
\fn{will...: Heb. I will not make void from him. to fall: Heb. to lie.}
\vs Psa 89:34 My covenant will I not break, nor alter the thing that is gone out of my lips.
\vs Psa 89:35 Once have I sworn by my holiness that I will not lie unto David.
\fn{that...: Heb. if I lie.}
\vs Psa 89:36 His seed shall endure for ever, and his throne as the sun before me.
\vs Psa 89:37 It shall be established for ever as the moon, and \bibemph{as} a faithful witness in heaven. Selah.
\vs Psa 89:38 But thou hast cast off and abhorred, thou hast been wroth with thine anointed.
\vs Psa 89:39 Thou hast made void the covenant of thy servant: thou hast profaned his crown \bibemph{by casting it} to the ground.
\vs Psa 89:40 Thou hast broken down all his hedges; thou hast brought his strong holds to ruin.
\vs Psa 89:41 All that pass by the way spoil him: he is a reproach to his neighbours.
\vs Psa 89:42 Thou hast set up the right hand of his adversaries; thou hast made all his enemies to rejoice.
\vs Psa 89:43 Thou hast also turned the edge of his sword, and hast not made him to stand in the battle.
\vs Psa 89:44 Thou hast made his glory to cease, and cast his throne down to the ground.
\fn{glory: Heb. brightness.}
\vs Psa 89:45 The days of his youth hast thou shortened: thou hast covered him with shame. Selah.
\vs Psa 89:46 How long, \textsc{Lord}? wilt thou hide thyself for ever? shall thy wrath burn like fire?
\vs Psa 89:47 Remember how short my time is: wherefore hast thou made all men in vain?
\vs Psa 89:48 What man \bibemph{is he that} liveth, and shall not see death? shall he deliver his soul from the hand of the grave? Selah.
\vs Psa 89:49 Lord, where \bibemph{are} thy former lovingkindnesses, \bibemph{which} thou swarest unto David in thy truth?
\vs Psa 89:50 Remember, Lord, the reproach of thy servants; \bibemph{how} I do bear in my bosom \bibemph{the reproach of} all the mighty people;
\vs Psa 89:51 Wherewith thine enemies have reproached, O \textsc{Lord}; wherewith they have reproached the footsteps of thine anointed.
\vs Psa 89:52 Blessed \bibemph{be} the \textsc{Lord} for evermore. Amen, and Amen.
\vs Psa 90:1 A Prayer of Moses the man of God. Lord, thou hast been our dwelling place in all generations.
\fn{A Prayer...: or, A Prayer, being a Psalm of Moses. in...: Heb. in generation and generation.}
\vs Psa 90:2 Before the mountains were brought forth, or ever thou hadst formed the earth and the world, even from everlasting to everlasting, thou \bibemph{art} God.
\vs Psa 90:3 Thou turnest man to destruction; and sayest, Return, ye children of men.
\vs Psa 90:4 For a thousand years in thy sight \bibemph{are but} as yesterday when it is past, and \bibemph{as} a watch in the night.
\fn{when...: or, when he hath passed them.}
\vs Psa 90:5 Thou carriest them away as with a flood; they are \bibemph{as} a sleep: in the morning \bibemph{they are} like grass \bibemph{which} groweth up.
\fn{groweth...: or, is changed.}
\vs Psa 90:6 In the morning it flourisheth, and groweth up; in the evening it is cut down, and withereth.
\vs Psa 90:7 For we are consumed by thine anger, and by thy wrath are we troubled.
\vs Psa 90:8 Thou hast set our iniquities before thee, our secret \bibemph{sins} in the light of thy countenance.
\vs Psa 90:9 For all our days are passed away in thy wrath: we spend our years as a tale \bibemph{that is told}.
\fn{passed...: Heb. turned away. as a...: or, as a meditation.}
\vs Psa 90:10 The days of our years \bibemph{are} threescore years and ten; and if by reason of strength \bibemph{they be} fourscore years, yet \bibemph{is} their strength labour and sorrow; for it is soon cut off, and we fly away.
\fn{The days...: Heb. As for the days of our years, in them are seventy years.}
\vs Psa 90:11 Who knoweth the power of thine anger? even according to thy fear, \bibemph{so is} thy wrath.
\vs Psa 90:12 So teach \bibemph{us} to number our days, that we may apply \bibemph{our} hearts unto wisdom.
\fn{apply: Heb. cause to come.}
\vs Psa 90:13 Return, O \textsc{Lord}, how long? and let it repent thee concerning thy servants.
\vs Psa 90:14 O satisfy us early with thy mercy; that we may rejoice and be glad all our days.
\vs Psa 90:15 Make us glad according to the days \bibemph{wherein} thou hast afflicted us, \bibemph{and} the years \bibemph{wherein} we have seen evil.
\vs Psa 90:16 Let thy work appear unto thy servants, and thy glory unto their children.
\vs Psa 90:17 And let the beauty of the \textsc{Lord} our God be upon us: and establish thou the work of our hands upon us; yea, the work of our hands establish thou it.
\vs Psa 91:1 He that dwelleth in the secret place of the most High shall abide under the shadow of the Almighty.
\fn{abide: Heb. lodge.}
\vs Psa 91:2 I will say of the \textsc{Lord}, \bibemph{He is} my refuge and my fortress: my God; in him will I trust.
\vs Psa 91:3 Surely he shall deliver thee from the snare of the fowler, \bibemph{and} from the noisome pestilence.
\vs Psa 91:4 He shall cover thee with his feathers, and under his wings shalt thou trust: his truth \bibemph{shall be thy} shield and buckler.
\vs Psa 91:5 Thou shalt not be afraid for the terror by night; \bibemph{nor} for the arrow \bibemph{that} flieth by day;
\vs Psa 91:6 \bibemph{Nor} for the pestilence \bibemph{that} walketh in darkness; \bibemph{nor} for the destruction \bibemph{that} wasteth at noonday.
\vs Psa 91:7 A thousand shall fall at thy side, and ten thousand at thy right hand; \bibemph{but} it shall not come nigh thee.
\vs Psa 91:8 Only with thine eyes shalt thou behold and see the reward of the wicked.
\vs Psa 91:9 Because thou hast made the \textsc{Lord}, \bibemph{which is} my refuge, \bibemph{even} the most High, thy habitation;
\vs Psa 91:10 There shall no evil befall thee, neither shall any plague come nigh thy dwelling.
\vs Psa 91:11 For he shall give his angels charge over thee, to keep thee in all thy ways.
\vs Psa 91:12 They shall bear thee up in \bibemph{their} hands, lest thou dash thy foot against a stone.
\vs Psa 91:13 Thou shalt tread upon the lion and adder: the young lion and the dragon shalt thou trample under feet.
\fn{adder: or, asp.}
\vs Psa 91:14 Because he hath set his love upon me, therefore will I deliver him: I will set him on high, because he hath known my name.
\vs Psa 91:15 He shall call upon me, and I will answer him: I \bibemph{will be} with him in trouble; I will deliver him, and honour him.
\vs Psa 91:16 With long life will I satisfy him, and shew him my salvation.
\fn{long...: Heb. length of days.}
\vs Psa 92:1 A Psalm \bibemph{or} Song for the sabbath day. \bibemph{It is a} good \bibemph{thing} to give thanks unto the \textsc{Lord}, and to sing praises unto thy name, O most High:
\vs Psa 92:2 To shew forth thy lovingkindness in the morning, and thy faithfulness every night,
\fn{every...: Heb. in the nights.}
\vs Psa 92:3 Upon an instrument of ten strings, and upon the psaltery; upon the harp with a solemn sound.
\fn{the harp...: or, the solemn sound with the harp. a solemn...: Heb. Higgaion.}
\vs Psa 92:4 For thou, \textsc{Lord}, hast made me glad through thy work: I will triumph in the works of thy hands.
\vs Psa 92:5 O \textsc{Lord}, how great are thy works! \bibemph{and} thy thoughts are very deep.
\vs Psa 92:6 A brutish man knoweth not; neither doth a fool understand this.
\vs Psa 92:7 When the wicked spring as the grass, and when all the workers of iniquity do flourish; \bibemph{it is} that they shall be destroyed for ever:
\vs Psa 92:8 But thou, \textsc{Lord}, \bibemph{art most} high for evermore.
\vs Psa 92:9 For, lo, thine enemies, O \textsc{Lord}, for, lo, thine enemies shall perish; all the workers of iniquity shall be scattered.
\vs Psa 92:10 But my horn shalt thou exalt like \bibemph{the horn of} an unicorn: I shall be anointed with fresh oil.
\vs Psa 92:11 Mine eye also shall see \bibemph{my desire} on mine enemies, \bibemph{and} mine ears shall hear \bibemph{my desire} of the wicked that rise up against me.
\vs Psa 92:12 The righteous shall flourish like the palm tree: he shall grow like a cedar in Lebanon.
\vs Psa 92:13 Those that be planted in the house of the \textsc{Lord} shall flourish in the courts of our God.
\vs Psa 92:14 They shall still bring forth fruit in old age; they shall be fat and flourishing;
\fn{flourishing: Heb. green.}
\vs Psa 92:15 To shew that the \textsc{Lord} \bibemph{is} upright: \bibemph{he is} my rock, and \bibemph{there is} no unrighteousness in him.
\vs Psa 93:1 The \textsc{Lord} reigneth, he is clothed with majesty; the \textsc{Lord} is clothed with strength, \bibemph{wherewith} he hath girded himself: the world also is stablished, that it cannot be moved.
\vs Psa 93:2 Thy throne \bibemph{is} established of old: thou \bibemph{art} from everlasting.
\fn{of old: Heb. from then.}
\vs Psa 93:3 The floods have lifted up, O \textsc{Lord}, the floods have lifted up their voice; the floods lift up their waves.
\vs Psa 93:4 The \textsc{Lord} on high \bibemph{is} mightier than the noise of many waters, \bibemph{yea, than} the mighty waves of the sea.
\vs Psa 93:5 Thy testimonies are very sure: holiness becometh thine house, O \textsc{Lord}, for ever.
\fn{for ever: Heb. to length of days.}
\vs Psa 94:1 O \textsc{Lord} God, to whom vengeance belongeth; O God, to whom vengeance belongeth, shew thyself.
\fn{God...: Heb. God of revenges. shew...: Heb. shine forth.}
\vs Psa 94:2 Lift up thyself, thou judge of the earth: render a reward to the proud.
\vs Psa 94:3 \textsc{Lord}, how long shall the wicked, how long shall the wicked triumph?
\vs Psa 94:4 \bibemph{How long} shall they utter \bibemph{and} speak hard things? \bibemph{and} all the workers of iniquity boast themselves?
\vs Psa 94:5 They break in pieces thy people, O \textsc{Lord}, and afflict thine heritage.
\vs Psa 94:6 They slay the widow and the stranger, and murder the fatherless.
\vs Psa 94:7 Yet they say, The \textsc{Lord} shall not see, neither shall the God of Jacob regard \bibemph{it}.
\vs Psa 94:8 Understand, ye brutish among the people: and \bibemph{ye} fools, when will ye be wise?
\vs Psa 94:9 He that planted the ear, shall he not hear? he that formed the eye, shall he not see?
\vs Psa 94:10 He that chastiseth the heathen, shall not he correct? he that teacheth man knowledge, \bibemph{shall not he know}?
\vs Psa 94:11 The \textsc{Lord} knoweth the thoughts of man, that they \bibemph{are} vanity.
\vs Psa 94:12 Blessed \bibemph{is} the man whom thou chastenest, O \textsc{Lord}, and teachest him out of thy law;
\vs Psa 94:13 That thou mayest give him rest from the days of adversity, until the pit be digged for the wicked.
\vs Psa 94:14 For the \textsc{Lord} will not cast off his people, neither will he forsake his inheritance.
\vs Psa 94:15 But judgment shall return unto righteousness: and all the upright in heart shall follow it.
\fn{shall follow...: Heb. shall be after it.}
\vs Psa 94:16 Who will rise up for me against the evildoers? \bibemph{or} who will stand up for me against the workers of iniquity?
\vs Psa 94:17 Unless the \textsc{Lord} \bibemph{had been} my help, my soul had almost dwelt in silence.
\fn{almost: or, quickly.}
\vs Psa 94:18 When I said, My foot slippeth; thy mercy, O \textsc{Lord}, held me up.
\vs Psa 94:19 In the multitude of my thoughts within me thy comforts delight my soul.
\vs Psa 94:20 Shall the throne of iniquity have fellowship with thee, which frameth mischief by a law?
\vs Psa 94:21 They gather themselves together against the soul of the righteous, and condemn the innocent blood.
\vs Psa 94:22 But the \textsc{Lord} is my defence; and my God \bibemph{is} the rock of my refuge.
\vs Psa 94:23 And he shall bring upon them their own iniquity, and shall cut them off in their own wickedness; \bibemph{yea}, the \textsc{Lord} our God shall cut them off.
\vs Psa 95:1 O come, let us sing unto the \textsc{Lord}: let us make a joyful noise to the rock of our salvation.
\vs Psa 95:2 Let us come before his presence with thanksgiving, and make a joyful noise unto him with psalms.
\fn{come...: Heb. prevent his face.}
\vs Psa 95:3 For the \textsc{Lord} \bibemph{is} a great God, and a great King above all gods.
\vs Psa 95:4 In his hand \bibemph{are} the deep places of the earth: the strength of the hills \bibemph{is} his also.
\fn{In...: Heb. In whose. the strength...: or, the heights of the hills are his.}
\vs Psa 95:5 The sea \bibemph{is} his, and he made it: and his hands formed the dry \bibemph{land}.
\fn{The sea...: Heb. Whose the sea is.}
\vs Psa 95:6 O come, let us worship and bow down: let us kneel before the \textsc{Lord} our maker.
\vs Psa 95:7 For he \bibemph{is} our God; and we \bibemph{are} the people of his pasture, and the sheep of his hand. To day if ye will hear his voice,
\vs Psa 95:8 Harden not your heart, as in the provocation, \bibemph{and} as \bibemph{in} the day of temptation in the wilderness:
\fn{provocation: Heb. contention.}
\vs Psa 95:9 When your fathers tempted me, proved me, and saw my work.
\vs Psa 95:10 Forty years long was I grieved with \bibemph{this} generation, and said, It \bibemph{is} a people that do err in their heart, and they have not known my ways:
\vs Psa 95:11 Unto whom I sware in my wrath that they should not enter into my rest.
\fn{that...: Heb. if they enter into my rest.}
\vs Psa 96:1 O sing unto the \textsc{Lord} a new song: sing unto the \textsc{Lord}, all the earth.
\vs Psa 96:2 Sing unto the \textsc{Lord}, bless his name; shew forth his salvation from day to day.
\vs Psa 96:3 Declare his glory among the heathen, his wonders among all people.
\vs Psa 96:4 For the \textsc{Lord} \bibemph{is} great, and greatly to be praised: he \bibemph{is} to be feared above all gods.
\vs Psa 96:5 For all the gods of the nations \bibemph{are} idols: but the \textsc{Lord} made the heavens.
\vs Psa 96:6 Honour and majesty \bibemph{are} before him: strength and beauty \bibemph{are} in his sanctuary.
\vs Psa 96:7 Give unto the \textsc{Lord}, O ye kindreds of the people, give unto the \textsc{Lord} glory and strength.
\vs Psa 96:8 Give unto the \textsc{Lord} the glory \bibemph{due unto} his name: bring an offering, and come into his courts.
\fn{due...: Heb. of his name.}
\vs Psa 96:9 O worship the \textsc{Lord} in the beauty of holiness: fear before him, all the earth.
\fn{in the...: or, in the glorious sanctuary.}
\vs Psa 96:10 Say among the heathen \bibemph{that} the \textsc{Lord} reigneth: the world also shall be established that it shall not be moved: he shall judge the people righteously.
\vs Psa 96:11 Let the heavens rejoice, and let the earth be glad; let the sea roar, and the fulness thereof.
\fn{the fulness...: or, all it containeth.}
\vs Psa 96:12 Let the field be joyful, and all that \bibemph{is} therein: then shall all the trees of the wood rejoice
\vs Psa 96:13 Before the \textsc{Lord}: for he cometh, for he cometh to judge the earth: he shall judge the world with righteousness, and the people with his truth.
\vs Psa 97:1 The \textsc{Lord} reigneth; let the earth rejoice; let the multitude of isles be glad \bibemph{thereof}.
\fn{multitude...: Heb. many, or, great isles.}
\vs Psa 97:2 Clouds and darkness \bibemph{are} round about him: righteousness and judgment \bibemph{are} the habitation of his throne.
\fn{habitation: or, establishment.}
\vs Psa 97:3 A fire goeth before him, and burneth up his enemies round about.
\vs Psa 97:4 His lightnings enlightened the world: the earth saw, and trembled.
\vs Psa 97:5 The hills melted like wax at the presence of the \textsc{Lord}, at the presence of the Lord of the whole earth.
\vs Psa 97:6 The heavens declare his righteousness, and all the people see his glory.
\vs Psa 97:7 Confounded be all they that serve graven images, that boast themselves of idols: worship him, all \bibemph{ye} gods.
\vs Psa 97:8 Zion heard, and was glad; and the daughters of Judah rejoiced because of thy judgments, O \textsc{Lord}.
\vs Psa 97:9 For thou, \textsc{Lord}, \bibemph{art} high above all the earth: thou art exalted far above all gods.
\vs Psa 97:10 Ye that love the \textsc{Lord}, hate evil: he preserveth the souls of his saints; he delivereth them out of the hand of the wicked.
\vs Psa 97:11 Light is sown for the righteous, and gladness for the upright in heart.
\vs Psa 97:12 Rejoice in the \textsc{Lord}, ye righteous; and give thanks at the remembrance of his holiness.
\fn{at...: or, to the memorial.}
\vs Psa 98:1 A Psalm. O sing unto the \textsc{Lord} a new song; for he hath done marvellous things: his right hand, and his holy arm, hath gotten him the victory.
\vs Psa 98:2 The \textsc{Lord} hath made known his salvation: his righteousness hath he openly shewed in the sight of the heathen.
\fn{openly...: or, revealed.}
\vs Psa 98:3 He hath remembered his mercy and his truth toward the house of Israel: all the ends of the earth have seen the salvation of our God.
\vs Psa 98:4 Make a joyful noise unto the \textsc{Lord}, all the earth: make a loud noise, and rejoice, and sing praise.
\vs Psa 98:5 Sing unto the \textsc{Lord} with the harp; with the harp, and the voice of a psalm.
\vs Psa 98:6 With trumpets and sound of cornet make a joyful noise before the \textsc{Lord}, the King.
\vs Psa 98:7 Let the sea roar, and the fulness thereof; the world, and they that dwell therein.
\fn{the fulness...: or, all it containeth.}
\vs Psa 98:8 Let the floods clap \bibemph{their} hands: let the hills be joyful together
\vs Psa 98:9 Before the \textsc{Lord}; for he cometh to judge the earth: with righteousness shall he judge the world, and the people with equity.
\vs Psa 99:1 The \textsc{Lord} reigneth; let the people tremble: he sitteth \bibemph{between} the cherubims; let the earth be moved.
\fn{be moved: Heb. stagger.}
\vs Psa 99:2 The \textsc{Lord} \bibemph{is} great in Zion; and he \bibemph{is} high above all the people.
\vs Psa 99:3 Let them praise thy great and terrible name; \bibemph{for} it \bibemph{is} holy.
\vs Psa 99:4 The king's strength also loveth judgment; thou dost establish equity, thou executest judgment and righteousness in Jacob.
\vs Psa 99:5 Exalt ye the \textsc{Lord} our God, and worship at his footstool; \bibemph{for} he \bibemph{is} holy.
\fn{he is...: or, it is holy.}
\vs Psa 99:6 Moses and Aaron among his priests, and Samuel among them that call upon his name; they called upon the \textsc{Lord}, and he answered them.
\vs Psa 99:7 He spake unto them in the cloudy pillar: they kept his testimonies, and the ordinance \bibemph{that} he gave them.
\vs Psa 99:8 Thou answeredst them, O \textsc{Lord} our God: thou wast a God that forgavest them, though thou tookest vengeance of their inventions.
\vs Psa 99:9 Exalt the \textsc{Lord} our God, and worship at his holy hill; for the \textsc{Lord} our God \bibemph{is} holy.
\vs Psa 100:1 A Psalm of praise. Make a joyful noise unto the \textsc{Lord}, all ye lands.
\fn{praise: or, thanksgiving. all...: Heb. all the earth.}
\vs Psa 100:2 Serve the \textsc{Lord} with gladness: come before his presence with singing.
\vs Psa 100:3 Know ye that the \textsc{Lord} he \bibemph{is} God: \bibemph{it is} he \bibemph{that} hath made us, and not we ourselves; \bibemph{we are} his people, and the sheep of his pasture.
\fn{and not...: or, and his we are.}
\vs Psa 100:4 Enter into his gates with thanksgiving, \bibemph{and} into his courts with praise: be thankful unto him, \bibemph{and} bless his name.
\vs Psa 100:5 For the \textsc{Lord} \bibemph{is} good; his mercy \bibemph{is} everlasting; and his truth \bibemph{endureth} to all generations.
\fn{to all...: Heb. to generation and generation.}
\vs Psa 101:1 A Psalm of David. I will sing of mercy and judgment: unto thee, O \textsc{Lord}, will I sing.
\vs Psa 101:2 I will behave myself wisely in a perfect way. O when wilt thou come unto me? I will walk within my house with a perfect heart.
\vs Psa 101:3 I will set no wicked thing before mine eyes: I hate the work of them that turn aside; \bibemph{it} shall not cleave to me.
\fn{wicked...: Heb. thing of Belial.}
\vs Psa 101:4 A froward heart shall depart from me: I will not know a wicked \bibemph{person}.
\vs Psa 101:5 Whoso privily slandereth his neighbour, him will I cut off: him that hath an high look and a proud heart will not I suffer.
\vs Psa 101:6 Mine eyes \bibemph{shall be} upon the faithful of the land, that they may dwell with me: he that walketh in a perfect way, he shall serve me.
\fn{in a...: or, perfect in the way.}
\vs Psa 101:7 He that worketh deceit shall not dwell within my house: he that telleth lies shall not tarry in my sight.
\fn{shall not tarry: Heb. shall not be established.}
\vs Psa 101:8 I will early destroy all the wicked of the land; that I may cut off all wicked doers from the city of the \textsc{Lord}.
\vs Psa 102:1 A Prayer of the afflicted, when he is overwhelmed, and poureth out his complaint before the \textsc{Lord}. Hear my prayer, O \textsc{Lord}, and let my cry come unto thee.
\fn{of: or, for.}
\vs Psa 102:2 Hide not thy face from me in the day \bibemph{when} I am in trouble; incline thine ear unto me: in the day \bibemph{when} I call answer me speedily.
\vs Psa 102:3 For my days are consumed like smoke, and my bones are burned as an hearth.
\fn{like...: or, (as some read) into smoke.}
\vs Psa 102:4 My heart is smitten, and withered like grass; so that I forget to eat my bread.
\vs Psa 102:5 By reason of the voice of my groaning my bones cleave to my skin.
\fn{skin: or, flesh.}
\vs Psa 102:6 I am like a pelican of the wilderness: I am like an owl of the desert.
\vs Psa 102:7 I watch, and am as a sparrow alone upon the house top.
\vs Psa 102:8 Mine enemies reproach me all the day; \bibemph{and} they that are mad against me are sworn against me.
\vs Psa 102:9 For I have eaten ashes like bread, and mingled my drink with weeping,
\vs Psa 102:10 Because of thine indignation and thy wrath: for thou hast lifted me up, and cast me down.
\vs Psa 102:11 My days \bibemph{are} like a shadow that declineth; and I am withered like grass.
\vs Psa 102:12 But thou, O \textsc{Lord}, shalt endure for ever; and thy remembrance unto all generations.
\vs Psa 102:13 Thou shalt arise, \bibemph{and} have mercy upon Zion: for the time to favour her, yea, the set time, is come.
\vs Psa 102:14 For thy servants take pleasure in her stones, and favour the dust thereof.
\vs Psa 102:15 So the heathen shall fear the name of the \textsc{Lord}, and all the kings of the earth thy glory.
\vs Psa 102:16 When the \textsc{Lord} shall build up Zion, he shall appear in his glory.
\vs Psa 102:17 He will regard the prayer of the destitute, and not despise their prayer.
\vs Psa 102:18 This shall be written for the generation to come: and the people which shall be created shall praise the \textsc{Lord}.
\vs Psa 102:19 For he hath looked down from the height of his sanctuary; from heaven did the \textsc{Lord} behold the earth;
\vs Psa 102:20 To hear the groaning of the prisoner; to loose those that are appointed to death;
\fn{those...: Heb. the children of death.}
\vs Psa 102:21 To declare the name of the \textsc{Lord} in Zion, and his praise in Jerusalem;
\vs Psa 102:22 When the people are gathered together, and the kingdoms, to serve the \textsc{Lord}.
\vs Psa 102:23 He weakened my strength in the way; he shortened my days.
\fn{weakened: Heb. afflicted.}
\vs Psa 102:24 I said, O my God, take me not away in the midst of my days: thy years \bibemph{are} throughout all generations.
\vs Psa 102:25 Of old hast thou laid the foundation of the earth: and the heavens \bibemph{are} the work of thy hands.
\vs Psa 102:26 They shall perish, but thou shalt endure: yea, all of them shall wax old like a garment; as a vesture shalt thou change them, and they shall be changed:
\fn{endure: Heb. stand.}
\vs Psa 102:27 But thou \bibemph{art} the same, and thy years shall have no end.
\vs Psa 102:28 The children of thy servants shall continue, and their seed shall be established before thee.
\vs Psa 103:1 \bibemph{A Psalm} of David. Bless the \textsc{Lord}, O my soul: and all that is within me, \bibemph{bless} his holy name.
\vs Psa 103:2 Bless the \textsc{Lord}, O my soul, and forget not all his benefits:
\vs Psa 103:3 Who forgiveth all thine iniquities; who healeth all thy diseases;
\vs Psa 103:4 Who redeemeth thy life from destruction; who crowneth thee with lovingkindness and tender mercies;
\vs Psa 103:5 Who satisfieth thy mouth with good \bibemph{things; so that} thy youth is renewed like the eagle's.
\vs Psa 103:6 The \textsc{Lord} executeth righteousness and judgment for all that are oppressed.
\vs Psa 103:7 He made known his ways unto Moses, his acts unto the children of Israel.
\vs Psa 103:8 The \textsc{Lord} \bibemph{is} merciful and gracious, slow to anger, and plenteous in mercy.
\vs Psa 103:9 He will not always chide: neither will he keep \bibemph{his anger} for ever.
\vs Psa 103:10 He hath not dealt with us after our sins; nor rewarded us according to our iniquities.
\vs Psa 103:11 For as the heaven is high above the earth, \bibemph{so} great is his mercy toward them that fear him.
\fn{as...: Heb. according to the height of the heaven.}
\vs Psa 103:12 As far as the east is from the west, \bibemph{so} far hath he removed our transgressions from us.
\vs Psa 103:13 Like as a father pitieth \bibemph{his} children, \bibemph{so} the \textsc{Lord} pitieth them that fear him.
\vs Psa 103:14 For he knoweth our frame; he remembereth that we \bibemph{are} dust.
\vs Psa 103:15 \bibemph{As for} man, his days \bibemph{are} as grass: as a flower of the field, so he flourisheth.
\vs Psa 103:16 For the wind passeth over it, and it is gone; and the place thereof shall know it no more.
\fn{it is...: Heb. it is not.}
\vs Psa 103:17 But the mercy of the \textsc{Lord} \bibemph{is} from everlasting to everlasting upon them that fear him, and his righteousness unto children's children;
\vs Psa 103:18 To such as keep his covenant, and to those that remember his commandments to do them.
\vs Psa 103:19 The \textsc{Lord} hath prepared his throne in the heavens; and his kingdom ruleth over all.
\vs Psa 103:20 Bless the \textsc{Lord}, ye his angels, that excel in strength, that do his commandments, hearkening unto the voice of his word.
\fn{that excel...: Heb. mighty in strength.}
\vs Psa 103:21 Bless ye the \textsc{Lord}, all \bibemph{ye} his hosts; \bibemph{ye} ministers of his, that do his pleasure.
\vs Psa 103:22 Bless the \textsc{Lord}, all his works in all places of his dominion: bless the \textsc{Lord}, O my soul.
\vs Psa 104:1 Bless the \textsc{Lord}, O my soul. O \textsc{Lord} my God, thou art very great; thou art clothed with honour and majesty.
\vs Psa 104:2 Who coverest \bibemph{thyself} with light as \bibemph{with} a garment: who stretchest out the heavens like a curtain:
\vs Psa 104:3 Who layeth the beams of his chambers in the waters: who maketh the clouds his chariot: who walketh upon the wings of the wind:
\vs Psa 104:4 Who maketh his angels spirits; his ministers a flaming fire:
\vs Psa 104:5 \bibemph{Who} laid the foundations of the earth, \bibemph{that} it should not be removed for ever.
\fn{Who...: Heb. He hath founded the earth upon her bases.}
\vs Psa 104:6 Thou coveredst it with the deep as \bibemph{with} a garment: the waters stood above the mountains.
\vs Psa 104:7 At thy rebuke they fled; at the voice of thy thunder they hasted away.
\vs Psa 104:8 They go up by the mountains; they go down by the valleys unto the place which thou hast founded for them.
\fn{They go up...: or, The mountains ascend, the valleys descend.}
\vs Psa 104:9 Thou hast set a bound that they may not pass over; that they turn not again to cover the earth.
\vs Psa 104:10 He sendeth the springs into the valleys, \bibemph{which} run among the hills.
\fn{He: Heb. Who. run: Heb. walk.}
\vs Psa 104:11 They give drink to every beast of the field: the wild asses quench their thirst.
\fn{quench: Heb. break.}
\vs Psa 104:12 By them shall the fowls of the heaven have their habitation, \bibemph{which} sing among the branches.
\fn{sing: Heb. give a voice.}
\vs Psa 104:13 He watereth the hills from his chambers: the earth is satisfied with the fruit of thy works.
\vs Psa 104:14 He causeth the grass to grow for the cattle, and herb for the service of man: that he may bring forth food out of the earth;
\vs Psa 104:15 And wine \bibemph{that} maketh glad the heart of man, \bibemph{and} oil to make \bibemph{his} face to shine, and bread \bibemph{which} strengtheneth man's heart.
\fn{oil...: Heb. to make his face shine with oil, or, more than oil.}
\vs Psa 104:16 The trees of the \textsc{Lord} are full \bibemph{of sap}; the cedars of Lebanon, which he hath planted;
\vs Psa 104:17 Where the birds make their nests: \bibemph{as for} the stork, the fir trees \bibemph{are} her house.
\vs Psa 104:18 The high hills \bibemph{are} a refuge for the wild goats; \bibemph{and} the rocks for the conies.
\vs Psa 104:19 He appointed the moon for seasons: the sun knoweth his going down.
\vs Psa 104:20 Thou makest darkness, and it is night: wherein all the beasts of the forest do creep \bibemph{forth}.
\fn{beasts...: Heb. beasts thereof do trample on the forest.}
\vs Psa 104:21 The young lions roar after their prey, and seek their meat from God.
\vs Psa 104:22 The sun ariseth, they gather themselves together, and lay them down in their dens.
\vs Psa 104:23 Man goeth forth unto his work and to his labour until the evening.
\vs Psa 104:24 O \textsc{Lord}, how manifold are thy works! in wisdom hast thou made them all: the earth is full of thy riches.
\vs Psa 104:25 \bibemph{So is} this great and wide sea, wherein \bibemph{are} things creeping innumerable, both small and great beasts.
\vs Psa 104:26 There go the ships: \bibemph{there is} that leviathan, \bibemph{whom} thou hast made to play therein.
\fn{made: Heb. formed.}
\vs Psa 104:27 These wait all upon thee; that thou mayest give \bibemph{them} their meat in due season.
\vs Psa 104:28 \bibemph{That} thou givest them they gather: thou openest thine hand, they are filled with good.
\vs Psa 104:29 Thou hidest thy face, they are troubled: thou takest away their breath, they die, and return to their dust.
\vs Psa 104:30 Thou sendest forth thy spirit, they are created: and thou renewest the face of the earth.
\vs Psa 104:31 The glory of the \textsc{Lord} shall endure for ever: the \textsc{Lord} shall rejoice in his works.
\fn{endure: Heb. be.}
\vs Psa 104:32 He looketh on the earth, and it trembleth: he toucheth the hills, and they smoke.
\vs Psa 104:33 I will sing unto the \textsc{Lord} as long as I live: I will sing praise to my God while I have my being.
\vs Psa 104:34 My meditation of him shall be sweet: I will be glad in the \textsc{Lord}.
\vs Psa 104:35 Let the sinners be consumed out of the earth, and let the wicked be no more. Bless thou the \textsc{Lord}, O my soul. Praise ye the \textsc{Lord}.
\vs Psa 105:1 O give thanks unto the \textsc{Lord}; call upon his name: make known his deeds among the people.
\vs Psa 105:2 Sing unto him, sing psalms unto him: talk ye of all his wondrous works.
\vs Psa 105:3 Glory ye in his holy name: let the heart of them rejoice that seek the \textsc{Lord}.
\vs Psa 105:4 Seek the \textsc{Lord}, and his strength: seek his face evermore.
\vs Psa 105:5 Remember his marvellous works that he hath done; his wonders, and the judgments of his mouth;
\vs Psa 105:6 O ye seed of Abraham his servant, ye children of Jacob his chosen.
\vs Psa 105:7 He \bibemph{is} the \textsc{Lord} our God: his judgments \bibemph{are} in all the earth.
\vs Psa 105:8 He hath remembered his covenant for ever, the word \bibemph{which} he commanded to a thousand generations.
\vs Psa 105:9 Which \bibemph{covenant} he made with Abraham, and his oath unto Isaac;
\vs Psa 105:10 And confirmed the same unto Jacob for a law, \bibemph{and} to Israel \bibemph{for} an everlasting covenant:
\vs Psa 105:11 Saying, Unto thee will I give the land of Canaan, the lot of your inheritance:
\fn{lot: Heb. cord.}
\vs Psa 105:12 When they were \bibemph{but} a few men in number; yea, very few, and strangers in it.
\vs Psa 105:13 When they went from one nation to another, from \bibemph{one} kingdom to another people;
\vs Psa 105:14 He suffered no man to do them wrong: yea, he reproved kings for their sakes;
\vs Psa 105:15 \bibemph{Saying}, Touch not mine anointed, and do my prophets no harm.
\vs Psa 105:16 Moreover he called for a famine upon the land: he brake the whole staff of bread.
\vs Psa 105:17 He sent a man before them, \bibemph{even} Joseph, \bibemph{who} was sold for a servant:
\vs Psa 105:18 Whose feet they hurt with fetters: he was laid in iron:
\fn{he...: Heb. his soul came into iron.}
\vs Psa 105:19 Until the time that his word came: the word of the \textsc{Lord} tried him.
\vs Psa 105:20 The king sent and loosed him; \bibemph{even} the ruler of the people, and let him go free.
\vs Psa 105:21 He made him lord of his house, and ruler of all his substance:
\fn{substance: Heb. possession.}
\vs Psa 105:22 To bind his princes at his pleasure; and teach his senators wisdom.
\vs Psa 105:23 Israel also came into Egypt; and Jacob sojourned in the land of Ham.
\vs Psa 105:24 And he increased his people greatly; and made them stronger than their enemies.
\vs Psa 105:25 He turned their heart to hate his people, to deal subtilly with his servants.
\vs Psa 105:26 He sent Moses his servant; \bibemph{and} Aaron whom he had chosen.
\vs Psa 105:27 They shewed his signs among them, and wonders in the land of Ham.
\fn{his...: Heb. words of his signs.}
\vs Psa 105:28 He sent darkness, and made it dark; and they rebelled not against his word.
\vs Psa 105:29 He turned their waters into blood, and slew their fish.
\vs Psa 105:30 Their land brought forth frogs in abundance, in the chambers of their kings.
\vs Psa 105:31 He spake, and there came divers sorts of flies, \bibemph{and} lice in all their coasts.
\vs Psa 105:32 He gave them hail for rain, \bibemph{and} flaming fire in their land.
\fn{them...: Heb. their rain hail.}
\vs Psa 105:33 He smote their vines also and their fig trees; and brake the trees of their coasts.
\vs Psa 105:34 He spake, and the locusts came, and caterpillers, and that without number,
\vs Psa 105:35 And did eat up all the herbs in their land, and devoured the fruit of their ground.
\vs Psa 105:36 He smote also all the firstborn in their land, the chief of all their strength.
\vs Psa 105:37 He brought them forth also with silver and gold: and \bibemph{there was} not one feeble \bibemph{person} among their tribes.
\vs Psa 105:38 Egypt was glad when they departed: for the fear of them fell upon them.
\vs Psa 105:39 He spread a cloud for a covering; and fire to give light in the night.
\vs Psa 105:40 \bibemph{The people} asked, and he brought quails, and satisfied them with the bread of heaven.
\vs Psa 105:41 He opened the rock, and the waters gushed out; they ran in the dry places \bibemph{like} a river.
\vs Psa 105:42 For he remembered his holy promise, \bibemph{and} Abraham his servant.
\vs Psa 105:43 And he brought forth his people with joy, \bibemph{and} his chosen with gladness:
\fn{gladness: Heb. singing.}
\vs Psa 105:44 And gave them the lands of the heathen: and they inherited the labour of the people;
\vs Psa 105:45 That they might observe his statutes, and keep his laws. Praise ye the \textsc{Lord}.
\fn{Praise...: Heb. Hallelujah.}
\vs Psa 106:1 Praise ye the \textsc{Lord}. O give thanks unto the \textsc{Lord}; for \bibemph{he is} good: for his mercy \bibemph{endureth} for ever.
\fn{Praise...: Heb. Hallelujah.}
\vs Psa 106:2 Who can utter the mighty acts of the \textsc{Lord}? \bibemph{who} can shew forth all his praise?
\vs Psa 106:3 Blessed \bibemph{are} they that keep judgment, \bibemph{and} he that doeth righteousness at all times.
\vs Psa 106:4 Remember me, O \textsc{Lord}, with the favour \bibemph{that thou bearest unto} thy people: O visit me with thy salvation;
\vs Psa 106:5 That I may see the good of thy chosen, that I may rejoice in the gladness of thy nation, that I may glory with thine inheritance.
\vs Psa 106:6 We have sinned with our fathers, we have committed iniquity, we have done wickedly.
\vs Psa 106:7 Our fathers understood not thy wonders in Egypt; they remembered not the multitude of thy mercies; but provoked \bibemph{him} at the sea, \bibemph{even} at the Red sea.
\vs Psa 106:8 Nevertheless he saved them for his name's sake, that he might make his mighty power to be known.
\vs Psa 106:9 He rebuked the Red sea also, and it was dried up: so he led them through the depths, as through the wilderness.
\vs Psa 106:10 And he saved them from the hand of him that hated \bibemph{them}, and redeemed them from the hand of the enemy.
\vs Psa 106:11 And the waters covered their enemies: there was not one of them left.
\vs Psa 106:12 Then believed they his words; they sang his praise.
\vs Psa 106:13 They soon forgat his works; they waited not for his counsel:
\fn{They soon...: Heb. They made haste, they forgat.}
\vs Psa 106:14 But lusted exceedingly in the wilderness, and tempted God in the desert.
\fn{lusted...: Heb. lusted a lust.}
\vs Psa 106:15 And he gave them their request; but sent leanness into their soul.
\vs Psa 106:16 They envied Moses also in the camp, \bibemph{and} Aaron the saint of the \textsc{Lord}.
\vs Psa 106:17 The earth opened and swallowed up Dathan, and covered the company of Abiram.
\vs Psa 106:18 And a fire was kindled in their company; the flame burned up the wicked.
\vs Psa 106:19 They made a calf in Horeb, and worshipped the molten image.
\vs Psa 106:20 Thus they changed their glory into the similitude of an ox that eateth grass.
\vs Psa 106:21 They forgat God their saviour, which had done great things in Egypt;
\vs Psa 106:22 Wondrous works in the land of Ham, \bibemph{and} terrible things by the Red sea.
\vs Psa 106:23 Therefore he said that he would destroy them, had not Moses his chosen stood before him in the breach, to turn away his wrath, lest he should destroy \bibemph{them}.
\vs Psa 106:24 Yea, they despised the pleasant land, they believed not his word:
\fn{the pleasant...: Heb. a land of desire.}
\vs Psa 106:25 But murmured in their tents, \bibemph{and} hearkened not unto the voice of the \textsc{Lord}.
\vs Psa 106:26 Therefore he lifted up his hand against them, to overthrow them in the wilderness:
\fn{to overthrow: Heb. to make them fall.}
\vs Psa 106:27 To overthrow their seed also among the nations, and to scatter them in the lands.
\fn{to overthrow: Heb. to make them fall.}
\vs Psa 106:28 They joined themselves also unto Baalpeor, and ate the sacrifices of the dead.
\vs Psa 106:29 Thus they provoked \bibemph{him} to anger with their inventions: and the plague brake in upon them.
\vs Psa 106:30 Then stood up Phinehas, and executed judgment: and \bibemph{so} the plague was stayed.
\vs Psa 106:31 And that was counted unto him for righteousness unto all generations for evermore.
\vs Psa 106:32 They angered \bibemph{him} also at the waters of strife, so that it went ill with Moses for their sakes:
\vs Psa 106:33 Because they provoked his spirit, so that he spake unadvisedly with his lips.
\vs Psa 106:34 They did not destroy the nations, concerning whom the \textsc{Lord} commanded them:
\vs Psa 106:35 But were mingled among the heathen, and learned their works.
\vs Psa 106:36 And they served their idols: which were a snare unto them.
\vs Psa 106:37 Yea, they sacrificed their sons and their daughters unto devils,
\vs Psa 106:38 And shed innocent blood, \bibemph{even} the blood of their sons and of their daughters, whom they sacrificed unto the idols of Canaan: and the land was polluted with blood.
\vs Psa 106:39 Thus were they defiled with their own works, and went a whoring with their own inventions.
\vs Psa 106:40 Therefore was the wrath of the \textsc{Lord} kindled against his people, insomuch that he abhorred his own inheritance.
\vs Psa 106:41 And he gave them into the hand of the heathen; and they that hated them ruled over them.
\vs Psa 106:42 Their enemies also oppressed them, and they were brought into subjection under their hand.
\vs Psa 106:43 Many times did he deliver them; but they provoked \bibemph{him} with their counsel, and were brought low for their iniquity.
\fn{brought...: or, impoverished, or, weakened.}
\vs Psa 106:44 Nevertheless he regarded their affliction, when he heard their cry:
\vs Psa 106:45 And he remembered for them his covenant, and repented according to the multitude of his mercies.
\vs Psa 106:46 He made them also to be pitied of all those that carried them captives.
\vs Psa 106:47 Save us, O \textsc{Lord} our God, and gather us from among the heathen, to give thanks unto thy holy name, \bibemph{and} to triumph in thy praise.
\vs Psa 106:48 Blessed \bibemph{be} the \textsc{Lord} God of Israel from everlasting to everlasting: and let all the people say, Amen. Praise ye the \textsc{Lord}.
\fn{Praise...: Heb. Hallelujah.}
\vs Psa 107:1 O give thanks unto the \textsc{Lord}, for \bibemph{he is} good: for his mercy \bibemph{endureth} for ever.
\vs Psa 107:2 Let the redeemed of the \textsc{Lord} say \bibemph{so}, whom he hath redeemed from the hand of the enemy;
\vs Psa 107:3 And gathered them out of the lands, from the east, and from the west, from the north, and from the south.
\fn{south: Heb. sea.}
\vs Psa 107:4 They wandered in the wilderness in a solitary way; they found no city to dwell in.
\vs Psa 107:5 Hungry and thirsty, their soul fainted in them.
\vs Psa 107:6 Then they cried unto the \textsc{Lord} in their trouble, \bibemph{and} he delivered them out of their distresses.
\vs Psa 107:7 And he led them forth by the right way, that they might go to a city of habitation.
\vs Psa 107:8 Oh that \bibemph{men} would praise the \textsc{Lord} \bibemph{for} his goodness, and \bibemph{for} his wonderful works to the children of men!
\vs Psa 107:9 For he satisfieth the longing soul, and filleth the hungry soul with goodness.
\vs Psa 107:10 Such as sit in darkness and in the shadow of death, \bibemph{being} bound in affliction and iron;
\vs Psa 107:11 Because they rebelled against the words of God, and contemned the counsel of the most High:
\vs Psa 107:12 Therefore he brought down their heart with labour; they fell down, and \bibemph{there was} none to help.
\vs Psa 107:13 Then they cried unto the \textsc{Lord} in their trouble, \bibemph{and} he saved them out of their distresses.
\vs Psa 107:14 He brought them out of darkness and the shadow of death, and brake their bands in sunder.
\vs Psa 107:15 Oh that \bibemph{men} would praise the \textsc{Lord} \bibemph{for} his goodness, and \bibemph{for} his wonderful works to the children of men!
\vs Psa 107:16 For he hath broken the gates of brass, and cut the bars of iron in sunder.
\vs Psa 107:17 Fools because of their transgression, and because of their iniquities, are afflicted.
\vs Psa 107:18 Their soul abhorreth all manner of meat; and they draw near unto the gates of death.
\vs Psa 107:19 Then they cry unto the \textsc{Lord} in their trouble, \bibemph{and} he saveth them out of their distresses.
\vs Psa 107:20 He sent his word, and healed them, and delivered \bibemph{them} from their destructions.
\vs Psa 107:21 Oh that \bibemph{men} would praise the \textsc{Lord} \bibemph{for} his goodness, and \bibemph{for} his wonderful works to the children of men!
\vs Psa 107:22 And let them sacrifice the sacrifices of thanksgiving, and declare his works with rejoicing.
\fn{rejoicing: Heb. singing.}
\vs Psa 107:23 They that go down to the sea in ships, that do business in great waters;
\vs Psa 107:24 These see the works of the \textsc{Lord}, and his wonders in the deep.
\vs Psa 107:25 For he commandeth, and raiseth the stormy wind, which lifteth up the waves thereof.
\fn{raiseth: Heb. maketh to stand.}
\vs Psa 107:26 They mount up to the heaven, they go down again to the depths: their soul is melted because of trouble.
\vs Psa 107:27 They reel to and fro, and stagger like a drunken man, and are at their wits' end.
\fn{are...: Heb. all their wisdom is swallowed up.}
\vs Psa 107:28 Then they cry unto the \textsc{Lord} in their trouble, and he bringeth them out of their distresses.
\vs Psa 107:29 He maketh the storm a calm, so that the waves thereof are still.
\vs Psa 107:30 Then are they glad because they be quiet; so he bringeth them unto their desired haven.
\vs Psa 107:31 Oh that \bibemph{men} would praise the \textsc{Lord} \bibemph{for} his goodness, and \bibemph{for} his wonderful works to the children of men!
\vs Psa 107:32 Let them exalt him also in the congregation of the people, and praise him in the assembly of the elders.
\vs Psa 107:33 He turneth rivers into a wilderness, and the watersprings into dry ground;
\vs Psa 107:34 A fruitful land into barrenness, for the wickedness of them that dwell therein.
\fn{barrenness: Heb. saltiness.}
\vs Psa 107:35 He turneth the wilderness into a standing water, and dry ground into watersprings.
\vs Psa 107:36 And there he maketh the hungry to dwell, that they may prepare a city for habitation;
\vs Psa 107:37 And sow the fields, and plant vineyards, which may yield fruits of increase.
\vs Psa 107:38 He blesseth them also, so that they are multiplied greatly; and suffereth not their cattle to decrease.
\vs Psa 107:39 Again, they are minished and brought low through oppression, affliction, and sorrow.
\vs Psa 107:40 He poureth contempt upon princes, and causeth them to wander in the wilderness, \bibemph{where there is} no way.
\fn{wilderness: or, void place.}
\vs Psa 107:41 Yet setteth he the poor on high from affliction, and maketh \bibemph{him} families like a flock.
\fn{from: or, after.}
\vs Psa 107:42 The righteous shall see \bibemph{it}, and rejoice: and all iniquity shall stop her mouth.
\vs Psa 107:43 Whoso \bibemph{is} wise, and will observe these \bibemph{things}, even they shall understand the lovingkindness of the \textsc{Lord}.
\vs Psa 108:1 A Song \bibemph{or} Psalm of David. O God, my heart is fixed; I will sing and give praise, even with my glory.
\vs Psa 108:2 Awake, psaltery and harp: I \bibemph{myself} will awake early.
\vs Psa 108:3 I will praise thee, O \textsc{Lord}, among the people: and I will sing praises unto thee among the nations.
\vs Psa 108:4 For thy mercy \bibemph{is} great above the heavens: and thy truth \bibemph{reacheth} unto the clouds.
\fn{clouds: or, skies.}
\vs Psa 108:5 Be thou exalted, O God, above the heavens: and thy glory above all the earth;
\vs Psa 108:6 That thy beloved may be delivered: save \bibemph{with} thy right hand, and answer me.
\vs Psa 108:7 God hath spoken in his holiness; I will rejoice, I will divide Shechem, and mete out the valley of Succoth.
\vs Psa 108:8 Gilead \bibemph{is} mine; Manasseh \bibemph{is} mine; Ephraim also \bibemph{is} the strength of mine head; Judah \bibemph{is} my lawgiver;
\vs Psa 108:9 Moab \bibemph{is} my washpot; over Edom will I cast out my shoe; over Philistia will I triumph.
\vs Psa 108:10 Who will bring me into the strong city? who will lead me into Edom?
\vs Psa 108:11 \bibemph{Wilt} not \bibemph{thou}, O God, \bibemph{who} hast cast us off? and wilt not thou, O God, go forth with our hosts?
\vs Psa 108:12 Give us help from trouble: for vain \bibemph{is} the help of man.
\vs Psa 108:13 Through God we shall do valiantly: for he \bibemph{it is that} shall tread down our enemies.
\vs Psa 109:1 To the chief Musician, A Psalm of David. Hold not thy peace, O God of my praise;
\vs Psa 109:2 For the mouth of the wicked and the mouth of the deceitful are opened against me: they have spoken against me with a lying tongue.
\fn{of the deceitful: Heb. of deceit. are opened: Heb. have opened themselves.}
\vs Psa 109:3 They compassed me about also with words of hatred; and fought against me without a cause.
\vs Psa 109:4 For my love they are my adversaries: but I \bibemph{give myself unto} prayer.
\vs Psa 109:5 And they have rewarded me evil for good, and hatred for my love.
\vs Psa 109:6 Set thou a wicked man over him: and let Satan stand at his right hand.
\fn{Satan: or, an adversary.}
\vs Psa 109:7 When he shall be judged, let him be condemned: and let his prayer become sin.
\fn{be condemned: Heb. go out guilty, or, wicked.}
\vs Psa 109:8 Let his days be few; \bibemph{and} let another take his office.
\fn{office: or, charge.}
\vs Psa 109:9 Let his children be fatherless, and his wife a widow.
\vs Psa 109:10 Let his children be continually vagabonds, and beg: let them seek \bibemph{their bread} also out of their desolate places.
\vs Psa 109:11 Let the extortioner catch all that he hath; and let the strangers spoil his labour.
\vs Psa 109:12 Let there be none to extend mercy unto him: neither let there be any to favour his fatherless children.
\vs Psa 109:13 Let his posterity be cut off; \bibemph{and} in the generation following let their name be blotted out.
\vs Psa 109:14 Let the iniquity of his fathers be remembered with the \textsc{Lord}; and let not the sin of his mother be blotted out.
\vs Psa 109:15 Let them be before the \textsc{Lord} continually, that he may cut off the memory of them from the earth.
\vs Psa 109:16 Because that he remembered not to shew mercy, but persecuted the poor and needy man, that he might even slay the broken in heart.
\vs Psa 109:17 As he loved cursing, so let it come unto him: as he delighted not in blessing, so let it be far from him.
\vs Psa 109:18 As he clothed himself with cursing like as with his garment, so let it come into his bowels like water, and like oil into his bones.
\fn{into his bowels: Heb. within him.}
\vs Psa 109:19 Let it be unto him as the garment \bibemph{which} covereth him, and for a girdle wherewith he is girded continually.
\vs Psa 109:20 \bibemph{Let} this \bibemph{be} the reward of mine adversaries from the \textsc{Lord}, and of them that speak evil against my soul.
\vs Psa 109:21 But do thou for me, O GOD the Lord, for thy name's sake: because thy mercy \bibemph{is} good, deliver thou me.
\vs Psa 109:22 For I \bibemph{am} poor and needy, and my heart is wounded within me.
\vs Psa 109:23 I am gone like the shadow when it declineth: I am tossed up and down as the locust.
\vs Psa 109:24 My knees are weak through fasting; and my flesh faileth of fatness.
\vs Psa 109:25 I became also a reproach unto them: \bibemph{when} they looked upon me they shaked their heads.
\vs Psa 109:26 Help me, O \textsc{Lord} my God: O save me according to thy mercy:
\vs Psa 109:27 That they may know that this \bibemph{is} thy hand; \bibemph{that} thou, \textsc{Lord}, hast done it.
\vs Psa 109:28 Let them curse, but bless thou: when they arise, let them be ashamed; but let thy servant rejoice.
\vs Psa 109:29 Let mine adversaries be clothed with shame, and let them cover themselves with their own confusion, as with a mantle.
\vs Psa 109:30 I will greatly praise the \textsc{Lord} with my mouth; yea, I will praise him among the multitude.
\vs Psa 109:31 For he shall stand at the right hand of the poor, to save \bibemph{him} from those that condemn his soul.
\fn{those...: Heb. the judges of.}
\vs Psa 110:1 A Psalm of David. The \textsc{Lord} said unto my Lord, Sit thou at my right hand, until I make thine enemies thy footstool.
\vs Psa 110:2 The \textsc{Lord} shall send the rod of thy strength out of Zion: rule thou in the midst of thine enemies.
\vs Psa 110:3 Thy people \bibemph{shall be} willing in the day of thy power, in the beauties of holiness from the womb of the morning: thou hast the dew of thy youth.
\fn{from...: or, more than the womb of the morning: thou shalt have, etc.}
\vs Psa 110:4 The \textsc{Lord} hath sworn, and will not repent, Thou \bibemph{art} a priest for ever after the order of Melchizedek.
\vs Psa 110:5 The Lord at thy right hand shall strike through kings in the day of his wrath.
\vs Psa 110:6 He shall judge among the heathen, he shall fill \bibemph{the places} with the dead bodies; he shall wound the heads over many countries.
\fn{many: or, great.}
\vs Psa 110:7 He shall drink of the brook in the way: therefore shall he lift up the head.
\vs Psa 111:1 Praise ye the \textsc{Lord}. I will praise the \textsc{Lord} with \bibemph{my} whole heart, in the assembly of the upright, and \bibemph{in} the congregation.
\fn{Praise ye...: Heb. Hallelujah.}
\vs Psa 111:2 The works of the \textsc{Lord} \bibemph{are} great, sought out of all them that have pleasure therein.
\vs Psa 111:3 His work \bibemph{is} honourable and glorious: and his righteousness endureth for ever.
\vs Psa 111:4 He hath made his wonderful works to be remembered: the \textsc{Lord} \bibemph{is} gracious and full of compassion.
\vs Psa 111:5 He hath given meat unto them that fear him: he will ever be mindful of his covenant.
\fn{meat: Heb. prey.}
\vs Psa 111:6 He hath shewed his people the power of his works, that he may give them the heritage of the heathen.
\vs Psa 111:7 The works of his hands \bibemph{are} verity and judgment; all his commandments \bibemph{are} sure.
\vs Psa 111:8 They stand fast for ever and ever, \bibemph{and are} done in truth and uprightness.
\fn{stand...: Heb. are established.}
\vs Psa 111:9 He sent redemption unto his people: he hath commanded his covenant for ever: holy and reverend \bibemph{is} his name.
\vs Psa 111:10 The fear of the \textsc{Lord} \bibemph{is} the beginning of wisdom: a good understanding have all they that do \bibemph{his commandments}: his praise endureth for ever.
\fn{a good...: or, good success. his commandments: Heb. them.}
\vs Psa 112:1 Praise ye the \textsc{Lord}. Blessed \bibemph{is} the man \bibemph{that} feareth the \textsc{Lord}, \bibemph{that} delighteth greatly in his commandments.
\fn{Praise...: Heb. Hallelujah.}
\vs Psa 112:2 His seed shall be mighty upon earth: the generation of the upright shall be blessed.
\vs Psa 112:3 Wealth and riches \bibemph{shall be} in his house: and his righteousness endureth for ever.
\vs Psa 112:4 Unto the upright there ariseth light in the darkness: \bibemph{he is} gracious, and full of compassion, and righteous.
\vs Psa 112:5 A good man sheweth favour, and lendeth: he will guide his affairs with discretion.
\fn{discretion: Heb. judgment.}
\vs Psa 112:6 Surely he shall not be moved for ever: the righteous shall be in everlasting remembrance.
\vs Psa 112:7 He shall not be afraid of evil tidings: his heart is fixed, trusting in the \textsc{Lord}.
\vs Psa 112:8 His heart \bibemph{is} established, he shall not be afraid, until he see \bibemph{his desire} upon his enemies.
\vs Psa 112:9 He hath dispersed, he hath given to the poor; his righteousness endureth for ever; his horn shall be exalted with honour.
\vs Psa 112:10 The wicked shall see \bibemph{it}, and be grieved; he shall gnash with his teeth, and melt away: the desire of the wicked shall perish.
\vs Psa 113:1 Praise ye the \textsc{Lord}. Praise, O ye servants of the \textsc{Lord}, praise the name of the \textsc{Lord}.
\fn{Praise ye...: Heb. Hallelujah.}
\vs Psa 113:2 Blessed be the name of the \textsc{Lord} from this time forth and for evermore.
\vs Psa 113:3 From the rising of the sun unto the going down of the same the \textsc{Lord's} name \bibemph{is} to be praised.
\vs Psa 113:4 The \textsc{Lord} \bibemph{is} high above all nations, \bibemph{and} his glory above the heavens.
\vs Psa 113:5 Who \bibemph{is} like unto the \textsc{Lord} our God, who dwelleth on high,
\fn{dwelleth...: Heb. exalteth himself to dwell.}
\vs Psa 113:6 Who humbleth \bibemph{himself} to behold \bibemph{the things that are} in heaven, and in the earth!
\vs Psa 113:7 He raiseth up the poor out of the dust, \bibemph{and} lifteth the needy out of the dunghill;
\vs Psa 113:8 That he may set \bibemph{him} with princes, \bibemph{even} with the princes of his people.
\vs Psa 113:9 He maketh the barren woman to keep house, \bibemph{and to be} a joyful mother of children. Praise ye the \textsc{Lord}.
\fn{to keep...: Heb. to dwell in an house.}
\vs Psa 114:1 When Israel went out of Egypt, the house of Jacob from a people of strange language;
\vs Psa 114:2 Judah was his sanctuary, \bibemph{and} Israel his dominion.
\vs Psa 114:3 The sea saw \bibemph{it}, and fled: Jordan was driven back.
\vs Psa 114:4 The mountains skipped like rams, \bibemph{and} the little hills like lambs.
\vs Psa 114:5 What \bibemph{ailed} thee, O thou sea, that thou fleddest? thou Jordan, \bibemph{that} thou wast driven back?
\vs Psa 114:6 Ye mountains, \bibemph{that} ye skipped like rams; \bibemph{and} ye little hills, like lambs?
\vs Psa 114:7 Tremble, thou earth, at the presence of the Lord, at the presence of the God of Jacob;
\vs Psa 114:8 Which turned the rock \bibemph{into} a standing water, the flint into a fountain of waters.
\vs Psa 115:1 Not unto us, O \textsc{Lord}, not unto us, but unto thy name give glory, for thy mercy, \bibemph{and} for thy truth's sake.
\vs Psa 115:2 Wherefore should the heathen say, Where \bibemph{is} now their God?
\vs Psa 115:3 But our God \bibemph{is} in the heavens: he hath done whatsoever he hath pleased.
\vs Psa 115:4 Their idols \bibemph{are} silver and gold, the work of men's hands.
\vs Psa 115:5 They have mouths, but they speak not: eyes have they, but they see not:
\vs Psa 115:6 They have ears, but they hear not: noses have they, but they smell not:
\vs Psa 115:7 They have hands, but they handle not: feet have they, but they walk not: neither speak they through their throat.
\vs Psa 115:8 They that make them are like unto them; \bibemph{so is} every one that trusteth in them.
\vs Psa 115:9 O Israel, trust thou in the \textsc{Lord}: he \bibemph{is} their help and their shield.
\vs Psa 115:10 O house of Aaron, trust in the \textsc{Lord}: he \bibemph{is} their help and their shield.
\vs Psa 115:11 Ye that fear the \textsc{Lord}, trust in the \textsc{Lord}: he \bibemph{is} their help and their shield.
\vs Psa 115:12 The \textsc{Lord} hath been mindful of us: he will bless \bibemph{us}; he will bless the house of Israel; he will bless the house of Aaron.
\vs Psa 115:13 He will bless them that fear the \textsc{Lord}, \bibemph{both} small and great.
\fn{and: Heb. with.}
\vs Psa 115:14 The \textsc{Lord} shall increase you more and more, you and your children.
\vs Psa 115:15 Ye \bibemph{are} blessed of the \textsc{Lord} which made heaven and earth.
\vs Psa 115:16 The heaven, \bibemph{even} the heavens, \bibemph{are} the \textsc{Lord's}: but the earth hath he given to the children of men.
\vs Psa 115:17 The dead praise not the \textsc{Lord}, neither any that go down into silence.
\vs Psa 115:18 But we will bless the \textsc{Lord} from this time forth and for evermore. Praise the \textsc{Lord}.
\vs Psa 116:1 I love the \textsc{Lord}, because he hath heard my voice \bibemph{and} my supplications.
\vs Psa 116:2 Because he hath inclined his ear unto me, therefore will I call upon \bibemph{him} as long as I live.
\fn{as long...: Heb. in my days.}
\vs Psa 116:3 The sorrows of death compassed me, and the pains of hell gat hold upon me: I found trouble and sorrow.
\fn{gat...: Heb. found me.}
\vs Psa 116:4 Then called I upon the name of the \textsc{Lord}; O \textsc{Lord}, I beseech thee, deliver my soul.
\vs Psa 116:5 Gracious \bibemph{is} the \textsc{Lord}, and righteous; yea, our God \bibemph{is} merciful.
\vs Psa 116:6 The \textsc{Lord} preserveth the simple: I was brought low, and he helped me.
\vs Psa 116:7 Return unto thy rest, O my soul; for the \textsc{Lord} hath dealt bountifully with thee.
\vs Psa 116:8 For thou hast delivered my soul from death, mine eyes from tears, \bibemph{and} my feet from falling.
\vs Psa 116:9 I will walk before the \textsc{Lord} in the land of the living.
\vs Psa 116:10 I believed, therefore have I spoken: I was greatly afflicted:
\vs Psa 116:11 I said in my haste, All men \bibemph{are} liars.
\vs Psa 116:12 What shall I render unto the \textsc{Lord} \bibemph{for} all his benefits toward me?
\vs Psa 116:13 I will take the cup of salvation, and call upon the name of the \textsc{Lord}.
\vs Psa 116:14 I will pay my vows unto the \textsc{Lord} now in the presence of all his people.
\vs Psa 116:15 Precious in the sight of the \textsc{Lord} \bibemph{is} the death of his saints.
\vs Psa 116:16 O \textsc{Lord}, truly I \bibemph{am} thy servant; I \bibemph{am} thy servant, \bibemph{and} the son of thine handmaid: thou hast loosed my bonds.
\vs Psa 116:17 I will offer to thee the sacrifice of thanksgiving, and will call upon the name of the \textsc{Lord}.
\vs Psa 116:18 I will pay my vows unto the \textsc{Lord} now in the presence of all his people,
\vs Psa 116:19 In the courts of the \textsc{Lord's} house, in the midst of thee, O Jerusalem. Praise ye the \textsc{Lord}.
\vs Psa 117:1 O praise the \textsc{Lord}, all ye nations: praise him, all ye people.
\vs Psa 117:2 For his merciful kindness is great toward us: and the truth of the \textsc{Lord} \bibemph{endureth} for ever. Praise ye the \textsc{Lord}.
\vs Psa 118:1 O give thanks unto the \textsc{Lord}; for \bibemph{he is} good: because his mercy \bibemph{endureth} for ever.
\vs Psa 118:2 Let Israel now say, that his mercy \bibemph{endureth} for ever.
\vs Psa 118:3 Let the house of Aaron now say, that his mercy \bibemph{endureth} for ever.
\vs Psa 118:4 Let them now that fear the \textsc{Lord} say, that his mercy \bibemph{endureth} for ever.
\vs Psa 118:5 I called upon the \textsc{Lord} in distress: the \textsc{Lord} answered me, \bibemph{and set me} in a large place.
\fn{in distress: Heb. out of distress.}
\vs Psa 118:6 The \textsc{Lord} \bibemph{is} on my side; I will not fear: what can man do unto me?
\fn{on...: Heb. for me.}
\vs Psa 118:7 The \textsc{Lord} taketh my part with them that help me: therefore shall I see \bibemph{my desire} upon them that hate me.
\vs Psa 118:8 \bibemph{It is} better to trust in the \textsc{Lord} than to put confidence in man.
\vs Psa 118:9 \bibemph{It is} better to trust in the \textsc{Lord} than to put confidence in princes.
\vs Psa 118:10 All nations compassed me about: but in the name of the \textsc{Lord} will I destroy them.
\fn{destroy...: Heb. cut them off.}
\vs Psa 118:11 They compassed me about; yea, they compassed me about: but in the name of the \textsc{Lord} I will destroy them.
\vs Psa 118:12 They compassed me about like bees; they are quenched as the fire of thorns: for in the name of the \textsc{Lord} I will destroy them.
\fn{destroy: Heb. cut down.}
\vs Psa 118:13 Thou hast thrust sore at me that I might fall: but the \textsc{Lord} helped me.
\vs Psa 118:14 The \textsc{Lord} \bibemph{is} my strength and song, and is become my salvation.
\vs Psa 118:15 The voice of rejoicing and salvation \bibemph{is} in the tabernacles of the righteous: the right hand of the \textsc{Lord} doeth valiantly.
\vs Psa 118:16 The right hand of the \textsc{Lord} is exalted: the right hand of the \textsc{Lord} doeth valiantly.
\vs Psa 118:17 I shall not die, but live, and declare the works of the \textsc{Lord}.
\vs Psa 118:18 The \textsc{Lord} hath chastened me sore: but he hath not given me over unto death.
\vs Psa 118:19 Open to me the gates of righteousness: I will go into them, \bibemph{and} I will praise the \textsc{Lord}:
\vs Psa 118:20 This gate of the \textsc{Lord}, into which the righteous shall enter.
\vs Psa 118:21 I will praise thee: for thou hast heard me, and art become my salvation.
\vs Psa 118:22 The stone \bibemph{which} the builders refused is become the head \bibemph{stone} of the corner.
\vs Psa 118:23 This is the \textsc{Lord's} doing; it \bibemph{is} marvellous in our eyes.
\fn{the \textsc{Lord's}...: Heb. from the \textsc{Lord}.}
\vs Psa 118:24 This \bibemph{is} the day \bibemph{which} the \textsc{Lord} hath made; we will rejoice and be glad in it.
\vs Psa 118:25 Save now, I beseech thee, O \textsc{Lord}: O \textsc{Lord}, I beseech thee, send now prosperity.
\vs Psa 118:26 Blessed \bibemph{be} he that cometh in the name of the \textsc{Lord}: we have blessed you out of the house of the \textsc{Lord}.
\vs Psa 118:27 God \bibemph{is} the \textsc{Lord}, which hath shewed us light: bind the sacrifice with cords, \bibemph{even} unto the horns of the altar.
\vs Psa 118:28 Thou \bibemph{art} my God, and I will praise thee: \bibemph{thou art} my God, I will exalt thee.
\vs Psa 118:29 O give thanks unto the \textsc{Lord}; for \bibemph{he is} good: for his mercy \bibemph{endureth} for ever.
\vs Psa 119:1 ALEPH. Blessed \bibemph{are} the undefiled in the way, who walk in the law of the \textsc{Lord}.
\fn{undefiled: or, perfect, or, sincere.}
\vs Psa 119:2 Blessed \bibemph{are} they that keep his testimonies, \bibemph{and that} seek him with the whole heart.
\vs Psa 119:3 They also do no iniquity: they walk in his ways.
\vs Psa 119:4 Thou hast commanded \bibemph{us} to keep thy precepts diligently.
\vs Psa 119:5 O that my ways were directed to keep thy statutes!
\vs Psa 119:6 Then shall I not be ashamed, when I have respect unto all thy commandments.
\vs Psa 119:7 I will praise thee with uprightness of heart, when I shall have learned thy righteous judgments.
\fn{thy...: Heb. judgments of thy righteousness.}
\vs Psa 119:8 I will keep thy statutes: O forsake me not utterly.
\vs Psa 119:9 BETH. Wherewithal shall a young man cleanse his way? by taking heed \bibemph{thereto} according to thy word.
\vs Psa 119:10 With my whole heart have I sought thee: O let me not wander from thy commandments.
\vs Psa 119:11 Thy word have I hid in mine heart, that I might not sin against thee.
\vs Psa 119:12 Blessed \bibemph{art} thou, O \textsc{Lord}: teach me thy statutes.
\vs Psa 119:13 With my lips have I declared all the judgments of thy mouth.
\vs Psa 119:14 I have rejoiced in the way of thy testimonies, as \bibemph{much as} in all riches.
\vs Psa 119:15 I will meditate in thy precepts, and have respect unto thy ways.
\vs Psa 119:16 I will delight myself in thy statutes: I will not forget thy word.
\vs Psa 119:17 GIMEL. Deal bountifully with thy servant, \bibemph{that} I may live, and keep thy word.
\vs Psa 119:18 Open thou mine eyes, that I may behold wondrous things out of thy law.
\fn{Open: Heb. Reveal.}
\vs Psa 119:19 I \bibemph{am} a stranger in the earth: hide not thy commandments from me.
\vs Psa 119:20 My soul breaketh for the longing \bibemph{that it hath} unto thy judgments at all times.
\vs Psa 119:21 Thou hast rebuked the proud \bibemph{that are} cursed, which do err from thy commandments.
\vs Psa 119:22 Remove from me reproach and contempt; for I have kept thy testimonies.
\vs Psa 119:23 Princes also did sit \bibemph{and} speak against me: \bibemph{but} thy servant did meditate in thy statutes.
\vs Psa 119:24 Thy testimonies also \bibemph{are} my delight \bibemph{and} my counsellors.
\fn{my counsellors: Heb. men of my counsel.}
\vs Psa 119:25 DALETH. My soul cleaveth unto the dust: quicken thou me according to thy word.
\vs Psa 119:26 I have declared my ways, and thou heardest me: teach me thy statutes.
\vs Psa 119:27 Make me to understand the way of thy precepts: so shall I talk of thy wondrous works.
\vs Psa 119:28 My soul melteth for heaviness: strengthen thou me according unto thy word.
\fn{melteth: Heb. droppeth.}
\vs Psa 119:29 Remove from me the way of lying: and grant me thy law graciously.
\vs Psa 119:30 I have chosen the way of truth: thy judgments have I laid \bibemph{before me}.
\vs Psa 119:31 I have stuck unto thy testimonies: O \textsc{Lord}, put me not to shame.
\vs Psa 119:32 I will run the way of thy commandments, when thou shalt enlarge my heart.
\vs Psa 119:33 HE. Teach me, O \textsc{Lord}, the way of thy statutes; and I shall keep it \bibemph{unto} the end.
\vs Psa 119:34 Give me understanding, and I shall keep thy law; yea, I shall observe it with \bibemph{my} whole heart.
\vs Psa 119:35 Make me to go in the path of thy commandments; for therein do I delight.
\vs Psa 119:36 Incline my heart unto thy testimonies, and not to covetousness.
\vs Psa 119:37 Turn away mine eyes from beholding vanity; \bibemph{and} quicken thou me in thy way.
\fn{Turn...: Heb. Make to pass.}
\vs Psa 119:38 Stablish thy word unto thy servant, who \bibemph{is devoted} to thy fear.
\vs Psa 119:39 Turn away my reproach which I fear: for thy judgments \bibemph{are} good.
\vs Psa 119:40 Behold, I have longed after thy precepts: quicken me in thy righteousness.
\vs Psa 119:41 VAU. Let thy mercies come also unto me, O \textsc{Lord}, \bibemph{even} thy salvation, according to thy word.
\vs Psa 119:42 So shall I have wherewith to answer him that reproacheth me: for I trust in thy word.
\fn{So...: or, So shall I answer him that reproveth me in a thing.}
\vs Psa 119:43 And take not the word of truth utterly out of my mouth; for I have hoped in thy judgments.
\vs Psa 119:44 So shall I keep thy law continually for ever and ever.
\vs Psa 119:45 And I will walk at liberty: for I seek thy precepts.
\fn{at liberty: Heb. at large.}
\vs Psa 119:46 I will speak of thy testimonies also before kings, and will not be ashamed.
\vs Psa 119:47 And I will delight myself in thy commandments, which I have loved.
\vs Psa 119:48 My hands also will I lift up unto thy commandments, which I have loved; and I will meditate in thy statutes.
\vs Psa 119:49 ZAIN. Remember the word unto thy servant, upon which thou hast caused me to hope.
\vs Psa 119:50 This \bibemph{is} my comfort in my affliction: for thy word hath quickened me.
\vs Psa 119:51 The proud have had me greatly in derision: \bibemph{yet} have I not declined from thy law.
\vs Psa 119:52 I remembered thy judgments of old, O \textsc{Lord}; and have comforted myself.
\vs Psa 119:53 Horror hath taken hold upon me because of the wicked that forsake thy law.
\vs Psa 119:54 Thy statutes have been my songs in the house of my pilgrimage.
\vs Psa 119:55 I have remembered thy name, O \textsc{Lord}, in the night, and have kept thy law.
\vs Psa 119:56 This I had, because I kept thy precepts.
\vs Psa 119:57 CHETH. \bibemph{Thou art} my portion, O \textsc{Lord}: I have said that I would keep thy words.
\vs Psa 119:58 I intreated thy favour with \bibemph{my} whole heart: be merciful unto me according to thy word.
\fn{favour: Heb. face.}
\vs Psa 119:59 I thought on my ways, and turned my feet unto thy testimonies.
\vs Psa 119:60 I made haste, and delayed not to keep thy commandments.
\vs Psa 119:61 The bands of the wicked have robbed me: \bibemph{but} I have not forgotten thy law.
\fn{bands: or, companies.}
\vs Psa 119:62 At midnight I will rise to give thanks unto thee because of thy righteous judgments.
\vs Psa 119:63 I \bibemph{am} a companion of all \bibemph{them} that fear thee, and of them that keep thy precepts.
\vs Psa 119:64 The earth, O \textsc{Lord}, is full of thy mercy: teach me thy statutes.
\vs Psa 119:65 TETH. Thou hast dealt well with thy servant, O \textsc{Lord}, according unto thy word.
\vs Psa 119:66 Teach me good judgment and knowledge: for I have believed thy commandments.
\vs Psa 119:67 Before I was afflicted I went astray: but now have I kept thy word.
\vs Psa 119:68 Thou \bibemph{art} good, and doest good; teach me thy statutes.
\vs Psa 119:69 The proud have forged a lie against me: \bibemph{but} I will keep thy precepts with \bibemph{my} whole heart.
\vs Psa 119:70 Their heart is as fat as grease; \bibemph{but} I delight in thy law.
\vs Psa 119:71 \bibemph{It is} good for me that I have been afflicted; that I might learn thy statutes.
\vs Psa 119:72 The law of thy mouth \bibemph{is} better unto me than thousands of gold and silver.
\vs Psa 119:73 JOD. Thy hands have made me and fashioned me: give me understanding, that I may learn thy commandments.
\vs Psa 119:74 They that fear thee will be glad when they see me; because I have hoped in thy word.
\vs Psa 119:75 I know, O \textsc{Lord}, that thy judgments \bibemph{are} right, and \bibemph{that} thou in faithfulness hast afflicted me.
\fn{right: Heb. righteousness.}
\vs Psa 119:76 Let, I pray thee, thy merciful kindness be for my comfort, according to thy word unto thy servant.
\fn{for...: Heb. to comfort me.}
\vs Psa 119:77 Let thy tender mercies come unto me, that I may live: for thy law \bibemph{is} my delight.
\vs Psa 119:78 Let the proud be ashamed; for they dealt perversely with me without a cause: \bibemph{but} I will meditate in thy precepts.
\vs Psa 119:79 Let those that fear thee turn unto me, and those that have known thy testimonies.
\vs Psa 119:80 Let my heart be sound in thy statutes; that I be not ashamed.
\vs Psa 119:81 CAPH. My soul fainteth for thy salvation: \bibemph{but} I hope in thy word.
\vs Psa 119:82 Mine eyes fail for thy word, saying, When wilt thou comfort me?
\vs Psa 119:83 For I am become like a bottle in the smoke; \bibemph{yet} do I not forget thy statutes.
\vs Psa 119:84 How many \bibemph{are} the days of thy servant? when wilt thou execute judgment on them that persecute me?
\vs Psa 119:85 The proud have digged pits for me, which \bibemph{are} not after thy law.
\vs Psa 119:86 All thy commandments \bibemph{are} faithful: they persecute me wrongfully; help thou me.
\fn{faithful: Heb. faithfulness.}
\vs Psa 119:87 They had almost consumed me upon earth; but I forsook not thy precepts.
\vs Psa 119:88 Quicken me after thy lovingkindness; so shall I keep the testimony of thy mouth.
\vs Psa 119:89 LAMED. For ever, O \textsc{Lord}, thy word is settled in heaven.
\vs Psa 119:90 Thy faithfulness \bibemph{is} unto all generations: thou hast established the earth, and it abideth.
\fn{unto...: Heb. to generation and generation. abideth: Heb. standeth.}
\vs Psa 119:91 They continue this day according to thine ordinances: for all \bibemph{are} thy servants.
\vs Psa 119:92 Unless thy law \bibemph{had been} my delights, I should then have perished in mine affliction.
\vs Psa 119:93 I will never forget thy precepts: for with them thou hast quickened me.
\vs Psa 119:94 I \bibemph{am} thine, save me; for I have sought thy precepts.
\vs Psa 119:95 The wicked have waited for me to destroy me: \bibemph{but} I will consider thy testimonies.
\vs Psa 119:96 I have seen an end of all perfection: \bibemph{but} thy commandment \bibemph{is} exceeding broad.
\vs Psa 119:97 MEM. O how love I thy law! it \bibemph{is} my meditation all the day.
\vs Psa 119:98 Thou through thy commandments hast made me wiser than mine enemies: for they \bibemph{are} ever with me.
\fn{they...: Heb. it is ever with me.}
\vs Psa 119:99 I have more understanding than all my teachers: for thy testimonies \bibemph{are} my meditation.
\vs Psa 119:100 I understand more than the ancients, because I keep thy precepts.
\vs Psa 119:101 I have refrained my feet from every evil way, that I might keep thy word.
\vs Psa 119:102 I have not departed from thy judgments: for thou hast taught me.
\vs Psa 119:103 How sweet are thy words unto my taste! \bibemph{yea, sweeter} than honey to my mouth!
\fn{taste: Heb. palate.}
\vs Psa 119:104 Through thy precepts I get understanding: therefore I hate every false way.
\vs Psa 119:105 NUN. Thy word \bibemph{is} a lamp unto my feet, and a light unto my path.
\fn{lamp: or, candle.}
\vs Psa 119:106 I have sworn, and I will perform \bibemph{it}, that I will keep thy righteous judgments.
\vs Psa 119:107 I am afflicted very much: quicken me, O \textsc{Lord}, according unto thy word.
\vs Psa 119:108 Accept, I beseech thee, the freewill offerings of my mouth, O \textsc{Lord}, and teach me thy judgments.
\vs Psa 119:109 My soul \bibemph{is} continually in my hand: yet do I not forget thy law.
\vs Psa 119:110 The wicked have laid a snare for me: yet I erred not from thy precepts.
\vs Psa 119:111 Thy testimonies have I taken as an heritage for ever: for they \bibemph{are} the rejoicing of my heart.
\vs Psa 119:112 I have inclined mine heart to perform thy statutes alway, \bibemph{even unto} the end.
\fn{to perform: Heb. to do.}
\vs Psa 119:113 SAMECH. I hate \bibemph{vain} thoughts: but thy law do I love.
\vs Psa 119:114 Thou \bibemph{art} my hiding place and my shield: I hope in thy word.
\vs Psa 119:115 Depart from me, ye evildoers: for I will keep the commandments of my God.
\vs Psa 119:116 Uphold me according unto thy word, that I may live: and let me not be ashamed of my hope.
\vs Psa 119:117 Hold thou me up, and I shall be safe: and I will have respect unto thy statutes continually.
\vs Psa 119:118 Thou hast trodden down all them that err from thy statutes: for their deceit \bibemph{is} falsehood.
\vs Psa 119:119 Thou puttest away all the wicked of the earth \bibemph{like} dross: therefore I love thy testimonies.
\fn{puttest...: Heb. causest to cease.}
\vs Psa 119:120 My flesh trembleth for fear of thee; and I am afraid of thy judgments.
\vs Psa 119:121 AIN. I have done judgment and justice: leave me not to mine oppressors.
\vs Psa 119:122 Be surety for thy servant for good: let not the proud oppress me.
\vs Psa 119:123 Mine eyes fail for thy salvation, and for the word of thy righteousness.
\vs Psa 119:124 Deal with thy servant according unto thy mercy, and teach me thy statutes.
\vs Psa 119:125 I \bibemph{am} thy servant; give me understanding, that I may know thy testimonies.
\vs Psa 119:126 \bibemph{It is} time for \bibemph{thee}, \textsc{Lord}, to work: \bibemph{for} they have made void thy law.
\vs Psa 119:127 Therefore I love thy commandments above gold; yea, above fine gold.
\vs Psa 119:128 Therefore I esteem all \bibemph{thy} precepts \bibemph{concerning} all \bibemph{things to be} right; \bibemph{and} I hate every false way.
\vs Psa 119:129 PE. Thy testimonies \bibemph{are} wonderful: therefore doth my soul keep them.
\vs Psa 119:130 The entrance of thy words giveth light; it giveth understanding unto the simple.
\vs Psa 119:131 I opened my mouth, and panted: for I longed for thy commandments.
\vs Psa 119:132 Look thou upon me, and be merciful unto me, as thou usest to do unto those that love thy name.
\fn{as thou...: Heb. according to the custom toward those, etc.}
\vs Psa 119:133 Order my steps in thy word: and let not any iniquity have dominion over me.
\vs Psa 119:134 Deliver me from the oppression of man: so will I keep thy precepts.
\vs Psa 119:135 Make thy face to shine upon thy servant; and teach me thy statutes.
\vs Psa 119:136 Rivers of waters run down mine eyes, because they keep not thy law.
\vs Psa 119:137 TZADDI. Righteous \bibemph{art} thou, O \textsc{Lord}, and upright \bibemph{are} thy judgments.
\vs Psa 119:138 Thy testimonies \bibemph{that} thou hast commanded \bibemph{are} righteous and very faithful.
\fn{righteous: Heb. righteousness. faithful: Heb. faithfulness.}
\vs Psa 119:139 My zeal hath consumed me, because mine enemies have forgotten thy words.
\fn{consumed...: Heb. cut me off.}
\vs Psa 119:140 Thy word \bibemph{is} very pure: therefore thy servant loveth it.
\fn{pure: Heb. tried, or, refined.}
\vs Psa 119:141 I \bibemph{am} small and despised: \bibemph{yet} do not I forget thy precepts.
\vs Psa 119:142 Thy righteousness \bibemph{is} an everlasting righteousness, and thy law \bibemph{is} the truth.
\vs Psa 119:143 Trouble and anguish have taken hold on me: \bibemph{yet} thy commandments \bibemph{are} my delights.
\fn{taken...: Heb. found me.}
\vs Psa 119:144 The righteousness of thy testimonies \bibemph{is} everlasting: give me understanding, and I shall live.
\vs Psa 119:145 KOPH. I cried with \bibemph{my} whole heart; hear me, O \textsc{Lord}: I will keep thy statutes.
\vs Psa 119:146 I cried unto thee; save me, and I shall keep thy testimonies.
\fn{and I...: or, that I may keep.}
\vs Psa 119:147 I prevented the dawning of the morning, and cried: I hoped in thy word.
\vs Psa 119:148 Mine eyes prevent the \bibemph{night} watches, that I might meditate in thy word.
\vs Psa 119:149 Hear my voice according unto thy lovingkindness: O \textsc{Lord}, quicken me according to thy judgment.
\vs Psa 119:150 They draw nigh that follow after mischief: they are far from thy law.
\vs Psa 119:151 Thou \bibemph{art} near, O \textsc{Lord}; and all thy commandments \bibemph{are} truth.
\vs Psa 119:152 Concerning thy testimonies, I have known of old that thou hast founded them for ever.
\vs Psa 119:153 RESH. Consider mine affliction, and deliver me: for I do not forget thy law.
\vs Psa 119:154 Plead my cause, and deliver me: quicken me according to thy word.
\vs Psa 119:155 Salvation \bibemph{is} far from the wicked: for they seek not thy statutes.
\vs Psa 119:156 Great \bibemph{are} thy tender mercies, O \textsc{Lord}: quicken me according to thy judgments.
\fn{Great: or, Many.}
\vs Psa 119:157 Many \bibemph{are} my persecutors and mine enemies; \bibemph{yet} do I not decline from thy testimonies.
\vs Psa 119:158 I beheld the transgressors, and was grieved; because they kept not thy word.
\vs Psa 119:159 Consider how I love thy precepts: quicken me, O \textsc{Lord}, according to thy lovingkindness.
\vs Psa 119:160 Thy word \bibemph{is} true \bibemph{from} the beginning: and every one of thy righteous judgments \bibemph{endureth} for ever.
\fn{Thy word...: Heb. The beginning of thy word is true.}
\vs Psa 119:161 SCHIN. Princes have persecuted me without a cause: but my heart standeth in awe of thy word.
\vs Psa 119:162 I rejoice at thy word, as one that findeth great spoil.
\vs Psa 119:163 I hate and abhor lying: \bibemph{but} thy law do I love.
\vs Psa 119:164 Seven times a day do I praise thee because of thy righteous judgments.
\vs Psa 119:165 Great peace have they which love thy law: and nothing shall offend them.
\fn{nothing...: Heb. they shall have no stumblingblock.}
\vs Psa 119:166 \textsc{Lord}, I have hoped for thy salvation, and done thy commandments.
\vs Psa 119:167 My soul hath kept thy testimonies; and I love them exceedingly.
\vs Psa 119:168 I have kept thy precepts and thy testimonies: for all my ways \bibemph{are} before thee.
\vs Psa 119:169 TAU. Let my cry come near before thee, O \textsc{Lord}: give me understanding according to thy word.
\vs Psa 119:170 Let my supplication come before thee: deliver me according to thy word.
\vs Psa 119:171 My lips shall utter praise, when thou hast taught me thy statutes.
\vs Psa 119:172 My tongue shall speak of thy word: for all thy commandments \bibemph{are} righteousness.
\vs Psa 119:173 Let thine hand help me; for I have chosen thy precepts.
\vs Psa 119:174 I have longed for thy salvation, O \textsc{Lord}; and thy law \bibemph{is} my delight.
\vs Psa 119:175 Let my soul live, and it shall praise thee; and let thy judgments help me.
\vs Psa 119:176 I have gone astray like a lost sheep; seek thy servant; for I do not forget thy commandments.
\vs Psa 120:1 A Song of degrees. In my distress I cried unto the \textsc{Lord}, and he heard me.
\vs Psa 120:2 Deliver my soul, O \textsc{Lord}, from lying lips, \bibemph{and} from a deceitful tongue.
\vs Psa 120:3 What shall be given unto thee? or what shall be done unto thee, thou false tongue?
\fn{What shall be given...: or, What shall the deceitful tongue give unto thee? or, what shall is profit thee?. done: Heb. added.}
\vs Psa 120:4 Sharp arrows of the mighty, with coals of juniper.
\fn{Sharp...: or, It is as the sharp arrows of the mighty man, with coals of juniper.}
\vs Psa 120:5 Woe is me, that I sojourn in Mesech, \bibemph{that} I dwell in the tents of Kedar!
\vs Psa 120:6 My soul hath long dwelt with him that hateth peace.
\vs Psa 120:7 I \bibemph{am for} peace: but when I speak, they \bibemph{are} for war.
\fn{for peace: or, a man of peace.}
\vs Psa 121:1 A Song of degrees. I will lift up mine eyes unto the hills, from whence cometh my help.
\fn{I will...: or, Shall I lift up mine eyes to the hills? whence should my help come?.}
\vs Psa 121:2 My help \bibemph{cometh} from the \textsc{Lord}, which made heaven and earth.
\vs Psa 121:3 He will not suffer thy foot to be moved: he that keepeth thee will not slumber.
\vs Psa 121:4 Behold, he that keepeth Israel shall neither slumber nor sleep.
\vs Psa 121:5 The \textsc{Lord} \bibemph{is} thy keeper: the \textsc{Lord} \bibemph{is} thy shade upon thy right hand.
\vs Psa 121:6 The sun shall not smite thee by day, nor the moon by night.
\vs Psa 121:7 The \textsc{Lord} shall preserve thee from all evil: he shall preserve thy soul.
\vs Psa 121:8 The \textsc{Lord} shall preserve thy going out and thy coming in from this time forth, and even for evermore.
\vs Psa 122:1 A Song of degrees of David. I was glad when they said unto me, Let us go into the house of the \textsc{Lord}.
\vs Psa 122:2 Our feet shall stand within thy gates, O Jerusalem.
\vs Psa 122:3 Jerusalem is builded as a city that is compact together:
\vs Psa 122:4 Whither the tribes go up, the tribes of the \textsc{Lord}, unto the testimony of Israel, to give thanks unto the name of the \textsc{Lord}.
\vs Psa 122:5 For there are set thrones of judgment, the thrones of the house of David.
\fn{are...: Heb. do sit.}
\vs Psa 122:6 Pray for the peace of Jerusalem: they shall prosper that love thee.
\vs Psa 122:7 Peace be within thy walls, \bibemph{and} prosperity within thy palaces.
\vs Psa 122:8 For my brethren and companions' sakes, I will now say, Peace \bibemph{be} within thee.
\vs Psa 122:9 Because of the house of the \textsc{Lord} our God I will seek thy good.
\vs Psa 123:1 A Song of degrees. Unto thee lift I up mine eyes, O thou that dwellest in the heavens.
\vs Psa 123:2 Behold, as the eyes of servants \bibemph{look} unto the hand of their masters, \bibemph{and} as the eyes of a maiden unto the hand of her mistress; so our eyes \bibemph{wait} upon the \textsc{Lord} our God, until that he have mercy upon us.
\vs Psa 123:3 Have mercy upon us, O \textsc{Lord}, have mercy upon us: for we are exceedingly filled with contempt.
\vs Psa 123:4 Our soul is exceedingly filled with the scorning of those that are at ease, \bibemph{and} with the contempt of the proud.
\vs Psa 124:1 A Song of degrees of David. If \bibemph{it had not been} the \textsc{Lord} who was on our side, now may Israel say;
\vs Psa 124:2 If \bibemph{it had not been} the \textsc{Lord} who was on our side, when men rose up against us:
\vs Psa 124:3 Then they had swallowed us up quick, when their wrath was kindled against us:
\vs Psa 124:4 Then the waters had overwhelmed us, the stream had gone over our soul:
\vs Psa 124:5 Then the proud waters had gone over our soul.
\vs Psa 124:6 Blessed \bibemph{be} the \textsc{Lord}, who hath not given us \bibemph{as} a prey to their teeth.
\vs Psa 124:7 Our soul is escaped as a bird out of the snare of the fowlers: the snare is broken, and we are escaped.
\vs Psa 124:8 Our help \bibemph{is} in the name of the \textsc{Lord}, who made heaven and earth.
\vs Psa 125:1 A Song of degrees. They that trust in the \textsc{Lord} \bibemph{shall be} as mount Zion, \bibemph{which} cannot be removed, \bibemph{but} abideth for ever.
\vs Psa 125:2 As the mountains \bibemph{are} round about Jerusalem, so the \textsc{Lord} \bibemph{is} round about his people from henceforth even for ever.
\vs Psa 125:3 For the rod of the wicked shall not rest upon the lot of the righteous; lest the righteous put forth their hands unto iniquity.
\fn{the wicked: Heb. wickedness.}
\vs Psa 125:4 Do good, O \textsc{Lord}, unto \bibemph{those that be} good, and \bibemph{to them that are} upright in their hearts.
\vs Psa 125:5 As for such as turn aside unto their crooked ways, the \textsc{Lord} shall lead them forth with the workers of iniquity: \bibemph{but} peace \bibemph{shall be} upon Israel.
\vs Psa 126:1 A Song of degrees. When the \textsc{Lord} turned again the captivity of Zion, we were like them that dream.
\fn{turned...: Heb. returned the returning.}
\vs Psa 126:2 Then was our mouth filled with laughter, and our tongue with singing: then said they among the heathen, The \textsc{Lord} hath done great things for them.
\fn{hath...: Heb. hath magnified to do with them.}
\vs Psa 126:3 The \textsc{Lord} hath done great things for us; \bibemph{whereof} we are glad.
\vs Psa 126:4 Turn again our captivity, O \textsc{Lord}, as the streams in the south.
\vs Psa 126:5 They that sow in tears shall reap in joy.
\fn{joy: or, singing.}
\vs Psa 126:6 He that goeth forth and weepeth, bearing precious seed, shall doubtless come again with rejoicing, bringing his sheaves \bibemph{with him}.
\fn{precious...: or, seed basket.}
\vs Psa 127:1 A Song of degrees for Solomon. Except the \textsc{Lord} build the house, they labour in vain that build it: except the \textsc{Lord} keep the city, the watchman waketh \bibemph{but} in vain.
\fn{for...: or, of Solomon. that...: Heb. that are builders of it in it.}
\vs Psa 127:2 \bibemph{It is} vain for you to rise up early, to sit up late, to eat the bread of sorrows: \bibemph{for} so he giveth his beloved sleep.
\vs Psa 127:3 Lo, children \bibemph{are} an heritage of the \textsc{Lord}: \bibemph{and} the fruit of the womb \bibemph{is his} reward.
\vs Psa 127:4 As arrows \bibemph{are} in the hand of a mighty man; so \bibemph{are} children of the youth.
\vs Psa 127:5 Happy \bibemph{is} the man that hath his quiver full of them: they shall not be ashamed, but they shall speak with the enemies in the gate.
\fn{his...: Heb. filled his quiver with them. speak...: or, subdue, or, destroy.}
\vs Psa 128:1 A Song of degrees. Blessed \bibemph{is} every one that feareth the \textsc{Lord}; that walketh in his ways.
\vs Psa 128:2 For thou shalt eat the labour of thine hands: happy \bibemph{shalt} thou \bibemph{be}, and \bibemph{it shall be} well with thee.
\vs Psa 128:3 Thy wife \bibemph{shall be} as a fruitful vine by the sides of thine house: thy children like olive plants round about thy table.
\vs Psa 128:4 Behold, that thus shall the man be blessed that feareth the \textsc{Lord}.
\vs Psa 128:5 The \textsc{Lord} shall bless thee out of Zion: and thou shalt see the good of Jerusalem all the days of thy life.
\vs Psa 128:6 Yea, thou shalt see thy children's children, \bibemph{and} peace upon Israel.
\vs Psa 129:1 A Song of degrees. Many a time have they afflicted me from my youth, may Israel now say:
\fn{Many...: or, Much.}
\vs Psa 129:2 Many a time have they afflicted me from my youth: yet they have not prevailed against me.
\fn{Many...: or, Much.}
\vs Psa 129:3 The plowers plowed upon my back: they made long their furrows.
\vs Psa 129:4 The \textsc{Lord} \bibemph{is} righteous: he hath cut asunder the cords of the wicked.
\vs Psa 129:5 Let them all be confounded and turned back that hate Zion.
\vs Psa 129:6 Let them be as the grass \bibemph{upon} the housetops, which withereth afore it groweth up:
\vs Psa 129:7 Wherewith the mower filleth not his hand; nor he that bindeth sheaves his bosom.
\vs Psa 129:8 Neither do they which go by say, The blessing of the \textsc{Lord} \bibemph{be} upon you: we bless you in the name of the \textsc{Lord}.
\vs Psa 130:1 A Song of degrees. Out of the depths have I cried unto thee, O \textsc{Lord}.
\vs Psa 130:2 Lord, hear my voice: let thine ears be attentive to the voice of my supplications.
\vs Psa 130:3 If thou, \textsc{Lord}, shouldest mark iniquities, O Lord, who shall stand?
\vs Psa 130:4 But \bibemph{there is} forgiveness with thee, that thou mayest be feared.
\vs Psa 130:5 I wait for the \textsc{Lord}, my soul doth wait, and in his word do I hope.
\vs Psa 130:6 My soul \bibemph{waiteth} for the Lord more than they that watch for the morning: \bibemph{I say, more than} they that watch for the morning.
\fn{I say...: or, which watch unto.}
\vs Psa 130:7 Let Israel hope in the \textsc{Lord}: for with the \textsc{Lord} \bibemph{there is} mercy, and with him \bibemph{is} plenteous redemption.
\vs Psa 130:8 And he shall redeem Israel from all his iniquities.
\vs Psa 131:1 A Song of degrees of David. \textsc{Lord}, my heart is not haughty, nor mine eyes lofty: neither do I exercise myself in great matters, or in things too high for me.
\fn{exercise...: Heb. walk. high: Heb. wonderful.}
\vs Psa 131:2 Surely I have behaved and quieted myself, as a child that is weaned of his mother: my soul \bibemph{is} even as a weaned child.
\fn{myself: Heb. my soul.}
\vs Psa 131:3 Let Israel hope in the \textsc{Lord} from henceforth and for ever.
\fn{henceforth: Heb. now.}
\vs Psa 132:1 A Song of degrees. \textsc{Lord}, remember David, \bibemph{and} all his afflictions:
\vs Psa 132:2 How he sware unto the \textsc{Lord}, \bibemph{and} vowed unto the mighty \bibemph{God} of Jacob;
\vs Psa 132:3 Surely I will not come into the tabernacle of my house, nor go up into my bed;
\vs Psa 132:4 I will not give sleep to mine eyes, \bibemph{or} slumber to mine eyelids,
\vs Psa 132:5 Until I find out a place for the \textsc{Lord}, an habitation for the mighty \bibemph{God} of Jacob.
\fn{an habitation: Heb. habitations.}
\vs Psa 132:6 Lo, we heard of it at Ephratah: we found it in the fields of the wood.
\vs Psa 132:7 We will go into his tabernacles: we will worship at his footstool.
\vs Psa 132:8 Arise, O \textsc{Lord}, into thy rest; thou, and the ark of thy strength.
\vs Psa 132:9 Let thy priests be clothed with righteousness; and let thy saints shout for joy.
\vs Psa 132:10 For thy servant David's sake turn not away the face of thine anointed.
\vs Psa 132:11 The \textsc{Lord} hath sworn \bibemph{in} truth unto David; he will not turn from it; Of the fruit of thy body will I set upon thy throne.
\fn{body: Heb. belly.}
\vs Psa 132:12 If thy children will keep my covenant and my testimony that I shall teach them, their children shall also sit upon thy throne for evermore.
\vs Psa 132:13 For the \textsc{Lord} hath chosen Zion; he hath desired \bibemph{it} for his habitation.
\vs Psa 132:14 This \bibemph{is} my rest for ever: here will I dwell; for I have desired it.
\vs Psa 132:15 I will abundantly bless her provision: I will satisfy her poor with bread.
\fn{abundantly: or, surely.}
\vs Psa 132:16 I will also clothe her priests with salvation: and her saints shall shout aloud for joy.
\vs Psa 132:17 There will I make the horn of David to bud: I have ordained a lamp for mine anointed.
\fn{lamp: or, candle.}
\vs Psa 132:18 His enemies will I clothe with shame: but upon himself shall his crown flourish.
\vs Psa 133:1 A Song of degrees of David. Behold, how good and how pleasant \bibemph{it is} for brethren to dwell together in unity!
\fn{together...: Heb. even together.}
\vs Psa 133:2 \bibemph{It is} like the precious ointment upon the head, that ran down upon the beard, \bibemph{even} Aaron's beard: that went down to the skirts of his garments;
\vs Psa 133:3 As the dew of Hermon, \bibemph{and as the dew} that descended upon the mountains of Zion: for there the \textsc{Lord} commanded the blessing, \bibemph{even} life for evermore.
\vs Psa 134:1 A Song of degrees. Behold, bless ye the \textsc{Lord}, all \bibemph{ye} servants of the \textsc{Lord}, which by night stand in the house of the \textsc{Lord}.
\vs Psa 134:2 Lift up your hands \bibemph{in} the sanctuary, and bless the \textsc{Lord}.
\fn{the sanctuary: or, holiness.}
\vs Psa 134:3 The \textsc{Lord} that made heaven and earth bless thee out of Zion.
\vs Psa 135:1 Praise ye the \textsc{Lord}. Praise ye the name of the \textsc{Lord}; praise \bibemph{him}, O ye servants of the \textsc{Lord}.
\vs Psa 135:2 Ye that stand in the house of the \textsc{Lord}, in the courts of the house of our God,
\vs Psa 135:3 Praise the \textsc{Lord}; for the \textsc{Lord} \bibemph{is} good: sing praises unto his name; for \bibemph{it is} pleasant.
\vs Psa 135:4 For the \textsc{Lord} hath chosen Jacob unto himself, \bibemph{and} Israel for his peculiar treasure.
\vs Psa 135:5 For I know that the \textsc{Lord} \bibemph{is} great, and \bibemph{that} our Lord \bibemph{is} above all gods.
\vs Psa 135:6 Whatsoever the \textsc{Lord} pleased, \bibemph{that} did he in heaven, and in earth, in the seas, and all deep places.
\vs Psa 135:7 He causeth the vapours to ascend from the ends of the earth; he maketh lightnings for the rain; he bringeth the wind out of his treasuries.
\vs Psa 135:8 Who smote the firstborn of Egypt, both of man and beast.
\fn{both...: Heb. from man unto beast.}
\vs Psa 135:9 \bibemph{Who} sent tokens and wonders into the midst of thee, O Egypt, upon Pharaoh, and upon all his servants.
\vs Psa 135:10 Who smote great nations, and slew mighty kings;
\vs Psa 135:11 Sihon king of the Amorites, and Og king of Bashan, and all the kingdoms of Canaan:
\vs Psa 135:12 And gave their land \bibemph{for} an heritage, an heritage unto Israel his people.
\vs Psa 135:13 Thy name, O \textsc{Lord}, \bibemph{endureth} for ever; \bibemph{and} thy memorial, O \textsc{Lord}, throughout all generations.
\fn{throughout...: Heb. to generation and generation.}
\vs Psa 135:14 For the \textsc{Lord} will judge his people, and he will repent himself concerning his servants.
\vs Psa 135:15 The idols of the heathen \bibemph{are} silver and gold, the work of men's hands.
\vs Psa 135:16 They have mouths, but they speak not; eyes have they, but they see not;
\vs Psa 135:17 They have ears, but they hear not; neither is there \bibemph{any} breath in their mouths.
\vs Psa 135:18 They that make them are like unto them: \bibemph{so is} every one that trusteth in them.
\vs Psa 135:19 Bless the \textsc{Lord}, O house of Israel: bless the \textsc{Lord}, O house of Aaron:
\vs Psa 135:20 Bless the \textsc{Lord}, O house of Levi: ye that fear the \textsc{Lord}, bless the \textsc{Lord}.
\vs Psa 135:21 Blessed be the \textsc{Lord} out of Zion, which dwelleth at Jerusalem. Praise ye the \textsc{Lord}.
\vs Psa 136:1 O give thanks unto the \textsc{Lord}; for \bibemph{he is} good: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:2 O give thanks unto the God of gods: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:3 O give thanks to the Lord of lords: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:4 To him who alone doeth great wonders: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:5 To him that by wisdom made the heavens: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:6 To him that stretched out the earth above the waters: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:7 To him that made great lights: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:8 The sun to rule by day: for his mercy \bibemph{endureth} for ever:
\fn{to rule...: Heb. for the rulings by day.}
\vs Psa 136:9 The moon and stars to rule by night: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:10 To him that smote Egypt in their firstborn: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:11 And brought out Israel from among them: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:12 With a strong hand, and with a stretched out arm: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:13 To him which divided the Red sea into parts: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:14 And made Israel to pass through the midst of it: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:15 But overthrew Pharaoh and his host in the Red sea: for his mercy \bibemph{endureth} for ever.
\fn{overthrew: Heb. shaked off.}
\vs Psa 136:16 To him which led his people through the wilderness: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:17 To him which smote great kings: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:18 And slew famous kings: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:19 Sihon king of the Amorites: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:20 And Og the king of Bashan: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:21 And gave their land for an heritage: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:22 \bibemph{Even} an heritage unto Israel his servant: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:23 Who remembered us in our low estate: for his mercy \bibemph{endureth} for ever:
\vs Psa 136:24 And hath redeemed us from our enemies: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:25 Who giveth food to all flesh: for his mercy \bibemph{endureth} for ever.
\vs Psa 136:26 O give thanks unto the God of heaven: for his mercy \bibemph{endureth} for ever.
\vs Psa 137:1 By the rivers of Babylon, there we sat down, yea, we wept, when we remembered Zion.
\vs Psa 137:2 We hanged our harps upon the willows in the midst thereof.
\vs Psa 137:3 For there they that carried us away captive required of us a song; and they that wasted us \bibemph{required of us} mirth, \bibemph{saying}, Sing us \bibemph{one} of the songs of Zion.
\fn{a song: Heb. the words of a song. wasted...: Heb. laid us on heaps.}
\vs Psa 137:4 How shall we sing the \textsc{Lord's} song in a strange land?
\fn{strange...: Heb. land of a stranger?.}
\vs Psa 137:5 If I forget thee, O Jerusalem, let my right hand forget \bibemph{her cunning}.
\vs Psa 137:6 If I do not remember thee, let my tongue cleave to the roof of my mouth; if I prefer not Jerusalem above my chief joy.
\fn{my chief...: Heb. the head of my joy.}
\vs Psa 137:7 Remember, O \textsc{Lord}, the children of Edom in the day of Jerusalem; who said, Rase \bibemph{it}, rase \bibemph{it, even} to the foundation thereof.
\fn{Rase it: Heb. Make bare.}
\vs Psa 137:8 O daughter of Babylon, who art to be destroyed; happy \bibemph{shall he be}, that rewardeth thee as thou hast served us.
\fn{destroyed: Heb. wasted. rewardeth...: Heb. recompenseth unto thee thy deed which thou didst to us.}
\vs Psa 137:9 Happy \bibemph{shall he be}, that taketh and dasheth thy little ones against the stones.
\fn{the stones: Heb. the rock.}
\vs Psa 138:1 \bibemph{A Psalm} of David. I will praise thee with my whole heart: before the gods will I sing praise unto thee.
\vs Psa 138:2 I will worship toward thy holy temple, and praise thy name for thy lovingkindness and for thy truth: for thou hast magnified thy word above all thy name.
\vs Psa 138:3 In the day when I cried thou answeredst me, \bibemph{and} strengthenedst me \bibemph{with} strength in my soul.
\vs Psa 138:4 All the kings of the earth shall praise thee, O \textsc{Lord}, when they hear the words of thy mouth.
\vs Psa 138:5 Yea, they shall sing in the ways of the \textsc{Lord}: for great \bibemph{is} the glory of the \textsc{Lord}.
\vs Psa 138:6 Though the \textsc{Lord} \bibemph{be} high, yet hath he respect unto the lowly: but the proud he knoweth afar off.
\vs Psa 138:7 Though I walk in the midst of trouble, thou wilt revive me: thou shalt stretch forth thine hand against the wrath of mine enemies, and thy right hand shall save me.
\vs Psa 138:8 The \textsc{Lord} will perfect \bibemph{that which} concerneth me: thy mercy, O \textsc{Lord}, \bibemph{endureth} for ever: forsake not the works of thine own hands.
\vs Psa 139:1 To the chief Musician, A Psalm of David. O \textsc{Lord}, thou hast searched me, and known \bibemph{me}.
\vs Psa 139:2 Thou knowest my downsitting and mine uprising, thou understandest my thought afar off.
\vs Psa 139:3 Thou compassest my path and my lying down, and art acquainted \bibemph{with} all my ways.
\fn{compassest: or, winnowest.}
\vs Psa 139:4 For \bibemph{there is} not a word in my tongue, \bibemph{but}, lo, O \textsc{Lord}, thou knowest it altogether.
\vs Psa 139:5 Thou hast beset me behind and before, and laid thine hand upon me.
\vs Psa 139:6 \bibemph{Such} knowledge \bibemph{is} too wonderful for me; it is high, I cannot \bibemph{attain} unto it.
\vs Psa 139:7 Whither shall I go from thy spirit? or whither shall I flee from thy presence?
\vs Psa 139:8 If I ascend up into heaven, thou \bibemph{art} there: if I make my bed in hell, behold, thou \bibemph{art there}.
\vs Psa 139:9 \bibemph{If} I take the wings of the morning, \bibemph{and} dwell in the uttermost parts of the sea;
\vs Psa 139:10 Even there shall thy hand lead me, and thy right hand shall hold me.
\vs Psa 139:11 If I say, Surely the darkness shall cover me; even the night shall be light about me.
\vs Psa 139:12 Yea, the darkness hideth not from thee; but the night shineth as the day: the darkness and the light \bibemph{are} both alike \bibemph{to thee}.
\fn{hideth...: Heb. darkeneth not. the darkness and...: Heb. as is the darkness, so is the light.}
\vs Psa 139:13 For thou hast possessed my reins: thou hast covered me in my mother's womb.
\vs Psa 139:14 I will praise thee; for I am fearfully \bibemph{and} wonderfully made: marvellous \bibemph{are} thy works; and \bibemph{that} my soul knoweth right well.
\fn{right...: Heb. greatly.}
\vs Psa 139:15 My substance was not hid from thee, when I was made in secret, \bibemph{and} curiously wrought in the lowest parts of the earth.
\fn{substance: or, strength, or, body.}
\vs Psa 139:16 Thine eyes did see my substance, yet being unperfect; and in thy book all \bibemph{my members} were written, \bibemph{which} in continuance were fashioned, when \bibemph{as yet there was} none of them.
\fn{all...: Heb. all of them. which...: or, what days they should be fashioned.}
\vs Psa 139:17 How precious also are thy thoughts unto me, O God! how great is the sum of them!
\vs Psa 139:18 \bibemph{If} I should count them, they are more in number than the sand: when I awake, I am still with thee.
\vs Psa 139:19 Surely thou wilt slay the wicked, O God: depart from me therefore, ye bloody men.
\vs Psa 139:20 For they speak against thee wickedly, \bibemph{and} thine enemies take \bibemph{thy name} in vain.
\vs Psa 139:21 Do not I hate them, O \textsc{Lord}, that hate thee? and am not I grieved with those that rise up against thee?
\vs Psa 139:22 I hate them with perfect hatred: I count them mine enemies.
\vs Psa 139:23 Search me, O God, and know my heart: try me, and know my thoughts:
\vs Psa 139:24 And see if \bibemph{there be any} wicked way in me, and lead me in the way everlasting.
\fn{wicked...: Heb. way of pain, or, grief.}
\vs Psa 140:1 To the chief Musician, A Psalm of David. Deliver me, O \textsc{Lord}, from the evil man: preserve me from the violent man;
\fn{violent...: Heb. man of violences.}
\vs Psa 140:2 Which imagine mischiefs in \bibemph{their} heart; continually are they gathered together \bibemph{for} war.
\vs Psa 140:3 They have sharpened their tongues like a serpent; adders' poison \bibemph{is} under their lips. Selah.
\vs Psa 140:4 Keep me, O \textsc{Lord}, from the hands of the wicked; preserve me from the violent man; who have purposed to overthrow my goings.
\vs Psa 140:5 The proud have hid a snare for me, and cords; they have spread a net by the wayside; they have set gins for me. Selah.
\vs Psa 140:6 I said unto the \textsc{Lord}, Thou \bibemph{art} my God: hear the voice of my supplications, O \textsc{Lord}.
\vs Psa 140:7 O GOD the Lord, the strength of my salvation, thou hast covered my head in the day of battle.
\vs Psa 140:8 Grant not, O \textsc{Lord}, the desires of the wicked: further not his wicked device; \bibemph{lest} they exalt themselves. Selah.
\fn{lest...: or, let them not be exalted.}
\vs Psa 140:9 \bibemph{As for} the head of those that compass me about, let the mischief of their own lips cover them.
\vs Psa 140:10 Let burning coals fall upon them: let them be cast into the fire; into deep pits, that they rise not up again.
\vs Psa 140:11 Let not an evil speaker be established in the earth: evil shall hunt the violent man to overthrow \bibemph{him}.
\fn{an...: or, an evil speaker (Heb. a man of tongue), a wicked man of violence, he established in the earth: let him be hunted to his overthrow.}
\vs Psa 140:12 I know that the \textsc{Lord} will maintain the cause of the afflicted, \bibemph{and} the right of the poor.
\vs Psa 140:13 Surely the righteous shall give thanks unto thy name: the upright shall dwell in thy presence.
\vs Psa 141:1 A Psalm of David. \textsc{Lord}, I cry unto thee: make haste unto me; give ear unto my voice, when I cry unto thee.
\vs Psa 141:2 Let my prayer be set forth before thee \bibemph{as} incense; \bibemph{and} the lifting up of my hands \bibemph{as} the evening sacrifice.
\fn{set...: Heb. directed.}
\vs Psa 141:3 Set a watch, O \textsc{Lord}, before my mouth; keep the door of my lips.
\vs Psa 141:4 Incline not my heart to \bibemph{any} evil thing, to practise wicked works with men that work iniquity: and let me not eat of their dainties.
\vs Psa 141:5 Let the righteous smite me; \bibemph{it shall be} a kindness: and let him reprove me; \bibemph{it shall be} an excellent oil, \bibemph{which} shall not break my head: for yet my prayer also \bibemph{shall be} in their calamities.
\fn{me: it shall be a...: or, me kindly, and reprove me; let not their precious oil break, etc.}
\vs Psa 141:6 When their judges are overthrown in stony places, they shall hear my words; for they are sweet.
\vs Psa 141:7 Our bones are scattered at the grave's mouth, as when one cutteth and cleaveth \bibemph{wood} upon the earth.
\vs Psa 141:8 But mine eyes \bibemph{are} unto thee, O GOD the Lord: in thee is my trust; leave not my soul destitute.
\fn{leave...: Heb. make not my soul bare.}
\vs Psa 141:9 Keep me from the snares \bibemph{which} they have laid for me, and the gins of the workers of iniquity.
\vs Psa 141:10 Let the wicked fall into their own nets, whilst that I withal escape.
\fn{escape: Heb. pass over.}
\vs Psa 142:1 Maschil of David; A Prayer when he was in the cave. I cried unto the \textsc{Lord} with my voice; with my voice unto the \textsc{Lord} did I make my supplication.
\fn{Maschil...: or, A Psalm of David, giving instruction.}
\vs Psa 142:2 I poured out my complaint before him; I shewed before him my trouble.
\vs Psa 142:3 When my spirit was overwhelmed within me, then thou knewest my path. In the way wherein I walked have they privily laid a snare for me.
\vs Psa 142:4 I looked on \bibemph{my} right hand, and beheld, but \bibemph{there was} no man that would know me: refuge failed me; no man cared for my soul.
\fn{I looked...: or, Look on the right hand, and see. failed...: Heb. perished from me. cared...: Heb. sought after.}
\vs Psa 142:5 I cried unto thee, O \textsc{Lord}: I said, Thou \bibemph{art} my refuge \bibemph{and} my portion in the land of the living.
\vs Psa 142:6 Attend unto my cry; for I am brought very low: deliver me from my persecutors; for they are stronger than I.
\vs Psa 142:7 Bring my soul out of prison, that I may praise thy name: the righteous shall compass me about; for thou shalt deal bountifully with me.
\vs Psa 143:1 A Psalm of David. Hear my prayer, O \textsc{Lord}, give ear to my supplications: in thy faithfulness answer me, \bibemph{and} in thy righteousness.
\vs Psa 143:2 And enter not into judgment with thy servant: for in thy sight shall no man living be justified.
\vs Psa 143:3 For the enemy hath persecuted my soul; he hath smitten my life down to the ground; he hath made me to dwell in darkness, as those that have been long dead.
\vs Psa 143:4 Therefore is my spirit overwhelmed within me; my heart within me is desolate.
\vs Psa 143:5 I remember the days of old; I meditate on all thy works; I muse on the work of thy hands.
\vs Psa 143:6 I stretch forth my hands unto thee: my soul \bibemph{thirsteth} after thee, as a thirsty land. Selah.
\vs Psa 143:7 Hear me speedily, O \textsc{Lord}: my spirit faileth: hide not thy face from me, lest I be like unto them that go down into the pit.
\fn{lest...: or, for I am become like, etc.}
\vs Psa 143:8 Cause me to hear thy lovingkindness in the morning; for in thee do I trust: cause me to know the way wherein I should walk; for I lift up my soul unto thee.
\vs Psa 143:9 Deliver me, O \textsc{Lord}, from mine enemies: I flee unto thee to hide me.
\fn{flee...: Heb. hide me with thee.}
\vs Psa 143:10 Teach me to do thy will; for thou \bibemph{art} my God: thy spirit \bibemph{is} good; lead me into the land of uprightness.
\vs Psa 143:11 Quicken me, O \textsc{Lord}, for thy name's sake: for thy righteousness' sake bring my soul out of trouble.
\vs Psa 143:12 And of thy mercy cut off mine enemies, and destroy all them that afflict my soul: for I \bibemph{am} thy servant.
\vs Psa 144:1 \bibemph{A Psalm} of David. Blessed \bibemph{be} the \textsc{Lord} my strength, which teacheth my hands to war, \bibemph{and} my fingers to fight:
\fn{strength: Heb. rock. to war...: Heb. to the war, etc.}
\vs Psa 144:2 My goodness, and my fortress; my high tower, and my deliverer; my shield, and \bibemph{he} in whom I trust; who subdueth my people under me.
\fn{My goodness: or, My mercy.}
\vs Psa 144:3 \textsc{Lord}, what \bibemph{is} man, that thou takest knowledge of him! \bibemph{or} the son of man, that thou makest account of him!
\vs Psa 144:4 Man is like to vanity: his days \bibemph{are} as a shadow that passeth away.
\vs Psa 144:5 Bow thy heavens, O \textsc{Lord}, and come down: touch the mountains, and they shall smoke.
\vs Psa 144:6 Cast forth lightning, and scatter them: shoot out thine arrows, and destroy them.
\vs Psa 144:7 Send thine hand from above; rid me, and deliver me out of great waters, from the hand of strange children;
\fn{hand from: Heb. hands from.}
\vs Psa 144:8 Whose mouth speaketh vanity, and their right hand \bibemph{is} a right hand of falsehood.
\vs Psa 144:9 I will sing a new song unto thee, O God: upon a psaltery \bibemph{and} an instrument of ten strings will I sing praises unto thee.
\vs Psa 144:10 \bibemph{It is he} that giveth salvation unto kings: who delivereth David his servant from the hurtful sword.
\fn{salvation: or, victory.}
\vs Psa 144:11 Rid me, and deliver me from the hand of strange children, whose mouth speaketh vanity, and their right hand \bibemph{is} a right hand of falsehood:
\vs Psa 144:12 That our sons \bibemph{may be} as plants grown up in their youth; \bibemph{that} our daughters \bibemph{may be} as corner stones, polished \bibemph{after} the similitude of a palace:
\fn{polished: Heb. cut.}
\vs Psa 144:13 \bibemph{That} our garners \bibemph{may be} full, affording all manner of store: \bibemph{that} our sheep may bring forth thousands and ten thousands in our streets:
\fn{all...: Heb. from kind to kind.}
\vs Psa 144:14 \bibemph{That} our oxen \bibemph{may be} strong to labour; \bibemph{that there be} no breaking in, nor going out; that \bibemph{there be} no complaining in our streets.
\fn{strong...: Heb. able to bear burdens, or, loaden with flesh.}
\vs Psa 144:15 Happy \bibemph{is that} people, that is in such a case: \bibemph{yea}, happy \bibemph{is that} people, whose God \bibemph{is} the \textsc{Lord}.
\vs Psa 145:1 David's \bibemph{Psalm} of praise. I will extol thee, my God, O king; and I will bless thy name for ever and ever.
\vs Psa 145:2 Every day will I bless thee; and I will praise thy name for ever and ever.
\vs Psa 145:3 Great \bibemph{is} the \textsc{Lord}, and greatly to be praised; and his greatness \bibemph{is} unsearchable.
\fn{and his...: Heb. and of his greatness there is no search.}
\vs Psa 145:4 One generation shall praise thy works to another, and shall declare thy mighty acts.
\vs Psa 145:5 I will speak of the glorious honour of thy majesty, and of thy wondrous works.
\fn{works: Heb. things, or, words.}
\vs Psa 145:6 And \bibemph{men} shall speak of the might of thy terrible acts: and I will declare thy greatness.
\fn{declare: Heb. declare it.}
\vs Psa 145:7 They shall abundantly utter the memory of thy great goodness, and shall sing of thy righteousness.
\vs Psa 145:8 The \textsc{Lord} \bibemph{is} gracious, and full of compassion; slow to anger, and of great mercy.
\fn{of great...: Heb. great in mercy.}
\vs Psa 145:9 The \textsc{Lord} \bibemph{is} good to all: and his tender mercies \bibemph{are} over all his works.
\vs Psa 145:10 All thy works shall praise thee, O \textsc{Lord}; and thy saints shall bless thee.
\vs Psa 145:11 They shall speak of the glory of thy kingdom, and talk of thy power;
\vs Psa 145:12 To make known to the sons of men his mighty acts, and the glorious majesty of his kingdom.
\vs Psa 145:13 Thy kingdom \bibemph{is} an everlasting kingdom, and thy dominion \bibemph{endureth} throughout all generations.
\fn{an...: Heb. a kingdom of all ages.}
\vs Psa 145:14 The \textsc{Lord} upholdeth all that fall, and raiseth up all \bibemph{those that be} bowed down.
\vs Psa 145:15 The eyes of all wait upon thee; and thou givest them their meat in due season.
\fn{wait...: or, look unto.}
\vs Psa 145:16 Thou openest thine hand, and satisfiest the desire of every living thing.
\vs Psa 145:17 The \textsc{Lord} \bibemph{is} righteous in all his ways, and holy in all his works.
\fn{holy: or, merciful, or, bountiful.}
\vs Psa 145:18 The \textsc{Lord} \bibemph{is} nigh unto all them that call upon him, to all that call upon him in truth.
\vs Psa 145:19 He will fulfil the desire of them that fear him: he also will hear their cry, and will save them.
\vs Psa 145:20 The \textsc{Lord} preserveth all them that love him: but all the wicked will he destroy.
\vs Psa 145:21 My mouth shall speak the praise of the \textsc{Lord}: and let all flesh bless his holy name for ever and ever.
\vs Psa 146:1 Praise ye the \textsc{Lord}. Praise the \textsc{Lord}, O my soul.
\fn{Praise ye...: Heb. Hallelujah.}
\vs Psa 146:2 While I live will I praise the \textsc{Lord}: I will sing praises unto my God while I have any being.
\vs Psa 146:3 Put not your trust in princes, \bibemph{nor} in the son of man, in whom \bibemph{there is} no help.
\fn{help: or, salvation.}
\vs Psa 146:4 His breath goeth forth, he returneth to his earth; in that very day his thoughts perish.
\vs Psa 146:5 Happy \bibemph{is he} that \bibemph{hath} the God of Jacob for his help, whose hope \bibemph{is} in the \textsc{Lord} his God:
\vs Psa 146:6 Which made heaven, and earth, the sea, and all that therein \bibemph{is}: which keepeth truth for ever:
\vs Psa 146:7 Which executeth judgment for the oppressed: which giveth food to the hungry. The \textsc{Lord} looseth the prisoners:
\vs Psa 146:8 The \textsc{Lord} openeth \bibemph{the eyes of} the blind: the \textsc{Lord} raiseth them that are bowed down: the \textsc{Lord} loveth the righteous:
\vs Psa 146:9 The \textsc{Lord} preserveth the strangers; he relieveth the fatherless and widow: but the way of the wicked he turneth upside down.
\vs Psa 146:10 The \textsc{Lord} shall reign for ever, \bibemph{even} thy God, O Zion, unto all generations. Praise ye the \textsc{Lord}.
\vs Psa 147:1 Praise ye the \textsc{Lord}: for \bibemph{it is} good to sing praises unto our God; for \bibemph{it is} pleasant; \bibemph{and} praise is comely.
\vs Psa 147:2 The \textsc{Lord} doth build up Jerusalem: he gathereth together the outcasts of Israel.
\vs Psa 147:3 He healeth the broken in heart, and bindeth up their wounds.
\fn{wounds: Heb. griefs.}
\vs Psa 147:4 He telleth the number of the stars; he calleth them all by \bibemph{their} names.
\vs Psa 147:5 Great \bibemph{is} our Lord, and of great power: his understanding \bibemph{is} infinite.
\fn{his...: Heb. of his understanding there is no number.}
\vs Psa 147:6 The \textsc{Lord} lifteth up the meek: he casteth the wicked down to the ground.
\vs Psa 147:7 Sing unto the \textsc{Lord} with thanksgiving; sing praise upon the harp unto our God:
\vs Psa 147:8 Who covereth the heaven with clouds, who prepareth rain for the earth, who maketh grass to grow upon the mountains.
\vs Psa 147:9 He giveth to the beast his food, \bibemph{and} to the young ravens which cry.
\vs Psa 147:10 He delighteth not in the strength of the horse: he taketh not pleasure in the legs of a man.
\vs Psa 147:11 The \textsc{Lord} taketh pleasure in them that fear him, in those that hope in his mercy.
\vs Psa 147:12 Praise the \textsc{Lord}, O Jerusalem; praise thy God, O Zion.
\vs Psa 147:13 For he hath strengthened the bars of thy gates; he hath blessed thy children within thee.
\vs Psa 147:14 He maketh peace \bibemph{in} thy borders, \bibemph{and} filleth thee with the finest of the wheat.
\fn{He...: Heb. Who maketh thy border peace. finest...: Heb. fat of wheat.}
\vs Psa 147:15 He sendeth forth his commandment \bibemph{upon} earth: his word runneth very swiftly.
\vs Psa 147:16 He giveth snow like wool: he scattereth the hoarfrost like ashes.
\vs Psa 147:17 He casteth forth his ice like morsels: who can stand before his cold?
\vs Psa 147:18 He sendeth out his word, and melteth them: he causeth his wind to blow, \bibemph{and} the waters flow.
\vs Psa 147:19 He sheweth his word unto Jacob, his statutes and his judgments unto Israel.
\fn{his word: Heb. his words.}
\vs Psa 147:20 He hath not dealt so with any nation: and \bibemph{as for his} judgments, they have not known them. Praise ye the \textsc{Lord}.
\vs Psa 148:1 Praise ye the \textsc{Lord}. Praise ye the \textsc{Lord} from the heavens: praise him in the heights.
\fn{Praise ye the \textsc{Lord}. Praise: Heb. Hallelujah, etc.}
\vs Psa 148:2 Praise ye him, all his angels: praise ye him, all his hosts.
\vs Psa 148:3 Praise ye him, sun and moon: praise him, all ye stars of light.
\vs Psa 148:4 Praise him, ye heavens of heavens, and ye waters that \bibemph{be} above the heavens.
\vs Psa 148:5 Let them praise the name of the \textsc{Lord}: for he commanded, and they were created.
\vs Psa 148:6 He hath also stablished them for ever and ever: he hath made a decree which shall not pass.
\vs Psa 148:7 Praise the \textsc{Lord} from the earth, ye dragons, and all deeps:
\vs Psa 148:8 Fire, and hail; snow, and vapour; stormy wind fulfilling his word:
\vs Psa 148:9 Mountains, and all hills; fruitful trees, and all cedars:
\vs Psa 148:10 Beasts, and all cattle; creeping things, and flying fowl:
\fn{flying...: Heb. birds of wing.}
\vs Psa 148:11 Kings of the earth, and all people; princes, and all judges of the earth:
\vs Psa 148:12 Both young men, and maidens; old men, and children:
\vs Psa 148:13 Let them praise the name of the \textsc{Lord}: for his name alone is excellent; his glory \bibemph{is} above the earth and heaven.
\fn{excellent: Heb. exalted.}
\vs Psa 148:14 He also exalteth the horn of his people, the praise of all his saints; \bibemph{even} of the children of Israel, a people near unto him. Praise ye the \textsc{Lord}.
\vs Psa 149:1 Praise ye the \textsc{Lord}. Sing unto the \textsc{Lord} a new song, \bibemph{and} his praise in the congregation of saints.
\fn{Praise...: Heb. Hallelujah.}
\vs Psa 149:2 Let Israel rejoice in him that made him: let the children of Zion be joyful in their King.
\vs Psa 149:3 Let them praise his name in the dance: let them sing praises unto him with the timbrel and harp.
\fn{in...: or, with the pipe.}
\vs Psa 149:4 For the \textsc{Lord} taketh pleasure in his people: he will beautify the meek with salvation.
\vs Psa 149:5 Let the saints be joyful in glory: let them sing aloud upon their beds.
\vs Psa 149:6 \bibemph{Let} the high \bibemph{praises} of God \bibemph{be} in their mouth, and a twoedged sword in their hand;
\fn{mouth: Heb. throat.}
\vs Psa 149:7 To execute vengeance upon the heathen, \bibemph{and} punishments upon the people;
\vs Psa 149:8 To bind their kings with chains, and their nobles with fetters of iron;
\vs Psa 149:9 To execute upon them the judgment written: this honour have all his saints. Praise ye the \textsc{Lord}.
\vs Psa 150:1 Praise ye the \textsc{Lord}. Praise God in his sanctuary: praise him in the firmament of his power.
\fn{Praise ye...: Heb. Hallelujah.}
\vs Psa 150:2 Praise him for his mighty acts: praise him according to his excellent greatness.
\vs Psa 150:3 Praise him with the sound of the trumpet: praise him with the psaltery and harp.
\fn{trumpet: or, cornet.}
\vs Psa 150:4 Praise him with the timbrel and dance: praise him with stringed instruments and organs.
\fn{dance: or, pipe.}
\vs Psa 150:5 Praise him upon the loud cymbals: praise him upon the high sounding cymbals.
\vs Psa 150:6 Let every thing that hath breath praise the \textsc{Lord}. Praise ye the \textsc{Lord}.
