\bibbook{the epistle of paul to\\\bibtitlemain{titus}}{titus}{Titus}{Tit}
\vs Tit 1:1 Paul, a servant of God, and an apostle of Jesus Christ, according to the faith of God's elect, and the acknowledging of the truth which is after godliness;
\vs Tit 1:2 In hope of eternal life, which God, that cannot lie, promised before the world began;
\fn{In: or, For.}
\vs Tit 1:3 But hath in due times manifested his word through preaching, which is committed unto me according to the commandment of God our Saviour;
\vs Tit 1:4 To Titus, \bibemph{mine} own son after the common faith: Grace, mercy, \bibemph{and} peace, from God the Father and the Lord Jesus Christ our Saviour.
\vs Tit 1:5 For this cause left I thee in Crete, that thou shouldest set in order the things that are wanting, and ordain elders in every city, as I had appointed thee:
\fn{wanting: or, left undone.}
\vs Tit 1:6 If any be blameless, the husband of one wife, having faithful children not accused of riot or unruly.
\vs Tit 1:7 For a bishop must be blameless, as the steward of God; not selfwilled, not soon angry, not given to wine, no striker, not given to filthy lucre;
\vs Tit 1:8 But a lover of hospitality, a lover of good men, sober, just, holy, temperate;
\fn{men: or, things.}
\vs Tit 1:9 Holding fast the faithful word as he hath been taught, that he may be able by sound doctrine both to exhort and to convince the gainsayers.
\fn{as...: or, in teaching.}
\vs Tit 1:10 For there are many unruly and vain talkers and deceivers, specially they of the circumcision:
\vs Tit 1:11 Whose mouths must be stopped, who subvert whole houses, teaching things which they ought not, for filthy lucre's sake.
\vs Tit 1:12 One of themselves, \bibemph{even} a prophet of their own, said, The Cretians \bibemph{are} alway liars, evil beasts, slow bellies.
\vs Tit 1:13 This witness is true. Wherefore rebuke them sharply, that they may be sound in the faith;
\vs Tit 1:14 Not giving heed to Jewish fables, and commandments of men, that turn from the truth.
\vs Tit 1:15 Unto the pure all things \bibemph{are} pure: but unto them that are defiled and unbelieving \bibemph{is} nothing pure; but even their mind and conscience is defiled.
\vs Tit 1:16 They profess that they know God; but in works they deny \bibemph{him}, being abominable, and disobedient, and unto every good work reprobate.
\fn{reprobate: or, void of judgment.}
\vs Tit 2:1 But speak thou the things which become sound doctrine:
\vs Tit 2:2 That the aged men be sober, grave, temperate, sound in faith, in charity, in patience.
\fn{sober: or, vigilant.}
\vs Tit 2:3 The aged women likewise, that \bibemph{they be} in behaviour as becometh holiness, not false accusers, not given to much wine, teachers of good things;
\fn{holiness: or, holy women. false...: or, one who foments strife.}
\vs Tit 2:4 That they may teach the young women to be sober, to love their husbands, to love their children,
\fn{sober: or, wise.}
\vs Tit 2:5 \bibemph{To be} discreet, chaste, keepers at home, good, obedient to their own husbands, that the word of God be not blasphemed.
\vs Tit 2:6 Young men likewise exhort to be sober minded.
\fn{sober...: or, discreet.}
\vs Tit 2:7 In all things shewing thyself a pattern of good works: in doctrine \bibemph{shewing} uncorruptness, gravity, sincerity,
\vs Tit 2:8 Sound speech, that cannot be condemned; that he that is of the contrary part may be ashamed, having no evil thing to say of you.
\vs Tit 2:9 \bibemph{Exhort} servants to be obedient unto their own masters, \bibemph{and} to please \bibemph{them} well in all \bibemph{things}; not answering again;
\fn{answering...: or, gainsaying.}
\vs Tit 2:10 Not purloining, but shewing all good fidelity; that they may adorn the doctrine of God our Saviour in all things.
\vs Tit 2:11 For the grace of God that bringeth salvation hath appeared to all men,
\fn{hath...: or, to all men, hath appeared.}
\vs Tit 2:12 Teaching us that, denying ungodliness and worldly lusts, we should live soberly, righteously, and godly, in this present world;
\vs Tit 2:13 Looking for that blessed hope, and the glorious appearing of the great God and our Saviour Jesus Christ;
\fn{glorious...: Gr. the appearance of the glory of the great God, and of our Saviour Jesus Christ.}
\vs Tit 2:14 Who gave himself for us, that he might redeem us from all iniquity, and purify unto himself a peculiar people, zealous of good works.
\vs Tit 2:15 These things speak, and exhort, and rebuke with all authority. Let no man despise thee.
\vs Tit 3:1 Put them in mind to be subject to principalities and powers, to obey magistrates, to be ready to every good work,
\vs Tit 3:2 To speak evil of no man, to be no brawlers, \bibemph{but} gentle, shewing all meekness unto all men.
\vs Tit 3:3 For we ourselves also were sometimes foolish, disobedient, deceived, serving divers lusts and pleasures, living in malice and envy, hateful, \bibemph{and} hating one another.
\vs Tit 3:4 But after that the kindness and love of God our Saviour toward man appeared,
\fn{love: or, pity.}
\vs Tit 3:5 Not by works of righteousness which we have done, but according to his mercy he saved us, by the washing of regeneration, and renewing of the Holy Ghost;
\vs Tit 3:6 Which he shed on us abundantly through Jesus Christ our Saviour;
\fn{abundantly: Gr. richly.}
\vs Tit 3:7 That being justified by his grace, we should be made heirs according to the hope of eternal life.
\vs Tit 3:8 \bibemph{This is} a faithful saying, and these things I will that thou affirm constantly, that they which have believed in God might be careful to maintain good works. These things are good and profitable unto men.
\vs Tit 3:9 But avoid foolish questions, and genealogies, and contentions, and strivings about the law; for they are unprofitable and vain.
\vs Tit 3:10 A man that is an heretick after the first and second admonition reject;
\vs Tit 3:11 Knowing that he that is such is subverted, and sinneth, being condemned of himself.
\vs Tit 3:12 When I shall send Artemas unto thee, or Tychicus, be diligent to come unto me to Nicopolis: for I have determined there to winter.
\vs Tit 3:13 Bring Zenas the lawyer and Apollos on their journey diligently, that nothing be wanting unto them.
\enlargethispage*{\baselineskip}
\vs Tit 3:14 And let ours also learn to maintain good works for necessary uses, that they be not unfruitful.
\fn{maintain...: or, profess honest trades.}
\vs Tit 3:15 All that are with me salute thee. Greet them that love us in the faith. Grace \bibemph{be} with you all. Amen. It was written to Titus, ordained the first bishop of the church of the Cretians, from Nicopolis of Macedonia.
